### 2014-09-20 CO_drift.tex
update

%\documentstyle[aas2pp4,epsf]{article}
%%%\documentstyle[aaspp4,epsf]{article}
%\documentstyle[12pt,aasms]{article}    % this is for a preprint
%(single-spaced)
%\documentstyle[aaspp4,epsf]{article} % this is for small print
%\documentstyle[12pt, aaspp4]{article}

%\documentstyle[11pt,aaspp]{article}
\documentclass[12pt, preprint]{aastex}

%\documentclass[manuscript]{aastex}
%\documentclass[apj]{emulateapj}

%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
%\documentstyle[12pt,aasms]{article}    % this is for submittal
                                       % (double-spaced)

%\documentstyle[12pt,aasms]{article}   \usepackage{emulateapj5}

\usepackage{graphicx} 
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{subeqnarray}
\usepackage{ulem}
%\bibliographystyle{apj}

\newcommand{\delad}{\nabla_{\rm ad}}
\newcommand{\delrad}{\nabla_{\rm rad}}
\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}

\newcommand{\ie}{i.e.\ }
\newcommand{\eg}{e.g.\ }
\newcommand{\p}{\partial}
\newcommand{\xv}{\vc{x}}
\newcommand{\kv}{\vc{k}}
\newcommand{\brak}[1]{\langle #1\rangle}

\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
\newcommand{\cm}{\; {\rm cm}}
\newcommand{\mm}{\; {\rm mm}}
%\newcommand{\ps}{\; {\rm s^{-1}}}
\newcommand{\km}{\; {\rm km}}
%\newcommand{\au}{\; \varpi_{\rm AU}}

\newcommand{\AU}{\; {\rm AU}}
\newcommand{\yr}{\; {\rm yr}}
\def\K{\; {\rm K}}

\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
\newcommand{\nab}{\vc{\nabla}}
\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}

\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
\newcommand{\Eqsss}[3]{Equations (\ref{#1}), (\ref{#2}) and~(\ref{#3})}
\newcommand{\App}[1]{Appendix~\ref{#1}}
\newcommand{\Sec}[1]{Sect.~\ref{#1}}
\newcommand{\Chap}[1]{Chapter~\ref{#1}}
\newcommand{\Fig}[1]{Fig.~\ref{#1}}
\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
\newcommand{\Tab}[1]{Table \ref{#1}}

\newenvironment{packed_item}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

%\newcommand{\delad}{\nabla_{\rm ad}}
%\newcommand{\delrad}{\nabla_{\rm rad}}
\newcommand{\Rg}{\mathcal{R}}
\newcommand{\RB}{R_{\rm B}}
\newcommand{\co}{_{\rm c}}
\newcommand{\di}{_{\rm d}}
\newcommand{\cb}{_{\rm RCB}}
\newcommand{\surf}{_M}
\newcommand{\mc}{m_{\rm c \oplus}}
\newcommand{\mcn}[1] { m_{ \rm c #1 \oplus} }
\newcommand{\MC}{M_{\rm crit}}
\newcommand{\au}{a_\oplus}
\newcommand{\aun}[1]{ a_{#1\oplus} }

\begin{document}
\bibliographystyle{apj}

\title{The Influence of Radial Drift and Gas Accretion on the C/O Ratio in Protoplanetary Disks}

The chemical composition of giant planet atmospheres can provide important constraints on the formation of these planets, as well as on their accretion and migration history. An important signature of atmospheric chemistry in giant planets is the C/O ratio. The C/O ratio has been estimated observationally in some giant planets, and it has been found to be different than the solar C/O ratio of the host star \citep{madhu11}. One explanation for this difference was proposed by \citet{oberg11}. Their basic idea was that the main carriers of carbon and oxygen, i.e. H$_2$O, CO$_2$ and CO, have different condensation temperatures. This results in variations in the abundances of C and O in solids and gas between the snow lines of the volatiles mentioned above. Figure \ref{fig:CtoO} shows the C/O ratios in the gas and in the dust as a function of semimajor axis, for a passively irradiated disk. As a fiducial disk model we choose the prescription of \citet{chiang10} for the gas temperature and surface density profile:

\begin{eqnarray}
\label{eq:diskmodel}
T_{\rm d} = 120 \times (a/\rm AU)^{-3/7} \,\,\, \rm K \\
\Sigma_{\rm d} = 2000 \times (a/\rm AU)^{-1} \,\,\, \rm{g \, cm}^{-2},
\end{eqnarray}
where $a$ is the semimajor axis.

\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{figs/C_O_ratio_2.pdf}
%\vspace{-0.5in}
\caption{C/O ratio in gas and in dust as a function of distance, for a passively irradiated protoplanetary disk (see Equation \ref{eq:diskmodel}). After \citet{oberg11}.} %  (See text for a description of evolution to yet higher masses.) 
\label{fig:CtoO}
\end{figure}

Starting from the premise of \citet{oberg11}, our long-term goal is to investigate the various chemical and dynamical processes that take place in protoplanetary disks, and understand which of this processes are relevant for the chemical variations in the atmospheres of gas giants, and under which conditions. We begin by studying the effect of the radial drift of solids on the location and shape of snowlines in disks.

\section{Radial Drift}

Solids planetesimals in disks orbit their host star at a Keplerian velocity. The gas, on the other hand, experiences an extra pressure gradient, which causes it to rotate at a sub-Keplerian velocity. Thus solid particles experience a headwind, which tends to remove angular momentum and cause the planetesimals to spiral inwards and fall into the host star. Small planetesimals are well-coupled to the gas, so they rotate with the gas at the sub-Keplerian frequency and their drift timescales are very large compared to the lifetime of the gas disk. Conversely, large planetesimals are decoupled from the gas, thus the influence of gas drag on them is negligible and their radial drift times also exceed the disk lifetime. On the other hand, planetesimals of intermediate sizes ($\sim$0.1cm - $\sim$10 km, depending on their location and the disk temperature and surface density profile) can have short drift timescales relative to the disk lifetime. This is shown in Figure \ref{fig:drift_times} at the H$_2$O, CO$_2$ and CO snowlines, respectively. The plot also shows the desorption (or evaporation) timescale as a function of particles size at the same stellocentric distances. We thus see that, if the drift timescale is shorter than the evaporation timescale, planetesimals might drift a significant distance past a given snowline before they desorb. This therefore changes the location and shape of snowlines, and thus the C/O ratio from the profile depicted in Figure \ref{fig:CtoO}.

\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{figs/drift_timescales_betaS1_gas_acc.pdf}
%\vspace{-0.5in}
\caption{Drift timescale (solid lines), desorption timescale (dashed lines), and gas accretion timescale for $\alpha=0.01$ (dotted lines) as a function of particle size, at three representative locations: the H$_2$O snowline (in blue), the CO$_2$ snowline (in green) and the CO snowline (in red). The horizontal dashed black line shows a typical disk lifetime of 3 Myr.} %  (See text for a description of evolution to yet higher masses.) 
\label{fig:drift_times}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{figs/desorption_distance_many_new_2.pdf}
%\vspace{-0.5in}
\caption{Desorption distance as a function on initial particle distance for H$_2$O (top panel), CO$_2$ (middle panel) and CO (bottom panel), for particle size between $10^{-3}$ and $10^8$ cm. The vertical and horizontal dashed lines show the respective snowlines as calculated based on \citet{hollenbach09}. Small particles desorb instantaneously, while large particles do not either drift or desorb before disk dissipation. Intermediate size planetesimals desorb at a fixed distance that depends on the particle size, regardless of their initial location in the disk.} %  (See text for a description of evolution to yet higher masses.) 
\label{fig:drift_dist}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{figs/desorption_distance_actual_vs_estimated.pdf}
%\vspace{-0.5in}
\caption{Desorption distance estimated analytically (when the drift and desorption times are equal) versus the desorption distance calculated numerically, for H$_2$O (top panel), CO$_2$ (middle panel) and CO (bottom panel), and for different initial particle sizes. The analytic and numerical results are in good agreement.} %  (See text for a description of evolution to yet higher masses.) 
\label{fig:an_vs_num}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{figs/nH2O_no_acc.pdf}
%\vspace{-0.5in}
\caption{Abundance of water vapor with respect to Hydrogen as a function of semimajor axis, for various particle sizes. Radial drift changes both the shape and the location of the H$_2$O snowline.} %  (See text for a description of evolution to yet higher masses.) 
\label{fig:nwater_drift}
\end{figure}

We calculate self-consistently the drift and desorption distances as a function of particle size for our fiducial disk. We calculate the desorption rates following the prescription of \citet{hollenbach09}, assuming that the solids are perfect spheres and composed of a single volatile material (either H$_2$O, CO$_2$ or CO). To calculate the drift timescale, we follow the analytic prescription of \citet{chiang10}. We integrate the drift and desorption equations self-consistently, and stop the evolution at $\tau=3$ Myr, the typical lifetime of a protoplanetary disk.

Figure \ref{fig:drift_dist} shows the desorption distance as a function of the initial planetesimal location, for H$_2$O, CO$_2$ and CO particles. As expected, small planetesimals desorb instantly, while large boulders do not have enough time to either desorb or drift before the disk dissipates. We notice, however, that planetesimals with sizes between $\sim 1$ cm and $\sim 10^3$ cm, desorb at a fixed distance from the host star, regardless of their initial location in the disk. This desorption distance depends on the planetesimal size. We have found that this distance can be well approximated analytically as the radius at which the desorption and drift timescales are equal (see Figure \ref{fig:an_vs_num}). Figure \ref{fig:nwater_drift} shows the abundance of H$_2$O with respect to Hydrogen as a function of semimajor axis. We see that particles evaporate most of their mass at their desorption distance (see Figure \ref{fig:drift_dist}).

Our drift calculation shows that the location of snowlines is not fixed, but rather dependent on planetesimal size. So far in our calculations we have assumed a passively irradiated disk. However, observations have shown that young Sun-like stars actively accrete gas from the disk at typical rates $\dot{M}_{\rm gas} \sim 10^{-8}$ $M_{\odot}$/year (e.g., \citealt{hartmann06}). The abundance of H$_2$O, CO$_2$ and CO in gas and grains, and therefore the C/O ratio in gas, will depend on the relative mass accretion rates of solids (due to drift) and of gas (due to viscosity).

For simplicity we first assume a constant gas accretion rate, $\dot{M}_{\rm gas}=-2 \pi a \dot{a_{\rm g}} \Sigma_{\rm d}$ , where $\dot{a}_{\rm g}$ is the radial velocity of the gas and $\Sigma_{\rm d}$ is given by Equation (\ref{eq:diskmodel}). The mass accretion rate of the drifting solids can be calculated as $\dot{M}_{\rm solids} = - 2 \pi a \dot{a}_{\rm drift} \Sigma_{\rm p}$, where $\dot{a}_{\rm drift}$ is the drift velocity of planetesimals (see \citealt{chiang10}, Equation 13 for an expression) and $\Sigma_{\rm p}$ is the surface density of solids. We assume a dust to gas ratio of 0.01, i.e. $\Sigma_{\rm d}=100 \Sigma_{\rm p}$.

\textbf{This is where it gets tricky, because my plot is obviously wrong (see below). Basically, I am starting with the simplest case: I'm assuming I have only one H$_2$O particle size (in this case $s=1$ cm), at an initial distance of 10 AU. My thinking was the following: the ratio between the abundance of H$_2$O in solids and the gas abundance in the disk should just be given by the ratio between the mass fluxes of solids and gas. So for a given $a$ and $s$, $n_{\rm H_2 O, solids}/n_{\rm H} = \dot{M}_{\rm solids}/\dot{M}_{\rm gas}$. Then $n_{\rm H_2 O, gas}/n_{\rm H} = 1-n_{\rm H_2 O, solids}/n_{\rm H}=1-\dot{M}_{\rm solids}/\dot{M}_{\rm gas}$. Then I multiply this by the total water abundance with respect to Hydrogen, i.e. $0.9 \times 10^{-4}$ from \citet{oberg11}. However, I am pretty sure this is wrong. First off, I don't get the results from Figure \ref{fig:nwater_drift} for $\dot{M}_{\rm gas}=0$. Figure \ref{fig:wrong} is also wrong, starting with the fact that $n_{\rm H_2 O} \neq 0$ at 10 AU, as it should. Hopefully we can discuss this during our Skype call --- I'm fairly sure it shouldn't be too complicated and I'm missing something obvious, but haven't figured out what so far.}

\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{figs/nH2O_acc_wrong.pdf}
%\vspace{-0.5in}
\caption{Attempt to show the abundance of water vapor with respect to Hydrogen as a function of semimajor axis for one particle size, when both drift and gas accretion are included. Obviously wrong.} %  (See text for a description of evolution to yet higher masses.) 
\label{fig:wrong}
\end{figure}

\section{TO DO NEXT}

\begin{itemize}
\item Calculate C/O ratios assuming a particle size distribution --- $n \sim s^{-2.5}$ seems reasonable, but see Til's 2011/2012 papers for better estimates.
\item Add particle growth into the drift calculation --- see Til's 2012 paper for an analytic estimate of growth rates; in principle it should be relatively straightforward, just need to add an extra differential equation in the drift + desorption evolution.
\item Calculate abundances and C/O ratios for different disk models; perhaps also use the temperature profile for an actively accreting disk, for consistency with the assumption of constant $\dot{M}_{\rm gas}$?
\item Discuss turbulence --- we had decided through qualitative arguments that it shouldn't matter, but we need to formulate it more rigorously.
\item Perhaps show snapshots of C/O ratio plots at different times in the disk evolution, not just after gas dissipation at 3 Myrs?
\item Discuss how good the sphere approximation is for the grains --- see the paper links that Til sent us.
\item Include also $\rm N_2$? Might be too much for this paper... 
\end{itemize}

\bibliographystyle{apj}
\bibliography{refs}

\end{document}

### 2014-09-23 python/drift_timescales_varying_sigma.py
update

@@ -13,7 +13,7 @@ def Tdisk(r, T0 = 120, betaT = 3./7):
      
 def Sigmadisk(r, Sigma0 = 2200, betaS = 3./2):
     """Disk gas surface density in g cm^-2 with r in AU"""
-    return Sigma0 * r**(-betaS)
+         return Sigma0 * r**(-betaS)
     
 def cdisk(r, T0 = 120, betaT = 3./7, mu = 2.35):
     """Sound speed in cm s^-1 with r in AU"""
@@ -33,14 +33,14 @@ def Hdisk(r, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun):
     
 def rhodisk(r, Sigma0 = 2200, betaS = 3./2, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun):
     """Disk gas density in g cm^-3 with r in AU"""
-    return Sigmadisk(r, Sigma0, betaS) / \
+    return Sigmadisk(r, Sigma0, betaS, dt) / \
         (np.sqrt(2 * np.pi) * Hdisk(r, T0, betaT, mu, Mstar))
         
 def lambdamfp(r, Sigma0 = 2200, betaS = 3./2, T0 = 120, betaT = 3./7, mu = 2.35, \
-    Mstar = Msun, sigma = 2 * 10**(-15)):
+    Mstar = Msun, sigma = 2 * 10**(-15), dt = 0):
     """Mean free path in cm with r in AU"""
     return 1 / (np.sqrt(2) * sigma * \
-        (rhodisk(r, Sigma0, betaS, T0, betaT, mu, Mstar) / (mu * mp)))
+        (rhodisk(r, Sigma0, betaS, T0, betaT, mu, Mstar, dt) / (mu * mp)))
 
 ##def vrel(r, s, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, Sigma0 = 2200, betaS = 3./2, \
 ##    Mstar = Msun, sigma = 2 * 10**(-15), eps = 0):
@@ -73,23 +73,23 @@ def lambdamfp(r, Sigma0 = 2200, betaS = 3./2, T0 = 120, betaT = 3./7, mu = 2.35,
 ####                 cdisk(r, T0, betaT, mu) * lambdamfp(r, Sigma0, betaS, T0, betaT, mu, Mstar, sigma))
 
 
-def n(betaS = 3./2, betaT = 3./7):
+def n(betaS = 3./2, betaT = 3./7, dt = 0):
[…]

### 2014-09-23 python/drift_timescales.py
update

@@ -87,7 +87,7 @@ def ts(r, s, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, Sigma0 = 2200, betaS
     """Stopping time in seconds with r in AU and s in cm"""
 
     if eps == 0:
-         if s <= 9 * lambdamfp(r, Sigma0, betaS, T0, betaT, mu) / 4:
+         if s <= 9 * lambdamfp(r, Sigma0, betaS, T0, betaT, mu, Mstar, sigma) / 4:
 
             return rhos * s / (rhodisk(r, Sigma0, betaS, T0, betaT, mu, Mstar) * \
                  vth(r, T0, betaT, mu))
@@ -104,7 +104,7 @@ def ts(r, s, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, Sigma0 = 2200, betaS
                   vrel = np.sqrt(vroverr**2 + vphioverr**2)
 
                   Re = 4 * vrel * s / \
-                        (lambdamfp(r, Sigma0, betaS, T0, betaT, mu) * vth(r, T0, betaT, mu))
+                        (lambdamfp(r, Sigma0, betaS, T0, betaT, mu, Mstar, sigma) * vth(r, T0, betaT, mu))
 
                   CD = 24.0/Re * (1.0+0.27*Re)**0.43 + 0.47 * (1.0 - np.exp(-0.04 * Re**0.38))
              

### 2014-09-23 python/drift_timescales_varying_sigma.py

@@ -92,7 +92,7 @@ def ts(r, s, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, Sigma0 = 2200, betaS
     """Stopping time in seconds with r in AU and s in cm"""
 
     if eps == 0:
-         if s <= 9 * lambdamfp(r, Sigma0, betaS, T0, betaT, mu, Sigmad) / 4:
+         if s <= 9 * lambdamfp(r, Sigma0, betaS, T0, betaT, mu, Mstar, sigma, Sigmad) / 4:
 
             return rhos * s / (rhodisk(r, Sigma0, betaS, T0, betaT, mu, Mstar, Sigmad) * \
                  vth(r, T0, betaT, mu))
@@ -109,7 +109,7 @@ def ts(r, s, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, Sigma0 = 2200, betaS
                   vrel = np.sqrt(vroverr**2 + vphioverr**2)
 
                   Re = 4 * vrel * s / \
-                        (lambdamfp(r, Sigma0, betaS, T0, betaT, mu, Sigmad) * vth(r, T0, betaT, mu))
+                        (lambdamfp(r, Sigma0, betaS, T0, betaT, mu, Mstar, sigma, Sigmad) * vth(r, T0, betaT, mu))
 
                   CD = 24.0/Re * (1.0+0.27*Re)**0.43 + 0.47 * (1.0 - np.exp(-0.04 * Re**0.38))
              
@@ -465,7 +465,7 @@ def dMdot(r1, r2, s, mx, Ex, t0, tmax, n, alpha, Nx = 1e15, rhos = 3.0, T0 = 120
 
           dMgas = dMgas_dt(r1, r2, alpha, T0, Sigma0, betaT, mu, Mstar, betaS, Sigmad, Sigmad1, Sigmad2) * (t[i + 1] - t[i]) * f
           dMsol = dMsol_dt(r1, r2, s, mx, Ex, Nx, rhos, T0, betaT, mu, Sigma0, betaS, Mstar, \
-                           sigma, npts, nptsin, tin, eps, vphi, dusttogas, Sigmad, Sigmap, Sigmap1, Sigmap2) * (t[i + 1] - t[i])
[…]

### 2014-09-23 python/drift_timescales_varying_sigma.py
update

@@ -458,7 +458,8 @@ def dMdot(r1, r2, s, mx, Ex, t0, tmax, n, alpha, Nx = 1e15, rhos = 3.0, T0 = 120
      Sigmap1 = Sigmadisk(r1, Sigma0, betaS) * dusttogas
      Sigmap2 = Sigmadisk(r2, Sigma0, betaS) * dusttogas
 
-     t = np.logspace(np.log10(t0), np.log10(tmax), n)
+     t = np.linspace(t0, tmax, n)
+     #t = np.logspace(np.log10(t0), np.log10(tmax), n)
      dM_gas, dM_sol = [], []
 
      for i in range(len(t) - 1):

### 2014-10-29 python/drift_timescales_varying_sigma.py
upd

@@ -186,13 +186,15 @@ def nu(r, alpha, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun):
 
      return alpha * cdisk(r, T0, betaT, mu) * Hdisk(r, T0, betaT, mu, Mstar)
 
-def Sigmadisk_act(r, t, alpha, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun, r1 = 100 * cmperau, C = 1):
+def Sigmadisk_act(r, t, alpha, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun, r1 = 100 * cmperau, C = 4.45e18):
 
      gammad = gamma(betaT)
      nu1 = nu(r1 / cmperau, alpha, T0, betaT, mu, Mstar)
      ts = 1. / (3 * (2 - gamma(betaT))**2) * r1**2 / nu1
      T = t / ts + 1
      rtild = r * cmperau / r1
+
+     C = 
 
      return C / (3 * np.pi * nu1 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
             * np.exp(- rtild**(2 - gammad) / T)

### 2014-10-29 python/drift_timescales_varying_sigma.py
update

@@ -194,10 +194,20 @@ def Sigmadisk_act(r, t, alpha, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun, 
      T = t / ts + 1
      rtild = r * cmperau / r1
 
-     C = 
-
      return C / (3 * np.pi * nu1 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
             * np.exp(- rtild**(2 - gammad) / T)
+
+def vacc_act(r, t, alpha, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun, r1 = 100 * cmperau, C = 4.45e18):
+    
+    gammad = gamma(betaT)
+    nu1 = nu(r1 / cmperau, alpha, T0, betaT, mu, Mstar)
+    ts = 1. / (3 * (2 - gamma(betaT))**2) * r1**2 / nu1
+    T = t / ts + 1
+    rtild = r * cmperau / r1
+
+    return 3 * alpha * kb * r**(-betaT) * rtild**(-1 - gammad) \
+            * (-cmperau**2 * (-2 + gammad) * r**2 + (-2 + betaT + gammad) * rtild**gammad * r1**2 * T) * T0 / \
+                    (mp * mu * Omegak(r, Mstar) * r1**3 * T)
 
      
 

### 2014-10-29 python/drift_timescales_varying_sigma_3.py
upd

@@ -1,0 +1,534 @@
+from utils.constants import G, kb, mp, Msun, cmperau
+import numpy as np
+from T_freeze import T_freeze, Rdes, tevap, vib_freq
+from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
+from scipy.integrate import odeint
+from scipy.optimize import brentq
+from scipy.interpolate import interp1d
+from utils.zbrac import zbrac
+
+def Tdisk(r, T0 = 120, betaT = 3./7):
+     """Disk temperature in K with r in AU"""
+     return T0 * r**(-betaT) 
+     
+def gamma(betaT = 3./7):
+     
+     return -betaT + 3./2
+
+def cdisk(r, T0 = 120, betaT = 3./7, mu = 2.35):
+    """Sound speed in cm s^-1 with r in AU"""
+    return np.sqrt(kb * Tdisk(r, T0, betaT) / (mu * mp))
+
+def vth(r, T0 = 120, betaT = 3./7, mu = 2.35):
+    """Mean thermal velocity for a Maxwellian distribution in cm s^-1 with r in AU"""
+    return np.sqrt(8 / np.pi) * cdisk(r, T0, betaT, mu)
+    
+def Omegak(r, Mstar = Msun):
+    """Keplerian angular frequency in s^-1 with r in AU"""
+    return np.sqrt(G * Mstar / (r * cmperau)**3)
+
+def Hdisk(r, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun):
+    """Disk scale height in cm with r in AU"""
+    return cdisk(r, T0, betaT, mu) / Omegak(r, Mstar)
+
+
+def nu(r, alpha, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun):
+
+     return alpha * cdisk(r, T0, betaT, mu) * Hdisk(r, T0, betaT, mu, Mstar)
+
+def Sigmadisk_act(r, t, alpha, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun, r1 = 100 * cmperau, C = 4.45e18):
+
+     gammad = gamma(betaT)
+     nu1 = nu(r1 / cmperau, alpha, T0, betaT, mu, Mstar)
+     ts = 1. / (3 * (2 - gamma(betaT))**2) * r1**2 / nu1
+     T = t / ts + 1
+     rtild = r * cmperau / r1
+
[…]

### 2014-10-29 python/drift_timescales_varying_sigma_3.py
update

@@ -63,8 +63,7 @@ def vacc_act(r, t, alpha, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun, r1 = 
     
 def rhodisk(r, t, alpha, T0 = 120, betaT = 3./7, mu = 2.35, Mstar = Msun, r1 = 100 * cmperau, C = 4.45e18):
     """Disk gas density in g cm^-3 with r in AU"""
-  
-     return Sigmadisk_act(r, t, alpha, T0, betaT, mu, Mstar, r1) / \
+    return Sigmadisk_act(r, t, alpha, T0, betaT, mu, Mstar, r1) / \
              (np.sqrt(2 * np.pi) * Hdisk(r, T0, betaT, mu, Mstar))
 
         
@@ -130,16 +129,15 @@ def taus(r, t, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, r1 = 100
 def tr(r, t, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, r1 = 100 * cmperau, C = 4.45e18, \
     Mstar = Msun, sigma = 2 * 10**(-15)):
     """Radial drift time in seconds with r in AU and s in cm"""
-
-   return (1 + taus(r, t, s, alpha, rhos, T0, betaT, mu, r1, C, Mstar, sigma)**2)/ \
+    return (1 + taus(r, t, s, alpha, rhos, T0, betaT, mu, r1, C, Mstar, sigma)**2)/ \
            taus(r, t, s, alpha, rhos, T0, betaT, mu, r1, C, Mstar, sigma) / \
            (2 * eta(r, T0, betaT, mu, Mstar) * Omegak(r, Mstar))
 
 
-def rdot(r, s, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, Sigma0 = 2200, betaS = 3./2, \
-    Mstar = Msun, sigma = 2 * 10**(-15), eps = 0, vphi = 0, Sigmad = 0):
+def rdot(r, t, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, r1 = 100 * cmperau, C = 4.45e18, \
+    Mstar = Msun, sigma = 2 * 10**(-15)):
 
-     return - (r * cmperau) / tr(r, s, rhos, T0, betaT, mu, Sigma0, betaS, Mstar, sigma, eps, vphi, Sigmad)
[…]

### 2014-10-29 tex/CO_drift.tex
update

--- 2014-09-20 CO_drift.tex
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figs/C_O_ratio_2.pdf}
+\includegraphics[width=0.8\textwidth]{../figs/C_O_ratio_2.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figs/drift_timescales_betaS1_gas_acc.pdf}
+\includegraphics[width=0.8\textwidth]{../figs/drift_timescales_betaS1_gas_acc.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figs/desorption_distance_many_new_2.pdf}
+\includegraphics[width=0.5\textwidth]{../figs/desorption_distance_many_new_2.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figs/desorption_distance_actual_vs_estimated.pdf}
+\includegraphics[width=0.5\textwidth]{../figs/desorption_distance_actual_vs_estimated.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figs/nH2O_no_acc.pdf}
+\includegraphics[width=0.8\textwidth]{../figs/nH2O_no_acc.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figs/nH2O_acc_wrong.pdf}
+\includegraphics[width=0.8\textwidth]{../figs/nH2O_acc_wrong.pdf}
 %\vspace{-0.5in}

### 2014-10-29 python/drift_timescales_varying_sigma_3.py

@@ -165,11 +165,14 @@ def tdes(mx, Ex, Tx, s, Nx = 1e15, rhos = 3.0):
 
 
 def Mdot_solids(r, t, s, alpha, Sigmap, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, r1 = 100 * cmperau, C = 4.45e18, \
-    Mstar = Msun, sigma = 2 * 10**(-15)):
+    Mstar = Msun, sigma = 2 * 10**(-15), vrw = 0):
+    
+    if vrw == 0:
+        v = - rdot(r, t, s, alpha, rhos, T0, betaT, mu, r1, C, Mstar, sigma)
+    else:
+        v = - rdot_with_acc(r, t, s, alpha, rhos, T0, betaT, mu, r1, C, Mstar, sigma)
 
-     v = - rdot(r, t, s, alpha, rhos, T0, betaT, mu, r1, C, Mstar, sigma)
-
-     return v * Sigmap * 2 * np.pi * (r * cmperau) #/ (Msun / (365 * 24 * 3600))
+    return v * Sigmap * 2 * np.pi * (r * cmperau) #/ (Msun / (365 * 24 * 3600))
 
 def Mdot_gas(r, t, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, r1 = 100 * cmperau, C = 4.45e18, \
     Mstar = Msun, sigma = 2 * 10**(-15)):
@@ -187,15 +190,15 @@ def dMgas_dt(rin, rout, t, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.
           
 
 def dMsol_dt(rin, rout, t, s, alpha, Sigmap_in, Sigmap_out, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, r1 = 100 * cmperau, C = 4.45e18, \
-    Mstar = Msun, sigma = 2 * 10**(-15)):
+    Mstar = Msun, sigma = 2 * 10**(-15), vrw = 0):
     
-    return Mdot_solids(rout, t, s, alpha, Sigmap_out, rhos, T0, betaT, mu, r1, C, Mstar, sigma) - \
-            Mdot_solids(rin, t, s, alpha, Sigmap_in, rhos, T0, betaT, mu, r1, C, Mstar, sigma)
+    return Mdot_solids(rout, t, s, alpha, Sigmap_out, rhos, T0, betaT, mu, r1, C, Mstar, sigma, vrw) - \
[…]

### 2014-12-01 tex/thesis_proposal_outline.tex
update

--- 2014-09-20 CO_drift.tex
@@
 %\documentstyle[11pt,aaspp]{article}
-\documentclass[12pt, preprint]{aastex}
+%\documentclass[12pt, preprint]{aastex}
 
-%\documentclass[manuscript]{aastex}
+\documentclass[manuscript]{aastex}
 %\documentclass[apj]{emulateapj}
@@
 \title{The Influence of Radial Drift and Gas Accretion on the C/O Ratio in Protoplanetary Disks}
-
-The chemical composition of giant planet atmospheres can provide important constraints on the formation of these planets, as well as on their accretion and migration history.
-An important signature of atmospheric chemistry in giant planets is the C/O ratio.
-The C/O ratio has been estimated observationally in some giant planets, and it has been found to be different than the solar C/O ratio of the host star \citep{madhu11}.
-One explanation for this difference was proposed by \citet{oberg11}.
-Their basic idea was that the main carriers of carbon and oxygen, i.e.
-H$_2$O, CO$_2$ and CO, have different condensation temperatures.
-This results in variations in the abundances of C and O in solids and gas between the snow lines of the volatiles mentioned above.
-Figure \ref{fig:CtoO} shows the C/O ratios in the gas and in the dust as a function of semimajor axis, for a passively irradiated disk.
-As a fiducial disk model we choose the prescription of \citet{chiang10} for the gas temperature and surface density profile:
-
-\begin{eqnarray}
-\label{eq:diskmodel}
-T_{\rm d} = 120 \times (a/\rm AU)^{-3/7} \,\,\, \rm K \\
-\Sigma_{\rm d} = 2000 \times (a/\rm AU)^{-1} \,\,\, \rm{g \, cm}^{-2},
-\end{eqnarray}
-where $a$ is the semimajor axis.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.8\textwidth]{figs/C_O_ratio_2.pdf}
-%\vspace{-0.5in}
-\caption{C/O ratio in gas and in dust as a function of distance, for a passively irradiated protoplanetary disk (see Equation \ref{eq:diskmodel}).
-After \citet{oberg11}.} %  (See text for a description of evolution to yet higher masses.)
-\label{fig:CtoO}
-\end{figure}
-
-Starting from the premise of \citet{oberg11}, our long-term goal is to investigate the various chemical and dynamical processes that take place in protoplanetary disks, and understand which of this processes are relevant for the chemical variations in the atmospheres of gas giants, and under which conditions.
-We begin by studying the effect of the radial drift of solids on the location and shape of snowlines in disks.
-
-\section{Radial Drift}
-
-Solids planetesimals in disks orbit their host star at a Keplerian velocity.
-The gas, on the other hand, experiences an extra pressure gradient, which causes it to rotate at a sub-Keplerian velocity.
-Thus solid particles experience a headwind, which tends to remove angular momentum and cause the planetesimals to spiral inwards and fall into the host star.
-Small planetesimals are well-coupled to the gas, so they rotate with the gas at the sub-Keplerian frequency and their drift timescales are very large compared to the lifetime of the gas disk.
-Conversely, large planetesimals are decoupled from the gas, thus the influence of gas drag on them is negligible and their radial drift times also exceed the disk lifetime.
-On the other hand, planetesimals of intermediate sizes ($\sim$0.1cm - $\sim$10 km, depending on their location and the disk temperature and surface density profile) can have short drift timescales relative to the disk lifetime.
-This is shown in Figure \ref{fig:drift_times} at the H$_2$O, CO$_2$ and CO snowlines, respectively.
-The plot also shows the desorption (or evaporation) timescale as a function of particles size at the same stellocentric distances.
-We thus see that, if the drift timescale is shorter than the evaporation timescale, planetesimals might drift a significant distance past a given snowline before they desorb.
-This therefore changes the location and shape of snowlines, and thus the C/O ratio from the profile depicted in Figure \ref{fig:CtoO}.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.8\textwidth]{figs/drift_timescales_betaS1_gas_acc.pdf}
-%\vspace{-0.5in}
-\caption{Drift timescale (solid lines), desorption timescale (dashed lines), and gas accretion timescale for $\alpha=0.01$ (dotted lines) as a function of particle size, at three representative locations: the H$_2$O snowline (in blue), the CO$_2$ snowline (in green) and the CO snowline (in red).
-The horizontal dashed black line shows a typical disk lifetime of 3 Myr.} %  (See text for a description of evolution to yet higher masses.)
-\label{fig:drift_times}
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.5\textwidth]{figs/desorption_distance_many_new_2.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance as a function on initial particle distance for H$_2$O (top panel), CO$_2$ (middle panel) and CO (bottom panel), for particle size between $10^{-3}$ and $10^8$ cm.
-The vertical and horizontal dashed lines show the respective snowlines as calculated based on \citet{hollenbach09}.
-Small particles desorb instantaneously, while large particles do not either drift or desorb before disk dissipation.
-Intermediate size planetesimals desorb at a fixed distance that depends on the particle size, regardless of their initial location in the disk.} %  (See text for a description of evolution to yet higher masses.)
-\label{fig:drift_dist}
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.5\textwidth]{figs/desorption_distance_actual_vs_estimated.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance estimated analytically (when the drift and desorption times are equal) versus the desorption distance calculated numerically, for H$_2$O (top panel), CO$_2$ (middle panel) and CO (bottom panel), and for different initial particle sizes.
-The analytic and numerical results are in good agreement.} %  (See text for a description of evolution to yet higher masses.)
-\label{fig:an_vs_num}
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.8\textwidth]{figs/nH2O_no_acc.pdf}
-%\vspace{-0.5in}
-\caption{Abundance of water vapor with respect to Hydrogen as a function of semimajor axis, for various particle sizes.
-Radial drift changes both the shape and the location of the H$_2$O snowline.} %  (See text for a description of evolution to yet higher masses.)
-\label{fig:nwater_drift}
-\end{figure}
-
-We calculate self-consistently the drift and desorption distances as a function of particle size for our fiducial disk.
-We calculate the desorption rates following the prescription of \citet{hollenbach09}, assuming that the solids are perfect spheres and composed of a single volatile material (either H$_2$O, CO$_2$ or CO).
-To calculate the drift timescale, we follow the analytic prescription of \citet{chiang10}.
-We integrate the drift and desorption equations self-consistently, and stop the evolution at $\tau=3$ Myr, the typical lifetime of a protoplanetary disk.
-
-Figure \ref{fig:drift_dist} shows the desorption distance as a function of the initial planetesimal location, for H$_2$O, CO$_2$ and CO particles.
-As expected, small planetesimals desorb instantly, while large boulders do not have enough time to either desorb or drift before the disk dissipates.
-We notice, however, that planetesimals with sizes between $\sim 1$ cm and $\sim 10^3$ cm, desorb at a fixed distance from the host star, regardless of their initial location in the disk.
-This desorption distance depends on the planetesimal size.
-We have found that this distance can be well approximated analytically as the radius at which the desorption and drift timescales are equal (see Figure \ref{fig:an_vs_num}).
-Figure \ref{fig:nwater_drift} shows the abundance of H$_2$O with respect to Hydrogen as a function of semimajor axis.
-We see that particles evaporate most of their mass at their desorption distance (see Figure \ref{fig:drift_dist}).
-
-Our drift calculation shows that the location of snowlines is not fixed, but rather dependent on planetesimal size.
-So far in our calculations we have assumed a passively irradiated disk.
-However, observations have shown that young Sun-like stars actively accrete gas from the disk at typical rates $\dot{M}_{\rm gas} \sim 10^{-8}$ $M_{\odot}$/year (e.g., \citealt{hartmann06}).
-The abundance of H$_2$O, CO$_2$ and CO in gas and grains, and therefore the C/O ratio in gas, will depend on the relative mass accretion rates of solids (due to drift) and of gas (due to viscosity).
-
-For simplicity we first assume a constant gas accretion rate, $\dot{M}_{\rm gas}=-2 \pi a \dot{a_{\rm g}} \Sigma_{\rm d}$ , where $\dot{a}_{\rm g}$ is the radial velocity of the gas and $\Sigma_{\rm d}$ is given by Equation (\ref{eq:diskmodel}).
-The mass accretion rate of the drifting solids can be calculated as $\dot{M}_{\rm solids} = - 2 \pi a \dot{a}_{\rm drift} \Sigma_{\rm p}$, where $\dot{a}_{\rm drift}$ is the drift velocity of planetesimals (see \citealt{chiang10}, Equation 13 for an expression) and $\Sigma_{\rm p}$ is the surface density of solids.
-We assume a dust to gas ratio of 0.01, i.e.
-$\Sigma_{\rm d}=100 \Sigma_{\rm p}$.
-
-\textbf{This is where it gets tricky, because my plot is obviously wrong (see below).
-Basically, I am starting with the simplest case: I'm assuming I have only one H$_2$O particle size (in this case $s=1$ cm), at an initial distance of 10 AU.
-My thinking was the following: the ratio between the abundance of H$_2$O in solids and the gas abundance in the disk should just be given by the ratio between the mass fluxes of solids and gas.
-So for a given $a$ and $s$, $n_{\rm H_2 O, solids}/n_{\rm H} = \dot{M}_{\rm solids}/\dot{M}_{\rm gas}$.
-Then $n_{\rm H_2 O, gas}/n_{\rm H} = 1-n_{\rm H_2 O, solids}/n_{\rm H}=1-\dot{M}_{\rm solids}/\dot{M}_{\rm gas}$.
-Then I multiply this by the total water abundance with respect to Hydrogen, i.e.
-$0.9 \times 10^{-4}$ from \citet{oberg11}.
-However, I am pretty sure this is wrong.
-First off, I don't get the results from Figure \ref{fig:nwater_drift} for $\dot{M}_{\rm gas}=0$.
-Figure \ref{fig:wrong} is also wrong, starting with the fact that $n_{\rm H_2 O} \neq 0$ at 10 AU, as it should.
-Hopefully we can discuss this during our Skype call --- I'm fairly sure it shouldn't be too complicated and I'm missing something obvious, but haven't figured out what so far.}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.8\textwidth]{figs/nH2O_acc_wrong.pdf}
-%\vspace{-0.5in}
-\caption{Attempt to show the abundance of water vapor with respect to Hydrogen as a function of semimajor axis for one particle size, when both drift and gas accretion are included.
-Obviously wrong.} %  (See text for a description of evolution to yet higher masses.)
-\label{fig:wrong}
-\end{figure}
-
-\section{TO DO NEXT}
-
-\begin{itemize}
-\item Calculate C/O ratios assuming a particle size distribution --- $n \sim s^{-2.5}$ seems reasonable, but see Til's 2011/2012 papers for better estimates.
-\item Add particle growth into the drift calculation --- see Til's 2012 paper for an analytic estimate of growth rates; in principle it should be relatively straightforward, just need to add an extra differential equation in the drift + desorption evolution.
-\item Calculate abundances and C/O ratios for different disk models; perhaps also use the temperature profile for an actively accreting disk, for consistency with the assumption of constant $\dot{M}_{\rm gas}$?
-\item Discuss turbulence --- we had decided through qualitative arguments that it shouldn't matter, but we need to formulate it more rigorously.
-\item Perhaps show snapshots of C/O ratio plots at different times in the disk evolution, not just after gas dissipation at 3 Myrs?
-\item Discuss how good the sphere approximation is for the grains --- see the paper links that Til sent us.
-\item Include also $\rm N_2$? Might be too much for this paper... 
-\end{itemize}
+\author{Ana-Maria A. Piso}
+\affil{Advisor: Karin I. \"Oberg}
+\affil{Suggested readers: Dimitar Sasselov, Sean Andrews, }
 

### 2014-12-02 tex/thesis_proposal_outline.tex
update

@@
 
-\title{The Influence of Radial Drift and Gas Accretion on the C/O Ratio in Protoplanetary Disks}
+\title{The role of disk volatile chemistry and dynamics in shaping the compositions of nascent planets}
 \author{Ana-Maria A. Piso}
 \affil{Advisor: Karin I. \"Oberg}
-\affil{Suggested readers: Dimitar Sasselov, Sean Andrews, }
+\affil{Suggested Thesis Advisory Committee members: Dimitar Sasselov, Sean Andrews, \textit{plus one of John Johnson, David Wilner, Dave Charbonneau, Matt Holman?}}
+
+\section{Introduction and Motivation}
+
+The chemical composition of giant planet atmospheres can provide important constraints on the formation of these planets, as well as on their accretion and migration history.
+An important signature of atmospheric chemistry in giant planets is the C/O ratio.
+The C/O ratio has been estimated observationally in some giant planets, and it has been found to be different than the solar C/O ratio of the host star \citep{madhu11}.
+One explanation for this difference was proposed by \citet{oberg11}.
+Their basic idea was that the main carriers of carbon and oxygen, i.e.
+H$_2$O, CO$_2$ and CO, have different condensation temperatures.
+This results in variations in the abundances of C and O in solids and gas between the snow lines of the volatiles mentioned above.
+Figure \ref{fig:CtoO} shows the C/O ratios in the gas and in the dust as a function of semimajor axis, for a passively irradiated disk.
+%As a fiducial disk model we choose the prescription of \citet{chiang10} for the gas temperature and surface density profile:
+
+%\begin{eqnarray}
+%\label{eq:diskmodel}
+%T_{\rm d} = 120 \times (a/\rm AU)^{-3/7} \,\,\, \rm K \\
+%\Sigma_{\rm d} = 2000 \times (a/\rm AU)^{-1} \,\,\, \rm{g \, cm}^{-2},
+%\end{eqnarray}
+%where $a$ is the semimajor axis.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.8\textwidth]{../figs/C_O_ratio_2.pdf}
+%\vspace{-0.5in}
+\caption{C/O ratio in gas and in dust as a function of distance, for a passively irradiated protoplanetary disk.
+After \citet{oberg11}.} %  (See text for a description of evolution to yet higher masses.)
+\label{fig:CtoO}
+\end{figure}
+
+The relative abundance of C and O in gas and dust across the protoplanetary disk is further affected by various chemical and dynamical process occurring in the disk throughout its lifetime.
+The chemical composition and distribution of the main C and O carriers evolves with time (e.g., \citealt{ciesla06}, \citealt{visser09}), as do the location and shape of the different snowlines (e.g., \citealt{garaud07}, \citealt{stevenson88}).
+Moreover, solids are redistributed throughout the protoplanetary disk due to radial drift \citep{chiang10} and planetary migration \citep{armitage10}.
+Observations have shown that young Sun-like stars actively accrete gas from the disk (e.g., \citealt{hartmann06}), which may further change the abundance of volatiles in solid and gaseous form in the disk.
+Exploring and quantifying the effect of all such processes on the atmospheric composition of giant planets is crucial for our understanding of their formation and overall evolution in the protoplanetary disk.
+
+\section{Proposed Research}
+
+In this thesis, I plan to investigate the various chemical and dynamical processes that take place in protoplanetary disks, and understand which of these processes are relevant for the chemical variations in the atmospheres of gas giants, and under which conditions.
+I aim to enhance the toy model of \citet{oberg11} by including dynamical effects, as well as a more complex chemical reaction network.
+The physical processes that I will consider, as well as the timeline of my thesis, can be summarized as follows.
+
+\subsection{Radial Drift of Solids}
+\label{sec:drift}
+
+Solid planetesimals in disks orbit their host star at a Keplerian velocity.
+The gas, on the other hand, experiences an extra pressure gradient, which causes it to rotate at a sub-Keplerian velocity.
+Thus solid particles experience a headwind, which tends to remove angular momentum and cause the planetesimals to spiral inwards and fall into the host star.
+Small planetesimals are well-coupled to the gas, so they rotate with the gas at the sub-Keplerian frequency and their drift timescales are very large compared to the lifetime of the gas disk.
+Conversely, large planetesimals are decoupled from the gas, thus the influence of gas drag on them is negligible and their radial drift times also exceed the disk lifetime.
+On the other hand, planetesimals of intermediate sizes ($\sim$0.1cm - $\sim$10 km, depending on their location and the disk temperature and surface density profile) can have short drift timescales relative to the disk lifetime.
+If  the drift timescale is shorter than the evaporation timescale, planetesimals might drift a significant distance past a given snowline before they desorb.
+This therefore changes the location and shape of snowlines, and thus the C/O ratio from the profile depicted in Figure \ref{fig:CtoO}.
+
+The calculation of \citet{oberg11} assumes a passively irradiated disk.
+However, observations have shown that young Sun-like stars actively accrete gas from the disk at typical rates $\dot{M}_{\rm gas} \sim 10^{-8}$ $M_{\odot}$/year (e.g., \citealt{hartmann06}).
+The abundance of H$_2$O, CO$_2$ and CO in gas and grains, and therefore the C/O ratio in gas, will depend on the relative mass accretion rates of solids (due to drift) and of gas (due to viscosity).
+The surface density of planetesimals (and thus the C/O ratio) is further affected by grain growth and particle fragmentation (e.g., \citealt{birnstiel12}).
+
+I will incorporate the effect of radial drift, particle growth and gas accretion, coupled with the desorption of ices, to calculate the C/O variation with semimajor axis in an actively accreting protoplanetary disk.
+\textbf{This will result in Paper I, estimated completion spring 2015}.
+
+\subsection{Nitrogen Abundance}
+\label{sec:nitrogen}
+
+Besides carbon and oxygen, nitrogen is another important volatile molecule that shapes the chemistry in protoplanetary disks.
+Nitrogen is primarily found as $\rm{N}_2$, although $\sim$$10$\% of the disk nitrogen abundance can be carried by $\rm{NH}_3$ (e.g., \citealt{lahuis00}).
+I plan to add nitrogen to the simple chemical model of section \ref{sec:drift} and apply the model summarized in section \ref{sec:drift} to study the effect of radial drift on the N/O ratio in protoplanetary disks.
+\textbf{This will result in Paper II, estimated completion summer 2015} \textit{(Is this too little content for just one paper?
+Perhaps this could be a letter?)}.
+
+\subsection{The Effect of Disk Chemistry on the Shape and Location of Snowlines}
+
+The calculations summarized in sections \ref{sec:drift} and \ref{sec:nitrogen} are based on a simplified, static chemical model.
+In reality, however, the gas-grain chemical network in protoplanetary disks is complex and evolves with time (e.g., \citealt{bergin09}, \citealt{fogel11}).
+In this part of my thesis, I plan to incorporate the detailed, time-dependent chemical reaction network developed by Merchantz et al.
+(in prep.) \textbf{(?)} in the radial drift calculation described in section \ref{sec:drift}.
+\textbf{This will result in Paper III, estimated completion late 2015 / early 2016} \textit{(Too much content for one paper?
+Should it be split into (Paper III) just the effect of evolving disk chemistry on snowline location, and (Paper IV) chemistry + drift?)}.
+
+\subsection{Dynamical Effects II: Core Accretion, Migration, Core Dredging}
+
+In this part of the thesis, I plan to incorporate further dynamical effects, such as atmospheric pollution of planetesimals during the atmospheric accretion phase, Type I and Type II migration, and core dredging (e.g., \citealt{lodders09}, \citealt{stevenson85}, \citealt{guillot04}).
+\textbf{This will result in Paper IV, estimated completion summer 2016} \textit{(Again, perhaps too little material for one paper?)}.
+
+\subsection{Potential Planet Formation Locations and Comparison with Observations}
+
+In the last part of my thesis, I plan to analyze the results obtained in the previous sections, asses the relevance of each chemical or dynamical effect considered, and combine the relevant processes to obtain a realistic framework for analyzing the relation between atmospheric chemical composition and  formation history for a given planet.
+I will then use this framework to explore a wide parameter space of initial disk conditions in order to generate model planet populations and explore potential planet formation locations.
+By comparing my results with present and future observations of atmospheric spectra of giant planets, I will be able to further constrain their formation, accretion and migration history.
+\textbf{This will result in Paper V, estimated completion spring 2017.}
+
+\section{Preliminary Results}
+
+As a first step, I investigate the effect of radial drift on the shape of snowlines in protoplanetary disks (see section \ref{sec:drift}).
+I calculate self-consistently the drift and desorption distances as a function of particle size for a passively irradiated disk.
+I calculate the desorption rates following the prescription of \citet{hollenbach09}, assuming that the solids are perfect spheres and composed of a single volatile material (either H$_2$O, CO$_2$ or CO).
+To calculate the drift timescale, I follow the analytic prescription of \citet{chiang10}.
+I integrate the drift and desorption equations self-consistently, and stop the evolution at $\tau=3$ Myr, the typical lifetime of a protoplanetary disk.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.47\textwidth]{../figs/desorption_distance_many_new_2.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance as a function on initial particle distance for H$_2$O (top panel), CO$_2$ (middle panel) and CO (bottom panel), for particle size between $10^{-3}$ and $10^8$ cm.
+The vertical and horizontal dashed lines show the respective snowlines as calculated based on \citet{hollenbach09}.
+Small particles desorb instantaneously, while large particles do not either drift or desorb before disk dissipation.
+Intermediate size planetesimals desorb at a fixed distance that depends on the particle size, regardless of their initial location in the disk.} %  (See text for a description of evolution to yet higher masses.)
+\label{fig:drift_dist}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.5\textwidth]{../figs/desorption_distance_actual_vs_estimated.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance estimated analytically (when the drift and desorption times are equal) versus the desorption distance calculated numerically, for H$_2$O (top panel), CO$_2$ (middle panel) and CO (bottom panel), and for different initial particle sizes.
+The analytic and numerical results are in good agreement.} %  (See text for a description of evolution to yet higher masses.)
+\label{fig:an_vs_num}
+\end{figure}
+
+Figure \ref{fig:drift_dist} shows the desorption distance as a function of the initial planetesimal location, for H$_2$O, CO$_2$ and CO particles.
+As expected, small planetesimals desorb instantly, while large boulders do not have enough time to either desorb or drift before the disk dissipates.
+Notice, however, that planetesimals with sizes between $\sim 1$ cm and $\sim 10^3$ cm desorb at a fixed distance from the host star, regardless of their initial location in the disk.
+This desorption distance depends on the planetesimal size.
+I have found that this distance can be well approximated analytically as the radius at which the desorption and drift timescales are equal (see Figure \ref{fig:an_vs_num}).
+%Figure \ref{fig:nwater_drift} shows the abundance of H$_2$O with respect to Hydrogen as a function of semimajor axis.
+We see that particles evaporate most of their mass at their desorption distance (see Figure \ref{fig:drift_dist}).
+
+\section{Discussion}
+
+\textbf{TBD}
 

### 2015-01-13 tex/thesis_proposal.tex
update
upd

--- 2014-12-02 tex/thesis_proposal_outline.tex
@@
 \affil{Advisor: Karin I. \"Oberg}
-\affil{Suggested Thesis Advisory Committee members: Dimitar Sasselov, Sean Andrews, \textit{plus one of John Johnson, David Wilner, Dave Charbonneau, Matt Holman?}}
+\affil{Suggested Thesis Advisory Committee members: \\ Dimitar Sasselov, Sean Andrews, Matthew Holman}
 
@@
 The chemical composition of giant planet atmospheres can provide important constraints on the formation of these planets, as well as on their accretion and migration history.
+In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11c}, \citealt{oberg11b}).
+Molecular line emissions from compounds such as CO$_2$, CO, H$_2$CO, CN, HCN, HNC etc.
+(see Table 1 from \citealt{henning13} for a more complete list) trace various chemical and physical processes in disks.
+Of particular importance are volatile molecules (i.e., compounds that have very low freezing temperatures), since the chemical composition of giant planet atmospheres strongly depends on whether such planets have formed inside or outside the snowlines of these volatiles.
+
 An important signature of atmospheric chemistry in giant planets is the C/O ratio.
@@
 This results in variations in the abundances of C and O in solids and gas between the snow lines of the volatiles mentioned above.
-Figure \ref{fig:CtoO} shows the C/O ratios in the gas and in the dust as a function of semimajor axis, for a passively irradiated disk.
+Figure \ref{fig:CtoO} shows the C/O ratios in the gas and in the dust as a function of semimajor axis, for a passive disk.
+The relative abundance of C and O in gas and dust across the protoplanetary disk is further affected by various chemical process occurring in the disk throughout its lifetime.
+The chemical composition and distribution of the main C and O carriers evolves with time (e.g., \citealt{ciesla06}, \citealt{visser09}), as do the location and shape of the different snowlines (e.g., \citealt{garaud07}, \citealt{stevenson88}).
+%The relative abundance of C and O in the atmospheres of nascent giant planets thus depends on where in the disk these planets have formed.
+
+%The model of \citet{oberg11} assumes that a static disk chemistry and
+
 %As a fiducial disk model we choose the prescription of \citet{chiang10} for the gas temperature and surface density profile:
@@
 %\vspace{-0.5in}
-\caption{C/O ratio in gas and in dust as a function of distance, for a passively irradiated protoplanetary disk.
-After \citet{oberg11}.} %  (See text for a description of evolution to yet higher masses.)
+\caption{C/O ratio in gas and in dust as a function of distance, for a passive protoplanetary disk. After \citet{oberg11}.} %  (See text for a description of evolution to yet higher masses.) 
 \label{fig:CtoO}
@@
 
-The relative abundance of C and O in gas and dust across the protoplanetary disk is further affected by various chemical and dynamical process occurring in the disk throughout its lifetime.
-The chemical composition and distribution of the main C and O carriers evolves with time (e.g., \citealt{ciesla06}, \citealt{visser09}), as do the location and shape of the different snowlines (e.g., \citealt{garaud07}, \citealt{stevenson88}).
-Moreover, solids are redistributed throughout the protoplanetary disk due to radial drift \citep{chiang10} and planetary migration \citep{armitage10}.
+In addition to the distribution of volatiles in protoplanetary disks, the compositions of planetary atmospheres are also affected by dynamical processes occurring in the disk.
+For example, solids are redistributed throughout the protoplanetary disk due to radial drift \citep{chiang10}.
 Observations have shown that young Sun-like stars actively accrete gas from the disk (e.g., \citealt{hartmann06}), which may further change the abundance of volatiles in solid and gaseous form in the disk.
-Exploring and quantifying the effect of all such processes on the atmospheric composition of giant planets is crucial for our understanding of their formation and overall evolution in the protoplanetary disk.
+
+The chemical composition and the C/O ratio in planetary atmospheres may also change in the late stages of planet formation, as well as due to planetary dynamics.
+Evaporating planetesimals may pollute the gaseous envelope during runaway atmospheric accretion and late stage accretion \citep{oberg11}.
+Other dynamical processes such as planetary migration \citep{armitage10} and core dredging (e.g., \citealt{lodders09}) may further affect the atmospheric composition of young planets.
+
+We can thus see that chemical and dynamical processes in protoplanetary disks are tightly connected and highly affect the composition of nascent planets.
+Exploring and quantifying the effect of all such processes is therefore crucial for our understanding of the formation of giant planets and their overall evolution in the protoplanetary disk.
+
+\section{Previous Research}
+
+For the past three years I worked on a different project related to planet formation, supervised by Prof.
+Ruth Murray-Clay and Dr.
+Andrew Youdin.
+The goal of my project was to determine the minimum (critical) core mass for a giant planet to form before the gas in the protoplanetary disk dissipates, across a range of semimajor axes between 5-100 AU.
+This critical core mass is typically quoted to be $\sim$$10$ $M_{\oplus}$.
+Several previous studies assume steady-state atmospheres heated by planetesimal accretion (e.g., \citealt{stevenson82}, \citealt{rafikov06}).
+In my work, I chose a complementary approach, in which planetesimal accretion is minimal and the gaseous envelope accretes gas while undergoing Kelvin-Helmholtz contraction.
+Within this framework, I calculated the critical core mass as a function of disk properties and stellocentric separation.
+I first assumed an ideal gas polytropic equation of state (EOS) for the nebular gas and an interstellar grain opacity \citep{piso14}.
+In the followup paper, \citet{piso15}, I used realistic EOS tables to calculate the critical core mass, and explored the effects of a realistic EOS and grain growth opacities on the minimum critical core mass.
+My results show that the critical core mass for giant planet formation is likely to be lower than 10 $M_{\oplus}$, and steadily decreases with stellocentric distance.
+
+This project has provided me with valuable insights into the process of giant planet formation and has given me the opportunity to extensively review the existing literature on this topic.
+I believe that the knowledge and experience that I have gained from this project have laid a solid foundation for my current research endeavors in my Ph.D.
+thesis.
+
+%and planetary migration \citep{armitage10}
 
@@
 
-\subsection{Radial Drift of Solids}
+\subsection{Radial Drift of Solids and Snowlines Location}
 \label{sec:drift}
 
-Solid planetesimals in disks orbit their host star at a Keplerian velocity.
+Solid particles in disks orbit their host star at a Keplerian velocity.
 The gas, on the other hand, experiences an extra pressure gradient, which causes it to rotate at a sub-Keplerian velocity.
-Thus solid particles experience a headwind, which tends to remove angular momentum and cause the planetesimals to spiral inwards and fall into the host star.
-Small planetesimals are well-coupled to the gas, so they rotate with the gas at the sub-Keplerian frequency and their drift timescales are very large compared to the lifetime of the gas disk.
+Thus dust grains experience a headwind, which tends to remove angular momentum and cause the solids to spiral inwards and fall into the host star.
+Small grains are well-coupled to the gas, so they rotate with the gas at the sub-Keplerian frequency and their drift timescales are very large compared to the lifetime of the gas disk.
 Conversely, large planetesimals are decoupled from the gas, thus the influence of gas drag on them is negligible and their radial drift times also exceed the disk lifetime.
 On the other hand, planetesimals of intermediate sizes ($\sim$0.1cm - $\sim$10 km, depending on their location and the disk temperature and surface density profile) can have short drift timescales relative to the disk lifetime.
-If  the drift timescale is shorter than the evaporation timescale, planetesimals might drift a significant distance past a given snowline before they desorb.
+If  the drift timescale is shorter than the evaporation timescale, solids might drift a significant distance past where a given snowline would be in a passive disk before they desorb.
 This therefore changes the location and shape of snowlines, and thus the C/O ratio from the profile depicted in Figure \ref{fig:CtoO}.
 
-The calculation of \citet{oberg11} assumes a passively irradiated disk.
+The calculation of \citet{oberg11} assumes a passive disk.
 However, observations have shown that young Sun-like stars actively accrete gas from the disk at typical rates $\dot{M}_{\rm gas} \sim 10^{-8}$ $M_{\odot}$/year (e.g., \citealt{hartmann06}).
@@
 \textbf{This will result in Paper I, estimated completion spring 2015}.
+In this part of my thesis, I will closely collaborate with Dr.
+Til Birnstiel and use some of his models for grain growth and particle drift as a starting point for my calculations.
+I also plan to discuss the theoretical aspects of this subproject with Prof.
+Ruth Murray-Clay.
 
-\subsection{Nitrogen Abundance}
+\subsubsection{Nitrogen Abundance}
 \label{sec:nitrogen}
@@
 I plan to add nitrogen to the simple chemical model of section \ref{sec:drift} and apply the model summarized in section \ref{sec:drift} to study the effect of radial drift on the N/O ratio in protoplanetary disks.
-\textbf{This will result in Paper II, estimated completion summer 2015} \textit{(Is this too little content for just one paper?
-Perhaps this could be a letter?)}.
+Since the main carries of nitrogen in disks are less well understood than those of carbon or oxygen, I will use abundance patterns both from the Solar system and from disk chemistry models (e.g., \citealt{schwarz14}) to define the range of abundances of different carries, then apply the dynamic framework developed in section \ref{sec:drift}.
+\textbf{This will result in Paper II, estimated completion summer 2015} .
 
-\subsection{The Effect of Disk Chemistry on the Shape and Location of Snowlines}
+\subsection{The Effect of Time-Dependent Disk Chemistry on the Shape and Location of Snowlines}
+\label{sec:chemical}
 
@@
 In reality, however, the gas-grain chemical network in protoplanetary disks is complex and evolves with time (e.g., \citealt{bergin09}, \citealt{fogel11}).
-In this part of my thesis, I plan to incorporate the detailed, time-dependent chemical reaction network developed by Merchantz et al.
+Comprehensive disk chemistry models have been developed by several groups, taking into account different radiation fields, dust grain sizes, disk viscosity, time-dependence and dynamics (see references in Table 3 of \citealt{henning13}).
+These studies have shown that disk chemistry is mostly regulated by disk temperature, density, stellar / interstellar radiation fields, and cosmic rays.
+Early chemical models of the Solar system were usually restricted to the inner disk regions (10-30 AU) and assumed thermodynamic equilibrium (e.g., \citealt{cameron95}).
+Due to the decrease in temperature with stellocentric distance, in more modern studies the disk chemistry is divided between the inner disk, $\lesssim$20 AU, and the outer disk beyond 20 AU \citep{henning13}.
+Due to the absence of intense sources of ionizing radiation, the chemistry in the inner disk approaches quasiequilibrium (e.g., \citealt{ilgner04}).
+In contrast, high-energy radiation and cosmic rays are key drivers of the chemistry in the outer disk (e.g., \citealt{vandishoeck}), and thus the chemistry is no longer in equilibrium.
+
+Chemical models of protoplanetary disks are based on detailed chemical kinetics models with hundreds or thousands of chemical reactions, which make them very intensive computationally.
+These models are thus usually decoupled from disk dynamics.
+Some studies include various heating and cooling processes, as well as grain evolution (e.g., \citealt{aikawa06}, \citealt{fogel11}, \citealt{vasyunin11}, \citealt{akimkin13}).
+Additionally, some chemo-dynamical models for protoplanetary disks have been developed --- for example, taking into account radial advective mass transport (e.g., \citealt{aikawa99}, \citealt{woods09}).
+\citet{semenov11} developed a 2D model using the mixing-length approximation and time-dependent chemistry, for a turbulent disk.
+They calculated two important timescales, the physical timescale, $t_{\rm phys}$, and the chemical timescale, $t_{\rm chem}$, and found that the chemical evolution is sensitive to changes in the physical conditions if $t_{\rm phys}<t_{\rm chem}$.
+
+It is thus clear that the chemical processes that occur in protoplanetary disks are very complex, and their coupling with disk dynamics is not always straightforward.
+I therefore plan to divide this part of my thesis in two steps.
+First, I will parametrize the detailed, time-dependent chemical reaction network developed by Merchantz et al.
+(in prep.) and use it in the radial drift calculation described in section \ref{sec:drift}.
+\textbf{This will result in Paper III, estimated completion fall 2015}.
+In the second step, I will self-consistently evolve the chemical model from Paper III and the disk dynamics.
+In order to couple chemistry and dynamics, I will use the tools currently being developed by Xu, Bai \& \"Oberg (in prep.) to reduce the chemical network.
+I will furthermore  collaborate closely with Ilse Cleeves, who is an expert in chemical evolution in protoplanetary disks.
+\textbf{This will result in Paper IV, estimated completion spring 2016}.
+%For this section of my thesis, I will collaborate closely with Ilse Cleeves, who is an expert in chemical evolution in protoplanetary disks.
+
+%In this part of my thesis, I plan to incorporate the detailed, time-dependent chemical reaction network developed by Merchantz et al.
 (in prep.) \textbf{(?)} in the radial drift calculation described in section \ref{sec:drift}.
@@
 
-In this part of the thesis, I plan to incorporate further dynamical effects, such as atmospheric pollution of planetesimals during the atmospheric accretion phase, Type I and Type II migration, and core dredging (e.g., \citealt{lodders09}, \citealt{stevenson85}, \citealt{guillot04}).
-\textbf{This will result in Paper IV, estimated completion summer 2016} \textit{(Again, perhaps too little material for one paper?)}.
+The calculations summarized in sections \ref{sec:drift} and \ref{sec:chemical} are focused on the relation between disk chemistry, gas and grain abundances, and disk dynamics.
+In this part of my thesis, I plan to understand how these disk chemical abundances translate into planetary abundances when taking into account different dynamic effects on defining the disk and the planet formation process.
+Specifically, I will incorporate dynamic effects such as atmospheric pollution of planetesimals during the atmospheric accretion phase, Type I and Type II migration, and core dredging (e.g., \citealt{lodders09}, \citealt{stevenson85}, \citealt{guillot04}).
+\textbf{This will result in Paper V, estimated completion summer 2016}.
+In this part of my thesis I will collaborate closely with Prof.
+Ruth Murray-Clay.
 
-\subsection{Potential Planet Formation Locations and Comparison with Observations}
+\subsection{Potential Planet Formation Locations} %and Comparison with Observations}
 
@@
 I will then use this framework to explore a wide parameter space of initial disk conditions in order to generate model planet populations and explore potential planet formation locations.
-By comparing my results with present and future observations of atmospheric spectra of giant planets, I will be able to further constrain their formation, accretion and migration history.
+\textbf{This will result in Paper VI, estimated completion spring 2017.}
+
+\section{Auxiliary Projects}
+
+While most of my thesis is focused on theoretical work, I aim to also partake in some observational side-projects.
+I plan to compare the results of Paper VI with with present and future observations of atmospheric spectra of giant planets, which will be instrumental in constraining their formation, accretion and migration history.
+For this project, I plan to collaborate closely with the Disk and Astrochemistry groups here at the CfA.
+
+%By comparing my results with present and future observations of atmospheric spectra of giant planets, I will be able to further constrain their formation, accretion and migration history.
 \textbf{This will result in Paper V, estimated completion spring 2017.}
@@
 As a first step, I investigate the effect of radial drift on the shape of snowlines in protoplanetary disks (see section \ref{sec:drift}).
-I calculate self-consistently the drift and desorption distances as a function of particle size for a passively irradiated disk.
+I calculate self-consistently the drift and desorption distances as a function of particle size for a passive disk.
 I calculate the desorption rates following the prescription of \citet{hollenbach09}, assuming that the solids are perfect spheres and composed of a single volatile material (either H$_2$O, CO$_2$ or CO).
@@
 
-\section{Discussion}
+\section{Timeline}
 
-\textbf{TBD}
+\begin{enumerate}
+\item Paper I: ApJ, estimated completion Spring 2015.
+\item Paper II: ApJ letter, estimated completion Summer 2015.
+\item Paper III: ApJ, estimated completion Fall 2015.
+\item Paper IV: ApJ, estimated completion Spring 2016.
+\item Paper V: ApJ letter, estimated completion Summer 2016.
+\item Paper VI: ApJ, estimated completion Spring 2017.
+\item Collaborative observational project papers: co-author, estimated completion Summer 2017+
+\end{enumerate}
 

### 2015-01-13 python/drift_timescales_varying_sigma_3.py
update

@@ -1,11 +1,131 @@
-from utils.constants import G, kb, mp, Msun, cmperau
+from utils.constants import G, kb, mp, Msun, cmperau, AU
 import numpy as np
+from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
 from T_freeze import T_freeze, Rdes, tevap, vib_freq
 from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
 from scipy.integrate import odeint
 from scipy.optimize import brentq
 from scipy.interpolate import interp1d
 from utils.zbrac import zbrac
+from utils.utilities import tridag
+#from advection_from_Til.two-pop-py.src
+
+###################################################################
+
+def impl_donorcell_adv_diff_delta(n_x,x,Diff,v,g,h,K,L,flim,u_in,dt,pl,pr,ql,qr,rl,rr,coagulation_method,A,B,C,D):
+    """
+    Implicit donor cell advection-diffusion scheme with piecewise constant values
+    
+        Perform one time step for the following PDE:
+    
+           du    d  /    \    d  /              d  /       u   \ \
+           -- + -- | u v | - -- | h(x) Diff(x) -- | g(x) ----  | | = K + L u
+           dt   dx \    /    dx \              dx \      h(x) / /
+    
+        with boundary conditions
+    
+            dgu/h |            |
+          p ----- |      + q u |       = r
+             dx   |x=xbc       |x=xbc
+    Arguments:
+          n_x   = # of grid points
+          x     = the grid
+          Diff  = value of Diff @ cell center
+          v     = the values for v @ interface (array[i] = value @ i-1/2)
+          g     = the values for g(x)
+          h     = the values for h(x)
+          K     = the values for K(x)
+          L     = the values for L(x)
+          flim  = diffusion flux limiting factor at interfaces
+          u     = the current values of u(x)
+          dt    = the time step
+    
+    OUTPUT:
+          u     = the updated values of u(x) after timestep dt
+    
[…]

### 2015-01-14 python/drift_timescales_varying_sigma_3.py
update

@@ -360,15 +360,16 @@ def dMsol_dt(rin, rout, t, s, alpha, Sigmap_in, Sigmap_out, rhos = 3.0, T0 = 120
 
 
 def Sigmap_act(rin, rout, nr, t, dt, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, r1 = 100 * cmperau, C = 4.45e18, \
-    Mstar = Msun, sigma = 2 * 10**(-15), D = 0, dusttogas = 0.01):
+    Mstar = Msun, sigma = 2 * 10**(-15), dusttogas = 0.01):
     
     r = np.logspace(np.log10(rin), np.log10(rout), nr)
     
-    v, Sigmad = [], []
+    v, Sigmad, D = [], [], []
     
     for i in range(nr):
         v = np.append(v, rdot_with_acc(r[i], t, s, alpha, rhos, T0, betaT, mu, r1, C, Mstar, sigma))
         Sigmad = np.append(Sigmad, Sigmadisk_act(r[i], t, alpha, T0, betaT, mu, Mstar, r1, C))
+        D = np.append(D, alpha * cdisk(r[i], T0, betaT, mu) * Hdisk(r[i], T0, betaT, mu, Mstar))
         
     h = Sigmad * r
     uin = r * Sigmad * dusttogas
@@ -378,13 +379,13 @@ def Sigmap_act(rin, rout, nr, t, dt, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./
     L = zeros(nr)
     flim = ones(nr)
     
-    A = zeros(nr)
-    B = zeros(nr)
-    C = zeros(nr)
-    D = zeros(nr)
+    A0 = zeros(nr)
+    B0 = zeros(nr)
+    C0 = zeros(nr)
+    D0 = zeros(nr)
     
     
-    uout = impl_donorcell_adv_diff_delta(nr, r * AU, D, v, g, h, K, L, flim, uin, dt, 1,1, 0, 0, 0, 0, 1, A, B, C, D)
+    uout = impl_donorcell_adv_diff_delta(nr, r * AU, D, v, g, h, K, L, flim, uin, dt, 1,1, 0, 0, 0, 0, 1, A0, B0, C0, D0)
     
     return uout / r
     

### 2015-01-14 python/drift_timescales_varying_sigma_3.py
update

@@ -359,6 +359,7 @@ def dMsol_dt(rin, rout, t, s, alpha, Sigmap_in, Sigmap_out, rhos = 3.0, T0 = 120
             Mdot_solids(rin, t, s, alpha, Sigmap_in, rhos, T0, betaT, mu, r1, C, Mstar, sigma, vrw)
 
 
+
 def Sigmap_act(rin, rout, nr, t, dt, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, r1 = 100 * cmperau, C = 4.45e18, \
     Mstar = Msun, sigma = 2 * 10**(-15), dusttogas = 0.01):
     
@@ -388,6 +389,38 @@ def Sigmap_act(rin, rout, nr, t, dt, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./
     uout = impl_donorcell_adv_diff_delta(nr, r * AU, D, v, g, h, K, L, flim, uin, dt, 1,1, 0, 0, 0, 0, 1, A0, B0, C0, D0)
     
     return uout / r
+
+
+
+#def Sigmap_act(rin, rout, nr, t, dt, s, alpha, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, r1 = 100 * cmperau, C = 4.45e18, \
+#    Mstar = Msun, sigma = 2 * 10**(-15), dusttogas = 0.01):
+#    
+#    r = np.logspace(np.log10(rin), np.log10(rout), nr)
+#    
+#    v, Sigmad, D = [], [], []
+#    
+#    for i in range(nr):
+#        v = np.append(v, rdot_with_acc(r[i], t, s, alpha, rhos, T0, betaT, mu, r1, C, Mstar, sigma))
+#        Sigmad = np.append(Sigmad, Sigmadisk_act(r[i], t, alpha, T0, betaT, mu, Mstar, r1, C))
+#        D = np.append(D, alpha * cdisk(r[i], T0, betaT, mu) * Hdisk(r[i], T0, betaT, mu, Mstar))
+#        
+#    h = Sigmad * r
+#    uin = r * Sigmad * dusttogas
+#    
+#    g = ones(nr)
+#    K = zeros(nr)
+#    L = zeros(nr)
+#    flim = ones(nr)
+#    
+#    A0 = zeros(nr)
+#    B0 = zeros(nr)
+#    C0 = zeros(nr)
+#    D0 = zeros(nr)
[…]

### 2015-01-14 python/advection_from_Til/two_pop_model.py
update

@@ -1,0 +1,796 @@
+from utilities import plotter, progress_bar, tridag
+from numpy import ones,zeros,shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
+from matplotlib.mlab import find
+from matplotlib.pyplot import semilogx,xlim,ylim,figure
+
+import sys
+
+def two_pop_model_run_ic(x,a_0,time,sig_g,sig_d,v_gas,T,alpha,m_star,V_FRAG,RHO_S,peak_position,E_drift,nogrowth=False):
+    """
+    this function evolves the two population model (all model settings
+    are stored in two_pop_velocity). It returns the important parameters of
+    the model.
+    
+    USAGE:
+    [time,solution,v_bar,v_0,v_1,a_dr,a_fr,a_df,a_t]  =  two_pop_model_run_ic ...
+    (x,a_0,time,sig_g,sig_d,v_gas,T,alpha,m_star,V_FRAG,RHO_S,peak_position,E_drift):
+    
+    WHERE:
+        x               = radial grid (nr)                [cm]
+        a_0             = monomer size                    [cm]
+        time            = time of snapshots (nt)          [s]
+        sig_g           = gas  surface density (nr)       [g cm^-2]
+        sig_d           = dust surface density (nr)       [g cm^-2]
+        v_gas           = gas velocity (nr)               [cm/s]
+        T               = temperature grid (nr)           [K]
+        alpha           = turbulence parameter (nr)       [-]
+        m_star          = stellar mass (nt)               [g]
+        V_FRAG          = fragmentation velocity          [cm s^-1]
+        RHO_S           = internal density of the dust    [g cm^-3]
+        peak_position   = index to level off the velocity [-]
+        E_drift         = drift efficiency                [-]
+        
+    KEYWORDS:
+        nogrowth        = true: particle size fixed to a0 [False]
+    
+    RETURNS:
+        time     = snap shot time           (nt,nr)       [s]
+        solution = dust surface density     (nt,nr)       [g cm^-2]
+        v_bar    = dust velocity            (nt,nr)       [cm s^-1]
[…]

### 2015-01-14 python/advection_from_Til/two_pop_test.py

@@ -1,0 +1,96 @@
+#!/usr/bin/env python
+from two_pop_model import two_pop_model_run
+from constants import AU,year,Grav,M_sun,k_b,mu,m_p,R_sun
+import os
+from utilities import plotter
+from matplotlib.pylab import * #@UnusedWildImport
+#
+# create the model
+#
+n_r       = 200
+n_t       = 100
+n_m       = 100
+x         = logspace(log10(0.05),log10(4e3),n_r)*AU
+
+alpha0        = 1e-3
+M_star        = M_sun
+T_star        = 4300.
+R_star        = 2.5*R_sun
+R_c           = 100*AU
+M_disk        = 0.1*M_star
+timesteps     = logspace(4,log10(3e6),n_t)*year
+alpha         = alpha0*ones([n_t,n_r])
+T             = ( (0.05**0.25*T_star * (x/R_star)**-0.5)**4 + 1e4)**0.25
+T             = tile(T,[n_t,1])
+peak_position = zeros(n_t)
+m_star        = M_star*ones(n_t)
+T_COAG_START  = 0.0
+RHO_S         = 1.6
+V_FRAG        = 1000
+a_0           = 1e-4
+E_drift       = 1.0
+#
+# set the initial surface density & velocity
+#
+profile     = maximum(M_disk/(2*pi*R_c**2)*(R_c/x)*exp(-x/R_c),1e-100)
+sigma_g     = tile(profile,[n_t,1])
+sigma_d     = tile(append(array(profile,ndmin=2)/100.,zeros([n_m-1,n_r]),0),[n_t,1])
+v_gas       = -3.0*alpha*k_b*T/mu/m_p/2./sqrt(Grav*M_star/x)*(1.+7./4.)
+#
+# call the model
+#
+[TI,SOL,V,V_0,V_1,A_DR,A_FR,A_DF,A_T] = two_pop_model_run(x,a_0,timesteps,sigma_g,sigma_d,v_gas,T,alpha,m_star,
+                              T_COAG_START,V_FRAG,RHO_S,peak_position,
+                              E_drift)
+#
+# remove the line which shouldn't be there (I should fix this)
+#         
+SOL  = delete(SOL,1,0)
+V    = delete(V,1,0)
[…]

### 2015-01-14 python/advection_from_Til/two_pop_test_ana.py

@@ -1,0 +1,96 @@
+#!/usr/bin/env python
+from two_pop_model import two_pop_model_run
+from constants import AU,year,Grav,M_sun,k_b,mu,m_p,R_sun
+import os
+from utilities import plotter
+from matplotlib.pylab import * #@UnusedWildImport
+#
+# create the model
+#
+n_r       = 200
+n_t       = 100
+n_m       = 100
+x         = logspace(-1,2,n_r)*AU
+
+alpha0        = 1e-3
+M_star        = M_sun
+T_star        = 4300.
+R_star        = 2.5 * R_sun
+R_c           = 100*AU
+M_disk        = 0.1*M_star
+timesteps     = logspace(4,log10(3e6),n_t)*year
+alpha         = alpha0*ones([n_t,n_r])
+T             = ( (0.05**0.25*T_star * (x/R_star)**-0.5)**4 + 1e4)**0.25
+T             = tile(T,[n_t,1])
+peak_position = zeros(n_t)
+m_star        = M_star*ones(n_t)
+T_COAG_START  = 0
+RHO_S         = 1.6
+V_FRAG        = 0 #1000
+a_0           = 1.0
+E_drift       = 1.0
+#
+# set the initial surface density & velocity
+#
+profile     = maximum(M_disk/(2*pi*R_c**2)*(R_c/x)*exp(-x/R_c),1e-100)
+sigma_g     = tile(profile,[n_t,1])
+sigma_d     = tile(append(array(profile,ndmin=2)/100.,zeros([n_m-1,n_r]),0),[n_t,1])
+v_gas       = -3.0*alpha*k_b*T/mu/m_p/2./sqrt(Grav*M_star/x)*(1.+7./4.)
+#
+# call the model
+#
+[TI,SOL,V,V_0,V_1,A_DR,A_FR,A_DF,A_T] = two_pop_model_run(x,a_0,timesteps,sigma_g,sigma_d,v_gas,T,alpha,m_star,
+                              T_COAG_START,V_FRAG,RHO_S,peak_position,
+                              E_drift)
+#
+# remove the line which shouldn't be there (I should fix this)
+#         
+SOL  = delete(SOL,1,0)
+V    = delete(V,1,0)
+V_0  = delete(V_0,1,0)
[…]

### 2015-01-14 python/advection_from_Til/two_pop_test_ana_2.py

@@ -1,0 +1,96 @@
+#!/usr/bin/env python
+from two_pop_model import two_pop_model_run
+from constants import AU,year,Grav,M_sun,k_b,mu,m_p,R_sun
+import os
+from utilities import plotter
+from matplotlib.pylab import * #@UnusedWildImport
+#
+# create the model
+#
+n_r       = 100
+n_t       = 20
+n_m       = 1
+x         = logspace(log10(0.1),log10(1e4),n_r)*AU
+
+alpha0        = 1e-2
+M_star        = M_sun
+T_star        = 4300.
+R_star        = 2.5*R_sun
+R_c           = 100*AU
+M_disk        = 0.1*M_star
+timesteps     = logspace(4,log10(3e6),n_t)*year
+alpha         = alpha0*ones([n_t,n_r])
+T             = 120 * (x/AU)**(-3./7) #( (0.05**0.25*T_star * (x/R_star)**-0.5)**4 + 1e4)**0.25
+T             = tile(T,[n_t,1])
+peak_position = zeros(n_t)
+m_star        = M_star*ones(n_t)
+T_COAG_START  = 0.0
+RHO_S         = 3.0
+V_FRAG        = 0.0 #1000
+a_0           = 100.0 #1e-4
+E_drift       = 1.0
+#
+# set the initial surface density & velocity
+#
+profile     = maximum(M_disk*20/(2*pi*R_c**2)*(R_c/x)*exp(-x/R_c),1e-100)
+sigma_g     = tile(profile,[n_t,1])
+sigma_d     = tile(append(array(profile,ndmin=2)/100.,zeros([n_m-1,n_r]),0),[n_t,1])
+v_gas       = -3.0*alpha*k_b*T/mu/m_p/2./sqrt(Grav*M_star/x)*(1.+7./4.)
+#
+# call the model
+#
+[TI,SOL,V,V_0,V_1,A_DR,A_FR,A_DF,A_T] = two_pop_model_run(x,a_0,timesteps,sigma_g,sigma_d,v_gas,T,alpha,m_star,
+                              T_COAG_START,V_FRAG,RHO_S,peak_position,
+                              E_drift)
+#
+# remove the line which shouldn't be there (I should fix this)
+#         
[…]

### 2015-03-16 python/advection_from_Til/two_pop_model.py
update

@@ -138,13 +138,13 @@ def two_pop_model_run_ic(x,a_0,time,sig_g,sig_d,v_gas,T,alpha,m_star,V_FRAG,RHO_
         #
         # try variable time step
         #
-        while any(u_out[2:-1][mask]/x[2:-1][mask]>=1e-30):
-            dt = dt/10.
-            if dt<1.0:
-                print('ERROR: time step got too short')
-                sys.exit(1)
-            u_out = impl_donorcell_adv_diff_delta(n_r,x,D,v,g,h,K,L,flim,u_in,dt,1,1,0,0,0,0,1,A0,B0,C0,D0)
-            mask = abs(u_out[2:-1]/u_in[2:-1]-1)>0.3
+        ##while any(u_out[2:-1][mask]/x[2:-1][mask]>=1e-30):
+        ##    dt = dt/10.
+        ##    if dt<1.0:
+        ##       print('ERROR: time step got too short')
+        ##        sys.exit(1)
+        ##    u_out = impl_donorcell_adv_diff_delta(n_r,x,D,v,g,h,K,L,flim,u_in,dt,1,1,0,0,0,0,1,A0,B0,C0,D0)
+        ##    mask = abs(u_out[2:-1]/u_in[2:-1]-1)>0.3
         #
         # update
         #
@@ -297,14 +297,14 @@ def two_pop_model_run(x_1,a_0,timesteps_1,sigma_g_1,sigma_d_1,v_gas_1,T_1,alpha_
         # set the time step
         #
         dt = timesteps_1[it_old]-t 
-        dt = min(dt*10,timesteps_1[it_old]-t)
-        if t != 0.0: dt = min(dt,t/200)
-        if dt==0:
-            print('ERROR:')
-            print('t      = %g years'%(t/year))
-            print('it_old = %g'%it_old)
-            print('dt = 0')
-            sys.exit(1)
+        ##dt = min(dt*10,timesteps_1[it_old]-t)
+        ##if t != 0.0: dt = min(dt,t/200)
+        ##if dt==0:
+        ##    print('ERROR:')
+        ##    print('t      = %g years'%(t/year))
+        ##    print('it_old = %g'%it_old)
+        ##    print('dt = 0')
+        ##    sys.exit(1)
         #
[…]

### 2015-03-16 python/advection_from_Til/two_pop_test_ana_0316_1.py

@@ -1,0 +1,109 @@
+#!/usr/bin/env python
+from two_pop_model import two_pop_model_run
+from constants import AU,year,Grav,M_sun,k_b,mu,m_p,R_sun
+import os
+from utilities import plotter
+from matplotlib.pylab import * #@UnusedWildImport
+#
+# create the model
+#
+n_r       = 500
+n_t       = 20
+n_m       = 1
+x         = logspace(log10(0.05),log10(4e3),n_r)*AU
+
+alpha0        = 1e-2
+M_star        = M_sun
+T_star        = 4300.
+R_star        = 2.5*R_sun
+R_c           = 60*AU
+M_disk        = 0.1*M_star
+timesteps     = linspace(1e4,3e6,n_t)*year
+alpha         = alpha0*ones([n_t,n_r])
+T             = ( (0.05**0.25*T_star * (x/R_star)**-0.5)**4 + 1e4)**0.25
+T             = tile(T,[n_t,1])
+peak_position = zeros(n_t)
+m_star        = M_star*ones(n_t)
+T_COAG_START  = 0.0
+RHO_S         = 1.6
+V_FRAG        = 0 #1000
+a_0           = 1e-3
+E_drift       = 1.0
+#
+# set the initial surface density & velocity
+#
+
+profile     = maximum(M_disk/(2*pi*R_c**2)*(R_c/x)*exp(-x/R_c),1e-100)
+
+        
+sigma_g     = tile(profile,[n_t,1])
+sigma_d     = tile(append(array(profile,ndmin=2)/100.,zeros([n_m-1,n_r]),0),[n_t,1])
+v_gas       = -3.0*alpha*k_b*T/mu/m_p/2./sqrt(Grav*M_star/x)*(1.+7./4.)
+
+#
+#sigma_d = np.ndarray(shape = (len(timesteps), len(x)), dtype = float)
+#for i in range(len(timesteps)):
+#    for j in range(len(x)):
+#        if x[j] >= 20 and x[j] <=22:
+#            sigma_d[i, j] = profile[j] / 100
+#        else:
+#            sigma_d[i, j] = 1e-100
+
+#
+# call the model
+#
[…]

### 2015-03-16 python/advection_from_Til/two_pop_test_ana_0316_2.py

@@ -1,0 +1,104 @@
+#!/usr/bin/env python
+from two_pop_model import two_pop_model_run
+from constants import AU,year,Grav,M_sun,k_b,mu,m_p,R_sun
+import os
+from utilities import plotter
+from matplotlib.pylab import * #@UnusedWildImport
+#
+# create the model
+#
+n_r       = 500
+n_t       = 200
+n_m       = 1
+x         = logspace(log10(0.05),log10(4e3),n_r)*AU
+
+alpha0        = 1e-3
+M_star        = M_sun
+T_star        = 4300.
+R_star        = 2.5*R_sun
+R_c           = 60*AU
+M_disk        = 0.1*M_star
+timesteps     = linspace(1e4,3e6,n_t)*year
+alpha         = alpha0*ones([n_t,n_r])
+T             = ( (0.05**0.25*T_star * (x/R_star)**-0.5)**4 + 1e4)**0.25
+T             = tile(T,[n_t,1])
+peak_position = zeros(n_t)
+m_star        = M_star*ones(n_t)
+T_COAG_START  = 0.0
+RHO_S         = 1.6
+V_FRAG        = 0 #1000
+a_0           = 1e1
+E_drift       = 1.0
+#
+# set the initial surface density & velocity
+#
+profile     = maximum(M_disk/(2*pi*R_c**2)*(R_c/x)*exp(-x/R_c),1e-100)
+sigma_g     = tile(profile,[n_t,1])
+#sigma_d     = tile(append(array(profile,ndmin=2)/100.,zeros([n_m-1,n_r]),0),[n_t,1])
+v_gas       = -3.0*alpha*k_b*T/mu/m_p/2./sqrt(Grav*M_star/x)*(1.+7./4.)
+
+sigma_d = np.ndarray(shape = (len(timesteps), len(x)), dtype = float)
+for i in range(len(timesteps)):
+    for j in range(len(x)):
+        if x[j] >= 20*AU and x[j] <=22*AU:
+            sigma_d[i, j] = profile[j] / 100
+        else:
+            sigma_d[i, j] = 1e-20
+
+# call the model
+#
+[TI,SOL,V,V_0,V_1,A_DR,A_FR,A_DF,A_T] = two_pop_model_run(x,a_0,timesteps,sigma_g,sigma_d,v_gas,T,alpha,m_star,
[…]

### 2015-03-16 python/drift_timescales_Til_parameters_ring_check.py

@@ -288,9 +288,12 @@ def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, Rstar, Tstar, rc, rho
         
     h = Sigmad * r * AU
     
+    
+    #uin = r * Sigmad * dusttogas * AU
+    
     uin = []
     for i in range(nr):
-        if r[i] >= 50 and r[i] <= 60:
+        if r[i] >= 20 and r[i] <= 22:
             uin = np.append(uin, r[i] * Sigmad[i] * dusttogas * AU)
         else:
             uin = np.append(uin, 1e-100)
@@ -306,10 +309,12 @@ def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, Rstar, Tstar, rc, rho
     D0 = zeros(nr)
     
     for i in range(nr):
-        sigarray[i, 0] = uin[i] / (r[i] * AU * dusttogas) #Sigmad[i] * dusttogas
+        sigarray[i, 0] = uin[i] / (r[i] * AU) 
+        #sigarray[i, 0] = Sigmad[i] * dusttogas
     
     for j in range(nt):
-        uout = impl_donorcell_adv_diff_delta(nr, r * AU, D, v, g, h, K, L, flim, uin, dt, 0, 0, 1, 1, 1e-100*r[0]*AU, 1e-100*r[-1]*AU, 1, A0, B0, C0, D0)
+        #uout = impl_donorcell_adv_diff_delta(nr, r * AU, D, v, g, h, K, L, flim, uin, dt, 0, 0, 1, 1, 1e-100*r[0]*AU, 1e-100*r[-1]*AU, 1, A0, B0, C0, D0)
+        uout = impl_donorcell_adv_diff_delta(nr, r * AU, D, v, g, h, K, L, flim, uin, dt, 1, 1, 0, 0, 0, 0, 1, A0, B0, C0, D0)
         
         for i in range(nr):
             sigarray[i, j + 1] = uout[i] / (r[i] * AU)

### 2015-03-17 python/advection_from_Til/two_pop_model.py
update

@@ -297,14 +297,14 @@ def two_pop_model_run(x_1,a_0,timesteps_1,sigma_g_1,sigma_d_1,v_gas_1,T_1,alpha_
         # set the time step
         #
         dt = timesteps_1[it_old]-t 
-        ##dt = min(dt*10,timesteps_1[it_old]-t)
-        ##if t != 0.0: dt = min(dt,t/200)
-        ##if dt==0:
-        ##    print('ERROR:')
-        ##    print('t      = %g years'%(t/year))
-        ##    print('it_old = %g'%it_old)
-        ##    print('dt = 0')
-        ##    sys.exit(1)
+        dt = min(dt*10,timesteps_1[it_old]-t)
+        if t != 0.0: dt = min(dt,t/200)
+        if dt==0:
+            print('ERROR:')
+            print('t      = %g years'%(t/year))
+            print('it_old = %g'%it_old)
+            print('dt = 0')
+            sys.exit(1)
         #
         # calculate the velocity
         #
@@ -333,13 +333,14 @@ def two_pop_model_run(x_1,a_0,timesteps_1,sigma_g_1,sigma_d_1,v_gas_1,T_1,alpha_
         ###
         ### try variable time step
         ###
-        ##while any(u_out[2:-1][mask]/x_1[2:-1][mask]>=1e-30):
-        ##    dt = dt/10.
-        ##    if dt<1.0:
-        ##        print('ERROR: time step got too short')
-        ##        sys.exit(1)
-        ##    u_out = impl_donorcell_adv_diff_delta(n_r,x_1,D,v,g,h,K,L,flim,u_in,dt,1,1,0,0,0,0,1,A0,B0,C0,D0)
-        ##    mask = abs(u_out[2:-1]/u_in[2:-1]-1)>0.3
+        while any(u_out[2:-1][mask]/x_1[2:-1][mask]>=1e-30):
+            dt = dt/10.
+            if dt<1.0:
+                print('ERROR: time step got too short')
+                sys.exit(1)
+            #u_out = impl_donorcell_adv_diff_delta(n_r,x_1,D,v,g,h,K,L,flim,u_in,dt,1,1,0,0,0,0,1,A0,B0,C0,D0)
+            u_out = impl_donorcell_adv_diff_delta(n_r,x_1,D,v,g,h,K,L,flim,u_in,dt,0,0,1,1,1e-100*x_1[0],1e-100*x_1[-1],1,A0,B0,C0,D0)
[…]

### 2015-03-17 python/advection_from_Til/two_pop_test_ana_0115_1.py

@@ -27,7 +27,7 @@ m_star        = M_star*ones(n_t)
 T_COAG_START  = 0.0
 RHO_S         = 1.6
 V_FRAG        = 0 #1000
-a_0           = 1e1
+a_0           = 1e-1
 E_drift       = 1.0
 #
 # set the initial surface density & velocity

### 2015-03-17 python/advection_from_Til/two_pop_test_ana_0316_2.py

@@ -7,12 +7,12 @@ from matplotlib.pylab import * #@UnusedWildImport
 #
 # create the model
 #
-n_r       = 500
-n_t       = 200
+n_r       = 200
+n_t       = 50
 n_m       = 1
 x         = logspace(log10(0.05),log10(4e3),n_r)*AU
 
-alpha0        = 1e-3
+alpha0        = 1e-2
 M_star        = M_sun
 T_star        = 4300.
 R_star        = 2.5*R_sun
@@ -27,7 +27,7 @@ m_star        = M_star*ones(n_t)
 T_COAG_START  = 0.0
 RHO_S         = 1.6
 V_FRAG        = 0 #1000
-a_0           = 1e1
+a_0           = 1e-3
 E_drift       = 1.0
 #
 # set the initial surface density & velocity
@@ -37,13 +37,18 @@ sigma_g     = tile(profile,[n_t,1])
 #sigma_d     = tile(append(array(profile,ndmin=2)/100.,zeros([n_m-1,n_r]),0),[n_t,1])
 v_gas       = -3.0*alpha*k_b*T/mu/m_p/2./sqrt(Grav*M_star/x)*(1.+7./4.)
 
-sigma_d = np.ndarray(shape = (len(timesteps), len(x)), dtype = float)
-for i in range(len(timesteps)):
-    for j in range(len(x)):
-        if x[j] >= 20*AU and x[j] <=22*AU:
-            sigma_d[i, j] = profile[j] / 100
-        else:
-            sigma_d[i, j] = 1e-20
+profiled = maximum(profile[110] / 100 * np.exp(-(x-x[110])**2/(2*(1*AU)**2)), 1e-100)
+sigma_d     = tile(append(array(profiled,ndmin=2),zeros([n_m-1,n_r]),0),[n_t,1])
+
+#sigma_d = np.ndarray(shape = (len(timesteps), len(x)), dtype = float)
+#for i in range(len(timesteps)):
+#    for j in range(len(x)):
+#        if x[j] >= 20*AU and x[j] <=22*AU:
+#            sigma_d[i, j] = profile[j] / 100
+#        elif x[j] <= 20*AU:
+#            sigma_d[i, j] = profile[j] / 100 * np.exp(-20*AU/x[j])
+#        elif x[j] >= 22*AU:
+#            sigma_d[i, j] = profile[j] / 100 * np.exp(-x[j]/(22*AU))
 
 # call the model
 #

### 2015-03-17 python/drift_timescales_Til_parameters_ring_check.py

@@ -288,6 +288,7 @@ def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, Rstar, Tstar, rc, rho
         
     h = Sigmad * r * AU
     
+    ##uin = r * maximum(Sigmad[110] / 100 * np.exp(-(r-r[110])**2/(2*(1)**2)), 1e-100) * AU
     
     #uin = r * Sigmad * dusttogas * AU
     

### 2015-03-17 python/drift_timescales_updated_sigmap.py

@@ -3,6 +3,7 @@ import numpy as np
 from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
 from T_freeze import T_freeze, Rdes, tevap, vib_freq
 from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
+import scipy
 from scipy.integrate import odeint
 from scipy.optimize import brentq
 from scipy.interpolate import interp1d
@@ -876,6 +877,7 @@ def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 2.
           #sf = y[:,1][i - 1]
           #npts = 10 * npts
      if i == len(y[:,0] - 1):
+        print "No desorption." 
         if returnall == True:
           return tv, y[:,0] / cmperau, y[:,1]
         else:
@@ -888,6 +890,8 @@ def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 2.
      	  fintt = interp1d(yint[:,1][::-1], tint)
 
      	  try:
+     	      print "It's desorbing."
+     	      print i
               af = float(finta(0))
               tf = float(fintt(0))
 
@@ -900,6 +904,8 @@ def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 2.
               else:
      	          return float(finta(0)) / cmperau
           except ValueError:
+              print "It's desorbing but there's an issue."
+              print i
               if returnall == True:
                   return tv, y[:,0] / cmperau, y[:,1]
               else:
@@ -911,7 +917,7 @@ def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 2.
 
 
 def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, rc, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, \
-    Mstar = Msun, sigma = 2 * 10**(-15), dusttogas = 0.01, gammadflag = 0, sigmad_dt = 1):
[…]

### 2015-03-17 tex/draft/ms.tex
update

%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
%\documentclass[12pt, preprint]{aastex}
\documentclass[apj]{emulateapj}

\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex

%\usepackage{pdfsync}
\usepackage{subeqnarray}
\usepackage{natbib}
\usepackage{color}

\bibliographystyle{apj}

\newcommand{\ie}{i.e.\ }
\newcommand{\eg}{e.g.\ }
\newcommand{\p}{\partial}
\newcommand{\brak}[1]{\langle #1\rangle}

\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
\newcommand{\cm}{\; {\rm cm}}
\newcommand{\mm}{\; {\rm mm}}
%\newcommand{\ps}{\; {\rm s^{-1}}}
\newcommand{\km}{\; {\rm km}}
\newcommand{\au}{\; \varpi_{\rm AU}}
\newcommand{\AU}{\; {\rm AU}}
\def\K{\; {\rm K}}

\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
\newcommand{\nab}{\vc{\nabla}}
\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}

\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
\newcommand{\App}[1]{Appendix~\ref{#1}}
\newcommand{\Sec}[1]{Sect.~\ref{#1}}
\newcommand{\Chap}[1]{Chapter~\ref{#1}}
\newcommand{\Fig}[1]{Fig.~\ref{#1}}
\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
\newcommand{\Tab}[1]{Table \ref{#1}}

\definecolor{gray}{gray}{0.5}
\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}

%\newenvironment{packed_item}{
%\begin{itemize}
%  \setlength{\itemsep}{1pt}
%  \setlength{\parskip}{0pt}
%  \setlength{\parsep}{0pt}
%}{\end{itemize}}

\begin{document}

%\slugcomment{Draft Modified \today}

%\shorttitle{Critical Core Mass at Wide Separations}
\shortauthors{Piso et al.}

\title{The Influence of Radial Drift and Gas Accretion on the C/O Ratio in Protoplanetary Disks}
%\title{Lower Limits on the Core Mass for Giant Planets}
%\title{The Critical Core Mass of Wide-Separation Giant Planets:  Lower Limits from Kelvin-Helmholtz Cooling}
\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}, Tilman Birnstiel\altaffilmark{1}}
\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}

%\begin{abstract}
%\end{abstract}

\section{Introduction}

\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al. 2011).}

\emgr{Describe goals and approach of this paper, i.e.: study the importance of radial drift on the location of snowlines in protoplanetary disks and how radial drift affects the C/O ratio. Mention that disks are viscously accreting in the early stages of planet formation, hence gas accretion has to be taken into account. The goal of this paper is two-fold: (1) estimate using both numerical models and analytic arguments the range of particle sizes for which radial drift changes snowline location; (2) calculate the C/O ratio for various particle sizes (and possibly a size distribution) and at different times throughout the evolution of the gas disk.}

This paper is organized as follows: \emgr{(section summaries)}.

\section{Radial Drift Model}
\subsection{Disk and Desorption Model}

\emgr{Describe disk model --- both passive MMSN and active self-similar solution with $\alpha$ prescription for viscosity. Describe the model for evolving the surface density profile of planetesimals following Birnstiel et al. (2012). Describe temperature profile --- currently power-law, but may change. Describe desorption model and parameters following Hollenbach et al. (2009). Mention that the solids are perfect spheres composed of a single volatile. Perhaps this subsection needs to be split into subsubsections.}

\subsection{Relevant timescales}

\emgr{Calculate timescale for radial drift following Chiang \& Youdin (2010). Calculate desorption timescale following Hollenbach et al. (2009). Estimate gas accretion timescale for a given $\alpha$. Important: mention that, for simplicity and illustrative purposes, these calculations are performed for a passive disk. Show plot with the timescales as a function of particle size at different snowlines to show the regime in which drift matters}.

\section{Snowline Locations}

\emgr{Present the equation set that you are solving, $dr/dt=\dot{r}$, $ds/dt=...$. I don't think it is necessary to describe in detail the numerical method of solving the equations since it's pretty straightforward. Present the pretty rainbow plots side by side for passive and active disks, and highlight the differences. Insert the plots that shows that for the intermediate size particles, the desorption distance can be estimated analytically with good accuracy. Maybe also include the plot showing the desorption distance as a function of particle size, both for passive and active disk. Split into subsections?}

\section{Results for the C/O Ratio}
\emgr{Present the flux equations used to keep track of the amount of C and O in gas and dust throughout the disk. Motivate the simplification of using a fixed particle sized, fixed timescale, and fixed desorption distance in the calculations by referring to the rainbow plot in the previous section and by inserting the plot that shows that particle desorb almost instantly at a fixed distance. Based on these assumptions, finally show the equivalent of Fig. 1 in \"Oberg et al. (2011) for different particle sizes, and at different times in the gas disk evolution (i.e., not just at 3 Myr) --- a multi-panel plot could be a good idea. Discuss the result, trends, etc. Ideally, have a final plot showing the results for a particle size distribution rather than for individual particles.}

\section{Discussion and Model Limitations}

\emgr{Include: non-inclusion of turbulence, assumption of perfect spheres when in fact they may have cracks, particles composed of a single volatile when in reality they are likely to be mixed, etc. Discuss uncertainty of initial conditions and estimate how much they matter. ....}

\section{Summary and Future Work}

\emgr{Summarize results. Mention inclusion of $N_2$ as a first expansion. Mention the implementation of time-dependent chemical models in the drift calculation.}

\appendix
\section{...}

\emgr{Right now it's unclear to me what could go in an appendix, if anything. Maybe discuss a bit the algorithm to evolve $\Sigma_{\rm p}$ (although it is already explained in detail in the appendix of Birnstiel et al. 2010). Maybe show some example profiles of $\Sigma_{\rm p}$ at different times and for different particle sizes.}

\if\bibinc n
\bibliography{refs}
\fi

\if\bibinc y
\begin{thebibliography}
\end{thebibliography}
\fi

\end{document}

### 2015-03-17 python/drift_timescales_updated_sigmap.py

@@ -856,7 +856,7 @@ def r_freeze(mx, Ex, nx, T0 = 120., betaT = 3./7):
      return brentq(f, 0.1, 100)
 
 
-def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 2.0, T0 = 120, betaT = 3./7, mu = 2.3, \
+def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.3, \
     Mstar = Msun, sigma = 2 * 10**(-15), npts = 1e6, nptsin = 1e4, tin = 1e-10, gammadflag = 0, returnall = False):
 
      def f(x, t):
@@ -1109,11 +1109,19 @@ def dMdot(rin, rout, t, dt, rstart, s, alpha, mx, Ex, Mdisk = 0.1*Msun, rc = 100
     tf, af, sf = rf(rstart, t, s, mx, Ex, alpha, dr, Mdisk, rc, Nx, rhos, T0, betaT, mu, \
         Mstar, sigma, npts, nptsin, tin, gammadflag, returnall)  
     
-    if sf[-1] != 0:
+    if sf[-1] != 0.0:
+        print "Not desorbing"
         return f * dMdotgas, f * dMdotsolids
         
-    elif af[-1] < rin or af[-1] > rout:
+        
+    elif af[-1] < rin:
+        print "Hasn't reached desorption distance"
         return f * dMdotgas, f * dMdotsolids
+    
+    elif af[-1] > rout:
+        print "Already desorbed"
+        return f * dMdotgas, 0
+        
         
     else:
         
@@ -1129,11 +1137,11 @@ def dMdot(rin, rout, t, dt, rstart, s, alpha, mx, Ex, Mdisk = 0.1*Msun, rc = 100
         Des = 3 * sigmapdes * (af[-1]*AU) * (0.001*af[-1])*AU * 12*np.pi / (2 * rhos * s) * \
             mu * mp * Nx * Rdes(mx, Ex, Tdisk(af[-1], T0, betaT)) 
         
-        dMdotgas_new = dMdotgas + Des
-        dMdotsolids_new = dMdotsolids - Des
+        dMdotgas_new = dMdotgas + f * Des
+        dMdotsolids_new = dMdotsolids - f * Des
[…]

### 2015-03-18 python/drift_timescales_updated_sigmap_ring_check.py
update

@@ -1,0 +1,1357 @@
+from utils.constants import G, kb, mp, Msun, cmperau, AU
+import numpy as np
+from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
+from T_freeze import T_freeze, Rdes, tevap, vib_freq
+from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
+import scipy
+from scipy.integrate import odeint
+from scipy.optimize import brentq
+from scipy.interpolate import interp1d
+from utils.zbrac import zbrac
+from utils.utilities import tridag
+#from advection_from_Til.two-pop-py.src
+
+###################################################################
+
+def impl_donorcell_adv_diff_delta(n_x,x,Diff,v,g,h,K,L,flim,u_in,dt,pl,pr,ql,qr,rl,rr,coagulation_method,A,B,C,D):
+    """
+    Implicit donor cell advection-diffusion scheme with piecewise constant values
+    
+        Perform one time step for the following PDE:
+    
+           du    d  /    \    d  /              d  /       u   \ \
+           -- + -- | u v | - -- | h(x) Diff(x) -- | g(x) ----  | | = K + L u
+           dt   dx \    /    dx \              dx \      h(x) / /
+    
+        with boundary conditions
+    
+            dgu/h |            |
+          p ----- |      + q u |       = r
+             dx   |x=xbc       |x=xbc
+    Arguments:
+          n_x   = # of grid points
+          x     = the grid
+          Diff  = value of Diff @ cell center
+          v     = the values for v @ interface (array[i] = value @ i-1/2)
+          g     = the values for g(x)
+          h     = the values for h(x)
+          K     = the values for K(x)
+          L     = the values for L(x)
+          flim  = diffusion flux limiting factor at interfaces
+          u     = the current values of u(x)
+          dt    = the time step
+    
+    OUTPUT:
+          u     = the updated values of u(x) after timestep dt
+    
+    """
+    D05=zeros(n_x)
+    h05=zeros(n_x)
+    rhs=zeros(n_x)
+    #
+    # calculate the arrays at the interfaces
+    #
+    for i in arange(1,n_x):
[…]

### 2015-03-18 python/drift_timescales_updated_sigmap.py
update

@@ -1111,7 +1111,7 @@ def dMdot(rin, rout, t, dt, rstart, s, alpha, mx, Ex, Mdisk = 0.1*Msun, rc = 100
     
     if sf[-1] != 0.0:
         print "Not desorbing"
-        return f * dMdotgas, f * dMdotsolids
+        return 0, f * dMdotsolids
         
         
     elif af[-1] < rin:

### 2015-03-18 python/drift_timescales_updated_sigmap.py
update

@@ -856,7 +856,7 @@ def r_freeze(mx, Ex, nx, T0 = 120., betaT = 3./7):
      return brentq(f, 0.1, 100)
 
 
-def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.3, \
+def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, \
     Mstar = Msun, sigma = 2 * 10**(-15), npts = 1e6, nptsin = 1e4, tin = 1e-10, gammadflag = 0, returnall = False):
 
      def f(x, t):
@@ -877,7 +877,7 @@ def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 3.
           #sf = y[:,1][i - 1]
           #npts = 10 * npts
      if i == len(y[:,0] - 1):
-        print "No desorption." 
+        #print "No desorption." 
         if returnall == True:
           return tv, y[:,0] / cmperau, y[:,1]
         else:
@@ -890,8 +890,8 @@ def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 3.
      	  fintt = interp1d(yint[:,1][::-1], tint)
 
      	  try:
-     	      print "It's desorbing."
-     	      print i
+     	      #print "It's desorbing."
+     	      #print i
               af = float(finta(0))
               tf = float(fintt(0))
 
@@ -904,8 +904,8 @@ def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 3.
               else:
      	          return float(finta(0)) / cmperau
           except ValueError:
-              print "It's desorbing but there's an issue."
-              print i
+              #print "It's desorbing but there's an issue."
+              #print i
               if returnall == True:
                   return tv, y[:,0] / cmperau, y[:,1]
               else:
@@ -1096,53 +1096,140 @@ def dMdot_gas(rin, rout, t, alpha, Mdisk = 0.1*Msun, mu = 2.35, T0 = 120, betaT 
[…]

### 2015-03-18 python/drift_timescales_updated_sigmap_ring_check.py

@@ -319,19 +319,21 @@ def Sigmadisk(r, t, alpha, Mdisk, rc, T0 = 120, betaT = 3./7, mu = 2.35, Mstar =
      gas surface density in g cm^-2
            
      """    
+    
+     return 2200. * r**(-3./2)
 
-     if gammadflag == 0:
-         gammad = gamma(betaT)
-     else:
-         gammad = 1
-         
-     nu1 = nu(rc / cmperau, alpha, T0, betaT, mu, Mstar)
-     ts = 1. / (3 * (2 - gammad)**2) * rc**2 / nu1
-     T = t / ts + 1
-     rtild = r * cmperau / rc
-
-     return Mdisk / (2 * np.pi * rc**2 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
-            * np.exp(- rtild**(2 - gammad) / T)
+#     if gammadflag == 0:
+#         gammad = gamma(betaT)
+#     else:
+#         gammad = 1
+#         
+#     nu1 = nu(rc / cmperau, alpha, T0, betaT, mu, Mstar)
+#     ts = 1. / (3 * (2 - gammad)**2) * rc**2 / nu1
+#     T = t / ts + 1
+#     rtild = r * cmperau / rc
+#
+#     return Mdisk / (2 * np.pi * rc**2 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
+#            * np.exp(- rtild**(2 - gammad) / T)
                 
 
 def vacc_act(r, t, alpha, mu = 2.35, T0 = 120, betaT = 3./7, rc = 100 * AU, Mstar = Msun, gammadflag = 0):
@@ -1003,7 +1005,7 @@ def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, rc, rhos = 3.0, T0 = 
     
     uin = []
     for i in range(nr):
-        if r[i] >= 9.4 and r[i] <= 9.6:
+        if r[i] >= 30 and r[i] <= 30.2:
             uin = np.append(uin, r[i] * Sigmad[i] * dusttogas * AU)
         else:
             uin = np.append(uin, 1e-100)
@@ -1033,25 +1035,25 @@ def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, rc, rhos = 3.0, T0 = 
         uin = uout #updating the new dust surface density to be used in the next time step
[…]

### 2015-03-18 python/utils/constants.py

@@ -22,7 +22,7 @@ sigma =  5.670373e-5
 #unit conversion
 cmperau = 14959788700000.0
 AU = 1.49597887e13
-year = 3.16e7
+yr = 365*24*3600
 Myr = 3.16e13
 barP = 1e6 #dyne/cm*2 to bar (not that `bar` in matplotlib is problem in pylab)
 

### 2015-03-20 python/drift_timescales.py
update

@@ -168,6 +168,20 @@ def rdot_with_acc(r, s, Mdotgas, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, 
          Mstar, sigma, eps, vphi) + rdot_gas(r, s, Mdotgas, Sigma0, betaS) / \
               (1 + taus(r, s, rhos , T0, betaT, mu, Sigma0, betaS, \
                    Mstar, sigma, eps)**2)
+                   
+                   
+def rdrift(rin, tin, tf, s, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, Sigma0 = 2200, betaS = 3./2, \
+    Mstar = Msun, sigma = 2 * 10**(-15), eps = 0, vphi = 0, npts = 50):
+    
+    def f(x, t):
+          
+        return rdot(x / cmperau, s, rhos, T0, betaT, mu, Sigma0, betaS, Mstar, sigma, eps, vphi)
+          
+
+    t = np.logspace(np.log10(tin), np.log10(tf), npts)
+    y = odeint(f, rin * cmperau, t)
+    
+    return y[-1][0] / cmperau
 
 
 ################################################################################

### 2015-03-20 python/drift_timescales_updated_sigmap.py

@@ -912,6 +912,22 @@ def rf(rin, tf, sin, mx, Ex, alpha, dr, Mdisk, rc = 100*AU, Nx = 1e15, rhos = 3.
                   return y[:,0][-1] / cmperau
 
 
+def rdrift(rin, tin, tf, s, alpha, dr, Mdisk, rc = 100*AU, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, \
+    Mstar = Msun, sigma = 2 * 10**(-15), npts = 100, gammadflag = 0):
+    
+    def f(x, t):
+          
+        return rdot_with_acc(x / cmperau, t, s, alpha, dr, Mdisk, rc, T0, betaT, rhos, mu, Mstar, gammadflag, sigma)
+          
+
+    t = np.logspace(np.log10(tin), np.log10(tf), npts)
+    y = odeint(f, rin * cmperau, t)
+    
+    return y[-1][0] / AU
+    
+    
+
+
 #######################################################################################
 
 
@@ -1002,7 +1018,7 @@ def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, rc, rhos = 3.0, T0 = 
     
     #uin = []
     #for i in range(nr):
-    #    if r[i] >= 27 and r[i] <= 29:
+    #    if r[i] >= 10. and r[i] <= 10.5:
     #        uin = np.append(uin, r[i] * Sigmad[i] * dusttogas * AU)
     #    else:
     #        uin = np.append(uin, 1e-100)
@@ -1119,6 +1135,10 @@ def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, Mdisk = 0.1*Msun, rc 
     Ms, Mg = [], []
     Desv, dMdotsv, dMdots_newv = [], [], []
     
+    tf, af, sf = rf(rstart, tf, s, mx, Ex, alpha, dr, Mdisk, rc, Nx, rhos, T0, betaT, mu, \
+            Mstar, sigma, npts, nptsin, ti, gammadflag, returnall)
+            
+    
     for i in range(nt - 1):
         
         
@@ -1134,8 +1154,8 @@ def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, Mdisk = 0.1*Msun, rc 
             Sigmadisk(rout, t[i], alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag)
         dMdotg = Mdotg_out - Mdotg_in
         
[…]

### 2015-03-20 python/drift_timescales_updated_sigmap_ring_check.py

@@ -533,6 +533,10 @@ def eta(r, t, alpha, dr, Mdisk, rc = 100*AU, T0=120, betaT=3./7, mu = 2.35, Msta
        
     """
     
+    return cdisk(r, T0, betaT, mu)**2 / \
+        (2 * (Omegak(r, Mstar) * r * cmperau)**2) 
+    
+    
     return - (r * AU) / (2 * rhodisk(r, t, alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag) * (r * AU * Omegak(r, Mstar))**2) \
             * (Pdisk(r+dr, t, alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag) - \
                     Pdisk(r, t, alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag)) / (dr * AU)
@@ -1005,7 +1009,7 @@ def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, rc, rhos = 3.0, T0 = 
     
     uin = []
     for i in range(nr):
-        if r[i] >= 30 and r[i] <= 30.2:
+        if r[i] >= 10 and r[i] <= 10.5:
             uin = np.append(uin, r[i] * Sigmad[i] * dusttogas * AU)
         else:
             uin = np.append(uin, 1e-100)

### 2015-04-01 python/drift_timescales_updated_sigmap.py
update

@@ -4,7 +4,7 @@ from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,i
 from T_freeze import T_freeze, Rdes, tevap, vib_freq
 from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
 import scipy
-from scipy.integrate import odeint
+from scipy.integrate import odeint, quad
 from scipy.optimize import brentq
 from scipy.interpolate import interp1d
 from utils.zbrac import zbrac
@@ -1132,65 +1132,67 @@ def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, Mdisk = 0.1*Msun, rc 
         
     func = scipy.interpolate.interp2d(t, rgrid, sigarray)
     
-    Ms, Mg = [], []
+    Ms, Mg, sig = [], [], []
     Desv, dMdotsv, dMdots_newv = [], [], []
     
     tf, af, sf = rf(rstart, tf, s, mx, Ex, alpha, dr, Mdisk, rc, Nx, rhos, T0, betaT, mu, \
             Mstar, sigma, npts, nptsin, ti, gammadflag, returnall)
-            
     
-    for i in range(nt - 1):
+    if sf[-1] != 0.0 or af[-1] < rin:        
+    
+        for i in range(nt):
         
         
-        Mdots_in = -2 * np.pi * (rin * AU) * rdot_with_acc(rin, t[i], s, alpha, dr, Mdisk, rc, T0, betaT, rhos, mu, \
-            Mstar, gammadflag, sigma) * func(t[i], rin)
-        Mdots_out = -2 * np.pi * (rout * AU) * rdot_with_acc(rout, t[i], s, alpha, dr, Mdisk, rc, T0, betaT, rhos, mu, \
-            Mstar, gammadflag, sigma) * func(t[i], rout)
-        dMdots = Mdots_out - Mdots_in
+            #Mdots_in = -2 * np.pi * (rin * AU) * rdot_with_acc(rin, t[i], s, alpha, dr, Mdisk, rc, T0, betaT, rhos, mu, \
+            #    Mstar, gammadflag, sigma) * func(t[i], rin)
+            #Mdots_out = -2 * np.pi * (rout * AU) * rdot_with_acc(rout, t[i], s, alpha, dr, Mdisk, rc, T0, betaT, rhos, mu, \
+            #    Mstar, gammadflag, sigma) * func(t[i], rout)
[…]

### 2015-04-01 python/drift_timescales_updated_sigmap_0401.py
update

@@ -1126,7 +1126,7 @@ def dMdot_gas(rin, rout, t, alpha, Mdisk = 0.1*Msun, mu = 2.35, T0 = 120, betaT 
                              
                                            
         
-def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, npts_at_des, Mdisk = 0.1*Msun, rc = 100*AU, dr = 1e-3, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, \
+def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, Mdisk = 0.1*Msun, rc = 100*AU, dr = 1e-3, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, \
     Mstar = Msun, sigma = 2 * 10**(-15), dusttogas = 0.01, gammadflag = 0, sigmad_dt = 1, f = 0.9*1e-4, Nx = 1e15, returnall = True, \
     npts = 1e6, nptsin = 1e4, tin = 1e-10, lin = 0):
     
@@ -1163,7 +1163,7 @@ def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, npts_at_des, Mdisk = 
             Ms = np.append(Ms, quad(fint, rin*AU, rout*AU)[0])
             
  
-        return Mg, Ms, Mg_tot
+        return Mg, Mg_tot, Ms
             
     elif af[-1] >= rin and af[-1] <= rout:
         print "It is desorbing."
@@ -1179,46 +1179,46 @@ def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, npts_at_des, Mdisk = 
         k = i
         print k
             
-        for i in range(k):
-            Mg = np.append(Mg, 0)
+        #for i in range(k):
+        #    Mg = np.append(Mg, 0)
         
-            def fint(x):
-                return 2 * np.pi * func(t[i], x/AU) * x
-            sig = np.append(sig, func(t[i], rin))
+        def fint(x):
+            return 2 * np.pi * func(tf[-1], x/AU) * x
+            #sig = np.append(sig, func(t[i], rin))
 
-            Ms = np.append(Ms, quad(fint, rin*AU, rout*AU)[0])
+        Ms_tot = quad(fint, rin*AU, rout*AU)[0]
             
             #sigmapdes = func(tf[-1], af[-1])
[…]

### 2015-04-03 python/drift_timescales_updated_sigmap_0402.py
update

@@ -1143,7 +1143,6 @@ def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, dtf = 0.001, Mdisk = 
     func = scipy.interpolate.interp2d(t, rgrid, sigarray)
     
     Ms, Mg, Mg_tot, Ms_tot, sig = [], [], [], [], []
-    #Desv, dMdotsv, dMdots_newv = [], [], []
     
     tf, af, sf = rf(rstart, tf, s, mx, Ex, alpha, dr, Mdisk, rc, Nx, rhos, T0, betaT, mu, \
             Mstar, sigma, npts, nptsin, ti, gammadflag, returnall)
@@ -1153,19 +1152,21 @@ def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, dtf = 0.001, Mdisk = 
         for i in range(nt):
             
             Mg = np.append(Mg, 0)
-            #Ms = np.append(Ms, dMdots * (t[i + 1] - t[i]) + func(t[i], rin) * np.pi * ((rout*AU)**2 - (rin*AU)**2))
+            
             def fint(x):
                 return 2 * np.pi * func(t[i], x/AU) * x
-            sig = np.append(sig, func(t[i], rin))
-            def fsig(x):
-                return 2 * np.pi * Sigmadisk(x/AU, t[i], alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag) * x
-            Mg_tot = np.append(Mg_tot, quad(fsig, rin*AU, rout*AU)[0])
             Ms = np.append(Ms, quad(fint, rin*AU, rout*AU)[0])
             
+            def fsig(x):
+                return 2 * np.pi * Sigmadisk(x/AU, t[i], alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag) * x
+            Mg_tot = np.append(Mg_tot, quad(fsig, rin*AU, rout*AU)[0])      
  
-        return Mg, Mg_tot, Ms
+        return Mg, Ms, t, Mg_tot
+        
+        
+        
             
-    elif af[-1] >= rin and af[-1] <= rout:
+    elif af[-1] >= rin:# and af[-1] <= rout:
         print "It is desorbing."
         
         sigmapdes = func(tf[-1], af[-1])
@@ -1179,39 +1180,27 @@ def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, dtf = 0.001, Mdisk = 
[…]

### 2015-04-03 python/drift_timescales_updated_sigmap_0402.py
update

@@ -1222,17 +1222,17 @@ def dMdot(rin, rout, ti, tf, nt, rstart, s, alpha, mx, Ex, dtf = 0.001, Mdisk = 
             #for i in range(npts_at_des, nt - 1):
                           
         if af[-1] > rout:
-            return np.append(Mg, Mg_after_des), \
-            np.append(Ms, Ms_after_des), \
-                t, \
-                np.append(Mg_tot, Mg_tot_bf_des)  
+            return np.append(np.append(Mg, Mg[-1]), Mg_after_des), \
+            np.append(np.append(Ms, Ms[-1]), Ms_after_des), \
+                np.append(np.append(t[:k + 1], tf[-1]), t[k+1:])#, \
+                    #np.append(Mg_tot, Mg_tot_bf_des)  
                 
         else:    
                               
             return np.append(np.append(Mg, Mg_at_des), Mg_after_des), \
                 np.append(np.append(Ms, Ms_at_des), Ms_after_des), \
-                    np.append(np.append(t[:k + 1], tf[-1]), t[k+1:]), \
-                        np.append(Mg_tot, Mg_tot_bf_des)  
+                    np.append(np.append(t[:k + 1], tf[-1]), t[k+1:])#, \
+                       # np.append(Mg_tot, Mg_tot_bf_des)  
         
                                         
 

### 2015-04-06 tex/draft/ms.tex
update

@@
 %\shorttitle{Critical Core Mass at Wide Separations}
-\shortauthors{Piso et al.}
+%\shortauthors{Piso et al.}
 
-\title{The Influence of Radial Drift and Gas Accretion on the C/O Ratio in Protoplanetary Disks}
-%\title{Lower Limits on the Core Mass for Giant Planets}
-%\title{The Critical Core Mass of Wide-Separation Giant Planets:  Lower Limits from Kelvin-Helmholtz Cooling}
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}, Tilman Birnstiel\altaffilmark{1}}
+\title{C/O in Protoplanetary Disks: The Effect of Radial Drift and Viscous Accretion}
+
+\author{Ana-Maria A. Piso\altaffilmark{1}, et al}. %Karin I. \"Oberg\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}, Tilman Birnstiel\altaffilmark{1}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
+%\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
 
-%\begin{abstract}
-%\end{abstract}
+\begin{abstract}
+...
+\end{abstract}
 
@@
 
-\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al.
+%\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al.
 2011).}
 
-\emgr{Describe goals and approach of this paper, i.e.: study the importance of radial drift on the location of snowlines in protoplanetary disks and how radial drift affects the C/O ratio.
+Main sequence stars commonly host giant planets (refs).
+The chemical composition of gas giant atmospheres can provide important constraints on their formation, accretion and migration history.
+In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11b}, \citealt{oberg11c}, ...).
+Of particular importance are volatile compounds, since the location of their snowlines determines their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
+
+Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio.
+Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}.
+One explanation for this discrepancy was proposed by \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e.
+H$_2$O, CO$_2$ and CO, have different condensation temperatures.
+This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
+\citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
+
+In order to obtain more realistic estimates for C/O ratios across protoplanetary disks, dynamical processes and the disk evolving chemistry have to be taken into account.
+In this paper, we enhance the model of \citet{oberg11} considering two additional dynamic effects: (1) the radial drift of solids throughout the protoplanetary disk, and (2) the viscous accretion of the disk gas onto the host star.
+Our goal is two-fold: (1) to quantify the effect of radial drift of solids of different sizes on the location and shape of H$_2$O, CO$_2$ and CO snowlines, and (2) to calculate the resulting C/O ratio in gaseous and solid form throughout an actively accreting protoplanetary disk as a function of the grain size distribution and the evolutionary time of the nebula.
+
+%\emgr{Describe goals and approach of this paper, i.e.: study the importance of radial drift on the location of snowlines in protoplanetary disks and how radial drift affects the C/O ratio.
 Mention that disks are viscously accreting in the early stages of planet formation, hence gas accretion has to be taken into account.
@@
 
-\section{Radial Drift Model}
-\subsection{Disk and Desorption Model}
+\section{Model Assumptions} 
+\label{sec:model}
+
+We describe our analytic model for the radial drift of solids in section \ref{sec:drift}.
+In section \ref{sec:disk}, we present our protoplanetary disk model, as well as our prescription for the gas and dust surface density evolution in an active disk.
+Finally, we summarize our ice desorption model in section \ref{sec:desorption}.
+
+\subsection{Radial Drift}
+\label{sec:drift}
+
+\subsection{Evolutionary Disk Model}
+\label{sec:disk}
+
+\subsection{Volatile Desorption}
+\label{sec:desorption}
+
+%\subsection{Disk and Desorption Model}
 

### 2015-04-06 python/advection_from_Til/two_pop_model.py
update

@@ -297,14 +297,14 @@ def two_pop_model_run(x_1,a_0,timesteps_1,sigma_g_1,sigma_d_1,v_gas_1,T_1,alpha_
         # set the time step
         #
         dt = timesteps_1[it_old]-t 
-        dt = min(dt*10,timesteps_1[it_old]-t)
-        if t != 0.0: dt = min(dt,t/200)
-        if dt==0:
-            print('ERROR:')
-            print('t      = %g years'%(t/year))
-            print('it_old = %g'%it_old)
-            print('dt = 0')
-            sys.exit(1)
+        #dt = min(dt*10,timesteps_1[it_old]-t)
+        #if t != 0.0: dt = min(dt,t/200)
+        #if dt==0:
+        #    print('ERROR:')
+        #    print('t      = %g years'%(t/year))
+        #    print('it_old = %g'%it_old)
+        #    print('dt = 0')
+        #    sys.exit(1)
         #
         # calculate the velocity
         #
@@ -333,14 +333,14 @@ def two_pop_model_run(x_1,a_0,timesteps_1,sigma_g_1,sigma_d_1,v_gas_1,T_1,alpha_
         ###
         ### try variable time step
         ###
-        while any(u_out[2:-1][mask]/x_1[2:-1][mask]>=1e-30):
-            dt = dt/10.
-            if dt<1.0:
-                print('ERROR: time step got too short')
-                sys.exit(1)
-            #u_out = impl_donorcell_adv_diff_delta(n_r,x_1,D,v,g,h,K,L,flim,u_in,dt,1,1,0,0,0,0,1,A0,B0,C0,D0)
-            u_out = impl_donorcell_adv_diff_delta(n_r,x_1,D,v,g,h,K,L,flim,u_in,dt,0,0,1,1,1e-100*x_1[0],1e-100*x_1[-1],1,A0,B0,C0,D0)
-            mask = abs(u_out[2:-1]/u_in[2:-1]-1)>0.3
+        #while any(u_out[2:-1][mask]/x_1[2:-1][mask]>=1e-30):
+        #    dt = dt/10.
+        #    if dt<1.0:
[…]

### 2015-04-06 python/advection_from_Til/two_pop_test_ana_0115_1.py

@@ -27,7 +27,7 @@ m_star        = M_star*ones(n_t)
 T_COAG_START  = 0.0
 RHO_S         = 1.6
 V_FRAG        = 0 #1000
-a_0           = 1e-1
+a_0           = 1e5
 E_drift       = 1.0
 #
 # set the initial surface density & velocity

### 2015-04-06 python/drift_timescales_Til_parameters.py

@@ -181,10 +181,15 @@ def lambdamfp(r, Mdisk, Rstar, Tstar, rc, mu = 2.35, Mstar = Msun, sigma = 2e-15
     return 1 / (np.sqrt(2) * sigma * \
           (rhodisk(r, Mdisk, Rstar, Tstar, rc, mu, Mstar) / (mu * mp)))
 
+#def eta(r, Mdisk, Rstar, Tstar, rc, dr, mu = 2.35, Mstar = Msun):
+#    """Power law coefficient in P \propto r^(-n)"""
+#    return - (r * AU) / (2 * rhodisk(r, Mdisk, Rstar, Tstar, rc, mu, Mstar) * (r * AU * Omegak(r, Mstar))**2) \
+#            * (Pdisk(r + dr, Mdisk, Rstar, Tstar, rc, mu, Mstar) - Pdisk(r, Mdisk, Rstar, Tstar, rc, mu, Mstar)) / (dr * AU)
+            
 def eta(r, Mdisk, Rstar, Tstar, rc, dr, mu = 2.35, Mstar = Msun):
     """Power law coefficient in P \propto r^(-n)"""
-    return - (r * AU) / (2 * rhodisk(r, Mdisk, Rstar, Tstar, rc, mu, Mstar) * (r * AU * Omegak(r, Mstar))**2) \
-            * (Pdisk(r + dr, Mdisk, Rstar, Tstar, rc, mu, Mstar) - Pdisk(r, Mdisk, Rstar, Tstar, rc, mu, Mstar)) / (dr * AU)
+    return  Pdisk(r, Mdisk, Rstar, Tstar, rc, mu, Mstar) / \
+        (2 * rhodisk(r, Mdisk, Rstar, Tstar, rc, mu, Mstar) * (r * AU * Omegak(r, Mstar))**2) * 2.75
        
             
 #def eta(r, Mdisk, Tstar, Rstar, dr, rc, mu = 2.35, Mstar = Msun):
@@ -283,7 +288,7 @@ def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, Rstar, Tstar, rc, rho
             dr = r[-1] - r[-2]
         v = np.append(v, rdot_with_acc(r[i], s, alpha, Mdisk, Rstar, Tstar, rc, dr, rhos, mu, Mstar, sigma))
         Sigmad = np.append(Sigmad, Sigmadisk(r[i], Mdisk, rc))
[…]

### 2015-04-06 python/drift_timescales_updated_sigmap_0402.py

@@ -534,7 +534,11 @@ def eta(r, t, alpha, dr, Mdisk, rc = 100*AU, T0=120, betaT=3./7, mu = 2.35, Msta
     return - (r * AU) / (2 * rhodisk(r, t, alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag) * (r * AU * Omegak(r, Mstar))**2) \
             * (Pdisk(r+dr, t, alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag) - \
                     Pdisk(r, t, alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag)) / (dr * AU)
-       
+
+
+    return  Pdisk(r, t, alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag) / \
+        (2 * rhodisk(r, t, alpha, Mdisk, rc, T0, betaT, mu, Mstar, gammadflag) * (r * AU * Omegak(r, Mstar))**2) * 2.75       
+                     
 
 def ts(r, t, s, alpha, dr, Mdisk, rc = 100*AU, T0=120, betaT=3./7, rhos = 3.0, mu = 2.35, \
     Mstar = Msun, gammadflag = 0, sigma = 2 * 10**(-15)):
@@ -854,6 +858,19 @@ def r_freeze(mx, Ex, nx, T0 = 120., betaT = 3./7):
           return Tdisk(x, T0, betaT) - T_freeze(mx, Ex, nx)
 
      return brentq(f, 0.1, 100)
+     
+
+def r_stop(t, s, alpha, dr, Mdisk, mx, Ex, rc = 100*AU, T0=120, betaT=3./7, rhos = 3.0, mu = 2.35, \
+    Mstar = Msun, gammadflag = 0, sigma = 2 * 10**(-15), Nx = 1e15):
+
+     def f(r):
+          return (r*AU) / (- rdot_with_acc(r, t, s, alpha, dr, Mdisk, rc, T0, betaT, rhos, mu, \
+            Mstar, gammadflag, sigma)) - tdes(mx, Ex, Tdisk(r, T0, betaT), s, Nx, rhos)
+
+     try:
+          return brentq(f, 1e-3, 1e2)
+     except ValueError:
+          return 1e-10
 
 
[…]

### 2015-06-05 tex/draft/ms.tex
update

@@
 
-%\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al.
+\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al.
 2011).}
@@
 \section{Results for the C/O Ratio}
-\emgr{Present the flux equations used to keep track of the amount of C and O in gas and dust throughout the disk.
+
+\emgr{Present the plots analogous to Fig.
+1 in \"Oberg et al.
+(2011) for different particle sizes, based on the snowline locations obtained in the previous section, both for passive and active disk.
+Perhaps show it at different times in the gas disk evolution (i.e., not just at 3 Myr) for the active disk.
+Then assume a particle size distribution and show the interpolated result for the C/O ratio.
+Generalize the result using a transition disk (this part might also fit in the discussion section).}
+
+%emgr{Present the flux equations used to keep track of the amount of C and O in gas and dust throughout the disk.
 Motivate the simplification of using a fixed particle sized, fixed timescale, and fixed desorption distance in the calculations by referring to the rainbow plot in the previous section and by inserting the plot that shows that particle desorb almost instantly at a fixed distance.
@@
 
-\emgr{Include: non-inclusion of turbulence, assumption of perfect spheres when in fact they may have cracks, particles composed of a single volatile when in reality they are likely to be mixed, etc.
+\emgr{Present the diagram that shows all the effects that can modify snowline location.
+For model limitations, include: non-inclusion of turbulence, assumption of perfect spheres when in fact they may have cracks, particles composed of a single volatile when in reality they are likely to be mixed, etc.
 Discuss uncertainty of initial conditions and estimate how much they matter.

### 2015-06-05 python/C_to_O_new.py

@@ -1,0 +1,133 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+from drift_timescales import read_drag_file_many
+
+def n_C_O(species):
+    
+    """
+        
+    Returns the abundance of C and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_C:
+        abundance of C with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 1.5, 1.5
+    elif species == 'CO2':
+        return 0.3, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0.6, 0
+    elif species == 'silicate':
+        return 0, 1.4
+
+
+#in the following we calculate T_freeze for various species
+s, a, arrayH2O = read_drag_file_many('s_neg3_8_varying_sigmap.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8_varying_sigmap.txt', 'CO2', 50, 50)
+s, a, arrayCO = read_drag_file_many('s_neg3_8_varying_sigmap.txt', 'CO', 50, 50)
+
+
+
+
+def n(index, r, elem):
+    
+    """
+    
+    Returns the gaseous and solid abundance of a given element (C, O)
+    as a function of temperature, e.g. between different snowlines
+    
+    Input
+    -----
+    T:
+        disk midplane temperature in units of K
+    elem:
+        string: 'C' or 'O'
+        
+    Output
+    ------
+    n_gas:
+        abundance of element 'elem' in gaseous form relative to H2 abundance
+    n_solid:
+        abundance of element 'elem' in solid form relative to H2 abundance
+    
+    """
+    
+    if elem == 'C':
+        i = 0 #flag to simplify the use of the n_C_O function; = 0 for C
+    elif elem == 'O':
+        i = 1 #flag to simplify the use of the n_C_O function; = 1 for O
+    
+    size = s[index]
+    aH2O = arrayH2O[index][-1]
[…]

### 2015-06-05 tex/draft/ms.tex
update

@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/drift_timescales_betaS1_gas_acc_new.pdf}
+%\vspace{-0.5in}
+\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm evap}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
+The timescales are calculated at three representative locations in the disk: ...
+AU (blue lines), ...
+AU (green lines) and ...
+AU(red lines).
+The horizontal dashed line represents a typical disk lifetime of 3 Myr.
+Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
+$t_{\rm r, drift}$ and $t_{\rm gas, acc}$ are comparable to the desorption timescale $t_{\rm evap}$.
+\textbf{Figure belongs to section 2.4.} (\emgr{We will probably want to change the notation on the plot to $t_{des}$ instead of $t_{\rm evap}$ for consistency.})}  %  (See text for a description of evolution to yet higher masses.)
+\label{fig:timescales}
+\end{figure}
+
 \emgr{Calculate timescale for radial drift following Chiang \& Youdin (2010).
@@
 \section{Snowline Locations}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_passive_active.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance as a function of a particle's initial location in the disk, for a range of particle sizes, and for both a passive disk (left panels) and an active disk (right panels).
+The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+The particle size increases from $10^{-3}$ cm to $10^8$ cm in a rainbow color scheme.
+For a range of particle sizes, the desorption distance is the same regardless of the particles' initial location.
+\textbf{Figure belongs to section 3.}}  %  (See text for a description of evolution to yet higher masses.)
+\label{fig:snowlines}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines} and text).
+The estimate is performed for a passive disk (left panels) and an active disk (right panels).
+The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+The analytic approximation is in good agreement with the numerical result for most cases.
+\textbf{Figure belongs to section 3.}}  %  (See text for a description of evolution to yet higher masses.)
+\label{fig:an_vs_actual}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_vs_s.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance as a function of initial particle size, for the range of particles that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines}).
+\textbf{Figure belongs to section 3.} \emgr{This is a figure that I'm not sure we should include, since it's not really provide that much useful inside.
+If we did include it though, there would be two panels both for the passive and the active disk.
+The caption is not complete since the active disk panel is not yet there.}}  %  (See text for a description of evolution to yet higher masses.)
+\label{fig:an_vs_actual}
+\end{figure}
 
@@
 
-\emgr{Present the plots analogous to Fig.
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/s_t_a.pdf}
+%\vspace{-0.5in}
+\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in a passive disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left).
+Particles desorb almost instantaneously.
+Right panel: fractional mass lost by the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
+Particles lose most of their mass very close to the distance at which they fully desorb.
+\textbf{Figure belongs to section 4.}}
+\label{fig:s_t_a}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_active_disk_many.pdf}
+%\vspace{-0.5in}
+\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (left panel \emgr{currently non-existent}) and for an active disk (right panel \emgr{currently just a single plot}), and for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
+The particle size increases from 0.001 cm to 700 cm in a rainbow color scheme.
+The horizontal dotted line represents the stellar value of 0.54.
+The black lines (\emgr{currently non-existent}) represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk.
+The snowline location moves inward as the particle size increases.
+\textbf{Figure belongs to section 4.}}
+\label{fig:CO_ratio}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{lala}
+%\vspace{-0.5in}
+\caption{C/O ratios for different particle size distributions...
+\textbf{Figure belongs to section 4.} (\emgr{It will probably be a multi-panel plot, for different particle size distributions, for both passive and active disk, and perhaps at different times for the active disk.
+Not yet unsure how that figure will look like since I don't have the results yet, so I'll leave the figure caption open for now.})}
+\label{fig:...}
+\end{figure}
+
+\emgr{Motivate the fact that we can calculate a sharp, fixed snowline for each particle size by inserting the plot that shows that particles desorb almost instantaneously at a fixed distance.
+Then present the plots analogous to Fig.
 1 in \"Oberg et al.

### 2015-06-05 python/C_to_O_new.py

@@ -83,20 +83,20 @@ def n(index, r, elem):
                 
     if r <= aH2O:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('H2O')[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i]])
+            n_C_O('C_grains')[i] + n_C_O('silicate')[i]]), size
             
     elif aH2O <= r <= aCO2:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i], \
-            n_C_O('H2O')[i] + n_C_O('C_grains')[i] + n_C_O('silicate')[i]])
+            n_C_O('H2O')[i] + n_C_O('C_grains')[i] + n_C_O('silicate')[i]]), size
             
     elif aCO2 < r < aCO:
         return np.array([n_C_O('CO')[i], \
             n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
-                n_C_O('CO2')[i]])
+                n_C_O('CO2')[i]]), size
                 
     elif aCO < r:
         return np.array([0, n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
-                n_C_O('CO2')[i] + n_C_O('CO')[i]])
+                n_C_O('CO2')[i] + n_C_O('CO')[i]]), size
             
             
 
@@ -126,7 +126,7 @@ def C_O_ratio(r, index):
     
     """
     
-    return n(index, r, 'C') / n(index, r, 'O')
+    return n(index, r, 'C')[0] / n(index, r, 'O')[0], n(index, r, 'O')[1]
                        
             
             

### 2015-06-06 tex/draft/ms.tex
update

@@
-%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
+\documentclass[12pt, preprint,numberedappendix]{emulateapj}
 %\documentclass[12pt, preprint]{aastex}
-\documentclass[apj]{emulateapj}
+%\documentclass[apj]{emulateapj}
 
@@
 Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
-$t_{\rm r, drift}$ and $t_{\rm gas, acc}$ are comparable to the desorption timescale $t_{\rm evap}$.
+$t_{\rm r, drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm evap}$.
 \textbf{Figure belongs to section 2.4.} (\emgr{We will probably want to change the notation on the plot to $t_{des}$ instead of $t_{\rm evap}$ for consistency.})}  %  (See text for a description of evolution to yet higher masses.)
@@
 \caption{Desorption distance as a function of initial particle size, for the range of particles that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines}).
-\textbf{Figure belongs to section 3.} \emgr{This is a figure that I'm not sure we should include, since it's not really provide that much useful inside.
+\textbf{Figure belongs to section 3.} \emgr{This is a figure that I'm not sure we should include, since it doesn't really provide that much useful inside.
 If we did include it though, there would be two panels both for the passive and the active disk.

### 2015-06-10 tex/draft/ms.tex
update

@@
-\documentclass[12pt, preprint,numberedappendix]{emulateapj}
+%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
 %\documentclass[12pt, preprint]{aastex}
-%\documentclass[apj]{emulateapj}
+\documentclass[apj]{emulateapj}
 
@@
 
-\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al.
+%\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al.
 2011).}
@@
 
-We describe our analytic model for the radial drift of solids in section \ref{sec:drift}.
-In section \ref{sec:disk}, we present our protoplanetary disk model, as well as our prescription for the gas and dust surface density evolution in an active disk.
-Finally, we summarize our ice desorption model in section \ref{sec:desorption}.
+We present our protoplanetary disk model for both a passive and an active disk in section \ref{sec:disk}.
+In section \ref{sec:drift}, we describe our analytic model for the radial drift of solids.
+We summarize our ice desorption model in section \ref{sec:desorption}.
+Finally, we discuss the relevant timescales for dynamical effects in the desorption process in section \ref{sec:timescales}.
+
+\subsection{Disk Model}
+\label{sec:disk}
+
+We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
+\begin{subeqnarray}
+\label{eq:disk}
+\Sigma_{\rm d}&=&2200\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
+T_{\rm d} &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
+\end{subeqnarray}
+where $r$ is the semimajor axis.
+Based on some observations of protoplanetary disks \citep{andrews10}, our surface density profile, $\Sigma_{\rm d} \propto r^{-1}$, is flatter than that of \citet{chiang10}, i.e.
+$\Sigma_{\rm d} \propto r^{-3/2}$.
+
+We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
+\begin{equation}
+\label{eq:nu}
+\nu=\alpha c_{\rm d} H_{\rm d}.
+\end{equation}
+Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient, and $c_{\rm d}$, $H_{\rm d}$ are the isothermal sound speed and disk scale height, respectively:
+\begin{subeqnarray}
+\label{eq:cdHd}
+c_{\rm d} &=& \sqrt{\frac{k_{\rm B} T_{\rm d}}{\mu m_{\rm p}}} \slabel{eq:cd} \\
+H_{\rm d}&=& \frac{c_{\rm d}}{\Omega_{\rm k}} \slabel{eq:Hd},
+\end{subeqnarray}
+where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k}$ is the Keplerian angular velocity, $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$, with $G$ the gravitational constant and $M_*$ the stellar mass.
+The temperature profile for an active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}).
+From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters.
+Following \citet{armitage10}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius.
+The gas surface density is then given by the self-similar solution
+\begin{equation}
+\label{eq:Sigmaact}
+\Sigma_{\rm d}(R, T) = \frac{M_{\rm d}}{2 \pi r_{\rm c}^2 R^{\gamma}} T^{(-5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{-(2-\gamma)}}{T}}\Big],
+\end{equation}
+where $M_{\rm d}$ is the total disk mass and
+\begin{subeqnarray}
+\label{eq:T}
+T & \equiv & \frac{t}{t_{\rm c}} + 1 \\
+t_{\rm c} & \equiv & \frac{1}{3(2-\gamma)} \frac{r_{\rm c}^2}{\nu_{\rm c}},
+\end{subeqnarray}
+where $t$ is time.
 
@@
 
-\subsection{Evolutionary Disk Model}
-\label{sec:disk}
+Solid particles in a disk orbit their host star at the Keplerian velocity $v_{\rm k} \equiv \Omega_{\rm k} r$.
+The gas, however, experiences an additional pressure gradient, which causes it to rotate at sub-Keplerian velocity \citep{weidenschilling77}.
+Dust grains thus experience a headwind, which removes angular momentum, causing the solids to spiral inwards and fall onto the host star.
+Small particles are well-coupled to the gas, while large planetesimals are decoupled from the gas.
+From \citet{chiang10}, the extent of coupling is quantified by the dimensionless stopping time, $\tau_{\rm s} \equiv \Omega_{\rm k} t_{\rm s}$, where $t_{\rm s}$ is
+\begin{equation}
+\label{eq:ts}
+t_{\rm s}= \left\{
+\begin{array}{l l}
+\rho_{\rm s} s / (\rho_{\rm d} c_{\rm d}), & \quad s < 9 \lambda/4 \,\,\,\ \text{Epstein drag} \\
+4 \rho_{\rm s} s^2 / (9 \rho_{\rm d} c_{\rm d} \lambda), & \quad s < 9 \lambda/4, \,\text{Re} \lesssim 1 \,\,\,\ \text{Stokes drag.}
+\end{array} 
+\right.
+\end{equation}
+Here $\rho_{\rm d}$ is the gas midplane density, $\rho_{\rm s}$ is the density of a solid particle, $s$ is the particle size, $\lambda$ is the mean free path, and Re is the Reynolds number.
+
+For a passive disk, the radial drift velocity can be approximated analytically as
+\begin{equation}
+\label{eq:rdotpas}
+\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big),
+\end{equation}
+where
+\begin{equation}
+\label{eq:eta}
+\eta \equiv - \frac{\partial P_{\rm d}/\partial \ln r}{2 \rho_{\rm d} v_{\rm k}^2} \approx \frac{c_{\rm d}^2}{2 v_k^2}
+\end{equation}
+and $P_{\rm d} = \rho_{\rm d} c_{\rm d}^2$ is the disk midplane pressure.
+
+For an active disk, the radial drift velocity has an additional term due to the radial movement of the gas, i.e.
+\begin{equation}
+\label{eq:rdotact}
+\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{v_{\rm gas}}{1+\tau_{\rm s}^2},
+\end{equation}
+where $v_{\rm gas}$ is the radial gas accretion velocity and can be expressed as (e.g., \citealt{fkr02})
+\begin{equation}
+\label{eq:vgas}
+v_{\rm gas}= - \frac{3}{\Sigma_{\rm d} \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma_{\rm d} \sqrt{r}) 
+\end{equation}
+with $\Sigma_{\rm d}$ from Equation (\ref{eq:Sigmaact}).
 
@@
 
-\emgr{Describe disk model --- both passive MMSN and active self-similar solution with $\alpha$ prescription for viscosity.
+In order for a volatile species to thermally desorb, it has to overcome the binding energy that keeps it on the grain surface.
+Following \citet{hollenbach09}, the desorption rate per molecule for a species $x$ can be expressed as
+\begin{equation}
+\label{eq:Rdes}
+R_{\rm{des}, x} = \nu_x \exp{(-E_x/T_{\rm grain})},
+\end{equation}
+where $E_x$ is the adsorption binding energy, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ is the molecule's vibrational frequency, with $\mu_x$ the mean molecular weight.
+We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T_{\rm d}$.
+For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively \citep{oberg11}.
+We use the desorption rate, $R_{\rm des}$, to estimate the desorption timescale for particles of different sizes as described in section \ref{sec:timescales}.
+
+%\emgr{Describe disk model --- both passive MMSN and active self-similar solution with $\alpha$ prescription for viscosity.
 Describe the model for evolving the surface density profile of planetesimals following Birnstiel et al.
@@
 \subsection{Relevant timescales}
+\label{sec:timescales}
+
+We can estimate the extent to which radial drift and gas accretion affect desorption by comparing the timescales for desorption, drift and accretion, for solids of different sizes and compositions.
+
+\textit{Desorption timescale.} We assume that the solid bodies are perfect spheres and are entirely composed of only one volatile species, i.e.
+either H$_2$O, CO$_2$ or CO \footnote{We discuss the validity of these simplifications in section \ref{sec:discussion}.}.
+The desorption timescale can then be estimated as
+\begin{equation}
+\label{eq:tdes}
+t_{\rm des}=\frac{\rho_{\rm s}}{3 \mu_x m_{\rm p}} \frac{s}{N_x R_{\rm des, x}},
+\end{equation}
+where $N_x \approx 10^{15}$ sites cm$^{-2}$ is the number of adsorption sites of volatile $x$ per cm$^2$ \citep{hollenbach09}.
+
+\textit{Radial drift timescale.} To order of magnitude, the radial drift timescale can be estimated as 
+\begin{equation}
+\label{eq:tdrift}
+t_{\rm r, drift} \sim \frac{r}{\dot{r}},
+\end{equation}
+where $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for a passive disk and by Equation (\ref{eq:rdotact}) for an active disk.
+
+\textit{Gas accretion timescale.} The timescale for gas accretion onto the central star for an active disk is (e.g., \citealt{armitage10})
+\begin{equation}
+\label{eq:tgas}
+t_{\rm gas, acc} \sim \frac{r^2}{\nu} \sim \frac{1}{2 \alpha \eta \Omega_{\rm k}},
+\end{equation}
+with the latter expression derived from Equations (\ref{eq:nu}) and (\ref{eq:eta}).
+
+For simplicity purposes, we calculate the radial drift timescale, $t_{\rm r, drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well.
+Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm r, drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, for $\rho_{\rm s}=2$ g cm$^{-3}$ and $\alpha=0.01$.
+As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years and their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude.
+Thus for small particles in a passive disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}) \footnote{This is not true for an active disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.}.
+At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well.
+This is true for both passive and active disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.
+
+Of particular interest for our purposes is the particle size regime for which $t_{\rm r, drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ or $t_{\rm gas, acc} \lesssim t_{\rm des} \lesssim t_{\rm d}$, where $t_{\rm d}=3$ Myr is the disk lifetime.
+In these cases, radial drift or gas accretion (or both) are faster than thermal desorption.
+Particles of sizes that satisfy these requirements will drift significantly due to radial drift or gas accretion before desorbing, thus moving the H$_2$O, CO$_2$ and CO snowlines closer towards the central star and changing the C/O ratio throughout the disk.
+We quantify these effects in sections \ref{sec:snowlines} and \ref{sec:COratio}.
 
@@
 
-\emgr{Calculate timescale for radial drift following Chiang \& Youdin (2010).
+%\emgr{Calculate timescale for radial drift following Chiang \& Youdin (2010).
 Calculate desorption timescale following Hollenbach et al.
@@
 \section{Snowline Locations}
+\label{sec:snowlines}
+
+In this section we use the model described in section \ref{sec:model} to quantify the effects of radial drift (passive disk) or radial drift and gas accretion (active disk) on the snowline location, for dust particles of different sizes composed of either H$_2$O, CO$_2$ or CO.
+Specifically, we determine a particle's final location (i.e., where the particle either fully desorbs or remains at its initial size due to a long desorption timescale) as a function of its initial position in the disk, after the gas disk has dissipated.
+The disk lifetime, $t_{\rm d}$, is particularly relevant since this is the timescale on which giant planets form.
+The snowline locations at $t=t_{\rm d}$ throughout the protoplanetary disk determine the disk C/O ratio in gas at this time, and thus the C/O ratio in giant planet atmospheres that have formed \textit{in situ}.
+
+As a fiducial model, we choose $M_*=M_{\odot}$, $M_{\rm d}=0.1 M_{\odot}$ and $r_{\rm c}=100$ AU.
+For each species $x$, we determine the final location in the disk of a particle of initial size $s_0$ by solving the following system of coupled differential equations:
+
+\begin{subeqnarray}
+\label{eq:ddt}
+\frac{ds}{dt} &= & - \frac{3 \mu_x m_{\rm p}}{\rho_{\rm s}} N_x R_{\rm des, x}  \slabel{eq:dsdt} \\
+\frac{dr}{dt} &=& \dot{r},
+\end{subeqnarray}
+where $R_{\rm des, x}$ is evaluated at $T=T_{\rm d}(r)$, and $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for a passive disk and Equation (\ref{eq:rdotact}) for an active disk.
+Equation (\ref{eq:dsdt}) can be derived straightforwardly from Equation (\ref{eq:tdes}).
+Our boundary conditions are $s(t_0)=s_0$, $r(t_0)=r_0$, and $s(t_{\rm d})=0$, where $t_0$ is the initial time at which we start the integration and $r_0$ is the initial location of the particle.
+We choose $t_0=1$ year, but our result is independent on the initial integration time as long as $t_0 \ll t_{\rm d}$.
+
+Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, and for both a passive and an active disk.
+The final distance of the dust particles is consistent with Figure \ref{fig:timescales}.
+Kilometer-sized bodies do not drift or desorb during the disk lifetime both for a passive and active disk.
+Micron- to mm-sized particles do not drift before desorbing in a passive disk, but they drift significantly in an active disk since they move at the same velocity as the accreting gas.
+For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in a passive disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an active disk, we notice that particles of initial size $s_0$ desorb at a fixed distance $r_{\rm des}$ regardless of their original location in the disk.
+We show in section \ref{sec:COratio} that this result is essential in determining the C/O ratio throughout the disk for different particle sizes.
 
@@
 
+Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm r, drift} \sim t_{\rm des}$, given an initial particle size.
+Figure \ref{fig:an_vs_actual} shows the validity of this approximation for the range of particle sizes that desorb at a fixed distance in a passive and an active disk.
+We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest.
+
 \begin{figure}[h!]
@@
 
+%\emgr{I didn't write the text for Figure 4 since we haven't yet decided if in the end we will include it in some form. We can }
+
 \begin{figure}[h!]
@@
 The caption is not complete since the active disk panel is not yet there.}}  %  (See text for a description of evolution to yet higher masses.)
-\label{fig:an_vs_actual}
+\label{fig:r_vs_s}
 \end{figure}
 
-\emgr{Present the equation set that you are solving, $dr/dt=\dot{r}$, $ds/dt=...$.
+%\emgr{Present the equation set that you are solving, $dr/dt=\dot{r}$, $ds/dt=...$.
 I don't think it is necessary to describe in detail the numerical method of solving the equations since it's pretty straightforward.
@@
 \section{Results for the C/O Ratio}
+\label{sec:COratio}
 
@@
 \caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (left panel \emgr{currently non-existent}) and for an active disk (right panel \emgr{currently just a single plot}), and for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
-The particle size increases from 0.001 cm to 700 cm in a rainbow color scheme.
+The particle size increases from 0.001 cm to $\sim$700 cm in a rainbow color scheme.
 The horizontal dotted line represents the stellar value of 0.54.
@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{lala}
-%\vspace{-0.5in}
-\caption{C/O ratios for different particle size distributions...
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=0.5\textwidth]{lala}
+%%\vspace{-0.5in}
+%\caption{C/O ratios for different particle size distributions...
 \textbf{Figure belongs to section 4.} (\emgr{It will probably be a multi-panel plot, for different particle size distributions, for both passive and active disk, and perhaps at different times for the active disk.
 Not yet unsure how that figure will look like since I don't have the results yet, so I'll leave the figure caption open for now.})}
-\label{fig:...}
-\end{figure}
+%\label{fig:...}
+%\end{figure}
 
@@
 \section{Discussion and Model Limitations}
+\label{sec:discussion}
 

### 2015-06-10 tex/draft/ms.tex
update

@@
 
-\begin{figure}[h!]
+\begin{figure*}[tb]
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_passive_active.pdf}
+\includegraphics[width=\textwidth]{../../figs/desorption_distance_passive_active_colorbar.pdf}
 %\vspace{-0.5in}
@@
 The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The particle size increases from $10^{-3}$ cm to $10^8$ cm in a rainbow color scheme.
+The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
 For a range of particle sizes, the desorption distance is the same regardless of the particles' initial location.
@@
 \label{fig:snowlines}
-\end{figure}
+\end{figure*}
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_active_disk_many.pdf}
+\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar.pdf}
 %\vspace{-0.5in}
-\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (left panel \emgr{currently non-existent}) and for an active disk (right panel \emgr{currently just a single plot}), and for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
-The particle size increases from 0.001 cm to $\sim$700 cm in a rainbow color scheme.
+\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel) and for an active disk (bottom panel), and for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
+The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
 The horizontal dotted line represents the stellar value of 0.54.
-The black lines (\emgr{currently non-existent}) represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk.
+The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk.
 The snowline location moves inward as the particle size increases.

### 2015-06-10 python/C_to_O_active.py

@@ -1,0 +1,133 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+from drift_timescales import read_drag_file_many
+
+def n_C_O(species):
+    
+    """
+        
+    Returns the abundance of C and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_C:
+        abundance of C with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 1.5, 1.5
+    elif species == 'CO2':
+        return 0.3, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0.6, 0
+    elif species == 'silicate':
+        return 0, 1.4
+
+
+#in the following we calculate T_freeze for various species
+s, a, arrayH2O = read_drag_file_many('s_neg3_8_varying_sigmap.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8_varying_sigmap.txt', 'CO2', 50, 50)
+s, a, arrayCO = read_drag_file_many('s_neg3_8_varying_sigmap.txt', 'CO', 50, 50)
+
+
+
+
+def n(index, r, elem):
+    
+    """
+    
+    Returns the gaseous and solid abundance of a given element (C, O)
+    as a function of temperature, e.g. between different snowlines
+    
+    Input
+    -----
+    T:
+        disk midplane temperature in units of K
+    elem:
+        string: 'C' or 'O'
+        
+    Output
+    ------
+    n_gas:
+        abundance of element 'elem' in gaseous form relative to H2 abundance
+    n_solid:
+        abundance of element 'elem' in solid form relative to H2 abundance
+    
+    """
+    
+    if elem == 'C':
+        i = 0 #flag to simplify the use of the n_C_O function; = 0 for C
+    elif elem == 'O':
+        i = 1 #flag to simplify the use of the n_C_O function; = 1 for O
+    
+    size = s[index]
+    aH2O = arrayH2O[index][-1]
[…]

### 2015-06-10 python/C_to_O_passive.py

@@ -1,0 +1,133 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+from drift_timescales import read_drag_file_many
+
+def n_C_O(species):
+    
+    """
+        
+    Returns the abundance of C and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_C:
+        abundance of C with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 1.5, 1.5
+    elif species == 'CO2':
+        return 0.3, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0.6, 0
+    elif species == 'silicate':
+        return 0, 1.4
+
+
+#in the following we calculate T_freeze for various species
+s, a, arrayH2O = read_drag_file_many('s_neg3_8.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8.txt', 'CO2', 50, 50)
+s, a, arrayCO = read_drag_file_many('s_neg3_8.txt', 'CO', 50, 50)
+
+
+
+
+def n(index, r, elem):
+    
+    """
+    
+    Returns the gaseous and solid abundance of a given element (C, O)
+    as a function of temperature, e.g. between different snowlines
+    
+    Input
+    -----
+    T:
+        disk midplane temperature in units of K
+    elem:
+        string: 'C' or 'O'
+        
+    Output
+    ------
+    n_gas:
+        abundance of element 'elem' in gaseous form relative to H2 abundance
+    n_solid:
+        abundance of element 'elem' in solid form relative to H2 abundance
+    
+    """
+    
+    if elem == 'C':
+        i = 0 #flag to simplify the use of the n_C_O function; = 0 for C
+    elif elem == 'O':
+        i = 1 #flag to simplify the use of the n_C_O function; = 1 for O
+    
+    size = s[index]
+    aH2O = arrayH2O[index][-1]
+    aCO2 = arrayCO2[index][-1]
[…]

### 2015-06-10 python/drift_timescales_updated_sigmap_0402_cavity.py

@@ -334,7 +334,8 @@ def Sigmadisk(r, t, alpha, Mdisk, rc, T0 = 120, betaT = 3./7, mu = 2.35, Mstar =
         return Mdisk / (2 * np.pi * rc**2 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
             * np.exp(- rtild**(2 - gammad) / T)
      else:
-         return 1e-100
+         return Mdisk / (2 * np.pi * rc**2 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
+            * np.exp(- rtild**(2 - gammad) / T) / 10
                 
 
 def vacc_act(r, t, alpha, mu = 2.35, T0 = 120, betaT = 3./7, rc = 100 * AU, Mstar = Msun, gammadflag = 0):

### 2015-06-12 tex/draft/ms.tex
update

@@
 \centering
-\includegraphics[width=\textwidth]{../../figs/desorption_distance_passive_active_colorbar.pdf}
+\includegraphics[width=\textwidth]{../../figs/desorption_distance_passive_active_colorbar_test.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar.pdf}
+\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar_test.pdf}
 %\vspace{-0.5in}

### 2015-07-13 tex/draft/ms.tex
update

@@
 
-We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
+\textit{Passive disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
 \begin{subeqnarray}
@@
 
-We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
+\textit{Actively accreting disk with passive temperature profile.} We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
 \begin{equation}
@@
 \end{equation}
-Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient, and $c_{\rm d}$, $H_{\rm d}$ are the isothermal sound speed and disk scale height, respectively:
+Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c_{\rm d}$, $H_{\rm d}$ are the isothermal sound speed and disk scale height, respectively:
 \begin{subeqnarray}
@@
 where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k}$ is the Keplerian angular velocity, $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$, with $G$ the gravitational constant and $M_*$ the stellar mass.
+We choose $\mu=2.35$, corresponding to the Solar composition go hydrogen and helium.
 The temperature profile for an active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}).
@@
 Following \citet{armitage10}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius.
+We choose $r_{\rm c}=100$ AU.
 The gas surface density is then given by the self-similar solution
@@
 \end{subeqnarray}
-where $t$ is time.
+where $t$ is time. We take $M_{\rm d}=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
+
+\textit{Active disk steady-state solution.} Calculating the midplane temperature self-consistently for an active disk is a non-trivial solution, so instead we use the steady state solution with a constant mass flux $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ \citep{armitage10}.
+For consistency we choose $\alpha=0.01$ as in the previous case.
+
+\textit{Static disk.} For comparison purposes with \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}), but that not takes into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
 
@@
 \end{equation}
-Here $\rho_{\rm d}$ is the gas midplane density, $\rho_{\rm s}$ is the density of a solid particle, $s$ is the particle size, $\lambda$ is the mean free path, and Re is the Reynolds number.
+Here $\rho_{\rm d}$ is the gas midplane density, $\rho_{\rm s}=2$ g cm$^{-3}$ is the density of a solid particle, $s$ is the particle size, $\lambda$ is the mean free path, and Re is the Reynolds number.
 
@@
 For simplicity purposes, we calculate the radial drift timescale, $t_{\rm r, drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well.
-Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm r, drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, for $\rho_{\rm s}=2$ g cm$^{-3}$ and $\alpha=0.01$.
+Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm r, drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk.
 As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years and their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude.

### 2015-07-13 python/drift_timescales_active_Tdisk.py

@@ -1,0 +1,1251 @@
+from utils.constants import G, kb, mp, Msun, cmperau, AU, sigma, yr
+import numpy as np
+from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
+from T_freeze import T_freeze, Rdes, tevap, vib_freq
+from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
+import scipy
+from scipy.integrate import odeint, quad
+from scipy.optimize import brentq
+from scipy.interpolate import interp1d
+from utils.zbrac import zbrac
+from utils.utilities import tridag
+#from advection_from_Til.two-pop-py.src
+
+###################################################################
+
+def impl_donorcell_adv_diff_delta(n_x,x,Diff,v,g,h,K,L,flim,u_in,dt,pl,pr,ql,qr,rl,rr,coagulation_method,A,B,C,D):
+    """
+    Implicit donor cell advection-diffusion scheme with piecewise constant values
+    
+        Perform one time step for the following PDE:
+    
+           du    d  /    \    d  /              d  /       u   \ \
+           -- + -- | u v | - -- | h(x) Diff(x) -- | g(x) ----  | | = K + L u
+           dt   dx \    /    dx \              dx \      h(x) / /
+    
+        with boundary conditions
+    
+            dgu/h |            |
+          p ----- |      + q u |       = r
+             dx   |x=xbc       |x=xbc
+    Arguments:
+          n_x   = # of grid points
+          x     = the grid
+          Diff  = value of Diff @ cell center
+          v     = the values for v @ interface (array[i] = value @ i-1/2)
+          g     = the values for g(x)
+          h     = the values for h(x)
+          K     = the values for K(x)
+          L     = the values for L(x)
+          flim  = diffusion flux limiting factor at interfaces
+          u     = the current values of u(x)
+          dt    = the time step
+    
+    OUTPUT:
+          u     = the updated values of u(x) after timestep dt
+    
+    """
+    D05=zeros(n_x)
+    h05=zeros(n_x)
+    rhs=zeros(n_x)
+    #
+    # calculate the arrays at the interfaces
+    #
+    for i in arange(1,n_x):
[…]

### 2015-07-13 tex/draft/ms.tex
update

@@
 
-\textit{Passive disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
+\textbf{Passive disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
 \begin{subeqnarray}
@@
 
-\textit{Actively accreting disk with passive temperature profile.} We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
+\textbf{Actively accreting disk with passive temperature profile.} We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
 \begin{equation}
@@
 where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k}$ is the Keplerian angular velocity, $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$, with $G$ the gravitational constant and $M_*$ the stellar mass.
-We choose $\mu=2.35$, corresponding to the Solar composition go hydrogen and helium.
+We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition go hydrogen and helium.
 The temperature profile for an active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}).
@@
 
-\textit{Active disk steady-state solution.} Calculating the midplane temperature self-consistently for an active disk is a non-trivial solution, so instead we use the steady state solution with a constant mass flux $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ \citep{armitage10}.
-For consistency we choose $\alpha=0.01$ as in the previous case.
+\textbf{Active disk steady-state solution.} Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):
 
-\textit{Static disk.} For comparison purposes with \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}), but that not takes into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
+\begin{subeqnarray}
+\label{eq:activeT}
+T_{\rm d}^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
+\dot{M} & = & 3 \pi \nu \Sigma_{\rm d} \slabel{eq:Mdot},
+\end{subeqnarray}
+where $\dot{M}$, the mass flux, is constant for steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}.
+Based on disk observations (e.g., \citealt{andrews10}), we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$.
+We can then easily determine the radial temperature profile from Equation (\ref{eq:Tdact}), $c_{\rm d}$ and $H_{\rm d}$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
+For consistency, we choose  $\alpha=0.01$ as in the previous case.
+Finally, we determine $\Sigma_{\rm d}$ from Equation (\ref{eq:Mdot}).
+
+\textbf{Static disk.} For comparison purposes with \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}).
+The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
 
@@
 \label{eq:rdotact}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{v_{\rm gas}}{1+\tau_{\rm s}^2},
+\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{\dot{r}_{\rm gas}}{1+\tau_{\rm s}^2},
 \end{equation}
-where $v_{\rm gas}$ is the radial gas accretion velocity and can be expressed as (e.g., \citealt{fkr02})
+where $\dot{r}_{\rm gas}$ is the radial gas accretion velocity and can be expressed as (e.g., \citealt{fkr02})
 \begin{equation}
 \label{eq:vgas}
-v_{\rm gas}= - \frac{3}{\Sigma_{\rm d} \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma_{\rm d} \sqrt{r}) 
+\dot{r}_{\rm gas} = - \frac{3}{\Sigma_{\rm d} \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma_{\rm d} \sqrt{r}) 
 \end{equation}
 with $\Sigma_{\rm d}$ from Equation (\ref{eq:Sigmaact}).
+For the active disk steady-state solution (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma_{\rm d}$, with $\dot{M}$ fixed and $\Sigma_{\rm d}$ obtained from Equation (\ref{eq:Mdot}).
 
@@
 \end{equation}
-where $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for a passive disk and by Equation (\ref{eq:rdotact}) for an active disk.
+where $\dot{r}$ is the radial drift velocity given by Equation (\ref{eq:rdotpas}) for a passive disk and by Equation (\ref{eq:rdotact}) for an active disk.
 
@@
 Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm r, drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk.
-As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years and their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude.
+As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence disk location (see Equation \ref{eq:Rdes}).
+On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas.
 Thus for small particles in a passive disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}) \footnote{This is not true for an active disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.}.
@@
 \caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm evap}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
-The timescales are calculated at three representative locations in the disk: ...
-AU (blue lines), ...
-AU (green lines) and ...
-AU(red lines).
+The timescales are calculated at three representative locations in the disk, i.e.
+the H$_2$O, CO$_2$ and CO snowlines.
+For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively.
 The horizontal dashed line represents a typical disk lifetime of 3 Myr.
 Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
-$t_{\rm r, drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm evap}$.
-\textbf{Figure belongs to section 2.4.} (\emgr{We will probably want to change the notation on the plot to $t_{des}$ instead of $t_{\rm evap}$ for consistency.})}  %  (See text for a description of evolution to yet higher masses.)
+$t_{\rm r, drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.}
 \label{fig:timescales}
@@
 
-As a fiducial model, we choose $M_*=M_{\odot}$, $M_{\rm d}=0.1 M_{\odot}$ and $r_{\rm c}=100$ AU.
 For each species $x$, we determine the final location in the disk of a particle of initial size $s_0$ by solving the following system of coupled differential equations:
@@
 \frac{ds}{dt} &= & - \frac{3 \mu_x m_{\rm p}}{\rho_{\rm s}} N_x R_{\rm des, x}  \slabel{eq:dsdt} \\
-\frac{dr}{dt} &=& \dot{r},
+\frac{dr}{dt} &=& \dot{r} \slabel{eq:drdt},
 \end{subeqnarray}
-where $R_{\rm des, x}$ is evaluated at $T=T_{\rm d}(r)$, and $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for a passive disk and Equation (\ref{eq:rdotact}) for an active disk.
-Equation (\ref{eq:dsdt}) can be derived straightforwardly from Equation (\ref{eq:tdes}).
+where the desorption rate $R_{\rm des, x}$ for each particle type (i.e., composed of H$_2$O, CO$_2$ or CO) is evaluated at $T=T_{\rm d}(r)$, and the radial drift velocity $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for a passive disk and Equation (\ref{eq:rdotact}) for an active disk.
+Equations (\ref{eq:dsdt}) and (\ref{eq:drdt}) describe the coupled desorption and radial drift, and can be derived straightforwardly from Equation (\ref{eq:tdes}).
 Our boundary conditions are $s(t_0)=s_0$, $r(t_0)=r_0$, and $s(t_{\rm d})=0$, where $t_0$ is the initial time at which we start the integration and $r_0$ is the initial location of the particle.
@@
 
+As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully desorbs.
+Thus a particle's final location will depend on whether a grain initially at a certain distance is completely desorbed or not within the disk lifetime of 3 Myr.
+For example, larger grains take longer to desorb and hence are more likely to not evaporate fully in a given timeframe.
+
 Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, and for both a passive and an active disk.
-The final distance of the dust particles is consistent with Figure \ref{fig:timescales}.
-Kilometer-sized bodies do not drift or desorb during the disk lifetime both for a passive and active disk.
-Micron- to mm-sized particles do not drift before desorbing in a passive disk, but they drift significantly in an active disk since they move at the same velocity as the accreting gas.
+We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}).
+Kilometer-sized bodies do not drift or desorb during the disk lifetime both for a passive and an active disk.
+Similarly, micron- to mm-sized particles neither drift nor desorb unless they are located inside the static disk snowlines.
+This is not in contradiction with the desorption timescales for small particles from Figure \ref{fig:timescales}, since small grains only desorb fast at or nearby their respective snowlines, as mentioned in Section \ref{sec:timescales}.
+In an active disk, however, micron-to mm-sized grains do drift significantly since they move at the same velocity as the accreting gas.
 For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in a passive disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an active disk, we notice that particles of initial size $s_0$ desorb at a fixed distance $r_{\rm des}$ regardless of their original location in the disk.
+In fact, the only grains that will both drift and evaporate are those that reach their fixed final location (represented by the horizontal curves in Figure \ref{fig:snowlines}) within the disk lifetime.
 We show in section \ref{sec:COratio} that this result is essential in determining the C/O ratio throughout the disk for different particle sizes.
@@
 The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
-For a range of particle sizes, the desorption distance is the same regardless of the particles' initial location.
-\textbf{Figure belongs to section 3.}}  %  (See text for a description of evolution to yet higher masses.)
+For the range of particle sizes that fully desorb during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particles' initial location.}
 \label{fig:snowlines}
@@
 
-\begin{figure}[h!]
+\begin{figure*}[h!]
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active.pdf}
+\includegraphics[width=0.8\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active.pdf}
 %\vspace{-0.5in}
@@
 \label{fig:an_vs_actual}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h!]
+\begin{figure*}[h!]
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/s_t_a.pdf}
+\includegraphics[width=0.8\textwidth]{../../figs/s_t_a.pdf}
 %\vspace{-0.5in}
@@
 \label{fig:s_t_a}
-\end{figure}
+\end{figure*}
+
+We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star.
+In Section \ref{sec:snowlines} we showed that solid particles which drift and fully desorb during the lifetime of the protoplanetary disk will do so at a fixed stellocentric distance, regardless of their initial location in the disk.
 

### 2015-07-14 tex/draft/ms.tex
update

@@
 
-\subsection{Relevant timescales}
+\subsection{Relevant Timescales}
 \label{sec:timescales}
@@
 
+\begin{figure*}[tb]
+\centering
+\includegraphics[width=\textwidth]{../../figs/desorption_distance_passive_active_colorbar_test.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance as a function of a particle's initial location in the disk, for a range of particle sizes, and for both a passive disk (left panels) and an active disk (right panels).
+The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
+For the range of particle sizes that fully desorb during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particles' initial location.}
+\label{fig:snowlines}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=0.7\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines} and text).
+The estimate is performed for a passive disk (left panels) and an active disk (right panels).
+The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+The analytic approximation is in good agreement with the numerical result for most cases.
+\textbf{Figure belongs to section 3.}}  %  (See text for a description of evolution to yet higher masses.)
+\label{fig:an_vs_actual}
+\end{figure*}
+
 In this section we use the model described in section \ref{sec:model} to quantify the effects of radial drift (passive disk) or radial drift and gas accretion (active disk) on the snowline location, for dust particles of different sizes composed of either H$_2$O, CO$_2$ or CO.
@@
 
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=\textwidth]{../../figs/desorption_distance_passive_active_colorbar_test.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance as a function of a particle's initial location in the disk, for a range of particle sizes, and for both a passive disk (left panels) and an active disk (right panels).
-The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
-For the range of particle sizes that fully desorb during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particles' initial location.}
-\label{fig:snowlines}
-\end{figure*}
-
 Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm r, drift} \sim t_{\rm des}$, given an initial particle size.
@@
 We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest.
-
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=0.8\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines} and text).
-The estimate is performed for a passive disk (left panels) and an active disk (right panels).
-The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The analytic approximation is in good agreement with the numerical result for most cases.
-\textbf{Figure belongs to section 3.}}  %  (See text for a description of evolution to yet higher masses.)
-\label{fig:an_vs_actual}
-\end{figure*}
 
@@
 Right panel: fractional mass lost by the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
-Particles lose most of their mass very close to the distance at which they fully desorb.
-\textbf{Figure belongs to section 4.}}
+Particles lose most of their mass very close to the distance at which they fully desorb.}
 \label{fig:s_t_a}
 \end{figure*}
-
-We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star.
-In Section \ref{sec:snowlines} we showed that solid particles which drift and fully desorb during the lifetime of the protoplanetary disk will do so at a fixed stellocentric distance, regardless of their initial location in the disk.
 
@@
 
+We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star.
+In Section \ref{sec:snowlines} we showed that solid particles which drift and fully desorb during the lifetime of the protoplanetary disk will do so (1) instantaneously and at (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
+Figure \ref{fig:s_t_a} confirms both of these effects.
+The left panels show the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
+Indeed, the solid H$_2$O particles evaporate almost instantly, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
+A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime.
+However, our model assumes that particles drift continuously at any location in the disk.
+Therefore, a particle that can fully desorb during the disk lifetime for at least one initial location will always desorb, at a fixed distance as discussed in Section \ref{sec:snowlines} and displayed in Figure \ref{fig:snowlines}.
+The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
+Figure \ref{fig:s_t_a} thus proves claims (1) and (2) above.
+It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active).
+The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane.
+
+We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
+\emgr{Is it necessary to reproduce that table?
+Seems sort of redundant.}.
+Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk with a passive temperature profile, and a steady-state active disk \emgr(not yet there).
+The C/O ratio for a static disk is shown as a guidline.
+The plot is consistent with Figure \ref{fig:snowlines}.
+For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus form a snowline.
+In contrast, even $\sim$micron-sized grains drift and desorb for the active disk, since the flow towards the host star together with the accreting gas.
+For the same particle size, the snowline locations are slightly closer to the central star in the active disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
+Equation \ref{eq:rdotact}).
+Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases.
+While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that even kilometer sized boulders are able to drift and desorb for both the passive and the active disk.
+However, bodies larger than $\sim$7 m will evaporate at the same location as the meter-sized planetesimals.
+Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how far close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
+Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
+Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution.
+
 %\begin{figure}[h!]
@@
 
-\emgr{Motivate the fact that we can calculate a sharp, fixed snowline for each particle size by inserting the plot that shows that particles desorb almost instantaneously at a fixed distance.
+%\emgr{Motivate the fact that we can calculate a sharp, fixed snowline for each particle size by inserting the plot that shows that particles desorb almost instantaneously at a fixed distance.
 Then present the plots analogous to Fig.
@@
 
+\subsection{Dependence on Disk Model and Initial Conditions}
+\label{sec:incond}
+
+\subsection{Neglected Effects}
+\label{sec:neglected}
+
 \emgr{Present the diagram that shows all the effects that can modify snowline location.
@@
 
-\appendix
-\section{...}
-
-\emgr{Right now it's unclear to me what could go in an appendix, if anything.
+%\appendix
+%\section{...}
+%
+%\emgr{Right now it's unclear to me what could go in an appendix, if anything.
 Maybe discuss a bit the algorithm to evolve $\Sigma_{\rm p}$ (although it is already explained in detail in the appendix of Birnstiel et al.

### 2015-07-14 tex/draft/ms.tex
update

@@
 
-Main sequence stars commonly host giant planets (refs).
-The chemical composition of gas giant atmospheres can provide important constraints on their formation, accretion and migration history.
+\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture .
+Gas giants are important.
+The chemical composition of their atmospheres constraints their formation and evolution.
+Is is therefore important to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) backtrack the planet formation location based on planet composition.}
+
+%Main sequence stars commonly host giant planets (refs). The chemical composition of gas giant atmospheres can provide important constraints on their formation, accretion and migration history.
+
+\emgr{Paragraph 2: disks are complex, a lot of dynamical and chemical processes going on.
+However, We have detections of organic molecules.
+We have detections of snowlines.
+We see complex chemistry and many molecules we can study.}
+
 In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11b}, \citealt{oberg11c}, ...).
 Of particular importance are volatile compounds, since the location of their snowlines determines their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
+\emgr{Add part about snowlines, complex chemistry, many molecules.}
 
-Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio.
+\emgr{Paragraph 3: let's focus on one (set of) molecules: the C/O ratio.
+It's an important signature of atmospheric chemistry.
+While we cannot detect it directly in disks, we have observations in giant planet atmospheres.
+It's different than stellar.
+Why?
+Possible explanation based on snowlines.}
+
+\emgr{Start with the fact that we can't see C/O in disks directly.
+Then transition to planet atmospheres and modify the next sentence to follow up logically.} Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio.
 Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}.
@@
 
-In order to obtain more realistic estimates for C/O ratios across protoplanetary disks, dynamical processes and the disk evolving chemistry have to be taken into account.
+\emgr{Paragraph 4: the static disk model is a great start, but there are additional dynamical and chemical processes to take into account.
+Several studies have done that in some form or another --- cite Ali-Dib+14, Madhusudhan+14, Thiabaud+15.
+Briefly discuss their assumptions.}
+
+\emgr{Paragraph 5: finally introduce our goal for this paper: study the dynamical effects such as drift and gas accretion.
+State our goals: understand the detailed qualitative and quantitative effect of drift and gas accretion on snowline locations, obtain a limit on how far in can the snowlines be pushed, and see how that affects the C/O ratio throughout the disk.
+State that our specific goals motivate the use of a simplified model (in other words, why what we are doing is different from the studies cited in the previous paragraph).
+}
+
+\emgr{This needs some restructuring and additions along the lines discussed above.} In order to obtain more realistic estimates for C/O ratios across protoplanetary disks, dynamical processes and the disk evolving chemistry have to be taken into account.
 In this paper, we enhance the model of \citet{oberg11} considering two additional dynamic effects: (1) the radial drift of solids throughout the protoplanetary disk, and (2) the viscous accretion of the disk gas onto the host star.
@@
 
-\begin{figure*}[h!]
+\begin{figure*}[t!]
 \centering
@@
 
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{../../figs/s_t_a.pdf}
+%\vspace{-0.5in}
+\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in a passive disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left).
+Particles desorb almost instantaneously.
+Right panel: fractional mass lost by the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
+Particles lose most of their mass very close to the distance at which they fully desorb.}
+\label{fig:s_t_a}
+\end{figure*}
+
 Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm r, drift} \sim t_{\rm des}$, given an initial particle size.
-Figure \ref{fig:an_vs_actual} shows the validity of this approximation for the range of particle sizes that desorb at a fixed distance in a passive and an active disk.
-We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest.
+We can calculate this location analytically by equating Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and solving for $r=r_{\rm des}(s)$ for a given particle size $s$.
+Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically,  for the range of particle sizes that desorb at a fixed distance in a passive and an active disk (see Figure \ref{fig:snowlines}).
+We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it slightly deviates for smaller particles.
+\emgr{I've yet to figure out why there is a discrepancy for the small grains.
+I'll keep thinking about it, but for now I'll leave it as it is since I don't want to give a false explanation just to have one.}
 
@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_vs_s.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance as a function of initial particle size, for the range of particles that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines}).
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_vs_s.pdf}
+%%\vspace{-0.5in}
+%\caption{Desorption distance as a function of initial particle size, for the range of particles that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines}).
 \textbf{Figure belongs to section 3.} \emgr{This is a figure that I'm not sure we should include, since it doesn't really provide that much useful inside.
@@
 The caption is not complete since the active disk panel is not yet there.}}  %  (See text for a description of evolution to yet higher masses.)
-\label{fig:r_vs_s}
-\end{figure}
+%\label{fig:r_vs_s}
+%\end{figure}
 
@@
 \label{sec:COratio}
-
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=0.8\textwidth]{../../figs/s_t_a.pdf}
-%\vspace{-0.5in}
-\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in a passive disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left).
-Particles desorb almost instantaneously.
-Right panel: fractional mass lost by the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
-Particles lose most of their mass very close to the distance at which they fully desorb.}
-\label{fig:s_t_a}
-\end{figure*}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar_test.pdf}
-%\vspace{-0.5in}
-\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel) and for an active disk (bottom panel), and for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
-The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
-The horizontal dotted line represents the stellar value of 0.54.
-The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk.
-The snowline location moves inward as the particle size increases.
-\textbf{Figure belongs to section 4.}}
-\label{fig:CO_ratio}
-\end{figure}
 
@@
 The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar_test.pdf}
+%\vspace{-0.5in}
+\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel) and for an active disk (bottom panel), and for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
+The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
+The horizontal dotted line represents the stellar value of 0.54.
+The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk.
+The snowline location moves inward as the particle size increases.
+\emgr{Some additional text for the currently non-existent third panel.}}
+\label{fig:CO_ratio}
+\end{figure}
 
@@
 Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution.
+\emgr{There will be 1-2 additional sentences for the third panel, the steady-state disk, once that finishes running.}
 
@@
 
-\subsection{Dependence on Disk Model and Initial Conditions}
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=\textwidth]{../../figs/time_plots.pdf}
+%\vspace{-0.5in}
+\caption{Desorption / final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ m (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve), and 3 Myr, the disk lifetime (blue curve).
+For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the timescale at which the simulation is stopped.}
+\label{fig:timeplots}
+\end{figure*}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_varying_Md.pdf}
+%\vspace{-0.5in}
+\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M_{\rm d}=0.1 M_{\odot}$ (solid lines) and $M_{\rm d}=0.01 M_{\odot}$ (dashed lines).
+The timescales of the simulations are the same as in Figure \ref{fig:timeplots}.
+A lower disk mass does not change the snowline location.}
+\label{fig:varMd}
+\end{figure}
+
+\subsection{Generality of Results: Dependence on Disk Parameters}
 \label{sec:incond}
+
+In this section we investigate the effect on our results of variations in our fiducial parameters, such as the total disk mass or the time in the disk evolution at which we determine the snowline locations and the C/O ratio.
+The disk lifetime $t_{\rm d}=3$ Myr is the most representative for our calculations, as this is the timeframe in which giant planets accrete their gaseous atmospheres (e.g., \citealt{pollack96}, \citealt{piso14}).
+However, protoplanetary cores start acquiring their envelope at earlier times, before the core is fully formed (e.g., \citealt{rafikov06}).
+Recent models such as aerodynamic pebble accretion \citep{lambrechts12} suggest rapid core growth on timescales of $10^5$ years, which implies that the time when cores accrete their massive envelope may be at least an order of magnitude shorter than the disk lifetime.
+The composition of giant planet atmospheres, and specifically their C/O ratio, can thus depend on the abundance of H$_2$O, CO$_2$ and CO in gas and dust forms at earlier times than $t_{\rm d}$ in the disk evolution.
+
+Figure \ref{fig:timeplots} shows the desorption or final distance as a function of a particle's initial location in the disk, for grains of initial sizes of 10 cm and 1 m, composed of either H$_2$O, CO$_2$ or CO.
+These sizes are representative since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
+We perform the simulations at four representative timescales in the disk evolution to see how the snowline locations evolve with time and affect the C/O ratio.
+Particles that start at large stellocentric distances are not able to desorb within the shorter timeframes, e.g.
+$10^4$ or $10^5$ years.
+However, they do evaporate if their initial location is closer to the host star.
+As stated in Section \ref{sec:COratio}, this implies that the grains form a snowline even after $10^4$ years.
+More importantly, for a given initial grain size, the snowline locations are independent of the time elapsed.
+Therefore, our results are not only valid at $t_{\rm d}$, but also throughout the earlier time evolution of the protoplanetary disk.
+
+We choose as a fiducial model a total disk mass $M_{\rm d}=0.1 M_{\odot}$, but this number is not universally valid.
+From observations of the dust continuum, \citet{andrews13} find that a linear scaling $M_{\rm d} \propto M_*$ is reasonable.
+Giant planets, however, have been detected around small stars (e.g., \citealt{montet14}), which can have masses as low as $M_* \sim 0.1 M_{\odot}$.
+We thus explore the effect of disk mass on the location of snowlines.
+Figure \ref{fig:varMd} shows the desorption or final distance as a function on the initial location of a H$_2$O particle with initial size of 1 m, for two total disk masses: $M_{\rm d}=0.1 M_{\odot}$, our fiducial model, and $M_{\rm d}=0.01 M_{\odot}$.
+The simulations are stopped after the same timeframes as those in Figure \ref{fig:timeplots}.
+The location of the H$_2$O snowline is the same for both disks (the same holds true for the CO$_2$ and CO snowlines).
+The C/O ratio is thus insensitive to the choice of $M_{\rm d}$.
+This result has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
+Since the snowline locations are independent of disk mass, we expect transition disks to exhibit the same drift-desorption behavior as standard disks under the simplified assumptions of our model.
+
+\begin{table}[t!]
+\caption{Dynamical and chemical processes that affect the snowline locations of the main C and O carriers, H$_2$O, CO$_2$ and CO.
+The arrows signify the direction in which a particular processes affects the snowlines --- $\leftarrow$=snowline is pushed closer to the host star, $\rightarrow$=snowline is pushed further from the host star.
+The presence of both arrows means that the process may have both effects on the snowline location.}
+\begin{center}
+\begin{tabular}{|l|l|}\hline
+\textbf{Process} & \textbf{Effect} \\\hline
+Radial drift & $\leftarrow$ \\\hline
+Gas accretion & $\leftarrow$ \\\hline
+Particle growth & $\rightarrow$ \\\hline
+Turbulent diffusion & $\rightarrow$ \\\hline
+Particle fragmentation & $\rightarrow$ $\leftarrow$ \\\hline
+Morphology & $\rightarrow$ \\\hline
+Particle composition & $\rightarrow$ $\leftarrow$ \\\hline
+Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-\emgr{Present the diagram that shows all the effects that can modify snowline location.
+Our goals in this paper were (1) to gain a detailed qualitative and quantitative understanding of the effect or radial drift and gas accretion on to the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
+We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical effects.
+In what follows, we discuss these limitations and their effects.
+We note that our future work will address some of these issues.
+
+We summarize in Table 1 the potential physical and chemical processes occurring in disks and their effect on snowline locations.
+For the sake of completion, Table 1 also includes the processes addressed in this paper, i.e.
+radial drift and gas accretion.
+The neglected effects are discussed in more detail below.
+
+\begin{enumerate}
+\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered fixed for a given grain throughout their drift and evolution.
+However, grain growth has been observed in protoplanetary disks (e.g., \citealt{ricci10}, \citealt{perez12}), as well as theoretically constrained (e.g., \citealt{birnstiel10}, \citealt{birnstiel12}).
+In Section \ref{sec:COratio} we have shown that larger grains move the snowline locations closer in, but those locations remain fixed above a certain particle size.
+Once the solids grow larger than km-sized, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
+It follows that grain growth will eventually push the snowline location outwards.
+
+\item \textbf{Turbulent diffusion.} The radial drift model presented in Section \ref{sec:drift} only considers a laminar flow and thus ignores turbulence.
+However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
+Turbulence causes eddies and vertical mixing, which is likely to reduce the radial gas accretion flow.
+Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially, causing it to move cross the snowline and refreeze, thus moving the snowline location outwards altogether.
+
+\item \textbf{Particle fragmentation.} Frequent particle collisions in disks causes them to fragment (e.g., \citealt{birnstiel12}).
+The fragmentation of meter- to km-sized particles will move the snowlines outwards, as smaller particles desorb faster and further out from the host star (cf.
+Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}).
+Large boulders, which neither drift nor desorb, may become e.g.
+meter-sized due to collisions and subsequent fragmentation, which will cause them to drift significantly before desorbing, pushing the snowlines inwards.
+Thus fragmentation can move the snowline locations in either radial direction.
+
+\item \textbf{Morphology.} Our model assumes that the ice particles are perfect, homogeneous spheres.
+However, this is not a very good approximation, since grain growth can be fractal rather than compact (\citealt{zsom10}, \citealt{okuzumi12}).
+The inhomogeneity due to cracks in the grain structure will cause the particles to desorb faster, hence drift less before evaporating and moving the snowlines outwards.
+
+\item \textbf{Particle composition.} The ice particles in our model are assumed to be fully formed of either H$_2$O, CO$_2$ or CO.
+In reality, the grains have a layered structure, such as an interior composed of non-volatile materials (e.g., sillicates) covered by an icy layer.
+The ice thus only constitutes a fraction of the total particle mass, which accelerates its desorption and pushes the snowlines outwards.
+The grains may also be composed of a mixture of H$_2$O, CO$_2$ and CO ices.
+The relative fraction of each volatile, as well as their degree of mixing, will determine their desorption timescale, and as a result whether the snowlines are moved towards or away from the central star.
+
+\item \textbf{Non-static chemistry.} As the goal of this paper was to explore the dynamical effects on snowline locations and the C/O ratio in disks, we have assumed a very simple, static chemical model.
+However, gas-grain chemistry is very complex and time-dependent.
+In the inner disk, chemistry approaches equilibrium due to intense sources of ionizing radiation (e.g., \citealt{ilgner04}), while in the outer disk high energy radiation and cosmic rays are the key drivers of chemistry, which is no longer in equilibrium (e.g., \citealt{vandishoeck06}).
+A multitude of chemical evolution models have been developed (see references in \citealt{henning13}), many of which contain tens or hundreds of reactions.
+Due to the complexity of these chemical models, most of them are decoupled from disk dynamics.
+Their effect on snowline locations, shape, time evolution, or the C/O ratio is therefore difficult to estimate.
+In a future paper, we plan to parametrize the chemical model developed by Merchantz et al.
+(in preparation) that traces the evolution of the main C and O oxygen carriers, and incorporate it in our radial drift calculation.
+
+\end{enumerate}
+
+%\emgr{Present the diagram that shows all the effects that can modify snowline location.
 For model limitations, include: non-inclusion of turbulence, assumption of perfect spheres when in fact they may have cracks, particles composed of a single volatile when in reality they are likely to be mixed, etc.

### 2015-07-15 tex/draft/ms.tex
update

@@
 \usepackage{color}
+\usepackage[utf8]{inputenc}
 
@@
 
-%\shorttitle{Critical Core Mass at Wide Separations}
-%\shortauthors{Piso et al.}
-
 \title{C/O in Protoplanetary Disks: The Effect of Radial Drift and Viscous Accretion}
@@
 
-\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture .
+\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture.
 Gas giants are important.
@@
 
-%Main sequence stars commonly host giant planets (refs). The chemical composition of gas giant atmospheres can provide important constraints on their formation, accretion and migration history.
+Main sequence stars commonly host giant planets (refs).
+The chemical composition of gas giant atmospheres can provide important constraints on their formation, accretion and migration history.
+\emgr{Add the rest as outlined above.} \\
 
 \emgr{Paragraph 2: disks are complex, a lot of dynamical and chemical processes going on.
-However, We have detections of organic molecules.
+However, we have detections of organic molecules.
 We have detections of snowlines.
@@
 Of particular importance are volatile compounds, since the location of their snowlines determines their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
-\emgr{Add part about snowlines, complex chemistry, many molecules.}
+\emgr{Add part about snowlines, complex chemistry, many molecules.} \\
 
@@
 \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
+\\
 
@@
 Several studies have done that in some form or another --- cite Ali-Dib+14, Madhusudhan+14, Thiabaud+15.
-Briefly discuss their assumptions.}
+Briefly discuss their assumptions.} \\
 
@@
 \end{subeqnarray}
-where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k}$ is the Keplerian angular velocity, $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$, with $G$ the gravitational constant and $M_*$ the stellar mass.
-We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition go hydrogen and helium.
+where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$ is the Keplerian angular velocity,  with $G$ the gravitational constant and $M_*$ the stellar mass.
+We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition of hydrogen and helium.
 The temperature profile for an active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}).
@@
 \end{subeqnarray}
-where $t$ is time. We take $M_{\rm d}=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
+where $t$ is time. We choose $M_{\rm d}=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
 
-\textbf{Active disk steady-state solution.} Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):
+\textbf{Active disk steady-state solution.} Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady-state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):
 
@@
 \end{subeqnarray}
-where $\dot{M}$, the mass flux, is constant for steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}.
+where $\dot{M}$, the mass flux, is constant in steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}.
 Based on disk observations (e.g., \citealt{andrews10}), we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$.
@@
 
-\textbf{Static disk.} For comparison purposes with \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}).
+\textbf{Static disk.} To compare our results with those of \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}).
 The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
@@
 Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm r, drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk.
-As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence disk location (see Equation \ref{eq:Rdes}).
+As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence on disk location (see Equation \ref{eq:Rdes}).
 On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas.
@@
 %\vspace{-0.5in}
-\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm evap}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
-The timescales are calculated at three representative locations in the disk, i.e.
-the H$_2$O, CO$_2$ and CO snowlines.
+\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
+The timescales are calculated at three representative locations, i.e.
+the H$_2$O, CO$_2$ and CO snowlines in the static disk.
 For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively.
@@
 %\vspace{-0.5in}
-\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines} and text).
+\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text).
 The estimate is performed for a passive disk (left panels) and an active disk (right panels).
 The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The analytic approximation is in good agreement with the numerical result for most cases.
-\textbf{Figure belongs to section 3.}}  %  (See text for a description of evolution to yet higher masses.)
+The analytic approximation is in good agreement with the numerical result for most cases.}
 \label{fig:an_vs_actual}
@@
 
-As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully desorbs.
+As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully evaporates.
 Thus a particle's final location will depend on whether a grain initially at a certain distance is completely desorbed or not within the disk lifetime of 3 Myr.
@@
 
-Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, and for both a passive and an active disk.
+Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both a passive and an active disk.
 We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}).
-Kilometer-sized bodies do not drift or desorb during the disk lifetime both for a passive and an active disk.
-Similarly, micron- to mm-sized particles neither drift nor desorb unless they are located inside the static disk snowlines.
+Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for a passive nor for an active disk.
+Similarly, micron- to mm-sized particles in the passive disk do not drift or desorb unless they are located inside the static snowlines.
 This is not in contradiction with the desorption timescales for small particles from Figure \ref{fig:timescales}, since small grains only desorb fast at or nearby their respective snowlines, as mentioned in Section \ref{sec:timescales}.
@@
 Particles desorb almost instantaneously.
-Right panel: fractional mass lost by the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
+Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
 Particles lose most of their mass very close to the distance at which they fully desorb.}
@@
 We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star.
-In Section \ref{sec:snowlines} we showed that solid particles which drift and fully desorb during the lifetime of the protoplanetary disk will do so (1) instantaneously and at (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
+In Section \ref{sec:snowlines}, we showed that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
 Figure \ref{fig:s_t_a} confirms both of these effects.
@@
 \emgr{Is it necessary to reproduce that table?
-Seems sort of redundant.}.
-Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk with a passive temperature profile, and a steady-state active disk \emgr(not yet there).
-The C/O ratio for a static disk is shown as a guidline.
+Seems a bit redundant.}.
+Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk with a passive temperature profile, and a steady-state active disk \emgr{(not yet there)}.
+The C/O ratio for a static disk is shown as a guideline.
 The plot is consistent with Figure \ref{fig:snowlines}.
 For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus form a snowline.
-In contrast, even $\sim$micron-sized grains drift and desorb for the active disk, since the flow towards the host star together with the accreting gas.
+In contrast, even $\sim$micron-sized grains drift and desorb for the active disk, since they flow towards the host star together with the accreting gas.
 For the same particle size, the snowline locations are slightly closer to the central star in the active disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
@@
 Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases.
-While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that even kilometer sized boulders are able to drift and desorb for both the passive and the active disk.
+While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk.
 However, bodies larger than $\sim$7 m will evaporate at the same location as the meter-sized planetesimals.
-Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how far close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
+Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
 Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
@@
 %\vspace{-0.5in}
-\caption{Desorption / final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ m (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+\caption{Desorption / final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
 The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve), and 3 Myr, the disk lifetime (blue curve).
-For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the timescale at which the simulation is stopped.}
+For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
 \label{fig:timeplots}
@@
 \caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M_{\rm d}=0.1 M_{\odot}$ (solid lines) and $M_{\rm d}=0.01 M_{\odot}$ (dashed lines).
-The timescales of the simulations are the same as in Figure \ref{fig:timeplots}.
+The timescales of the simulations and their color code are the same as in Figure \ref{fig:timeplots}.
 A lower disk mass does not change the snowline location.}
@@
 
-In this section we investigate the effect on our results of variations in our fiducial parameters, such as the total disk mass or the time in the disk evolution at which we determine the snowline locations and the C/O ratio.
+In this section we investigate how variations in our fiducial parameters, such as the total disk mass or the time in the disk evolution at which we determine the snowline locations and the C/O ratio, affect our results.
 The disk lifetime $t_{\rm d}=3$ Myr is the most representative for our calculations, as this is the timeframe in which giant planets accrete their gaseous atmospheres (e.g., \citealt{pollack96}, \citealt{piso14}).
@@
 We perform the simulations at four representative timescales in the disk evolution to see how the snowline locations evolve with time and affect the C/O ratio.
-Particles that start at large stellocentric distances are not able to desorb within the shorter timeframes, e.g.
+Particles that start at large stellocentric distances do not desorb within the shorter timeframes, e.g.
 $10^4$ or $10^5$ years.
@@
 The C/O ratio is thus insensitive to the choice of $M_{\rm d}$.
-This result has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
+This result also has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
 Since the snowline locations are independent of disk mass, we expect transition disks to exhibit the same drift-desorption behavior as standard disks under the simplified assumptions of our model.
+
+\subsection{Neglected Effects}
+\label{sec:neglected}
+
+Our goals in this paper were (1) to gain a detailed qualitative and quantitative understanding of the effect of radial drift and gas accretion on to the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
+We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical processes.
+In what follows, we discuss these limitations and their effects.
+We note that our future work will address some of these issues.
 
@@
 \caption{Dynamical and chemical processes that affect the snowline locations of the main C and O carriers, H$_2$O, CO$_2$ and CO.
-The arrows signify the direction in which a particular processes affects the snowlines --- $\leftarrow$=snowline is pushed closer to the host star, $\rightarrow$=snowline is pushed further from the host star.
+The arrows signify the direction in which a particular process affects the snowlines: $\leftarrow$ means that the snowline is pushed closer to the host star, $\rightarrow$ means that the snowline is pushed further from the host star.
 The presence of both arrows means that the process may have both effects on the snowline location.}
@@
 
-\subsection{Neglected Effects}
-\label{sec:neglected}
-
-Our goals in this paper were (1) to gain a detailed qualitative and quantitative understanding of the effect or radial drift and gas accretion on to the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
-We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical effects.
-In what follows, we discuss these limitations and their effects.
-We note that our future work will address some of these issues.
-
 We summarize in Table 1 the potential physical and chemical processes occurring in disks and their effect on snowline locations.
@@
 \begin{enumerate}
-\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered fixed for a given grain throughout their drift and evolution.
+\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered fixed for a given grain throughout its drift and evolution.
 However, grain growth has been observed in protoplanetary disks (e.g., \citealt{ricci10}, \citealt{perez12}), as well as theoretically constrained (e.g., \citealt{birnstiel10}, \citealt{birnstiel12}).
@@
 However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
-Turbulence causes eddies and vertical mixing, which is likely to reduce the radial gas accretion flow.
-Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially, causing it to move cross the snowline and refreeze, thus moving the snowline location outwards altogether.
+Turbulence causes eddies and vertical mixing, which are likely to reduce the radial gas accretion velocity.
+Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially, causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.
 
-\item \textbf{Particle fragmentation.} Frequent particle collisions in disks causes them to fragment (e.g., \citealt{birnstiel12}).
+\item \textbf{Particle fragmentation.} Frequent particle collisions in disks cause them to fragment (e.g., \citealt{birnstiel12}).
 The fragmentation of meter- to km-sized particles will move the snowlines outwards, as smaller particles desorb faster and further out from the host star (cf.
@@
 However, this is not a very good approximation, since grain growth can be fractal rather than compact (\citealt{zsom10}, \citealt{okuzumi12}).
-The inhomogeneity due to cracks in the grain structure will cause the particles to desorb faster, hence drift less before evaporating and moving the snowlines outwards.
+The inhomogeneity due to cracks in the grain structure will cause the particles to desorb faster.
+They will therefore drift less before evaporating and will move the snowlines outwards.
 
@@
 
-\item \textbf{Non-static chemistry.} As the goal of this paper was to explore the dynamical effects on snowline locations and the C/O ratio in disks, we have assumed a very simple, static chemical model.
+\item \textbf{Non-static chemistry.} As the goal of this paper was to explore only the dynamical effects on snowline locations and the C/O ratio in disks, we have assumed a simple, static chemical model.
 However, gas-grain chemistry is very complex and time-dependent.
 In the inner disk, chemistry approaches equilibrium due to intense sources of ionizing radiation (e.g., \citealt{ilgner04}), while in the outer disk high energy radiation and cosmic rays are the key drivers of chemistry, which is no longer in equilibrium (e.g., \citealt{vandishoeck06}).
-A multitude of chemical evolution models have been developed (see references in \citealt{henning13}), many of which contain tens or hundreds of reactions.
+A multitude of chemical evolution models have been developed (see references in \citealt{henning13}), many of which contain tens or hundreds of chemical reactions.
 Due to the complexity of these chemical models, most of them are decoupled from disk dynamics.
-Their effect on snowline locations, shape, time evolution, or the C/O ratio is therefore difficult to estimate.
+The effect of disk chemistry on snowline locations, shape, time evolution, or the C/O ratio is therefore difficult to estimate.
 In a future paper, we plan to parametrize the chemical model developed by Merchantz et al.
@@
 
-\section{Summary and Future Work}
+\section{Summary}
 
-\emgr{Summarize results. Mention inclusion of $N_2$ as a first expansion. Mention the implementation of time-dependent chemical models in the drift calculation.}
+\emgr{To be done soon.}
 

### 2015-07-15 python/drift_timescales_active_Tdisk.py
update

@@ -652,6 +652,17 @@ def rdrift(rin, tin, tf, s, alpha, dr, Mdisk, rc = 100*AU, rhos = 3.0, T0 = 120,
     return y[-1][0] / AU
     
     
+    
+def r_stop(s, alpha, Mdot, mx, Ex, dr = 1e-3, rhos = 3.0, mu = 2.35, Mstar = Msun, sigma = 2 * 10**(-15), Nx = 1e15):
+
+     def f(r):
+          return (r*AU) / (- rdot_with_acc(r, s, alpha, Mdot, dr, rhos, mu, Mstar, sigma)) - \
+            tdes(mx, Ex, Tdisk(r, Mdot, Mstar), s, Nx, rhos)
+
+     try:
+          return brentq(f, 1e-3, 1e2)
+     except ValueError:
+          return 1e-10     
 
 
 #######################################################################################

### 2015-07-15 python/drift_timescales_updated_sigmap.py

@@ -1477,7 +1477,19 @@ def n_with_acc(rin, sin, mx, Ex, Mdotgas, Nx = 1e15, rhos = 3.0, T0 = 120, betaT
 
      t, a, s = rfin
      
-     
+
+def r_stop(t, s, alpha, dr, Mdisk, mx, Ex, rc = 100*AU, T0=120, betaT=3./7, rhos = 3.0, mu = 2.35, \
+    Mstar = Msun, gammadflag = 0, sigma = 2 * 10**(-15), Nx = 1e15):
+
+     def f(r):
+          return (r*AU) / (- rdot_with_acc(r, t, s, alpha, dr, Mdisk, rc, T0, betaT, rhos, mu, \
+            Mstar, gammadflag, sigma)) - tdes(mx, Ex, Tdisk(r, T0, betaT), s, Nx, rhos)
+
+     try:
+          return brentq(f, 1e-3, 1e2)
+     except ValueError:
+          return 1e-10     
+               
      
 #def r_stop(s, mx, Ex, Nx = 1e15, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, Sigma0 = 2200, betaS = 3./2, \
 #    Mstar = Msun, sigma = 2 * 10**(-15)):

### 2015-07-16 python/drift_timescales_updated_sigmap_0402_cavity.py
update

@@ -335,7 +335,7 @@ def Sigmadisk(r, t, alpha, Mdisk, rc, T0 = 120, betaT = 3./7, mu = 2.35, Mstar =
             * np.exp(- rtild**(2 - gammad) / T)
      else:
          return Mdisk / (2 * np.pi * rc**2 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
-            * np.exp(- rtild**(2 - gammad) / T) / 10
+            * np.exp(- rtild**(2 - gammad) / T) / 100
                 
 
 def vacc_act(r, t, alpha, mu = 2.35, T0 = 120, betaT = 3./7, rc = 100 * AU, Mstar = Msun, gammadflag = 0):

### 2015-07-16 python/drift_timescales_updated_sigmap_0402_cavity.py
update

@@ -335,7 +335,7 @@ def Sigmadisk(r, t, alpha, Mdisk, rc, T0 = 120, betaT = 3./7, mu = 2.35, Mstar =
             * np.exp(- rtild**(2 - gammad) / T)
      else:
          return Mdisk / (2 * np.pi * rc**2 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
-            * np.exp(- rtild**(2 - gammad) / T) / 100
+            * np.exp(- rtild**(2 - gammad) / T) / 10
                 
 
 def vacc_act(r, t, alpha, mu = 2.35, T0 = 120, betaT = 3./7, rc = 100 * AU, Mstar = Msun, gammadflag = 0):

### 2015-07-16 tex/draft/ms.tex
update

@@
 \centering
-\includegraphics[width=\textwidth]{../../figs/desorption_distance_passive_active_colorbar_test.pdf}
+\includegraphics[width=0.7\textwidth]{../../figs/desorption_distance_passive_active_colorbar_test.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active.pdf}
+\includegraphics[width=0.7\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active_new.pdf}
 %\vspace{-0.5in}

### 2015-07-16 python/drift_timescales_active_Tdisk.py
update

@@ -128,7 +128,7 @@ def impl_donorcell_adv_diff_delta(n_x,x,Diff,v,g,h,K,L,flim,u_in,dt,pl,pr,ql,qr,
 ###########################
 
 
-def Tdisk(r, Mdot, Mstar = Msun):
+def Tsurf(r, Mdot, Mstar = Msun):
     
     """
     
@@ -150,6 +150,8 @@ def Tdisk(r, Mdot, Mstar = Msun):
     """
     
     return (3 * G * Mstar * (Mdot * Mstar / yr) / (8 * np.pi * sigma * (r * AU)**3))**(1./4)
+    
+def tau(r, )
      
 def gamma(betaT = 3./7):
     

### 2015-07-16 python/drift_timescales_updated_sigmap_0402_cavity.py
update

@@ -335,7 +335,7 @@ def Sigmadisk(r, t, alpha, Mdisk, rc, T0 = 120, betaT = 3./7, mu = 2.35, Mstar =
             * np.exp(- rtild**(2 - gammad) / T)
      else:
          return Mdisk / (2 * np.pi * rc**2 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
-            * np.exp(- rtild**(2 - gammad) / T) / 10
+            * np.exp(- rtild**(2 - gammad) / T) / 1000
                 
 
 def vacc_act(r, t, alpha, mu = 2.35, T0 = 120, betaT = 3./7, rc = 100 * AU, Mstar = Msun, gammadflag = 0):

### 2015-07-16 python/drift_timescales_active_Tdisk.py
update

@@ -128,30 +128,14 @@ def impl_donorcell_adv_diff_delta(n_x,x,Diff,v,g,h,K,L,flim,u_in,dt,pl,pr,ql,qr,
 ###########################
 
 
-def Tsurf(r, Mdot, Mstar = Msun):
-    
-    """
-    
-    Calculates the disk temperature as a function of semimajor axis for a passive disk.
-    
-    Input
-    -----
-    r:
-        semimajor axis in AU
-    T0:
-        temperature at 1 AU in K
-    betaT:
-        powerlaw coefficient in T = T0 * r**(-betaT)
-        
-    Output
-    ------
-    disk temperature in K
-    
-    """
-    
-    return (3 * G * Mstar * (Mdot * Mstar / yr) / (8 * np.pi * sigma * (r * AU)**3))**(1./4)
-    
-def tau(r, )
+
+
+
+
+
+
+
+###############################
      
 def gamma(betaT = 3./7):
     

### 2015-07-17 python/drift_timescales_active_Tdisk.py
update

@@ -1,4 +1,4 @@
-from utils.constants import G, kb, mp, Msun, cmperau, AU, sigma, yr
+from utils.constants import G, kb, mp, Msun, cmperau, AU, yr
 import numpy as np
 from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
 from T_freeze import T_freeze, Rdes, tevap, vib_freq
@@ -9,6 +9,7 @@ from scipy.optimize import brentq
 from scipy.interpolate import interp1d
 from utils.zbrac import zbrac
 from utils.utilities import tridag
+from drift_timescales_updated_sigmap import Tdisk as Tpas
 #from advection_from_Til.two-pop-py.src
 
 ###################################################################
@@ -128,346 +129,110 @@ def impl_donorcell_adv_diff_delta(n_x,x,Diff,v,g,h,K,L,flim,u_in,dt,pl,pr,ql,qr,
 ###########################
 
 
-
-
-
-
-
-
-
-###############################
-     
-def gamma(betaT = 3./7):
+def Tc(r, alpha, Mdot, k0, mu, Mstar = Msun):
     
-    """
-    Powerlaw coefficient in the viscosity dependendce,
-    nu \propto r^gamma
+    Mdot1 = Mdot * Mstar / yr
+    r1 = r * AU
     
-    Input
-    -----
-    betaT:
-        powerlaw coefficient in the temperature dependence
+    return 3**(1./3)*G**(1./3)*k0**(1./3)*Mdot1**(2./3)*mp**(1./3)*Msun**(1./3)*mu**(1./3)*(G*Mstar/r1**3)**(1./6) / \
+        (4*alpha**(1./3)*kb**(1./3)*np.pi**(2./3)*r1*sigma**(1./3))
         
-    Output
-    ------
-    gamma (dimensionless)
-    
-    """
-     
-    return -betaT + 3./2
 
-def cdisk(r, Mdot, Mstar = Msun, mu = 2.35):
+def Tdisk(r, alpha, Mdot, k0, mu, Mstar = Msun, T0 = 120, betaT = 3./7):
     
-    """
-    Sound speed as a function of semimajor axis
-    
-    Input
-    -----
-    r:
-        semimajor axis in AU
-    T0, betaT:
-        normalization temperature and powerlaw in the disk temperature dependence
-    mu:
-        mean molecular weight of the nebular gas
-        
[…]

### 2015-07-17 python/drift_timescales_active_Tdisk.py
update

@@ -1,4 +1,4 @@
-from utils.constants import G, kb, mp, Msun, cmperau, AU, yr
+from utils.constants import G, kb, mp, Msun, cmperau, AU, yr, sigmaB
 import numpy as np
 from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
 from T_freeze import T_freeze, Rdes, tevap, vib_freq
@@ -135,7 +135,7 @@ def Tc(r, alpha, Mdot, k0, mu, Mstar = Msun):
     r1 = r * AU
     
     return 3**(1./3)*G**(1./3)*k0**(1./3)*Mdot1**(2./3)*mp**(1./3)*Msun**(1./3)*mu**(1./3)*(G*Mstar/r1**3)**(1./6) / \
-        (4*alpha**(1./3)*kb**(1./3)*np.pi**(2./3)*r1*sigma**(1./3))
+        (4*alpha**(1./3)*kb**(1./3)*np.pi**(2./3)*r1*sigmaB**(1./3))
         
 
 def Tdisk(r, alpha, Mdot, k0, mu, Mstar = Msun, T0 = 120, betaT = 3./7):

### 2015-07-17 python/drift_timescales_active_Tdisk.py
update

@@ -1,4 +1,4 @@
-from utils.constants import G, kb, mp, Msun, cmperau, AU, yr, sigmaB
+from utils.constants import G, kb, mp, Msun, cmperau, AU, yr, sigma
 import numpy as np
 from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
 from T_freeze import T_freeze, Rdes, tevap, vib_freq
@@ -135,7 +135,7 @@ def Tc(r, alpha, Mdot, k0, mu, Mstar = Msun):
     r1 = r * AU
     
     return 3**(1./3)*G**(1./3)*k0**(1./3)*Mdot1**(2./3)*mp**(1./3)*Msun**(1./3)*mu**(1./3)*(G*Mstar/r1**3)**(1./6) / \
-        (4*alpha**(1./3)*kb**(1./3)*np.pi**(2./3)*r1*sigmaB**(1./3))
+        (4*alpha**(1./3)*kb**(1./3)*np.pi**(2./3)*r1*sigma**(1./3))
         
 
 def Tdisk(r, alpha, Mdot, k0, mu, Mstar = Msun, T0 = 120, betaT = 3./7):

### 2015-07-19 tex/draft/ms.tex
update

@@
 
-\section{Model Assumptions} 
+\section{Model Framework} 
 \label{sec:model}
@@
 
-\textbf{Passive disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
+\textbf{Passive disk.} A passive disk is only irradiated by the central star and does not experience accretion heating.
+We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}.
+The gas surface density and midplane temperature are
 \begin{subeqnarray}
@@
 
-\textbf{Actively accreting disk with passive temperature profile.} We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
+\textbf{Active disk with passive temperature profile.} The gas is accreting onto the central star in an active disk, which causes the gas surface density to decrease with time.
+We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
 \begin{equation}
@@
 From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters.
-Following \citet{armitage10}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius.
+Following \citet{hartmann98}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius.
 We choose $r_{\rm c}=100$ AU.
@@
 
-\textbf{Active disk steady-state solution.} Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady-state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):
+\textbf{Active disk steady-state.} A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma_{\rm d}$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
+Since calculating the midplane temperature self-consistently for an active disk is non-trivial, we use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm d, act}$.
+The equations governing the evolution of the steady-state disk are listed in \App{app:steadystate}.
+We assume an interstellar opacity for the dust grains given by \citet{bell94}, but reduced by a factor of 100.
+This reduction is due to the fact that disk opacities are lower than the interstellar one.
+The scaling is consistent with more detailed models of grain opacities in disks (e.g., \citealt{mordasini14}).
+Our opacity law is thus
 
-\begin{subeqnarray}
+\begin{equation}
+\label{eq:opacity}
+\kappa=\kappa_0 T_{\rm d, act}^2,
+\end{equation}
+where $\kappa_0=2 \times 10^{-6}$. By solving the Equation set (\ref{eq:diskeq}) we find
+\begin{equation}
+\label{eq:Tdact}
+T_{\rm d, act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p}. \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}
+\end{equation}
+However, both accretion heating and stellar irradiation contribute to the thermal evolution of the disk. We thus compute the midplane temperature for our steady-state active disk as
+\begin{equation}
 \label{eq:activeT}
-T_{\rm d}^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
-\dot{M} & = & 3 \pi \nu \Sigma_{\rm d} \slabel{eq:Mdot},
-\end{subeqnarray}
-where $\dot{M}$, the mass flux, is constant in steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}.
+T_{\rm d}^4 = T_{\rm d, acc}^4 + T_{\rm d, pas}^4,
+\end{equation}
+where to avoid notation confusion $T_{\rm d, pas}=T_{\rm d}$ from Equation (\ref{eq:diskT}), the temperature profile for an active disk.
+We can then easily determine  $c_{\rm d}$ and $H_{\rm d}$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
+For consistency, we choose  $\alpha=0.01$ as in the previous case.
+Finally, we determine $\Sigma_{\rm d}$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}).
+We note that $\Sigma_{\rm d}$ is no longer constant when stellar irradiation is taken into account in the disk evolution.
+
+%Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady-state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):
+
+%\begin{subeqnarray}
+%\label{eq:activeT}
+%T_{\rm d}^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
+%\dot{M} & = & 3 \pi \nu \Sigma_{\rm d} \slabel{eq:Mdot},
+%\end{subeqnarray}
+%where $\dot{M}$, the mass flux, is constant in steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}.
 Based on disk observations (e.g., \citealt{andrews10}), we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$.
@@
 
-Solid particles in a disk orbit their host star at the Keplerian velocity $v_{\rm k} \equiv \Omega_{\rm k} r$.
+Solid particles in orbit their host star at the Keplerian velocity $v_{\rm k} \equiv \Omega_{\rm k} r$.
 The gas, however, experiences an additional pressure gradient, which causes it to rotate at sub-Keplerian velocity \citep{weidenschilling77}.
-Dust grains thus experience a headwind, which removes angular momentum, causing the solids to spiral inwards and fall onto the host star.
+Dust grains that are large enough thus experience a headwind, which removes angular momentum, causing the solids to spiral inwards and fall onto the host star.
 Small particles are well-coupled to the gas, while large planetesimals are decoupled from the gas.
@@
 \end{equation}
-where $E_x$ is the adsorption binding energy, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ is the molecule's vibrational frequency, with $\mu_x$ the mean molecular weight.
+where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ is the molecule's vibrational frequency, with $\mu_x$ the mean molecular weight.
 We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T_{\rm d}$.
@@
 \label{eq:tdrift}
-t_{\rm r, drift} \sim \frac{r}{\dot{r}},
+t_{\rm r, drift} \sim \Big|\frac{r}{\dot{r}}\Big|,
 \end{equation}
@@
 
-Of particular interest for our purposes is the particle size regime for which $t_{\rm r, drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ or $t_{\rm gas, acc} \lesssim t_{\rm des} \lesssim t_{\rm d}$, where $t_{\rm d}=3$ Myr is the disk lifetime.
+Of particular interest for our purposes is the particle size regime for which (1) $t_{\rm r, drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
+for $\sim$$0.5$ cm $\lesssim s \lesssim$ $1000$ cm, or (2) $t_{\rm gas, acc} \lesssim t_{\rm des} \lesssim t_{\rm d}$, i.e.
+for $\sim$$0.1$ cm $\lesssim s \lesssim$ $10$ cm.
 In these cases, radial drift or gas accretion (or both) are faster than thermal desorption.
-Particles of sizes that satisfy these requirements will drift significantly due to radial drift or gas accretion before desorbing, thus moving the H$_2$O, CO$_2$ and CO snowlines closer towards the central star and changing the C/O ratio throughout the disk.
+We note that $t_{\rm gas, acc}<t_{\rm d}$ always holds true.
+Particles of sizes that satisfy the requirements above will drift significantly due to radial drift or gas accretion before desorbing, thus moving the H$_2$O, CO$_2$ and CO snowlines closer towards the central star and changing the C/O ratio throughout the disk.
 We quantify these effects in sections \ref{sec:snowlines} and \ref{sec:COratio}.
@@
 %\vspace{-0.5in}
-\caption{Desorption distance as a function of a particle's initial location in the disk, for a range of particle sizes, and for both a passive disk (left panels) and an active disk (right panels).
+\caption{Desorption distance (if a grain fully desorbs) or final distance (if a grain does not fully desorb), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both a passive disk (left panels) and an active disk (right panels).
 The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
 The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
-For the range of particle sizes that fully desorb during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particles' initial location.}
+For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.}
 \label{fig:snowlines}
@@
 The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The analytic approximation is in good agreement with the numerical result for most cases.}
+The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
 \label{fig:an_vs_actual}
@@
 
-As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully evaporates.
-Thus a particle's final location will depend on whether a grain initially at a certain distance is completely desorbed or not within the disk lifetime of 3 Myr.
+%As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully evaporates.
+
+%Thus a particle's final location will depend on whether a grain initially at a certain distance is completely desorbed or not within the disk lifetime of 3 Myr.
 For example, larger grains take longer to desorb and hence are more likely to not evaporate fully in a given timeframe.
 
+We define the final position of a grain as the disk location it has reached after $t_{\rm d}=3$ Myr, or the radius at which it completely desorbs if that happens after a time shorter than 3 Myr.
 Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both a passive and an active disk.
-We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}).
+We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state disk results in Section \ref{COratio}.
 Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for a passive nor for an active disk.
@@
 Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance in a passive and an active disk (see Figure \ref{fig:snowlines}).
-We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it slightly deviates for smaller particles.
-\emgr{I've yet to figure out why there is a discrepancy for the small grains.
+We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it slightly deviates for particles larger than $s \gtrsim10$ cm.
+For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm r, drift}$ is a power-law in $r$ (for our parameters, $t_{\rm r, drift} \propto r^{-1/14}$ for the passive disk), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
+Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm r, drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
+
+%\emgr{I've yet to figure out why there is a discrepancy for the small grains.
 I'll keep thinking about it, but for now I'll leave it as it is since I don't want to give a false explanation just to have one.}
@@
 We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star.
-In Section \ref{sec:snowlines}, we showed that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
+Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
+Solid H$_2$O particles evaporate almost instantly, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
+A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
+The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
+Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
+It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active).
+The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane.
+
+%In Section \ref{sec:snowlines}, we showed that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
 Figure \ref{fig:s_t_a} confirms both of these effects.
@@
 Figure \ref{fig:s_t_a} thus proves claims (1) and (2) above.
-It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active).
-The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar_test.pdf}
+\includegraphics[width=0.55\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar_complete.pdf}
 %\vspace{-0.5in}
@@
 We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
-\emgr{Is it necessary to reproduce that table?
-Seems a bit redundant.}.
-Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk with a passive temperature profile, and a steady-state active disk \emgr{(not yet there)}.
+Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk with a passive temperature profile, and a steady-state active disk.
 The C/O ratio for a static disk is shown as a guideline.
 The plot is consistent with Figure \ref{fig:snowlines}.
-For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus form a snowline.
+For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus move the snowline compared to the static disk.
 In contrast, even $\sim$micron-sized grains drift and desorb for the active disk, since they flow towards the host star together with the accreting gas.
-For the same particle size, the snowline locations are slightly closer to the central star in the active disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
+For the same particle size, the snowline locations are slightly closer to the central star in the active disk with a passive temperature profile, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
 Equation \ref{eq:rdotact}).
+The addition of accretional heating in the steady-state active disk moves the H$_2$O snowline outwards.
+This is due to the fact that accretional heating dominates in the inner disk, where high temperatures cause the grains to evaporate further away from the star.
+Once $r\gtrsim1-2$  AU, stellar irradiation dominates the thermal evolution of the disk, and therefore the CO$_2$ and CO snowlines locations are the same as in the active disk with a passive temperature profile.
+
 Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases.
 While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk.
-However, bodies larger than $\sim$7 m will evaporate at the same location as the meter-sized planetesimals.
+Bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals, though only the kilometer-sized boulders that are very close to the snownlines will contribute to determining the snowline location.
 Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
@@
 Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution.
-\emgr{There will be 1-2 additional sentences for the third panel, the steady-state disk, once that finishes running.}
+
+For our choice of parameters, the snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk with a passive temperature profile and $r_{\rm{H_2O}} \approx 0.63$ AU for the steady-state disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; and $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
+For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
+Radial drift and gas accretion therefore push the snowline locations by up to $\sim$$60$ \% for H$_2$O and CO$_2$, and by up to $\sim$$50$ \% for CO.
+
+%\emgr{There will be 1-2 additional sentences for the third panel, the steady-state disk, once that finishes running.}
 
@@
 
-\section{Discussion and Model Limitations}
+\section{Discussion}
 \label{sec:discussion}
@@
 These sizes are representative since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
-We perform the simulations at four representative timescales in the disk evolution to see how the snowline locations evolve with time and affect the C/O ratio.
-Particles that start at large stellocentric distances do not desorb within the shorter timeframes, e.g.
-$10^4$ or $10^5$ years.
-However, they do evaporate if their initial location is closer to the host star.
-As stated in Section \ref{sec:COratio}, this implies that the grains form a snowline even after $10^4$ years.
-More importantly, for a given initial grain size, the snowline locations are independent of the time elapsed.
+We stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively, to see how the snowline locations evolve with time and affect the C/O ratio.
+Particles of a given size that start at large stellocentric distances do not desorb within the shorter timeframes, e.g.
+$10^4$ or $10^5$ years, but they do evaporate at a fixed radius if their initial location is closer to the host star.
+While the amount of material that moves through the disk changes with time, the radius at which particles desorb and thus the snowline locations are independent of the time elapsed.
 Therefore, our results are not only valid at $t_{\rm d}$, but also throughout the earlier time evolution of the protoplanetary disk.
@@
 The C/O ratio is thus insensitive to the choice of $M_{\rm d}$.
-This result also has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
+%This result also has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
 Since the snowline locations are independent of disk mass, we expect transition disks to exhibit the same drift-desorption behavior as standard disks under the simplified assumptions of our model.
 
-\subsection{Neglected Effects}
+\subsection{Model Extensions}
 \label{sec:neglected}
@@
 \begin{table}[t!]
-\caption{Dynamical and chemical processes that affect the snowline locations of the main C and O carriers, H$_2$O, CO$_2$ and CO.
-The arrows signify the direction in which a particular process affects the snowlines: $\leftarrow$ means that the snowline is pushed closer to the host star, $\rightarrow$ means that the snowline is pushed further from the host star.
-The presence of both arrows means that the process may have both effects on the snowline location.}
+\caption{The effects of dynamical and chemical processes on snowline shapes and locations}
 \begin{center}
@@
 \textbf{Process} & \textbf{Effect} \\\hline
-Radial drift & $\leftarrow$ \\\hline
+Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star, $\rightarrow$ means that the snowline is pushed further from the host star.
+The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
 Gas accretion & $\leftarrow$ \\\hline
 Particle growth & $\rightarrow$ \\\hline
-Turbulent diffusion & $\rightarrow$ \\\hline
+Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
 Particle fragmentation & $\rightarrow$ $\leftarrow$ \\\hline
-Morphology & $\rightarrow$ \\\hline
+Grain morphology & $\rightarrow$ \\\hline
 Particle composition & $\rightarrow$ $\leftarrow$ \\\hline
+Disk gaps and holes & $\rightarrow$ \\\hline
 Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
 \end{tabular}
+
 \end{center}
@@
 In Section \ref{sec:COratio} we have shown that larger grains move the snowline locations closer in, but those locations remain fixed above a certain particle size.
-Once the solids grow larger than km-sized, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
+Particle growth will thus initially push the snowlines inwards.
+However, once the solids grow larger than km-sized, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
 It follows that grain growth will eventually push the snowline location outwards.
@@
 Turbulence causes eddies and vertical mixing, which are likely to reduce the radial gas accretion velocity.
-Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially, causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.
+Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially.
+Back-diffusion across the snowline will change the shape of the snowline, as well as the C/O ratio in gas and dust both inside and outside of the snowline, due to the reduction of gas-phase volatile abundance interior to the snowline.
+
+%causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.
 
@@
 
-\item \textbf{Morphology.} Our model assumes that the ice particles are perfect, homogeneous spheres.
+\item \textbf{Grain morphology.} Our model assumes that the ice particles are perfect, homogeneous spheres.
 However, this is not a very good approximation, since grain growth can be fractal rather than compact (\citealt{zsom10}, \citealt{okuzumi12}).
@@
 The grains may also be composed of a mixture of H$_2$O, CO$_2$ and CO ices.
-The relative fraction of each volatile, as well as their degree of mixing, will determine their desorption timescale, and as a result whether the snowlines are moved towards or away from the central star.
+In this case, the desorption energies are largest for the most volatile species, which will move the snowlines inwards.
+
+%The relative fraction of each volatile, as well as their degree of mixing, will determine their desorption timescale, and as a result whether the snowlines are moved towards or away from the central star.
+
+\item \textbf{Disk gaps and holes.} The snowline locations will be different for transition disks, which have inner cavities significantly depleted of gas (e.g., \citealt{espaillat12}), or pre-transitional disks, which have a gap between an inner and outer full disk (e.g., \citealt{kraus11}).
+The decrease in gas pressure in these gaps or holes will reduce the particles' drift velocity close to the gap edge, thus slowing them down and pushing the snowline outwards.
 
@@
 The effect of disk chemistry on snowline locations, shape, time evolution, or the C/O ratio is therefore difficult to estimate.
-In a future paper, we plan to parametrize the chemical model developed by Merchantz et al.
-(in preparation) that traces the evolution of the main C and O oxygen carriers, and incorporate it in our radial drift calculation.
+In a future paper, we plan to use a parametrized chemical model and incorporate it in our radial drift calculation.
+
+%parametrize the chemical model developed by Merchantz et al. (in preparation) that traces the evolution of the main C and O oxygen carriers,
 
@@
 
-\emgr{To be done soon.}
+We study the effect of radial drift of solids and viscous gas accretion onto the central star on the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in a protoplanetary disk, assuming static chemistry.
+We develop a simplified model to describe the coupled drift-desorption process and determine the time evolution of particles of different sizes throughout the disk.
+We assume that the solid particles are perfect, homogeneous spheres, fully composed of either H$_2$O, CO$_2$ or CO.
+We apply our model to a passive disk, an active disk with a passive temperature profile, and a steady-state active disk that also takes into account stellar irradiation.
+We determine the desorption or final location of drifting particles after a time equal to the disk lifetime, and use this result to set an inner limit for the location of the H$_2$O, CO$_2$ and CO snowlines.
+Our results can be summarized as follows:
 
-%\appendix
-%\section{...}
+\begin{enumerate}
+\item Radial drift and gas accretion affect desorption and move the snowline locations compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk.
+Moreover, boulders up to almost kilometer-size contribute to setting the snowline locations if they are very close to the snowlines.
+
+\item Particles with sizes in the above range desorb instantaneously, and at a fixed location in the disk that only depends on the particle size.
+Thus for each particle size there is a fixed and uniquely determined H$_2$O, CO$_2$ or CO snowline.
+
+\item The results of the numerical simulation are in agreement with the analytic solution of the drift-desorption system of differential equations if the stopping time $\tau_{\rm s} \ll 1$.
+We present an explicit analytic solution for the desorption distance in this regime.
+
+\item The snowline locations move inwards as the particle size increases; the innermost snowline is set by particles with initial size $s \sim 7$ m.
+Gas accretion causes even micron-sized particles to drift, desorb and move the snowline location compared to a static disk.
+A steady-state disk that includes accretion heating moves the H$_2$O snowline outwards compared to an active disk with a passive temperature profile, but has no effect on the CO$_2$ and CO snowline locations.
+
+\item Since realistic grain size distributions are dominated by the largest particles, the H$_2$O, CO$_2$ and CO snowlines are those created by the largest drifting particles in our model.
+This corresponds to the innermost snowlines that we determine.
+Our model thus sets a limit on how close to the central star the snowlines can be pushed by radial drift and gas accretion.
+
+\item For our fiducial model, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for a passive disk, 0.26 AU, 3.4 AU and 30 AU for an active disk with a passive temperature profile, and 0.63 AU, 3.4 AU and 30 AU for a steady-state active disk.
+Compared to a static disk, radial drift and gas accretion move the snowlines by up to 60 \% for H$_2$O and CO$_2$, and by up to 50 \% for CO.
+
+\item Our model finds that the C/O ratio is enhanced compared to the stellar value throughput most of the disk, with the C/O ratio reaching unity between the CO$_2$ and CO snowlines.
+This is consistent with observations of C/O ratios in some exoplanet atmospheres.
+
+\item The snowline locations are independent of the time at which we stop our simulation and of the total disk mass.
+Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.
+
+\end{enumerate}
+
+Our model does not address additional effects, such as gas diffusion, grain composition and morphology, or complex time-dependent chemical processes.
+Future work will address some of these dynamical and chemical processes, with the goal of obtaining more realistic results for the snowline locations, shapes and time evolution, and the resulting effect on the C/O ratio.
+%Nevertheless,
+
+\appendix
+\section{Steady-state active disk solution} \label{app:steadystate}
+
+Following \citet{shakura73} and \citet{armitage10}, the steady-state solution for a thin actively accreting disk with an $\alpha$-prescription for viscosity is governed by the following set of equations:
+
+\begin{subeqnarray}
+\label{eq:diskeq}
+\nu & = & \alpha c_{\rm d} H_{\rm d} \slabel{eq:nueq} \\
+c_{\rm d}^2 & = & \frac{k_{\rm B} T_{\rm d, act}}{\mu m_{\rm p}} \slabel{eq:cdeq} \\
+\rho_{\rm d} & = & \frac{1}{\sqrt{2 \pi}} \frac{\Sigma_{\rm d}}{H_{\rm d}} \slabel{eq:rhoeq} \\
+H_{\rm d} & = & \frac{c_{\rm d}}{\Omega_{k}} \slabel{eq:Heq} \\
+T_{\rm d, act}^4 & = & \frac{3}{4} \tau T_{\rm d, surf}^4 \slabel{eq:Teq} \\
+\tau & = & \frac{1}{2} \Sigma_{\rm d} \kappa \slabel{eq:taueq} \\
+\nu \Sigma_{\rm d} & = & \frac{\dot{M}}{3 \pi} \slabel{eq:Mdot} \\
+\sigma T_{\rm d, surf}^4 & = & \frac{9}{8} \nu \Sigma_{\rm d} \Omega_{\rm k}^2 \slabel{eq:nusigeq} \\
+\kappa & = & \kappa_0 T_{\rm d,act}^2 \slabel{eq:kappaeq},
+\end{subeqnarray}
+where $T_{\rm d, surf}$ is the surface temperature of the disk and the other quantities are defined in the main text.
+This is a system of nine equations with nine unknowns ($\nu$, $c_{\rm d}$, $H_{\rm d}$, $T_{\rm d, act}$, $\rho_{\rm d}$, $\Sigma_{\rm d}$, $\tau$, $T_{\rm d, surf}$, $\kappa$) that can be solved numerically once $\alpha$ and $\kappa_0$ are specified.
+
 %

### 2015-07-19 python/C_to_O_active_steady.py

@@ -1,0 +1,133 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+from drift_timescales import read_drag_file_many
+
+def n_C_O(species):
+    
+    """
+        
+    Returns the abundance of C and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_C:
+        abundance of C with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 1.5, 1.5
+    elif species == 'CO2':
+        return 0.3, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0.6, 0
+    elif species == 'silicate':
+        return 0, 1.4
+
+
+#in the following we calculate T_freeze for various species
+s, a, arrayH2O = read_drag_file_many('s_neg3_8_steady_correct.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO2', 50, 50)
+s, a, arrayCO = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO', 50, 50)
+
+
+
+
+def n(index, r, elem):
+    
+    """
+    
+    Returns the gaseous and solid abundance of a given element (C, O)
+    as a function of temperature, e.g. between different snowlines
+    
+    Input
+    -----
+    T:
+        disk midplane temperature in units of K
+    elem:
+        string: 'C' or 'O'
+        
+    Output
+    ------
+    n_gas:
+        abundance of element 'elem' in gaseous form relative to H2 abundance
+    n_solid:
+        abundance of element 'elem' in solid form relative to H2 abundance
+    
+    """
+    
+    if elem == 'C':
+        i = 0 #flag to simplify the use of the n_C_O function; = 0 for C
+    elif elem == 'O':
+        i = 1 #flag to simplify the use of the n_C_O function; = 1 for O
+    
+    size = s[index]
+    aH2O = arrayH2O[index][-1]
[…]

### 2015-07-19 python/drift_timescales_updated_sigmap_0402_cavity.py

@@ -335,7 +335,7 @@ def Sigmadisk(r, t, alpha, Mdisk, rc, T0 = 120, betaT = 3./7, mu = 2.35, Mstar =
             * np.exp(- rtild**(2 - gammad) / T)
      else:
          return Mdisk / (2 * np.pi * rc**2 * rtild**gammad) * T**(- (2.5 - gammad) / (2 - gammad)) \
-            * np.exp(- rtild**(2 - gammad) / T) / 1000
+            * np.exp(- rtild**(2 - gammad) / T) / 100
                 
 
 def vacc_act(r, t, alpha, mu = 2.35, T0 = 120, betaT = 3./7, rc = 100 * AU, Mstar = Msun, gammadflag = 0):

### 2015-07-20 tex/draft/ms.tex
update

@@
 
-\title{C/O in Protoplanetary Disks: The Effect of Radial Drift and Viscous Accretion}
+\title{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Accretion}
 
@@
 ...
+%We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the H$_2$O, CO$_2$ and CO snowline locations in a protoplanetary disk, and the resulting C/O ratio in gas and dust.
 \end{abstract}
@@
 
-\textbf{Passive disk.} A passive disk is only irradiated by the central star and does not experience accretion heating.
+\textbf{Passive disk.} A passive disk is only irradiated by the central star and does not experience radial movement of gas or accretion heating.
 We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}.
@@
 \label{eq:disk}
-\Sigma_{\rm d}&=&2200\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
+\Sigma_{\rm d}&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
 T_{\rm d} &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
@@
 \label{eq:Tdact}
-T_{\rm d, act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p}. \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}
+T_{\rm d, act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
 \end{equation}
@@
 \label{eq:activeT}
-T_{\rm d}^4 = T_{\rm d, acc}^4 + T_{\rm d, pas}^4,
+T_{\rm d}^4 = T_{\rm d, act}^4 + T_{\rm d, pas}^4,
 \end{equation}
@@
 Finally, we determine $\Sigma_{\rm d}$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}).
-We note that $\Sigma_{\rm d}$ is no longer constant when stellar irradiation is taken into account in the disk evolution.
+We note that $\Sigma_{\rm d}$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.
 
@@
 \end{equation}
-where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ is the molecule's vibrational frequency, with $\mu_x$ the mean molecular weight.
+where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency, with $\mu_x$ the mean molecular weight.
 We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T_{\rm d}$.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/drift_timescales_betaS1_gas_acc_new.pdf}
+%\vspace{-0.5in}
+\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
+The timescales are calculated at three representative locations, i.e.
+the H$_2$O, CO$_2$ and CO snowlines in the static disk.
+For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively.
+The horizontal dashed line represents a typical disk lifetime of 3 Myr.
+Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
+$t_{\rm r, drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.}
+\label{fig:timescales}
+\end{figure}
+
 For simplicity purposes, we calculate the radial drift timescale, $t_{\rm r, drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well.
@@
 We quantify these effects in sections \ref{sec:snowlines} and \ref{sec:COratio}.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{../../figs/drift_timescales_betaS1_gas_acc_new.pdf}
-%\vspace{-0.5in}
-\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
-The timescales are calculated at three representative locations, i.e.
-the H$_2$O, CO$_2$ and CO snowlines in the static disk.
-For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively.
-The horizontal dashed line represents a typical disk lifetime of 3 Myr.
-Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
-$t_{\rm r, drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.}
-\label{fig:timescales}
-\end{figure}
 
@@
 Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both a passive and an active disk.
-We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state disk results in Section \ref{COratio}.
+We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state active disk results in Section \ref{sec:COratio}.
 Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for a passive nor for an active disk.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.55\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar_complete.pdf}
+%\vspace{-0.5in}
+\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel), an active disk with a passive temperature profile (middle panel) and a steady-state active disk (bottom panel), for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
+The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
+The horizontal dotted line represents the stellar value of 0.54.
+The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk.
+The snowline location moves inward as the particle size increases.}
+\label{fig:CO_ratio}
+\end{figure}
+
 We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star.
@@
 Figure \ref{fig:s_t_a} thus proves claims (1) and (2) above.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.55\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar_complete.pdf}
-%\vspace{-0.5in}
-\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel) and for an active disk (bottom panel), and for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
-The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
-The horizontal dotted line represents the stellar value of 0.54.
-The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk.
-The snowline location moves inward as the particle size increases.
-\emgr{Some additional text for the currently non-existent third panel.}}
-\label{fig:CO_ratio}
-\end{figure}
 
@@
 
-For our choice of parameters, the snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk with a passive temperature profile and $r_{\rm{H_2O}} \approx 0.63$ AU for the steady-state disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; and $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
+For our choice of parameters, the snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk with a passive temperature profile and $r_{\rm{H_2O}} \approx 0.63$ AU for the steady-state disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
 For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
@@
 
-Our goals in this paper were (1) to gain a detailed qualitative and quantitative understanding of the effect of radial drift and gas accretion on to the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
+Our goals in this paper were (1) to gain a detailed qualitative and quantitative understanding of the effect of radial drift and gas accretion onto the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
 We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical processes.
@@
 We note that our future work will address some of these issues.
+
+We summarize in Table 1 the potential physical and chemical processes occurring in disks and their effect on snowline locations.
+For the sake of completion, Table 1 also includes the processes addressed in this paper, i.e.
+radial drift and gas accretion.
+The neglected effects are discussed in more detail below.
+
+\begin{enumerate}
+\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered fixed for a given grain throughout its drift and evolution.
+However, grain growth has been observed in protoplanetary disks (e.g., \citealt{ricci10}, \citealt{perez12}), as well as theoretically constrained (e.g., \citealt{birnstiel10}, \citealt{birnstiel12}).
+In Section \ref{sec:COratio} we have shown that larger grains move the snowline locations closer in, but those locations remain fixed above a certain particle size.
+Particle growth will thus initially push the snowlines inwards.
+However, once the solids grow larger than km-sized, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
+It follows that grain growth will eventually push the snowline location outwards.
+
+\item \textbf{Turbulent diffusion.} The radial drift model presented in Section \ref{sec:drift} only considers a laminar flow and thus ignores turbulence.
+However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
+Turbulence causes eddies and vertical mixing, which are likely to reduce the radial gas accretion velocity.
+Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially.
+Back-diffusion across the snowline will change the shape of the snowline, as well as the C/O ratio in gas and dust both inside and outside of the snowline, due to the reduction of gas-phase volatile abundance interior to the snowline.
 
@@
 \end{table}
-
-We summarize in Table 1 the potential physical and chemical processes occurring in disks and their effect on snowline locations.
-For the sake of completion, Table 1 also includes the processes addressed in this paper, i.e.
-radial drift and gas accretion.
-The neglected effects are discussed in more detail below.
-
-\begin{enumerate}
-\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered fixed for a given grain throughout its drift and evolution.
-However, grain growth has been observed in protoplanetary disks (e.g., \citealt{ricci10}, \citealt{perez12}), as well as theoretically constrained (e.g., \citealt{birnstiel10}, \citealt{birnstiel12}).
-In Section \ref{sec:COratio} we have shown that larger grains move the snowline locations closer in, but those locations remain fixed above a certain particle size.
-Particle growth will thus initially push the snowlines inwards.
-However, once the solids grow larger than km-sized, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
-It follows that grain growth will eventually push the snowline location outwards.
-
-\item \textbf{Turbulent diffusion.} The radial drift model presented in Section \ref{sec:drift} only considers a laminar flow and thus ignores turbulence.
-However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
-Turbulence causes eddies and vertical mixing, which are likely to reduce the radial gas accretion velocity.
-Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially.
-Back-diffusion across the snowline will change the shape of the snowline, as well as the C/O ratio in gas and dust both inside and outside of the snowline, due to the reduction of gas-phase volatile abundance interior to the snowline.
-
 %causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.
@@
 
-\item Our model finds that the C/O ratio is enhanced compared to the stellar value throughput most of the disk, with the C/O ratio reaching unity between the CO$_2$ and CO snowlines.
+\item Our model finds that the C/O ratio is enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching unity between the CO$_2$ and CO snowlines.
 This is consistent with observations of C/O ratios in some exoplanet atmospheres.
@@
 
+\section{Desorption distance analytic solution} \label{app:tdriftan}
+
+For a particle of size $s$ that desorbs and satisfies $\tau_{\rm s} \ll 1$ ($\tau_{\rm s}$ is the dimensionless stopping time, defined in Section \ref{sec:drift}), we can derive an explicit analytic solution for the particle's desorption distance in a passive disk.
+For $\tau_{\rm s} \ll 1$, a particle is in the Epstein drag regime (see Equation \ref{eq:ts}) and its drift velocity $\dot{r}$ (Equation \ref{eq:rdotpas}) can be approximated as
+\begin{equation}
+\label{eq:rdotapp}
+\dot{r} \approx -2 \eta \Omega_{\rm k} r \tau_{\rm s}.
+\end{equation}
+By using Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and setting $t_{\rm r, drift}=t_{\rm des}$, we can express a particle's desorption distance as
+
+\begin{equation}
+\label{eq:tan}
+r_{\rm des}=\Bigg(\frac{d\,\, \mathcal{W}\Big[\frac{(B/A)^{-q/d} q\,C}{d} q C\Big]}{q \,C}\Bigg)^{\frac{1}{q}},
+\end{equation}
+where $\mathcal{W}$ is the Lambert-W function, $q=3/7$ is the power-law coefficient in Equation (\ref{eq:diskT}), $d=-\frac{1}{2}+p-q$ with $p=1$ the power-law coefficient in Equation (\ref{eq:disksigma}), and
+\begin{subeqnarray}
+%\label{eq:tan}
+A & = &  \frac{\rho_0}{\rho_{\rm s}} \frac{r_0^2}{s c_0} r_0^d \\ 
+B & = & \frac{\rho_{\rm s} s}{3 \mu_{\rm x} N_{\rm x} \nu_{\rm x}} \\
+C & = & \frac{E_{\rm x}}{T_0} r_0^{-q}, \\
+\end{subeqnarray}
+where $r_0=1$ AU, $\rho_0=\rho_{\rm d}(r_0)$ and $c_0=c_{\rm d}(r_0)$.
+
 %

### 2015-07-20 python/drift_timescales.py

@@ -1,10 +1,11 @@
-from utils.constants import G, kb, mp, Msun, cmperau
+from utils.constants import G, kb, mp, Msun, cmperau, AU
 import numpy as np
 from T_freeze import T_freeze, Rdes, tevap, vib_freq
 from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
 from scipy.integrate import odeint
 from scipy.optimize import brentq
 from scipy.interpolate import interp1d
+from scipy.special import lambertw
 from utils.zbrac import zbrac
 
 def Tdisk(r, T0 = 120, betaT = 3./7):
@@ -305,6 +306,20 @@ def r_stop(s, mx, Ex, Nx = 1e15, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, 
           return brentq(f, 1e-3, 1e2)
      except ValueError:
           return 1e-10
+
+
+def ran(s, mx, Ex, Nx = 1e15, rhos = 3.0, T0 = 120, betaT = 3./7, mu = 2.35, Sigma0 = 2200, betaS = 3./2, \
+    Mstar = Msun, sigma = 2 * 10**(-15), r0 = AU):
+    
+    d = -1./2 + betaS - betaT
+    c = Ex*r0**(-betaT)/T0
+    B = rhos*s/(3*mx*mp*Nx*vib_freq(Ex, mx))
+    
+
+    A = np.sqrt(8/np.pi)*np.sqrt(G*Mstar)*mu*mp*Sigma0*r0**(1./2+d) / (np.sqrt(2*np.pi)*kb*rhos*s*T0)
+
+    dist = (d * lambertw((B/A)**(-betaT/d) * betaT * c / d) / (betaT*c))**(1./betaT)
+    return np.real(dist)/AU
 
 
 def r_freeze(mx, Ex, nx, T0 = 120., betaT = 3./7):

### 2015-07-20 tex/draft/ms.tex
update

@@
 
-\author{Ana-Maria A. Piso\altaffilmark{1}, et al}. %Karin I. \"Oberg\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}, Tilman Birnstiel\altaffilmark{1}}
+\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
@@
 \begin{abstract}
-...
-%We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the H$_2$O, CO$_2$ and CO snowline locations in a protoplanetary disk, and the resulting C/O ratio in gas and dust.
+The C/O ratio is an important signature of giant planet atmospheres and disk chemistry.
+We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and the resulting C/O ratio in gas and dust throughout the disk.
+We determine the snowline locations for a range of fixed initial particle sizes, in both an active and a passive disk.
+We find that grains with sizes sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk desorb instantaneously and at a fixed size-dependent location in the disk regardless of the particle's initial position.
+The snowline location move inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
+Compared to a static disk, we find that radial drift and gas accretion move the H$_2$O and CO$_2$ snowlines inwards by up to 60 \%, and the CO snowline by up to 50 \%.
+Our results thus set an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
+Our results
+have direct implications on the C/O ratio in giant planet atmospheres that have
+formed in situ. 
 \end{abstract}
@@
 
-\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture.
+%\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture.
 Gas giants are important.
@@
 
-Main sequence stars commonly host giant planets (refs).
-The chemical composition of gas giant atmospheres can provide important constraints on their formation, accretion and migration history.
-\emgr{Add the rest as outlined above.} \\
+The chemical composition of the atmospheres of giant planets can provide important constraints on their formation, accretion and migration history.
+These processes are tightly linked to the structure and evolution of the protoplanetary disk in which such planets form.
+It is thus essential to understand the disk chemistry and dynamics well enough to predict the types of planet compositions that result from planet formation in different parts of the disk, as well as backtrack the planet formation location based on planet compositions.
 
-\emgr{Paragraph 2: disks are complex, a lot of dynamical and chemical processes going on.
+%\emgr{Add the rest as outlined above.} \\
+
+%\emgr{Paragraph 2: disks are complex, a lot of dynamical and chemical processes going on.
 However, we have detections of organic molecules.
@@
 
-In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11b}, \citealt{oberg11c}, ...).
-Of particular importance are volatile compounds, since the location of their snowlines determines their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
-\emgr{Add part about snowlines, complex chemistry, many molecules.} \\
+In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11c}, \citealt{oberg11b}).
+Of particular importance are volatile compounds, such as H$_2$O, CO$_2$, CO, N$_2$ etc.
+Recent detections of the disk snowlines (such as the CO snowline, \citealt{qi11}) are significant, since the snowline locations of volatile molecules determine their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
+%Recent detections of the CO snowline in disks (e.g., \citealt{qi11}) provide the framework...
 
-\emgr{Paragraph 3: let's focus on one (set of) molecules: the C/O ratio.
+%\emgr{Add part about snowlines, complex chemistry, many molecules.} \\
+
+An important signature of disk chemistry is the carbon-to-oxygen (C/O) ratio, as it provides information about the relative abundances of the main C and O carrier species.
+While the gas C/O ratio in disks may be potentially observable \citep{du15}, such detections do not yet exist.
+However, we can indirectly infer the relative abundance of disk C and O using detections of C/O ratios in exoplanet atmospheres.
+Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}.
+One explanation for this discrepancy was proposed by \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e.
+H$_2$O, CO$_2$ and CO, have different condensation temperatures.
+This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
+\citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
+
+%\emgr{Paragraph 3: let's focus on one (set of) molecules: the C/O ratio.
 It's an important signature of atmospheric chemistry.
@@
 
-\emgr{Start with the fact that we can't see C/O in disks directly.
-Then transition to planet atmospheres and modify the next sentence to follow up logically.} Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio.
-Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}.
-One explanation for this discrepancy was proposed by \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e.
-H$_2$O, CO$_2$ and CO, have different condensation temperatures.
-This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
-\citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-\\
+%\emgr{Start with the fact that we can't see C/O in disks directly. Then transition to planet atmospheres and modify the next sentence to follow up logically.} 
+%Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio.
 
-\emgr{Paragraph 4: the static disk model is a great start, but there are additional dynamical and chemical processes to take into account.
+\citet{oberg11} assume a static disk with passive chemistry.
+In reality, additional dynamical and chemical processes affect the snowline locations and the resulting C/O ratio.
+Several works have addressed some of these effects.
+\citet{madhu14} use a steady-state active disk model that includes planetary migration and use the C/O ratio to constrain migration mechanisms.
+\citealt{alidib14} calculate the C/O ratio throughout the disk by incorporating the evolution of solids, i.e.
+radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors.
+\citet{thiabaud15} consider additional volatile species in their chemical network, such as N$_2$ or NH$_3$, and find that the C/O ratio may be enriched by up to four times the Solar value in certain parts of the disk.
+
+Most studies of the C/O ratio in disks are particularly focused on the inner disk, in order to explain observations of super-stellar C/O ratios in the atmospheres of close-in exoplanets \citep{madhu11}.
+In addition, the simultaneous inclusion of various dynamical and/or chemical processes makes it difficult to quantify the separate effect of each of these processes on snowline locations and the C/O ratio.
+In this paper, we focus on two particular dynamical effects i.e.
+the radial drift of solids and viscous gas accretion.
+We enhance the model of \citet{oberg11} by considering these two additional processes.
+We perform a systematic study to understand the detailed qualitative and quantitative effects of radial drift and gas accretion on the H$_2$O, CO$_2$ and CO snowline locations, and the resulting C/O ratio in gas and dust throughout the protoplanetary disk.
+More importantly, we obtain a limit on how close to the star the snowline locations can be pushed by radial drift and gas accretion.
+
+%\emgr{Paragraph 4: the static disk model is a great start, but there are additional dynamical and chemical processes to take into account.
 Several studies have done that in some form or another --- cite Ali-Dib+14, Madhusudhan+14, Thiabaud+15.
@@
 
-\emgr{Paragraph 5: finally introduce our goal for this paper: study the dynamical effects such as drift and gas accretion.
+%\emgr{Paragraph 5: finally introduce our goal for this paper: study the dynamical effects such as drift and gas accretion.
 State our goals: understand the detailed qualitative and quantitative effect of drift and gas accretion on snowline locations, obtain a limit on how far in can the snowlines be pushed, and see how that affects the C/O ratio throughout the disk.
@@
 
-\emgr{This needs some restructuring and additions along the lines discussed above.} In order to obtain more realistic estimates for C/O ratios across protoplanetary disks, dynamical processes and the disk evolving chemistry have to be taken into account.
+%\emgr{This needs some restructuring and additions along the lines discussed above.} In order to obtain more realistic estimates for C/O ratios across protoplanetary disks, dynamical processes and the disk evolving chemistry have to be taken into account.
 In this paper, we enhance the model of \citet{oberg11} considering two additional dynamic effects: (1) the radial drift of solids throughout the protoplanetary disk, and (2) the viscous accretion of the disk gas onto the host star.
@@
 
-This paper is organized as follows: \emgr{(section summaries)}.
+This paper is organized as follows.
+In Section \ref{sec:model}, we present our disk, radial drift and desorption models, as well as the timescales relevant to the coupled drift-desorption process.
+We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for a passive and an active disk in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}.
+In Section \ref{sec:discussion}, we discuss the generality of our results, as well as additional effects on the snowline locations.
+Finally, we summarize our findings in Section \ref{sec:summary}.
 
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_transition_disk_1000.pdf}
+%\vspace{-0.5in}
+\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines).
+The timescales of the simulations and their color code are the same as in Figure \ref{fig:timeplots}.
+Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge.}
+\label{fig:cavity}
+\end{figure}
+
 \subsection{Generality of Results: Dependence on Disk Parameters}
@@
 Since the snowline locations are independent of disk mass, we expect transition disks to exhibit the same drift-desorption behavior as standard disks under the simplified assumptions of our model.
+
+We also apply our active disk model to a transition disk, i.e.
+a protoplanetary disk with an inner cavity significantly depleted of gas.
+We choose a disk with an inner gap of radius $r_0=4$ AU, consistent with observations of TW Hya \citep{zhang13}, and with the surface density in the gap reduced by a factor of 1000.
+Figure \ref{fig:cavity} shows the desorption or final distance for a H$_2$O particle of initial size of 1 m, with the simulation stopped at the same timescales as in Figures \ref{fig:timeplots} and \ref{fig:varMd}.
+Particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the edge gap, due to the decrease in gas pressure inside the cavity, thus forming a snowline at $\sim$$3.8$ AU.
+This is consistent with the observations of \citet{zhang13} for the H$_2$O snowline being pushed outwards in a transition disk compared to a full disk.
 
@@
 \section{Summary}
+\label{sec:summary}
 

### 2015-07-20 tex/draft/ms.tex
update

@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/drift_timescales_betaS1_gas_acc_new.pdf}
+\includegraphics[width=0.5\textwidth]{drift_timescales_betaS1_gas_acc_new.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{../../figs/desorption_distance_passive_active_colorbar_test.pdf}
+\includegraphics[width=0.7\textwidth]{desorption_distance_passive_active_colorbar_test.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active_new.pdf}
+\includegraphics[width=0.7\textwidth]{desorption_distance_actual_vs_estimated_passive_active_new.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{../../figs/s_t_a.pdf}
+\includegraphics[width=0.8\textwidth]{s_t_a.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.55\textwidth]{../../figs/C_O_ratio_passive_active_disk_many_colorbar_complete.pdf}
+\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=\textwidth]{../../figs/time_plots.pdf}
+\includegraphics[width=\textwidth]{time_plots.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_varying_Md.pdf}
+\includegraphics[width=0.5\textwidth]{desorption_distance_varying_Md.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_transition_disk_1000.pdf}
+\includegraphics[width=0.5\textwidth]{desorption_distance_transition_disk_1000.pdf}
 %\vspace{-0.5in}

### 2015-07-20 tex/draft/ms.tex
update

@@
 a protoplanetary disk with an inner cavity significantly depleted of gas.
-We choose a disk with an inner gap of radius $r_0=4$ AU, consistent with observations of TW Hya \citep{zhang13}, and with the surface density in the gap reduced by a factor of 1000.
+We choose a disk with an inner gap of radius $r_0=4$ AU, consistent with observations of TW Hya \citep{zhang13}, and with the gas surface density in the gap reduced by a factor of 1000.
 Figure \ref{fig:cavity} shows the desorption or final distance for a H$_2$O particle of initial size of 1 m, with the simulation stopped at the same timescales as in Figures \ref{fig:timeplots} and \ref{fig:varMd}.
-Particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the edge gap, due to the decrease in gas pressure inside the cavity, thus forming a snowline at $\sim$$3.8$ AU.
-This is consistent with the observations of \citet{zhang13} for the H$_2$O snowline being pushed outwards in a transition disk compared to a full disk.
+Particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the gap edge, due to the decrease in gas pressure inside the cavity, thus forming a snowline at $\sim$$3.8$ AU.
+This is consistent with the observations of \citet{zhang13}, which show that the H$_2$O snowline is pushed outwards in a transition disk compared to a full disk.
 

### 2015-07-20 tex/draft/ms.tex
update

@@
 
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
+\author{Ana-Maria A. Piso\altaffilmark{1}, et al.}%, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
@@
 \begin{abstract}
-The C/O ratio is an important signature of giant planet atmospheres and disk chemistry.
+The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry.
 We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and the resulting C/O ratio in gas and dust throughout the disk.
@@
 Compared to a static disk, we find that radial drift and gas accretion move the H$_2$O and CO$_2$ snowlines inwards by up to 60 \%, and the CO snowline by up to 50 \%.
-Our results thus set an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
+We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
 Our results
@@
 Of particular importance are volatile compounds, such as H$_2$O, CO$_2$, CO, N$_2$ etc.
-Recent detections of the disk snowlines (such as the CO snowline, \citealt{qi11}) are significant, since the snowline locations of volatile molecules determine their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
+Recent detections of the snowlines of volatiles (such as the CO snowline, \citealt{qi11}) are significant, since the snowline locations of these molecules determine their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
 %Recent detections of the CO snowline in disks (e.g., \citealt{qi11}) provide the framework...
@@
 In addition, the simultaneous inclusion of various dynamical and/or chemical processes makes it difficult to quantify the separate effect of each of these processes on snowline locations and the C/O ratio.
-In this paper, we focus on two particular dynamical effects i.e.
+In this paper, we focus on two particular dynamical effects, i.e.
 the radial drift of solids and viscous gas accretion.

### 2015-07-22 tex/draft/ms_kedit.tex
draft update

--- 2015-07-20 tex/draft/ms.tex
@@
 
-\author{Ana-Maria A. Piso\altaffilmark{1}, et al.}%, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
+\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
@@
 \begin{abstract}
-The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry.
-We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and the resulting C/O ratio in gas and dust throughout the disk.
+%The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry. 
+The C/O ratio is a defining feature of both gas giant atmospheric and protoplanetary disk chemistry.
+In disks, the C/O ratio is regulated by the presence of snowlines of major volatiles at different distances from the central star.
+We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO.
+We calculate the resulting C/O ratio in gas and dust throughout the disk.
 We determine the snowline locations for a range of fixed initial particle sizes, in both an active and a passive disk.
-We find that grains with sizes sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk desorb instantaneously and at a fixed size-dependent location in the disk regardless of the particle's initial position.
-The snowline location move inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
+We find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk desorb %instantaneously and [I think instantaneously is too strong]
+at a %fixed 
+size-dependent location in the disk, which is independent %regardless 
+of the particle's initial position.
+The snowline radius decreases for larger particles, up to sizes of $\sim$7~m.
+%location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
 Compared to a static disk, we find that radial drift and gas accretion move the H$_2$O and CO$_2$ snowlines inwards by up to 60 \%, and the CO snowline by up to 50 \%.
 We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
-Our results
-have direct implications on the C/O ratio in giant planet atmospheres that have
-formed in situ. 
+%Our results have direct implications on the C/O ratio in giant planet atmospheres that have formed in situ.
+[I don't think we discuss this enough to warrant it as a conclusion
 \end{abstract}
@@
 
-The chemical composition of the atmospheres of giant planets can provide important constraints on their formation, accretion and migration history.
-These processes are tightly linked to the structure and evolution of the protoplanetary disk in which such planets form.
-It is thus essential to understand the disk chemistry and dynamics well enough to predict the types of planet compositions that result from planet formation in different parts of the disk, as well as backtrack the planet formation location based on planet compositions.
+The chemical composition of the atmospheres of giant planets can provide important constraints on %their [unclear association]
+the planet formation, accretion and migration history. 
+%[Need explicit statement that gas giants accrete their atmospheres from the nebular gas.] 
+Gas giants accrete their envelopes from the nebular gas.
+As such, planet compositions are tightly linked to the structure and evolution of the protoplanetary disk in which they form.
+It is thus essential to understand the disk chemistry and dynamics well enough to (1) predict the types of planet compositions that result from planet formation in different parts of the disk, and (2) backtrack the planet formation location based on planet compositions.
+%[split into two sentences]
 
@@
 
-In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11c}, \citealt{oberg11b}).
-Of particular importance are volatile compounds, such as H$_2$O, CO$_2$, CO, N$_2$ etc.
+%In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11c}, \citealt{oberg11b}).
+[the molecules relevant for this paper are not organics]
+%[Topic sentence on that the most important volatiles are ... based on protostellar and comet abundances.]
+
+The structure of protoplanetary disks in complex, and affected by a multitude of chemical and dynamical processes (see review by \citealt{henning13}).
+From the chemistry perspective, volatile compounds are particularly important.
+Their snowline locations determine their relative abundance in gaseous and solid form in the disk, and thus the chemical composition of nascent giant planets.
+Based on protostellar and comet abundances, some of the most important volatile molecules are H$_2$O, CO$_2$, CO, N$_2$.
+Recent observations of protoplanetary disks have provided valuable information about the abundance and snowline locations of some of these compounds.
+For example, the CO snowline has been detected in the disks around HD 163296 \citep{qi11} and TW Hya \citep{qi13}.
+Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}.
+With the advent of ALMA, more such detections are expected.
+An interpretive framework of these observations is thus crucial regardless of their importance for planet compositions.
+
+%Of particular importance are volatile compounds, such as H$_2$O, CO$_2$, CO, N$_2$ etc.
 Recent detections of the snowlines of volatiles (such as the CO snowline, \citealt{qi11}) are significant, since the snowline locations of these molecules determine their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
-%Recent detections of the CO snowline in disks (e.g., \citealt{qi11}) provide the framework...
+[I think this paragraph belongs further down after you have discussed planets since that is the topic of your intro paragraph above.
+Also need a bit more meat here.
+So far a single CO snowline detection and a single H2O snowline detection.
+More are expected and an interpretive framework is key regardless of the importance for planets.]%Recent detections of the CO snowline in disks (e.g., \citealt{qi11}) provide the framework...
 
@@
 
-An important signature of disk chemistry is the carbon-to-oxygen (C/O) ratio, as it provides information about the relative abundances of the main C and O carrier species.
-While the gas C/O ratio in disks may be potentially observable \citep{du15}, such detections do not yet exist.
-However, we can indirectly infer the relative abundance of disk C and O using detections of C/O ratios in exoplanet atmospheres.
+An important signature of disk chemistry is the carbon-to-oxygen (C/O) ratio.
+%, as it provides information about the relative abundances of the main C and O carrier species. [the preceding sentence doesn't really say anything...] 
+Disk C/O ratios may be inferred by observing molecules that trace the C/O ratio (e.g., \citealt{du15}), but such observations do not yet exist.
+%[You cannot detect a C/O ratio.
+You can try to observe molecules that trace the C/O ratio and infer a C/O ratio based on that.]
+Given the tight connection between planet and disk compositions, detections of C/O ratios in exoplanet atmospheres can provide some insight into the relative abundance of C and O in disks.
 Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}.
@@
 \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
+%[This is not a very natural approach I think, i.e.
+to use a dicy eco-planet C/O detection to infer something about the disk chemical composition.]
 
@@
 In addition, the simultaneous inclusion of various dynamical and/or chemical processes makes it difficult to quantify the separate effect of each of these processes on snowline locations and the C/O ratio.
+%[The preceding paragraph needs to be made clearer.
+Currently it is not clear what effects these studies have had on our understanding of C/O ratio sin disks and why your study is needed.
+I.e.
+why are the processes you focus on the most important ones?
+Introduce separately papers that deal with C/O and water snowline locations.]
+
 In this paper, we focus on two particular dynamical effects, i.e.
 the radial drift of solids and viscous gas accretion.
-We enhance the model of \citet{oberg11} by considering these two additional processes.
+%We enhance the model of \citet{oberg11} by considering these two additional processes.
 We perform a systematic study to understand the detailed qualitative and quantitative effects of radial drift and gas accretion on the H$_2$O, CO$_2$ and CO snowline locations, and the resulting C/O ratio in gas and dust throughout the protoplanetary disk.
@@
 We summarize our ice desorption model in section \ref{sec:desorption}.
-Finally, we discuss the relevant timescales for dynamical effects in the desorption process in section \ref{sec:timescales}.
+Finally, we discuss the relevant timescales for dynamical effects that affect snowline locations %in the desorption process
+in section \ref{sec:timescales}.
 
@@
 
-\textbf{Passive disk.} A passive disk is only irradiated by the central star and does not experience radial movement of gas or accretion heating.
-We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}.
-The gas surface density and midplane temperature are
+To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state}, for which the gas surface density $\Sigma_{\rm d}$ and mass flux $\dot{M}$ are constant in time and independent of semimajor axis \footnote{However, we also include stellar irradiation in the disk thermal structure, which changes the steady-state solution as we explain further in this section.}; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
+
+%[Perhaps define your four disk models up here and motivate why you are exploring several different ones, and then go through and describe them in detail.
+That would save you the slightly awkward Active disk with passive temperature profile.]
+
+\textbf{Passive disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
 \begin{subeqnarray}
@@
 
-\textbf{Active disk with passive temperature profile.} The gas is accreting onto the central star in an active disk, which causes the gas surface density to decrease with time.
-We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
+\textbf{Active disk.} We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
 \begin{equation}
@@
 \end{subeqnarray}
+
 where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$ is the Keplerian angular velocity,  with $G$ the gravitational constant and $M_*$ the stellar mass.
 We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition of hydrogen and helium.
-The temperature profile for an active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}).
+The temperature profile for the %an
+active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}).
 From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters.
@@
 The gas surface density is then given by the self-similar solution
+
 \begin{equation}
@@
 
-\textbf{Active disk steady-state.} A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma_{\rm d}$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
-Since calculating the midplane temperature self-consistently for an active disk is non-trivial, we use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm d, act}$.
+\textbf{Active disk at steady-state.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma_{\rm d}$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
+Since
+Calculating the midplane temperature self-consistently for an active disk is non-trivial.
+We thus use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm d, act}$.
 The equations governing the evolution of the steady-state disk are listed in \App{app:steadystate}.
@@
 \textbf{Static disk.} To compare our results with those of \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}).
-The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
+%The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
 
@@
 Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both a passive and an active disk.
-We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state active disk results in Section \ref{sec:COratio}.
+The results for the steady-state active disk are qualitatively similar to the active disk with a passive disk temperature gradient and therefore not shown.
+%We do not show the equivalent result for the steady-state active disk since The trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state active disk results in Section \ref{sec:COratio}.
 Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for a passive nor for an active disk.
-Similarly, micron- to mm-sized particles in the passive disk do not drift or desorb unless they are located inside the static snowlines.
-This is not in contradiction with the desorption timescales for small particles from Figure \ref{fig:timescales}, since small grains only desorb fast at or nearby their respective snowlines, as mentioned in Section \ref{sec:timescales}.
+Similarly, micron- to mm-sized particles in the passive disk do not drift and only %or
+desorb if %unless 
+they are located inside the static snowlines.
+%This is not in contradiction with the desorption timescales for small particles from Figure \ref{fig:timescales}, since small grains only desorb fast at or nearby their respective snowlines, as mentioned in Section \ref{sec:timescales}.
+[don't think this is needed, but Til or Ruth might think otherwise]
 In an active disk, however, micron-to mm-sized grains do drift significantly since they move at the same velocity as the accreting gas.
-For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in a passive disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an active disk, we notice that particles of initial size $s_0$ desorb at a fixed distance $r_{\rm des}$ regardless of their original location in the disk.
+For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in a passive disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an active disk, we notice that particles of initial size $s_0$ desorb at a %fixed distance
+particle size specific radius $r_{\rm des}$ regardless of their original location in the disk.
 In fact, the only grains that will both drift and evaporate are those that reach their fixed final location (represented by the horizontal curves in Figure \ref{fig:snowlines}) within the disk lifetime.
@@
 We can calculate this location analytically by equating Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and solving for $r=r_{\rm des}(s)$ for a given particle size $s$.
-Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance in a passive and an active disk (see Figure \ref{fig:snowlines}).
+Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically.
+We display this result for the range of particle sizes that desorb at a fixed distance in a passive and an active disk (see Figure \ref{fig:snowlines}).
 We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it slightly deviates for particles larger than $s \gtrsim10$ cm.
@@
 %\vspace{-0.5in}
-\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel), an active disk with a passive temperature profile (middle panel) and a steady-state active disk (bottom panel), for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
+\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel), an active disk (middle panel) and a steady-state active disk (bottom panel), for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
 The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
@@
 
-We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star.
+We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
+Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then calculate the C/O ratio throughout the disk.
+
+%[I think a sentence is missing here outlining what you need to establish to calculate the C/O ratio, i.e. the connection between the preceding and the next sentence is not obvious.]
+
 Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
-Solid H$_2$O particles evaporate almost instantly, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
+Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
+%, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
+%A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
+[not sure you need to explain this]
 The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
@@
 It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active).
-The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane.
+The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} on the disk age when only considering drift, accretion and desorption.
 
@@
 We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
-Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk with a passive temperature profile, and a steady-state active disk.
+Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and a steady-state active disk.
 The C/O ratio for a static disk is shown as a guideline.
-The plot is consistent with Figure \ref{fig:snowlines}.
+%The plot is consistent with Figure \ref{fig:snowlines}.
 For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus move the snowline compared to the static disk.
 In contrast, even $\sim$micron-sized grains drift and desorb for the active disk, since they flow towards the host star together with the accreting gas.
-For the same particle size, the snowline locations are slightly closer to the central star in the active disk with a passive temperature profile, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
+For the same particle size, the snowline locations are slightly closer to the central star in the active disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
 Equation \ref{eq:rdotact}).
@@
 This is due to the fact that accretional heating dominates in the inner disk, where high temperatures cause the grains to evaporate further away from the star.
-Once $r\gtrsim1-2$  AU, stellar irradiation dominates the thermal evolution of the disk, and therefore the CO$_2$ and CO snowlines locations are the same as in the active disk with a passive temperature profile.
+Once $r\gtrsim1-2$  AU, stellar irradiation dominates the thermal evolution of the disk, and therefore the CO$_2$ and CO snowlines locations are the same as in the active disk.
 
 Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases.
-While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk.
-Bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals, though only the kilometer-sized boulders that are very close to the snownlines will contribute to determining the snowline location.
+While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that %almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk.
+B
+bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals.
+However, the contribution of kilometer-sized bodies to the snowline location is modest, since they only drift if they are located very close to the snowline.
+%[new sentence explaining that their contribution may not be very big due to their modest drift.] though 
+%only the kilometer-sized boulders that are very close to the snownlines will contribute to determining the snowline location. 
 Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
 Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
-Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution.
+Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from this plot.
+%Our model thus sets an inner limit to the snowline locations.
+%Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution --- for a given grain size .
 
-For our choice of parameters, the snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk with a passive temperature profile and $r_{\rm{H_2O}} \approx 0.63$ AU for the steady-state disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
+% [the last sentence needs to be made more explicit, i.e.
+for a given grain size distribution with a maximum grain size a reader can now pick out the appropriate minimum snowline location from this plot]
+
+For our choice of parameters, the minimum snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk and $r_{\rm{H_2O}} \approx 0.63$ AU for the steady-state disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
 For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
@@
 
-In this section we investigate how variations in our fiducial parameters, such as the total disk mass or the time in the disk evolution at which we determine the snowline locations and the C/O ratio, affect our results.
-The disk lifetime $t_{\rm d}=3$ Myr is the most representative for our calculations, as this is the timeframe in which giant planets accrete their gaseous atmospheres (e.g., \citealt{pollack96}, \citealt{piso14}).
-However, protoplanetary cores start acquiring their envelope at earlier times, before the core is fully formed (e.g., \citealt{rafikov06}).
-Recent models such as aerodynamic pebble accretion \citep{lambrechts12} suggest rapid core growth on timescales of $10^5$ years, which implies that the time when cores accrete their massive envelope may be at least an order of magnitude shorter than the disk lifetime.
+In this section we investigate how variations in our fiducial parameters, the total disk mass, disk age, and disk structure, %the time in the disk evolution at which we determine 
+affect the calculated snowline locations and the C/O ratio. %, affect our results. 
+All previous results assumed a disk lifetime $t_{\rm d}=3$ Myr%is the most representative for our calculations, as this is the timeframe in which giant planets accrete their gaseous atmospheres 
+, the typical disk life time and the expected time scale for giant planets to accrete their gaseous atmopsheres (e.g., \citealt{pollack96}, \citealt{piso14}).
+Some gas accretion may occur at earlier times, however,%However, protoplanetary cores start acquiring their envelope at earlier times,
+before the core is fully formed (e.g., \citealt{rafikov06}).
+Recent models such as aerodynamic pebble accretion \citep{lambrechts12} suggest that rapid core growth on timescales of $10^5$ years is possible.
+% which implies that the time when cores accrete their massive envelope may be at least an order of magnitude shorter than the disk lifetime.
 The composition of giant planet atmospheres, and specifically their C/O ratio, can thus depend on the abundance of H$_2$O, CO$_2$ and CO in gas and dust forms at earlier times than $t_{\rm d}$ in the disk evolution.
 
-Figure \ref{fig:timeplots} shows the desorption or final distance as a function of a particle's initial location in the disk, for grains of initial sizes of 10 cm and 1 m, composed of either H$_2$O, CO$_2$ or CO.
+Figure \ref{fig:timeplots} shows the particle desorption or final distance as a function of a particle's initial location in the disk, for ice particles %grains 
+of initial sizes of 10 cm and 1 m, composed of either H$_2$O, CO$_2$ or CO.
 These sizes are representative since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
-We stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively, to see how the snowline locations evolve with time and affect the C/O ratio.
-Particles of a given size that start at large stellocentric distances do not desorb within the shorter timeframes, e.g.
-$10^4$ or $10^5$ years, but they do evaporate at a fixed radius if their initial location is closer to the host star.
-While the amount of material that moves through the disk changes with time, the radius at which particles desorb and thus the snowline locations are independent of the time elapsed.
-Therefore, our results are not only valid at $t_{\rm d}$, but also throughout the earlier time evolution of the protoplanetary disk.
+We stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively.
+%, to see how the snowline locations evolve with time and affect the C/O ratio.
+The most important result of these plots is that particles of a given size always desorb at the same disk radii, the 3 Myr snowline, regardless of simulation stopping time.
+Particles %of a given size
+that start at large stellocentric distances do not desorb within the shorter timeframes, e.g.
+$10^4$ or $10^5$ years.
+%, but they do evaporate at a fixed radius if their initial location is closer to the host star.
+While the amount of material that moves through the disk changes with time, the radius at which particles desorb and
+The snowline locations are thus independent of the time elapsed, and %. Therefore, 
+our results are %not only valid at $t_{\rm d}$, but also throughout the earlier 
+valid throughout the time evolution of the protoplanetary disk.
 
-We choose as a fiducial model a total disk mass $M_{\rm d}=0.1 M_{\odot}$, but this number is not universally valid.
-From observations of the dust continuum, \citet{andrews13} find that a linear scaling $M_{\rm d} \propto M_*$ is reasonable.
+We choose as a fiducial model a total disk mass $M_{\rm d}=0.1 M_{\odot}$.
+Observationally, disk masses span at least an order of magnitude around Solar type stars %, but this number is not universally valid.
+From observations of the dust continuum,
+\citep{andrews13}.
+% find that a linear scaling $M_{\rm d} \propto M_*$ is reasonable.
 Giant planets, however, have been detected around small stars (e.g., \citealt{montet14}), which can have masses as low as $M_* \sim 0.1 M_{\odot}$.
@@
 Since the snowline locations are independent of disk mass, we expect transition disks to exhibit the same drift-desorption behavior as standard disks under the simplified assumptions of our model.
+%[Comment on differences, i.e. short time frames there is a difference because of smaller drift velocity in the lighter disk I presume.]
 
@@
 Particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the gap edge, due to the decrease in gas pressure inside the cavity, thus forming a snowline at $\sim$$3.8$ AU.
-This is consistent with the observations of \citet{zhang13}, which show that the H$_2$O snowline is pushed outwards in a transition disk compared to a full disk.
+This is qualitatively consistent with the observations of \citet{zhang13}, which show that the H$_2$O snowline is pushed outwards in a transition disk compared to a full disk.
+Our model framework is thus generally valid for more complicated disk structures as well.
+
+%[Need a sentence about how this makes the model framework also generally valid for more complicated disk structures.]
 
@@
 The ice thus only constitutes a fraction of the total particle mass, which accelerates its desorption and pushes the snowlines outwards.
-The grains may also be composed of a mixture of H$_2$O, CO$_2$ and CO ices.
-In this case, the desorption energies are largest for the most volatile species, which will move the snowlines inwards.
+The grains may also be composed of a mixture of H$_2$O, CO$_2$ and CO ices, which will increase the binding energies of the more volatiles species, moving %.
+In this case, the desorption energies are largest for the most volatile species, which will move
+the snowlines inwards.
 
@@
 
-\item \textbf{Non-static chemistry.} As the goal of this paper was to explore only the dynamical effects on snowline locations and the C/O ratio in disks, we have assumed a simple, static chemical model.
-However, gas-grain chemistry is very complex and time-dependent.
+\item \textbf{Time dependent chemistry.} As the goal of this paper was to explore only the dynamical effects on snowline locations and the C/O ratio in disks, we have assumed a simple, static chemical model.
+In reality, the chemistry in most of the disk is expected to be time-dependent.%However, gas-grain chemistry is very complex and time-dependent.
 In the inner disk, chemistry approaches equilibrium due to intense sources of ionizing radiation (e.g., \citealt{ilgner04}), while in the outer disk high energy radiation and cosmic rays are the key drivers of chemistry, which is no longer in equilibrium (e.g., \citealt{vandishoeck06}).
@@
 The effect of disk chemistry on snowline locations, shape, time evolution, or the C/O ratio is therefore difficult to estimate.
-In a future paper, we plan to use a parametrized chemical model and incorporate it in our radial drift calculation.
+%In a future paper, we plan to use a parametrized chemical model and incorporate it in our radial drift calculation.
+[let's not promise this until we have figured out how to do it.]
 
@@
 We assume that the solid particles are perfect, homogeneous spheres, fully composed of either H$_2$O, CO$_2$ or CO.
-We apply our model to a passive disk, an active disk with a passive temperature profile, and a steady-state active disk that also takes into account stellar irradiation.
+We apply our model to a passive disk, an active disk with a passive temperature profile (hereafter active disk), and a steady-state active disk that also takes into account stellar irradiation.
 We determine the desorption or final location of drifting particles after a time equal to the disk lifetime, and use this result to set an inner limit for the location of the H$_2$O, CO$_2$ and CO snowlines.
@@
 \item Radial drift and gas accretion affect desorption and move the snowline locations compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk.
-Moreover, boulders up to almost kilometer-size contribute to setting the snowline locations if they are very close to the snowlines.
+%Moreover, boulders up to almost kilometer-size contribute to setting the snowline locations if they are very close to the snowlines.
+[I find this conclusion a bit iffy and also it sounds like it is in contradiction with the previous sentence.]
 
-\item Particles with sizes in the above range desorb instantaneously, and at a fixed location in the disk that only depends on the particle size.
+\item Particles with sizes in the above range desorb almost instantaneously once desorption has begun, and at a fixed location in the disk that only depends on the particle size.
 Thus for each particle size there is a fixed and uniquely determined H$_2$O, CO$_2$ or CO snowline.
@@
 Gas accretion causes even micron-sized particles to drift, desorb and move the snowline location compared to a static disk.
-A steady-state disk that includes accretion heating moves the H$_2$O snowline outwards compared to an active disk with a passive temperature profile, but has no effect on the CO$_2$ and CO snowline locations.
+A steady-state disk that includes accretion heating moves the H$_2$O snowline outwards compared to an active disk, but has no effect on the CO$_2$ and CO snowline locations.
 
@@
 
-\item For our fiducial model, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for a passive disk, 0.26 AU, 3.4 AU and 30 AU for an active disk with a passive temperature profile, and 0.63 AU, 3.4 AU and 30 AU for a steady-state active disk.
+\item For our fiducial model, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for a passive disk, 0.26 AU, 3.4 AU and 30 AU for an active disk, and 0.63 AU, 3.4 AU and 30 AU for a steady-state active disk.
 Compared to a static disk, radial drift and gas accretion move the snowlines by up to 60 \% for H$_2$O and CO$_2$, and by up to 50 \% for CO.

### 2015-07-23 tex/draft/ms_kedit.tex
update

@@
 
-\title{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Accretion}
+\title{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}
 
@@
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-%\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
+\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
 
@@
 
-The chemical composition of the atmospheres of giant planets can provide important constraints on %their [unclear association]
-the planet formation, accretion and migration history. 
+The chemical composition of protoplanerary disks affects planet formation efficiencies and the composition of nascent planets.
 %[Need explicit statement that gas giants accrete their atmospheres from the nebular gas.] 
@@
 
-The structure of protoplanetary disks in complex, and affected by a multitude of chemical and dynamical processes (see review by \citealt{henning13}).
+The structures of protoplanetary disks are complex, and affected by a multitude of chemical and dynamical processes (see review by \citealt{henning13}).
 From the chemistry perspective, volatile compounds are particularly important.
@@
 Based on protostellar and comet abundances, some of the most important volatile molecules are H$_2$O, CO$_2$, CO, N$_2$.
-Recent observations of protoplanetary disks have provided valuable information about the abundance and snowline locations of some of these compounds.
+Recent observations of protoplanetary disks have provided valuable information about the abundances and snowline locations of some of these compounds.
 For example, the CO snowline has been detected in the disks around HD 163296 \citep{qi11} and TW Hya \citep{qi13}.
-Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}.
-With the advent of ALMA, more such detections are expected.
-An interpretive framework of these observations is thus crucial regardless of their importance for planet compositions.
+Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}, and more such snowline detections are expected in future ALMA cycles.
+These observations are currently lacking an interpretive framework that takes into account all important dynamical and chemical processes.
+Furthermore, such a framework is crucial to connect observed snowline locations to planet formation.
 
@@
 
-An important signature of disk chemistry is the carbon-to-oxygen (C/O) ratio.
-%, as it provides information about the relative abundances of the main C and O carrier species. [the preceding sentence doesn't really say anything...] 
-Disk C/O ratios may be inferred by observing molecules that trace the C/O ratio (e.g., \citealt{du15}), but such observations do not yet exist.
-%[You cannot detect a C/O ratio.
-You can try to observe molecules that trace the C/O ratio and infer a C/O ratio based on that.]
-Given the tight connection between planet and disk compositions, detections of C/O ratios in exoplanet atmospheres can provide some insight into the relative abundance of C and O in disks.
+An important consequence of snowline formations in disks is that disks are expected to present different carbon-to-oxygen (C/O) ratios in the gas and in icy dust mantles at different disk radii.
+This effect was quantified by  \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e.
+H$_2$O, CO$_2$ and CO, have different condensation temperatures.
+This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
+\citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
+This effect have also been observed in some exoplanet atmospheres, where superstellar C/O ratios have been detected (e.g., WASP-12b, \citealt{madhu11}).
+%Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.
+
+%Given the tight connection between planet and disk compositions, detections of C/O ratios in exoplanet atmospheres can provide some insight into the relative abundance of C and O in disks.
 Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}.
@@
 
-Most studies of the C/O ratio in disks are particularly focused on the inner disk, in order to explain observations of super-stellar C/O ratios in the atmospheres of close-in exoplanets \citep{madhu11}.
+%Most studies of the C/O ratio in disks are particularly focused on the inner disk, in order to explain observations of super-stellar C/O ratios in the atmospheres of close-in exoplanets \citep{madhu11}.
 In addition, the simultaneous inclusion of various dynamical and/or chemical processes makes it difficult to quantify the separate effect of each of these processes on snowline locations and the C/O ratio.
@@
 Introduce separately papers that deal with C/O and water snowline locations.]
+Each of these studies have considered a specific combination of dynamical and chemical effects.
+One combination that has not yet been considered is the combination of radial drift and viscous gas accretion.
+Studying these two dynamical processes makes it possible to quantify their separate effect on snowline locations and the C/O ratio at various disk radii.
 
-In this paper, we focus on two particular dynamical effects, i.e.
+ %In this paper, we focus on these two particular dynamical processes.
+%i.e.
 the radial drift of solids and viscous gas accretion.
 %We enhance the model of \citet{oberg11} by considering these two additional processes.
-We perform a systematic study to understand the detailed qualitative and quantitative effects of radial drift and gas accretion on the H$_2$O, CO$_2$ and CO snowline locations, and the resulting C/O ratio in gas and dust throughout the protoplanetary disk.
+In this paper, we perform a systematic study to understand the detailed qualitative and quantitative effects of radial drift and gas accretion on the H$_2$O, CO$_2$ and CO snowline locations, and the resulting C/O ratio in gas and dust throughout the protoplanetary disk.
 More importantly, we obtain a limit on how close to the star the snowline locations can be pushed by radial drift and gas accretion.

### 2015-07-27 tex/draft/ms_kedit_tedit.tex
update

--- 2015-07-23 tex/draft/ms_kedit.tex
@@
 \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-This effect have also been observed in some exoplanet atmospheres, where superstellar C/O ratios have been detected (e.g., WASP-12b, \citealt{madhu11}).
+This effect has also been observed in some exoplanet atmospheres, where superstellar C/O ratios have been detected (e.g., WASP-12b, \citealt{madhu11}).
 %Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.
@@
 
-\citet{oberg11} assume a static disk with passive chemistry.
+\citet{oberg11} assumed a static disk with passive chemistry.
 In reality, additional dynamical and chemical processes affect the snowline locations and the resulting C/O ratio.
@@
 
-To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state}, for which the gas surface density $\Sigma_{\rm d}$ and mass flux $\dot{M}$ are constant in time and independent of semimajor axis \footnote{However, we also include stellar irradiation in the disk thermal structure, which changes the steady-state solution as we explain further in this section.}; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
+To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state}, for which the gas surface density $\Sigma_{\rm d}$ and mass flux $\dot{M}$ are constant in time and independent of semimajor axis\footnote{However, we also include stellar irradiation in the disk thermal structure, which changes the steady-state solution as we explain further in this section.}; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
 
@@
 
-Solid particles in orbit their host star at the Keplerian velocity $v_{\rm k} \equiv \Omega_{\rm k} r$.
+Solid particles orbit their host star at the Keplerian velocity $v_{\rm k} \equiv \Omega_{\rm k} r$.
 The gas, however, experiences an additional pressure gradient, which causes it to rotate at sub-Keplerian velocity \citep{weidenschilling77}.
@@
 
-For a passive disk, the radial drift velocity can be approximated analytically as
+For a passive disk, the radial drift velocity can be approximated as
 \begin{equation}
@@
 \label{eq:tdrift}
-t_{\rm r, drift} \sim \Big|\frac{r}{\dot{r}}\Big|,
+t_{\rm drift} \sim \Big|\frac{r}{\dot{r}}\Big|,
 \end{equation}
@@
 %\vspace{-0.5in}
-\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
+\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
 The timescales are calculated at three representative locations, i.e.
@@
 Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
-$t_{\rm r, drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.}
+$t_{\rm drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.}
 \label{fig:timescales}
@@
 
-For simplicity purposes, we calculate the radial drift timescale, $t_{\rm r, drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well.
-Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm r, drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk.
+For simplicity, we calculate the radial drift timescale, $t_{\rm drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well.
+Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk.
 As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence on disk location (see Equation \ref{eq:Rdes}).
@@
 
-Of particular interest for our purposes is the particle size regime for which (1) $t_{\rm r, drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
+Of particular interest for our purposes is the particle size regime for which (1) $t_{\rm drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
 for $\sim$$0.5$ cm $\lesssim s \lesssim$ $1000$ cm, or (2) $t_{\rm gas, acc} \lesssim t_{\rm des} \lesssim t_{\rm d}$, i.e.
@@
 \centering
-\includegraphics[width=0.7\textwidth]{desorption_distance_passive_active_colorbar_test.pdf}
+\includegraphics[width=0.7\textwidth]{desorption_distance_passive_active_colorbar_test2.pdf}
 %\vspace{-0.5in}
@@
 Equations (\ref{eq:dsdt}) and (\ref{eq:drdt}) describe the coupled desorption and radial drift, and can be derived straightforwardly from Equation (\ref{eq:tdes}).
-Our boundary conditions are $s(t_0)=s_0$, $r(t_0)=r_0$, and $s(t_{\rm d})=0$, where $t_0$ is the initial time at which we start the integration and $r_0$ is the initial location of the particle.
+Our initial conditions are $s(t_0)=s_0$ and $r(t_0)=r_0$, where $t_0$ is the initial time at which we start the integration and $r_0$ is the initial location of the particle.
 We choose $t_0=1$ year, but our result is independent on the initial integration time as long as $t_0 \ll t_{\rm d}$.
+The desorption timescale $t_{\rm des}$ will then satisfy $s(t_{\rm des})=0$, from which we can determine the desorption distance $r_{\rm des}=r(t_{\rm des})$.
 
@@
 For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in a passive disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an active disk, we notice that particles of initial size $s_0$ desorb at a %fixed distance
-particle size specific radius $r_{\rm des}$ regardless of their original location in the disk.
+particle size dependent radius $r_{\rm des}$ regardless of their original location in the disk.
 In fact, the only grains that will both drift and evaporate are those that reach their fixed final location (represented by the horizontal curves in Figure \ref{fig:snowlines}) within the disk lifetime.
@@
 Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
-Particles lose most of their mass very close to the distance at which they fully desorb.}
+Particles lose most of their mass very close to the distance at which they fully desorb.
+The static H$_2$O snowline is shown for reference (dashed vertical lines).}
 \label{fig:s_t_a}
@@
 
-Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm r, drift} \sim t_{\rm des}$, given an initial particle size.
+Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm drift} \sim t_{\rm des}$, given an initial particle size.
 We can calculate this location analytically by equating Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and solving for $r=r_{\rm des}(s)$ for a given particle size $s$.
@@
 We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it slightly deviates for particles larger than $s \gtrsim10$ cm.
-For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm r, drift}$ is a power-law in $r$ (for our parameters, $t_{\rm r, drift} \propto r^{-1/14}$ for the passive disk), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
-Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm r, drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
+For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the passive disk), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
+Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
 
@@
 
-Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
+Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk\footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
 Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
@@
 It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active).
-The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} on the disk age when only considering drift, accretion and desorption.
+The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
 
@@
 \caption{Desorption / final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve), and 3 Myr, the disk lifetime (blue curve).
+The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve).
 For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
@@
 \caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M_{\rm d}=0.1 M_{\odot}$ (solid lines) and $M_{\rm d}=0.01 M_{\odot}$ (dashed lines).
-The timescales of the simulations and their color code are the same as in Figure \ref{fig:timeplots}.
+The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve).
 A lower disk mass does not change the snowline location.}
@@
 \caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines).
-The timescales of the simulations and their color code are the same as in Figure \ref{fig:timeplots}.
+The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}.
 Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge.}
@@
 % which implies that the time when cores accrete their massive envelope may be at least an order of magnitude shorter than the disk lifetime.
-The composition of giant planet atmospheres, and specifically their C/O ratio, can thus depend on the abundance of H$_2$O, CO$_2$ and CO in gas and dust forms at earlier times than $t_{\rm d}$ in the disk evolution.
+The composition of giant planet atmospheres, and specifically their C/O ratio, can thus depend on the abundance of H$_2$O, CO$_2$ and CO in gas at earlier times than $t_{\rm d}$ in the disk evolution.
 
@@
 
-We summarize in Table 1 the potential physical and chemical processes occurring in disks and their effect on snowline locations.
-For the sake of completion, Table 1 also includes the processes addressed in this paper, i.e.
+We summarize in Table 1 the potential physical and chemical processes occurring in disks and their effect on snowline locations compared to a static disk.
+For the sake of completeness, Table 1 also includes the processes addressed in this paper, i.e.
 radial drift and gas accretion.
@@
 Particle growth will thus initially push the snowlines inwards.
-However, once the solids grow larger than km-sized, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
+This is consistent with particle growth models, which predict a maximum particle size often around or below the particle sizes that drift the fastest \citep{birnstiel12}.
+As the largest grains contain most of the solid mass, grain growth models should produce snowlines corresponding to our snowline location estimates for the largest grains in the particle size distribution.
+However, once the solids grow larger than km-sized and form planetesimals, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
 It follows that grain growth will eventually push the snowline location outwards.
@@
 \textbf{Process} & \textbf{Effect} \\\hline
-Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star, $\rightarrow$ means that the snowline is pushed further from the host star.
+Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star compared to the static snowline, $\rightarrow$ means that the snowline is pushed further from the host star compared to the static snowline.
 The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
@@
 meter-sized due to collisions and subsequent fragmentation, which will cause them to drift significantly before desorbing, pushing the snowlines inwards.
-Thus fragmentation can move the snowline locations in either radial direction.
+Thus fragmentation can move the snowline locations in either radial direction --- specifically, fragmentation leads to a certain grain size distribution, and the largest particles in this size distribution are the ones that determine the position of the snowline.
 
 \item \textbf{Grain morphology.} Our model assumes that the ice particles are perfect, homogeneous spheres.
-However, this is not a very good approximation, since grain growth can be fractal rather than compact (\citealt{zsom10}, \citealt{okuzumi12}).
+However, this is not a very good approximation, since grain growth can be fractal rather than compact (\citealt{zsom10}, \citealt{okuzumi12}, \citealt{krijt15}).
 The inhomogeneity due to cracks in the grain structure will cause the particles to desorb faster.
@@
 
-\item \textbf{Particle composition.} The ice particles in our model are assumed to be fully formed of either H$_2$O, CO$_2$ or CO.
-In reality, the grains have a layered structure, such as an interior composed of non-volatile materials (e.g., sillicates) covered by an icy layer.
+\item \textbf{Particle composition.} The ice particles in our model are assumed to be fully made of either H$_2$O, CO$_2$ or CO.
+More realistically, grains may have a layered structure, such as an interior composed of non-volatile materials (e.g., sillicates) covered by an icy layer.
 The ice thus only constitutes a fraction of the total particle mass, which accelerates its desorption and pushes the snowlines outwards.
@@
 
-\item The snowline locations move inwards as the particle size increases; the innermost snowline is set by particles with initial size $s \sim 7$ m.
+\item The snowline locations move inwards as the particle size increases; the innermost snowline is set by particles with initial size $s \sim 7$ m in our model.
 Gas accretion causes even micron-sized particles to drift, desorb and move the snowline location compared to a static disk.
@@
 \end{equation}
-By using Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and setting $t_{\rm r, drift}=t_{\rm des}$, we can express a particle's desorption distance as
+By using Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and setting $t_{\rm drift}=t_{\rm des}$, we can express a particle's desorption distance as
 
@@
 \label{eq:tan}
-r_{\rm des}=\Bigg(\frac{d\,\, \mathcal{W}\Big[\frac{(B/A)^{-q/d} q\,C}{d} q C\Big]}{q \,C}\Bigg)^{\frac{1}{q}},
+r_{\rm des}=\Bigg(\frac{d}{q \,C} \,\, \mathcal{W}\Big[\frac{(B/A)^{-q/d} q\,C}{d}\Big]\Bigg)^{\frac{1}{q}},
 \end{equation}

### 2015-07-27 _Jamila/classDisk.py

@@ -1,0 +1,158 @@
+#FILE: classDisk
+#FUNCTION: This file contains the class Disk, which is meant to simulate a protoplanetary disk with particular density and temperature profiles and given molecular species.
+
+
+######
+###BACKGROUND
+#Below Section: Imports necessary functions
+import numpy as np
+import matplotlib.pyplot as graph
+import astropy.constants as const
+
+import diskutils as utils
+from classMol import Mol
+
+
+
+class Disk(object):
+#FUNCTION: __init__
+#PURPOSE: This function is meant to initialize a disk with given temperature and density profiles.
+	def __init__(self, mols=None):
+		#Below Section: Initializes molecular species in disk
+		if mols == None: #If no mols given
+			self.mols = []
+		elif mols != None: #If list of mols given
+			self.mols = mols
+			
+			
+
+#FUNCTION: addMol
+#PURPOSE: This function is meant to add a mol to the current disk.
+	def addMol(self, thismol):
+		#Below Section: Adds given mol to disk's mols
+		if thismol not in self.mols: #If not in disk yet
+			self.mols.append(thismol)
+		elif thismol in self.mols: #If mol already in disk
+			print("This molecular species is already in the disk.  Nothing was changed.")
+			
+			
+
+#FUNCTION: delMol
+#PURPOSE: This function is meant to delete a mol from the current disk.
+	def delMol(self, thismol):
+		#Below Section: Deletes given mol from disk's mols
+		if thismol in self.mols: #If in disk
+			self.mols.remove(thismol)
+		elif thismol not in self.mols: #If mol not in disk
+			print("This molecular species was not in the disk to start with.  Nothing was changed.")
+		
+
+	
+#FUNCTION: printMol
+#PURPOSE: This function is meant to return a string listing the mols currently contained in the disk.
+	def stringMol(self):
+		#Below Section: Builds a string of molecular species within the disk
+		names = '['
+		for molhere in self.mols: #Iterates through mols
+			names = names + molhere.name + ', '
+		names = names[0:-2] + ']'	
+		
+		#Below Section: Returns finished string
+		return names
+		
+		
+		
+#FUNCTION: profT
+#PURPOSE: This function is meant to calculate a temperature profile for the given disk using the given type.
+	def profT(self, R, z, type='basic'):
[…]

### 2015-07-27 _Jamila/classMol.py

@@ -1,0 +1,96 @@
+#FILE: classMol.py
+#PURPOSE: This file contains the class Mol, which models a particular molecular species.
+
+class Mol(object):
+#FUNCTION: __init__
+#PURPOSE: This function initializes the current Mol with given parameters.
+#INPUTS: energy = molecular bond breaking energy (K); mass = molecule mass (kg); xtot = xmolgr + xmolgas = total molecular fraction for molecule (over both grain and gas) of base nH; name = molecule chemical formula
+	def __init__(self, xtot, mass, energy, name):
+		#Automatic attributes
+		self.xtot = xtot
+		self.mass = mass
+		self.energy = energy
+		self.name = name
+		self.elements = self.getElements(name)
+
+		
+#FUNCTION:
+#PURPOSE: This function is meant to take the chemical formula of a molecule and separate it into element components with counted frequencies.
+#EXAMPLE: 'H2O' would return the dictionary {'H':2, 'O':1}
+#INPUTS: name = molecular species's chemical formula as string	
+	def getElements(self, name):
+		#Below Section: Records the elements in molecule
+		elements = {}
+		elhere = name[0]
+		dighere = ''
+		for a in range(0, len(name)):
+			#Below handles if current letter in alphabet
+			if name[a].isalpha():
+				#If continued name of current element
+				if name[a].islower():
+					elhere = elhere + name[a]
+										
+				#Else if new element started
+				elif name[a].isupper() and a > 0:
+					#If a repeated element within this molecule
+					if elhere in elements:
+						if dighere == '':
+							elements[elhere] = elements[elhere] + 1
+						elif dighere != '':
+							elements[elhere] = elements[elhere] + int(dighere)
+					
+					#If first time element added for molecule
+					elif dighere == '':
+						elements[elhere] = 1
+					elif dighere != '':
+						elements[elhere] = int(dighere)
+					
+					#Below resets elements
+					elhere = name[a]
+					dighere = ''
+					
+			#Else if number encountered
+			elif name[a].isdigit():
+				dighere = dighere + name[a]
+					
+			#Else if error encountered
+			else:
+				raise ValueError("Weird molecular name encountered at " + str(a) + "th index: " + name)
+
+			#Below if end of name string reached
+			if (a+1) >= len(name):
+				#Records last element
+				#If a repeated element within this molecule
+				if elhere in elements:
+					if dighere == '':
[…]

### 2015-07-27 _Jamila/diskutils.py

@@ -1,0 +1,1190 @@
+#FILE: diskutils.py
+#PURPOSE: This function is meant to provide certain base calculations and utilities for calculating various disk properties.
+#USES: classMol, classDisk
+
+
+
+
+######
+###BACKGROUND
+#Below Section: Imports necessary functions
+import numpy as np
+import matplotlib.pyplot as graph
+import matplotlib as mpl
+import astropy.constants as const
+
+from classDisk import Disk
+from classMol import Mol
+
+#Below Section: Records necessary constants
+pi = np.pi #Pi
+kB = const.k_B.value #Boltzmann Constant
+Gconst = const.G.value #Gravitational Constant
+kBerg = const.k_B.cgs.value #Boltzmann Constant in cgs
+au = const.au.value #Astronomical Unit (m)
+mH = 1.661e-27 #Mass of H atom (kg)
+msun = const.M_sun.value #Mass of Sun (kg)
+
+
+
+
+
+
+######
+###PLOTS
+#FUNCTION: plotgradxel
+#PURPOSE: This function is meant to plot the x_profile of an element within a given disk.  If two elements are given, will plot the ratio betwixt the two.
+def plotgradxel(R, z=None, el1=None,
+				el2=None, disk=None, cmap=graph.cm.rainbow,
+				profTtype='basic', profnHtype='gaus-basic',
+				xgr=1.0e-12, adsign=-1.0, design=1.0,
+				rgr=((0.1e-6)*100), s=1, ratetot=0,
+				snowlinemols=['H2O','CO2','CO','NH3','N2'],
+				showsnowlines=False, ratioplot=False,
+				root='PlotGrad-'):
+			
+	#Below Section: Makes sure disks and mols as correct inputs
+	if not isinstance(disk, Disk) or not isinstance(el1, str):
+		raise ValueError("Oh no!  Please make sure you input a Disk for the keyword 'disk' and the name of the element for the keyword 'el1'.")
+
+	#Below sets up location matrix
+	#Below sets up radius and height matrices
+	Rmatr = np.resize(R, (len(z), len(R)))
+	zmatr = np.resize(z, (len(R), len(z))).T
+	
+	#Below Section: Calculates for first element
[…]

### 2015-07-28 python/Nitrogen/N_to_O.py
update

@@ -1,0 +1,151 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+
+def n_N_O(species, NH3mid = 0, NH3max = 0):
+    
+    """
+        
+    Returns the abundance of N and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_N:
+        abundance of N with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 0, 1.5
+    elif species == 'CO2':
+        return 0, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0, 0
+    elif species == 'silicate':
+        return 0, 1.4
+    elif species == 'N2' and NH3mid == 0 and NH3max == 0:
+        return 1.12 * 2, 0
+    elif species == 'N2' and NH3mid == 1 and NH3max == 0:
+        return (1.12 - 0.9 * 0.055) * 2, 0
+    elif species == 'N2' and NH3max == 1 and NH3mid == 0:
+        return (1.12 - 0.9 * 0.1537) * 2, 0
+    elif species == 'NH3' and NH3mid == 0 and NH3max == 0:
+        return 0, 0
+    elif species == 'NH3' and NH3mid == 1 and NH3max == 0:
+        return 0.055 * 0.9, 0
+    elif species == 'NH3' and NH3max == 1 and NH3mid == 0:
+        return 0.1537 * 0.9, 0
+
+
+#in the following we calculate T_freeze for various species
+T_freeze_H20 = T_freeze(18., 5800, 1e6)
+T_freeze_NH3 = T_freeze(17., 2965., 5.5e4)
+T_freeze_CO2 = T_freeze(44., 2000, 3e5)
+T_freeze_CO = T_freeze(28., 1389., 1e6)
+T_freeze_N2 = T_freeze(28., 1053., 1.12e6)
+#T_freeze_NH3 = T_freeze(17., 2965., 1.5371e5)
+
+
+
+def n(T, elem, NH3mid = 0, NH3max = 0):
+    
+    """
+    
[…]

### 2015-07-28 python/Nitrogen/T_freeze.py

@@ -1,0 +1,95 @@
+"""This script is used to calculate the vibrational frequency, desorption rate, 
+desorption time, and freezing temperature of a given volatile"""
+
+
+import numpy as np
+from utils.constants import mp, kb
+from scipy.optimize import brentq
+
+def vib_freq(Ex, mx):
+    
+    """
+    Input
+    -----
+    Ex: 
+        binding energy in K
+    mx:
+        molecular weight in proton masses
+        
+    Output
+    ------
+    nu:
+        vibrational frequency in s^-1
+    """
+    
+    return 1.6 * 1e11 * np.sqrt(Ex / mx)
+
+def Rdes(mx, Ex, Tx):
+    
+    return vib_freq(Ex, mx) * np.exp(- Ex / Tx) 
+    
+def tevap(mx, Ex, Tx):
+    
+    return 1 / Rdes(mx, Ex, Tx)
+    
+def v_thermal(T, mu):
+    
+    """
+    Input
+    -----
+    T:
+        temperature in K
+    mu:
+        molecular weight in proton masses
+        
+    Output
+    ------
+    v_thermal:
+        thermal velocity in cm s^-1
+    """
+    
+    return np.sqrt(kb * T / (mu * mp))        
+                
+    
+def T_freeze(mx, Ex, nx, Nx = 1e15, fx = 1):
+    
+    """
+    Input
+    -----
+    mx:
+        molecular weight in proton masses
+    Ex:
+        binding energy in K
+    nx:
+        number density in cm^-3
+    Nx:
+        number of adsorption sites per cm^2 (default 1e15)
+    fx:
+        fraction of occupied adsorption sites (default 1)
+        
+    Output
+    ------
+    Tfreeze:
+        freezing temperature in K
+    """
+    
+    def Tfreezefn(x):
+        
+        """
+        Auxiliary function, defined as Tfreezefn(x) = f(x) - x,
+        where "x" stands for temperature and f(x) is the expression
+        on the RHS of equation (4) in Hollenbach et al. (2009).
+        Needed to find the freezing temperature Tf from 
+        Tfreezefn(Tf) = 0 (i.e., f(Tf) = Tf), because the expression 
+        for T_freeze in equation (4) also depends on Tf through the
+        thermal velocity. 
+        """
+        
+        return Ex * (np.log(4 * Nx * fx * vib_freq(Ex, mx) / \
+            (nx * v_thermal(x, mx))))**(-1) - x
+            
[…]

### 2015-07-28 python/C_to_O_CH4.py
update

@@ -1,0 +1,141 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+
+def n_C_O(species, CH4mid = 0, CH4max = 0):
+    
+    """
+        
+    Returns the abundance of C and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_C:
+        abundance of C with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 1.5, 1.5
+    elif species == 'CO2':
+        return 0.3, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0.6, 0
+    elif species == 'silicate':
+        return 0, 1.4
+    elif species == 'CH4' and CH4mid == 0 and CH4max == 0:
+        return 0, 0
+    elif species == 'CH4' and CH4mid == 1 and CH4max == 0:
+        return 0.0555 * 0.9, 0 #5.55% H2O (Oberg+08 Spitzer paper) 
+    elif species == 'CH4' and CH4mid == 0 and CH4max == 1:
+        return 0.13, 0 #13% H2O (Oberg+08 c2d paper)
+     
+
+
+#in the following we calculate T_freeze for various species
+T_freeze_H20 = T_freeze(18., 5800, 1e6)
+T_freeze_CO2 = T_freeze(44., 2000, 3e5)
+T_freeze_CH4 = T_freeze(16., 1300., 0.0555*1e6) #binding energy from Garrod+06 A&A paper
+T_freeze_CO = T_freeze(28., 850, 1e6)
+#T_freeze_CH4 = T_freeze(16., 1300., 0.13*1e6)
+
+
+def n(T, elem, CH4mid, CH4max):
+    
+    """
+    
+    Returns the gaseous and solid abundance of a given element (C, O)
+    as a function of temperature, e.g. between different snowlines
+    
+    Input
+    -----
+    T:
+        disk midplane temperature in units of K
+    elem:
+        string: 'C' or 'O'
+        
+    Output
+    ------
[…]

### 2015-07-28 python/Nitrogen/N_to_O.py

@@ -46,9 +46,9 @@ def n_N_O(species, NH3mid = 0, NH3max = 0):
     elif species == 'NH3' and NH3mid == 0 and NH3max == 0:
         return 0, 0
     elif species == 'NH3' and NH3mid == 1 and NH3max == 0:
-        return 0.055 * 0.9, 0
+        return 0.055 * 0.9, 0 #0.055 * n_H2O from Oberg+11 Spitzer paper
     elif species == 'NH3' and NH3max == 1 and NH3mid == 0:
-        return 0.1537 * 0.9, 0
+        return 0.1537 * 0.9, 0 #0.1537 * n_H2O from Bottinelli+10 Spitzer paper
 
 
 #in the following we calculate T_freeze for various species

### 2015-07-29 tex/draft/ms_kedit_tedit_updated.tex
update

--- 2015-07-27 tex/draft/ms_kedit_tedit.tex
@@
 
-To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state}, for which the gas surface density $\Sigma_{\rm d}$ and mass flux $\dot{M}$ are constant in time and independent of semimajor axis\footnote{However, we also include stellar irradiation in the disk thermal structure, which changes the steady-state solution as we explain further in this section.}; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
+To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state}, for which the gas surface density $\Sigma$ and mass flux $\dot{M}$ are constant in time and independent of semimajor axis\footnote{However, we also include stellar irradiation in the disk thermal structure, which changes the steady-state solution as we explain further in this section.}; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
 
@@
 \label{eq:disk}
-\Sigma_{\rm d}&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
-T_{\rm d} &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
+\Sigma&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
+T &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
 \end{subeqnarray}
 where $r$ is the semimajor axis.
-Based on some observations of protoplanetary disks \citep{andrews10}, our surface density profile, $\Sigma_{\rm d} \propto r^{-1}$, is flatter than that of \citet{chiang10}, i.e.
-$\Sigma_{\rm d} \propto r^{-3/2}$.
+Based on some observations of protoplanetary disks \citep{andrews10}, our surface density profile, $\Sigma \propto r^{-1}$, is flatter than that of \citet{chiang10}, i.e.
+$\Sigma \propto r^{-3/2}$.
 
@@
 \label{eq:nu}
-\nu=\alpha c_{\rm d} H_{\rm d}.
+\nu=\alpha c H.
 \end{equation}
-Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c_{\rm d}$, $H_{\rm d}$ are the isothermal sound speed and disk scale height, respectively:
+Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c$, $H$ are the isothermal sound speed and disk scale height, respectively:
 \begin{subeqnarray}
 \label{eq:cdHd}
-c_{\rm d} &=& \sqrt{\frac{k_{\rm B} T_{\rm d}}{\mu m_{\rm p}}} \slabel{eq:cd} \\
-H_{\rm d}&=& \frac{c_{\rm d}}{\Omega_{\rm k}} \slabel{eq:Hd},
+c &=& \sqrt{\frac{k_{\rm B} T}{\mu m_{\rm p}}} \slabel{eq:cd} \\
+H&=& \frac{c}{\Omega_{\rm k}} \slabel{eq:Hd},
 \end{subeqnarray}
-
 where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$ is the Keplerian angular velocity,  with $G$ the gravitational constant and $M_*$ the stellar mass.
@@
 \label{eq:Sigmaact}
-\Sigma_{\rm d}(R, T) = \frac{M_{\rm d}}{2 \pi r_{\rm c}^2 R^{\gamma}} T^{(-5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{-(2-\gamma)}}{T}}\Big],
+\Sigma(R, T) = \frac{M}{2 \pi r_{\rm c}^2 R^{\gamma}} T^{(-5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{-(2-\gamma)}}{T}}\Big],
 \end{equation}
-where $M_{\rm d}$ is the total disk mass and
+where $M$ is the total disk mass and
 \begin{subeqnarray}
@@
 \end{subeqnarray}
-where $t$ is time. We choose $M_{\rm d}=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
+where $t$ is time. We choose $M=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
 
-\textbf{Active disk at steady-state.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma_{\rm d}$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
+\textbf{Active disk at steady-state.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
 Since
 Calculating the midplane temperature self-consistently for an active disk is non-trivial.
-We thus use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm d, act}$.
+We thus use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm act}$.
 The equations governing the evolution of the steady-state disk are listed in \App{app:steadystate}.
@@
 \label{eq:opacity}
-\kappa=\kappa_0 T_{\rm d, act}^2,
+\kappa=\kappa_0 T_{\rm act}^2,
 \end{equation}
@@
 \label{eq:Tdact}
-T_{\rm d, act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
+T_{\rm act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
 \end{equation}
@@
 \label{eq:activeT}
-T_{\rm d}^4 = T_{\rm d, act}^4 + T_{\rm d, pas}^4,
+T^4 = T_{\rm act}^4 + T_{\rm pas}^4,
 \end{equation}
-where to avoid notation confusion $T_{\rm d, pas}=T_{\rm d}$ from Equation (\ref{eq:diskT}), the temperature profile for an active disk.
-We can then easily determine  $c_{\rm d}$ and $H_{\rm d}$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
+where to avoid notation confusion $T_{\rm pas}=T$ from Equation (\ref{eq:diskT}), the temperature profile for an active disk.
+We can then easily determine  $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
 For consistency, we choose  $\alpha=0.01$ as in the previous case.
-Finally, we determine $\Sigma_{\rm d}$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}).
-We note that $\Sigma_{\rm d}$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.
+Finally, we determine $\Sigma$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}).
+We note that $\Sigma$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.
 
@@
 %\label{eq:activeT}
-%T_{\rm d}^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
-%\dot{M} & = & 3 \pi \nu \Sigma_{\rm d} \slabel{eq:Mdot},
+%T^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
+%\dot{M} & = & 3 \pi \nu \Sigma \slabel{eq:Mdot},
 %\end{subeqnarray}
@@
 Based on disk observations (e.g., \citealt{andrews10}), we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$.
-We can then easily determine the radial temperature profile from Equation (\ref{eq:Tdact}), $c_{\rm d}$ and $H_{\rm d}$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
+We can then easily determine the radial temperature profile from Equation (\ref{eq:Tdact}), $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
 For consistency, we choose  $\alpha=0.01$ as in the previous case.
-Finally, we determine $\Sigma_{\rm d}$ from Equation (\ref{eq:Mdot}).
+Finally, we determine $\Sigma$ from Equation (\ref{eq:Mdot}).
 
-\textbf{Static disk.} To compare our results with those of \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}).
+\textbf{Static disk.} To compare our results with those of \citet{oberg11}, we also use a static disk model, with $\Sigma$ and $T$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}).
 %The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
@@
 \begin{array}{l l}
-\rho_{\rm s} s / (\rho_{\rm d} c_{\rm d}), & \quad s < 9 \lambda/4 \,\,\,\ \text{Epstein drag} \\
-4 \rho_{\rm s} s^2 / (9 \rho_{\rm d} c_{\rm d} \lambda), & \quad s < 9 \lambda/4, \,\text{Re} \lesssim 1 \,\,\,\ \text{Stokes drag.}
+\rho_{\rm s} s / (\rho c), & \quad s < 9 \lambda/4 \,\,\,\ \text{Epstein drag} \\
+4 \rho_{\rm s} s^2 / (9 \rho c \lambda), & \quad s < 9 \lambda/4, \,\text{Re} \lesssim 1 \,\,\,\ \text{Stokes drag.}
 \end{array} 
@@
 \end{equation}
-Here $\rho_{\rm d}$ is the gas midplane density, $\rho_{\rm s}=2$ g cm$^{-3}$ is the density of a solid particle, $s$ is the particle size, $\lambda$ is the mean free path, and Re is the Reynolds number.
+Here $\rho$ is the gas midplane density, $\rho_{\rm s}=2$ g cm$^{-3}$ is the density of a solid particle, $s$ is the particle size, $\lambda$ is the mean free path, and Re is the Reynolds number.
 
@@
 \label{eq:eta}
-\eta \equiv - \frac{\partial P_{\rm d}/\partial \ln r}{2 \rho_{\rm d} v_{\rm k}^2} \approx \frac{c_{\rm d}^2}{2 v_k^2}
+\eta \equiv - \frac{\partial P/\partial \ln r}{2 \rho v_{\rm k}^2} \approx \frac{c^2}{2 v_k^2}
 \end{equation}
-and $P_{\rm d} = \rho_{\rm d} c_{\rm d}^2$ is the disk midplane pressure.
+and $P = \rho c^2$ is the disk midplane pressure.
 
@@
 \label{eq:vgas}
-\dot{r}_{\rm gas} = - \frac{3}{\Sigma_{\rm d} \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma_{\rm d} \sqrt{r}) 
+\dot{r}_{\rm gas} = - \frac{3}{\Sigma \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma \sqrt{r}) 
 \end{equation}
-with $\Sigma_{\rm d}$ from Equation (\ref{eq:Sigmaact}).
-For the active disk steady-state solution (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma_{\rm d}$, with $\dot{M}$ fixed and $\Sigma_{\rm d}$ obtained from Equation (\ref{eq:Mdot}).
+with $\Sigma$ from Equation (\ref{eq:Sigmaact}).
+For the active disk steady-state solution (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}).
 
@@
 where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency, with $\mu_x$ the mean molecular weight.
-We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T_{\rm d}$.
+We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T$.
 For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively \citep{oberg11}.
@@
 \end{subeqnarray}
-where the desorption rate $R_{\rm des, x}$ for each particle type (i.e., composed of H$_2$O, CO$_2$ or CO) is evaluated at $T=T_{\rm d}(r)$, and the radial drift velocity $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for a passive disk and Equation (\ref{eq:rdotact}) for an active disk.
+where the desorption rate $R_{\rm des, x}$ for each particle type (i.e., composed of H$_2$O, CO$_2$ or CO) is evaluated at $T=T(r)$, and the radial drift velocity $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for a passive disk and Equation (\ref{eq:rdotact}) for an active disk.
 Equations (\ref{eq:dsdt}) and (\ref{eq:drdt}) describe the coupled desorption and radial drift, and can be derived straightforwardly from Equation (\ref{eq:tdes}).
@@
 %\vspace{-0.5in}
-\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M_{\rm d}=0.1 M_{\odot}$ (solid lines) and $M_{\rm d}=0.01 M_{\odot}$ (dashed lines).
+\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines).
 The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve).
@@
 
-We choose as a fiducial model a total disk mass $M_{\rm d}=0.1 M_{\odot}$.
+We choose as a fiducial model a total disk mass $M=0.1 M_{\odot}$.
 Observationally, disk masses span at least an order of magnitude around Solar type stars %, but this number is not universally valid.
@@
 \citep{andrews13}.
-% find that a linear scaling $M_{\rm d} \propto M_*$ is reasonable.
+% find that a linear scaling $M \propto M_*$ is reasonable.
 Giant planets, however, have been detected around small stars (e.g., \citealt{montet14}), which can have masses as low as $M_* \sim 0.1 M_{\odot}$.
 We thus explore the effect of disk mass on the location of snowlines.
-Figure \ref{fig:varMd} shows the desorption or final distance as a function on the initial location of a H$_2$O particle with initial size of 1 m, for two total disk masses: $M_{\rm d}=0.1 M_{\odot}$, our fiducial model, and $M_{\rm d}=0.01 M_{\odot}$.
+Figure \ref{fig:varMd} shows the desorption or final distance as a function on the initial location of a H$_2$O particle with initial size of 1 m, for two total disk masses: $M=0.1 M_{\odot}$, our fiducial model, and $M=0.01 M_{\odot}$.
 The simulations are stopped after the same timeframes as those in Figure \ref{fig:timeplots}.
 The location of the H$_2$O snowline is the same for both disks (the same holds true for the CO$_2$ and CO snowlines).
-The C/O ratio is thus insensitive to the choice of $M_{\rm d}$.
+The C/O ratio is thus insensitive to the choice of $M$.
 %This result also has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
@@
 Disk gaps and holes & $\rightarrow$ \\\hline
+Accretion rate evolution & $\rightarrow$ $\leftarrow$ \\\hline
+Stellar luminosity evolution & $\leftarrow$ \\\hline
 Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
@@
 The decrease in gas pressure in these gaps or holes will reduce the particles' drift velocity close to the gap edge, thus slowing them down and pushing the snowline outwards.
+
+\item \textbf{Accretion rate evolution.} Our steady-state active disk model assumes a constant mass accretion rate $\dot{M}$.
+However, $\dot{M}$ decreases over time, which lowers the accretional component of the disk temperature (Equation \ref{eq:Tdact}), thus pushing the snowline location inwards if the disk is optically thick \citep{garaud07}.
+Once $\dot{M}$ reaches low enough values for the snowline to become optically thin, the snowline location moves outwards \citep{garaud07}.
+Our choice of $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ roughly corresponds to the mass accretion rate during the quasi-static accretion phase of the disk evolution predicted by \citet{garaud07}.
+This is the timescale on which giant planets accrete most of their gaseous envelope; thus the implications of our results for the atmospheric composition of gas giants are robust.
+
+%our results for the snowline location and C/O ratio in a steady-state active disk will only be weakly affected by
+
+\item \textbf{Stellar luminosity evolution.} As the host star contracts during its pre-main sequence phase, its luminosity decreases, which reduces the disk temperature and pushes the snowline locations inwards.
+\citet{kennedy06} found that the snowline may move inward by a factor of $\sim$$15-20$ due to the stellar contraction.
 
@@
 \label{eq:diskeq}
-\nu & = & \alpha c_{\rm d} H_{\rm d} \slabel{eq:nueq} \\
-c_{\rm d}^2 & = & \frac{k_{\rm B} T_{\rm d, act}}{\mu m_{\rm p}} \slabel{eq:cdeq} \\
-\rho_{\rm d} & = & \frac{1}{\sqrt{2 \pi}} \frac{\Sigma_{\rm d}}{H_{\rm d}} \slabel{eq:rhoeq} \\
-H_{\rm d} & = & \frac{c_{\rm d}}{\Omega_{k}} \slabel{eq:Heq} \\
-T_{\rm d, act}^4 & = & \frac{3}{4} \tau T_{\rm d, surf}^4 \slabel{eq:Teq} \\
-\tau & = & \frac{1}{2} \Sigma_{\rm d} \kappa \slabel{eq:taueq} \\
-\nu \Sigma_{\rm d} & = & \frac{\dot{M}}{3 \pi} \slabel{eq:Mdot} \\
-\sigma T_{\rm d, surf}^4 & = & \frac{9}{8} \nu \Sigma_{\rm d} \Omega_{\rm k}^2 \slabel{eq:nusigeq} \\
-\kappa & = & \kappa_0 T_{\rm d,act}^2 \slabel{eq:kappaeq},
+\nu & = & \alpha c H \slabel{eq:nueq} \\
+c^2 & = & \frac{k_{\rm B} T_{\rm act}}{\mu m_{\rm p}} \slabel{eq:cdeq} \\
+\rho & = & \frac{1}{\sqrt{2 \pi}} \frac{\Sigma}{H} \slabel{eq:rhoeq} \\
+H & = & \frac{c}{\Omega_{k}} \slabel{eq:Heq} \\
+T_{\rm act}^4 & = & \frac{3}{4} \tau T_{\rm surf}^4 \slabel{eq:Teq} \\
+\tau & = & \frac{1}{2} \Sigma \kappa \slabel{eq:taueq} \\
+\nu \Sigma & = & \frac{\dot{M}}{3 \pi} \slabel{eq:Mdot} \\
+\sigma T_{\rm surf}^4 & = & \frac{9}{8} \nu \Sigma \Omega_{\rm k}^2 \slabel{eq:nusigeq} \\
+\kappa & = & \kappa_0 T_{\rm act}^2 \slabel{eq:kappaeq},
 \end{subeqnarray}
-where $T_{\rm d, surf}$ is the surface temperature of the disk and the other quantities are defined in the main text.
-This is a system of nine equations with nine unknowns ($\nu$, $c_{\rm d}$, $H_{\rm d}$, $T_{\rm d, act}$, $\rho_{\rm d}$, $\Sigma_{\rm d}$, $\tau$, $T_{\rm d, surf}$, $\kappa$) that can be solved numerically once $\alpha$ and $\kappa_0$ are specified.
+where $T_{\rm surf}$ is the surface temperature of the disk and the other quantities are defined in the main text.
+This is a system of nine equations with nine unknowns ($\nu$, $c$, $H$, $T_{\rm act}$, $\rho$, $\Sigma$, $\tau$, $T_{\rm surf}$, $\kappa$) that can be solved numerically once $\alpha$ and $\kappa_0$ are specified.
 
@@
 \end{subeqnarray}
-where $r_0=1$ AU, $\rho_0=\rho_{\rm d}(r_0)$ and $c_0=c_{\rm d}(r_0)$.
+where $r_0=1$ AU, $\rho_0=\rho(r_0)$ and $c_0=c(r_0)$.
 

### 2015-07-29 python/C_to_O_CH4.py

@@ -42,7 +42,7 @@ def n_C_O(species, CH4mid = 0, CH4max = 0):
     elif species == 'CH4' and CH4mid == 1 and CH4max == 0:
         return 0.0555 * 0.9, 0 #5.55% H2O (Oberg+08 Spitzer paper) 
     elif species == 'CH4' and CH4mid == 0 and CH4max == 1:
-        return 0.13, 0 #13% H2O (Oberg+08 c2d paper)
+        return 0.13 * 0.9, 0 #13% H2O (Oberg+08 c2d paper)
      
 
 

### 2015-07-29 tex/draft2/ms.tex
update

--- 2015-07-20 tex/draft/ms.tex
@@
 
-\title{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Accretion}
+\title{TBD}
 
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
+\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, et al. } %Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
@@
 \begin{abstract}
-The C/O ratio is an important signature of giant planet atmospheres and disk chemistry.
-We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and the resulting C/O ratio in gas and dust throughout the disk.
-We determine the snowline locations for a range of fixed initial particle sizes, in both an active and a passive disk.
-We find that grains with sizes sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk desorb instantaneously and at a fixed size-dependent location in the disk regardless of the particle's initial position.
-The snowline location move inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
-Compared to a static disk, we find that radial drift and gas accretion move the H$_2$O and CO$_2$ snowlines inwards by up to 60 \%, and the CO snowline by up to 50 \%.
-Our results thus set an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
-Our results
-have direct implications on the C/O ratio in giant planet atmospheres that have
-formed in situ. 
+...
 \end{abstract}
@@
 
-%\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al.
-2011).}
-
-%\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture.
-Gas giants are important.
-The chemical composition of their atmospheres constraints their formation and evolution.
-Is is therefore important to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) backtrack the planet formation location based on planet composition.}
-
-The chemical composition of the atmospheres of giant planets can provide important constraints on their formation, accretion and migration history.
-These processes are tightly linked to the structure and evolution of the protoplanetary disk in which such planets form.
-It is thus essential to understand the disk chemistry and dynamics well enough to predict the types of planet compositions that result from planet formation in different parts of the disk, as well as backtrack the planet formation location based on planet compositions.
-
-%\emgr{Add the rest as outlined above.} \\
-
-%\emgr{Paragraph 2: disks are complex, a lot of dynamical and chemical processes going on.
-However, we have detections of organic molecules.
-We have detections of snowlines.
-We see complex chemistry and many molecules we can study.}
-
-In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11c}, \citealt{oberg11b}).
-Of particular importance are volatile compounds, such as H$_2$O, CO$_2$, CO, N$_2$ etc.
-Recent detections of the disk snowlines (such as the CO snowline, \citealt{qi11}) are significant, since the snowline locations of volatile molecules determine their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
-%Recent detections of the CO snowline in disks (e.g., \citealt{qi11}) provide the framework...
-
-%\emgr{Add part about snowlines, complex chemistry, many molecules.} \\
-
-An important signature of disk chemistry is the carbon-to-oxygen (C/O) ratio, as it provides information about the relative abundances of the main C and O carrier species.
-While the gas C/O ratio in disks may be potentially observable \citep{du15}, such detections do not yet exist.
-However, we can indirectly infer the relative abundance of disk C and O using detections of C/O ratios in exoplanet atmospheres.
-Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}.
-One explanation for this discrepancy was proposed by \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e.
-H$_2$O, CO$_2$ and CO, have different condensation temperatures.
-This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
-\citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-
-%\emgr{Paragraph 3: let's focus on one (set of) molecules: the C/O ratio.
-It's an important signature of atmospheric chemistry.
-While we cannot detect it directly in disks, we have observations in giant planet atmospheres.
-It's different than stellar.
-Why?
-Possible explanation based on snowlines.}
-
-%\emgr{Start with the fact that we can't see C/O in disks directly. Then transition to planet atmospheres and modify the next sentence to follow up logically.} 
-%Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio.
-
-\citet{oberg11} assume a static disk with passive chemistry.
-In reality, additional dynamical and chemical processes affect the snowline locations and the resulting C/O ratio.
-Several works have addressed some of these effects.
-\citet{madhu14} use a steady-state active disk model that includes planetary migration and use the C/O ratio to constrain migration mechanisms.
-\citealt{alidib14} calculate the C/O ratio throughout the disk by incorporating the evolution of solids, i.e.
-radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors.
-\citet{thiabaud15} consider additional volatile species in their chemical network, such as N$_2$ or NH$_3$, and find that the C/O ratio may be enriched by up to four times the Solar value in certain parts of the disk.
-
-Most studies of the C/O ratio in disks are particularly focused on the inner disk, in order to explain observations of super-stellar C/O ratios in the atmospheres of close-in exoplanets \citep{madhu11}.
-In addition, the simultaneous inclusion of various dynamical and/or chemical processes makes it difficult to quantify the separate effect of each of these processes on snowline locations and the C/O ratio.
-In this paper, we focus on two particular dynamical effects i.e.
-the radial drift of solids and viscous gas accretion.
-We enhance the model of \citet{oberg11} by considering these two additional processes.
-We perform a systematic study to understand the detailed qualitative and quantitative effects of radial drift and gas accretion on the H$_2$O, CO$_2$ and CO snowline locations, and the resulting C/O ratio in gas and dust throughout the protoplanetary disk.
-More importantly, we obtain a limit on how close to the star the snowline locations can be pushed by radial drift and gas accretion.
-
-%\emgr{Paragraph 4: the static disk model is a great start, but there are additional dynamical and chemical processes to take into account.
-Several studies have done that in some form or another --- cite Ali-Dib+14, Madhusudhan+14, Thiabaud+15.
-Briefly discuss their assumptions.} \\
-
-%\emgr{Paragraph 5: finally introduce our goal for this paper: study the dynamical effects such as drift and gas accretion.
-State our goals: understand the detailed qualitative and quantitative effect of drift and gas accretion on snowline locations, obtain a limit on how far in can the snowlines be pushed, and see how that affects the C/O ratio throughout the disk.
-State that our specific goals motivate the use of a simplified model (in other words, why what we are doing is different from the studies cited in the previous paragraph).
-}
-
-%\emgr{This needs some restructuring and additions along the lines discussed above.} In order to obtain more realistic estimates for C/O ratios across protoplanetary disks, dynamical processes and the disk evolving chemistry have to be taken into account.
-In this paper, we enhance the model of \citet{oberg11} considering two additional dynamic effects: (1) the radial drift of solids throughout the protoplanetary disk, and (2) the viscous accretion of the disk gas onto the host star.
-Our goal is two-fold: (1) to quantify the effect of radial drift of solids of different sizes on the location and shape of H$_2$O, CO$_2$ and CO snowlines, and (2) to calculate the resulting C/O ratio in gaseous and solid form throughout an actively accreting protoplanetary disk as a function of the grain size distribution and the evolutionary time of the nebula.
-
-%\emgr{Describe goals and approach of this paper, i.e.: study the importance of radial drift on the location of snowlines in protoplanetary disks and how radial drift affects the C/O ratio.
-Mention that disks are viscously accreting in the early stages of planet formation, hence gas accretion has to be taken into account.
-The goal of this paper is two-fold: (1) estimate using both numerical models and analytic arguments the range of particle sizes for which radial drift changes snowline location; (2) calculate the C/O ratio for various particle sizes (and possibly a size distribution) and at different times throughout the evolution of the gas disk.}
-
-This paper is organized as follows.
-In Section \ref{sec:model}, we present our disk, radial drift and desorption models, as well as the timescales relevant to the coupled drift-desorption process.
-We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for a passive and an active disk in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}.
-In Section \ref{sec:discussion}, we discuss the generality of our results, as well as additional effects on the snowline locations.
-Finally, we summarize our findings in Section \ref{sec:summary}.
-
-\section{Model Framework} 
-\label{sec:model}
-
-We present our protoplanetary disk model for both a passive and an active disk in section \ref{sec:disk}.
-In section \ref{sec:drift}, we describe our analytic model for the radial drift of solids.
-We summarize our ice desorption model in section \ref{sec:desorption}.
-Finally, we discuss the relevant timescales for dynamical effects in the desorption process in section \ref{sec:timescales}.
-
-\subsection{Disk Model}
-\label{sec:disk}
-
-\textbf{Passive disk.} A passive disk is only irradiated by the central star and does not experience radial movement of gas or accretion heating.
-We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}.
-The gas surface density and midplane temperature are
-\begin{subeqnarray}
-\label{eq:disk}
-\Sigma_{\rm d}&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
-T_{\rm d} &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
-\end{subeqnarray}
-where $r$ is the semimajor axis.
-Based on some observations of protoplanetary disks \citep{andrews10}, our surface density profile, $\Sigma_{\rm d} \propto r^{-1}$, is flatter than that of \citet{chiang10}, i.e.
-$\Sigma_{\rm d} \propto r^{-3/2}$.
-
-\textbf{Active disk with passive temperature profile.} The gas is accreting onto the central star in an active disk, which causes the gas surface density to decrease with time.
-We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
-\begin{equation}
-\label{eq:nu}
-\nu=\alpha c_{\rm d} H_{\rm d}.
-\end{equation}
-Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c_{\rm d}$, $H_{\rm d}$ are the isothermal sound speed and disk scale height, respectively:
-\begin{subeqnarray}
-\label{eq:cdHd}
-c_{\rm d} &=& \sqrt{\frac{k_{\rm B} T_{\rm d}}{\mu m_{\rm p}}} \slabel{eq:cd} \\
-H_{\rm d}&=& \frac{c_{\rm d}}{\Omega_{\rm k}} \slabel{eq:Hd},
-\end{subeqnarray}
-where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$ is the Keplerian angular velocity,  with $G$ the gravitational constant and $M_*$ the stellar mass.
-We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition of hydrogen and helium.
-The temperature profile for an active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}).
-From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters.
-Following \citet{hartmann98}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius.
-We choose $r_{\rm c}=100$ AU.
-The gas surface density is then given by the self-similar solution
-\begin{equation}
-\label{eq:Sigmaact}
-\Sigma_{\rm d}(R, T) = \frac{M_{\rm d}}{2 \pi r_{\rm c}^2 R^{\gamma}} T^{(-5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{-(2-\gamma)}}{T}}\Big],
-\end{equation}
-where $M_{\rm d}$ is the total disk mass and
-\begin{subeqnarray}
-\label{eq:T}
-T & \equiv & \frac{t}{t_{\rm c}} + 1 \\
-t_{\rm c} & \equiv & \frac{1}{3(2-\gamma)} \frac{r_{\rm c}^2}{\nu_{\rm c}},
-\end{subeqnarray}
-where $t$ is time. We choose $M_{\rm d}=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
-
-\textbf{Active disk steady-state.} A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma_{\rm d}$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
-Since calculating the midplane temperature self-consistently for an active disk is non-trivial, we use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm d, act}$.
-The equations governing the evolution of the steady-state disk are listed in \App{app:steadystate}.
-We assume an interstellar opacity for the dust grains given by \citet{bell94}, but reduced by a factor of 100.
-This reduction is due to the fact that disk opacities are lower than the interstellar one.
-The scaling is consistent with more detailed models of grain opacities in disks (e.g., \citealt{mordasini14}).
-Our opacity law is thus
-
-\begin{equation}
-\label{eq:opacity}
-\kappa=\kappa_0 T_{\rm d, act}^2,
-\end{equation}
-where $\kappa_0=2 \times 10^{-6}$. By solving the Equation set (\ref{eq:diskeq}) we find
-\begin{equation}
-\label{eq:Tdact}
-T_{\rm d, act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
-\end{equation}
-However, both accretion heating and stellar irradiation contribute to the thermal evolution of the disk. We thus compute the midplane temperature for our steady-state active disk as
-\begin{equation}
-\label{eq:activeT}
-T_{\rm d}^4 = T_{\rm d, act}^4 + T_{\rm d, pas}^4,
-\end{equation}
-where to avoid notation confusion $T_{\rm d, pas}=T_{\rm d}$ from Equation (\ref{eq:diskT}), the temperature profile for an active disk.
-We can then easily determine  $c_{\rm d}$ and $H_{\rm d}$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
-For consistency, we choose  $\alpha=0.01$ as in the previous case.
-Finally, we determine $\Sigma_{\rm d}$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}).
-We note that $\Sigma_{\rm d}$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.
-
-%Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady-state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):
-
-%\begin{subeqnarray}
-%\label{eq:activeT}
-%T_{\rm d}^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
-%\dot{M} & = & 3 \pi \nu \Sigma_{\rm d} \slabel{eq:Mdot},
-%\end{subeqnarray}
-%where $\dot{M}$, the mass flux, is constant in steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}.
-Based on disk observations (e.g., \citealt{andrews10}), we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$.
-We can then easily determine the radial temperature profile from Equation (\ref{eq:Tdact}), $c_{\rm d}$ and $H_{\rm d}$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
-For consistency, we choose  $\alpha=0.01$ as in the previous case.
-Finally, we determine $\Sigma_{\rm d}$ from Equation (\ref{eq:Mdot}).
-
-\textbf{Static disk.} To compare our results with those of \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}).
-The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
-
-\subsection{Radial Drift}
-\label{sec:drift}
-
-Solid particles in orbit their host star at the Keplerian velocity $v_{\rm k} \equiv \Omega_{\rm k} r$.
-The gas, however, experiences an additional pressure gradient, which causes it to rotate at sub-Keplerian velocity \citep{weidenschilling77}.
-Dust grains that are large enough thus experience a headwind, which removes angular momentum, causing the solids to spiral inwards and fall onto the host star.
-Small particles are well-coupled to the gas, while large planetesimals are decoupled from the gas.
-From \citet{chiang10}, the extent of coupling is quantified by the dimensionless stopping time, $\tau_{\rm s} \equiv \Omega_{\rm k} t_{\rm s}$, where $t_{\rm s}$ is
-\begin{equation}
-\label{eq:ts}
-t_{\rm s}= \left\{
-\begin{array}{l l}
-\rho_{\rm s} s / (\rho_{\rm d} c_{\rm d}), & \quad s < 9 \lambda/4 \,\,\,\ \text{Epstein drag} \\
-4 \rho_{\rm s} s^2 / (9 \rho_{\rm d} c_{\rm d} \lambda), & \quad s < 9 \lambda/4, \,\text{Re} \lesssim 1 \,\,\,\ \text{Stokes drag.}
-\end{array} 
-\right.
-\end{equation}
-Here $\rho_{\rm d}$ is the gas midplane density, $\rho_{\rm s}=2$ g cm$^{-3}$ is the density of a solid particle, $s$ is the particle size, $\lambda$ is the mean free path, and Re is the Reynolds number.
-
-For a passive disk, the radial drift velocity can be approximated analytically as
-\begin{equation}
-\label{eq:rdotpas}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big),
-\end{equation}
-where
-\begin{equation}
-\label{eq:eta}
-\eta \equiv - \frac{\partial P_{\rm d}/\partial \ln r}{2 \rho_{\rm d} v_{\rm k}^2} \approx \frac{c_{\rm d}^2}{2 v_k^2}
-\end{equation}
-and $P_{\rm d} = \rho_{\rm d} c_{\rm d}^2$ is the disk midplane pressure.
-
-For an active disk, the radial drift velocity has an additional term due to the radial movement of the gas, i.e.
-\begin{equation}
-\label{eq:rdotact}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{\dot{r}_{\rm gas}}{1+\tau_{\rm s}^2},
-\end{equation}
-where $\dot{r}_{\rm gas}$ is the radial gas accretion velocity and can be expressed as (e.g., \citealt{fkr02})
-\begin{equation}
-\label{eq:vgas}
-\dot{r}_{\rm gas} = - \frac{3}{\Sigma_{\rm d} \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma_{\rm d} \sqrt{r}) 
-\end{equation}
-with $\Sigma_{\rm d}$ from Equation (\ref{eq:Sigmaact}).
-For the active disk steady-state solution (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma_{\rm d}$, with $\dot{M}$ fixed and $\Sigma_{\rm d}$ obtained from Equation (\ref{eq:Mdot}).
-
-\subsection{Volatile Desorption}
-\label{sec:desorption}
-
-%\subsection{Disk and Desorption Model}
-
-In order for a volatile species to thermally desorb, it has to overcome the binding energy that keeps it on the grain surface.
-Following \citet{hollenbach09}, the desorption rate per molecule for a species $x$ can be expressed as
-\begin{equation}
-\label{eq:Rdes}
-R_{\rm{des}, x} = \nu_x \exp{(-E_x/T_{\rm grain})},
-\end{equation}
-where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency, with $\mu_x$ the mean molecular weight.
-We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T_{\rm d}$.
-For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively \citep{oberg11}.
-We use the desorption rate, $R_{\rm des}$, to estimate the desorption timescale for particles of different sizes as described in section \ref{sec:timescales}.
-
-%\emgr{Describe disk model --- both passive MMSN and active self-similar solution with $\alpha$ prescription for viscosity.
-Describe the model for evolving the surface density profile of planetesimals following Birnstiel et al.
-(2012).
-Describe temperature profile --- currently power-law, but may change.
-Describe desorption model and parameters following Hollenbach et al.
-(2009).
-Mention that the solids are perfect spheres composed of a single volatile.
-Perhaps this subsection needs to be split into subsubsections.}
-
-\subsection{Relevant Timescales}
-\label{sec:timescales}
-
-We can estimate the extent to which radial drift and gas accretion affect desorption by comparing the timescales for desorption, drift and accretion, for solids of different sizes and compositions.
-
-\textit{Desorption timescale.} We assume that the solid bodies are perfect spheres and are entirely composed of only one volatile species, i.e.
-either H$_2$O, CO$_2$ or CO \footnote{We discuss the validity of these simplifications in section \ref{sec:discussion}.}.
-The desorption timescale can then be estimated as
-\begin{equation}
-\label{eq:tdes}
-t_{\rm des}=\frac{\rho_{\rm s}}{3 \mu_x m_{\rm p}} \frac{s}{N_x R_{\rm des, x}},
-\end{equation}
-where $N_x \approx 10^{15}$ sites cm$^{-2}$ is the number of adsorption sites of volatile $x$ per cm$^2$ \citep{hollenbach09}.
-
-\textit{Radial drift timescale.} To order of magnitude, the radial drift timescale can be estimated as 
-\begin{equation}
-\label{eq:tdrift}
-t_{\rm r, drift} \sim \Big|\frac{r}{\dot{r}}\Big|,
-\end{equation}
-where $\dot{r}$ is the radial drift velocity given by Equation (\ref{eq:rdotpas}) for a passive disk and by Equation (\ref{eq:rdotact}) for an active disk.
-
-\textit{Gas accretion timescale.} The timescale for gas accretion onto the central star for an active disk is (e.g., \citealt{armitage10})
-\begin{equation}
-\label{eq:tgas}
-t_{\rm gas, acc} \sim \frac{r^2}{\nu} \sim \frac{1}{2 \alpha \eta \Omega_{\rm k}},
-\end{equation}
-with the latter expression derived from Equations (\ref{eq:nu}) and (\ref{eq:eta}).
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{drift_timescales_betaS1_gas_acc_new.pdf}
-%\vspace{-0.5in}
-\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
-The timescales are calculated at three representative locations, i.e.
-the H$_2$O, CO$_2$ and CO snowlines in the static disk.
-For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively.
-The horizontal dashed line represents a typical disk lifetime of 3 Myr.
-Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
-$t_{\rm r, drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.}
-\label{fig:timescales}
-\end{figure}
-
-For simplicity purposes, we calculate the radial drift timescale, $t_{\rm r, drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well.
-Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm r, drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk.
-As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence on disk location (see Equation \ref{eq:Rdes}).
-On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas.
-Thus for small particles in a passive disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}) \footnote{This is not true for an active disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.}.
-At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well.
-This is true for both passive and active disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.
-
-Of particular interest for our purposes is the particle size regime for which (1) $t_{\rm r, drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
-for $\sim$$0.5$ cm $\lesssim s \lesssim$ $1000$ cm, or (2) $t_{\rm gas, acc} \lesssim t_{\rm des} \lesssim t_{\rm d}$, i.e.
-for $\sim$$0.1$ cm $\lesssim s \lesssim$ $10$ cm.
-In these cases, radial drift or gas accretion (or both) are faster than thermal desorption.
-We note that $t_{\rm gas, acc}<t_{\rm d}$ always holds true.
-Particles of sizes that satisfy the requirements above will drift significantly due to radial drift or gas accretion before desorbing, thus moving the H$_2$O, CO$_2$ and CO snowlines closer towards the central star and changing the C/O ratio throughout the disk.
-We quantify these effects in sections \ref{sec:snowlines} and \ref{sec:COratio}.
-
-%\emgr{Calculate timescale for radial drift following Chiang \& Youdin (2010).
-Calculate desorption timescale following Hollenbach et al.
-(2009).
-Estimate gas accretion timescale for a given $\alpha$.
-Important: mention that, for simplicity and illustrative purposes, these calculations are performed for a passive disk.
-Show plot with the timescales as a function of particle size at different snowlines to show the regime in which drift matters}.
-
-\section{Snowline Locations}
-\label{sec:snowlines}
-
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=0.7\textwidth]{desorption_distance_passive_active_colorbar_test.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance (if a grain fully desorbs) or final distance (if a grain does not fully desorb), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both a passive disk (left panels) and an active disk (right panels).
-The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
-For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.}
-\label{fig:snowlines}
-\end{figure*}
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=0.7\textwidth]{desorption_distance_actual_vs_estimated_passive_active_new.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text).
-The estimate is performed for a passive disk (left panels) and an active disk (right panels).
-The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
-\label{fig:an_vs_actual}
-\end{figure*}
-
-In this section we use the model described in section \ref{sec:model} to quantify the effects of radial drift (passive disk) or radial drift and gas accretion (active disk) on the snowline location, for dust particles of different sizes composed of either H$_2$O, CO$_2$ or CO.
-Specifically, we determine a particle's final location (i.e., where the particle either fully desorbs or remains at its initial size due to a long desorption timescale) as a function of its initial position in the disk, after the gas disk has dissipated.
-The disk lifetime, $t_{\rm d}$, is particularly relevant since this is the timescale on which giant planets form.
-The snowline locations at $t=t_{\rm d}$ throughout the protoplanetary disk determine the disk C/O ratio in gas at this time, and thus the C/O ratio in giant planet atmospheres that have formed \textit{in situ}.
-
-For each species $x$, we determine the final location in the disk of a particle of initial size $s_0$ by solving the following system of coupled differential equations:
-
-\begin{subeqnarray}
-\label{eq:ddt}
-\frac{ds}{dt} &= & - \frac{3 \mu_x m_{\rm p}}{\rho_{\rm s}} N_x R_{\rm des, x}  \slabel{eq:dsdt} \\
-\frac{dr}{dt} &=& \dot{r} \slabel{eq:drdt},
-\end{subeqnarray}
-where the desorption rate $R_{\rm des, x}$ for each particle type (i.e., composed of H$_2$O, CO$_2$ or CO) is evaluated at $T=T_{\rm d}(r)$, and the radial drift velocity $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for a passive disk and Equation (\ref{eq:rdotact}) for an active disk.
-Equations (\ref{eq:dsdt}) and (\ref{eq:drdt}) describe the coupled desorption and radial drift, and can be derived straightforwardly from Equation (\ref{eq:tdes}).
-Our boundary conditions are $s(t_0)=s_0$, $r(t_0)=r_0$, and $s(t_{\rm d})=0$, where $t_0$ is the initial time at which we start the integration and $r_0$ is the initial location of the particle.
-We choose $t_0=1$ year, but our result is independent on the initial integration time as long as $t_0 \ll t_{\rm d}$.
-
-%As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully evaporates.
-
-%Thus a particle's final location will depend on whether a grain initially at a certain distance is completely desorbed or not within the disk lifetime of 3 Myr.
-For example, larger grains take longer to desorb and hence are more likely to not evaporate fully in a given timeframe.
-
-We define the final position of a grain as the disk location it has reached after $t_{\rm d}=3$ Myr, or the radius at which it completely desorbs if that happens after a time shorter than 3 Myr.
-Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both a passive and an active disk.
-We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state active disk results in Section \ref{sec:COratio}.
-Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for a passive nor for an active disk.
-Similarly, micron- to mm-sized particles in the passive disk do not drift or desorb unless they are located inside the static snowlines.
-This is not in contradiction with the desorption timescales for small particles from Figure \ref{fig:timescales}, since small grains only desorb fast at or nearby their respective snowlines, as mentioned in Section \ref{sec:timescales}.
-In an active disk, however, micron-to mm-sized grains do drift significantly since they move at the same velocity as the accreting gas.
-For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in a passive disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an active disk, we notice that particles of initial size $s_0$ desorb at a fixed distance $r_{\rm des}$ regardless of their original location in the disk.
-In fact, the only grains that will both drift and evaporate are those that reach their fixed final location (represented by the horizontal curves in Figure \ref{fig:snowlines}) within the disk lifetime.
-We show in section \ref{sec:COratio} that this result is essential in determining the C/O ratio throughout the disk for different particle sizes.
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=0.8\textwidth]{s_t_a.pdf}
-%\vspace{-0.5in}
-\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in a passive disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left).
-Particles desorb almost instantaneously.
-Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
-Particles lose most of their mass very close to the distance at which they fully desorb.}
-\label{fig:s_t_a}
-\end{figure*}
-
-Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm r, drift} \sim t_{\rm des}$, given an initial particle size.
-We can calculate this location analytically by equating Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and solving for $r=r_{\rm des}(s)$ for a given particle size $s$.
-Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance in a passive and an active disk (see Figure \ref{fig:snowlines}).
-We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it slightly deviates for particles larger than $s \gtrsim10$ cm.
-For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm r, drift}$ is a power-law in $r$ (for our parameters, $t_{\rm r, drift} \propto r^{-1/14}$ for the passive disk), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
-Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm r, drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
-
-%\emgr{I've yet to figure out why there is a discrepancy for the small grains.
-I'll keep thinking about it, but for now I'll leave it as it is since I don't want to give a false explanation just to have one.}
-
-%\emgr{I didn't write the text for Figure 4 since we haven't yet decided if in the end we will include it in some form. We can }
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_vs_s.pdf}
-%%\vspace{-0.5in}
-%\caption{Desorption distance as a function of initial particle size, for the range of particles that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines}).
-\textbf{Figure belongs to section 3.} \emgr{This is a figure that I'm not sure we should include, since it doesn't really provide that much useful inside.
-If we did include it though, there would be two panels both for the passive and the active disk.
-The caption is not complete since the active disk panel is not yet there.}}  %  (See text for a description of evolution to yet higher masses.)
-%\label{fig:r_vs_s}
-%\end{figure}
-
-%\emgr{Present the equation set that you are solving, $dr/dt=\dot{r}$, $ds/dt=...$.
-I don't think it is necessary to describe in detail the numerical method of solving the equations since it's pretty straightforward.
-Present the pretty rainbow plots side by side for passive and active disks, and highlight the differences.
-Insert the plots that shows that for the intermediate size particles, the desorption distance can be estimated analytically with good accuracy.
-Maybe also include the plot showing the desorption distance as a function of particle size, both for passive and active disk.
-Split into subsections?}
-
-\section{Results for the C/O Ratio}
-\label{sec:COratio}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete.pdf}
-%\vspace{-0.5in}
-\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel), an active disk with a passive temperature profile (middle panel) and a steady-state active disk (bottom panel), for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
-The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
-The horizontal dotted line represents the stellar value of 0.54.
-The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk.
-The snowline location moves inward as the particle size increases.}
-\label{fig:CO_ratio}
-\end{figure}
-
-We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star.
-Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
-Solid H$_2$O particles evaporate almost instantly, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
-The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
-Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active).
-The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane.
-
-%In Section \ref{sec:snowlines}, we showed that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-Figure \ref{fig:s_t_a} confirms both of these effects.
-The left panels show the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
-Indeed, the solid H$_2$O particles evaporate almost instantly, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime.
-However, our model assumes that particles drift continuously at any location in the disk.
-Therefore, a particle that can fully desorb during the disk lifetime for at least one initial location will always desorb, at a fixed distance as discussed in Section \ref{sec:snowlines} and displayed in Figure \ref{fig:snowlines}.
-The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
-Figure \ref{fig:s_t_a} thus proves claims (1) and (2) above.
-
-We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
-Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk with a passive temperature profile, and a steady-state active disk.
-The C/O ratio for a static disk is shown as a guideline.
-The plot is consistent with Figure \ref{fig:snowlines}.
-For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus move the snowline compared to the static disk.
-In contrast, even $\sim$micron-sized grains drift and desorb for the active disk, since they flow towards the host star together with the accreting gas.
-For the same particle size, the snowline locations are slightly closer to the central star in the active disk with a passive temperature profile, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
-Equation \ref{eq:rdotact}).
-The addition of accretional heating in the steady-state active disk moves the H$_2$O snowline outwards.
-This is due to the fact that accretional heating dominates in the inner disk, where high temperatures cause the grains to evaporate further away from the star.
-Once $r\gtrsim1-2$  AU, stellar irradiation dominates the thermal evolution of the disk, and therefore the CO$_2$ and CO snowlines locations are the same as in the active disk with a passive temperature profile.
-
-Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases.
-While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk.
-Bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals, though only the kilometer-sized boulders that are very close to the snownlines will contribute to determining the snowline location.
-Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
-Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
-Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution.
-
-For our choice of parameters, the snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk with a passive temperature profile and $r_{\rm{H_2O}} \approx 0.63$ AU for the steady-state disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
-For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
-Radial drift and gas accretion therefore push the snowline locations by up to $\sim$$60$ \% for H$_2$O and CO$_2$, and by up to $\sim$$50$ \% for CO.
-
-%\emgr{There will be 1-2 additional sentences for the third panel, the steady-state disk, once that finishes running.}
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.5\textwidth]{lala}
-%%\vspace{-0.5in}
-%\caption{C/O ratios for different particle size distributions...
-\textbf{Figure belongs to section 4.} (\emgr{It will probably be a multi-panel plot, for different particle size distributions, for both passive and active disk, and perhaps at different times for the active disk.
-Not yet unsure how that figure will look like since I don't have the results yet, so I'll leave the figure caption open for now.})}
-%\label{fig:...}
-%\end{figure}
-
-%\emgr{Motivate the fact that we can calculate a sharp, fixed snowline for each particle size by inserting the plot that shows that particles desorb almost instantaneously at a fixed distance.
-Then present the plots analogous to Fig.
-1 in \"Oberg et al.
-(2011) for different particle sizes, based on the snowline locations obtained in the previous section, both for passive and active disk.
-Perhaps show it at different times in the gas disk evolution (i.e., not just at 3 Myr) for the active disk.
-Then assume a particle size distribution and show the interpolated result for the C/O ratio.
-Generalize the result using a transition disk (this part might also fit in the discussion section).}
-
-%emgr{Present the flux equations used to keep track of the amount of C and O in gas and dust throughout the disk.
-Motivate the simplification of using a fixed particle sized, fixed timescale, and fixed desorption distance in the calculations by referring to the rainbow plot in the previous section and by inserting the plot that shows that particle desorb almost instantly at a fixed distance.
-Based on these assumptions, finally show the equivalent of Fig.
-1 in \"Oberg et al.
-(2011) for different particle sizes, and at different times in the gas disk evolution (i.e., not just at 3 Myr) --- a multi-panel plot could be a good idea.
-Discuss the result, trends, etc.
-Ideally, have a final plot showing the results for a particle size distribution rather than for individual particles.}
-
-\section{Discussion}
-\label{sec:discussion}
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=\textwidth]{time_plots.pdf}
-%\vspace{-0.5in}
-\caption{Desorption / final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve), and 3 Myr, the disk lifetime (blue curve).
-For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
-\label{fig:timeplots}
-\end{figure*}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{desorption_distance_varying_Md.pdf}
-%\vspace{-0.5in}
-\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M_{\rm d}=0.1 M_{\odot}$ (solid lines) and $M_{\rm d}=0.01 M_{\odot}$ (dashed lines).
-The timescales of the simulations and their color code are the same as in Figure \ref{fig:timeplots}.
-A lower disk mass does not change the snowline location.}
-\label{fig:varMd}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{desorption_distance_transition_disk_1000.pdf}
-%\vspace{-0.5in}
-\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines).
-The timescales of the simulations and their color code are the same as in Figure \ref{fig:timeplots}.
-Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge.}
-\label{fig:cavity}
-\end{figure}
-
-\subsection{Generality of Results: Dependence on Disk Parameters}
-\label{sec:incond}
-
-In this section we investigate how variations in our fiducial parameters, such as the total disk mass or the time in the disk evolution at which we determine the snowline locations and the C/O ratio, affect our results.
-The disk lifetime $t_{\rm d}=3$ Myr is the most representative for our calculations, as this is the timeframe in which giant planets accrete their gaseous atmospheres (e.g., \citealt{pollack96}, \citealt{piso14}).
-However, protoplanetary cores start acquiring their envelope at earlier times, before the core is fully formed (e.g., \citealt{rafikov06}).
-Recent models such as aerodynamic pebble accretion \citep{lambrechts12} suggest rapid core growth on timescales of $10^5$ years, which implies that the time when cores accrete their massive envelope may be at least an order of magnitude shorter than the disk lifetime.
-The composition of giant planet atmospheres, and specifically their C/O ratio, can thus depend on the abundance of H$_2$O, CO$_2$ and CO in gas and dust forms at earlier times than $t_{\rm d}$ in the disk evolution.
-
-Figure \ref{fig:timeplots} shows the desorption or final distance as a function of a particle's initial location in the disk, for grains of initial sizes of 10 cm and 1 m, composed of either H$_2$O, CO$_2$ or CO.
-These sizes are representative since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
-We stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively, to see how the snowline locations evolve with time and affect the C/O ratio.
-Particles of a given size that start at large stellocentric distances do not desorb within the shorter timeframes, e.g.
-$10^4$ or $10^5$ years, but they do evaporate at a fixed radius if their initial location is closer to the host star.
-While the amount of material that moves through the disk changes with time, the radius at which particles desorb and thus the snowline locations are independent of the time elapsed.
-Therefore, our results are not only valid at $t_{\rm d}$, but also throughout the earlier time evolution of the protoplanetary disk.
-
-We choose as a fiducial model a total disk mass $M_{\rm d}=0.1 M_{\odot}$, but this number is not universally valid.
-From observations of the dust continuum, \citet{andrews13} find that a linear scaling $M_{\rm d} \propto M_*$ is reasonable.
-Giant planets, however, have been detected around small stars (e.g., \citealt{montet14}), which can have masses as low as $M_* \sim 0.1 M_{\odot}$.
-We thus explore the effect of disk mass on the location of snowlines.
-Figure \ref{fig:varMd} shows the desorption or final distance as a function on the initial location of a H$_2$O particle with initial size of 1 m, for two total disk masses: $M_{\rm d}=0.1 M_{\odot}$, our fiducial model, and $M_{\rm d}=0.01 M_{\odot}$.
-The simulations are stopped after the same timeframes as those in Figure \ref{fig:timeplots}.
-The location of the H$_2$O snowline is the same for both disks (the same holds true for the CO$_2$ and CO snowlines).
-The C/O ratio is thus insensitive to the choice of $M_{\rm d}$.
-%This result also has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
-Since the snowline locations are independent of disk mass, we expect transition disks to exhibit the same drift-desorption behavior as standard disks under the simplified assumptions of our model.
-
-We also apply our active disk model to a transition disk, i.e.
-a protoplanetary disk with an inner cavity significantly depleted of gas.
-We choose a disk with an inner gap of radius $r_0=4$ AU, consistent with observations of TW Hya \citep{zhang13}, and with the surface density in the gap reduced by a factor of 1000.
-Figure \ref{fig:cavity} shows the desorption or final distance for a H$_2$O particle of initial size of 1 m, with the simulation stopped at the same timescales as in Figures \ref{fig:timeplots} and \ref{fig:varMd}.
-Particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the edge gap, due to the decrease in gas pressure inside the cavity, thus forming a snowline at $\sim$$3.8$ AU.
-This is consistent with the observations of \citet{zhang13} for the H$_2$O snowline being pushed outwards in a transition disk compared to a full disk.
-
-\subsection{Model Extensions}
-\label{sec:neglected}
-
-Our goals in this paper were (1) to gain a detailed qualitative and quantitative understanding of the effect of radial drift and gas accretion onto the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
-We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical processes.
-In what follows, we discuss these limitations and their effects.
-We note that our future work will address some of these issues.
-
-We summarize in Table 1 the potential physical and chemical processes occurring in disks and their effect on snowline locations.
-For the sake of completion, Table 1 also includes the processes addressed in this paper, i.e.
-radial drift and gas accretion.
-The neglected effects are discussed in more detail below.
-
-\begin{enumerate}
-\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered fixed for a given grain throughout its drift and evolution.
-However, grain growth has been observed in protoplanetary disks (e.g., \citealt{ricci10}, \citealt{perez12}), as well as theoretically constrained (e.g., \citealt{birnstiel10}, \citealt{birnstiel12}).
-In Section \ref{sec:COratio} we have shown that larger grains move the snowline locations closer in, but those locations remain fixed above a certain particle size.
-Particle growth will thus initially push the snowlines inwards.
-However, once the solids grow larger than km-sized, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
-It follows that grain growth will eventually push the snowline location outwards.
-
-\item \textbf{Turbulent diffusion.} The radial drift model presented in Section \ref{sec:drift} only considers a laminar flow and thus ignores turbulence.
-However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
-Turbulence causes eddies and vertical mixing, which are likely to reduce the radial gas accretion velocity.
-Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially.
-Back-diffusion across the snowline will change the shape of the snowline, as well as the C/O ratio in gas and dust both inside and outside of the snowline, due to the reduction of gas-phase volatile abundance interior to the snowline.
-
-\begin{table}[t!]
-\caption{The effects of dynamical and chemical processes on snowline shapes and locations}
-\begin{center}
-\begin{tabular}{|l|l|}\hline
-\textbf{Process} & \textbf{Effect} \\\hline
-Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star, $\rightarrow$ means that the snowline is pushed further from the host star.
-The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
-Gas accretion & $\leftarrow$ \\\hline
-Particle growth & $\rightarrow$ \\\hline
-Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
-Particle fragmentation & $\rightarrow$ $\leftarrow$ \\\hline
-Grain morphology & $\rightarrow$ \\\hline
-Particle composition & $\rightarrow$ $\leftarrow$ \\\hline
-Disk gaps and holes & $\rightarrow$ \\\hline
-Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
-\end{tabular}
-
-\end{center}
-\end{table}
-%causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.
-
-\item \textbf{Particle fragmentation.} Frequent particle collisions in disks cause them to fragment (e.g., \citealt{birnstiel12}).
-The fragmentation of meter- to km-sized particles will move the snowlines outwards, as smaller particles desorb faster and further out from the host star (cf.
-Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}).
-Large boulders, which neither drift nor desorb, may become e.g.
-meter-sized due to collisions and subsequent fragmentation, which will cause them to drift significantly before desorbing, pushing the snowlines inwards.
-Thus fragmentation can move the snowline locations in either radial direction.
-
-\item \textbf{Grain morphology.} Our model assumes that the ice particles are perfect, homogeneous spheres.
-However, this is not a very good approximation, since grain growth can be fractal rather than compact (\citealt{zsom10}, \citealt{okuzumi12}).
-The inhomogeneity due to cracks in the grain structure will cause the particles to desorb faster.
-They will therefore drift less before evaporating and will move the snowlines outwards.
-
-\item \textbf{Particle composition.} The ice particles in our model are assumed to be fully formed of either H$_2$O, CO$_2$ or CO.
-In reality, the grains have a layered structure, such as an interior composed of non-volatile materials (e.g., sillicates) covered by an icy layer.
-The ice thus only constitutes a fraction of the total particle mass, which accelerates its desorption and pushes the snowlines outwards.
-The grains may also be composed of a mixture of H$_2$O, CO$_2$ and CO ices.
-In this case, the desorption energies are largest for the most volatile species, which will move the snowlines inwards.
-
-%The relative fraction of each volatile, as well as their degree of mixing, will determine their desorption timescale, and as a result whether the snowlines are moved towards or away from the central star.
-
-\item \textbf{Disk gaps and holes.} The snowline locations will be different for transition disks, which have inner cavities significantly depleted of gas (e.g., \citealt{espaillat12}), or pre-transitional disks, which have a gap between an inner and outer full disk (e.g., \citealt{kraus11}).
-The decrease in gas pressure in these gaps or holes will reduce the particles' drift velocity close to the gap edge, thus slowing them down and pushing the snowline outwards.
-
-\item \textbf{Non-static chemistry.} As the goal of this paper was to explore only the dynamical effects on snowline locations and the C/O ratio in disks, we have assumed a simple, static chemical model.
-However, gas-grain chemistry is very complex and time-dependent.
-In the inner disk, chemistry approaches equilibrium due to intense sources of ionizing radiation (e.g., \citealt{ilgner04}), while in the outer disk high energy radiation and cosmic rays are the key drivers of chemistry, which is no longer in equilibrium (e.g., \citealt{vandishoeck06}).
-A multitude of chemical evolution models have been developed (see references in \citealt{henning13}), many of which contain tens or hundreds of chemical reactions.
-Due to the complexity of these chemical models, most of them are decoupled from disk dynamics.
-The effect of disk chemistry on snowline locations, shape, time evolution, or the C/O ratio is therefore difficult to estimate.
-In a future paper, we plan to use a parametrized chemical model and incorporate it in our radial drift calculation.
-
-%parametrize the chemical model developed by Merchantz et al. (in preparation) that traces the evolution of the main C and O oxygen carriers,
-
-\end{enumerate}
-
-%\emgr{Present the diagram that shows all the effects that can modify snowline location.
-For model limitations, include: non-inclusion of turbulence, assumption of perfect spheres when in fact they may have cracks, particles composed of a single volatile when in reality they are likely to be mixed, etc.
-Discuss uncertainty of initial conditions and estimate how much they matter.
-....}
-
-\section{Summary}
-\label{sec:summary}
-
-We study the effect of radial drift of solids and viscous gas accretion onto the central star on the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in a protoplanetary disk, assuming static chemistry.
-We develop a simplified model to describe the coupled drift-desorption process and determine the time evolution of particles of different sizes throughout the disk.
-We assume that the solid particles are perfect, homogeneous spheres, fully composed of either H$_2$O, CO$_2$ or CO.
-We apply our model to a passive disk, an active disk with a passive temperature profile, and a steady-state active disk that also takes into account stellar irradiation.
-We determine the desorption or final location of drifting particles after a time equal to the disk lifetime, and use this result to set an inner limit for the location of the H$_2$O, CO$_2$ and CO snowlines.
-Our results can be summarized as follows:
-
-\begin{enumerate}
-\item Radial drift and gas accretion affect desorption and move the snowline locations compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk.
-Moreover, boulders up to almost kilometer-size contribute to setting the snowline locations if they are very close to the snowlines.
-
-\item Particles with sizes in the above range desorb instantaneously, and at a fixed location in the disk that only depends on the particle size.
-Thus for each particle size there is a fixed and uniquely determined H$_2$O, CO$_2$ or CO snowline.
-
-\item The results of the numerical simulation are in agreement with the analytic solution of the drift-desorption system of differential equations if the stopping time $\tau_{\rm s} \ll 1$.
-We present an explicit analytic solution for the desorption distance in this regime.
-
-\item The snowline locations move inwards as the particle size increases; the innermost snowline is set by particles with initial size $s \sim 7$ m.
-Gas accretion causes even micron-sized particles to drift, desorb and move the snowline location compared to a static disk.
-A steady-state disk that includes accretion heating moves the H$_2$O snowline outwards compared to an active disk with a passive temperature profile, but has no effect on the CO$_2$ and CO snowline locations.
-
-\item Since realistic grain size distributions are dominated by the largest particles, the H$_2$O, CO$_2$ and CO snowlines are those created by the largest drifting particles in our model.
-This corresponds to the innermost snowlines that we determine.
-Our model thus sets a limit on how close to the central star the snowlines can be pushed by radial drift and gas accretion.
-
-\item For our fiducial model, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for a passive disk, 0.26 AU, 3.4 AU and 30 AU for an active disk with a passive temperature profile, and 0.63 AU, 3.4 AU and 30 AU for a steady-state active disk.
-Compared to a static disk, radial drift and gas accretion move the snowlines by up to 60 \% for H$_2$O and CO$_2$, and by up to 50 \% for CO.
-
-\item Our model finds that the C/O ratio is enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching unity between the CO$_2$ and CO snowlines.
-This is consistent with observations of C/O ratios in some exoplanet atmospheres.
-
-\item The snowline locations are independent of the time at which we stop our simulation and of the total disk mass.
-Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.
-
-\end{enumerate}
-
-Our model does not address additional effects, such as gas diffusion, grain composition and morphology, or complex time-dependent chemical processes.
-Future work will address some of these dynamical and chemical processes, with the goal of obtaining more realistic results for the snowline locations, shapes and time evolution, and the resulting effect on the C/O ratio.
-%Nevertheless,
-
-\appendix
-\section{Steady-state active disk solution} \label{app:steadystate}
-
-Following \citet{shakura73} and \citet{armitage10}, the steady-state solution for a thin actively accreting disk with an $\alpha$-prescription for viscosity is governed by the following set of equations:
-
-\begin{subeqnarray}
-\label{eq:diskeq}
-\nu & = & \alpha c_{\rm d} H_{\rm d} \slabel{eq:nueq} \\
-c_{\rm d}^2 & = & \frac{k_{\rm B} T_{\rm d, act}}{\mu m_{\rm p}} \slabel{eq:cdeq} \\
-\rho_{\rm d} & = & \frac{1}{\sqrt{2 \pi}} \frac{\Sigma_{\rm d}}{H_{\rm d}} \slabel{eq:rhoeq} \\
-H_{\rm d} & = & \frac{c_{\rm d}}{\Omega_{k}} \slabel{eq:Heq} \\
-T_{\rm d, act}^4 & = & \frac{3}{4} \tau T_{\rm d, surf}^4 \slabel{eq:Teq} \\
-\tau & = & \frac{1}{2} \Sigma_{\rm d} \kappa \slabel{eq:taueq} \\
-\nu \Sigma_{\rm d} & = & \frac{\dot{M}}{3 \pi} \slabel{eq:Mdot} \\
-\sigma T_{\rm d, surf}^4 & = & \frac{9}{8} \nu \Sigma_{\rm d} \Omega_{\rm k}^2 \slabel{eq:nusigeq} \\
-\kappa & = & \kappa_0 T_{\rm d,act}^2 \slabel{eq:kappaeq},
-\end{subeqnarray}
-where $T_{\rm d, surf}$ is the surface temperature of the disk and the other quantities are defined in the main text.
-This is a system of nine equations with nine unknowns ($\nu$, $c_{\rm d}$, $H_{\rm d}$, $T_{\rm d, act}$, $\rho_{\rm d}$, $\Sigma_{\rm d}$, $\tau$, $T_{\rm d, surf}$, $\kappa$) that can be solved numerically once $\alpha$ and $\kappa_0$ are specified.
-
-\section{Desorption distance analytic solution} \label{app:tdriftan}
-
-For a particle of size $s$ that desorbs and satisfies $\tau_{\rm s} \ll 1$ ($\tau_{\rm s}$ is the dimensionless stopping time, defined in Section \ref{sec:drift}), we can derive an explicit analytic solution for the particle's desorption distance in a passive disk.
-For $\tau_{\rm s} \ll 1$, a particle is in the Epstein drag regime (see Equation \ref{eq:ts}) and its drift velocity $\dot{r}$ (Equation \ref{eq:rdotpas}) can be approximated as
-\begin{equation}
-\label{eq:rdotapp}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \tau_{\rm s}.
-\end{equation}
-By using Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and setting $t_{\rm r, drift}=t_{\rm des}$, we can express a particle's desorption distance as
-
-\begin{equation}
-\label{eq:tan}
-r_{\rm des}=\Bigg(\frac{d\,\, \mathcal{W}\Big[\frac{(B/A)^{-q/d} q\,C}{d} q C\Big]}{q \,C}\Bigg)^{\frac{1}{q}},
-\end{equation}
-where $\mathcal{W}$ is the Lambert-W function, $q=3/7$ is the power-law coefficient in Equation (\ref{eq:diskT}), $d=-\frac{1}{2}+p-q$ with $p=1$ the power-law coefficient in Equation (\ref{eq:disksigma}), and
-\begin{subeqnarray}
-%\label{eq:tan}
-A & = &  \frac{\rho_0}{\rho_{\rm s}} \frac{r_0^2}{s c_0} r_0^d \\ 
-B & = & \frac{\rho_{\rm s} s}{3 \mu_{\rm x} N_{\rm x} \nu_{\rm x}} \\
-C & = & \frac{E_{\rm x}}{T_0} r_0^{-q}, \\
-\end{subeqnarray}
-where $r_0=1$ AU, $\rho_0=\rho_{\rm d}(r_0)$ and $c_0=c_{\rm d}(r_0)$.
-
-%
-%\emgr{Right now it's unclear to me what could go in an appendix, if anything.
-Maybe discuss a bit the algorithm to evolve $\Sigma_{\rm p}$ (although it is already explained in detail in the appendix of Birnstiel et al.
-2010).
-Maybe show some example profiles of $\Sigma_{\rm p}$ at different times and for different particle sizes.}
+\emgr{Background info.
+Importance of volatiles in disks and planetary atmospheres, detections of snowlines in disks, C/O ratios etc.
+State again the importance of radial drift and gas accretion on the snowlines location, and that a systematic study of the combination of these two particular effects across the disk has not been done before.
+Then transition to the fact that we provide such a systematic study in Paper I and in this paper: we enhance the model of Paper I by making two additions: (1) we add nitrogen in the static chemistry model, and (2) we explore how the abundance }
 

### 2015-07-30 tex/draft2/ms.tex
update

@@
 
-\title{TBD}
+\title{TBD (\textit{Effect of Radial Drift and Viscous Gas Accretion on Snowline Locations in Protoplanetary Disks: Volatile Abundances and Binding Energies?})}
 
@@
 State again the importance of radial drift and gas accretion on the snowlines location, and that a systematic study of the combination of these two particular effects across the disk has not been done before.
-Then transition to the fact that we provide such a systematic study in Paper I and in this paper: we enhance the model of Paper I by making two additions: (1) we add nitrogen in the static chemistry model, and (2) we explore how the abundance }
+Then transition to the fact that we provide such a systematic study in Paper I and in this paper: here we enhance the model of Paper I by making three additions: (1) we add N and CH4 in the static chemistry model, (2) we explore how different abundances of CH4 and of the N main carriers (N2 and NH3) affect the C/O  and N/O ratios, and (3) we explore how different binding energies of CO, N2 and (perhaps) CH4 affect the snowline locations.}
+
+\section{Model Review}
+
+\emgr{Review the four disk models, desorption model, relevant timescales.
+Summarize the findings of Paper I, i.e.
+particles of certain sizes desorb instantaneously and at a fixed particle size dependent location.
+Probably split into 2 subsections.}
+
+\section{Volatile Abundances and Binding Energies: Effect on C/O and N/O Ratios in a Static Disk}
+
+\subsection{Nitrogen and CH4}
+
+\emgr{Discuss that nitrogen is abundant in the solar system and should be abundant in disks as well, but that its dominant form is largely unknown.
+Discuss that the main carrier of nitrogen is N2, but some fraction of it can be NH3 as well.
+Present and motivate the choices that we make for NH3 abundances.
+Along the same lines, discuss CH4, and the choices that we make for CH4 abundances.}
+
+\subsection{Volatile Desorption Energies}
+
+\emgr{State that desorption energies for H2O and CO2 are well constrained experimentally, and that the CO2 binding energy is only weakly dependent on whether it's pure CO2 or combined with H2O, but that is not the case for CO, N2 (and perhaps CH4 and NH3?).
+Briefly discuss CO-CO, N2-N2, CO-H2O and N2-H2O (and perhaps the same for CH4 and NH3 if we find literature on that) experimental results for binding energies, and motivate the choices that we make.}
+
+\subsection{Results for C/O and N/O in a Static Disk}
+
+\emgr{For each of them (C/O and N/O), show a 3-panel plot as follows: each panel has a specific CH4/NH3 abundance (top: none, middle: median abundance, bottom: maximum abundance); for a given panel, have multiple curves for C/O or N/O, depending on the choice of binding energies, so that it's clear visually how the binding energy changes the snowline location.
+Discuss how different abundances  and binding energies affect snowline locations and C/O or N/O ratios.}
+\section{Results}
+
+\subsection{N2, NH3 and CH4 Snowline Locations}
+
+\emgr{One multipanel 3x3 (or 3x2) rainbow plot similar to the snowline plots from Paper I, for \textit{one} choice for the binding energies (perhaps the largest ones, since we want a limit on how far in we can push the snowlines?).
+Rows: snowlines as a function or particle size for passive, active and (maybe) steady-state disk.
+Columns: N2, NH3 and CH4.
+Not entirely certain how necessary these plots/subsection actually are, since it is exactly the same qualitatively as in Paper I...}
+
+\subsection{C/O and N/O Ratios}
+
+\emgr{For each of them (C/O and N/O) show a 3x3 multipanel plot similar to the C/O plot from Paper I, for the same choice of binding energies as in the previous subsection.
+Columns: C/O or N/O in passive, active and steady-state disk.
+Rows: No CH4/NH3, median CH4/NH3, maximum CH4/NH3.
+Again, not sure if we need/want this for all disk choices, at least in the case of C/O since we already have that in Paper I.
+For N/O, quantify how the snowline location changes from a static disk due to drift and gas accretion.}
+
+\section{Discussion}
+
+\emgr{Discuss how entrapment of volatiles by H2O affects volatile abundances and C/O ratios. More TBD.}
+
+\section{Summary}
+
+\emgr{Maybe we can include the summary in the discussion section?}
 

### 2015-08-06 tex/draft/ms_kedit_tedit_updated.tex
update

@@
 
-The chemical composition of protoplanerary disks affects planet formation efficiencies and the composition of nascent planets.
+The chemical composition of protoplanetary disks affects planet formation efficiencies and the composition of nascent planets.
 %[Need explicit statement that gas giants accrete their atmospheres from the nebular gas.] 
@@
 \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-This effect has also been observed in some exoplanet atmospheres, where superstellar C/O ratios have been detected (e.g., WASP-12b, \citealt{madhu11}).
+Claims of detections of superstellar C/O ratios also exist for exoplanet atmospheres
+%This effect may have also been observed in some exoplanet atmospheres, where superstellar C/O ratios may have been detected 
+(e.g., WASP-12b, \citealt{madhu11}).
 %Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.
@@
 \citealt{alidib14} calculate the C/O ratio throughout the disk by incorporating the evolution of solids, i.e.
-radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors.
-\citet{thiabaud15} consider additional volatile species in their chemical network, such as N$_2$ or NH$_3$, and find that the C/O ratio may be enriched by up to four times the Solar value in certain parts of the disk.
+radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors; they find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.
+\citet{thiabaud15} consider additional volatile species in their chemical network, such as N$_2$, NH$_3$ or CH$_4$, and find that the gas C/O ratio may be enriched by up to four times the Solar value in the outer parts of the disk where CH$_4$ and CO are the only gaseous carriers of C and O.
 
@@
 
-To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state}, for which the gas surface density $\Sigma$ and mass flux $\dot{M}$ are constant in time and independent of semimajor axis\footnote{However, we also include stellar irradiation in the disk thermal structure, which changes the steady-state solution as we explain further in this section.}; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
+To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state with accretion heating} (hereafter \textit{active disk accretionally heated}), for which the gas surface density $\Sigma$ and mass flux $\dot{M}$ are constant in time and independent of semimajor axis, for our choice of opacity (see Equation \ref{eq:opacity}) --- however, we also include stellar irradiation in the disk thermal structure, which changes the steady-state solution as we explain further in this section; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
 
@@
 We choose $r_{\rm c}=100$ AU.
-The gas surface density is then given by the self-similar solution
+The gas surface density is given by the self-similar solution
 
@@
 \label{eq:Sigmaact}
-\Sigma(R, T) = \frac{M}{2 \pi r_{\rm c}^2 R^{\gamma}} T^{(-5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{-(2-\gamma)}}{T}}\Big],
+\Sigma(R, \overline T) = \frac{M}{2 \pi r_{\rm c}^2 R^{\gamma}} \overline T^{-(5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{(2-\gamma)}}{\overline T}}\Big],
 \end{equation}
@@
 \label{eq:T}
-T & \equiv & \frac{t}{t_{\rm c}} + 1 \\
-t_{\rm c} & \equiv & \frac{1}{3(2-\gamma)} \frac{r_{\rm c}^2}{\nu_{\rm c}},
+\overline T & \equiv & \frac{t}{t_{\rm c}} + 1 \\
+t_{\rm c} & \equiv & \frac{1}{3(2-\gamma)^2} \frac{r_{\rm c}^2}{\nu_{\rm c}},
 \end{subeqnarray}
@@
 
-\textbf{Active disk at steady-state.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
+\textbf{Active disk accretionally heated.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
 Since
@@
 Small particles are well-coupled to the gas, while large planetesimals are decoupled from the gas.
-From \citet{chiang10}, the extent of coupling is quantified by the dimensionless stopping time, $\tau_{\rm s} \equiv \Omega_{\rm k} t_{\rm s}$, where $t_{\rm s}$ is
+From the review by \citet{chiang10}, the extent of coupling is quantified by the dimensionless stopping time, $\tau_{\rm s} \equiv \Omega_{\rm k} t_{\rm s}$, where $t_{\rm s}$ is
 \begin{equation}
@@
 
-For an active disk, the radial drift velocity has an additional term due to the radial movement of the gas, i.e.
+For an active disk, the radial drift velocity has an additional term due to the radial movement of the gas \citep{birnstiel12}, i.e.
 \begin{equation}
@@
 \end{equation}
-where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency, with $\mu_x$ the mean molecular weight.
+where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well, with $\mu_x$ the dimensionless mean molecular weight.
 We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T$.
@@
 either H$_2$O, CO$_2$ or CO \footnote{We discuss the validity of these simplifications in section \ref{sec:discussion}.}.
-The desorption timescale can then be estimated as
+The timescale to desorb a single layer of molecules can then be estimated as
 \begin{equation}
@@
 On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas.
-Thus for small particles in a passive disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}) \footnote{This is not true for an active disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.}.
+Thus for small particles in a passive disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}).
+This is not true for an active disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.
 At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well.
@@
 %\vspace{-0.5in}
-\caption{Desorption distance (if a grain fully desorbs) or final distance (if a grain does not fully desorb), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both a passive disk (left panels) and an active disk (right panels).
+\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both a passive disk (left panels) and an active disk (right panels).
 The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
@@
 In this section we use the model described in section \ref{sec:model} to quantify the effects of radial drift (passive disk) or radial drift and gas accretion (active disk) on the snowline location, for dust particles of different sizes composed of either H$_2$O, CO$_2$ or CO.
-Specifically, we determine a particle's final location (i.e., where the particle either fully desorbs or remains at its initial size due to a long desorption timescale) as a function of its initial position in the disk, after the gas disk has dissipated.
+Specifically, we determine a particle's final location (i.e., where the particle either fully desorbs or remains at its initial size due to having a desorption timescale longer than the time at which we stop the simulation) as a function of its initial position in the disk, after the gas disk has dissipated.
 The disk lifetime, $t_{\rm d}$, is particularly relevant since this is the timescale on which giant planets form.
-The snowline locations at $t=t_{\rm d}$ throughout the protoplanetary disk determine the disk C/O ratio in gas at this time, and thus the C/O ratio in giant planet atmospheres that have formed \textit{in situ}.
+The snowline locations at $t=t_{\rm d}$ throughout the protoplanetary disk determine the disk C/O ratio in gas at this time, and thus the C/O ratio in giant planet atmospheres that have formed \textit{in situ}, before planetesimal accretion or core dredging.
 
@@
 Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both a passive and an active disk.
-The results for the steady-state active disk are qualitatively similar to the active disk with a passive disk temperature gradient and therefore not shown.
+The results for the accretionally heated active disk are quantitatively similar with those of the active disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
 %We do not show the equivalent result for the steady-state active disk since The trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state active disk results in Section \ref{sec:COratio}.
@@
 We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
-Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then calculate the C/O ratio throughout the disk.
+Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then calculate the C/O ratio throughout the disk as a function of the particle size.
+Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
+Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.
 
@@
 
-Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk\footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
+Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk.
 Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
@@
 It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active).
+\footnote{Both of these conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
 The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
@@
 Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
-Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
+%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
 Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from this plot.
@@
 
-For our choice of parameters, the minimum snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk and $r_{\rm{H_2O}} \approx 0.63$ AU for the steady-state disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
+For our choice of parameters, the minimum snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk and $r_{\rm{H_2O}} \approx 0.63$ AU for the active disk accretionally heated; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
 For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
 Radial drift and gas accretion therefore push the snowline locations by up to $\sim$$60$ \% for H$_2$O and CO$_2$, and by up to $\sim$$50$ \% for CO.
+We note that the H$_2$O snowline in all disks is significantly closer to the host star compared with Solar system models, which place the H$_2$O snowline between $\sim$$2.7$ to $\sim$$3.1$ AU (\citealt{hayashi81}, \citealt{podolak04}, \citealt{martin12}).
+This is partially because we choose a colder disk model, as well as the fact that gas accretion rates decrease over time, moving the snowline location outwards (see also \citealt{garaud07} and Section \ref{sec:neglected}).
 
@@
 The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}.
-Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge.}
+Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge, due to being trapped in a pressure maximum.}
 \label{fig:cavity}
@@
 of initial sizes of 10 cm and 1 m, composed of either H$_2$O, CO$_2$ or CO.
-These sizes are representative since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
+These sizes are important since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
 We stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively.
@@
 that start at large stellocentric distances do not desorb within the shorter timeframes, e.g.
-$10^4$ or $10^5$ years.
-%, but they do evaporate at a fixed radius if their initial location is closer to the host star.
+$10^4$ or $10^5$ years, but they do evaporate at a fixed radius if their initial location is closer to the host star.
 While the amount of material that moves through the disk changes with time, the radius at which particles desorb and
-The snowline locations are thus independent of the time elapsed, and %. Therefore, 
+the snowline locations are thus independent of the time elapsed, and %. Therefore, 
 our results are %not only valid at $t_{\rm d}$, but also throughout the earlier 
@@
 However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
-Turbulence causes eddies and vertical mixing, which are likely to reduce the radial gas accretion velocity.
+Turbulence causes eddies and vertical mixing, which are likely to reduce the radial drift velocity of the solids.
 Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially.
@@
 Gas accretion & $\leftarrow$ \\\hline
-Particle growth & $\rightarrow$ \\\hline
+Particle growth & $\rightarrow$ $\leftarrow$ \\\hline
 Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
@@
 
-\item Particles with sizes in the above range desorb almost instantaneously once desorption has begun, and at a fixed location in the disk that only depends on the particle size.
+\item Particles with sizes in the above range desorb almost instantaneously once desorption has begun, and at a fixed location in the disk that only depends on the particle size and the gas accretion rate.
 Thus for each particle size there is a fixed and uniquely determined H$_2$O, CO$_2$ or CO snowline.
@@
 
-\item Since realistic grain size distributions are dominated by the largest particles, the H$_2$O, CO$_2$ and CO snowlines are those created by the largest drifting particles in our model.
+\item Since realistic grain size distributions are dominated in mass by the largest particles, the H$_2$O, CO$_2$ and CO snowlines are those created by the largest drifting particles in our model.
 This corresponds to the innermost snowlines that we determine.
@@
 \item Our model finds that the C/O ratio is enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching unity between the CO$_2$ and CO snowlines.
-This is consistent with observations of C/O ratios in some exoplanet atmospheres.
+This is consistent with possible detections of superstellar C/O ratios in some exoplanet atmospheres.
 
-\item The snowline locations are independent of the time at which we stop our simulation and of the total disk mass.
+\item The snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk remains optically thick .
 Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.
@@
 
-Following \citet{shakura73} and \citet{armitage10}, the steady-state solution for a thin actively accreting disk with an $\alpha$-prescription for viscosity is governed by the following set of equations:
+Following \citet{shakura73} and \citet{armitage10}, the steady-state solution for a geometrically thin, optically thick actively accreting disk with an $\alpha$-prescription for viscosity is governed by the following set of equations:
 

### 2015-08-06 python/C_to_O_active_steady_new.py

@@ -1,0 +1,187 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+from drift_timescales import read_drag_file_many
+from drift_timescales_active_Tdisk import Mdots, Sigmadisk
+from utils.constants import Msun
+
+def n_C_O(species):
+    
+    """
+        
+    Returns the abundance of C and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_C:
+        abundance of C with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 1.5, 1.5
+    elif species == 'CO2':
+        return 0.3, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0.6, 0
+    elif species == 'silicate':
+        return 0, 1.4
+mC = 12
+mO = 16
+mH = 1
+mSi = 28
+
+f = 0.01
+
+Mtot = (0.9*mC + 0.9*mO + 0.3*mC + 0.6*mO + 0.45*mH + 0.9*mO + 0.6*mC + 1.4*mO + 0.7*mSi)
+fCO = (0.9*mC + 0.9*mO) * f / Mtot
+fCO2 = (0.3*mC + 0.6*mO) * f / Mtot
+fH2O = (0.9*mO + 0.45*mH) * f / Mtot
+fCgr = (0.6*mC) * f / Mtot
+fSiO2 = (1.4*mO + 0.7*mSi) * f / Mtot
+
+#in the following we calculate T_freeze for various species
+s, a, arrayH2O = read_drag_file_many('s_neg3_8_steady_correct.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO2', 50, 50)
+s, a, arrayCO = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO', 50, 50)
+
+
+
+
[…]

### 2015-08-06 python/drift_timescales_active_Tdisk.py

@@ -274,6 +274,11 @@ def rdot_with_acc(r, s, alpha, Mdot, k0, dr = 1e-3, rhos = 3.0, mu = 2.35, T0 = 
 
 #######################################################################################
 
+def Mdots(r, s, alpha, Mdot, k0, rhos = 3.0, dr = 1e-3, Mstar = Msun, mu = 2.35, T0 = 120, betaT = 3./7):
+    
+    return -2 * np.pi * Sigmadisk(r, alpha, Mdot, k0, Mstar, mu, T0, betaT) * (r*AU) * \
+        rdot_with_acc(r, s, alpha, Mdot, k0, dr, rhos, mu, T0, betaT, Mstar, sigma) / (Mstar/yr)
+
 def tdes(mx, Ex, Tx, s, Nx = 1e15, rhos = 3.0):
   
      
@@ -520,6 +525,7 @@ def Sigmap_act(rin, rout, nr, ti, tf, nt, s, alpha, Mdisk, rc, rhos = 3.0, T0 = 
         
     
     return sigarray
+
 
 
     

### 2015-08-07 tex/draft/ms_kedit_tedit_updated.tex
update

@@
 We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO.
-We calculate the resulting C/O ratio in gas and dust throughout the disk.
+We estimate the resulting C/O ratio in gas and dust throughout the disk.
 We determine the snowline locations for a range of fixed initial particle sizes, in both an active and a passive disk.
@@
 For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively.
-The horizontal dashed line represents a typical disk lifetime of 3 Myr.
+The horizontal dot-dashed line represents a typical disk lifetime of 3 Myr.
 Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
@@
 We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it slightly deviates for particles larger than $s \gtrsim10$ cm.
-For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the passive disk), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
+For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the passive disk in the Epstein drag regime), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
 Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
@@
 
-\section{Results for the C/O Ratio}
+\section{C/O Ratio Estimates}
 \label{sec:COratio}
@@
 \centering
-\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete.pdf}
+\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new.pdf}
 %\vspace{-0.5in}
-\caption{C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel), an active disk (middle panel) and a steady-state active disk (bottom panel), for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
+\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel), an active disk (middle panel) and an active disk accretionally heated (bottom panel), for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
 The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
 The horizontal dotted line represents the stellar value of 0.54.
-The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk.
-The snowline location moves inward as the particle size increases.}
+The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk, with the temperature profile given by Equation (\ref{eq:diskT}) for the top two panels and by Equation (\ref{eq:activeT}).
+The snowline location moves inward as the particle size increases.
+The curves exterior to the to static snowline represent the desorption distance for particles in the size range corresponding to the color bar, but they do not form snowlines since readsorption is not taken into account.
+For both active disks, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, with the decrease more significant interior to the H$_2$O snowline.}
 \label{fig:CO_ratio}
@@
 We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
-Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then calculate the C/O ratio throughout the disk as a function of the particle size.
+Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then estimate the C/O ratio throughout the disk as a function of the particle size.
 Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
@@
 Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active).
-\footnote{Both of these conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
+It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active)\footnote{Both of these conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
 The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
@@
 We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
-Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and a steady-state active disk.
+Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and an active disk accretionally heated.
 The C/O ratio for a static disk is shown as a guideline.
+We note that particles desorbing outside the static snowline will not form snowlines, since our calculation of the desorption distance does not account for readsorption  \citep{hollenbach09}.
+
+Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the active disks ignore the movement of the desorbed ices with the accreting gas --- the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
+As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
+Specifically, the gas C/O ratios may be significantly lower inside the H$_2$O and CO$_2$ snowlines for the active disks, depending on the gas accretion rate, due to the additional oxygen contained by H$_2$O and CO$_2$ vapor that moves together with the accreting gas and crosses their respective snowlines.
+
+We thus present this result for the C/O ratio primarily as a visual comparison with previous work \citep{oberg11} and note that our estimates for the C/O ratio in the active disks are not quantitatively accurate at all disk locations.
+Future work will address this issue.
+
  %The plot is consistent with Figure \ref{fig:snowlines}. 
@@
 bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals.
+This is due to the fact that large bodies lose their mass as they desorb until they reach the same size as the smaller, meter-sized particles, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
 However, the contribution of kilometer-sized bodies to the snowline location is modest, since they only drift if they are located very close to the snowline.
@@
 This is partially because we choose a colder disk model, as well as the fact that gas accretion rates decrease over time, moving the snowline location outwards (see also \citealt{garaud07} and Section \ref{sec:neglected}).
+Observations of the CO snowline in TW Hya \citep{qi13} have found its location at a disk midplane temperature of 17 K (at 30 AU for the TW Hya specific temperature profile).
+The inferred  desorption temperature corresponds to the CO desorption temperature in a static disk, or to desorption from very small grains in an active disk, i.e.
+from grains that are too small to drift substantially.
+This suggests that the outer TW Hya disk is dominated by small grains, since larger particles would push the snowline location inwards, and therefore to higher desorption temperatures.
+This may seem contradictory to observations of grain growth in disks in general and in TW Hya in particular \citep{wilner00}.
+However, recent observations have revealed that grain growth is concentrated to the inner disk \citep{perez12} and outer disk snowlines may therefore be close to the ones expected in a static disk.
 
@@
 Gas accretion causes even micron-sized particles to drift, desorb and move the snowline location compared to a static disk.
-A steady-state disk that includes accretion heating moves the H$_2$O snowline outwards compared to an active disk, but has no effect on the CO$_2$ and CO snowline locations.
+A steady-state disk that includes accretion heating moves the H$_2$O snowline outwards compared to an active disk, but has no effect on the CO$_2$ and CO snowline locations, for our particular choice of mass accretion rate $\dot{M}$ and midplane opacity $\kappa$.
+%However, the magnitude of the effect of accretional heating on snowline locations will depend on .
 
@@
 
-\item Our model finds that the C/O ratio is enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching unity between the CO$_2$ and CO snowlines.
+\item Our C/O estimates find that the C/O ratio may be enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching its maximum value between the CO$_2$ and CO snowlines.
 This is consistent with possible detections of superstellar C/O ratios in some exoplanet atmospheres.
+We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the active disks.
 
-\item The snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk remains optically thick .
+\item The snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk midplane remains optically thick .
 Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.

### 2015-08-07 python/C_to_O.py

@@ -4,6 +4,8 @@ to the H2O, CO2 and CO snowlines"""
 
 import numpy as np
 from T_freeze import T_freeze
+from drift_timescales_active_Tdisk import Tdisk
+from utils.constants import Msun
 
 def n_C_O(species):
     
@@ -92,7 +94,7 @@ def n(T, elem):
             
             
 
-def C_O_ratio(a, T0 = 200, q = 0.62):
+def C_O_ratio(a, alpha, Mdot, k0, mu, Mstar = Msun, T0 = 200, q = 0.62, acc = 0):
     
     """
     
@@ -117,8 +119,10 @@ def C_O_ratio(a, T0 = 200, q = 0.62):
     C/O ratio in grains
     
     """
-    
-    T = T0 * a**(-q)
+    if acc == 0:
+        T = T0 * a**(-q)
+    else:
+        T = Tdisk(a, alpha, Mdot, k0, mu, Mstar, T0, q)
     
     return n(T, 'C') / n(T, 'O')
                        

### 2015-08-07 python/drift_timescales_active_Tdisk.py

@@ -2,7 +2,7 @@ from utils.constants import G, kb, mp, Msun, cmperau, AU, yr, sigma
 import numpy as np
 from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
 from T_freeze import T_freeze, Rdes, tevap, vib_freq
-from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
+#$from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
 import scipy
 from scipy.integrate import odeint, quad
 from scipy.optimize import brentq

### 2015-08-07 python/drift_timescales_updated_sigmap.py

@@ -2,7 +2,7 @@ from utils.constants import G, kb, mp, Msun, cmperau, AU
 import numpy as np
 from numpy import ones, zeros, shape,arange,Inf,maximum,minimum,exp,array,sqrt,invert
 from T_freeze import T_freeze, Rdes, tevap, vib_freq
-from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
+#from C_to_O import T_freeze_H20, T_freeze_CO2, T_freeze_CO
 import scipy
 from scipy.integrate import odeint, quad
 from scipy.optimize import brentq

### 2015-08-07 tex/draft/ms_kedit_tedit_updated.tex
update

@@
 \citet{madhu14} use a steady-state active disk model that includes planetary migration and use the C/O ratio to constrain migration mechanisms.
-\citealt{alidib14} calculate the C/O ratio throughout the disk by incorporating the evolution of solids, i.e.
-radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors; they find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.
-\citet{thiabaud15} consider additional volatile species in their chemical network, such as N$_2$, NH$_3$ or CH$_4$, and find that the gas C/O ratio may be enriched by up to four times the Solar value in the outer parts of the disk where CH$_4$ and CO are the only gaseous carriers of C and O.
+\citet{alidib14} calculate the C/O ratio throughout the disk by incorporating the evolution of solids, i.e.
+radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors.
+\citet{alidib14} use the 1+1D $\alpha$-disk model of \citet{hughes10}, in which the gas drifts outwards in the disk midplane, and thus small particles that are well-coupled to the gas will also advect outward.
+Their model assumes a cyclical conversion between H$_2$O or CO dust and vapor: large enough particles that are decoupled from the gas drift inwards and start desorbing.
+Once their sublimation is complete, back-diffusion moves the H$_2$O or CO vapor ouwards to their respective snowlines, where they instantly condense into mm-sized particles that diffuse outwards with the gas while coagulating into larger particles.
+Once the grains become large enough to decouple from the gas and drift inwards, the cycle restarts.
+This approach leads \citet{alidib14} to find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.
+\citet{thiabaud15} consider additional carbon carrier volatile species in their chemical network, such as CH$_4$, and find that the gas C/O ratio may be enriched by up to four times the Solar value in the outer parts of the disk where CH$_4$ and CO are the only gaseous carriers of C and O.
+They also include nitrogen carriers such as N$_2$ or NH$_3$, and perform similar calculations for nitrogen.
 
@@
 
-To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state with accretion heating} (hereafter \textit{active disk accretionally heated}), for which the gas surface density $\Sigma$ and mass flux $\dot{M}$ are constant in time and independent of semimajor axis, for our choice of opacity (see Equation \ref{eq:opacity}) --- however, we also include stellar irradiation in the disk thermal structure, which changes the steady-state solution as we explain further in this section; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
+To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state with accretion heating} (hereafter \textit{active disk accretionally heated}), for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis and the temperature profile is calculated using both accretional heating and stellar irradiation; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
 
@@
 where $r$ is the semimajor axis.
-Based on some observations of protoplanetary disks \citep{andrews10}, our surface density profile, $\Sigma \propto r^{-1}$, is flatter than that of \citet{chiang10}, i.e.
-$\Sigma \propto r^{-3/2}$.
+Our surface density profile is flatter than the $\Sigma \propto r^{-3/2}$ used by \citet{chiang10}.
+Our choice is inspired by observations of protoplanetary disks at radii larger than $\sim$20 AU (e.g., \citealt{andrews10}), which suggest that typical disks may have surface density profiles with $\Sigma \propto r^{-1}$.
+
+%Based on some observations of protoplanetary disks \citep{andrews10}, our surface density profile, $\Sigma \propto r^{-1}$, is flatter than that of \citet{chiang10}, i.e. $\Sigma \propto r^{-3/2}$.
 
@@
 This reduction is due to the fact that disk opacities are lower than the interstellar one.
-The scaling is consistent with more detailed models of grain opacities in disks (e.g., \citealt{mordasini14}).
+While this scaling is consistent with more detailed models of grain opacities in disks (e.g., \citealt{mordasini14}), realistic disk opacities are much less sensitive to changes in temperature than the interstellar opacity if substantial grain growth has occurred.
+However, the disk temperature does not vary significantly across the small region of the disk where accretion heating is important ($r \lesssim 1$ AU).
+Moreover, using an analytic opacity formula is more convenient since it results in a constant gas surface density in the inner disk region (see below).
 Our opacity law is thus
@@
 \end{equation}
-However, both accretion heating and stellar irradiation contribute to the thermal evolution of the disk. We thus compute the midplane temperature for our steady-state active disk as
+Since both accretion heating and stellar irradiation contribute to the thermal evolution of the disk, we compute the midplane temperature for our steady-state active disk as
 \begin{equation}
@@
 \end{equation}
-where to avoid notation confusion $T_{\rm pas}=T$ from Equation (\ref{eq:diskT}), the temperature profile for an active disk.
+where to avoid notation confusion $T_{\rm pas}=T$ from Equation (\ref{eq:diskT}), the temperature profile for a passive disk.
 We can then easily determine  $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
@@
 Finally, we determine $\Sigma$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}).
-We note that $\Sigma$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.
+In the inner portion of our disk ($r  \lesssim 1$ AU for our fiducial model with $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$), our choice of opacity (Equation \ref{eq:opacity}) implies that the disk has a constant surface density with radius (see Equations \ref{eq:diskeq}).
+%We note that $\Sigma$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.
 
@@
 The horizontal dot-dashed line represents a typical disk lifetime of 3 Myr.
+The particle size ordering at the minimum $t_{\rm drift}$ is not monotonic in snowline distance due to different drag regimes for those particle sizes at the snowline locations (Epstein drag at the H$_2$O and CO$_2$ snowlines, and Stokes drag at the CO snowline).
+Similarly, the ordering of $t_{\rm des}$ is not monotonic in snowline distance due to the non-monotony in mean molecular weight between H$_2$O, CO$_2$ and CO (18 $m_{\rm p}$, 44 $m_{\rm p}$ and 28 $m_{\rm p}$, respectively).
 Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.

### 2015-08-13 tex/draft/ms_kedit_tedit_updated.tex
update

@@
 In disks, the C/O ratio is regulated by the presence of snowlines of major volatiles at different distances from the central star.
-We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO.
-We estimate the resulting C/O ratio in gas and dust throughout the disk.
+We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk.
 We determine the snowline locations for a range of fixed initial particle sizes, in both an active and a passive disk.
@@
 Once the grains become large enough to decouple from the gas and drift inwards, the cycle restarts.
+This ```conveyor belt''' model is also used by \citet{ciesla06} for the evolution of H$_2$O in a viscous disk.
 This approach leads \citet{alidib14} to find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.
@@
 \end{equation}
-where $N_x \approx 10^{15}$ sites cm$^{-2}$ is the number of adsorption sites of volatile $x$ per cm$^2$ \citep{hollenbach09}.
+where $N_x \approx 10^{15}$ sites cm$^{-2}$ is the number of adsorption sites of volatile $x$ per cm$^2$, assuming that the particle has a smooth surface \citep{hollenbach09}.
 
@@
 The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+The desorption distance for a static disk is shown for comparison (dashed vertical and horizontal lines).
 The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
@@
 Specifically, we determine a particle's final location (i.e., where the particle either fully desorbs or remains at its initial size due to having a desorption timescale longer than the time at which we stop the simulation) as a function of its initial position in the disk, after the gas disk has dissipated.
-The disk lifetime, $t_{\rm d}$, is particularly relevant since this is the timescale on which giant planets form.
+The disk lifetime, $t_{\rm d}$, is particularly relevant since the timescale for giant planet formation must be less than $t_{\rm d}$, and plausibly $t_{\rm d}$.
 The snowline locations at $t=t_{\rm d}$ throughout the protoplanetary disk determine the disk C/O ratio in gas at this time, and thus the C/O ratio in giant planet atmospheres that have formed \textit{in situ}, before planetesimal accretion or core dredging.
@@
 The results for the accretionally heated active disk are quantitatively similar with those of the active disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
+We also show the static snowlines for comparison, which are calculated by balancing adsorption and desorption \citep{hollenbach09}.
 %We do not show the equivalent result for the steady-state active disk since The trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state active disk results in Section \ref{sec:COratio}.
@@
 
+Another interesting feature of Figure \ref{fig:snowlines} is that particles above a certain size ($\sim$7 m for our choice of parameters) all desorb at the same distance.
+This is due to the fact that once the large bodies pass the static snowline, they first lose mass, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
+
 \begin{figure*}[t!]
@@
 
+%This is due to the fact that large bodies lose their mass as they desorb until they reach the same size as the smaller, meter-sized particles, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
+
+Given $r_{\rm des}$, we need to only calculate the distance over which particles desorb to determine the location of a snowline.
+Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk.
+Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
+%, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
+%A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
+[not sure you need to explain this]
+The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
+Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
+It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active)\footnote{Both of these conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
+The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
 %\emgr{I've yet to figure out why there is a discrepancy for the small grains.
@@
 The snowline location moves inward as the particle size increases.
-The curves exterior to the to static snowline represent the desorption distance for particles in the size range corresponding to the color bar, but they do not form snowlines since readsorption is not taken into account.
+The curves exterior to the static snowline represent desorption distances in the absence of readsorption.
+Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline.
 For both active disks, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, with the decrease more significant interior to the H$_2$O snowline.}
@@
 
-We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
+Given our results in Section \ref{sec:snowlines}, a disk's C/O ratio in mainly affected by the snowline location for the particle size housing the most mass in ice.
+%We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
 Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then estimate the C/O ratio throughout the disk as a function of the particle size.
 Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
-Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.
+In Figure \ref{fig:CO_ratio}, we display the H$_2$O, CO$_2$, and CO snowline locations as a function of particle size for disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
+The minimum snowline distance for a disk is given by the curve corresponding to the maximum particle size it hosts.
+If grains have grown to radii larger than $\sim$7 m, then the 7m snowline applies (see Section 3).
+%Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.
 
@@
 
-Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk.
+%Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk.
 Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
@@
 [not sure you need to explain this]
-The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
+%The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
 Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active)\footnote{Both of these conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
+%It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active)\footnote{Both of these conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
 The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
@@
 
+Drift and gas accretion affect the C/O ratio in a disk both because they move the snowline locations of the main C and O carriers  and because they cause solids and gas---which contain different proportions of C and O---to move inward at different rates.
+As shown in Section 3, the snowline locations depend on disk age only indirectly, through changes in disk properties and grain size.
+The C/O ratio is a function of the locations of the snowlines and the abundances of H$_2$O, CO$_2$, and CO relative to the H$_2$ abundance in the disk midplane.
+These abundances evolve over time as solids and gas move inward at different rates.
+
+Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and an active disk accretionally heated, under the simplifying assumption that the abundance relative to hydrogen for each volatile is fixed, so that drift and accretion affect only the locations of the snowlines.
 We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
+Snowline locations correspond to $r_{\rm des}$ in Figure \ref{fig:snowlines}, representing the location at which particles desorb in the absence of readsorption.
+The C/O ratio for a static disk, where desorption and readsorption balance \citep{hollenbach09}, is shown as a guideline.
+We note that the true snowline for particles with $r_{\rm des}$ outside the static snowline is the static snowline itself---curves plotted at larger distances are not true snowlines.
+
+Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the active disks ignore the movement of the desorbed ices with the accreting gas---the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
+As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
+For example, for the disk parameters and particle sizes displayed in Figure 5, water molecules in solid particles drift up to $\sim$1000 times faster across the H2O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.
+This differential inward motion will result in an increased oxygen gas abundance inside the H$_2$O snowline, and thus a (in some cases much) lower gaseous C/O ratio in this region.
+Conversely, oxygen gas inside the water snowline will be depleted compared to the static disk if H$_2$O particles grow to planetesimal size and stall their migration between the H$_2$O and CO$_2$ snowlines, leaving only gaseous CO and CO$_2$ to accrete inward.
+Growth of large planetesimals can therefore increase the C/O ratio in the inner disk.
+
+Figure \ref{fig:CO_ratio} plots snowline curves for particle sizes less than $\sim$ 7m.
+In the outermost disk, H$_2$O, CO$_2$, and CO all solidify.
+Hence, relative drift across the CO snowline can alter only the abundances of volatiles between the CO$_2$ and CO snowlines, but not the C/O ratio in this region.
+Interior to the CO$_2$ snowline, however, relative drift is important.
+We have found that the largest drifting particles in our model ($\sim$7 m) drift faster than the gas at both the H$_2$O and CO$_2$ snowlines.
+We thus conclude that the C/O ratio interior to the H$_2$O and CO$_2$ snowlines in our active disks will be lower than in the static disk, due to the additional oxygen added to the gas by desorbing H$_2$O and CO$_2$.
+For these particle sizes, our calculated C/O ratio is an upper limit,  as indicated by the arrows in Figure \ref{fig:CO_ratio}.
+
+%We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
 Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and an active disk accretionally heated.
@@
 
-Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the active disks ignore the movement of the desorbed ices with the accreting gas --- the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
+%Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the active disks ignore the movement of the desorbed ices with the accreting gas --- the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
 As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
@@
 
-We thus present this result for the C/O ratio primarily as a visual comparison with previous work \citep{oberg11} and note that our estimates for the C/O ratio in the active disks are not quantitatively accurate at all disk locations.
+%e thus present this result for the C/O ratio primarily as a visual comparison with previous work \citep{oberg11} and note that our estimates for the C/O ratio in the active disks are not quantitatively accurate at all disk locations.
 Future work will address this issue.
@@
  %The plot is consistent with Figure \ref{fig:snowlines}. 
+The snowline locations in these disks exhibit several interesting features.
 For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus move the snowline compared to the static disk.
@@
 B
-bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals.
-This is due to the fact that large bodies lose their mass as they desorb until they reach the same size as the smaller, meter-sized particles, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
+bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals (see Section \ref{sec:snowlines}).
 However, the contribution of kilometer-sized bodies to the snowline location is modest, since they only drift if they are located very close to the snowline.
@@
 %\vspace{-0.5in}
-\caption{Desorption / final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+\caption{Desorption or final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
 The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve).
@@
 %\vspace{-0.5in}
-\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines).
+\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines).
 The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve).
@@
 %\vspace{-0.5in}
-\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines).
+\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines).
 The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}.
@@
 These sizes are important since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
-We stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively.
+We choose the active disk without accretion heating as a disk model, and we stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively.
 %, to see how the snowline locations evolve with time and affect the C/O ratio.
@@
 the snowline locations are thus independent of the time elapsed, and %. Therefore, 
-our results are %not only valid at $t_{\rm d}$, but also throughout the earlier 
+our results for the snowline locations are %not only valid at $t_{\rm d}$, but also throughout the earlier 
 valid throughout the time evolution of the protoplanetary disk.
@@
 Figure \ref{fig:varMd} shows the desorption or final distance as a function on the initial location of a H$_2$O particle with initial size of 1 m, for two total disk masses: $M=0.1 M_{\odot}$, our fiducial model, and $M=0.01 M_{\odot}$.
+Similarly to Figure \ref{fig:timeplots}, we perform our calculations for an active disk without accretion heating.
 The simulations are stopped after the same timeframes as those in Figure \ref{fig:timeplots}.
@@
 
-Our goals in this paper were (1) to gain a detailed qualitative and quantitative understanding of the effect of radial drift and gas accretion onto the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
+Our goals in this paper were (1) to gain a qualitative and quantitative understanding of the effect of radial drift and gas accretion onto the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
 We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical processes.
@@
 \begin{enumerate}
-\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered fixed for a given grain throughout its drift and evolution.
+\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered fixed for a given grain throughout its drift and evolution, since we do not take into account particle coagulation.
 However, grain growth has been observed in protoplanetary disks (e.g., \citealt{ricci10}, \citealt{perez12}), as well as theoretically constrained (e.g., \citealt{birnstiel10}, \citealt{birnstiel12}).
@@
 The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
-Gas accretion & $\leftarrow$ \\\hline
+Gas accretion & $\leftarrow$ \footnote{Gas accretion pushes the snowlines inwards compared to the snowline locations in a static disk.
+However, accretional heating may push the snowline outwards compared to an active disk without accretion heating.} \\\hline
 Particle growth & $\rightarrow$ $\leftarrow$ \\\hline
@@
 Our choice of $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ roughly corresponds to the mass accretion rate during the quasi-static accretion phase of the disk evolution predicted by \citet{garaud07}.
-This is the timescale on which giant planets accrete most of their gaseous envelope; thus the implications of our results for the atmospheric composition of gas giants are robust.
+This is the timescale on which giant planets accrete most of their gaseous envelope.%; thus the implications of our results for the atmospheric composition of gas giants are robust.
 
@@
 
-\item The snowline locations move inwards as the particle size increases; the innermost snowline is set by particles with initial size $s \sim 7$ m in our model.
+\item Since realistic grain size distributions are dominated in mass by the largest particles, the H$_2$O, CO$_2$ and CO snowlines are those created by the largest drifting particles in our model.
+This corresponds to the innermost snowlines that we determine.
+Our model thus sets a limit on how close to the central star the snowlines can be pushed by radial drift and gas accretion.
+
+\item The snowline locations move inwards as the particle size increases; the innermost snowline is set by particles with initial size $s \sim 7$ m in our model --- bigger particles drift too slowly to make it further in before desorbing (see Section \ref{sec:snowlines}).
 Gas accretion causes even micron-sized particles to drift, desorb and move the snowline location compared to a static disk.
@@
 
-\item Since realistic grain size distributions are dominated in mass by the largest particles, the H$_2$O, CO$_2$ and CO snowlines are those created by the largest drifting particles in our model.
-This corresponds to the innermost snowlines that we determine.
-Our model thus sets a limit on how close to the central star the snowlines can be pushed by radial drift and gas accretion.
-
-\item For our fiducial model, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for a passive disk, 0.26 AU, 3.4 AU and 30 AU for an active disk, and 0.63 AU, 3.4 AU and 30 AU for a steady-state active disk.
+\item For our fiducial model, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for a passive disk, 0.26 AU, 3.4 AU and 30 AU for an active disk, and 0.63 AU, 3.4 AU and 30 AU for a steady-state active disk with accretion heating.
 Compared to a static disk, radial drift and gas accretion move the snowlines by up to 60 \% for H$_2$O and CO$_2$, and by up to 50 \% for CO.
@@
 
-\item The snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk midplane remains optically thick .
-Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.
+\item The snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk midplane remains optically thick.
+%Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.
 

### 2015-08-14 tex/draft/ms_kedit_tedit_updated.tex
update

@@
 From the chemistry perspective, volatile compounds are particularly important.
-Their snowline locations determine their relative abundance in gaseous and solid form in the disk, and thus the chemical composition of nascent giant planets.
+Their snowline locations determine their relative abundance in gaseous and solid form in the disk,.
 Based on protostellar and comet abundances, some of the most important volatile molecules are H$_2$O, CO$_2$, CO, N$_2$.
 Recent observations of protoplanetary disks have provided valuable information about the abundances and snowline locations of some of these compounds.
-For example, the CO snowline has been detected in the disks around HD 163296 \citep{qi11} and TW Hya \citep{qi13}.
+For example, the CO snowline has been detected in the disk around TW Hya \citep{qi13}.
 Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}, and more such snowline detections are expected in future ALMA cycles.
@@
 \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-Claims of detections of superstellar C/O ratios also exist for exoplanet atmospheres
-%This effect may have also been observed in some exoplanet atmospheres, where superstellar C/O ratios may have been detected 
-(e.g., WASP-12b, \citealt{madhu11}).
+This effect was used to explain claims of detections of superstellar C/O ratios in exoplanet atmospheres (e.g., WASP-12b, \citealt{madhu11}).
 %Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.
@@
 
-\citet{oberg11} assumed a static disk with passive chemistry.
-In reality, additional dynamical and chemical processes affect the snowline locations and the resulting C/O ratio.
+\citet{oberg11} assumed a static disk with no chemical evolution.
+In reality, dynamical and chemical processes affect the snowline locations and the resulting C/O ratio.
 Several works have addressed some of these effects.
@@
 We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T$.
-For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively \citep{oberg11}.
+For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively (\citealt{oberg11} and references therein).
 We use the desorption rate, $R_{\rm des}$, to estimate the desorption timescale for particles of different sizes as described in section \ref{sec:timescales}.
@@
 Specifically, we determine a particle's final location (i.e., where the particle either fully desorbs or remains at its initial size due to having a desorption timescale longer than the time at which we stop the simulation) as a function of its initial position in the disk, after the gas disk has dissipated.
-The disk lifetime, $t_{\rm d}$, is particularly relevant since the timescale for giant planet formation must be less than $t_{\rm d}$, and plausibly $t_{\rm d}$.
+The disk lifetime, $t_{\rm d}$, is particularly relevant since the timescale for giant planet formation must be less than or equal to $t_{\rm d}$.
 The snowline locations at $t=t_{\rm d}$ throughout the protoplanetary disk determine the disk C/O ratio in gas at this time, and thus the C/O ratio in giant planet atmospheres that have formed \textit{in situ}, before planetesimal accretion or core dredging.
@@
 We display this result for the range of particle sizes that desorb at a fixed distance in a passive and an active disk (see Figure \ref{fig:snowlines}).
-We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it slightly deviates for particles larger than $s \gtrsim10$ cm.
+We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it deviates for particles larger than $s \gtrsim10$ cm.
 For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the passive disk in the Epstein drag regime), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
@@
 As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
-For example, for the disk parameters and particle sizes displayed in Figure 5, water molecules in solid particles drift up to $\sim$1000 times faster across the H2O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.
+For example, for the disk parameters and particle sizes displayed in Figure 5, water molecules in solid particles drift up to $\sim$1000 times faster across the H$_2$O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.
 This differential inward motion will result in an increased oxygen gas abundance inside the H$_2$O snowline, and thus a (in some cases much) lower gaseous C/O ratio in this region.
@@
 \begin{enumerate}
-\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered fixed for a given grain throughout its drift and evolution, since we do not take into account particle coagulation.
+\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered initially fixed for a given grain before it drifts and desorbs, since we do not take into account particle coagulation.
 However, grain growth has been observed in protoplanetary disks (e.g., \citealt{ricci10}, \citealt{perez12}), as well as theoretically constrained (e.g., \citealt{birnstiel10}, \citealt{birnstiel12}).

### 2015-08-14 python/C_to_O_CH4_COCO.py

@@ -50,7 +50,7 @@ def n_C_O(species, CH4mid = 0, CH4max = 0):
 T_freeze_H20 = T_freeze(18., 5800, 1e6)
 T_freeze_CO2 = T_freeze(44., 2000, 3e5)
 T_freeze_CH4 = T_freeze(16., 1300., 0.0555*1e6) #binding energy from Garrod+06 A&A paper
-T_freeze_CO = T_freeze(28., 1388, 1e6)
+T_freeze_CO = T_freeze(28., 834, 1e6)
 #T_freeze_CH4 = T_freeze(16., 1300., 0.13*1e6)
 
 
@@ -90,17 +90,17 @@ def n(T, elem, CH4mid, CH4max):
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
             n_C_O('H2O')[i] + n_C_O('C_grains')[i] + n_C_O('silicate')[i]])
 
-    elif T_freeze_CO2 >= T > T_freeze_CO:
+    elif T_freeze_CO2 > T > T_freeze_CH4:
         return np.array([n_C_O('CO')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
             n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
                 n_C_O('CO2')[i]])            
             
-    elif T_freeze_CO >= T > T_freeze_CH4:
-        return np.array([n_C_O('CH4', CH4mid, CH4max)[i], \
+    elif T_freeze_CH4 > T > T_freeze_CO:
+        return np.array([n_C_O('CO')[i], \
             n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
-                n_C_O('CO2')[i] + +n_C_O('CO')[i]])
+                n_C_O('CO2')[i] + n_C_O('CH4', CH4mid, CH4max)[i]])
                 
-    elif T_freeze_CH4 >= T:
+    elif T_freeze_CO > T:
         return np.array([0, n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
                 n_C_O('CO2')[i] + n_C_O('CO')[i] + n_C_O('CH4', CH4mid, CH4max)[i]])
             

### 2015-08-14 python/Nitrogen/N_to_O_N2N2.py

@@ -55,8 +55,8 @@ def n_N_O(species, NH3mid = 0, NH3max = 0):
 T_freeze_H20 = T_freeze(18., 5800, 1e6)
 T_freeze_NH3 = T_freeze(17., 2965., 5.5e4)
 T_freeze_CO2 = T_freeze(44., 2000, 3e5)
-T_freeze_CO = T_freeze(28., 1388., 1e6)
-T_freeze_N2 = T_freeze(28., 1266., 1.12e6)
+T_freeze_CO = T_freeze(28., 834., 1e6)
+T_freeze_N2 = T_freeze(28., 767., 1.12e6)
 #T_freeze_NH3 = T_freeze(17., 2965., 1.5371e5)
 
 

### 2015-08-24 Postdoc Applications/AnaMariaPisoResume.tex
update

\documentclass[margin,line]{res}
\usepackage{hyperref}

\oddsidemargin -.5in
\evensidemargin -.5in
\textwidth=6.0in
\itemsep=0in
\parsep=0in

\newenvironment{list1}{
  \begin{list}{\ding{113}}{%
      \setlength{\itemsep}{0in}
      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
      \setlength{\topsep}{0in} \setlength{\partopsep}{0in} 
      \setlength{\leftmargin}{0.17in}}}{\end{list}}
\newenvironment{list2}{
  \begin{list}{$\bullet$}{%
      \setlength{\itemsep}{0in}
      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
      \setlength{\topsep}{0in} \setlength{\partopsep}{0in} 
      \setlength{\leftmargin}{0.2in}}}{\end{list}}

\begin{document}

\name{Ana-Maria A. Piso \vspace*{.1in}}

\begin{resume}
\section{\sc Contact Information}
\vspace{.05in}
\begin{tabular}{@{}p{3.5in}p{4in}}

Harvard-Smithsonian Center for Astrophysics & {\it Phone:}    (617) 818-6780 \\         
60 Garden Street, MS-10 & {\it E-mail:}  apiso@cfa.harvard.edu\\       
Cambridge, MA  02138  & {\it WWW:} www.cfa.harvard.edu/$\sim$apiso \\     
\end{tabular}

%\section{\sc Research Interests}
%Bayesian statistics, spatial statistics, nonparametric regression,
%statistical methods for large datasets, statistics for public policy

\section{\sc Education}
{\bf Harvard University}, Cambridge, MA \\
\begin{list1}
\vspace{-0.1in}
\item[] Ph.D., Astronomy \& Astrophysics, May 2016 (expected)
\item[] Advisor: Dr. Karin \"Oberg 
\item[] Thesis Topic: ``The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets''
\end{list1}
{\bf Harvard University}, Cambridge, MA \\
\begin{list1}
\vspace{-0.1in}
\item[] A.M., Astronomy \& Astrophysics, May 2013
\item[] Advisor: Dr. Ruth Murray-Clay
\item[] Research Exam Project: ``On the Minimum Core Mass for Giant Planet Formation''
\end{list1}
{\bf Massachusetts Institute of Technology}, Cambridge, MA\\
%{\em Department of Statistics} 
\vspace*{-.1in}
\begin{list1}
\item[] S.B., Physics,  June 2010 \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, Major GPA: 4.6/5.0
\item[] S.B., Mathematics, June 2010\,\,\,\, Major GPA: 4.8/5.0
\end{list1}

\section{\sc Research Experience \& Employment}
\begin{list2}
\item[] {\bf Research assistant} \hfill {\bf August 2010 - July 2011}
%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
\begin{list1}
%\vspace{-0.1in}
\item[] MIT, EAPS Department \hfill \textit{Cambridge, MA}
\item[] Project: The Magnetic Field Signature of Super Earths
\item[] Advisor: Prof. Sara Seager \\
\end{list1}

\item[] {\bf Undergraduate researcher} \hfill {\bf June 2008 - June 2010}
%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
\begin{list1}
%\vspace{-0.1in}
\item[] MIT, Kavli Institute or Astrophysics \hfill \textit{Cambridge, MA}
\item[] Project: The Solar Wind (2008) \& Structure of Accretion Disks (2009 - 2010)
\item[] Advisors: Dr. Paola Rebusco \& Prof. Edmund Bertschinger \\
\end{list1}

\item[] {\bf Research assistant} \hfill {\bf June 2009 - August 2009}
%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
\begin{list1}
%\vspace{-0.1in}
\item[] Viena University of Technology (TU Wien) \hfill \textit{Vienna, Austria}
\item[] Project: Exact relativistic viscous fluid solutions in near horizon extremal Kerr background
\item[] Advisor: Dr. Daniel Grumiller \\
\end{list1}

\item[] {\bf Undergraduate researcher} \hfill {\bf January 2007 - August 2007}
%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
\begin{list1}
%\vspace{-0.1in}
\item[] MIT, Laboratory of Nuclear Science \hfill \textit{Cambridge, MA}
\item[] Project: Dark Matter Direct Detection
\item[] Advisors: Prof. Gabriela Sciolla \& Dr. Denis Dujmic \\
\end{list1}

\item[] {\bf Assistant manager} \hfill {\bf November 2005 - June 2006}
%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
\begin{list1}
%\vspace{-0.1in}
\item[] Neuron Group S.R.L. Software Company \hfill \textit{Bucharest, Romania}
\item[] Digital map designer and database manager for the '112 Emergency Call Center' national project
%\item[] Advisors: Dr. Paola Rebusco \& Prof. Edmund Bertschinger
\end{list1}
\end{list2}

%\vspace{-.4cm}
%\textbf {\emph{Research Assistant}} \hfill {\bf August 2011 - present}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item Conducted research in the area of exoplanetary magnetic fields
%\item Modeled the planet---star magnetic interaction between super Earths and their host stars
%\item Submitted results for publication as a first author to the Astrophysical Journal
%\end{list2}
%
%{\bf MIT, EAPS Department} \hfill \textit{Cambridge, MA}
%
%\vspace{-.4cm}
%\textbf {\emph{Research Assistant}} \hfill {\bf August 2010 - July 2011}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item Conducted research in the area of exoplanetary magnetic fields
%\item Modeled the planet---star magnetic interaction between super Earths and their host stars
%\item Submitted results for publication as a first author to the Astrophysical Journal
%\end{list2}
%
%{\bf Vienna University of Technology (TU Wien)} \hfill \textit{Vienna, Austria}
%
%\vspace{-.4cm}
%\textbf {\emph{Researcher}} \hfill {\bf June 2009 - August 2009}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item Conducted supervised research on extremal Kerr black holes 
%\item Presented on solar wind at a seminar aimed at graduate students and postdoctoral fellows
%\item Submitted results for publication in Physical Review D journal
%\end{list2}
%
%{\bf MIT, Kavli Institute for Astrophysics} \hfill \textit{Cambridge, MA}
%
%\vspace{-.4cm}
%\textbf {\emph{Researcher in the UROP \footnote{Undegraduate Research Opportunities Program}}} \hfill {\bf June 2008 - June 2010}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item  Conducted supervised research in the fields of solar wind and accretion disks
%\item Performed simulations using the Mathematica software
%\item Designed and published two Mathematica Demonstration Projects
%\end{list2}
%
%{\bf MIT, Laboratory of Nuclear Science} \hfill \textit{Cambridge, MA}
%
%\vspace{-.4cm}
%\textbf {\emph{Researcher in the UROP}} \hfill {\bf January 2007 - August 2007}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item Collaborated with a group of five researchers on directional detection of dark matter 
%\item Collected experimental data from a CCD camera and analyzed it using the ROOT framework 
%\item Built flash animations for presentations and updated the LNS website
%\end{list2}
%
%{\bf NEURON GROUP S.R.L. Software Company} \hfill \textit{Bucharest, Romania}
%
%\vspace{-.4cm}
%\textbf {\emph{Assistant Manager}} \hfill {\bf November 2005 - June 2006}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item Worked on the '112 Emergency Call Center' national project 
%\item Created digital maps and associated databases for Romanian cities and counties
%\item Prepared analyses and reviews of the company's projects for several conferences and fairs
%\end{list2}

\section{\sc Teaching \& Outreach}
\begin{list2}
\item[] {\bf Teaching Fellow} \hfill {\bf February 2012 - May 2012}
%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
\begin{list1}
%\vspace{-0.1in}
\item[] Harvard College class SPU 30: Life as a Planetary Phenomenon \hfill \textit{Cambridge, MA}
\item[] Course Head: Prof. Dimitar Sasselov
\item[] Held two weekly two-hour sections
\end{list1}

\vspace{0.2in}

\item[] {\bf Science Club For Girls Mentor Scientist} \hfill { \bf September 2014 - May 2015}
\begin{list1}
\item[] Taught second grade girls at the Amigos School the class ``Sound \& Light'' \hfill \textit{Cambridge, MA}
\end{list1}

\vspace{0.2in}

\item[] {\bf WISTEM Program Mentor} \hfill {\bf September 2013 - present}
\begin{list1}
\item[] Mentor for a Harvard College undergraduate \hfill \textit{Cambridge, MA}
\end{list1}

\vspace{0.2in}

\item[] {\bf CfA Summer Mentor} \hfill {\bf June 2014 - August 2014}
\begin{list1}
\item[] Co-mentored an REU summer student \hfill \textit{Cambridge, MA}
\end{list1}

\vspace{0.2in}

\item[] {\bf Co-Organizer of Harvard Graduate Student Prospective Visits} \hfill {\bf March 2013}
\begin{list1}
\item[] Organized and coordinated meetings and activities for two groups of ~10 prospective graduate students each \hfill \textit{Cambridge, MA}
\end{list1}

\end{list2}

%
%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
%
%\vspace{-.4cm}
%\textbf {\emph{Teaching Fellow}} \hfill {\bf February 2012 - May 2012}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item Teaching assistant for SPU 30: Life as a Planetary Phenomenon (Course Head: Dimitar Sasselov)
%\item Held two weekly two-hour sections
%\item Led discussions, participated in TF meetings and graded weekly problem sets
%\end{list2}

\section{\sc Refereed Publications}

\begin{list2}
\item[] {\bf Piso, A.-M. A.}, \"Oberg, K. I., Birnstiel, T., \& Murray-Clay, R. A. \textit{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}. ApJ, under review \\
\end{list2}

\begin{list2}
\item[] {\bf Piso, A.-M. A.}, Youdin, A. N., \& Murray-Clay, R. A. \textit{Minimum Core Masses for Giant Planet Formation with Realistic Equations of State and Opacities}. ApJ, 2015, 800, 82 \\
\end{list2}

\begin{list2}
\item[] {\bf Piso, A.-M. A.} \& Youdin, A. N. \textit{On the Minimum Core Mass for Giant Planet Formation at Wide Separations}. ApJ, 2014, 786, 21 \\
\end{list2}
%\end{list1}

\section{\sc Publications in Preparation}

\begin{list2}
\item[] {\bf Piso, A.-M. A.}, \"Oberg, K.I., \& Pegues, J. \textit{The Role of Ice Compositions and Morphology For Snowlines and the C/N/O Ratios in Active Disks} \\
\end{list2}

%\begin{list1}
%\item[] {\bf On the Minimum Core Mass for Giant Planet Formation} (2013, in prep.) \\
%(http://web.mit.edu/apiso/www/MagFieldFinal.pdf) - submitted to the Astrophysical Journal\\
%Authors: Ana-Maria A. Piso and Andrew N. Youdin  \\
%\end{list1}

%\begin{list2}
%\item {\bf Magnetic field signature of Super Earths} \\
%(http://web.mit.edu/apiso/www/MagFieldFinal.pdf) - submitted to the Astrophysical Journal\\
%Authors: Ana-Maria Piso, Paola Rebusco and Sara Seager\\
%\end{list2}
%
%\vspace*{-.13in}
%\begin{list2}
%\item {\bf Exact relativistic viscous fluid solutions in near horizon extremal Kerr background} \\
%(arxiv.org: 0909.2041 [astro-ph.SR]) - submitted to Physical Review D\\
%Authors: Daniel Grumiller and Ana-Maria Piso\\
%\end{list2}

%\vspace{0.2in}

\section{\sc Online Publications \& Educational Material}

%\vspace*{-.13in}
\begin{list2}
\item[] {\bf The Solar Wind} \\
(Mathematica Demonstration Project: http://demonstrations.wolfram.com/TheSolarWind/) \\
Author: Ana-Maria Piso \\
\end{list2}

\vspace*{-.13in}
\begin{list2}
\item[] {\bf The Interplanetary Magnetic Field (Parker Spiral)} \\
(Mathematica Demonstration Project: \\
http://demonstrations.wolfram.com/TheInterplanetaryMagneticFieldParkerSpiral/ \\
Author: Ana-Maria Piso \\
\end{list2}

\section{\sc Conferences and Seminars}

\begin{list2}
\item[] {\bf Minimum Core Masses for Giant Planet Formation} \\
CfA Exoplanet Pizza Lunch, Cambridge, MA, May 2015  \\
Internal department talk \\
\end{list2}

\begin{list2}
\item[] {\bf Minimum Core Masses for Giant Planet Formation} \\
Star and Planet Formation in the Southwest, Oracle, AZ, March 2015  \\
Contributed talk \\
\end{list2}

\begin{list2}
\item[] {\bf On the Minimum Core Mass for Giant Planet Formation} \\
CfA Exoplanet Pizza Lunch, Cambridge, MA, November 2013  \\
Internal department talk \\
\end{list2}

\begin{list2}
\item[] {\bf On the Minimum Core Mass for Giant Planet Formation} \\
Protostars and Planets VI, Heidelberg, Germany, July 2013  \\
Poster \\
\end{list2}

\begin{list2}
\item[] {\bf On the Minimum Core Mass for Giant Planet Formation} \\
IAUS 299: Exploring the Formation and Evolution of Planetary Systems, Victoria, BC, June 2013  \\
Contributed talk \\
\end{list2}

\begin{list2}
\item[] {\bf The Structure and Stability of Atmospheres Accreting around Protoplanetary Cores} \\
Exoplanets in Multi-body Systems in the Kepler Era, Aspen, CO, February 2013 \\
Poster \\
\end{list2}

\begin{list2}
\item[] {\bf Magnetic field signature of Super Earths} \\
AAS 217$^{th}$ Meeting, Washington, Seattle, January 2011  \\
Poster \\
\end{list2}

\vspace*{-.13in}
\begin{list2}
\item[] {\bf Exact relativistic viscous fluid solutions in NHEK background} \\
APS April Meeting, Washington, DC, February 2010  \\
Poster \\
\end{list2}

\vspace*{-.13in}
\begin{list2}
\item[] {\bf The Solar Wind}\\ 
Vienna Theory Lunch Club, TU Wien, Vienna, Austria, June 2009  \\
Invited talk \\
\end{list2}

%\section{\sc Leadership}
%{\bf Delta Psi Co-ed Fraternity} \hfill \textit{Cambridge, MA}
%
%\vspace{-.4cm}
%\textbf {\emph{President/House Manager}} \hfill {\bf February 2008 - February 2009}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item Coordinated a 5-day work week at the beginning of the fall semester
%\item Supervised the assembling of wooden floors in the house library and acquired furniture for the house
%\item Increased members' involvement in the fraternity by assigning maintenance duties
%\end{list2}
%
%{\bf MIT International Students Association} \hfill \textit{Cambridge, MA}
%
%\vspace{-.4cm}
%\textbf {\emph{President}} \hfill {\bf May 2008 - May 2009}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item Organized the MIT-Harvard-Wellesley Annual International Cruise and study breaks 
%\item Designed the setup and decorations of the MIT International Fair 2008 
%\item Improved awareness of international cultures in the MIT community
%\end{list2}

\section{\sc Professional Activities \& Service}
%\vspace*{-.17in}
\begin{list2}
\item[] American Physical Society member
\item[] American Astronomical Society member
%\item Romanian Student Association treasurer (2007-2010)
%\item MIT International Student Orientation coordinator (2008, 2009)
\end{list2}

\section{\sc Skills}
%\vspace*{-.17in}
\begin{list2}
\item[] Languages: Fluent in Romanian, English and Spanish, Conversant in German, Basics in French
\item[] Computer: Python, Mathematica, Matlab, LaTeX, C++, ROOT, Mac OS, Windows 2000/XP/Vista, Microsoft Office, Corel, Database Desktop
\end{list2}

%\section{\sc Honors and Awards}
%%\vspace*{-.17in}
%\begin{list2}
%\item Selected in the 20-person Romanian representative team for the International Physics Olympiad
%\item Awards in the Physics National Olympiad, Math and Chemistry Municipal Olympiads (2002-2005)
%\end{list2}

%\section{\sc Other Interests/Hobbies}
%\begin{list2}
%\item Dancing, swimming, shopping, fashion, reading, foreign languages, traveling
%\end{list2}

%\vspace{-.4cm}
%{\em Research Assistant} \hfill {\bf August, 2010 - present}\\  
%\vspace*{-.17in}
%\begin{list2}
%\item TBD1
%\item TBD2
%\item TBD3
%\end{list2}
%%\vspace{-.1cm}
%{\em NSF VIGRE Teaching Fellow} \hfill {\bf January - May, 2001}\\
%Head teaching assistant.   
%Duties included  shared administrative responsibilities with faculty
%instructor, fielding of all student inquiries, and oversight of
%graduate student teaching assistants and graders.
%\vspace*{.05in}  
%\begin{list2}
%\item 36-217 Probability Theory and Random Processes, Spring 2001.
%\end{list2}
%
%%\vspace{-.1cm}
%{\em Teaching Assistant} \hfill {\bf August, 2001  - present}\\
%Duties at various times have included 
%office hours and leading weekly computer lab exercises.
%
%
%
%\section{\sc Publications}
%Paciorek, C.J., J.S. Risbey, V. Ventura, and R.D.Rosen. 2002. Multiple indices of Northern Hemisphere Cyclone
%Activity, Winters 1949-1999. Journal of Climate 15:1573-1590.
%
%Paciorek, C.J., R. Condit, S.P. Hubbell, and R.B. Foster.  2000.
%The demographics of resprouting in tree and shrub species of a moist
%tropical forest.  Journal of Ecology 88:765-777.
%
%Paciorek, C.J., B.R. Moyer, R.A. Levin, and S.L. Halpern.  1995.
%Pollen consumption by hummingbird flower mite {\it Proctolaelaps
%  kirmsei} and possible fitness effects on {\it Hamelia patens}.
%Biotropica 27:258-262.  (author order determined by lot)     
%
%\section{\sc Papers in preparation}
%
%Ventura, V., C.J. Paciorek, and J.S. Risbey.  Controlling the proportion of falsely-rejected hypotheses when conducting multiple tests with geophysical data.
%
%Ickes, K., C.J. Paciorek, and S. Thomas.  Effects of wild pigs on
%forest demographic processes in Malaysia.
%
%\section{\sc Conference Presentations}
%Paciorek, C.J., J.S. Risbey, V. Ventura, and R.D.Rosen.  2001.  Changes in Northern Hemisphere winter storm activity (1949-1999) based
%on a comparison of cyclone indices.  8th International Meeting on
%Statistical Climatology, Luneberg, Germany, March, 2001.
%
%Paciorek, C.J. and R. Rosenfeld.  2000.  Minimum classification error
%training in exponential language models.  2000 Spring Transcription
%Workshop, College Park, Maryland.
%\vspace*{-.25in}  
%\begin{verbatim}http://www.nist.gov/speech/publications/tw00/html/abstract.htm#cp1-50\end{verbatim}
%
%\section{\sc Professional Experience}
%{\bf Bureau of Transportation Statistics, U.S. Department of
%  Transportation}, Washington, District of Columbia USA
%
%\vspace{-.3cm}
%{\em Summer researcher} \hfill {\bf May, 2000 - August, 2000}\\
%Carried out several consulting projects, including modelling of
%injuries to cadavers in crash test experiments, analysis of airline
%delay data, and advice on analysis of airline economics data.
%
%{\bf Abt Associates}, Bethesda, Maryland USA
%
%\vspace{-.3cm}
%{\em Associate Programmer Analyst and Research Assistant} \hfill {\bf
%  October, 1994 - August, 1996}\\
%Researcher and computer model developer for U.S. EPA Regulatory Impact
%Analysis of Section 403 Lead Paint Hazard Rule.  Other projects
%included database analysis, literature reviews, and cost-benefit analysis.
%
%\section{\sc Computer Skills} 
%\begin{list2}
%\item Statistical Packages:  R, S-Plus, BUGS; some experience
%  with SAS; extensive use of C and Fortran statistical libraries.
%\item Languages:  C++, Perl, Pascal, some use of Unix shell scripts,
%  MPI parallel processing library.
%\item Applications: Generic Mapping Tools (GMT) - Unix mapping software, \LaTeX, common Windows
%  database, spreadsheet, and presentation software
%\item Algorithms: Experience programming Markov Chain Monte Carlo
%  simulations of Bayesian posterior distributions
%\item Operating Systems:  Unix/Linux, Windows.\\ 
%\end{list2}

\end{resume}
\end{document}

### 2015-09-30 Postdoc Applications/AnaMariaPisoResume.tex
update

@@
 \item[] Advisor: Dr. Karin \"Oberg 
-\item[] Thesis Topic: ``The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets''
+\item[] Thesis Topic: ``The Effect of Disk Chemistry and Dynamics in Shaping Volatile Snowlines''   
 \end{list1}
@@
 %\vspace{-0.1in}
-\item[] Viena University of Technology (TU Wien) \hfill \textit{Vienna, Austria}
+\item[] Vienna University of Technology (TU Wien) \hfill \textit{Vienna, Austria}
 \item[] Project: Exact relativistic viscous fluid solutions in near horizon extremal Kerr background

### 2015-10-03 tex/draft/ms_final_revised.tex
update

--- 2015-08-14 tex/draft/ms_kedit_tedit_updated.tex
@@
 %location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
-Compared to a static disk, we find that radial drift and gas accretion move the H$_2$O and CO$_2$ snowlines inwards by up to 60 \%, and the CO snowline by up to 50 \%.
+Compared to a static disk, we find that radial drift and gas accretion \textbf{in a steady-state accretionally heated disk} move the H$_2$O snowline inwards by up to \textbf{40 \%}, the CO$_2$ snowlines by up to 60 \%, and the CO snowline by up to 50 \%.
 We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 For the active disk steady-state solution (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}).
+\textbf{For our choice of parameters for both active disks, we have found that the radial flow of gas is always directed inward, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}).
+We note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations.
+Since drifting particles larger than a few cm are only modestly affected by gas accretion, an outward gas flow would move the CO$_2$ and CO locations further away from the star only for the smallest particles in our model, which are well-coupled to the gas.}
 
@@
 For our choice of parameters, the minimum snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk and $r_{\rm{H_2O}} \approx 0.63$ AU for the active disk accretionally heated; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
-For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
-Radial drift and gas accretion therefore push the snowline locations by up to $\sim$$60$ \% for H$_2$O and CO$_2$, and by up to $\sim$$50$ \% for CO.
+For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU\footnote{\textbf{For the viscous disk, we calculated the static snowline location using the same temperature profile as that of the viscous disk, for consistency purposes.
+Thus $r_{\rm{H_2O}} \approx 0.98$ AU for the static disk in this scenario.}}, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
+\textbf{For the viscous disk model, which is the most realistic,} radial drift and gas accretion push the snowline locations inwards by up to \textbf{$\sim$$40$ \%} for H$_2$O, by up to  $\sim$$60$ \% for CO$_2$, and by up to $\sim$$50$ \% for CO.
 We note that the H$_2$O snowline in all disks is significantly closer to the host star compared with Solar system models, which place the H$_2$O snowline between $\sim$$2.7$ to $\sim$$3.1$ AU (\citealt{hayashi81}, \citealt{podolak04}, \citealt{martin12}).

### 2015-10-05 tex/draft/ms_final_revised.tex
update

@@
 
-Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and an active disk accretionally heated, under the simplifying assumption that the abundance relative to hydrogen for each volatile is fixed, so that drift and accretion affect only the locations of the snowlines.
+Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and an active disk accretionally heated, under the simplifying assumption that the abundance relative to hydrogen for each volatile is \textbf{initially} fixed, so that drift and accretion affect only the locations of the snowlines.
 We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
@@
 We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the active disks.
+\textbf{We plan to address this issue in a future paper.}
 

### 2015-10-05 tex/draft/ms_final_revised.tex
update

@@
 We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk.
-We determine the snowline locations for a range of fixed initial particle sizes, in both an active and a passive disk.
-We find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk desorb %instantaneously and [I think instantaneously is too strong]
+We determine the snowline locations for a range of fixed initial particle sizes, in both \textbf{evolving (active)} and \textbf{irradiated (passive) disks}.
+\textbf{For our fiducial disk model,} we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk desorb %instantaneously and [I think instantaneously is too strong]
 at a %fixed 
@@
 %location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
-Compared to a static disk, we find that radial drift and gas accretion \textbf{in a steady-state accretionally heated disk} move the H$_2$O snowline inwards by up to \textbf{40 \%}, the CO$_2$ snowlines by up to 60 \%, and the CO snowline by up to 50 \%.
+Compared to a static disk, we find that radial drift and gas accretion \textbf{in a viscous disk} move the H$_2$O snowline inwards by up to \textbf{40 \%}, the CO$_2$ snowlines by up to 60 \%, and the CO snowline by up to 50 \%.
 We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
@@
 Recent observations of protoplanetary disks have provided valuable information about the abundances and snowline locations of some of these compounds.
-For example, the CO snowline has been detected in the disk around TW Hya \citep{qi13}.
+For example, the CO snowline has been detected in the disk around TW Hya \citep{qi13}, \textbf{as well as in the disk around HD 163296 \citep{mathews13} using line emissions from DCO$^+$}.
 Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}, and more such snowline detections are expected in future ALMA cycles.
@@
 \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-This effect was used to explain claims of detections of superstellar C/O ratios in exoplanet atmospheres (e.g., WASP-12b, \citealt{madhu11}).
+This effect was used to explain claims of detections of superstellar C/O ratios in exoplanet atmospheres (e.g., WASP-12b, \citealt{madhu11}), \textbf{which however have been unambiguously refuted (\citealt{stevenson14}, \citealt{kreidberg15})}.
 %Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.
@@
 Once the grains become large enough to decouple from the gas and drift inwards, the cycle restarts.
-This ```conveyor belt''' model is also used by \citet{ciesla06} for the evolution of H$_2$O in a viscous disk.
+This ```conveyor belt''' model is \textbf{based on the pioneering work by \citet{cuzzi04}}, and \citet{ciesla06} for the evolution of H$_2$O in a viscous disk.
 This approach leads \citet{alidib14} to find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.
@@
 Each of these studies have considered a specific combination of dynamical and chemical effects.
-One combination that has not yet been considered is the combination of radial drift and viscous gas accretion.
+One \textbf{scenario} that has not yet been considered is the combination of radial drift and viscous gas accretion \textbf{in isolation}.
 Studying these two dynamical processes makes it possible to quantify their separate effect on snowline locations and the C/O ratio at various disk radii.
@@
 In Section \ref{sec:model}, we present our disk, radial drift and desorption models, as well as the timescales relevant to the coupled drift-desorption process.
-We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for a passive and an active disk in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}.
+We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for \textbf{an irradiated and an evolving disk} in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}.
 In Section \ref{sec:discussion}, we discuss the generality of our results, as well as additional effects on the snowline locations.
@@
 
-We present our protoplanetary disk model for both a passive and an active disk in section \ref{sec:disk}.
+We present our protoplanetary disk model for \textbf{an irradiated, an evolving, and a viscous disk} in section \ref{sec:disk}.
 In section \ref{sec:drift}, we describe our analytic model for the radial drift of solids.
@@
 
-To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state with accretion heating} (hereafter \textit{active disk accretionally heated}), for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis and the temperature profile is calculated using both accretional heating and stellar irradiation; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.
+To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{static disk}, which is solely irradiated by the host star and does not take into account gas accretion onto the star or radial drift; \textit{\textbf{irradiated} disk}, which has the same temperature profile as the static disk and does not experience gas accretion or accretional heating, but it takes into account radial drift of solids; \textit{\textbf{evolving} disk}, in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; and \textit{\textbf{viscous} disk}, for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis and the temperature profile is calculated using both accretional heating and stellar irradiation.
 
@@
 
-\textbf{Passive disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
+\textbf{Irradiated disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a irradiated disk similar to the prescription of \citet{chiang10}.
+The gas surface density and midplane temperature are
 \begin{subeqnarray}
@@
 
-\textbf{Active disk.} We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
+\textbf{Evolving disk.} We model the evolving disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
 \begin{equation}
@@
 The temperature profile for the %an
-active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}).
+evolving disk is assumed to be the same as for the irradiated disk and given by Equation (\ref{eq:diskT}).
 From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters.
@@
 
-\textbf{Active disk accretionally heated.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
+\textbf{Viscous disk.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
 Since
-Calculating the midplane temperature self-consistently for an active disk is non-trivial.
+Calculating the midplane temperature self-consistently for an evolving disk is non-trivial.
 We thus use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm act}$.
@@
 \end{equation}
-Since both accretion heating and stellar irradiation contribute to the thermal evolution of the disk, we compute the midplane temperature for our steady-state active disk as
+Since both accretion heating and stellar irradiation contribute to the thermal evolution of the disk, we compute the midplane temperature for our viscous disk as
 \begin{equation}
@@
 \end{equation}
-where to avoid notation confusion $T_{\rm pas}=T$ from Equation (\ref{eq:diskT}), the temperature profile for a passive disk.
+where to avoid notation confusion $T_{\rm pas}=T$ from Equation (\ref{eq:diskT}), the temperature profile for an irradiated disk.
 We can then easily determine  $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
@@
 
-For a passive disk, the radial drift velocity can be approximated as
+For an irradiated disk, the radial drift velocity can be approximated as
 \begin{equation}
@@
 
-For an active disk, the radial drift velocity has an additional term due to the radial movement of the gas \citep{birnstiel12}, i.e.
+For an evolving disk, the radial drift velocity has an additional term due to the radial movement of the gas \citep{birnstiel12}, i.e.
 \begin{equation}
@@
 with $\Sigma$ from Equation (\ref{eq:Sigmaact}).
-For the active disk steady-state solution (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}).
-\textbf{For our choice of parameters for both active disks, we have found that the radial flow of gas is always directed inward, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}).
+For the viscous disk (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}).
+\textbf{For our choice of parameters for both the evolving and the viscous disks, we have found that the radial flow of gas is always directed inward, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}).
 We note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations.
@@
 \end{equation}
-where $\dot{r}$ is the radial drift velocity given by Equation (\ref{eq:rdotpas}) for a passive disk and by Equation (\ref{eq:rdotact}) for an active disk.
-
-\textit{Gas accretion timescale.} The timescale for gas accretion onto the central star for an active disk is (e.g., \citealt{armitage10})
+where $\dot{r}$ is the radial drift velocity given by Equation (\ref{eq:rdotpas}) for an irradiated disk and by Equation (\ref{eq:rdotact}) for an evolving disk.
+
+\textit{Gas accretion timescale.} The timescale for gas accretion onto the central star for an evolving disk is (e.g., \citealt{armitage10})
 \begin{equation}
@@
 
-For simplicity, we calculate the radial drift timescale, $t_{\rm drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well.
+For simplicity, we calculate the radial drift timescale, $t_{\rm drift}$, for an irradiated disk in this section, but most of our conclusions hold true for an evolving disk as well.
 Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk.
@@
 On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas.
-Thus for small particles in a passive disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}).
-This is not true for an active disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.
+Thus for small particles in an irradiated disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}).
+This is not true for an evolving disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.
 At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well.
-This is true for both passive and active disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.
+This is true for both irradiated and evolving disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.
 
@@
 %\vspace{-0.5in}
-\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both a passive disk (left panels) and an active disk (right panels).
+\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both an irradiated disk (left panels) and an evolving disk (right panels).
 The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
@@
 \caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text).
-The estimate is performed for a passive disk (left panels) and an active disk (right panels).
+The estimate is performed for an irradiated disk (left panels) and an evolving disk (right panels).
 The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
@@
 
-In this section we use the model described in section \ref{sec:model} to quantify the effects of radial drift (passive disk) or radial drift and gas accretion (active disk) on the snowline location, for dust particles of different sizes composed of either H$_2$O, CO$_2$ or CO.
+In this section we use the model described in section \ref{sec:model} to quantify the effects of radial drift (irradiated disk) or radial drift and gas accretion (evolving disk) on the snowline location, for dust particles of different sizes composed of either H$_2$O, CO$_2$ or CO.
 Specifically, we determine a particle's final location (i.e., where the particle either fully desorbs or remains at its initial size due to having a desorption timescale longer than the time at which we stop the simulation) as a function of its initial position in the disk, after the gas disk has dissipated.
@@
 \end{subeqnarray}
-where the desorption rate $R_{\rm des, x}$ for each particle type (i.e., composed of H$_2$O, CO$_2$ or CO) is evaluated at $T=T(r)$, and the radial drift velocity $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for a passive disk and Equation (\ref{eq:rdotact}) for an active disk.
+where the desorption rate $R_{\rm des, x}$ for each particle type (i.e., composed of H$_2$O, CO$_2$ or CO) is evaluated at $T=T(r)$, and the radial drift velocity $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for an irradiated disk and Equation (\ref{eq:rdotact}) for an evolving disk.
 Equations (\ref{eq:dsdt}) and (\ref{eq:drdt}) describe the coupled desorption and radial drift, and can be derived straightforwardly from Equation (\ref{eq:tdes}).
@@
 We define the final position of a grain as the disk location it has reached after $t_{\rm d}=3$ Myr, or the radius at which it completely desorbs if that happens after a time shorter than 3 Myr.
-Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both a passive and an active disk.
-The results for the accretionally heated active disk are quantitatively similar with those of the active disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
+Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both an irradiated and an evolving disk.
+The results for the viscous disk are quantitatively similar with those of the evolving disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
 We also show the static snowlines for comparison, which are calculated by balancing adsorption and desorption \citep{hollenbach09}.
 %We do not show the equivalent result for the steady-state active disk since The trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state active disk results in Section \ref{sec:COratio}.
-Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for a passive nor for an active disk.
-Similarly, micron- to mm-sized particles in the passive disk do not drift and only %or
+Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for an irradiated nor for an evolving disk.
+Similarly, micron- to mm-sized particles in the irradiated disk do not drift and only %or
 desorb if %unless 
@@
 [don't think this is needed, but Til or Ruth might think otherwise]
-In an active disk, however, micron-to mm-sized grains do drift significantly since they move at the same velocity as the accreting gas.
-For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in a passive disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an active disk, we notice that particles of initial size $s_0$ desorb at a %fixed distance
+In an evolving disk, however, micron-to mm-sized grains do drift significantly since they move at the same velocity as the accreting gas.
+For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in an irradiated disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an evolving disk, we notice that particles of initial size $s_0$ desorb at a %fixed distance
 particle size dependent radius $r_{\rm des}$ regardless of their original location in the disk.
@@
 %\vspace{-0.5in}
-\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in a passive disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left).
+\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in an irradiated disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left).
 Particles desorb almost instantaneously.
@@
 Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically.
-We display this result for the range of particle sizes that desorb at a fixed distance in a passive and an active disk (see Figure \ref{fig:snowlines}).
+We display this result for the range of particle sizes that desorb at a fixed distance in an irradiated and an evolving disk (see Figure \ref{fig:snowlines}).
 We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it deviates for particles larger than $s \gtrsim10$ cm.
-For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the passive disk in the Epstein drag regime), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
+For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the irradiated disk in the Epstein drag regime), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
 Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
@@
 Given $r_{\rm des}$, we need to only calculate the distance over which particles desorb to determine the location of a snowline.
-Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk.
+Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in an irradiated disk.
 Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
@@
 Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active)\footnote{Both of these conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
+It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (irradiated or evolving)\footnote{Both of these conclusions remain valid for an evolving disk and for particles composed of CO$_2$ or CO.}.
 The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
@@
 %\vspace{-0.5in}
-\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for a passive disk (top panel), an active disk (middle panel) and an active disk accretionally heated (bottom panel), for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
+\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for an irradiated disk (top panel), an evolving disk (middle panel) and a viscous disk (bottom panel), for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
 The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
@@
 Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline.
-For both active disks, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, with the decrease more significant interior to the H$_2$O snowline.}
+For both the evolving and the viscous disk, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, with the decrease more significant interior to the H$_2$O snowline.}
 \label{fig:CO_ratio}
@@
 
-Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and an active disk accretionally heated, under the simplifying assumption that the abundance relative to hydrogen for each volatile is \textbf{initially} fixed, so that drift and accretion affect only the locations of the snowlines.
+Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for an irradiated disk, an evolving disk, and a viscous disk, under the simplifying assumption that the abundance relative to hydrogen for each volatile is \textbf{initially} fixed, so that drift and accretion affect only the locations of the snowlines.
 We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
@@
 
-Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the active disks ignore the movement of the desorbed ices with the accreting gas---the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
+Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the evolving and viscous disks ignore the movement of the desorbed ices with the accreting gas---the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
 As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
@@
 We have found that the largest drifting particles in our model ($\sim$7 m) drift faster than the gas at both the H$_2$O and CO$_2$ snowlines.
-We thus conclude that the C/O ratio interior to the H$_2$O and CO$_2$ snowlines in our active disks will be lower than in the static disk, due to the additional oxygen added to the gas by desorbing H$_2$O and CO$_2$.
+We thus conclude that the C/O ratio interior to the H$_2$O and CO$_2$ snowlines in our evolving and viscous disks will be lower than in the static disk, due to the additional oxygen added to the gas by desorbing H$_2$O and CO$_2$.
 For these particle sizes, our calculated C/O ratio is an upper limit,  as indicated by the arrows in Figure \ref{fig:CO_ratio}.
@@
 The snowline locations in these disks exhibit several interesting features.
-For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus move the snowline compared to the static disk.
-In contrast, even $\sim$micron-sized grains drift and desorb for the active disk, since they flow towards the host star together with the accreting gas.
-For the same particle size, the snowline locations are slightly closer to the central star in the active disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
+For the irradiated disk, only grains larger than $\sim$0.5 cm drift, desorb and thus move the snowline compared to the static disk.
+In contrast, even $\sim$micron-sized grains drift and desorb for the evolving disk, since they flow towards the host star together with the accreting gas.
+For the same particle size, the snowline locations are slightly closer to the central star in the evolving disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
 Equation \ref{eq:rdotact}).
-The addition of accretional heating in the steady-state active disk moves the H$_2$O snowline outwards.
+The addition of accretional heating in the viscous disk moves the H$_2$O snowline outwards.
 This is due to the fact that accretional heating dominates in the inner disk, where high temperatures cause the grains to evaporate further away from the star.
-Once $r\gtrsim1-2$  AU, stellar irradiation dominates the thermal evolution of the disk, and therefore the CO$_2$ and CO snowlines locations are the same as in the active disk.
+Once $r\gtrsim1-2$  AU, stellar irradiation dominates the thermal evolution of the disk, and therefore the CO$_2$ and CO snowlines locations are the same as in the evolving and viscous disks.
 
@@
 Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
-%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
+\textbf{For a grain size distribution with a maximum particle size different than our model, one can pick out the appropriate minimum snowline locations from this plot.}%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
 Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from this plot.
@@
 
-For our choice of parameters, the minimum snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the passive disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the active disk and $r_{\rm{H_2O}} \approx 0.63$ AU for the active disk accretionally heated; $r_{\rm{CO_2}} \approx 3.7$ AU for the passive disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both active disks; $r_{\rm{CO}} \approx 30$ AU for the passive and both active disks.
+For our choice of parameters, the minimum snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the irradiated disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the evolving disk and $r_{\rm{H_2O}} \approx 0.63$ AU for the viscous disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the irradiated disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both the evolving and the viscous disks; $r_{\rm{CO}} \approx 30$ AU for the irradiated and both the evolving and the viscous disks.
 For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU\footnote{\textbf{For the viscous disk, we calculated the static snowline location using the same temperature profile as that of the viscous disk, for consistency purposes.
@@
 Observations of the CO snowline in TW Hya \citep{qi13} have found its location at a disk midplane temperature of 17 K (at 30 AU for the TW Hya specific temperature profile).
-The inferred  desorption temperature corresponds to the CO desorption temperature in a static disk, or to desorption from very small grains in an active disk, i.e.
+The inferred  desorption temperature corresponds to the CO desorption temperature in a static disk, or to desorption from very small grains in an evolving disk, i.e.
 from grains that are too small to drift substantially.
@@
 These sizes are important since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
-We choose the active disk without accretion heating as a disk model, and we stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively.
+We choose the evolving disk as a disk model, and we stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively.
 %, to see how the snowline locations evolve with time and affect the C/O ratio.
@@
 Figure \ref{fig:varMd} shows the desorption or final distance as a function on the initial location of a H$_2$O particle with initial size of 1 m, for two total disk masses: $M=0.1 M_{\odot}$, our fiducial model, and $M=0.01 M_{\odot}$.
-Similarly to Figure \ref{fig:timeplots}, we perform our calculations for an active disk without accretion heating.
+Similarly to Figure \ref{fig:timeplots}, we perform our calculations for an evolving disk.
 The simulations are stopped after the same timeframes as those in Figure \ref{fig:timeplots}.
@@
 
-We also apply our active disk model to a transition disk, i.e.
+We also apply our evolving disk model to a transition disk, i.e.
 a protoplanetary disk with an inner cavity significantly depleted of gas.
@@
 Gas accretion & $\leftarrow$ \footnote{Gas accretion pushes the snowlines inwards compared to the snowline locations in a static disk.
-However, accretional heating may push the snowline outwards compared to an active disk without accretion heating.} \\\hline
+However, accretional heating may push the snowline outwards compared to an evolving disk without accretion heating.} \\\hline
 Particle growth & $\rightarrow$ $\leftarrow$ \\\hline
@@
 
-\item \textbf{Accretion rate evolution.} Our steady-state active disk model assumes a constant mass accretion rate $\dot{M}$.
+\item \textbf{Accretion rate evolution.} Our viscous disk model assumes a constant mass accretion rate $\dot{M}$.
 However, $\dot{M}$ decreases over time, which lowers the accretional component of the disk temperature (Equation \ref{eq:Tdact}), thus pushing the snowline location inwards if the disk is optically thick \citep{garaud07}.
@@
 We assume that the solid particles are perfect, homogeneous spheres, fully composed of either H$_2$O, CO$_2$ or CO.
-We apply our model to a passive disk, an active disk with a passive temperature profile (hereafter active disk), and a steady-state active disk that also takes into account stellar irradiation.
+We apply our model to an irradiated disk, an evolving disk, and a viscous disk that also takes into account stellar irradiation.
 We determine the desorption or final location of drifting particles after a time equal to the disk lifetime, and use this result to set an inner limit for the location of the H$_2$O, CO$_2$ and CO snowlines.
@@
 \begin{enumerate}
-\item Radial drift and gas accretion affect desorption and move the snowline locations compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk.
+\item Radial drift and gas accretion affect desorption and move the snowline locations compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk.
 %Moreover, boulders up to almost kilometer-size contribute to setting the snowline locations if they are very close to the snowlines.
@@
 Gas accretion causes even micron-sized particles to drift, desorb and move the snowline location compared to a static disk.
-A steady-state disk that includes accretion heating moves the H$_2$O snowline outwards compared to an active disk, but has no effect on the CO$_2$ and CO snowline locations, for our particular choice of mass accretion rate $\dot{M}$ and midplane opacity $\kappa$.
+A viscous disk that includes accretion heating moves the H$_2$O snowline outwards compared to an evolving disk, but has no effect on the CO$_2$ and CO snowline locations, for our particular choice of mass accretion rate $\dot{M}$ and midplane opacity $\kappa$.
 %However, the magnitude of the effect of accretional heating on snowline locations will depend on .
 
-\item For our fiducial model, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for a passive disk, 0.26 AU, 3.4 AU and 30 AU for an active disk, and 0.63 AU, 3.4 AU and 30 AU for a steady-state active disk with accretion heating.
+\item For our fiducial model, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for an irradiated disk, 0.26 AU, 3.4 AU and 30 AU for an evolving disk, and 0.63 AU, 3.4 AU and 30 AU for a viscous disk with accretion heating.
 Compared to a static disk, radial drift and gas accretion move the snowlines by up to 60 \% for H$_2$O and CO$_2$, and by up to 50 \% for CO.
@@
 This is consistent with possible detections of superstellar C/O ratios in some exoplanet atmospheres.
-We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the active disks.
+We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the evolving and viscous disks.
 \textbf{We plan to address this issue in a future paper.}
@@
 
-For a particle of size $s$ that desorbs and satisfies $\tau_{\rm s} \ll 1$ ($\tau_{\rm s}$ is the dimensionless stopping time, defined in Section \ref{sec:drift}), we can derive an explicit analytic solution for the particle's desorption distance in a passive disk.
+For a particle of size $s$ that desorbs and satisfies $\tau_{\rm s} \ll 1$ ($\tau_{\rm s}$ is the dimensionless stopping time, defined in Section \ref{sec:drift}), we can derive an explicit analytic solution for the particle's desorption distance in an irradiated disk.
 For $\tau_{\rm s} \ll 1$, a particle is in the Epstein drag regime (see Equation \ref{eq:ts}) and its drift velocity $\dot{r}$ (Equation \ref{eq:rdotpas}) can be approximated as

### 2015-10-06 tex/draft/ms_final_revised.tex
update

@@
 In the inner portion of our disk ($r  \lesssim 1$ AU for our fiducial model with $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$), our choice of opacity (Equation \ref{eq:opacity}) implies that the disk has a constant surface density with radius (see Equations \ref{eq:diskeq}).
+
+\textbf{Before we proceed forward, we note that our disk models assume a constant stellar luminosity $L_*$, as well as a constant mass accretion rate $\dot{M}$ for the viscous disk.
+In reality, the stellar luminosity decreases as the host star contracts, which will reduce the disk temperature and push the snowlines inward, as we explain in Section \ref{sec:neglected}.
+For a Solar type star, as our fiducial model assumes, $L_*$ remains relatively constant during the star's pre-main sequence evolution of $\sim$ 10 Myr \citep{kennedy06}, which is larger than the giant planet formation timescale.
+Thus the midplane temperature will not change significantly for our model due to variations in stellar luminosity, but it may decrease substantially for smaller stars, pushing the snowline inward (see Section \ref{sec:neglected} for details).
+Realistic mass accretion rates, $\dot{M}$, may vary between $\sim$$10^{-7}$ and $\sim$$10^{-9}$ $M_{\odot}$ yr$^{-1}$ as the disk evolves (e.g., \citealt{garaud07}).
+For $\dot{M}\lesssim 10^{-9} M_{\odot}$ yr$^{-1}$, the disk becomes optically thin and hence depleted of gas, which means giant planets must have formed before $\dot{M}$ becomes too low.
+\citet{garaud07} find that the snowline locations scale as $r_{\rm snow} \propto \dot{M}^{1/3}$.
+A factor of 100 reduction in the mass accretion rate will thus move the snowlines inwards by a factor of $\sim$$4$.
+However, \citet{garaud07} find that the mass accretion rate is close to $10^{-8} M_{\odot}$ yr$^{-1}$ during the quasi-static accretion phase of the disk (see also Section \ref{sec:neglected}), hence the snowline locations are likely to move inward only in the very early stages of the disk evolution.
+We thus claim that changes in $L_*$ and $\dot{M}$ throughout time may only modestly affect our results, but they should still be taken into account when drawing more robust conclusions, as well as for different host star and disk properties.}
+
  %We note that $\Sigma$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.
@@
 \item \textbf{Stellar luminosity evolution.} As the host star contracts during its pre-main sequence phase, its luminosity decreases, which reduces the disk temperature and pushes the snowline locations inwards.
-\citet{kennedy06} found that the snowline may move inward by a factor of $\sim$$15-20$ due to the stellar contraction.
+\citet{kennedy06} found that the snowline \textbf{is unlikely to move significantly during the pre-main sequence phase for Solar type stars, but it} may move inward by a factor of $\sim$$15-20$ \textbf{for $M_* \sim 0.25 M_{\odot}$} due to the stellar contraction.
 

### 2015-10-09 Postdoc Applications/AnaMariaPisoResume.tex
update

@@
 \begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+MIT Exoplanet Tea, Cambridge, MA, October 2015 \\
+Seminar speaker \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+Center for Integrative Planetary Science Planet and Star Formation Seminar, Berkeley, CA, September 2015 \\
+Seminar speaker \\
+\end{list2}
+
+\begin{list2}
 \item[] {\bf Minimum Core Masses for Giant Planet Formation} \\

### 2015-10-09 tex/draft/ms_final_revised.tex
update

@@
 We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk.
-We determine the snowline locations for a range of fixed initial particle sizes, in both \textbf{evolving (active)} and \textbf{irradiated (passive) disks}.
-\textbf{For our fiducial disk model,} we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk desorb %instantaneously and [I think instantaneously is too strong]
+We determine the snowline locations for a range of fixed initial particle sizes \textbf{and disk types}.
+\textbf{For our fiducial disk model,} we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk, and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving and viscous disk, desorb %instantaneously and [I think instantaneously is too strong]
 at a %fixed 
@@
 %location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
-Compared to a static disk, we find that radial drift and gas accretion \textbf{in a viscous disk} move the H$_2$O snowline inwards by up to \textbf{40 \%}, the CO$_2$ snowlines by up to 60 \%, and the CO snowline by up to 50 \%.
+Compared to a static disk, we find that radial drift and gas accretion \textbf{in a viscous disk} move the H$_2$O snowline inwards by up to \textbf{40 \%}, the CO$_2$ snowline by up to 60 \%, and the CO snowline by up to 50 \%.
 We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
@@
 
-We present our protoplanetary disk model for \textbf{an irradiated, an evolving, and a viscous disk} in section \ref{sec:disk}.
+We present our protoplanetary disk model for \textbf{a static, an irradiated, an evolving, and a viscous disk} in section \ref{sec:disk}.
 In section \ref{sec:drift}, we describe our analytic model for the radial drift of solids.
@@
 
-To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{static disk}, which is solely irradiated by the host star and does not take into account gas accretion onto the star or radial drift; \textit{\textbf{irradiated} disk}, which has the same temperature profile as the static disk and does not experience gas accretion or accretional heating, but it takes into account radial drift of solids; \textit{\textbf{evolving} disk}, in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; and \textit{\textbf{viscous} disk}, for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis and the temperature profile is calculated using both accretional heating and stellar irradiation.
+To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{static disk}, which is solely irradiated by the host star and does not take into account gas accretion onto the star or radial drift; \textit{\textbf{irradiated} disk}, which has the same temperature profile as the static disk and does not experience gas accretion or accretional heating, but it takes into account radial drift of solids; \textit{\textbf{evolving} disk}, in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; and \textit{\textbf{viscous} disk}, for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis, and the temperature profile is calculated using both accretional heating and stellar irradiation.
 
@@
 
-\textbf{Irradiated disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a irradiated disk similar to the prescription of \citet{chiang10}.
+\textbf{Static and Irradiated disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a static and and an irradiated disk similar to the prescription of \citet{chiang10}.
 The gas surface density and midplane temperature are
@@
 Our choice is inspired by observations of protoplanetary disks at radii larger than $\sim$20 AU (e.g., \citealt{andrews10}), which suggest that typical disks may have surface density profiles with $\Sigma \propto r^{-1}$.
+\textbf{A slope flatter than $\Sigma \propto r^{-3/2}$ is also more consistent with the temperature profile for a steady-state gas disk (see the Viscous disk heading below and \App{app:steadystate}).} We use the static disk model to compare our results with those of \citet{oberg11}.
 
@@
 \label{eq:Sigmaact}
-\Sigma(R, \overline T) = \frac{M}{2 \pi r_{\rm c}^2 R^{\gamma}} \overline T^{-(5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{(2-\gamma)}}{\overline T}}\Big],
+\Sigma(R, \tilde t) = \frac{M (2 - \gamma)}{2 \pi r_{\rm c}^2 R^{\gamma}} \tilde t^{-(5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{(2-\gamma)}}{\tilde t}}\Big],
 \end{equation}
@@
 \label{eq:T}
-\overline T & \equiv & \frac{t}{t_{\rm c}} + 1 \\
+\tilde t & \equiv & \frac{t}{t_{\rm c}} + 1 \\
 t_{\rm c} & \equiv & \frac{1}{3(2-\gamma)^2} \frac{r_{\rm c}^2}{\nu_{\rm c}},
 \end{subeqnarray}
-where $t$ is time. We choose $M=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
+where $t$ is time.
+We choose $M=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
+\textbf{The irradiated and evolving disk surface densities match at $t \approx 5 \times 10^4$ years in the inner disk, but they diverge at distances larger than a few AU due the exponential cutoff in radius of the surface density of the evolving disk (Equation \ref{eq:Sigmaact}).}
 
@@
 Since
-Calculating the midplane temperature self-consistently for an evolving disk is non-trivial.
-We thus use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm act}$.
+Calculating the midplane temperature self-consistently for an evolving disk \textbf{that is also actively heated, and thus whose thermal evolution is dominated both by accretion heating and stellar irradiation}, is non-trivial.
+We therefore use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm act}$.
 The equations governing the evolution of the steady-state disk are listed in \App{app:steadystate}.
@@
 \label{eq:activeT}
-T^4 = T_{\rm act}^4 + T_{\rm pas}^4,
-\end{equation}
-where to avoid notation confusion $T_{\rm pas}=T$ from Equation (\ref{eq:diskT}), the temperature profile for an irradiated disk.
+T^4 = T_{\rm act}^4 + T_{\rm irr}^4,
+\end{equation}
+where to avoid notation confusion $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}), the temperature profile for an irradiated disk.
 We can then easily determine  $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
@@
 
-\textbf{Static disk.} To compare our results with those of \citet{oberg11}, we also use a static disk model, with $\Sigma$ and $T$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}).
  %The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
@@
 We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T$.
-For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively (\citealt{oberg11} and references therein).
+For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively (\textbf{\citealt{collings04}, \citealt{fraser01}, \citealt{aikawa96}}).
 We use the desorption rate, $R_{\rm des}$, to estimate the desorption timescale for particles of different sizes as described in section \ref{sec:timescales}.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{drift_timescales_betaS1_gas_acc_new.pdf}
+\includegraphics[width=0.5\textwidth]{drift_timescales_betaS1_gas_acc_new2.pdf}
 %\vspace{-0.5in}
@@
 
-Of particular interest for our purposes is the particle size regime for which (1) $t_{\rm drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
+\textbf{In the particle size regime} for which (1) $t_{\rm drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
 for $\sim$$0.5$ cm $\lesssim s \lesssim$ $1000$ cm, or (2) $t_{\rm gas, acc} \lesssim t_{\rm des} \lesssim t_{\rm d}$, i.e.
-for $\sim$$0.1$ cm $\lesssim s \lesssim$ $10$ cm.
-In these cases, radial drift or gas accretion (or both) are faster than thermal desorption.
-We note that $t_{\rm gas, acc}<t_{\rm d}$ always holds true.
+for $\sim$$0.1$ cm $\lesssim s \lesssim$ $10$ cm, radial drift or gas accretion (or both) are faster than thermal desorption, \textbf{ which is of particular interest for our purposes.} We note that $t_{\rm gas, acc}<t_{\rm d}$ always holds true.
 Particles of sizes that satisfy the requirements above will drift significantly due to radial drift or gas accretion before desorbing, thus moving the H$_2$O, CO$_2$ and CO snowlines closer towards the central star and changing the C/O ratio throughout the disk.
@@
 Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both an irradiated and an evolving disk.
-The results for the viscous disk are quantitatively similar with those of the evolving disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
+\textbf{We do not show the results for the viscous disk as they would complicate the plot without adding any qualitative insight --- } the results for the viscous disk are quantitatively similar with those of the evolving disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
 We also show the static snowlines for comparison, which are calculated by balancing adsorption and desorption \citep{hollenbach09}.
@@
 Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (irradiated or evolving)\footnote{Both of these conclusions remain valid for an evolving disk and for particles composed of CO$_2$ or CO.}.
-The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
+It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model.
+\textbf{Both of these conclusions remain valid for an evolving and a viscous disk, as well as for particles composed of CO$_2$ or CO, but the snowline locations will vary between the three disks for a given initial particle size (see Section \ref{sec:COratio}).} \textbf{If we do not take into account the time dependence of the mass accretion rate and stellar luminosity (see Section \ref{sec:disk}),} the C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
 %\emgr{I've yet to figure out why there is a discrepancy for the small grains.
@@
 %\vspace{-0.5in}
-\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for an irradiated disk (top panel), an evolving disk (middle panel) and a viscous disk (bottom panel), for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk.
+\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for an irradiated disk (top panel), an evolving disk (middle panel) and a viscous disk (bottom panel).
+%, for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk. 
 The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
 The horizontal dotted line represents the stellar value of 0.54.
-The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk, with the temperature profile given by Equation (\ref{eq:diskT}) for the top two panels and by Equation (\ref{eq:activeT}).
-The snowline location moves inward as the particle size increases.
+The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk, with the temperature profile given by Equation (\ref{eq:diskT}) for the top two panels and by Equation (\ref{eq:activeT}) for the bottom panel.
+%The snowline location moves inward as the particle size increases. 
 The curves exterior to the static snowline represent desorption distances in the absence of readsorption.
-Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline.
-For both the evolving and the viscous disk, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, with the decrease more significant interior to the H$_2$O snowline.}
+%Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline.
+For both the evolving and the viscous disk, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, as shown by the arrows.}
+%, with the decrease more significant interior to the H$_2$O snowline.}
 \label{fig:CO_ratio}
@@
 The minimum snowline distance for a disk is given by the curve corresponding to the maximum particle size it hosts.
-If grains have grown to radii larger than $\sim$7 m, then the 7m snowline applies (see Section 3).
+\textbf{For} grains that have grown to radii larger than $\sim$7 m \textbf{and are able to drift and desorb}, the $\sim$7 m snowline applies (see Section 3).
 %Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.
@@
 We note that the H$_2$O snowline in all disks is significantly closer to the host star compared with Solar system models, which place the H$_2$O snowline between $\sim$$2.7$ to $\sim$$3.1$ AU (\citealt{hayashi81}, \citealt{podolak04}, \citealt{martin12}).
-This is partially because we choose a colder disk model, as well as the fact that gas accretion rates decrease over time, moving the snowline location outwards (see also \citealt{garaud07} and Section \ref{sec:neglected}).
+This is partially because we choose a colder disk model, as well as the fact that gas accretion rates decrease over time, moving the snowline location \textbf{inwards} (see also \citealt{garaud07} and Section \ref{sec:neglected}).
+\textbf{\citet{min11} find that the location of the H$_2$O snowline is highly sensitive to the gas mass accretion rate $\dot{M}$ (equal to $10^{-8} M_{\odot}$ yr$^{-1}$ in our model) and the dust opacity $\kappa$ (Equation \ref{eq:opacity}).
+Higher values of $\dot{M}$ and $\kappa$ would increase the accretional component of the disk temperature (cf.
+Equation \ref{eq:Tdact}), which would push the H$_2$O snowline in the viscous disk outwards to match the Solar system snowline.
+At the same time, the snowline location in Solar type stars may be as close as $\sim$1 AU \citep{mulders15}, further in than the H$_2$O snowline in our Solar system.}
+
  Observations of the CO snowline in TW Hya \citep{qi13} have found its location at a disk midplane temperature of 17 K (at 30 AU for the TW Hya specific temperature profile).
@@
 All previous results assumed a disk lifetime $t_{\rm d}=3$ Myr%is the most representative for our calculations, as this is the timeframe in which giant planets accrete their gaseous atmospheres 
-, the typical disk life time and the expected time scale for giant planets to accrete their gaseous atmopsheres (e.g., \citealt{pollack96}, \citealt{piso14}).
+, the typical disk lifetime and the expected time scale for giant planets to accrete their gaseous atmopsheres (e.g., \citealt{pollack96}, \citealt{piso14}).
 Some gas accretion may occur at earlier times, however,%However, protoplanetary cores start acquiring their envelope at earlier times,
@@
 The C/O ratio is thus insensitive to the choice of $M$.
+\textbf{We note that our conclusions regarding the disk age and total mass are only valid if the snowline itself does not move with time or disk mass (see also Sections \ref{sec:disk} and \ref{sec:neglected}).}
+
 %This result also has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
@@
 However, once the solids grow larger than km-sized and form planetesimals, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
-It follows that grain growth will eventually push the snowline location outwards.
+%It follows that grain growth will eventually push the snowline location outwards.
 
@@
 However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
-Turbulence causes eddies and vertical mixing, which are likely to reduce the radial drift velocity of the solids.
+Turbulence causes eddies and vertical mixing, which are likely to reduce the radial drift velocity of the solids \textbf{(e.g., \citealt{youdin07})}.
 Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially.
@@
 However, accretional heating may push the snowline outwards compared to an evolving disk without accretion heating.} \\\hline
-Particle growth & $\rightarrow$ $\leftarrow$ \\\hline
+Particle growth & $\leftarrow$ \footnote{\textbf{As stated in the main text, if particles grow to km-sizes and above, the snowline remains the same as that of a static disk.}} \\\hline
 Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
@@
 The inhomogeneity due to cracks in the grain structure will cause the particles to desorb faster.
-They will therefore drift less before evaporating and will move the snowlines outwards.
+They will therefore drift less before evaporating and will move the snowlines \textbf{less far inward}.
 
@@
 
-\item \textbf{Disk gaps and holes.} The snowline locations will be different for transition disks, which have inner cavities significantly depleted of gas (e.g., \citealt{espaillat12}), or pre-transitional disks, which have a gap between an inner and outer full disk (e.g., \citealt{kraus11}).
+\item \textbf{Disk gaps and holes.} The snowline locations will be different for transition disks, which have inner cavities significantly depleted of gas (e.g., \citealt{espaillat12}, \textbf{\citealt{vandermarel15}}), or pre-transitional disks, which have a gap between an inner and outer full disk (e.g., \citealt{kraus11}).
 The decrease in gas pressure in these gaps or holes will reduce the particles' drift velocity close to the gap edge, thus slowing them down and pushing the snowline outwards.
@@
 \begin{enumerate}
-\item Radial drift and gas accretion affect desorption and move the snowline locations compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk.
+\item Radial drift and gas accretion affect desorption and move the snowline locations \textbf{inward} compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk.
 %Moreover, boulders up to almost kilometer-size contribute to setting the snowline locations if they are very close to the snowlines.
@@
 
-\item Particles with sizes in the above range desorb almost instantaneously once desorption has begun, and at a fixed location in the disk that only depends on the particle size and the gas accretion rate.
+\item \textbf{For our simplified model that does not account for the effects outlined in Section \ref{sec:neglected},} particles with sizes in the above range desorb almost instantaneously once desorption has begun, and at a fixed location in the disk that only depends on the particle size and the gas accretion rate.
 Thus for each particle size there is a fixed and uniquely determined H$_2$O, CO$_2$ or CO snowline.
 
-\item The results of the numerical simulation are in agreement with the analytic solution of the drift-desorption system of differential equations if the stopping time $\tau_{\rm s} \ll 1$.
+\item The results of \textbf{our} numerical simulation are in agreement with the analytic solution of the drift-desorption system of differential equations if the stopping time $\tau_{\rm s} \ll 1$.
 We present an explicit analytic solution for the desorption distance in this regime.
@@
 
-\item For our fiducial model, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for an irradiated disk, 0.26 AU, 3.4 AU and 30 AU for an evolving disk, and 0.63 AU, 3.4 AU and 30 AU for a viscous disk with accretion heating.
+\item For our fiducial model, \textbf{which considers particles with sizes between $10^{-3}$ and $10^8$ cm}, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for an irradiated disk, 0.26 AU, 3.4 AU and 30 AU for an evolving disk, and 0.63 AU, 3.4 AU and 30 AU for a viscous disk with accretion heating.
 Compared to a static disk, radial drift and gas accretion move the snowlines by up to 60 \% for H$_2$O and CO$_2$, and by up to 50 \% for CO.
-
-\item Our C/O estimates find that the C/O ratio may be enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching its maximum value between the CO$_2$ and CO snowlines.
-This is consistent with possible detections of superstellar C/O ratios in some exoplanet atmospheres.
-We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the evolving and viscous disks.
-\textbf{We plan to address this issue in a future paper.}
-
-\item The snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk midplane remains optically thick.
+\textbf{For the viscous disk, however, which is the most realistic of the three models since it takes into account accretion heating, the H$_2$O snowline location moves inwards by up to 40 \%.}
+
+\item Our C/O estimates \textbf{confirm the conclusions of \citet{oberg11}} that the C/O ratio \textbf{in gas} may be enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching its maximum value between the CO$_2$ and CO snowlines.
+%This is consistent with possible detections of superstellar C/O ratios in some exoplanet atmospheres.
+We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the evolving and viscous disks, \textbf{which may significantly decrease the C/O ratio in gas inside the H$_2$O and CO$_2$ snowlines.} \textbf{We plan to address this issue in a future paper.}
+
+\item \textbf{For a constant gas mass accretion rate $\dot{M}$ and stellar luminosity $L_*$,} the snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk midplane remains optically thick.
 %Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.

### 2015-10-21 tex/draft/ms_final_revised.tex
update

@@
 \centering
-\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new.pdf}
+\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
 %\vspace{-0.5in}
@@
 %, for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk. 
-The particle size increases from 0.001 cm to $\sim$700 cm as indicated by the color bar.
+The particle size increases from \textbf{$\sim$0.05 cm} to $\sim$700 cm as indicated by the color bar.
 The horizontal dotted line represents the stellar value of 0.54.
@@
 %The snowline location moves inward as the particle size increases. 
-The curves exterior to the static snowline represent desorption distances in the absence of readsorption.
+%The curves exterior to the static snowline represent desorption distances in the absence of readsorption.
 %Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline.
@@
 The C/O ratio for a static disk, where desorption and readsorption balance \citep{hollenbach09}, is shown as a guideline.
-We note that the true snowline for particles with $r_{\rm des}$ outside the static snowline is the static snowline itself---curves plotted at larger distances are not true snowlines.
+We note that the true snowline for particles with $r_{\rm des}$ outside the static snowline is the static snowline itself---\textbf{thus only particles with initial sizes larger than $\sim$0.05 cm are plotted in the three panels, as particles that form snowlines at larger distances (cf.
+Figure \ref{fig:snowlines}) are not true snowlines}.
 
@@
 
-Figure \ref{fig:CO_ratio} plots snowline curves for particle sizes less than $\sim$ 7m.
+Figure \ref{fig:CO_ratio} plots snowline curves for particle sizes \textbf{$\sim0.5$ cm $\lesssim $s$ \lesssim$ 7m}.
 In the outermost disk, H$_2$O, CO$_2$, and CO all solidify.

### 2015-11-04 Postdoc Applications/hubble_proposal.tex
update

%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
%\documentclass[12pt, preprint]{aastex}
%\documentclass[apj]{emulateapj}
\documentclass[12pt, letterpaper]{article}

%\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
%\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex

%\usepackage{pdfsync}
%\usepackage{subeqnarray}
\usepackage[top=0.9in, bottom=0.8in, left=1in, right=1in]{geometry}
\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage{titling}
\usepackage{sectsty}
\usepackage{sidecap}
\usepackage{placeins}
\usepackage{indentfirst}
\setlength{\droptitle}{-7em}
\setlength{\abovecaptionskip}{-0.2ex}
\setlength{\belowcaptionskip}{-0.2ex}
%\pagenumbering{gobble}
\pagestyle{fancy}
\lhead{Ana-Maria Piso}
\rhead{Hubble Fellowship Research Proposal}
\date{}
%\rhead{\thepage}

%\bibliographystyle{apj}

\title{Hubble Fellowship Research Statement: \\
The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets}
\author{Ana-Maria Piso}

%\newenvironment{packed_item}{
%\begin{itemize}
%  \setlength{\itemsep}{1pt}
%  \setlength{\parskip}{0pt}
%  \setlength{\parsep}{0pt}
%}{\end{itemize}}

\begin{document}
\maketitle

%\slugcomment{Draft Modified \today}

\vspace{-0.9cm}

Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered $[2]$. Their diversity in terms of mass, radius, location and composition $[6]$ provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore and understand how planets obtain their compositions. Observations of Earth-like planets that can provide useful insight about their composition are challenging --- the solid interior structure of terrestrial planets cannot be detected, and their gaseous envelopes are small by comparison (both in mass and radius), which makes it difficult to obtain atmospheric spectra and find out what chemical compounds they are made of. We therefore turn to giant planets, which have provided a rich and intriguing research area for decades. Gas giants contain most of their mass in their atmosphere, hence their chemical composition is determined by that of their envelopes. The last few years have seen a substantial increase in the number of giant planets with observed atmospheric spectra (e.g., $[3]$, $[5]$), which has enhanced our understanding of these planets' chemical structure, and has provided us with quantitative information about the abundances of various compounds in their envelopes besides hydrogen and helium. Finally, gas giants shape the architecture of planetary systems and affect the delivery of volatile compunds to terrestrial planets, which has direct consequences for the habitability of worlds similar to our own.

Both terrestrial and giant planets are born in protoplanetary disks, which implies that their \textbf{composition is determined by and tightly linked to the structure and composition of the disk}. The chemical and dynamical evolution of disks, as well the formation of giant planets have both been previously investigated. However, the coupled chemo-dynamical disk evolution, planet compositions, and most importantly the disk-planet connection have not yet been considered in detail. As shown in my work on the minimum core mass of gas giants ($[8]$, $[9]$), planet formation depends sensitively on disk physics and chemistry. \textbf{I propose to develop a holistic chemo-dynamical framework to explore how disk dynamics and chemistry, as well as the dynamics of nascent planets and planetesimals, regulate the compositions of mature giant planets.} Such a model will enhance our understanding of planetary structures by enabling us to predict what kind of planet compositions result from planet formation in different parts of the disk. Furthermore, this work provides essential context for characterizing the gas giants that instruments such as the James Webb Space Telescope (JWST) and the Transiting Exoplanet Survey Satellite (TESS) will one day discover. Characterizing the composition of planetary atmospheres is therefore an essential component for understanding the observations of NASA's Cosmic Origins Program. 
\vspace{0.15in}

%\section{Coupled Chemical and Dynamical Disk Evolution} 
\underline{\textbf{1. Coupled Chemical and Dynamical Disk Evolution}}

Chemical and dynamical processes in a protoplanetary disk affect the disk structure and composition, and thus the composition of nascent planets. As shown in Figure 1, the timescales for various disk chemical and dynamical processes may be comparable at least in some parts of the disk. It follows that these two effects cannot be decoupled, and that the chemo-dynamical evolution of the disk has to be studied simultaneously and self-consistently. Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research $[10]$. Such effects include \textbf{particle growth and fragmentation}, as well as \textbf{accretion rate and stellar luminosity evolution}.

From the chemistry perspective, molecular abundances vary significantly across a typical disk, due to steep gradients in temperature, density and radiation. 
%Figure 1 shows a theoretical example of how both changes in disk temperature, as well as time evolution, decrease the abundance of carbon monoxide (CO) by several orders of magnitude \cite{aikawa96}. 
The complexity of disk chemistry means that coupling it with dynamical processes, while necessary, is non-trivial.  I will couple the dynamical framework outlined above with \textbf{time-dependent chemical models of increasing complexity}, informed by results from state-of-the-art disk chemistry models (that can only be run on static disks). This will show how the \textbf{snowline locations of volatiles}, as well as the \textbf{chemical composition of the disk gas and dust evolve}, which has \textbf{direct implications on the compositions of young planets}. Moreover, since I am primarily interested in processes that affect the disk mid-plane where planets form, my results could be simplified by the fact that certain chemical processes may not be important in this disk parameter space.

%this dynamical model  develop a (chemical models of increasing complexity) simplified time-dependent chemistry,

\vspace{0.15in}
%\section{Planet and Planetesimal Migration}
\underline{\textbf{2. Planet and Planetesimal Migration}}

The complex processes outlined in part 1 will directly impact the composition and dynamics of forming gas giants, as the latter are born and evolve simultaneously with the disk. The chemical evolution of the gas disk molecular abundances, as well as disk dynamics, determine how a young planet's atmospheric composition changes in time and with the planet's location, and therefore the final chemical structure of mature planets. \textbf{It is thus clear that disks and planets are deeply connected, and that this relation needs to be explored}. %I will add planetary embryos in the code developed in step 1, and

Giant planets can migrate through the disk while still accumulating gas. Figure 2 shows an exciting ALMA observation of a planetary gap in the disk around HL Tau $[1]$. This will change a planet's atmospheric composition since the disk chemical abundances are different at different disk locations. Additionally, giant planets may still accumulate planetesimals while accreting nebular gas [7]. The final composition of a planet's atmosphere will thus depend on how much gas and solids are accreted in this stage. \textbf{I will add planet dynamical effects such as migration and planetesimal accretion in the chemical and dynamical model developed in part 1, and quantify how these processes affect the chemical composition of gas giant envelopes}.

\vspace{0.15in}

%\section{Model Planet Populations} 
\underline{\textbf{3. Model Planet Populations}}

The results from parts 1 and 2 will feed into a \textbf{large planet synthesis model}, in which I will use a grid of different \textbf{initial disk and planetary embryo conditions}. I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[4]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations. In contrast, for this computationally expensive step, I will only include the processes that I have identified to be the most important in the local simulations from steps 1 and 2. This will allow me to \textbf{constrain a planet's formation location based on its chemical composition}. Comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[5]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.

\section*{Proposed Host Institutions}

\noindent \textbf{1. Massachusetts Institute of Technology (MIT)}. I believe MIT to be the best place for me to pursue my postdoctoral research, due to its vibrant community of experts in exoplanet atmospheres and dynamics, both in the Physics and Earth and Planetary Sciences (EAPS) departments. In particular, I would like to collaborate with Sara Seager, who is a co-investigator of the TESS mission and a world-leading theorist in atmospheric chemistry, as well as with members of her group, who specialize in a broad range of exoplanet theory and computational topics. I would also love to work with \textbf{Hilke Schlichting} (proposed faculty host), a leader in planet formation theory and dynamics. Margaret Pan, an excellent theorist in planetary dynamics and portoplanetary disk, will join the EAPS department as a research scientist in early 2016, so it would be a pleasure to work with her as well. Finally, in order to connect my theoretical work with observations, I would like to collaborate with Joshua Winn, who is an expert in discovering and characterizing exoplanets. \\

\noindent \textbf{2. University of Chicago and University of Michigan} The University of Chicago, my second choice for a host institution, would also be an ideal place for me to undertake my proposed research, due to its opportunities for valuable collaborations with experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences. \textbf{Fred Ciesla} (proposed faculty host) is an expert in protoplanetary disk dynamics, as well as chemical composition and evolution. Leslie Rogers, who will begin her faculty appointment in Fall 2016, is a leader in exoplanet theory, specifically planetary interiors and atmospheres, so I would love the opportunity to collaborate with her on my postdoctoral work. Additionally, the University of Chicago Department of Astronomy hosts leaders in detecting and characterizing worlds outside the Solar system, such as Dan Fabrycky and Jacob Bean, which presents great prospects in connecting my theoretical research work with observations. Finally, The University of Michigan, my third choice for a host institution, would be an excellent place to pursue my postdodcotral research due to its plethora of protoplanetary disks and exoplanet experts. \textbf{Edwin Bergin} (proposed faculty host) is an expert in modeling protplanetary disk chemistry, so I would greatly benefit from collaborating with him on this part of my proposed research. I would also love to collaborate with Emily Rauscher, who is a leader in theoretical studies of exoplanet atmospheres. Additionally, the University of Michigan Department of Astronomy has strong ties with the Physics Department, where I would like to collaborate with Fred Adams, an expert in circumstellar disk dynamics and exoplanet theory.

%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.7\textwidth]{CO_abundances}
%%\vspace{-0.5in}
%\caption{TBD}
%\label{fig:CO}
%\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chemical_timescales}
%\vspace{-0.5in}
\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height. The numbered curves are $\log_{10} (\tau)$ in million years (from $[11]$).}
\label{fig:chemical}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.4\textwidth]{HLTau_nrao}
%\vspace{-0.5in}
\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[1]$).}
\label{fig:HLTau}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.9\textwidth]{water_spectrum}
%\vspace{-0.5in}
\caption{The transmission spectrum of exoplanet WASP-12b measured with the Hubble Space Telescope/Wide Field Camera 3. The white dots are the measured transit depths, the blue squares and dark blue line are the best fit model, while the shaded regions represent 1- and 2 $\sigma$ intervals in the retrieved spectrum. The increase in transit depth around $\sim$1.4 $\mu$m is evidence for a water feature in the atmosphere (from $[5]$).}
\label{fig:HLTau}
\end{figure}

%\if\bibinc n
%\bibliography{refs}
%\fi
\FloatBarrier
%\def\bibfont{\footnotesize}
%\setlength{\bibsep}{0.0pt}

\section*{References}
\footnotesize
%\begin{thebibliography} {9}
\noindent $[1]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
$[2]$ Batalha, N. M. 2014, Proceedings of the National Academy of Science, 111, 12647 \\
$[3]$ Debes, J. H., Jang-Condell, H., Weinberger, A. J., Roberge, A., \& Schneider, G. 2013, ApJ, 771, 45 \\
$[4]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
$[5]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ArXiv e-prints, arXiv:1504.05586 \\
$[6]$ Lissauer J. J., Dawson R. I., Tremaine S., 2014, Nature, 513, 336 \\
$[7]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
$[8]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21 \\
$[9]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
$[10]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, resubmitted after referee report \\
$[11]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25
%\end{thebibliography}

%\bibliographystyle{abbrv}
%\bibliography{refs}

%\if\bibinc y
%\begin{thebibliography}
%\end{thebibliography}
%\fi

\end{document}

### 2015-11-04 Postdoc Applications/sagan_proposal.tex
update

--- 2015-11-04 Postdoc Applications/hubble_proposal.tex
@@
 %\usepackage{subeqnarray}
-\usepackage[top=0.9in, bottom=0.8in, left=1in, right=1in]{geometry}
+\usepackage[top=1in, bottom=0.8in, left=1in, right=1in]{geometry}
 \usepackage{natbib}
@@
 \setlength{\droptitle}{-7em}
-\setlength{\abovecaptionskip}{-0.2ex}
-\setlength{\belowcaptionskip}{-0.2ex}
+\setlength{\abovecaptionskip}{-0.4ex}
+\setlength{\belowcaptionskip}{-0.4ex}
 %\pagenumbering{gobble}
@@
 \lhead{Ana-Maria Piso}
-\rhead{Hubble Fellowship Research Proposal}
+\rhead{Sagan Fellowship Research Proposal}
 \date{}
@@
 
-\title{Hubble Fellowship Research Statement: \\
+\title{Sagan Fellowship Research Statement: \\
 The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets}
@@
 
-\vspace{-0.9cm}
+\vspace{-0.8cm}
 
@@
 Furthermore, this work provides essential context for characterizing the gas giants that instruments such as the James Webb Space Telescope (JWST) and the Transiting Exoplanet Survey Satellite (TESS) will one day discover.
-Characterizing the composition of planetary atmospheres is therefore an essential component for understanding the observations of NASA's Cosmic Origins Program.
-\vspace{0.15in}
+My proposed work has therefore direct relevance to NASA's Exoplanet Exploration Program.
+
+\vspace{0.2in}
 
@@
 
-\vspace{0.15in}
+\vspace{0.2in}
 %\section{Planet and Planetesimal Migration}
@@
 
-\vspace{0.15in}
+\vspace{0.2in}
 
@@
 \noindent \textbf{2.
-University of Chicago and University of Michigan} The University of Chicago, my second choice for a host institution, would also be an ideal place for me to undertake my proposed research, due to its opportunities for valuable collaborations with experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
+University of Chicago.} The University of Chicago would be an ideal place for me to undertake my proposed research, due to its opportunities for valuable collaborations with experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
 \textbf{Fred Ciesla} (proposed faculty host) is an expert in protoplanetary disk dynamics, as well as chemical composition and evolution.
@@
 Additionally, the University of Chicago Department of Astronomy hosts leaders in detecting and characterizing worlds outside the Solar system, such as Dan Fabrycky and Jacob Bean, which presents great prospects in connecting my theoretical research work with observations.
-Finally, The University of Michigan, my third choice for a host institution, would be an excellent place to pursue my postdodcotral research due to its plethora of protoplanetary disks and exoplanet experts.
-\textbf{Edwin Bergin} (proposed faculty host) is an expert in modeling protplanetary disk chemistry, so I would greatly benefit from collaborating with him on this part of my proposed research.
-I would also love to collaborate with Emily Rauscher, who is a leader in theoretical studies of exoplanet atmospheres.
-Additionally, the University of Michigan Department of Astronomy has strong ties with the Physics Department, where I would like to collaborate with Fred Adams, an expert in circumstellar disk dynamics and exoplanet theory.
 

### 2015-11-04 Postdoc Applications/previous_research.tex
update

%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
%\documentclass[12pt, preprint]{aastex}
%\documentclass[apj]{emulateapj}
\documentclass[12pt, letterpaper]{article}

%\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
%\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex

%\usepackage{pdfsync}
%\usepackage{subeqnarray}
\usepackage[top=1in, bottom=0.8in, left=1in, right=1in]{geometry}
\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage{titling}
\usepackage{sectsty}
\usepackage{sidecap}
\usepackage{placeins}
\usepackage{indentfirst}
\usepackage{wrapfig}
\setlength{\droptitle}{-7em}
\setlength{\abovecaptionskip}{-0.4ex}
\setlength{\belowcaptionskip}{-0.4ex}
%\pagenumbering{gobble}
\pagestyle{fancy}
\lhead{Ana-Maria Piso}
\rhead{Sagan Fellowship Current and Previous Research Statement}
\date{}
%\rhead{\thepage}

%\bibliographystyle{apj}

\title{Sagan Fellowship Previous and Current Research Statement: \\
Giant Planet Formation and Snowlines in Protoplanetary Disks}
\author{Ana-Maria Piso}

%\newenvironment{packed_item}{
%\begin{itemize}
%  \setlength{\itemsep}{1pt}
%  \setlength{\parskip}{0pt}
%  \setlength{\parsep}{0pt}
%}{\end{itemize}}

\begin{document}
\maketitle

%\slugcomment{Draft Modified \today}

\vspace{-0.8cm}

Planets are born in protoplanetary disks, which means that their structure and composition are determined by and highly connected to the chemical composition and structure of the disk in which they form. In my current and previous research, I have approached this intricate disk-planet link from two directions: (1) by exploring the core accretion mechanism, specifically calculating the minimum required core mass to form a gas giant before the dissipation of the gas in the protoplanetary disk, and (2) by understanding how disk chemistry and dynamics shape the snowline locations of volatiles in disks, which has direct implications for the chemical composition of extrasolar planet (exoplanet) atmospheres.

\vspace{0.2in}

%\section{Coupled Chemical and Dynamical Disk Evolution} 
\underline{\textbf{Minimum Core Masses for Giant Planet Formation}}

Gas giants are widely believed to form through core accretion (Stevenson 1982, Pollack et al. 1996), a theory in which solid protoplanetary cores grow large enough to accumulate a massive atmosphere. This is particularly challenging in the outer parts of a disk, where long dynamical times make it difficult for a core to grow fast enough before the gas disk dissipates on a timescale of a few million years. At the same time, however, giant planets on wide orbits have been discovered in recent years, such as the HR 8799 system (Marois et al. 2008). This poses an intriguing question: how did these planets form? In this part of my research, I attempted to address this issue ? specifically, what is the lowest possible core mass required to form a gas giant before disk dissipation? We start investigating this issue in Piso & Youdin (2014), where we develop a core accretion model complementary to most state-of-the-art planet formation models, which assume that a giant planet?s core and atmosphere grow simultaneously (Bodenheimer & Pollack 1986, Wuchterl 1993, Rafikov 2006). Specifically, we assume that planetesimal accretion is negligible, and instead the envelope accretes gas while undergoing Kelvin-Helmholtz contraction. This gives a lower limit on the minimum (critical) core mass, Mcrit, to form a giant planet, since additional planetesimal accretion in this stage would heat up the atmosphere, inhibit its ability to cool, thus increasing the critical core mass. Moreover, while Mcrit has been systematically computed as a function of disk properties and stellocentric separation for steady-state atmospheres heated by planetesimal accretion (Rafikov 2006), no equivalent systematic study is available for this minimum value of Mcrit. We develop such a study by considering envelopes accreting around fully formed cores at distances between 5 and 100 astronomical units (AU) from their host star. We use a simple model that assumes the nebular gas to be ideal and polytropic, and that the dust opacity is the standard interstellar opacity (ISM; Bell & Lin 94). We find Mcrit to decrease between 8.5 Earth Masses (ME) at 5 AU to 3.5 ME at 100 AU. These results are lower than those of standard studies, which typically assume Mcrit ~ 10 ME (Stevenson 1982). Our study hence makes great strides in understanding the formation of giant planets at large separations. Moreover, our numerical technique has been used by Lee et al. (2014) and Lee et al. (2015) to understand planet formation in the inner disk.

\begin{wrapfigure}{r}{0.5\textwidth}
  \begin{center}
    \includegraphics[width=0.5\textwidth]{Mcrit_vs_a_gg}
  \end{center}
  \caption{A gull}
\end{wrapfigure}

The idealized model developed in Piso & Youdin (2014) neglects realistic effects, such as the fact that the nebular gas is non-ideal (Saumon et al. 1995), as well as that the dust opacity is likely to be lower than the ISM one due to grain growth (Pollack 1985). In Piso, Youdin & Murray-Clay (2015), we improve our model by including these effects. We use the Saumon et al. (1995) EOS tables, and we extend them to lower temperatures and pressures as suitable for the parameter space we are exploring. These tables are available upon request and may prove useful to understand planet formation in the outer disk. Moreover, we use realistic opacity tables that take into account grain growth (D?Alessio et al. 2001) to calculate Mcrit. We find that a realistic EOS increases Mcrit, while grain growth opacities, in contrast, decrease Mcrit. Taking these two competing effects together, we calculate Mcrit ~ 8 ME at 5 AU, decreasing to 5 ME at 100 AU. Our results are still lower than 10 ME, the typically quoted value, and may be up to one order of magnitude lower if grain coagulation is taken into account (see Piso, Youdin & Murray-Clay 2015 for details). Our work provides further insight into giant planet formation, and is a rich field of research to pursue.

\vspace{0.2in}
%\section{Planet and Planetesimal Migration}
\underline{\textbf{Snowline Locations in Protoplanetary Disks}}

A. H2O, CO2, CO Snowlines and the C/O ratio	
The locations of volatile snowlines in protoplanetary disks are a defining feature of both gas giant and protoplanetary disk chemistry, as they provide vital information about the abundance of these molecules in gas and dust throughout the disk. In this part of my dissertation and beyond, I want to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) back-track the planet formation location based on planet composition. While disk structure is very complex, we do have valuable observational evidence of molecular abundance detections in disks (Henning & Semenov 2013), as well as volatile snowlines (Qi et al. 2013, Zhang et al. 2013). One important signature of disk and exoplanet chemistry is the carbon-to-oxygen (C/O) ratio. Specifically, an important consequence of snowline formation in disks is that disks are expected to present different C/O ratios in the gas and in icy dust mantles at different disk radii. This effect was quantified by Öberg et al. (2011), who considered the fact that the main carries of carbon and oxygen, i.e. H2O, CO2 and CO, have different condensation temperatures. This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above. In Piso, Öberg, et al. (2015), we improve on this model by including the effect of radial drift of solids and viscous gas accretion on the volatile snowline locations. We use a simple, semi-analytical model to solve for the coupled drift-desorbed equations for particles of different initial sizes, as well as for various disk models (irradiated, evolving, viscous). For all disks, we find that particles within an initial size range desorb at a size-dependent location in the disk, which is independent of the particle's initial position. From this, we can estimate upper limits for the C/O ratio in disks (see Piso, Öberg, et al. 2015 for details). The largest desorbing particles in our model have initial sizes of ~7 m, and they are representative for the snowline locations, as realistic grain size distributions are dominated in mass by the largest particles. We thus obtain a very powerful result: radial drift and gas accretion may move the H2O, CO2 and CO snowlines by up to 40-60\% compared to a static disk, which is a huge difference. For a transition disk, i.e. a disk with an inner cavity significantly depleted of gas, we find that H2O particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the gap edge, due to the decrease in gas pressure inside the cavity. This is qualitatively consistent with the observations of Zhang et al. (2013), which show that the H2O snowline is pushed outwards in a transition disk compared to a full disk. Our model framework is thus generally valid for more complicated disk structures as well. We plan to further improve this model by adding several disk processes, such as turbulent diffusion, particle composition and grain morphology.

\begin{wrapfigure}{l}{0.5\textwidth}
  \begin{center}
    \includegraphics[width=0.5\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new2}
  \end{center}
  \caption{A gull}
\end{wrapfigure}

B. Nitrogen Abundance and Ice Binding Energies
Aside from the main C and O carriers, nitrogen is an important molecule to study, as it is highly abundant in the Solar system (Lodders 2009) and disks, and primarily found as N2. In Piso et al. (in prep), we show that the nitrogen to oxygen (N/O) ratio in gas is significantly larger than the Solar abundance. This implies that the N/O ratio may be used as an additional signature of disk and exoplanet atmospheric chemistry, aside from the C/O ratio, which may give us another clue into the chemical composition of protoplanetary disks and, implicitly, of planets. Moreover, we use new results from Fayolle et al. (in prep) for the binding energies of CO and N2, mixed with H2O or as pure ices. We find that the binding energies of water-ice mixtures are significantly larger than the pure ice ones, and move the CO and N2 snowlines by at least a few AU.

%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.7\textwidth]{CO_abundances}
%%\vspace{-0.5in}
%\caption{TBD}
%\label{fig:CO}
%\end{figure}

%\if\bibinc n
%\bibliography{refs}
%\fi
\FloatBarrier
%\def\bibfont{\footnotesize}
%\setlength{\bibsep}{0.0pt}

\section*{References}
\footnotesize
%\begin{thebibliography} {9}
\noindent $[1]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
$[2]$ Batalha, N. M. 2014, Proceedings of the National Academy of Science, 111, 12647 \\
$[3]$ Debes, J. H., Jang-Condell, H., Weinberger, A. J., Roberge, A., \& Schneider, G. 2013, ApJ, 771, 45 \\
$[4]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
$[5]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ArXiv e-prints, arXiv:1504.05586 \\
$[6]$ Lissauer J. J., Dawson R. I., Tremaine S., 2014, Nature, 513, 336 \\
$[7]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
$[8]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21 \\
$[9]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
$[10]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, resubmitted after referee report \\
$[11]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25
%\end{thebibliography}

%\bibliographystyle{abbrv}
%\bibliography{refs}

%\if\bibinc y
%\begin{thebibliography}
%\end{thebibliography}
%\fi

\end{document}

### 2015-11-04 Postdoc Applications/AnaMariaPisoResume.tex
update

@@
 \item[] Advisor: Dr. Karin \"Oberg 
-\item[] Thesis Topic: ``The Effect of Disk Chemistry and Dynamics in Shaping Volatile Snowlines''   
+\item[] Thesis Topic: ``Dynamics and Chemistry in Protoplanetary Disks''   
 \end{list1}
@@
 \textit{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}.
-ApJ, under review \\
+ApJ, resubmitted after referee report \\
 \end{list2}
@@
 \section{\sc Conferences and Seminars}
+
+\begin{list2}
+\item[] {\bf C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion} \\
+Extreme Solar Systems III, Waikoloa, HI, December 2015 \\
+Poster \\
+\end{list2}
 

### 2015-11-04 Postdoc Applications/previous_research.tex
update

@@
 %\usepackage{subeqnarray}
-\usepackage[top=1in, bottom=0.8in, left=1in, right=1in]{geometry}
+\usepackage[top=0.9in, bottom=0.6in, left=1in, right=1in]{geometry}
 \usepackage{natbib}
@@
 \usepackage{wrapfig}
+\usepackage[font={footnotesize}]{caption}
+\usepackage{multicol}
 \setlength{\droptitle}{-7em}
-\setlength{\abovecaptionskip}{-0.4ex}
-\setlength{\belowcaptionskip}{-0.4ex}
+\setlength{\abovecaptionskip}{-2.5ex}
+\setlength{\belowcaptionskip}{-2.5ex}
 %\pagenumbering{gobble}
@@
 \lhead{Ana-Maria Piso}
-\rhead{Sagan Fellowship Current and Previous Research Statement}
+\rhead{Sagan Fellowship Previous and Current Research Statement}
 \date{}
@@
 
-\title{Sagan Fellowship Previous and Current Research Statement: \\
-Giant Planet Formation and Snowlines in Protoplanetary Disks}
-\author{Ana-Maria Piso}
+\title{\Large Sagan Fellowship Previous and Current Research Statement: \\
+Giant Planet Formation and Snowlines in Protoplanetary Disks\vspace{-2ex}}
+\author{Ana-Maria Piso \vspace{-2ex}}
 
@@
 
-\vspace{-0.8cm}
+\vspace{-1.5cm}
 
@@
 In my current and previous research, I have approached this intricate disk-planet link from two directions: (1) by exploring the core accretion mechanism, specifically calculating the minimum required core mass to form a gas giant before the dissipation of the gas in the protoplanetary disk, and (2) by understanding how disk chemistry and dynamics shape the snowline locations of volatiles in disks, which has direct implications for the chemical composition of extrasolar planet (exoplanet) atmospheres.
+\\
+\underline{\textbf{1. Minimum Core Masses for Giant Planet Formation}}
 
-\vspace{0.2in}
+Gas giants are widely believed to form through core accretion $[18]$, a theory in which solid protoplanetary cores grow large enough to accumulate a massive atmosphere.
+In this model, planetesimals in a disk grow larger through collisions, eventually forming a planetary embryo, which continues to grow by attracting planetesimals in its neighborhood.
+Once an embryo becomes large enough so that its escape velocity exceeds the thermal velocity of the nebular gas in its vicinity, it starts accumulating a gaseous envelope.
+From this point on, the accretion of gas is regulated by the pressure support within the envelope --- the amount of gas a core can accumulate is limited by the atmosphere's ability to radiate away the energy due to the incoming planetesimals, as well as by envelope contraction.
 
-%\section{Coupled Chemical and Dynamical Disk Evolution} 
-\underline{\textbf{Minimum Core Masses for Giant Planet Formation}}
-
-Gas giants are widely believed to form through core accretion (Stevenson 1982, Pollack et al.
-1996), a theory in which solid protoplanetary cores grow large enough to accumulate a massive atmosphere.
-This is particularly challenging in the outer parts of a disk, where long dynamical times make it difficult for a core to grow fast enough before the gas disk dissipates on a timescale of a few million years.
+Core accretion is particularly challenging in the outer parts of a disk, where long dynamical times make it difficult for a core to grow fast enough before the gas disk dissipates on a timescale of a few million years.
 At the same time, however, giant planets on wide orbits have been discovered in recent years, such as the HR 8799 system (Marois et al.
@@
 This poses an intriguing question: how did these planets form?
-In this part of my research, I attempted to address this issue ?
-specifically, what is the lowest possible core mass required to form a gas giant before disk dissipation?
-We start investigating this issue in Piso & Youdin (2014), where we develop a core accretion model complementary to most state-of-the-art planet formation models, which assume that a giant planet?s core and atmosphere grow simultaneously (Bodenheimer & Pollack 1986, Wuchterl 1993, Rafikov 2006).
-Specifically, we assume that planetesimal accretion is negligible, and instead the envelope accretes gas while undergoing Kelvin-Helmholtz contraction.
-This gives a lower limit on the minimum (critical) core mass, Mcrit, to form a giant planet, since additional planetesimal accretion in this stage would heat up the atmosphere, inhibit its ability to cool, thus increasing the critical core mass.
+I addressed this issue in two papers, $[13]$ and $[14]$.
+Specifically, I answered the following question: what is the lowest possible core mass required to form a gas giant before disk dissipation?
+This minimum core mass applies when  solid cores no longer accrete planetesimals, as this would heat the gaseous envelope and inhibit the atmosphere's ability to cool and contract .
+Most state-of-the-art models, however, assume that a giant planet's core and atmosphere grow simultaneously ($[2]$, $[20]$), and thus have not been constructed to effectively explore the value of the true minimum core mass required to form a gas giant.
+We start investigating this minimum (critical) core mass $M_{\rm crit}$ in $[13]$, where we develop a core accretion model which assumes that planetesimal accretion is negligible, and instead the envelope accretes gas while undergoing Kelvin-Helmholtz contraction.
+%This gives a lower limit on the minimum (critical) core mass, Mcrit, to form a giant planet, since additional planetesimal accretion in this stage would heat up the atmosphere, inhibit its ability to cool, thus increasing the critical core mass.
 Moreover, while Mcrit has been systematically computed as a function of disk properties and stellocentric separation for steady-state atmospheres heated by planetesimal accretion (Rafikov 2006), no equivalent systematic study is available for this minimum value of Mcrit.
-We develop such a study by considering envelopes accreting around fully formed cores at distances between 5 and 100 astronomical units (AU) from their host star.
-We use a simple model that assumes the nebular gas to be ideal and polytropic, and that the dust opacity is the standard interstellar opacity (ISM; Bell & Lin 94).
-We find Mcrit to decrease between 8.5 Earth Masses (ME) at 5 AU to 3.5 ME at 100 AU.
-These results are lower than those of standard studies, which typically assume Mcrit ~ 10 ME (Stevenson 1982).
-Our study hence makes great strides in understanding the formation of giant planets at large separations.
-Moreover, our numerical technique has been used by Lee et al.
-(2014) and Lee et al.
-(2015) to understand planet formation in the inner disk.
+We consider envelopes accreting around fully formed cores at distances between 5 and 100 astronomical units (AU) from their host star.
+We use a simple model that assumes the nebular gas to be ideal and polytropic, and that the dust opacity is the standard interstellar opacity (ISM; $[1]$).
+We find $M_{\rm crit}$ to decrease between 8.5 Earth Masses ($M_{\oplus}$) at 5 AU to 3.5 $M_{\oplus}$ at 100 AU.
+%These results are lower than those of standard studies, which typically assume $M_{\rm crit} \sim 10 M_{\oplus}$. 
+%(Stevenson 1982). Our study hence makes great strides in understanding the formation of giant planets at large separations. Moreover, 
+We note that our numerical technique has been used by $[7]$ and $[8]$ to understand planet formation in the inner disk.
 
-\begin{wrapfigure}{r}{0.5\textwidth}
+The simplified model developed in $[13]$ assumes an ideal equation of state (EOS) and an ISM opacity.
+To obtain more robust quantitative results for $M_{\rm crit}$, we expand this model in $[14]$ by making two important additions: (1) a realistic EOS, and (2) realistic dust opacities.
+We use the $[21]$ EOS tables, and we extend them to lower temperatures and pressures as suitable for the parameter space we are exploring.
+Moreover, we use realistic opacity tables that take into account grain growth $[4]$ to calculate $M_{\rm crit}$.
+We find that a realistic EOS increases $M_{\rm crit}$, while grain growth opacities, in contrast, decrease $M_{\rm crit}$.
+Taking these two competing effects together, we calculate $M_{\rm crit} \sim 8 M_{\oplus}$ at 5 AU, decreasing to 5 $M_{\oplus}$ at 100 AU (Figure 1).
+Our results are still lower than 10 $M_{\oplus}$, the typically quoted value, and may be up to one order of magnitude lower if grain coagulation is taken into account (see $[14]$ for details).
+Thus our studies clearly challenge previous claims that core accretion cannot operate in the outer parts of protoplanetary disks, reopening the case for in situ formation of wide-separation gas giants.
+\begin{wrapfigure}{r}{0.4\textwidth}
   \begin{center}
-    \includegraphics[width=0.5\textwidth]{Mcrit_vs_a_gg}
+    \includegraphics[width=0.4\textwidth]{Mcrit_vs_a_gg}
   \end{center}
-  \caption{A gull}
+  %\vspace{-0.1in}
+  \caption{$M_{\rm crit}$ as a function of semi-major axis for a realistic EOS, and both ISM and grain growth opacities. Grain growth significantly decreases $M_{\rm crit}$ (from $[14]$).}
+  %\vspace{-0.3in}
 \end{wrapfigure}
+\underline{\textbf{2. Snowline Locations in Protoplanetary Disks}}
 
-The idealized model developed in Piso & Youdin (2014) neglects realistic effects, such as the fact that the nebular gas is non-ideal (Saumon et al.
-1995), as well as that the dust opacity is likely to be lower than the ISM one due to grain growth (Pollack 1985).
-In Piso, Youdin & Murray-Clay (2015), we improve our model by including these effects.
-We use the Saumon et al.
-(1995) EOS tables, and we extend them to lower temperatures and pressures as suitable for the parameter space we are exploring.
-These tables are available upon request and may prove useful to understand planet formation in the outer disk.
-Moreover, we use realistic opacity tables that take into account grain growth (D?Alessio et al.
-2001) to calculate Mcrit.
-We find that a realistic EOS increases Mcrit, while grain growth opacities, in contrast, decrease Mcrit.
-Taking these two competing effects together, we calculate Mcrit ~ 8 ME at 5 AU, decreasing to 5 ME at 100 AU.
-Our results are still lower than 10 ME, the typically quoted value, and may be up to one order of magnitude lower if grain coagulation is taken into account (see Piso, Youdin & Murray-Clay 2015 for details).
-Our work provides further insight into giant planet formation, and is a rich field of research to pursue.
+The locations of volatile snowlines in protoplanetary disks are a defining feature of both gas giant and disk chemistry, as they provide vital information about the abundance of these molecules in gas and dust throughout the disk.
+In this part of my dissertation and beyond, I want to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) back-track the planet formation location based on planet composition.
+While disk structure is very complex, we do have valuable observational evidence of molecular abundance detections in disks $[6]$, as well as volatile snowlines ($[19]$, $[22]$).
+\\
+\underline{\textbf{A. H$_2$O, CO$_2$, CO Snowlines and the C/O ratio.}}
 
-\vspace{0.2in}
-%\section{Planet and Planetesimal Migration}
-\underline{\textbf{Snowline Locations in Protoplanetary Disks}}
+One important signature of disk and exoplanet chemistry is the carbon-to-oxygen (C/O) ratio, as small variations of the C/O ratio may affect the abundance of other volatiles by several orders of magnitudes (e.g., $[11]$ for the case of exoplanet atmospheres).
+Specifically, an important consequence of volatile condensation and sublimation in disks is that disks are expected to present different C/O ratios in the gas and in the icy dust mantles at different disk radii.
+This effect was quantified by $[12]$, who considered the fact that the main carries of carbon and oxygen, i.e.
+H$_2$O, CO$_2$ and CO, have different condensation temperatures.
+This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
 
-A. H2O, CO2, CO Snowlines and the C/O ratio	
-The locations of volatile snowlines in protoplanetary disks are a defining feature of both gas giant and protoplanetary disk chemistry, as they provide vital information about the abundance of these molecules in gas and dust throughout the disk.
-In this part of my dissertation and beyond, I want to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) back-track the planet formation location based on planet composition.
-While disk structure is very complex, we do have valuable observational evidence of molecular abundance detections in disks (Henning & Semenov 2013), as well as volatile snowlines (Qi et al.
-2013, Zhang et al.
-2013).
-One important signature of disk and exoplanet chemistry is the carbon-to-oxygen (C/O) ratio.
-Specifically, an important consequence of snowline formation in disks is that disks are expected to present different C/O ratios in the gas and in icy dust mantles at different disk radii.
-This effect was quantified by Öberg et al.
-(2011), who considered the fact that the main carries of carbon and oxygen, i.e.
-H2O, CO2 and CO, have different condensation temperatures.
-This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
-In Piso, Öberg, et al.
-(2015), we improve on this model by including the effect of radial drift of solids and viscous gas accretion on the volatile snowline locations.
-We use a simple, semi-analytical model to solve for the coupled drift-desorbed equations for particles of different initial sizes, as well as for various disk models (irradiated, evolving, viscous).
+In $[15]$, we expand this model to include the effect of radial drift of solids and viscous gas accretion on the volatile snowline locations.
+We use a simple, semi-analytical model to solve for the coupled drift-desorption equations for particles of different initial sizes, as well as for various disk models (irradiated, evolving, viscous).
 For all disks, we find that particles within an initial size range desorb at a size-dependent location in the disk, which is independent of the particle's initial position.
-From this, we can estimate upper limits for the C/O ratio in disks (see Piso, Öberg, et al.
-2015 for details).
-The largest desorbing particles in our model have initial sizes of ~7 m, and they are representative for the snowline locations, as realistic grain size distributions are dominated in mass by the largest particles.
-We thus obtain a very powerful result: radial drift and gas accretion may move the H2O, CO2 and CO snowlines by up to 40-60\% compared to a static disk, which is a huge difference.
+From this, we can estimate upper limits for the C/O ratio in disks (Figure 2).
+%The largest desorbing particles in our model have initial sizes of ~7 m, and they are representative for the snowline locations, as realistic grain size distributions are dominated in mass by the largest particles.
+By comparing the innermost snowlines in Figure 2, which are set by the largest desorbing particles ($\sim$7 m in our model), we obtain a very powerful result: radial drift and gas accretion may move the H$_2$O, CO$_2$ and CO snowlines by up to 40-60\% compared to a static disk, which is a huge difference.
 For a transition disk, i.e.
-a disk with an inner cavity significantly depleted of gas, we find that H2O particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the gap edge, due to the decrease in gas pressure inside the cavity.
-This is qualitatively consistent with the observations of Zhang et al.
-(2013), which show that the H2O snowline is pushed outwards in a transition disk compared to a full disk.
+a disk with an inner cavity significantly depleted of gas, we find that H$_2$O particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the gap edge, due to the decrease in gas pressure inside the cavity.
+This is qualitatively consistent with the observations of $[22]$, which show that the H$_2$O snowline is pushed outwards in a transition disk compared to a full disk.
 Our model framework is thus generally valid for more complicated disk structures as well.
 We plan to further improve this model by adding several disk processes, such as turbulent diffusion, particle composition and grain morphology.
+\begin{wrapfigure}{r}{0.45\textwidth}
+  \begin{center}
+    \includegraphics[width=0.46\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new2}
+  \end{center}
+  %\vspace{-0.1in}
+  \caption{Snowline locations and C/O ratios upper limits in three disk models.
+The color bar corresponds to the initial particle size in cm.
+Compared to a static disk (black lines), radial drift and gas accretion may push the snowlines inwards by 40-60\%.
+The arrows show that the C/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in disks that take into account gas accretion, as the relative fluxes of the desorbed icy particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (from $[15]$).}
+  \vspace{-0.1in}
+\end{wrapfigure}
+%\vspace{-0.05in}
+\underline{\textbf{B. Nitrogen Abundance and Ice Binding Energies}}
 
-\begin{wrapfigure}{l}{0.5\textwidth}
-  \begin{center}
-    \includegraphics[width=0.5\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new2}
-  \end{center}
-  \caption{A gull}
-\end{wrapfigure}
+Aside from the main C and O carriers, nitrogen (N) is an important molecule to study, as it is highly abundant in the Solar system $[9]$ and disks, and primarily found as N$_2$.
+Gas phase enhancements in the nitrogen-to-oxygen ratio (N/O) are predicted to greatly exceed the C/O ratios in the outer disk regions.
+Giant planets that form at wide separations should thus have an excess of N in their atmospheres, which could be used to trace their formation origin.
+In $[16]$, we quantify this effect in disks.
+We find that the N/O ratio in gas is significantly larger than the Solar abundance, and indeed exceeds the C/O ratio in the outer disk, in agreement with predictions.
+We can thus set the stage to answer one of our main questions, i.e.
+back-track the planet formation location based on planet composition.
+Moreover, we use new results from $[5]$ for the binding energies of CO and N$_2$, mixed with H$_2$O or as pure ices.
+We find that the binding energies of water-ice mixtures are significantly larger than the pure ice ones, and move the CO and N$_2$ snowlines by at least a few tens of AU.
+\\
+\underline{\textbf{C. Self-consistent C/N/O Calculation}}
 
-B. Nitrogen Abundance and Ice Binding Energies
-Aside from the main C and O carriers, nitrogen is an important molecule to study, as it is highly abundant in the Solar system (Lodders 2009) and disks, and primarily found as N2.
-In Piso et al.
-(in prep), we show that the nitrogen to oxygen (N/O) ratio in gas is significantly larger than the Solar abundance.
-This implies that the N/O ratio may be used as an additional signature of disk and exoplanet atmospheric chemistry, aside from the C/O ratio, which may give us another clue into the chemical composition of protoplanetary disks and, implicitly, of planets.
-Moreover, we use new results from Fayolle et al.
-(in prep) for the binding energies of CO and N2, mixed with H2O or as pure ices.
-We find that the binding energies of water-ice mixtures are significantly larger than the pure ice ones, and move the CO and N2 snowlines by at least a few AU.
+As shown in Figure 2, our current drift-desorption model can only provide us with upper estimates for C/O (or N/O) ratios.
+We plan to enhance this model by performing the time-dependent calculation that accounts for the relative motion of the desorbed ices and the overall nebular gas, the time evolution of the dust surface density,  and radial gas diffusion; all these effects will change the volatile abundances and the C/O ratio in time.
+We have developed a code that evolves the dust surface density for a particle of fixed size.
+Next, we will treat  each species in gaseous and solid forms as two fluids that are interchanging.
+Following $[3]$, we will use advection-like equations to solve for their separate time-dependent abundance.
 
@@
 %\fi
-\FloatBarrier
+%\FloatBarrier
 %\def\bibfont{\footnotesize}
@@
 
-\section*{References}
 \footnotesize
+
+%\textbf{References} \\
 %\begin{thebibliography} {9}
-\noindent $[1]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
-$[2]$ Batalha, N. M. 2014, Proceedings of the National Academy of Science, 111, 12647 \\
-$[3]$ Debes, J. H., Jang-Condell, H., Weinberger, A. J., Roberge, A., \& Schneider, G. 2013, ApJ, 771, 45 \\
-$[4]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
-$[5]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ArXiv e-prints, arXiv:1504.05586 \\
-$[6]$ Lissauer J. J., Dawson R. I., Tremaine S., 2014, Nature, 513, 336 \\
-$[7]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
-$[8]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21 \\
-$[9]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
-$[10]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, resubmitted after referee report \\
-$[11]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25
+\begin{multicols}{2}
+\noindent 
+$[1]$ Bell, K. R. \& Lin, D. N. C. 1994, ApJ, 427, 987 \\
+$[2]$ Bodenheimer, P.\&Pollack, J. 1986, Icarus, 67, 391 \\
+$[3]$ Ciesla, F. J. \& Cuzzi, J. N. 2006, Icarus, 181, 178 \\
+$[4]$ D'Alessio, P., et al. 2001, ApJ, 553, 321 \\
+$[5]$ Fayolle, E., et al. (in prep) \\
+$[6]$ Henning, T.,\&Semenov, D. 2013, ChRv, 113, 9016 \\
+$[7]$ Lee, E. J., et al. 2014, ApJ, 797, 95 \\ 
+$[8]$ Lee, E. J., \& Chiang, E. 2015, ApJ, 811, 41 \\
+$[9]$ Lodders, K. 2009, arXiv:0910.0811 \\
+$[10]$ Marois, C.,  et al.. 2008, Science, 322, 1348 \\
+$[11]$ Molli{\`e}re, P., et al. 2015, ApJ, 813, 1 \\
+$[12]$ \"Oberg, K. I., et al. 2011, ApJ, 743, L16 \\
+$[13]$ \textbf{Piso, A.-M. A.},\&Youdin,A.N. 2014,ApJ,786,21 \\
+$[14]$ \textbf{Piso, A.-M. A.}, et al. 2015, ApJ, 800, 82 \\
+$[15]$ \textbf{Piso, A.-M. A.}, et al. ApJ (under review) \\
+$[16]$ \textbf{Piso, A.-M. A.}, et al. (in prep) \\
+$[17]$ Pollack, et al. 1985, Icarus, 64, 471 \\
+$[18]$ Pollack, et al. 1996, Icarus, 124, 62 \\
+$[19]$ Qi, C., et al. 2013, Science, 341, 630 \\
+$[20]$ Rafikov, R. R. 2006, ApJ, 648, 666 \\
+$[21]$ Saumon, et al. 1995, ApJS, 99, 713 \\
+%Stevenson, D. J. 1982, Planet. Space Sci., 30, 755 \\
+$[22]$ Zhang, K., et al. 2013, ApJ, 766, 82
+\end{multicols}
 %\end{thebibliography}

### 2015-11-05 Postdoc Applications/hubble_proposal.tex
update

@@
 
-\title{Hubble Fellowship Research Statement: \\
+\title{Hubble Fellowship Research Proposal: \\
 The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets}
@@
 
-\vspace{-0.9cm}
+\vspace{-0.5cm}
 
-Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered $[2]$.
-Their diversity in terms of mass, radius, location and composition $[6]$ provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life.
+Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered $[1]$.
+Their diversity in terms of mass, radius, location and composition $[2]$ provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life.
 For this purpose, it is thus crucial to explore and understand how planets obtain their compositions.
@@
 Gas giants contain most of their mass in their atmosphere, hence their chemical composition is determined by that of their envelopes.
-The last few years have seen a substantial increase in the number of giant planets with observed atmospheric spectra (e.g., $[3]$, $[5]$), which has enhanced our understanding of these planets' chemical structure, and has provided us with quantitative information about the abundances of various compounds in their envelopes besides hydrogen and helium.
+The last few years have seen a substantial increase in the number of giant planets with observed atmospheric spectra (e.g., $[3]$, $[4]$), which has enhanced our understanding of these planets' chemical structure, and has provided us with quantitative information about the abundances of various compounds in their envelopes besides hydrogen and helium.
 Finally, gas giants shape the architecture of planetary systems and affect the delivery of volatile compunds to terrestrial planets, which has direct consequences for the habitability of worlds similar to our own.
+Thus testing theories of planet formation against gas giant compositions will help constrain planet formation theories more generally.
 
-Both terrestrial and giant planets are born in protoplanetary disks, which implies that their \textbf{composition is determined by and tightly linked to the structure and composition of the disk}.
-The chemical and dynamical evolution of disks, as well the formation of giant planets have both been previously investigated.
+Both terrestrial and giant planets are born in protoplanetary disks, which implies that their \textbf{compositions are determined by and tightly linked to the structure and composition of the disk}.
+The chemical and dynamical evolution of disks, as well the formation of giant planets have both been previously investigated in isolation.
 However, the coupled chemo-dynamical disk evolution, planet compositions, and most importantly the disk-planet connection have not yet been considered in detail.
-As shown in my work on the minimum core mass of gas giants ($[8]$, $[9]$), planet formation depends sensitively on disk physics and chemistry.
+As shown in my work on the minimum core mass of gas giants ($[5]$, $[6]$), planet formation depends sensitively on disk physics and chemistry.
 \textbf{I propose to develop a holistic chemo-dynamical framework to explore how disk dynamics and chemistry, as well as the dynamics of nascent planets and planetesimals, regulate the compositions of mature giant planets.} Such a model will enhance our understanding of planetary structures by enabling us to predict what kind of planet compositions result from planet formation in different parts of the disk.
@@
 It follows that these two effects cannot be decoupled, and that the chemo-dynamical evolution of the disk has to be studied simultaneously and self-consistently.
-Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research $[10]$.
-Such effects include \textbf{particle growth and fragmentation}, as well as \textbf{accretion rate and stellar luminosity evolution}.
+Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research $[7]$.
+Such effects include particle growth and fragmentation, as well as accretion rate and stellar luminosity evolution.
 
@@
 I will couple the dynamical framework outlined above with \textbf{time-dependent chemical models of increasing complexity}, informed by results from state-of-the-art disk chemistry models (that can only be run on static disks).
-This will show how the \textbf{snowline locations of volatiles}, as well as the \textbf{chemical composition of the disk gas and dust evolve}, which has \textbf{direct implications on the compositions of young planets}.
+This will show how the {snowline locations of volatiles, as well as the chemical composition of the disk gas and dust evolve, which has \textbf{direct implications on the compositions of young planets}.
 Moreover, since I am primarily interested in processes that affect the disk mid-plane where planets form, my results could be simplified by the fact that certain chemical processes may not be important in this disk parameter space.
@@
 Giant planets can migrate through the disk while still accumulating gas.
-Figure 2 shows an exciting ALMA observation of a planetary gap in the disk around HL Tau $[1]$.
+Figure 2 shows an exciting ALMA observation of a planetary gap in the disk around HL Tau $[8]$.
 This will change a planet's atmospheric composition since the disk chemical abundances are different at different disk locations.
-Additionally, giant planets may still accumulate planetesimals while accreting nebular gas [7].
+Additionally, giant planets may still accumulate planetesimals while accreting nebular gas $[9]$.
 The final composition of a planet's atmosphere will thus depend on how much gas and solids are accreted in this stage.
@@
 The results from parts 1 and 2 will feed into a \textbf{large planet synthesis model}, in which I will use a grid of different \textbf{initial disk and planetary embryo conditions}.
-I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[4]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations.
+I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[10]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations.
 In contrast, for this computationally expensive step, I will only include the processes that I have identified to be the most important in the local simulations from steps 1 and 2.
 This will allow me to \textbf{constrain a planet's formation location based on its chemical composition}.
-Comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[5]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.
+Comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[4]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.
 
@@
 I would also love to work with \textbf{Hilke Schlichting} (proposed faculty host), a leader in planet formation theory and dynamics.
-Margaret Pan, an excellent theorist in planetary dynamics and portoplanetary disk, will join the EAPS department as a research scientist in early 2016, so it would be a pleasure to work with her as well.
+Margaret Pan, an excellent theorist in planetary dynamics and protoplanetary disks, will join the EAPS department as a research scientist in early 2016, so it would be a pleasure to work with her as well.
 Finally, in order to connect my theoretical work with observations, I would like to collaborate with Joshua Winn, who is an expert in discovering and characterizing exoplanets.
@@
 \noindent \textbf{2.
-University of Chicago and University of Michigan} The University of Chicago, my second choice for a host institution, would also be an ideal place for me to undertake my proposed research, due to its opportunities for valuable collaborations with experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
+University of Chicago}.
+The University of Chicago would also be an ideal place for me to undertake my proposed research, due to its opportunities for valuable collaborations with experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
 \textbf{Fred Ciesla} (proposed faculty host) is an expert in protoplanetary disk dynamics, as well as chemical composition and evolution.
@@
 Additionally, the University of Chicago Department of Astronomy hosts leaders in detecting and characterizing worlds outside the Solar system, such as Dan Fabrycky and Jacob Bean, which presents great prospects in connecting my theoretical research work with observations.
-Finally, The University of Michigan, my third choice for a host institution, would be an excellent place to pursue my postdodcotral research due to its plethora of protoplanetary disks and exoplanet experts.
+\\
+
+\noindent \textbf{3.
+University of Michigan}.
+The University of Michigan would be an excellent place to pursue my postdodcotral research due to its plethora of protoplanetary disks and exoplanet experts.
 \textbf{Edwin Bergin} (proposed faculty host) is an expert in modeling protplanetary disk chemistry, so I would greatly benefit from collaborating with him on this part of my proposed research.
@@
 %\vspace{-0.5in}
-\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[1]$).}
+\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[8]$).}
 \label{fig:HLTau}
@@
 The white dots are the measured transit depths, the blue squares and dark blue line are the best fit model, while the shaded regions represent 1- and 2 $\sigma$ intervals in the retrieved spectrum.
-The increase in transit depth around $\sim$1.4 $\mu$m is evidence for a water feature in the atmosphere (from $[5]$).}
+The increase in transit depth around $\sim$1.4 $\mu$m is evidence for a water feature in the atmosphere (from $[4]$).}
 \label{fig:HLTau}
@@
 %\begin{thebibliography} {9}
-\noindent $[1]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
-$[2]$ Batalha, N. M. 2014, Proceedings of the National Academy of Science, 111, 12647 \\
+\noindent $[1]$ Batalha, N. M. 2014, Proceedings of the National Academy of Science, 111, 12647 \\
+$[2]$ Lissauer J. J., Dawson R. I., Tremaine S., 2014, Nature, 513, 336 \\
 $[3]$ Debes, J. H., Jang-Condell, H., Weinberger, A. J., Roberge, A., \& Schneider, G. 2013, ApJ, 771, 45 \\
-$[4]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
-$[5]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ArXiv e-prints, arXiv:1504.05586 \\
-$[6]$ Lissauer J. J., Dawson R. I., Tremaine S., 2014, Nature, 513, 336 \\
-$[7]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
-$[8]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21 \\
-$[9]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
-$[10]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, resubmitted after referee report \\
+$[4]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ArXiv e-prints, arXiv:1504.05586 \\
+$[5]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21 \\
+$[6]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
+$[7]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, resubmitted after referee report \\
+$[8]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
+$[9]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
+$[10]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
 $[11]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25

### 2015-11-05 Postdoc Applications/previous_research_hubble.tex
update

--- 2015-11-04 Postdoc Applications/previous_research.tex
@@
 \lhead{Ana-Maria Piso}
-\rhead{Sagan Fellowship Previous and Current Research Statement}
+\rhead{Hubble Fellowship Previous and Current Research Statement}
 \date{}
@@
 
-\title{\Large Sagan Fellowship Previous and Current Research Statement: \\
+\title{\Large Hubble Fellowship Previous and Current Research Statement: \\
 Giant Planet Formation and Snowlines in Protoplanetary Disks\vspace{-2ex}}
@@
 
-Gas giants are widely believed to form through core accretion $[18]$, a theory in which solid protoplanetary cores grow large enough to accumulate a massive atmosphere.
+Gas giants are widely believed to form through core accretion $[1]$, a theory in which solid protoplanetary cores grow large enough to accumulate a massive atmosphere.
 In this model, planetesimals in a disk grow larger through collisions, eventually forming a planetary embryo, which continues to grow by attracting planetesimals in its neighborhood.
@@
 Core accretion is particularly challenging in the outer parts of a disk, where long dynamical times make it difficult for a core to grow fast enough before the gas disk dissipates on a timescale of a few million years.
-At the same time, however, giant planets on wide orbits have been discovered in recent years, such as the HR 8799 system (Marois et al.
-2008).
+At the same time, however, giant planets on wide orbits have been discovered in recent years, such as the HR 8799 system $[2]$.
 This poses an intriguing question: how did these planets form?
-I addressed this issue in two papers, $[13]$ and $[14]$.
+I addressed this issue in two papers, $[3]$ and $[4]$.
 Specifically, I answered the following question: what is the lowest possible core mass required to form a gas giant before disk dissipation?
 This minimum core mass applies when  solid cores no longer accrete planetesimals, as this would heat the gaseous envelope and inhibit the atmosphere's ability to cool and contract .
-Most state-of-the-art models, however, assume that a giant planet's core and atmosphere grow simultaneously ($[2]$, $[20]$), and thus have not been constructed to effectively explore the value of the true minimum core mass required to form a gas giant.
-We start investigating this minimum (critical) core mass $M_{\rm crit}$ in $[13]$, where we develop a core accretion model which assumes that planetesimal accretion is negligible, and instead the envelope accretes gas while undergoing Kelvin-Helmholtz contraction.
+Most state-of-the-art models, however, assume that a giant planet's core and atmosphere grow simultaneously ($[5]$, $[6]$), and thus have not been constructed to effectively explore the value of the true minimum core mass required to form a gas giant.
+We start investigating this minimum (critical) core mass $M_{\rm crit}$ in $[3]$, where we develop a core accretion model which assumes that planetesimal accretion is negligible, and instead the envelope accretes gas while undergoing Kelvin-Helmholtz contraction.
 %This gives a lower limit on the minimum (critical) core mass, Mcrit, to form a giant planet, since additional planetesimal accretion in this stage would heat up the atmosphere, inhibit its ability to cool, thus increasing the critical core mass.
@@
 We consider envelopes accreting around fully formed cores at distances between 5 and 100 astronomical units (AU) from their host star.
-We use a simple model that assumes the nebular gas to be ideal and polytropic, and that the dust opacity is the standard interstellar opacity (ISM; $[1]$).
+We use a simple model that assumes the nebular gas to be ideal and polytropic, and that the dust opacity is the standard interstellar opacity (ISM; $[7]$).
 We find $M_{\rm crit}$ to decrease between 8.5 Earth Masses ($M_{\oplus}$) at 5 AU to 3.5 $M_{\oplus}$ at 100 AU.
@@
 %(Stevenson 1982). Our study hence makes great strides in understanding the formation of giant planets at large separations. Moreover, 
-We note that our numerical technique has been used by $[7]$ and $[8]$ to understand planet formation in the inner disk.
+We note that our numerical technique has been used by $[8]$ and $[9]$ to understand planet formation in the inner disk.
 
-The simplified model developed in $[13]$ assumes an ideal equation of state (EOS) and an ISM opacity.
-To obtain more robust quantitative results for $M_{\rm crit}$, we expand this model in $[14]$ by making two important additions: (1) a realistic EOS, and (2) realistic dust opacities.
-We use the $[21]$ EOS tables, and we extend them to lower temperatures and pressures as suitable for the parameter space we are exploring.
-Moreover, we use realistic opacity tables that take into account grain growth $[4]$ to calculate $M_{\rm crit}$.
+The simplified model developed in $[3]$ assumes an ideal equation of state (EOS) and an ISM opacity.
+To obtain more robust quantitative results for $M_{\rm crit}$, we expand this model in $[4]$ by making two important additions: (1) a realistic EOS, and (2) realistic dust opacities.
+We use the $[10]$ EOS tables, and we extend them to lower temperatures and pressures as suitable for the parameter space we are exploring.
+Moreover, we use realistic opacity tables that take into account grain growth $[11]$ to calculate $M_{\rm crit}$.
 We find that a realistic EOS increases $M_{\rm crit}$, while grain growth opacities, in contrast, decrease $M_{\rm crit}$.
 Taking these two competing effects together, we calculate $M_{\rm crit} \sim 8 M_{\oplus}$ at 5 AU, decreasing to 5 $M_{\oplus}$ at 100 AU (Figure 1).
-Our results are still lower than 10 $M_{\oplus}$, the typically quoted value, and may be up to one order of magnitude lower if grain coagulation is taken into account (see $[14]$ for details).
+Our results are still lower than 10 $M_{\oplus}$, the typically quoted value, and may be up to one order of magnitude lower if grain coagulation is taken into account (see $[4]$ for details).
 Thus our studies clearly challenge previous claims that core accretion cannot operate in the outer parts of protoplanetary disks, reopening the case for in situ formation of wide-separation gas giants.
@@
   %\vspace{-0.1in}
-  \caption{$M_{\rm crit}$ as a function of semi-major axis for a realistic EOS, and both ISM and grain growth opacities. Grain growth significantly decreases $M_{\rm crit}$ (from $[14]$).}
+  \caption{$M_{\rm crit}$ as a function of semi-major axis for a realistic EOS, and both ISM and grain growth opacities. Grain growth significantly decreases $M_{\rm crit}$ (from $[4]$).}
   %\vspace{-0.3in}
@@
 In this part of my dissertation and beyond, I want to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) back-track the planet formation location based on planet composition.
-While disk structure is very complex, we do have valuable observational evidence of molecular abundance detections in disks $[6]$, as well as volatile snowlines ($[19]$, $[22]$).
+While disk structure is very complex, we do have valuable observational evidence of molecular abundance detections in disks $[12]$, as well as volatile snowlines ($[13]$, $[14]$).
 \\
@@
 
-One important signature of disk and exoplanet chemistry is the carbon-to-oxygen (C/O) ratio, as small variations of the C/O ratio may affect the abundance of other volatiles by several orders of magnitudes (e.g., $[11]$ for the case of exoplanet atmospheres).
+One important signature of disk and exoplanet chemistry is the carbon-to-oxygen (C/O) ratio, as small variations of the C/O ratio may affect the abundance of other volatiles by several orders of magnitudes (e.g., $[15]$ for the case of exoplanet atmospheres).
 Specifically, an important consequence of volatile condensation and sublimation in disks is that disks are expected to present different C/O ratios in the gas and in the icy dust mantles at different disk radii.
-This effect was quantified by $[12]$, who considered the fact that the main carries of carbon and oxygen, i.e.
+This effect was quantified by $[16]$, who considered the fact that the main carriers of carbon and oxygen, i.e.
 H$_2$O, CO$_2$ and CO, have different condensation temperatures.
@@
 
-In $[15]$, we expand this model to include the effect of radial drift of solids and viscous gas accretion on the volatile snowline locations.
+In $[17]$, we expand this model to include the effect of radial drift of solids and viscous gas accretion on the volatile snowline locations.
 We use a simple, semi-analytical model to solve for the coupled drift-desorption equations for particles of different initial sizes, as well as for various disk models (irradiated, evolving, viscous).
@@
 %The largest desorbing particles in our model have initial sizes of ~7 m, and they are representative for the snowline locations, as realistic grain size distributions are dominated in mass by the largest particles.
-By comparing the innermost snowlines in Figure 2, which are set by the largest desorbing particles ($\sim$7 m in our model), we obtain a very powerful result: radial drift and gas accretion may move the H$_2$O, CO$_2$ and CO snowlines by up to 40-60\% compared to a static disk, which is a huge difference.
+By comparing the innermost snowlines in Figure 2, which are set by the largest desorbing particles ($\sim$7 m in our model), we obtain a very powerful result: radial drift and gas accretion may move the H$_2$O, CO$_2$ and CO snowlines by up to 40-60\% compared to a static disk.
 For a transition disk, i.e.
 a disk with an inner cavity significantly depleted of gas, we find that H$_2$O particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the gap edge, due to the decrease in gas pressure inside the cavity.
-This is qualitatively consistent with the observations of $[22]$, which show that the H$_2$O snowline is pushed outwards in a transition disk compared to a full disk.
+This is qualitatively consistent with the observations of $[14]$, which show that the H$_2$O snowline is pushed outwards in a transition disk compared to a full disk.
 Our model framework is thus generally valid for more complicated disk structures as well.
-We plan to further improve this model by adding several disk processes, such as turbulent diffusion, particle composition and grain morphology.
+We plan to expand this model by adding several disk processes, such as turbulent diffusion, particle composition and grain morphology.
 \begin{wrapfigure}{r}{0.45\textwidth}
@@
 Compared to a static disk (black lines), radial drift and gas accretion may push the snowlines inwards by 40-60\%.
-The arrows show that the C/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in disks that take into account gas accretion, as the relative fluxes of the desorbed icy particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (from $[15]$).}
+The arrows show that the C/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in disks that take into account gas accretion, as the relative fluxes of the desorbed icy particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (from $[17]$).}
   \vspace{-0.1in}
@@
 
-Aside from the main C and O carriers, nitrogen (N) is an important molecule to study, as it is highly abundant in the Solar system $[9]$ and disks, and primarily found as N$_2$.
-Gas phase enhancements in the nitrogen-to-oxygen ratio (N/O) are predicted to greatly exceed the C/O ratios in the outer disk regions.
+Aside from the main C and O carriers, nitrogen (N) bearing species are important to study.
+Nitrogen is highly abundant in the Solar system $[18]$ and disks, and primarily found as N$_2$.
+Because of the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio.
 Giant planets that form at wide separations should thus have an excess of N in their atmospheres, which could be used to trace their formation origin.
-In $[16]$, we quantify this effect in disks.
-We find that the N/O ratio in gas is significantly larger than the Solar abundance, and indeed exceeds the C/O ratio in the outer disk, in agreement with predictions.
+In $[19]$, we quantify this effect in disks.
+We find that the N/O ratio in gas is significantly larger than the Solar abundance, and indeed exceeds the C/O ratio in the outer disk.
 We can thus set the stage to answer one of our main questions, i.e.
 back-track the planet formation location based on planet composition.
-Moreover, we use new results from $[5]$ for the binding energies of CO and N$_2$, mixed with H$_2$O or as pure ices.
-We find that the binding energies of water-ice mixtures are significantly larger than the pure ice ones, and move the CO and N$_2$ snowlines by at least a few tens of AU.
+In this study we also explore the effects of assuming that N$_2$ is bound to water ice instead of pure N$_2$, and the presence of some N in the form of NH$_3$.
+In either case, the general result of a large N/O enhancement in the outer disk is preserved.
 \\
@@
 Next, we will treat  each species in gaseous and solid forms as two fluids that are interchanging.
-Following $[3]$, we will use advection-like equations to solve for their separate time-dependent abundance.
+Following $[20]$, we will use advection-like equations to solve for their separate time-dependent abundance.
 
@@
 \noindent 
-$[1]$ Bell, K. R. \& Lin, D. N. C. 1994, ApJ, 427, 987 \\
-$[2]$ Bodenheimer, P.\&Pollack, J. 1986, Icarus, 67, 391 \\
-$[3]$ Ciesla, F. J. \& Cuzzi, J. N. 2006, Icarus, 181, 178 \\
-$[4]$ D'Alessio, P., et al. 2001, ApJ, 553, 321 \\
-$[5]$ Fayolle, E., et al. (in prep) \\
-$[6]$ Henning, T.,\&Semenov, D. 2013, ChRv, 113, 9016 \\
-$[7]$ Lee, E. J., et al. 2014, ApJ, 797, 95 \\ 
-$[8]$ Lee, E. J., \& Chiang, E. 2015, ApJ, 811, 41 \\
-$[9]$ Lodders, K. 2009, arXiv:0910.0811 \\
-$[10]$ Marois, C.,  et al.. 2008, Science, 322, 1348 \\
-$[11]$ Molli{\`e}re, P., et al. 2015, ApJ, 813, 1 \\
-$[12]$ \"Oberg, K. I., et al. 2011, ApJ, 743, L16 \\
-$[13]$ \textbf{Piso, A.-M. A.},\&Youdin,A.N. 2014,ApJ,786,21 \\
-$[14]$ \textbf{Piso, A.-M. A.}, et al. 2015, ApJ, 800, 82 \\
-$[15]$ \textbf{Piso, A.-M. A.}, et al. ApJ (under review) \\
-$[16]$ \textbf{Piso, A.-M. A.}, et al. (in prep) \\
-$[17]$ Pollack, et al. 1985, Icarus, 64, 471 \\
-$[18]$ Pollack, et al. 1996, Icarus, 124, 62 \\
-$[19]$ Qi, C., et al. 2013, Science, 341, 630 \\
-$[20]$ Rafikov, R. R. 2006, ApJ, 648, 666 \\
-$[21]$ Saumon, et al. 1995, ApJS, 99, 713 \\
+$[1]$ Pollack, et al. 1996, Icarus, 124, 62 \\
+$[2]$ Marois, C.,  et al.. 2008, Science, 322, 1348 \\
+$[3]$ \textbf{Piso, A.-M. A.},\&Youdin,A.N. 2014,ApJ,786,21 \\
+$[4]$ \textbf{Piso, A.-M. A.}, et al. 2015, ApJ, 800, 82 \\
+$[5]$ Bodenheimer, P.\&Pollack, J. 1986, Icarus, 67, 391 \\
+$[6]$ Rafikov, R. R. 2006, ApJ, 648, 666 \\
+$[7]$ Bell, K. R. \& Lin, D. N. C. 1994, ApJ, 427, 987 \\
+$[8]$ Lee, E. J., et al. 2014, ApJ, 797, 95 \\ 
+$[9]$ Lee, E. J., \& Chiang, E. 2015, ApJ, 811, 41 \\
+$[10]$ Saumon, et al. 1995, ApJS, 99, 713 \\
+$[11]$ D'Alessio, P., et al. 2001, ApJ, 553, 321 \\
+$[12]$ Henning,T.,\&Semenov, D. 2013, ChRv, 113, 9016 \\
+$[13]$ Qi, C., et al. 2013, Science, 341, 630 \\
+$[14]$ Zhang, K., et al. 2013, ApJ, 766, 82 \\
+$[15]$ Molli{\`e}re, P., et al. 2015, ApJ, 813, 1 \\
+$[16]$ \"Oberg, K. I., et al. 2011, ApJ, 743, L16 \\
+$[17]$ \textbf{Piso, A.-M. A.}, et al. ApJ (under review) \\
+$[18]$ Lodders, K. 2009, arXiv:0910.0811 \\
+$[19]$ \textbf{Piso, A.-M. A.}, et al. (in prep) \\
+$[20]$ Ciesla, F. J. \& Cuzzi, J. N. 2006, Icarus, 181, 178
+
 %Stevenson, D. J. 1982, Planet. Space Sci., 30, 755 \\
-$[22]$ Zhang, K., et al. 2013, ApJ, 766, 82
+
 \end{multicols}

### 2015-11-05 Postdoc Applications/sagan_proposal.tex
update

@@
 %\usepackage{subeqnarray}
-\usepackage[top=1in, bottom=0.8in, left=1in, right=1in]{geometry}
+\usepackage[top=0.9in, bottom=0.8in, left=1in, right=1in]{geometry}
 \usepackage{natbib}
@@
 
-\title{Sagan Fellowship Research Statement: \\
+\title{Sagan Fellowship Research Proposal: \\
 The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets}
@@
 
-\vspace{-0.8cm}
+\vspace{-0.9cm}
 
-Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered $[2]$.
-Their diversity in terms of mass, radius, location and composition $[6]$ provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life.
+Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered $[1]$.
+Their diversity in terms of mass, radius, location and composition $[2]$ provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life.
 For this purpose, it is thus crucial to explore and understand how planets obtain their compositions.
@@
 I will couple the dynamical framework outlined above with \textbf{time-dependent chemical models of increasing complexity}, informed by results from state-of-the-art disk chemistry models (that can only be run on static disks).
-This will show how the \textbf{snowline locations of volatiles}, as well as the \textbf{chemical composition of the disk gas and dust evolve}, which has \textbf{direct implications on the compositions of young planets}.
+This will show how the snowline locations of volatiles, as well as the chemical composition of the disk gas and dust evolve, which has \textbf{direct implications on the compositions of young planets}.
 Moreover, since I am primarily interested in processes that affect the disk mid-plane where planets form, my results could be simplified by the fact that certain chemical processes may not be important in this disk parameter space.
@@
 \noindent \textbf{2.
-University of Chicago.} The University of Chicago would be an ideal place for me to undertake my proposed research, due to its opportunities for valuable collaborations with experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
+University of Chicago.} The University of Chicago would also be an ideal place for me to undertake my proposed research, due to its opportunities for valuable collaborations with experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
 \textbf{Fred Ciesla} (proposed faculty host) is an expert in protoplanetary disk dynamics, as well as chemical composition and evolution.

### 2015-11-05 Postdoc Applications/AnaMariaPisoResume.tex
update

@@
 \begin{list2}
-\item[] {\bf Teaching Fellow} \hfill {\bf February 2012 - May 2012}
-%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
+
+\item[] {\bf WISTEM Program Mentor} \hfill {\bf September 2013 - present}
 \begin{list1}
-%\vspace{-0.1in}
-\item[] Harvard College class SPU 30: Life as a Planetary Phenomenon \hfill \textit{Cambridge, MA}
-\item[] Course Head: Prof. Dimitar Sasselov
-\item[] Held two weekly two-hour sections
+\item[] Mentor for a Harvard College undergraduate \hfill \textit{Cambridge, MA}
+\end{list1}
+
+\vspace{0.2in}
+
+\item[] {\bf MIT Educational Counselor} \hfill {\bf December 2011 - present}
+\begin{list1}
+\item[] Interviewer for prospective undergraduate students \hfill \textit{Cambridge, MA}
 \end{list1}
@@
 \item[] Taught second grade girls at the Amigos School the class ``Sound \& Light'' \hfill \textit{Cambridge, MA}
-\end{list1}
-
-\vspace{0.2in}
-
-\item[] {\bf WISTEM Program Mentor} \hfill {\bf September 2013 - present}
-\begin{list1}
-\item[] Mentor for a Harvard College undergraduate \hfill \textit{Cambridge, MA}
 \end{list1}
@@
 \item[] Organized and coordinated meetings and activities for two groups of ~10 prospective graduate students each \hfill \textit{Cambridge, MA}
+\end{list1}
+
+\vspace{0.2in}
+
+\item[] {\bf Teaching Fellow} \hfill {\bf February 2012 - May 2012}
+%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
+\begin{list1}
+%\vspace{-0.1in}
+\item[] Harvard College class SPU 30: Life as a Planetary Phenomenon \hfill \textit{Cambridge, MA}
+\item[] Course Head: Prof. Dimitar Sasselov
+\item[] Held two weekly two-hour sections
 \end{list1}
@@
 \item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+University of Chicago Exoplanet Journal Club, Chicago, IL, November 2015 \\
+Seminar speaker \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
 MIT Exoplanet Tea, Cambridge, MA, October 2015 \\
@@
 Center for Integrative Planetary Science Planet and Star Formation Seminar, Berkeley, CA, September 2015 \\
-Seminar speaker \\
+Invited talk \\
 \end{list2}

### 2015-11-05 Postdoc Applications/previous_research.tex
update

@@
 Core accretion is particularly challenging in the outer parts of a disk, where long dynamical times make it difficult for a core to grow fast enough before the gas disk dissipates on a timescale of a few million years.
-At the same time, however, giant planets on wide orbits have been discovered in recent years, such as the HR 8799 system (Marois et al.
-2008).
+At the same time, however, giant planets on wide orbits have been discovered in recent years, such as the HR 8799 system $[2]$.
 This poses an intriguing question: how did these planets form?
-I addressed this issue in two papers, $[13]$ and $[14]$.
+I addressed this issue in two papers, $[3]$ and $[4]$.
 Specifically, I answered the following question: what is the lowest possible core mass required to form a gas giant before disk dissipation?
-This minimum core mass applies when  solid cores no longer accrete planetesimals, as this would heat the gaseous envelope and inhibit the atmosphere's ability to cool and contract .
-Most state-of-the-art models, however, assume that a giant planet's core and atmosphere grow simultaneously ($[2]$, $[20]$), and thus have not been constructed to effectively explore the value of the true minimum core mass required to form a gas giant.
-We start investigating this minimum (critical) core mass $M_{\rm crit}$ in $[13]$, where we develop a core accretion model which assumes that planetesimal accretion is negligible, and instead the envelope accretes gas while undergoing Kelvin-Helmholtz contraction.
+This minimum core mass applies when  solid cores no longer accrete planetesimals, as this would heat the gaseous envelope and inhibit the atmosphere's ability to cool and contract.
+Most state-of-the-art models, however, assume that a giant planet's core and atmosphere grow simultaneously ($[5]$, $[6]$), and thus have not been constructed to effectively explore the value of the true minimum core mass required to form a gas giant.
+We start investigating this minimum (critical) core mass $M_{\rm crit}$ in $[3]$, where we develop a core accretion model which assumes that planetesimal accretion is negligible, and instead the envelope accretes gas while undergoing Kelvin-Helmholtz contraction.
 %This gives a lower limit on the minimum (critical) core mass, Mcrit, to form a giant planet, since additional planetesimal accretion in this stage would heat up the atmosphere, inhibit its ability to cool, thus increasing the critical core mass.

### 2015-11-05 Postdoc Applications/previous_research_hubble.tex
update

@@
 Specifically, I answered the following question: what is the lowest possible core mass required to form a gas giant before disk dissipation?
-This minimum core mass applies when  solid cores no longer accrete planetesimals, as this would heat the gaseous envelope and inhibit the atmosphere's ability to cool and contract .
+This minimum core mass applies when  solid cores no longer accrete planetesimals, as this would heat the gaseous envelope and inhibit the atmosphere's ability to cool and contract.
 Most state-of-the-art models, however, assume that a giant planet's core and atmosphere grow simultaneously ($[5]$, $[6]$), and thus have not been constructed to effectively explore the value of the true minimum core mass required to form a gas giant.

### 2015-11-07 Postdoc Applications/AnaMariaPisoResearchStatement.tex
update

%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
%\documentclass[12pt, preprint]{aastex}
%\documentclass[apj]{emulateapj}
\documentclass[12pt, letterpaper]{article}

%\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
%\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex

%\usepackage{pdfsync}
%\usepackage{subeqnarray}
\usepackage[top=0.9in, bottom=0.6in, left=1in, right=1in]{geometry}
\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage{titling}
\usepackage{sectsty}
\usepackage{sidecap}
\usepackage{placeins}
\usepackage{indentfirst}
\usepackage{wrapfig}
\usepackage[font={footnotesize}]{caption}
\usepackage{multicol}
\setlength{\droptitle}{-7em}
\setlength{\abovecaptionskip}{-2.5ex}
\setlength{\belowcaptionskip}{-2.5ex}
%\pagenumbering{gobble}
\pagestyle{fancy}
\lhead{Ana-Maria Piso}
\rhead{Sagan Fellowship Previous and Current Research Statement}
\date{}
%\rhead{\thepage}

%\bibliographystyle{apj}

\title{\large PCEPS Research Statement: \\
Dynamics and Chemistry: The Disk-Planet Connection \vspace{-2.8ex}}
\author{\normalsize Ana-Maria Piso \vspace{-2.8ex}}

%\newenvironment{packed_item}{
%\begin{itemize}
%  \setlength{\itemsep}{1pt}
%  \setlength{\parskip}{0pt}
%  \setlength{\parsep}{0pt}
%}{\end{itemize}}

\begin{document}
\maketitle

%\slugcomment{Draft Modified \today}

\vspace{-1.5cm}

Within the last two decades, more than one thousand extrasolar planets have been discovered. Their diversity in terms of mass, radius, location and composition provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore and understand how planets obtain their compositions. Both terrestrial and giant planets are born in protoplanetary disks, which implies that their composition is determined by and tightly linked to the structure and composition of the disk. In my past, current and future research, I approach this intricate disk-planet link from three directions: (1) by exploring the core accretion mechanism, specifically calculating the minimum required core mass to form a gas giant before the dissipation of the gas in the protoplanetary disk, (2) by understanding how disk chemistry and dynamics affect the snowline locations of volatiles in disks, and (3) by investigating the role of disk volatile chemistry and dynamics in shaping the compositions of nascent planets.

\noindent \underline{\textbf{Previous Research}}

%Planets are born in protoplanetary disks, which means that their structure and composition are determined by and highly connected to the chemical composition and structure of the disk in which they form. In my current and previous research, I have approached this intricate disk-planet link from two directions: (1) by exploring the core accretion mechanism, specifically calculating the minimum required core mass to form a gas giant before the dissipation of the gas in the protoplanetary disk, and (2) by understanding how disk chemistry and dynamics shape the snowline locations of volatiles in disks, which has direct implications for the chemical composition of extrasolar planet (exoplanet) atmospheres.

\textbf{1. Minimum Core Masses for Giant Planet Formation.} Gas giants are widely believed to form through core accretion, a theory in which solid protoplanetary cores grow large enough to accumulate a massive atmosphere. Core accretion is particularly challenging in the outer parts of a disk, where long dynamical times make it difficult for a core to grow fast enough before the gas disk dissipates on a timescale of a few million years. At the same time, however, giant planets on wide orbits have been discovered in recent years, which poses the question of how these planets have formed. I addressed this issue by calculating the minimum (critical) core mass $M_{\rm crit}$ required to form a giant planet during the lifetime of the protoplanetary disk. This minimum applies when solid cores no longer accrete planetesimals, as this would heat the gaseous envelope and inhibit the atmosphere's ability to cool and contract. I thus considered envelopes accreting around fully formed cores at distances between 5 and 100 astronomical units (AU) from their host star. To obtain robust quantitative results for $M_{\rm crit}$, I assumed a realistic equation of state (EOS) for the nebular gas and realistic opacities that take into account grain growth. I found that $M_{\rm crit}$ decreases with semimajor axis, from 8 Earth masses ($M_{\oplus}$) at 5 AU to 5 $M_{\oplus}$ at 100 AU. These results are lower than the typically quoted value of 10 $M_{\oplus}$, and may be up to one order of magnitude lower if grain coagulation is taken into account. Thus our studies clearly challenge previous claims that core accretion cannot operate in the outer parts of protoplanetary disks, reopening the case for in situ formation of wide-separation gas giants.

In the second part of my talk, I will transition to protoplanetary disks, and to how their composition and evolution may affect the formation and chemical composition of giant planets. As the C/O ratio is an important signature of giant planet atmospheres, I will discuss how the snowline locations of the main C and O carriers, i.e. H2O, CO2 and CO, are affected by radial drift of solids and viscous gas accretion. Compared to a static disk, these two effects move the snowlines inwards by ~50%. This affects the C/O ratio in gas and dust throughout the disk, and thus has direct implications in shaping the composition of nascent giant planets.
\underline{\textbf{B. Nitrogen Abundance and Ice Binding Energies}}

Aside from the main C and O carriers, nitrogen (N) bearing species are important to study. Nitrogen is highly abundant in the Solar system $[18]$ and disks, and primarily found as N$_2$. Because of the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio. Giant planets that form at wide separations should thus have an excess of N in their atmospheres, which could be used to trace their formation origin. In $[19]$, we quantify this effect in disks. We find that the N/O ratio in gas is significantly larger than the Solar abundance, and indeed exceeds the C/O ratio in the outer disk. We can thus set the stage to answer one of our main questions, i.e. back-track the planet formation location based on planet composition. In this study we also explore the effects of assuming that N$_2$ is bound to water ice instead of pure N$_2$, and the presence of some N in the form of NH$_3$. In either case, the general result of a large N/O enhancement in the outer disk is preserved. \\
\underline{\textbf{C. Self-consistent C/N/O Calculation}}

As shown in Figure 2, our current drift-desorption model can only provide us with upper estimates for C/O (or N/O) ratios. We plan to enhance this model by performing the time-dependent calculation that accounts for the relative motion of the desorbed ices and the overall nebular gas, the time evolution of the dust surface density,  and radial gas diffusion; all these effects will change the volatile abundances and the C/O ratio in time. We have developed a code that evolves the dust surface density for a particle of fixed size. Next, we will treat  each species in gaseous and solid forms as two fluids that are interchanging. Following $[20]$, we will use advection-like equations to solve for their separate time-dependent abundance.

\noindent \underline{\textbf{Proposed Research}}

%Within the last two decades, more than one thousand extrasolar planets have been discovered. Their diversity in terms of mass, radius, location and composition provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore and understand how planets obtain their compositions. Both terrestrial and giant planets are born in protoplanetary disks, which implies that their composition is determined by and tightly linked to the structure and composition of the disk. 
As shown in my work on the minimum core mass of gas giants and snowlines in disks, planet formation depends sensitively on disk physics and chemistry. I propose to develop a holistic chemo-dynamical framework to explore how disk dynamics and chemistry, as well as the dynamics of nascent planets and planetesimals, regulate the compositions of mature giant planets. Such a model will enhance our understanding of planetary structures by enabling us to predict what kind of planet compositions result from planet formation in different parts of the disk. Furthermore, this work provides essential context for characterizing the planets that instruments such as the James Webb Space Telescope (JWST) and the Transiting Exoplanet Survey Satellite (TESS) will one day discover.

\textbf{1. Coupled Chemical and Dynamical Disk Evolution.} Chemical and dynamical processes in a protoplanetary disk affect the disk structure and composition, and thus the composition of nascent planets. Chemical abundances vary significantly across a typical disk, due to steep gradients in temperature, density and radiation. The complexity of disk chemistry means that coupling it with dynamical processes is non-trivial. Through analytical and numerical calculations, I will first explore a range of dynamical processes that may affect the distribution of volatiles in disks, expanding and generalizing the framework I developed during my dissertation research. I will couple this dynamical model with a simple chemical network, then use more complex chemical networks to develop a simplified time-dependent chemistry, informed by results from state-of-the-art disk chemistry models (that can only be run on static disks). This will show how the snowline locations of volatiles, as well as the chemical composition of the disk gas and dust evolve, which has direct implications on the compositions of young planets.

\textbf{2. Planet and Planetesimal Migration.} Giant planets can migrate through the disk while still accumulating gas, which will change their atmospheric composition since the disk chemical abundances are different at different disk locations. Additionally, giant planets may still accumulate planetesimals while accreting nebular gas. The final composition of a planet?s atmosphere will thus depend on how much gas and solids are accreted in this stage. I will add planet dynamical effects such as migration and planetesimal accretion in the chemical and dynamical model developed in part 1, and quantify how these processes affect the chemical composition of gas giant envelopes.

\textbf{3. Model Planet Populations.} The results from parts 1 and 2 will feed into a large planet synthesis model, in which I will use a grid of different initial disk and planetary embryo conditions. For this computationally expensive step, I will only include the processes that I have identified to be the most important in the local simulations from steps 1 and 2. This will allow me to constrain a planet?s formation location based on its chemical composition. Comparing my results with future JWST observations of atmospheric spectra will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.

%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.7\textwidth]{CO_abundances}
%%\vspace{-0.5in}
%\caption{TBD}
%\label{fig:CO}
%\end{figure}

%\if\bibinc n
%\bibliography{refs}
%\fi
%\FloatBarrier
%\def\bibfont{\footnotesize}
%\setlength{\bibsep}{0.0pt}

\footnotesize

%\textbf{References} \\
%\begin{thebibliography} {9}
%\begin{multicols}{2}
%\noindent 
%$[1]$ Pollack, et al. 1996, Icarus, 124, 62 \\
%$[2]$ Marois, C.,  et al.. 2008, Science, 322, 1348 \\
%$[3]$ \textbf{Piso, A.-M. A.},\&Youdin,A.N. 2014,ApJ,786,21 \\
%$[4]$ \textbf{Piso, A.-M. A.}, et al. 2015, ApJ, 800, 82 \\
%$[5]$ Bodenheimer, P.\&Pollack, J. 1986, Icarus, 67, 391 \\
%$[6]$ Rafikov, R. R. 2006, ApJ, 648, 666 \\
%$[7]$ Bell, K. R. \& Lin, D. N. C. 1994, ApJ, 427, 987 \\
%$[8]$ Lee, E. J., et al. 2014, ApJ, 797, 95 \\ 
%$[9]$ Lee, E. J., \& Chiang, E. 2015, ApJ, 811, 41 \\
%$[10]$ Saumon, et al. 1995, ApJS, 99, 713 \\
%$[11]$ D'Alessio, P., et al. 2001, ApJ, 553, 321 \\
%$[12]$ Henning,T.,\&Semenov, D. 2013, ChRv, 113, 9016 \\
%$[13]$ Qi, C., et al. 2013, Science, 341, 630 \\
%$[14]$ Zhang, K., et al. 2013, ApJ, 766, 82 \\
%$[15]$ Molli{\`e}re, P., et al. 2015, ApJ, 813, 1 \\
%$[16]$ \"Oberg, K. I., et al. 2011, ApJ, 743, L16 \\
%$[17]$ \textbf{Piso, A.-M. A.}, et al. ApJ (under review) \\
%$[18]$ Lodders, K. 2009, arXiv:0910.0811 \\
%$[19]$ \textbf{Piso, A.-M. A.}, et al. (in prep) \\
%$[20]$ Ciesla, F. J. \& Cuzzi, J. N. 2006, Icarus, 181, 178
%
%
%%Stevenson, D. J. 1982, Planet. Space Sci., 30, 755 \\
%
%\end{multicols}
%\end{thebibliography}

%\bibliographystyle{abbrv}
%\bibliography{refs}

%\if\bibinc y
%\begin{thebibliography}
%\end{thebibliography}
%\fi

\end{document}

### 2015-11-09 Postdoc Applications/AnaMariaPisoResearchProposal.tex
update

--- 2015-11-05 Postdoc Applications/sagan_proposal.tex
@@
 %\usepackage{subeqnarray}
-\usepackage[top=0.9in, bottom=0.8in, left=1in, right=1in]{geometry}
+\usepackage[top=0.9in, bottom=0.6in, left=1in, right=1in]{geometry}
 \usepackage{natbib}
@@
 \usepackage{indentfirst}
+\usepackage{wrapfig}
+\usepackage[font={footnotesize}]{caption}
 \setlength{\droptitle}{-7em}
-\setlength{\abovecaptionskip}{-0.4ex}
-\setlength{\belowcaptionskip}{-0.4ex}
+%\setlength{\abovecaptionskip}{-2.5ex}
+\setlength{\belowcaptionskip}{-1ex}
 %\pagenumbering{gobble}
@@
 \lhead{Ana-Maria Piso}
-\rhead{Sagan Fellowship Research Proposal}
+\rhead{Postdoctoral Position in Exoplanet Theory Research Proposal}
 \date{}
@@
 
-\title{Sagan Fellowship Research Proposal: \\
-The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets}
-\author{Ana-Maria Piso}
+\title{\Large Postdoctoral Position in Exoplanet Theory Research Proposal: \\
+The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets\vspace{-2ex}}
+\author{Ana-Maria Piso \vspace{-2ex}}
 
@@
 
-\vspace{-0.9cm}
+\vspace{-1.5cm}
 
@@
 Furthermore, this work provides essential context for characterizing the gas giants that instruments such as the James Webb Space Telescope (JWST) and the Transiting Exoplanet Survey Satellite (TESS) will one day discover.
-My proposed work has therefore direct relevance to NASA's Exoplanet Exploration Program.
+%My proposed work has therefore direct relevance to NASA's Exoplanet Exploration Program.
 
@@
 Such effects include particle growth and fragmentation, as well as accretion rate and stellar luminosity evolution.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chemical_timescales}
+\vspace{-0.3in}
+\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height.
+The numbered curves are $\log_{10} (\tau)$ in million years (from $[11]$).}
+\label{fig:chemical}
+\end{figure}
 
@@
 The complex processes outlined in part 1 will directly impact the composition and dynamics of forming gas giants, as the latter are born and evolve simultaneously with the disk. 
+\begin{wrapfigure}{r}{0.42\textwidth}
+\vspace{-0.4in}
+\begin{center}
+\includegraphics[width=0.42\textwidth]{HLTau_nrao}
+\end{center}
+%\vspace{-0.5in}
+\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[8]$).}
+\vspace{-0.25in}
+\end{wrapfigure}
 The chemical evolution of the gas disk molecular abundances, as well as disk dynamics, determine how a young planet's atmospheric composition changes in time and with the planet's location, and therefore the final chemical structure of mature planets.
@@
 
-\section*{Proposed Host Institutions}
+\begin{figure}[h!]
+\centering
+\vspace{-0.2in}
+\includegraphics[width=0.9\textwidth]{water_spectrum}
+\vspace{-0.2in}
+\caption{The transmission spectrum of exoplanet WASP-12b measured with the Hubble Space Telescope/Wide Field Camera 3.
+The white dots are the measured transit depths, the blue squares and dark blue line are the best fit model, while the shaded regions represent 1- and 2 $\sigma$ intervals in the retrieved spectrum.
+The increase in transit depth around $\sim$1.4 $\mu$m is evidence for a water feature in the atmosphere (from $[4]$).}
+\label{fig:HLTau}
+\end{figure}
 
-\noindent \textbf{1.
-Massachusetts Institute of Technology (MIT)}.
-I believe MIT to be the best place for me to pursue my postdoctoral research, due to its vibrant community of experts in exoplanet atmospheres and dynamics, both in the Physics and Earth and Planetary Sciences (EAPS) departments.
-In particular, I would like to collaborate with Sara Seager, who is a co-investigator of the TESS mission and a world-leading theorist in atmospheric chemistry, as well as with members of her group, who specialize in a broad range of exoplanet theory and computational topics.
-I would also love to work with \textbf{Hilke Schlichting} (proposed faculty host), a leader in planet formation theory and dynamics.
-Margaret Pan, an excellent theorist in planetary dynamics and protoplanetary disks, will join the EAPS department as a research scientist in early 2016, so it would be a pleasure to work with her as well.
-Finally, in order to connect my theoretical work with observations, I would like to collaborate with Joshua Winn, who is an expert in discovering and characterizing exoplanets.
-\\
+\textbf{The University of Chicago is the ideal place for me to pursue my postdoctoral research}, due to its vibrant community of experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
+I would love the opportunity to work with \textbf{Prof.
+Leslie Rogers}, an expert in planetary interiors and atmospheres --- such a collaboration would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards.
+In the Department of Geophysical Sciences, I would like to collaborate with \textbf{Prof.
+Fred Ciesla}, a leader in protplanetary disk dynamics, chemical composition and evolution.
+Additionally, the Department of Astronomy hosts experts in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof.
+Dan Fabrycky} and \textbf{Prof.
+Jacob Bean}, which presents great prospects in connecting my theoretical research work with observations.
 
-\noindent \textbf{2.
-University of Chicago.} The University of Chicago would also be an ideal place for me to undertake my proposed research, due to its opportunities for valuable collaborations with experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
-\textbf{Fred Ciesla} (proposed faculty host) is an expert in protoplanetary disk dynamics, as well as chemical composition and evolution.
+%\textbf{Fred Ciesla} (proposed faculty host) is an expert in protoplanetary disk dynamics, as well as chemical composition and evolution.
 Leslie Rogers, who will begin her faculty appointment in Fall 2016, is a leader in exoplanet theory, specifically planetary interiors and atmospheres, so I would love the opportunity to collaborate with her on my postdoctoral work.
@@
 %\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chemical_timescales}
-%\vspace{-0.5in}
-\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height.
-The numbered curves are $\log_{10} (\tau)$ in million years (from $[11]$).}
-\label{fig:chemical}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.4\textwidth]{HLTau_nrao}
-%\vspace{-0.5in}
-\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[8]$).}
-\label{fig:HLTau}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.9\textwidth]{water_spectrum}
-%\vspace{-0.5in}
-\caption{The transmission spectrum of exoplanet WASP-12b measured with the Hubble Space Telescope/Wide Field Camera 3.
-The white dots are the measured transit depths, the blue squares and dark blue line are the best fit model, while the shaded regions represent 1- and 2 $\sigma$ intervals in the retrieved spectrum.
-The increase in transit depth around $\sim$1.4 $\mu$m is evidence for a water feature in the atmosphere (from $[4]$).}
-\label{fig:HLTau}
-\end{figure}
 

### 2015-11-09 Postdoc Applications/TAC_proposal.tex
update

--- 2015-11-05 Postdoc Applications/sagan_proposal.tex
@@
 \lhead{Ana-Maria Piso}
-\rhead{Sagan Fellowship Research Proposal}
+\rhead{TAC Postdoctoral Fellowship Research Proposal}
 \date{}
@@
 
-\title{Sagan Fellowship Research Proposal: \\
+\title{TAC Postdoctoral Fellowship Research Proposal: \\
 The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets}
@@
 Furthermore, this work provides essential context for characterizing the gas giants that instruments such as the James Webb Space Telescope (JWST) and the Transiting Exoplanet Survey Satellite (TESS) will one day discover.
-My proposed work has therefore direct relevance to NASA's Exoplanet Exploration Program.
 
@@
 
-\section*{Proposed Host Institutions}
-
-\noindent \textbf{1.
-Massachusetts Institute of Technology (MIT)}.
-I believe MIT to be the best place for me to pursue my postdoctoral research, due to its vibrant community of experts in exoplanet atmospheres and dynamics, both in the Physics and Earth and Planetary Sciences (EAPS) departments.
-In particular, I would like to collaborate with Sara Seager, who is a co-investigator of the TESS mission and a world-leading theorist in atmospheric chemistry, as well as with members of her group, who specialize in a broad range of exoplanet theory and computational topics.
-I would also love to work with \textbf{Hilke Schlichting} (proposed faculty host), a leader in planet formation theory and dynamics.
-Margaret Pan, an excellent theorist in planetary dynamics and protoplanetary disks, will join the EAPS department as a research scientist in early 2016, so it would be a pleasure to work with her as well.
-Finally, in order to connect my theoretical work with observations, I would like to collaborate with Joshua Winn, who is an expert in discovering and characterizing exoplanets.
-\\
-
-\noindent \textbf{2.
-University of Chicago.} The University of Chicago would also be an ideal place for me to undertake my proposed research, due to its opportunities for valuable collaborations with experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
-\textbf{Fred Ciesla} (proposed faculty host) is an expert in protoplanetary disk dynamics, as well as chemical composition and evolution.
-Leslie Rogers, who will begin her faculty appointment in Fall 2016, is a leader in exoplanet theory, specifically planetary interiors and atmospheres, so I would love the opportunity to collaborate with her on my postdoctoral work.
-Additionally, the University of Chicago Department of Astronomy hosts leaders in detecting and characterizing worlds outside the Solar system, such as Dan Fabrycky and Jacob Bean, which presents great prospects in connecting my theoretical research work with observations.
+\textbf{The Theoretical Astrophysics Center at UC Berkeley is the ideal place for me to pursue my postdoctoral research}, due to its vibrant community of expert theorists in protoplanetary disks and exoplanets.
+In particular, I would love to collaborate with \textbf{Prof.
+Eugene Chiang}, a world-leading planet formation theorist.
+Prof.
+Chiang's expertise in various topics related to protoplanetary disks and planet dynamics would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards.
+I would also like to collaborate with \textbf{Prof.
+Alfred Glassgold}, an expert in protoplanetary disk chemistry.
+Additionally, the larger UC Berkeley Department of Astronomy hosts leaders in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof.
+Paul Kalas} and \textbf{Prof.
+James Graham}, Project Leads of The Gemini Planet Imager Exoplanet Survey.
+Fostering such collaborations presents great prospects in connecting my
+theoretical research work with observations.
 

### 2015-11-10 Postdoc Applications/ITC_proposal.tex
update

--- 2015-11-09 Postdoc Applications/TAC_proposal.tex
@@
 \lhead{Ana-Maria Piso}
-\rhead{TAC Postdoctoral Fellowship Research Proposal}
+\rhead{ITC Postdoctoral Fellowship Research Proposal}
 \date{}
@@
 
-\title{TAC Postdoctoral Fellowship Research Proposal: \\
+\title{ITC Postdoctoral Fellowship Research Proposal: \\
 The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets}
@@
 
-\textbf{The Theoretical Astrophysics Center at UC Berkeley is the ideal place for me to pursue my postdoctoral research}, due to its vibrant community of expert theorists in protoplanetary disks and exoplanets.
-In particular, I would love to collaborate with \textbf{Prof.
-Eugene Chiang}, a world-leading planet formation theorist.
-Prof.
-Chiang's expertise in various topics related to protoplanetary disks and planet dynamics would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards.
-I would also like to collaborate with \textbf{Prof.
-Alfred Glassgold}, an expert in protoplanetary disk chemistry.
-Additionally, the larger UC Berkeley Department of Astronomy hosts leaders in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof.
-Paul Kalas} and \textbf{Prof.
-James Graham}, Project Leads of The Gemini Planet Imager Exoplanet Survey.
-Fostering such collaborations presents great prospects in connecting my
-theoretical research work with observations.
+\textbf{The Institute for Theory and Computation is the ideal place for me to pursue my postdoctoral work.} Its research facilities and academic excellence are incomparable.
+Even though I am completing my Ph.D.
+at Harvard, I believe I could continue benefitting as a postdoctoral fellow from the plethora of resources the ITC has to offer, especially given the number of experts in protoplanetary disks and exoplanets that the ITC hosts.
+As an ITC Fellow, I will have the opportunity to foster new collaborations and thrive as a scientist, both during my postdoctoral tenure and beyond.
+I would love to collaborate with \textbf{Prof.
+Dimitar Sasselov}, a world-leading theorist in many aspects of planet formation, atmospheric chemistry and astrobiology.
+I would also like to work with \textbf{Dr.
+Matt Holman}, an expert in planetary dynamics.
+Additionally, both the ITC and the larger CfA research community include leaders in protoplanetary disk observations, such as \textbf{Dr.
+Sean Andrews} and \textbf{Dr.
+David Wilner}, as well as distinguished experts in exoplanet detection and characterization, such as \textbf{Prof.
+Dave Charbonneau} and \textbf{Prof.
+John Johnson}.
+Forming such collaborations presents great prospects in connecting my theoretical research work with observations.
 

### 2015-11-11 Postdoc Applications/AnaMariaPisoResume.tex
update

@@
 \item[] {\bf C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion} \\
-Extreme Solar Systems III, Waikoloa, HI, December 2015 \\
+Extreme Solar Systems III, Waikoloa Village, HI, December 2015 \\
 Poster \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+University of Michigan Astronomy Lunch Talk, Ann Arbor, MI, November 2015 \\
+Seminar speaker \\
 \end{list2}

### 2015-11-11 python/C_to_O_CH4_COCO.py

@@ -4,6 +4,8 @@ to the H2O, CO2 and CO snowlines"""
 
 import numpy as np
 from T_freeze import T_freeze
+from drift_timescales_active_Tdisk import Tdisk
+from utils.constants import Msun
 
 def n_C_O(species, CH4mid = 0, CH4max = 0):
     
@@ -106,7 +108,7 @@ def n(T, elem, CH4mid, CH4max):
             
             
 
-def C_O_ratio(a, T0 = 200, q = 0.62, CH4mid = 0, CH4max = 0):
+def C_O_ratio(a, alpha, Mdot, k0, mu, Mstar = Msun, T0 = 200, q = 0.62, acc = 0, CH4mid = 0, CH4max = 0):
     
     """
     
@@ -131,8 +133,10 @@ def C_O_ratio(a, T0 = 200, q = 0.62, CH4mid = 0, CH4max = 0):
     C/O ratio in grains
     
     """
-    
-    T = T0 * a**(-q)
+    if acc == 0:
+        T = T0 * a**(-q)
+    else:
+        T = Tdisk(a, alpha, Mdot, k0, mu, Mstar, T0, q)
     
     return n(T, 'C', CH4mid, CH4max) / n(T, 'O', CH4mid, CH4max)
                        

### 2015-11-11 python/C_to_O_steady_CO.py

@@ -4,8 +4,8 @@ to the H2O, CO2 and CO snowlines"""
 
 import numpy as np
 from T_freeze import T_freeze
+from drift_timescales import read_drag_file_many
 from drift_timescales_active_Tdisk import Tdisk
-from utils.constants import Msun
 
 def n_C_O(species, CH4mid = 0, CH4max = 0):
     
@@ -45,18 +45,18 @@ def n_C_O(species, CH4mid = 0, CH4max = 0):
         return 0.0555 * 0.9, 0 #5.55% H2O (Oberg+08 Spitzer paper) 
     elif species == 'CH4' and CH4mid == 0 and CH4max == 1:
         return 0.13 * 0.9, 0 #13% H2O (Oberg+08 c2d paper)
-     
 
 
 #in the following we calculate T_freeze for various species
-T_freeze_H20 = T_freeze(18., 5800, 1e6)
-T_freeze_CO2 = T_freeze(44., 2000, 3e5)
-T_freeze_CH4 = T_freeze(16., 1300., 0.0555*1e6) #binding energy from Garrod+06 A&A paper
-T_freeze_CO = T_freeze(28., 1388, 1e6)
-#T_freeze_CH4 = T_freeze(16., 1300., 0.13*1e6)
+s, a, arrayH2O = read_drag_file_many('s_neg3_8_steady_correct.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO2', 50, 50)
+s, a, arrayCO = read_drag_file_many('s_neg3_8_steady_correct_CO_CO.txt', 'CO', 50, 50)
+s, a, arrayCH4 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CH4', 50, 50)
 
 
-def n(T, elem, CH4mid, CH4max):
+
+
+def n(index, r, elem, CH4mid = 0, CH4max = 0):
     
     """
     
@@ -84,31 +84,36 @@ def n(T, elem, CH4mid, CH4max):
     elif elem == 'O':
         i = 1 #flag to simplify the use of the n_C_O function; = 1 for O
         
-    if T >= T_freeze_H20:
+    size = s[index]
+    aH2O = arrayH2O[index][-1]
+    aCO2 = arrayCO2[index][-1]
+    aCO = arrayCO[index][-1] 
+    aCH4 = arrayCH4[index][-1]
+        
+    if r <= aH2O:
[…]

### 2015-11-11 python/C_to_O_steady_water.py

@@ -4,8 +4,8 @@ to the H2O, CO2 and CO snowlines"""
 
 import numpy as np
 from T_freeze import T_freeze
+from drift_timescales import read_drag_file_many
 from drift_timescales_active_Tdisk import Tdisk
-from utils.constants import Msun
 
 def n_C_O(species, CH4mid = 0, CH4max = 0):
     
@@ -45,18 +45,18 @@ def n_C_O(species, CH4mid = 0, CH4max = 0):
         return 0.0555 * 0.9, 0 #5.55% H2O (Oberg+08 Spitzer paper) 
     elif species == 'CH4' and CH4mid == 0 and CH4max == 1:
         return 0.13 * 0.9, 0 #13% H2O (Oberg+08 c2d paper)
-     
 
 
 #in the following we calculate T_freeze for various species
-T_freeze_H20 = T_freeze(18., 5800, 1e6)
-T_freeze_CO2 = T_freeze(44., 2000, 3e5)
-T_freeze_CH4 = T_freeze(16., 1300., 0.0555*1e6) #binding energy from Garrod+06 A&A paper
-T_freeze_CO = T_freeze(28., 1388, 1e6)
-#T_freeze_CH4 = T_freeze(16., 1300., 0.13*1e6)
+s, a, arrayH2O = read_drag_file_many('s_neg3_8_steady_correct.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO2', 50, 50)
+s, a, arrayCO = read_drag_file_many('s_neg3_8_steady_correct_CO_water.txt', 'CO', 50, 50)
+s, a, arrayCH4 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CH4', 50, 50)
 
 
-def n(T, elem, CH4mid, CH4max):
+
+
+def n(index, r, elem, CH4mid = 0, CH4max = 0):
     
     """
     
@@ -84,31 +84,36 @@ def n(T, elem, CH4mid, CH4max):
     elif elem == 'O':
         i = 1 #flag to simplify the use of the n_C_O function; = 1 for O
         
-    if T >= T_freeze_H20:
+    size = s[index]
+    aH2O = arrayH2O[index][-1]
+    aCO2 = arrayCO2[index][-1]
+    aCO = arrayCO[index][-1] 
+    aCH4 = arrayCH4[index][-1]
+        
+    if r < aH2O:
[…]

### 2015-11-11 python/Nitrogen/N_to_O.py

@@ -38,11 +38,11 @@ def n_N_O(species, NH3mid = 0, NH3max = 0):
     elif species == 'silicate':
         return 0, 1.4
     elif species == 'N2' and NH3mid == 0 and NH3max == 0:
-        return 1.12 * 2, 0
+        return 0.8, 0
     elif species == 'N2' and NH3mid == 1 and NH3max == 0:
-        return (1.12 - 0.9 * 0.055) * 2, 0
+        return (0.8 - 0.9 * 0.055), 0
     elif species == 'N2' and NH3max == 1 and NH3mid == 0:
-        return (1.12 - 0.9 * 0.1537) * 2, 0
+        return (0.8 - 0.9 * 0.1537), 0
     elif species == 'NH3' and NH3mid == 0 and NH3max == 0:
         return 0, 0
     elif species == 'NH3' and NH3mid == 1 and NH3max == 0:
@@ -56,7 +56,7 @@ T_freeze_H20 = T_freeze(18., 5800, 1e6)
 T_freeze_NH3 = T_freeze(17., 2965., 5.5e4)
 T_freeze_CO2 = T_freeze(44., 2000, 3e5)
 T_freeze_CO = T_freeze(28., 1388., 1e6)
-T_freeze_N2 = T_freeze(28., 1266., 1.12e6)
+T_freeze_N2 = T_freeze(28., 1266., 0.8e6/2)
 #T_freeze_NH3 = T_freeze(17., 2965., 1.5371e5)
 
 

### 2015-11-11 python/Nitrogen/N_to_O_N2N2.py

@@ -38,11 +38,11 @@ def n_N_O(species, NH3mid = 0, NH3max = 0):
     elif species == 'silicate':
         return 0, 1.4
     elif species == 'N2' and NH3mid == 0 and NH3max == 0:
-        return 1.12 * 2, 0
+        return 0.8, 0
     elif species == 'N2' and NH3mid == 1 and NH3max == 0:
-        return (1.12 - 0.9 * 0.055) * 2, 0
+        return (0.8 - 0.9 * 0.055), 0
     elif species == 'N2' and NH3max == 1 and NH3mid == 0:
-        return (1.12 - 0.9 * 0.1537) * 2, 0
+        return (0.8 - 0.9 * 0.1537), 0
     elif species == 'NH3' and NH3mid == 0 and NH3max == 0:
         return 0, 0
     elif species == 'NH3' and NH3mid == 1 and NH3max == 0:
@@ -56,7 +56,7 @@ T_freeze_H20 = T_freeze(18., 5800, 1e6)
 T_freeze_NH3 = T_freeze(17., 2965., 5.5e4)
 T_freeze_CO2 = T_freeze(44., 2000, 3e5)
 T_freeze_CO = T_freeze(28., 834., 1e6)
-T_freeze_N2 = T_freeze(28., 767., 1.12e6)
+T_freeze_N2 = T_freeze(28., 767., 0.8e6/2)
 #T_freeze_NH3 = T_freeze(17., 2965., 1.5371e5)
 
 

### 2015-11-11 python/N_to_O.py
update

@@ -1,0 +1,155 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+from drift_timescales_active_Tdisk import Tdisk
+from utils.constants import Msun
+
+def n_N_O(species, NH3mid = 0, NH3max = 0):
+    
+    """
+        
+    Returns the abundance of N and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_N:
+        abundance of N with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 0, 1.5
+    elif species == 'CO2':
+        return 0, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0, 0
+    elif species == 'silicate':
+        return 0, 1.4
+    elif species == 'N2' and NH3mid == 0 and NH3max == 0:
+        return 0.8, 0
+    elif species == 'N2' and NH3mid == 1 and NH3max == 0:
+        return (0.8 - 0.9 * 0.055), 0
+    elif species == 'N2' and NH3max == 1 and NH3mid == 0:
+        return (0.8 - 0.9 * 0.1537), 0
+    elif species == 'NH3' and NH3mid == 0 and NH3max == 0:
+        return 0, 0
+    elif species == 'NH3' and NH3mid == 1 and NH3max == 0:
+        return 0.055 * 0.9, 0 #0.055 * n_H2O from Oberg+11 Spitzer paper
+    elif species == 'NH3' and NH3max == 1 and NH3mid == 0:
+        return 0.1537 * 0.9, 0 #0.1537 * n_H2O from Bottinelli+10 Spitzer paper
+
+
+#in the following we calculate T_freeze for various species
+T_freeze_H20 = T_freeze(18., 5800, 1e6)
+T_freeze_NH3 = T_freeze(17., 2965., 5.5e4)
+T_freeze_CO2 = T_freeze(44., 2000, 3e5)
[…]

### 2015-11-11 python/N_to_O_N2N2.py

@@ -4,6 +4,8 @@ to the H2O, CO2 and CO snowlines"""
 
 import numpy as np
 from T_freeze import T_freeze
+from drift_timescales_active_Tdisk import Tdisk
+from utils.constants import Msun
 
 def n_N_O(species, NH3mid = 0, NH3max = 0):
     
@@ -55,8 +57,8 @@ def n_N_O(species, NH3mid = 0, NH3max = 0):
 T_freeze_H20 = T_freeze(18., 5800, 1e6)
 T_freeze_NH3 = T_freeze(17., 2965., 5.5e4)
 T_freeze_CO2 = T_freeze(44., 2000, 3e5)
-T_freeze_CO = T_freeze(28., 1388., 1e6)
-T_freeze_N2 = T_freeze(28., 1266., 0.8e6/2)
+T_freeze_CO = T_freeze(28., 834., 1e6)
+T_freeze_N2 = T_freeze(28., 767., 0.8e6/2)
 #T_freeze_NH3 = T_freeze(17., 2965., 1.5371e5)
 
 
@@ -116,11 +118,11 @@ def n(T, elem, NH3mid = 0, NH3max = 0):
             
             
 
-def N_O_ratio(a, T0 = 200, q = 0.62, NH3mid = 0, NH3max = 0):
+def N_O_ratio(a, alpha, Mdot, k0, mu, Mstar = Msun, T0 = 200, q = 0.62, acc = 0, NH3mid = 0, NH3max = 0):
     
     """
     
-    Returns the N/O ratio as a function of semimajor axis for a given disk
+    Returns the C/O ratio as a function of semimajor axis for a given disk
     temperature profile
     
     Input
@@ -137,12 +139,14 @@ def N_O_ratio(a, T0 = 200, q = 0.62, NH3mid = 0, NH3max = 0):
      
     Output
     ------
-    N/O ratio in gas
-    N/O ratio in grains
+    C/O ratio in gas
+    C/O ratio in grains
     
     """
-    
-    T = T0 * a**(-q)
+    if acc == 0:
+        T = T0 * a**(-q)
+    else:
+        T = Tdisk(a, alpha, Mdot, k0, mu, Mstar, T0, q)    
     
     return n(T, 'N', NH3mid, NH3max) / n(T, 'O', NH3mid, NH3max)
                        

### 2015-11-11 python/N_to_O_steady_N2.py

@@ -4,6 +4,9 @@ to the H2O, CO2 and CO snowlines"""
 
 import numpy as np
 from T_freeze import T_freeze
+from drift_timescales_active_Tdisk import Tdisk
+from utils.constants import Msun
+from drift_timescales import read_drag_file_many
 
 def n_N_O(species, NH3mid = 0, NH3max = 0):
     
@@ -51,17 +54,14 @@ def n_N_O(species, NH3mid = 0, NH3max = 0):
         return 0.1537 * 0.9, 0 #0.1537 * n_H2O from Bottinelli+10 Spitzer paper
 
 
-#in the following we calculate T_freeze for various species
-T_freeze_H20 = T_freeze(18., 5800, 1e6)
-T_freeze_NH3 = T_freeze(17., 2965., 5.5e4)
-T_freeze_CO2 = T_freeze(44., 2000, 3e5)
-T_freeze_CO = T_freeze(28., 834., 1e6)
-T_freeze_N2 = T_freeze(28., 767., 0.8e6/2)
-#T_freeze_NH3 = T_freeze(17., 2965., 1.5371e5)
+s, a, arrayH2O = read_drag_file_many('s_neg3_8_steady_correct.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO2', 50, 50)
+s, a, arrayCO = read_drag_file_many('s_neg3_8_steady_correct_CO_CO.txt', 'CO', 50, 50)
+s, a, arrayN2 = read_drag_file_many('s_neg3_8_steady_correct_N2_N2.txt', 'N2', 50, 50)
+s, a, arrayNH3 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'NH3', 50, 50)
 
 
-
-def n(T, elem, NH3mid = 0, NH3max = 0):
+def n(index, r, elem, NH3mid = 0, NH3max = 0):
     
     """
     
@@ -89,38 +89,45 @@ def n(T, elem, NH3mid = 0, NH3max = 0):
     elif elem == 'O':
         i = 1 #flag to simplify the use of the n_N_O function; = 1 for O
         
-    if T >= T_freeze_H20:
+    size = s[index]
+    aH2O = arrayH2O[index][-1]
+    aCO2 = arrayCO2[index][-1]
+    aCO = arrayCO[index][-1]
+    aN2 = arrayN2[index][-1]
+    aNH3 = arrayNH3[index][-1]
+        
[…]

### 2015-11-11 python/N_to_O_steady_water.py

@@ -1,0 +1,156 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+from drift_timescales_active_Tdisk import Tdisk
+from utils.constants import Msun
+from drift_timescales import read_drag_file_many
+
+def n_N_O(species, NH3mid = 0, NH3max = 0):
+    
+    """
+        
+    Returns the abundance of N and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_N:
+        abundance of N with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 0, 1.5
+    elif species == 'CO2':
+        return 0, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0, 0
+    elif species == 'silicate':
+        return 0, 1.4
+    elif species == 'N2' and NH3mid == 0 and NH3max == 0:
+        return 0.8, 0
+    elif species == 'N2' and NH3mid == 1 and NH3max == 0:
+        return (0.8 - 0.9 * 0.055), 0
+    elif species == 'N2' and NH3max == 1 and NH3mid == 0:
+        return (0.8 - 0.9 * 0.1537), 0
+    elif species == 'NH3' and NH3mid == 0 and NH3max == 0:
+        return 0, 0
+    elif species == 'NH3' and NH3mid == 1 and NH3max == 0:
+        return 0.055 * 0.9, 0 #0.055 * n_H2O from Oberg+11 Spitzer paper
+    elif species == 'NH3' and NH3max == 1 and NH3mid == 0:
+        return 0.1537 * 0.9, 0 #0.1537 * n_H2O from Bottinelli+10 Spitzer paper
+
+
+s, a, arrayH2O = read_drag_file_many('s_neg3_8_steady_correct.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO2', 50, 50)
[…]

### 2015-11-11 python/Nitrogen/T_freeze.py

@@ -1,95 +1,0 @@
-"""This script is used to calculate the vibrational frequency, desorption rate, 
-desorption time, and freezing temperature of a given volatile"""
-
-
-import numpy as np
-from utils.constants import mp, kb
-from scipy.optimize import brentq
-
-def vib_freq(Ex, mx):
-    
-    """
-    Input
-    -----
-    Ex: 
-        binding energy in K
-    mx:
-        molecular weight in proton masses
-        
-    Output
-    ------
-    nu:
-        vibrational frequency in s^-1
-    """
-    
-    return 1.6 * 1e11 * np.sqrt(Ex / mx)
-
-def Rdes(mx, Ex, Tx):
-    
-    return vib_freq(Ex, mx) * np.exp(- Ex / Tx) 
-    
-def tevap(mx, Ex, Tx):
-    
-    return 1 / Rdes(mx, Ex, Tx)
-    
-def v_thermal(T, mu):
-    
-    """
-    Input
-    -----
-    T:
-        temperature in K
-    mu:
-        molecular weight in proton masses
-        
-    Output
-    ------
-    v_thermal:
-        thermal velocity in cm s^-1
-    """
-    
-    return np.sqrt(kb * T / (mu * mp))        
-                
-    
-def T_freeze(mx, Ex, nx, Nx = 1e15, fx = 1):
-    
-    """
-    Input
-    -----
-    mx:
-        molecular weight in proton masses
-    Ex:
-        binding energy in K
-    nx:
-        number density in cm^-3
-    Nx:
-        number of adsorption sites per cm^2 (default 1e15)
-    fx:
-        fraction of occupied adsorption sites (default 1)
-        
-    Output
-    ------
-    Tfreeze:
-        freezing temperature in K
-    """
-    
-    def Tfreezefn(x):
-        
-        """
-        Auxiliary function, defined as Tfreezefn(x) = f(x) - x,
-        where "x" stands for temperature and f(x) is the expression
-        on the RHS of equation (4) in Hollenbach et al. (2009).
-        Needed to find the freezing temperature Tf from 
-        Tfreezefn(Tf) = 0 (i.e., f(Tf) = Tf), because the expression 
-        for T_freeze in equation (4) also depends on Tf through the
-        thermal velocity. 
-        """
-        
-        return Ex * (np.log(4 * Nx * fx * vib_freq(Ex, mx) / \
-            (nx * v_thermal(x, mx))))**(-1) - x
-            
[…]

### 2015-11-11 tex/draft2/ms.tex
update

@@
 %\documentclass[12pt, preprint]{aastex}
+%\documentclass[dvips,12pt]{article}
 \documentclass[apj]{emulateapj}
@@
 
-\title{TBD (\textit{Effect of Radial Drift and Viscous Gas Accretion on Snowline Locations in Protoplanetary Disks: Volatile Abundances and Binding Energies?})}
+\title{TBD (\textit{The role of ice compositions and morphology for snowlines the C/N/O ratios in active disks})}
 
@@
 State again the importance of radial drift and gas accretion on the snowlines location, and that a systematic study of the combination of these two particular effects across the disk has not been done before.
-Then transition to the fact that we provide such a systematic study in Paper I and in this paper: here we enhance the model of Paper I by making three additions: (1) we add N and CH4 in the static chemistry model, (2) we explore how different abundances of CH4 and of the N main carriers (N2 and NH3) affect the C/O  and N/O ratios, and (3) we explore how different binding energies of CO, N2 and (perhaps) CH4 affect the snowline locations.}
+Then transition to the fact that we provide such a systematic study in Paper I and in this paper: here we expand the model of Paper I by making three additions: (1) we add N and CH4 in the static chemistry model, and explore how different abundances of CH4 and of the N main carriers (N2 and NH3) affect the C/O  and N/O ratios, (2) we quantify the effect of radial drift and gas accretion on N2, CH4 and NH3 snowline locations, and (3) we explore how different binding energies of CO, N2, affect the snowline locations.}
 
@@
 
-\emgr{Review the four disk models, desorption model, relevant timescales.
+\emgr{Review disk models, desorption model, relevant timescales.
+State that we use a steady-state disk for the coupled drift-desorption evolution, since it's the most realistic, therefore only summarize the static and steady-state disk.
 Summarize the findings of Paper I, i.e.
-particles of certain sizes desorb instantaneously and at a fixed particle size dependent location.
-Probably split into 2 subsections.}
+particles of certain sizes desorb instantaneously and at a fixed particle size dependent location}
+
+\section{CH4 and C/O Ratios}
+
+\emgr{Discuss observed abundances for CH4 and the choices that we make (no CH4, median value, maximum value).
+Present new binding energies for CO as pure ice and mixed with water.
+Show Figure 1 and discuss how different CH4 abundances and binding energies affect snowline locations and C/O ratio: CO-H2O mixture (though I think it's rather CO layered on top of H2O) moves the CO snowline inward by $\sim$40 AU (will calculate percentages too); the maximum reasonable abundance of CH4 changes the C/O ratio by less than 10\%.
+Show Figure 2 and quantify the effect of drift and accretion on the CH4 snowline compared to a static disk.
+While CH4 has only a modest effect on the C/O ratio in a static disk, this effect may be larger in a viscous disk, as the C gas abundances inside the CH4 snowline may be enhanced due to the differential motion of the desorbed ices and overall nebular gas (refer to Paper I).
+In this study, however, we neglect these effects and therefore do not include CH4 in estimating the C/O ratio.
+Show Figure 2.
+Estimate how much }
+
+\section{Nitrogen Carriers and N/O Ratios}
+
+\emgr{Discuss that nitrogen is abundant in the solar system and disks and primarily found as N2.
+Due to the high volatility of N2, the gas phase N/O ratio in the outer disk may be even more enhanced than the C/O ratio.
+A fraction of the nitrogen abundance may be also carried by NH3.
+Discuss NH3 observed abundances and the choices that we make (no NH3, median, maximum).
+Show }
 

### 2015-11-11 tex/draft2/ms.tex
update

--- 2015-07-29 tex/draft2/ms.tex
@@
 
-\title{TBD}
+\title{The Role of Ice Compositions and Morphology for Snowlines and the C/N/O Ratios in Active Disks}
 
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, et al. } %Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
+\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, \& Jamila Pegues\altaffilmark{2}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-%\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
+\altaffiltext{2}{Department of Astrophysical Sciences, Princeton University}
 
@@
 Importance of volatiles in disks and planetary atmospheres, detections of snowlines in disks, C/O ratios etc.
-State again the importance of radial drift and gas accretion on the snowlines location, and that a systematic study of the combination of these two particular effects across the disk has not been done before.
-Then transition to the fact that we provide such a systematic study in Paper I and in this paper: we enhance the model of Paper I by making two additions: (1) we add nitrogen in the static chemistry model, and (2) we explore how the abundance }
+State again the importance of radial drift and gas accretion on the snowline locations, and that a systematic study of the combination of these two particular effects across the disk has not been done before.
+Then transition to the fact that we provide such a systematic study in Paper I and in this paper.
+Here, we expand the model of Paper I by making three additions: (1) we add N and CH4 in the static chemistry model, and explore how different abundances of CH4 and of the N main carriers (N2 and NH3) affect the C/O  and N/O ratios, (2) we quantify the effect of radial drift and gas accretion on the N2, CH4 and NH3 snowline locations, and (3) we explore how different binding energies of CO and N2 affect their snowline locations.}
+
+\section{Model Review}
+
+\emgr{Review disk models, desorption model, relevant timescales.
+State that we use a steady-state disk for the coupled drift-desorption evolution, since it is the most realistic, therefore only summarize the static and steady-state (viscous) disk.
+Summarize the findings of Paper I, i.e.
+particles of certain sizes desorb instantaneously and at a fixed particle size dependent location.}
+
+\section{CH4 and C/O Ratios}
+
+\emgr{Discuss observed abundances for CH4 and the choices that we make (no CH4, median value, maximum value).
+State that desorption energies for H2O, CO2 and CH4 are well constrained experimentally, and that the CO2 and CH4 binding energies are only weakly dependent on whether it's pure CO2/CH4 or combined with H2O, but that is not the case for CO (and N2 as we will show in the next section).
+Present new binding energies for CO as pure ice and mixed with water.
+Show Figure 1 and discuss how different CH4 abundances and binding energies affect snowline locations and C/O ratio: CO-H2O mixture (though I think it's rather CO layered on top of H2O) moves the CO snowline inward by $\sim$40 AU (will calculate percentages too); the maximum reasonable abundance of CH4 changes the C/O ratio by less than 10\%.
+Show Figure 2 and quantify the effect of drift and accretion on the CH4 snowline compared to a static disk.
+While CH4 has only a modest effect on the C/O ratio in a static disk, this effect may be larger in a viscous disk, as the C gas abundances inside the CH4 snowline may be enhanced due to the differential motion of the desorbed ices and overall nebular gas (refer to Paper I).
+In this study, however, we neglect these effects and therefore do not include CH4 in estimating the C/O ratio (as an aside, the figures that include CH4 in the C/O ratio with drift and desorption are quite messy due to snowlines overlapping).
+Show Figure 3 and estimate the difference between CO-H2O and CO pure ice snowlines in the case of drift and accretion, as well as the comparisons for the static disk for the CO snowline.}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_CH4_test.pdf}
+%\vspace{-0.5in}
+\caption{C/O ratio in a static disk for different CH4 abundances and CO binding energies...} 
+\label{fig:COstatic}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_CH4_steady.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance as a function of initial distance for CH4.... Drift and gas accretion move the CH4 snowline inward by x\%.} 
+\label{fig:CH4}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/C_O_water_ice.pdf}
+%\vspace{-0.5in}
+\caption{C/O ratio as function of semimajor axis for CO combined with H2O (top panel) and pure CO ice (bottom panel)....
+Drift and gas accretion move the CO snowlines inward by x\% and y\%, respectively.}
+\label{fig:CO_ratio}
+\end{figure}
+
+\section{Nitrogen Carriers and N/O Ratios}
+
+\emgr{Similar to the previous section, but with more details.
+Discuss that nitrogen is abundant in the solar system and disks and primarily found as N2.
+Due to the high volatility of N2, the gas phase N/O ratio in the outer disk may be even more enhanced than the C/O ratio.
+A fraction of the nitrogen abundance may be also carried by NH3.
+Discuss NH3 observed abundances and the choices that we make (no NH3, median, maximum).
+State that the NH3 desorption energy is only weakly dependent on whether it's pure NH3 or combined with H2O, but that is not the case for N2.
+Present new binding energies for N2 as pure ice and combined with water.
+Show Figure 4 and discuss how different nitrogen abundances and binding energies affect snowline locations and N/O ratio: N2 combined with H2O moves the N2 snowline inward by $\sim$50 AU (will calculate percentages too); the maximum reasonable abundance of NH3 changes the N/O ratio by $\sim$15\%.
+In the outer disk, the N/O ratio is enhanced by a factor of $\sim$4 compared to the solar value, twice as much as the C/O enhancement.
+Show Figure 5 and quantify the effect of drift and accretion on the NH3 snowline compared to a static disk.
+While NH3 does not have a significant effect on the N/O ratio in a static disk, this effect may be larger in a viscous disk, as the N gas abundance inside the NH3 snowline may be enhanced due to the differential motion of the desorbed ices and overall nebular gas (refer to Paper I).
+In this study, however, we neglect these effects and therefore do not include NH3 in estimating the N/O ratio (again, N/O ratio figure with drift is quite messy when including NH3 and does not add any information that is not already shown in Figures 4 and 5).
+Show Figure 6 and estimate the difference between N2-H2O and N2 pure ice snowlines in the case of drift, as well as the comparisons for the static disk for the N2 snowline.
+State that there will be an overabundance of gas-phase N/O between the CO and N2 snowlines, as there is no oxygen gas in this region.}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/N_O_ratio_test.pdf}
+%\vspace{-0.5in}
+\caption{N/O ratio in a static disk for different NH3 abundances and N2 binding energies...} 
+\label{fig:Nstatic}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_NH3_steady.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance as a function of initial distance for NH3.... Drift and gas accretion move the NH3 snowline inward by x\%.} 
+\label{fig:NH3}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/N_O_water_ice.pdf}
+%\vspace{-0.5in}
+\caption{N/O ratio as function of semimajor axis for N2 combined with H2O (top panel) and pure N2 ice (bottom panel)....
+Drift and gas accretion move the N2 snowlines inward by x\% and y\%, respectively.
+Overabundance of gas-phase N/O between the CO and N2 snowlines, marked by the vertical red dash-dotted lines for the largest drifting particles in our model.}
+\label{fig:NO_ratio}
+\end{figure}
+
+\section{Discussion}
+
+\emgr{Discuss how entrapment of volatiles by H2O affects volatile abundances and C/O ratios.
+Re-emphasize the fact that the C/O and N/O ratios are upper estimates, and that CH4 and NH3 might matter in a viscous disk.
+State that we plan to address this in a future paper.
+More TBD.}
+
+\section{Summary}
+
+\emgr{Maybe we can include the summary in the discussion section?}
+
+%\section{Volatile Abundances and Binding Energies: Effect on C/O and N/O Ratios in a Static Disk}
+%
+%\subsection{Nitrogen and CH4}
+%
+%\emgr{Discuss that nitrogen is abundant in the solar system and should be abundant in disks as well, but that its dominant form is largely unknown.
+Discuss that the main carrier of nitrogen is N2, but some fraction of it can be NH3 as well.
+Present and motivate the choices that we make for NH3 abundances.
+Along the same lines, discuss CH4, and the choices that we make for CH4 abundances.}
+%
+%\subsection{Volatile Desorption Energies}
+%
+%\emgr{State that desorption energies for H2O and CO2 are well constrained experimentally, and that the CO2 binding energy is only weakly dependent on whether it's pure CO2 or combined with H2O, but that is not the case for CO, N2 (and perhaps CH4 and NH3?).
+Briefly discuss CO-CO, N2-N2, CO-H2O and N2-H2O (and perhaps the same for CH4 and NH3 if we find literature on that) experimental results for binding energies, and motivate the choices that we make.}
+%
+%\subsection{Results for C/O and N/O in a Static Disk}
+%
+%\emgr{For each of them (C/O and N/O), show a 3-panel plot as follows: each panel has a specific CH4/NH3 abundance (top: none, middle: median abundance, bottom: maximum abundance); for a given panel, have multiple curves for C/O or N/O, depending on the choice of binding energies, so that it's clear visually how the binding energy changes the snowline location.
+Discuss how different abundances  and binding energies affect snowline locations and C/O or N/O ratios.}
+%\section{Results}
+%
+%\subsection{N2, NH3 and CH4 Snowline Locations}
+%
+%\emgr{One multipanel 3x3 (or 3x2) rainbow plot similar to the snowline plots from Paper I, for \textit{one} choice for the binding energies (perhaps the largest ones, since we want a limit on how far in we can push the snowlines?).
+Rows: snowlines as a function or particle size for passive, active and (maybe) steady-state disk.
+Columns: N2, NH3 and CH4.
+Not entirely certain how necessary these plots/subsection actually are, since it is exactly the same qualitatively as in Paper I...}
+%
+%\subsection{C/O and N/O Ratios}
+%
+%\emgr{For each of them (C/O and N/O) show a 3x3 multipanel plot similar to the C/O plot from Paper I, for the same choice of binding energies as in the previous subsection.
+Columns: C/O or N/O in passive, active and steady-state disk.
+Rows: No CH4/NH3, median CH4/NH3, maximum CH4/NH3.
+Again, not sure if we need/want this for all disk choices, at least in the case of C/O since we already have that in Paper I.
+For N/O, quantify how the snowline location changes from a static disk due to drift and gas accretion.}
 

### 2015-11-12 tex/draft2/ms.tex
update

@@
 
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, \& Jamila Pegues\altaffilmark{2}}
+\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Jamila Pegues\altaffilmark{2}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
@@
 
-\section{Model Review}
+\section{Coupled Drift-Desorption Model Review}
+
+We begin with a brief review of Paper I's model for the effect of radial drift and viscous gas accretion on volatile snowline locations.
+We review our disk model assumptions and relevant timescales in Section \ref{sec:disktime}, and summarize our findings in Section \ref{sec:driftdes}.
+
+\subsection{Disk Model and Important Timescales}
+\label{sec:disktime}
+
+We first assume a static disk, which is only irradiated by the central star and does not experience redistribution of solids or radial movement of the nebular gas.
+To quantify the effects of radial drift and gas accretion, we use a viscous disk with a spatially and temporally constant mass flux, $\dot{M}$.
+The viscous disk takes into account radial drift, gas accretion onto the central star, as well as accretion heating.
+We prefer this disk model to an irradiated or evolving disk (see Paper I) because it includes all the dynamical and thermal processes we are interested in for the scope of this paper, and therefore it is the most realistic one.
+
+We model the static disk as a minimum mass solar nebula (MMSN), using a prescription for the gas surface density, $\Sigma$, and disk midplane temperature, $T$, similar to that of \citet{chiang10}:
+\begin{subeqnarray}
+\label{eq:disk}
+\Sigma&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
+T &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
+\end{subeqnarray}
+where $r$ is the semimajor axis.
+Based on observations of protoplanetary disks \citep{andrews10}, we choose a flatter surface density than the one assumed by \citet{chiang10}, where $\Sigma \propto r^{-3/2}$.
+
+We use the \citet{shakura73}  steady-state disk solution to model the viscous disk.
+Solving the Equation set of Appendix A in Paper I yields an expression for the temperature profile in a steady-state disk:
+\begin{equation}
+\label{eq:Tdact}
+T_{\rm act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
+\end{equation}
+Here $G$ is the gravitational constant, $\kappa_0=2 \times 10^{-6}$ is a dimensionless coefficient in the opacity law $\kappa=\kappa_0 T_{\rm act}^2$, $M_*=M_{\odot}$ is the mass of the central star, $\mu=2.35$ is the mean molecular weight of the nebular gas, $m_{\rm p}$ is the proton mass, $\Omega_{\rm k}=\sqrt{G M_{\odot}/r^3}$ is the Keplerian angular velocity, $\alpha=0.01$ is a dimensionless coefficient (see below for details), $k_{\rm B}$ is the Boltzmann constant, and $\sigma$ is the Stefan-Boltzmann constant.
+The final midplane temperature profile is computed as
+\begin{equation}
+\label{eq:activeT}
+T^4 = T_{\rm act}^4 + T_{\rm irr}^4,
+\end{equation}
+where $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}). We use this expression because in addition to accretion heating, stellar irradiation also contributes to the disk thermal structure.
+
+The steady-state disk has an $\alpha$-viscosity prescription, where the kinematic viscosity is $\mu=\alpha c H$.
+Here $c \equiv \sqrt{k_{\rm B} T /(\mu m_{\rm p})}$ is the isothermal sound speed (with $T$ from Equation \ref{eq:activeT}), and $H \equiv c/\Omega_{\rm k}$ is the disk scale height.
+We can then determine the gas surface density for a viscous disk as (\citealt{shakura73}; see also Paper I for a more detailed explanation of these calculations):
+\begin{equation}
+\label{eq:Sigmaact}
+\Sigma=\frac{\dot{M}}{3 \pi \nu}.
+\end{equation}
+We choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$, consistent with mass flux observations in disks (e.g., \citealt{andrews10}).
+
+In what follows, we summarize the
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_CH4_test.pdf}
+\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_CH4.pdf}
 %\vspace{-0.5in}
@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_CH4_steady.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance as a function of initial distance for CH4.... Drift and gas accretion move the CH4 snowline inward by x\%.} 
-\label{fig:CH4}
-\end{figure}
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_CH4_steady.pdf}
+%%\vspace{-0.5in}
+%\caption{Desorption distance as a function of initial distance for CH4.... Drift and gas accretion move the CH4 snowline inward by x\%.} 
+%\label{fig:CH4}
+%\end{figure}
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/N_O_ratio_test.pdf}
+\includegraphics[width=0.5\textwidth]{../../figs/N_O_ratio.pdf}
 %\vspace{-0.5in}
@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_NH3_steady.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance as a function of initial distance for NH3.... Drift and gas accretion move the NH3 snowline inward by x\%.} 
-\label{fig:NH3}
-\end{figure}
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_NH3_steady.pdf}
+%%\vspace{-0.5in}
+%\caption{Desorption distance as a function of initial distance for NH3.... Drift and gas accretion move the NH3 snowline inward by x\%.} 
+%\label{fig:NH3}
+%\end{figure}
 

### 2015-11-13 Postdoc Applications/AnaMariaPisoResearchProposal.tex
update

@@
 I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[10]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations.
-In contrast, for this computationally expensive step, I will only include the processes that I have identified to be the most important in the local simulations from steps 1 and 2.
-This will allow me to \textbf{constrain a planet's formation location based on its chemical composition}.
+To minimize computational expenses, I will build a program from bottom-up that includes select physical and chemical processes based on the local simulations from steps 1 and 2.
+This will allow me to predict the \textbf{range of planet compositions that can form under reasonable disk conditions}, as well as \textbf{constrain a planet's formation location based on its chemical composition}.
 Comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[4]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.
@@
 Jacob Bean}, which presents great prospects in connecting my theoretical research work with observations.
+\\
 
@@
 
-\section*{References}
+%\section*{References}
 \footnotesize

### 2015-11-13 tex/draft/ms_final_revised_2.tex
update

--- 2015-10-21 tex/draft/ms_final_revised.tex
@@
 We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk.
-We determine the snowline locations for a range of fixed initial particle sizes \textbf{and disk types}.
-\textbf{For our fiducial disk model,} we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk, and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving and viscous disk, desorb %instantaneously and [I think instantaneously is too strong]
+We determine the snowline locations for a range of fixed initial particle sizes and disk types.
+For our fiducial disk model, we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk, and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving and viscous disk, desorb %instantaneously and [I think instantaneously is too strong]
 at a %fixed 
@@
 %location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
-Compared to a static disk, we find that radial drift and gas accretion \textbf{in a viscous disk} move the H$_2$O snowline inwards by up to \textbf{40 \%}, the CO$_2$ snowline by up to 60 \%, and the CO snowline by up to 50 \%.
+Compared to a static disk, we find that radial drift and gas accretion in a viscous disk move the H$_2$O snowline inwards by up to 40 \%, the CO$_2$ snowline by up to 60 \%, and the CO snowline by up to 50 \%.
 We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
@@
 Recent observations of protoplanetary disks have provided valuable information about the abundances and snowline locations of some of these compounds.
-For example, the CO snowline has been detected in the disk around TW Hya \citep{qi13}, \textbf{as well as in the disk around HD 163296 \citep{mathews13} using line emissions from DCO$^+$}.
+For example, the CO snowline has been detected in the disk around TW Hya \citep{qi13}, as well as in the disk around HD 163296 \citep{mathews13} using line emissions from DCO$^+$.
 Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}, and more such snowline detections are expected in future ALMA cycles.
@@
 \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-This effect was used to explain claims of detections of superstellar C/O ratios in exoplanet atmospheres (e.g., WASP-12b, \citealt{madhu11}), \textbf{which however have been unambiguously refuted (\citealt{stevenson14}, \citealt{kreidberg15})}.
+This effect was used to explain claims of detections of superstellar C/O ratios in exoplanet atmospheres (e.g., WASP-12b, \citealt{madhu11}), which however have been unambiguously refuted (\citealt{stevenson14}, \citealt{kreidberg15}).
 %Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.
@@
 Once the grains become large enough to decouple from the gas and drift inwards, the cycle restarts.
-This ```conveyor belt''' model is \textbf{based on the pioneering work by \citet{cuzzi04}}, and \citet{ciesla06} for the evolution of H$_2$O in a viscous disk.
+This ```conveyor belt''' model is based on the pioneering work by \citet{cuzzi04}, and \citet{ciesla06} for the evolution of H$_2$O in a viscous disk.
 This approach leads \citet{alidib14} to find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.
@@
 Each of these studies have considered a specific combination of dynamical and chemical effects.
-One \textbf{scenario} that has not yet been considered is the combination of radial drift and viscous gas accretion \textbf{in isolation}.
+One scenario that has not yet been considered is the combination of radial drift and viscous gas accretion in isolation.
 Studying these two dynamical processes makes it possible to quantify their separate effect on snowline locations and the C/O ratio at various disk radii.
@@
 In Section \ref{sec:model}, we present our disk, radial drift and desorption models, as well as the timescales relevant to the coupled drift-desorption process.
-We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for \textbf{an irradiated and an evolving disk} in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}.
+We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for an irradiated and an evolving disk in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}.
 In Section \ref{sec:discussion}, we discuss the generality of our results, as well as additional effects on the snowline locations.
@@
 
-We present our protoplanetary disk model for \textbf{a static, an irradiated, an evolving, and a viscous disk} in section \ref{sec:disk}.
+We present our protoplanetary disk model for a static, an irradiated, an evolving, and a viscous disk in section \ref{sec:disk}.
 In section \ref{sec:drift}, we describe our analytic model for the radial drift of solids.
@@
 
-To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{static disk}, which is solely irradiated by the host star and does not take into account gas accretion onto the star or radial drift; \textit{\textbf{irradiated} disk}, which has the same temperature profile as the static disk and does not experience gas accretion or accretional heating, but it takes into account radial drift of solids; \textit{\textbf{evolving} disk}, in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; and \textit{\textbf{viscous} disk}, for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis, and the temperature profile is calculated using both accretional heating and stellar irradiation.
+To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{static disk}, which is solely irradiated by the host star and does not take into account gas accretion onto the star or radial drift; \textit{irradiated disk}, which has the same temperature profile as the static disk and does not experience gas accretion or accretional heating, but it takes into account radial drift of solids; \textit{evolving disk}, in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; and \textit{viscous disk}, for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis, and the temperature profile is calculated using both accretional heating and stellar irradiation.
 
@@
 Our choice is inspired by observations of protoplanetary disks at radii larger than $\sim$20 AU (e.g., \citealt{andrews10}), which suggest that typical disks may have surface density profiles with $\Sigma \propto r^{-1}$.
-\textbf{A slope flatter than $\Sigma \propto r^{-3/2}$ is also more consistent with the temperature profile for a steady-state gas disk (see the Viscous disk heading below and \App{app:steadystate}).} We use the static disk model to compare our results with those of \citet{oberg11}.
+A slope flatter than $\Sigma \propto r^{-3/2}$ is also more consistent with the temperature profile for a steady-state gas disk (see the Viscous disk heading below and \App{app:steadystate}).
+We use the static disk model to compare our results with those of \citet{oberg11}.
 
@@
 We choose $M=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
-\textbf{The irradiated and evolving disk surface densities match at $t \approx 5 \times 10^4$ years in the inner disk, but they diverge at distances larger than a few AU due the exponential cutoff in radius of the surface density of the evolving disk (Equation \ref{eq:Sigmaact}).}
+The irradiated and evolving disk surface densities match at \textbf{$t \approx 5 \times 10^5$} years in the inner disk, but they diverge at distances larger than a few AU due the exponential cutoff in radius of the surface density of the evolving disk (Equation \ref{eq:Sigmaact}).
 
@@
 Since
-Calculating the midplane temperature self-consistently for an evolving disk \textbf{that is also actively heated, and thus whose thermal evolution is dominated both by accretion heating and stellar irradiation}, is non-trivial.
+Calculating the midplane temperature self-consistently for an evolving disk that is also actively heated, and thus whose thermal evolution is dominated both by accretion heating and stellar irradiation, is non-trivial.
 We therefore use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm act}$.
@@
 
-\textbf{Before we proceed forward, we note that our disk models assume a constant stellar luminosity $L_*$, as well as a constant mass accretion rate $\dot{M}$ for the viscous disk.
+Before we proceed forward, we note that our disk models assume a constant stellar luminosity $L_*$, as well as a constant mass accretion rate $\dot{M}$ for the viscous disk.
 In reality, the stellar luminosity decreases as the host star contracts, which will reduce the disk temperature and push the snowlines inward, as we explain in Section \ref{sec:neglected}.
@@
 Thus the midplane temperature will not change significantly for our model due to variations in stellar luminosity, but it may decrease substantially for smaller stars, pushing the snowline inward (see Section \ref{sec:neglected} for details).
-Realistic mass accretion rates, $\dot{M}$, may vary between $\sim$$10^{-7}$ and $\sim$$10^{-9}$ $M_{\odot}$ yr$^{-1}$ as the disk evolves (e.g., \citealt{garaud07}).
+Realistic mass accretion rates, $\dot{M}$, may vary between $\sim$$10^{-7}$ and $\sim$$10^{-9}$ $M_{\odot}$ yr$^{-1}$ as the disk evolves \textbf{(e.g., \citealt{chambers09}, \citealt{sicilia10})}.
 For $\dot{M}\lesssim 10^{-9} M_{\odot}$ yr$^{-1}$, the disk becomes optically thin and hence depleted of gas, which means giant planets must have formed before $\dot{M}$ becomes too low.
 \citet{garaud07} find that the snowline locations scale as $r_{\rm snow} \propto \dot{M}^{1/3}$.
-A factor of 100 reduction in the mass accretion rate will thus move the snowlines inwards by a factor of $\sim$$4$.
-However, \citet{garaud07} find that the mass accretion rate is close to $10^{-8} M_{\odot}$ yr$^{-1}$ during the quasi-static accretion phase of the disk (see also Section \ref{sec:neglected}), hence the snowline locations are likely to move inward only in the very early stages of the disk evolution.
+A factor of 100 reduction in the mass accretion rate will thus move the H$_2$O snowline inwards by a factor of $\sim$$4$ \textbf{--- since accretion heating is dominant only in the inner disk, the CO$_2$ and CO snowline locations are unlikely to be affected by changes in mass accretion rate.
+The inward movement of the H$_2$O snowline due to the decrease in $\dot{M}$ may be even larger, by up to one order of magnitude \citep{chambers09}}.
+We thus conclude that changes in $L_*$ throughout time may only modestly affect our results, \textbf{while changes in $\dot{M}$ may significantly affect our results for the H$_2$O snowline, as its location may be determined by the decline in mass accretion rate rather than radial drift.} The time variability of $L_*$ and $\dot{M}$ should be taken into account when drawing more robust conclusions, as well as for different host star and disk properties.
+
+%However, \citet{garaud07} find that the mass accretion rate is close to $10^{-8} M_{\odot}$ yr$^{-1}$ during the quasi-static accretion phase of the disk (see also Section \ref{sec:neglected}), hence the snowline locations are likely to move inward only in the very early stages of the disk evolution.
 We thus claim that changes in $L_*$ and $\dot{M}$ throughout time may only modestly affect our results, but they should still be taken into account when drawing more robust conclusions, as well as for different host star and disk properties.}
@@
 For the viscous disk (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}).
-\textbf{For our choice of parameters for both the evolving and the viscous disks, we have found that the radial flow of gas is always directed inward, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}).
-We note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations.
-Since drifting particles larger than a few cm are only modestly affected by gas accretion, an outward gas flow would move the CO$_2$ and CO locations further away from the star only for the smallest particles in our model, which are well-coupled to the gas.}
+For our choice of parameters for both the evolving and the viscous disks, we have found that the radial flow of gas \textbf{(calculated from Equation \ref{eq:vgas} for the evolving disk and from $\dot{r}_{\rm gas}=-\dot{M}/(2 \pi r \Sigma)$ for the viscous disk)} is always directed inward \textbf{for our parameter space of interest}, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}).
+\textbf{For our evolving disk, the gas starts drifting outwards at a radius $r_{\rm switch} \approx 200 AU$, which is however well outside the CO snowline in our model.} We \textbf{thus} note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations.
+Since drifting particles larger than a few cm are only modestly affected by gas accretion, an outward gas flow would move the CO$_2$ and CO locations further away from the star only for the smallest particles in our model, which are well-coupled to the gas.
 
@@
 We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T$.
-For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively (\textbf{\citealt{collings04}, \citealt{fraser01}, \citealt{aikawa96}}).
+For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively (\citealt{collings04}, \citealt{fraser01}, \citealt{aikawa96}).
 We use the desorption rate, $R_{\rm des}$, to estimate the desorption timescale for particles of different sizes as described in section \ref{sec:timescales}.
@@
 
-\textbf{In the particle size regime} for which (1) $t_{\rm drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
+In the particle size regime for which (1) $t_{\rm drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
 for $\sim$$0.5$ cm $\lesssim s \lesssim$ $1000$ cm, or (2) $t_{\rm gas, acc} \lesssim t_{\rm des} \lesssim t_{\rm d}$, i.e.
-for $\sim$$0.1$ cm $\lesssim s \lesssim$ $10$ cm, radial drift or gas accretion (or both) are faster than thermal desorption, \textbf{ which is of particular interest for our purposes.} We note that $t_{\rm gas, acc}<t_{\rm d}$ always holds true.
+for $\sim$$0.1$ cm $\lesssim s \lesssim$ $10$ cm, radial drift or gas accretion (or both) are faster than thermal desorption, which is of particular interest for our purposes.
+We note that $t_{\rm gas, acc}<t_{\rm d}$ always holds true.
 Particles of sizes that satisfy the requirements above will drift significantly due to radial drift or gas accretion before desorbing, thus moving the H$_2$O, CO$_2$ and CO snowlines closer towards the central star and changing the C/O ratio throughout the disk.
@@
 Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both an irradiated and an evolving disk.
-\textbf{We do not show the results for the viscous disk as they would complicate the plot without adding any qualitative insight --- } the results for the viscous disk are quantitatively similar with those of the evolving disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
+We do not show the results for the viscous disk as they would complicate the plot without adding any qualitative insight ---  the results for the viscous disk are quantitatively similar with those of the evolving disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
 We also show the static snowlines for comparison, which are calculated by balancing adsorption and desorption \citep{hollenbach09}.
@@
 It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model.
-\textbf{Both of these conclusions remain valid for an evolving and a viscous disk, as well as for particles composed of CO$_2$ or CO, but the snowline locations will vary between the three disks for a given initial particle size (see Section \ref{sec:COratio}).} \textbf{If we do not take into account the time dependence of the mass accretion rate and stellar luminosity (see Section \ref{sec:disk}),} the C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
+Both of these conclusions remain valid for an evolving and a viscous disk, as well as for particles composed of CO$_2$ or CO, but the snowline locations will vary between the three disks for a given initial particle size (see Section \ref{sec:COratio}).
+If we do not take into account the time dependence of the mass accretion rate and stellar luminosity (see Section \ref{sec:disk}), the C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
 %\emgr{I've yet to figure out why there is a discrepancy for the small grains.
@@
 %, for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk. 
-The particle size increases from \textbf{$\sim$0.05 cm} to $\sim$700 cm as indicated by the color bar.
+The particle size increases from $\sim$0.05 cm to $\sim$700 cm as indicated by the color bar.
 The horizontal dotted line represents the stellar value of 0.54.
@@
 The minimum snowline distance for a disk is given by the curve corresponding to the maximum particle size it hosts.
-\textbf{For} grains that have grown to radii larger than $\sim$7 m \textbf{and are able to drift and desorb}, the $\sim$7 m snowline applies (see Section 3).
+For grains that have grown to radii larger than $\sim$7 m and are able to drift and desorb, the $\sim$7 m snowline applies (see Section 3).
 %Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.
@@
 
-Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for an irradiated disk, an evolving disk, and a viscous disk, under the simplifying assumption that the abundance relative to hydrogen for each volatile is \textbf{initially} fixed, so that drift and accretion affect only the locations of the snowlines.
+Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for an irradiated disk, an evolving disk, and a viscous disk, under the simplifying assumption that the abundance relative to hydrogen for each volatile is fixed, so that drift and accretion affect only the locations of the snowlines.
 We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
@@
 The C/O ratio for a static disk, where desorption and readsorption balance \citep{hollenbach09}, is shown as a guideline.
-We note that the true snowline for particles with $r_{\rm des}$ outside the static snowline is the static snowline itself---\textbf{thus only particles with initial sizes larger than $\sim$0.05 cm are plotted in the three panels, as particles that form snowlines at larger distances (cf.
-Figure \ref{fig:snowlines}) are not true snowlines}.
+We note that the true snowline for particles with $r_{\rm des}$ outside the static snowline is the static snowline itself---thus only particles with initial sizes larger than $\sim$0.05 cm are plotted in the three panels, as particles that form snowlines at larger distances (cf.
+Figure \ref{fig:snowlines}) are not true snowlines.
 
@@
 As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
-For example, for the disk parameters and particle sizes displayed in Figure 5, water molecules in solid particles drift up to $\sim$1000 times faster across the H$_2$O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.
+For example, for the disk parameters and particle sizes displayed in Figure \ref{fig:CO_ratio}, water molecules in solid particles drift up to $\sim$1000 times faster across the H$_2$O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.
 This differential inward motion will result in an increased oxygen gas abundance inside the H$_2$O snowline, and thus a (in some cases much) lower gaseous C/O ratio in this region.
@@
 
-Figure \ref{fig:CO_ratio} plots snowline curves for particle sizes \textbf{$\sim0.5$ cm $\lesssim $s$ \lesssim$ 7m}.
+\textbf{Our premise that each volatile has a fixed abundance at every location in the disk yields good approximations for the C/O ratio in an irradiated disk --- our model assumes a constant influx of particles at any given radius, implying that the solid surface density of a given volatile, and therefore the amount of desorbed vapor, remain constant with semimajor axis.
+For the evolving and viscous disks, the C/O ratio in gas may decrease due to the decline in the surface density of solids with time.
+Thus both the relative fluxes of the volatiles and the decrease in solid ice abundances will lower the C/O ratio in evolving and viscous disks, as indicated by the arrows in Figure \ref{fig:CO_ratio}.
+However, a high gas-phase C/O ratio is maintained further in for all disks compared to a static disk, since drift and gas accretion cause the ices to desorb closer to the host star.}
+
+Figure \ref{fig:CO_ratio} plots snowline curves for particle sizes $\sim0.5$ cm $\lesssim $s$ \lesssim$ 7m.
 In the outermost disk, H$_2$O, CO$_2$, and CO all solidify.
@@
 Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
-\textbf{For a grain size distribution with a maximum particle size different than our model, one can pick out the appropriate minimum snowline locations from this plot.}%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
+For a grain size distribution with a maximum particle size different than our model, one can pick out the appropriate minimum snowline locations from this plot.%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
 Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from this plot.
@@
 For our choice of parameters, the minimum snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the irradiated disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the evolving disk and $r_{\rm{H_2O}} \approx 0.63$ AU for the viscous disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the irradiated disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both the evolving and the viscous disks; $r_{\rm{CO}} \approx 30$ AU for the irradiated and both the evolving and the viscous disks.
-For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU\footnote{\textbf{For the viscous disk, we calculated the static snowline location using the same temperature profile as that of the viscous disk, for consistency purposes.
-Thus $r_{\rm{H_2O}} \approx 0.98$ AU for the static disk in this scenario.}}, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
-\textbf{For the viscous disk model, which is the most realistic,} radial drift and gas accretion push the snowline locations inwards by up to \textbf{$\sim$$40$ \%} for H$_2$O, by up to  $\sim$$60$ \% for CO$_2$, and by up to $\sim$$50$ \% for CO.
+For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU\footnote{For the viscous disk, we calculated the static snowline location using the same temperature profile as that of the viscous disk, for consistency purposes.
+Thus $r_{\rm{H_2O}} \approx 0.98$ AU for the static disk in this scenario.}, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
+For the viscous disk model, which is the most realistic, radial drift and gas accretion push the snowline locations inwards by up to $\sim$$40$ \% for H$_2$O, by up to  $\sim$$60$ \% for CO$_2$, and by up to $\sim$$50$ \% for CO.
 We note that the H$_2$O snowline in all disks is significantly closer to the host star compared with Solar system models, which place the H$_2$O snowline between $\sim$$2.7$ to $\sim$$3.1$ AU (\citealt{hayashi81}, \citealt{podolak04}, \citealt{martin12}).
-This is partially because we choose a colder disk model, as well as the fact that gas accretion rates decrease over time, moving the snowline location \textbf{inwards} (see also \citealt{garaud07} and Section \ref{sec:neglected}).
-\textbf{\citet{min11} find that the location of the H$_2$O snowline is highly sensitive to the gas mass accretion rate $\dot{M}$ (equal to $10^{-8} M_{\odot}$ yr$^{-1}$ in our model) and the dust opacity $\kappa$ (Equation \ref{eq:opacity}).
+This is partially because we choose a colder disk model, as well as the fact that gas accretion rates decrease over time, moving the snowline location inwards (see also \citealt{garaud07} and Section \ref{sec:neglected}).
+\citet{min11} find that the location of the H$_2$O snowline is highly sensitive to the gas mass accretion rate $\dot{M}$ (equal to $10^{-8} M_{\odot}$ yr$^{-1}$ in our model) and the dust opacity $\kappa$ (Equation \ref{eq:opacity}).
 Higher values of $\dot{M}$ and $\kappa$ would increase the accretional component of the disk temperature (cf.
 Equation \ref{eq:Tdact}), which would push the H$_2$O snowline in the viscous disk outwards to match the Solar system snowline.
-At the same time, the snowline location in Solar type stars may be as close as $\sim$1 AU \citep{mulders15}, further in than the H$_2$O snowline in our Solar system.}
+At the same time, the snowline location in Solar type stars may be as close as $\sim$1 AU \citep{mulders15}, further in than the H$_2$O snowline in our Solar system.
 
@@
 The C/O ratio is thus insensitive to the choice of $M$.
-\textbf{We note that our conclusions regarding the disk age and total mass are only valid if the snowline itself does not move with time or disk mass (see also Sections \ref{sec:disk} and \ref{sec:neglected}).}
+We note that our conclusions regarding the disk age and total mass are only valid if the snowline itself does not move with time or disk mass (see also Sections \ref{sec:disk} and \ref{sec:neglected}).
 
@@
 However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
-Turbulence causes eddies and vertical mixing, which are likely to reduce the radial drift velocity of the solids \textbf{(e.g., \citealt{youdin07})}.
+Turbulence causes eddies and vertical mixing, which are likely to reduce the radial drift velocity of the solids (e.g., \citealt{youdin07}).
 Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially.
@@
 However, accretional heating may push the snowline outwards compared to an evolving disk without accretion heating.} \\\hline
-Particle growth & $\leftarrow$ \footnote{\textbf{As stated in the main text, if particles grow to km-sizes and above, the snowline remains the same as that of a static disk.}} \\\hline
+Particle growth & $\leftarrow$ \footnote{As stated in the main text, if particles grow to km-sizes and above, the snowline remains the same as that of a static disk.} \\\hline
 Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
@@
 The inhomogeneity due to cracks in the grain structure will cause the particles to desorb faster.
-They will therefore drift less before evaporating and will move the snowlines \textbf{less far inward}.
+They will therefore drift less before evaporating and will move the snowlines less far inward.
 
@@
 
-\item \textbf{Disk gaps and holes.} The snowline locations will be different for transition disks, which have inner cavities significantly depleted of gas (e.g., \citealt{espaillat12}, \textbf{\citealt{vandermarel15}}), or pre-transitional disks, which have a gap between an inner and outer full disk (e.g., \citealt{kraus11}).
+\item \textbf{Disk gaps and holes.} The snowline locations will be different for transition disks, which have inner cavities significantly depleted of gas (e.g., \citealt{espaillat12}, \citealt{vandermarel15}), or pre-transitional disks, which have a gap between an inner and outer full disk (e.g., \citealt{kraus11}).
 The decrease in gas pressure in these gaps or holes will reduce the particles' drift velocity close to the gap edge, thus slowing them down and pushing the snowline outwards.
@@
 \item \textbf{Accretion rate evolution.} Our viscous disk model assumes a constant mass accretion rate $\dot{M}$.
-However, $\dot{M}$ decreases over time, which lowers the accretional component of the disk temperature (Equation \ref{eq:Tdact}), thus pushing the snowline location inwards if the disk is optically thick \citep{garaud07}.
+However, $\dot{M}$ decreases over time, which lowers the accretional component of the disk temperature (Equation \ref{eq:Tdact}), thus pushing the snowline location inwards if the disk is optically thick (\citealt{garaud07}).
 Once $\dot{M}$ reaches low enough values for the snowline to become optically thin, the snowline location moves outwards \citep{garaud07}.
-Our choice of $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ roughly corresponds to the mass accretion rate during the quasi-static accretion phase of the disk evolution predicted by \citet{garaud07}.
+\textbf{During the giant planet formation stage of a few Myr, however, $\dot{M}$ steadily decreases with time \citep{chambers09}, which may result in the inward movement of the H$_2$O snowline by up to one order of magnitude, significantly larger than the inward movement caused by radial drift (cf.
+Section \ref{sec:COratio}).
+We thus acknowledge that the location of the H$_2$O snowline may be set by the mass accretion rate evolution rather than the drift of solids.}
+
+%Our choice of $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ roughly corresponds to the mass accretion rate during the quasi-static accretion phase of the disk evolution predicted by \citet{garaud07}.
 This is the timescale on which giant planets accrete most of their gaseous envelope.%; thus the implications of our results for the atmospheric composition of gas giants are robust.
@@
 \item \textbf{Stellar luminosity evolution.} As the host star contracts during its pre-main sequence phase, its luminosity decreases, which reduces the disk temperature and pushes the snowline locations inwards.
-\citet{kennedy06} found that the snowline \textbf{is unlikely to move significantly during the pre-main sequence phase for Solar type stars, but it} may move inward by a factor of $\sim$$15-20$ \textbf{for $M_* \sim 0.25 M_{\odot}$} due to the stellar contraction.
+\citet{kennedy06} found that the snowline is unlikely to move significantly during the pre-main sequence phase for Solar type stars, but it may move inward by a factor of $\sim$$15-20$ for $M_* \sim 0.25 M_{\odot}$ due to the stellar contraction.
 
@@
 \begin{enumerate}
-\item Radial drift and gas accretion affect desorption and move the snowline locations \textbf{inward} compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk.
+\item Radial drift and gas accretion affect desorption and move the snowline locations inward compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk.
 %Moreover, boulders up to almost kilometer-size contribute to setting the snowline locations if they are very close to the snowlines.
@@
 
-\item \textbf{For our simplified model that does not account for the effects outlined in Section \ref{sec:neglected},} particles with sizes in the above range desorb almost instantaneously once desorption has begun, and at a fixed location in the disk that only depends on the particle size and the gas accretion rate.
+\item For our simplified model that does not account for the effects outlined in Section \ref{sec:neglected}, particles with sizes in the above range desorb almost instantaneously once desorption has begun, and at a fixed location in the disk that only depends on the particle size and the gas accretion rate.
 Thus for each particle size there is a fixed and uniquely determined H$_2$O, CO$_2$ or CO snowline.
 
-\item The results of \textbf{our} numerical simulation are in agreement with the analytic solution of the drift-desorption system of differential equations if the stopping time $\tau_{\rm s} \ll 1$.
+\item The results of our numerical simulation are in agreement with the analytic solution of the drift-desorption system of differential equations if the stopping time $\tau_{\rm s} \ll 1$.
 We present an explicit analytic solution for the desorption distance in this regime.
@@
 
-\item For our fiducial model, \textbf{which considers particles with sizes between $10^{-3}$ and $10^8$ cm}, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for an irradiated disk, 0.26 AU, 3.4 AU and 30 AU for an evolving disk, and 0.63 AU, 3.4 AU and 30 AU for a viscous disk with accretion heating.
+\item For our fiducial model, which considers particles with sizes between $10^{-3}$ and $10^8$ cm, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for an irradiated disk, 0.26 AU, 3.4 AU and 30 AU for an evolving disk, and 0.63 AU, 3.4 AU and 30 AU for a viscous disk with accretion heating.
 Compared to a static disk, radial drift and gas accretion move the snowlines by up to 60 \% for H$_2$O and CO$_2$, and by up to 50 \% for CO.
-\textbf{For the viscous disk, however, which is the most realistic of the three models since it takes into account accretion heating, the H$_2$O snowline location moves inwards by up to 40 \%.}
-
-\item Our C/O estimates \textbf{confirm the conclusions of \citet{oberg11}} that the C/O ratio \textbf{in gas} may be enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching its maximum value between the CO$_2$ and CO snowlines.
+For the viscous disk, however, which is the most realistic of the three models since it takes into account accretion heating, the H$_2$O snowline location moves inwards by up to 40 \%.
+
+\item Our C/O estimates confirm the conclusions of \citet{oberg11} that the C/O ratio \textbf{in gas} may be enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching its maximum value between the CO$_2$ and CO snowlines.
 %This is consistent with possible detections of superstellar C/O ratios in some exoplanet atmospheres.
-We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the evolving and viscous disks, \textbf{which may significantly decrease the C/O ratio in gas inside the H$_2$O and CO$_2$ snowlines.} \textbf{We plan to address this issue in a future paper.}
-
-\item \textbf{For a constant gas mass accretion rate $\dot{M}$ and stellar luminosity $L_*$,} the snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk midplane remains optically thick.
+We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the evolving and viscous disks, which may significantly decrease the C/O ratio in gas inside the H$_2$O and CO$_2$ snowlines.
+We plan to address this issue in a future paper.
+
+\item For a constant gas mass accretion rate $\dot{M}$ and stellar luminosity $L_*$, the snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk midplane remains optically thick.
 %Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.

### 2015-11-14 python/C_to_O_CH4.py
update

@@ -35,8 +35,12 @@ def n_C_O(species, CH4mid = 0, CH4max = 0):
         return 0.3, 0.6
     elif species == 'H2O':
         return 0, 0.9
-    elif species == 'C_grains':
+    elif species == 'C_grains' and CH4mid == 0 and CH4max == 0:
         return 0.6, 0
+    elif species == 'C_grains' and CH4mid == 1 and CH4max == 0:
+        return 0.6 - 0.0555 * 0.9, 0
+    elif species == 'C_grains' and CH4mid == 0 and CH4max == 1:
+        return 0.6 - 0.13 * 0.9, 0
     elif species == 'silicate':
         return 0, 1.4
     elif species == 'CH4' and CH4mid == 0 and CH4max == 0:
@@ -86,24 +90,24 @@ def n(T, elem, CH4mid, CH4max):
         
     if T >= T_freeze_H20:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('H2O')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i]])
+            n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i]])
             
     elif T_freeze_H20 >= T >= T_freeze_CO2:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('H2O')[i] + n_C_O('C_grains')[i] + n_C_O('silicate')[i]])
+            n_C_O('H2O')[i] + n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i]])
 
     elif T_freeze_CO2 >= T > T_freeze_CO:
         return np.array([n_C_O('CO')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
+            n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
                 n_C_O('CO2')[i]])            
             
     elif T_freeze_CO >= T > T_freeze_CH4:
         return np.array([n_C_O('CH4', CH4mid, CH4max)[i], \
[…]

### 2015-11-14 python/C_to_O_CH4_COCO.py

@@ -35,8 +35,12 @@ def n_C_O(species, CH4mid = 0, CH4max = 0):
         return 0.3, 0.6
     elif species == 'H2O':
         return 0, 0.9
-    elif species == 'C_grains':
+    elif species == 'C_grains' and CH4mid == 0 and CH4max == 0:
         return 0.6, 0
+    elif species == 'C_grains' and CH4mid == 1 and CH4max == 0:
+        return 0.6 - 0.0555 * 0.9, 0
+    elif species == 'C_grains' and CH4mid == 0 and CH4max == 1:
+        return 0.6 - 0.13 * 0.9, 0
     elif species == 'silicate':
         return 0, 1.4
     elif species == 'CH4' and CH4mid == 0 and CH4max == 0:
@@ -86,24 +90,24 @@ def n(T, elem, CH4mid, CH4max):
         
     if T >= T_freeze_H20:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('H2O')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i]])
+            n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i]])
             
     elif T_freeze_H20 >= T >= T_freeze_CO2:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('H2O')[i] + n_C_O('C_grains')[i] + n_C_O('silicate')[i]])
+            n_C_O('H2O')[i] + n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i]])
 
     elif T_freeze_CO2 > T > T_freeze_CH4:
         return np.array([n_C_O('CO')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
+            n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
                 n_C_O('CO2')[i]])            
             
     elif T_freeze_CH4 > T > T_freeze_CO:
         return np.array([n_C_O('CO')[i], \
[…]

### 2015-11-14 python/C_to_O_steady_CO.py

@@ -35,8 +35,12 @@ def n_C_O(species, CH4mid = 0, CH4max = 0):
         return 0.3, 0.6
     elif species == 'H2O':
         return 0, 0.9
-    elif species == 'C_grains':
+    elif species == 'C_grains' and CH4mid == 0 and CH4max == 0:
         return 0.6, 0
+    elif species == 'C_grains' and CH4mid == 1 and CH4max == 0:
+        return 0.6 - 0.0555 * 0.9, 0
+    elif species == 'C_grains' and CH4mid == 0 and CH4max == 1:
+        return 0.6 - 0.13 * 0.9, 0        
     elif species == 'silicate':
         return 0, 1.4
     elif species == 'CH4' and CH4mid == 0 and CH4max == 0:
@@ -92,24 +96,24 @@ def n(index, r, elem, CH4mid = 0, CH4max = 0):
         
     if r <= aH2O:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('H2O')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i]]), size
+            n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i]]), size
             
     elif aH2O <= r <= aCO2:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('H2O')[i] + n_C_O('C_grains')[i] + n_C_O('silicate')[i]]), size
+            n_C_O('H2O')[i] + n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i]]), size
 
     elif aCO2 < r <= aCH4:
         return np.array([n_C_O('CO')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
+            n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
                 n_C_O('CO2')[i]]), size            
             
     elif aCH4 < r <= aCO:
         return np.array([n_C_O('CO')[i], \
[…]

### 2015-11-14 python/C_to_O_steady_water.py

@@ -92,24 +92,24 @@ def n(index, r, elem, CH4mid = 0, CH4max = 0):
         
     if r < aH2O:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('H2O')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i]]), size
+            n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i]]), size
             
     elif aH2O <= r <= aCO2:
         return np.array([n_C_O('CO')[i] + n_C_O('CO2')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('H2O')[i] + n_C_O('C_grains')[i] + n_C_O('silicate')[i]]), size
+            n_C_O('H2O')[i] + n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i]]), size
 
     elif aCO2 <= r < aCO:
         return np.array([n_C_O('CO')[i] + n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
+            n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
                 n_C_O('CO2')[i]]), size            
             
     elif aCO <= r < aCH4:
         return np.array([n_C_O('CH4', CH4mid, CH4max)[i], \
-            n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
+            n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
                 n_C_O('CO2')[i] + +n_C_O('CO')[i]]), size
                 
     elif aCH4 <= r:
-        return np.array([0, n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
+        return np.array([0, n_C_O('C_grains', CH4mid, CH4max)[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
                 n_C_O('CO2')[i] + n_C_O('CO')[i] + n_C_O('CH4', CH4mid, CH4max)[i]]), size
             
 

### 2015-11-14 tex/draft/ms_final_revised_2.tex
update

@@
 
-\textbf{Our premise that each volatile has a fixed abundance at every location in the disk yields good approximations for the C/O ratio in an irradiated disk --- our model assumes a constant influx of particles at any given radius, implying that the solid surface density of a given volatile, and therefore the amount of desorbed vapor, remain constant with semimajor axis.
+%\textbf{Our premise that each volatile has a fixed abundance at every location in the disk yields good approximations for the C/O ratio in an irradiated disk --- our model assumes a constant influx of particles at any given radius, implying that the solid surface density of a given volatile, and therefore the amount of desorbed vapor, remain constant with semimajor axis.
 For the evolving and viscous disks, the C/O ratio in gas may decrease due to the decline in the surface density of solids with time.
@@
 However, a high gas-phase C/O ratio is maintained further in for all disks compared to a static disk, since drift and gas accretion cause the ices to desorb closer to the host star.}
+
+%\textbf{Desorption is taken into account in the figure, and fundamentally the elevated C/O ratio interior to the static CO2 and H2O snowlines are simply due to the inward mobility of the snowlines (desorption fronts) due to drift and accretion flows.
+Qualitatively this scenario should be robust to changes in total abundances throughout the disk, i.e.
+at e.g.
+the ‘dynamic’ CO2 snowlines, the rapid return of CO2 into the gas-phase (during CO2 desorption) will reduce the C/O ratio interior to the CO2 dynamic snowline, while no major change in gas-phase composition and therefore C/O ratio is expected between the static and dynamic snowlines.
+
+%As the referee notes, we do operate under the simplifying assumption that the total (ice+gas) abundances are the same at each radius after ices have migrated. This is a good
+%approximation for the irradiated disk, given that this model by definition presents a constant influx of particles at any given radius while the gas is static, and thus the ice+gas surface density should remain constant.
+For the evolving disk, we agree with the referee that this is not a good approximation.
+In evolving disks, the gas-phase C/O ratio may decrease everywhere interior to the CO2 and H2O desorption fronts due to the decrease in the surface density of solids with time at any given radius.
+In the steady-state viscous disk, the solid abundances at a fixed radius are constant, given that this model is
+%not time-dependent, but the solid/gas ratio is not constant, which can result in a substantially lower C/O ratio interior to the H2O and CO2 snowlines compared to the static case (as indicated by the arrows in the figure).
+
+%These caveats are important and we have explained and clarified them further in the next.
+We have also clarified that the main purpose of this figure is to show the different snowline locations in static and dynamic disks, and thus where in the disk C/O is reduced or increased rather than providing a quantitative estimate of how big or small that increase/decrease is.}
 
@@
 For these particle sizes, our calculated C/O ratio is an upper limit,  as indicated by the arrows in Figure \ref{fig:CO_ratio}.
+
+\textbf{Fundamentally, the elevated C/O ratios interior to the static H$_2$O and CO$_2$ snowlines are simply caused by the inward movement of the snowlines due to radial drift and gas accretion.
+Qualitatively, this scenario should be robust to changes in total abundances throughout the disk --- for example, at the dynamic (non-static) CO$_2$ snowlines, the rapid return of CO$_2$ into gas-phase during CO$_2$ desorption will reduce the C/O ratio interior to the CO$_2$ dynamic snowline, while no major change in gas-phase composition, and therefore C/O ratio, is expected between the static and dynamic snowlines.}
+
+\textbf{As noted earlier, we assume that the total (ice and gas) abundance of each volatile is the same at every radius after ices have migrated.
+This is a good approximation for the irradiated disk, given that this model, by definition, presents a constant influx of particles at any given radius while the gas is static, and thus the ice and gas surface density remain constant.
+For the evolving disk, the gas-phase C/O ratio may decrease everywhere interior to the H$_2$O and CO$_2$ snowlines due to the decline in the surface density of solids with time at any given radius.
+For the viscous disk, the solid abundances at a fixed radius are constant, given that this model is not time-dependent, but the gas-to-solid ratio is not constant, which can result in a substantially lower C/O ratio interior to the H$_2$O and CO$_2$ snowlines compared to the static case (as indicated by the arrows in Figure \ref{fig:CO_ratio}).
+The main goal of Figure \ref{fig:CO_ratio}, however, is to show the different snowline radii in static and non-static disks, and therefore the locations in the disk where the gas-phase C/O ratio is reduced or increased, rather than provide a quantitative estimate of the magnitude of the C/O increase or decrease.}
 

### 2015-11-17 Postdoc Applications/TAC_proposal.tex
update

@@
 $[6]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
-$[7]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, resubmitted after referee report \\
+$[7]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, accepted \\
 $[8]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\

### 2015-11-17 tex/draft/ms_final_revised_2.tex
update

@@
 
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
+\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1, 3}, Ruth A. Murray-Clay\altaffilmark{2}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
 \altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
+\altaffiltext{3}{Max Planck Institute for Astronomy, K\"onigstuhl 17, 69117 Heidelberg, Germany}
 
@@
 We determine the snowline locations for a range of fixed initial particle sizes and disk types.
-For our fiducial disk model, we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk, and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving and viscous disk, desorb %instantaneously and [I think instantaneously is too strong]
-at a %fixed 
-size-dependent location in the disk, which is independent %regardless 
-of the particle's initial position.
+For our fiducial disk model, we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk, and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving and viscous disk, desorb at a size-dependent location in the disk, which is independent of the particle's initial position.
 The snowline radius decreases for larger particles, up to sizes of $\sim$7~m.
-%location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
 Compared to a static disk, we find that radial drift and gas accretion in a viscous disk move the H$_2$O snowline inwards by up to 40 \%, the CO$_2$ snowline by up to 60 \%, and the CO snowline by up to 50 \%.
 We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
-%Our results have direct implications on the C/O ratio in giant planet atmospheres that have formed in situ.
-[I don't think we discuss this enough to warrant it as a conclusion
 \end{abstract}
@@
 We choose $M=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
-The irradiated and evolving disk surface densities match at \textbf{$t \approx 5 \times 10^5$} years in the inner disk, but they diverge at distances larger than a few AU due the exponential cutoff in radius of the surface density of the evolving disk (Equation \ref{eq:Sigmaact}).
+The irradiated and evolving disk surface densities match at $t \approx 5 \times 10^5$ years in the inner disk, but they diverge at distances larger than a few AU due the exponential cutoff in radius of the surface density of the evolving disk (Equation \ref{eq:Sigmaact}).
 
@@
 Thus the midplane temperature will not change significantly for our model due to variations in stellar luminosity, but it may decrease substantially for smaller stars, pushing the snowline inward (see Section \ref{sec:neglected} for details).
-Realistic mass accretion rates, $\dot{M}$, may vary between $\sim$$10^{-7}$ and $\sim$$10^{-9}$ $M_{\odot}$ yr$^{-1}$ as the disk evolves \textbf{(e.g., \citealt{chambers09}, \citealt{sicilia10})}.
+Realistic mass accretion rates, $\dot{M}$, may vary between $\sim$$10^{-7}$ and $\sim$$10^{-9}$ $M_{\odot}$ yr$^{-1}$ as the disk evolves (e.g., \citealt{chambers09}, \citealt{sicilia10}).
 For $\dot{M}\lesssim 10^{-9} M_{\odot}$ yr$^{-1}$, the disk becomes optically thin and hence depleted of gas, which means giant planets must have formed before $\dot{M}$ becomes too low.
 \citet{garaud07} find that the snowline locations scale as $r_{\rm snow} \propto \dot{M}^{1/3}$.
-A factor of 100 reduction in the mass accretion rate will thus move the H$_2$O snowline inwards by a factor of $\sim$$4$ \textbf{--- since accretion heating is dominant only in the inner disk, the CO$_2$ and CO snowline locations are unlikely to be affected by changes in mass accretion rate.
-The inward movement of the H$_2$O snowline due to the decrease in $\dot{M}$ may be even larger, by up to one order of magnitude \citep{chambers09}}.
-We thus conclude that changes in $L_*$ throughout time may only modestly affect our results, \textbf{while changes in $\dot{M}$ may significantly affect our results for the H$_2$O snowline, as its location may be determined by the decline in mass accretion rate rather than radial drift.} The time variability of $L_*$ and $\dot{M}$ should be taken into account when drawing more robust conclusions, as well as for different host star and disk properties.
+A factor of 100 reduction in the mass accretion rate will thus move the H$_2$O snowline inwards by a factor of $\sim$$4$ --- since accretion heating is dominant only in the inner disk, the CO$_2$ and CO snowline locations are unlikely to be affected by changes in mass accretion rate.
+The inward movement of the H$_2$O snowline due to the decrease in $\dot{M}$ may be even larger, by up to one order of magnitude \citep{chambers09}.
+We thus conclude that changes in $L_*$ throughout time may only modestly affect our results, while changes in $\dot{M}$ may significantly affect our results for the H$_2$O snowline, as its location may be determined by the decline in mass accretion rate rather than radial drift.
+The time variability of $L_*$ and $\dot{M}$ should be taken into account when drawing more robust conclusions, as well as for different host star and disk properties.
 
@@
 For the viscous disk (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}).
-For our choice of parameters for both the evolving and the viscous disks, we have found that the radial flow of gas \textbf{(calculated from Equation \ref{eq:vgas} for the evolving disk and from $\dot{r}_{\rm gas}=-\dot{M}/(2 \pi r \Sigma)$ for the viscous disk)} is always directed inward \textbf{for our parameter space of interest}, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}).
-\textbf{For our evolving disk, the gas starts drifting outwards at a radius $r_{\rm switch} \approx 200 AU$, which is however well outside the CO snowline in our model.} We \textbf{thus} note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations.
+For our choice of parameters for both the evolving and the viscous disks, we have found that the radial flow of gas(calculated from Equation \ref{eq:vgas} for the evolving disk and from $\dot{r}_{\rm gas}=-\dot{M}/(2 \pi r \Sigma)$ for the viscous disk) is always directed inward for our parameter space of interest, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}).
+For our evolving disk, the gas starts drifting outwards at a radius $r_{\rm switch} \approx 200 AU$, which is however well outside the CO snowline in our model.
+We thus note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations.
 Since drifting particles larger than a few cm are only modestly affected by gas accretion, an outward gas flow would move the CO$_2$ and CO locations further away from the star only for the smallest particles in our model, which are well-coupled to the gas.
@@
 
-\textbf{Fundamentally, the elevated C/O ratios interior to the static H$_2$O and CO$_2$ snowlines are simply caused by the inward movement of the snowlines due to radial drift and gas accretion.
-Qualitatively, this scenario should be robust to changes in total abundances throughout the disk --- for example, at the dynamic (non-static) CO$_2$ snowlines, the rapid return of CO$_2$ into gas-phase during CO$_2$ desorption will reduce the C/O ratio interior to the CO$_2$ dynamic snowline, while no major change in gas-phase composition, and therefore C/O ratio, is expected between the static and dynamic snowlines.}
-
-\textbf{As noted earlier, we assume that the total (ice and gas) abundance of each volatile is the same at every radius after ices have migrated.
+Fundamentally, the elevated C/O ratios interior to the static H$_2$O and CO$_2$ snowlines are simply caused by the inward movement of the snowlines due to radial drift and gas accretion.
+Qualitatively, this scenario should be robust to changes in total abundances throughout the disk --- for example, at the dynamic (non-static) CO$_2$ snowlines, the rapid return of CO$_2$ into gas-phase during CO$_2$ desorption will reduce the C/O ratio interior to the CO$_2$ dynamic snowline, while no major change in gas-phase composition, and therefore C/O ratio, is expected between the static and dynamic snowlines.
+
+As noted earlier, we assume that the total (ice and gas) abundance of each volatile is the same at every radius after ices have migrated.
 This is a good approximation for the irradiated disk, given that this model, by definition, presents a constant influx of particles at any given radius while the gas is static, and thus the ice and gas surface density remain constant.
@@
 For the viscous disk, the solid abundances at a fixed radius are constant, given that this model is not time-dependent, but the gas-to-solid ratio is not constant, which can result in a substantially lower C/O ratio interior to the H$_2$O and CO$_2$ snowlines compared to the static case (as indicated by the arrows in Figure \ref{fig:CO_ratio}).
-The main goal of Figure \ref{fig:CO_ratio}, however, is to show the different snowline radii in static and non-static disks, and therefore the locations in the disk where the gas-phase C/O ratio is reduced or increased, rather than provide a quantitative estimate of the magnitude of the C/O increase or decrease.}
+The main goal of Figure \ref{fig:CO_ratio}, however, is to show the different snowline radii in static and non-static disks, and therefore the locations in the disk where the gas-phase C/O ratio is reduced or increased, rather than provide a quantitative estimate of the magnitude of the C/O increase or decrease.
 
@@
 Once $\dot{M}$ reaches low enough values for the snowline to become optically thin, the snowline location moves outwards \citep{garaud07}.
-\textbf{During the giant planet formation stage of a few Myr, however, $\dot{M}$ steadily decreases with time \citep{chambers09}, which may result in the inward movement of the H$_2$O snowline by up to one order of magnitude, significantly larger than the inward movement caused by radial drift (cf.
+During the giant planet formation stage of a few Myr, however, $\dot{M}$ steadily decreases with time \citep{chambers09}, which may result in the inward movement of the H$_2$O snowline by up to one order of magnitude, significantly larger than the inward movement caused by radial drift (cf.
 Section \ref{sec:COratio}).
-We thus acknowledge that the location of the H$_2$O snowline may be set by the mass accretion rate evolution rather than the drift of solids.}
+We thus acknowledge that the location of the H$_2$O snowline may be set by the mass accretion rate evolution rather than the drift of solids.
 
@@
 
-\item Our C/O estimates confirm the conclusions of \citet{oberg11} that the C/O ratio \textbf{in gas} may be enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching its maximum value between the CO$_2$ and CO snowlines.
+\item Our C/O estimates confirm the conclusions of \citet{oberg11} that the C/O ratio in gas may be enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching its maximum value between the CO$_2$ and CO snowlines.
 %This is consistent with possible detections of superstellar C/O ratios in some exoplanet atmospheres.
@@
 %Nevertheless,
+
+\acknowledgements{We thank the referee for a helpful and detailed report.
+We also thank Sean Andrews for useful conversations and suggestions.
+T.
+B.
+is grateful for support from the NASA Origins of Solar Systems grant NNX12AJ04G and the Smithsonian Institution Pell Grant program.}
 

### 2015-12-11 Postdoc Applications/DTM_proposal.tex
update

--- 2015-11-13 Postdoc Applications/ITC_proposal.tex
@@
 \lhead{Ana-Maria Piso}
-\rhead{ITC Postdoctoral Fellowship Research Proposal}
+\rhead{Postdoctoral Fellowship in Astronomy and Planetary Science}
 \date{}
@@
 
-\title{ITC Postdoctoral Fellowship Research Proposal: \\
+\title{Postdoctoral Fellowship in Astronomy and Planetary Science Research Proposal: \\
 The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets}
@@
 
-\textbf{The Institute for Theory and Computation is the ideal place for me to pursue my postdoctoral work.} Its research facilities and academic excellence are incomparable.
-Even though I am completing my Ph.D.
-at Harvard, I believe I could continue benefitting as a postdoctoral fellow from the plethora of resources the ITC has to offer, especially given the number of experts in protoplanetary disks and exoplanets that the ITC hosts.
-As an ITC Fellow, I will have the opportunity to foster new collaborations and thrive as a scientist, both during my postdoctoral tenure and beyond.
-I would love to collaborate with \textbf{Prof.
-Dimitar Sasselov}, a world-leading theorist in many aspects of planet formation, atmospheric chemistry and astrobiology.
-I would also like to work with \textbf{Dr.
-Matt Holman}, an expert in planetary dynamics.
-Additionally, both the ITC and the larger CfA research community include leaders in protoplanetary disk observations, such as \textbf{Dr.
-Sean Andrews} and \textbf{Dr.
-David Wilner}, as well as distinguished experts in exoplanet detection and characterization, such as \textbf{Prof.
-Dave Charbonneau} and \textbf{Prof.
-John Johnson}.
-Forming such collaborations presents great prospects in connecting my theoretical research work with observations.
+\textbf{Carnegie DTM is the ideal place for me to pursue my postdoctoral research}, due to is vibrant community of planetary scientists and cosmochemists.
+I believe that my proposed research lies at the intersection of both these exciting fields, and would mesh well with ongoing research at DTM.
+In particular, I would love the opportunity to collaborate with \textbf{Dr.
+Alan Boss} and \textbf{Dr.
+John Chambers}, world leading planet formation theorists.
+For the chemical aspect of my proposed project, I would like to collaborate with and learn from \textbf{Dr.
+Conlel M.
+O'D.
+Alexandrer}, an expert in cosmochemistry and the origin of interstellar grains.
+Additionally, Carnegie DTM hosts leaders in observations of prototplanetary disks and planet formation, such as \textbf{Dr.
+Alycia Weinberger}.
+Fostering such collaborations presents great prospects in eventually connecting my theoretical research work with observations.
+
+%\textbf{The Theoretical Astrophysics Center at UC Berkeley is the ideal place for me to pursue my postdoctoral research}, due to its vibrant community of expert theorists in protoplanetary disks and exoplanets.
+In particular, I would love to collaborate with \textbf{Prof.
+Eugene Chiang}, a world-leading planet formation theorist.
+Prof.
+Chiang's expertise in various topics related to protoplanetary disks and planet dynamics would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards.
+I would also like to collaborate with \textbf{Prof.
+Alfred Glassgold}, an expert in protoplanetary disk chemistry.
+Additionally, the larger UC Berkeley Department of Astronomy hosts leaders in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof.
+Paul Kalas} and \textbf{Prof.
+James Graham}, Project Leads of The Gemini Planet Imager Exoplanet Survey.
+Fostering such collaborations presents great prospects in connecting my
+%theoretical research work with observations.
 
@@
 $[6]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
-$[7]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, resubmitted after referee report \\
+$[7]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, accepted, arXiv:1511.05563 \\
 $[8]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\

### 2015-12-11 Postdoc Applications/previous_research_torres.tex
update

--- 2015-11-05 Postdoc Applications/previous_research_hubble.tex
@@
 \lhead{Ana-Maria Piso}
-\rhead{Hubble Fellowship Previous and Current Research Statement}
+\rhead{Juan Carlos Torres Fellowship Research Statement}
 \date{}
@@
 
-\title{\Large Hubble Fellowship Previous and Current Research Statement: \\
+\title{\Large Juan Carlos Torres Fellowship Research Statement: \\
 Giant Planet Formation and Snowlines in Protoplanetary Disks\vspace{-2ex}}

### 2015-12-11 Postdoc Applications/torres_proposal.tex
update

--- 2015-11-13 Postdoc Applications/AnaMariaPisoResearchProposal.tex
@@
 \lhead{Ana-Maria Piso}
-\rhead{Postdoctoral Position in Exoplanet Theory Research Proposal}
+\rhead{Juan Carlos Torres Fellowship Research Proposal}
 \date{}
@@
 
-\title{\Large Postdoctoral Position in Exoplanet Theory Research Proposal: \\
+\title{\Large Juan Carlos Torres Fellowship Research Proposal \\
 The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets\vspace{-2ex}}
@@
 
-\textbf{The University of Chicago is the ideal place for me to pursue my postdoctoral research}, due to its vibrant community of experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
-I would love the opportunity to work with \textbf{Prof.
-Leslie Rogers}, an expert in planetary interiors and atmospheres --- such a collaboration would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards.
-In the Department of Geophysical Sciences, I would like to collaborate with \textbf{Prof.
-Fred Ciesla}, a leader in protplanetary disk dynamics, chemical composition and evolution.
-Additionally, the Department of Astronomy hosts experts in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof.
-Dan Fabrycky} and \textbf{Prof.
-Jacob Bean}, which presents great prospects in connecting my theoretical research work with observations.
+\textbf{The Massachusetts Institute of Technology (MIT) is the best place for me to pursue my postdoctoral research}, due to its vibrant community of experts in exoplanet atmospheres and dynamics, both in the Physics and Earth, Atmospheric and Planetary Sciences (EAPS) departments.
+In particular, I would like to collaborate with \textbf{Prof.
+Sara Seager}, who is a co-investigator of the \textbf{TESS mission} and a world-leading theorist in atmospheric chemistry, as well as with members of her group, who specialize in a broad range of exoplanet theory and computational topics.
+I would also love to work with \textbf{Prof.
+Hilke Schlichting}, a leader in planet formation theory and dynamics.
+\textbf{Dr.
+Margaret Pan}, an excellent theorist in planetary dynamics and protoplanetary disks, will join the EAPS department as a research scientist in early 2016, so it would be a pleasure to work with her as well.
+Finally, in order to connect my theoretical work with observations, I would like to collaborate with \textbf{Prof.
+Joshua Winn}, who is an expert in discovering and characterizing exoplanets.
 \\

### 2015-12-15 tex/draft2/ms.tex
update

@@
 \begin{abstract}
-...
+The elemental compositions of planets define their chemistry, and could potentially be used as beacons for their formation location if the elemental gas and grain ratios of planet birth environments, i.e.
+protoplanetary disks, are well understood.
+In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, the ice environment in which these carriers reside, and the presence of snowlines of major volatiles at different distances from the central star.
+We explore the effects of dynamical processes, molecular compositions and abundances, and the ice morphology of dust grains in disks on the snowline locations of the main C, O and N carriers, and their consequences for the C/N/O ratio in gas and dust throughout the disk.
+We find that radial drift and accretion alone can reduce the snowline radii of the main C, O and N carriers, i.e.
+H$_2$O, CO$_2$, CO and N$_2$, by 40-60\% compared to static disks.
+If CO and N$_2$ are bound to water ice instead of pure ices, their snowlines move inward by $\sim$$70$\%.
+Both of these effects substantially change the disk regions where C/O and N/O are enhanced over the stellar value.
+In the outer disk, the gaseous C/O and N/O are enhanced by factors of $\sim$$2$ and $\sim$$3$, respectively.
+Our estimates for the C/N/O ratios are only modestly affected by the presence of some C in the form of CH$_4$ and of some N in the form of NH$_3$.
+
+%We note that N/O enhancements in disk gas can be even more extreme than C/O in the outer disk due to the low volatility of N2 compared to all major C and O carriers.
+I will discuss these results together with the effects of additional dynamical processes, and outline a path toward a coupled drift-desorption-chemistry model that will provide robust quantitative results for volatile snowline locations and C/N/O abundance ratios as the disk evolves in time.
 \end{abstract}
@@
 
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=\textwidth]{../../figs/CNO_and_snowlines.pdf}
+%\vspace{-0.5in}
+\caption{CNO abundance in grains...} 
+\label{fig:CNOstatic}
+\end{figure*}
+
 \section{Coupled Drift-Desorption Model Review}
+\label{sec:review}
 
 We begin with a brief review of Paper I's model for the effect of radial drift and viscous gas accretion on volatile snowline locations.
-We review our disk model assumptions and relevant timescales in Section \ref{sec:disktime}, and summarize our findings in Section \ref{sec:driftdes}.
+We review our disk model in Section \ref{sec:disktime}, and summarize our numerical method and results in Section \ref{sec:driftdes}.
 
-\subsection{Disk Model and Important Timescales}
+\subsection{Disk Model}
 \label{sec:disktime}
@@
 We choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$, consistent with mass flux observations in disks (e.g., \citealt{andrews10}).
+As acknowledged in Paper I, the mass flux rate $\dot{M}$ and stellar luminosity $L_*$ will vary throughout the disk lifetime (\citealt{kennedy06}, \citealt{chambers09}), in contrast with our simplified model which assumes that both quantities are constant.
+This effect will be most pronounced in the inner disk ($\lesssim$ few AU), where accretion heating dominates.
+We thus acknowledge that the location of the H$_2$O snowline may be determined by the decline in $\dot{M}$ or $L_*$ with time, rather than radial drift (see Paper I, Section 2.1 for a more detailed explanation).
 
-In what follows, we summarize the
+%In what follows, we summarize the relevant timescales that affect volatile snowline locations. 
+%
+%\textit{Desorption timescale.} Following \citet{hollenbach09} and Paper I, the timescale to desorb a single layer of molecules of a volatile $x$ can be approximated as
+%\begin{equation}
+%\label{eq:tdes}
+%t_{\rm des}=\frac{\rho_{\rm s}}{3 \mu_x m_{\rm p}} \frac{s}{N_x R_{\rm des, x}},
+%\end{equation} 
+%where $\rho_{\rm s}=2$ g cm$^{-3}$ is the density of an icy particle, $s$ is the particle size, $\mu_x$ is the mean molecular weight of volatile $x$, and $N_x \approx 10^{15}$ sites cm $^{-2}$ is the number of adsorption sites of molecule $x$ per cm$^{-2}$.
+The desorption rate $R_{\rm des, x}$ (per molecule) is \citep{hollenbach09}
+%\begin{equation}
+%\label{eq:Rdes}
+%R_{\rm{des}, x} = \nu_x \exp{(-E_x/T_{\rm grain})},
+%\end{equation}
+%where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}=T$ is the grain temperature (assumed to be the same as the disk temperature, see Paper I), and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well.
+We discuss our choices for $E_x$ for the different volatile species in Sections \ref{sec:CH4} and \ref{sec:N}.
 
-\emgr{Review disk models, desorption model, relevant timescales.
+\subsection{Desorption-Drift Equations and Results}
+\label{sec:driftdes}
+
+For a range of initial icy grain sizes composed of a single volatile, we showed in Paper I that the timescale on which these particles desorb is comparable to their radial drift time, as well as to the accretion timescale of the nebular gas onto the central star.
+We thus have to take into account both drift and gas accretion when we calculate the disk location at which a particle desorbs, since that location may be different from the snowline position in a static disk for a given volatile (see Figure \ref{fig:CNOstatic} and \citealt{oberg11}).
+We determine a particle's final location in the disk by solving the following coupled differential equations:
+\begin{subeqnarray}
+\label{eq:ddt}
+\frac{ds}{dt} &= & - \frac{3 \mu_x m_{\rm p}}{\rho_{\rm s}} N_x R_{\rm des, x}  \slabel{eq:dsdt} \\
+\frac{dr}{dt} &=& \dot{r} \slabel{eq:drdt},
+\end{subeqnarray}
+where $s$ is the particle size, $t$ is time, $\mu_x$ is the mean molecular weight of volatile $x$, $\rho_{\rm s}=2$ g cm $^{-3}$ is the density of an icy particle, $N_x \approx 10^{15}$ sites cm $^{-2}$ is the number of adsorption sites of molecule $x$ per cm$^{-2}$, $R_{\rm des, x}$ is the desorption rate of species $x$, and $\dot{r}$ is the particle's radial drift velocity.
+We calculate $R_{\rm, des}$ and $\dot{r}$ as follows.
+
+The desorption rate $R_{\rm des, x}$ (per molecule) is \citep{hollenbach09}
+\begin{equation}
+\label{eq:Rdes}
+R_{\rm{des}, x} = \nu_x \exp{(-E_x/T_{\rm grain})},
+\end{equation}
+where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}=T$ is the grain temperature (assumed to be the same as the disk temperature, see Paper I), and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well.
+We discuss our choices for $E_x$ for the different volatile species in Sections \ref{sec:CH4} and \ref{sec:N}.
+
+Following \citet{chiang10} and \citet{birnstiel12}, a particle's radial drift velocity can be approximated as 
+\begin{equation}
+\label{eq:rdotact}
+\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{\dot{r}_{\rm gas}}{1+\tau_{\rm s}^2},
+\end{equation}
+where the first term is the drift velocity in a non-accreting disk and the second term accounts for the radial movement of the gas.
+Here $\eta \approx c^2/(2 v_{\rm k}^2)$, where $v_{\rm k}$ is the Keplerian velocity, and $\tau_{\rm s} \equiv \Omega_{\rm k} t_{\rm s}$ is the dimensionless stopping time:
+\begin{equation}
+\label{eq:ts}
+t_{\rm s}= \left\{
+\begin{array}{l l}
+\rho_{\rm s} s / (\rho c), & \quad s < 9 \lambda/4 \,\,\,\ \text{Epstein drag} \\
+4 \rho_{\rm s} s^2 / (9 \rho c \lambda), & \quad s < 9 \lambda/4, \,\text{Re} \lesssim 1 \,\,\,\ \text{Stokes drag,}
+\end{array} 
+\right.
+\end{equation}
+where $\rho$ is the disk mid-plane density, $\lambda$ is the mean free path and Re is the Reynolds number.
+The gas accretion velocity $\dot{r}_{\rm gas}$ is determined from $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, for a fixed $\dot{M}$ and with $\Sigma$ given by Equation (\ref{eq:Sigmaact}).
+
+For a particle of initial size $s_0$, we solve the Equation set (\ref{eq:ddt}) with the initial conditions $s(t_0)=s_0$ and $r(t_0)=r_0$, where $t_0$ is the time at which we start the integration and $r_0$ is the particle's initial location.
+We stop our simulation after $t_{\rm d}=3$ Myr, the disk lifetime, since this is roughly the timescale on which planets form \citep{}, and determine the desorption timescale $t_{\rm des}$ from $s(t_{\rm des})=0$, and thus a particle's desorption distance $r_{\rm des}=r(t_{\rm des})$.
+Our results are insensitive to our choice of $t_0$ as long as $t_0 \ll t_{\rm d}$.
+We note that a particle's size is initially fixed and only changes due to desorption.
+We thus do not take into account processes such as grain coagulation or fragmentation, which nonetheless occur in disks (e.g., \citealt{birnstiel12}, \citealt{perez12}).
+We discuss the effect of these processes on snowline locations in Paper I.
+
+As we show in Paper I, a particle of initial size $s_0$ can experience three outcomes after $t_{\rm d}=3$ Myr: (1) it can remain at its initial location, (2) it can drift towards the host star, then stop without evaporating significantly, and (3) it can completely desorb on a timescale shorter than 3 Myr.
+Particles in scenarios (1) and (2) are thus not affected by radial drift or gas accretion, and the snowline locations are those for a static disk.
+In contrast, the grains in case (3) desorb \textit{instantaneously} and \textit{at a fixed particle-size dependent location} in the disk, regardless of their initial position.
+The snowline locations for these particles will thus be fixed for a given initial particle size and disk model.
+We have found that grains with sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 satisfy this condition for our fiducial disk.
+
+%\emgr{Review disk models, desorption model, relevant timescales.
 State that we use a steady-state disk for the coupled drift-desorption evolution, since it is the most realistic, therefore only summarize the static and steady-state (viscous) disk.
@@
 
-\section{CH4 and C/O Ratios}
+\section{CH$_4$ and C/O Ratios}
+\label{sec:CH4}
 
-\emgr{Discuss observed abundances for CH4 and the choices that we make (no CH4, median value, maximum value).
-State that desorption energies for H2O, CO2 and CH4 are well constrained experimentally, and that the CO2 and CH4 binding energies are only weakly dependent on whether it's pure CO2/CH4 or combined with H2O, but that is not the case for CO (and N2 as we will show in the next section).
-Present new binding energies for CO as pure ice and mixed with water.
-Show Figure 1 and discuss how different CH4 abundances and binding energies affect snowline locations and C/O ratio: CO-H2O mixture (though I think it's rather CO layered on top of H2O) moves the CO snowline inward by $\sim$40 AU (will calculate percentages too); the maximum reasonable abundance of CH4 changes the C/O ratio by less than 10\%.
-Show Figure 2 and quantify the effect of drift and accretion on the CH4 snowline compared to a static disk.
-While CH4 has only a modest effect on the C/O ratio in a static disk, this effect may be larger in a viscous disk, as the C gas abundances inside the CH4 snowline may be enhanced due to the differential motion of the desorbed ices and overall nebular gas (refer to Paper I).
-In this study, however, we neglect these effects and therefore do not include CH4 in estimating the C/O ratio (as an aside, the figures that include CH4 in the C/O ratio with drift and desorption are quite messy due to snowlines overlapping).
-Show Figure 3 and estimate the difference between CO-H2O and CO pure ice snowlines in the case of drift and accretion, as well as the comparisons for the static disk for the CO snowline.}
+Both in Solar system comets and in protoplanetary disks, carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{rodgers02}, \citealt{lodders03}, \citealt{pontoppidan06}).
+However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{mumma96}), which may change the C/O ratio in gas and in dust throughout the disk.
+To quantify the magnitude of this effect, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
+We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max).
+Thus $n_{\rm CH_4-mid}=0.0555 \times n_{\rm H_2O}$ \citep{oberg11a} and $n_{\rm CH_4-mid}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}.
+Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}.
+Since the abundance of carbon grains is uncertain, we assume that all the carbon that is not in the form of CH$_4$ is found in carbon grains, so that we reproduce the Solar C/O ratio (gas+dust) of 0.54.
+
+We determine the location of the H$_2$O, CO$_2$, CO and CH$_4$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}.
+The binding energies of H$_2$O, CO$_2$, CO and CH$_4$ as pure ices are 5800 K, 2000 K, 1388 K and 1300 K, respectively (\citealt{fraser01}, \citealt{collings04}, Fayolle et al., submittted, \citealt{garrod06}).
 
@@
 
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_CH4_steady.pdf}
-%%\vspace{-0.5in}
-%\caption{Desorption distance as a function of initial distance for CH4.... Drift and gas accretion move the CH4 snowline inward by x\%.} 
-%\label{fig:CH4}
-%\end{figure}
+Figure \ref{fig:COstatic} shows the C/O ratio in gas and dust as a function of semimajor axis in a static disk, for different CH$_4$ abundances as outlined above.
+As in \citet{oberg11} and Paper I, a gaseous C/O ratio of unity can be achieved beyond the CO$_2$ snowline, where oxygen gas is significantly depleted (top panel).
+The gas-phase C/O ratio may be further enhanced between the CO$_2$ and CH$_4$ snowlines due to the presence of additional carbon gas from CH$_4$.
+In this region, the C/O ratio increases by 3\% for CH$_4$-mid and by 8\% for  CH$_4$-max, as displayed in the middle and bottom panel of Figure \ref{fig:COstatic}.
+Based on current observations of CH$_4$ abundances, its presence in the disk only modestly affects the C/O ratio.
+Given the larger uncertainties in overall volatile abundances, we can neglect CH$_4$ when estimating the C/O ratio in static disks %\footnote{This may not be the case, however, in a viscous disk, as we show below.}.
+
+As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the environment in which the icy grains reside.
+If CO ice is layered on top of a water ice substrate, its binding energy will be larger than in the pure ice case (1388 K) due to the higher H$_2$O binding energy.
+Fayolle et al.
+(submitted) find a CO binding energy of 834 K in the layered ice scenario.
+We use the model of Section \ref{sec:review} to estimate the movement of the CO snowline for different grain morphologies in a viscous disk.
+
+Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.06$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above.
+The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.06$ cm do not form true snowlines.
+As calculated in Paper I, drift and gas accretion may move the snowlines inwards by up to 40-60\% compared to a static disk, and specifically by up to $\sim50$ \% in the case of the CO snowline.
+This result is preserved for the updated CO binding energies, both for pure ice and on a water ice substrate.
+However, a layered grain structure moves the CO snowline inwards significantly: for our fiducial disk model, $r_{\rm CO, layered} \approx 8.7$ AU.
+Thus if CO is bound to water ice instead of pure ices, the CO snowline may move inward by up to 70\%.
+This large inward movement of the CO snowline implies that C/O ratios of order unity may be reached much closer to the host star if CO is layered on a water ice substrate, and may be inside 10 AU for certain disk parameters.
+
+%\emgr{Discuss observed abundances for CH4 and the choices that we make (no CH4, median value, maximum value).
+State that desorption energies for H2O, CO2 and CH4 are well constrained experimentally, and that the CO2 and CH4 binding energies are only weakly dependent on whether it's pure CO2/CH4 or combined with H2O, but that is not the case for CO (and N2 as we will show in the next section).
+Present new binding energies for CO as pure ice and mixed with water.
+Show Figure 1 and discuss how different CH4 abundances and binding energies affect snowline locations and C/O ratio: CO-H2O mixture (though I think it's rather CO layered on top of H2O) moves the CO snowline inward by $\sim$40 AU (will calculate percentages too); the maximum reasonable abundance of CH4 changes the C/O ratio by less than 10\%.
+Show Figure 2 and quantify the effect of drift and accretion on the CH4 snowline compared to a static disk.
+While CH4 has only a modest effect on the C/O ratio in a static disk, this effect may be larger in a viscous disk, as the C gas abundances inside the CH4 snowline may be enhanced due to the differential motion of the desorbed ices and overall nebular gas (refer to Paper I).
+In this study, however, we neglect these effects and therefore do not include CH4 in estimating the C/O ratio (as an aside, the figures that include CH4 in the C/O ratio with drift and desorption are quite messy due to snowlines overlapping).
+Show Figure 3 and estimate the difference between CO-H2O and CO pure ice snowlines in the case of drift and accretion, as well as the comparisons for the static disk for the CO snowline.}
 
@@
 \section{Nitrogen Carriers and N/O Ratios}
+\label{sec:N}
 
@@
 \section{Discussion}
+\label{sec:discussion}
 

### 2015-12-17 tex/draft2/ms.tex
update

@@
 
+\begin {itemize}
+\item  
+\end {itemize}
+
 \emgr{Background info.
@@
 
-We model the static disk as a minimum mass solar nebula (MMSN), using a prescription for the gas surface density, $\Sigma$, and disk midplane temperature, $T$, similar to that of \citet{chiang10}:
-\begin{subeqnarray}
-\label{eq:disk}
-\Sigma&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
-T &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
-\end{subeqnarray}
-where $r$ is the semimajor axis.
-Based on observations of protoplanetary disks \citep{andrews10}, we choose a flatter surface density than the one assumed by \citet{chiang10}, where $\Sigma \propto r^{-3/2}$.
-
+Following \citet{chiang10}, the temperature profile for a static disk is
+\begin{equation}
+\label{eq:diskT}
+T = 120\, (r/\text{AU})^{-3/7} \,\,\text{K},
+\end{equation}
+%We model the static disk as a minimum mass solar nebula (MMSN), using a prescription for the gas surface density, $\Sigma$, and disk midplane temperature, $T$, similar to that of \citet{chiang10}:
+%\begin{subeqnarray}
+%\label{eq:disk}
+%\Sigma&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
+%T &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
+%\end{subeqnarray}
+where $r$ is the semimajor axis.% Based on observations of protoplanetary disks \citep{andrews10}, we choose a flatter surface density than the one assumed by \citet{chiang10}, where $\Sigma \propto r^{-3/2}$.
 We use the \citet{shakura73}  steady-state disk solution to model the viscous disk.
-Solving the Equation set of Appendix A in Paper I yields an expression for the temperature profile in a steady-state disk:
+From Paper I, the viscous disk temperature profile is computed as % Solving the Equation set of Appendix A in Paper I yields an expression for the temperature profile in a steady-state disk:
 \begin{equation}
@@
 
-\section{Nitrogen Carriers and N/O Ratios}
+\section{Nitrogen and N/O Ratios}
 \label{sec:N}
 
-\emgr{Similar to the previous section, but with more details.
+In addition to carbon and oxygen, nitrogen is
+
+%\emgr{Similar to the previous section, but with more details.
 Discuss that nitrogen is abundant in the solar system and disks and primarily found as N2.

### 2015-12-30 Postdoc Applications/research_proposal_hilke.tex
update

--- 2015-11-13 Postdoc Applications/AnaMariaPisoResearchProposal.tex
@@
 \lhead{Ana-Maria Piso}
-\rhead{Postdoctoral Position in Exoplanet Theory Research Proposal}
+\rhead{Postdoctoral Position Research Proposal}
 \date{}
@@
 
-\title{\Large Postdoctoral Position in Exoplanet Theory Research Proposal: \\
+\title{\Large Postdoctoral Position Research Proposal: \\
 The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets\vspace{-2ex}}
@@
 
-\textbf{The University of Chicago is the ideal place for me to pursue my postdoctoral research}, due to its vibrant community of experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
+\textbf{MIT is the ideal place for me to pursue my postdoctoral work.} Its research facilities and academic excellence are unparalleled.
+I would love the opportunity to work with \textbf{Prof.
+Hilke Schlichting}, an expert in planet formation theory and dynamics --- such a collaboration would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards.
+I would also like to collaborate with \textbf{Prof.
+Sara Seager}, a co-investigator of the TESS mission and a leading theorist in atmospheric chemistry, as well as with members of her group, who specialize in a broad range of exoplanet theory and computational topics.
+\textbf{Dr.
+Margaret Pan}, an excellent theorist in planetary dynamics and protoplanetary disks, will join the MIT EAPS department in early 2016 --- it would be a pleasure to work with her as well.
+Additionally, the MIT Physics department hosts leaders in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof.
+Joshua Winn}, which presents great prospects in connecting my theoretical research work with observations.
+\\
+
+%Finally, in order to connect my theoretical work with observations, I would like to collaborate with \textbf{Prof. Joshua Winn}, who is an expert in discovering and characterizing exoplanets. \\
+
+%\textbf{The University of Chicago is the ideal place for me to pursue my postdoctoral research}, due to its vibrant community of experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences.
 I would love the opportunity to work with \textbf{Prof.
@@
 In the Department of Geophysical Sciences, I would like to collaborate with \textbf{Prof.
-Fred Ciesla}, a leader in protplanetary disk dynamics, chemical composition and evolution.
+Fred Ciesla}, a leader in protoplanetary disk dynamics, chemical composition and evolution.
 Additionally, the Department of Astronomy hosts experts in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof.
@@
 $[3]$ Debes, J. H., Jang-Condell, H., Weinberger, A. J., Roberge, A., \& Schneider, G. 2013, ApJ, 771, 45 \\
-$[4]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ArXiv e-prints, arXiv:1504.05586 \\
+$[4]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ApJ, 814, 66 \\
 $[5]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21 \\
 $[6]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
-$[7]$ Piso, A.-M. A., Öberg, K.I., Birnstiel, T, \& Murray-Clay, R.M., ApJ, resubmitted after referee report \\
+$[7]$ Piso, A.-M. A., \"Oberg, K. I. , Birnstiel, T., \& Murray-Clay, R. A. 2015, ApJ, 815, 109 \\
 $[8]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\

### 2016-01-25 tex/draft2/ms.tex
update

@@
 \centering
-\includegraphics[width=\textwidth]{../../figs/CNO_and_snowlines.pdf}
+\includegraphics[width=\textwidth]{../../figs/CNO_and_snowlines_2.pdf}
 %\vspace{-0.5in}
@@
 We determine the location of the H$_2$O, CO$_2$, CO and CH$_4$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}.
-The binding energies of H$_2$O, CO$_2$, CO and CH$_4$ as pure ices are 5800 K, 2000 K, 1388 K and 1300 K, respectively (\citealt{fraser01}, \citealt{collings04}, Fayolle et al., submittted, \citealt{garrod06}).
+The binding energies of H$_2$O, CO$_2$, CO and CH$_4$ as pure ices are 5800 K, 2000 K, 834 K and 1300 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}).
 
@@
 As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the environment in which the icy grains reside.
-If CO ice is layered on top of a water ice substrate, its binding energy will be larger than in the pure ice case (1388 K) due to the higher H$_2$O binding energy.
-Fayolle et al.
-(submitted) find a CO binding energy of 834 K in the layered ice scenario.
+If CO ice is layered on top of a water ice substrate, its binding energy will be larger than in the pure ice case (834 K) due to the higher H$_2$O binding energy.
+\citet{fayolle16} find a CO binding energy of 1388 K in the layered ice scenario.
 We use the model of Section \ref{sec:review} to estimate the movement of the CO snowline for different grain morphologies in a viscous disk.
@@
 
-In addition to carbon and oxygen, nitrogen is
+In addition to carbon and oxygen, nitrogen is another abundant volatile in the Solar system and in disks.
+Chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
+Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}).
+Moreover, the Rosetta spacecraft has recently made the first direct measurement of N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
+In addition to N$_2$, a fraction of the nitrogen abundances may be also carried by NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).
+
+Because of the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk
+may be even more enhanced than the C/O ratio compared to its average value in the disk.
+Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
+In this study, we quantify this effect in protoplanetary disks.
+We assume that the main nitrogen-bearing species are N$_2$ and NH$_3$, since other volatiles that contain nitrogen have significantly lower abundances in comparison (e.g., \citealt{mumma11}).
+We use the measured total nitrogen abundance in the Solar system, $n_{\rm N}=8 \times 10^{-5} n_{\rm H}$ \citep{lodders03}, where $n_{\rm H}$ is the hydrogen abundance in the disk midplane.
+Similarly to the case of CH$_4$ (see Section \ref{sec:CH4}), we explore the parameter space of possible NH$_3$ abundances using data from the Spitzer c2d Legacy ice survey, as follows: (1) no NH$_3$, (2) the median NH$_3$ observed abundance $n_{\rm NH_3-mid}=0.055 \times n_{\rm H_2O}$ \citep{oberg11a}, and (3) the maximum observed NH$_3$ abundance $n_{\rm NH_3-max}=0.1537 \times n_{\rm H_2O}$ \citep{bottinelli10}.
+In each case, the N$_2$ abundance then simply follows as $n_{\rm N_2}=(n_{\rm N}-n_{\rm NH_3})/2$.
+We determine the locations of the N$_2$ and NH$_3$ snowlines by balancing desorption with readsorption \citep{hollenbach09}, with N$_2$ and NH$_3$ pure ice binding energies of  767 K and  2965 K, respectively (\citealt{fayolle16}, \citealt{martin14}).
+
+Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk for our three choices of the NH$_3$ abundance.
+For comparison, the horizontal dashed line shows the average N/O ratio in the disk.
+As expected, the gaseous N/O ratio generally exhibits an increasing trend towards the outer disk as more oxygen gas is depleted, with small decreases between the NH$_3$ and CO$_2$ snowlines (by x\% for NH$_3$-mid and by x\% for NH$_3$-max, respectively) due to NH$_3$ freeze-out.
+More importantly, the gas-phase N/O ratio in the outer disk is enhanced by more than a factor of three compared to its average value.
+This enhancement is more pronounced than the C/O gas-phase enhancement of a factor of two in the outer disk (see Figure \ref{fig:COstatic}).
+The N/O ratio reaches particularly high values between the CO and N$_2$ snowlines, where all the oxygen is now contained in grains.
 
@@
 %\end{figure}
+
+As is the case for CO, the N$_2$ binding energy is also strongly dependent on the ice environment in which it resides.
+If N$_2$ is layered on a water ice substrate, the N$_2$ binding energy is 1266 K \citep{fayolle16}.
+Figure \ref{fig:NO_ratio} shows the H$_2$O, CO$_2$, CO and N$_2$ snowline locations in a viscous disk for particles with initial size $\sim0.06$ cm $\lesssim s \lesssim$ 7 m, and with the CO and N$_2$ snowlines calculated assuming different grain morphologies as explained above, as well as upper estimates for the N/O ratio in gas throughout the disk.
+The N$_2$ snowline moves inward by x\% in the pure ice scenario and by x\%....
 

### 2016-01-25 tex/draft2/ms.tex
update

@@
 \begin{equation}
-\label{eq:Tdact}
-T_{\rm act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
+\label{eq:activeT}
+T^4 =\Big[\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}\Big]^4 + T_{\rm irr}^4,
 \end{equation}
-Here $G$ is the gravitational constant, $\kappa_0=2 \times 10^{-6}$ is a dimensionless coefficient in the opacity law $\kappa=\kappa_0 T_{\rm act}^2$, $M_*=M_{\odot}$ is the mass of the central star, $\mu=2.35$ is the mean molecular weight of the nebular gas, $m_{\rm p}$ is the proton mass, $\Omega_{\rm k}=\sqrt{G M_{\odot}/r^3}$ is the Keplerian angular velocity, $\alpha=0.01$ is a dimensionless coefficient (see below for details), $k_{\rm B}$ is the Boltzmann constant, and $\sigma$ is the Stefan-Boltzmann constant.
-The final midplane temperature profile is computed as
-\begin{equation}
-\label{eq:activeT}
-T^4 = T_{\rm act}^4 + T_{\rm irr}^4,
-\end{equation}
-where $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}). We use this expression because in addition to accretion heating, stellar irradiation also contributes to the disk thermal structure.
+where $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}).
+Here $G$ is the gravitational constant, $\kappa_0=2 \times 10^{-6}$ is a dimensionless opacity coefficient, $M_*=M_{\odot}$ is the mass of the central star, $\mu=2.35$ is the mean molecular weight of the nebular gas, $m_{\rm p}$ is the proton mass, $\Omega_{\rm k}=\sqrt{G M_{\odot}/r^3}$ is the Keplerian angular velocity, $\alpha=0.01$ is a dimensionless coefficient (see below for details), $k_{\rm B}$ is the Boltzmann constant, and $\sigma$ is the Stefan-Boltzmann constant.
 
-The steady-state disk has an $\alpha$-viscosity prescription, where the kinematic viscosity is $\mu=\alpha c H$.
+%The final midplane temperature profile is computed as 
+%\begin{equation}
+%\label{eq:activeT}
+%T^4 = T_{\rm act}^4 + T_{\rm irr}^4,
+%\end{equation}
+%where $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}). We use this expression because in addition to accretion heating, stellar irradiation also contributes to the disk thermal structure.
+
+The steady-state disk has an $\alpha$-viscosity prescription, where the kinematic viscosity is $\nu=\alpha c H$.
 Here $c \equiv \sqrt{k_{\rm B} T /(\mu m_{\rm p})}$ is the isothermal sound speed (with $T$ from Equation \ref{eq:activeT}), and $H \equiv c/\Omega_{\rm k}$ is the disk scale height.
@@
 
-\section{Discussion}
-\label{sec:discussion}
+%\section{Discussion}
+%\label{sec:discussion}
 
-\emgr{Discuss how entrapment of volatiles by H2O affects volatile abundances and C/O ratios.
+%\emgr{Discuss how entrapment of volatiles by H2O affects volatile abundances and C/O ratios.
 Re-emphasize the fact that the C/O and N/O ratios are upper estimates, and that CH4 and NH3 might matter in a viscous disk.
@@
 
-\emgr{Maybe we can include the summary in the discussion section?}
+In this paper we explore the role of icy grain morphology and disk dynamics on the snowline locations of major volatile carrier molecules and the C/N/O ratios in protoplanetary disks.
+We enhance the coupled drift-desorption model developed in \citet{piso15b} by adding more carbon- and nitrogen-bearing species into our framework and by considering different environments in which the icy grains reside.
+Our results can be summarized as follows:
+
+\begin{enumerate}
+
+\item Due to the high volatility of N$_2$, the gaseous N/O ratio in the outer disk is enhanced by more than a factor of three compared to its average value.
+This enhancement is more pronounced than in the case of the gas-phase C/O ratio, which is increased by a factor of two compared to the stellar value.
+Moreover, the N/O ratio in gas is expected to be very large between the CO and N$_2$ snowlines due to the complete depletion of oxygen gas in this region
+
+\item The presence of some carbon in the form of CH$_4$ and of some nitrogen in the form of NH$_3$ only modestly affects our results for the C/O and N/O ratios, respectively.
+In both cases, large C/O and N/O ratios in the outer disk are preserved.
+
+\item Grain composition sensitively affects the CO and N$_2$ snowline locations.
+If CO and N$_2$ are layered on a water-ice substrate rather than existing as pure ices, their snowlines move inward by up to $\sim$70 \%.
+This effect is separate from that of radial drift and viscous gas accretion, which also cause an inward movement of the snowlines by up to $\sim$50 \%.
+
+\end{enumerate}
+
+Our results have direct consequences for the composition of nascent giant planets.
+The considerable inward movement of the CO and N$_2$ snowlines due to the ice grains residing in a water ice environment rather than as pure ices implies than giant planets with high C/O and/or N/O ratios in their atmospheres may form closer in than previously predicted by theoretical models.
+Moreover, our model shows that wide separation gas giants may have an excess of nitrogen in their envelopes, which may be used to trace their origins.
+In future work, we plan to add new levels of complexity to our model in terms of disk chemistry, dynamics, and planetary dynamics, thus forming a solid framework for understanding the origins of gas giants.
+
+%\emgr{Maybe we can include the summary in the discussion section?}
 

### 2016-01-25 tex/draft2/ms.tex
update

@@
 
-\begin {itemize}
-\item  
-\end {itemize}
+\begin {enumerate}
+\item The chemical composition of protoplanetary disks is largely dictated by the freeze-out of volatile species, such as carbon, oxygen and nitrogen carriers.
 
-\emgr{Background info.
+\item Carbon and oxygen bearing molecules, such as H$_2$O, CO$_2$ and CO, as well as the C/O ratio in protoplanetary disks/giant planet atmospheres have been extensively studied from a theoretical standpoint (refs, Oberg+11, Ali-Dib+14, Madhusudhan+14, Molliere+15, etc.), and volatile snowlines have been detected in disks, such as H$_2$O and CO.
+However, both observations and chemical models (refs) have shown that other volatiles are abundant in disks, for example nitrogen bearing species and hydrocarbons, thus shaping the disk composition.
+
+\item Moreover, the snowline locations of these volatiles strongly depend on the grain morphology and the ice environment in which they reside.
+Laboratory experiments (cite Fayolle+16 and reference therein) have shown that volatiles such as CO and N$_2$ have significantly different binding energies depending on whether they are pure ices or layered on a water-ice substrate (reference Sections 3 and 4 for details).
+
+\item Thus, there are other important snowlines in disks.
+Discuss Figure 1.
+Re-emphasize how this directly changes chemical abundances and C/N/O ratios throughout the disk, and thus the compositions of forming giant planets.
+
+\item We thus need a more thorough theoretical framework to understand how the processes above affect the C/O and N/O ratios in disks.
+To address this issue, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, in addition to disk dynamical processes.
+
+\end {enumerate}
+
+%\emgr{Background info.
 Importance of volatiles in disks and planetary atmospheres, detections of snowlines in disks, C/O ratios etc.

### 2016-01-25 tex/draft2/ms.tex
update

@@
 In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, the ice environment in which these carriers reside, and the presence of snowlines of major volatiles at different distances from the central star.
-We explore the effects of dynamical processes, molecular compositions and abundances, and the ice morphology of dust grains in disks on the snowline locations of the main C, O and N carriers, and their consequences for the C/N/O ratio in gas and dust throughout the disk.
+We explore the effects of dynamical processes, molecular compositions and abundances, and the ice morphology of dust grains in disks on the snowline locations of the main C, O and N carriers, and their consequences for the C/N/O ratios in gas and dust throughout the disk.
 We find that radial drift and accretion alone can reduce the snowline radii of the main C, O and N carriers, i.e.
@@
 
-\item Carbon and oxygen bearing molecules, such as H$_2$O, CO$_2$ and CO, as well as the C/O ratio in protoplanetary disks/giant planet atmospheres have been extensively studied from a theoretical standpoint (refs, Oberg+11, Ali-Dib+14, Madhusudhan+14, Molliere+15, etc.), and volatile snowlines have been detected in disks, such as H$_2$O and CO.
+\item Carbon and oxygen bearing molecules, such as H$_2$O, CO$_2$ and CO, as well as the C/O ratio in protoplanetary disks/giant planet atmospheres have been extensively studied from a theoretical standpoint (refs, Oberg+11, Ali-Dib+14, Madhusudhan+14, Molliere+15, etc.), and volatile snowlines have been detected in disks, such as H$_2$O and CO (refs).
 However, both observations and chemical models (refs) have shown that other volatiles are abundant in disks, for example nitrogen bearing species and hydrocarbons, thus shaping the disk composition.
@@
 Discuss Figure 1.
-Re-emphasize how this directly changes chemical abundances and C/N/O ratios throughout the disk, and thus the compositions of forming giant planets.
+Re-emphasize how this directly changes chemical abundances and C/N/O ratios throughout the disk, and thus the compositions of nascent giant planets.
 
@@
 %\vspace{-0.5in}
-\caption{CNO abundance in grains...} 
+\caption{The total carbon, nitrogen and oxygen abundance as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and layered on a water ice substrate (bottom panel).
+Relevant volatile snowlines are marked by the vertical dashed lines.
+The grain abundances are calculated as a function of the observed median CH$_4$ and NH$_3$ abundances in protostellar cores (see Sections \ref{sec:CH4} and \ref{sec:N}).
+The total grain abundance increases with semimajor axis as more and more species freeze out.}
 \label{fig:CNOstatic}
@@
 %\end{subeqnarray}
-where $r$ is the semimajor axis.% Based on observations of protoplanetary disks \citep{andrews10}, we choose a flatter surface density than the one assumed by \citet{chiang10}, where $\Sigma \propto r^{-3/2}$.
+where $r$ is the semimajor axis.
 We use the \citet{shakura73}  steady-state disk solution to model the viscous disk.
@@
 For a particle of initial size $s_0$, we solve the Equation set (\ref{eq:ddt}) with the initial conditions $s(t_0)=s_0$ and $r(t_0)=r_0$, where $t_0$ is the time at which we start the integration and $r_0$ is the particle's initial location.
-We stop our simulation after $t_{\rm d}=3$ Myr, the disk lifetime, since this is roughly the timescale on which planets form \citep{}, and determine the desorption timescale $t_{\rm des}$ from $s(t_{\rm des})=0$, and thus a particle's desorption distance $r_{\rm des}=r(t_{\rm des})$.
+We stop our simulation after $t_{\rm d}=3$ Myr, the disk lifetime, since this is roughly the timescale on which planets form, and determine the desorption timescale $t_{\rm des}$ from $s(t_{\rm des})=0$, and thus a particle's desorption distance $r_{\rm des}=r(t_{\rm des})$.
 Our results are insensitive to our choice of $t_0$ as long as $t_0 \ll t_{\rm d}$.
@@
 The snowline locations for these particles will thus be fixed for a given initial particle size and disk model.
-We have found that grains with sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 satisfy this condition for our fiducial disk.
+We have found that grains with sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m satisfy this condition for our fiducial disk.
 
@@
 We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max).
-Thus $n_{\rm CH_4-mid}=0.0555 \times n_{\rm H_2O}$ \citep{oberg11a} and $n_{\rm CH_4-mid}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}.
+Thus $n_{\rm CH_4-mid}=0.0555 \times n_{\rm H_2O}$ \citep{oberg11a} and $n_{\rm CH_4-max}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}, where $n_{\rm H_2O}$ is the total H$_2$O abundance.
 Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}.
@@
 %\vspace{-0.5in}
-\caption{C/O ratio in a static disk for different CH4 abundances and CO binding energies...} 
+\caption{The C/O ratio in gas (solid lines) and dust (dashed lines) as a function of semimajor axis in a static disk, assuming no carbon is present in the form of CH$_4$ (top panel), the median observed CH$_4$ abundance is assumed (middle panel), and the maximum observed CH$_4$ abundance is assumed (bottom panel).
+The C/O estimates are performed assuming that the CO ices are in pure form.
+The vertical dashed lines mark the snowline locations of the main C and O carriers.
+The horizontal dashed lines represent the stellar C/O value.
+The presence of methane only modestly increases the C/O ratio in gas between the CO$_2$ and CH$_4$ snowlines.}
 \label{fig:COstatic}
@@
 Based on current observations of CH$_4$ abundances, its presence in the disk only modestly affects the C/O ratio.
-Given the larger uncertainties in overall volatile abundances, we can neglect CH$_4$ when estimating the C/O ratio in static disks %\footnote{This may not be the case, however, in a viscous disk, as we show below.}.
+Given the larger uncertainties in overall volatile abundances, we can neglect CH$_4$ when estimating the C/O ratio in static disks.
+%\footnote{This may not be the case, however, in a viscous disk, as we show below.}.
 
 As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the environment in which the icy grains reside.
-If CO ice is layered on top of a water ice substrate, its binding energy will be larger than in the pure ice case (834 K) due to the higher H$_2$O binding energy.
+If CO ice is layered on a water ice substrate, its binding energy will be larger than in the pure ice case (834 K) due to the higher H$_2$O binding energy.
 \citet{fayolle16} find a CO binding energy of 1388 K in the layered ice scenario.
@@
 Thus if CO is bound to water ice instead of pure ices, the CO snowline may move inward by up to 70\%.
-This large inward movement of the CO snowline implies that C/O ratios of order unity may be reached much closer to the host star if CO is layered on a water ice substrate, and may be inside 10 AU for certain disk parameters.
+This large inward movement of the CO snowline implies that C/O ratios of order unity may be reached much closer to the host star if CO is layered on a water ice substrate, and the CO snowline may be inside 10 AU for certain disk parameters.
 
@@
 %\vspace{-0.5in}
-\caption{C/O ratio as function of semimajor axis for CO combined with H2O (top panel) and pure CO ice (bottom panel)....
-Drift and gas accretion move the CO snowlines inward by x\% and y\%, respectively.}
+\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO residing in a water ice environment (top panel) or as pure CO ice (bottom panel).
+The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
+The C/O ratio in a static disk (black lines) is shown for comparison.
+The arrows show that the C/O ratio in gas will decrease inside the H$_2$O and CO$_2$
+snowlines in the viscous disk, as the relative fluxes of the desorbed icy
+particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details).
+Radial drift and gas accretion move the snowlines inward by 40-60\%.
+The presence of CO in a water ice environment rather than as pure ice moves the CO snowline significantly inward by $\sim$70\%.}
 \label{fig:CO_ratio}
@@
 Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}).
-Moreover, the Rosetta spacecraft has recently made the first direct measurement of N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
+Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
 In addition to N$_2$, a fraction of the nitrogen abundances may be also carried by NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).
@@
 
-Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk for our three choices of the NH$_3$ abundance.
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/N_O_ratio.pdf}
+%\vspace{-0.5in}
+\caption{The N/O ratio in gas (solid lines) and dust (dashed lines) as a function of semimajor axis in a static disk, assuming no nitrogen is present in the form of NH$_3$ (top panel), the median observed NH$_3$ abundance is assumed (middle panel), and the maximum observed NH$_3$ abundance is assumed (bottom panel).
+The N/O estimates are performed assuming that the CO and N$_2$ ices are in pure form.
+The vertical dashed lines mark the snowline locations of the main C,O and N carriers.
+The horizontal dashed lines represent the average N/O value in the disk.
+The gas-phase N/O ratio is highly enhanced in the outer disk (by more than a factor of three) compared to its average value.
+The arrows mark a highly elevated N/O ratio in gas between the CO and N$_2$ snowlines due to the depletion of oxygen gas in this region.
+The presence of NH$_3$ moderately decreases the N/O ratio in gas between the NH$_3$ and CO$_2$ snowlines.}
+\label{fig:Nstatic}
+\end{figure}
+
+Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk, for our three choices of the NH$_3$ abundance.
 For comparison, the horizontal dashed line shows the average N/O ratio in the disk.
-As expected, the gaseous N/O ratio generally exhibits an increasing trend towards the outer disk as more oxygen gas is depleted, with small decreases between the NH$_3$ and CO$_2$ snowlines (by x\% for NH$_3$-mid and by x\% for NH$_3$-max, respectively) due to NH$_3$ freeze-out.
+As expected, the gaseous N/O ratio generally exhibits an increasing trend towards the outer disk as more oxygen gas is depleted, with small decreases between the NH$_3$ and CO$_2$ snowlines (by 16\% for NH$_3$-mid and by 18\% for NH$_3$-max, respectively) due to NH$_3$ freeze-out.
 More importantly, the gas-phase N/O ratio in the outer disk is enhanced by more than a factor of three compared to its average value.
@@
 The N/O ratio reaches particularly high values between the CO and N$_2$ snowlines, where all the oxygen is now contained in grains.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{../../figs/N_O_water_ice.pdf}
+%\vspace{-0.5in}
+\caption{N/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO and N$_2$ residing in a water ice environment (top panel) or as pure ices (bottom panel).
+The H$_2$O, CO$_2$, CO and N$_2$ snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
+The N/O ratio in a static disk (black lines) is shown for comparison.
+The arrows show that the N/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in the viscous disk, as the relative fluxes of the desorbed icy
+particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details).
+Radial drift and gas accretion move the N$_2$ snowline inward by up to $\sim$50\% compared to a static disk.
+The presence of N$_2$ in a water ice environment rather than as pure ice moves the N$_2$ snowline significantly inward by $\sim$70\%.
+The results of highly enhanced gas-phase N/O ratios in the outer disk compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}
+\label{fig:NO_ratio}
+\end{figure}
 
@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{../../figs/N_O_ratio.pdf}
-%\vspace{-0.5in}
-\caption{N/O ratio in a static disk for different NH3 abundances and N2 binding energies...} 
-\label{fig:Nstatic}
-\end{figure}
-
 %\begin{figure}[h!]
@@
 
-As is the case for CO, the N$_2$ binding energy is also strongly dependent on the ice environment in which it resides.
+As in the case for CO, the N$_2$ binding energy is also strongly dependent on the ice environment in which the grain resides.
 If N$_2$ is layered on a water ice substrate, the N$_2$ binding energy is 1266 K \citep{fayolle16}.
-Figure \ref{fig:NO_ratio} shows the H$_2$O, CO$_2$, CO and N$_2$ snowline locations in a viscous disk for particles with initial size $\sim0.06$ cm $\lesssim s \lesssim$ 7 m, and with the CO and N$_2$ snowlines calculated assuming different grain morphologies as explained above, as well as upper estimates for the N/O ratio in gas throughout the disk.
-The N$_2$ snowline moves inward by x\% in the pure ice scenario and by x\%....
+Figure \ref{fig:NO_ratio} shows the H$_2$O, CO$_2$, CO and N$_2$ snowline locations in a viscous disk for particles with initial size $\sim0.06$ cm $\lesssim s \lesssim$ 7 m, and with the CO and N$_2$ snowlines calculated assuming different grain morphologies as explained above, as well as estimates for the N/O ratio throughout the disk.
+For simplicity, we assume that all nitrogen is bound in N$_2$.
+This choice is justified since the presence of some NH$_3$ only moderately changes the N/O ratio (see Figure \ref{fig:Nstatic}), and since we are primarily interested in the N$_2$ snowline locations rather than exact values for the N/O ratio.
+The innermost N$_2$ snowlines in the viscous disk, created by particles with $s \sim 7$ m for our fiducial model, are located at $r_{\rm N_2, pure} \approx 42$ AU for N$_2$ as pure ice and at $r_{\rm N_2, layered} \approx 11$ AU for N$_2$ layered on water ice.
+Compared to the static snowlines, i.e.
+$r_{\rm N_2, pure, static} \approx 79$ AU and $r_{\rm N_2, layered, static} \approx 25$ AU, drift and gas accretion move the snowlines inward by up to $\sim$50 \%, similar to the case for CO (see Section \ref{sec:CH4}).
+By comparing our results in the viscous disk, we find that the N$_2$ snowline moves inward by more than 70\% if N$_2$ resides in a water ice environment rather than as pure ice.
+Similarly to the CO case, the N$_2$ snowline may be close to or inward of 10 AU for certain disk models.
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{../../figs/N_O_water_ice.pdf}
-%\vspace{-0.5in}
-\caption{N/O ratio as function of semimajor axis for N2 combined with H2O (top panel) and pure N2 ice (bottom panel)....
-Drift and gas accretion move the N2 snowlines inward by x\% and y\%, respectively.
-Overabundance of gas-phase N/O between the CO and N2 snowlines, marked by the vertical red dash-dotted lines for the largest drifting particles in our model.}
-\label{fig:NO_ratio}
-\end{figure}
+%The N$_2$ snowline moves inward by x\% in the pure ice scenario and by x\%....
 
@@
 If CO and N$_2$ are layered on a water-ice substrate rather than existing as pure ices, their snowlines move inward by up to $\sim$70 \%.
-This effect is separate from that of radial drift and viscous gas accretion, which also cause an inward movement of the snowlines by up to $\sim$50 \%.
+This effect is separate from that of radial drift and viscous gas accretion, which also cause an inward movement of the CO and N$_2$ snowlines by up to $\sim$50 \%.
 

### 2016-02-21 tex/draft2/ms.tex
update

@@
 If CO and N$_2$ are bound to water ice instead of pure ices, their snowlines move inward by $\sim$$70$\%.
-Both of these effects substantially change the disk regions where C/O and N/O are enhanced over the stellar value.
+Both of these effects substantially change the disk regions where C/O and N/O are enhanced over the stellar value, and taken together they may move the CO and N$_2$ snowlines by a factor of $\sim$7 inward.
 In the outer disk, the gaseous C/O and N/O are enhanced by factors of $\sim$$2$ and $\sim$$3$, respectively.
@@
 
-\begin {enumerate}
-\item The chemical composition of protoplanetary disks is largely dictated by the freeze-out of volatile species, such as carbon, oxygen and nitrogen carriers.
+%\begin {enumerate}
+The chemical composition of protoplanetary disks is largely dictated by the freeze-out of volatile species, such as carbon, oxygen and nitrogen carriers.
+The snowline locations of volatile molecules are crucial in determining disk chemical abundances in gas and dust, as well as planet compositions.
 
-\item Carbon and oxygen bearing molecules, such as H$_2$O, CO$_2$ and CO, as well as the C/O ratio in protoplanetary disks/giant planet atmospheres have been extensively studied from a theoretical standpoint (refs, Oberg+11, Ali-Dib+14, Madhusudhan+14, Molliere+15, etc.), and volatile snowlines have been detected in disks, such as H$_2$O and CO (refs).
-However, both observations and chemical models (refs) have shown that other volatiles are abundant in disks, for example nitrogen bearing species and hydrocarbons, thus shaping the disk composition.
+Carbon and oxygen bearing molecules, such as H$_2$O, CO$_2$ and CO, as well as the carbon-to-oxygen (C/O) ratio in protoplanetary disks and in giant planet atmospheres have been extensively studied from a theoretical standpoint (\citealt{oberg11}, \citealt{alidib14}, \citealt{madhu14}, \citealt{molliere15}), and the snowlines of volatiles such as H$_2$O and CO have been detected (\citealt{zhang13}, \citealt{qi13}).
+However, disk chemistry is complex, and it involves many species besides carbon and oxygen carriers \citep{henning13}.
+In particular, several other volatiles have been detected in disks, such as nitrogen bearing species and hydrocarbons (e.g., \citealt{mandell12}), which adds further complexity to the disk composition.
 
-\item Moreover, the snowline locations of these volatiles strongly depend on the grain morphology and the ice environment in which they reside.
-Laboratory experiments (cite Fayolle+16 and reference therein) have shown that volatiles such as CO and N$_2$ have significantly different binding energies depending on whether they are pure ices or layered on a water-ice substrate (reference Sections 3 and 4 for details).
+ %However, both observations and chemical models (refs) have shown that other volatiles are abundant in disks, for example nitrogen bearing species and hydrocarbons, thus shaping the disk composition.
 
-\item Thus, there are other important snowlines in disks.
-Discuss Figure 1.
-Re-emphasize how this directly changes chemical abundances and C/N/O ratios throughout the disk, and thus the compositions of nascent giant planets.
+Moreover, volatile snowline locations strongly depend on the grain morphology and the ice environment in which they reside.
+Laboratory experiments (\citealt{fayolle16} and reference therein) have shown that volatiles such as CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
+This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Sections \ref{sec:CH4} and \ref{sec:N}).
 
-\item We thus need a more thorough theoretical framework to understand how the processes above affect the C/O and N/O ratios in disks.
-To address this issue, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, in addition to disk dynamical processes.
+ It follows that there are several important snowlines in disks, determined both by the disk composition and ice morphology.
+Figure \ref{fig:CNOstatic} shows an example of the H$_2$O, CO$_2$, CO, CH$_4$, NH$_3$ and N$_2$ snowlines in a static disk, with CO and N$_2$ as pure and water dominated ices, in the top and bottom panel, respectively.
+The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as function of the hydrogen abundance.
+We note that we display this Figure here for qualitative purposes; the quantitative choices that we make for volatile abundances are discussed in Sections \ref{sec:CH4} and \ref{sec:N}.
+As expected, the total grain abundance increases with semimajor axis, as more and more species sublimate.
+More importantly, the CO and N$_2$ snowlines move several tens of AU inward if the ices are water dominated rather than pure.
+This changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ.
 
-\end {enumerate}
+% Disk dynamical processes such as radial drift of solids and viscous gas accretion onto the central store also change the position of the snowlines, as demonstrated by \citet{piso15b} (hereafter Paper I), which introduces another element of uncertainty in the snowline locations.
+
+In order to explain this added chemical and physical complexity, a more thorough theoretical framework is needed.
+In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, in addition to disk dynamical processes.
+
+This paper is organized as follows.
+In Section \ref{sec:review}, we review the drift-desorption model developed in Paper I.
+We discuss the effect of different abundances of the main carbon and oxygen carriers, grain morphology and dynamical processes on the snowline locations of carbon and oxygen bearing species and the C/O ratio in Section \ref{sec:CH4}.
+In Section \ref{sec:N}, we perform the same analyses for the main nitrogen carriers and the nitrogen-to-oxygen (N/O) ratio.
+We summarize our findings in Section \ref{sec:summary}.
+
+%\end {enumerate}
 
@@
 
-\section{CH$_4$ and C/O Ratios}
+\section{The Effect of CO Ice Morphology and CH$_4$ on the C/O Ratio}
 \label{sec:CH4}
+
+\subsection{Role of CH$_4$}
 
@@
 
+\subsection{CO Ice Environment and Disk Dynamics}
+
 As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the environment in which the icy grains reside.
@@
 Thus if CO is bound to water ice instead of pure ices, the CO snowline may move inward by up to 70\%.
+By taking into account both disk dynamics and of CO in a water dominated environment, we find that the CO snowline may move inward by a factor of $\sim$$7$ compared to a static disk with CO as pure ice.
 This large inward movement of the CO snowline implies that C/O ratios of order unity may be reached much closer to the host star if CO is layered on a water ice substrate, and the CO snowline may be inside 10 AU for certain disk parameters.
@@
 
-\section{Nitrogen and N/O Ratios}
+\section{The Effect of N$_2$ Ice Morphology and NH$_3$ on the N/O Ratio}
 \label{sec:N}
+
+\subsection{The Role of NH$_3$}
 
@@
 
+\subsection{N$_2$ Ice Environment and Disk Dynamics}
+
 As in the case for CO, the N$_2$ binding energy is also strongly dependent on the ice environment in which the grain resides.
@@
 By comparing our results in the viscous disk, we find that the N$_2$ snowline moves inward by more than 70\% if N$_2$ resides in a water ice environment rather than as pure ice.
-Similarly to the CO case, the N$_2$ snowline may be close to or inward of 10 AU for certain disk models.
+Similarly to the CO case, the N$_2$ snowline may move by a factor of $\sim$$7$ inward if we account for both disk dynamics and ice morphology (water dominated compared to pure), and may be close to or inward of 10 AU for certain disk models.
 
@@
 This effect is separate from that of radial drift and viscous gas accretion, which also cause an inward movement of the CO and N$_2$ snowlines by up to $\sim$50 \%.
+
+\item The locations of the CO and N$_2$ snowlines are uncertain when we take consider both viscous versus static disk, and pure versus water dominated ices.
+The snowlines in a viscous disk with CO in a water environment are by up to a factor of $\sim$7 closer to the host star that in a static disk with CO as pure ice.
 

### 2016-02-22 tex/draft2/ms.tex
update

@@
 H$_2$O, CO$_2$, CO and N$_2$, by 40-60\% compared to static disks.
-If CO and N$_2$ are bound to water ice instead of pure ices, their snowlines move inward by $\sim$$70$\%.
-Both of these effects substantially change the disk regions where C/O and N/O are enhanced over the stellar value, and taken together they may move the CO and N$_2$ snowlines by a factor of $\sim$7 inward.
+If CO and N$_2$ are water dominated instead of pure ices, their snowlines move inward by $\sim$$70$\%.
+Both of these effects substantially change the disk regions where gas phase C/O and N/O are enhanced over the stellar value, and taken together they may move the CO and N$_2$ snowlines by a factor of $\sim$7 inward.
 In the outer disk, the gaseous C/O and N/O are enhanced by factors of $\sim$$2$ and $\sim$$3$, respectively.
@@
 
-Carbon and oxygen bearing molecules, such as H$_2$O, CO$_2$ and CO, as well as the carbon-to-oxygen (C/O) ratio in protoplanetary disks and in giant planet atmospheres have been extensively studied from a theoretical standpoint (\citealt{oberg11}, \citealt{alidib14}, \citealt{madhu14}, \citealt{molliere15}), and the snowlines of volatiles such as H$_2$O and CO have been detected (\citealt{zhang13}, \citealt{qi13}).
+Carbon and oxygen bearing molecules, such as H$_2$O, CO$_2$ and CO, as well as the carbon-to-oxygen (C/O) ratio in protoplanetary disks and in giant planet atmospheres have been extensively studied from a theoretical standpoint (\citealt{oberg11}, \citealt{alidib14}, \citealt{madhu14}, \citealt{molliere15}), and snowlines of volatiles such as H$_2$O and CO have been detected (\citealt{zhang13}, \citealt{qi13}).
 However, disk chemistry is complex, and it involves many species besides carbon and oxygen carriers \citep{henning13}.
@@
 
-Moreover, volatile snowline locations strongly depend on the grain morphology and the ice environment in which they reside.
-Laboratory experiments (\citealt{fayolle16} and reference therein) have shown that volatiles such as CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
+Moreover, snowline locations strongly depend on the grain morphology and the ice environment in which the grains reside.
+Laboratory experiments (\citealt{fayolle16}) have shown that volatiles such as CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
 This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Sections \ref{sec:CH4} and \ref{sec:N}).
@@
 Figure \ref{fig:CNOstatic} shows an example of the H$_2$O, CO$_2$, CO, CH$_4$, NH$_3$ and N$_2$ snowlines in a static disk, with CO and N$_2$ as pure and water dominated ices, in the top and bottom panel, respectively.
-The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as function of the hydrogen abundance.
+The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as a function of the hydrogen abundance.
 We note that we display this Figure here for qualitative purposes; the quantitative choices that we make for volatile abundances are discussed in Sections \ref{sec:CH4} and \ref{sec:N}.
@@
 In order to explain this added chemical and physical complexity, a more thorough theoretical framework is needed.
-In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, in addition to disk dynamical processes.
+In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, and ice morphology, in addition to disk dynamical processes.
 
@@
 We discuss the effect of different abundances of the main carbon and oxygen carriers, grain morphology and dynamical processes on the snowline locations of carbon and oxygen bearing species and the C/O ratio in Section \ref{sec:CH4}.
-In Section \ref{sec:N}, we perform the same analyses for the main nitrogen carriers and the nitrogen-to-oxygen (N/O) ratio.
+In Section \ref{sec:N}, we perform the same analysis for the main nitrogen carriers and the nitrogen-to-oxygen (N/O) ratio.
 We summarize our findings in Section \ref{sec:summary}.
@@
 \centering
-\includegraphics[width=\textwidth]{../../figs/CNO_and_snowlines_2.pdf}
+\includegraphics[width=\textwidth]{CNO_and_snowlines_2.pdf}
 %\vspace{-0.5in}
-\caption{The total carbon, nitrogen and oxygen abundance as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and layered on a water ice substrate (bottom panel).
+\caption{The total carbon, nitrogen and oxygen abundance as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and water dominated ices (bottom panel).
 Relevant volatile snowlines are marked by the vertical dashed lines.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_CH4.pdf}
+\includegraphics[width=0.5\textwidth]{C_O_ratio_CH4.pdf}
 %\vspace{-0.5in}
@@
 As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the environment in which the icy grains reside.
-If CO ice is layered on a water ice substrate, its binding energy will be larger than in the pure ice case (834 K) due to the higher H$_2$O binding energy.
-\citet{fayolle16} find a CO binding energy of 1388 K in the layered ice scenario.
+If CO ice is in a water dominated environment, its binding energy will be larger than in the pure ice case (834 K) due to the higher H$_2$O binding energy.
+\citet{fayolle16} find a CO binding energy of 1388 K in the water dominated ice scenario.
 We use the model of Section \ref{sec:review} to estimate the movement of the CO snowline for different grain morphologies in a viscous disk.
@@
 As calculated in Paper I, drift and gas accretion may move the snowlines inwards by up to 40-60\% compared to a static disk, and specifically by up to $\sim50$ \% in the case of the CO snowline.
-This result is preserved for the updated CO binding energies, both for pure ice and on a water ice substrate.
-However, a layered grain structure moves the CO snowline inwards significantly: for our fiducial disk model, $r_{\rm CO, layered} \approx 8.7$ AU.
-Thus if CO is bound to water ice instead of pure ices, the CO snowline may move inward by up to 70\%.
+This result is preserved for the updated CO binding energies, both for pure and water dominated ices.
+However, a water dominated environment moves the CO snowline inwards significantly: for our fiducial disk model, $r_{\rm CO, water} \approx 8.7$ AU.
+Thus if CO is water dominated instead of pure ice, the CO snowline may move inward by up to 70\%.
 By taking into account both disk dynamics and of CO in a water dominated environment, we find that the CO snowline may move inward by a factor of $\sim$$7$ compared to a static disk with CO as pure ice.
-This large inward movement of the CO snowline implies that C/O ratios of order unity may be reached much closer to the host star if CO is layered on a water ice substrate, and the CO snowline may be inside 10 AU for certain disk parameters.
+This large inward movement of the CO snowline implies that C/O ratios of order unity may be reached much closer to the host star if CO is a water dominated ice, and the CO snowline may be inside 10 AU for certain disk parameters.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/C_O_water_ice.pdf}
+\includegraphics[width=0.5\textwidth]{C_O_water_ice.pdf}
 %\vspace{-0.5in}
@@
 
-\section{The Effect of N$_2$ Ice Morphology and NH$_3$ on the N/O Ratio}
+\section{The Effect of N$_2$ Ice Morphology on the N/O Ratio}
 \label{sec:N}
 
-\subsection{The Role of NH$_3$}
+\subsection{Nitrogen Carriers and the Role of NH$_3$}
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/N_O_ratio.pdf}
+\includegraphics[width=0.5\textwidth]{N_O_ratio.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{../../figs/N_O_water_ice.pdf}
+\includegraphics[width=0.5\textwidth]{N_O_water_ice.pdf}
 %\vspace{-0.5in}
@@
 As in the case for CO, the N$_2$ binding energy is also strongly dependent on the ice environment in which the grain resides.
-If N$_2$ is layered on a water ice substrate, the N$_2$ binding energy is 1266 K \citep{fayolle16}.
+If N$_2$ is a water dominated ice, the N$_2$ binding energy is 1266 K \citep{fayolle16}.
 Figure \ref{fig:NO_ratio} shows the H$_2$O, CO$_2$, CO and N$_2$ snowline locations in a viscous disk for particles with initial size $\sim0.06$ cm $\lesssim s \lesssim$ 7 m, and with the CO and N$_2$ snowlines calculated assuming different grain morphologies as explained above, as well as estimates for the N/O ratio throughout the disk.
@@
 This choice is justified since the presence of some NH$_3$ only moderately changes the N/O ratio (see Figure \ref{fig:Nstatic}), and since we are primarily interested in the N$_2$ snowline locations rather than exact values for the N/O ratio.
-The innermost N$_2$ snowlines in the viscous disk, created by particles with $s \sim 7$ m for our fiducial model, are located at $r_{\rm N_2, pure} \approx 42$ AU for N$_2$ as pure ice and at $r_{\rm N_2, layered} \approx 11$ AU for N$_2$ layered on water ice.
+The innermost N$_2$ snowlines in the viscous disk, created by particles with $s \sim 7$ m for our fiducial model, are located at $r_{\rm N_2, pure} \approx 42$ AU for N$_2$ as pure ice and at $r_{\rm N_2, water} \approx 11$ AU for N$_2$ as water dominated ices.
 Compared to the static snowlines, i.e.
-$r_{\rm N_2, pure, static} \approx 79$ AU and $r_{\rm N_2, layered, static} \approx 25$ AU, drift and gas accretion move the snowlines inward by up to $\sim$50 \%, similar to the case for CO (see Section \ref{sec:CH4}).
+$r_{\rm N_2, pure, static} \approx 79$ AU and $r_{\rm N_2, water, static} \approx 25$ AU, drift and gas accretion move the snowlines inward by up to $\sim$50 \%, similar to the case for CO (see Section \ref{sec:CH4}).
 By comparing our results in the viscous disk, we find that the N$_2$ snowline moves inward by more than 70\% if N$_2$ resides in a water ice environment rather than as pure ice.
-Similarly to the CO case, the N$_2$ snowline may move by a factor of $\sim$$7$ inward if we account for both disk dynamics and ice morphology (water dominated compared to pure), and may be close to or inward of 10 AU for certain disk models.
+Similarly to the CO case, the N$_2$ snowline may move inward by a factor of $\sim$$7$ if we account for both disk dynamics and ice morphology (water dominated compared to pure), and may be close to or inward of 10 AU for certain disk models.
+
+\subsection{Implications for the Solar System}
+
+%Our results for the N/O ratio imply that both gas phase and solid nitrogen abundances are enhanced in the outer disk, and that the regions in which these enhancements occur depend on disk dynamics and particle morphology.
 
@@
 \item Grain composition sensitively affects the CO and N$_2$ snowline locations.
-If CO and N$_2$ are layered on a water-ice substrate rather than existing as pure ices, their snowlines move inward by up to $\sim$70 \%.
+If CO and N$_2$ are water dominated rather than pure ices, their snowlines move inward by up to $\sim$70 \%.
 This effect is separate from that of radial drift and viscous gas accretion, which also cause an inward movement of the CO and N$_2$ snowlines by up to $\sim$50 \%.
 
-\item The locations of the CO and N$_2$ snowlines are uncertain when we take consider both viscous versus static disk, and pure versus water dominated ices.
-The snowlines in a viscous disk with CO in a water environment are by up to a factor of $\sim$7 closer to the host star that in a static disk with CO as pure ice.
+\item The locations of the CO and N$_2$ snowlines are uncertain when we consider both viscous versus static disks, and pure versus water dominated ices.
+The snowlines in a viscous disk with CO or N$_2$ in a water environment are by up to a factor of $\sim$7 closer to the host star that in a static disk with CO or N$_2$ as pure ices.
 

### 2016-02-29 tex/draft2/ms_edited.tex
update

--- 2016-02-22 tex/draft2/ms.tex
@@
 We explore the effects of dynamical processes, molecular compositions and abundances, and the ice morphology of dust grains in disks on the snowline locations of the main C, O and N carriers, and their consequences for the C/N/O ratios in gas and dust throughout the disk.
-We find that radial drift and accretion alone can reduce the snowline radii of the main C, O and N carriers, i.e.
-H$_2$O, CO$_2$, CO and N$_2$, by 40-60\% compared to static disks.
-If CO and N$_2$ are water dominated instead of pure ices, their snowlines move inward by $\sim$$70$\%.
+%We find that radial drift and accretion alone can reduce the snowline radii of the main C, O and N carriers, i.e. H$_2$O, CO$_2$, CO and N$_2$, by 40-60\% compared to static disks. 
+We find that if CO and N$_2$ are water dominated instead of pure ices, their snowlines move inward by $\sim$$70$\%.
 Both of these effects substantially change the disk regions where gas phase C/O and N/O are enhanced over the stellar value, and taken together they may move the CO and N$_2$ snowlines by a factor of $\sim$7 inward.
-In the outer disk, the gaseous C/O and N/O are enhanced by factors of $\sim$$2$ and $\sim$$3$, respectively.
+Between the CO$_2$ and CO snowlines, the gaseous C/O and N/O are enhanced by factors of $\sim$$2$ and $\sim$$3$, respectively.
+The gas-phase N/O ratio is enhanced by many orders of magnitude between the CO and N$_2$ snowlines due to the depletion of oxygen gas in this region.
 Our estimates for the C/N/O ratios are only modestly affected by the presence of some C in the form of CH$_4$ and of some N in the form of NH$_3$.
@@
 %\begin {enumerate}
-The chemical composition of protoplanetary disks is largely dictated by the freeze-out of volatile species, such as carbon, oxygen and nitrogen carriers.
-The snowline locations of volatile molecules are crucial in determining disk chemical abundances in gas and dust, as well as planet compositions.
+The chemical composition of protoplanetary disks is largely dictated by the freeze-out of volatile species.
+The snowline locations of volatile molecules are therefore crucial in determining disk chemical abundances in gas and dust, as well as planet compositions.
 
 Carbon and oxygen bearing molecules, such as H$_2$O, CO$_2$ and CO, as well as the carbon-to-oxygen (C/O) ratio in protoplanetary disks and in giant planet atmospheres have been extensively studied from a theoretical standpoint (\citealt{oberg11}, \citealt{alidib14}, \citealt{madhu14}, \citealt{molliere15}), and snowlines of volatiles such as H$_2$O and CO have been detected (\citealt{zhang13}, \citealt{qi13}).
-However, disk chemistry is complex, and it involves many species besides carbon and oxygen carriers \citep{henning13}.
+However, disk chemistry is complex, and it involves many other molecular compounds \citep{henning13}.
 In particular, several other volatiles have been detected in disks, such as nitrogen bearing species and hydrocarbons (e.g., \citealt{mandell12}), which adds further complexity to the disk composition.
@@
 
-Moreover, snowline locations strongly depend on the grain morphology and the ice environment in which the grains reside.
-Laboratory experiments (\citealt{fayolle16}) have shown that volatiles such as CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
+Moreover, snowline locations strongly depend on the ice grain morphology.
+Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that volatiles such as CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
 This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Sections \ref{sec:CH4} and \ref{sec:N}).
@@
 Figure \ref{fig:CNOstatic} shows an example of the H$_2$O, CO$_2$, CO, CH$_4$, NH$_3$ and N$_2$ snowlines in a static disk, with CO and N$_2$ as pure and water dominated ices, in the top and bottom panel, respectively.
-The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as a function of the hydrogen abundance.
+The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as a function of the hydrogen total abundance.
 We note that we display this Figure here for qualitative purposes; the quantitative choices that we make for volatile abundances are discussed in Sections \ref{sec:CH4} and \ref{sec:N}.
@@
 In order to explain this added chemical and physical complexity, a more thorough theoretical framework is needed.
-In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, and ice morphology, in addition to disk dynamical processes.
+In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, as well as nitrogen-to-oxygen (N/O) and nitrogen-to-carbon (N/C) ratios.
 
@@
 
-\section{Coupled Drift-Desorption Model Review}
+\section{Coupled Drift-Desorption Model}
 \label{sec:review}
@@
 
-We first assume a static disk, which is only irradiated by the central star and does not experience redistribution of solids or radial movement of the nebular gas.
+In this work we consider both a static and a viscous disk.
+The static disk is irradiated by the central star and does not experience redistribution of solids or radial movement of the nebular gas.
 To quantify the effects of radial drift and gas accretion, we use a viscous disk with a spatially and temporally constant mass flux, $\dot{M}$.
 The viscous disk takes into account radial drift, gas accretion onto the central star, as well as accretion heating.
-We prefer this disk model to an irradiated or evolving disk (see Paper I) because it includes all the dynamical and thermal processes we are interested in for the scope of this paper, and therefore it is the most realistic one.
+We focus on this disk model which includes all the dynamical and thermal processes we are interested in for the scope of this paper, and do not further consider the other disk models presented in Paper I.
 
@@
 We choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$, consistent with mass flux observations in disks (e.g., \citealt{andrews10}).
-As acknowledged in Paper I, the mass flux rate $\dot{M}$ and stellar luminosity $L_*$ will vary throughout the disk lifetime (\citealt{kennedy06}, \citealt{chambers09}), in contrast with our simplified model which assumes that both quantities are constant.
+As described in Paper I, the mass flux rate $\dot{M}$ and stellar luminosity $L_*$ will vary throughout the disk lifetime (\citealt{kennedy06}, \citealt{chambers09}), in contrast with our simplified model which assumes that both quantities are constant.
 This effect will be most pronounced in the inner disk ($\lesssim$ few AU), where accretion heating dominates.
@@
 
+The model is described in full in Paper I, here we review and summarize key concepts and results.
 For a range of initial icy grain sizes composed of a single volatile, we showed in Paper I that the timescale on which these particles desorb is comparable to their radial drift time, as well as to the accretion timescale of the nebular gas onto the central star.
@@
 Particles in scenarios (1) and (2) are thus not affected by radial drift or gas accretion, and the snowline locations are those for a static disk.
-In contrast, the grains in case (3) desorb \textit{instantaneously} and \textit{at a fixed particle-size dependent location} in the disk, regardless of their initial position.
+In contrast, the grains in case (3) desorb practically \textit{instantaneously} and \textit{at a fixed particle-size dependent location} in the disk, regardless of their initial position.
 The snowline locations for these particles will thus be fixed for a given initial particle size and disk model.
@@
 Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}.
-Since the abundance of carbon grains is uncertain, we assume that all the carbon that is not in the form of CH$_4$ is found in carbon grains, so that we reproduce the Solar C/O ratio (gas+dust) of 0.54.
+Since the abundance of carbon grains is uncertain, we assume that all the carbon that is not in the form of CH$_4$, CO and CO$_2$ is in carbon grains, so that we reproduce the Solar C/O ratio (gas+dust) of 0.54.
 
@@
 Based on current observations of CH$_4$ abundances, its presence in the disk only modestly affects the C/O ratio.
-Given the larger uncertainties in overall volatile abundances, we can neglect CH$_4$ when estimating the C/O ratio in static disks.
+Given the larger uncertainties in overall volatile abundances, we can neglect CH$_4$ when estimating the C/O ratio in disks.
 %\footnote{This may not be the case, however, in a viscous disk, as we show below.}.
@@
 
-As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the environment in which the icy grains reside.
-If CO ice is in a water dominated environment, its binding energy will be larger than in the pure ice case (834 K) due to the higher H$_2$O binding energy.
+As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the morphology of the icy grains.
+If CO ice is in a water dominated environment, its binding energy will be larger than in the pure ice case (834 K) due to the higher binding energy of CO to H$_2$O than to other CO molecules.
 \citet{fayolle16} find a CO binding energy of 1388 K in the water dominated ice scenario.

### 2016-03-03 tex/draft2/ms_edited.tex
update

@@
 protoplanetary disks, are well understood.
-In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, the ice environment in which these carriers reside, and the presence of snowlines of major volatiles at different distances from the central star.
+In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, their ice binding environment, and the presence of snowlines of major volatiles at different distances from the central star.
 We explore the effects of dynamical processes, molecular compositions and abundances, and the ice morphology of dust grains in disks on the snowline locations of the main C, O and N carriers, and their consequences for the C/N/O ratios in gas and dust throughout the disk.
@@
 We find that if CO and N$_2$ are water dominated instead of pure ices, their snowlines move inward by $\sim$$70$\%.
-Both of these effects substantially change the disk regions where gas phase C/O and N/O are enhanced over the stellar value, and taken together they may move the CO and N$_2$ snowlines by a factor of $\sim$7 inward.
+Both ice morphology and disk dynamics substantially change the disk regions where gas phase C/O and N/O are enhanced over the stellar value, and taken together they may move the CO and N$_2$ snowlines by a factor of $\sim$7 inward.
 Between the CO$_2$ and CO snowlines, the gaseous C/O and N/O are enhanced by factors of $\sim$$2$ and $\sim$$3$, respectively.
 The gas-phase N/O ratio is enhanced by many orders of magnitude between the CO and N$_2$ snowlines due to the depletion of oxygen gas in this region.
-Our estimates for the C/N/O ratios are only modestly affected by the presence of some C in the form of CH$_4$ and of some N in the form of NH$_3$.
+Our estimates for the C/N/O ratios are moderately affected by the presence of some C in the form of CH$_4$ and of some N in the form of NH$_3$.
 
@@
 
- %However, both observations and chemical models (refs) have shown that other volatiles are abundant in disks, for example nitrogen bearing species and hydrocarbons, thus shaping the disk composition.
+Both in Solar system comets and in protoplanetary disks carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{rodgers02}, \citealt{lodders03}, \citealt{pontoppidan06}).
+However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{mumma96}), which may change the C/O ratio in gas and in dust throughout the disk.
+Nitrogen is another abundant volatile in the Solar system and in disks.
+Chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
+Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}).
+Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
+In addition to N$_2$, a fraction of the nitrogen abundances may be also carried by NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).
+Because of the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio compared to its average value in the disk.
+Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
 
@@
 Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that volatiles such as CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
-This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Sections \ref{sec:CH4} and \ref{sec:N}).
+This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{sec:static}).
 
- It follows that there are several important snowlines in disks, determined both by the disk composition and ice morphology.
+%It follows that there are several important snowlines in disks, determined both by the disk composition and ice morphology.
 Figure \ref{fig:CNOstatic} shows an example of the H$_2$O, CO$_2$, CO, CH$_4$, NH$_3$ and N$_2$ snowlines in a static disk, with CO and N$_2$ as pure and water dominated ices, in the top and bottom panel, respectively.
@@
 In Section \ref{sec:review}, we review the drift-desorption model developed in Paper I.
-We discuss the effect of different abundances of the main carbon and oxygen carriers, grain morphology and dynamical processes on the snowline locations of carbon and oxygen bearing species and the C/O ratio in Section \ref{sec:CH4}.
-In Section \ref{sec:N}, we perform the same analysis for the main nitrogen carriers and the nitrogen-to-oxygen (N/O) ratio.
-We summarize our findings in Section \ref{sec:summary}.
+We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain morphology and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{sec:results}.
+We discuss the implications of our results in Section \ref{sec:discussion} and summarize our findings in Section \ref{sec:summary}.
 
@@
 Here, we expand the model of Paper I by making three additions: (1) we add N and CH4 in the static chemistry model, and explore how different abundances of CH4 and of the N main carriers (N2 and NH3) affect the C/O  and N/O ratios, (2) we quantify the effect of radial drift and gas accretion on the N2, CH4 and NH3 snowline locations, and (3) we explore how different binding energies of CO and N2 affect their snowline locations.}
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=\textwidth]{CNO_and_snowlines_2.pdf}
-%\vspace{-0.5in}
-\caption{The total carbon, nitrogen and oxygen abundance as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and water dominated ices (bottom panel).
-Relevant volatile snowlines are marked by the vertical dashed lines.
-The grain abundances are calculated as a function of the observed median CH$_4$ and NH$_3$ abundances in protostellar cores (see Sections \ref{sec:CH4} and \ref{sec:N}).
-The total grain abundance increases with semimajor axis as more and more species freeze out.}
-\label{fig:CNOstatic}
-\end{figure*}
 
@@
 
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=\textwidth]{CNO_and_snowlines_2.pdf}
+%\vspace{-0.5in}
+\caption{The total carbon, nitrogen and oxygen abundance as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and water dominated ices (bottom panel).
+Relevant volatile snowlines are marked by the vertical dashed lines.
+The grain abundances are calculated as a function of the observed median CH$_4$ and NH$_3$ abundances in protostellar cores (see Sections \ref{sec:CH4} and \ref{sec:N}).
+The total grain abundance increases with semimajor axis as more and more species freeze out.}
+\label{fig:CNOstatic}
+\end{figure*}
+
 Following \citet{chiang10} and \citet{birnstiel12}, a particle's radial drift velocity can be approximated as 
@@
 
-\section{The Effect of CO Ice Morphology and CH$_4$ on the C/O Ratio}
-\label{sec:CH4}
+\section{Results}
+\label{sec:results}
 
-\subsection{Role of CH$_4$}
+\subsection{Snowlines in a Static Disk: The Importance of Ice Morphology}
+\label{sec:snowlines}
 
-Both in Solar system comets and in protoplanetary disks, carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{rodgers02}, \citealt{lodders03}, \citealt{pontoppidan06}).
-However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{mumma96}), which may change the C/O ratio in gas and in dust throughout the disk.
-To quantify the magnitude of this effect, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
+As we note in Section \ref{sec:intro}, the disk volatile composition and the ice morphology determine the location of several important snowlines.
+In this work we focus on the primary carbon, oxygen and nitrogen carriers, i.e.
+H$_2$O, CO$_2$, CO, N$_2$, and to a lesser extent, CH$_4$ and NH$_3$.
+To qualify the effect of the presence of some carbon in the form of CH$_4$, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
 We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max).
@@
 
-We determine the location of the H$_2$O, CO$_2$, CO and CH$_4$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}.
-The binding energies of H$_2$O, CO$_2$, CO and CH$_4$ as pure ices are 5800 K, 2000 K, 834 K and 1300 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}).
+We assume that the main nitrogen-bearing species are N$_2$ and NH$_3$, since other volatiles that contain nitrogen have significantly lower abundances in comparison (e.g., \citealt{mumma11}).
+We use the measured total nitrogen abundance in the Solar system, $n_{\rm N}=8 \times 10^{-5} n_{\rm H}$ \citep{lodders03}, where $n_{\rm H}$ is the hydrogen abundance in the disk midplane.
+Similarly to the case of CH$_4$, we explore the parameter space of possible NH$_3$ abundances using data from the Spitzer c2d Legacy ice survey, as follows: (1) no NH$_3$, (2) the median NH$_3$ observed abundance $n_{\rm NH_3-mid}=0.055 \times n_{\rm H_2O}$ \citep{oberg11a}, and (3) the maximum observed NH$_3$ abundance $n_{\rm NH_3-max}=0.1537 \times n_{\rm H_2O}$ \citep{bottinelli10}.
+In each case, the N$_2$ abundance then simply follows as $n_{\rm N_2}=(n_{\rm N}-n_{\rm NH_3})/2$.
+
+We determine the location of the H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}.
+The binding energies of H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ as pure ices are 5800 K, 2000 K, 834 K, 1300 K, 767 K and 2965 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}, \citealt{martin14}).
+For CO and N$_2$ as water dominated ices the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}.
+Figure \ref{fig:CNOstatic} shows the resulting snowline locations, with CO and N$_2$ pure ices (top panel) and water dominated ices (bottom panel).
+The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as a function of the hydrogen total abundance.
+We assume that the CH$_4$ and NH$_3$ abundances are $n_{\rm CH_4-mid}$ and $n_{\rm NH_3-mid}$.
+As expected, the total grain abundance increases with semimajor axis, as more and more species sublimate.
+More importantly, the CO and N$_2$ snowlines move several tens of AU inward if the ices are water dominated rather than pure.
+This changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ.
+
+%It follows that there are several important snowlines in disks, determined both by the disk composition and ice morphology.
+Figure \ref{fig:CNOstatic} shows an example of the H$_2$O, CO$_2$, CO, CH$_4$, NH$_3$ and N$_2$ snowlines in a static disk, with CO and N$_2$ as pure and water dominated ices, in the top and bottom panel, respectively.
+
+\subsection{C/N/O Ratios in Static Disks}
+\label{sec:static}
+
+In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and the extent by which they are affected by the presence of CH$_4$ and NH$_3$.
+Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed extensively in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.
+
+%Both in Solar system comets and in protoplanetary disks, carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{rodgers02}, \citealt{lodders03}, \citealt{pontoppidan06}).
+However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{mumma96}), which may change the C/O ratio in gas and in dust throughout the disk.
+%To quantify the magnitude of this effect, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
+We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max).
+Thus $n_{\rm CH_4-mid}=0.0555 \times n_{\rm H_2O}$ \citep{oberg11a} and $n_{\rm CH_4-max}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}, where $n_{\rm H_2O}$ is the total H$_2$O abundance.
+Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}.
+Since the abundance of carbon grains is uncertain, we assume that all the carbon that is not in the form of CH$_4$, CO and CO$_2$ is in carbon grains, so that we reproduce the Solar C/O ratio (gas+dust) of 0.54.
+
+%We determine the location of the H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}.
+The binding energies of H$_2$O, CO$_2$, CO, CH$_4$ and N$_2$ as pure ices are 5800 K, 2000 K, 834 K, 1300 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}).
 
@@
 
-Figure \ref{fig:COstatic} shows the C/O ratio in gas and dust as a function of semimajor axis in a static disk, for different CH$_4$ abundances as outlined above.
-As in \citet{oberg11} and Paper I, a gaseous C/O ratio of unity can be achieved beyond the CO$_2$ snowline, where oxygen gas is significantly depleted (top panel).
+Figure \ref{fig:COstatic} shows the C/O ratio in gas and dust as a function of semimajor axis in a static disk, for three different CH$_4$ abundances as outlined above: no CH$_4$ (top panel), CH$_4$-mid (middle panel) and CH$_4$-max (bottom panel).
+As in \citet{oberg11} and Paper I, a gaseous C/O ratio of unity can be achieved between the CO$_2$ and CO snowlines, where oxygen gas is significantly depleted.
 The gas-phase C/O ratio may be further enhanced between the CO$_2$ and CH$_4$ snowlines due to the presence of additional carbon gas from CH$_4$.
-In this region, the C/O ratio increases by 3\% for CH$_4$-mid and by 8\% for  CH$_4$-max, as displayed in the middle and bottom panel of Figure \ref{fig:COstatic}.
-Based on current observations of CH$_4$ abundances, its presence in the disk only modestly affects the C/O ratio.
+In this region, the C/O ratio increases by 3\% for CH$_4$-mid and by 8\% for CH$_4$-max, as displayed in the middle and bottom panels of Figure \ref{fig:COstatic}.
+Based on the range of observed CH$_4$ protostellar abundances, its presence in the disk only modestly affects the C/O ratio.
 Given the larger uncertainties in overall volatile abundances, we can neglect CH$_4$ when estimating the C/O ratio in disks.
-%\footnote{This may not be the case, however, in a viscous disk, as we show below.}.
 
-\subsection{CO Ice Environment and Disk Dynamics}
-
-As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the morphology of the icy grains.
-If CO ice is in a water dominated environment, its binding energy will be larger than in the pure ice case (834 K) due to the higher binding energy of CO to H$_2$O than to other CO molecules.
-\citet{fayolle16} find a CO binding energy of 1388 K in the water dominated ice scenario.
-We use the model of Section \ref{sec:review} to estimate the movement of the CO snowline for different grain morphologies in a viscous disk.
-
-Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.06$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above.
-The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.06$ cm do not form true snowlines.
-As calculated in Paper I, drift and gas accretion may move the snowlines inwards by up to 40-60\% compared to a static disk, and specifically by up to $\sim50$ \% in the case of the CO snowline.
-This result is preserved for the updated CO binding energies, both for pure and water dominated ices.
-However, a water dominated environment moves the CO snowline inwards significantly: for our fiducial disk model, $r_{\rm CO, water} \approx 8.7$ AU.
-Thus if CO is water dominated instead of pure ice, the CO snowline may move inward by up to 70\%.
-By taking into account both disk dynamics and of CO in a water dominated environment, we find that the CO snowline may move inward by a factor of $\sim$$7$ compared to a static disk with CO as pure ice.
-This large inward movement of the CO snowline implies that C/O ratios of order unity may be reached much closer to the host star if CO is a water dominated ice, and the CO snowline may be inside 10 AU for certain disk parameters.
-
-%\emgr{Discuss observed abundances for CH4 and the choices that we make (no CH4, median value, maximum value).
-State that desorption energies for H2O, CO2 and CH4 are well constrained experimentally, and that the CO2 and CH4 binding energies are only weakly dependent on whether it's pure CO2/CH4 or combined with H2O, but that is not the case for CO (and N2 as we will show in the next section).
-Present new binding energies for CO as pure ice and mixed with water.
-Show Figure 1 and discuss how different CH4 abundances and binding energies affect snowline locations and C/O ratio: CO-H2O mixture (though I think it's rather CO layered on top of H2O) moves the CO snowline inward by $\sim$40 AU (will calculate percentages too); the maximum reasonable abundance of CH4 changes the C/O ratio by less than 10\%.
-Show Figure 2 and quantify the effect of drift and accretion on the CH4 snowline compared to a static disk.
-While CH4 has only a modest effect on the C/O ratio in a static disk, this effect may be larger in a viscous disk, as the C gas abundances inside the CH4 snowline may be enhanced due to the differential motion of the desorbed ices and overall nebular gas (refer to Paper I).
-In this study, however, we neglect these effects and therefore do not include CH4 in estimating the C/O ratio (as an aside, the figures that include CH4 in the C/O ratio with drift and desorption are quite messy due to snowlines overlapping).
-Show Figure 3 and estimate the difference between CO-H2O and CO pure ice snowlines in the case of drift and accretion, as well as the comparisons for the static disk for the CO snowline.}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{C_O_water_ice.pdf}
-%\vspace{-0.5in}
-\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO residing in a water ice environment (top panel) or as pure CO ice (bottom panel).
-The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
-The C/O ratio in a static disk (black lines) is shown for comparison.
-The arrows show that the C/O ratio in gas will decrease inside the H$_2$O and CO$_2$
-snowlines in the viscous disk, as the relative fluxes of the desorbed icy
-particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details).
-Radial drift and gas accretion move the snowlines inward by 40-60\%.
-The presence of CO in a water ice environment rather than as pure ice moves the CO snowline significantly inward by $\sim$70\%.}
-\label{fig:CO_ratio}
-\end{figure}
-
-\section{The Effect of N$_2$ Ice Morphology on the N/O Ratio}
-\label{sec:N}
-
-\subsection{Nitrogen Carriers and the Role of NH$_3$}
-
-In addition to carbon and oxygen, nitrogen is another abundant volatile in the Solar system and in disks.
+%Nitrogen is another abundant volatile in the Solar system and in disks.
 Chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
@@
 
-Because of the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk
-may be even more enhanced than the C/O ratio compared to its average value in the disk.
+%Because of the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk
+%may be even more enhanced than the C/O ratio compared to its average value in the disk.
 Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
-In this study, we quantify this effect in protoplanetary disks.
+%In this study, we quantify this effect in protoplanetary disks.
 We assume that the main nitrogen-bearing species are N$_2$ and NH$_3$, since other volatiles that contain nitrogen have significantly lower abundances in comparison (e.g., \citealt{mumma11}).
@@
 
-Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk, for our three choices of the NH$_3$ abundance.
-For comparison, the horizontal dashed line shows the average N/O ratio in the disk.
+Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk, for our three choices of the NH$_3$ abundance (see Section \ref{sec:snowlines}): no NH$_3$ (top panel), NH$_3$-mid (middle panel) and NH$_3$-max (bottom panel).
+For comparison, the horizontal dashed lines show the average N/O ratio in the disk.
 As expected, the gaseous N/O ratio generally exhibits an increasing trend towards the outer disk as more oxygen gas is depleted, with small decreases between the NH$_3$ and CO$_2$ snowlines (by 16\% for NH$_3$-mid and by 18\% for NH$_3$-max, respectively) due to NH$_3$ freeze-out.
-More importantly, the gas-phase N/O ratio in the outer disk is enhanced by more than a factor of three compared to its average value.
+While the presence of NH$_3$ only moderately affects our results for the N/O ratio, NH$_3$ is important since otherwise there would be no solid nitrogen inside the N$_2$ snowline, which would be contrary to asteroid and comet observations (\citealt{wyckoff91}, \citealt{mumma11}, \citealt{bergin15}).
+
+The gas-phase N/O ratio is enhanced by a factor of two outside the H$_2$O snowline compared to its average value, by more than a factor of three between the CO$_2$ and CO snowlines, and by orders of magnitude between the CO and N$_2$ snowlines, which can span tens of AU depending on disk parameters and the relative CO and N$_2$ ice binding environment.
 This enhancement is more pronounced than the C/O gas-phase enhancement of a factor of two in the outer disk (see Figure \ref{fig:COstatic}).
-The N/O ratio reaches particularly high values between the CO and N$_2$ snowlines, where all the oxygen is now contained in grains.
+%The N/O ratio reaches particularly high values between the CO and N$_2$ snowlines, where all the oxygen is now contained in grains.
+
+\subsection{C/N/O Ratios in Dynamic Disks}
+\label{sec:dynamic}
+
+%As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the morphology of the icy grains.
+If CO ice is in a water dominated environment, its binding energy will be larger than in the pure ice case (834 K) due to the higher binding energy of CO to H$_2$O than to other CO molecules.
+\citet{fayolle16} find a CO binding energy of 1388 K in the water dominated ice scenario.
+
+Here we use the model of Section \ref{sec:review} to estimate the movement of the CO and N$_2$ snowlines for different grain morphologies in a viscous disk.
+Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.06$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above.
+The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.06$ cm do not form true snowlines.
+If the CO binding environment is known, the CO snowline moves inward by up to $\sim$50 \% compared to a static disk for each case (pure and water dominated ices) due to disk dynamics.
+The full range of potential CO snowlines taking into account both ice morphology and disk dynamics span $\sim$8.7 AU to $\sim$61 AU, which is a factor of $\sim$7 difference.
+This implies that gas phase C/O ratios of order unity may be reached in the giant planet forming zone, and the CO snowline may be inside 10 AU for certain disk parameters.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{C_O_water_ice.pdf}
+%\vspace{-0.5in}
+\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO residing in a water ice environment (top panel) or as pure CO ice (bottom panel).
+The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
+The C/O ratio in a static disk (black lines) is shown for comparison.
+The arrows show that the C/O ratio in gas will decrease inside the H$_2$O and CO$_2$
+snowlines in the viscous disk, as the relative fluxes of the desorbed icy
+particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details).
+Radial drift and gas accretion move the snowlines inward by 40-60\%.
+The presence of CO in a water ice environment rather than as pure ice moves the CO snowline significantly inward by $\sim$70\%.}
+\label{fig:CO_ratio}
+\end{figure}
+
+%As in the case for CO, the N$_2$ binding energy is also strongly dependent on the ice environment in which the grain resides.
+If N$_2$ is a water dominated ice, the N$_2$ binding energy is 1266 K \citep{fayolle16}.
+
+Figure \ref{fig:NO_ratio} shows the H$_2$O, CO$_2$, CO and N$_2$ snowline locations in a viscous disk for the same range of initial particle sizes as in Figure \ref{fig:CO_ratio}, and with the CO and N$_2$ snowlines calculated assuming different grain morphologies as explained above, as well as estimates for the N/O ratio throughout the disk.
+For simplicity, we assume that all nitrogen is bound in N$_2$.
+This choice is justified since the presence of some NH$_3$ only moderately changes the N/O ratio (see Figure \ref{fig:Nstatic}), and since we are primarily interested in the N$_2$ snowline locations rather than exact values for the N/O ratio.
+The innermost N$_2$ snowlines in the viscous disk, created by particles with $s \sim 7$ m for our fiducial model, are located at $r_{\rm N_2, pure} \approx 42$ AU for N$_2$ as pure ice and at $r_{\rm N_2, water} \approx 11$ AU for N$_2$ in water dominated ices.
+Thus for each case (pure versus water dominated ices), the N$_2$ snowline moves inward by up to 50\% due to disk dynamics.
+By taking into account both ice morphology and disk dynamics, the full range of potential N$_2$ snowlines span $\sim$11 to $\sim$79 AU, which is a factor of $\sim$7 difference.
+Similarly to the case for CO, the N$_2$ snowline may be close to or inward of 10 AU for certain disk models.
 
@@
 
-%\emgr{Similar to the previous section, but with more details.
-Discuss that nitrogen is abundant in the solar system and disks and primarily found as N2.
-Due to the high volatility of N2, the gas phase N/O ratio in the outer disk may be even more enhanced than the C/O ratio.
-A fraction of the nitrogen abundance may be also carried by NH3.
-Discuss NH3 observed abundances and the choices that we make (no NH3, median, maximum).
-State that the NH3 desorption energy is only weakly dependent on whether it's pure NH3 or combined with H2O, but that is not the case for N2.
-Present new binding energies for N2 as pure ice and combined with water.
-Show Figure 4 and discuss how different nitrogen abundances and binding energies affect snowline locations and N/O ratio: N2 combined with H2O moves the N2 snowline inward by $\sim$50 AU (will calculate percentages too); the maximum reasonable abundance of NH3 changes the N/O ratio by $\sim$15\%.
-In the outer disk, the N/O ratio is enhanced by a factor of $\sim$4 compared to the solar value, twice as much as the C/O enhancement.
-Show Figure 5 and quantify the effect of drift and accretion on the NH3 snowline compared to a static disk.
-While NH3 does not have a significant effect on the N/O ratio in a static disk, this effect may be larger in a viscous disk, as the N gas abundance inside the NH3 snowline may be enhanced due to the differential motion of the desorbed ices and overall nebular gas (refer to Paper I).
-In this study, however, we neglect these effects and therefore do not include NH3 in estimating the N/O ratio (again, N/O ratio figure with drift is quite messy when including NH3 and does not add any information that is not already shown in Figures 4 and 5).
-Show Figure 6 and estimate the difference between N2-H2O and N2 pure ice snowlines in the case of drift, as well as the comparisons for the static disk for the N2 snowline.
-State that there will be an overabundance of gas-phase N/O between the CO and N2 snowlines, as there is no oxygen gas in this region.}
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_NH3_steady.pdf}
-%%\vspace{-0.5in}
-%\caption{Desorption distance as a function of initial distance for NH3.... Drift and gas accretion move the NH3 snowline inward by x\%.} 
-%\label{fig:NH3}
-%\end{figure}
-
-\subsection{N$_2$ Ice Environment and Disk Dynamics}
-
-As in the case for CO, the N$_2$ binding energy is also strongly dependent on the ice environment in which the grain resides.
-If N$_2$ is a water dominated ice, the N$_2$ binding energy is 1266 K \citep{fayolle16}.
-Figure \ref{fig:NO_ratio} shows the H$_2$O, CO$_2$, CO and N$_2$ snowline locations in a viscous disk for particles with initial size $\sim0.06$ cm $\lesssim s \lesssim$ 7 m, and with the CO and N$_2$ snowlines calculated assuming different grain morphologies as explained above, as well as estimates for the N/O ratio throughout the disk.
-For simplicity, we assume that all nitrogen is bound in N$_2$.
-This choice is justified since the presence of some NH$_3$ only moderately changes the N/O ratio (see Figure \ref{fig:Nstatic}), and since we are primarily interested in the N$_2$ snowline locations rather than exact values for the N/O ratio.
-The innermost N$_2$ snowlines in the viscous disk, created by particles with $s \sim 7$ m for our fiducial model, are located at $r_{\rm N_2, pure} \approx 42$ AU for N$_2$ as pure ice and at $r_{\rm N_2, water} \approx 11$ AU for N$_2$ as water dominated ices.
-Compared to the static snowlines, i.e.
-$r_{\rm N_2, pure, static} \approx 79$ AU and $r_{\rm N_2, water, static} \approx 25$ AU, drift and gas accretion move the snowlines inward by up to $\sim$50 \%, similar to the case for CO (see Section \ref{sec:CH4}).
-By comparing our results in the viscous disk, we find that the N$_2$ snowline moves inward by more than 70\% if N$_2$ resides in a water ice environment rather than as pure ice.
-Similarly to the CO case, the N$_2$ snowline may move inward by a factor of $\sim$$7$ if we account for both disk dynamics and ice morphology (water dominated compared to pure), and may be close to or inward of 10 AU for certain disk models.
-
-\subsection{Implications for the Solar System}
-
 %Our results for the N/O ratio imply that both gas phase and solid nitrogen abundances are enhanced in the outer disk, and that the regions in which these enhancements occur depend on disk dynamics and particle morphology.
 
-%The N$_2$ snowline moves inward by x\% in the pure ice scenario and by x\%....
-
-%\section{Discussion}
-%\label{sec:discussion}
+\section{Discussion}
+\label{sec:discussion}
 

### 2016-03-03 tex/draft2/ms_edited.tex
update

@@
 In order to explain this added chemical and physical complexity, a more thorough theoretical framework is needed.
-In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, as well as nitrogen-to-oxygen (N/O) and nitrogen-to-carbon (N/C) ratios.
+In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, as well as nitrogen-to-oxygen (N/O) ratios.
 
@@
 We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain morphology and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{sec:results}.
-We discuss the implications of our results in Section \ref{sec:discussion} and summarize our findings in Section \ref{sec:summary}.
+We address the implications of our results in Section \ref{sec:discussion} and summarize our findings in Section \ref{sec:summary}.
 
@@
 where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}=T$ is the grain temperature (assumed to be the same as the disk temperature, see Paper I), and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well.
-We discuss our choices for $E_x$ for the different volatile species in Sections \ref{sec:CH4} and \ref{sec:N}.
+We discuss our choices for $E_x$ for the different volatile species in Section \ref{sec:snowlines}.
 
@@
 Relevant volatile snowlines are marked by the vertical dashed lines.
-The grain abundances are calculated as a function of the observed median CH$_4$ and NH$_3$ abundances in protostellar cores (see Sections \ref{sec:CH4} and \ref{sec:N}).
+The grain abundances are calculated as a function of the observed median CH$_4$ and NH$_3$ abundances in protostellar cores.
 The total grain abundance increases with semimajor axis as more and more species freeze out.}
@@
 H$_2$O, CO$_2$, CO, N$_2$, and to a lesser extent, CH$_4$ and NH$_3$.
-To qualify the effect of the presence of some carbon in the form of CH$_4$, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
+To quantify the effect of the presence of some carbon in the form of CH$_4$, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
 We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max).
@@
 The binding energies of H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ as pure ices are 5800 K, 2000 K, 834 K, 1300 K, 767 K and 2965 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}, \citealt{martin14}).
-For CO and N$_2$ as water dominated ices the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}.
+For CO and N$_2$ as water dominated ices, the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}.
 Figure \ref{fig:CNOstatic} shows the resulting snowline locations, with CO and N$_2$ pure ices (top panel) and water dominated ices (bottom panel).
@@
 
-In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and the extent by which they are affected by the presence of CH$_4$ and NH$_3$.
+In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and the extent to which they are affected by the presence of CH$_4$ and NH$_3$.
 Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed extensively in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.
@@
 The horizontal dashed lines represent the average N/O value in the disk.
-The gas-phase N/O ratio is highly enhanced in the outer disk (by more than a factor of three) compared to its average value.
+The gas-phase N/O ratio is highly enhanced by a factor of three between the CO$_2$ and CO snowlines compared to its average value.
 The arrows mark a highly elevated N/O ratio in gas between the CO and N$_2$ snowlines due to the depletion of oxygen gas in this region.
@@
 The gas-phase N/O ratio is enhanced by a factor of two outside the H$_2$O snowline compared to its average value, by more than a factor of three between the CO$_2$ and CO snowlines, and by orders of magnitude between the CO and N$_2$ snowlines, which can span tens of AU depending on disk parameters and the relative CO and N$_2$ ice binding environment.
-This enhancement is more pronounced than the C/O gas-phase enhancement of a factor of two in the outer disk (see Figure \ref{fig:COstatic}).
+This enhancement is more pronounced than the C/O gas phase enhancement of a factor of two in the outer disk (see Figure \ref{fig:COstatic}).
 %The N/O ratio reaches particularly high values between the CO and N$_2$ snowlines, where all the oxygen is now contained in grains.
@@
 %\vspace{-0.5in}
-\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO residing in a water ice environment (top panel) or as pure CO ice (bottom panel).
+\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO as water dominated ices (top panel) or as pure ice (bottom panel).
 The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
@@
 particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details).
-Radial drift and gas accretion move the snowlines inward by 40-60\%.
-The presence of CO in a water ice environment rather than as pure ice moves the CO snowline significantly inward by $\sim$70\%.}
+The presence of CO in a water ice environment rather than as pure ice moves the CO snowline significantly inward by $\sim$70\%.
+Taken together, disk dynamics and ice morphology move the CO snowline inward by a factor of $\sim$7.}
 \label{fig:CO_ratio}
@@
 %\vspace{-0.5in}
-\caption{N/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO and N$_2$ residing in a water ice environment (top panel) or as pure ices (bottom panel).
+\caption{N/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO and N$_2$ as water dominated ices (top panel) or as pure ices (bottom panel).
 The H$_2$O, CO$_2$, CO and N$_2$ snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
@@
 The presence of N$_2$ in a water ice environment rather than as pure ice moves the N$_2$ snowline significantly inward by $\sim$70\%.
-The results of highly enhanced gas-phase N/O ratios in the outer disk compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}
+Taken together, disk dynamics and ice morphology move the CO snowline inward by a factor of $\sim$7.
+The results of highly enhanced gas-phase N/O ratios between the CO$_2$ and CO snowlines compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}
 \label{fig:NO_ratio}
@@
 More TBD.}
+This study shows that the gas-phase N/O ratio in protoplanetary disks is considerably enhanced throughout most of the disk midplane compared to its average value.
+As demonstrated in Figure \ref{fig:enhance}, the gaseous N/O ratio is enhanced by a factor of two beyond the H$_2$O snowline, by more than a factor of three between the CO$_2$ and CO snowlines, and by several orders of magnitude between the CO and N$_2$ snowlines.
+Thus constraining the N/O ratio in a giant planet atmosphere could be used to trace its formation origins.
+Due to disk dynamics and ice morphology, however, the disk regions with highly elevated gas phase N/O ratios are uncertain and may span tens of AU.
+It is thus crucial to build a quantitative framework to explore the range of gaseous N/O ratios across disks with different dynamics and grain morphologies, as such models are currently lacking.
+The theoretical expectations for N/O ratios in disks and giant planets may be further constrained by future JWST observations of exoplanet atmospheric spectra and detections of nitrogen bearing species (e.g., NH$_3$, \citealt{greene16}).
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{CN_norm.pdf}
+%\vspace{-0.5in}
+\caption{Gas phase C/O (blue curve) and N/O (red curve) ratios normalized to the average C/O and N/O ratio in the disk, assuming CO and N$_2$ are pure ices, and there is no CH$_4$ or NH$_3$.
+The dashed vertical lines mark the H$_2$O, CO$_2$, CO and N$_2$ snowlines.
+The arrow indicates that the N/O ratio is enhanced by orders of magnitude compared to its average value between the CO and N$_2$ snowlines.
+The gaseous N/O ratio is enhanced throughout most of the disk, and more enhanced than the C/O ratio.}
+\label{fig:enhance}
+\end{figure}
+
+The N/O ratio enhancement is larger than that of the gas phase C/O ratio throughout most of the disk.
+Thus measurements of an enhanced C/O ratio in an exoplanet atmosphere could be corroborated (disproved) by measurements of enhanced (non-enhanced) N/O ratios.
+Moreover, Figure \ref{fig:enhance} shows that giant planets that have formed in situ between the H$_2$O and CO snowlines are expected to present elevated both C/O and N/O ratios in their atmospheres, whereas planets between the CO and N$_2$ snowlines will have a highly enhanced N/O ratio in their atmospheres, but not C/O.
+
+Our results could also be applied to the Solar system.
+Recent measurements of nitrogen abundance in comets found a N$_2$/CO ratio $\sim 10^{-3}$ \citep{rubin15}.
+A low N$_2$/CO ratio is consistent with comets having formed inside the N$_2$ snowline where solid N$_2$ is depleted.
+It is thus possible, in principle, to use measurements of the N$_2$ abundance in comets to determine where the N$_2$ snowline was located in our Solar system.
+However due to the uncertainty in the calculated location of the N$_2$ snowline (see Section \ref{sec:dynamic}), more detailed modeling is needed.
+
+Finally, detections of snowlines in disks that are reasonably well characterized could help further constrain the dynamics and ice morphology in the disks.
 
@@
 In this paper we explore the role of icy grain morphology and disk dynamics on the snowline locations of major volatile carrier molecules and the C/N/O ratios in protoplanetary disks.
-We enhance the coupled drift-desorption model developed in \citet{piso15b} by adding more carbon- and nitrogen-bearing species into our framework and by considering different environments in which the icy grains reside.
+We enhance the coupled drift-desorption model developed in \citet{piso15b} by adding more carbon- and nitrogen-bearing species into our framework and by considering different binding ice environments.
 Our results can be summarized as follows:
@@
 
-\item Due to the high volatility of N$_2$, the gaseous N/O ratio in the outer disk is enhanced by more than a factor of three compared to its average value.
-This enhancement is more pronounced than in the case of the gas-phase C/O ratio, which is increased by a factor of two compared to the stellar value.
-Moreover, the N/O ratio in gas is expected to be very large between the CO and N$_2$ snowlines due to the complete depletion of oxygen gas in this region
+\item Due to the high volatility of N$_2$, the gaseous N/O ratio outside the H$_2$O snowline is enhanced by a factor of two compared to its average value, by more than a factor of three between the CO$_2$ and CO snowlines, and by many orders of magnitude between the CO and N$_2$ snowlines doe to the complete depletion of oxygen gas in this region.
+This enhancement is more pronounced than in the case of the gas-phase C/O ratio, which is increased by at most a factor of two compared to the stellar value.
+%Moreover, the N/O ratio in gas is expected to be very large between the CO and N$_2$ snowlines due to the complete depletion of oxygen gas in this region
 
-\item The presence of some carbon in the form of CH$_4$ and of some nitrogen in the form of NH$_3$ only modestly affects our results for the C/O and N/O ratios, respectively.
-In both cases, large C/O and N/O ratios in the outer disk are preserved.
+\item The presence of some carbon in the form of CH$_4$ only modestly affects our results for the gas phase C/O ratio, increasing it by up to 8\% between the CO$_2$ and CH$_4$ snowlines.
+The presence of some nitrogen in the form of NH$_3$ moderately affects our results and may decrease the gaseous N/O ratio by up to 18\% between the NH$_3$ and CO$_2$ snowlines.
+In both cases, large gas phase C/O and N/O ratios in the outer disk are preserved.
 
@@
 Our results have direct consequences for the composition of nascent giant planets.
-The considerable inward movement of the CO and N$_2$ snowlines due to the ice grains residing in a water ice environment rather than as pure ices implies than giant planets with high C/O and/or N/O ratios in their atmospheres may form closer in than previously predicted by theoretical models.
+The considerable inward movement of the CO and N$_2$ snowlines due to the ice grains being water dominated rather than pure ices implies than giant planets with high C/O and/or N/O ratios in their atmospheres may form closer in than previously predicted by theoretical models.
 Moreover, our model shows that wide separation gas giants may have an excess of nitrogen in their envelopes, which may be used to trace their origins.

### 2016-03-06 tex/draft2/ms_edited.tex
update

@@
 %We find that radial drift and accretion alone can reduce the snowline radii of the main C, O and N carriers, i.e. H$_2$O, CO$_2$, CO and N$_2$, by 40-60\% compared to static disks. 
-We find that if CO and N$_2$ are water dominated instead of pure ices, their snowlines move inward by $\sim$$70$\%.
-Both ice morphology and disk dynamics substantially change the disk regions where gas phase C/O and N/O are enhanced over the stellar value, and taken together they may move the CO and N$_2$ snowlines by a factor of $\sim$7 inward.
+The gas-phase N/O ratio enhancement in the outer disk (exterior to the H$_2$O snowline) exceeds the C/O ratio enhancement for all reasonable volatile compositions.
+Ice morphology and disk dynamics individually change the snowline location of N$_2$, the main nitrogen carrier, by a factor of 2-3, and when considered together the range of possible N$_2$ snowline locations is $\sim$11-$\sim$79 AU in a standard disk model.
+The CO snowline location is similarly uncertain.
+This snowline location uncertainty limits the utility of C/N/O compositional ratios in planets to trace their formation zone.
+Observations that anchor snowline locations at different stages of planet formation are therefore key.
+
+%Both ice morphology and disk dynamics substantially change the disk regions where gas phase C/O and N/O are enhanced over the stellar value, and taken together they may move the CO and N$_2$ snowlines by a factor of $\sim$7 inward.
 Between the CO$_2$ and CO snowlines, the gaseous C/O and N/O are enhanced by factors of $\sim$$2$ and $\sim$$3$, respectively.
@@
 Carbon and oxygen bearing molecules, such as H$_2$O, CO$_2$ and CO, as well as the carbon-to-oxygen (C/O) ratio in protoplanetary disks and in giant planet atmospheres have been extensively studied from a theoretical standpoint (\citealt{oberg11}, \citealt{alidib14}, \citealt{madhu14}, \citealt{molliere15}), and snowlines of volatiles such as H$_2$O and CO have been detected (\citealt{zhang13}, \citealt{qi13}).
-However, disk chemistry is complex, and it involves many other molecular compounds \citep{henning13}.
-In particular, several other volatiles have been detected in disks, such as nitrogen bearing species and hydrocarbons (e.g., \citealt{mandell12}), which adds further complexity to the disk composition.
+However, disk chemistry involves many other molecular compounds \citep{henning13} including nitrogen bearing species and hydrocarbons (e.g., \citealt{mandell12}), which may affect the compositions of nascent planets.
 
-Both in Solar system comets and in protoplanetary disks carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{rodgers02}, \citealt{lodders03}, \citealt{pontoppidan06}).
-However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{mumma96}), which may change the C/O ratio in gas and in dust throughout the disk.
-Nitrogen is another abundant volatile in the Solar system and in disks.
-Chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
+Both in Solar system comets and in protoplanetary disks carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{lodders03}, \citealt{mumma11}, \citealt{oberg11}, \citealt{boogert15}).
+However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{oberg08}), which may change the C/O ratio in gas and in dust throughout the disk.
+In the case of nitrogen, chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
 Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}).
@@
 Moreover, snowline locations strongly depend on the ice grain morphology.
-Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that volatiles such as CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
+Very volatile species, such as CO and N$_2$, present binding energies, and therefore snowline locations, that are sensitive to the details of the morphology of the icy grain mantles.
+Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
 This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{sec:static}).
@@
 
-In order to explain this added chemical and physical complexity, a more thorough theoretical framework is needed.
+%In order to explain this added chemical and physical complexity, a more thorough theoretical framework is needed.
+
 In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, as well as nitrogen-to-oxygen (N/O) ratios.
@@
 H$_2$O, CO$_2$, CO, N$_2$, and to a lesser extent, CH$_4$ and NH$_3$.
-To quantify the effect of the presence of some carbon in the form of CH$_4$, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
-We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max).
-Thus $n_{\rm CH_4-mid}=0.0555 \times n_{\rm H_2O}$ \citep{oberg11a} and $n_{\rm CH_4-max}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}, where $n_{\rm H_2O}$ is the total H$_2$O abundance.
-Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}.
-Since the abundance of carbon grains is uncertain, we assume that all the carbon that is not in the form of CH$_4$, CO and CO$_2$ is in carbon grains, so that we reproduce the Solar C/O ratio (gas+dust) of 0.54.
-
-We assume that the main nitrogen-bearing species are N$_2$ and NH$_3$, since other volatiles that contain nitrogen have significantly lower abundances in comparison (e.g., \citealt{mumma11}).
-We use the measured total nitrogen abundance in the Solar system, $n_{\rm N}=8 \times 10^{-5} n_{\rm H}$ \citep{lodders03}, where $n_{\rm H}$ is the hydrogen abundance in the disk midplane.
-Similarly to the case of CH$_4$, we explore the parameter space of possible NH$_3$ abundances using data from the Spitzer c2d Legacy ice survey, as follows: (1) no NH$_3$, (2) the median NH$_3$ observed abundance $n_{\rm NH_3-mid}=0.055 \times n_{\rm H_2O}$ \citep{oberg11a}, and (3) the maximum observed NH$_3$ abundance $n_{\rm NH_3-max}=0.1537 \times n_{\rm H_2O}$ \citep{bottinelli10}.
-In each case, the N$_2$ abundance then simply follows as $n_{\rm N_2}=(n_{\rm N}-n_{\rm NH_3})/2$.
+To quantify the effect of the presence of some carbon in the form of CH$_4$ and of some nitrogen in the form of NH$_3$, we use measured median CH$_4$ and NH$_3$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
 
@@
 The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as a function of the hydrogen total abundance.
-We assume that the CH$_4$ and NH$_3$ abundances are $n_{\rm CH_4-mid}$ and $n_{\rm NH_3-mid}$.
 As expected, the total grain abundance increases with semimajor axis, as more and more species sublimate.
+Freeze-out at the CO$_2$ and CO snowlines pulls more heavy elements into the grains than in the case of the H$_2$O snowline.
 More importantly, the CO and N$_2$ snowlines move several tens of AU inward if the ices are water dominated rather than pure.
@@
 In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and the extent to which they are affected by the presence of CH$_4$ and NH$_3$.
-Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed extensively in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.
+Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.
 
-%Both in Solar system comets and in protoplanetary disks, carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{rodgers02}, \citealt{lodders03}, \citealt{pontoppidan06}).
-However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{mumma96}), which may change the C/O ratio in gas and in dust throughout the disk.
-%To quantify the magnitude of this effect, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
 We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max).
@@
 
-%We determine the location of the H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}.
-The binding energies of H$_2$O, CO$_2$, CO, CH$_4$ and N$_2$ as pure ices are 5800 K, 2000 K, 834 K, 1300 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}).
+Figure \ref{fig:COstatic} shows the C/O ratio in gas and dust as a function of semimajor axis in a static disk: no CH$_4$ (top panel), CH$_4$-mid (middle panel) and CH$_4$-max (bottom panel).
+As in \citet{oberg11} and Paper I, a gaseous C/O ratio of unity can be achieved between the CO$_2$ and CO snowlines, where oxygen gas is significantly depleted.
+The gas-phase C/O ratio may be further enhanced between the CO$_2$ and CH$_4$ snowlines due to the presence of additional carbon gas from CH$_4$.
+In this region, the C/O ratio increases by 3\% for CH$_4$-mid and by 8\% for CH$_4$-max, as displayed in the middle and bottom panels of Figure \ref{fig:COstatic}.
+Based on the range of observed CH$_4$ protostellar abundances, its presence in the disk only modestly affects the C/O ratio.
+%Given the larger uncertainties in overall volatile abundances, we can neglect CH$_4$ when estimating the C/O ratio in disks.
 
@@
 
-Figure \ref{fig:COstatic} shows the C/O ratio in gas and dust as a function of semimajor axis in a static disk, for three different CH$_4$ abundances as outlined above: no CH$_4$ (top panel), CH$_4$-mid (middle panel) and CH$_4$-max (bottom panel).
-As in \citet{oberg11} and Paper I, a gaseous C/O ratio of unity can be achieved between the CO$_2$ and CO snowlines, where oxygen gas is significantly depleted.
-The gas-phase C/O ratio may be further enhanced between the CO$_2$ and CH$_4$ snowlines due to the presence of additional carbon gas from CH$_4$.
-In this region, the C/O ratio increases by 3\% for CH$_4$-mid and by 8\% for CH$_4$-max, as displayed in the middle and bottom panels of Figure \ref{fig:COstatic}.
-Based on the range of observed CH$_4$ protostellar abundances, its presence in the disk only modestly affects the C/O ratio.
-Given the larger uncertainties in overall volatile abundances, we can neglect CH$_4$ when estimating the C/O ratio in disks.
+We assume that the main nitrogen-bearing species are N$_2$ and NH$_3$, since other volatiles that contain nitrogen have significantly lower abundances in comparison (e.g., \citealt{mumma11}).
+We use the measured total nitrogen abundance in the Solar system, $n_{\rm N}=8 \times 10^{-5} n_{\rm H}$ \citep{lodders03}, where $n_{\rm H}$ is the hydrogen abundance in the disk midplane.
+Similarly to the case of CH$_4$, we explore the parameter space of possible NH$_3$ abundances using data from the Spitzer c2d Legacy ice survey, as follows: (1) no NH$_3$, (2) the median NH$_3$ observed abundance $n_{\rm NH_3-mid}=0.055 \times n_{\rm H_2O}$ \citep{oberg11a}, and (3) the maximum observed NH$_3$ abundance $n_{\rm NH_3-max}=0.15 \times n_{\rm H_2O}$ \citep{bottinelli10}.
+In each case, the N$_2$ abundance then simply follows as $n_{\rm N_2}=(n_{\rm N}-n_{\rm NH_3})/2$.
+
+%Both in Solar system comets and in protoplanetary disks, carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{rodgers02}, \citealt{lodders03}, \citealt{pontoppidan06}).
+However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{mumma96}), which may change the C/O ratio in gas and in dust throughout the disk.
+%To quantify the magnitude of this effect, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
+We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max).
+Thus $n_{\rm CH_4-mid}=0.0555 \times n_{\rm H_2O}$ \citep{oberg11a} and $n_{\rm CH_4-max}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}, where $n_{\rm H_2O}$ is the total H$_2$O abundance.
+Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}.
+Since the abundance of carbon grains is uncertain, we assume that all the carbon that is not in the form of CH$_4$, CO and CO$_2$ is in carbon grains, so that we reproduce the Solar C/O ratio (gas+dust) of 0.54.
+
+%We determine the location of the H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}.
+The binding energies of H$_2$O, CO$_2$, CO, CH$_4$ and N$_2$ as pure ices are 5800 K, 2000 K, 834 K, 1300 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}).
 
@@
 
-Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk, for our three choices of the NH$_3$ abundance (see Section \ref{sec:snowlines}): no NH$_3$ (top panel), NH$_3$-mid (middle panel) and NH$_3$-max (bottom panel).
+Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk, for our three choices of the NH$_3$ abundance: no NH$_3$ (top panel), NH$_3$-mid (middle panel) and NH$_3$-max (bottom panel).
 For comparison, the horizontal dashed lines show the average N/O ratio in the disk.
-As expected, the gaseous N/O ratio generally exhibits an increasing trend towards the outer disk as more oxygen gas is depleted, with small decreases between the NH$_3$ and CO$_2$ snowlines (by 16\% for NH$_3$-mid and by 18\% for NH$_3$-max, respectively) due to NH$_3$ freeze-out.
-While the presence of NH$_3$ only moderately affects our results for the N/O ratio, NH$_3$ is important since otherwise there would be no solid nitrogen inside the N$_2$ snowline, which would be contrary to asteroid and comet observations (\citealt{wyckoff91}, \citealt{mumma11}, \citealt{bergin15}).
+As expected, the gaseous N/O ratio generally exhibits an increasing trend towards the outer disk as more oxygen gas is depleted, with small decreases between the NH$_3$ and CO$_2$ snowlines (by 6\% for NH$_3$-mid and by 18\% for NH$_3$-max, respectively) due to NH$_3$ freeze-out.
+While the presence of NH$_3$ only moderately affects our results for the N/O ratio, NH$_3$ is important since otherwise the nitrogen content in solid bodies would be more depleted than is observed for comets and asteroids (\citealt{wyckoff91}, \citealt{mumma11}, \citealt{bergin15}).
 
-The gas-phase N/O ratio is enhanced by a factor of two outside the H$_2$O snowline compared to its average value, by more than a factor of three between the CO$_2$ and CO snowlines, and by orders of magnitude between the CO and N$_2$ snowlines, which can span tens of AU depending on disk parameters and the relative CO and N$_2$ ice binding environment.
+The gas-phase N/O ratio is enhanced by a factor of two outside the H$_2$O snowline compared to its average value, by more than a factor of three between the CO$_2$ and CO snowlines, and by orders of magnitude between the CO and N$_2$ snowlines.
+This latter region can span tens of AU depending on disk parameters and the relative CO and N$_2$ ice binding environment.
 This enhancement is more pronounced than the C/O gas phase enhancement of a factor of two in the outer disk (see Figure \ref{fig:COstatic}).
@@
 Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.06$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above.
+We assume there is no carbon in the form of CH$_4$.
 The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.06$ cm do not form true snowlines.
@@
 By taking into account both ice morphology and disk dynamics, the full range of potential N$_2$ snowlines span $\sim$11 to $\sim$79 AU, which is a factor of $\sim$7 difference.
-Similarly to the case for CO, the N$_2$ snowline may be close to or inward of 10 AU for certain disk models.
+Similarly to the case for CO, the N$_2$ snowline may be close to 10 AU for certain disk models.
 
@@
 Thus constraining the N/O ratio in a giant planet atmosphere could be used to trace its formation origins.
-Due to disk dynamics and ice morphology, however, the disk regions with highly elevated gas phase N/O ratios are uncertain and may span tens of AU.
-It is thus crucial to build a quantitative framework to explore the range of gaseous N/O ratios across disks with different dynamics and grain morphologies, as such models are currently lacking.
-The theoretical expectations for N/O ratios in disks and giant planets may be further constrained by future JWST observations of exoplanet atmospheric spectra and detections of nitrogen bearing species (e.g., NH$_3$, \citealt{greene16}).
+Theoretical models of the magnitude and role of N/O (and N/C) ratios in exoplanet atmospheres are therefore key in order to use these ratios as probes for a planet's formation location.
+Measurements of the N/O ratio in planet atmospheres are essential to further constrain the theoretical expectations --- while nitrogen-bearing molecules have not been targeted so far due to lack of instrument sensitivity, such observations and detections are likely in the near future with the advent of JWST (e.g., NH$_3$, \citealt{greene16}).
 
@@
 %\vspace{-0.5in}
-\caption{Gas phase C/O (blue curve) and N/O (red curve) ratios normalized to the average C/O and N/O ratio in the disk, assuming CO and N$_2$ are pure ices, and there is no CH$_4$ or NH$_3$.
+\caption{Gas phase C/O (blue curve) and N/O (red curve) ratios divided by the average C/O and N/O ratio in the disk, assuming CO and N$_2$ are pure ices, and there is no CH$_4$ or NH$_3$.
 The dashed vertical lines mark the H$_2$O, CO$_2$, CO and N$_2$ snowlines.
@@
 
+Due to disk dynamics and ice morphology, the disk regions with highly elevated gas phase N/O and C/O ratios are uncertain and may span tens of AU.
+In particular, one can imagine a scenario where CO is in a water binding environment and N$_2$ is not.
+This could be attributed to the fact that H$_2$O may bind preferentially to CO than N$_2$, since both H$_2$O and CO are polar molecules while N$_2$ is not.
+It is also possible for N$_2$ ices to form later than CO (e.g., \citealt{pagani12}), and thus be deposited on the outer layers of the icy mantles which are typically water poor (e.g., \citealt{garrod11}).
+Detections of CO and N$_2$ snowlines in the same disk could potentially verify this scenario, since CO and N$_2$ desorb at substantially different temperature, and therefore locations, if the former is water dominated and the latter is in a pure ice environment.
+Changes in stellar luminosity (e.g., \citealt{kennedy06}) and gas mass accretion rate (e.g., \citealt{chambers09}) may introduce additional uncertainties in the snowline locations, and thus the C/N/O ratios (see also Paper I).
+The impact of the ice environment on the snowline location is much smaller in the case of CO$_2$ and NH$_3$, as their binding energies and behavior are closer to that of H$_2$O.
+No detailed measurements for the CH$_4$ binding energy in a water environment exist so far, but due to its low desorption temperature a similar behavior to that of CO and N$_2$ would be expected.
+While the presence of some carbon in the form of CH$_4$ only modestly affects our results, CH$_4$ may become important in disks where a large fraction of the CO abundance has been converted into hydrocarbons (e.g., \citealt{du15}).
+%It is thus crucial to build a quantitative framework to explore the range of snowline locations and thus gaseous C/N/O ratios across disks with different dynamics and grain morphologies, as such models are currently lacking.
+%The theoretical expectations for N/O ratios in disks and giant planets may be further constrained by future JWST observations of exoplanet atmospheric spectra and detections of nitrogen bearing species (e.g., NH$_3$, \citealt{greene16}).
+
 Our results could also be applied to the Solar system.
-Recent measurements of nitrogen abundance in comets found a N$_2$/CO ratio $\sim 10^{-3}$ \citep{rubin15}.
+Recent measurements of nitrogen abundance in comet 67P/Churyumov-Gerasimenko found a N$_2$/CO ratio $\sim 10^{-3}$ \citep{rubin15}.
 A low N$_2$/CO ratio is consistent with comets having formed inside the N$_2$ snowline where solid N$_2$ is depleted.
-It is thus possible, in principle, to use measurements of the N$_2$ abundance in comets to determine where the N$_2$ snowline was located in our Solar system.
+Theoretical models suggest that Jupiter-family comets, such as 67P, originate from the Kuiper belt \citep{duncan97} (but see \citealt{rubin15} for alternative formation scenarios for 67P).
+It is thus possible, in principle, to use measurements of the N$_2$ abundance in Jupiter-family comets to determine where the N$_2$ snowline was located in our Solar system.
 However due to the uncertainty in the calculated location of the N$_2$ snowline (see Section \ref{sec:dynamic}), more detailed modeling is needed.

### 2016-03-07 tex/draft2/ms_edited.tex
update

@@
 %\vspace{-0.5in}
-\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO as water dominated ices (top panel) or as pure ice (bottom panel).
+\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO as pure ice (top panel) or as water dominated ices (bottom panel).
 The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
@@
 %\vspace{-0.5in}
-\caption{N/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO and N$_2$ as water dominated ices (top panel) or as pure ices (bottom panel).
-The H$_2$O, CO$_2$, CO and N$_2$ snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
+\caption{N/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO and N$_2$ as pure ices (top panel) or as water dominated ices (bottom panel).
+The H$_2$O, CO$_2$ and N$_2$ snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
+The CO snowline (inner vertical dashed line in both panels) is only shown for the 7 m particles.
 The N/O ratio in a static disk (black lines) is shown for comparison.
-The arrows show that the N/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in the viscous disk, as the relative fluxes of the desorbed icy
+The downward arrows show that the N/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in the viscous disk, as the relative fluxes of the desorbed icy
 particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details).

### 2016-03-07 Dissertate-Harvard-LaTeX/chapters/chapter1.tex
update

%\begin{savequote}[75mm]
%Nulla facilisi. In vel sem. Morbi id urna in diam dignissim feugiat. Proin molestie tortor eu velit. Aliquam erat volutpat. Nullam ultrices, diam tempus vulputate egestas, eros pede varius leo.
%\qauthor{Quoteauthor Lastname}
%\end{savequote}
%
\chapter{The title of chapter one}
%
%\newthought{There's something to be said} for having a good opening line. Morbi commodo, ipsum sed pharetra gravida, orci  $x = 1/\alpha$ magna rhoncus neque, id pulvinar odio lorem non turpis \cite{Eigen1971, Knuth1968}. Nullam sit amet enim. Suspendisse id velit vitae ligula volutpat condimentum. Aliquam erat volutpat. Sed quis velit. Nulla facilisi. Nulla libero. Vivamus pharetra posuere sapien. Nam consectetuer. Sed aliquam, nunc eget euismod ullamcorper, lectus nunc ullamcorper orci, fermentum bibendum enim nibh eget ipsum. Donec porttitor ligula eu dolor. Maecenas vitae nulla consequat libero cursus venenatis. Nam magna enim, accumsan eu, blandit sed, blandit a, eros.
%$$\zeta = \frac{1039}{\pi}$$
%

\section{Introduction}
\label{intro}

Models of giant planet formation fall in two categories:  core accretion or gravitational instability \citep{dangelo11, youdin13}. In core accretion models, a solid core grows until it becomes massive enough to rapidly accrete gas \citep{PerCam74, mizuno78}.  Gravitational instability (GI) theories investigate the fragmentation of the protoplanetary disk into bound clumps \citep{cameron78, boss97}.

Forming giant planets at wide separations in the protoplanetary disk poses theoretical challenges for both models.  While disks cannot fragment close to their host star, GI is more promising in the outer disk \citep{matzner05, rafikov05}.   However,  for GI to form planetary mass objects versus brown dwarfs, both instantaneous disk conditions and the history of gas infall must be finely tuned \citep{kratter10}.  The main concern for core accretion models, which are successful in the inner disk, is that they operate too slowly in the outer disk. The few Myr lifetime of gas disks sets the constraint \citep{williams11}.  However, mechanisms for the rapid accretion of solids, even at large separations, exist \citep{dones93, kenyon09, lambrechts12}. We thus consider the complementary problem of gas accretion timescales in the outer disk.

Core accretion models fall in several categories: static, quasi-static evolutionary and hydrodynamic.
In static models, the accretion of planetesimals provides a steady luminosity that determines the structure and mass of the atmosphere \citep{stevenson82}.    For a given planetesimal accretion rate, static solutions only exist up to a maximum core mass, referred to as the ``critical core mass",  $\MC$. Above this mass, the atmosphere is assumed to collapse and rapidly accrete disk gas, a process known as the ``core accretion instability."

Many static studies reproduce a canonical $\MC \sim 10 M_\oplus$.  However, \citet[hereafter R06]{Raf11, rafikov06} finds a wide range of values, $0.1 M_\oplus \lesssim \MC \lesssim 100 M_\oplus$, depending on the planetesimal accretion rate and disk parameters. % \citep[see also][]{Raf11}.  
Section \ref{sec:placc} compares our results to R06.  A general limitation of static models is the neglect of heat generated by atmospheric collapse, which can transform the core accretion instability into a slower process of Kelvin-Helmholtz (KH) contraction.

Quasi-static models include time-dependent atmospheric evolution due to KH contraction, and typically include a time-varying planetesimal accretion rate \citep{boden86, alibert05}.   Successful quasi-static models describe three phases of giant planet formation \citep{pollack96}.  In phase 1, rapid planetesimal accretion gives significant core growth, but prevents the accretion of a massive atmosphere.  When planetesimal accretion abates, as the core's feeding zone is depleted of solids, phase 2 begins.  During phase 2, the atmosphere grows gradually, cooling by KH contraction.  An ongoing trickle of planetesimal accretion can heat the atmosphere and slow this contraction.  Eventually, the atmosphere reaches the ``crossover mass", when it equals the mass of the now slowly growing core.  Phase 3, the runaway growth of the atmosphere, begins around the crossover mass.  This runaway occurs quasi-statically, i.e.\ in hydrostatic balance, but very rapidly compared to disk lifetimes.

Dynamical models are needed to understand how runaway growth ends and final planet masses are determined.  Relevant processes -- including gap opening in disks and the transition of accretion from spherical to planar -- can be simulated and then added to atmospheric evolution calculations \citep{LisHub09}.

This work develops quasi-static models in which the core has a fixed mass and is no longer accreting solids.  By isolating the process of KH contraction, we determine the minimum timescale for runaway atmospheric growth, and thus for giant planet formation by core accretion.  Previous studies have considered this type of growth \citep[hereafter I00; PN05, respectively]{ikoma00, pn05}, which is an end-member case of phases 2 and 3 in more detailed quasi-static models.  Especially since phase 2 of atmospheric growth is often the slowest stage of core accretion, understanding the fastest allowed rates of gas accretion is of crucial importance.  Our study develops simplified cooling models for the purposes of broadly exploring parameter space -- especially in the outer disk -- and developing a more detailed understanding of atmospheric accretion.

This paper is organized as follows. Section \S\ref{sec:model} describes our model of atmospheric growth.  In Section \S\ref{sec:coolingan}, we develop a simplified analytic version of our atmospheric model.  We describe the structure and evolution of our atmosphere solutions in Section \S\ref{sec:KH}.   Section \S\ref{sec:critical} gives our final results for growth timescales and critical core masses.  We discuss some neglected effects in Section \S\ref{sec:neglected} and summarize our findings in Section \S\ref{sec:conclusions}.  Some detailed derivations are presented in the appendices.

\section{Atmospheric Accretion Model} \label{sec:model}

To model the accretion of gas by a solid protoplanet, we develop a simplified two-layer model for time-dependent atmospheric growth via cooling, i.e. KH contraction.  With a convective interior and radiative exterior, this model is motivated by similar models of hot Jupiters \citep{ab06, ym10}.

Our model can  treat atmospheric growth up to the early stages of runaway growth, around the crossover mass.  At higher masses, our approximate treatment of the radiative zone (explained below) breaks down.   Since evolution after the onset of runaway growth contributes minimally to the total planet formation timescale, we can model growth times to good accuracy. 
%Our treatment is also inappropriate for hot, short-period planets, where dust sublimation gives deeper radiative zones that require more detailed models.

Our main assumptions are summarized as follows:
\begin{enumerate}
\item The atmosphere is spherically symmetric, remains in hydrostatic balance, and matches onto the disk's midplane temperature and pressure at the planet's Hill radius.
\item The core mass and radius are fixed in evolutionary calculations, neglecting ongoing planetesimal or dust accretion.
%\item At the planet's Hill radius, the atmospheric temperature and pressure match the conditions in the disk midplane.
\item Gravitational contraction of the atmosphere is the only source of planetary luminosity.  
\item Cooling of the atmosphere is dominated by the convective interior.  Thus the luminosity in the radiative exterior is held spatially constant. We justify this assumption in \S\ref{sec:twolayer} and confirm its validity in \S\ref{sec:endoftime}.
%\item A global cooling model connects independent static solutions into a time-dependent sequence.
\item The atmosphere obeys a polytropic equation of state (EOS), with adiabatic index $\gamma = 7/5$ for an ideal diatomic gas.  Corrections from a realistic EOS are discussed in \S\ref{sec:EOS} and deferred to future work.
\item Dust grains provide the opacity in the radiative zone. For the cool temperatures in the outer disk, the radiative layer remains cool enough to avoid dust sublimation. Other opacity choices are discussed in \S\ref{sec:op}.
%\item Because gas accretion accelerates after the crossover mass, the time to reach the crossover mass, i.e. the crossover time, is a good approximation of the total time to form a gas giant.
\end{enumerate}
%The remainder of this section develops our model in more detail.

The first assumption of spherical accretion breaks down early in the inner disk, as explained below.  However, our focus is on the outer disk, where this standard approximation is better justified.

\subsection{Disk and Opacity Model}\label{sec:disk}

We adopt a minimum mass solar nebula (MMSN) model for a passively irradiated disk \citep{chiang10}. With the semimajor axis $a$ normalized to the outer disk as $\aun{10} = a/(10 \text{ AU})$, the gas surface density and mid-plane temperature are  
\begin{subeqnarray} \label{eq:diskparam}
\varSigma\di  &=& 70 \,F_\varSigma \aun{10}^{-3/2} ~{\rm g~cm}^{-2} \\
T\di &=& 45  \,F_T\, \aun{10}^{-3/7} ~{\rm K} \, .
%\Sigma\di&=&2200 F_{\Sigma} a^{-3/2}\,\, \text{g cm}^{-2} \slabel{eq:diska}\\
%T\di &=& 120 F_T a^{-3/7} \, \text{K}, \slabel{eq:diskb}
\end{subeqnarray}
The normalization factors $F_{\Sigma}$ and $F_T$  adjust the model relative to the fiducial MMSN.  We fix $F_{\varSigma}=F_T=1$ unless noted otherwise.  Some observations support a flatter surface density profile, $\varSigma\di \propto a^{-1}$ \citep{andrews10}.  We find that gas density and pressure are weak corrections to atmospheric growth (see Section \S\ref{sec:critical}).  However, a greater surface density of solids in the outer disk could favor the rapid growth of cores \citep{bromley11}.

For a vertically isothermal disk in hydrostatic balance (with no self-gravity), the mid-plane pressure of disk gas is 
\begin{equation}
\label{eq:Pd}
P\di = 6.9 \times 10^{-3} F_\varSigma \sqrt{F_T} \, \aun{10}^{-45/14}~{\rm dyn~cm^{-2}}
%P\di=1.1 \times 10^{-4} F_{\Sigma} \sqrt{F_T} a^{-45/14} \,\, \text{dyne cm}^{-2} %\times \sqrt{m_\ast}
\end{equation}
for a molecular weight of $\mu=2.35$ proton masses and a Solar mass star.  %Changes to the stellar mass (not considered here) would affect both the dynamic mass and, by heating the disk, $T\di$.
%The fiducial pressure is a paltry 7 nanobars.

The (thermodynamically isothermal) sound speed in the disk is
\begin{equation}
c\di = \sqrt{\Rg T\di} = 0.4 \sqrt{F_T} \aun{10}^{3/14} ~\text{km s}^{-1} \,,
\end{equation}  
in terms of the specific gas constant $\Rg$.  The disk scale height is 
\begin{equation}
H\di = {c\di / \varOmega} = 0.42 \sqrt{F_T}  \, \aun{10}^{9/7} \AU\, ,
\end{equation} 
in terms of the Keplerian frequency $\varOmega = \sqrt{G M_\ast/a^3}$, with $G$ the gravitational constant and $M_\ast$ the stellar (in this work Solar) mass.

We assume a dust opacity characteristic of interstellar grains, following \citet{bell94}:
\begin{equation}
\label{eq:opacitylaw}
\kappa= 2 F_\kappa  \left(\frac{T}{100\; \rm{K}}\right)^{\beta} \; \mathrm{cm^2 ~ g^{-1}},
\end{equation}
with a power-law index $\beta = 2$ and normalization $F_\kappa = 1$ unless noted otherwise. Grain growth tends to lower both $F_\kappa$ and $\beta$, while dust abundance scales with $F_\kappa$.   Section \S\ref{sec:op} discusses dust sublimation and more realistic opacity laws.

\subsection{Length Scales}
\label{sec:scales}

The characteristic length scales for protoplanetary atmospheres are crucial for choosing boundary conditions and for understanding the validity of  spherical symmetry in a gas disk of scale height $H\di$.

The solid core has a radius
\begin{equation}
\label{eq:rc}
R\co \equiv \left(\frac{3 M\co}{4 \pi \rho\co}\right)^{1/3} \approx 10^{-4} \mcn{10}^{1/3} ~\text{AU},
\end{equation}
where the core mass, $M\co$, is normalized to 10 Earth masses as $\mcn{10} \equiv M\co/(10~M_\oplus)$. The core density is held fixed at $\rho\co=3.2$ g cm$^{-3}$, representing a mixture of ice and rocky material \citep{pap99}.  We thus neglect  the detailed EOS of the solid core \citep{fortney07}.

A planet can bind a dense atmosphere if its escape velocity exceeds the sound speed.  This criterion is satisfied inside the Bondi radius
\begin{equation}
\label{eq:RB}
\RB \equiv \frac{G M\pla}{c\di^2} \approx 0.17 \, {\mpn{10}  \, \aun{10}^{3/7} \over F_T} ~\AU,
\end{equation}
where the enclosed planet mass, $M\pla = M\co + M_\mathrm{atm}$, includes the core and any atmosphere within the Bondi radius and $\mpn{10} \equiv M\pla /(10~M_\oplus)$.   The contribution of the atmospheric mass is small in early evolutionary stages.

Stellar tides dominate the planet's gravity beyond the Hill radius
\begin{equation}
\label{eq:RHill}
R_{\rm H} = \left(M\pla \over 3 M_\ast \right)^{1/3}a \approx 0.22 \, {\mpn{10}^{1/3} \, \aun{10} }~\AU ,
\end{equation}
where spherical symmetry and hydrostatic balance break down.  (Here we use the same symbol, $M\pla$, to now mean the mass within $\RH$.)

The relevant length scales of the atmosphere and disk satisfy the relation $\RB H\di^2 = 3 R_{\rm H}^3$.  The length scales are roughly equal at the ``thermal mass'' (e.g., \citealt{menou04})
\begin{equation}\label{eq:Mth}
M_{\rm th} > {c\di^{3} \over G \varOmega} \approx 25 \, {F_T^{3/2} \over \sqrt{m_\ast} } \, \aun{10}^{6/7}~ M_\oplus \, .
\end{equation}

In the low mass regime, $M\pla < M_{\rm th}/\sqrt{3}$, the length scales order as $\RB< \RH<H\di$.  In this regime, many studies assume the atmosphere matches the disk conditions at $\RB$.  We use $\RH$ as the matching radius, i.e.\ outer boundary, in all regimes. We discuss this choice further below.  
%\textbf{This choice reflects the fact that, even outside the Bondi radius, hydrostatic densities exceed the background density (see R06). Nevertheless, since the density contrast is modest outside $\RB$, these two boundary conditions give similar results.  We do not extend our models past $\RH$, where tidal forces dominate. }

 %We are accounting for the fact that, in hydrostatic models, the density at 
 %$\RB$ is higher than the background density (see R06). Nevertheless, since the density change is modest, these two boundary conditions give similar results.

For a finite range of intermediate masses, $M_{\rm th}/\sqrt{3} < M\pla < 3 M_{\rm th}$, the Hill radius is the smallest scale, satisfying both $\RH < \RB$ and $\RH < H\di$.  Spherical symmetry remains a good, if imperfect, approximation because the disk is only weakly vertically stratified on  scales $\lesssim H\di$.

At higher planet masses where $M\pla > 3 M_{\rm th}$ and $H\di < \RH < \RB$, spherical symmetry is no longer a good approximation, due to both the vertical stratification of the disk and gap opening.  See \S\ref{sec:hydro} for further discussion of neglected hydrodynamic effects.

We quote planet masses as the enclosed mass within the smaller of $\RB$ and $\RH$.  Thus when $\RB < \RH$, our computational domain -- which always ends at $\RH$ -- extends further than the location where we define atmosphere mass, within $\RB$.  We emphasize that this mass definition does not affect the evolutionary calculation, which consistently includes the enclosed mass at all radii.   We justify our mass convention as follows.  First, we want to conservatively define planet mass in a way that does not exaggerate atmospheric growth and is more consistent with previous works that use $\RB$ as the outer boundary in this regime (e.g., \citealt{ikoma00}, \citealt{pollack96}).  Second, most of the gas outside $\RB$ is weakly bound, and unlikely to remain attached to the planet if the disk suddenly dissipates.

The choice of $\RH$ as the outer boundary is based on this being the largest radius where spherical hydrostatic balance is approximately (but not exactly, see \S\ref{sec:hydro}) valid.  When $\RB < \RH$, we thus include the fact that the density at $\RB$ slightly exceeds the background disk density (see R06).  In practice, this effect is not very significant.  We would get similar results by choosing the outer boundary at $\RB$ in this regime, as previous studies have done.

 %  This conservative choice in quoting planet masses is usually a minor distinction because (when $\RB < \RH$) the gas between $\RB$ and $\RH$ is weakly compressed.

%\textbf{Finally, the mass at which core growth rapidly decreases due to depletion of planetesimals from the core's feeding zone is called the isolation mass. For our fiducial disk model and for an assumed dust-to-gas ratio of 0.01, the isolation mass is given by}
%
%\begin{equation}
%\label{eq:Misol}
%M_{\rm isol} \sim 2 \pi a \times 2 \RH \times \Sigma_{\rm p} \approx 0.2 a_{10}^{3/4} ~ M_{\oplus},
%\end{equation}
%\textbf{where $\Sigma_p$ is the surface density of solids. Since in our regime the core is no longer accreting solids, we require $M_{\rm isol}$ to always be smaller than the core mass at which runaway gas accretion commences, i.e. the critical core mass.}

\subsection{Structure Equations and Boundary Conditions}
\label{sec:struct}

Our atmosphere calculations use the standard structure equations of mass conservation, hydrostatic balance, thermal gradients, and energy conservation:
\begin{subeqnarray}
\label{eq:struct}
\frac{dm}{dr}&=&4 \pi r^2 \rho\slabel{eq:structb} \\
\frac{dP}{dr}&=&-\frac{G m}{r^2}\rho \slabel{eq:structa} \\
\frac{dT}{dr}&=&\nabla \frac{T}{P}\frac{dP}{dr}\slabel{eq:structc} \\
\frac{dL}{dr}&=&4 \pi r^2 \rho \left(\epsilon - \left. T {\partial S \over \partial t} \right|_m \right)\slabel{eq:structd}, 
\end{subeqnarray}
\noindent where $r$ is the radial coordinate, $L$ is the luminosity, and $P$, $T$, $\rho$  and $S$ are the gas pressure, temperature, density and entropy, respectively.  The enclosed mass  at radius $r$ is $m$. \Eq{eq:structc} simply defines the temperature gradient  $\nabla \equiv d \ln T/d \ln P$.

Radiation diffusion gives a temperature gradient
\begin{equation}
\label{eq:delrad}
\delrad \equiv \frac{3 \kappa P}{64 \pi G m \sigma T^4} L\, ,
\end{equation}
with $\sigma$ the Stefan-Boltzmann constant.  In our models, optically thick diffusion is a good approximation throughout the radiative zones.  In convectively unstable regions, efficient convection gives an isentropic temperature gradient with $\nabla = \delad$, the adiabatic gradient

\begin{equation}
\label{eq:delad}
\delad \equiv \Big(\frac{d \ln T}{d \ln P}\Big)_{\rm{ad}}.
\end{equation}
According to the Schwarzschild criterion, convective instability occurs when $\delrad > \delad$.  Thus $\nabla = \min(\delrad, \delad)$ sets the temperature gradient.

In the energy equation (\ref{eq:structd}), $\epsilon$ represents all local sources of heat input, except for the motion of the atmosphere itself.  From stellar structure, $\epsilon$ may be familiar as a nuclear burning term.  In a protoplanetary atmosphere, dissipative drag on planetesimals contributes to $\epsilon$.  Our models set $\epsilon = 0$, consistent with our neglect of planetesimal accretion.

The energy input from gravitational contraction, $\epsilon_{\rm g} = -T \partial S / \partial t$, is crucial for a cooling model.\footnote{In general, any motion is accounted for by this term.  The partial time derivative is performed on shells of fixed mass.}  The partial time derivative would normally require our radial derivative to be partial as well.  However, our subsequent developments will replace the local energy equation (\ref{eq:structd}) with global energy balance (see section \S\ref{cooling}), reverting the structure equations to time-independent ordinary differential equations (ODEs). %\textbf{}

To solve the equation set (\ref{eq:struct}), an EOS is required for closure. In our study, we adopt an ideal gas law with a polytropic EOS 
\begin{subeqnarray}\label{eq:idealEOS}
P &=& \rho \Rg T \, ,\slabel{eq:idealgas} \\
S &=& \Rg \ln \left(T^{1/\delad} \over P \right) \, ,\slabel{eq:polyEOS}  % P &=&K \rho^{\gamma} \, , 
\end{subeqnarray}
with $S$ a relative entropy.  This work uses $\delad=2/7$, i.e.\  a polytropic index $\gamma \equiv 1/(1 - \delad) = 7/5$, for an ideal diatomic gas.\footnote{\Eq{eq:polyEOS} is equivalent to the standard polytropic relation,  $P =K \rho^{\gamma}$, if the polytropic index $K$ replaces $S$.}    We thus neglect the presence of monatomic Helium in our choice of polytropes, a common practice in idealized studies.  We do, however, account for Helium in our reference value of $\mu = 2.35$ proton masses.

Boundary conditions must be satisfied at both the base and the top of the atmosphere, with $m(R\co) = M\co$, $T(\RH) = T\di$ and $P(\RH) = P\di$.  Our model atmospheres also obey $L(R\co) = 0$. Along with \Eq{eq:structd}, this boundary condition is incorporated in the global cooling model described below.

\subsection{Global Cooling of an Embedded Planet}\label{cooling}

We now consider the global energy balance of a planet embedded in a gas disk.  More generally, our derivation applies to any spherical, hydrostatic object in pressure equilibrium with a background medium.  The total atmospheric energy includes gravitational and internal energies, $E = E_G + U$, with
\begin{subeqnarray}
E_G&=&-\int_{M\co}^M \frac{G m}{r} dm \, , \label{eq:Eg} \\
U&=&\int_{M\co}^M u dm \, .\slabel{eq:U}
\end{subeqnarray}
The specific internal energy is $u = C_V T = \Rg (\delad^{-1} -1) T$ for a polytropic EOS.  For a star or coreless planet, $M\co = 0$.

We start with a standard result, global energy balance for an isolated, i.e.\ not embedded, planet:
\begin{equation}
\label{eq:coolingstar}
L_M = L\co + \Gamma - \dot{E}.
\end{equation}
The surface luminosity, $L_M$, includes the core luminosity $L\co$ from e.g.\ planetesimal accretion or radioactive decay.   The total heat generation $\Gamma$  is the integral of $\epsilon$ over the atmosphere.  The rate of change of atmospheric energy, $\dot{E}$, is a loss term.

For an object with no core luminosity (or no core) and no internal heat sources, the energy equation $L_M = -\dot{E}$ describes KH contraction in its simplest form.  When internal heat sources dominate, $L_M = \Gamma$, e.g.\ for nuclear burning in a main sequence star.

A protoplanetary atmosphere embedded in a gas disk lacks a free surface.  For objects without a free surface (or interior to a free surface), the full energy equation, 
\begin{equation}
\label{eq:coolingglobal}
L_M=L\co+\Gamma-\dot{E}+e_M \dot{M} - P_M \left. \frac{\partial V_M}{\partial t}\right|_M \, ,
\end{equation}
acquires surface terms as derived in  \App{sec:globalderiv}.  The surface can be at any mass level $M$, where the instantaneous radius is $R$.  Surface quantities are labeled by $M$ subscripts (except for $R$).  The energy accreted across the surface is given by the specific energy, $e_M = u_M-G M/R$, and the mass accretion rate of gas, $\dot{M}$.  The work done by the surface is $P_M \partial V_M/ \partial t$, with the partial derivative performed at fixed mass.

For static solutions, which are not the focus of this paper, the surface terms (and also $\dot{E})$ vanish.  Static solutions are valid when imposed heat sources, i.e.\ $L\co$ and $\Gamma$, exceed the atmospheric losses.  Quantitatively, static solutions apply when the KH timescale,
\begin{equation}
\tau_{\rm KH} \sim {|E| \over L_M}, 
\end{equation} 
is shorter than the actual evolutionary timescale.  Thus, $\tau_{\rm KH}$, which our models calculate, gives strong lower limits on the time to form giant planets by core accretion.

\subsection{The Two-Layer Model} \label{sec:twolayer}

To simplify our evolutionary calculations, we develop a two-layer atmospheric model with a convective interior and a radiative exterior.   The existence of this layered structure is well known from previous studies (e.g., R06) and can be readily understood.  Before the protoplanetary atmosphere can cool, it has the entropy of the disk.  As the atmosphere cools, the deep interior remains convective.  Convective interiors are a common feature of low mass cool objects (brown dwarfs and planets) that results from the behavior of $\delrad$ for realistic opacity laws.  However, the entropy of the deep interior decreases as the atmosphere cools.  A region of outwardly increasing entropy, i.e.\ a radiative layer, is required to connect the convective interior to the disk.  A more complicated structure, with radiative windows in the convection zone, is possible as discussed in \S\ref{sec:op}.

In convective regions, the adiabatic structure is independent of luminosity and can be calculated without local energy balance, \Eq{eq:structd}.  Thus, for fully convective objects, a cooling sequence can be established by connecting a series of adiabatic solutions using a global energy equation, $L_M = -\dot{E}$ or \Eq{eq:coolingstar}.  Such methods are commonly used for their computational efficiency and are sometime referred to as ``following the adiabats," since the steady state solutions evolve in order of decreasing entropy \citep{marleau13}.

In the radiative zone,  local energy balance, \Eq{eq:structd}, does affect the atmospheric structure.  We proceed by assuming that the majority of energy is lost from the convective interior, and thus the luminosity can be treated as constant in the outer radiative zone, i.e. the RHS of \Eq{eq:structd} is set to zero.  This assumption greatly simplifies the numerical problem.  Instead of solving time dependent partial differential equations, we can solve for  static solutions to a set of ODEs, Equations (\ref{eq:struct}a -- c), which we connect in a time series as described below. We show in \S\ref{sec:endoftime} that the approximation of constant luminosity in the outer layer holds for our regime of interest.

%{Under our assumptions, the  convective interior fully sets the atmospheric luminosity. We note, however, that the location and extent of the convective region is set by the matching of the upper radiative layer to the nebula.} %is set by the convective interior, and the radiative upper layer   luminosity  emerging from the convective interior   Since most of the luminosity emerges from the inner convective region of the atmosphere,

To obtain a single atmosphere solution (indexed by $i$), we choose a planet mass $M_i$.  At the outer boundary, at $\RH(M_i)$, the temperature and pressure are set to the disk values.  A luminosity value is required to compute $\delrad$ and integrate Equations (\ref{eq:struct}a--c).  The correct value of the luminosity is not known in advance, and is the eigenvalue of the problem.  The boundary conditions can only be satisfied for the correct value of the luminosity eigenvalue, which we find by the shooting method.  Specifically, we alter the luminosity until the integrated value of mass at the core, $m(R\co)$, matches the actual core mass, $M\co$. Physically, the luminosity determines the location of the radiative-convective boundary (RCB), consistent with the structure equations and the Schwarzschild criterion.

To understand time evolution, we construct an array of solutions to Equations (\ref{eq:struct}a -- c) in order of increasing atmospheric mass.  We then use global energy balance, \Eq{eq:coolingglobal}, to ``follow the mass" and place these solutions in a cooling sequence.  To establish the time difference between neighboring solutions, we apply \Eq{eq:coolingglobal} at the RCB.  In principle, energy balance could be evaluated at any level.  Our approximate treatment of the radiative zone makes the RCB  the preferred location.     The elapsed time $\Delta t$  between states $i$ and $i +1$ is given by the finite difference
\begin{equation}
\label{eq:dti}
\Delta t = %\left(
{ -\Delta E + \brak{e} \Delta M - \brak{P}  \Delta V_{\brak{M}} \over \brak{L} }\, ,
\end{equation} 
using \Eq{eq:coolingglobal} with $\Gamma = L\co = 0$.
Brackets indicate an average of, and $\Delta$ indicates a difference between, the two states.  All values are evaluated at the RCB.  Due to the partial derivative in \Eq{eq:coolingglobal}, the volume difference $\Delta V_{<M>}$ is performed at fixed mass, here the average of the masses at the RCB.

\section{Analytic Cooling Model}
\label{sec:coolingan}

This section develops the analytic version of our two-layer model. This analytic model is less accurate than our numerical model, primarily because it neglects self-gravity.  We show (in \S\ref{sec:KH}) that self-gravity becomes important at rather low atmospheric masses, $M_{\rm atm} \gtrsim 0.1 M\co$.  Nevertheless, the analytic model is useful for understanding atmospheric evolution and interpreting the numerical results.

The analytic model also assumes that the upper radiative layer is thick enough that $P\cb \gg P\di$.  This approximation ignores the earliest stages of cooling, which are rapid enough to be of minor importance.  The analytic model also ignores the surface terms in \Eq{eq:coolingglobal}, which we show to be a modest correction in \S\ref{sec:endoftime}.  Finally, while the numerical models set the outer boundary at $\RH$, we simplify the analytic calculations by setting the outer boundary to infinity, effectively neglecting the finite scale-height of the disk.   As shown in \App{iso}, the effect of this approximation is minor for $\RB \lesssim \RH$, the primary case of interest (see also R06).

Of all the approximations, the neglect of self-gravity is by far the most significant, as we have verified by comparison to numerical integrations that only neglect self-gravity.

\subsection{Two Layer Structure}
In order to apply the two-layer cooling model analytically, we require expressions for the atmospheric structure.  Conditions at the RCB are crucial as they set the interior adiabat and the radiative losses from the interior. (Recall that luminosity generation in the radiative zone is neglected.)  We express the temperature and pressure of the RCB, at the radius $R\cb$, as 
\begin{subeqnarray}\label{eq:cb2}
T\cb &=& \chi T\di \slabel{eq:Tcb} \\
P\cb &=& \theta P_{\rm d} e^{R_{\rm B}/R\cb} \, .\slabel{eq:PcbRcb}
\end{subeqnarray}
The leading constants would be unity, $\chi = \theta = 1$, if the radiative zone were replaced by an isothermal layer.  In practice, deviations from unity are modest.    Standard radiative structure calculations (see \App{RCBcorr} for details) give 
\begin{equation}
\label{eq:chi}
\chi \simeq \Big(1-\frac{\delad}{\nabla_{\infty}} \Big)^{-\frac{1}{4-\beta}} \simeq 1.53 \, ,
\end{equation}
for $P\cb \gg P\di$, our regime of interest, and with the radiative temperature gradient at depth, $\nabla_\infty = 1/2$, for our dust opacity.  The radiative zones are thus nearly isothermal, as found by R06.  A numerical integration gives $\theta \simeq 0.556$ for our parameters.   
%According to \Eq{eq:PcbRcb} our pressure criterion also gives $R\cb \ll \RB$, but the inequality is logarithmically weaker.

Given the conditions at the RCB,  the density and temperature profiles along the interior adiabat,
\begin{subeqnarray}
\rho &=& \rho\cb \left[ 1 + {\RB' \over r} - {\RB' \over R\cb}  \right]^{1/(\gamma -1)} \slabel{eq:rhoconv}  \\
T	&=& T\cb \left[ 1 + {\RB' \over r} - {\RB' \over R\cb}  \right] \slabel{eq:Tconv}\, ,
\end{subeqnarray} 
%Note (to self) that the "1+" is neglibible in Rcb << RB limit taken elsewhere. 
follow from hydrostatic balance.  We introduce an effective Bondi radius
\begin{equation}
\RB' \equiv {G M\co \over C_PT\cb} = {\delad \over \chi} \RB
\end{equation} 
to simplify expressions, with $C_P = \Rg / \delad$ the specific heat capacity at constant pressure.

Deep in the adiabatic interior, where $r \ll R\cb \ll \RB'$, the profiles follow simple power laws,
\begin{subeqnarray}\label{eq:deep}
\rho &\simeq& \rho\cb \left[ {\RB' \over r}\right]^{1/(\gamma -1)} \propto r^{-5/2} \slabel{eq:rhoconvdeep}  \\
T	&\simeq& T\cb {\RB' \over r} = {G M\co \over C_P r} \slabel{eq:Tconvdeep}\, .
\end{subeqnarray} 
While the radial density profile depends on the adiabatic index, the $r^{-1}$ temperature scaling is universal.  In self-gravitating models, the temperature gradient,
\begin{equation} \label{eq:TPsg}
{dT \over dr} = - {G m(r) \over C_P r^2}\, ,
\end{equation} 
gives a profile that is flatter than $T \propto r^{-1}$.

Returning to our non-self-gravitating model, the total specific energy at depth,
\begin{equation}\label{eq:ean}
e = e_g + u = -\delad {GM\co \over r} \, ,
\end{equation} 
is simply proportional to the gravitational potential, $e_g = -GM\co/r$.

\subsection{Mass, Energy and Luminosity}
\label{MELan}
The most relevant quantities for global cooling are the integrated energy, luminosity, and atmospheric mass.  In our non-self-gravitating limit, the mass of our nearly isothermal radiative zones is less than the convective interior, as shown in \App{iso}.

The atmospheric mass is thus given by the integration of \Eq{eq:rhoconv}, 
\begin{eqnarray} 
\label{eq:Matman}
M_{\rm atm} %&=& 4 \pi \int_{R\co}^{R\cb} \rho r^2 dr \\  %AY: any interested reader knows this step, right?
&=& {5 \pi^2 \over 4} \rho\cb {\RB'}^{5/2} \sqrt{R\cb}, 
%AY: alternate forms
% \approx 0.54 \rho\cb {\RB'}^{5/2} \sqrt{r\cb} \\
%&=& 0.54{\RB'^{3} \sqrt{\chi} \over  \sqrt{ \ln[P\cb/(\theta P\di)]}} \, .
\end{eqnarray}
in the relevant limit $R\co \ll R\cb \ll \RB'$ and for $\gamma =  7/5$.  Mass is concentrated near the outer regions of the convective zone, a result that holds for $\gamma > 4/3$.

Using \Eq{eq:PcbRcb} to eliminate $R\cb$, the ratio of atmosphere to core mass becomes
\begin{equation} \label{eq:crit}
{M_{\rm atm} \over M\co} = {P\cb \over \xi P_M}\, ,
\end{equation} 
where we define a characteristic pressure and a logarithmic factor:
\begin{subeqnarray} 
P_M &\equiv& {4 \delad^{3/2} \over 5 \pi^2 \sqrt{\chi} } {G M\co^2 \over {\RB'}^4} \,  \label{eq:PM} \\
\xi &\equiv& \sqrt{\ln[ P\cb/(\theta P_{\rm d})]} \slabel{eq:xidef}\, .%P_M \approx {1.85 \over \sqrt{\chi}} {G M\co^2 \over {\RB'}^4}\, .
\end{subeqnarray}

The atmosphere mass increases as radiative losses lower the internal adiabat and increase $P\cb$.  The crossover mass, $M_{\rm atm} = M\co$, is reached when
\begin{equation} \label{eq:Pcbc}
P\cb = \xi P_M ,
\end{equation} 
i.e.\ near the characteristic pressure $P_M$.  The critical value of the order unity factor $\xi$ is found by eliminating $P\cb$ from \Eqs{eq:xidef}{eq:Pcbc}.  This logarithmic factor complicates our analytic description.  Since it remains order unity, we simply hold it fixed in our scalings. %, but not when plotting the results.

The total energy is concentrated towards the core if $|e| \rho  r^3 \propto \rho r^2$ drops with increasing $r$.  This condition requires $\gamma < 3/2$, which our choice of $\gamma = 7/5$ satisfies, but $\gamma = 5/3$ (monatomic gas) would not.

Integration of \Eq{eq:ean} over the mass of the atmosphere thus gives
\begin{subeqnarray} 
E &=& - 4 \pi \nabla_{\rm ad} G M\co \int_{R\co}^{R\cb} \rho r dr \\
&\approx& - 4 \pi P\cb {\RB'}^{1 \over \nabla_{\rm ad}} \left(\gamma-1 \over 3 - 2 \gamma\right)  R\co^{2\gamma-3\over \gamma-1}  \slabel{eq:Ean} \\ 
&\approx& - 8 \pi P\cb {\RB'^{7/2} \over \sqrt{R\co}} \slabel{eq:Eus} \, ,
%&\approx& -0.31 P\cb {(\RB / \chi)^{7/2} \over \sqrt{R\co}}
%&\approx& -0.31 P\cb {\RB'^{7/2} \over \sqrt{R\co}}
\end{subeqnarray} 
with $\gamma < 3/2$ and $\gamma = 7/5$ in \Eqs{eq:Ean}{eq:Eus}, respectively. % and then revert to the standard Bondi radius $\RB = G M\co/(\Rg T\di)$ in terms of the disk temperature.

The emergent luminosity from the RCB, 
\begin{equation} \label{eq:Lcb}
L\cb = {64 \pi G M\cb \sigma T\cb^4 \over 3 \kappa P\cb } \nabla_{\rm ad} \approx L\di {P_{\rm d} \over P\cb}\, , % \left({P\di \over P\cb}\right)%^{1+\alpha}
\end{equation} 
follows from \Eq{eq:delrad} and marginal convective stability, $\delrad = \delad$, where we define 
\begin{equation} 
L\di \equiv {64 \pi G M\cb \sigma T_{\rm d}^4 \over 3 \kappa(T_{\rm d}) P_{\rm d}} \nabla_{\rm ad}\chi^{4-\beta}.
\end{equation} 
The scaling $L\cb \propto 1/P\cb$ shows that luminosity drops as the atmosphere cools and $P\cb$ deepens.  This result relies on the pressure independence of  dust opacities.  For fully non-self-gravitating results, we replace $M\cb$, the mass up to the RCB, with the core mass, but the mass of the convective atmosphere can be included for a slightly higher order estimate.

\subsection{Cooling Times \& Core Masses}
\label{coolingan}

Our analytic cooling model uses $L = -\dot{E}$,  neglecting the surface terms in \Eq{eq:coolingglobal}.\footnote{\App{surfterms} shows that these terms are negligible for a non-self-gravitating model.}  %They do become relevant when self-gravity is important and included, as shown in \S\ref{sec:endoftime}.}  
Applying \Eqs{eq:Eus}{eq:Lcb}, the time it takes to cool the atmosphere until the RCB reaches a given pressure depth, $P\cb$, is 
\begin{subeqnarray} 
t_{\rm  cool} &=&- \int_{P\di}^{P\cb } {d E/d P\cb \over L \cb} dP\cb \\
&\approx& 4 \pi {P\cb^{2} \over P\di} {\RB'^{7/2} \over L\di \sqrt{R\co}} \label{eq:tcool}\, .
%&\approx& {0.16 }{P\cb^{2} \over P\di} {\RB'^{7/2} \over L\di \sqrt{R\co}}
\end{subeqnarray} 
The initial RCB depth is set to $P\di$ as a formality.  The cooling slows as it proceeds with $t_{\rm cool} \propto P\cb^2$.

We expect runaway growth to begin around the crossover mass, $M_{\rm{atm}} = M_{\rm c}$. \Eqs{eq:tcool}{eq:Pcbc} give the time to crossover as
\begin{eqnarray} 
\label{eq:tcoolan}
t_{\rm co} &\approx & 2 \times 10^8 {F_T^{5/2}  F_\kappa \left(\xi \over 3.4\right)^2  \over \mcn{10}^{5/3} \aun{10}^{15 / 14}} \yr \, .
\end{eqnarray} 
We estimate the critical core mass, $\MC$, by equating the crossover time with a typical protoplanetary disk lifetime,
\begin{equation}
t\di = 3 \times 10^6 ~{\rm yr} \,.
\end{equation} 
Setting $t_{\rm co} = t\di$ gives
\begin{equation}\label{eq:Mcrit}
\MC \approx 100 {F_T^{3/2} F_\kappa^{3/5}   \left(\xi \over 2.6 \right)^{6/5} \over \aun{10}^{9 / 14}} \; M_\oplus.
\end{equation}

Both $t_{\rm co}$ and $\MC$ are too large to be interesting, or to be correct based on previous results and the numerical results in this paper.  The main reason for this discrepancy is the neglect of self-gravity.  Section \ref{sec:KH} explores in detail the effects of self-gravity on atmospheric structure, luminosity and evolution.  One might expect self-gravity to be only a modest correction for $M_{\rm atm} \leq M\co$.   This seemingly reasonable expectation can be misleading, an interesting result in itself.

Aside from this insight, the analytic model is useful because, despite the amplitude error, it explains the basic scaling of numerical cooling models.  As is well known \citep{HubBod05}, a lower opacity, here scaling with $F_\kappa$, allows faster cooling and gives a lower critical core mass.  The cooling timescale and critical core mass depend only weakly on the disk pressure, via the logarithmic factor $\xi$.   The almost exponential rise in radiative zone pressure with depth, see \Eq{eq:PcbRcb}, explains why disk pressure has only a weak effect on cooling at the RCB.  %AY: add density explanation?

Lower disk temperatures decrease both $t_{\rm co}$ and $\MC$.  The decline in both quantities with semimajor axis is completely explained by the temperature  profile (ignoring the logarithmic factor $\xi$).  The temperature dependence is a competition between two main effects.   The Bondi radius, which quantifies the strength of gravity, decreases for lower temperatures. On the other hand, the luminosity, $\propto T^{4-\beta}$, is lower at colder temperatures, which slows cooling and opposes the overall effect.  With $t_{\rm co} \propto F_T^{ \beta + 1/2}$, the slope of the dust opacity,  $\beta$, is an important factor in regulating the resulting growth time.

To facilitate comparison with our numerical results, we rescale the analytic results as follows.   Since growth times are too slow in the non-self-gravitating analytic model, we modify the runaway growth criterion to occur at an effective crossover mass $M_{\rm atm} = f M\co$, with $f < 1$.  Specifically, we choose $f = 0.13$, because with this value the analytic model gives the same critical core mass at 10 AU as the numerical model, for our standard choices of other parameters.   This prescription does not mean that runaway growth physically occurs at low atmosphere masses, nor does it cause perfect agreement between the models.  Rather by artificially accelerating growth in the analytic model, we can more easily compare the parameter scalings of the two models.  Mathematically, the modified crossover mass amounts to replacing $\xi \rightarrow f\xi$ in our analysis.  With $f = 0.13$, the revised scalings are
\begin{subeqnarray}
t_{\rm run} & \approx & 3 \times 10^6 {F_T^{5/2}  F_\kappa \left(\xi \over 3.4\right)^2  \over \mcn{10}^{5/3} \aun{10}^{15 / 14}} \yr  \slabel{eq:tcoolanf} \\
\MC & \approx & 8\, {F_T^{3/2} F_\kappa^{3/5}   \left(\xi \over 2.6 \right)^{6/5} \over \aun{10}^{9 / 14}} \; M_\oplus\ \, ,  \slabel{eq:Mcritf}
\end{subeqnarray} 
now using $t_{\rm run}$, defined as the time when $M_{\rm atm} = f M\co$,  instead of $t_{\rm co}$ for the runaway growth timescale.  When plotting these rescaled analytic results, we include the variations in $\xi$ from \Eq{eq:xidef}.

%To facilitate comparison with our numerical results, we rescale the analytic results as a rough correction for the neglect of self-gravity.  The goal of the rescaling is not to get a perfect agreement, but to aid in the interpretation of general trends.  To rescale, we simply assume that runaway occurs at a modified crossover mass, $M_{\rm atm} = f M\co$.
%\textbf{Here $f<1$, as growth is slower in our analytic model due to the neglect of self-gravity. For the purpose of this scaling, we are only interested in atmospheric growth up to the modified crossover mass defined above. We are thus not concerned about evolution to yet higher masses where the analytic approximation becomes inaccurate.} Mathematically, this choice of crossover mass amounts to replacing $\xi \rightarrow f\xi$ (cf. Equations \ref{eq:crit} and \ref{eq:Pcbc}).  We choose $f = 0.13$, which normalizes the analytic and numerical models at 10 AU, for standard parameters.  The revised scalings are
%\begin{subeqnarray}
%t_{\rm run} & \approx & 3 \times 10^6 {F_T^{5/2}  F_\kappa \left(\xi \over 3.4\right)^2  \over \mcn{10}^{5/3} \aun{10}^{15 / 14}} \yr  \slabel{eq:tcoolanf} \\
%\MC & \approx & 8\, {F_T^{3/2} F_\kappa^{3/5}   \left(\xi \over 2.6 \right)^{6/5} \over \aun{10}^{9 / 14}} \; M_\oplus\ \, ,  \slabel{eq:Mcritf}
%\end{subeqnarray} 
%now using $t_{\rm run}$ instead of $t_{\rm co}$ for the runaway growth timescale, which is no longer tied to the crossover mass. \textbf{Here $t_{\rm run}$ is the time at which $M_{\rm atm}=f M\co$.} The numerical models are more accurate and self-consistent because they measure runaway growth directly, instead of relying on any assumed $M_{\rm atm}/M\co$ threshold.  When plotting the analytic scalings, we include the variations in $\xi$ from \Eq{eq:xidef}.

\section{Quasi-Static Kelvin-Helmholtz Contraction}
\label{sec:KH}

We examine the structure and evolution of our model atmospheres, calculated as described in \S\ref{sec:model}.  For comparison, we also show results of the non-self-gravitating analytic model of \S\ref{sec:coolingan}.  Radial structure is presented in \S\ref{sec:profiles}, time evolution is described in \S\ref{sec:timeev}, and the validity of our two-layer cooling model is examined in \S\ref{sec:endoftime}.

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/PTm_profiles_v2_smallM.pdf}
%\vspace{-0.5in}
\caption{Radial profiles of atmospheric pressure, temperature and enclosed mass (core included) for a $5 M_{\oplus}$ core at $60$ AU.   Solid, dotted, dashed and dot-dashed lines correspond to solutions with total mass (core and atmosphere) of $5.01 M_{\oplus}$, $5.10 M_{\oplus}$, $6.00 M_{\oplus}$ and $8.99 M_{\oplus}$, respectively (see text for significance of these masses).  Circles and triangles mark the locations of the Bondi radii and of the radiative-convective boundaries, respectively.  The radial profiles extend from the core to the Hill radius.} %  (See text for a description of evolution to yet higher masses.) 
\label{fig:profiles}
\end{figure}

\subsection{Atmospheric Structure}
\label{sec:profiles}
Figure \ref{fig:profiles} shows radial profiles at different stages of atmospheric growth around a $5 M_{\oplus}$ core at $60$ AU.  Quoted mass values include the core plus atmosphere within the smaller of $\RB$ or $\RH$, which for these cases is $\RB$.  The 8.99 $M_{\oplus}$ solution is the mass 
that satisfies our runaway growth criteria (described in \S\ref{sec:timeev}). % we can reach in our evolutionary sequence, as we explain further in \S\ref{sec:endoftime}.

The lowest mass atmosphere -- which we take as our initial state -- is fully convective and shares the disk's entropy.  In \Fig{fig:profiles} this state is the 5.01 $M_{\oplus}$ solution with no radiative zone.

Cooling and contraction allow the atmosphere to accrete more gas.  In the convective zones, higher mass solutions have lower entropy and higher pressures.  A radiative zone emerges to connect the lower entropy interior to the higher entropy disk.  \Fig{fig:profiles} shows that this radiative zone is already fairly deep in the 5.10 $M_\oplus$ solution.

The atmospheric structure is well approximated by our non-self-gravitating, analytic solutions.  Deep in convective zones, thermal energy is a fixed fraction of the gravitational potential energy, giving $T \propto r^{-1}$ and $P\propto r^{-1/\delad}$  as in \Eq{eq:deep}.  This behavior is seen in \Fig{fig:profiles} for $r\ll R\cb$. Near the core, \Eq{eq:Tconvdeep} gives the core temperature, $T\co= G M\co / (C_{\rm P} R\co)$, unaffected by  the overlying atmosphere mass which does not contribute to the gravitational potential.  Closer to the RCB, self-gravity is no longer negligible, particularly for large envelope masses. Instead, these high mass solutions show a slightly flatter profile in $T$ and also in $P \propto T^{1/\delad}$, as explained by \Eq{eq:TPsg}.

In agreement with \Eq{eq:cb2}, the radiative zones remain nearly isothermal, even for the higher masses.  Consequently, the pressure increases nearly exponentially with depth.

\subsection{Time Evolution}\label{sec:timeev}

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/Lt_profiles_v2.pdf}
%\vspace{-0.5in}
\caption{Evolution of the luminosity and elapsed time during atmospheric growth around a $5 M_{\oplus}$ core at $60$ AU.  The luminosity is initially high, then decreases as the atmosphere grows in mass and the radiative zone becomes optically thicker.  Due to the neglect of self-gravity, the analytic model (\emph{dashed curve}) gives luminosities that are too low and evolution times that are too long.}
\label{fig:Ltplot}
\end{figure}

The cooling model of \S\ref{sec:twolayer} is used to connect solutions with different atmospheric masses into an evolutionary sequence.  \Fig{fig:Ltplot} shows the luminosity evolution and the elapsed time as a function of atmospheric mass for the same parameters as in \Fig{fig:profiles}.

During the early stages of atmospheric growth, the luminosity drops sharply.  This behavior is seen in both the full numerical solutions and the analytic model.  With increasing atmospheric mass, the pressure depth of the RCB increases, along with the optical depth ($\propto \kappa P\cb)$.  Consequently, the radiative luminosity decreases.  This behavior is described in \Eqs{eq:crit}{eq:Lcb}.

At later stages of evolution, the numerical model in \Fig{fig:Ltplot} shows a flat luminosity with increasing mass and also time (not shown).  By contrast, the non-self-gravitating analytic model gives a luminosity that continues to drop as the atmosphere becomes more massive.  To understand this difference, consider the scaling of \Eq{eq:Lcb}, $L\cb \propto M\cb T\cb^4/ (\kappa\cb  P\cb)$, which holds in both cases.   Accounting for the higher enclosed mass in the self-gravitating model gives a somewhat higher luminosity, as desired.  However, the main effect is that \Eq{eq:crit} -- which describes a nearly linear relation between atmospheric mass and RCB pressure --  breaks down for self-gravitating solutions.  This behavior can be seen in the top panel of \Fig{fig:profiles} where the $P\cb$ increases significantly from 5.10 to 6.0 $M_\oplus$, but only increases relatively modestly with further growth to 8.99 $M_\oplus$.   In the higher mass solutions, the relatively low $P\cb$ values (and thus the relatively high luminosities) require an outward shift in $R\cb$, as shown in \Fig{fig:profiles}. This shift does not occur in the analytic solution, where $R\cb$ continually decreases with atmospheric mass (cf. Equations \ref{eq:PcbRcb} and \ref{eq:crit}).

The accelerated growth in the numerical model, as shown in the bottom panel of \Fig{fig:Ltplot}, is also a direct result of the higher cooling luminosities with self-gravity included.  Even when $M_{\rm atm} / M\co$ is only few percent, i.e.\ well before the crossover mass, the effect of self-gravity is  quite evident.  Closer to the star, the effects of self-gravity are not as strong for low atmosphere masses.  Nevertheless, all our models show that self-gravity noticeably accelerates growth for $M_{\rm atm} > 0.1 M\co$.

In \Fig{fig:LtvsMopacity}, our standard dust opacity, \Eq{eq:opacitylaw}, is reduced by factors of 10 and 100.  Lower opacities result in higher luminosities and faster evolution.  Our model thus confirms a well-established result \citep{HubBod05}.  While clearly an important effect, atmospheric dust opacities are difficult to robustly predict. Ablation of infalling solids is a dust source.  Sinks include the sequestration of solids in the core and dust settling through the radiative zone.  Grain growth both reduces dust opacities per unit mass and favors settling.  Our scenario of negligible ongoing particle accretion tends to favor low dust opacities.   To be conservative, however, our reference case considers full Solar abundances. The effect of opacity reduction on the critical core mass is described in \S\ref{sec:critical}.

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/opacity_effect.pdf}
%\vspace{-0.5in}
\caption{The effect of dust abundance on atmospheric evolution.   Reducing dust opacities by factors of 10 and 100 from standard Solar abundances gives higher luminosities and faster atmospheric growth.  Plotted quantities are similar to \Fig{fig:Ltplot}, but for a $5 M_{\oplus}$ core at 10 AU. }  %The actual magnitude of dust opacities is difficult to robustly predict.}
\label{fig:LtvsMopacity}
\end{figure}

\Fig{fig:growthtime} plots the evolution of the atmospheric growth timescale, $M_{\rm atm}/\dot{M}$,  around a $5 M_{\oplus}$ core at several locations in our reference disk model.  This instantaneous growth time shows clearly that the atmosphere spends the bulk of its time growing though intermediate atmospheric masses, $\sim 1 -3$ $M_\oplus$ in this case.  Growth times are short both early -- when the radiative zone is transparent -- and late -- when self-gravity accelerates growth.

The fact that growth times have a well defined maximum is a characteristic of accelerating growth.  Unlike our analytic model, which must assume that runaway growth begins near the crossover mass, our numerical model allows us to measure when runaway accretion starts.  Runaway growth does not begin at a universal value of $M_{\rm atm}/M\co$.  Further from the star, runaway growth begins at smaller $M_{\rm atm}/M\co$, as \Fig{fig:growthtime} shows.  \Fig{fig:tvsM} (described in the next section) shows how the onset of runaway growth depends on core mass.

We quantify the runaway growth timescale, $t_{\rm run}$, as the time when $M_{\rm atm}/\dot{M}$ drops to 10\% of its maximum value.  The choice of 10\% is arbitrary; the precise threshold chosen is relatively unimportant because growth continues to accelerate.

%The onset of runaway growth at high masses is also evident in \Figs{fig:Ltplot}{fig:LtvsMopacity} (bottom panels) as the flattening of the $t$ vs.\ $M_{\rm atm}$ curve.

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/Mt_profile_temp.pdf}
%\vspace{-0.5in}
\caption{Evolution of the atmospheric growth timescale with mass around a $5 M_{\oplus}$ solid core  located at 10, 30 or 60 AU, for standard Solar opacities.  Growth is slowest for $M_{\mathrm{atm}} \sim 1 - 3 M_{\oplus}$, i.e.\ before the crossover mass at $M_{\rm atm} = M\co$.}
\label{fig:growthtime}
\end{figure}

\subsection{Validity of the Two-Layer Cooling Model}
\label{sec:endoftime}

We examine the  validity of our cooling model by comparing our model luminosity to the neglected luminosity, $L_{\rm negl}$,  that a more detailed model would generate in the radiative zone.  We compute $L_{\rm negl}$ from the entropy difference between successive radiative zone solutions.  We then integrate the energy equation, $\p L / \p m = - T \p S/ \p t$, over the average depth of the radiative zone.\footnote{While useful as a diagnostic, the neglected luminosity cannot reliably correct the global cooling model because the effects of $L_{\rm negl}$ on the structure of the radiative zone are still ignored.}

 \Fig{fig:coolingterms} shows that the neglected luminosity is indeed negligible during the early stages of evolution.  However, $L_{\rm negl}$ exceeds the model luminosity, $L$, at high masses, $M_{\rm atm} > 3 M_\oplus$ in this case.  Our cooling model is thus inaccurate at higher masses.  However, the model remains reasonably accurate up to the beginning stages of runaway growth, which is sufficient for our purposes of widely exploring parameter space and exploring trends.

The individual terms in the global cooling model of \Eq{eq:coolingglobal}, evaluated at the RCB, are also plotted in \Fig{fig:coolingterms}.  At low masses, the change in energy, $- \dot{E}$, makes the dominant contribution to luminosity.  As the mass increases, the surface terms become more significant, led by the accretion energy.  However, the surface terms are everywhere smaller than $L_{\rm negl}$.  Thus wherever our model is accurate -- including the crucial early phases of growth -- surface terms are a minor correction.  The neglect of surface terms in the analytic model is thus not a serious omission.

%At even higher masses than shown in \Fig{fig:coolingterms}, our cooling model breaks down.  Mathematically, the breakdown manifests as a negative time step (due to an increase in internal entropy) when advancing to a higher atmospheric mass.  This unphysical result has no deep significance, as the neglected luminosity overwhelms the model luminosity by this point.    Effectively the model is telling us when to ignore it.  Our results are unaffected by the breakdown, which occurs comfortably after $t_{\rm run}$, our threshold for runaway growth.

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/cooling_a10_Mc5_rcb.pdf}
%\vspace{-0.5in}
\caption{Individual terms in the atmospheric cooling model of \Eq{eq:coolingglobal}, for a $5 M_\oplus$ core at 10 AU.  The dashed curve for accretion energy indicates a negative contribution.  All quantities are evaluated at the RCB, except for $L_{\rm negl}$, the extra luminosity that would have been generated in the radiative zone, but is neglected in our model. The neglected luminosity is a small correction to the model luminosity $L$ for $M_{\rm atm} \lesssim 3 M_{\oplus}$.   Since these low masses dominate growth times, our model is roughly accurate.}
\label{fig:coolingterms}
\end{figure}

\section{Results for Giant Planet Formation}
\label{sec:critical}

%We now use our structure and evolution models to calculate the necessary conditions for gas giant formation by core accretion.   
We now use our structure and evolution models to estimate the timescales and minimum core masses for giant planet formation for a range of disk conditions and other model parameters.  Our results for atmosphere growth times -- the time for a core of fixed mass to undergo runaway gas accretion -- are presented in \S\ref{sec:tcross}.  Section \ref{sec:critcore} gives our results for critical core masses, the minimum values that trigger runaway atmospheric growth within a plausible disk lifetime, here 3 Myr.

Our models focus on giant planet formation between 5 and 100 AU, as the outer disk is of particular interest for direct imaging searches.   The growth of atmospheres close to the star is also important, but spherical accretion models (including ours) are less applicable here.  In the inner disk, critical core masses increase, yet lower mass planets start to open gaps and outgrow the disk scale height, see \Eq{eq:Mth}.  These concerns prevent us from applying our model to the inner disk.

\subsection{Runaway Growth Timescale}
\label{sec:tcross}
The time to undergo atmospheric runaway growth, $t_{\rm run}$, sets a minimum timescale for the formation of giant planets.  Due to the accelerating nature of runaway growth, the precise threshold chosen for $t_{\rm run}$ (explained in \S\ref{sec:timeev}) is of minor significance.

\begin{figure}[tb]
%\centering
\hspace{-.1in}
\includegraphics[width=0.5\textwidth]{figures/cumul_coolingtime_vs_Matm_10au_mu235.pdf}
%\vspace{-0.5in}
\caption{Time to grow an atmosphere of mass $M_{\rm{atm}}$ for cores with fixed masses between $5 M_{\oplus}$ and $14 M_{\oplus}$ (as labeled) at $10$ AU in our fiducial disk. Circles mark the runaway growth time, $t_{\rm run}$, which occurs at roughly the crossover mass, $M_{\rm{atm}} = M\co$.  Both the time to reach a fixed atmosphere mass and the runaway growth time are shorter for larger cores. For larger $M\co$, runaway growth commences at higher $M_{\rm atm}/M\co$ values.}
\label{fig:tvsM}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{figures/coolingtime_vs_Mc_10au.pdf}
%\vspace{-0.5in}
\caption{Runaway growth time, $t_{\rm{run}}$, vs.\ core mass at $10$ AU, for two values of the mean molecular weight.  Our numerical model (\emph{solid curves}) is compared to our non-self-gravitating analytic model (\emph{dashed curves}, from Equation \ref{eq:tcoolanf}).  A typical protoplanetary disk life time of $3$ Myr is plotted for comparison. The runaway growth time is larger for a lower mean molecular weight.}
\label{fig:tvsMcomp}
\end{figure}

\subsubsection{Effects of Core Mass}
\label{Mct}

\Fig{fig:tvsM} shows the growth of atmospheric mass with time for several core masses at $10$ AU in our fiducial disk.  Atmospheres grow faster around more massive cores due to stronger gravitational binding.  The endpoint of each curve marks $t_{\rm run}$.   Runaway growth occurs near the crossover mass, when $M_{\rm atm} \sim M\co$, in agreement with previous studies. Lower core masses undergo runaway accretion at fractionally smaller atmosphere masses. %Thus, when compared to the assumption that runaway occurs at a universal $M_{\rm atm}/M\co$ ratio,  effects that lower the critical core mass are amplified by the earlier onset of runaway growth.

\Fig{fig:tvsMcomp} shows how $t_{\rm run}$ varies with core mass, also at 10 AU.  The numerical results are plotted against our non-self-gravitating analytic model, described in \S\ref{sec:coolingan}.  The analytic model reproduces the general decline in $t_{\rm run}$ with core mass.  The numerical model, which includes self-gravity, has a somewhat steeper mass dependence.  A modest correction due to self-gravity is unsurprising, and consistent with the above-mentioned trend in $M_{\rm atm}/M\co$ ratios.  Moreover, in the analytic theory, crucial quantities like $M_{\rm atm}$ and $L$ (both roughly $\propto M\co^3$ near crossover) have non-linear dependence on core mass, offering plenty of opportunity for self-gravitational corrections.

The effect of mean molecular weight is also shown in \Fig{fig:tvsMcomp}.  A lower $\mu$ gives longer growth times, because more cooling is required to compress the atmosphere.  This effect is both well established in core accretion studies \citep{stevenson82} and intuitive since the atmospheric scale height $\propto 1/\mu$ is more extended for lower $\mu$.   Moreover, the Bondi radius decreases as $\RB \propto \mu$, giving a smaller gravitational sphere of influence (and a weaker compression at $\RH$ when that is the more relevant scale).  To see how this affects cooling times, note that the characteristic RCB depth near runaway scales as $P\cb \sim P_{\rm M} \propto \mu^{-4}$ from \Eq{eq:PM}.  Thus $L \propto 1/P\cb \propto \mu^4$ explains the trend of slower cooling for lower $\mu$.
%he larger atmospheric scale height due to a lower $\mu$ also results in a larger pressure depth at the RCB, $P\cb$, and thus a lower $R\cb$, cf. \Eq{eq:PcbRcb}.}

For $\mu = 2.0$, which represents the idealized case of an H$_2$ atmosphere completely devoid of Helium, $t_{\rm run}$ increases by factors of $\sim 2 - 3$.  Thus fairly drastic changes in atmospheric composition are required for $\mu$ to significantly affect core accretion timescales. In principle, changes in the mean molecular weight of the gas also affect the EOS, including $\delad$, but such effects are not considered here (see \S\ref{sec:EOS}).

\begin{figure*}[tb]
\centering
\includegraphics[width=1.\textwidth]{figures/TdPd_effect.pdf}
\vspace{-0.3in}
\caption{Runaway growth time as a function of disk temperature (\emph{left}) and pressure (\emph{right}) around a $M_{\rm c} = 5 M_{\oplus}$ core.   The disk pressure or temperature (\emph{left} or \emph{right}, respectively) are fixed at values for 10 AU in our disk model.   The analytic scalings given by \Eq{eq:tcoolanf} are plotted for comparison, as described in the text.  Gas accretion slows down significantly at higher temperatures, but only speeds up modestly as the disk pressure or density increase.} 
\label{fig:TPeffects}
\end{figure*}

\subsubsection{Effects of Disk Temperature and Pressure}
\label{sec:TPeffects}

\Fig{fig:TPeffects} shows how the runaway accretion time varies with disk temperature, $T\di$, or pressure, $P\di$, holding the other quantity fixed.   The analytic model roughly reproduces the temperature and pressure scalings, again with some discrepancies due mainly to the neglect of self-gravity.  Temperature variations are much more significant than pressure variations (note the difference in logarithmic and linear axes).  Since midplane disk conditions depend only on temperature and pressure in our model,\footnote{See \Eq{eq:diskparam}.  When gap opening is considered in models of later growth stages, the orbital frequency and effective viscosity become relevant as well.} the dominant effect of disk location is temperature.

The decline in growth times with lower temperatures arises from a balance of competing effects.  The cooling luminosity is inherently smaller at lower temperatures.  Overpowering this effect, the larger Bondi radius and lower dust opacity act to accelerate growth at lower temperatures.

Growth times depend only weakly on, but do fall slightly with, pressure.   This result may be surprising, given that the disk is the source of atmospheric mass and the atmosphere must match onto the disk's density and pressure.  The nearly exponential increase in pressure with depth through the radiative zone explains this effect.  Cooling is largely regulated at the RCB, and a modest change in RCB depth compensates for large variations in disk pressure.

Section \ref{sec:coolingan} shows how these temperature and pressure effects arise in our analytic model.

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{figures/Mcrit_vs_a_3Myrs_new.pdf}
%\vspace{-0.5in}
\caption{The critical core mass as a function of semimajor axis, for a disk lifetime of $3$ Myrs and two values of the mean molecular weight ($\mu = 2.35$ is for Solar abundances). The decline in $\MC$ with distance is a robust result for standard disk models.  The analytic model, which neglects self-gravity, over-predicts the steepness of the decline.}
\label{fig:Mcvsa}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{figures/Mcrit_vs_a_3Myrs_opacity.pdf}
%\vspace{-0.5in}
\caption{Critical core masses vs. distance for standard and reduced (by a factor of ten) dust opacities.  Lower opacities give significantly lower $\MC$ values.  Atmospheric opacity remains a large uncertainty in core accretion models.}
\label{fig:Mcritopacity}
\end{figure}

\subsection{Critical Core Mass}
\label{sec:critcore}

The critical core mass declines with distance from the star, as shown in \Fig{fig:Mcvsa} for our standard disk model.   The main reason for the decline, as explained above, is that atmospheres grow faster at lower temperatures.  The lower densities and pressures in the outer disk have a much smaller effect.  Since the vast majority of disk models have a disk temperature that declines with $a$, our qualitative result is robust.   We also show that higher $\mu$ values give lower values of $\MC$.

The average power-law decline from 1 to 100 AU (not plotted) is $\MC \propto a^{-0.3}$, for both choices of $\mu$.   While not a drastic decline, the ability of distant low mass cores to accrete gas efficiently is significant for the interpretation of direct imaging surveys.  However, our model offers no guarantee of copious giant planets at large distances.  Many histories of solid accretion are possible, and solid cores may grow too slowly to allow the rapid gas accretion that we model.

Our non-self-gravitating analytic model (dashed curves in \Fig{fig:Mcvsa}) over-predicts the steepness of the decline in $\MC$ with $a$.  This discrepancy is not surprising, as we have shown that self-gravity strongly affects evolution, even before runaway growth begins and the crossover mass is reached.

\Fig{fig:Mcritopacity} shows that reducing the opacity by an order of magnitude significantly reduces $\MC$.  Furthermore, this opacity effect is stronger at larger distances.  The opacity reduction by a factor of 10 lowers $\MC$ at 5 AU by a factor of $\sim$$2.5$ and at 100 AU by a factor of $\sim$$3.5$.  Atmospheric opacity is a dominant uncertainty in core accretion modeling.  However, unless atmospheric opacity varies significantly with disk radius, the general decline in $\MC$ with increasing $a$ should hold.  For our scenario of negligible ongoing planetesimal accretion, it is tempting to think that dust could settle out of the radiative zone, lowering the opacity \citep{podolak03}.  Nonetheless, since opacity near the RCB is a crucial factor, we speculate that convective overshoot may prevent very low opacities.

A larger disk lifetime further reduces $M_{\rm crit}$. Recent studies have shown that gas disks may live up to $\propto$$10-12$ Myr \citep{bell13}. As $M_{\rm crit} \sim t\di^{-3/5}$ (cf. Equation \ref{eq:tcoolan}), a disk lifetime of 10 Myrs decreases the critical core mass by a factor of two.  Of course, some fraction of the disk lifetime must be allocated to the growth and possibly migration of the core.

%\textbf{As the core is assumed to form through accretion of solids, we require that $M_{\rm crit}>M_{\rm isol}$. Since $M_{\rm crit}$ decreases with semi major axis, while $M_{\rm isol}$ increases with $a$ (cf. Equation \ref{eq:Misol}), it is possible that the isolation mass may become larger than the critical core mass at large orbital distances. From \Eq{eq:Misol}, $M_{\rm isol} \approx 1 M_{\oplus}$ at 100 AU, which is smaller than $M_{\rm crit}$ under standard opacity assumptions, but comparable for a reduced opacity (see Figure \ref{fig:Mcritopacity}). A flatter surface density profile (see \S\ref{sec:disk}) may further increase $M_{\rm isol}$. We note, however, that the estimate of \Eq{eq:Misol} assumes an initial uniform dust-to-gas ratio throughout the disk, although the dust-to-gas ratio may be distance dependent. A lower dust-to-gas ratio in the outer disk reduces $M_{\rm isol}$. As stated in \S\ref{sec:scales}, we assume that the isolation mass is always lower than our calculated $M_{\rm crit}$.}

%\textbf{Finally, the mass at which core growth rapidly decreases due to depletion of planetesimals from the core's feeding zone is called the isolation mass. For our fiducial disk model and for an assumed dust-to-gas ratio of 0.01, the isolation mass is given by}
%
%\begin{equation}
%\label{eq:Misol}
%M_{\rm isol} \sim 2 \pi a \times 2 \RH \times \Sigma_{\rm p} \approx 0.2 a_{10}^{3/4} ~ M_{\oplus},
%\end{equation}
%\textbf{where $\Sigma_p$ is the surface density of solids. Since in our regime the core is no longer accreting solids, we require $M_{\rm isol}$ to always be smaller than the core mass at which runaway gas accretion commences, i.e. the critical core mass.}

\subsection{Comparison with Previous Studies}
\label{sec:comp}

We can directly compare our results with other studies of protoplanetary atmospheric growth in the absence of planetesimal accretion, notably those in I00 and PN05.  A major distinguishing feature of our study is that we explore a range of disk conditions, while I00 and PN05 focus on growth at the location of Jupiter, 5.2 AU, in their disk models.  Moreover, both I00 and PN05 solve the full time dependent energy equation and consider more detailed EOSs, in addition to other detailed differences in model parameters.  Despite these differences, the qualitative and quantitative agreement with our study is good.

I00 give an approximate analytic fit to their models' envelope formation time (the equivalent of the runaway accretion time $t_{\rm run}$ in our models), at $a=5.2$ AU, 
\begin{equation}
\label{eq:tauenv}
\tau_{\rm env} \sim 3 \times 10^8 \Big(\frac{M\co}{M_{\oplus}}\Big)^{-2.5} \Big(\frac{\kappa}{1 \text{cm$^2$ s$^{-1}$}}\Big) \,\,\, \text{yr}\, .
\end{equation}
Our model reproduces the linear opacity dependence (see \S\ref{sec:timeev}), even though we use a temperature dependent opacity, \Eq{eq:opacitylaw}, instead of a constant $\kappa$.  Our growth times at 5 AU,
\begin{equation}
\label{eq:trun5au}
t_{\rm run} \sim 5 \times 10^8 \Big(\frac{M\co}{M_{\oplus}}\Big)^{-2.4} \,\,\, \text{yr},
\end{equation} 
agree well with the I00 results, both in magnitude and scaling with core mass.

One goal of our study was to explain the extended luminosity minimum in evolutionary models that characterizes phase 2 in core accretion models, as described in the introduction.  Previous work (see Figure 2 in I00 and Figures 3 and 4 in PN05) shows that this luminosity minimum is an intrinsic feature of atmospheric cooling even in the absence of planetesimal accretion.  We not only reproduce this effect in our simplified model (see \Figs{fig:Ltplot}{fig:LtvsMopacity}), but we also show that self-gravity is the essential ingredient to produce a broad luminosity minimum well before the crossover mass.

Time-dependent studies that incorporate planetesimal accretion generally find larger formation timescales and critical core masses than those of this work.  This result is expected since additional energy from planetesimal accretion limits the ability of the atmosphere to cool. \citet{pollack96} find an evolutionary time and crossover mass of $\sim$$7.5$ Myrs and $\sim$$16 M_{\oplus}$, respectively, for a MMSN disk model at 5.2 AU and interstellar grain opacity. These values, however, decrease to $\sim$$3$ Myrs and $\sim$$12 M_{\oplus}$ if planetesimal accretion is entirely shut off during the gas accretion phase. The core accretion model of \citet{HubBod05} for Jupiter's formation predicts an evolutionary time of 3.3 Myrs for a $10 M_{\oplus}$ core, for an interstellar opacity, which is also consistent with our result at 5 AU.

A direct comparison with studies that only include planetesimal accretion, and neglect atmospheric cooling, is difficult.  Note, however, that I00 establishes the correspondence between the minimum luminosity in evolutionary models and the minimum planetesimal accretion rate needed in static models, which gives rise to the classical critical core mass of static models \citep{mizuno78, stevenson82}.

In terms of modern static studies, our results complement (and borrow some tools from) R06 and \citet{Raf11}.  These studies consider the disk radius dependence of core accretion for protoplanets that continuously accrete planetesimals.  We show that the limits on core accretion at large radial distance claimed by \cite{Raf11} (and references therein) can be overcome if planetesimal accretion shuts off and the atmosphere is allowed to cool.  Nevertheless, the plausibility of rapid core growth followed by negligible subsequent solid accretion admittedly remains uncertain, as discussed in more detail in \S\ref{sec:conclusions}.

\section{Neglected Effects}\label{sec:neglected}
Since one goal of this paper was to obtain a detailed understanding of atmospheric evolution with a simple model, we have necessarily ignored   effects of potential significance.  We briefly address the most important of these and note that a followup work, Piso, Youdin \& Murray-Clay (in prep., hereafter Paper II), will extend our current models to address some of these effects.

\subsection{Hydrodynamic Effects}\label{sec:hydro}
The neglect of hydrodynamical effects in our model is best discussed in terms of the thermal mass, $M_{\rm th}$, and the length scales introduced in \S\ref{sec:scales}.  In the low mass regime, $M\pla < M_{\rm th}/ \sqrt{3}$, where $\RB < \RH$, we assume that hydrostatic balance holds out to the outer boundary at $\RH$.   In this low mass regime, \citet{Orm13} calculated the 2D (radial and azimuthal) flow patterns driven by stellar tides and disk headwinds.     On scales $\gtrsim \RB$ the flows no longer circulate the planet: they belong to the disk.  While the density structure still appears roughly spherical and hydrostatic, these flows could affect the planet's cooling.  We expect such effects to be weak, as heat losses at greater depths dominate planetary cooling, but more study is needed, especially in 3D.

 %Nevertheless, the density structure outside $\RB$ remains spherically symmetric and hydrostatic.

At higher masses, non-hydrostatic effects become more severe.  At $M\pla \gtrsim M_{\rm th}$ planets can open significant gaps \citep{zhu13}.  At yet higher masses accretion instabilities could occur \citep{AylBat12}.  However, in this high mass regime, the spherically symmetric approximation has already broken down.

Thus by restricting our attention to low masses, neglected hydrodynamic effects should be minor.   Moreover, since $M_{\rm th} \propto a^{6/7}$ increases with disk radius, spherical hydrostatic models like ours have a greater range of applicability in the outer regions of disks.

\subsection{Realistic Opacities}\label{sec:op}

The importance of envelope opacity is an established factor in core accretion calculations (e.g., \citealt{stevenson82}, \citealt{ikoma00}, R06), and several studies explore the influence of opacity on the timescales for giant planet formation in detail (e.g., \citealt{HubBod05}). Treating dust (and total) opacities as a power law in temperature is a simplification.  Opacities drop by order unity when ice grains sublimate for $T \gtrsim 150$ K and they drop by orders of magnitude when silicate grains evaporate above $T \gtrsim 1500$ K \citep{semenov03, FerAle05}.  Grain growth and composition also affect opacities. Our scenario of no ongoing accretion of solids may result in a grain-free atmosphere, which could substantially reduce the critical core mass ($\sim$$1M_{\oplus}$ in the case of Jupiter, \citealt{hori10}). Paper II explores more realistic opacity laws.

For this work, we justify a simplified dust opacity by the cool temperatures of both the outer disk and our nearly isothermal radiative zones (see \S\ref{sec:profiles}).  While  convective interiors get significantly hotter than 1500 K, opacity does not affect the structure of adiabatic convecting regions.   A possible caveat is the existence of  radiative zones sandwiched inside the convection interior.  Such ``radiative windows" arise if the opacity drop from ice and metal grain sublimation occurs at sufficiently low pressures.  Our two-layer model ignores this possibility, but radiative windows are known to exist in hot Jupiter models \citep{burrows97, ab06} and have been seen in some core accretion models (Lissauer, personal communication).  The role of radiative windows in core accretion models remains to be explored in detail.

\subsection{Equation of State}
\label{sec:EOS}

Our model uses an ideal gas law and a polytropic EOS, given by \Eq{eq:idealEOS}.  However, non-ideal effects can affect atmospheric structure and evolution.  In the lower atmosphere, H$_2$ can partially dissociate at high temperatures.  In the upper atmosphere, H$_2$ rotational levels can become depopulated at lower temeratures.   Paper II uses the \citet{saumon95} EOS, with extensions to lower pressures and temperatures as needed, to explore these effects.

\subsection{Core growth}
\label{sec:placc}

Our model purposefully neglects the accretion of solids to study the fastest rates of gas accretion.  Our $\MC$ values thus differ from the $\MC$ values in static models, such as R06.  For similar parameters, we generally obtain lower $\MC$ values than R06 because our atmospheres are not heated by ongoing planetesimal accretion.  Paper II presents a quantitative comparison.

%For giant planet formation to occur with a core mass close to our $M_{\rm crit}$, \textit{in-situ} formation of the core requires $M_{\rm crit}$ to be smaller than the isolation mass, $M_{\rm iso}$. }

%\textbf{As the core is assumed to form through accretion of solids, we require that $M_{\rm crit}>M_{\rm isol}$. Since $M_{\rm crit}$ decreases with semi major axis, while $M_{\rm isol}$ increases with $a$ (cf. Equation \ref{eq:Misol}), it is possible that the isolation mass may become larger than the critical core mass at large orbital distances. From \Eq{eq:Misol}, $M_{\rm isol} \approx 1 M_{\oplus}$ at 100 AU, which is smaller than $M_{\rm crit}$ under standard opacity assumptions, but comparable for a reduced opacity (see Figure \ref{fig:Mcritopacity}). A flatter surface density profile (see \S\ref{sec:disk}) may further increase $M_{\rm isol}$. We note, however, that the estimate of \Eq{eq:Misol} assumes an initial uniform dust-to-gas ratio throughout the disk, although the dust-to-gas ratio may be distance dependent. A lower dust-to-gas ratio in the outer disk reduces $M_{\rm isol}$. As stated in \S\ref{sec:scales}, we assume that the isolation mass is always lower than our calculated $M_{\rm crit}$.}

At low planetesimal accretion rates, $\dot{M}\co$, a static model could formally give lower $\MC$ values than our evolutionary calculations.  R06 shows that $\MC \propto \dot{M}\co^{3/5}$ in static models.  The underlying assumption of static models, a negligibly short KH contraction time, fails whenever static models give a lower $\MC$.  Thus our results give a firm lower limit on $\MC$ which complement the results of static models with planetesimal accretion.

\section{Summary} \label{sec:conclusions}

We study the formation of giant planets by the core accretion mechanism.  Our models start with a solid core that is embedded in a gas disk and no longer accreting solids.  We determine -- as a function of disk location, core mass, and the atmosphere's mean molecular weight and opacity -- whether runaway atmospheric growth can occur within a typical disk lifetime of 3 Myr.  By neglecting the accretion luminosity of planetesimals and smaller solids, we obtain the fastest allowed rate of gas accretion.

We address core accretion in the outer disk, as it is relevant to direct imaging surveys.  Our model approximations, including spherical accretion and low mass radiative exteriors with opacities dominated by dust, are tuned to conditions in the outer disk.   Our main findings are as follows:

\begin{enumerate}
\item The minimum or critical core mass, $\MC$, for giant planet formation declines with stellocentric distance in standard protoplanetary disk models.  For our reference case, the critical mass is $\sim$$8.5 M_{\oplus}$ at $a = 5$ AU, decreasing to $\sim$$3.5 M_{\oplus}$ at $a =100$ AU.  This decline roughly follows $\MC \propto a^{-0.3}$.

\item The drop in disk temperature with radial distance explains the decrease in critical core masses.  The lower pressures and densities in the outer disk only weakly suppress atmospheric growth.

\item Reducing dust opacities by a factor of 10 reduces critical core masses by a factor of $\sim$$3$.  This reduction is somewhat stronger (weaker) at larger (smaller) separations from the star.

\item A larger mean molecular weight reduces critical core masses, in agreement with \citet{HorIko11}.  If enrichment in heavy elements correlates with increased dust opacity, then the stronger opacity effect will dominate, increasing $\MC$.

\item Runaway growth begins roughly at the crossover mass, when atmosphere and core masses are equal, $M_{\rm atm} \sim M\co$, in agreement with previous work \citep{pollack96}.  Further from the star, runaway growth begins at smaller $M_{\rm atm} / M\co$ ratios.  For larger core masses, runaway growth begins at larger values of $M_{\rm atm} / M\co$.

\item Self-gravity affects atmospheric evolution before crossover.  Significant self-gravitational corrections appear when the atmosphere is only $\sim$$10 \%$ as massive as the core.

\end{enumerate}

Rapid gas accretion onto low mass cores could explain the origin of distant directly imaged giant planets \citep{marois08, lagrange10}.  However, our model does not address the details of how solid cores grow, as many possibilities exist and many uncertainties remain. %\textbf{Our brief treatment of core formation is purposeful, as there are large uncertainties in both planetesimal formation and the subsequent processes of core growth, particularly at large orbital distances. Core formation in the outer disk is challenging due to large dynamical times, and cores are unlikely to reach critical masses at $\gtrsim 40-50$ AU under classic core accretion scenarios \citep{Raf11}. However, modern models of aerodynamic pebble accretion (\citealt{ormel10}; \citealt{lambrechts12}) can give faster growth. Moreover, cores can form closer to the star, then migrate or be scattered outwards by other giants in the system \citep{ida13}.} 
For a giant planet to form with a core near the minimum masses we derive, core growth must  first be rapid and then slow significantly, as in phases 1 and 2, respectively, of \cite{pollack96}.

Initial core growth must be fast, compared to the disk lifetime, to get a sufficiently massive core.  Such rapid core growth is possible in a variety of scenarios, including the fastest gas-free planetesimal accretion rates \citep{dones93} and -- probably more relevantly for gas rich disks --  the aerodynamic accretion of mm-m sized ``pebbles" and ``boulders" in gas disks \citep{ormel10, lambrechts12}.  Moreover, cores could form rapidly closer to the star, then migrate or be scattered outwards by already formed giants \citep{ida13}.

A stronger constraint is that core growth subsequently slow severely, to allow the atmosphere to cool and contract.  To be more quantitative, the minimum cooling luminosity in \Fig{fig:Ltplot}, $L\approx 3.5 \times 10^{24}\;\mathrm{erg s}^{-1}$, could be cancelled by low levels of heating from solid accretion.   A core mass doubling timescale of $\sim$$400$ Myr, or faster, would thus provide enough heating to stall atmospheric cooling and growth.   An additional concern is that isolation masses tend to grow with disk radius, as $M_{\rm iso} \propto \varSigma_{\rm p}^{3/2} a^2 \propto a^{3/4}$ under the approximation that the surface density of accreted planetesimals ($\varSigma_{\rm p}$) scales with the gas \citep{youdin13}.   While this behavior is nominally inconsistent with final core masses that decline with distance, the predictive power of the isolation mass is imperfect.   For starters, the efficiency of planetesimal formation remains uncertain.  Moreover, the locality of core growth, which underlies the isolation mass, disappears when accreted solids drift and/or cores migrate significantly.

Thus while our calculations show that low mass cores can grow into gas giants in the outer disk, ongoing solid accretion could prevent significant atmospheric growth.  In the Solar System, the ice giants Uranus and Neptune,  with core (here ice and rock) masses of $\sim 13 -15$ M$_\oplus$, argue for the latter possibility.  Ongoing exoplanet imaging surveys and their successors \citep{hinz12, macintosh12, close14} will help discriminate among the various planet formation pathways in the outskirts of protoplanetary disks.

% For an example of a full page figure, see Fig.~\ref{fig:myFullPageFigure}.

### 2016-03-07 Dissertate-Harvard-LaTeX/frontmatter/abstract.tex
update

% the abstract
\doublespacing

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Morbi commodo, ipsum sed pharetra gravida, orci magna rhoncus neque, id pulvinar odio lorem non turpis. Nullam sit amet enim. Suspendisse id velit vitae ligula volutpat condimentum. Aliquam erat volutpat. Sed quis velit. Nulla facilisi. Nulla libero. Vivamus pharetra posuere sapien. Nam consectetuer. Sed aliquam, nunc eget euismod ullamcorper, lectus nunc ullamcorper orci, fermentum bibendum enim nibh eget ipsum. Donec porttitor ligula eu dolor. Maecenas vitae nulla consequat libero cursus venenatis. Nam magna enim, accumsan eu, blandit sed, blandit a, eros.

Quisque facilisis erat a dui. Nam malesuada ornare dolor. Cras gravida, diam sit amet rhoncus ornare, erat elit consectetuer erat, id egestas pede nibh eget odio. Proin tincidunt, velit vel porta elementum, magna diam molestie sapien, non aliquet massa pede eu diam. Aliquam iaculis. Fusce et ipsum et nulla tristique facilisis. Donec eget sem sit amet ligula viverra gravida. Etiam vehicula urna vel turpis. Suspendisse sagittis ante a urna. Morbi a est quis orci consequat rutrum. Nullam egestas feugiat felis. Integer adipiscing semper ligula. Nunc molestie, nisl sit amet cursus convallis, sapien lectus pretium metus, vitae pretium enim wisi id lectus. Donec vestibulum. Etiam vel nibh. Nulla facilisi. Mauris pharetra. Donec augue. Fusce ultrices, neque id dignissim ultrices, tellus mauris dictum elit, vel lacinia enim metus eu nunc.

### 2016-03-09 tex/draft2/ms_edited.tex
update

@@
 
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Jamila Pegues\altaffilmark{2}}
+\author{Ana-Maria A. Piso\altaffilmark{1}, Jamila Pegues\altaffilmark{2}, Karin I. \"Oberg\altaffilmark{1}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
@@
 
-Both in Solar system comets and in protoplanetary disks carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{lodders03}, \citealt{mumma11}, \citealt{oberg11}, \citealt{boogert15}).
-However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{oberg08}), which may change the C/O ratio in gas and in dust throughout the disk.
+Both in Solar system comets and in protoplanetary disks, volatile carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{lodders03}, \citealt{mumma11}, \citealt{oberg11}, \citealt{boogert15}).
+However, some fraction of carbon may also be carried by CH$_4$ (e.g., \citealt{oberg08}), which may change the C/O ratio in gas and in dust throughout the disk.
 In the case of nitrogen, chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
@@
 
-Moreover, snowline locations strongly depend on the ice grain morphology.
+The snowline locations of the main carbon, oxygen and nitrogen carriers strongly depend on the ice grain morphology.
 Very volatile species, such as CO and N$_2$, present binding energies, and therefore snowline locations, that are sensitive to the details of the morphology of the icy grain mantles.
@@
 In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, as well as nitrogen-to-oxygen (N/O) ratios.
-
 This paper is organized as follows.
@@
 
-As we note in Section \ref{sec:intro}, the disk volatile composition and the ice morphology determine the location of several important snowlines.
+As we note in Section \ref{sec:intro}, the disk volatile composition and the ice morphology determine the location of important snowlines.
 In this work we focus on the primary carbon, oxygen and nitrogen carriers, i.e.
 H$_2$O, CO$_2$, CO, N$_2$, and to a lesser extent, CH$_4$ and NH$_3$.
-To quantify the effect of the presence of some carbon in the form of CH$_4$ and of some nitrogen in the form of NH$_3$, we use measured median CH$_4$ and NH$_3$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
+%To quantify the effect of the presence of some carbon in the form of CH$_4$ and of some nitrogen in the form of NH$_3$, we use measured median CH$_4$ and NH$_3$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}.
+Our standard model is based on the median ice abundances observed toward Solar-type protostars \citep{oberg11a}, which are $n_{\rm CO_2}=0.29 \times n_{\rm H_2O}$, $n_{\rm CO}=0.38 \times n_{\rm H_2O}$, $n_{\rm CH_4}=0.0555 \times n_{\rm H_2O}$ (hereafter CH$_4$-mid) and $n_{\rm NH_3}=0.055 \times n_{\rm H_2O}$ (hereafter NH$_3$-mid).
+Here $n_{\rm H_2O} \approx 10^{-4} \times n_{\rm H}$ is the total water abundance \citep{vandishoeck06}, with $n_{\rm H}$ the hydrogen abundance in the disk midplane.
+For CO, we also take into account that the observed CO ice only traces some of the CO reservoir due to its high volatility, and similarly to \citet{oberg11} and Paper I we set the total CO abundance to $0.9 \times 10^{-4} n_{\rm H}$.
+Finally, we assume that all nitrogen not found in NH$_3$ is in N$_2$ and assume a Solar nitrogen abundance, $n_{\rm N}=8 \times 10^{-5} n_{\rm H}$ \citep{lodders03}.
+In effect, this model assumes no chemical evolution between the protostellar and disk midplane stages.
+This is reasonable for material that accretes onto the disk at large radii \citep{visser09}, but may overestimate the contribution of the original volatiles to the total volatile budget in the innermost disk.
 
@@
 The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as a function of the hydrogen total abundance.
-As expected, the total grain abundance increases with semimajor axis, as more and more species sublimate.
+As expected, the total grain abundance increases with semimajor axis, as more and more species freeze out.
 Freeze-out at the CO$_2$ and CO snowlines pulls more heavy elements into the grains than in the case of the H$_2$O snowline.
-More importantly, the CO and N$_2$ snowlines move several tens of AU inward if the ices are water dominated rather than pure.
+The CO and N$_2$ snowlines move several tens of AU inward if the ices are water dominated rather than pure.
 This changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ.
@@
 
-In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and the extent to which they are affected by the presence of CH$_4$ and NH$_3$.
+In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and to what extent they are affected by the presence of CH$_4$ and NH$_3$ over the full range of observed CH$_4$ and NH$_3$ abundances.
 Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.
 
-We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max).
-Thus $n_{\rm CH_4-mid}=0.0555 \times n_{\rm H_2O}$ \citep{oberg11a} and $n_{\rm CH_4-max}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}, where $n_{\rm H_2O}$ is the total H$_2$O abundance.
-Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}.
+We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) CH$_4$-mid, and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max), $n_{\rm CH_4-max}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}, where $n_{\rm H_2O}$ is the total H$_2$O abundance.
+%Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}.
 Since the abundance of carbon grains is uncertain, we assume that all the carbon that is not in the form of CH$_4$, CO and CO$_2$ is in carbon grains, so that we reproduce the Solar C/O ratio (gas+dust) of 0.54.
@@
 We assume that the main nitrogen-bearing species are N$_2$ and NH$_3$, since other volatiles that contain nitrogen have significantly lower abundances in comparison (e.g., \citealt{mumma11}).
-We use the measured total nitrogen abundance in the Solar system, $n_{\rm N}=8 \times 10^{-5} n_{\rm H}$ \citep{lodders03}, where $n_{\rm H}$ is the hydrogen abundance in the disk midplane.
-Similarly to the case of CH$_4$, we explore the parameter space of possible NH$_3$ abundances using data from the Spitzer c2d Legacy ice survey, as follows: (1) no NH$_3$, (2) the median NH$_3$ observed abundance $n_{\rm NH_3-mid}=0.055 \times n_{\rm H_2O}$ \citep{oberg11a}, and (3) the maximum observed NH$_3$ abundance $n_{\rm NH_3-max}=0.15 \times n_{\rm H_2O}$ \citep{bottinelli10}.
+Similarly to the case of CH$_4$, we explore the parameter space of possible NH$_3$ abundances using observations of protostellar cores, as follows: (1) no NH$_3$, (2) NH$_3$-mid, and (3) the maximum observed NH$_3$ abundance $n_{\rm NH_3-max}=0.15 \times n_{\rm H_2O}$ \citep{bottinelli10}.
 In each case, the N$_2$ abundance then simply follows as $n_{\rm N_2}=(n_{\rm N}-n_{\rm NH_3})/2$.
@@
 The horizontal dashed lines represent the average N/O value in the disk.
-The gas-phase N/O ratio is highly enhanced by a factor of three between the CO$_2$ and CO snowlines compared to its average value.
+The gas-phase N/O ratio is enhanced by a factor of two between the H$_2$O and CO$_2$ snowlines compared to its average value, and by a factor of three between the CO$_2$ and CO snowlines.
 The arrows mark a highly elevated N/O ratio in gas between the CO and N$_2$ snowlines due to the depletion of oxygen gas in this region.
@@
 Figure \ref{fig:NO_ratio} shows the H$_2$O, CO$_2$, CO and N$_2$ snowline locations in a viscous disk for the same range of initial particle sizes as in Figure \ref{fig:CO_ratio}, and with the CO and N$_2$ snowlines calculated assuming different grain morphologies as explained above, as well as estimates for the N/O ratio throughout the disk.
-For simplicity, we assume that all nitrogen is bound in N$_2$.
+For simplicity, we assume that all nitrogen is the form of N$_2$.
 This choice is justified since the presence of some NH$_3$ only moderately changes the N/O ratio (see Figure \ref{fig:Nstatic}), and since we are primarily interested in the N$_2$ snowline locations rather than exact values for the N/O ratio.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{N_O_water_ice.pdf}
+\includegraphics[width=0.5\textwidth]{N_O_water_ice_many.pdf}
 %\vspace{-0.5in}
 \caption{N/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO and N$_2$ as pure ices (top panel) or as water dominated ices (bottom panel).
-The H$_2$O, CO$_2$ and N$_2$ snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
-The CO snowline (inner vertical dashed line in both panels) is only shown for the 7 m particles.
+The H$_2$O, CO$_2$, CO and N$_2$ snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
 The N/O ratio in a static disk (black lines) is shown for comparison.
@@
 Taken together, disk dynamics and ice morphology move the CO snowline inward by a factor of $\sim$7.
-The results of highly enhanced gas-phase N/O ratios between the CO$_2$ and CO snowlines compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}
+The results of an enhanced gas-phase N/O ratio between the H$_2$O and CO snowlines compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}
 \label{fig:NO_ratio}
@@
 Thus constraining the N/O ratio in a giant planet atmosphere could be used to trace its formation origins.
+
 Theoretical models of the magnitude and role of N/O (and N/C) ratios in exoplanet atmospheres are therefore key in order to use these ratios as probes for a planet's formation location.
-Measurements of the N/O ratio in planet atmospheres are essential to further constrain the theoretical expectations --- while nitrogen-bearing molecules have not been targeted so far due to lack of instrument sensitivity, such observations and detections are likely in the near future with the advent of JWST (e.g., NH$_3$, \citealt{greene16}).
+Models that explore the effect of varying the C/O ratio in exoplanet atmospheres exist in literature, and they display a large and observable effect on gas giant envelope chemistry (\citealt{lodders09}, \citealt{molliere15}).
+However, no similar model explorations exist for the effect of N/O and C/N/O ratios, and both are needed to exploit this potential constraint.
+Given the existence of such theoretical models, measurements of the N/O ratio in planetary envelopes may be possible to infer from atmospheric compositions of nitrogen versus carbon and oxygen bearing species.
+Nitrogen carriers have not been targeted so far due to lack of instrument sensitivity, but such observations and detections are likely in the near future with the advent of JWST (e.g., NH$_3$, \citealt{greene16}).
+The N/O ratio enhancement is larger than that of the gas phase C/O ratio throughout most of the disk.
+Thus measurements of an enhanced C/O ratio in an exoplanet atmosphere could be corroborated (disproved) by measurements of enhanced (non-enhanced) N/O ratios.
+Moreover, Figure \ref{fig:enhance} shows that giant planets that have formed in situ between the H$_2$O and CO snowlines are expected to present elevated both C/O and N/O ratios in their atmospheres, whereas planets between the CO and N$_2$ snowlines will have a highly enhanced N/O ratio in their atmospheres, but not C/O.
 
@@
 
-The N/O ratio enhancement is larger than that of the gas phase C/O ratio throughout most of the disk.
-Thus measurements of an enhanced C/O ratio in an exoplanet atmosphere could be corroborated (disproved) by measurements of enhanced (non-enhanced) N/O ratios.
-Moreover, Figure \ref{fig:enhance} shows that giant planets that have formed in situ between the H$_2$O and CO snowlines are expected to present elevated both C/O and N/O ratios in their atmospheres, whereas planets between the CO and N$_2$ snowlines will have a highly enhanced N/O ratio in their atmospheres, but not C/O.
-
 Due to disk dynamics and ice morphology, the disk regions with highly elevated gas phase N/O and C/O ratios are uncertain and may span tens of AU.
-In particular, one can imagine a scenario where CO is in a water binding environment and N$_2$ is not.
+Both ice morphologies discussed in this study, pure and water dominated ices, are plausible in protoplanetary disks and depend on whether H$_2$O and CO ices formed on similar timescales or successively (e.g., \citealt{garrod11}).
+Observations of protostellar cores show that a large fraction of CO is bound in a pure ice multilayer \citep{pontoppidan03}, but theoretical models also suggest an icy mantle structure where CO resides on a H$_2$O ice layer (e.g., \citealt{collings03}).
+One can also imagine a scenario where CO is in a water binding environment and N$_2$ is not.
 This could be attributed to the fact that H$_2$O may bind preferentially to CO than N$_2$, since both H$_2$O and CO are polar molecules while N$_2$ is not.
 It is also possible for N$_2$ ices to form later than CO (e.g., \citealt{pagani12}), and thus be deposited on the outer layers of the icy mantles which are typically water poor (e.g., \citealt{garrod11}).
-Detections of CO and N$_2$ snowlines in the same disk could potentially verify this scenario, since CO and N$_2$ desorb at substantially different temperature, and therefore locations, if the former is water dominated and the latter is in a pure ice environment.
-Changes in stellar luminosity (e.g., \citealt{kennedy06}) and gas mass accretion rate (e.g., \citealt{chambers09}) may introduce additional uncertainties in the snowline locations, and thus the C/N/O ratios (see also Paper I).
+Detections of CO and N$_2$ snowlines in the same disk could potentially verify this scenario, since CO and N$_2$ desorb at substantially different temperatures, and therefore locations, if the former is water dominated and the latter is in a pure ice environment.
 The impact of the ice environment on the snowline location is much smaller in the case of CO$_2$ and NH$_3$, as their binding energies and behavior are closer to that of H$_2$O.
+No detailed measurements for the CH$_4$ binding energy in a water environment exist so far, but due to its low desorption temperature a similar behavior to that of CO and N$_2$ would be expected.
+While the presence of some carbon in the form of CH$_4$ only modestly affects our results, CH$_4$ may become important in disks where a large fraction of the CO abundance has been converted into hydrocarbons (e.g., \citealt{du15}).
+
+Changes in stellar luminosity (e.g., \citealt{kennedy06}) and gas mass accretion rate (e.g., \citealt{chambers09}), as well as the evolution of icy dust particles due to grain growth and fragmentation (e.g., \citealt{birnstiel12}), may introduce additional uncertainties in the snowline locations, and thus the C/N/O ratios.
+Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints.
+Current detections include the H$_2$O snowline in TW Hya \citep{zhang13} and the CO snowline in TW Hya \citep{qi13} and HD 163296 \citep{qi15}.
+However, more snowline detections in disks of different ages are expected in future ALMA cycles.
+
+ %The impact of the ice environment on the snowline location is much smaller in the case of CO$_2$ and NH$_3$, as their binding energies and behavior are closer to that of H$_2$O.
 No detailed measurements for the CH$_4$ binding energy in a water environment exist so far, but due to its low desorption temperature a similar behavior to that of CO and N$_2$ would be expected.
@@
 
-Our results could also be applied to the Solar system.
+Uncertainties in snowline locations of this magnitude also affect interpretations of Solar system observations.
 Recent measurements of nitrogen abundance in comet 67P/Churyumov-Gerasimenko found a N$_2$/CO ratio $\sim 10^{-3}$ \citep{rubin15}.
-A low N$_2$/CO ratio is consistent with comets having formed inside the N$_2$ snowline where solid N$_2$ is depleted.
+A low N$_2$/CO ratio is consistent with comets having formed inside the N$_2$ snowline where N$_2$ is still in the gas phase.
 Theoretical models suggest that Jupiter-family comets, such as 67P, originate from the Kuiper belt \citep{duncan97} (but see \citealt{rubin15} for alternative formation scenarios for 67P).
@@
 
-Finally, detections of snowlines in disks that are reasonably well characterized could help further constrain the dynamics and ice morphology in the disks.
+%Finally, detections of snowlines in disks that are reasonably well characterized could help further constrain the dynamics and ice morphology in the disks.
 
@@
 
-\item The presence of some carbon in the form of CH$_4$ only modestly affects our results for the gas phase C/O ratio, increasing it by up to 8\% between the CO$_2$ and CH$_4$ snowlines.
-The presence of some nitrogen in the form of NH$_3$ moderately affects our results and may decrease the gaseous N/O ratio by up to 18\% between the NH$_3$ and CO$_2$ snowlines.
+\item The effect of CH$_4$ and NH$_3$ on the C/O and N/O ratios is small, even when we consider the maximum observed CH$_4$ and NH$_3$ abundances in protostellar cores.
+In this scenario, the gas phase C/O ratio increases by 8\% between the CO$_2$ and CH$_4$ snowlines, and  the gaseous N/O ratio increases by up to 18\% between the NH$_3$ and CO$_2$ snowlines.
 In both cases, large gas phase C/O and N/O ratios in the outer disk are preserved.

### 2016-03-09 tex/draft2/ms_edited.tex
update

@@
 Current detections include the H$_2$O snowline in TW Hya \citep{zhang13} and the CO snowline in TW Hya \citep{qi13} and HD 163296 \citep{qi15}.
-However, more snowline detections in disks of different ages are expected in future ALMA cycles.
+More snowline detections in disks of different ages are expected in future ALMA cycles.
+Snowline observations may also be used to constrain the degree of turbulence in protoplanetary disks \citep{owen14}.
+We note, however, that observations of snowlines may not be enough to break the degeneracies in snowline locations produced by disk dynamics, ice morphology, and other effects outlined above.
+For example, a detection of the CO snowline at a temperature higher than the assumed desorption temperature in a static disk with CO as pure ice ($\sim$17 K, e.g.
+\citealt{qi13}) could be caused either by CO being in a water dominated environment or by dynamical effects that push the CO snowline inward.
+This degeneracy can be broken in extreme cases --- for our fiducial disk model, this implies a detection of a CO snowline close to $\sim$9 AU or $\sim$61 AU (see Section \ref{sec:dynamic}).
 

### 2016-03-09 python/C_to_O_active_steady_time.py

@@ -1,0 +1,141 @@
+"""This script is used to calculate the C/O ratio in gas and in dust
+as a function of temperature, i.e. the location in the disk relative 
+to the H2O, CO2 and CO snowlines"""
+
+import numpy as np
+from T_freeze import T_freeze
+from drift_timescales import read_drag_file_many
+
+def n_C_O(species):
+    
+    """
+        
+    Returns the abundance of C and O in various molecular species
+    (numbers are taken from Oberg+11b, Table 1)
+    
+    
+    Input
+    -----
+    species:
+        string: 'CO', 'CO2' etc.
+    
+    Output
+    ------
+    n_C:
+        abundance of C with respect to H2
+    n_O:
+        abundance of O with respect to H2
+        
+    """
+    
+    if species == 'CO':
+        return 1.5, 1.5
+    elif species == 'CO2':
+        return 0.3, 0.6
+    elif species == 'H2O':
+        return 0, 0.9
+    elif species == 'C_grains':
+        return 0.6, 0
+    elif species == 'silicate':
+        return 0, 1.4
+
+
+#in the following we calculate T_freeze for various species
+s, a, arrayH2O = read_drag_file_many('s_neg3_8_steady_correct.txt', 'H2O', 50, 50)
+s, a, arrayCO2 = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO2', 50, 50)
+s, a, arrayCO = read_drag_file_many('s_neg3_8_steady_correct.txt', 'CO', 50, 50)
+
+
+
+
+def n(index, r, elem):
+    
+    """
+    
+    Returns the gaseous and solid abundance of a given element (C, O)
+    as a function of temperature, e.g. between different snowlines
+    
+    Input
+    -----
+    T:
+        disk midplane temperature in units of K
+    elem:
+        string: 'C' or 'O'
+        
+    Output
+    ------
+    n_gas:
+        abundance of element 'elem' in gaseous form relative to H2 abundance
+    n_solid:
+        abundance of element 'elem' in solid form relative to H2 abundance
+    
+    """
+    
+    if elem == 'C':
+        i = 0 #flag to simplify the use of the n_C_O function; = 0 for C
+    elif elem == 'O':
+        i = 1 #flag to simplify the use of the n_C_O function; = 1 for O
+    
+    size = s[index]
+    aH2O = arrayH2O[index][-1]
[…]

### 2016-03-10 tex/draft2/ms_edited.tex
update

--- 2016-03-09 tex/draft2/ms_edited.tex
@@
 We explore the effects of dynamical processes, molecular compositions and abundances, and the ice morphology of dust grains in disks on the snowline locations of the main C, O and N carriers, and their consequences for the C/N/O ratios in gas and dust throughout the disk.
-%We find that radial drift and accretion alone can reduce the snowline radii of the main C, O and N carriers, i.e. H$_2$O, CO$_2$, CO and N$_2$, by 40-60\% compared to static disks. 
 The gas-phase N/O ratio enhancement in the outer disk (exterior to the H$_2$O snowline) exceeds the C/O ratio enhancement for all reasonable volatile compositions.
@@
 
-Due to disk dynamics and ice morphology, the disk regions with highly elevated gas phase N/O and C/O ratios are uncertain and may span tens of AU.
+Due to disk dynamics and ice morphology, the locations of the CO and N$_2$ snowlines, and thus the disk regions with highly elevated gas phase N/O and C/O ratios, are uncertain and may span tens of AU.
 Both ice morphologies discussed in this study, pure and water dominated ices, are plausible in protoplanetary disks and depend on whether H$_2$O and CO ices formed on similar timescales or successively (e.g., \citealt{garrod11}).
@@
 It is also possible for N$_2$ ices to form later than CO (e.g., \citealt{pagani12}), and thus be deposited on the outer layers of the icy mantles which are typically water poor (e.g., \citealt{garrod11}).
-Detections of CO and N$_2$ snowlines in the same disk could potentially verify this scenario, since CO and N$_2$ desorb at substantially different temperatures, and therefore locations, if the former is water dominated and the latter is in a pure ice environment.
 The impact of the ice environment on the snowline location is much smaller in the case of CO$_2$ and NH$_3$, as their binding energies and behavior are closer to that of H$_2$O.
@@
 Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints.
-Current detections include the H$_2$O snowline in TW Hya \citep{zhang13} and the CO snowline in TW Hya \citep{qi13} and HD 163296 \citep{qi15}.
-However, more snowline detections in disks of different ages are expected in future ALMA cycles.
+The uncertainties in snowline locations caused by disk dynamics, ice morphology, and other effects outlined above can be resolved in extreme cases, such as a detection of a CO snowline at a temperature corresponding to pure CO ice desorption in a static disk (e.g., \citealt{qi13} at $\sim$ 17 K).
+In other cases, however, the degeneracies in snowline locations cannot be broken.
+For example, the CO snowline in HD 163296 is at a higher temperature of $\sim$ 25 K \citep{qi15}, which could be caused either by CO being in a water dominated environment or by dynamical effects that push the CO snowline inward.
+Detections of both the CO and N$_2$ snowlines in the same disk could potentially break this degeneracy, since CO and N$_2$ desorb at substantially different temperatures, and therefore locations, if the former is water dominated and the latter is in a pure ice environment.
+
+%Current detections include the H$_2$O snowline in TW Hya \citep{zhang13} and the CO snowline in TW Hya \citep{qi13} and HD 163296 \citep{qi15}.
+More snowline detections in disks of different ages are expected in future ALMA cycles.
+Snowline observations may also be used to constrain the degree of turbulence in protoplanetary disks \citep{owen14}.
+We note, however, that observations of snowlines may not be enough to break the degeneracies in snowline locations produced by disk dynamics, ice morphology, and other effects outlined above.
+For example, a detection of the CO snowline at a temperature higher than the assumed desorption temperature in a static disk with CO as pure ice ($\sim$17 K, e.g.
+\citealt{qi13}) could be caused either by CO being in a water dominated environment or by dynamical effects that push the CO snowline inward.
+This degeneracy can be broken in extreme cases --- for our fiducial disk model, this implies a detection of a CO snowline close to $\sim$9 AU or $\sim$61 AU (see Section \ref{sec:dynamic}).
 

### 2016-03-11 tex/draft2/ms_edited.tex
update

@@
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-\altaffiltext{2}{Department of Astrophysical Sciences, Princeton University}
+\altaffiltext{2}{Department of Astrophysical Sciences, Princeton University, Princeton, NJ 08544}
 
@@
 In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, their ice binding environment, and the presence of snowlines of major volatiles at different distances from the central star.
-We explore the effects of dynamical processes, molecular compositions and abundances, and the ice morphology of dust grains in disks on the snowline locations of the main C, O and N carriers, and their consequences for the C/N/O ratios in gas and dust throughout the disk.
+We explore the effects of disk dynamical processes, molecular compositions and abundances, and ice morphology on the snowline locations of the main C, O and N carriers, and the C/N/O ratios in gas and dust throughout the disk.
 The gas-phase N/O ratio enhancement in the outer disk (exterior to the H$_2$O snowline) exceeds the C/O ratio enhancement for all reasonable volatile compositions.
 Ice morphology and disk dynamics individually change the snowline location of N$_2$, the main nitrogen carrier, by a factor of 2-3, and when considered together the range of possible N$_2$ snowline locations is $\sim$11-$\sim$79 AU in a standard disk model.
-The CO snowline location is similarly uncertain.
-This snowline location uncertainty limits the utility of C/N/O compositional ratios in planets to trace their formation zone.
-Observations that anchor snowline locations at different stages of planet formation are therefore key.
+Observations that anchor snowline locations at different stages of planet formation are therefore key to develop C/N/O ratios as a probe of planet formation zones.
 
@@
 Both in Solar system comets and in protoplanetary disks, volatile carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{lodders03}, \citealt{mumma11}, \citealt{oberg11}, \citealt{boogert15}).
-However, some fraction of carbon may also be carried by CH$_4$ (e.g., \citealt{oberg08}), which may change the C/O ratio in gas and in dust throughout the disk.
+However, some fraction of carbon may also be carried by CH$_4$ (e.g., \citealt{oberg08}), which may change the C/O ratio in gas and in dust at some disk locations.
 In the case of nitrogen, chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
@@
 Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
-In addition to N$_2$, a fraction of the nitrogen abundances may be also carried by NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).
 Because of the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio compared to its average value in the disk.
 Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
+In addition to N$_2$, a fraction of the nitrogen abundance may also be carried by less volatile species such as NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).
 
@@
 \centering
-\includegraphics[width=\textwidth]{CNO_and_snowlines_2.pdf}
+\includegraphics[width=0.8\textwidth]{CNO_and_snowlines_2.pdf}
 %\vspace{-0.5in}
@@
 For CO and N$_2$ as water dominated ices, the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}.
-Figure \ref{fig:CNOstatic} shows the resulting snowline locations, with CO and N$_2$ pure ices (top panel) and water dominated ices (bottom panel).
+Figure \ref{fig:CNOstatic} shows the resulting snowline locations, assuming CO and N$_2$ pure ices (top panel), and CO and N$_2$ in water dominated ices (bottom panel).
 The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as a function of the hydrogen total abundance.
@@
 
-In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and to what extent they are affected by the presence of CH$_4$ and NH$_3$ over the full range of observed CH$_4$ and NH$_3$ abundances.
-Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.
+In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and to what extent they are affected by the presence of CH$_4$ and NH$_3$ over the full range of observed CH$_4$ and NH$_3$ abundances toward low-mass protostars.
+In this section we only consider pure ices.
+%Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.
 
@@
 We assume that the main nitrogen-bearing species are N$_2$ and NH$_3$, since other volatiles that contain nitrogen have significantly lower abundances in comparison (e.g., \citealt{mumma11}).
-Similarly to the case of CH$_4$, we explore the parameter space of possible NH$_3$ abundances using observations of protostellar cores, as follows: (1) no NH$_3$, (2) NH$_3$-mid, and (3) the maximum observed NH$_3$ abundance $n_{\rm NH_3-max}=0.15 \times n_{\rm H_2O}$ \citep{bottinelli10}.
+Similarly to the case of CH$_4$, we explore the parameter space of possible NH$_3$ abundances using observations toward low-mass protostars, as follows: (1) no NH$_3$, (2) NH$_3$-mid, and (3) the maximum observed NH$_3$ abundance $n_{\rm NH_3-max}=0.15 \times n_{\rm H_2O}$ \citep{bottinelli10}.
 In each case, the N$_2$ abundance then simply follows as $n_{\rm N_2}=(n_{\rm N}-n_{\rm NH_3})/2$.
@@
 This latter region can span tens of AU depending on disk parameters and the relative CO and N$_2$ ice binding environment.
-This enhancement is more pronounced than the C/O gas phase enhancement of a factor of two in the outer disk (see Figure \ref{fig:COstatic}).
+This N/O  enhancement is more pronounced than the C/O gas phase enhancement of a factor of two in the outer disk (see Figure \ref{fig:COstatic}).
 %The N/O ratio reaches particularly high values between the CO and N$_2$ snowlines, where all the oxygen is now contained in grains.
@@
 
-Theoretical models of the magnitude and role of N/O (and N/C) ratios in exoplanet atmospheres are therefore key in order to use these ratios as probes for a planet's formation location.
+Theoretical models of the magnitude and role of N/O (and N/C) ratios in exoplanet atmospheres are needed in order to use these ratios as probes for a planet's formation location.
 Models that explore the effect of varying the C/O ratio in exoplanet atmospheres exist in literature, and they display a large and observable effect on gas giant envelope chemistry (\citealt{lodders09}, \citealt{molliere15}).
@@
 Changes in stellar luminosity (e.g., \citealt{kennedy06}) and gas mass accretion rate (e.g., \citealt{chambers09}), as well as the evolution of icy dust particles due to grain growth and fragmentation (e.g., \citealt{birnstiel12}), may introduce additional uncertainties in the snowline locations, and thus the C/N/O ratios.
-Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints.
-The uncertainties in snowline locations caused by disk dynamics, ice morphology, and other effects outlined above can be resolved in extreme cases, such as a detection of a CO snowline at a temperature corresponding to pure CO ice desorption in a static disk (e.g., \citealt{qi13} at $\sim$ 17 K).
-In other cases, however, the degeneracies in snowline locations cannot be broken.
+Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints on the relative importance of ice morphology and disk dynamics in setting snowline locations.
+The uncertainties in snowline locations caused by disk dynamics, ice morphology, and other effects outlined above can be resolved in extreme cases, such as a detection of a CO snowline at a temperature corresponding to pure CO ice desorption in a static disk (e.g., \citealt{qi13} at $\sim$ 17 K) or CO desorption from a water dominated ice in a dynamic disk.
+In intermediate cases it is more difficult to resolve the relative importance of ice morphology and disk dynamics.
 For example, the CO snowline in HD 163296 is at a higher temperature of $\sim$ 25 K \citep{qi15}, which could be caused either by CO being in a water dominated environment or by dynamical effects that push the CO snowline inward.
-Detections of both the CO and N$_2$ snowlines in the same disk could potentially break this degeneracy, since CO and N$_2$ desorb at substantially different temperatures, and therefore locations, if the former is water dominated and the latter is in a pure ice environment.
+Detections of multiple snowlines in the same disk could potentially break this degeneracy.
+%, since CO and N$_2$ desorb at substantially different temperatures, and therefore locations, if the former is water dominated and the latter is in a pure ice environment.
 
@@
 
-\item Due to the high volatility of N$_2$, the gaseous N/O ratio outside the H$_2$O snowline is enhanced by a factor of two compared to its average value, by more than a factor of three between the CO$_2$ and CO snowlines, and by many orders of magnitude between the CO and N$_2$ snowlines doe to the complete depletion of oxygen gas in this region.
+\item Due to the high volatility of N$_2$, the gaseous N/O ratio outside the H$_2$O snowline is enhanced by a factor of two compared to its average value, by more than a factor of three between the CO$_2$ and CO snowlines, and by many orders of magnitude between the CO and N$_2$ snowlines due to the complete depletion of oxygen gas in this region.
 This enhancement is more pronounced than in the case of the gas-phase C/O ratio, which is increased by at most a factor of two compared to the stellar value.
@@
 \item The effect of CH$_4$ and NH$_3$ on the C/O and N/O ratios is small, even when we consider the maximum observed CH$_4$ and NH$_3$ abundances in protostellar cores.
-In this scenario, the gas phase C/O ratio increases by 8\% between the CO$_2$ and CH$_4$ snowlines, and  the gaseous N/O ratio increases by up to 18\% between the NH$_3$ and CO$_2$ snowlines.
+In this scenario, the gas phase C/O ratio increases by 8\% between the CO$_2$ and CH$_4$ snowlines, and  the gaseous N/O ratio decreases by up to 18\% between the NH$_3$ and CO$_2$ snowlines.
 In both cases, large gas phase C/O and N/O ratios in the outer disk are preserved.
@@
 \item Grain composition sensitively affects the CO and N$_2$ snowline locations.
-If CO and N$_2$ are water dominated rather than pure ices, their snowlines move inward by up to $\sim$70 \%.
+If CO and N$_2$ reside in water dominated rather than pure ices, their snowlines move inward by up to $\sim$70 \%.
 This effect is separate from that of radial drift and viscous gas accretion, which also cause an inward movement of the CO and N$_2$ snowlines by up to $\sim$50 \%.
@@
 Moreover, our model shows that wide separation gas giants may have an excess of nitrogen in their envelopes, which may be used to trace their origins.
-In future work, we plan to add new levels of complexity to our model in terms of disk chemistry, dynamics, and planetary dynamics, thus forming a solid framework for understanding the origins of gas giants.
+%In future work, we plan to add new levels of complexity to our model in terms of disk chemistry, dynamics, and planetary dynamics, thus forming a solid framework for understanding the origins of gas giants.
+
+\acknowledgments{This work is supported by a Simons Collaboration on the Origins of Life (SCOL) investigator award to KIO. JP was supported by the Banneker institute.}
 

### 2016-03-11 Dissertate-Harvard-LaTeX/chapters/chapter1.tex
update

@@
 %\end{savequote}
-%
-\chapter{The title of chapter one}
+
+\chapter{On the Minimum Core Mass For Giant Planet Formation at Wide Separations}
 %
@@
 %
+
+\section*{Abstract}
+
+In the core accretion hypothesis, giant planets form by gas accretion onto solid protoplanetary cores.
+The minimum (or critical) core mass to form a gas giant is typically quoted as $10 M_{\oplus}$.
+The actual value depends on several factors: the location in the protoplanetary disk, atmospheric opacity, and the accretion rate of solids.
+Motivated by ongoing direct imaging searches for giant planets, this study investigates core mass requirements in the outer disk.
+To determine the fastest allowed rates of gas accretion, we consider solid cores that no longer accrete planetesimals, as this would heat the gaseous envelope.
+Our spherical, two-layer atmospheric cooling model includes an inner convective region and an outer radiative zone that matches onto the disk.
+We determine the minimum core mass for a giant planet to form within a typical disk lifetime of 3 Myr.
+The minimum core mass declines with disk radius, from $\sim$$8.5 M_{\oplus}$ at 5 AU to $\sim$$3.5 M_{\oplus}$ at 100 AU, with standard interstellar grain opacities.
+Lower temperatures in the outer disk explain this trend, while variations in disk density are less influential.
+At all distances, a lower dust opacity or higher mean molecular weight reduces the critical core mass.
+Our non-self-gravitating, analytic cooling model reveals that self-gravity significantly affects early atmospheric evolution, starting when the atmosphere is only $\sim$$10\%$ as massive as the core.
+
+\noindent The contents of this chapter are published in \textit{Piso, A.-M. A \& Youdin, A. N. Y}
 

### 2016-03-11 tex/draft2/ms_edited.tex
update

@@
 
-In this work, we expand the coupled drift-desorption model developed in \citet{piso15b} (hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, as well as nitrogen-to-oxygen (N/O) ratios.
+In this work, we expand the coupled drift-desorption model developed in (\citealt{piso15b}; hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, as well as nitrogen-to-oxygen (N/O) ratios.
 This paper is organized as follows.
@@
 \centering
-\includegraphics[width=0.8\textwidth]{CNO_and_snowlines_2.pdf}
+\includegraphics[width=0.95\textwidth]{CNO_and_snowlines_2.eps}
 %\vspace{-0.5in}
-\caption{The total carbon, nitrogen and oxygen abundance as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and water dominated ices (bottom panel).
+\caption{The total carbon, nitrogen and oxygen abundance in solids as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and water dominated ices (bottom panel).
 Relevant volatile snowlines are marked by the vertical dashed lines.
@@
 
-We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) CH$_4$-mid, and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max), $n_{\rm CH_4-max}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}, where $n_{\rm H_2O}$ is the total H$_2$O abundance.
+We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) CH$_4$-mid, and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max), $n_{\rm CH_4-max}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}.
 %Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{C_O_ratio_CH4.pdf}
+\includegraphics[width=0.5\textwidth]{C_O_ratio_CH4.eps}
 %\vspace{-0.5in}
@@
 The C/O estimates are performed assuming that the CO ices are in pure form.
-The vertical dashed lines mark the snowline locations of the main C and O carriers.
-The horizontal dashed lines represent the stellar C/O value.
+The vertical dotted lines mark the snowline locations of the main C and O carriers.
+The horizontal dotted lines represent the stellar C/O value.
 The presence of methane only modestly increases the C/O ratio in gas between the CO$_2$ and CH$_4$ snowlines.}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{N_O_ratio.pdf}
+\includegraphics[width=0.5\textwidth]{N_O_ratio.eps}
 %\vspace{-0.5in}
@@
 The N/O estimates are performed assuming that the CO and N$_2$ ices are in pure form.
-The vertical dashed lines mark the snowline locations of the main C,O and N carriers.
-The horizontal dashed lines represent the average N/O value in the disk.
+The vertical dotted lines mark the snowline locations of the main C,O and N carriers.
+The horizontal dotted lines represent the average N/O value in the disk.
 The gas-phase N/O ratio is enhanced by a factor of two between the H$_2$O and CO$_2$ snowlines compared to its average value, and by a factor of three between the CO$_2$ and CO snowlines.
@@
 Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk, for our three choices of the NH$_3$ abundance: no NH$_3$ (top panel), NH$_3$-mid (middle panel) and NH$_3$-max (bottom panel).
-For comparison, the horizontal dashed lines show the average N/O ratio in the disk.
+For comparison, the horizontal dotted lines show the average N/O ratio in the disk.
 As expected, the gaseous N/O ratio generally exhibits an increasing trend towards the outer disk as more oxygen gas is depleted, with small decreases between the NH$_3$ and CO$_2$ snowlines (by 6\% for NH$_3$-mid and by 18\% for NH$_3$-max, respectively) due to NH$_3$ freeze-out.
@@
 Here we use the model of Section \ref{sec:review} to estimate the movement of the CO and N$_2$ snowlines for different grain morphologies in a viscous disk.
-Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.06$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above.
+Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above.
 We assume there is no carbon in the form of CH$_4$.
-The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.06$ cm do not form true snowlines.
+The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.05$ cm do not form true snowlines.
 If the CO binding environment is known, the CO snowline moves inward by up to $\sim$50 \% compared to a static disk for each case (pure and water dominated ices) due to disk dynamics.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{C_O_water_ice.pdf}
+\includegraphics[width=0.5\textwidth]{C_O_water_ice.eps}
 %\vspace{-0.5in}
 \caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO as pure ice (top panel) or as water dominated ices (bottom panel).
-The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
+The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
 The C/O ratio in a static disk (black lines) is shown for comparison.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{N_O_water_ice_many.pdf}
+\includegraphics[width=0.5\textwidth]{N_O_water_ice_many.eps}
 %\vspace{-0.5in}
 \caption{N/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO and N$_2$ as pure ices (top panel) or as water dominated ices (bottom panel).
-The H$_2$O, CO$_2$, CO and N$_2$ snowlines are shown for particles with initial sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
+The H$_2$O, CO$_2$, CO and N$_2$ snowlines are shown for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar.
 The N/O ratio in a static disk (black lines) is shown for comparison.
-The downward arrows show that the N/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in the viscous disk, as the relative fluxes of the desorbed icy
+The arrows show that the N/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in the viscous disk, as the relative fluxes of the desorbed icy
 particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details).
@@
 The presence of N$_2$ in a water ice environment rather than as pure ice moves the N$_2$ snowline significantly inward by $\sim$70\%.
-Taken together, disk dynamics and ice morphology move the CO snowline inward by a factor of $\sim$7.
+Taken together, disk dynamics and ice morphology move the N$_2$ snowline inward by a factor of $\sim$7.
 The results of an enhanced gas-phase N/O ratio between the H$_2$O and CO snowlines compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{CN_norm.pdf}
+\includegraphics[width=0.5\textwidth]{CN_norm.eps}
 %\vspace{-0.5in}
-\caption{Gas phase C/O (blue curve) and N/O (red curve) ratios divided by the average C/O and N/O ratio in the disk, assuming CO and N$_2$ are pure ices, and there is no CH$_4$ or NH$_3$.
+\caption{Gas phase C/O (blue curve) and N/O (red curve) ratios divided by the average C/O and N/O ratio in a static disk, assuming CO and N$_2$ are pure ices, and there is no CH$_4$ or NH$_3$.
 The dashed vertical lines mark the H$_2$O, CO$_2$, CO and N$_2$ snowlines.
@@
 A low N$_2$/CO ratio is consistent with comets having formed inside the N$_2$ snowline where N$_2$ is still in the gas phase.
-Theoretical models suggest that Jupiter-family comets, such as 67P, originate from the Kuiper belt \citep{duncan97} (but see \citealt{rubin15} for alternative formation scenarios for 67P).
+Theoretical models suggest that Jupiter-family comets, such as 67P, originate from the Kuiper belt (\citealt{duncan97}; but see \citealt{rubin15} for alternative formation scenarios for 67P).
 It is thus possible, in principle, to use measurements of the N$_2$ abundance in Jupiter-family comets to determine where the N$_2$ snowline was located in our Solar system.
@@
 In this paper we explore the role of icy grain morphology and disk dynamics on the snowline locations of major volatile carrier molecules and the C/N/O ratios in protoplanetary disks.
-We enhance the coupled drift-desorption model developed in \citet{piso15b} by adding more carbon- and nitrogen-bearing species into our framework and by considering different binding ice environments.
+We enhance the coupled drift-desorption model developed in \citet{piso15b} by adding more carbon- and nitrogen-bearing species into our framework, and by considering different binding ice environments.
 Our results can be summarized as follows:
@@
 \item The effect of CH$_4$ and NH$_3$ on the C/O and N/O ratios is small, even when we consider the maximum observed CH$_4$ and NH$_3$ abundances in protostellar cores.
-In this scenario, the gas phase C/O ratio increases by 8\% between the CO$_2$ and CH$_4$ snowlines, and  the gaseous N/O ratio decreases by up to 18\% between the NH$_3$ and CO$_2$ snowlines.
+In this scenario, the gas phase C/O ratio increases by 8\% between the CO$_2$ and CH$_4$ snowlines, and  the gaseous N/O ratio decreases by 18\% between the NH$_3$ and CO$_2$ snowlines.
 In both cases, large gas phase C/O and N/O ratios in the outer disk are preserved.

### 2016-03-11 python/C_to_O_active_steady_time.py

@@ -82,20 +82,20 @@ def n(index, r, elem):
     aCO = arrayCO[index][-1]    
                 
     if r <= aH2O:
-        return np.array([n_C_O('CO')[i] / (1 - (r/aCO)**(3./2)) + \
-                        n_C_O('CO2')[i] / (1 - (r/aCO2)**(3./2)) + \
-                        n_C_O('H2O')[i] / (1 - (r/aH2O)**(3./2)), \
+        return np.array([n_C_O('CO')[i] * (1 - (r/aCO)**(3./2)) + \
+                        n_C_O('CO2')[i] * (1 - (r/aCO2)**(3./2)) + \
+                        n_C_O('H2O')[i] * (1 - (r/aH2O)**(3./2)), \
             #n_C_O('C_grains')[i] + n_C_O('silicate')[i]]), 
                         size])
             
     elif aH2O <= r <= aCO2:
-        return np.array([n_C_O('CO')[i] / (1 - (r/aCO)**(3./2)) + \
-                        n_C_O('CO2')[i] / (1 - (r/aCO2)**(3./2)), \
+        return np.array([n_C_O('CO')[i] * (1 - (r/aCO)**(3./2)) + \
+                        n_C_O('CO2')[i] * (1 - (r/aCO2)**(3./2)), \
             #n_C_O('H2O')[i] + n_C_O('C_grains')[i] + n_C_O('silicate')[i]]), 
                         size])
             
     elif aCO2 < r < aCO:
-        return np.array([n_C_O('CO')[i] / (1 - (r/aCO)**(3./2)), \
+        return np.array([n_C_O('CO')[i] * (1 - (r/aCO)**(3./2)), \
             #n_C_O('C_grains')[i] + n_C_O('silicate')[i] + n_C_O('H2O')[i] + \
                 #n_C_O('CO2')[i]]), 
                         size])

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/appendixA.tex
upd

\chapter{Derivation of the Global Energy Equation}
\label{sec:globalderiv}

%\section{Derivation of the Global Energy Equation}\label{sec:globalderiv}

To derive  the global energy equation (\ref{eq:coolingglobal}) for an embedded protoplanet, we generalize the analogous calculations in stellar structure theory, e.g.\ in \S4.3 of \citet{kippenhahn90}.  For our problem, we add the effects of finite core radius, surface pressure and mass accretion. We start with the local energy equation (\ref{eq:structd}), whose more natural form in Lagrangian (mass) coordinates is $\p L/ \p m = \epsilon - T \p S /\p t$.  Integrating from the core to a higher shell with enclosed mass $M$ gives:
\begin{subeqnarray}
L - L\co &=& \int_{M\co}^M {\p L \over \p m} dm \\
&=& \int_{M\co}^M \left(\epsilon - T {\p S \over \p t} \right)dm \\
&=& \Gamma  - \int_{M\co}^M{\p u \over \p t} dm +  \int_{M\co}^M {P \over \rho^2} {\p \rho \over \p t} dm\slabel{eq:DLc}\, ,
\end{subeqnarray} 
with $\Gamma = \int \epsilon dm$ the integral of the direct heating rate, and applying the first law of thermodynamics in the final step.

The global energy equation is derived by eliminating the partial time derivatives in \Eq{eq:DLc}, which are performed at a fixed mass,
in favor of total time derivatives, denoted with overdots.  %The physical distinction is that total derivatives include mass  accreted through the outer boundary.  
For instance, the surface radius $R$ of the shell with enclosed mass $M$ evolves as  
\begin{equation}\label{eq:Rdot}
 \dot{R} = {\p R \over \p t} + {\dot{M} \over 4 \pi R^2 \rho_M},
\end{equation} 
where $\p R/\p t$ gives the Lagrangian contraction of the ``original" shell, and mass accretion through the upper boundary at rate $\dot{M}$ also changes the shell location.  
%The subscript $M$  denotes quantities at the upper boundary of total mass $M$ (though it is omitted from $M$ and $R$).  
Similarly, the volume $V = (4 \pi/3)R^3$ and pressure at the outer shell evolve as
\begin{subeqnarray}\label{eq:dot}
\dot{V}_M &=&  {\p V_{\rm M} \over \p t} + {\dot{M} \over \rho_{\rm M}}  \\
 \dot{P}_M &=& {\p P_{\rm M} \over \p t} + {\p P_M \over \p m}\dot{M} =  {\p P_{\rm M} \over \p t} - {G M  \over 4 \pi R^4} \dot{M}\, .
\end{subeqnarray} 
This derivation holds the core mass and radius fixed, $\dot{M}\co = \dot{R}\co = 0$.  Therefore, the core pressure satisfies
\begin{equation}\label{eq:Pcdot}
 \dot{P}\co = \p P\co / \p t \, .
\end{equation}
The internal energy integral follows simply from  Leibniz's rule as
\begin{equation}\label{eq:udot}
\int_{M\co}^{M(t)}{\p u \over \p t} dm = \dot{U}  -  \dot{M}u_M\, .
\end{equation} 
To make further progress, we use the virial theorem:
\begin{equation}
\label{eq:virial}
E_G=-3 \int_{M\co}^M \frac{P}{\rho} dm + 4 \pi (R^3 P_M-R\co^3 P\co),
\end{equation}
which follows from \Eqsss{eq:structb}{eq:structa}{eq:Eg} by integrating hydrostatic balance in Lagrangian coordinates.  As an aside, the integral in equation (\ref{eq:virial}) can be evaluated for a polytropic EOS to give simple expressions for the total energy:
\begin{subeqnarray}
E&=&(1-\zeta)U+4 \pi (R^3 P_M-R\co^3 P\co) \slabel{eq:vira} \\
&=&\frac{\zeta-1}{\zeta}E_G+\frac{4 \pi}{\zeta} (R^3 P_M-R\co^3 P\co) \slabel{eq:virb} \, ,
\end{subeqnarray}
where $\zeta \equiv 3(\gamma - 1)$.  We will not make this assumption and will keep the EOS general.

To express the work integral, i.e. the final term in \Eq{eq:DLc}, in terms of changes to gravitational energy, we first take the
 time derivative of \Eq{eq:virial}:
\begin{eqnarray}\label{eq:EGdot}
\dot{E}_G = 3  \int_{M\co}^M {P \over \rho^2} {\p \rho \over \p t} dm -3 \int_{M\co}^M {\p P\over \p t}{dm \over \rho} 
 -  3{P_M \over \rho_M} \dot{M}+ 3 \dot{P}_M V_M -3 \dot{P}\co V\co  + 3  P_M {\dot{ V}_M} \, . 
\end{eqnarray} 
%where the volumes, $V_M = 4 \pi R^3/3$ and $V\co = 4 \pi R\co^3/3$. 
The first integral in \Eq{eq:EGdot} is the one we want, but the second one must be eliminated.  The time derivative of \Eq{eq:Eg} (times four) gives
\begin{subeqnarray}
 4 \dot{E}_G &=&  -4 {G M \dot{M} \over R} + 4 \int_{M\co}^M {G m \over r^2}{\p r \over \p t} dm\\ 
&=&   -4 {G M \dot{M} \over R} + 4 \pi \int_{M\co}^M r^3{\p \over \p m}{\p P \over \p t} dm \slabel{eq:4EGb} \\
&=&  -4 {G M \dot{M} \over R} -3  \int_{M\co}^M {\p P\over \p t}{dm \over \rho}  + 3 V_M {\p P_M \over \p t} -3 V\co {\p P\co \over \p t} \slabel{eq:4EGc} \, ,
\end{subeqnarray} 
where \Eqs{eq:4EGb}{eq:4EGc} use hydrostatic balance  and integration by parts.

%To eliminate the time derivates of pressure, we take the time derivative of the hydrostatic balance equation for $\p^2 P / \p m\p t$ and integrate over $4\pi r^3 dm$ (as in the virial equation derivation) to get
%\begin{equation}\label{eq:dHBdt}
%3 \dot{P}_M V_M -3 \dot{P}\co V\co -3 \int_{M\co}^M {\p P\over \p t}{dm \over \rho}  = 4 \dot{E}_G + 4{G M \over R} \dot{M}  \, .
%\end{equation} 
%Combining \Eqs{eq:EGdot}{eq:dHBdt} gives 
%\begin{eqnarray}\label{eq:rhodot}
%\int_{M\co}^M {P \over \rho^2} {\p \rho \over \p t} dm  &=& - \dot{E}_G - {4 \over 3}{G M\over R} \dot{M} + {P_M \over \rho_M} \dot{M} -  P_M \dot{V}_M  \, , \nonumber \\
%&=&- \dot{E}_G - {4 \over 3}{G M\over R} \dot{M}  -  P_M {\p V_M \over \p t}  \, ,
%\end{eqnarray} 
%where the final step uses \Eq{eq:Rdot}.

Subtracting \Eqs{eq:udot}{eq:4EGc} and rearranging terms with the help of \Eqsss{eq:Rdot}{eq:dot}{eq:Pcdot} gives
\begin{eqnarray}\label{eq:PdVint}
\int_{M\co}^M {P \over \rho^2} {\p \rho \over \p t} dm  &=&  - \dot{E}_G - {G M \dot{M} \over R} - P_M {\p V_M \over \p t} \,  .
\end{eqnarray} 
Combining \Eqsss{eq:DLc}{eq:udot}{eq:PdVint}, we reproduce \Eq{eq:coolingglobal} with the accreted specific energy $e_M \equiv u_M - GM/R$.

%\section{Analytic Cooling Model Details}\label{sec:analytic}
%TODO (academic): Figure out if theta < 1 is actually required? Should be but need to check

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/appendixB.tex
upd

\chapter{Analytic Cooling Model Details}
\label{sec:analytic}
%TODO (academic): Figure out if theta < 1 is actually required? Should be but need to check

\subsection{Isothermal Atmosphere}
\label{iso}

We consider the structure of a non-self-gravitating, isothermal atmosphere that extends outward from the radiative-convective boundary (RCB) and matches onto the disk density, $\rho_{\rm d}$, at a distance $r_{\rm fit} = n_{\rm fit} \RB$, where $R_{\rm B}$ is the Bondi radius defined in equation (\ref{eq:RB}). From equation (\ref{eq:structa}) the resulting density profile is
\begin{equation} \label{eq:rhoiso}
\rho = \rho_{\rm d} \exp \left({R_{\rm B} \over r} - {1 \over n_{\rm{fit}}} \right) \approx   \rho_{\rm{d}} \exp \left(R_{\rm B} \over r  \right),
\end{equation} 
where the approximate inequality is valid deep inside the atmosphere ($r \ll \RB$) for any $n_{\rm fit} \gtrsim 1$.  However, the choice of boundary condition does have an order unity effect on the density near the Bondi radius.

The mass of the atmosphere is determined by integrating equation (\ref{eq:structb}) from the RCB to the Bondi radius using the density profile (\ref{eq:rhoiso}) and can be approximated as
\begin{equation} \label{eq:MatmISO}
M_{\rm iso} \approx 4 \pi \rho\di {R\cb^4 \over R_{\rm B}} e^{R_B/R\cb} = 4\pi \rho\cb \frac{R\cb^4}{R_{\rm B}} \, ,
\end{equation}
with $\rho\cb$ the density at the RCB.
This result is the leading order term in a series expansion. By comparing the expression above and \Eq{eq:Matman} under the assumption that $R\cb \ll R_{\rm B}$, we see that the mass of the outer radiative region (which is nearly isothermal) is negligible when compared with the atmosphere mass in the convective layer, as stated in Section \ref{sec:coolingan}.
% Furthermore, because the atmospheric scale-height at $R\co$ is $H_\rho = |dr /d\ln\rho| = R\co^2/R_B$, the result is intuitively the correct order of magnitude.
%Planets can attract massive atmospheres if $\theta\co \equiv R_B/R\co \gg 1$.

\subsection{Temperature and Pressure Corrections at the Radiative-Convective Boundary}
\label{RCBcorr}

%\subsection{Temperature Contrast at Convective Boundary}
We estimate the temperature and pressure corrections at the RCB due to the fact that the radiative region is not purely isothermal. From equation (\ref{eq:delrad}), we express the radiative lapse rate
\begin{equation}\label{eq:delradan}
\delrad = {3 \kappa P \over 64 \pi  G M \sigma T^4} L = \nabla\di {P/P_{\rm d} \over (T/T_{\rm d})^{4-\beta}},
\end{equation}

\noindent where the second equality follows from the opacity law (\ref{eq:opacitylaw}) and $\nabla_{\rm d}$ is the radiative temperature gradient at the disk:

\begin{equation}
\label{eq:delo}
\nabla \di \equiv \frac{3 \kappa(T{\di}) P{\di}}{64 \pi G M \sigma T_{\rm d}^4} L.
\end{equation}
Here $M$ is the total planet mass. Since our analytic model neglects self-gravity, $M=M\co$ and therefore $\nabla\di$ is constant. From equation (\ref{eq:delradan}) and $\delrad=d \ln T / d\ln P$, the temperature profile in the radiative region integrates to
\begin{equation}\label{eq:radTP}
\left(T \over T_{\rm d}\right)^{4-\beta} - 1 = {\nabla\di \over \nabla_\infty} \left( {P \over P_{\rm d}} - 1 \right) \, ,
\end{equation} 
where $\nabla_\infty = 1/(4-\beta)$ is the radiative temperature gradient for $T ,P \rightarrow \infty$.
Applying \Eqs{eq:delradan}{eq:radTP} at the RCB (where $\delrad = \delad$) under the assumption that $P\cb \gg P_{\rm d}$ results in  $T\cb=\chi T\di$ as in \Eq{eq:Tcb}, with $\chi$ defined in \Eq{eq:chi}.

The pressure at the RCB follows from \Eqs{eq:radTP}{eq:Tcb} as
\begin{equation}
\label{eq:Pcbapprox}
{P\cb\over P_{\rm d}} \simeq {\delad/\nabla\di \over 1 - \delad/\nabla_\infty}.
\end{equation} 
%This pressure contrast can be quite large since $\nabla_{\rm d} \ll 1$.
 We can eliminate $\nabla\di$ from equation (\ref{eq:Pcbapprox}) to obtain a relation between temperature and pressure in the radiative zone as a function of the RCB pressure $P\cb$. From \Eq{eq:radTP}, it follows that
 \begin{equation}\label{eq:TP}
{T \over T_{\rm d}} = \left[1 + {1 \over {\nabla_\infty \over \delad} - 1} \left({P \over P\cb} -  {P_{\rm d} \over P\cb}\right) \right]^{1 \over 4-\beta}\, .
\end{equation} 
 We can then determine the RCB radius $R\cb$ from \Eq{eq:structa} as 
\begin{equation}\label{eq:RCBint}
{R_B \over R\cb} = \int_{P\di}^{P\cb} {T \over T_{\rm d}} {dP \over P}\, .
\end{equation}
Evaluating the integral leads to 
\begin{equation}\label{eq:Rcb}
{R_B \over r\cb} = \ln \left(P\cb \over P\di \right) - \ln \theta \, ,
\end{equation} 
with an extra correction term $\theta < 1$, when compared to an isothermal atmosphere (see Equation \ref{eq:rhoiso}). From this we arrive at the relation between $P\cb$ and $P\di$ given by Equation (\ref{eq:PcbRcb}). As opposed from the temperature correction factor $\chi$, an analytic expression for  $\theta$ cannot be obtained. Estimates for $\chi$ and $\theta$ for different values of the exponent $\beta$ in the opacity law (\ref{eq:opacitylaw}) are presented in Table 1.

%As shown in section \S\ref{iso, an isothermal atmosphere gives a simple logarithmic dependence on $P\cb$.  However, using \Eq{eq:TP} in the integral gives

% The form we chose for the correction term allows us to relate the disk and radiative-convective boundary pressures as :
% \begin{equation}\label{eq:PcbRcb}
%P\cb = \theta P_{\rm d} e^{R_B/R\cb} \, .
%\end{equation}   
%In the $P\cb \gg P_{\rm d}$ limit, the correction term is an order unity constant that depends on $\alpha$, $\beta$ and $\delad$. Similarly to the temperature correction factor $\chi$,  $\theta$ accounts for the fact that the radiative region is not perfectly isothermal. For $\delad=2/7$, $\alpha=0$ and $\beta=2$, we find $\theta \approx 0.556$. Values of $\theta$ for other choices of $\beta$ are summarized in \App{sec:analytic}.  A simple analytic expression for $\theta$ is not possible.

\subsection{The Opacity Effect}
\label{opacityan}
A  lower opacity  decreases the critical core mass.  Reducing the opacity by a factor of one hundred results in a critical core mass one order of magnitude lower than in the standard ISM case, for our analytic model. The reduction is not  as strong as the nominal scaling would imply, $0.01^{3/5} \approx 0.06$, because $\xi$ increases.

Even with significantly lower opacities, radiative diffusion remains a good approximation at the RCB. For $\beta = 2$, we estimate the optical depth as
\begin{equation}
\tau\cb \sim {\kappa\cb P\cb \over g} \sim 7 \times 10^4 {F_T^4 F_\kappa \over \left(\mc \over 10 \right) \left(\au \over 10\right)^{12 \over 7}}, 
\end{equation} 
where $P\cb \sim P_M$ for a self-gravitating atmosphere and $g \sim G M\co/R_B^2$, with both approximations good to within the order unity factor $\xi$.  We see that $\tau\cb \gg 1$ even for $F_\kappa \lesssim 0.01$ out to very wide separations, hence the atmosphere remains optically thick at the RCB.

%A hotter disk would increase core masses.  Instead of our passive disk model, adopting the standard Hayashi temperature profile would increase core masses by $\sim 50\%$.  A hotter accretion phase would further increase core masses, but such phases are presumably short lived.

\subsection{Surface Terms}
\label{surfterms}
In this section we check the relevance of the neglected surface terms in \Eq{eq:coolingglobal}.  We first show that accretion energy is only a small correction at the RCB, which is where we apply our cooling model. A rough comparison (ignoring terms of order $\xi$) of  accretion luminosity vs. $\dot{E}$ gives
\begin{equation}
{G M \dot{M} \over R \dot{E}} = {G M  \over R}{dM \over dE} \sim {G M\co^2 \over R_B E}{P\cb \over P_M} \sim \sqrt{R\co \over R_B} \ll 1,
\end{equation} 
where we assume $P\cb \sim P_M$  for a massive atmosphere.  Accretion energy at the protoplanetary surface is thus very weak for marginally self-gravitating atmospheres, and even weaker for lower mass atmospheres.  A similar scaling analysis shows that the work term $P_M \p V_M/\p t$ is similarly weak.  Nevertheless, our numerical calculations include these surface terms in a more realistic and complete model of self-gravitating atmospheres.

%\subsection{Hydrogen Dissociation}
%The dissociation of molecular hydrogen deep in the atmospheres of accreting protoplanets plays a significant role in the energetics of core accretion.  In the high density regions $r  \ll R\cb$ of a convective atmosphere, the thermal plus gravitational energy scales as
%\begin{equation}
%dE = -4 \pi \nabla_{\rm ad}^{1/\nabla_{\rm ad}} \rho\cb R_B'^{1/(\gamma-1)} r^{\frac{2\gamma - 3}{\gamma - 1}} {dr \over r}
%\end{equation} 
%If $\gamma < 3/2$ then the main contribution to the energy is at the bottom of the atmosphere, i.e.\ the core.  This is the case for a diatomic ideal gas ($\gamma = 1.4$) or a solar mixture of hydrogen and helium ($\gamma \approx 1.43$).  However a monatomic gas has $\gamma = 5/3 > 3/2$.  In this case, the atmosphere's energy is no longer concentrated near the bottom, but will be concentrated near the top of the convective zone.
%
%A likely structure is an atmosphere that is dissociated near the base, but becomes molecular near the top of the convective zone.  In this case the atmosphere's energy budget would be concentrated near the atomic to molecular transition.
%
%The energy required to dissociate a hydrogen molecule, $I = 4.467$ eV can be significant to the overall energy budget.
%Since
%\begin{equation}
%{I \over \nabla_{\rm ad} G M\co (2 m_{\rm H})/r} \approx 3 \left(M\co \over 10 M_\oplus \right)^{-2/3} {r \over R_{\rm c}}
%\end{equation} 
%we see that this energy is always relevant.
%
%We can use the Saha equation to determine where dissociation is significant,
%\begin{equation}
%{n_{\rm H}^2 \over n_{\rm H_2}} = \left(\pi m_{\rm H} k T \over h \right)^{3/2} e^{-I/(kT)}
%\end{equation} 
%We introduce a reaction coordinate $\delta$ so that $n_{\rm H} = 2 \delta n_o$ and $n_{\rm H_2} = (1-\delta) n_o$ with $n_o = \rho X/(2 m_{\rm H})$ the number density when all hydrogen is molecular.  We express equilibrium as
%\begin{equation}
%{\delta^2 \over 1-\delta} = f_\mu {P_{\rm diss}(T) \over P}
%\end{equation} 
%with the characteristic pressure below which dissociation occurs is
%\begin{equation}
%P_{\rm diss} = {\left(kT\right)^{5/2} \over 4} \left( \pi m_{\rm H} \over h^2\right)^{3/2}  e^{-I/(kT)}
%\end{equation} 
%and the order unity factor
%\begin{equation}
%f_\mu = 2\delta + (1-\delta) + Y/2 + Z/\mu_Z
%\end{equation} 
%accounts for variations in the mean molecular weight with dissociation.  (Take $\mu_Z = 31/2$, but not too significant.)
%
%Thus dissociation occurs where $P \lesssim P_{\rm diss}(T)$.  At disk temperatures (say 150 K) the dissociation pressure is negligibly small ($\sim 10^{-141}$ dyne cm$^{-2}$) and no dissociation occurs.  However at core temperatures the dissociation pressure is quite large especially for massive cores.  Dissociation is guaranteed.

\begin{table}{cccccc}  % <--- column justification (center/left/right)
\gdef \numcols {6}
\tablecolumns{\numcols}
\tablecaption{Parameters Describing Structure of Radiative Zone.}
\tablehead{   \multicolumn{\numcols}{c} {$\gamma = 7/5$ ($\delad = 2/7$)} }  
\startdata
 $\beta$   		 &1/2  	& 3/4 &1   		& 3/2  		& 2   \\
 $\nabla_\infty$ & 2/7 \tablenotemark{a}  	&  4/13	& 1/3 	& 2/5 	 	& 1/2 \\
 $\chi$ 		 & \nodata &  2.25245 &1.91293 	& 1.65054 	& 1.52753 \\
 $\theta $  		 &\nodata   & 0.145032	&0.285824   &0.456333   & 0.556069   \\
\enddata
\tablenotetext{a}{Since $\delad = \nabla_\infty$ there is no convective transition at depth for this case.}
\end{table}

%\subsection{Estimate of Atmosphere Mass Outside the Bondi Radius}
%
%Here we consider the mass exterior to the Bondi radius.  For a meaningful evaluation we only include the mass coming from the overdensity relative to the background density.  The resulting external mass for an isothermal atmosphere is
%\begin{subeqnarray}
%M_{\rm ext} &=& 4 \pi \int_{\RB}^{r_{\rm fit}} (\rho - \rho\di) r^2 dr \\
%&=& M\co \theta\co \int_1^{n_{\rm fit}} 3 \left[ \exp \left({1 \over x} - {1 \over n_{\rm fit}}\right) - 1 \right] x^2 dx  \nonumber \\
%&\equiv& M\co \theta\co I(n_{\rm fit})
%\end{subeqnarray} 
%where $\theta_c=R_B/R_{\rm c}$ and the dimensionless integral $I(n_{\rm fit})$ obeys the limits $I(1) = 0$ and $I \rightarrow n_{\rm fit}^2/2$ as $n_{\rm fit} \rightarrow \infty$.  Since this external mass does not converge, the choice of an outer boundary does matter in principle.  In practice, however, the assumption that   $r_{\rm fit} = R_{\rm H}$ limits $n_{\rm fit}$ to modest values
%\begin{equation}
%n_{\rm fit} = {R_{\rm H} \over \RB} \approx 1.3 {\aun{10}^{4/7} \over \mcn{10}^{2/3}}{F_T \over  m_\ast^{1/3}}.
%\end{equation} 
%Since for instance $I(2) = 1.1$, these modest $n_{\rm fit}$ values will only produce a small external mass.

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/appendixC.tex
upd

\section{Equation of State Table Extension}\label{EOStables}

In this study we consider atmosphere growth in the outer parts of protoplanetary disks ($5<a<100$ AU), where temperature and pressure drop to as little as $T \sim 20$ K and $P \sim4\times10^{-6}$ dyn cm$^{-2}$ for our fiducial disk model (see equations \ref{eq:diskb} and \ref{eq:Pd}). We model the nebular gas using the EOS tables of \citet{saumon95}. However, these tables only cover the relatively high temperature and pressure ranges $2.1 < \log_{10} T(\rm{K})<7.06$ and $4.0<\log_{10}P$(dyn cm$^{-2})<19.0$. We thus need to extend the tables to lower $T$ and $P$. We calculate $\delad$ for

\begin{eqnarray}
1.0 & < & \log_{10} T <2.1 \\ 
-5.4& < & \log_{10} P<4.0 
\end{eqnarray} 
using the following method.

%In this section we explain the procedure for extending and interpolating the \cite{saumon95} EOS tables. The EOS takes into account non ideal interactions, and includes physical treatments of dissociation and ionization. However, the \cite{saumon95} EOS tables only cover a relatively high range of temperatures and pressures: $2.10 < \log_{10} T(\rm{K})<7.06$ and $4<\log_{10}P$(dyn cm$^{-2})<19$. We consider cold disks, where the temperature and pressure drop to $\sim 20$ K and $\sim 10^{-4}$ dyn cm$^{-2}$, respectively (see equations (\ref{eq:diskb}) and (\ref{eq:Pd})). It is therefore necessary to extend the \cite{saumon95} EOS tables to lower temperature and pressure values.

%We choose $\log_{10} T (\rm{K})=1$ and $ \log_{10}P$(dyn cm$^{-2})=-4.4$ as our lower boundaries for temperature and pressure, respectively. Our temperature and pressure grid becomes: $1 < \log_{10} T(\rm{K})<7.06$ and $-4.4<\log_{10}P$(dyn cm$^{-2})<19$. The other thermodynamic variables in the tables are calculated as follows.

\subsection{Hydrogen}

\label{hydrogen}

Following \citet{kittel}, we calculate $\delad$ from the partition function for the internal energy of a system of hydrogen gas molecules (see also \citealt{dangelo13} for EOS calculations that take into account hydrogen isomers).  We begin by writing the partition function $Z$ of a gas molecule of mass $m$ as the product of the partition functions associated with each type of internal energy:

%\begin{equation}
%\label{eq:z}
%Z=Z_t Z_r Z_v Z_e Z_n,
%\end{equation}

\begin{equation}
\label{eq:zagain}
Z=Z_t Z_r Z_v \;\;,
\end{equation}

\noindent where $Z_t$, $Z_r$, $Z_v$ are associated with translation, rotation, and vibration, respectively.\footnote{We ignore electronic and nuclear excitation as they are only important at temperatures much higher than our regime of interest.}  %For hydrogen, electronic and nuclear excitation are only significant at temperatures higher than our region of interest ($\theta_e \approx 12000$ K and $\theta_n >> \theta_e$, where $\theta_e$ and $\theta_n$ are the characteristic temperatures for electronic and nuclear excitation, respectively). As such, we will only take into account the translation, rotation and vibration of the hydrogen molecule:

%In what follows we present and briefly derive expressions for thermodynamic variables based on quantum mechanics principles. More details on the derivations can be found in \citet{kittel}.

In the classical limit, the molecule's center of mass motion generates

\begin{equation}
\label{eq:Zt}
Z_t=(m k_B T/2 \pi \hbar^2)^{3/2} V,
\end{equation}
where  $T$ and $V$ are the gas temperature and volume, respectively, and $\hbar$ is the reduced Planck constant. The rotational partition function is

\begin{equation}
\label{eq:Zr}
Z_r=\sum_{j=0}^\infty (2 j+1) \exp{\Big[\frac{-j (j+1)\Theta_r}{T}\Big]},
\end{equation}

\noindent where the characteristic temperature for rotational motion $\Theta_r \approx 85$ K  for hydrogen. However, molecular hydrogen occurs in two isomeric forms: parahydrogen with a symmetric (even) rotational wavefunction, and orthohydrogen with an antisymmetric (odd) wavefunction (see Section \ref{deladtable}). %Parahydrogen with an even rotational wave function, while orthohydrogen can only have an antisymmetric (odd) wave function. 
The rotational partition functions for ortho- and parahydrogen are thus
\begin{equation}
\label{eq:Zpara}
Z_{\rm{r,para}}=\sum_{j=0}^\infty \frac{1+(-1)^j}{2} (2 j +1) \exp\Big[-\frac{j(j+1)\Theta_r}{T}\Big]
\end{equation}
and
\begin{equation}
\label{eq:Zortho}
Z_{\rm{r,ortho}}=3\sum_{j=0}^\infty \frac{1-(-1)^j}{2} (2 j +1) \exp\Big[-\frac{j(j+1)\Theta_r}{T}\Big] \;\;.
\end{equation}
The factor of 3 in Equation (\ref{eq:Zortho}) accounts for the three-fold degeneracy of the ortho state.

 In thermal equilibrium, the spin isomers have a combined partition function $Z_{\rm r}=Z_{\rm{r, ortho}}+Z_{\rm{r,para}}$, which can be written

\begin{equation}
\label{eq:Zrspin}
Z_r=\sum_{j=0}^\infty (2-(-1)^j) (2j+1) \exp{\Big[\frac{-j (j+1) \Theta_r}{T}\Big]} \;\;.
\end{equation}
For a fixed 3:1 ortho-to-para ratio, the combined partition function is $Z_{\rm r}=Z_{\rm r,para}^{1/4} Z_{\rm r, ortho}^{3/4}$. In our range of temperatures of interest, we find that $Z_r$ converges after about 25 terms in the series, for both the equilibrium and 3:1 cases.

Note that $Z_{\rm r, ortho} \rightarrow 0$ and $Z_{\rm r,para} \rightarrow 1$ as $T \rightarrow 0$. This is inconsistent with a fixed 3:1 ortho-to-para ratio, since $Z_{\rm r, ortho} \rightarrow 0$ implies that there is no orthohydrogen in the system. \citet{boley07} and \citet{dangelo13} ensure that this requirement is not violated by using a normalized orthohydrogen partition function,  $Z'_{\rm r, ortho}=Z_{\rm r, ortho} \exp(2 \theta_{\rm r}/T)$, which reduces the energy of the lowest rotational state for orthohydrogen from $2k_{\rm B}\theta_{\rm r}$ per molecule to zero. This decreases the total internal energy of the system by a constant factor, but does not change $c_{\rm v}$, $\delad$, or the relative internal energies of static atmospheric profiles, and therefore does not affect atmospheric evolution.

Finally, the partition function for vibrational motion is given by:

\begin{equation}
\label{eq:Zv}
Z_v=[1-\exp{(\theta_v/T)}]^{-1},
\end{equation}

\noindent where the characteristic temperature for vibrational motion is $\theta_v  \approx 6140$ K for hydrogen.

For a system of $N$ particles of mass $m$, the partition function of the ensemble is $Z_{\rm N}=(1/N!)Z^N$. Given $Z_{\rm N}$ as a function of $(V, T)$, the internal energy per mass, entropy per mass and specific heat capacity can be written as

%\begin{equation}
%\label{eq:U}
%U_N=k T^2 \Big(\frac{\partial \log{Z}}{\partial T}\Big)_{V, N}
%\end{equation}
%
%\begin{equation}
%\label{eq:S}
%S_N=k \log{Z} + \frac{U_N}{T}
%\end{equation}
%
%The energy, and entropy per mass and specific heat capacity will subsequently be:

\begin{equation}
\label{eq:u}
u=\mathcal{R}T^2 \Big(\frac{\partial \ln{Z}}{\partial T}\Big)_{V}
\end{equation}
%{\bf CHANGED BIG U TO LITTLE u IN EQUATIONS BELOW: CONFIRM}
\begin{equation}
\label{eq:s}
S=\mathcal{R} \ln{Z} + \frac{u}{T}-\frac{\mathcal{R}}{N} \ln N!
\end{equation}
\begin{equation}
\label{eq:cv}
c_{\rm V}=\Big(\frac{\partial u}{\partial T}\Big)_{V}.
\end{equation}
Note that, following the convention of \citet{saumon95}, we use $S$ to denote entropy per mass ($[S]=$erg K$^{-1}$ g$^{-1}$).  Since $Z=Z_t Z_r Z_v$, we may write $u=u_t+u_r+u_v$ and $S=S_t+S_r+S_v$, where variables subscripted $t$, $r$, and $v$ are the quantities corresponding to the individual translation, rotation and vibration partition functions, respectively. We include the term $\mathcal{R}/N \ln N!$ from Equation (\ref{eq:s}) in $S_{\rm t}$.

%\textbf{Using the translational partition function, internal energy and entropy expressions for an ensemble of $N$ identical particles of mass $m$, i.e. $Z_N=(1/N!) Z_t^N$, $U_N=k_B T^2 (\partial \ln{Z_N}/\partial T)_{V, N}$, and $S_N=k_B \ln{Z_N} + U_N/T$, Stirling's approximation $\ln N! \approx N \ln N-N,$ and the ideal gas law $P=\rho \mathcal{R} T$, the {\bf resulting} entropy per mass due to translational motion can be expressed as:}

 In the temperature regime for which the rotational states of $\rm{H}_2$ are selectively occupied, the total number of particles in the system is constant and we may use the ideal gas law, $P=\rho \mathcal{R} T$. With Equations (\ref{eq:u}), (\ref{eq:s}), and Stirling's approximation, $\ln N! \approx N \ln N-N$, the resulting entropy per mass due to translational motion is

\begin{equation}
\label{eq:st}
S_{\rm t}=\mathcal{R} \Big[ \frac{5}{2} \ln{T} - \ln{P} + \ln \Big( \frac{(2 \pi)^{3/2} \mathcal{R}^{5/2} m^4}{h^3}\Big) +\frac{5}{2} \Big]
\end{equation}
Equation (\ref{eq:st}) is known as the Sackur-Tetrode formula. % \textbf{and is only applicable to an ideal gas REALLY? WE'RE USING IT FOR THE TRANSLATIONAL COMPONENT IN A NON-IDEAL GAS}. 
%\textbf{We can use the ideal gas law to derive Equation (\ref{eq:st}) for the $\rm{H}_2$ molecule because the total number of particles in the system, and hence the mean molecular weight and specific gas constant $\mathcal{R}$, do not change with temperature in the regime where the rotational states are selectively occupied.}

The internal energy per mass due to translational motion is given by:

\begin{equation}
\label{eq:ut}
u_t=\frac{3}{2} \mathcal{R} T
\end{equation}

Putting all of the above together, we can now evaluate the thermodynamic quantities needed to extend the \cite{saumon95} EOS tables to low temperatures and pressures.

\begin{enumerate}

\item{\textbf{Density.}} In the low temperature, low pressure regime, $\rho$ is related to $T$ and $P$ by the ideal gas law. %hydrogen is molecular and behaves like an ideal gas. As such, the density in this region follows the ideal gas law $P=\rho \mathcal{R} T$.
\item{\textbf{Internal energy per mass.}} $u=u_t+u_r+u_v$, where $u_t$ is given by Equation (\ref{eq:ut}), and $u_r$ and $u_v$ are determined using Equations (\ref{eq:Zrspin}), (\ref{eq:Zv}) and (\ref{eq:u}) above.
\item{\textbf{Entropy per unit mass}}. Similarly, $S=S_t+S_r+S_v$, where $S_t$ is given by Equation (\ref{eq:st}), and $S_r$ and $S_v$ can be determined from Equation (\ref{eq:s}) and the calculated expressions for $u_r$ and $u_v$, respectively.
\item{\textbf{Entropy logarithmic derivatives}}. The logarithmic derivatives $S_T$ and $S_P$ are given by:

\begin{equation}
\label{eq:sT}
S_T=\frac{\partial \ln{S}}{\partial \ln{T}} \Big |_P
\end{equation}
and
\begin{equation}
\label{eq:sP}
S_P=\frac{\partial \ln{S}}{\partial \ln{P}} \Big |_T
\end{equation}
We calculate $S_T$ and $S_P$ through finite differencing.

\item{\textbf{Adiabatic gradient $\delad$}}. The adiabatic gradient is defined as:

\begin{equation}
\label{eq:deladSP}
\delad=\frac{\partial \ln{T}}{\partial \ln{P}} \Big |_S = -\frac{S_P}{S_T}
\end{equation}

We evaluate $\delad$ from the tabulated values for $S_T$ and $S_P$ determined above. Figure \ref{fig:deladH} shows a contour plot of $\delad$ as a function of temperature and pressure for the extended EOS table for hydrogen, assuming thermal equilibrium between the spin isomers. For the 3:1 ortho-to-para ratio, $\delad$ decreases continuously with $T$ for $T \lesssim 200$ K, in contrast with the equilibrium case, in which $\delad$ sharply decreases, then increases as $T$ goes down. Our extension is only valid for $T \lesssim 2000$ K, since it does not take into account hydrogen dissociation. We choose $T=1500$ K as a conservative temperature cutoff. While we account for vibrational motion for completeness, its contribution is negligible in the temperature regime of interest. \citet{saumon95} do not compute the EOS at very high pressures, since hydrogen is solid or may form a Coulomb lattice in this regime, and thus their EOS treatment is no longer valid. While the boundaries of the region in which the free-energy EOS treatment fails can be determined from fundamental thermodynamic constraints, such calculations are not the object of this work. Instead, we choose as boundary a constant entropy curve ($\log(S)=8.4$) above the region in which the \citet{saumon95} model fails. The expressions derived above are sufficient to give good results for the colored regions of the extended map, which fully cover the temperature and pressures ranges required by our models.

 %Our extension smoothly matches the original tables for $8.80<\log{S}$(K g$^{-1})<9.07$ (\textbf{numbers are wrong, change once you have the final figure version}).

\end{enumerate}

\begin{figure}[h!]
\centering
%\includegraphics[scale=.8]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/delad_S_H.pdf}
\includegraphics[scale=.8]{figures/delad_S_H.pdf}
\caption{Contour plot of the hydrogen adiabatic gradient $\delad$ as a function of gas temperature and pressure. The upper right rectangle encloses the region described by the original \citet{saumon95} EOS tables, while the rest of the plot is our extension to lower temperatures and pressures for an equilibrium mixture of ortho- and parahydrogen. The black curves represent constant entropy adiabats with labels $\log_{10}(S)$, where $S$ [erg K$^{-1}$ g$^{-1}$] is the absolute entropy per unit mass.  At high temperatures, hydrogen dissociates and ionizes, while at low temperatures the rotational states of the hydrogen molecule are only partially excited and it no longer behaves like an ideal diatomic gas. Regions in which the EOS is invalid or has not been computed are masked in white (see text).}
\label{fig:deladH}
\end{figure}

\subsection{Helium}

We extend the helium EOS tables based on a similar procedure. Since helium is primarily neutral and atomic at low temperatures and pressures, we treat it as an ideal monoatomic gas and  only take into account the translational component of the partition function (\ref{eq:Zt}). Figure \ref{fig:deladHe} shows $\delad$ as a function of temperature and pressure for the extended EOS table. %The original and extended table join smoothly for entropy curves between $8.29<\log{S}$(K g$^{-1})<8.77$ in this case (\textbf{numbers are wrong, change once you have the final figure version}).

\begin{figure}[h!]
\centering
%\includegraphics[scale=.8]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/delad_S_He.pdf}
\includegraphics[scale=.8]{figures/delad_S_He.pdf}
\caption{Same as Figure \ref{fig:deladH} but for pure helium. Helium ionizes at $T \gtrsim 10,000$ K, but behaves as an ideal monatomic gas otherwise. We choose $T=7,000$ K as a conservative temperature cutoff above which our  extension is no longer valid (masked in white). The EOS has not been computed in the lower-right region of the plot (see text).}
\label{fig:deladHe}
\end{figure}

\vspace{0.2in}

Lastly, we combine Figures \ref{fig:deladH} and \ref{fig:deladHe} to obtain EOS tables for a hydrogen-helium mixture using the procedure described in \citet{saumon95}.  Figure \ref{fig:deladmap} displays results for helium mass fraction $Y=0.3$.

%Using equation (\ref{eq:upartition}) we therefore recover the standard result $U_{\rm r}=\mathcal{R} T$ (refs). Furthermore, we know that the internal energy and entropy per unit mass associated with translation are given by $U_{\rm t}=\frac{3}{2} \mathcal{R}$ and $C_{\rm{v,t}}=\frac{3}{2}\mathcal{R}$, respectively, and so we are able to calculate the total internal energy and specific heat of a diatomic molecule as a function of temperature. An example of the variation of heat capacity with temperature is shown in \citet{kittel}, chapter 3.

%The partition function associated with rotation is generally written as:
%
%\begin{equation}
%\label{eq:Zr}
%Z_{\rm r}=\sum_0^\infty (2 j +1) \exp\Big[-\frac{j(j+1)\Theta_r}{T}\Big],
%\end{equation}
%with $j$ the angular momentum quantum number \citep{kittel}. Various thermodynamic quantities can be derived from the partition function. For example, the internal energy per unit mass can be written as:
%
%\begin{equation}
%\label{eq:upartition}
%u_{\rm r}=\mathcal{R}T^2 \frac{\partial \log Z}{\partial T}
%\end{equation}
%
%The specific heat at constant volume then easily follows as
%
%\begin{equation}
%\label{eq:cvpartition}
%c_{\rm{v,r}}=\Big(\frac{\partial u}{\partial T}\Big)_V
%\end{equation}

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/appendixD.tex
upd

\section{Adiabatic Gradient Variations}
\label{alldelad}

\subsection{Adiabatic Gradient during Partial Dissociation}\label{deladdiss}

%The adiabatic gradient during dissociation or ionization is a function of temperature and the dissociation or ionization fraction $x$. The Saha equation (e.g., \citealt{kippenhahn90}) relates $x$ to the gas

The total internal energy of a partially dissociated gas includes contributions from the individual internal energies of the molecules and atoms, as well as from the dissociation energy. The dissociation energy depends on the dissociation fraction $x$ (i.e., the fraction of molecules that have dissociated), which can be found from the Saha equation (see e.g., \citealt{kippenhahn90}, Chapter 14) as a function of temperature and density,

\begin{equation}
\label{eq:saha}
\frac{x^2}{1-x} \propto \frac{T^{3/2}}{\rho} e^{-\chi/k_B T},
\end{equation} 
where $\chi$=4.48 eV is the dissociation energy for molecular hydrogen \citep{blanksby03}.

The above also holds true for ionization, with the dissociation energy replaced by ionization energy $\chi=13.6$ eV for atomic hydrogen \citep{mandl89}. From the Saha equation one can find an expression for $\rho$ as a function of $T$ and $x$, then derive the adiabatic gradient directly from its definition (Equation \ref{eq:delad}), taking into account the fact that the mean molecular weight in the ideal gas law varies with $x$, hence the pressure will not only be a function of $T$ and $\rho$ but also of $x$ (see \citealt{kippenhahn90}, Chapter 14.3 for a detailed derivation). The final expression for the adiabatic gradient during ionization is

\begin{equation}
\label{eq:deladioniz}
\delad=\frac{2+x (1-x) \Phi_H}{5+x(1-x)\Phi_H^2},
\end{equation}
with $\Phi_H \equiv \frac{5}{2}+\frac{\chi}{k_B T}$. The derivation of $\delad$ during dissociation is more involved mathematically (see, e.g., \citealt{vardya60}) and leads to a slightly more complicated final expression,

%One can derive a simple expression for the adiabatic gradient as a function of $x$ and $T$ from it's definition (equation \ref{eq:delad}), using the

%The adiabatic gradient follows as (see \citealt{vardya60} for a detailed derivation)

\begin{equation}
\label{eq:deladdiss}
\delad=\frac{1+x+ \frac{x(1-x^2)}{2} \frac{\chi}{k_B T}}{5 x + \frac{7(1-x)}{2} + \frac{x(1-x^2)}{2} \Big(\frac{\chi}{k_B T}\Big)^2} \, .
\end{equation} 
Using Equation (\ref{eq:deladdiss}), we recover $\delad=2/7$ for $x=0$ (no ongoing dissociation hence hydrogen is purely molecular and diatomic) and $\delad=2/5$ for $x=1$ (hydrogen is fully dissociated into atoms and hence monatomic). Figure \ref{fig:deladdiss} shows the dependence of $\delad$ on the dissociation fraction, for $T=3000$ K, the temperature at which dissociation typically occurs \citep{langmuir12}. The adiabatic gradient drops substantially during partial dissociation, since part of the internal energy is used in dissociation rather than in increasing the temperature of the system.

\begin{figure}[h]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/delad_dissociation.pdf}
\includegraphics[width=0.5\textwidth]{figures/delad_dissociation.pdf}
%%\vspace{-0.5in}
\caption{Adiabatic gradient as a function of the hydrogen dissociation fraction $x$. The adiabatic gradient is $\delad=2/7$ for pure molecular hydrogen ($x=0$) and $\delad=2/5$ for fully atomic hydrogen ($x=1$), and drops to low values during partial dissociation.}
\label{fig:deladdiss}
\end{figure}

\subsection{Adiabatic Gradient during Conversion of Spin Isomers}
\label{deladspin}

The adiabatic gradient scales as $\delad \sim 1/c_{\rm V}$.
%  where $c_{\rm V} \sim c_{\rm V,t}+c_{\rm V,r}$.
% is the specific heat capacity at constant volume and $c_{\rm V, t}$ and $c_{\rm V,r}$ are the translational and rotational components, respectively.  
The translational component of the heat capacity, $c_{\rm V, t}=3\mathcal{R}/2$, is independent of temperature.  As $c_{\rm V,r}$ increases, $\delad$ declines. We can therefore understand how conversion between spin isomers affects $\delad$ by studying the dependence on temperature of $c_{\rm V, r}$ of the ortho-para mixture.

The internal energy per unit mass and specific heat capacity associated with rotation for the individual isomers, for the equilibrium mixture, and for a fixed ortho-para ratio of 3:1 can be derived from their partition functions (see \App{EOStables} for details), and are plotted in Figure  \ref{fig:ucvr} (after \citealt{farkas35}, Figure 1). At low temperatures, parahydrogen is in the $j=0$ state and has no rotational energy, while orthohydrogen is in the $j=1$ state and has the energy of its first rotational level. Both para- and orthohydrogen, as well as their equilibrium mixture, behave like monatomic gases at low temperatures and thus have zero rotational heat capacity. This is consistent with $\delad=2/5$ at low temperatures as seen in Figure \ref{fig:deladmap}. As the temperature increases, the energetically higher-lying rotational states of para- and orthohydrogen are populated and the heat capacity of both spin isomers increases as a result. We note that the heat capacity of the equilibrium mixture is not a weighted average of the heat capacities of the individual components because it takes into account both the rotational energy uptake of para- and orthohydrogen, and also the shift in their equilibrium concentrations with temperature. This results in a peak in the heat capacity of the mixture around $\sim$$50$ K, as seen in the bottom plot of Figure \ref{fig:ucvr}. It follows that the adiabatic gradient has to reduce, reach a minimum, then increase as the temperature rises, as shown in Figure \ref{fig:deladmap}. In contrast, the heat capacity for the 3:1 ortho-para ratio mixture will be a weighted average between the individual ortho- and para- components, and will hence have intermediate values between the two, as displayed in Figure \ref{fig:ucvr}.

Figure \ref{fig:Lt_31}, bottom panel, shows that the atmospheric growth time may increase by a factor of $\sim$$3$ if a fixed 3:1 ortho-para ratio is assumed instead of thermal equilibrium between the hydrogen spin isomers.  This enhanced growth time  increases $M_{\rm crit}$.

% As the temperature increases, the energetically higher-lying ($j=1$) ortho-hydrogen is formed, and the concomitant energy increase is seen as a peak in the heat capacity of the equilibrium mixture.

%There are two significant maxima in the heat capacities of parahydrogen and of the mixture. At very low temperatures, the heat capacity of parahydrogen is zero because only the lowest accessible energy level $j=0$ is occupied and a temperature increase does not provide enough energy to populate the next higher level. When the temperature becomes sufficiently high to populate the second lowest level $j=2$, the heat capacity rapidly increases, passes through a maximum and starts to decrease when the second lowest level becomes saturated.

%the adiabatic gradient is inversely proportional to the heat capacity, it

%means that  the former has to first decrease from 2/5 as the temperature increases, reach a minimum around 50 K ($\delad \approx 0.25$ from Fig. \ref{fig:deladmap}), then gradually increase to 2/7 as for a diatomic gas. This behavior is illustrated in Figure \ref{fig:deladmap}.

%The maxima in the ortho-para mixture appears when parahydrogen starts converting into orthohydrogen. The heat capacity of the equilibrium mixture is not a weighted average of the heat capacities of the individual components because it takes into account both the rotational energy uptake of para- and orthohydrogen, and also the shift in their equilibrium concentrations with temperature. At $T=0$, only parahydrogen is present in the equilibrium mixture; as the temperature is increased, the energetically higher-lying ($j=1$) ortho-hydrogen is formed, and the concomitant energy increase is seen as a peak in the heat capacity. As the adiabatic gradient is inversely proportional to the heat capacity, it means that  the former has to first decrease from 2/5 as the temperature increases, reach a minimum around 50 K ($\delad \approx 0.25$ from Fig. \ref{fig:deladmap}), then gradually increase to 2/7 as for a diatomic gas. This behavior is illustrated in Figure \ref{fig:deladmap}.

\begin{figure}[h]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/ortho_para_energy.pdf}
\includegraphics[width=0.5\textwidth]{figures/ortho_para_energy.pdf}
%%\vspace{-0.5in}
\caption{Internal energy per unit mass and specific heat capacity associated with rotation for parahydrogen (dashed blue), orthohydrogen (dotted green), the equilibrium mixture (solid red) and a fixed 3:1 ortho-to-para ratio (dash-dotted light blue) as a function of temperature. After \citet{farkas35}, Figure 1.}
\label{fig:ucvr}
\end{figure}

\begin{figure}[h]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/Ltplot_31.pdf}
\includegraphics[width=0.5\textwidth]{figures/Ltplot_31.pdf}
%%\vspace{-0.5in}
\caption{Evolution of the luminosity and elapsed time during atmospheric growth around a $8 M_{\oplus}$ core at 100 AU, for a realistic EOS with hydrogen spin isomers in thermal equilibrium (solid line), and with a fixed ortho-to-para ratio 3:1 (dashed line). The assumption of a fixed ortho-to-para ratio increases the runaway accretion time $t_{\rm run}$ by a factor of $\sim$$3$ compared to the equilibrium mixture.}o
\label{fig:Lt_31}
\end{figure}

%with $\chi$ and $T$ the dissociation energy and temperature, respectively, equal to $\chi=4.48$ eV\citep{blanksby03} and $T\sim3000$ K \citep{langmuir12} for molecular hydrogen.

%Specifically, if we denote the internal energies of neutral hydrogen, protons and electrons as $U_{H}$, $U_+$ and $U_e$, respectively, then the total internal energy of the gas is given by:

%\begin{equation}
%U=U_H+U_+ + U_e + x \chi,
%\end{equation}
%
%\noindent where $x$ is the ionization fraction and $\chi$ is the ionization energy (equal to -13.6 eV for hydrogen). The ionization fraction can be determined from the Saha equation (see e.g., \citealt{kippenhahn90}).
%
%\begin{equation}
%\label{eq:saha}
%\frac{x^2}{1-x} \frac{\rho}{m_H}=\frac{(2 \pi m_e k_B T)^{3/2}}{h^3} e^{-\chi/k_B T},
%\end{equation}
%
%\noindent where $m_e$ is the mass of the electron and $h$ is Planck's constant. It can be seen from the Saha equation that the ionization fraction depends only on the gas temperature and density: $x=x(T, \rho)$. As such, all the thermodynamic quantities also depend only on the gas temperature and density, and hence on the equation of state. The adiabatic gradient is given by (see \citealt{kippenhahn90}, chapter 14 for a derivation):
%
%\begin{equation}
%\delad=\frac{2+x (1-x) \Phi_H}{5+x (1-x) \Phi_H^2},
%\end{equation} 
%with $\Phi_H=\frac{5}{2}+\frac{\chi}{k T}$. Figure \ref{fig:deladion} shows the behavior of $\delad$ for partially ionized hydrogen. We recover $\delad=2/5$ for $x=0$ (pure atomic hydrogen) and $x=1$ (fully ionized plasma). The adiabatic gradient decreases significantly for intermediate values of $x$, becoming smaller than 0.1 at its minimum (for $x=0.5$).

%\section{Location of radiative-convective boundary for polytropes with varying $\delad$} \label{vardelad}

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/appendixE.tex
upd

\section{Grain growth opacity and radiative windows} \label{radwindow}

The opacity of the interstellar medium is reasonably well constrained and approximate analytic expressions for the Rosseland mean opacity as a function of temperature and density are derived in \citet{bell94}. For low temperatures ($T \lesssim 100$ K) at which ice grains are present, opacity scales with temperature as $\kappa \sim T^2$. Sublimation of ice grains at $\sim$$150$ K and of metal grains at $\sim$$1000$ K results in sharp opacity drops. This is shown in Figure \ref{fig:opacity} for a gas density $\rho=10^{-8}$ g cm$^{-3}$, which is typical for the outer regions of protoplanetary disks. \citet{semenov03} calculate Rosseland mean opacities in protoplanetary disks for grains of different sizes and structure. As shown in Figure \ref{fig:opacity}, their results are in good agreement with \citet{bell94}. However, \citet{semenov03} do not take grain growth into account, which is likely to occur in protoplanetary disks, particularly at the late times when cores form. \citet{dalessio01} compute wavelength dependent opacities for a range of maximum particle sizes and different size distributions. Figure \ref{fig:opacity} shows the integrated Rosseland mean opacity for a maximum particle size of 1 cm and a power law differential size distribution $dN/ds \propto s^{-p}$, with $s$ the grain size and $p=3.5$ for a standard collisional cascade and $p=2.5$ when coagulation is taken into account. We see in Figure \ref{fig:opacity} that this yields a mean opacity that is both lower and less sensitive to temperature, when compared to \citet{bell94} or \citet{semenov03}. However, \citet{dalessio01} only computes opacities for temperatures less than the dust sublimation temperature, appropriate for current observations of dust in protoplanetary disks. As we see in Figure \ref{fig:opacity}, the opacity dramatically decreases during dust sublimation. We thus use the \citet{bell94} opacities for $T \gtrsim 1000$ K, ensuring that they smoothly match the \citet{dalessio01} opacities for lower temperatures.

\begin{figure}[h!]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/kappa_grain_growth_SBL_paper.pdf}
\includegraphics[width=0.5\textwidth]{figures/kappa_grain_growth_SBL_paper.pdf}
%%\vspace{-0.5in}
\caption{Rosseland mean opacity of dust grains as a function of temperature for different opacity assumptions. The dashed black curve shows the \citet{bell94} analytic ISM opacity for $\rho=10^{-8}$ g cm$^{-3}$. The solid black curve shows the tabulated opacity of \citet{semenov03} for a dust composition of 'normal' silicates. The dashed red curve shows the \citet{dalessio01} opacity, which takes grain growth into account, for a maximum particle size of 1 cm and a standard collisional cascade grain size distribution ($p=3.5$). The solid red curve is the same as the dashed red curve, but it accounts for coagulation ($p=2.5$).}
\label{fig:opacity}
\end{figure}

The significant opacity drop due to the sublimation of ice and metal grains lowers the radiative temperature gradient $\delrad$, which may result in one or more inner radiative layers inside the atmosphere of a protoplanet. This is displayed in Figure \ref{fig:delvsr}: depending on the semimajor axis and core mass, the opacity drop will generate no radiative window (top panel), one radiative window (middle panel), or two radiative windows (bottom panel).

\begin{figure}[h!]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/del_vs_r.pdf}
\includegraphics[width=0.5\textwidth]{figures/del_vs_r.pdf}
%%\vspace{-0.5in}
\caption{Snapshots of the radiative and adiabatic gradient as a function of the radial coordinate, for planets with different core masses forming at various locations in the disk. The nebular gas is described by a realistic EOS, with a standard collisional cascade size distribution ($p=3.5$). The sharp drop in opacity due to dust sublimation may generate one or more radiative windows. Top panel: no radiative window for $a=100$ AU and $M\co=3 M_{\oplus}$. Middle panel: the sharp opacity decrease produces one radiative window for $a=50$ AU and $M\co=5 M_{\oplus}$. Bottom panel: the decrease in opacity results in two radiative windows for $a=20$ AU and $M\co=3 M_{\oplus}$.}
\label{fig:delvsr}
\end{figure}

%This is not a problem, however, if most atmospheric luminosity is generated in the innermost convective region of the envelope (see \S\ref{sec:opacity}). We can check this \textit{a posteriori} by using the local energy equation,
%\begin{equation}
%\label{eq:localen}
%\frac{\partial L}{\partial m}=-T \frac{\partial S}{\partial t},
%\end{equation}
%assuming that $\partial S/\partial t$ is constant for a given atmospheric profile. Our model is then valid if $\Delta I \ll I$, where
%\begin{equation}
%\label{eq:I}
%I = \int_{M_{\rm c}}^{M_{\rm{RCB_1}}} T d m
%\end{equation}
%and
%\begin{equation}
%\label{eq:I}
%\Delta I = \int_{M_{\rm{RCB_1}}}^{M_{\rm p}}T d m,
%\end{equation}
%with $\rm{RCB_1}$ the innermost RCB and $M_{\rm p}$ the total planet mass. We find $\Delta I /I \lesssim 30\%$ for all our models of interest (see, however, \S\ref{sec:opacity} and \S\ref{critical} for exceptions).

%\textit{Radiative windows could, in principle, change atmospheric structure and thus affect evolution if most luminosity is not generated in the innermost convective region of the envelope. }

{Our idealization of a constant $L$ with radius may be challenged by the presence of radiative windows. While the structure of convective regions is unaffected by $L$, the structure of radiative windows depends on $L$. Fortunately, the assumption of constant luminosity remains reasonable if most of the luminosity is generated in the innermost convective region of the envelope. We can check whether this is true \textit{a posteriori} by using the local energy equation,
\begin{equation}
\label{eq:localen}
\frac{\partial L}{\partial m}=-T \frac{\partial S}{\partial t},
\end{equation}
and integrating it between $M\co$ and $M_{\rm{RCB_1}}$, where $\rm{RCB_1}$ is the innermost RCB. This luminosity can be as little as half of the assumed fixed atmospheric luminosity $L$.

Self-consistently calculating $L(r)$ is not feasible for our code. Instead, we note that since  $\partial S/\partial t$ is fixed in convective regions \citep{arras06}, the luminosity profile is more centrally concentrated than $L \propto m$, which can be implemented simply.  We calculated example profiles using $L \propto m$ and found that though this luminosity scaling may move the location of the outermost RCB, it does not substantially change the luminosity emerging at the top of the atmosphere or the time evolution. %This result may be understood by noting that small changes in the relative normalization of $\delad$ and $\delrad$ in Figure \ref{fig:delvsr} can change the location of the RCB without affecting structure significantly.

%We therefore expect our results to approximate those for the true $L(r)$.

%assuming that $\partial S/\partial t$ is constant for a given atmospheric profile. Our model is then valid if $\Delta I \ll I$, where
%\begin{equation}
%\label{eq:I}
%I = \int_{M_{\rm c}}^{M_{\rm{RCB_1}}} T d m
%\end{equation}
%and
%\begin{equation}
%\label{eq:I}
%\Delta I = \int_{M_{\rm{RCB_1}}}^{M_{\rm p}}T d m,
%\end{equation}
%with $\rm{RCB_1}$ the innermost RCB and $M_{\rm p}$ the total planet mass. We find $\Delta I /I \lesssim 30\%$ for all our models of interest (see, however, \S\ref{sec:opacity} and \S\ref{critical} for exceptions).

%If radiative windows exist, our model requires most of the atmospheric luminosity to be generated in the innermost convective region. By design, the luminosity in the radiative windows, $L_{\rm radw}$, must satisfy $L_{\rm radw}=L$, where $L$ is the assumed fixed luminosity at the top of the atmosphere. Non-negligible extra luminosity generated in the outer convective layers would, however, yield $L_{\rm radw}<L$, in conflict with our assumptions, and could change atmospheric structure. A simple way to check this \textit{a posteriori} is by rewriting the local energy equation (\ref{eq:structd}) as $\partial L/\partial m=-T \partial S/\partial t$, and integrating it throughout the atmosphere assuming that $\partial S/\partial t$ is fixed (e.g., \citealt{arras06}). If the value of this integral, $I$\footnote{This integral does not have units of luminosity, as we drop the constant $\partial S/\partial t$ term.}, throughout the innermost convective region is significantly larger than its value throughout the rest of the envelope, $\Delta I$, then our assumptions hold. For all the models for which $\Delta L \ll L$ holds (see paragraph above), we have found $\Delta I / I \lesssim 30\%$, and typically $\lesssim 10\%$ (see \App{radwindow} for additional details).

%A more rigorous check would be comparing our models with constant $L$ throughout with example profiles for which $L \propto \int T dm$. In practice, this calculation is not feasible for our code. However, we can make a simple estimate of how a non-constant $L$ affects atmospheric structure by assuming $L \propto m$ throughout the atmosphere. Note that $L \propto \int T dm$ makes the luminosity more centrally concentrated than if $L \propto m$, so our simple estimate is conservative. We have found that $L \propto m$ may change atmospheric structure by moving the outermost RCB closer to the top of the envelope, but this does not change the atmospheric luminosity or time evolution.

%As an additional check that the atmospheric structure and evolution are not changed due to the radiative windows, we

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/appendixF.tex
upd

--- 2015-11-13 tex/draft/ms_final_revised_2.tex
@@
-%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
-%\documentclass[12pt, preprint]{aastex}
-\documentclass[apj]{emulateapj}
-
-\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
-\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex
-
-%\usepackage{pdfsync}
-\usepackage{subeqnarray}
-\usepackage{natbib}
-\usepackage{color}
-\usepackage[utf8]{inputenc}
-
-\bibliographystyle{apj}
-
-\newcommand{\ie}{i.e.\ }
-\newcommand{\eg}{e.g.\ }
-\newcommand{\p}{\partial}
-\newcommand{\brak}[1]{\langle #1\rangle}
-
-\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
-\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
-\newcommand{\cm}{\; {\rm cm}}
-\newcommand{\mm}{\; {\rm mm}}
-%\newcommand{\ps}{\; {\rm s^{-1}}}
-\newcommand{\km}{\; {\rm km}}
-\newcommand{\au}{\; \varpi_{\rm AU}}
-\newcommand{\AU}{\; {\rm AU}}
-\def\K{\; {\rm K}}
-
-\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
-\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
-\newcommand{\nab}{\vc{\nabla}}
-\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
-\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
-\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}
-
-\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
-\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
-\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
-\newcommand{\App}[1]{Appendix~\ref{#1}}
-\newcommand{\Sec}[1]{Sect.~\ref{#1}}
-\newcommand{\Chap}[1]{Chapter~\ref{#1}}
-\newcommand{\Fig}[1]{Fig.~\ref{#1}}
-\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
-\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
-\newcommand{\Tab}[1]{Table \ref{#1}}
-
-\definecolor{gray}{gray}{0.5}
-\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}
-
-%\newenvironment{packed_item}{
-%\begin{itemize}
-%  \setlength{\itemsep}{1pt}
-%  \setlength{\parskip}{0pt}
-%  \setlength{\parsep}{0pt}
-%}{\end{itemize}}
-
-\begin{document}
-
-%\slugcomment{Draft Modified \today}
-
-\title{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}
-
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
-\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
-
-\begin{abstract}
-%The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry. 
-The C/O ratio is a defining feature of both gas giant atmospheric and protoplanetary disk chemistry.
-In disks, the C/O ratio is regulated by the presence of snowlines of major volatiles at different distances from the central star.
-We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk.
-We determine the snowline locations for a range of fixed initial particle sizes and disk types.
-For our fiducial disk model, we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk, and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving and viscous disk, desorb %instantaneously and [I think instantaneously is too strong]
-at a %fixed 
-size-dependent location in the disk, which is independent %regardless 
-of the particle's initial position.
-The snowline radius decreases for larger particles, up to sizes of $\sim$7~m.
-%location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
-Compared to a static disk, we find that radial drift and gas accretion in a viscous disk move the H$_2$O snowline inwards by up to 40 \%, the CO$_2$ snowline by up to 60 \%, and the CO snowline by up to 50 \%.
-We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
-%Our results have direct implications on the C/O ratio in giant planet atmospheres that have formed in situ.
-[I don't think we discuss this enough to warrant it as a conclusion
-\end{abstract}
-
-\section{Introduction}
-\label{sec:intro}
-
-%\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al.
-2011).}
-
-%\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture.
-Gas giants are important.
-The chemical composition of their atmospheres constraints their formation and evolution.
-Is is therefore important to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) backtrack the planet formation location based on planet composition.}
-
-The chemical composition of protoplanetary disks affects planet formation efficiencies and the composition of nascent planets.
-%[Need explicit statement that gas giants accrete their atmospheres from the nebular gas.] 
-Gas giants accrete their envelopes from the nebular gas.
-As such, planet compositions are tightly linked to the structure and evolution of the protoplanetary disk in which they form.
-It is thus essential to understand the disk chemistry and dynamics well enough to (1) predict the types of planet compositions that result from planet formation in different parts of the disk, and (2) backtrack the planet formation location based on planet compositions.
-%[split into two sentences]
-
-%\emgr{Add the rest as outlined above.} \\
-
-%\emgr{Paragraph 2: disks are complex, a lot of dynamical and chemical processes going on.
-However, we have detections of organic molecules.
-We have detections of snowlines.
-We see complex chemistry and many molecules we can study.}
-
-%In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11c}, \citealt{oberg11b}).
-[the molecules relevant for this paper are not organics]
-%[Topic sentence on that the most important volatiles are ... based on protostellar and comet abundances.]
-
-The structures of protoplanetary disks are complex, and affected by a multitude of chemical and dynamical processes (see review by \citealt{henning13}).
-From the chemistry perspective, volatile compounds are particularly important.
-Their snowline locations determine their relative abundance in gaseous and solid form in the disk,.
-Based on protostellar and comet abundances, some of the most important volatile molecules are H$_2$O, CO$_2$, CO, N$_2$.
-Recent observations of protoplanetary disks have provided valuable information about the abundances and snowline locations of some of these compounds.
-For example, the CO snowline has been detected in the disk around TW Hya \citep{qi13}, as well as in the disk around HD 163296 \citep{mathews13} using line emissions from DCO$^+$.
-Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}, and more such snowline detections are expected in future ALMA cycles.
-These observations are currently lacking an interpretive framework that takes into account all important dynamical and chemical processes.
-Furthermore, such a framework is crucial to connect observed snowline locations to planet formation.
-
-%Of particular importance are volatile compounds, such as H$_2$O, CO$_2$, CO, N$_2$ etc.
-Recent detections of the snowlines of volatiles (such as the CO snowline, \citealt{qi11}) are significant, since the snowline locations of these molecules determine their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
-[I think this paragraph belongs further down after you have discussed planets since that is the topic of your intro paragraph above.
-Also need a bit more meat here.
-So far a single CO snowline detection and a single H2O snowline detection.
-More are expected and an interpretive framework is key regardless of the importance for planets.]%Recent detections of the CO snowline in disks (e.g., \citealt{qi11}) provide the framework...
-
-%\emgr{Add part about snowlines, complex chemistry, many molecules.} \\
-
-An important consequence of snowline formations in disks is that disks are expected to present different carbon-to-oxygen (C/O) ratios in the gas and in icy dust mantles at different disk radii.
-This effect was quantified by  \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e.
-H$_2$O, CO$_2$ and CO, have different condensation temperatures.
-This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
-\citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-This effect was used to explain claims of detections of superstellar C/O ratios in exoplanet atmospheres (e.g., WASP-12b, \citealt{madhu11}), which however have been unambiguously refuted (\citealt{stevenson14}, \citealt{kreidberg15}).
-%Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.
-
-%Given the tight connection between planet and disk compositions, detections of C/O ratios in exoplanet atmospheres can provide some insight into the relative abundance of C and O in disks.
-Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}.
-One explanation for this discrepancy was proposed by \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e.
-H$_2$O, CO$_2$ and CO, have different condensation temperatures.
-This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
-\citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-%[This is not a very natural approach I think, i.e.
-to use a dicy eco-planet C/O detection to infer something about the disk chemical composition.]
-
-%\emgr{Paragraph 3: let's focus on one (set of) molecules: the C/O ratio.
-It's an important signature of atmospheric chemistry.
-While we cannot detect it directly in disks, we have observations in giant planet atmospheres.
-It's different than stellar.
-Why?
-Possible explanation based on snowlines.}
-
-%\emgr{Start with the fact that we can't see C/O in disks directly. Then transition to planet atmospheres and modify the next sentence to follow up logically.} 
-%Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio.
-
-\citet{oberg11} assumed a static disk with no chemical evolution.
-In reality, dynamical and chemical processes affect the snowline locations and the resulting C/O ratio.
-Several works have addressed some of these effects.
-\citet{madhu14} use a steady-state active disk model that includes planetary migration and use the C/O ratio to constrain migration mechanisms.
-\citet{alidib14} calculate the C/O ratio throughout the disk by incorporating the evolution of solids, i.e.
-radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors.
-\citet{alidib14} use the 1+1D $\alpha$-disk model of \citet{hughes10}, in which the gas drifts outwards in the disk midplane, and thus small particles that are well-coupled to the gas will also advect outward.
-Their model assumes a cyclical conversion between H$_2$O or CO dust and vapor: large enough particles that are decoupled from the gas drift inwards and start desorbing.
-Once their sublimation is complete, back-diffusion moves the H$_2$O or CO vapor ouwards to their respective snowlines, where they instantly condense into mm-sized particles that diffuse outwards with the gas while coagulating into larger particles.
-Once the grains become large enough to decouple from the gas and drift inwards, the cycle restarts.
-This ```conveyor belt''' model is based on the pioneering work by \citet{cuzzi04}, and \citet{ciesla06} for the evolution of H$_2$O in a viscous disk.
-This approach leads \citet{alidib14} to find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.
-\citet{thiabaud15} consider additional carbon carrier volatile species in their chemical network, such as CH$_4$, and find that the gas C/O ratio may be enriched by up to four times the Solar value in the outer parts of the disk where CH$_4$ and CO are the only gaseous carriers of C and O.
-They also include nitrogen carriers such as N$_2$ or NH$_3$, and perform similar calculations for nitrogen.
-
-%Most studies of the C/O ratio in disks are particularly focused on the inner disk, in order to explain observations of super-stellar C/O ratios in the atmospheres of close-in exoplanets \citep{madhu11}.
-In addition, the simultaneous inclusion of various dynamical and/or chemical processes makes it difficult to quantify the separate effect of each of these processes on snowline locations and the C/O ratio.
-%[The preceding paragraph needs to be made clearer.
-Currently it is not clear what effects these studies have had on our understanding of C/O ratio sin disks and why your study is needed.
-I.e.
-why are the processes you focus on the most important ones?
-Introduce separately papers that deal with C/O and water snowline locations.]
-Each of these studies have considered a specific combination of dynamical and chemical effects.
-One scenario that has not yet been considered is the combination of radial drift and viscous gas accretion in isolation.
-Studying these two dynamical processes makes it possible to quantify their separate effect on snowline locations and the C/O ratio at various disk radii.
-
- %In this paper, we focus on these two particular dynamical processes.
-%i.e.
-the radial drift of solids and viscous gas accretion.
-%We enhance the model of \citet{oberg11} by considering these two additional processes.
-In this paper, we perform a systematic study to understand the detailed qualitative and quantitative effects of radial drift and gas accretion on the H$_2$O, CO$_2$ and CO snowline locations, and the resulting C/O ratio in gas and dust throughout the protoplanetary disk.
-More importantly, we obtain a limit on how close to the star the snowline locations can be pushed by radial drift and gas accretion.
-
-%\emgr{Paragraph 4: the static disk model is a great start, but there are additional dynamical and chemical processes to take into account.
-Several studies have done that in some form or another --- cite Ali-Dib+14, Madhusudhan+14, Thiabaud+15.
-Briefly discuss their assumptions.} \\
-
-%\emgr{Paragraph 5: finally introduce our goal for this paper: study the dynamical effects such as drift and gas accretion.
-State our goals: understand the detailed qualitative and quantitative effect of drift and gas accretion on snowline locations, obtain a limit on how far in can the snowlines be pushed, and see how that affects the C/O ratio throughout the disk.
-State that our specific goals motivate the use of a simplified model (in other words, why what we are doing is different from the studies cited in the previous paragraph).
-}
-
-%\emgr{This needs some restructuring and additions along the lines discussed above.} In order to obtain more realistic estimates for C/O ratios across protoplanetary disks, dynamical processes and the disk evolving chemistry have to be taken into account.
-In this paper, we enhance the model of \citet{oberg11} considering two additional dynamic effects: (1) the radial drift of solids throughout the protoplanetary disk, and (2) the viscous accretion of the disk gas onto the host star.
-Our goal is two-fold: (1) to quantify the effect of radial drift of solids of different sizes on the location and shape of H$_2$O, CO$_2$ and CO snowlines, and (2) to calculate the resulting C/O ratio in gaseous and solid form throughout an actively accreting protoplanetary disk as a function of the grain size distribution and the evolutionary time of the nebula.
-
-%\emgr{Describe goals and approach of this paper, i.e.: study the importance of radial drift on the location of snowlines in protoplanetary disks and how radial drift affects the C/O ratio.
-Mention that disks are viscously accreting in the early stages of planet formation, hence gas accretion has to be taken into account.
-The goal of this paper is two-fold: (1) estimate using both numerical models and analytic arguments the range of particle sizes for which radial drift changes snowline location; (2) calculate the C/O ratio for various particle sizes (and possibly a size distribution) and at different times throughout the evolution of the gas disk.}
-
-This paper is organized as follows.
-In Section \ref{sec:model}, we present our disk, radial drift and desorption models, as well as the timescales relevant to the coupled drift-desorption process.
-We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for an irradiated and an evolving disk in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}.
-In Section \ref{sec:discussion}, we discuss the generality of our results, as well as additional effects on the snowline locations.
-Finally, we summarize our findings in Section \ref{sec:summary}.
-
-\section{Model Framework} 
-\label{sec:model}
-
-We present our protoplanetary disk model for a static, an irradiated, an evolving, and a viscous disk in section \ref{sec:disk}.
-In section \ref{sec:drift}, we describe our analytic model for the radial drift of solids.
-We summarize our ice desorption model in section \ref{sec:desorption}.
-Finally, we discuss the relevant timescales for dynamical effects that affect snowline locations %in the desorption process
-in section \ref{sec:timescales}.
-
-\subsection{Disk Model}
-\label{sec:disk}
-
-To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{static disk}, which is solely irradiated by the host star and does not take into account gas accretion onto the star or radial drift; \textit{irradiated disk}, which has the same temperature profile as the static disk and does not experience gas accretion or accretional heating, but it takes into account radial drift of solids; \textit{evolving disk}, in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; and \textit{viscous disk}, for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis, and the temperature profile is calculated using both accretional heating and stellar irradiation.
-
-%[Perhaps define your four disk models up here and motivate why you are exploring several different ones, and then go through and describe them in detail.
-That would save you the slightly awkward Active disk with passive temperature profile.]
-
-\textbf{Static and Irradiated disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a static and and an irradiated disk similar to the prescription of \citet{chiang10}.
-The gas surface density and midplane temperature are
-\begin{subeqnarray}
-\label{eq:disk}
-\Sigma&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
-T &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
-\end{subeqnarray}
-where $r$ is the semimajor axis.
-Our surface density profile is flatter than the $\Sigma \propto r^{-3/2}$ used by \citet{chiang10}.
-Our choice is inspired by observations of protoplanetary disks at radii larger than $\sim$20 AU (e.g., \citealt{andrews10}), which suggest that typical disks may have surface density profiles with $\Sigma \propto r^{-1}$.
-A slope flatter than $\Sigma \propto r^{-3/2}$ is also more consistent with the temperature profile for a steady-state gas disk (see the Viscous disk heading below and \App{app:steadystate}).
-We use the static disk model to compare our results with those of \citet{oberg11}.
-
-%Based on some observations of protoplanetary disks \citep{andrews10}, our surface density profile, $\Sigma \propto r^{-1}$, is flatter than that of \citet{chiang10}, i.e. $\Sigma \propto r^{-3/2}$.
-
-\textbf{Evolving disk.} We model the evolving disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
-\begin{equation}
-\label{eq:nu}
-\nu=\alpha c H.
-\end{equation}
-Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c$, $H$ are the isothermal sound speed and disk scale height, respectively:
-\begin{subeqnarray}
-\label{eq:cdHd}
-c &=& \sqrt{\frac{k_{\rm B} T}{\mu m_{\rm p}}} \slabel{eq:cd} \\
-H&=& \frac{c}{\Omega_{\rm k}} \slabel{eq:Hd},
-\end{subeqnarray}
-where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$ is the Keplerian angular velocity,  with $G$ the gravitational constant and $M_*$ the stellar mass.
-We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition of hydrogen and helium.
-The temperature profile for the %an
-evolving disk is assumed to be the same as for the irradiated disk and given by Equation (\ref{eq:diskT}).
-From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters.
-Following \citet{hartmann98}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius.
-We choose $r_{\rm c}=100$ AU.
-The gas surface density is given by the self-similar solution
-
-\begin{equation}
-\label{eq:Sigmaact}
-\Sigma(R, \tilde t) = \frac{M (2 - \gamma)}{2 \pi r_{\rm c}^2 R^{\gamma}} \tilde t^{-(5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{(2-\gamma)}}{\tilde t}}\Big],
-\end{equation}
-where $M$ is the total disk mass and
-\begin{subeqnarray}
-\label{eq:T}
-\tilde t & \equiv & \frac{t}{t_{\rm c}} + 1 \\
-t_{\rm c} & \equiv & \frac{1}{3(2-\gamma)^2} \frac{r_{\rm c}^2}{\nu_{\rm c}},
-\end{subeqnarray}
-where $t$ is time.
-We choose $M=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
-The irradiated and evolving disk surface densities match at \textbf{$t \approx 5 \times 10^5$} years in the inner disk, but they diverge at distances larger than a few AU due the exponential cutoff in radius of the surface density of the evolving disk (Equation \ref{eq:Sigmaact}).
-
-\textbf{Viscous disk.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
-Since
-Calculating the midplane temperature self-consistently for an evolving disk that is also actively heated, and thus whose thermal evolution is dominated both by accretion heating and stellar irradiation, is non-trivial.
-We therefore use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm act}$.
-The equations governing the evolution of the steady-state disk are listed in \App{app:steadystate}.
-We assume an interstellar opacity for the dust grains given by \citet{bell94}, but reduced by a factor of 100.
-This reduction is due to the fact that disk opacities are lower than the interstellar one.
-While this scaling is consistent with more detailed models of grain opacities in disks (e.g., \citealt{mordasini14}), realistic disk opacities are much less sensitive to changes in temperature than the interstellar opacity if substantial grain growth has occurred.
-However, the disk temperature does not vary significantly across the small region of the disk where accretion heating is important ($r \lesssim 1$ AU).
-Moreover, using an analytic opacity formula is more convenient since it results in a constant gas surface density in the inner disk region (see below).
-Our opacity law is thus
-
-\begin{equation}
-\label{eq:opacity}
-\kappa=\kappa_0 T_{\rm act}^2,
-\end{equation}
-where $\kappa_0=2 \times 10^{-6}$. By solving the Equation set (\ref{eq:diskeq}) we find
-\begin{equation}
-\label{eq:Tdact}
-T_{\rm act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
-\end{equation}
-Since both accretion heating and stellar irradiation contribute to the thermal evolution of the disk, we compute the midplane temperature for our viscous disk as
-\begin{equation}
-\label{eq:activeT}
-T^4 = T_{\rm act}^4 + T_{\rm irr}^4,
-\end{equation}
-where to avoid notation confusion $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}), the temperature profile for an irradiated disk.
-We can then easily determine  $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
-For consistency, we choose  $\alpha=0.01$ as in the previous case.
-Finally, we determine $\Sigma$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}).
-In the inner portion of our disk ($r  \lesssim 1$ AU for our fiducial model with $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$), our choice of opacity (Equation \ref{eq:opacity}) implies that the disk has a constant surface density with radius (see Equations \ref{eq:diskeq}).
-
-Before we proceed forward, we note that our disk models assume a constant stellar luminosity $L_*$, as well as a constant mass accretion rate $\dot{M}$ for the viscous disk.
-In reality, the stellar luminosity decreases as the host star contracts, which will reduce the disk temperature and push the snowlines inward, as we explain in Section \ref{sec:neglected}.
-For a Solar type star, as our fiducial model assumes, $L_*$ remains relatively constant during the star's pre-main sequence evolution of $\sim$ 10 Myr \citep{kennedy06}, which is larger than the giant planet formation timescale.
-Thus the midplane temperature will not change significantly for our model due to variations in stellar luminosity, but it may decrease substantially for smaller stars, pushing the snowline inward (see Section \ref{sec:neglected} for details).
-Realistic mass accretion rates, $\dot{M}$, may vary between $\sim$$10^{-7}$ and $\sim$$10^{-9}$ $M_{\odot}$ yr$^{-1}$ as the disk evolves \textbf{(e.g., \citealt{chambers09}, \citealt{sicilia10})}.
-For $\dot{M}\lesssim 10^{-9} M_{\odot}$ yr$^{-1}$, the disk becomes optically thin and hence depleted of gas, which means giant planets must have formed before $\dot{M}$ becomes too low.
-\citet{garaud07} find that the snowline locations scale as $r_{\rm snow} \propto \dot{M}^{1/3}$.
-A factor of 100 reduction in the mass accretion rate will thus move the H$_2$O snowline inwards by a factor of $\sim$$4$ \textbf{--- since accretion heating is dominant only in the inner disk, the CO$_2$ and CO snowline locations are unlikely to be affected by changes in mass accretion rate.
-The inward movement of the H$_2$O snowline due to the decrease in $\dot{M}$ may be even larger, by up to one order of magnitude \citep{chambers09}}.
-We thus conclude that changes in $L_*$ throughout time may only modestly affect our results, \textbf{while changes in $\dot{M}$ may significantly affect our results for the H$_2$O snowline, as its location may be determined by the decline in mass accretion rate rather than radial drift.} The time variability of $L_*$ and $\dot{M}$ should be taken into account when drawing more robust conclusions, as well as for different host star and disk properties.
-
-%However, \citet{garaud07} find that the mass accretion rate is close to $10^{-8} M_{\odot}$ yr$^{-1}$ during the quasi-static accretion phase of the disk (see also Section \ref{sec:neglected}), hence the snowline locations are likely to move inward only in the very early stages of the disk evolution.
-We thus claim that changes in $L_*$ and $\dot{M}$ throughout time may only modestly affect our results, but they should still be taken into account when drawing more robust conclusions, as well as for different host star and disk properties.}
-
- %We note that $\Sigma$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.
-
-%Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady-state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):
-
-%\begin{subeqnarray}
-%\label{eq:activeT}
-%T^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
-%\dot{M} & = & 3 \pi \nu \Sigma \slabel{eq:Mdot},
-%\end{subeqnarray}
-%where $\dot{M}$, the mass flux, is constant in steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}.
-Based on disk observations (e.g., \citealt{andrews10}), we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$.
-We can then easily determine the radial temperature profile from Equation (\ref{eq:Tdact}), $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
-For consistency, we choose  $\alpha=0.01$ as in the previous case.
-Finally, we determine $\Sigma$ from Equation (\ref{eq:Mdot}).
-
- %The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
-
-\subsection{Radial Drift}
-\label{sec:drift}
-
-Solid particles orbit their host star at the Keplerian velocity $v_{\rm k} \equiv \Omega_{\rm k} r$.
-The gas, however, experiences an additional pressure gradient, which causes it to rotate at sub-Keplerian velocity \citep{weidenschilling77}.
-Dust grains that are large enough thus experience a headwind, which removes angular momentum, causing the solids to spiral inwards and fall onto the host star.
-Small particles are well-coupled to the gas, while large planetesimals are decoupled from the gas.
-From the review by \citet{chiang10}, the extent of coupling is quantified by the dimensionless stopping time, $\tau_{\rm s} \equiv \Omega_{\rm k} t_{\rm s}$, where $t_{\rm s}$ is
-\begin{equation}
-\label{eq:ts}
-t_{\rm s}= \left\{
-\begin{array}{l l}
-\rho_{\rm s} s / (\rho c), & \quad s < 9 \lambda/4 \,\,\,\ \text{Epstein drag} \\
-4 \rho_{\rm s} s^2 / (9 \rho c \lambda), & \quad s < 9 \lambda/4, \,\text{Re} \lesssim 1 \,\,\,\ \text{Stokes drag.}
-\end{array} 
-\right.
-\end{equation}
-Here $\rho$ is the gas midplane density, $\rho_{\rm s}=2$ g cm$^{-3}$ is the density of a solid particle, $s$ is the particle size, $\lambda$ is the mean free path, and Re is the Reynolds number.
-
-For an irradiated disk, the radial drift velocity can be approximated as
-\begin{equation}
-\label{eq:rdotpas}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big),
-\end{equation}
-where
-\begin{equation}
-\label{eq:eta}
-\eta \equiv - \frac{\partial P/\partial \ln r}{2 \rho v_{\rm k}^2} \approx \frac{c^2}{2 v_k^2}
-\end{equation}
-and $P = \rho c^2$ is the disk midplane pressure.
-
-For an evolving disk, the radial drift velocity has an additional term due to the radial movement of the gas \citep{birnstiel12}, i.e.
-\begin{equation}
-\label{eq:rdotact}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{\dot{r}_{\rm gas}}{1+\tau_{\rm s}^2},
-\end{equation}
-where $\dot{r}_{\rm gas}$ is the radial gas accretion velocity and can be expressed as (e.g., \citealt{fkr02})
-\begin{equation}
-\label{eq:vgas}
-\dot{r}_{\rm gas} = - \frac{3}{\Sigma \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma \sqrt{r}) 
-\end{equation}
-with $\Sigma$ from Equation (\ref{eq:Sigmaact}).
-For the viscous disk (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}).
-For our choice of parameters for both the evolving and the viscous disks, we have found that the radial flow of gas \textbf{(calculated from Equation \ref{eq:vgas} for the evolving disk and from $\dot{r}_{\rm gas}=-\dot{M}/(2 \pi r \Sigma)$ for the viscous disk)} is always directed inward \textbf{for our parameter space of interest}, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}).
-\textbf{For our evolving disk, the gas starts drifting outwards at a radius $r_{\rm switch} \approx 200 AU$, which is however well outside the CO snowline in our model.} We \textbf{thus} note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations.
-Since drifting particles larger than a few cm are only modestly affected by gas accretion, an outward gas flow would move the CO$_2$ and CO locations further away from the star only for the smallest particles in our model, which are well-coupled to the gas.
-
-\subsection{Volatile Desorption}
-\label{sec:desorption}
-
-%\subsection{Disk and Desorption Model}
-
-In order for a volatile species to thermally desorb, it has to overcome the binding energy that keeps it on the grain surface.
-Following \citet{hollenbach09}, the desorption rate per molecule for a species $x$ can be expressed as
-\begin{equation}
-\label{eq:Rdes}
-R_{\rm{des}, x} = \nu_x \exp{(-E_x/T_{\rm grain})},
-\end{equation}
-where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well, with $\mu_x$ the dimensionless mean molecular weight.
-We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T$.
-For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively (\citealt{collings04}, \citealt{fraser01}, \citealt{aikawa96}).
-We use the desorption rate, $R_{\rm des}$, to estimate the desorption timescale for particles of different sizes as described in section \ref{sec:timescales}.
-
-%\emgr{Describe disk model --- both passive MMSN and active self-similar solution with $\alpha$ prescription for viscosity.
-Describe the model for evolving the surface density profile of planetesimals following Birnstiel et al.
-(2012).
-Describe temperature profile --- currently power-law, but may change.
-Describe desorption model and parameters following Hollenbach et al.
-(2009).
-Mention that the solids are perfect spheres composed of a single volatile.
-Perhaps this subsection needs to be split into subsubsections.}
-
-\subsection{Relevant Timescales}
-\label{sec:timescales}
-
-We can estimate the extent to which radial drift and gas accretion affect desorption by comparing the timescales for desorption, drift and accretion, for solids of different sizes and compositions.
-
-\textit{Desorption timescale.} We assume that the solid bodies are perfect spheres and are entirely composed of only one volatile species, i.e.
-either H$_2$O, CO$_2$ or CO \footnote{We discuss the validity of these simplifications in section \ref{sec:discussion}.}.
-The timescale to desorb a single layer of molecules can then be estimated as
-\begin{equation}
-\label{eq:tdes}
-t_{\rm des}=\frac{\rho_{\rm s}}{3 \mu_x m_{\rm p}} \frac{s}{N_x R_{\rm des, x}},
-\end{equation}
-where $N_x \approx 10^{15}$ sites cm$^{-2}$ is the number of adsorption sites of volatile $x$ per cm$^2$, assuming that the particle has a smooth surface \citep{hollenbach09}.
-
-\textit{Radial drift timescale.} To order of magnitude, the radial drift timescale can be estimated as 
-\begin{equation}
-\label{eq:tdrift}
-t_{\rm drift} \sim \Big|\frac{r}{\dot{r}}\Big|,
-\end{equation}
-where $\dot{r}$ is the radial drift velocity given by Equation (\ref{eq:rdotpas}) for an irradiated disk and by Equation (\ref{eq:rdotact}) for an evolving disk.
-
-\textit{Gas accretion timescale.} The timescale for gas accretion onto the central star for an evolving disk is (e.g., \citealt{armitage10})
-\begin{equation}
-\label{eq:tgas}
-t_{\rm gas, acc} \sim \frac{r^2}{\nu} \sim \frac{1}{2 \alpha \eta \Omega_{\rm k}},
-\end{equation}
-with the latter expression derived from Equations (\ref{eq:nu}) and (\ref{eq:eta}).
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{drift_timescales_betaS1_gas_acc_new2.pdf}
-%\vspace{-0.5in}
-\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
-The timescales are calculated at three representative locations, i.e.
-the H$_2$O, CO$_2$ and CO snowlines in the static disk.
-For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively.
-The horizontal dot-dashed line represents a typical disk lifetime of 3 Myr.
-The particle size ordering at the minimum $t_{\rm drift}$ is not monotonic in snowline distance due to different drag regimes for those particle sizes at the snowline locations (Epstein drag at the H$_2$O and CO$_2$ snowlines, and Stokes drag at the CO snowline).
-Similarly, the ordering of $t_{\rm des}$ is not monotonic in snowline distance due to the non-monotony in mean molecular weight between H$_2$O, CO$_2$ and CO (18 $m_{\rm p}$, 44 $m_{\rm p}$ and 28 $m_{\rm p}$, respectively).
-Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
-$t_{\rm drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.}
-\label{fig:timescales}
-\end{figure}
-
-For simplicity, we calculate the radial drift timescale, $t_{\rm drift}$, for an irradiated disk in this section, but most of our conclusions hold true for an evolving disk as well.
-Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk.
-As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence on disk location (see Equation \ref{eq:Rdes}).
-On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas.
-Thus for small particles in an irradiated disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}).
-This is not true for an evolving disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.
-At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well.
-This is true for both irradiated and evolving disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.
-
-In the particle size regime for which (1) $t_{\rm drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
-for $\sim$$0.5$ cm $\lesssim s \lesssim$ $1000$ cm, or (2) $t_{\rm gas, acc} \lesssim t_{\rm des} \lesssim t_{\rm d}$, i.e.
-for $\sim$$0.1$ cm $\lesssim s \lesssim$ $10$ cm, radial drift or gas accretion (or both) are faster than thermal desorption, which is of particular interest for our purposes.
-We note that $t_{\rm gas, acc}<t_{\rm d}$ always holds true.
-Particles of sizes that satisfy the requirements above will drift significantly due to radial drift or gas accretion before desorbing, thus moving the H$_2$O, CO$_2$ and CO snowlines closer towards the central star and changing the C/O ratio throughout the disk.
-We quantify these effects in sections \ref{sec:snowlines} and \ref{sec:COratio}.
-
-%\emgr{Calculate timescale for radial drift following Chiang \& Youdin (2010).
-Calculate desorption timescale following Hollenbach et al.
-(2009).
-Estimate gas accretion timescale for a given $\alpha$.
-Important: mention that, for simplicity and illustrative purposes, these calculations are performed for a passive disk.
-Show plot with the timescales as a function of particle size at different snowlines to show the regime in which drift matters}.
-
-\section{Snowline Locations}
-\label{sec:snowlines}
-
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=0.7\textwidth]{desorption_distance_passive_active_colorbar_test2.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both an irradiated disk (left panels) and an evolving disk (right panels).
-The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The desorption distance for a static disk is shown for comparison (dashed vertical and horizontal lines).
-The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
-For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.}
-\label{fig:snowlines}
-\end{figure*}
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=0.7\textwidth]{desorption_distance_actual_vs_estimated_passive_active_new.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text).
-The estimate is performed for an irradiated disk (left panels) and an evolving disk (right panels).
-The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
-\label{fig:an_vs_actual}
-\end{figure*}
-
-In this section we use the model described in section \ref{sec:model} to quantify the effects of radial drift (irradiated disk) or radial drift and gas accretion (evolving disk) on the snowline location, for dust particles of different sizes composed of either H$_2$O, CO$_2$ or CO.
-Specifically, we determine a particle's final location (i.e., where the particle either fully desorbs or remains at its initial size due to having a desorption timescale longer than the time at which we stop the simulation) as a function of its initial position in the disk, after the gas disk has dissipated.
-The disk lifetime, $t_{\rm d}$, is particularly relevant since the timescale for giant planet formation must be less than or equal to $t_{\rm d}$.
-The snowline locations at $t=t_{\rm d}$ throughout the protoplanetary disk determine the disk C/O ratio in gas at this time, and thus the C/O ratio in giant planet atmospheres that have formed \textit{in situ}, before planetesimal accretion or core dredging.
-
-For each species $x$, we determine the final location in the disk of a particle of initial size $s_0$ by solving the following system of coupled differential equations:
-
-\begin{subeqnarray}
-\label{eq:ddt}
-\frac{ds}{dt} &= & - \frac{3 \mu_x m_{\rm p}}{\rho_{\rm s}} N_x R_{\rm des, x}  \slabel{eq:dsdt} \\
-\frac{dr}{dt} &=& \dot{r} \slabel{eq:drdt},
-\end{subeqnarray}
-where the desorption rate $R_{\rm des, x}$ for each particle type (i.e., composed of H$_2$O, CO$_2$ or CO) is evaluated at $T=T(r)$, and the radial drift velocity $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for an irradiated disk and Equation (\ref{eq:rdotact}) for an evolving disk.
-Equations (\ref{eq:dsdt}) and (\ref{eq:drdt}) describe the coupled desorption and radial drift, and can be derived straightforwardly from Equation (\ref{eq:tdes}).
-Our initial conditions are $s(t_0)=s_0$ and $r(t_0)=r_0$, where $t_0$ is the initial time at which we start the integration and $r_0$ is the initial location of the particle.
-We choose $t_0=1$ year, but our result is independent on the initial integration time as long as $t_0 \ll t_{\rm d}$.
-The desorption timescale $t_{\rm des}$ will then satisfy $s(t_{\rm des})=0$, from which we can determine the desorption distance $r_{\rm des}=r(t_{\rm des})$.
-
-%As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully evaporates.
-
-%Thus a particle's final location will depend on whether a grain initially at a certain distance is completely desorbed or not within the disk lifetime of 3 Myr.
-For example, larger grains take longer to desorb and hence are more likely to not evaporate fully in a given timeframe.
-
-We define the final position of a grain as the disk location it has reached after $t_{\rm d}=3$ Myr, or the radius at which it completely desorbs if that happens after a time shorter than 3 Myr.
-Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both an irradiated and an evolving disk.
-We do not show the results for the viscous disk as they would complicate the plot without adding any qualitative insight ---  the results for the viscous disk are quantitatively similar with those of the evolving disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
-We also show the static snowlines for comparison, which are calculated by balancing adsorption and desorption \citep{hollenbach09}.
-%We do not show the equivalent result for the steady-state active disk since The trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state active disk results in Section \ref{sec:COratio}.
-Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for an irradiated nor for an evolving disk.
-Similarly, micron- to mm-sized particles in the irradiated disk do not drift and only %or
-desorb if %unless 
-they are located inside the static snowlines.
-%This is not in contradiction with the desorption timescales for small particles from Figure \ref{fig:timescales}, since small grains only desorb fast at or nearby their respective snowlines, as mentioned in Section \ref{sec:timescales}.
-[don't think this is needed, but Til or Ruth might think otherwise]
-In an evolving disk, however, micron-to mm-sized grains do drift significantly since they move at the same velocity as the accreting gas.
-For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in an irradiated disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an evolving disk, we notice that particles of initial size $s_0$ desorb at a %fixed distance
-particle size dependent radius $r_{\rm des}$ regardless of their original location in the disk.
-In fact, the only grains that will both drift and evaporate are those that reach their fixed final location (represented by the horizontal curves in Figure \ref{fig:snowlines}) within the disk lifetime.
-We show in section \ref{sec:COratio} that this result is essential in determining the C/O ratio throughout the disk for different particle sizes.
-
-Another interesting feature of Figure \ref{fig:snowlines} is that particles above a certain size ($\sim$7 m for our choice of parameters) all desorb at the same distance.
-This is due to the fact that once the large bodies pass the static snowline, they first lose mass, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=0.8\textwidth]{s_t_a.pdf}
-%\vspace{-0.5in}
-\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in an irradiated disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left).
-Particles desorb almost instantaneously.
-Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
-Particles lose most of their mass very close to the distance at which they fully desorb.
-The static H$_2$O snowline is shown for reference (dashed vertical lines).}
-\label{fig:s_t_a}
-\end{figure*}
-
-Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm drift} \sim t_{\rm des}$, given an initial particle size.
-We can calculate this location analytically by equating Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and solving for $r=r_{\rm des}(s)$ for a given particle size $s$.
-Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically.
-We display this result for the range of particle sizes that desorb at a fixed distance in an irradiated and an evolving disk (see Figure \ref{fig:snowlines}).
-We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it deviates for particles larger than $s \gtrsim10$ cm.
-For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the irradiated disk in the Epstein drag regime), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
-Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
-
-%This is due to the fact that large bodies lose their mass as they desorb until they reach the same size as the smaller, meter-sized particles, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
-
-Given $r_{\rm des}$, we need to only calculate the distance over which particles desorb to determine the location of a snowline.
-Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in an irradiated disk.
-Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
-%, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-%A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
-[not sure you need to explain this]
-The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
-Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model.
-Both of these conclusions remain valid for an evolving and a viscous disk, as well as for particles composed of CO$_2$ or CO, but the snowline locations will vary between the three disks for a given initial particle size (see Section \ref{sec:COratio}).
-If we do not take into account the time dependence of the mass accretion rate and stellar luminosity (see Section \ref{sec:disk}), the C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
-%\emgr{I've yet to figure out why there is a discrepancy for the small grains.
-I'll keep thinking about it, but for now I'll leave it as it is since I don't want to give a false explanation just to have one.}
-
-%\emgr{I didn't write the text for Figure 4 since we haven't yet decided if in the end we will include it in some form. We can }
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_vs_s.pdf}
-%%\vspace{-0.5in}
-%\caption{Desorption distance as a function of initial particle size, for the range of particles that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines}).
-\textbf{Figure belongs to section 3.} \emgr{This is a figure that I'm not sure we should include, since it doesn't really provide that much useful inside.
-If we did include it though, there would be two panels both for the passive and the active disk.
-The caption is not complete since the active disk panel is not yet there.}}  %  (See text for a description of evolution to yet higher masses.)
-%\label{fig:r_vs_s}
-%\end{figure}
-
-%\emgr{Present the equation set that you are solving, $dr/dt=\dot{r}$, $ds/dt=...$.
-I don't think it is necessary to describe in detail the numerical method of solving the equations since it's pretty straightforward.
-Present the pretty rainbow plots side by side for passive and active disks, and highlight the differences.
-Insert the plots that shows that for the intermediate size particles, the desorption distance can be estimated analytically with good accuracy.
-Maybe also include the plot showing the desorption distance as a function of particle size, both for passive and active disk.
-Split into subsections?}
-
-\section{C/O Ratio Estimates}
-\label{sec:COratio}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
-%\vspace{-0.5in}
-\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for an irradiated disk (top panel), an evolving disk (middle panel) and a viscous disk (bottom panel).
-%, for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk. 
-The particle size increases from $\sim$0.05 cm to $\sim$700 cm as indicated by the color bar.
-The horizontal dotted line represents the stellar value of 0.54.
-The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk, with the temperature profile given by Equation (\ref{eq:diskT}) for the top two panels and by Equation (\ref{eq:activeT}) for the bottom panel.
-%The snowline location moves inward as the particle size increases. 
-%The curves exterior to the static snowline represent desorption distances in the absence of readsorption.
-%Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline.
-For both the evolving and the viscous disk, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, as shown by the arrows.}
-%, with the decrease more significant interior to the H$_2$O snowline.}
-\label{fig:CO_ratio}
-\end{figure}
-
-Given our results in Section \ref{sec:snowlines}, a disk's C/O ratio in mainly affected by the snowline location for the particle size housing the most mass in ice.
-%We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
-Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then estimate the C/O ratio throughout the disk as a function of the particle size.
-Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
-In Figure \ref{fig:CO_ratio}, we display the H$_2$O, CO$_2$, and CO snowline locations as a function of particle size for disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
-The minimum snowline distance for a disk is given by the curve corresponding to the maximum particle size it hosts.
-For grains that have grown to radii larger than $\sim$7 m and are able to drift and desorb, the $\sim$7 m snowline applies (see Section 3).
-%Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.
-
-%[I think a sentence is missing here outlining what you need to establish to calculate the C/O ratio, i.e. the connection between the preceding and the next sentence is not obvious.]
-
-%Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk.
-Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
-%, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-%A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
-[not sure you need to explain this]
-%The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
-Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-%It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active)\footnote{Both of these conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
-The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
-
-%In Section \ref{sec:snowlines}, we showed that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-Figure \ref{fig:s_t_a} confirms both of these effects.
-The left panels show the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
-Indeed, the solid H$_2$O particles evaporate almost instantly, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime.
-However, our model assumes that particles drift continuously at any location in the disk.
-Therefore, a particle that can fully desorb during the disk lifetime for at least one initial location will always desorb, at a fixed distance as discussed in Section \ref{sec:snowlines} and displayed in Figure \ref{fig:snowlines}.
-The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
-Figure \ref{fig:s_t_a} thus proves claims (1) and (2) above.
-
-Drift and gas accretion affect the C/O ratio in a disk both because they move the snowline locations of the main C and O carriers  and because they cause solids and gas---which contain different proportions of C and O---to move inward at different rates.
-As shown in Section 3, the snowline locations depend on disk age only indirectly, through changes in disk properties and grain size.
-The C/O ratio is a function of the locations of the snowlines and the abundances of H$_2$O, CO$_2$, and CO relative to the H$_2$ abundance in the disk midplane.
-These abundances evolve over time as solids and gas move inward at different rates.
-
-Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for an irradiated disk, an evolving disk, and a viscous disk, under the simplifying assumption that the abundance relative to hydrogen for each volatile is fixed, so that drift and accretion affect only the locations of the snowlines.
-We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
-Snowline locations correspond to $r_{\rm des}$ in Figure \ref{fig:snowlines}, representing the location at which particles desorb in the absence of readsorption.
-The C/O ratio for a static disk, where desorption and readsorption balance \citep{hollenbach09}, is shown as a guideline.
-We note that the true snowline for particles with $r_{\rm des}$ outside the static snowline is the static snowline itself---thus only particles with initial sizes larger than $\sim$0.05 cm are plotted in the three panels, as particles that form snowlines at larger distances (cf.
-Figure \ref{fig:snowlines}) are not true snowlines.
-
-Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the evolving and viscous disks ignore the movement of the desorbed ices with the accreting gas---the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
-As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
-For example, for the disk parameters and particle sizes displayed in Figure \ref{fig:CO_ratio}, water molecules in solid particles drift up to $\sim$1000 times faster across the H$_2$O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.
-This differential inward motion will result in an increased oxygen gas abundance inside the H$_2$O snowline, and thus a (in some cases much) lower gaseous C/O ratio in this region.
-Conversely, oxygen gas inside the water snowline will be depleted compared to the static disk if H$_2$O particles grow to planetesimal size and stall their migration between the H$_2$O and CO$_2$ snowlines, leaving only gaseous CO and CO$_2$ to accrete inward.
-Growth of large planetesimals can therefore increase the C/O ratio in the inner disk.
-
-\textbf{Our premise that each volatile has a fixed abundance at every location in the disk yields good approximations for the C/O ratio in an irradiated disk --- our model assumes a constant influx of particles at any given radius, implying that the solid surface density of a given volatile, and therefore the amount of desorbed vapor, remain constant with semimajor axis.
-For the evolving and viscous disks, the C/O ratio in gas may decrease due to the decline in the surface density of solids with time.
-Thus both the relative fluxes of the volatiles and the decrease in solid ice abundances will lower the C/O ratio in evolving and viscous disks, as indicated by the arrows in Figure \ref{fig:CO_ratio}.
-However, a high gas-phase C/O ratio is maintained further in for all disks compared to a static disk, since drift and gas accretion cause the ices to desorb closer to the host star.}
-
-Figure \ref{fig:CO_ratio} plots snowline curves for particle sizes $\sim0.5$ cm $\lesssim $s$ \lesssim$ 7m.
-In the outermost disk, H$_2$O, CO$_2$, and CO all solidify.
-Hence, relative drift across the CO snowline can alter only the abundances of volatiles between the CO$_2$ and CO snowlines, but not the C/O ratio in this region.
-Interior to the CO$_2$ snowline, however, relative drift is important.
-We have found that the largest drifting particles in our model ($\sim$7 m) drift faster than the gas at both the H$_2$O and CO$_2$ snowlines.
-We thus conclude that the C/O ratio interior to the H$_2$O and CO$_2$ snowlines in our evolving and viscous disks will be lower than in the static disk, due to the additional oxygen added to the gas by desorbing H$_2$O and CO$_2$.
-For these particle sizes, our calculated C/O ratio is an upper limit,  as indicated by the arrows in Figure \ref{fig:CO_ratio}.
-
-%We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
-Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and an active disk accretionally heated.
-The C/O ratio for a static disk is shown as a guideline.
-We note that particles desorbing outside the static snowline will not form snowlines, since our calculation of the desorption distance does not account for readsorption  \citep{hollenbach09}.
-
-%Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the active disks ignore the movement of the desorbed ices with the accreting gas --- the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
-As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
-Specifically, the gas C/O ratios may be significantly lower inside the H$_2$O and CO$_2$ snowlines for the active disks, depending on the gas accretion rate, due to the additional oxygen contained by H$_2$O and CO$_2$ vapor that moves together with the accreting gas and crosses their respective snowlines.
-
-%e thus present this result for the C/O ratio primarily as a visual comparison with previous work \citep{oberg11} and note that our estimates for the C/O ratio in the active disks are not quantitatively accurate at all disk locations.
-Future work will address this issue.
-
- %The plot is consistent with Figure \ref{fig:snowlines}. 
-The snowline locations in these disks exhibit several interesting features.
-For the irradiated disk, only grains larger than $\sim$0.5 cm drift, desorb and thus move the snowline compared to the static disk.
-In contrast, even $\sim$micron-sized grains drift and desorb for the evolving disk, since they flow towards the host star together with the accreting gas.
-For the same particle size, the snowline locations are slightly closer to the central star in the evolving disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
-Equation \ref{eq:rdotact}).
-The addition of accretional heating in the viscous disk moves the H$_2$O snowline outwards.
-This is due to the fact that accretional heating dominates in the inner disk, where high temperatures cause the grains to evaporate further away from the star.
-Once $r\gtrsim1-2$  AU, stellar irradiation dominates the thermal evolution of the disk, and therefore the CO$_2$ and CO snowlines locations are the same as in the evolving and viscous disks.
-
-Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases.
-While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that %almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk.
-B
-bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals (see Section \ref{sec:snowlines}).
-However, the contribution of kilometer-sized bodies to the snowline location is modest, since they only drift if they are located very close to the snowline.
-%[new sentence explaining that their contribution may not be very big due to their modest drift.] though 
-%only the kilometer-sized boulders that are very close to the snownlines will contribute to determining the snowline location. 
-Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
-For a grain size distribution with a maximum particle size different than our model, one can pick out the appropriate minimum snowline locations from this plot.%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
-Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from this plot.
-%Our model thus sets an inner limit to the snowline locations.
-%Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution --- for a given grain size .
-
-% [the last sentence needs to be made more explicit, i.e.
-for a given grain size distribution with a maximum grain size a reader can now pick out the appropriate minimum snowline location from this plot]
-
-For our choice of parameters, the minimum snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the irradiated disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the evolving disk and $r_{\rm{H_2O}} \approx 0.63$ AU for the viscous disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the irradiated disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both the evolving and the viscous disks; $r_{\rm{CO}} \approx 30$ AU for the irradiated and both the evolving and the viscous disks.
-For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU\footnote{For the viscous disk, we calculated the static snowline location using the same temperature profile as that of the viscous disk, for consistency purposes.
-Thus $r_{\rm{H_2O}} \approx 0.98$ AU for the static disk in this scenario.}, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
-For the viscous disk model, which is the most realistic, radial drift and gas accretion push the snowline locations inwards by up to $\sim$$40$ \% for H$_2$O, by up to  $\sim$$60$ \% for CO$_2$, and by up to $\sim$$50$ \% for CO.
-We note that the H$_2$O snowline in all disks is significantly closer to the host star compared with Solar system models, which place the H$_2$O snowline between $\sim$$2.7$ to $\sim$$3.1$ AU (\citealt{hayashi81}, \citealt{podolak04}, \citealt{martin12}).
-This is partially because we choose a colder disk model, as well as the fact that gas accretion rates decrease over time, moving the snowline location inwards (see also \citealt{garaud07} and Section \ref{sec:neglected}).
-\citet{min11} find that the location of the H$_2$O snowline is highly sensitive to the gas mass accretion rate $\dot{M}$ (equal to $10^{-8} M_{\odot}$ yr$^{-1}$ in our model) and the dust opacity $\kappa$ (Equation \ref{eq:opacity}).
-Higher values of $\dot{M}$ and $\kappa$ would increase the accretional component of the disk temperature (cf.
-Equation \ref{eq:Tdact}), which would push the H$_2$O snowline in the viscous disk outwards to match the Solar system snowline.
-At the same time, the snowline location in Solar type stars may be as close as $\sim$1 AU \citep{mulders15}, further in than the H$_2$O snowline in our Solar system.
-
- Observations of the CO snowline in TW Hya \citep{qi13} have found its location at a disk midplane temperature of 17 K (at 30 AU for the TW Hya specific temperature profile).
-The inferred  desorption temperature corresponds to the CO desorption temperature in a static disk, or to desorption from very small grains in an evolving disk, i.e.
-from grains that are too small to drift substantially.
-This suggests that the outer TW Hya disk is dominated by small grains, since larger particles would push the snowline location inwards, and therefore to higher desorption temperatures.
-This may seem contradictory to observations of grain growth in disks in general and in TW Hya in particular \citep{wilner00}.
-However, recent observations have revealed that grain growth is concentrated to the inner disk \citep{perez12} and outer disk snowlines may therefore be close to the ones expected in a static disk.
-
-%\emgr{There will be 1-2 additional sentences for the third panel, the steady-state disk, once that finishes running.}
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.5\textwidth]{lala}
-%%\vspace{-0.5in}
-%\caption{C/O ratios for different particle size distributions...
-\textbf{Figure belongs to section 4.} (\emgr{It will probably be a multi-panel plot, for different particle size distributions, for both passive and active disk, and perhaps at different times for the active disk.
-Not yet unsure how that figure will look like since I don't have the results yet, so I'll leave the figure caption open for now.})}
-%\label{fig:...}
-%\end{figure}
-
-%\emgr{Motivate the fact that we can calculate a sharp, fixed snowline for each particle size by inserting the plot that shows that particles desorb almost instantaneously at a fixed distance.
-Then present the plots analogous to Fig.
-1 in \"Oberg et al.
-(2011) for different particle sizes, based on the snowline locations obtained in the previous section, both for passive and active disk.
-Perhaps show it at different times in the gas disk evolution (i.e., not just at 3 Myr) for the active disk.
-Then assume a particle size distribution and show the interpolated result for the C/O ratio.
-Generalize the result using a transition disk (this part might also fit in the discussion section).}
-
-%emgr{Present the flux equations used to keep track of the amount of C and O in gas and dust throughout the disk.
-Motivate the simplification of using a fixed particle sized, fixed timescale, and fixed desorption distance in the calculations by referring to the rainbow plot in the previous section and by inserting the plot that shows that particle desorb almost instantly at a fixed distance.
-Based on these assumptions, finally show the equivalent of Fig.
-1 in \"Oberg et al.
-(2011) for different particle sizes, and at different times in the gas disk evolution (i.e., not just at 3 Myr) --- a multi-panel plot could be a good idea.
-Discuss the result, trends, etc.
-Ideally, have a final plot showing the results for a particle size distribution rather than for individual particles.}
-
-\section{Discussion}
-\label{sec:discussion}
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=\textwidth]{time_plots.pdf}
-%\vspace{-0.5in}
-\caption{Desorption or final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve).
-For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
-\label{fig:timeplots}
-\end{figure*}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{desorption_distance_varying_Md.pdf}
-%\vspace{-0.5in}
-\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines).
-The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve).
-A lower disk mass does not change the snowline location.}
-\label{fig:varMd}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{desorption_distance_transition_disk_1000.pdf}
-%\vspace{-0.5in}
-\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines).
-The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}.
-Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge, due to being trapped in a pressure maximum.}
-\label{fig:cavity}
-\end{figure}
-
-\subsection{Generality of Results: Dependence on Disk Parameters}
-\label{sec:incond}
-
-In this section we investigate how variations in our fiducial parameters, the total disk mass, disk age, and disk structure, %the time in the disk evolution at which we determine 
-affect the calculated snowline locations and the C/O ratio. %, affect our results. 
-All previous results assumed a disk lifetime $t_{\rm d}=3$ Myr%is the most representative for our calculations, as this is the timeframe in which giant planets accrete their gaseous atmospheres 
-, the typical disk lifetime and the expected time scale for giant planets to accrete their gaseous atmopsheres (e.g., \citealt{pollack96}, \citealt{piso14}).
-Some gas accretion may occur at earlier times, however,%However, protoplanetary cores start acquiring their envelope at earlier times,
-before the core is fully formed (e.g., \citealt{rafikov06}).
-Recent models such as aerodynamic pebble accretion \citep{lambrechts12} suggest that rapid core growth on timescales of $10^5$ years is possible.
-% which implies that the time when cores accrete their massive envelope may be at least an order of magnitude shorter than the disk lifetime.
-The composition of giant planet atmospheres, and specifically their C/O ratio, can thus depend on the abundance of H$_2$O, CO$_2$ and CO in gas at earlier times than $t_{\rm d}$ in the disk evolution.
-
-Figure \ref{fig:timeplots} shows the particle desorption or final distance as a function of a particle's initial location in the disk, for ice particles %grains 
-of initial sizes of 10 cm and 1 m, composed of either H$_2$O, CO$_2$ or CO.
-These sizes are important since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
-We choose the evolving disk as a disk model, and we stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively.
-%, to see how the snowline locations evolve with time and affect the C/O ratio.
-The most important result of these plots is that particles of a given size always desorb at the same disk radii, the 3 Myr snowline, regardless of simulation stopping time.
-Particles %of a given size
-that start at large stellocentric distances do not desorb within the shorter timeframes, e.g.
-$10^4$ or $10^5$ years, but they do evaporate at a fixed radius if their initial location is closer to the host star.
-While the amount of material that moves through the disk changes with time, the radius at which particles desorb and
-the snowline locations are thus independent of the time elapsed, and %. Therefore, 
-our results for the snowline locations are %not only valid at $t_{\rm d}$, but also throughout the earlier 
-valid throughout the time evolution of the protoplanetary disk.
-
-We choose as a fiducial model a total disk mass $M=0.1 M_{\odot}$.
-Observationally, disk masses span at least an order of magnitude around Solar type stars %, but this number is not universally valid.
-From observations of the dust continuum,
-\citep{andrews13}.
-% find that a linear scaling $M \propto M_*$ is reasonable.
-Giant planets, however, have been detected around small stars (e.g., \citealt{montet14}), which can have masses as low as $M_* \sim 0.1 M_{\odot}$.
-We thus explore the effect of disk mass on the location of snowlines.
-Figure \ref{fig:varMd} shows the desorption or final distance as a function on the initial location of a H$_2$O particle with initial size of 1 m, for two total disk masses: $M=0.1 M_{\odot}$, our fiducial model, and $M=0.01 M_{\odot}$.
-Similarly to Figure \ref{fig:timeplots}, we perform our calculations for an evolving disk.
-The simulations are stopped after the same timeframes as those in Figure \ref{fig:timeplots}.
-The location of the H$_2$O snowline is the same for both disks (the same holds true for the CO$_2$ and CO snowlines).
-The C/O ratio is thus insensitive to the choice of $M$.
-We note that our conclusions regarding the disk age and total mass are only valid if the snowline itself does not move with time or disk mass (see also Sections \ref{sec:disk} and \ref{sec:neglected}).
-
-%This result also has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
-Since the snowline locations are independent of disk mass, we expect transition disks to exhibit the same drift-desorption behavior as standard disks under the simplified assumptions of our model.
-%[Comment on differences, i.e. short time frames there is a difference because of smaller drift velocity in the lighter disk I presume.]
-
-We also apply our evolving disk model to a transition disk, i.e.
-a protoplanetary disk with an inner cavity significantly depleted of gas.
-We choose a disk with an inner gap of radius $r_0=4$ AU, consistent with observations of TW Hya \citep{zhang13}, and with the gas surface density in the gap reduced by a factor of 1000.
-Figure \ref{fig:cavity} shows the desorption or final distance for a H$_2$O particle of initial size of 1 m, with the simulation stopped at the same timescales as in Figures \ref{fig:timeplots} and \ref{fig:varMd}.
-Particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the gap edge, due to the decrease in gas pressure inside the cavity, thus forming a snowline at $\sim$$3.8$ AU.
-This is qualitatively consistent with the observations of \citet{zhang13}, which show that the H$_2$O snowline is pushed outwards in a transition disk compared to a full disk.
-Our model framework is thus generally valid for more complicated disk structures as well.
-
-%[Need a sentence about how this makes the model framework also generally valid for more complicated disk structures.]
-
-\subsection{Model Extensions}
-\label{sec:neglected}
-
-Our goals in this paper were (1) to gain a qualitative and quantitative understanding of the effect of radial drift and gas accretion onto the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
-We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical processes.
-In what follows, we discuss these limitations and their effects.
-We note that our future work will address some of these issues.
-
-We summarize in Table 1 the potential physical and chemical processes occurring in disks and their effect on snowline locations compared to a static disk.
-For the sake of completeness, Table 1 also includes the processes addressed in this paper, i.e.
-radial drift and gas accretion.
-The neglected effects are discussed in more detail below.
-
-\begin{enumerate}
-\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered initially fixed for a given grain before it drifts and desorbs, since we do not take into account particle coagulation.
-However, grain growth has been observed in protoplanetary disks (e.g., \citealt{ricci10}, \citealt{perez12}), as well as theoretically constrained (e.g., \citealt{birnstiel10}, \citealt{birnstiel12}).
-In Section \ref{sec:COratio} we have shown that larger grains move the snowline locations closer in, but those locations remain fixed above a certain particle size.
-Particle growth will thus initially push the snowlines inwards.
-This is consistent with particle growth models, which predict a maximum particle size often around or below the particle sizes that drift the fastest \citep{birnstiel12}.
-As the largest grains contain most of the solid mass, grain growth models should produce snowlines corresponding to our snowline location estimates for the largest grains in the particle size distribution.
-However, once the solids grow larger than km-sized and form planetesimals, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
-%It follows that grain growth will eventually push the snowline location outwards.
-
-\item \textbf{Turbulent diffusion.} The radial drift model presented in Section \ref{sec:drift} only considers a laminar flow and thus ignores turbulence.
-However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
-Turbulence causes eddies and vertical mixing, which are likely to reduce the radial drift velocity of the solids (e.g., \citealt{youdin07}).
-Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially.
-Back-diffusion across the snowline will change the shape of the snowline, as well as the C/O ratio in gas and dust both inside and outside of the snowline, due to the reduction of gas-phase volatile abundance interior to the snowline.
-
-\begin{table}[t!]
-\caption{The effects of dynamical and chemical processes on snowline shapes and locations}
-\begin{center}
-\begin{tabular}{|l|l|}\hline
-\textbf{Process} & \textbf{Effect} \\\hline
-Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star compared to the static snowline, $\rightarrow$ means that the snowline is pushed further from the host star compared to the static snowline.
-The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
-Gas accretion & $\leftarrow$ \footnote{Gas accretion pushes the snowlines inwards compared to the snowline locations in a static disk.
-However, accretional heating may push the snowline outwards compared to an evolving disk without accretion heating.} \\\hline
-Particle growth & $\leftarrow$ \footnote{As stated in the main text, if particles grow to km-sizes and above, the snowline remains the same as that of a static disk.} \\\hline
-Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
-Particle fragmentation & $\rightarrow$ $\leftarrow$ \\\hline
-Grain morphology & $\rightarrow$ \\\hline
-Particle composition & $\rightarrow$ $\leftarrow$ \\\hline
-Disk gaps and holes & $\rightarrow$ \\\hline
-Accretion rate evolution & $\rightarrow$ $\leftarrow$ \\\hline
-Stellar luminosity evolution & $\leftarrow$ \\\hline
-Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
-\end{tabular}
-
-\end{center}
-\end{table}
-%causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.
-
-\item \textbf{Particle fragmentation.} Frequent particle collisions in disks cause them to fragment (e.g., \citealt{birnstiel12}).
-The fragmentation of meter- to km-sized particles will move the snowlines outwards, as smaller particles desorb faster and further out from the host star (cf.
-Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}).
-Large boulders, which neither drift nor desorb, may become e.g.
-meter-sized due to collisions and subsequent fragmentation, which will cause them to drift significantly before desorbing, pushing the snowlines inwards.
-Thus fragmentation can move the snowline locations in either radial direction --- specifically, fragmentation leads to a certain grain size distribution, and the largest particles in this size distribution are the ones that determine the position of the snowline.
-
-\item \textbf{Grain morphology.} Our model assumes that the ice particles are perfect, homogeneous spheres.
-However, this is not a very good approximation, since grain growth can be fractal rather than compact (\citealt{zsom10}, \citealt{okuzumi12}, \citealt{krijt15}).
-The inhomogeneity due to cracks in the grain structure will cause the particles to desorb faster.
-They will therefore drift less before evaporating and will move the snowlines less far inward.
-
-\item \textbf{Particle composition.} The ice particles in our model are assumed to be fully made of either H$_2$O, CO$_2$ or CO.
-More realistically, grains may have a layered structure, such as an interior composed of non-volatile materials (e.g., sillicates) covered by an icy layer.
-The ice thus only constitutes a fraction of the total particle mass, which accelerates its desorption and pushes the snowlines outwards.
-The grains may also be composed of a mixture of H$_2$O, CO$_2$ and CO ices, which will increase the binding energies of the more volatiles species, moving %.
-In this case, the desorption energies are largest for the most volatile species, which will move
-the snowlines inwards.
-
-%The relative fraction of each volatile, as well as their degree of mixing, will determine their desorption timescale, and as a result whether the snowlines are moved towards or away from the central star.
-
-\item \textbf{Disk gaps and holes.} The snowline locations will be different for transition disks, which have inner cavities significantly depleted of gas (e.g., \citealt{espaillat12}, \citealt{vandermarel15}), or pre-transitional disks, which have a gap between an inner and outer full disk (e.g., \citealt{kraus11}).
-The decrease in gas pressure in these gaps or holes will reduce the particles' drift velocity close to the gap edge, thus slowing them down and pushing the snowline outwards.
-
-\item \textbf{Accretion rate evolution.} Our viscous disk model assumes a constant mass accretion rate $\dot{M}$.
-However, $\dot{M}$ decreases over time, which lowers the accretional component of the disk temperature (Equation \ref{eq:Tdact}), thus pushing the snowline location inwards if the disk is optically thick (\citealt{garaud07}).
-Once $\dot{M}$ reaches low enough values for the snowline to become optically thin, the snowline location moves outwards \citep{garaud07}.
-\textbf{During the giant planet formation stage of a few Myr, however, $\dot{M}$ steadily decreases with time \citep{chambers09}, which may result in the inward movement of the H$_2$O snowline by up to one order of magnitude, significantly larger than the inward movement caused by radial drift (cf.
-Section \ref{sec:COratio}).
-We thus acknowledge that the location of the H$_2$O snowline may be set by the mass accretion rate evolution rather than the drift of solids.}
-
-%Our choice of $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ roughly corresponds to the mass accretion rate during the quasi-static accretion phase of the disk evolution predicted by \citet{garaud07}.
-This is the timescale on which giant planets accrete most of their gaseous envelope.%; thus the implications of our results for the atmospheric composition of gas giants are robust.
-
-%our results for the snowline location and C/O ratio in a steady-state active disk will only be weakly affected by
-
-\item \textbf{Stellar luminosity evolution.} As the host star contracts during its pre-main sequence phase, its luminosity decreases, which reduces the disk temperature and pushes the snowline locations inwards.
-\citet{kennedy06} found that the snowline is unlikely to move significantly during the pre-main sequence phase for Solar type stars, but it may move inward by a factor of $\sim$$15-20$ for $M_* \sim 0.25 M_{\odot}$ due to the stellar contraction.
-
-\item \textbf{Time dependent chemistry.} As the goal of this paper was to explore only the dynamical effects on snowline locations and the C/O ratio in disks, we have assumed a simple, static chemical model.
-In reality, the chemistry in most of the disk is expected to be time-dependent.%However, gas-grain chemistry is very complex and time-dependent.
-In the inner disk, chemistry approaches equilibrium due to intense sources of ionizing radiation (e.g., \citealt{ilgner04}), while in the outer disk high energy radiation and cosmic rays are the key drivers of chemistry, which is no longer in equilibrium (e.g., \citealt{vandishoeck06}).
-A multitude of chemical evolution models have been developed (see references in \citealt{henning13}), many of which contain tens or hundreds of chemical reactions.
-Due to the complexity of these chemical models, most of them are decoupled from disk dynamics.
-The effect of disk chemistry on snowline locations, shape, time evolution, or the C/O ratio is therefore difficult to estimate.
-%In a future paper, we plan to use a parametrized chemical model and incorporate it in our radial drift calculation.
-[let's not promise this until we have figured out how to do it.]
-
-%parametrize the chemical model developed by Merchantz et al. (in preparation) that traces the evolution of the main C and O oxygen carriers,
-
-\end{enumerate}
-
-%\emgr{Present the diagram that shows all the effects that can modify snowline location.
-For model limitations, include: non-inclusion of turbulence, assumption of perfect spheres when in fact they may have cracks, particles composed of a single volatile when in reality they are likely to be mixed, etc.
-Discuss uncertainty of initial conditions and estimate how much they matter.
-....}
-
-\section{Summary}
-\label{sec:summary}
-
-We study the effect of radial drift of solids and viscous gas accretion onto the central star on the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in a protoplanetary disk, assuming static chemistry.
-We develop a simplified model to describe the coupled drift-desorption process and determine the time evolution of particles of different sizes throughout the disk.
-We assume that the solid particles are perfect, homogeneous spheres, fully composed of either H$_2$O, CO$_2$ or CO.
-We apply our model to an irradiated disk, an evolving disk, and a viscous disk that also takes into account stellar irradiation.
-We determine the desorption or final location of drifting particles after a time equal to the disk lifetime, and use this result to set an inner limit for the location of the H$_2$O, CO$_2$ and CO snowlines.
-Our results can be summarized as follows:
-
-\begin{enumerate}
-\item Radial drift and gas accretion affect desorption and move the snowline locations inward compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk.
-%Moreover, boulders up to almost kilometer-size contribute to setting the snowline locations if they are very close to the snowlines.
-[I find this conclusion a bit iffy and also it sounds like it is in contradiction with the previous sentence.]
-
-\item For our simplified model that does not account for the effects outlined in Section \ref{sec:neglected}, particles with sizes in the above range desorb almost instantaneously once desorption has begun, and at a fixed location in the disk that only depends on the particle size and the gas accretion rate.
-Thus for each particle size there is a fixed and uniquely determined H$_2$O, CO$_2$ or CO snowline.
-
-\item The results of our numerical simulation are in agreement with the analytic solution of the drift-desorption system of differential equations if the stopping time $\tau_{\rm s} \ll 1$.
-We present an explicit analytic solution for the desorption distance in this regime.
-
-\item Since realistic grain size distributions are dominated in mass by the largest particles, the H$_2$O, CO$_2$ and CO snowlines are those created by the largest drifting particles in our model.
-This corresponds to the innermost snowlines that we determine.
-Our model thus sets a limit on how close to the central star the snowlines can be pushed by radial drift and gas accretion.
-
-\item The snowline locations move inwards as the particle size increases; the innermost snowline is set by particles with initial size $s \sim 7$ m in our model --- bigger particles drift too slowly to make it further in before desorbing (see Section \ref{sec:snowlines}).
-Gas accretion causes even micron-sized particles to drift, desorb and move the snowline location compared to a static disk.
-A viscous disk that includes accretion heating moves the H$_2$O snowline outwards compared to an evolving disk, but has no effect on the CO$_2$ and CO snowline locations, for our particular choice of mass accretion rate $\dot{M}$ and midplane opacity $\kappa$.
-%However, the magnitude of the effect of accretional heating on snowline locations will depend on .
-
-\item For our fiducial model, which considers particles with sizes between $10^{-3}$ and $10^8$ cm, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for an irradiated disk, 0.26 AU, 3.4 AU and 30 AU for an evolving disk, and 0.63 AU, 3.4 AU and 30 AU for a viscous disk with accretion heating.
-Compared to a static disk, radial drift and gas accretion move the snowlines by up to 60 \% for H$_2$O and CO$_2$, and by up to 50 \% for CO.
-For the viscous disk, however, which is the most realistic of the three models since it takes into account accretion heating, the H$_2$O snowline location moves inwards by up to 40 \%.
-
-\item Our C/O estimates confirm the conclusions of \citet{oberg11} that the C/O ratio \textbf{in gas} may be enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching its maximum value between the CO$_2$ and CO snowlines.
-%This is consistent with possible detections of superstellar C/O ratios in some exoplanet atmospheres.
-We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the evolving and viscous disks, which may significantly decrease the C/O ratio in gas inside the H$_2$O and CO$_2$ snowlines.
-We plan to address this issue in a future paper.
-
-\item For a constant gas mass accretion rate $\dot{M}$ and stellar luminosity $L_*$, the snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk midplane remains optically thick.
-%Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.
-
-\end{enumerate}
-
-Our model does not address additional effects, such as gas diffusion, grain composition and morphology, or complex time-dependent chemical processes.
-Future work will address some of these dynamical and chemical processes, with the goal of obtaining more realistic results for the snowline locations, shapes and time evolution, and the resulting effect on the C/O ratio.
-%Nevertheless,
-
-\appendix
 \section{Steady-state active disk solution} \label{app:steadystate}
@@
 This is a system of nine equations with nine unknowns ($\nu$, $c$, $H$, $T_{\rm act}$, $\rho$, $\Sigma$, $\tau$, $T_{\rm surf}$, $\kappa$) that can be solved numerically once $\alpha$ and $\kappa_0$ are specified.
-
-\section{Desorption distance analytic solution} \label{app:tdriftan}
-
-For a particle of size $s$ that desorbs and satisfies $\tau_{\rm s} \ll 1$ ($\tau_{\rm s}$ is the dimensionless stopping time, defined in Section \ref{sec:drift}), we can derive an explicit analytic solution for the particle's desorption distance in an irradiated disk.
-For $\tau_{\rm s} \ll 1$, a particle is in the Epstein drag regime (see Equation \ref{eq:ts}) and its drift velocity $\dot{r}$ (Equation \ref{eq:rdotpas}) can be approximated as
-\begin{equation}
-\label{eq:rdotapp}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \tau_{\rm s}.
-\end{equation}
-By using Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and setting $t_{\rm drift}=t_{\rm des}$, we can express a particle's desorption distance as
-
-\begin{equation}
-\label{eq:tan}
-r_{\rm des}=\Bigg(\frac{d}{q \,C} \,\, \mathcal{W}\Big[\frac{(B/A)^{-q/d} q\,C}{d}\Big]\Bigg)^{\frac{1}{q}},
-\end{equation}
-where $\mathcal{W}$ is the Lambert-W function, $q=3/7$ is the power-law coefficient in Equation (\ref{eq:diskT}), $d=-\frac{1}{2}+p-q$ with $p=1$ the power-law coefficient in Equation (\ref{eq:disksigma}), and
-\begin{subeqnarray}
-%\label{eq:tan}
-A & = &  \frac{\rho_0}{\rho_{\rm s}} \frac{r_0^2}{s c_0} r_0^d \\ 
-B & = & \frac{\rho_{\rm s} s}{3 \mu_{\rm x} N_{\rm x} \nu_{\rm x}} \\
-C & = & \frac{E_{\rm x}}{T_0} r_0^{-q}, \\
-\end{subeqnarray}
-where $r_0=1$ AU, $\rho_0=\rho(r_0)$ and $c_0=c(r_0)$.
-
-%
-%\emgr{Right now it's unclear to me what could go in an appendix, if anything.
-Maybe discuss a bit the algorithm to evolve $\Sigma_{\rm p}$ (although it is already explained in detail in the appendix of Birnstiel et al.
-2010).
-Maybe show some example profiles of $\Sigma_{\rm p}$ at different times and for different particle sizes.}
-
-\if\bibinc n
-\bibliography{refs}
-\fi
-
-\if\bibinc y
-\begin{thebibliography}
-\end{thebibliography}
-\fi
-
-\end{document}

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/appendixG.tex
upd

--- 2015-11-13 tex/draft/ms_final_revised_2.tex
@@
-%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
-%\documentclass[12pt, preprint]{aastex}
-\documentclass[apj]{emulateapj}
-
-\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
-\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex
-
-%\usepackage{pdfsync}
-\usepackage{subeqnarray}
-\usepackage{natbib}
-\usepackage{color}
-\usepackage[utf8]{inputenc}
-
-\bibliographystyle{apj}
-
-\newcommand{\ie}{i.e.\ }
-\newcommand{\eg}{e.g.\ }
-\newcommand{\p}{\partial}
-\newcommand{\brak}[1]{\langle #1\rangle}
-
-\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
-\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
-\newcommand{\cm}{\; {\rm cm}}
-\newcommand{\mm}{\; {\rm mm}}
-%\newcommand{\ps}{\; {\rm s^{-1}}}
-\newcommand{\km}{\; {\rm km}}
-\newcommand{\au}{\; \varpi_{\rm AU}}
-\newcommand{\AU}{\; {\rm AU}}
-\def\K{\; {\rm K}}
-
-\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
-\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
-\newcommand{\nab}{\vc{\nabla}}
-\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
-\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
-\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}
-
-\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
-\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
-\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
-\newcommand{\App}[1]{Appendix~\ref{#1}}
-\newcommand{\Sec}[1]{Sect.~\ref{#1}}
-\newcommand{\Chap}[1]{Chapter~\ref{#1}}
-\newcommand{\Fig}[1]{Fig.~\ref{#1}}
-\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
-\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
-\newcommand{\Tab}[1]{Table \ref{#1}}
-
-\definecolor{gray}{gray}{0.5}
-\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}
-
-%\newenvironment{packed_item}{
-%\begin{itemize}
-%  \setlength{\itemsep}{1pt}
-%  \setlength{\parskip}{0pt}
-%  \setlength{\parsep}{0pt}
-%}{\end{itemize}}
-
-\begin{document}
-
-%\slugcomment{Draft Modified \today}
-
-\title{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}
-
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1}, Ruth A. Murray-Clay\altaffilmark{2}}
-\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
-
-\begin{abstract}
-%The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry. 
-The C/O ratio is a defining feature of both gas giant atmospheric and protoplanetary disk chemistry.
-In disks, the C/O ratio is regulated by the presence of snowlines of major volatiles at different distances from the central star.
-We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk.
-We determine the snowline locations for a range of fixed initial particle sizes and disk types.
-For our fiducial disk model, we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk, and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving and viscous disk, desorb %instantaneously and [I think instantaneously is too strong]
-at a %fixed 
-size-dependent location in the disk, which is independent %regardless 
-of the particle's initial position.
-The snowline radius decreases for larger particles, up to sizes of $\sim$7~m.
-%location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m.
-Compared to a static disk, we find that radial drift and gas accretion in a viscous disk move the H$_2$O snowline inwards by up to 40 \%, the CO$_2$ snowline by up to 60 \%, and the CO snowline by up to 50 \%.
-We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
-%Our results have direct implications on the C/O ratio in giant planet atmospheres that have formed in situ.
-[I don't think we discuss this enough to warrant it as a conclusion
-\end{abstract}
-
-\section{Introduction}
-\label{sec:intro}
-
-%\emgr{Background topics: importance of atmospheric chemistry in providing constraints on the formation of giant planets; C/O ratio as important signature of atmospheric chemistry; C/O ratios observationally determined are different from interstellar --- one explanation is the different abundance in gas and dust form of the main C and O carriers, H$_2$O, CO$_2$ and CO, between their respective snowlines (cite \"Oberg et al.
-2011).}
-
-%\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture.
-Gas giants are important.
-The chemical composition of their atmospheres constraints their formation and evolution.
-Is is therefore important to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) backtrack the planet formation location based on planet composition.}
-
-The chemical composition of protoplanetary disks affects planet formation efficiencies and the composition of nascent planets.
-%[Need explicit statement that gas giants accrete their atmospheres from the nebular gas.] 
-Gas giants accrete their envelopes from the nebular gas.
-As such, planet compositions are tightly linked to the structure and evolution of the protoplanetary disk in which they form.
-It is thus essential to understand the disk chemistry and dynamics well enough to (1) predict the types of planet compositions that result from planet formation in different parts of the disk, and (2) backtrack the planet formation location based on planet compositions.
-%[split into two sentences]
-
-%\emgr{Add the rest as outlined above.} \\
-
-%\emgr{Paragraph 2: disks are complex, a lot of dynamical and chemical processes going on.
-However, we have detections of organic molecules.
-We have detections of snowlines.
-We see complex chemistry and many molecules we can study.}
-
-%In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11c}, \citealt{oberg11b}).
-[the molecules relevant for this paper are not organics]
-%[Topic sentence on that the most important volatiles are ... based on protostellar and comet abundances.]
-
-The structures of protoplanetary disks are complex, and affected by a multitude of chemical and dynamical processes (see review by \citealt{henning13}).
-From the chemistry perspective, volatile compounds are particularly important.
-Their snowline locations determine their relative abundance in gaseous and solid form in the disk,.
-Based on protostellar and comet abundances, some of the most important volatile molecules are H$_2$O, CO$_2$, CO, N$_2$.
-Recent observations of protoplanetary disks have provided valuable information about the abundances and snowline locations of some of these compounds.
-For example, the CO snowline has been detected in the disk around TW Hya \citep{qi13}, as well as in the disk around HD 163296 \citep{mathews13} using line emissions from DCO$^+$.
-Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}, and more such snowline detections are expected in future ALMA cycles.
-These observations are currently lacking an interpretive framework that takes into account all important dynamical and chemical processes.
-Furthermore, such a framework is crucial to connect observed snowline locations to planet formation.
-
-%Of particular importance are volatile compounds, such as H$_2$O, CO$_2$, CO, N$_2$ etc.
-Recent detections of the snowlines of volatiles (such as the CO snowline, \citealt{qi11}) are significant, since the snowline locations of these molecules determine their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets.
-[I think this paragraph belongs further down after you have discussed planets since that is the topic of your intro paragraph above.
-Also need a bit more meat here.
-So far a single CO snowline detection and a single H2O snowline detection.
-More are expected and an interpretive framework is key regardless of the importance for planets.]%Recent detections of the CO snowline in disks (e.g., \citealt{qi11}) provide the framework...
-
-%\emgr{Add part about snowlines, complex chemistry, many molecules.} \\
-
-An important consequence of snowline formations in disks is that disks are expected to present different carbon-to-oxygen (C/O) ratios in the gas and in icy dust mantles at different disk radii.
-This effect was quantified by  \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e.
-H$_2$O, CO$_2$ and CO, have different condensation temperatures.
-This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
-\citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-This effect was used to explain claims of detections of superstellar C/O ratios in exoplanet atmospheres (e.g., WASP-12b, \citealt{madhu11}), which however have been unambiguously refuted (\citealt{stevenson14}, \citealt{kreidberg15}).
-%Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.
-
-%Given the tight connection between planet and disk compositions, detections of C/O ratios in exoplanet atmospheres can provide some insight into the relative abundance of C and O in disks.
-Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}.
-One explanation for this discrepancy was proposed by \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e.
-H$_2$O, CO$_2$ and CO, have different condensation temperatures.
-This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above.
-\citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.
-%[This is not a very natural approach I think, i.e.
-to use a dicy eco-planet C/O detection to infer something about the disk chemical composition.]
-
-%\emgr{Paragraph 3: let's focus on one (set of) molecules: the C/O ratio.
-It's an important signature of atmospheric chemistry.
-While we cannot detect it directly in disks, we have observations in giant planet atmospheres.
-It's different than stellar.
-Why?
-Possible explanation based on snowlines.}
-
-%\emgr{Start with the fact that we can't see C/O in disks directly. Then transition to planet atmospheres and modify the next sentence to follow up logically.} 
-%Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio.
-
-\citet{oberg11} assumed a static disk with no chemical evolution.
-In reality, dynamical and chemical processes affect the snowline locations and the resulting C/O ratio.
-Several works have addressed some of these effects.
-\citet{madhu14} use a steady-state active disk model that includes planetary migration and use the C/O ratio to constrain migration mechanisms.
-\citet{alidib14} calculate the C/O ratio throughout the disk by incorporating the evolution of solids, i.e.
-radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors.
-\citet{alidib14} use the 1+1D $\alpha$-disk model of \citet{hughes10}, in which the gas drifts outwards in the disk midplane, and thus small particles that are well-coupled to the gas will also advect outward.
-Their model assumes a cyclical conversion between H$_2$O or CO dust and vapor: large enough particles that are decoupled from the gas drift inwards and start desorbing.
-Once their sublimation is complete, back-diffusion moves the H$_2$O or CO vapor ouwards to their respective snowlines, where they instantly condense into mm-sized particles that diffuse outwards with the gas while coagulating into larger particles.
-Once the grains become large enough to decouple from the gas and drift inwards, the cycle restarts.
-This ```conveyor belt''' model is based on the pioneering work by \citet{cuzzi04}, and \citet{ciesla06} for the evolution of H$_2$O in a viscous disk.
-This approach leads \citet{alidib14} to find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.
-\citet{thiabaud15} consider additional carbon carrier volatile species in their chemical network, such as CH$_4$, and find that the gas C/O ratio may be enriched by up to four times the Solar value in the outer parts of the disk where CH$_4$ and CO are the only gaseous carriers of C and O.
-They also include nitrogen carriers such as N$_2$ or NH$_3$, and perform similar calculations for nitrogen.
-
-%Most studies of the C/O ratio in disks are particularly focused on the inner disk, in order to explain observations of super-stellar C/O ratios in the atmospheres of close-in exoplanets \citep{madhu11}.
-In addition, the simultaneous inclusion of various dynamical and/or chemical processes makes it difficult to quantify the separate effect of each of these processes on snowline locations and the C/O ratio.
-%[The preceding paragraph needs to be made clearer.
-Currently it is not clear what effects these studies have had on our understanding of C/O ratio sin disks and why your study is needed.
-I.e.
-why are the processes you focus on the most important ones?
-Introduce separately papers that deal with C/O and water snowline locations.]
-Each of these studies have considered a specific combination of dynamical and chemical effects.
-One scenario that has not yet been considered is the combination of radial drift and viscous gas accretion in isolation.
-Studying these two dynamical processes makes it possible to quantify their separate effect on snowline locations and the C/O ratio at various disk radii.
-
- %In this paper, we focus on these two particular dynamical processes.
-%i.e.
-the radial drift of solids and viscous gas accretion.
-%We enhance the model of \citet{oberg11} by considering these two additional processes.
-In this paper, we perform a systematic study to understand the detailed qualitative and quantitative effects of radial drift and gas accretion on the H$_2$O, CO$_2$ and CO snowline locations, and the resulting C/O ratio in gas and dust throughout the protoplanetary disk.
-More importantly, we obtain a limit on how close to the star the snowline locations can be pushed by radial drift and gas accretion.
-
-%\emgr{Paragraph 4: the static disk model is a great start, but there are additional dynamical and chemical processes to take into account.
-Several studies have done that in some form or another --- cite Ali-Dib+14, Madhusudhan+14, Thiabaud+15.
-Briefly discuss their assumptions.} \\
-
-%\emgr{Paragraph 5: finally introduce our goal for this paper: study the dynamical effects such as drift and gas accretion.
-State our goals: understand the detailed qualitative and quantitative effect of drift and gas accretion on snowline locations, obtain a limit on how far in can the snowlines be pushed, and see how that affects the C/O ratio throughout the disk.
-State that our specific goals motivate the use of a simplified model (in other words, why what we are doing is different from the studies cited in the previous paragraph).
-}
-
-%\emgr{This needs some restructuring and additions along the lines discussed above.} In order to obtain more realistic estimates for C/O ratios across protoplanetary disks, dynamical processes and the disk evolving chemistry have to be taken into account.
-In this paper, we enhance the model of \citet{oberg11} considering two additional dynamic effects: (1) the radial drift of solids throughout the protoplanetary disk, and (2) the viscous accretion of the disk gas onto the host star.
-Our goal is two-fold: (1) to quantify the effect of radial drift of solids of different sizes on the location and shape of H$_2$O, CO$_2$ and CO snowlines, and (2) to calculate the resulting C/O ratio in gaseous and solid form throughout an actively accreting protoplanetary disk as a function of the grain size distribution and the evolutionary time of the nebula.
-
-%\emgr{Describe goals and approach of this paper, i.e.: study the importance of radial drift on the location of snowlines in protoplanetary disks and how radial drift affects the C/O ratio.
-Mention that disks are viscously accreting in the early stages of planet formation, hence gas accretion has to be taken into account.
-The goal of this paper is two-fold: (1) estimate using both numerical models and analytic arguments the range of particle sizes for which radial drift changes snowline location; (2) calculate the C/O ratio for various particle sizes (and possibly a size distribution) and at different times throughout the evolution of the gas disk.}
-
-This paper is organized as follows.
-In Section \ref{sec:model}, we present our disk, radial drift and desorption models, as well as the timescales relevant to the coupled drift-desorption process.
-We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for an irradiated and an evolving disk in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}.
-In Section \ref{sec:discussion}, we discuss the generality of our results, as well as additional effects on the snowline locations.
-Finally, we summarize our findings in Section \ref{sec:summary}.
-
-\section{Model Framework} 
-\label{sec:model}
-
-We present our protoplanetary disk model for a static, an irradiated, an evolving, and a viscous disk in section \ref{sec:disk}.
-In section \ref{sec:drift}, we describe our analytic model for the radial drift of solids.
-We summarize our ice desorption model in section \ref{sec:desorption}.
-Finally, we discuss the relevant timescales for dynamical effects that affect snowline locations %in the desorption process
-in section \ref{sec:timescales}.
-
-\subsection{Disk Model}
-\label{sec:disk}
-
-To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{static disk}, which is solely irradiated by the host star and does not take into account gas accretion onto the star or radial drift; \textit{irradiated disk}, which has the same temperature profile as the static disk and does not experience gas accretion or accretional heating, but it takes into account radial drift of solids; \textit{evolving disk}, in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; and \textit{viscous disk}, for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis, and the temperature profile is calculated using both accretional heating and stellar irradiation.
-
-%[Perhaps define your four disk models up here and motivate why you are exploring several different ones, and then go through and describe them in detail.
-That would save you the slightly awkward Active disk with passive temperature profile.]
-
-\textbf{Static and Irradiated disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a static and and an irradiated disk similar to the prescription of \citet{chiang10}.
-The gas surface density and midplane temperature are
-\begin{subeqnarray}
-\label{eq:disk}
-\Sigma&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
-T &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
-\end{subeqnarray}
-where $r$ is the semimajor axis.
-Our surface density profile is flatter than the $\Sigma \propto r^{-3/2}$ used by \citet{chiang10}.
-Our choice is inspired by observations of protoplanetary disks at radii larger than $\sim$20 AU (e.g., \citealt{andrews10}), which suggest that typical disks may have surface density profiles with $\Sigma \propto r^{-1}$.
-A slope flatter than $\Sigma \propto r^{-3/2}$ is also more consistent with the temperature profile for a steady-state gas disk (see the Viscous disk heading below and \App{app:steadystate}).
-We use the static disk model to compare our results with those of \citet{oberg11}.
-
-%Based on some observations of protoplanetary disks \citep{andrews10}, our surface density profile, $\Sigma \propto r^{-1}$, is flatter than that of \citet{chiang10}, i.e. $\Sigma \propto r^{-3/2}$.
-
-\textbf{Evolving disk.} We model the evolving disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
-\begin{equation}
-\label{eq:nu}
-\nu=\alpha c H.
-\end{equation}
-Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c$, $H$ are the isothermal sound speed and disk scale height, respectively:
-\begin{subeqnarray}
-\label{eq:cdHd}
-c &=& \sqrt{\frac{k_{\rm B} T}{\mu m_{\rm p}}} \slabel{eq:cd} \\
-H&=& \frac{c}{\Omega_{\rm k}} \slabel{eq:Hd},
-\end{subeqnarray}
-where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$ is the Keplerian angular velocity,  with $G$ the gravitational constant and $M_*$ the stellar mass.
-We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition of hydrogen and helium.
-The temperature profile for the %an
-evolving disk is assumed to be the same as for the irradiated disk and given by Equation (\ref{eq:diskT}).
-From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters.
-Following \citet{hartmann98}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius.
-We choose $r_{\rm c}=100$ AU.
-The gas surface density is given by the self-similar solution
-
-\begin{equation}
-\label{eq:Sigmaact}
-\Sigma(R, \tilde t) = \frac{M (2 - \gamma)}{2 \pi r_{\rm c}^2 R^{\gamma}} \tilde t^{-(5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{(2-\gamma)}}{\tilde t}}\Big],
-\end{equation}
-where $M$ is the total disk mass and
-\begin{subeqnarray}
-\label{eq:T}
-\tilde t & \equiv & \frac{t}{t_{\rm c}} + 1 \\
-t_{\rm c} & \equiv & \frac{1}{3(2-\gamma)^2} \frac{r_{\rm c}^2}{\nu_{\rm c}},
-\end{subeqnarray}
-where $t$ is time.
-We choose $M=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).
-The irradiated and evolving disk surface densities match at \textbf{$t \approx 5 \times 10^5$} years in the inner disk, but they diverge at distances larger than a few AU due the exponential cutoff in radius of the surface density of the evolving disk (Equation \ref{eq:Sigmaact}).
-
-\textbf{Viscous disk.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis.
-Since
-Calculating the midplane temperature self-consistently for an evolving disk that is also actively heated, and thus whose thermal evolution is dominated both by accretion heating and stellar irradiation, is non-trivial.
-We therefore use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm act}$.
-The equations governing the evolution of the steady-state disk are listed in \App{app:steadystate}.
-We assume an interstellar opacity for the dust grains given by \citet{bell94}, but reduced by a factor of 100.
-This reduction is due to the fact that disk opacities are lower than the interstellar one.
-While this scaling is consistent with more detailed models of grain opacities in disks (e.g., \citealt{mordasini14}), realistic disk opacities are much less sensitive to changes in temperature than the interstellar opacity if substantial grain growth has occurred.
-However, the disk temperature does not vary significantly across the small region of the disk where accretion heating is important ($r \lesssim 1$ AU).
-Moreover, using an analytic opacity formula is more convenient since it results in a constant gas surface density in the inner disk region (see below).
-Our opacity law is thus
-
-\begin{equation}
-\label{eq:opacity}
-\kappa=\kappa_0 T_{\rm act}^2,
-\end{equation}
-where $\kappa_0=2 \times 10^{-6}$. By solving the Equation set (\ref{eq:diskeq}) we find
-\begin{equation}
-\label{eq:Tdact}
-T_{\rm act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
-\end{equation}
-Since both accretion heating and stellar irradiation contribute to the thermal evolution of the disk, we compute the midplane temperature for our viscous disk as
-\begin{equation}
-\label{eq:activeT}
-T^4 = T_{\rm act}^4 + T_{\rm irr}^4,
-\end{equation}
-where to avoid notation confusion $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}), the temperature profile for an irradiated disk.
-We can then easily determine  $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
-For consistency, we choose  $\alpha=0.01$ as in the previous case.
-Finally, we determine $\Sigma$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}).
-In the inner portion of our disk ($r  \lesssim 1$ AU for our fiducial model with $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$), our choice of opacity (Equation \ref{eq:opacity}) implies that the disk has a constant surface density with radius (see Equations \ref{eq:diskeq}).
-
-Before we proceed forward, we note that our disk models assume a constant stellar luminosity $L_*$, as well as a constant mass accretion rate $\dot{M}$ for the viscous disk.
-In reality, the stellar luminosity decreases as the host star contracts, which will reduce the disk temperature and push the snowlines inward, as we explain in Section \ref{sec:neglected}.
-For a Solar type star, as our fiducial model assumes, $L_*$ remains relatively constant during the star's pre-main sequence evolution of $\sim$ 10 Myr \citep{kennedy06}, which is larger than the giant planet formation timescale.
-Thus the midplane temperature will not change significantly for our model due to variations in stellar luminosity, but it may decrease substantially for smaller stars, pushing the snowline inward (see Section \ref{sec:neglected} for details).
-Realistic mass accretion rates, $\dot{M}$, may vary between $\sim$$10^{-7}$ and $\sim$$10^{-9}$ $M_{\odot}$ yr$^{-1}$ as the disk evolves \textbf{(e.g., \citealt{chambers09}, \citealt{sicilia10})}.
-For $\dot{M}\lesssim 10^{-9} M_{\odot}$ yr$^{-1}$, the disk becomes optically thin and hence depleted of gas, which means giant planets must have formed before $\dot{M}$ becomes too low.
-\citet{garaud07} find that the snowline locations scale as $r_{\rm snow} \propto \dot{M}^{1/3}$.
-A factor of 100 reduction in the mass accretion rate will thus move the H$_2$O snowline inwards by a factor of $\sim$$4$ \textbf{--- since accretion heating is dominant only in the inner disk, the CO$_2$ and CO snowline locations are unlikely to be affected by changes in mass accretion rate.
-The inward movement of the H$_2$O snowline due to the decrease in $\dot{M}$ may be even larger, by up to one order of magnitude \citep{chambers09}}.
-We thus conclude that changes in $L_*$ throughout time may only modestly affect our results, \textbf{while changes in $\dot{M}$ may significantly affect our results for the H$_2$O snowline, as its location may be determined by the decline in mass accretion rate rather than radial drift.} The time variability of $L_*$ and $\dot{M}$ should be taken into account when drawing more robust conclusions, as well as for different host star and disk properties.
-
-%However, \citet{garaud07} find that the mass accretion rate is close to $10^{-8} M_{\odot}$ yr$^{-1}$ during the quasi-static accretion phase of the disk (see also Section \ref{sec:neglected}), hence the snowline locations are likely to move inward only in the very early stages of the disk evolution.
-We thus claim that changes in $L_*$ and $\dot{M}$ throughout time may only modestly affect our results, but they should still be taken into account when drawing more robust conclusions, as well as for different host star and disk properties.}
-
- %We note that $\Sigma$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.
-
-%Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady-state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):
-
-%\begin{subeqnarray}
-%\label{eq:activeT}
-%T^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
-%\dot{M} & = & 3 \pi \nu \Sigma \slabel{eq:Mdot},
-%\end{subeqnarray}
-%where $\dot{M}$, the mass flux, is constant in steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}.
-Based on disk observations (e.g., \citealt{andrews10}), we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$.
-We can then easily determine the radial temperature profile from Equation (\ref{eq:Tdact}), $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$.
-For consistency, we choose  $\alpha=0.01$ as in the previous case.
-Finally, we determine $\Sigma$ from Equation (\ref{eq:Mdot}).
-
- %The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).
-
-\subsection{Radial Drift}
-\label{sec:drift}
-
-Solid particles orbit their host star at the Keplerian velocity $v_{\rm k} \equiv \Omega_{\rm k} r$.
-The gas, however, experiences an additional pressure gradient, which causes it to rotate at sub-Keplerian velocity \citep{weidenschilling77}.
-Dust grains that are large enough thus experience a headwind, which removes angular momentum, causing the solids to spiral inwards and fall onto the host star.
-Small particles are well-coupled to the gas, while large planetesimals are decoupled from the gas.
-From the review by \citet{chiang10}, the extent of coupling is quantified by the dimensionless stopping time, $\tau_{\rm s} \equiv \Omega_{\rm k} t_{\rm s}$, where $t_{\rm s}$ is
-\begin{equation}
-\label{eq:ts}
-t_{\rm s}= \left\{
-\begin{array}{l l}
-\rho_{\rm s} s / (\rho c), & \quad s < 9 \lambda/4 \,\,\,\ \text{Epstein drag} \\
-4 \rho_{\rm s} s^2 / (9 \rho c \lambda), & \quad s < 9 \lambda/4, \,\text{Re} \lesssim 1 \,\,\,\ \text{Stokes drag.}
-\end{array} 
-\right.
-\end{equation}
-Here $\rho$ is the gas midplane density, $\rho_{\rm s}=2$ g cm$^{-3}$ is the density of a solid particle, $s$ is the particle size, $\lambda$ is the mean free path, and Re is the Reynolds number.
-
-For an irradiated disk, the radial drift velocity can be approximated as
-\begin{equation}
-\label{eq:rdotpas}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big),
-\end{equation}
-where
-\begin{equation}
-\label{eq:eta}
-\eta \equiv - \frac{\partial P/\partial \ln r}{2 \rho v_{\rm k}^2} \approx \frac{c^2}{2 v_k^2}
-\end{equation}
-and $P = \rho c^2$ is the disk midplane pressure.
-
-For an evolving disk, the radial drift velocity has an additional term due to the radial movement of the gas \citep{birnstiel12}, i.e.
-\begin{equation}
-\label{eq:rdotact}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{\dot{r}_{\rm gas}}{1+\tau_{\rm s}^2},
-\end{equation}
-where $\dot{r}_{\rm gas}$ is the radial gas accretion velocity and can be expressed as (e.g., \citealt{fkr02})
-\begin{equation}
-\label{eq:vgas}
-\dot{r}_{\rm gas} = - \frac{3}{\Sigma \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma \sqrt{r}) 
-\end{equation}
-with $\Sigma$ from Equation (\ref{eq:Sigmaact}).
-For the viscous disk (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}).
-For our choice of parameters for both the evolving and the viscous disks, we have found that the radial flow of gas \textbf{(calculated from Equation \ref{eq:vgas} for the evolving disk and from $\dot{r}_{\rm gas}=-\dot{M}/(2 \pi r \Sigma)$ for the viscous disk)} is always directed inward \textbf{for our parameter space of interest}, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}).
-\textbf{For our evolving disk, the gas starts drifting outwards at a radius $r_{\rm switch} \approx 200 AU$, which is however well outside the CO snowline in our model.} We \textbf{thus} note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations.
-Since drifting particles larger than a few cm are only modestly affected by gas accretion, an outward gas flow would move the CO$_2$ and CO locations further away from the star only for the smallest particles in our model, which are well-coupled to the gas.
-
-\subsection{Volatile Desorption}
-\label{sec:desorption}
-
-%\subsection{Disk and Desorption Model}
-
-In order for a volatile species to thermally desorb, it has to overcome the binding energy that keeps it on the grain surface.
-Following \citet{hollenbach09}, the desorption rate per molecule for a species $x$ can be expressed as
-\begin{equation}
-\label{eq:Rdes}
-R_{\rm{des}, x} = \nu_x \exp{(-E_x/T_{\rm grain})},
-\end{equation}
-where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature, and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well, with $\mu_x$ the dimensionless mean molecular weight.
-We assume that the dust and gas have the same temperature in the disk midplane, hence $T_{\rm grain}=T$.
-For H$_2$O, CO$_2$ and CO, the binding energies $E_x$ are assumed to be 5800 K, 2000 K and 850 K, respectively (\citealt{collings04}, \citealt{fraser01}, \citealt{aikawa96}).
-We use the desorption rate, $R_{\rm des}$, to estimate the desorption timescale for particles of different sizes as described in section \ref{sec:timescales}.
-
-%\emgr{Describe disk model --- both passive MMSN and active self-similar solution with $\alpha$ prescription for viscosity.
-Describe the model for evolving the surface density profile of planetesimals following Birnstiel et al.
-(2012).
-Describe temperature profile --- currently power-law, but may change.
-Describe desorption model and parameters following Hollenbach et al.
-(2009).
-Mention that the solids are perfect spheres composed of a single volatile.
-Perhaps this subsection needs to be split into subsubsections.}
-
-\subsection{Relevant Timescales}
-\label{sec:timescales}
-
-We can estimate the extent to which radial drift and gas accretion affect desorption by comparing the timescales for desorption, drift and accretion, for solids of different sizes and compositions.
-
-\textit{Desorption timescale.} We assume that the solid bodies are perfect spheres and are entirely composed of only one volatile species, i.e.
-either H$_2$O, CO$_2$ or CO \footnote{We discuss the validity of these simplifications in section \ref{sec:discussion}.}.
-The timescale to desorb a single layer of molecules can then be estimated as
-\begin{equation}
-\label{eq:tdes}
-t_{\rm des}=\frac{\rho_{\rm s}}{3 \mu_x m_{\rm p}} \frac{s}{N_x R_{\rm des, x}},
-\end{equation}
-where $N_x \approx 10^{15}$ sites cm$^{-2}$ is the number of adsorption sites of volatile $x$ per cm$^2$, assuming that the particle has a smooth surface \citep{hollenbach09}.
-
-\textit{Radial drift timescale.} To order of magnitude, the radial drift timescale can be estimated as 
-\begin{equation}
-\label{eq:tdrift}
-t_{\rm drift} \sim \Big|\frac{r}{\dot{r}}\Big|,
-\end{equation}
-where $\dot{r}$ is the radial drift velocity given by Equation (\ref{eq:rdotpas}) for an irradiated disk and by Equation (\ref{eq:rdotact}) for an evolving disk.
-
-\textit{Gas accretion timescale.} The timescale for gas accretion onto the central star for an evolving disk is (e.g., \citealt{armitage10})
-\begin{equation}
-\label{eq:tgas}
-t_{\rm gas, acc} \sim \frac{r^2}{\nu} \sim \frac{1}{2 \alpha \eta \Omega_{\rm k}},
-\end{equation}
-with the latter expression derived from Equations (\ref{eq:nu}) and (\ref{eq:eta}).
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{drift_timescales_betaS1_gas_acc_new2.pdf}
-%\vspace{-0.5in}
-\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines).
-The timescales are calculated at three representative locations, i.e.
-the H$_2$O, CO$_2$ and CO snowlines in the static disk.
-For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively.
-The horizontal dot-dashed line represents a typical disk lifetime of 3 Myr.
-The particle size ordering at the minimum $t_{\rm drift}$ is not monotonic in snowline distance due to different drag regimes for those particle sizes at the snowline locations (Epstein drag at the H$_2$O and CO$_2$ snowlines, and Stokes drag at the CO snowline).
-Similarly, the ordering of $t_{\rm des}$ is not monotonic in snowline distance due to the non-monotony in mean molecular weight between H$_2$O, CO$_2$ and CO (18 $m_{\rm p}$, 44 $m_{\rm p}$ and 28 $m_{\rm p}$, respectively).
-Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e.
-$t_{\rm drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.}
-\label{fig:timescales}
-\end{figure}
-
-For simplicity, we calculate the radial drift timescale, $t_{\rm drift}$, for an irradiated disk in this section, but most of our conclusions hold true for an evolving disk as well.
-Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk.
-As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence on disk location (see Equation \ref{eq:Rdes}).
-On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas.
-Thus for small particles in an irradiated disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}).
-This is not true for an evolving disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.
-At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well.
-This is true for both irradiated and evolving disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.
-
-In the particle size regime for which (1) $t_{\rm drift} \lesssim t_{\rm des} \lesssim t_{\rm d}$ ($t_{\rm d}=3$ Myr is the disk lifetime), i.e.
-for $\sim$$0.5$ cm $\lesssim s \lesssim$ $1000$ cm, or (2) $t_{\rm gas, acc} \lesssim t_{\rm des} \lesssim t_{\rm d}$, i.e.
-for $\sim$$0.1$ cm $\lesssim s \lesssim$ $10$ cm, radial drift or gas accretion (or both) are faster than thermal desorption, which is of particular interest for our purposes.
-We note that $t_{\rm gas, acc}<t_{\rm d}$ always holds true.
-Particles of sizes that satisfy the requirements above will drift significantly due to radial drift or gas accretion before desorbing, thus moving the H$_2$O, CO$_2$ and CO snowlines closer towards the central star and changing the C/O ratio throughout the disk.
-We quantify these effects in sections \ref{sec:snowlines} and \ref{sec:COratio}.
-
-%\emgr{Calculate timescale for radial drift following Chiang \& Youdin (2010).
-Calculate desorption timescale following Hollenbach et al.
-(2009).
-Estimate gas accretion timescale for a given $\alpha$.
-Important: mention that, for simplicity and illustrative purposes, these calculations are performed for a passive disk.
-Show plot with the timescales as a function of particle size at different snowlines to show the regime in which drift matters}.
-
-\section{Snowline Locations}
-\label{sec:snowlines}
-
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=0.7\textwidth]{desorption_distance_passive_active_colorbar_test2.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both an irradiated disk (left panels) and an evolving disk (right panels).
-The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The desorption distance for a static disk is shown for comparison (dashed vertical and horizontal lines).
-The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
-For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.}
-\label{fig:snowlines}
-\end{figure*}
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=0.7\textwidth]{desorption_distance_actual_vs_estimated_passive_active_new.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text).
-The estimate is performed for an irradiated disk (left panels) and an evolving disk (right panels).
-The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
-\label{fig:an_vs_actual}
-\end{figure*}
-
-In this section we use the model described in section \ref{sec:model} to quantify the effects of radial drift (irradiated disk) or radial drift and gas accretion (evolving disk) on the snowline location, for dust particles of different sizes composed of either H$_2$O, CO$_2$ or CO.
-Specifically, we determine a particle's final location (i.e., where the particle either fully desorbs or remains at its initial size due to having a desorption timescale longer than the time at which we stop the simulation) as a function of its initial position in the disk, after the gas disk has dissipated.
-The disk lifetime, $t_{\rm d}$, is particularly relevant since the timescale for giant planet formation must be less than or equal to $t_{\rm d}$.
-The snowline locations at $t=t_{\rm d}$ throughout the protoplanetary disk determine the disk C/O ratio in gas at this time, and thus the C/O ratio in giant planet atmospheres that have formed \textit{in situ}, before planetesimal accretion or core dredging.
-
-For each species $x$, we determine the final location in the disk of a particle of initial size $s_0$ by solving the following system of coupled differential equations:
-
-\begin{subeqnarray}
-\label{eq:ddt}
-\frac{ds}{dt} &= & - \frac{3 \mu_x m_{\rm p}}{\rho_{\rm s}} N_x R_{\rm des, x}  \slabel{eq:dsdt} \\
-\frac{dr}{dt} &=& \dot{r} \slabel{eq:drdt},
-\end{subeqnarray}
-where the desorption rate $R_{\rm des, x}$ for each particle type (i.e., composed of H$_2$O, CO$_2$ or CO) is evaluated at $T=T(r)$, and the radial drift velocity $\dot{r}$ is given by Equation (\ref{eq:rdotpas}) for an irradiated disk and Equation (\ref{eq:rdotact}) for an evolving disk.
-Equations (\ref{eq:dsdt}) and (\ref{eq:drdt}) describe the coupled desorption and radial drift, and can be derived straightforwardly from Equation (\ref{eq:tdes}).
-Our initial conditions are $s(t_0)=s_0$ and $r(t_0)=r_0$, where $t_0$ is the initial time at which we start the integration and $r_0$ is the initial location of the particle.
-We choose $t_0=1$ year, but our result is independent on the initial integration time as long as $t_0 \ll t_{\rm d}$.
-The desorption timescale $t_{\rm des}$ will then satisfy $s(t_{\rm des})=0$, from which we can determine the desorption distance $r_{\rm des}=r(t_{\rm des})$.
-
-%As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully evaporates.
-
-%Thus a particle's final location will depend on whether a grain initially at a certain distance is completely desorbed or not within the disk lifetime of 3 Myr.
-For example, larger grains take longer to desorb and hence are more likely to not evaporate fully in a given timeframe.
-
-We define the final position of a grain as the disk location it has reached after $t_{\rm d}=3$ Myr, or the radius at which it completely desorbs if that happens after a time shorter than 3 Myr.
-Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both an irradiated and an evolving disk.
-We do not show the results for the viscous disk as they would complicate the plot without adding any qualitative insight ---  the results for the viscous disk are quantitatively similar with those of the evolving disk for the CO$_2$ and CO particles, but they are different for the H$_2$O grains, since accretion heating will push the H$_2$O snowline outwards (see Section \ref{sec:COratio}).
-We also show the static snowlines for comparison, which are calculated by balancing adsorption and desorption \citep{hollenbach09}.
-%We do not show the equivalent result for the steady-state active disk since The trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}), but we discuss the C/O ratio steady-state active disk results in Section \ref{sec:COratio}.
-Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for an irradiated nor for an evolving disk.
-Similarly, micron- to mm-sized particles in the irradiated disk do not drift and only %or
-desorb if %unless 
-they are located inside the static snowlines.
-%This is not in contradiction with the desorption timescales for small particles from Figure \ref{fig:timescales}, since small grains only desorb fast at or nearby their respective snowlines, as mentioned in Section \ref{sec:timescales}.
-[don't think this is needed, but Til or Ruth might think otherwise]
-In an evolving disk, however, micron-to mm-sized grains do drift significantly since they move at the same velocity as the accreting gas.
-For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in an irradiated disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an evolving disk, we notice that particles of initial size $s_0$ desorb at a %fixed distance
-particle size dependent radius $r_{\rm des}$ regardless of their original location in the disk.
-In fact, the only grains that will both drift and evaporate are those that reach their fixed final location (represented by the horizontal curves in Figure \ref{fig:snowlines}) within the disk lifetime.
-We show in section \ref{sec:COratio} that this result is essential in determining the C/O ratio throughout the disk for different particle sizes.
-
-Another interesting feature of Figure \ref{fig:snowlines} is that particles above a certain size ($\sim$7 m for our choice of parameters) all desorb at the same distance.
-This is due to the fact that once the large bodies pass the static snowline, they first lose mass, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=0.8\textwidth]{s_t_a.pdf}
-%\vspace{-0.5in}
-\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in an irradiated disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left).
-Particles desorb almost instantaneously.
-Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel.
-Particles lose most of their mass very close to the distance at which they fully desorb.
-The static H$_2$O snowline is shown for reference (dashed vertical lines).}
-\label{fig:s_t_a}
-\end{figure*}
-
-Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm drift} \sim t_{\rm des}$, given an initial particle size.
-We can calculate this location analytically by equating Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and solving for $r=r_{\rm des}(s)$ for a given particle size $s$.
-Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically.
-We display this result for the range of particle sizes that desorb at a fixed distance in an irradiated and an evolving disk (see Figure \ref{fig:snowlines}).
-We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it deviates for particles larger than $s \gtrsim10$ cm.
-For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the irradiated disk in the Epstein drag regime), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
-Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
-
-%This is due to the fact that large bodies lose their mass as they desorb until they reach the same size as the smaller, meter-sized particles, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
-
-Given $r_{\rm des}$, we need to only calculate the distance over which particles desorb to determine the location of a snowline.
-Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in an irradiated disk.
-Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
-%, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-%A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
-[not sure you need to explain this]
-The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
-Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model.
-Both of these conclusions remain valid for an evolving and a viscous disk, as well as for particles composed of CO$_2$ or CO, but the snowline locations will vary between the three disks for a given initial particle size (see Section \ref{sec:COratio}).
-If we do not take into account the time dependence of the mass accretion rate and stellar luminosity (see Section \ref{sec:disk}), the C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
-%\emgr{I've yet to figure out why there is a discrepancy for the small grains.
-I'll keep thinking about it, but for now I'll leave it as it is since I don't want to give a false explanation just to have one.}
-
-%\emgr{I didn't write the text for Figure 4 since we haven't yet decided if in the end we will include it in some form. We can }
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_vs_s.pdf}
-%%\vspace{-0.5in}
-%\caption{Desorption distance as a function of initial particle size, for the range of particles that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines}).
-\textbf{Figure belongs to section 3.} \emgr{This is a figure that I'm not sure we should include, since it doesn't really provide that much useful inside.
-If we did include it though, there would be two panels both for the passive and the active disk.
-The caption is not complete since the active disk panel is not yet there.}}  %  (See text for a description of evolution to yet higher masses.)
-%\label{fig:r_vs_s}
-%\end{figure}
-
-%\emgr{Present the equation set that you are solving, $dr/dt=\dot{r}$, $ds/dt=...$.
-I don't think it is necessary to describe in detail the numerical method of solving the equations since it's pretty straightforward.
-Present the pretty rainbow plots side by side for passive and active disks, and highlight the differences.
-Insert the plots that shows that for the intermediate size particles, the desorption distance can be estimated analytically with good accuracy.
-Maybe also include the plot showing the desorption distance as a function of particle size, both for passive and active disk.
-Split into subsections?}
-
-\section{C/O Ratio Estimates}
-\label{sec:COratio}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
-%\vspace{-0.5in}
-\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for an irradiated disk (top panel), an evolving disk (middle panel) and a viscous disk (bottom panel).
-%, for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk. 
-The particle size increases from $\sim$0.05 cm to $\sim$700 cm as indicated by the color bar.
-The horizontal dotted line represents the stellar value of 0.54.
-The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk, with the temperature profile given by Equation (\ref{eq:diskT}) for the top two panels and by Equation (\ref{eq:activeT}) for the bottom panel.
-%The snowline location moves inward as the particle size increases. 
-%The curves exterior to the static snowline represent desorption distances in the absence of readsorption.
-%Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline.
-For both the evolving and the viscous disk, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, as shown by the arrows.}
-%, with the decrease more significant interior to the H$_2$O snowline.}
-\label{fig:CO_ratio}
-\end{figure}
-
-Given our results in Section \ref{sec:snowlines}, a disk's C/O ratio in mainly affected by the snowline location for the particle size housing the most mass in ice.
-%We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
-Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then estimate the C/O ratio throughout the disk as a function of the particle size.
-Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
-In Figure \ref{fig:CO_ratio}, we display the H$_2$O, CO$_2$, and CO snowline locations as a function of particle size for disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
-The minimum snowline distance for a disk is given by the curve corresponding to the maximum particle size it hosts.
-For grains that have grown to radii larger than $\sim$7 m and are able to drift and desorb, the $\sim$7 m snowline applies (see Section 3).
-%Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.
-
-%[I think a sentence is missing here outlining what you need to establish to calculate the C/O ratio, i.e. the connection between the preceding and the next sentence is not obvious.]
-
-%Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk.
-Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
-%, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-%A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
-[not sure you need to explain this]
-%The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
-Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-%It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active)\footnote{Both of these conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
-The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.
-
-%In Section \ref{sec:snowlines}, we showed that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk.
-Figure \ref{fig:s_t_a} confirms both of these effects.
-The left panels show the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.
-Indeed, the solid H$_2$O particles evaporate almost instantly, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime.
-However, our model assumes that particles drift continuously at any location in the disk.
-Therefore, a particle that can fully desorb during the disk lifetime for at least one initial location will always desorb, at a fixed distance as discussed in Section \ref{sec:snowlines} and displayed in Figure \ref{fig:snowlines}.
-The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
-Figure \ref{fig:s_t_a} thus proves claims (1) and (2) above.
-
-Drift and gas accretion affect the C/O ratio in a disk both because they move the snowline locations of the main C and O carriers  and because they cause solids and gas---which contain different proportions of C and O---to move inward at different rates.
-As shown in Section 3, the snowline locations depend on disk age only indirectly, through changes in disk properties and grain size.
-The C/O ratio is a function of the locations of the snowlines and the abundances of H$_2$O, CO$_2$, and CO relative to the H$_2$ abundance in the disk midplane.
-These abundances evolve over time as solids and gas move inward at different rates.
-
-Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for an irradiated disk, an evolving disk, and a viscous disk, under the simplifying assumption that the abundance relative to hydrogen for each volatile is fixed, so that drift and accretion affect only the locations of the snowlines.
-We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
-Snowline locations correspond to $r_{\rm des}$ in Figure \ref{fig:snowlines}, representing the location at which particles desorb in the absence of readsorption.
-The C/O ratio for a static disk, where desorption and readsorption balance \citep{hollenbach09}, is shown as a guideline.
-We note that the true snowline for particles with $r_{\rm des}$ outside the static snowline is the static snowline itself---thus only particles with initial sizes larger than $\sim$0.05 cm are plotted in the three panels, as particles that form snowlines at larger distances (cf.
-Figure \ref{fig:snowlines}) are not true snowlines.
-
-Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the evolving and viscous disks ignore the movement of the desorbed ices with the accreting gas---the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
-As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
-For example, for the disk parameters and particle sizes displayed in Figure \ref{fig:CO_ratio}, water molecules in solid particles drift up to $\sim$1000 times faster across the H$_2$O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.
-This differential inward motion will result in an increased oxygen gas abundance inside the H$_2$O snowline, and thus a (in some cases much) lower gaseous C/O ratio in this region.
-Conversely, oxygen gas inside the water snowline will be depleted compared to the static disk if H$_2$O particles grow to planetesimal size and stall their migration between the H$_2$O and CO$_2$ snowlines, leaving only gaseous CO and CO$_2$ to accrete inward.
-Growth of large planetesimals can therefore increase the C/O ratio in the inner disk.
-
-\textbf{Our premise that each volatile has a fixed abundance at every location in the disk yields good approximations for the C/O ratio in an irradiated disk --- our model assumes a constant influx of particles at any given radius, implying that the solid surface density of a given volatile, and therefore the amount of desorbed vapor, remain constant with semimajor axis.
-For the evolving and viscous disks, the C/O ratio in gas may decrease due to the decline in the surface density of solids with time.
-Thus both the relative fluxes of the volatiles and the decrease in solid ice abundances will lower the C/O ratio in evolving and viscous disks, as indicated by the arrows in Figure \ref{fig:CO_ratio}.
-However, a high gas-phase C/O ratio is maintained further in for all disks compared to a static disk, since drift and gas accretion cause the ices to desorb closer to the host star.}
-
-Figure \ref{fig:CO_ratio} plots snowline curves for particle sizes $\sim0.5$ cm $\lesssim $s$ \lesssim$ 7m.
-In the outermost disk, H$_2$O, CO$_2$, and CO all solidify.
-Hence, relative drift across the CO snowline can alter only the abundances of volatiles between the CO$_2$ and CO snowlines, but not the C/O ratio in this region.
-Interior to the CO$_2$ snowline, however, relative drift is important.
-We have found that the largest drifting particles in our model ($\sim$7 m) drift faster than the gas at both the H$_2$O and CO$_2$ snowlines.
-We thus conclude that the C/O ratio interior to the H$_2$O and CO$_2$ snowlines in our evolving and viscous disks will be lower than in the static disk, due to the additional oxygen added to the gas by desorbing H$_2$O and CO$_2$.
-For these particle sizes, our calculated C/O ratio is an upper limit,  as indicated by the arrows in Figure \ref{fig:CO_ratio}.
-
-%We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
-Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk, and an active disk accretionally heated.
-The C/O ratio for a static disk is shown as a guideline.
-We note that particles desorbing outside the static snowline will not form snowlines, since our calculation of the desorption distance does not account for readsorption  \citep{hollenbach09}.
-
-%Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the active disks ignore the movement of the desorbed ices with the accreting gas --- the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk.
-As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
-Specifically, the gas C/O ratios may be significantly lower inside the H$_2$O and CO$_2$ snowlines for the active disks, depending on the gas accretion rate, due to the additional oxygen contained by H$_2$O and CO$_2$ vapor that moves together with the accreting gas and crosses their respective snowlines.
-
-%e thus present this result for the C/O ratio primarily as a visual comparison with previous work \citep{oberg11} and note that our estimates for the C/O ratio in the active disks are not quantitatively accurate at all disk locations.
-Future work will address this issue.
-
- %The plot is consistent with Figure \ref{fig:snowlines}. 
-The snowline locations in these disks exhibit several interesting features.
-For the irradiated disk, only grains larger than $\sim$0.5 cm drift, desorb and thus move the snowline compared to the static disk.
-In contrast, even $\sim$micron-sized grains drift and desorb for the evolving disk, since they flow towards the host star together with the accreting gas.
-For the same particle size, the snowline locations are slightly closer to the central star in the evolving disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf.
-Equation \ref{eq:rdotact}).
-The addition of accretional heating in the viscous disk moves the H$_2$O snowline outwards.
-This is due to the fact that accretional heating dominates in the inner disk, where high temperatures cause the grains to evaporate further away from the star.
-Once $r\gtrsim1-2$  AU, stellar irradiation dominates the thermal evolution of the disk, and therefore the CO$_2$ and CO snowlines locations are the same as in the evolving and viscous disks.
-
-Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases.
-While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that %almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk.
-B
-bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals (see Section \ref{sec:snowlines}).
-However, the contribution of kilometer-sized bodies to the snowline location is modest, since they only drift if they are located very close to the snowline.
-%[new sentence explaining that their contribution may not be very big due to their modest drift.] though 
-%only the kilometer-sized boulders that are very close to the snownlines will contribute to determining the snowline location. 
-Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
-For a grain size distribution with a maximum particle size different than our model, one can pick out the appropriate minimum snowline locations from this plot.%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
-Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from this plot.
-%Our model thus sets an inner limit to the snowline locations.
-%Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution --- for a given grain size .
-
-% [the last sentence needs to be made more explicit, i.e.
-for a given grain size distribution with a maximum grain size a reader can now pick out the appropriate minimum snowline location from this plot]
-
-For our choice of parameters, the minimum snowline radii are: $r_{\rm{H_2O}} \approx 0.3$ AU for the irradiated disk, $r_{\rm{H_2O}} \approx 0.26$ AU for the evolving disk and $r_{\rm{H_2O}} \approx 0.63$ AU for the viscous disk; $r_{\rm{CO_2}} \approx 3.7$ AU for the irradiated disk, $r_{\rm{CO_2}} \approx 3.4$ AU for both the evolving and the viscous disks; $r_{\rm{CO}} \approx 30$ AU for the irradiated and both the evolving and the viscous disks.
-For comparison, $r_{\rm{H_2O}} \approx 0.67$ AU\footnote{For the viscous disk, we calculated the static snowline location using the same temperature profile as that of the viscous disk, for consistency purposes.
-Thus $r_{\rm{H_2O}} \approx 0.98$ AU for the static disk in this scenario.}, $r_{\rm{CO_2}} \approx 8.6$ AU and $r_{\rm{CO}} \approx 59$ AU for the static disk.
-For the viscous disk model, which is the most realistic, radial drift and gas accretion push the snowline locations inwards by up to $\sim$$40$ \% for H$_2$O, by up to  $\sim$$60$ \% for CO$_2$, and by up to $\sim$$50$ \% for CO.
-We note that the H$_2$O snowline in all disks is significantly closer to the host star compared with Solar system models, which place the H$_2$O snowline between $\sim$$2.7$ to $\sim$$3.1$ AU (\citealt{hayashi81}, \citealt{podolak04}, \citealt{martin12}).
-This is partially because we choose a colder disk model, as well as the fact that gas accretion rates decrease over time, moving the snowline location inwards (see also \citealt{garaud07} and Section \ref{sec:neglected}).
-\citet{min11} find that the location of the H$_2$O snowline is highly sensitive to the gas mass accretion rate $\dot{M}$ (equal to $10^{-8} M_{\odot}$ yr$^{-1}$ in our model) and the dust opacity $\kappa$ (Equation \ref{eq:opacity}).
-Higher values of $\dot{M}$ and $\kappa$ would increase the accretional component of the disk temperature (cf.
-Equation \ref{eq:Tdact}), which would push the H$_2$O snowline in the viscous disk outwards to match the Solar system snowline.
-At the same time, the snowline location in Solar type stars may be as close as $\sim$1 AU \citep{mulders15}, further in than the H$_2$O snowline in our Solar system.
-
- Observations of the CO snowline in TW Hya \citep{qi13} have found its location at a disk midplane temperature of 17 K (at 30 AU for the TW Hya specific temperature profile).
-The inferred  desorption temperature corresponds to the CO desorption temperature in a static disk, or to desorption from very small grains in an evolving disk, i.e.
-from grains that are too small to drift substantially.
-This suggests that the outer TW Hya disk is dominated by small grains, since larger particles would push the snowline location inwards, and therefore to higher desorption temperatures.
-This may seem contradictory to observations of grain growth in disks in general and in TW Hya in particular \citep{wilner00}.
-However, recent observations have revealed that grain growth is concentrated to the inner disk \citep{perez12} and outer disk snowlines may therefore be close to the ones expected in a static disk.
-
-%\emgr{There will be 1-2 additional sentences for the third panel, the steady-state disk, once that finishes running.}
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.5\textwidth]{lala}
-%%\vspace{-0.5in}
-%\caption{C/O ratios for different particle size distributions...
-\textbf{Figure belongs to section 4.} (\emgr{It will probably be a multi-panel plot, for different particle size distributions, for both passive and active disk, and perhaps at different times for the active disk.
-Not yet unsure how that figure will look like since I don't have the results yet, so I'll leave the figure caption open for now.})}
-%\label{fig:...}
-%\end{figure}
-
-%\emgr{Motivate the fact that we can calculate a sharp, fixed snowline for each particle size by inserting the plot that shows that particles desorb almost instantaneously at a fixed distance.
-Then present the plots analogous to Fig.
-1 in \"Oberg et al.
-(2011) for different particle sizes, based on the snowline locations obtained in the previous section, both for passive and active disk.
-Perhaps show it at different times in the gas disk evolution (i.e., not just at 3 Myr) for the active disk.
-Then assume a particle size distribution and show the interpolated result for the C/O ratio.
-Generalize the result using a transition disk (this part might also fit in the discussion section).}
-
-%emgr{Present the flux equations used to keep track of the amount of C and O in gas and dust throughout the disk.
-Motivate the simplification of using a fixed particle sized, fixed timescale, and fixed desorption distance in the calculations by referring to the rainbow plot in the previous section and by inserting the plot that shows that particle desorb almost instantly at a fixed distance.
-Based on these assumptions, finally show the equivalent of Fig.
-1 in \"Oberg et al.
-(2011) for different particle sizes, and at different times in the gas disk evolution (i.e., not just at 3 Myr) --- a multi-panel plot could be a good idea.
-Discuss the result, trends, etc.
-Ideally, have a final plot showing the results for a particle size distribution rather than for individual particles.}
-
-\section{Discussion}
-\label{sec:discussion}
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=\textwidth]{time_plots.pdf}
-%\vspace{-0.5in}
-\caption{Desorption or final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve).
-For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
-\label{fig:timeplots}
-\end{figure*}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{desorption_distance_varying_Md.pdf}
-%\vspace{-0.5in}
-\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines).
-The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve).
-A lower disk mass does not change the snowline location.}
-\label{fig:varMd}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{desorption_distance_transition_disk_1000.pdf}
-%\vspace{-0.5in}
-\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines).
-The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}.
-Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge, due to being trapped in a pressure maximum.}
-\label{fig:cavity}
-\end{figure}
-
-\subsection{Generality of Results: Dependence on Disk Parameters}
-\label{sec:incond}
-
-In this section we investigate how variations in our fiducial parameters, the total disk mass, disk age, and disk structure, %the time in the disk evolution at which we determine 
-affect the calculated snowline locations and the C/O ratio. %, affect our results. 
-All previous results assumed a disk lifetime $t_{\rm d}=3$ Myr%is the most representative for our calculations, as this is the timeframe in which giant planets accrete their gaseous atmospheres 
-, the typical disk lifetime and the expected time scale for giant planets to accrete their gaseous atmopsheres (e.g., \citealt{pollack96}, \citealt{piso14}).
-Some gas accretion may occur at earlier times, however,%However, protoplanetary cores start acquiring their envelope at earlier times,
-before the core is fully formed (e.g., \citealt{rafikov06}).
-Recent models such as aerodynamic pebble accretion \citep{lambrechts12} suggest that rapid core growth on timescales of $10^5$ years is possible.
-% which implies that the time when cores accrete their massive envelope may be at least an order of magnitude shorter than the disk lifetime.
-The composition of giant planet atmospheres, and specifically their C/O ratio, can thus depend on the abundance of H$_2$O, CO$_2$ and CO in gas at earlier times than $t_{\rm d}$ in the disk evolution.
-
-Figure \ref{fig:timeplots} shows the particle desorption or final distance as a function of a particle's initial location in the disk, for ice particles %grains 
-of initial sizes of 10 cm and 1 m, composed of either H$_2$O, CO$_2$ or CO.
-These sizes are important since radial drift timescales are shortest for particles within this size range (see Figure \ref{fig:timescales}) --- these are the particles whose drift and desorption evolution should be most strongly affected by variations in disk conditions.
-We choose the evolving disk as a disk model, and we stop the simulations after $10^4$ yr, $10^5$ yr, 1 Myr and $t_{\rm d}=3$ Myr, respectively.
-%, to see how the snowline locations evolve with time and affect the C/O ratio.
-The most important result of these plots is that particles of a given size always desorb at the same disk radii, the 3 Myr snowline, regardless of simulation stopping time.
-Particles %of a given size
-that start at large stellocentric distances do not desorb within the shorter timeframes, e.g.
-$10^4$ or $10^5$ years, but they do evaporate at a fixed radius if their initial location is closer to the host star.
-While the amount of material that moves through the disk changes with time, the radius at which particles desorb and
-the snowline locations are thus independent of the time elapsed, and %. Therefore, 
-our results for the snowline locations are %not only valid at $t_{\rm d}$, but also throughout the earlier 
-valid throughout the time evolution of the protoplanetary disk.
-
-We choose as a fiducial model a total disk mass $M=0.1 M_{\odot}$.
-Observationally, disk masses span at least an order of magnitude around Solar type stars %, but this number is not universally valid.
-From observations of the dust continuum,
-\citep{andrews13}.
-% find that a linear scaling $M \propto M_*$ is reasonable.
-Giant planets, however, have been detected around small stars (e.g., \citealt{montet14}), which can have masses as low as $M_* \sim 0.1 M_{\odot}$.
-We thus explore the effect of disk mass on the location of snowlines.
-Figure \ref{fig:varMd} shows the desorption or final distance as a function on the initial location of a H$_2$O particle with initial size of 1 m, for two total disk masses: $M=0.1 M_{\odot}$, our fiducial model, and $M=0.01 M_{\odot}$.
-Similarly to Figure \ref{fig:timeplots}, we perform our calculations for an evolving disk.
-The simulations are stopped after the same timeframes as those in Figure \ref{fig:timeplots}.
-The location of the H$_2$O snowline is the same for both disks (the same holds true for the CO$_2$ and CO snowlines).
-The C/O ratio is thus insensitive to the choice of $M$.
-We note that our conclusions regarding the disk age and total mass are only valid if the snowline itself does not move with time or disk mass (see also Sections \ref{sec:disk} and \ref{sec:neglected}).
-
-%This result also has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
-Since the snowline locations are independent of disk mass, we expect transition disks to exhibit the same drift-desorption behavior as standard disks under the simplified assumptions of our model.
-%[Comment on differences, i.e. short time frames there is a difference because of smaller drift velocity in the lighter disk I presume.]
-
-We also apply our evolving disk model to a transition disk, i.e.
-a protoplanetary disk with an inner cavity significantly depleted of gas.
-We choose a disk with an inner gap of radius $r_0=4$ AU, consistent with observations of TW Hya \citep{zhang13}, and with the gas surface density in the gap reduced by a factor of 1000.
-Figure \ref{fig:cavity} shows the desorption or final distance for a H$_2$O particle of initial size of 1 m, with the simulation stopped at the same timescales as in Figures \ref{fig:timeplots} and \ref{fig:varMd}.
-Particles that start at an initial distance interior to the gap drift towards the original snowline, while grains located exterior to the gap stop shortly after crossing the gap edge, due to the decrease in gas pressure inside the cavity, thus forming a snowline at $\sim$$3.8$ AU.
-This is qualitatively consistent with the observations of \citet{zhang13}, which show that the H$_2$O snowline is pushed outwards in a transition disk compared to a full disk.
-Our model framework is thus generally valid for more complicated disk structures as well.
-
-%[Need a sentence about how this makes the model framework also generally valid for more complicated disk structures.]
-
-\subsection{Model Extensions}
-\label{sec:neglected}
-
-Our goals in this paper were (1) to gain a qualitative and quantitative understanding of the effect of radial drift and gas accretion onto the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion.
-We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical processes.
-In what follows, we discuss these limitations and their effects.
-We note that our future work will address some of these issues.
-
-We summarize in Table 1 the potential physical and chemical processes occurring in disks and their effect on snowline locations compared to a static disk.
-For the sake of completeness, Table 1 also includes the processes addressed in this paper, i.e.
-radial drift and gas accretion.
-The neglected effects are discussed in more detail below.
-
-\begin{enumerate}
-\item \textbf{Particle growth.} While our model assumes a range of particle sizes, each size is considered initially fixed for a given grain before it drifts and desorbs, since we do not take into account particle coagulation.
-However, grain growth has been observed in protoplanetary disks (e.g., \citealt{ricci10}, \citealt{perez12}), as well as theoretically constrained (e.g., \citealt{birnstiel10}, \citealt{birnstiel12}).
-In Section \ref{sec:COratio} we have shown that larger grains move the snowline locations closer in, but those locations remain fixed above a certain particle size.
-Particle growth will thus initially push the snowlines inwards.
-This is consistent with particle growth models, which predict a maximum particle size often around or below the particle sizes that drift the fastest \citep{birnstiel12}.
-As the largest grains contain most of the solid mass, grain growth models should produce snowlines corresponding to our snowline location estimates for the largest grains in the particle size distribution.
-However, once the solids grow larger than km-sized and form planetesimals, they are no longer affected by drift or desorption, and the snowline reduces to that of a static disk.
-%It follows that grain growth will eventually push the snowline location outwards.
-
-\item \textbf{Turbulent diffusion.} The radial drift model presented in Section \ref{sec:drift} only considers a laminar flow and thus ignores turbulence.
-However, the disk gas also experiences turbulent diffusion (e.g., \citealt{birnstiel12}, \citealt{alidib14}).
-Turbulence causes eddies and vertical mixing, which are likely to reduce the radial drift velocity of the solids (e.g., \citealt{youdin07}).
-Additionally, the flow of H$_2$O, CO$_2$ and CO vapor will diffuse radially.
-Back-diffusion across the snowline will change the shape of the snowline, as well as the C/O ratio in gas and dust both inside and outside of the snowline, due to the reduction of gas-phase volatile abundance interior to the snowline.
-
-\begin{table}[t!]
-\caption{The effects of dynamical and chemical processes on snowline shapes and locations}
-\begin{center}
-\begin{tabular}{|l|l|}\hline
-\textbf{Process} & \textbf{Effect} \\\hline
-Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star compared to the static snowline, $\rightarrow$ means that the snowline is pushed further from the host star compared to the static snowline.
-The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
-Gas accretion & $\leftarrow$ \footnote{Gas accretion pushes the snowlines inwards compared to the snowline locations in a static disk.
-However, accretional heating may push the snowline outwards compared to an evolving disk without accretion heating.} \\\hline
-Particle growth & $\leftarrow$ \footnote{As stated in the main text, if particles grow to km-sizes and above, the snowline remains the same as that of a static disk.} \\\hline
-Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
-Particle fragmentation & $\rightarrow$ $\leftarrow$ \\\hline
-Grain morphology & $\rightarrow$ \\\hline
-Particle composition & $\rightarrow$ $\leftarrow$ \\\hline
-Disk gaps and holes & $\rightarrow$ \\\hline
-Accretion rate evolution & $\rightarrow$ $\leftarrow$ \\\hline
-Stellar luminosity evolution & $\leftarrow$ \\\hline
-Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
-\end{tabular}
-
-\end{center}
-\end{table}
-%causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.
-
-\item \textbf{Particle fragmentation.} Frequent particle collisions in disks cause them to fragment (e.g., \citealt{birnstiel12}).
-The fragmentation of meter- to km-sized particles will move the snowlines outwards, as smaller particles desorb faster and further out from the host star (cf.
-Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}).
-Large boulders, which neither drift nor desorb, may become e.g.
-meter-sized due to collisions and subsequent fragmentation, which will cause them to drift significantly before desorbing, pushing the snowlines inwards.
-Thus fragmentation can move the snowline locations in either radial direction --- specifically, fragmentation leads to a certain grain size distribution, and the largest particles in this size distribution are the ones that determine the position of the snowline.
-
-\item \textbf{Grain morphology.} Our model assumes that the ice particles are perfect, homogeneous spheres.
-However, this is not a very good approximation, since grain growth can be fractal rather than compact (\citealt{zsom10}, \citealt{okuzumi12}, \citealt{krijt15}).
-The inhomogeneity due to cracks in the grain structure will cause the particles to desorb faster.
-They will therefore drift less before evaporating and will move the snowlines less far inward.
-
-\item \textbf{Particle composition.} The ice particles in our model are assumed to be fully made of either H$_2$O, CO$_2$ or CO.
-More realistically, grains may have a layered structure, such as an interior composed of non-volatile materials (e.g., sillicates) covered by an icy layer.
-The ice thus only constitutes a fraction of the total particle mass, which accelerates its desorption and pushes the snowlines outwards.
-The grains may also be composed of a mixture of H$_2$O, CO$_2$ and CO ices, which will increase the binding energies of the more volatiles species, moving %.
-In this case, the desorption energies are largest for the most volatile species, which will move
-the snowlines inwards.
-
-%The relative fraction of each volatile, as well as their degree of mixing, will determine their desorption timescale, and as a result whether the snowlines are moved towards or away from the central star.
-
-\item \textbf{Disk gaps and holes.} The snowline locations will be different for transition disks, which have inner cavities significantly depleted of gas (e.g., \citealt{espaillat12}, \citealt{vandermarel15}), or pre-transitional disks, which have a gap between an inner and outer full disk (e.g., \citealt{kraus11}).
-The decrease in gas pressure in these gaps or holes will reduce the particles' drift velocity close to the gap edge, thus slowing them down and pushing the snowline outwards.
-
-\item \textbf{Accretion rate evolution.} Our viscous disk model assumes a constant mass accretion rate $\dot{M}$.
-However, $\dot{M}$ decreases over time, which lowers the accretional component of the disk temperature (Equation \ref{eq:Tdact}), thus pushing the snowline location inwards if the disk is optically thick (\citealt{garaud07}).
-Once $\dot{M}$ reaches low enough values for the snowline to become optically thin, the snowline location moves outwards \citep{garaud07}.
-\textbf{During the giant planet formation stage of a few Myr, however, $\dot{M}$ steadily decreases with time \citep{chambers09}, which may result in the inward movement of the H$_2$O snowline by up to one order of magnitude, significantly larger than the inward movement caused by radial drift (cf.
-Section \ref{sec:COratio}).
-We thus acknowledge that the location of the H$_2$O snowline may be set by the mass accretion rate evolution rather than the drift of solids.}
-
-%Our choice of $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ roughly corresponds to the mass accretion rate during the quasi-static accretion phase of the disk evolution predicted by \citet{garaud07}.
-This is the timescale on which giant planets accrete most of their gaseous envelope.%; thus the implications of our results for the atmospheric composition of gas giants are robust.
-
-%our results for the snowline location and C/O ratio in a steady-state active disk will only be weakly affected by
-
-\item \textbf{Stellar luminosity evolution.} As the host star contracts during its pre-main sequence phase, its luminosity decreases, which reduces the disk temperature and pushes the snowline locations inwards.
-\citet{kennedy06} found that the snowline is unlikely to move significantly during the pre-main sequence phase for Solar type stars, but it may move inward by a factor of $\sim$$15-20$ for $M_* \sim 0.25 M_{\odot}$ due to the stellar contraction.
-
-\item \textbf{Time dependent chemistry.} As the goal of this paper was to explore only the dynamical effects on snowline locations and the C/O ratio in disks, we have assumed a simple, static chemical model.
-In reality, the chemistry in most of the disk is expected to be time-dependent.%However, gas-grain chemistry is very complex and time-dependent.
-In the inner disk, chemistry approaches equilibrium due to intense sources of ionizing radiation (e.g., \citealt{ilgner04}), while in the outer disk high energy radiation and cosmic rays are the key drivers of chemistry, which is no longer in equilibrium (e.g., \citealt{vandishoeck06}).
-A multitude of chemical evolution models have been developed (see references in \citealt{henning13}), many of which contain tens or hundreds of chemical reactions.
-Due to the complexity of these chemical models, most of them are decoupled from disk dynamics.
-The effect of disk chemistry on snowline locations, shape, time evolution, or the C/O ratio is therefore difficult to estimate.
-%In a future paper, we plan to use a parametrized chemical model and incorporate it in our radial drift calculation.
-[let's not promise this until we have figured out how to do it.]
-
-%parametrize the chemical model developed by Merchantz et al. (in preparation) that traces the evolution of the main C and O oxygen carriers,
-
-\end{enumerate}
-
-%\emgr{Present the diagram that shows all the effects that can modify snowline location.
-For model limitations, include: non-inclusion of turbulence, assumption of perfect spheres when in fact they may have cracks, particles composed of a single volatile when in reality they are likely to be mixed, etc.
-Discuss uncertainty of initial conditions and estimate how much they matter.
-....}
-
-\section{Summary}
-\label{sec:summary}
-
-We study the effect of radial drift of solids and viscous gas accretion onto the central star on the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in a protoplanetary disk, assuming static chemistry.
-We develop a simplified model to describe the coupled drift-desorption process and determine the time evolution of particles of different sizes throughout the disk.
-We assume that the solid particles are perfect, homogeneous spheres, fully composed of either H$_2$O, CO$_2$ or CO.
-We apply our model to an irradiated disk, an evolving disk, and a viscous disk that also takes into account stellar irradiation.
-We determine the desorption or final location of drifting particles after a time equal to the disk lifetime, and use this result to set an inner limit for the location of the H$_2$O, CO$_2$ and CO snowlines.
-Our results can be summarized as follows:
-
-\begin{enumerate}
-\item Radial drift and gas accretion affect desorption and move the snowline locations inward compared to a static disk for particles with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk.
-%Moreover, boulders up to almost kilometer-size contribute to setting the snowline locations if they are very close to the snowlines.
-[I find this conclusion a bit iffy and also it sounds like it is in contradiction with the previous sentence.]
-
-\item For our simplified model that does not account for the effects outlined in Section \ref{sec:neglected}, particles with sizes in the above range desorb almost instantaneously once desorption has begun, and at a fixed location in the disk that only depends on the particle size and the gas accretion rate.
-Thus for each particle size there is a fixed and uniquely determined H$_2$O, CO$_2$ or CO snowline.
-
-\item The results of our numerical simulation are in agreement with the analytic solution of the drift-desorption system of differential equations if the stopping time $\tau_{\rm s} \ll 1$.
-We present an explicit analytic solution for the desorption distance in this regime.
-
-\item Since realistic grain size distributions are dominated in mass by the largest particles, the H$_2$O, CO$_2$ and CO snowlines are those created by the largest drifting particles in our model.
-This corresponds to the innermost snowlines that we determine.
-Our model thus sets a limit on how close to the central star the snowlines can be pushed by radial drift and gas accretion.
-
-\item The snowline locations move inwards as the particle size increases; the innermost snowline is set by particles with initial size $s \sim 7$ m in our model --- bigger particles drift too slowly to make it further in before desorbing (see Section \ref{sec:snowlines}).
-Gas accretion causes even micron-sized particles to drift, desorb and move the snowline location compared to a static disk.
-A viscous disk that includes accretion heating moves the H$_2$O snowline outwards compared to an evolving disk, but has no effect on the CO$_2$ and CO snowline locations, for our particular choice of mass accretion rate $\dot{M}$ and midplane opacity $\kappa$.
-%However, the magnitude of the effect of accretional heating on snowline locations will depend on .
-
-\item For our fiducial model, which considers particles with sizes between $10^{-3}$ and $10^8$ cm, the innermost H$_2$O, CO$_2$ and CO snowlines are located at 0.3 AU, 3.7 AU and 30 AU for an irradiated disk, 0.26 AU, 3.4 AU and 30 AU for an evolving disk, and 0.63 AU, 3.4 AU and 30 AU for a viscous disk with accretion heating.
-Compared to a static disk, radial drift and gas accretion move the snowlines by up to 60 \% for H$_2$O and CO$_2$, and by up to 50 \% for CO.
-For the viscous disk, however, which is the most realistic of the three models since it takes into account accretion heating, the H$_2$O snowline location moves inwards by up to 40 \%.
-
-\item Our C/O estimates confirm the conclusions of \citet{oberg11} that the C/O ratio \textbf{in gas} may be enhanced compared to the stellar value throughout most of the disk, with the C/O ratio reaching its maximum value between the CO$_2$ and CO snowlines.
-%This is consistent with possible detections of superstellar C/O ratios in some exoplanet atmospheres.
-We note, however, that our results for the C/O ratio do not take into account the radial movement of the desorbed ices with the accreting gas in the evolving and viscous disks, which may significantly decrease the C/O ratio in gas inside the H$_2$O and CO$_2$ snowlines.
-We plan to address this issue in a future paper.
-
-\item For a constant gas mass accretion rate $\dot{M}$ and stellar luminosity $L_*$, the snowline locations are independent of the time at which we stop our simulation and of the total disk mass, as long as the disk midplane remains optically thick.
-%Our results are thus valid throughout the evolution of the gas disk, not only after the disk has dissipated, which has implications on the composition of nascent giant planets.
-
-\end{enumerate}
-
-Our model does not address additional effects, such as gas diffusion, grain composition and morphology, or complex time-dependent chemical processes.
-Future work will address some of these dynamical and chemical processes, with the goal of obtaining more realistic results for the snowline locations, shapes and time evolution, and the resulting effect on the C/O ratio.
-%Nevertheless,
-
-\appendix
-\section{Steady-state active disk solution} \label{app:steadystate}
-
-Following \citet{shakura73} and \citet{armitage10}, the steady-state solution for a geometrically thin, optically thick actively accreting disk with an $\alpha$-prescription for viscosity is governed by the following set of equations:
-
-\begin{subeqnarray}
-\label{eq:diskeq}
-\nu & = & \alpha c H \slabel{eq:nueq} \\
-c^2 & = & \frac{k_{\rm B} T_{\rm act}}{\mu m_{\rm p}} \slabel{eq:cdeq} \\
-\rho & = & \frac{1}{\sqrt{2 \pi}} \frac{\Sigma}{H} \slabel{eq:rhoeq} \\
-H & = & \frac{c}{\Omega_{k}} \slabel{eq:Heq} \\
-T_{\rm act}^4 & = & \frac{3}{4} \tau T_{\rm surf}^4 \slabel{eq:Teq} \\
-\tau & = & \frac{1}{2} \Sigma \kappa \slabel{eq:taueq} \\
-\nu \Sigma & = & \frac{\dot{M}}{3 \pi} \slabel{eq:Mdot} \\
-\sigma T_{\rm surf}^4 & = & \frac{9}{8} \nu \Sigma \Omega_{\rm k}^2 \slabel{eq:nusigeq} \\
-\kappa & = & \kappa_0 T_{\rm act}^2 \slabel{eq:kappaeq},
-\end{subeqnarray}
-where $T_{\rm surf}$ is the surface temperature of the disk and the other quantities are defined in the main text.
-This is a system of nine equations with nine unknowns ($\nu$, $c$, $H$, $T_{\rm act}$, $\rho$, $\Sigma$, $\tau$, $T_{\rm surf}$, $\kappa$) that can be solved numerically once $\alpha$ and $\kappa_0$ are specified.
-
 \section{Desorption distance analytic solution} \label{app:tdriftan}
@@
 where $r_0=1$ AU, $\rho_0=\rho(r_0)$ and $c_0=c(r_0)$.
-
-%
-%\emgr{Right now it's unclear to me what could go in an appendix, if anything.
-Maybe discuss a bit the algorithm to evolve $\Sigma_{\rm p}$ (although it is already explained in detail in the appendix of Birnstiel et al.
-2010).
-Maybe show some example profiles of $\Sigma_{\rm p}$ at different times and for different particle sizes.}
-
-\if\bibinc n
-\bibliography{refs}
-\fi
-
-\if\bibinc y
-\begin{thebibliography}
-\end{thebibliography}
-\fi
-
-\end{document}

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/chapter2.tex
upd

%\begin{savequote}[75mm]
%This is some random quote to start off the chapter.
%\qauthor{Firstname lastname}
%\end{savequote}

\chapter{Minimum Core Masses for Giant Planet Formation With Realistic Equations of State and Opacities}

\section*{Abstract}

Giant planet formation by core accretion requires a core that is sufficiently massive to trigger runaway gas accretion in less that the typical lifetime of protoplanetary disks. We explore how the minimum required core mass, $M_{\rm crit}$, depends on a non-ideal equation of state and on opacity changes due to grain growth, across a range of stellocentric distances from 5-100 AU. This minimum $M_{\rm crit}$ applies when planetesimal accretion does not substantially heat the atmosphere.  Compared to an ideal gas polytrope, the inclusion of molecular hydrogen (H$_2$) dissociation and variable occupation of H$_2$ rotational states increases $M_{\rm crit}$.  Specifically, $M_{\rm crit}$ increases by a factor of $\sim$$2$ if the H$_2$ spin isomers, ortho- and parahydrogen, are in thermal equilibrium, and by a factor of $\sim$$2-4$ if the ortho-to-para ratio is fixed at 3:1.  Lower opacities due to grain growth reduce $M_{\rm crit}$. For a standard disk model around a Solar mass star, we calculate $M_{\rm crit} \sim 8 M_{\oplus}$ at 5 AU, decreasing to $\sim$$5 M_{\oplus}$ at 100 AU, for a realistic EOS with an equilibrium ortho-to-para ratio and for grain growth to cm-sizes. If grain coagulation is taken into account, $M_{\rm crit}$ may further reduce by up to one order of magnitude. These results for the minimum critical core mass are useful for the interpretation of surveys that find exoplanets at a range of orbital distances.

%Giant planet formation by core accretion requires a core that is sufficiently massive to trigger runaway gas accretion in less that the typical lifetime of protoplanetary disks. We explore how the minimum required core mass, $M_{\rm crit}$, depends on a non-ideal equation of state and on opacity changes due to grain growth, across a range of stellocentric distances from 5--100 AU. This minimum $M_{\rm crit}$ applies when planetesimal accretion does not substantially heat the atmosphere.  Compared to an ideal gas polytrope, including molecular hydrogen (H$_2$) dissociation and variable occupation of H$_2$ spin states increase $M_{\rm crit}$.  We analyze how these EOS effects alter both the atmospheric energy and cooling luminosity, which together determine the atmospheric growth timescale.  Specifically, $M_{\rm crit}$ increases by more than a factor of two if the H$_2$ spin isomers, the ortho and para states, are in thermal equilibrium, and by up to a factor of $\sim$$3.5$ if the ortho-to-para ratio is fixed at 3:1. This latter increase is most pronounced for disk temperatures below $\sim$ xxK.  Lower opacities due to grain growth reduce $M_{\rm crit}$, especially at higher temperatures (check temp dependence, note possibly by an order of magnitude).
 %For a standard disk model around a Solar mass star, we calculate $M_{\rm crit} \sim 7 M_{\oplus}$ at 5 AU, decreasing to $\sim$$4.5 M_{\oplus}$ at 100 AU, for a realistic EOS with an equilibrium ortho-para ratio and for grain growth to cm-sizes.   Without grain growth, the rise in $M_{\rm crit}$ towards shorter separations is steeper, due to the greater temperature dependence of the opacity. These results for the minimum critical core mass are useful for the interpretation of surveys that find exoplanets at a range of orbital distances.

%The descriptive result for Mcrit and p = 2.5  probably doesn't belong in the abstract as the actual result isn't included in the paper/figures.  There's a $t_run$ for this case, but not $M_crit$.

 %, an important parameter for understanding the origin of gas giants directly imaged at wide separations.

%We build on the idealized model of Piso \& Youdin (ApJ, accepted), which derives the evolution of atmospheres accreting around protoplanetary cores and calculates $M_{\rm crit}$ in the limiting regime in which core growth is halted and the luminosity evolution of a core's atmosphere is dominated by Kelvin-Helmholtz contraction. We enhance this model by employing a more realistic equation of state and dust opacity.

%The drop in $M_{\rm crit}$ with semimajor axis is less sharp when grain growth opacities are used than in the case of interstellar opacities. We show that giant planets may grow from smaller cores if core growth is halted, which is particularly relevant for understanding the origin of gas giants directly imaged at wide separations.

%The core accretion model postulates that giant planets form through gas accretion onto a solid core. To accumulate a massive atmosphere, this core's mass must exceed a minimum value $M_{\rm crit}$ often quoted as $10 M_{\oplus}$. The value of $M_{\rm crit}$, however, depends on the rate at which planetesimals accrete onto the core and heat its atmosphere.

 %In this study we calculate the minimum core mass for atmospheric collapse in the limiting regime in which core growth is halted and the luminosity evolution of a core's atmosphere is dominated by Kelvin-Helmholtz contraction.

%Thus halting a core's growth may facilitate giant planet formation in the outer disk, which helps understand the origin of giant planets directly imaged at wide separations.

%It is thus easier to form gas giants from fully grown cores, which may facilitate giant planet formation through core accretion in the outer disk and may be helpful for understanding the origin of directly imaged giant planets at wide separations.

 %However, this minimum core mass decreases back to $M_{\rm{c, crit}}\sim5 M_{\oplus}$ once we assume realistic dust opacities that take into account grain growth. Our results yield lower critical core masses than corresponding studies for large planetesimal accretion rates. We therefore conclude that it is easier to form a planet by growing the core first, then accreting a massive gaseous envelope, rather than forming the core and atmosphere simultaneously.

%requires rapid planetesimal accretion, which deposits energy into the atmosphere.

%Heating increases the core mass required for collapse. However, the planetesimal accretion rate need not be constant --- in particular, it may decline over time. In this study we consider the limiting regime

 %When a core embedded in a protoplanetary disk accretes an atmosphere with mass comparable to itself, the envelope is no longer in hydrostatic balance and unstable atmosphere collapse occurs.

 %The core accretion model proposes that giant planets form by the accretion of gas onto a solid protoplanetary core. Previous studies have found that there exists a ``critical core mass'' past which hydrostatic solutions can no longer be found and unstable atmosphere collapse occurs. In standard calculations of the critical core mass, planetesimal accretion deposits enough heat to alter the luminosity of the atmosphere, increasing the core mass required for the atmosphere to collapse. In this study we consider the extreme case in which planetesimal accretion is negligible and Kelvin-Helmholtz contraction dominates the luminosity evolution of the planet. We develop a two-layer atmosphere model with an inner convective region and an outer radiative zone that matches onto the protoplanetary disk, and we determine the minimum core mass for a giant planet to form within the typical disk life timescale for a variety of disk conditions, which we denote as  ``critical core mass''.  We find that the absolute minimum core mass required to nucleate atmosphere collapse within the disk lifetime is smaller for planets forming further away from their host stars. Moreover, the critical core mass is strongly dependent on disk temperature, opacity and mean molecular weight of the gas

\section{Introduction}
\label{intro}

Core accretion --- a prominent theory of giant planet formation --- stipulates that Jupiter-sized planets form when planetesimal accretion produces a solid core large enough to attract a massive atmosphere   (e.g., \citealt{mizuno78}, \citealt{stevenson82}, \citealt{boden86}, \citealt{wuchterl93}, \citealt{dangelo11}). Because protoplanetary disks dissipate on timescales of a few Myrs \citep[e.g.,][]{jay99}, cores must grow quickly to accrete disk gas. Fast  growth is particularly challenging far from a core's host star, where dynamical times are long. %, making the production of massive cores within the disk lifetime less likely.
Understanding whether core accretion can work in the outer disk may provide valuable insights into the formation of wide-separation planets such as the directly-imaged giants orbiting HR 8799 \citep{marois08}.

Production of such planets \textit{in situ} by core accretion requires faster core growth rates than typically assumed.  However, even if fast  growth could be achieved, it produces an additional difficulty.  Though the minimum, or critical, core mass to form a giant planet is often quoted as $M_{\rm crit}\sim 10M_{\oplus}$, the actual value depends on how quickly the core accretes planetesimals  (e.g.,  \citealt{pollack96}, \citealt{ikoma00}, \citealt{rafikov06}).  Fast accretion heats a core's atmosphere, increasing pressure support and hence $M_{\rm crit}$.  \citet{rafikov11} finds that beyond 40-50 AU, a core cannot  reach $M_{\rm crit}$ while accreting at the rate it requires to grow during its host disk's lifetime.

%Given standard core accretion rates, the $M_{\rm crit}$ you need to grow an atmosphere during the disk lifetime while the core is growing is too large, leading people to believe that core accretion doesn't work in the outer disk.

This limit on the distance at which core accretion operates may be overcome if core growth does not proceed at a constant rate. Time-dependent models \citep[e.g.,][]{pollack96,ikoma00} suggest that a planets's feeding zone may be depleted of planetesimals before disk dissipation, causing core growth to stall. Planetesimal accretion no longer deposits energy into the atmosphere which, without this balance for radiative losses, cannot maintain a steady state.  The envelope accretes gas while  undergoing Kelvin-Helmholtz (KH) contraction.  In this regime, a core of any mass would evolve into a giant planet if given an infinite amount of time.  In practice, a core produces a giant planet if it has enough time to accumulate an atmosphere of approximately its own mass before its host disk dissipates.  The minimum core mass satisfying this criterion is $M_{\rm crit}$.

Because extra heating increases an atmosphere's pressure, opposing runaway envelope growth,  $M_{\rm crit}$ is smallest in the absence of planetesimal accretion.
 While $M_{\rm crit}$ has been systematically computed as a function of disk properties and stellocentric separation for steady-state atmospheres heated by planetesimal accretion \citep{rafikov06}, no equivalent systematic study is available for this minimum value of $M_{\rm crit}$.

In \citet[hereafter Paper I]{piso14} and this paper, we provide such a systematic study.   Given a fiducial disk model, we calculate the $M_{\rm crit}$ required to nucleate runaway atmospheric growth for a fully formed---and no longer accreting---core.  We model the planet's evolution using a series of quasi-static two-layer atmospheres embedded in a protoplanetary disk. Here, we build on the results of Paper I by making two important additions: (1) a realistic equation of state (EOS), and (2) realistic dust opacities.  Our aim is twofold: (1) to explain how hydrogen and helium's non-ideal EOS affects atmospheric growth when compared to an ideal gas, and (2) to provide realistic estimates for the minimum critical core mass required to form a giant planet over a range of semimajor axes.

 %, as additional heating due to accretion of solids would limit the envelope's ability to cool and increase the core mass required for collapse.
%if core growth is stalled during the atmospheric accretion phase, the target $M_{\rm crit}$ decreases, conceivably making runaway atmospheric growth possible in the outer disk.

%the envelope mass remains small during this initial phase of core growth, and runaway gas accretion cannot commence. Once the planet's feeding zone has been depleted of solids, planetesimal accretion slows enough that the resulting energy deposition can no longer balance radiative losses.

%However, the planetesimal accretion rate, $\dot{M_{\rm c}}$, need not be constant over time and may significantly reduce as the planet's feeding zone is depleted of solids.

%While a core is accreting the majority of its mass, $\dot{M_{\rm c}}$ is large, and the energy its atmosphere receives from infalling planetesimals balances its luminosity. The core's envelope is in steady state, and $M_{\rm crit}$ is uniquely determined for a fixed $\dot{M_{\rm c}}$ and a set of disk conditions. Time-dependent core accretion models, such as \citet{pollack96} and \citet{ikoma00}, suggest that the envelope mass remains small during this initial phase of core growth, and runaway gas accretion cannot commence. Once the planet's feeding zone has been depleted of solids, planetesimal accretion slows enough that the resulting energy deposition can no longer balance radiative losses. Core growth stalls, and the atmosphere cools, undergoes Kelvin-Helmholtz (KH) contraction, and steadily accretes gas. In this accretion regime,  $M_{\rm crit}$ is no longer unique ---  any core may produce a giant planet if it has enough time to accumulate an atmosphere of comparable mass to itself before disk dissipation. Given the accretion history employed by \cite{pollack96}, which neglects drift of planetesimals through the gas disk, the atmospheric cooling phase lasts much longer than the phase of core growth, and so the the time requires to form a giant planet is set by KH contraction.

%It is not clear \textit{a priori} which regime --- atmosphere evolution dominated by planetesimal accretion versus KH contraction --- is the most relevant, as the conclusion of \citet{pollack96} depends on their assumptions about planetesimal accretion history. In particular, a core's feeding zone may be replenished by migration of solids through the protoplanetary disk.

%This study does not model an accretion history, but instead calculates a minimum value for $M_{\rm crit}$. We accept the two-stage growth suggested by time-dependent models, in which the core is fully formed before it accumulates a significant atmosphere. During the slow phase of

%In this paper, we calculate the $M_{\rm crit}$ required to nucleate runaway atmospheric growth for a fully formed core undergoing minimal planetesimal accretion. We assume that the luminosity of the envelope is solely due to KH contraction. As additional heating due to accretion of solids would limit the envelope's ability to cool, and thus increase the core mass required for collapse, our results give a robust minimum for the core mass needed to form a giant planet before disk dissipation.

%Our results yield lower core masses than studies that assume large planetesimal accretion rates, confirming that if cores stop growing, they can accrete massive atmospheres more effectively. Moreover, 
%However, planetesimal accretion rates need not be constant over time. This is supported by time-dependent core accretion models, such as \citet{pollack96}, \citet{ikoma00}, which suggest that the evolution of a giant planet can be separated into three phases. In phase 1, the core forms via runaway planetesimal accretion, while the envelope mass stays small. Once the planet's feeding zone has been depleted of solids, planetesimal accretion is significantly reduced and can no longer balance radiative losses. As a result, core growth is stalled and the atmosphere cools while undergoing Kelvin-Helmholtz (KH) contraction; this is phase 2. The envelope mass steadily increases. When it becomes comparable to the core mass, runaway gas accretion begins in phase 3. \citet{pollack96} find that phase 2 lasts much longer than phases 1 and 3, and so the evolutionary timescale of the planet is set by KH contraction.

%It is not clear \textit{a priori} which regime --- atmosphere evolution dominated by planetesimal accretion versus KH contraction --- is the most relevant, as the conclusion of \citet{pollack96} depends on their assumptions about planetesimal accretion history. In particular, a core's feeding zone may be replenished by migration of solids through the protoplanetary disk. This study does not model an accretion history, but instead calculates a minimum value for $M_{\rm crit}$. To this purpose, we must first understand how $M_{\rm crit}$ depends on the planetesimal accretion rate, $\dot{M_{\rm c}}$. For large $\dot{M_{\rm c}}$, the energy due to planetesimals balances the planet's luminosity and the core's envelope is in steady state. In this case, $M_{\rm crit}$ is uniquely determined for a fixed $\dot{M_{\rm c}}$ and a set of disk conditions. For small $\dot{M_{\rm c}}$, this energy balance is no longer maintained and the gaseous envelope loses energy via KH contraction. Moreover, $M_{\rm crit}$ is no longer unique ---  any core may initiate runaway gas accretion if it can accumulate a massive atmosphere before disk dissipation.

%In this study, we accept the three-stage growth, in which the core is fully formed before it accumulates a significant atmosphere. During the slow phase of atmospheric growth we thus assume that the luminosity of the envelope is only due to KH contraction and neglect ongoing planetesimal accretion. Our results yield lower core masses than studies that assume large planetesimal accretion rates, confirming that if cores stop growing, they can accrete massive atmospheres more effectively. Moreover, as additional heating due to accretion of solids limits the envelope's ability to cool and thus increases the core mass required for collapse, our results give a robust minimum for the core mass needed to form a giant planet before disk dissipation.

%Studying atmosphere growth onto a solid core of fixed mass (e.g., \citealt{pn05}) can thus provide accurate estimates for the timescale required to form a giant planet.

%However, the planetesimal accretion rates, $\dot{M_{\rm c}}$, need not be constant over time, and may significantly reduce as the planet's feeding zone is depleted of solids.  If  accretion slows before the nebular gas dissipates, the target $M_{\rm crit}$ for successful giant planet formation is substantially reduced. 
%
%To calculate the minimum value of $M_{\rm crit}$, we must first understand how $M_{\rm crit}$ depends on $\dot{M_{\rm c}}$. For large $\dot{M_{\rm c}}$, planetesimals deposit sufficient energy to balance the planet's luminosity and the core's envelope is in steady state. In this case,  $M_{\rm crit}$ is uniquely determined for a fixed accretion rate of solids and a set of disk conditions. For small $\dot{M_{\rm c}}$, this energy balance cannot be maintained and the gaseous envelope loses energy via Kelvin-Helmholtz contraction, resulting in gas accretion. Moreover, $M_{\rm crit}$ is no longer uniquely determined ---  any core has the potential to initiate runaway gas accretion, provided it can accumulate a massive atmosphere before disk dissipation.  
%
%Time-dependent core accretion models, such as \citet{pollack96}, \citet{ikoma00}, suggest that the evolution of a giant planet can be separated into three phases. The core forms during phase 1 due to runaway planetesimal accretion, while the envelope mass stays small. Once the planet's feeding zone has been depleted of solids, planetesimal accretion is significantly reduced and can no longer balance radiative losses. As a result, the atmosphere cools while undergoing KH contraction; this is phase 2. The envelope mass now steadily increases while core growth is stalled. Runaway gas accretion begins once the atmosphere mass is comparable to the core mass in phase 3. \citet{pollack96} find that phase 2 lasts much longer than phases 1 and 3, and so the evolutionary timescale of the planet is set by KH contraction. %Studying atmosphere growth onto a solid core of fixed mass (e.g., \citealt{pn05}) can thus provide accurate estimates for the timescale required to form a giant planet.  
%
%It is not clear \textit{a priori} which of the two regimes --- atmosphere evolution dominated by planetesimal accretion versus KH contraction --- is the most relevant, as the conclusion of \citet{pollack96} depends on their assumptions about planetesimal accretion history. In particular, a core's feeding zone may be replenished by migration of solids through the protoplanetary disk. In this study, we accept the three-stage growth, in which the core is fully formed before it accumulates a significant atmosphere. During the slow phase of atmospheric growth we thus assume that the luminosity of the envelope is only due to KH contraction and neglect ongoing planetesimal accretion. Our results yield lower core masses than studies that assume large planetesimal accretion rates, confirming that if cores stop growing, they can accrete massive atmospheres more effectively. Moreover, as additional heating due to accretion of solids limits the envelope's ability to cool and thus increases the core mass required for collapse, our results give a robust minimum for the core mass needed to form a giant planet before disk dissipation.
%

%\newpage 
\pagebreak

\subsection{Equation of State}
\label{sec:introeos}

We use the EOS of hydrogen and helium mixtures calculated by \citet{saumon95}, which captures non-ideal effects such as dissociation, ionization, and selective occupation of quantum states at low temperatures.  We extend these tables to the very low temperatures and pressures required for planets forming in the outer regions of disks (\App{EOStables}).
The EOS tables from \citet{saumon95} have often been used to model the interiors of giant planets (e.g., \citealt{pollack96}, \citealt{ikoma00}, \citealt{alibert05}, \citealt{hubickyj05}, \citealt{pn05}, \citealt{mordasini12}), as well as in complex stellar evolution simulations involving low temperatures (e.g., \citealt{paxton11}, \citealt{paxton13}, which use the code MESA). More recent EOS tables (\citealt{nettelmann08}, \citealt{nettelmann12}, \citealt{militzer13}) are based on ab initio molecular dynamics simulations and thus avoid some of the approximations of the \citet{saumon95} semi-analytical approach. However, though these newer tables are sufficient to model the internal structures of Jupiter and of close-in extrasolar planets,  they do not extend to the low temperatures and pressures required for wide-separation planets ($T \lesssim 500$ K, $P \lesssim 1$ GPa, e.g. \citealt{militzer13}). Moreover, the \citet{militzer13} EOS tables and the \citet{saumon95} tables are in good agreement for entropies, $S$, such that $\log_{10}(S) \gtrsim 8.75$ erg g$^{-1}$ K$^{-1}$.  All models presented here satisfy this constraint.

\subsection{Opacity}
\label{sec:introopacity}

Opacities in protoplanetary disks are unlikely to be interstellar, and are lowered by grain growth and dust settling.  Numerous studies have demonstrated that atmospheric evolution depends strongly on opacity---lower opacities accelerate envelope growth and reduce $M_{\rm crit}$.  For example, the analytic model of \citet{stevenson82} showed that $M_{\rm crit} \propto \kappa^{3/4}$ for a fully radiative envelope with constant opacity $\kappa$. A series of more modern studies conclude that: Jupiter may have formed in 1 Myr with a core of $10 M_{\oplus}$ for an opacity arbitrarily reduced to 2\% of the interstellar value \citep{hubickyj05};  $M_{\rm crit}$ for Jupiter may be as low as $\sim$$1 M_{\oplus}$ for a grain free envelope \citep{hori10}; and the use of grain opacities that take into account coagulation and settling reduces the formation time for Jupiter by up to 80\%  \citep{movshovitz10}. A recent study by \citet{mordasini14} investigates the effect of opacity on the formation of populations of synthetic planets, and finds from comparisons with observations that opacities are likely to be much smaller than interstellar. Lastly, Paper I shows that reducing the opacity from interstellar to 1\% of interstellar reduces $M_{\rm crit}$ by a factor of $\sim$$2$.

Thus, to provide realistic estimates for $M_{\rm crit}$, we must employ realistic opacities.    For temperatures below dust sublimation, we employ opacity tables from \citet{dalessio01}, which are used to model observations of protoplanetary disks.  We present results for both a standard collisional cascade grain size distribution and for a shallower distribution, appropriate for coagulation (see Section \ref{sec:opacity} for details).  At high temperatures, where dust sublimates, we employ the analytic opacity law of \citet{bell94}.  For our purposes, this expression sufficiently represents the more detailed opacity tables of \citet{semenov03} with reduced computational complexity (see \App{radwindow}).  
We note that most previous works that study the quantitative effects of opacity reductions on $M_{\rm crit}$ are primarily focused on the formation of Jupiter at 5.2 AU.  In contrast,
our aim is to provide realistic estimates for $M_{\rm crit}$ for a larger parameter space, and to analyze the dependence of $M_{\rm crit}$ on semimajor axis.

%We use the EOS calculated by \citet{saumon95}, which captures non-ideal effects such as dissociation, ionization, and selective occupation of quantum states at low temperatures.  Opacities in protoplanetary disks are unlikely to be interstellar, and are lowered by grain growth and dust settling. We use the \citet{dalessio01} opacity tables, which are used to model observations of protoplanetary disks.

\subsection{Paper Plan}

After reviewing the quasi-static and cooling models derived in Paper I (Section \ref{sec2}), we add a non-ideal EOS.  Because the  adiabatic gradient, $\nabla_{\rm ad}$, is an important determinant of atmospheric structure, we first explain how dissociation and variable occupation of low-energy rotational states change $\nabla_{\rm ad}$ (Section \ref{deladtable}).  Section \ref{EOSeffects} presents the impact of these effects on atmosphere evolution.  We introduce realistic opacities in Section \ref{sec:opacity} and determine $M_{\rm crit}$ in Section \ref{critical}. Section \ref{acc}  compares our results to those obtained by studies employing high planetesimal accretion rates. We summarize our findings in Section \ref{conclusions}.
%-- which is particularly relevant to our regime of low planetesimal accretion in which solids have been segregated from the gas and are not being added back in ---

%Fast accretion heats a core's  atmosphere, increasing pressure support and hence $M_{\rm crit}$.  For high planetesimal accretion rates, atmospheres are primarily heated by planetesimals (e.g., \citealt{rafikov06}). In this case,  $M_{\rm crit}$ is uniquely determined for a fixed accretion rate of solids and a set of disk conditions.

 %However, planetesimal accretion rates need not be constant over time.  If  accretion slows before the nebular gas dissipates, the target $M_{\rm crit}$ for successful giant planet formation is substantially reduced.
%At distances $\gtrsim40-50$ AU, simultaneous core and atmosphere growth to critical masses may not be possible within a disk lifetime \citep{rafikov11}. This problem may be circumvented if planetesimal accretion slows before the nebular gas dissipates, $M_{\rm crit}$ may be substantially reduced if . %  An alternative solution is core growth at moderate rates and thus to smaller masses, followed by a halt to planetesimal accretion. This study explore the latter scenario.

%Fast accretion generates an additional difficulty.

%An \textit{in situ} solution to this problem requires accelerating core growth to rates faster than those typically assumed. However, the minimum, or critical, core mass to form a giant planet, though typically quoted to be $M_{\rm crit}\sim 10M_{\oplus}$, depends on how quickly planetesimals accrete (e.g., \citealt{ikoma00}, \citealt{pollack96}, \citealt{rafikov06}).  Fast accretion increases $M_{\rm crit}$.  However, $M_{\rm crit}$ may be substantially reduced if planetesimal accretion slows before the nebular gas dissipates.

%or fast  to grow the core and atmosphere simultaneously, much faster rates would be required ---  fast planetesimal accretion increases the core mass necessary for atmospheric collapse. during the era of fast accretion requires larger critical core masses.

%Core accretion at wide separations thus requires either growth of cores at rates faster than typically assumed in the first regime of high planetesimal accretion or, perhaps counterintuitively, moderate core growth rates followed by a halt to planetesimal accretion in the second regime --- i.e., the three-phase regime assumed by \citet{pollack96}.

 %If cores and their atmospheres grow simultaneously, fast core growth requires large planetesimal accretion rates and hence large critical core masses. Core growth to these critical masses is particularly challenging in the outer disk, where long dynamical times slow core growth, making production of a massive core within the disk lifetime less likely.

%Fast accretion implies large critical core masses.

%In order to calculate the minimum value of $M_{\rm crit}$, we need to understand how $M_{\rm crit}$ depends on the planetesimal accretion rate. For high planetesimal accretion rates, atmospheres are primarily heated by planetesimals (e.g., \citealt{rafikov06}). As a result, the core's envelope is in a steady state at all times, in which all the incoming energy due to planetesimals is radiated away.

%the evolution of a core's atmosphere depends on the rate at which planetesimals are accreted.

%As the envelope and core become comparable in mass, hydrostatic balance no longer holds and runaway gas accretion commences. In this case, $M_{\rm crit}$ is the maximum core mass for which the atmosphere is still in hydrostatic equilibrium. For a fixed planetesimal accretion rate and a set of disk conditions, $M_{\rm crit}$ is uniquely determined.

%In models that assume a high planetesimal accretion rate (e.g., \citealt{rafikov06}), the gaseous envelope is heated solely by planetesimals and is therefore in a steady state at all times, in which all the incoming energy due to planetesimals is radiated away. In this regime, the core and atmosphere grow simultaneously, with the atmosphere increasing in mass faster than the core.

%Moreover, $M_{\rm crit}$ is no longer uniquely determined ---  any core has the potential to initiate runaway gas accretion, provided it can accumulate a massive atmosphere before disk dissipation.

%However, planetesimal accretion rates need not be constant at a given location in the protoplanetary disk throughout the disk's lifetime --- in particular, planetesimal accretion may significantly reduce as the planet's feeding zone is depleted of solids.

%For high planetesimal accretion rates, the core's envelope is in steady state at all times, in which all the incoming energy due to planetesimals is radiated away. On the other hand, time-dependent core accretion models (e.g., \citealt{pollack96}, \citealt{ikoma00}) include variable planetesimal accretion rates which allow for the Kelvin-Helmholtz (KH) contraction of the gaseous envelope to become important. In this regime, the envelope is no longer in a steady state, but rather it accretes gas as it cools.

%Fast core growth to critical masses is particularly challenging in the outer disk. Long dynamical times at these distances slow core growth, making production of a massive core within the disk lifetime less likely. We thus investigate the possibility of forming gas giant at wide separations from smaller cores. %Core accretion at wide separations thus requires either growth of cores at rates faster than typically assumed or, perhaps counterintuitively, moderate core growth rates followed by a halt to planetesimal accretion.

%We note, however, that long dynamical timescales in the outer disk slow core growth. Core accretion at wide separations thus requires either growth of cores at rates faster than typically assumed in the first regime of high planetesimal accretion or, perhaps counterintuitively, moderate core growth rates followed by a halt to planetesimal accretion in the second regime --- i.e., the three-phase regime assumed by \citet{pollack96}.

%Both the steady state regime, where planetesimal accretion dominates the atmosphere evolution, and the KH regime, Therefore, it may be easier to grow a giant planet from a fully formed core, which is particularly relevant in the outer disk, where low planetesimal accretion rates prevent the growth of large cores.

 %Paper I assumes that the nebular gas is ideal, diatomic and can be described by a polytropic equation of state (EOS). However, non-ideal effects such as dissociation and ionization have to be taken into account in order to obtain better quantitative results. In this study, we consider atmosphere growth assuming a realistic gas with an EOS prescribed by the \citet{saumon95} tables. %We find that the critical core mass is more than twice as large than in the ideal gas case.
 %We find that the resulting critical core mass may be significantly lower when ISM opacities are used, i.e. by more than a factor of three at 5 AU, and around 1.5 times lower at 100 AU. %We therefore see that it is less sensitive to disk temperature and hence semimajor axis.

% making two important additions: we include realistic equation of state (EOS) effects, as well as realistic dust opacities.

%Fast accretion implies large critical masses. Because protoplanetary disks dissipate on timescales of a few Myrs, such cores must grow quickly to have access to the nebular gas. If cores and their atmospheres grow simultaneously, fast core growth requires large planetesimal accretion rates and hence large critical core masses. Fast core growth to large critical masses poses particular challenges in the outer disk, where dynamical times are long. Core accretion at wide separations thus requires either growth of cores at rates faster than typically assumed or, perhaps counterintuitively, moderate core growth rates followed by a halt to planetesimal accretion. In this study we explore the latter possibility.

\section{Atmospheric Model Review}
\label{sec2}

%\textbf{Describe the assumptions of the model: disk, BCs, structure equations, cooling model etc., but with less detail than in paper I (and obviously refer to paper I for more details); for the BCs, emphasize that for a given core, the atmosphere profile and evolution are determined by the outer boundary conditions, i.e. Pout, Tout, Rout --- this will be relevant for section 3.2, i.e. outer boundary effects.}

We begin with a brief review of Paper I's model for the structure and evolution of a planetary atmosphere embedded in a protoplanetary disk. We summarize assumptions of the model and  properties of our assumed  disk in \S\ref{model} and list expressions for the atmosphere's structure and time evolution in \S\ref{struct}.

\subsection{Assumptions and Disk Model}
\label{model}

We assume that the planet consists of a solid core of fixed mass and a two-layer atmosphere composed of an inner convective region and an outer radiative zone that matches smoothly onto the disk. The two regions are separated by the Schwarzschild criterion for convective instability (see \S\ref{struct}) at radius $r=R\cb$, known as the radiative-convective boundary (RCB). We assume that the luminosity is constant throughout the radiative region (see Section \S\ref{critical} for additional discussion). Note that a similar method is used by \citet{pn05} and \citet{mordasini12},  who find that it agrees with more complex models.

Our model applies when planetesimal accretion is minimal, so that the envelope's evolution is dominated by KH contraction (see \S\ref{acc} for a discussion of the physical conditions required for a core to be in this regime). The atmosphere is spherically symmetric, self-gravitating and in hydrostatic balance. We apply our model at semimajor axes $a\gtrsim5$ AU. Here the disk scale height is larger than the radius at which the planet matches onto the disk (see Paper I for further details), and hence spherical symmetry holds. The nebular gas is composed of a hydrogen-helium mixture, with hydrogen and helium mass fractions of 0.7 and 0.3, respectively. Because the envelope grows slowly, we calculate its evolution using a series of linked quasi-static equilibrium models.

%The time dependence of the atmosphere structure equations may be neglected or explicitly taken into account. Some previous studies of atmosphere accretion (e.g., \citealt{stevenson82}, \citealt{wuchterl93}, \citealt{rafikov06}) consider static envelopes, in which the luminosity is solely supplied by planetesimal accretion and fully radiated away by the atmosphere. In other studies, the time evolution is explicitly taken into account and full time dependent models are developed (e.g., \citealt{ikoma00}). We follow an intermediate approach and consider quasi-static evolution. Our model for the atmosphere growth time is described in section \ref{cooling}.

The temperature and pressure at the outer boundary of the atmosphere are given by the nebular temperature and pressure. We use the minimum mass, passively irradiated disk model of  \citet{chiang10}. The surface density, mid-plane temperature and mid-plane pressure are

\begin{subeqnarray}
\label{eq:diskparam}
\Sigma\di&=&2200\, (a/\text{AU})^{-3/2}\,\, \text{g cm}^{-2} \slabel{eq:diska}\\
T\di &=& 120\, (a/\text{AU})^{-3/7} \,\,K \slabel{eq:diskb} \\
P\di&=&11\,  (a/\text{AU})^{-45/14} \,\, \text{dyn cm}^{-2} \slabel{eq:Pd} \, ,
\end{subeqnarray}
for a mean molecular weight $\mu=2.35$.

\subsection{Structure Equations and Cooling Model}
\label{struct}

The structure of a static atmosphere is described by the standard equations of hydrostatic balance and thermal equilibrium:

\begin{subeqnarray}
\label{eq:struct}
\frac{d P}{d r}&=&-\frac{G m}{r^2}\rho \slabel{eq:structa} \\
\frac{d m}{d r}&=&4 \pi r^2 \rho\slabel{eq:structb} \\
\frac{d T}{d r}&=&\nabla \frac{T}{P}\frac{d P}{d r}\slabel{eq:structc} \\
\frac{d L}{d r}&=&4 \pi r^2 \rho (\epsilon + \epsilon_g)\slabel{eq:structd}, 
\end{subeqnarray}

\noindent where $r$ is the radial coordinate, $P$, $T$ and $\rho$ are the gas pressure, temperature, and density, respectively, $m$ is the mass enclosed by radius $r$, $L$  is the luminosity from the surface of radius $r$, and $G$ is the gravitational constant. The gas is heated at a rate $\epsilon_g \equiv -T dS/dt$ per unit mass due to gravitational contraction, where $S$ is the specific gas entropy, while  $\epsilon$ represents the rate at which internal heat is generated per unit mass. We do not take into account any internal energy sources and set $\epsilon=0$. The temperature gradient $\nabla \equiv d \ln T/d \ln P$ depends on whether energy is transported throughout the atmosphere by radiation or convection. In the case of radiative diffusion for an optically thick gas, the temperature gradient is

\begin{equation}
\label{eq:delrad}
\nabla = \delrad \equiv \frac{3 \kappa P}{64 \pi G m \sigma T^4} L,
\end{equation}

\noindent where $\sigma$ is the Stefan-Boltzmann constant and $\kappa$ is the dust opacity. In our models the atmosphere is optically thick throughout the outer boundary. Where energy is transported by convection, the temperature gradient is

\begin{equation}
\label{eq:delad}
\nabla = \delad \equiv \Big(\frac{d \ln T}{d \ln P}\Big)_{\mathrm{ad}},
\end{equation}
with $\delad$ the adiabatic temperature gradient. The convective and radiative layers of the envelope are separated by the Schwarzschild criterion (e.g., \citealt{thompson06}): the atmosphere is stable against convection when $\nabla < \delad$ and convectively unstable when $\nabla > \delad$. Since convective energy transport is highly efficient,  $\nabla \approx \delad$ in convecting regions. The temperature gradient is thus given by $\nabla=\mathrm{min}(\delad, \delrad)$.

Equation set (\ref{eq:struct}) is supplemented by an equation of state (EOS) relating pressure, temperature and density, as well as an opacity law.  As summarized in Sections \ref{sec:introeos} and \ref{sec:introopacity}, this paper improves on the ideal gas polytropic EOS and interstellar medium dust opacity employed in Paper I.  We discuss the impact of our choices in Sections  \ref{deladtable}, \ref{EOSeffects}, and \ref{sec:opacity}.

%{\bf We briefly mention our choices of EOS and opacity law here.  Because understanding their import is a focus of this paper, we defer further discussion to Sections \ref{deladtable}, \ref{EOSeffects}, and \ref{sec:opacity}.}  Paper I assumes an ideal gas polytropic EOS, $P=K\rho^{\gamma}$, with $K$ the adiabatic constant and $\gamma$ the adiabatic index. In this study we use the EOS tables of \citet{saumon95}\textbf{, which we extend to lower temperatures and pressures}. Furthermore, Paper I assumes a power-law opacity given by
%
%\begin{equation}
%\label{eq:opacitylaw}
%\kappa=2 F_{\kappa} \Big(\frac{T}{T_{\rm ref}}\Big)^{\beta},
%\end{equation}  
%\noindent with $\beta$ and $F_{\kappa}$ constants, and $T_{\rm{ref}}$ a normalizing temperature. Paper I sets $\beta=2$, $F_{\kappa}=1$ and $T_{\rm ref}=100$ K, appropriate for ice grains in the interstellar medium (ISM) \citep{bell94}. However, these values are valid only for low disk temperatures, $T\di \lesssim 100 K$, where all dust components are present. Furthermore, dust settling and grain growth lower both the normalization factor $F_{\kappa}$ and the power-law coefficient $\beta$. In this paper, we explore more realistic opacity laws derived from the power-law size distributions typically used to interpret observations of protoplanetary disks (see Section \ref{sec:opacity}). 
%
%and their effect on the critical core mass when compared to the ISM power-law, as discussed in Section \S\ref{sec:opacity}.

%In addition to the \citet{bell94} opacity, we also explore more realistic opacity laws, as discussed in section \S\ref{critical}.

%We further discuss our choice of core parameters and boundary conditions.

We assume that the atmosphere forms around a solid core of fixed mass $M_{\rm c}$ with a radius $R_{\rm c}=(3 M_{\rm c}/4 \pi \rho_{\rm c})^{1/3}$, where $\rho_{\rm c}$ is the core density. We choose $\rho_{\rm c}=3.2$ g cm$^{-3}$ (e.g., \citealt{pap99}). Two radial scales determine the extent of the atmosphere: the Hill radius, $R_{\rm H} \equiv a [M_{\rm p} / (3 M_{\odot})]^{1/3}$, where the gravitational attraction of the planet and the tidal gravity due to the host star are equal, and the Bondi radius, $R_{\rm B} \equiv G M_{\rm p}/c_{\rm s}^2=G M_{\rm p} / (\mathcal{R} T\di)$, where the thermal energy of the nebular gas is approximately the gravitational energy of the planet. Here, $M_{\rm p}$ is the total planet mass, $c_{\rm s}$ is the isothermal sound speed, $\mathcal{R}=k_B/(\mu m_p)$ is the reduced gas constant, $k_B$ is the Boltzmann constant, and $m_p$ the proton mass. We define the planet mass as the mass enclosed inside the smaller of $R_{\rm B}$ or $R_{\rm H}$. For $R_{\rm B}<R_{\rm H}$, several studies assume that the atmosphere matches onto the disk at $R_{\rm B}$ (e.g., \citealt{ikoma00}, \citealt{pollack96}). In all cases, we choose the Hill radius as our outer boundary because the temperature and pressure at $R_{\rm H}$ are those of the disk,  $T(R_{\rm H})=T_{\rm d}$ and $P(R_{\rm H})=P_{\rm d}$ (see Paper I for more details). Outside $R_{\rm B}$, gas flows no longer circulate the planet, but rather belong to the disk. However, when $R_{\rm B}<R_{\rm H}$, the density structure between $R_{\rm B}$ and $R_{\rm H}$ remains spherical and is still well described by hydrostatic balance \citep{ormel13}.

%The flows outside $R_{\rm B}$ could affect the atmosphere's cooling; however, we expect this effect to be weak, as heat losses predominantly occur deep in the planetary atmosphere. This justifies our outer boundary choice.
%, which justifies our outer boundary choice.

%One reason we always choose $R_{\rm H}$ as the matching radius is that, for hydrostatic solutions, the density at $R_B$ is larger than the disk density by an order unity factor \citep{rafikov06}. Additionally, even though gas flows no longer circulate the planet outside $R_{\rm B}$, the density structure between $R_{\rm B}$ and $R_{\rm H}$ remains spherical and is still well described by hydrostatic balance \citep{ormel13}, which justifies our outer boundary choice. At the Hill radius, the temperature and pressure are given by the nebular temperature and pressure: $T(R_{\rm H})=T_{\rm d}$ and $P(R_{\rm H})=P_{\rm d}$. We note, however, that we define the planet mass as the mass enclosed inside the smaller of $R_{\rm B}$ or $R_{\rm H}$.

%Outside the Bondi sphere, the gravity of the planet is too weak to significantly affect the nebular gas, which justifies the choice of Bondi radius as the relevant radius of the atmosphere. However, the nebular gas is still perturbed outside the Bondi sphere, and therefore the Hill radius is the correct scale for matching on to the disk. This choice for the atmosphere boundary applies only when the Bondi radius is smaller than the Hill radius; when $R_{\rm B}>R_{\rm H}$, the atmosphere only extends out to the Hill radius, since material cannot be gravitationally bound to the protoplanet outside the Hill sphere.

%\subsection{Standards Methods of Solution}

%Analogously to the stellar evolution case, simply integrating the structure equations (\ref{eq:struct}) from one boundary to the other is not possible, since the boundary conditions are given both at the center and at the surface. In this case, the standard procedures for numerical integration are the shooting method or the Henyey method (\citealt{kippenhahn90}). The shooting method solves the boundary value problem by reducing it to an initial value problem: trial values are chosen for the parameters at one of the boundaries, then the equations are integrated and the resulting values at the other boundary are compared to the actual boundary conditions. The procedure is repeated until convergence is achieved. Alternatively, inward and outward integrations are carried to an intermediate fitting point, where they are fitted smoothly to each other. In the Henyey method, a trial solution for the whole interval is initially guessed, then gradually adjusted through subsequent iterations until the desired level of accuracy is achieved. In this study we use the shooting method --- we integrate inwards from the disk, and match at the core. The detailed numerical procedure is described in section \ref{twolayer}.

Finally, we employ a cooling model developed in Paper I to determine the time evolution of the atmosphere between subsequent static models. A protoplanetary atmosphere embedded in a gas disk emits a total luminosity

\begin{equation}
\label{eq:coolingglobal}
L=L_c+\Gamma-\dot{E}+e_{\mathrm{acc}}\dot{M}-P_M \frac{\partial V_M}{\partial t}.
\end{equation}
Here, $L_{\rm c}$ is the luminosity from the solid core, which may include planetesimal accretion and radioactive decay, and $\Gamma$ is the rate of internal heat generation. We set $L\co=\Gamma=0$. The $\dot{E}$ term is the rate at which total energy (internal and gravitational) is lost.  Gas accretes at a rate  $\dot{M}(r)$ with a specific energy $e(r)=u(r) - G M(r) / r$, where $u(r)$ is the internal energy per unit mass and $M(r)$ is the mass enclosed by radius $r$. To evaluate Equation \ref{eq:coolingglobal}, we choose a boundary radius $r=R$, so that  $e_{\mathrm{acc}} = e(R)$.
%, and $e_{\mathrm{acc}}$ is the specific total energy brought in by gas accreting at rate $\dot{M}$: $e_{\mathrm{acc}}=u-G M/R$, where $u$ is the internal energy per unit mass. 
Finally, the last term in Equation \ref{eq:coolingglobal} represents the work done on a surface mass element.

%As a consequence of the equations above, both the atmosphere structure and the gas accretion rate are uniquely determined by the current atmosphere mass. As this mass accretion rate is slow compared to the time it takes to relax to this solution, we can make a quasi-static model of the atmosphere growth.

We obtain an evolutionary series for the atmosphere by connecting sets of subsequent static atmospheres through the cooling Equation (\ref{eq:coolingglobal}). Details of our numerical procedure are described in Paper I. In contrast with Paper I, we evaluate Equation (\ref{eq:coolingglobal}) at $R_{\rm B}$ rather than $R\cb$. This ensures that our calculations are consistent,  because realistic opacities introduce inner radiative windows in the atmospheric structure and thus multiple RCBs (see Section \ref{sec:opacity}).

%Paper I finds that atmosphere growth continuously accelerates after self-gravity becomes important, which happens at relatively low atmosphere masses, $M_{\rm atm} \lessim 0.1 M\co$.

\section{Adiabatic Gradient for the Tabulated Equation of State}
\label{deladtable}

%\textbf{Explain the effects separately: dissociation vs. spin effects; show plots that explore these effects separately.}

%\subsection{Interpretation of Adiabatic Gradient Table}
%\label{deladinterp}

For our gas equation of state, we use the interpolated EOS tables of \citet{saumon95} for a helium mass fraction $Y=0.3$, and extend them to lower temperatures and pressures corresponding to the conditions in our fiducial disk. \App{EOStables} describes our extension procedure.
%useful parameter for understanding atmospheric structure, we choose to represent the EOS through $\delad$.

%devote this section to the impact on $\delad$ of the realistic EOS.

%More details on our extension procedure and on the methodology of combining the separate tables for hydrogen and helium are presented in \App{EOStables}.

 For ease of discussion, we choose to represent the EOS through the adiabatic gradient $\delad$ (defined in Equation \ref{eq:delad}). In contrast with an ideal gas polytrope, the EOS of a realistic gas cannot be fully specified by $\delad$. The \citet{saumon95} EOS tables are obtained using free energy minimization  (see, e.g., \citealt{graboske69}), which provides additional information about the relationship between the thermodynamic variables. Nevertheless, $\delad$ is a useful parameter for understanding atmospheric structure. For an ideal gas polytropic EOS, the adiabatic gradient is constant. Non-ideal effects such as dissociation or ionization produce temperature-dependent variations in $\delad$. Figure \ref{fig:deladmap} shows a contour plot of $\delad$  as a function of gas temperature and pressure. We distinguish three separate temperature regimes:

%, which was obtained by interpolating and extending the \citet{saumon95} EOS tables as described in \App{EOStables}.

%In this section we aim to explain how the variable adiabatic index of the hydrogen-helium mixture described by a real equation of state affects the atmosphere evolution when compared to an ideal gas with constant $\delad$. A contour plot of the adiabatic gradient as extrapolated from the \citet{saumon95} EOS tables is shown in Figure \ref{fig:deladmap}. We distinguish three separate regimes:

\begin{enumerate}
\item Intermediate temperature regime (300 K $\lesssim T \lesssim 2000$ K), where the hydrogen-helium mixture behaves like an ideal gas with a polytropic EOS.
\item High temperature regime ($T \gtrsim$ 2000 K), where dissociation of molecular hydrogen occurs, followed by ionization of atomic hydrogen for $T \gtrsim 10,000$ K.
\item Low temperature regime ($T \lesssim 300$ K), where the rotational states of the hydrogen molecule are not fully excited. % temperature is low enough for rotational motion to reduce.
\end{enumerate}

We note that helium behaves like an ideal monatomic gas with $\delad=2/5$ in our regime of interest.  Its presence in the atmosphere thus only causes a small, constant upward shift in the adiabatic gradient of the mixture.

\begin{figure}[h]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/delad_S_mixt.pdf}
\includegraphics[width=0.62\textwidth]{figures/delad_S_mixt.pdf}
%%\vspace{-0.5in}
\caption{Contour plot of the adiabatic gradient $\delad$ for a hydrogen-helium mixture in thermodynamic equilibrium as a function of gas temperature and pressure. The upper-right rectangle encloses the region described by the original \citet{saumon95} EOS tables, while the rest of the plot is our extension. The black curves represent constant entropy adiabats, with the labels $\log_{10}(S)$ [erg K$^{-1}$ g$^{-1}$]. The regions in which the EOS is either invalid or not computed are masked in white. Further details about the masked regions can be found in \App{hydrogen}.}
\label{fig:deladmap}
\end{figure}
%but not the white ones. We note, however, that the temperature and pressure ranges required by our models are fully covered by the colored regions, where our expressions are valid

%If follows that the expressions derived in \App{EOStables} cannot be used to extend the original tables in this temperature and pressure regime.

%At high temperatures, hydrogen dissociates and ionizes, while at low temperatures the rotational states of the hydrogen molecule are only partially excited and it no longer behaves like an ideal diatomic gas.

%In what follows we explain the behavior of the adiabatic gradient in the three temperature regimes separately.

%\vspace{0.2in}

%\textbf{1. Intermediate T: Ideal Gas}

\subsection{Intermediate $T$: Ideal Gas}
\label{sec:intT}

For $300$ K $\lesssim T \lesssim 2000$ K, the hydrogen molecule is not energetic enough to dissociate and hydrogen behaves as an ideal diatomic gas with constant $\delad$ (Figure \ref{fig:deladmap}). The monatomic helium component increases the adiabatic index slightly, so that $\delad \approx 0.3$ rather than 2/7 for a pure diatomic gas.\footnote{Recall that for an ideal gas, $\delad \equiv \frac{\gamma-1}{\gamma}$, where the adiabatic index  $\gamma=7/5$ for a diatomic gas and 5/3 for a monatomic gas.}

%\vspace{0.2in}

%\textbf{2. High T: Dissociation and Ionization of Hydrogen.}

\subsection{High $T$: Dissociation and Ionization of Hydrogen}

 Hydrogen is molecular at low temperatures, dissociates at $T \gtrsim 2000-3000$ K, and ionizes at  $T \gtrsim 10,000$ K. In stellar and giant planet interiors there is little overlap between the two processes: hydrogen is almost entirely dissociated into atoms by the time ionization becomes important.

As displayed in Figure \ref{fig:deladmap}, the adiabatic gradient decreases significantly in regions of partial dissociation and partial ionization.  This reduction occurs because dissociation and ionization act as energy sinks.  When internal energy is input into a partially dissociated gas, a portion is used to break down molecules, reducing the amount available to increase the temperature of the system. %The energy required to dissociate depends on the dissociation fraction, which can be determined from the Saha equation (see e.g., \citealt{kippenhahn90}). The dissociation fraction only depends on gas temperature and density, and hence only on the EOS.

 %This behavior is different than that of a mixture of molecular and atomic hydrogen, for which $2/7<\delad<2/5$, or a mixture of protons and electrons, for which $\delad=2/5$.

An expression for $\delad$ as a function of the dissociation fraction is presented in \App{deladdiss}. As expected,  $\delad$ is $2/7$ for pure molecular hydrogen and $2/5$ when hydrogen is fully dissociated, but decreases significantly during partial dissociation and is smallest when half of the gas is dissociated. Note that this behavior differs from a mixture of molecular and atomic hydrogen, for which $2/7<\delad<2/5$. %This behavior is shown in Figure \ref{fig:deladmap}. Ionization generates an analogous dip in $\delad$ at higher temperatures.

%The ionization of atomic hydrogen is also dictated by the Saha equation, with the dissociation energy replaced by ionization energy, hence the adiabatic gradient behaves analogously, consistent with Fig. \ref{fig:deladmap}.

%For a mixture of ideal gases, the total internal energy is given by the sum of the internal energies of the individual gases.

%\vspace{0.2in}

%\textbf{3. Low T: Hydrogen Rotation and Spin Isomers}

\subsection{Low $T$: Hydrogen Rotation and Spin Isomers}

The (diatomic) hydrogen molecule has five degrees of freedom, three associated with translational motion and two associated with rotation. At $T$ well above the molecule's excitation temperature for rotation, $\Theta_r \approx 85$ K \citep{kittel}, its rotational states are fully excited and its EOS is that of an ideal diatomic gas (Section \ref{sec:intT}).   As $T \rightarrow 0$, rotation  ceases entirely and $\delad \approx 2/5$, the value for an ideal monatomic gas. %We note that at temperatures larger than $\gtrsim 6000$ K, vibrational motions also become important. While at $T\lesssim2000$ K, where our extension of the \citet{saumon95} EOS tables is valid, vibrational motion is negligible, we include these effects in our extension of the EOS for completion (see \App{EOStables} for details).
At  temperatures comparable to $\Theta_r$, the rotational states of $H_2$ are  are partially excited, and its EOS depends on the relative occupation of two isomeric forms, parahydrogen and orthohydrogen, distinguished by the spin symmetry of the molecule's two protons.

Because protons are fermions, the Pauli exclusion principle implies that the total wave function of $H_2$ must be antisymmetric with respect to proton exchange.  Two components of the wave function may provide this asymmetry---proton spins and the molecule's rotational state. Parahydrogen has antiparallel proton spins and thus can only occupy symmetric rotational states with even angular quantum number $j$ \citep{farkas35}. In contrast, orthohydrogen has parallel proton spins, and can only occupy states with odd $j$. In thermal equilibrium at $T \rightarrow 0$ all hydrogen molecules are in the ground state with $j=0$, which corresponds to parahydrogen.  As temperature increases, parahydrogen starts converting into orthohydrogen.  Because the proton spin state is a triplet for orthohydrogen and a singlet for parahydrogen, this conversion plateaus at an ortho-para equilibrium ratio of 3:1 for $T \gtrsim 150$ K.

Spin conversion requires energy.  In thermal equilibrium, a portion of any internal energy input at $20 \lesssim T \lesssim 150$K is used to convert para- to orthohydrogen, reducing the amount available to increase $T$.  As a result, $\delad$ declines (Figure \ref{fig:deladmap}; see \App{deladspin} for further discussion). %At higher temperatures, the ortho-para 3:1 equilibrium ratio is reached; no further isomer conversion occurs, and thus $\delad$ remains relatively constant until dissociation temperatures are reached.

\subsubsection{Thermodynamic equilibrium of spin isomers}

We conduct the majority of our calculations using the thermal equilibrium EOS illustrated in Figure \ref{fig:deladmap}.  
%In contrast, studies such as \citet{dangelo13} assume a fixed ortho-to-para ratio of 3:1 in their evolutionary calculations.  We include results for a fixed 3:1 ratio in Section \ref{critical} for reference (see \App{EOStables} and \ref{deladspin} for discussion of the EOS).
For reference, we also include results for a fixed 3:1 ortho-para ratio (used, e.g., by \citealt{dangelo13}).\footnote{When interconversion timescales are longer than the system evolution time, populations of the two isomers evolve independently, with a fixed abundance ratio.  Even at low temperature, formation of $H_2$ on grains produces an ortho-para ratio of 3:1 since the formation energy 4.48 eV, equipartitioned into vibrational, rotational, and translational energy, yields a rotation temperature of 9200K
%, which falls in the high temperature limit 
(e.g., \citealt{takahashi01}; c.f. \citealt{fukutani13}).  In the absence of thermal equilibrium, a fixed ratio of 3:1 is likely.}

Ortho- and parahydrogen remain in thermodynamic equilibrium if the isomers can interconvert on a timescale shorter than the atmosphere's KH contraction time.  In isolation, isomeric conversion requires a forbidden transition and has a timescale longer than the age of the universe \citep[e.g.,][]{pachucki08}.  Fast conversion requires a magnetic catalyst to aid the transition between the triplet and singlet spin states.

In astrophysical contexts, this catalyst is typically provided by collisions with ions such as $H^+$ or $H_3^+$ \citep[e.g.,][]{lique12,lique14}.  
%For example, collisions between $H_2$ and $H^+$ keep interstellar clouds in equilibrium \citep[e.g.,][]{dalgarno73}.  
At high densities, collisions with $H_2$ also contribute to conversion through interactions between the spin state and the magnetic dipole of the $H_2$ molecule \citep{huestis08}.

%The low temperature EOS affects atmospheric evolution throughout the radiative envelope. It sets the scale height at the RCB, which determines the envelope's radiative diffusion time and hence the atmospheric luminosity (Equation \ref{eq:delrad} with $\nabla_{\rm rad} = \delad$).  The EOS above the RCB determines the scale length and hence the radial extent of the radiative region (see Section \ref{sec:roleofL}).  
For a core of mass $M_{\rm crit}$, the KH contraction time during the slowest phase of growth is approximately the disk lifetime, $t_{\rm disk} \sim 3\times 10^6$ years.  
The thermodynamic equilibrium EOS is thus appropriate if the ortho-para equilibration time, $t_{\rm equil}$ is smaller than $t_{\rm disk}$.
% in the outer atmosphere where $T \lesssim 150$ K.  
To determine whether this condition is met throughout the outer atmosphere (where $T \lesssim 150$ K), we check $t_{\rm equil}$ at the RCB and in the disk.
%To decide whether to use the thermodynamic equilibrium EOS, we therefore ask whether the ortho-para equilibration time $t_{\rm equil}$ is smaller than $t_{\rm disk}$ for $H_2$ number densities $n$ in a range spanning properties at the RCB and at the disk midplane, where the outer portion of the atmosphere matches smoothly.

%In our models, RCB densities span the range $n  = 10^{13}$ to $5\times 10^{14}$ all times, cores 6--30 $M_\oplus$. Roughly dependent on the critical core mass.  Lower masses give higher densities. 1.5--5$\times 10^{14}$ during longest evolution   
For core masses 6--30$M_\oplus$, our models produce RCB densities in the range $n = $1.5--5$\times 10^{14}$ during the longest phase of atmospheric evolution.  The integrated column through the atmosphere at the RCB is $\sim$$10^{25}$cm$^{-2}$, larger than the $\sim$$10^{22}$--$10^{23}$ cm$^{-2}$ required to be optically thick to X-rays \citep{glassgold97}, so that ionizations in this interior portion of the atmosphere are attenuated. 
However, densities at the RCB are high enough that 
%even in the absence of an ion to act as a catalyst, 
$H_2$ collisions cause substantial isomeric conversion.
%can speed conversion to timescales shorter than the KH contraction time. 
Extrapolating from calculations of catalysis by $O_2$,  \citet{conrath84} estimate a rate coefficient for conversion of $k_{H_2} = 8\times 10^{-29} (T/125 {\rm K})^{1/2}$ cm$^3$ s$^{-1}$, 
%$k_{H_2} = (C/n) Z$, with $Z=4\times 10^{-19}$ and $(C/n) = 2\times 10^{-10} (T/125 {\rm K})^{1/2}$ cm$^3$ s$^{-1}$, 
marginally consistent with the production of non-equilibrium ortho-para ratios in solar system giants by upwellings and flows
 \citep[e.g.][]{fouchet03}. \citet{huestis08} estimates a faster rate coefficient given by $\log_{10} k_{H_2} [$cm$^3$ s$^{-1}] = 10^{-28}[1.56+12.2\exp(-173{\rm K}/T )]$, consistent with laboratory experiments in liquid and gaseous $H_2$ \citep{farkas35, milenko97}.  
 %The temperature in the radiative region only varies from the disk temperature by an order unity factor (see Section \ref{sec:roleofL}).  
 At $T = 50$K ($T$ in the radiative region only varies from the disk temperature by an order unity factor; see Section \ref{sec:roleofL}), the equilibration time is
\begin{equation}
t_{\rm equil} = (k_{H_2} n_{H_2})^{-1} = \text{1.5--6} \times 10^6 {\rm yr} \left(\frac{n}{10^{14} {\rm cm}^{-3}}\right)^{-1} \,\; ,
\end{equation}
which is comparable to or shorter than $t_{\rm disk}$.
%which, at stellocentric distance larger than FILL IN AU, is shorter than $t_{\rm disk}$ at the RCB for cores with mass $M_{\rm crit}$ during their slowest phase of atmospheric evolution.

Where the outer atmosphere matches conditions in the disk, we turn to calculations for protoplanetary disks for guidance.  \citet{boley07}  estimate a minimum $t_{\rm equil} \sim 300$ yr.  Isomeric conversion primarily results from collisions between $H_2$ and ionic species, with $t_{\rm equil} \sim (k_{\rm ion} n_{\rm ion})^{-1} \sim 3\times 10^6$ yr $(n_{\rm ion}/10^{-4} {\rm cm}^{-3})^{-1}$, where $k_{\rm ion} \sim 10^{-10}$ cm$^3$ s$^{-1}$ \cite{walmsley04}.  Typical ion abundances in disks are small, and calculations require involved photochemical networks. 
%and are sensitive to assumptions about the properties of disk dust.  
Detailed work suggests that ion densities of $10^{-4}$ cm$^{-3}$ are achieved beyond 30 AU and, depending on the disk's dust complement, may be present throughout our region of interest \citep[e.g.,][]{glassgold97,bai09, turner10, perez11}. 
Inside $\sim$10AU, our atmospheric profiles are less sensitive to whether spin isomers reach thermodynamic equilibrium since disk temperatures exceed the peak conversion temperature of $\sim$50K.

We conclude that during the longest phase of atmospheric growth, to which $M_{\rm crit}$ is most sensitive, the equilibrium EOS is appropriate for the majority of and possibly all of our parameter space.

%use the simplified assumption that all ionizations lead to the production of $H_3^+$, which is depleted by collisions with grains.  They estimate an $H_3^+$ number density of $n_{ion} = 0.1 $cm$^{-3}(s/0.1 \mu m)(\zeta/10^{17})$s$^{-1})^{-1}$, where $s$ is the typical grain size and $\zeta$ is the ionization rate. 
%This corresponds to a reasonable ion fraction in more complicated models (FIX STATEMENT CITE). The resulting ortho-para equilibration time in the disk is $t_{\rm equil} \sim (k_{\rm ion} n_{\rm ion})^{-1} \sim 3000$ years $(s/0.1 \mu m)(\zeta/10^{17}$s$^{-1})^{-1}$, where $k_{\rm ion} \sim 10^{-10}$ \cite{walmsley04}.  Beyond 10 AU, where the disk is optically thin to X-rays, the ortho-para ratio equilibrates on timescales less than $t_{\rm disk}$.  Interior to 10 AU (FIX WITH ABOVECITE). 
 %\citet{glassgold04} has the optically thin ionization rate  $\zeta_{\rm X} = 6\times 10^{-9}$ s$^{-1}$ at 1AU at the top of the disk atmosphere, which is greater than or equal to  $10^{-17} s^{-1}$ out to 100 AU.

%take out $n_+$ estimate equation, scale to $n_+$ and say it's okay in extreme case at 30 AU (Perez Becker), throughout our range of interest (Glassgold)
%Bai, Turner refs in Perez-Becker  $10^{-4}$ ion number density gives $t_{\rm disk}$
%\citep{bai09,turner10}

%Inside $\sim$10AU, less sensitive to equil vs 3:1 b/c outer temps high enough that avoid the peak conversion temp of 50K.

%Early in the atmosphere's growth, interconversion timescales are longer than the system evolution time, and populations of the two isomers will instead evolve independently, with a fixed abundance ratio.  Even at low temperature, formation of $H_2$ on grains produces an ortho-para ratio of 3:1 since the formation energy 4.48 eV, equipartitioned into vibrational, rotational, and translational energy, yields a rotation temperature of 9200K, which falls in the high temperature limit (e.g., \citealt{takahashi01}; c.f. \citealt{fukutani13}).  In the absence of thermal equilibrium, a fixed ratio of 3:1 is likely.

For our calculations, the difference between thermal equilibrium and a fixed ortho-para ratio of 3:1 is more prominent at larger stellocentric distances, where disk temperatures are lower. We find  that a fixed ortho-para ratio may increase the atmospheric evolutionary time by up to a factor of $\sim$$3$ (corresponding to 100 AU in our fiducial disk; see \App{deladspin}, Figure \ref{fig:Lt_31}), and hence $M_{\rm crit}$ by up to a factor of 2. This effect is much smaller in the inner parts of the disk --- the atmospheric growth time only increases by $\sim$$25$\% at 10 AU.

 %At equilibrium, the relative abundance of the ortho- and para- states is given by the ratio of their partition functions, described in \App{EOStables}.
%During the para-to-ortho conversion, part of the internal energy of the hydrogen molecule is used

\section{Role of the Equation of State}
\label{EOSeffects}

Variations in the EOS, and hence $\delad$, due to partial dissociation and fractional occupation of $\rm{H}_2$ rotational states affect atmospheric evolution by yielding: (1) a lower envelope luminosity $L$, and (2) a larger amount of radiated energy per unit of accreted mass $-dE/dM$, when compared to the polytropic EOS considered in Paper I. As a result, the rate of change in atmospheric mass,

\begin{equation}
\label{eq:dMdt}
\dot{M} = -\frac{L}{dE/dM},
\end{equation}
(cf. Equation \ref{eq:coolingglobal} and ignoring surface terms, which only become significant in the late stages of atmospheric evolution) is lower than in the polytropic case. Slower gas accretion increases the growth time of the atmosphere. Since envelope growth is faster for larger cores, we calculate a larger $M_{\rm crit}$. % --- the minimum core mass required to accumulate a massive envelope during a fixed disk lifetime.%, i.e. $M_{\rm crit}$, increases when realistic EOS effects are included.

Modifications in the EOS affect $L$ and $-dE/dM$ because both of these quantities depend on the global structure of the envelope. From \Eq{eq:delrad} applied at the RCB where $\delad=\delrad$, and for a fixed atmospheric mass, the luminosity emerging from the convective interior scales as $L \propto T\cb^4/P\cb$. We have shown in Paper I that the outer radiative layer of the atmosphere is nearly isothermal, and that $T\cb$ is only a factor unity correction from the disk temperature, while the pressure in the radiative region increases exponentially with depth. It follows that $L$ primarily scales as $1/P\cb$. This pressure depth at the RCB depends on both the interior structure of the atmosphere at high temperatures, and on the matching to the nebula through the radiative region at low temperatures. Similarly, the EOS affects the distribution of total energy $E(r)$ throughout the envelope, and thus $-dE/dM$.

Since both dissociation deep in the convective interior and variable occupation of rotation states in the outer envelope affect atmospheric structure and evolution, it is helpful to separate these effects and study them independently.

%
%the amount of energy radiated per unit of accreted mass, $-dE/dM$, depends on how the total energy $E(r)$ is distributed throughout the entire envelope, and thus on the 
%
%
% is thus affected by both dissociation in the deep interior and the $\rm{H}_2$ partially excited rotational states at the top of the envelope. 
%
%In order to understand how $L$ and 
%
%It follows that both $L$ and $-dE/dM$ depend on the global structure of the atmosphere, and therefore on the EOS.  
%
% Understanding how variations in the adiabatic gradient due to a realistic EOS affect the energy behavior thus requires

%To understand the separate effects of dissociation and rotation on atmospheric growth, we consider the resulting EOS deviations from an ideal gas independently.  
%%{\bf In what follows it may be clearer to start with the motivation: isolating the two EOS effects, i.e. dissociation and rotation.  Then introduce the two patched EOSs as how we do this.  Currently it's less clear and more drawn out where this is going.}
%In order to explain how quantum rotational states at low temperatures affect atmospheric growth, we 
%% {\bf (I don't see how this is a thought experiment, or why we have to imagine.  I would just describe the EOSs)}: 
%imagine \textbf{(I used 'imagine' to make it clear that this is just an exercise to isolate the EOS effects, rather than a physically motivated assumption)} that the EOS deviates from a polytrope only in the upper atmosphere where temperatures are low, and that the gas is ideal and polytropic deep in the envelope. Conversely, we study the effects of dissociation at high temperatures by assuming that the EOS is realistic at the bottom of the atmosphere where temperatures are high, but that the gas is polytropic in the outer regions. We compare the behavior of each EOS described above with that of an ideal gas polytrope. Lastly, we use the realistic EOS at all temperatures to study the combined effects of dissociation and rotation.

%{{\bf Finish the description of the EOSs first.  Then describe and interpret Fig. 2 starting in a new paragraph.  Starting the fig description at the end of a methods discussion lost me.)}

Figure \ref{fig:tplotall} shows the time evolution of atmospheres forming at 5 AU around cores of mass $M\co=10 M_{\oplus}$ and described by various equations of state, as follows:
%{\bf I would make this list numbered and omit the reference to the curve styles, which is clear in figures.}
\begin{enumerate}
\item Ideal gas polytrope with $\delad=0.3$. % (hereafter EOS 1).
\item Ideal gas polytrope with $\delad=0.3$ for $T>500$ K and realistic EOS for $T<500$ K, which includes the effects of fractional occupation of $\rm H_2$ rotational states.
\item Ideal gas polytrope with $\delad=0.3$ for $T<500$ and realistic EOS for $T>500$ K , which includes the effects of hydrogen dissociation.
\item Realistic EOS at all $T$. 
%({\bf Here and elsewhere, I don't like ``fully" realistic, as it implies a perfect EOS.  I would just say realistic at all $T$.})
\end{enumerate}
We choose $T=500$ K as the cutoff temperature because the hydrogen-helium mixture behaves like an ideal gas, with $\delad=0.3$, in this temperature regime (see Figure \ref{fig:deladmap}). 
%{\bf (I find the references to curve numbers distracting in general.  I would try to do less of this and focus on describing the physical effects.  The reader still needs to look at the figure, and there the curves are clearly labelled.)} \sout{As such}, 
%By separately comparing curves (1) and (2), and (1) and (3)  while the difference between curves (1) and (3) accounts for hydrogen dissociation. 
Both EOS 2 and EOS 3 yield slower gas accretion when compared to EOS 1. Noting the logarithmic scale in Figure \ref{fig:tplotall}, we see that the combined effect of fractional occupation of rotational states and dissociation is significantly greater than either individually. We explore these contributions in Sections  \ref{sec:dEdM} and \ref{sec:roleofL}.

% --- due to dissociation for the high-$T$ EOS, and due to fractional occupation of $\rm H_2$ rotational states for the low-$T$ EOS.
%Both dissociation and fractional occupation of $\rm H_2$ rotational states result in slower cooling for both the high-$T$ and low-$T$ realistic EOSs. 
%decrease the atmospheric accretion rate $\dot{M}$ and thus
%This follows from Equation (\ref{eq:dMdt}), with rotation decreasing $L$ and dissociation increasing $-dE/dM$, as we show further. 
%consistent with Equation \ref{eq:dMdt}, since dissociation increases $-dE/dM$ while molecular rotation decreases $L$.

 %From Equation \ref{eq:dMdt}, the atmospheric growth time is dependent on

%The growth time is dependent on both the total energy that must be released (i.e., the rate at which energy is released) and on the luminosity of the atmosphere. {\bf `Rate of energy release' and luminosity are the same thing.  You may be calling $dE/dM$ a `rate' which is confusing.  Again, I think if you introduce and physically explain $\dot{M} = -L/(dE/dM)$ at the outset, then the resulting explanations can more compactly and clearly refer to the terms that matter.}  We explore the \sout{relative} influence of these two factors separately.  {\bf Haven't you already started exploring these factors separately?  Why say this again? Also perhaps worth a note (if not an explanation) that the combined effect is significantly greater than either individually (again noting log scale).}

\begin{figure}[h]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/tplot.pdf}
\includegraphics[width=0.5\textwidth]{figures/tplot.pdf}
%%\vspace{-0.5in}
\caption{Elapsed time to grow a planet of total mass (core + atmosphere) for a variety of EOS combinations (see text), for a planet forming at 5 AU and with a fixed core mass $M_{\rm c}=10 M_{\oplus}$. Both hydrogen dissociation at high temperatures deep in the atmosphere and fractional occupation of $\rm H_2$ rotational states at low temperatures in the outer envelope result in slower gas accretion when compared to an ideal gas polytrope.}
\label{fig:tplotall}
\end{figure}

\begin{figure*}[tb]
\centering
%\includegraphics[width=\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/all_plot_test.pdf}
\includegraphics[width=\textwidth]{figures/all_plot_test.pdf}
%%\vspace{-0.5in}
\caption{
%{\bf (Left and right panels could be different figs.  The left panel could benefit from showing $P$ or $\rho$.  Also $E_{\rm tot}$ is not yet well defined; the local quantity $e \rho r^3$ might be more instructive.  A greater range of $E$ would show what the polytrope is doing better.  The right panel could benefit from adding $\dot{M}$.  We call $E_{\rm tot} \rightarrow E$ elsewhere, so should be consistent.  Why don't x-axes line up in right panel, are masses different?  What is the discontinuity in $dE/dM$ for the polytrope around 14 $M_\oplus$?  (A kink would be course resolution, but a discontinuity is harder to understand.)  I don't think you need an abs.\ val. around $-dE/dM$.)} 
We explore atmosphere growth around a core of $M\co=10 M_{\oplus}$ forming at 5 AU in our fiducial disk, for the EOS choices in Figure \ref{fig:tplotall}. Left panels show radial profiles of pressure (upper), $-e \rho r^3$ (middle) and temperature (lower), for a total mass (core + atmosphere) of $12 M_{\oplus}$. Right panels display evolution with total mass of $L$ (upper), $-dE/dM$ (middle) and $\dot{M}$ (lower). Upper-left and upper-right: The variable occupation of $\rm H_2$ rotational states in the outer atmosphere results in a deeper RCB and a lower luminosity for the realistic EOS. Middle-left and middle-right: the (negative) total energy of the atmosphere is more concentrated at the bottom of the envelope when compared to an ideal gas polytrope due to $H_2$ dissociation. This increases the amount of energy per unit mass, $-dE/dM$, that needs to be radiated away to accrete the next parcel of gas. Lower-left: $H_2$ dissociation in the inner atmosphere decreases the temperature near the core. Lower-right: Dissociation and fractional occupation of rotational states reduce $\dot{M}$ and thus slow down atmospheric growth.}
\label{fig:all_plot}
\end{figure*}

%In Figure \ref{fig:all_plot}, middle-right and upper-right panels, we show an example evolution with planet mass of $-dE/dM$ and $L$, respectively, for the EOSs described above. The realistic EOS increases $-dE/dM$ by a factor of $\sim$$3$ and decreases $L$ by a factor of $\sim$$1.5$, when compared to the ideal gas polytrope. 
%Atmospheric growth is therefore primarily slowed by changes in the energy profile $E(r)$, with the luminosity $L$ having a secondary effect. 
%In what follows we explore each contribution separately.

Throughout Section \ref{EOSeffects}, we use a power-law opacity given by
\begin{equation}
\label{eq:opacitylaw}
\kappa=2 F_{\kappa} \Big(\frac{T}{T_{\rm ref}}\Big)^{\beta},
\end{equation}  
\noindent with $\beta =2$, $F_{\kappa} = 1$, and $T_{\rm{ref}}=100$K, appropriate for ice grains in the interstellar medium (ISM) \citep{bell94}.  We improve our treatment of opacity in Section \ref{sec:opacity}.

\subsection{Role of $-dE/dM$}\label{sec:dEdM}

Variations in $\delad$ due to dissociation increase $-dE/dM$ (Figure \ref{fig:all_plot}, middle-right). 
The atmosphere's total (negative) energy at scale $r$, $-e \rho r^3$ 
%Figure \ref{fig:all_plot}, middle-left panel, shows a radial profile of the atmosphere's total (negative) energy at scale $r$, $-e \rho r^3$, where $e$ is the total specific energy.   %the cumulative total energy (internal and gravitational) profile as a function of the radial coordinate, \sout{for the same example planet}. {\bf Two choices: (1) define this cumulative energy more clearly and give it a different symbol than $E_{\rm tot}$ that's already used, maybe $E(r)$, or (2) plot the local (noncumulative) quantity $e \rho r^3$, which should peak in the interesting places.} 
%The bulk of the energy 
is concentrated in the atmospheric interior for all EOSs (Figure \ref{fig:all_plot}, middle-left). However, $-e \rho r^3$ for EOS 3 and EOS 4 is much larger in magnitude near the core when compared to the others. This is due to the fact that $\delad$ decreases in dissociation regions (Figure \ref{fig:deladmap}). Dissociation adds particles to the gas and hence increases its entropy. In order for entropy to stay constant with radius in the convective interior, the temperature must drop, lowering $\delad$ and $|dT/dr|$ and resulting in lower temperatures near the core (cf. Figure \ref{fig:all_plot}, bottom-left). Because  the specific internal energy $u \propto T$, dissociation decreases the total energy deep in the interior.  Note that this loss of thermal energy due to dissociation can be large enough to trigger dynamical instabilities and eventual collapse in higher mass objects such as protostars \citep{larson69} or during the runaway growth of giant planets \citep{bodenheimer80}.

The reduced total energy of an atmosphere with a realistic EOS can also be explained qualitatively using $\delad$.  The density profile in an adiabatic, non-self-gravitating atmosphere composed of an ideal gas scales as $\rho(r) \propto r^{-1/\delad+1}$, and the total specific energy is $e(r) \propto 1/r$ (see Paper I). Thus $- e \rho r^3 \sim r^{-1/\delad+3}$.
 The total energy is concentrated near the core if $\delad<1/3$, and at the outer boundary otherwise. 
 %In our regime of interest, $\delad<1/3$ for all EOS choices, so the total energy  is concentrated near the core (cf. Figure \ref{fig:all_plot}, middle-left).  
Dissociation reduces $\delad$, and for smaller $\delad$, the total energy is more tightly packed in the envelope's interior.

%Dissociation reduces $\delad$, causing $-e \rho r^3$ to rise more sharply with radius. Smaller $\delad$ 
%More importantly, the dependence of $-e \rho r^3$ on $\delad$ also implies that adiabats with lower $\delad$ have their energy more tightly packed towards the interior of the envelope, which is the case during dissociation.

An atmosphere with more negative total energy (EOS 3 and 4) requires a larger amount of energy to be radiated away to bind the next batch of gas. This implies a larger $-dE/dM$ during atmospheric evolution (Figure \ref{fig:all_plot}, middle-right), and thus a lower $\dot{M}$ (Figure \ref{fig:all_plot}, lower-right).

%A more negative total energy in the interior of a planet's atmosphere, as is the case for EOS 3, requires a larger amount of energy to be radiated away to bind the next batch of gas. This implies a larger $-dE/dM$ during atmospheric evolution (middle-right panel of Figure \ref{fig:all_plot}), and thus a lower $\dot{M}$ (lower-right panel of Figure \ref{fig:all_plot}).

%We note that the amount of energy that needs to be released in order to accrete gas at the outer boundary is going to be roughly equal for all EOSs, as it has to match the outer boundary conditions of the nebula. However,

%While dissociation at high temperatures is primarily responsible for the larger $-dE/dM$ for the realistic EOS, the variable occupation of $\rm H_2$ rotational states at low temperatures also increases $-e \rho r^3$ (Figure \ref{fig:all_plot}, middle-left panel), and hence $-dE/dM$ (Figure \ref{fig:all_plot}, middle-right panel), albeit less significantly. %As both dissociation and spin isomers modify the energy profile throughout the envelope, the rate of change in atmospheric mass $\dot{M}$, and thus the atmospheric cooling time, is primarily affected by changes in $-dE/dM$ rather than $L$ (see Equation \ref{eq:dMdt}).

%However, the fact that $-dE/dM$ is affected by both dissociation and spin isomers results in it

%The high-$T$ realistic EOS has a significantly higher binding energy than the ideal gas polytrope.

%which increases the amount of energy that needs to be radiated away in order to accrete more gas, i.e. $-dE/dM$.

 %Dissociation deep in the atmosphere reduces $\delad$, and thus the temperature gradient. Figure \ref{fig:all_plot}, bottom-left panel shows the radial temperature profile (for $M\co=10 M_{\oplus}$ and $M_{\rm tot}=12 M_{\oplus}$). A lower adiabatic gradient results in lower temperatures near the core for the high-$T$ EOS when compared to the ideal gas polytrope. As the internal energy $U \propto T$, dissociation decreases $U$. The total energy $E$ thus becomes more negative

%The bulk of the energy for the high-$T$ realistic EOS and the realistic EOS at all $T$ is , which shows that hydrogen dissociation in the inner atmosphere dictates the energy behavior of the envelope. In contrast, energy is concentrated towards the outer boundary for the $\delad=0.3$ polytrope. 
%{\bf The key question here is how does the slope of the energy affect the quantity that matters, $|dE/dM|$?  This is not clear yet.  I think the shift to lower temperatures, seen in the $T(r)$ plot is helpful, but perhaps the EOS plots should also include an internal energy map.}
%Qualitatively, this can be explained through a simple analytic argument: the density profile in an adiabatic, non-self-gravitating atmosphere composed of an ideal gas scales as $\rho(r) \propto r^{-1/\delad+1}$, and the total specific energy is $e(r) \propto 1/r$ (see Paper I). Thus the total energy scales as $E(r) \propto e(r) \rho(r) r^3 \sim r^{-1/\delad+3}$. It follows that adiabats with lower $\delad$ have their energy more tightly packed towards the interior of the envelope, which is the case during dissociation when $\delad$ decreases significantly (see Section \S\ref{deladtable} and Figure \ref{fig:deladmap}). Dissociation thus causes the atmospheric interior to dominate the total binding energy, which increases the amount of energy that needs to be radiated away in order to accrete more gas, i.e. $-dE/dM$.

%The larger $-dE/dM$ due to dissociation can also be understood directly from the $\delad$ behavior. Dissociation deep in the atmosphere reduces $\delad$, and thus the temperature gradient. Figure \ref{fig:all_plot}, bottom-left panel shows the radial temperature profile (for $M\co=10 M_{\oplus}$ and $M_{\rm tot}=12 M_{\oplus}$). A lower adiabatic gradient results in lower temperatures near the core for the high-$T$ EOS when compared to the ideal gas polytrope. As the internal energy $U \propto T$, dissociation decreases $U$. The total energy $E$ thus becomes more negative, which translates into a larger $-dE/dM$ during atmospheric evolution (middle-right panel of Figure \ref{fig:all_plot}), and thus a lower $\dot{M}$ for the high-T EOS (lower-right panel of Figure \ref{fig:all_plot}). Note that, in some astrophysical contexts, this loss of thermal energy due to dissociation can be so large as to trigger dynamical instabilities and eventual collapse in higher mass objects, such as in protostars \citep{larson69} or during the runaway growth of giant planets \citep{bodenheimer80}.

% It takes more energy to bring gas deep into the atmosphere for an envelope that has the bulk of its energy concentrated towards the bottom,  {\bf (This argument isn't clear, new gas isn't brought deep into the atmosphere, it is added at the surface.)} which increases the amount of energy per unit mass that needs to be radiated, i.e. $|dE/dM|$. The cooling time is therefore larger {\bf with dissociation} \sout{for the realistic EOS at high temperatures} (3) when compared to the ideal gas polytrope (1). This increase in $|dE/dM|$ is shown in the bottom-right panel of Figure \ref{fig:all_plot}.

\begin{figure*}[tb]
\centering
%\includegraphics[width=\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/varying_delad_4panel_3.pdf}
\includegraphics[width=\textwidth]{figures/varying_delad_4panel_3.pdf}
%%\vspace{-0.5in}
\caption{As an illustrative numerical experiment, we explore atmosphere growth around a core with $M\co=10 M_{\oplus}$ forming at 5 AU, and with total mass (core + atmosphere) 12 $M_{\oplus}$, for the EOS choice of \Eq{eq:deladvar} and various $\nabla_{\rm ad,1}$ and $\nabla_{\rm ad,2}$. In each panel, the circles mark the RCB location. The horizontal dash-dotted lines show the temperature $T_{\rm switch}=500$ K at which the adiabatic gradient changes. We show radial profiles of temperature (upper panels) and pressure (lower panels). Upper-left: fixed $\nabla_{\rm ad,1}$, varying  $\nabla_{\rm ad,2}$. A larger $\nabla_{\rm ad,2}$ yields a deeper radiative region. Upper-right: varying $\nabla_{\rm ad,1}$, fixed $\nabla_{\rm ad,2}$. A lower $\nabla_{\rm ad,1}$ yields a more shallow radiative region. In both upper panels, the analytic profile of Equation (\ref{eq:Tdeep}) is over-plotted for comparison (\textit{dotted line}) in regions of agreement. Lower-left: fixed $\nabla_{\rm ad,1}$, varying  $\nabla_{\rm ad,2}$. Lower-right: varying $\nabla_{\rm ad,1}$, fixed $\nabla_{\rm ad,2}$. In both lower panels, the dashed black line is the analytic prediction for the pressure profile in the radiative zone, which agrees with the numerical results.}
\label{fig:varying_delad}
\end{figure*}

\subsection{Role of $L$}\label{sec:roleofL}

%The radiative and convective profiles must match at the RCB, constrained by the known total mass of the atmosphere

%The EOS variations in the cold outer envelope affect $\dot{M}$ by decreasing the atmosphere's luminosity compared to an ideal gas polytrope (Figure \ref{fig:all_plot}, upper-right panel, for the same parameters as Figure \ref{fig:tplotall}). From Figures \ref{fig:deladmap} and \ref{fig:all_plot}, bottom-left panel, we see that the adiabatic gradient of the low-$T$ realistic EOS increases with radius (i.e., with decreasing temperature) throughout the convective zone when $T<500$ K, then becomes lower near the RCB. This results in a larger adiabatic gradient, on average, for the low-$T$ realistic EOS in the outer parts of the convective region of the atmosphere.

%The variations in $\delad$ at low temperatures

The increase in $\delad$ at low temperatures, where $H_2$ rotational states are not fully excited (Figure \ref{fig:all_plot}, upper-right) decreases the atmosphere's luminosity.
%compared to an ideal gas polytrope 
This result may be understood as consequence of matching between the atmosphere's interior convective and exterior radiative profiles.  The profiles must match at the RCB, constrained by the known total mass.  This match determines the pressure at the RCB, $P\cb \propto \exp(\RB/R\cb)$. Higher $P\cb$ implies lower luminosity (Equation \ref{eq:delrad}).

The adiabatic gradient of EOS 2 is larger, on average, than for EOS 1 in the outer part of the convective region.\footnote {$\delad$ for EOS 2 increases with radius (i.e., with decreasing temperature) throughout the convective zone, then becomes lower near the RCB.} To understand how a variable, but overall larger $\delad$ in the outer atmosphere affects evolution, we study the simplified problem of atmospheres composed of an ideal gas with adiabatic gradient

\begin{equation}
\label{eq:deladvar}
\delad = \left\{
\begin{array}{l l}
\nabla_{\rm ad, 1}, & \quad T > 500 \text{ K} \\
\nabla_{\rm ad, 2}, & \quad T < 500 \text{ K}
\end{array} 
\right.
\end{equation}   
where $\nabla_{\rm ad, 1}$ and $\nabla_{\rm ad,2}$ are constant (Figure \ref{fig:varying_delad}).
%We explore the independent effects of varying $\nabla_{\rm ad,1}$ and $\nabla_{\rm ad,2}$ by alternately fixing one and varying the other   (Figure \ref{fig:varying_delad}). 
%The adiabatic gradient $\nabla_{\rm ad,1}$ dictates the temperature profile deep in the convective interior, while $\nabla_{\rm ad,2}$ sets the RCB temperature $T\cb$. 
%Given these inputs, the RCB location and its pressure $P\cb \propto \exp(\RB/R\cb)$ (see Figure \ref{fig:varying_delad}, lower panels) are determined by matching the interior and exterior profiles. Higher $P\cb$ implies lower luminosity (Equation \ref{eq:delrad}).

%the location of the RCB, is set by the matching between the convective interior and the outer radiative zone. As $P(r) \propto \exp(\RB/r)$ in the radiative region,

%As $P(r) \propto \exp(\RB/r)$ in the outer radiative region,

%Figure \ref{fig:varying_delad}, lower-left panel, shows the radial pressure profile for $\nabla_{\rm ad,1}=0.3$ and various $\nabla_{\rm ad,2}$, for the same parameters as Figures \ref{fig:tplotall} and \ref{fig:all_plot}, which are in agreement with the analytic expectation of an exponential profile.

%The numerical solutions reproduce the analytic expectation of a nearly exponential pressure profile in the radiative region, $P(r) \propto \exp(\RB/r)$. Thus $L \propto 1/P\cb \propto 1/\exp(\RB/R\cb)$ is directly correlated to the physical depth of the RCB.

%The location of the RCB is set by the matching between the convective interior and the outer radiative zone.

The adiabatic gradient $\nabla_{\rm ad,1}$ dictates the temperature profile deep in the convective interior.  Deep in the atmosphere, where $r \ll R\cb$, virial equilibrium yields a temperature profile 
\begin{equation}
\label{eq:Tdeep}
T(r) \approx \frac{\nabla_{\rm ad,1} G M\co}{\mathcal{R} r}.
\end{equation}   
Even though $\nabla_{\rm ad,1}$ is an order unity coefficient, the temperature scaling with $\nabla_{\rm ad,1}$ in \Eq{eq:Tdeep} is exact (Figure \ref{fig:varying_delad}, upper-right panel). Because the radial temperature profile in the atmospheric interior is set by $\nabla_{\rm ad,1}$, the radius $R_{\rm switch}$ at which the adiabatic gradient changes shows little variation with $\nabla_{\rm ad,2}$. %We can therefore match the fixed interior temperature profile to the exterior in order to determine $R\cb$ as a function of $\nabla_{\rm ad,2}$. Note, however, that \Eq{eq:Tdeep} is only valid to order of magnitude at $R_{\rm switch}$.

%n contrast with the deep interior, however, we cannot rigorously apply \Eq{eq:Tdeep} at $R_{\rm switch}$, although the approximation is still valid to order unity.

%We can use 
The RCB temperature, $T\cb$, is an order unity correction from the disk temperature that depends on the adiabatic gradient in the outer envelope, $\nabla_{\rm ad,2}$.  We have shown in Paper I that $T\cb \simeq T\di (1- 2 \nabla_{\rm ad, 2})^{-1/2}$ (for $\beta=2$ in Equation \ref{eq:opacitylaw}). This approximation agrees with the numerically calculated $T\cb$ shown in Figure \ref{fig:varying_delad} to within 5\%. Note that $T\cb$ modestly increases with $\nabla_{\rm ad,2}$ (upper-left panel), and is constant for fixed $\nabla_{\rm ad,2}$  regardless of $\nabla_{\rm ad,1}$ (upper-right panel).

% modestly increases with $\nabla_{\rm ad,2}$, consistent with the upper-left panel of Figure \ref{fig:varying_delad}. Moreover, a constant $\nabla_{\rm ad,2}$ in the exterior yields the same value for $T\cb$, independent of $\nabla_{\rm ad,1}$ (Figure \ref{fig:varying_delad}, upper-right panel), again consistent with the analytic prediction. Quantitatively, we have found that the analytic approximation for $T\cb$ agrees with the numerical result to within 5\%.
When these temperature conditions are matched for a fixed atmosphere mass, Figure \ref{fig:varying_delad} shows that a larger adiabatic gradient in the outer atmosphere results in a deeper radiative region. We can understand this effect by assuming that $R_{\rm switch}$ is fixed. A steeper adiabat (i.e., with a larger adiabatic gradient) that starts at a given depth (here $R_{\rm switch}$) will match the fixed RCB temperature at a smaller radius. A quantitative scaling of $R\cb$ is not possible because Equation (\ref{eq:Tdeep})  is only valid to order of magnitude at $R_{\rm switch}$ and relatively small variations in the physical depth of the RCB can impact the RCB pressure significantly. It follows that $L \propto 1/P\cb$ decreases with increasing outer $\nabla_{\rm ad}$; thus the overall larger adiabatic gradient due to the variable occupation of rotation states at low temperatures decreases $L$ and slows down growth.

 %which dictates the RCB location, is only an order unity correction to the (also order unity) adiabatic gradient in the outer atmosphere. 
%ignoring the order unity variations in    discussed above due to changes in the adiabatic gradient. For

%coefficients in $R_{\rm switch}$ and $T\cb$ discussed above that arise due to $\delad$ variations. For a fixed $R_{\rm switch}$, a steeper adiabat (i.e., larger $\nabla_{\rm ad,2}$) will match the (assumed fixed) $T\cb$ at a smaller $R\cb$. A quantitative scaling of $R\cb$ with $\nabla_{\rm ad,2}$ is not possible, as all the relevant quantities ($R_{\rm switch}$, $T\cb$, and $R\cb$) are only order unity corrections to the (also order unity) variable $\nabla_{\rm ad,2}$. However, relatively small changes in $R\cb$ can impact the RCB pressure significantly since $P\cb$ is an exponential function of $R\cb$, as explained above. As $L \propto 1/P\cb$, the atmospheric luminosity decreases with increasing $\nabla_{\rm ad,2}$; thus the overall larger adiabatic gradient due to the variable occupation of rotation states at low temperatures decreases $L$ and slows down growth.

%From Equation (\ref{eq:dMdt}), the lower $L$ and larger $-dE/dM$ caused by non-ideal EOS effects reduce the atmospheric accretion rate $\dot{M}$, resulting in slower growth and thus larger $M_{\rm crit}$ (see \S\ref{critical}).

%realistic EOS effects reduce the atmosphere accretion rate $\dot{M}$ This results in slower atmospheric growth and therefore a larger critical core mass, as we show in Section \S\ref{critical}.

%Figure \ref{fig:varying_delad}, top panel, shows the resulting radial temperature profile, for various values of $\nabla_{\rm ad, 2}$. The core and atmosphere masses are the same as in Figure \ref{fig:all_plot}. The temperature profiles overlap in the inner atmosphere, where all polytropes have the same adiabatic gradient $\nabla_{\rm ad,1}=0.3$. A larger $\nabla_{\rm ad,2}$ in the outer envelope results in a larger RCB temperature and a deeper radiative region. The radius $R_{\rm sw}$ at which $T(R_{\rm sw})=T_{\rm sw}=500$ K modestly increases with $\nabla_{\rm ad,2}$.

%\begin{figure}[h]
%\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/varying_delad.pdf}
%%%\vspace{-0.5in}
%\caption{Top panel: Radial temperature profiles for polytropes with the same adiabatic gradient in the interior and varying $\delad$ at low temperatures. The circles mark the RCB location. The dashed black line is the temperature $T_{\rm sw}=500$ K at which $\delad$ changes. A larger $\delad$ in the outer envelope increases the RCB temperature and the depth of the radiative region of the atmosphere. Bottom panel: Radial pressure profiles for polytropes with the same adiabatic gradient in the interior and varying $\delad$ at low temperatures. The circles mark the RCB location. The pressure profiles in the radiative region are nearly exponential and overlap for the different polytropes.}
%\label{fig:varying_delad}
%\end{figure}

% We can explain these effects analytically. By substituting Equation (\ref{eq:structa}) into Equation (\ref{eq:structc}) in the convective region of the atmosphere, and assuming the nebular gas obeys the ideal gas law, we can write hydrostatic balance along an adiabat as 
%
%\begin{equation}
%\label{eq:dTdr}
%\frac{d T}{d r}=-\frac{\delad G m(r)}{\mathcal{R} r^2},
%\end{equation}
%where $\mathcal{R}$ is the specific gas constant. For a non-self-gravitating atmosphere with an EOS described by Equation (\ref{eq:deladvar}), an analytic expression for the temperature profile can be obtained by integrating Equation (\ref{eq:dTdr}) piece-wise from the core to $R_{\rm sw}$ and from $R_{\rm sw}$ to the RCB, respectively. We have showed in Paper I that the temperature dependence deep in a non-self-gravitating atmosphere (where $r \ll R\cb$) with constant $\delad$ throughout can be expressed as
%
%
%Figure \ref{fig:varying_delad} shows that the $R_{\rm sw} \ll R\cb$ approximation, and thus the
%\begin{equation}
%\label{eq:Tswapprox}
%T_{\rm sw} \approx \frac{\nabla_{\rm ad,1} G M\co}{\mathcal{R} R_{\rm sw}}
%\end{equation}  
%approximation, become less accurate as $\nabla_{\rm ad, 2}$ increases. However, the variations in $R_{\rm sw}$ are small enough that we can assume that $R_{\rm sw}$ is in principle the same for all polytropes. Integrating Equation (\ref{eq:dTdr}) from $R_{\rm sw}$ to $R\cb$ thus gives
%
%\begin{equation}
%\label{eq:rcbteor}
%\frac{1}{R\cb} = \frac{1}{R_{\rm sw}} - \frac{\mathcal{R}}{\nabla_{\rm ad,2} G M\co} (T_{\rm sw}-T\cb).
%\end{equation}
%An approximate expression for $T\cb$ as a function of the adiabatic gradient $\nabla_{\rm ad,2}$ and disk temperature can be derived using Equation (\ref{eq:delrad}) applied at the RCB (where $\delad=\delrad$) and $\delrad=d \ln T/ d \ln P$ (see Paper I for details). For $\beta=2$ in the opacity law (\ref{eq:opacitylaw}), $T\cb \simeq T\di (1- 2 \nabla_{\rm ad, 2})^{-1/2}$, which increases with $\nabla_{\rm ad,2}$ as shown in Figure \ref{fig:varying_delad}. Both $T\cb$ and $\nabla_{\rm ad, 2}$ thus reduce the second term on the right-hand side of Equation (\ref{eq:rcbteor}) when $\nabla_{\rm ad,2}$ is larger, and hence $R\cb$. The slightly larger $R_{\rm sw}$ cannot make up for this effect, and therefore a larger adiabatic gradient in the outer atmosphere increases the thickness of the radiative region, consistent with Figures \ref{fig:varying_delad}, top panel, and \ref{fig:all_plot}, middle-left panel. It follows that the RCB location is set by the interior structure of the atmosphere, rather than the EOS in the outer envelope. This is also shown Figure \ref{fig:varying_delad_in}, where polytropes with the same $\delad$ at the top of the atmosphere have different RCB locations as the interior adiabatic gradient varies.  
%
%
%
%An overall larger $\delad$ in the outer atmosphere due to molecular rotation thus increases the physical extent of the radiative zone. As the radiative region is nearly isothermal, the density and pressure increase exponentially with depth, as shown in Figure \ref{fig:varying_delad}, bottom panel.  The lower $R\cb$ for the low-$T$ realistic EOS thus results in a larger pressure depth of the RCB, as shown in the upper-left panel of Figure \ref{fig:all_plot}. As $L \propto 1/P\cb$ (cf. Equation \ref{eq:delrad}), the deeper RCB results in a lower $L$ and thus $\dot{M}$ (Figure \ref{fig:all_plot}, lower-right panel). \footnote{The $\delad$ and $T\cb$ terms in Equation (\ref{eq:delrad}) applied at the RCB only modestly affect $L$ as $\delad$ varies.}

%Qualitatively, this can be explained through a simple analytic argument: for a non-self-gravitating atmosphere composed of an ideal gas, the temperature profile in convective regions scales as $T(r) \propto \delad/r$ (see Paper I for details; it can also be derived from Equations \ref{eq:structa} and \ref{eq:structc} applied in convective regions). Applying the expression above at the RCB yields $T\cb \propto \delad\cb/ R\cb$. As the outer radiative layer of the atmosphere is nearly isothermal, $T\cb \approx T\di=const.$, to order unity corrections. It easily follows that $R\cb \propto \deladcb$   Figure \ref{fig:all_plot_temp}, middle-left panel shows 
%This is directly correlated to

%directly correlated to the depth of the radiative region of the envelope, which is shown in the upper-left panel of Figure \ref{fig:all_plot} for a total planet mass of $12 M_{\oplus}$. We see that the realistic EOS for low temperatures (2) results in a deeper location of the RCB when compared to the ideal gas polytrope (1). {\bf This is an accurate observation.  
%The deeper question is why does the EOS change result in a deeper RCB?  This should be possible to understand, right?}  This deeper RCB translates into a larger number of steps that photons need to diffuse to escape from the RCB, and therefore a lower luminosity (cf. Equation \ref{eq:delrad}).  {\bf Better to just say it's an optical depth effect than reexplaining how radiative diffusion works.  Also it's the pressure depth (not the radius) that correlates more closely with optical depth.  A plot of $P(r)$ might thus help as noted in fig caption.}

%We see that the lower luminosity of the fully realistic EOS (4) when compared to the polytrope (1) is due to the realistic EOS in the outer atmosphere (2). T

%We found that the RCB for the fully realistic EOS (4) is roughly at the same location as the RCB for the realistic EOS at low temperatures (2) for low atmosphere masses, and it moves further out relatively as the atmosphere mass increases. This explains our choice of $M_{\rm tot}=12 M_{\oplus}$ as a representative case to illustrate the effect of spin isomers. Envelope growth is slowest at this relative atmosphere-to-core mass (see PY13)...

%Dissociation deep in the planet's atmosphere increases the amount of energy per unit mass that needs to be radiated by the envelope, which further slows down growth.

%{\bf Why is this true?  Dissociation absorbs energy which by itself decreases the amount of energy that needs to be radiated.}

%hence  The $\delad$ variations given by Equation (\ref{eq:deladvar})

%This result is valid for the $\delad=0.3$ polytrope. 
%
%As the low$-T$ realistic EOS only slightly (see Figure \ref{fig:deladmap}) deviates from an ideal gas in the outermost convective layers, we can assume that the gas described by the low-$T$ EOS can be approximated as ideal throughout the convective interior. The low-$T$ EOS thus also obeys Equation (\ref{eq:dTdr}) with $\delad \approx const.$. We further neglect self-gravity in Equation (\ref{eq:dTdr}) and set $m(r)=M\co$. Integrating Equation (\ref{eq:dTdr}) from the core to the RCB under these assumptions yields
%\begin{equation}
%\label{eq:dTdrint}
%T\cb=T\co - \frac{\delad G M\co}{\mathcal{R}} \Big(\frac{1}{R\co} - \frac{1}{R\cb}\Big).
%\end{equation}  
%Deep in the convective interior, $T \gg T\cb$ and $r \ll R\cb$. From virial equilibrium, the temperature in the inner atmosphere can thus be expressed as
%\begin{equation}
%\label{eq:virialT}
%T(r) \simeq \frac{\nabla_{\rm ad} G M\co}{\mathcal{R} r}.
%\end{equation}
%%The temperature scale height deep in the convective interior is $H_T \sim r$, with $dT/dr \sim -T/H_T \sim -T/r$. From Equation (\ref{eq:structc}) it immediately follows that the pressure scale height, for which $dP/dr \sim -P/H_P$, is $H_P \sim r \delad$. Deep in the interior and for $m(r)=M\co$, hydrostatic balance (\ref{eq:structa}) thus becomes
%
%%\begin{equation}
%%\label{eq:structint}
%%\frac{P}{r \delad} \sim -\frac{G M\co}{r^2}. 
%%\end{equation}
%%From this and the definition of the isothermal sound speed
%%\begin{equation}
%%\label{eq:cs}
%%c_s^2 = \frac{P}{\rho} =\mathcal{R} T, 
%%\end{equation} 
%%with both expressions applied at the core where $r=R\co$, $T=T\co$ and $\delad=\nabla_{\rm ad, c}$, we find
%Applying equation (\ref{eq:virialT}) at the core, where $T=T\co$, $r=R\co$ and $\delad=\nabla_{\rm ad, c}$, gives
%\begin{equation}
%\label{eq:Tc}
%T\co \simeq \frac{\nabla_{\rm ad, c} G M\co}{\mathcal{R} R\co}.
%\end{equation}
%As the adiabatic gradient at the core is the same for both the ideal gas polytrope and the low$-T$ realistic EOS, $T\co$ has the same value for both EOSs. This is shown in Figure \ref{fig:all_plot}, lower-left panel. Substituting Equation (\ref{eq:Tc}) into Equation (\ref{eq:dTdrint}) applied at the RCB gives
%\begin{equation}
%\label{eq:Tcb}
%T\cb \simeq \frac{\nabla_{\rm ad, RCB} G M\co}{\mathcal{R} R\cb},
%\end{equation}
%%We note that the expression above and the definition of the pressure scale height $H_P \equiv c_s^2 r^2/(G M\co)$ imply that the pressure scale height at the RCB is the same as in the deep interior, $H_P \sim \delad r$.
%and thus
%\begin{equation}
%\label{eq:rcb}
%R\cb \propto \frac{\nabla_{\rm ad, RCB}}{T\cb}.
%\end{equation}
%An approximate expression for $T\cb$ as a function of the adiabatic gradient and disk temperature can be derived using Equation (\ref{eq:delrad}) applied at the RCB (where $\delad=\delrad$) and $\delrad=d \ln T/ d \ln P$ (see Paper I for details). For $\beta=2$ in the opacity law (\ref{eq:opacitylaw}), $T\cb \simeq T\di (1- 2 \nabla_{\rm ad, RCB})^{-1/2}$, and thus
%\begin{equation}
%\label{eq:rcbdel}
%R\cb \propto \frac{\nabla_{\rm ad, RCB}}{(1- 2 \nabla_{\rm ad, RCB})^{-1/2}}.
%\end{equation}
%For the ideal gas polytrope $\nabla_{\rm ad, RCB}=0.3$, and for the low-$T$ realistic EOS $\nabla_{\rm ad, RCB}\approx 0.25$. For these particular choices, Equation (\ref{eq:rcbdel}) yields a smaller $R\cb$ for the low$-T$ realistic EOS.

%Molecular rotational effects at the top of the atmosphere increase the luminosity $L$, 
%
%We have seen that quantum rotational states transitions due to the low temperatures at the top of the atmosphere {\bf (already too long and convoluted)} increase the thickness of the radiative zone, and decrease the luminosity of the envelope, while dissociation at high temperatures deep in the atmosphere increases the amount of energy needed to increase the atmosphere mass by a fixed amount. {\bf Way too long a sentence!} Overall, both effects result in a longer time for the atmosphere to evolve.

%we find that only polytropes with $\gamma<3/2$, i.e. $\delad<1/3$, have the total energy concentrated towards the core. This is satisfied by $\delad=2/7$ but not by $\delad=2/5$, in agreement with the results in the bottom panel of Figure \ref{fig:ETrplotpoly}. 
%
%Similarly to section \ref{deladpoly}, we use instantaneous atmosphere profiles to explain the differences. The left panel of Figure \ref{fig:TLrplot} shows the instantaneous temperature profile and the location of the radiative-convective boundary for a total fixed mass (core + atmosphere) $M_{\rm{tot}}=11.8 M_{\oplus}$. The realistic equation of state for low temperatures is characterized by a lower adiabatic index in the outer regions, due to the spin effects, and is therefore dominant in the radiative zone. As a result, it generates a deeper radiative zone with a lower luminosity, which explains the results in the left panel of Figure \ref{fig:TLrplot}. Moreover, since the cooling time is inversely proportional to the luminosity, the spin effect will result in a longer cooling time.
%
%The energy behavior is shown in Figure \ref{fig:Erplot}. The realistic EOS for high temperatures has a low adiabatic index deep in the atmosphere, due to hydrogen dissociation, and thus the bulk of its energy concentrated at the bottom of the atmosphere, for the reasons described in section \ref{deladpoly}. It takes more energy to add mass more mass deep in the atmosphere, and $|dE/dM|$ is larger as a result. 
%%
%
%We have seen that the spin effect at the outer boundary dictates the location of the radiative zone, and therefore the luminosity behavior, while dissociation deep in the atmosphere dictates the energy behavior. Overall, both effects result in a longer time for the atmosphere to evolve. 
%
%\subsection{Ideal Gas Polytropes with Different Adiabatic Gradient}
%\label{deladpoly}
%
%In this section we investigate the differences in luminosity and $dE/dM$, and the resulting time evolution, between ideal gas polytropes with different adiabatic gradients: $\delad=2/7$ (diatomic gas) and $\delad=2/5$ (monatomic gas). We assume both gases have the same mean molecular weight. %We use these results to explain the separate effects of dissociation and spin on the time evolution of the atmosphere.
%
%We generate atmosphere profiles for the two different adiabatic indices at $a=10$ AU and for a core mass $M_{\rm c}=10 M_{\oplus}$, and estimate the luminosity and cooling time evolution as described in section \ref{sec2}. The results are shown in Figure \ref{fig:Ltplotpoly}. We find that the polytrope with the lower adiabatic gradient has both a higher luminosity and a longer cooling time. We use instantaneous atmosphere profiles to explain these effects. 
%
%%We first discuss the effect of the variable adiabatic gradient on luminosity. 
%
%Figure \ref{fig:ETrplotpoly}, top panel, shows the temperature profile for the two polytropes at a fixed total mass $M_{\rm{tot}}=15 M_{\oplus}$. The polytrope with a larger adiabatic gradient, $\delad=2/5$, has a more shallow convective zone, and hence a deeper radiative region, since a larger temperature gradient delays the onset of convection. A deeper radiative layer increases the number of steps that photons need to diffuse, resulting in a lower luminosity as seen in the top panel of Figure \ref{fig:Ltplotpoly}. 
%
%\begin{figure}[h]
%\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/Ltplot_poly.pdf}
%%%\vspace{-0.5in}
%\caption{Luminosity and time evolution as a function of total mass (core + atmosphere) for polytropes with different adiabatic indices, for a planet forming at 10 AU and with a fixed core mass $M_{\rm c}=10 M_{\oplus}$. A larger adiabatic index results in both a lower luminosity and a shorter cooling time.}
%\label{fig:Ltplotpoly}
%\end{figure}
%
%\begin{figure}[h]
%\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/TErplot_poly.pdf}
%%%\vspace{-0.5in}
%\caption{Instantaneous temperature and total energy profiles as a function of radius for polytropes with different adiabatic indices, for a planet forming at 10 AU and with a fixed core mass $M_{\rm c}=10 M_{\oplus}$. The total mass (core + atmosphere) is $15 M_{\oplus}$. The location of the radiative-convective boundary is marked. A lower adiabatic gradient results in a more shallow radiative region (upper panel), and in the total energy being concentrated at the bottom of the atmosphere (lower panel).}
%\label{fig:ETrplotpoly}
%\end{figure}
%
%%In spite of the low luminosity of the atmosphere with $\delad=2/5$, the envelope still grows faster than in the case of a diatomic gas, as shown in the bottom panel of Figure \ref{fig:Ltplotpoly}. This is due to fact that the amount of energy per unit mass that needs to be radiated away, i.e. $|dE/dM|$, is lower for the $\delad=2/5$ polytrope. We have shown in PY13 that polytropes with $\delad=2/7$ have most of the energy concentrated at the bottom of the atmosphere, while polytropes with $\delad=2/5$ have the bulk of the energy towards the outer boundary. 
%
%
%The bottom panel of Figure \ref{fig:ETrplotpoly} shows an instantaneous energy profile for the same total mass $M_{\rm{tot}} = 15 M_{\oplus}$.  The bulk of the energy is concentrated deep in the atmosphere for the $\delad=2/7$ polytrope and towards the outer boundary for the $\delad=2/5$ polytrope. Qualitatively, this can be explained through a simple analytic argument: the density profile in an adiabatic, non-self gravitating atmosphere composed of an ideal gas scales as $\rho(r) \sim r^{1/(\gamma-1)}$ with $\gamma$ the adiabatic index (see PY13). Since the energy per unit mass scales as $e(r) \sim \rho(r) r^2$, we find that only polytropes with $\gamma<3/2$, i.e. $\delad<1/3$, have the total energy concentrated towards the core. This is satisfied by $\delad=2/7$ but not by $\delad=2/5$, in agreement with the results in the bottom panel of Figure \ref{fig:ETrplotpoly}.  It takes more energy to bring in gas deep in the atmosphere for an envelope that has the bulk of its energy concentrated towards the bottom,  which increases $|dE/dM|$ for the $\delad=2/7$ polytrope. 
%
%We find that the energy effect prevails over the luminosity effect, resulting in a longer cooling time for the polytrope with the lower adiabatic gradient (i.e., $\delad=2/7$), as shown in the bottom panel of Figure \ref{fig:Ltplotpoly}.

%From the adiabatic gradient table shown in Figure \ref{fig:deladmap} we find that $\delad$ has the flattest behavior around $T=500$ K. In this region, the gas behaves like an ideal gas with constant polytropic index $\delad \approx 0.3$, with the shift from diatomic gas caused by the helium in the mixture. We generate three sets of atmosphere profiles. The first one corresponds to an ideal gas of constant adiabatic index $\delad=0.3$. The second one is described by the real EOS for temperatures larger than 500 K and by an ideal gas with $\delad=0.3$ for $T<500$ K. Finally, the third profile consists of an ideal gas polytrope with $\delad=0.3$ for $T>500$ K and a real gas in the low temperature regime. We compare the first two profiles to show the dissociation effects, and the first and third profile to show the effects of ortho- and parahydrogen. The resulting time evolution is shown in Figure \ref{fig:Ltplotall}. We see that both dissociation and spin isomers have a comparable effect on the atmosphere growth, and result in slower cooling, and therefore a longer crossover time, when compared to the polytropic ideal gas equation of state. In what follows we explore the two effects separately.

%As compared to the polytrope, the real EOS therefore generates a deeper radiative zone with a lower luminosity, due to the lower adiabatic index in the outer regions, as well as an atmosphere with the bulk of its energy concentrated at the bottom, due to the low $\delad$ caused by dissociation. As shown above for the ideal gas polytropes, and remembering that $\Delta t \sim -\Delta E/L$, we see that both effects result in a longer time for the atmosphere to evolve.

% for an atmosphere described by an ideal gas EOS with $\delad=0.3$ (dashed curve), an atmosphere composed of an ideal gas with $\delad=0.3$ for $T>500$ K and a realistic gas

%In this section we use the results obtained in section \ref{deladpoly} to explain the effects on the atmosphere evolution of a realistic equation of state. We explore the effect of hydrogen dissociation at the high temperatures in the inner part of the atmosphere, and the effect of hydrogen spin isomers at the low temperatures at the top of the atmosphere separately. In order to do this, we generate atmosphere profiles in which the nebular gas is assumed to be described by a combination of ideal and realistic equations of state, depending on the temperature. To explain the effects of the hydrogen spin isomers in the outer parts of the atmosphere, we assume that the realistic EOS effects only matter at low temperatures; conversely, we assume that the realistic EOS effects are important only at high temperatures in order to study the effects of hydrogen dissociation deep in the atmosphere. As such, we generate three sets of atmosphere profiles, choosing $a=10$ AU and $M\co=5 M_{\oplus}$. The first one is described by a realistic EOS for temperatures larger than 500 K and by an ideal gas with $\delad=0.3$ for $T<500$ K. The second profile consists of an ideal gas polytrope with $\delad=0.3$ for $T>500$ K and a realistic gas in the low temperature regime. Finally, the last profile corresponds to an ideal gas with $\delad=0.3$, the adiabatic gradient of an ideal hydrogen-helium mixture.  We choose $T=500$ K as the reference temperature since the adiabatic gradient of the realistic gas is roughly constant around this temperature (see Figure \ref{fig:deladmap}). The differences in atmosphere structure and evolution between the ideal gas and the realistic gas at low (high) temperatures highlight the effects of hydrogen spin isomers (hydrogen dissociation). Figure \ref{fig:tplotall} illustrates the differences in time evolution between these profiles. We see that both dissociation and spin isomers have a comparable effect on the atmosphere growth, and result in slower cooling, and therefore a longer crossover time, when compared to the polytropic ideal gas equation of state. The cooling time is dependent on both the total energy released due to the contraction of the envelope and the luminosity of the atmosphere. In what follows we explore the relative influence of these two factors separately.

%The resulting time evolution is shown in Figure \ref{fig:tplotall}. We see that both dissociation and spin isomers have a comparable effect on the atmosphere growth, and result in slower cooling, and therefore a longer crossover time, when compared to the polytropic ideal gas equation of state. The cooling time is dependent on both the total energy released due to the contraction of the envelope and the luminosity of the atmosphere. In what follows we explore the relative influence of these two factors separately.

%We now discuss the differences in luminosity and $dE/dM$ between an ideal gas with constant adiabatic gradient and atmospheres with variations in $\delad$ as prescribed by the equation of state discussed in section \ref{EOSeffects}. In what follows we describe our choices of equation of state combinations. From the adiabatic gradient table shown in Figure \ref{fig:deladmap} we find that $\delad$ has the flattest behavior around $T=500$ K. In this region, the gas behaves like an ideal gas with constant polytropic index $\delad \approx 0.3$, with the shift from diatomic gas caused by the helium in the mixture. We generate three sets of atmosphere profiles. The first one corresponds to an ideal gas of constant adiabatic index $\delad=0.3$. The second one is described by the realistic EOS for temperatures larger than 500 K and by an ideal gas with $\delad=0.3$ for $T<500$ K. Finally, the third profile consists of an ideal gas polytrope with $\delad=0.3$ for $T>500$ K and a realistic gas in the low temperature regime. We compare the first two profiles to show the dissociation effects, and the first and third profile to show the effects of ortho- and parahydrogen. The resulting time evolution is shown in Figure \ref{fig:tplotall}. We see that both dissociation and spin isomers have a comparable effect on the atmosphere growth, and result in slower cooling, and therefore a longer crossover time, when compared to the polytropic ideal gas equation of state. The cooling time is dependent on both the total energy released due to the contraction of the envelope and the luminosity of the atmosphere. In what follows we explore the relative influence of these two factors separately.

%We now investigate the effect of the low adiabatic index caused by hydrogen dissociation, on the one hand, and spin isomers on the other hand, on the luminosity and cooling time evolution of the atmosphere, in light of the discussion in section \ref{deladpoly}. The left panel of Figure (x) shows the luminosity evolution with mass for the four combinations of equations of state described in section

%
%
%\begin{figure}[h]
%\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/Er_plot.pdf}
%%%\vspace{-0.5in}
%\caption{Instantaneous energy profiles as a function on radius for a variety of EOS combinations, for a planet forming at 10 AU and with a fixed core mass $M_{\rm c}=10 M_{\oplus}$. The total mass (core + atmosphere) is $11.8 M_{\oplus}$. Hydrogen dissociation deep in the atmospheres causes the bulk of the energy to be concentrated at the bottom of the atmosphere. This increases the amount of energy per unit mass that needs to be radiated away, i.e. $|dE/dM|$, resulting in a longer crossover time for the realistic EOS when compared to the polytrope.}
%\label{fig:Erplot}
%\end{figure}

\section{Impact of Opacity on Atmosphere Evolution}
\label{sec:opacity}

Our calculations so far have assumed that the dust opacity in the radiative region of the atmosphere is given by the standard ISM opacity. However, our scenario of low planetesimal accretion is likely to favor lower dust opacities, due to grain growth and dust settling. Grain growth, in particular, lowers the absolute value of the opacity, and may change the particle size distribution when compared to the standard ISM size distribution (e.g., \citealt{pollack85}). Enhanced metallicity due to planetesimal accretion, by construction not present during our atmosphere's growth, cannot make up for this reduction.

%Observations of dust in protoplanetary disks (e.g., \citealt{beckwith90}, \citealt{beckwith91}, \citealt{perez12}) have shown evidence for grain growth and that the particle size distribution is not interstellar.

Although grain growth and evidence for a non-ISM size distribution have been observed in protoplanetary disks (e.g., \citealt{beckwith90}, \citealt{beckwith91}, \citealt{perez12}), the size distribution of dust particles has not been tightly constrained. Typically, the differential grain size distribution is assumed to be a power-law:

\begin{equation}
\label{eq:graindistr}
\frac{dN}{ds} \propto s^{-p},
\end{equation}
where $dN$ is the number of particles with sizes between $s$ and $s + ds$ and $p=3.5$ (a standard \citealt{dohnanyi69} collisional cascade; appropriate for the ISM) or $p=2.5$ (an approximation for coagulation). In this work we use the  \citet{dalessio01} frequency-dependent opacity tables to obtain the temperature-dependent Rosseland mean opacity $\kappa$. We take as a fiducial case a maximum particle size $s_{\rm max}=1$ cm and a grain size distribution given by Equation (\ref{eq:graindistr}) with $p=3.5$. Other choices for the power-law coefficient $p$ are discussed later in this section. 
Though we choose our opacities to reflect ambient conditions in the disk, we note that 
%even if the opacity were close to that of the ISM in the disk, 
grains can also grow within an accreting atmosphere, decreasing its opacity at depth \citep{movshovitz10,mordasini14b,ormel14}.

The \citet{dalessio01} opacities are only relevant at temperatures that are sufficiently low for dust grains to remain solid ($T \lesssim 1000$ K). At higher temperatures, we use the \citet{bell94} analytic opacity laws, ensuring smooth transition from the grain growth opacities, as illustrated in Figure \ref{fig:opacity}.

The sharp drop in opacity ($\kappa \sim T^{-24}$, see Figure \ref{fig:delvsr}) due to dust sublimation lowers the radiative temperature gradient significantly (see Equation \ref{eq:delrad}), and may thus generate radiative layers within the inner region of the atmosphere (see \App{radwindow}, Figure \ref{fig:delvsr}). Additionally, the weak temperature dependence of grain growth opacities may cause larger outer radiative zones than when ISM opacities are used. This could pose two challenges for our model:

\begin{enumerate}

 \item The additional luminosity, $\Delta L$, generated in the outer radiative layer of the envelope may not satisfy $\Delta L \ll L$, where $L$ is the assumed fixed atmospheric luminosity. For $p=3.5$ in Equation (\ref{eq:graindistr}), we have checked that $\Delta L \ll L$ in all the cases presented in this study. For $p=2.5$, however, this approximation breaks down at low core masses, as we show in \S\ref{critical}.

 %In practice, however, the inner radiative windows are either very narrow compared to the height of the convective regions (Figure \ref{fig:delvsr}, middle panel), or  have $\delad \approx \delrad$ throughout, which makes the distinction between convective and radiative layers less  pronounced (Figure \ref{fig:delvsr}, bottom panel).

%This results in a negligible extra luminosity generated in the radiative windows. However, the radiative windows may become non-negligible for other opacity choices and sufficiently low core masses, as we show later in this section.

\item As little as half of the atmosphere's luminosity is generated in the innermost convective layer when radiative windows exist.   We must therefore check that our assumption of constant $L$ does not substantially change the structure of the atmosphere in the region of the radiative windows. Fortunately, these radiative windows are either very narrow compared to the height of the convective regions (Figure \ref{fig:delvsr}, middle panel), or  have $\delad \approx \delrad$ throughout, which makes the distinction between convective and radiative layers less  pronounced (Figure \ref{fig:delvsr}, bottom panel). This implies that the entropy drop across the radiative windows is small. We investigate the luminosity structure in greater detail in \App{radwindow} and conclude that our model remains a reasonably good approximation even in the presence of radiative windows.
%\textit{We have verified using an extreme luminosity profile that our results reasonably approximate the atmosphere's structure interior to the outermost radiative layer (see \App{radwindow} for additional details)}.

%By design, the luminosity in the radiative windows, $L_{\rm radw}$, must satisfy $L_{\rm radw}=L$, where $L$ is the assumed fixed luminosity at the top of the atmosphere. Non-negligible extra luminosity generated in the outer convective layers would, however, yield $L_{\rm radw}<L$, in conflict with our assumptions, and could change atmospheric structure. A simple way to check this \textit{a posteriori} is by rewriting the local energy equation (\ref{eq:structd}) as $\partial L/\partial m=-T \partial S/\partial t$, and integrating it throughout the atmosphere assuming that $\partial S/\partial t$ is fixed (e.g., \citealt{arras06}). If the value of this integral, $I$\footnote{This integral does not have units of luminosity, as we drop the constant $\partial S/\partial t$ term.}, throughout the innermost convective region is significantly larger than its value throughout the rest of the envelope, $\Delta I$, then our assumptions hold. For all the models for which $\Delta L \ll L$ holds (see paragraph above), we have found $\Delta I / I \lesssim 30\%$, and typically $\lesssim 10\%$ (see \App{radwindow} for additional details).

\end{enumerate}

% If radiative windows exist, our model requires most of the atmospheric luminosity to be generated in the innermost convective region. By design, the luminosity in the radiative windows, $L_{\rm radw}$, must satisfy $L_{\rm radw}=L$, where $L$ is the assumed fixed luminosity at the top of the atmosphere. Non-negligible extra luminosity generated in the outer layers would, however, yield $L_{\rm radw}<L$, in conflict with our assumptions, and could change atmospheric structure. A simple way to check this \textit{a posteriori} is by rewriting the local energy equation (\ref{eq:structd}) as $\partial L/\partial m=-T \partial S/\partial t$, and integrating it throughout the atmosphere assuming that $\partial S/\partial t$ is constant throughout the convective zone (e.g., \citealt{arras06}). If the value of this integral, $I$\footnote{This integral does not have units of luminosity, as we drop the constant $\partial S/\partial t$ term.}, throughout the innermost convective region is significantly larger than its value throughout the rest of the envelope, $\Delta I$, then our assumptions hold. For $p=3.5$ in the power-law (\ref{eq:graindistr}), we have found $\Delta I / I \lesssim 30\%$, and typically $\lesssim 10\%$ (see \App{radwindow} for additional details). For $p=2.5$, however, this approximation breaks down at low core masses, as we show in \S\ref{critical} --- specifically, the additional luminosity generated in either the radiative windows or the outer radiative layer is no longer non-negligible compared to the fixed atmospheric luminosity $L$.

%For all our models of interest, we have found $\Delta I / I \lesssim 30\%$, and typically $\lesssim 10\%$ (see \App{radwindow} for additional details).

%Additionally,

%two challenges for our model, as follows:

%\begin{enumerate}
%
% \item The additional luminosity, $\Delta L$, generated in these radiative windows may not satisfy $\Delta L \ll L$, where $L$ is the assumed fixed atmospheric luminosity. This would make our assumption of constant luminosity throughout radiative regions invalid. In practice, however, the inner radiative windows are either very narrow compared to the height of the convective regions (Figure \ref{fig:delvsr}, middle panel), or  have $\delad \approx \delrad$ throughout, which makes the distinction between convective and radiative layers less  pronounced (Figure \ref{fig:delvsr}, bottom panel). For $p=3.5$ in the power-law (\ref{eq:graindistr}), we have checked that $\Delta L \ll L$ in all the cases presented in this study. For $p=2.5$, however, this approximation breaks down at low core masses, as we show in \S\ref{critical}.
%
%%This results in a negligible extra luminosity generated in the radiative windows. However, the radiative windows may become non-negligible for other opacity choices and sufficiently low core masses, as we show later in this section.
%
%\item If radiative windows exist, our model requires most of the atmospheric luminosity to be generated in the innermost convective region. By design, the luminosity in the radiative windows, $L_{\rm radw}$, must satisfy $L_{\rm radw}=L$, where $L$ is the assumed fixed luminosity at the top of the atmosphere. Non-negligible extra luminosity generated in the outer convective layers would, however, yield $L_{\rm radw}<L$, in conflict with our assumptions, and could change atmospheric structure. A simple way to check this \textit{a posteriori} is by rewriting the local energy equation (\ref{eq:structd}) as $\partial L/\partial m=-T \partial S/\partial t$, and integrating it throughout the atmosphere assuming that $\partial S/\partial t$ is fixed (e.g., \citealt{arras06}). If the value of this integral, $I$\footnote{This integral does not have units of luminosity, as we drop the constant $\partial S/\partial t$ term.}, throughout the innermost convective region is significantly larger than its value throughout the rest of the envelope, $\Delta I$, then our assumptions hold. For all the models for which $\Delta L \ll L$ holds (see paragraph above), we have found $\Delta I / I \lesssim 30\%$, and typically $\lesssim 10\%$ (see \App{radwindow} for additional details).
%
%\end{enumerate}

 %throughout the innermost convective layer and ensuring it is much larger than its value throughout the rest of the atmosphere. We have performed this check and found that

%most of the atmospheric luminosity be generated in the innermost convective region of the envelope, since we

%In practice, however, the inner radiative windows are either very thin (Figure \ref{fig:delvsr}, middle panel), or the temperature gradient is flat enough to result in a roughly constant luminosity throughout the region (Figure \ref{fig:delvsr}, bottom panel; also see Equation \ref{eq:structc}). This results in a negligible extra luminosity generated in the radiative windows. However, the radiative windows may become non-negligible for other opacity choices and sufficiently low core masses, as we show later in this section. Despite the existence of one or more radiative windows in the planet's atmosphere, the luminosity that emerges at the outer boundary is set by the outer radiative layer. To order of magnitude, Equation (\ref{eq:delrad}) can be expressed as
%\begin{equation}
%\label{eq:sigmatau4}
%\mathcal{F} \sim \frac{\sigma T\cb^4}{\tau},
%\end{equation}
%where $\mathcal{F}$ is the radiative flux through the RCB and $\tau \sim \kappa\cb \rho\cb H\cb$ is the optical depth of the first scale height of the radiative region, $H\cb \equiv k_{\rm b} T R^2/({\mu m_{\rm p}} G M)$, with all quantities evaluated at the RCB. While Equation (\ref{eq:delrad}) is always true as part of the structure Equations set (\ref{eq:struct}), Equation (\ref{eq:sigmatau4}) shows that the atmospheric luminosity is set by the diffusive flux through the outermost RCB. We have verified that Equation (\ref{eq:sigmatau4}) is valid for our model atmospheres.

%A simple estimate of the luminosity emerging at the RCB can be found as follows. 
%
%The equation of radiative diffusion (\ref{eq:structc}) can be approximated in terms of the radiative flux $\mathcal{F}$ as 
%
%\begin{equation}
%\label{eq:sigmatau4}
%\mathcal{F} \sim \frac{\sigma T\cb^4}{\tau},
%\end{equation}
%where $\tau$ is the optical depth of the upper radiative layer. Furthermore, we find that $\tau$ is dominated by the optical depth of the first scale height, i.e. 
%
%\begin{equation}
%\label{eq:taucb}
%\tau \sim \tau\cb=\kappa\cb \rho\cb H\cb,
%\end{equation}
%with
%\begin{equation}
%\label{eq:Hcb}
%H\cb=\frac{k_B T\cb}{\mu m_p} \frac{R\cb^2}{G M\cb},
%\end{equation}
%where $\mu$ is the mean molecular weight of the gas, $m_p$ the proton mass, and all other quantities are evaluated at the RCB. By substituting Equations (\ref{eq:Hcb}) and (\ref{eq:taucb}) into Equation (\ref{eq:sigmatau4}), and noting that $\mathcal{F}=L/(4 \pi (R\cb+H\cb)^2) \approx L / (4 \pi R\cb^2)$, we recover Equation (\ref{eq:delrad}) applied at the boundary between the uppermost convective region of the atmosphere and the outer radiative layer (where $\delrad=\delad$) as
%
%\begin{equation}
%\label{eq:Lapprox}
%L \sim \frac{4 \pi \sigma T\cb^3 G M\cb \mu m_p}{k_B \kappa\cb \rho\cb}.
%\end{equation}
%Equation (\ref{eq:Lapprox}) can be used to estimate the luminosity emerging at the top of the convective layer of the atmosphere. We have verified that the luminosity emerging at the bottom of the outer radiative layer in our models can indeed be approximated by Equation (\ref{eq:Lapprox}). %As $T\cb$ only differs from the disk temperature by an order unity factor, $L\sim 1/\rho\cb$. The change in EOS thus results in a deeper RCB and a lower luminosity.

We thus find that our atmospheric and cooling model is valid in our regions of interest, with some exceptions discussed in \S\ref{critical}. %Due to the variable number and position of radiative windows, and therefore radiative-convective boundaries, within the planet atmosphere, we cannot consistently calculate the time evolution of different atmospheres if we evaluate our cooling Equation (\ref{eq:coolingglobal}) at the RCB, as we do in our standard model. We choose to evaluate the cooling time at the Bondi radius instead (since our cooling model applies at any radius $R$, see section \S\ref{struct}). %We note that our choice of $R$ does not change the estimate of the atmosphere evolution time, to order of magnitude, since the additional luminosity generated in all radiative regions is negligible for our opacity choice (also see Paper I).

%\subsection{Outer Boundary Effects}

%\textbf{Reemphasize the fact that the atmosphere structure is determined by your outer boundary conditions: T_{out}, P_{out}, R_{out} $\rightarrow$ explore the separate effects of pressure, temperature and a (since the Hill radius is determined by a); show how temperature is the strongest effect. }

\section{Critical Core Mass}
\label{critical}

%\textbf{Define what it is (refer again to paper I also); show the Mcrit vs a for fixed disk life plot, for both real EOS, gamma 7/5 and gamma 5/3; justify the differences in terms of the EOS effects from 4.2; show that using a real EOS makes a significant difference to the results; however, the core masses we get are still doable. Again, a lot of text below will be used but needs reorganizing/rephrasing.}

%\textbf{Define what it is (also refer to paper I) and emphasize how it's different from $M_{crit}$ in standard calculations. Define the crossover mass and crossover time.}

In this section we put together the results obtained in Sections \S\ref{EOSeffects}  and \S\ref{sec:opacity}, and determine the minimum core mass, $M_{\rm crit}$, to initiate runaway gas accretion during a typical protoplanetary disk lifetime, $t= 3$ Myr. As in Paper I, we quantify the runaway accretion time $t_{\rm run}$ as the time at which the atmosphere growth timescale $M_{\rm atm}/\dot{M}$ drops to 10\% of its maximum value (see Paper I for details).

%We first explore the dependence of $t_{\rm run}$ on the core mass for a fixed semimajor axis. We then determine the critical core mass $M_{\rm crit}$ to form a giant planet from a gas composed of a realistic hydrogen-helium mixture, and we compare this with the results from Paper I for an ideal diatomic gas. Finally, we determine $M_{\rm crit}$ under more realistic opacity assumptions.

%\subsection{Crossover Time as a Function of Core Mass}
%\label{tvsM}

%\textbf{t vs. M at fixed distance, similar to the plot from paper I. Compare scalings.}

Figure \ref{fig:tvsMplot} displays the time evolution and the runaway growth time for atmospheres forming around cores with masses between 15 $M_{\oplus}$ and 25 $M_{\oplus}$ at $a=10$ AU in our fiducial disk, for a realistic EOS with an equilibrium ortho-to-para ratio and standard ISM opacity. Higher mass cores have shorter $t_{\rm run}$, consistent with the results of Paper I. We also note that $M_{\rm atm}$ at $t_{\rm run}$ is larger for the realistic EOS than for the polytropic EOS considered in Paper I, for the same core mass.

\begin{figure}[h!]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/t_vs_M_10au.pdf}
\includegraphics[width=0.5\textwidth]{figures/t_vs_M_10au.pdf}
%%\vspace{-0.5in}
\caption{Elapsed time as a function of atmosphere mass, for cores with fixed masses between $15 M_{\oplus}$ and $25 M_{\oplus}$ at $a=10$ AU in our fiducial disk, for a realistic EOS with an equilibrium ortho-to-para ratio and standard ISM opacity. The circles mark the runaway growth time. The numbers label the core mass in Earth masses. A larger core mass results in a lower $t_{\rm run}$.}% and a higher $M_{\rm atm}/M\co$ at runaway.}
\label{fig:tvsMplot}
\end{figure}

%\subsection{Critical Core Mass}
%\label{Mcrit}

%\textbf{Mcrit vs. a plot, realistic EOS and polytrope. Discuss the larger critical core mass for the real EOS in light of the effects from section 4.}

Figure \ref{fig:Mvsaplot}, upper panel, displays $M_{\rm crit}$
%(e.g., \citealt{jay99}), 
for a gas described by a realistic EOS and an ISM dust opacity. The results of Paper I for an ideal diatomic gas are plotted for comparison. When compared to an ideal gas polytrope, the inclusion of realistic EOS effects increases $M_{\rm crit}$ by a factor of $\sim$2 if the $H_2$ spin isomers are in equilibrium, and by a factor of $\sim$$2-4$ for a fixed 3:1 ortho-to-para ratio. This latter increase is more significant at larger stellocentric distances. In Figure \ref{fig:Mvsaplot}, bottom panel, we compare our results with those for a disk with a gas surface density an order of magnitude larger than $\Sigma_{\rm d}$ of our fiducial disk (see Equation \ref{eq:diska}), and find that $M_{\rm crit}$ reduces by $\sim$15-25$\%$.

 %As such, non-ideal effects substantially increase the core mass needed to form a giant planet  before the dissipation of the protoplanetary disk.

%Figure \ref{fig:Mvsaplot} shows the critical core mass for a massive atmosphere to form during a typical lifetime of a protoplanetary disk $t=3$ Myrs 
%(e.g., \citealt{jay99}), 
%for a gas described by a realistic EOS and an ISM dust opacity. The results of Paper I for an ideal diatomic gas are plotted for comparison. The inclusion of realistic EOS effects increases $M_{\rm crit}$ by more than a factor of two when compared to an ideal gas polytrope. %As such, non-ideal effects substantially increase the core mass needed to form a giant planet  before the dissipation of the protoplanetary disk.

\begin{figure}[h!]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/Mc_vs_a_poly_real_paper.pdf}
\includegraphics[width=0.5\textwidth]{figures/Mc_vs_a_poly_real_paper.pdf}
%%\vspace{-0.5in}
\caption{Top panel: The minimum core mass for an atmosphere to initiate runaway gas accretion within the lifetime of a typical protoplanetary disk $t \sim 3$ Myrs as a function of semimajor axis, for a realistic hydrogen-helium mixture and a standard ISM opacity. The results of Paper I for an ideal diatomic gas are plotted for comparison. The realistic EOS yields core masses larger by a factor of $\sim$2 when compared to the polytrope, for an equilibrium ortho-to-para ratio. The critical core mass is $\sim$$2-4$ times larger than the polytrope case for a fixed 3:1 ratio between the $H_2$ spin isomers. The increase is more pronounced at larger stellocentric distances. Bottom panel: Critical core mass as a function of semimajor axis for a disk gas surface density 10 times larger than that of our fiducial disk. A larger $\Sigma_{\rm d}$ reduces $M_{\rm crit}$ by $\sim$$15-25 \%$.}
\label{fig:Mvsaplot}
\end{figure}

\begin{figure}[h!]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/Mcrit_vs_a_gg.pdf}
\includegraphics[width=0.5\textwidth]{figures/Mcrit_vs_a_gg.pdf}
%%\vspace{-0.5in}
\caption{Critical core mass as a function of semimajor axis for a realistic EOS with an equilibrium ortho-to-para ratio and radiative opacities that account for grain growth (purple triangles, with $p=3.5$ and $a_{\rm max}=1$ cm; see text for details). The critical core mass is lower than it would be if dust grains had an ISM-like size distribution (blue circles).}
\label{fig:Mcritvsagg}
\end{figure}

Figure \ref{fig:Mcritvsagg} shows $M_{\rm crit}$ as a function of semimajor axis, for a realistic EOS with an equilibrium ortho-to-para ratio and grain growth opacity with a size distribution given by Equation (\ref{eq:graindistr}) with $p=3.5$ and maximum particle size $s_{\rm max}=1$ cm. The critical core mass is lower than in the standard interstellar opacity case, and less sensitive to location in the disk. Location primarily affects the atmosphere through the opacity in the outer envelope, which depends on disk temperature. For the simplified analytic model developed in Paper I, we approximated $t_{\rm run}$ by  $t_{\rm co}$, the time when $M_{\rm atm}=M\co$, and found that $t_{\rm co} \sim T_{\rm d}^{\beta+1/2}$, with $\beta$ the power-law exponent in Equation (\ref{eq:opacitylaw}). Opacity is less sensitive to temperature variations for larger grains and has an almost flat profile (see Figure \ref{fig:opacity}), which results in $\beta \ll 1$ and a much weaker temperature (and therefore semimajor axis) dependence of $M_{\rm crit}$, as seen in Figure \ref{fig:Mcritvsagg}. Moreover, grain growth reduces the absolute value of the opacity, which also lowers $M_{\rm crit}$. % As noted in Section \S\ref{sec:opacity}, the critical core mass may be significantly lower if coagulation is taken into account, i.e. $p=2.5$.

\begin{figure}[h!]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/tco_vs_a_Mc4_comp.pdf}
\includegraphics[width=0.5\textwidth]{figures/tco_vs_a_Mc4_comp.pdf}
%%\vspace{-0.5in}
\caption{Runaway accretion times for a realistic EOS with an equilibrium ortho-to-para ratio and different grain size distributions, for an atmosphere forming around a core with $M\co=4 M_{\oplus}$. The lines marked by purple and orange triangles have grain growth opacities with $a_{\rm max}=1$ cm, and $p=3.5$ and $p=2.5$, respectively (see text for details). The blue circle line has an ISM power-law opacity. The runaway accretion time is more than one order of magnitude lower when coagulation is accounted for, i.e. $p=2.5$.}
\label{fig:p25p35}
\end{figure}

In Equation (\ref{eq:graindistr}), the coefficient $p=3.5$ corresponds to a standard collisional cascade. If coagulation is taken into account, the exponent $p$ can be approximated as $p=2.5$ \citep{dalessio01}. This results in a flatter and significantly lower opacity (see \App{radwindow}), which may substantially reduce $M_{\rm crit}$. However, we have found that our model breaks down for low core masses ($M\co \lesssim 3 M_{\oplus}$) under our assumption of constant luminosity in the outer radiative layer. Figure \ref{fig:p25p35} shows the runaway accretion time $t_{\rm run}$ as a function of semimajor axis for the lowest core mass for which our model is valid, $M\co=4 M_{\oplus}$. The runaway accretion time is more than one order of magnitude lower for $p=2.5$, which implies that $M_{\rm crit}$ may be, in fact, significantly lower than presented in Figure \ref{fig:Mcritvsagg}. In other words, grain growth can yield critical core masses up to an order of magnitude lower than in the case where interstellar opacities are used.

In summary, we have found $M_{\rm crit}\sim 30 M_{\oplus}$ at 5 AU, steadily decreasing to $\sim$$7 M_{\oplus}$ at 100 AU, for a realistic EOS with the $H_2$ spin isomers in equilibrium and interstellar opacity. For a fixed 3:1 ortho-to-para ratio and interstellar opacity, $M_{\rm crit}$ is $\sim$$32 M_{\oplus}$ at 5 AU and decreases to $\sim$$15 M_{\oplus}$ at 100 AU. For a grain growth opacity with a size distribution given by Equation (\ref{eq:graindistr}) with $p=3.5$ and an equilibrium ortho-to-para ratio, $M_{\rm crit}$ significantly drops to $\sim$$8 M_{\oplus}$ at 5 AU and $\sim$$5 M_{\oplus}$ at 100 AU. Accounting for coagulation (i.e., $p=2.5$) $M_{\rm crit}$ is less than $4M_\oplus$ and may be up to an order of magnitude smaller.

%Figure \ref{fig:p25p35} shows the crossover time $t_{\rm co}$ as a function of semimajor axis for a core of mass $M\co=4 M_{\oplus}$ and for two choices of the exponent $p$ in the power-law (\ref{eq:graindistr}): $p=3.5$, which corresponds to a normal collisional cascade, and $p=2.5$, which accounts for coagulation \citep{dalessio01}. We choose $M\co=4 M_{\oplus}$ as our model breaks down for lower core masses ($M\co \lesssim 3 M_{\oplus}$) under our assumption of constant luminosity in the outer radiative layer. The crossover time is lower for both choices of $p$ when compared to $t_{\rm co}$ when an ISM opacity is assumed. Moreover, $t_{\rm co}$ is more than one order of magnitude lower for $p=2.5$ than for $p=3.5$. As $t_{\rm co}$ is lower than the typical disk lifetime of a few Myrs for $M\co=4 M_{\oplus}$ and $p=2.5$, the critical core mass under these assumptions is lower than the core mass for which our model is valid. We thus cannot accurately calculate $M_{\rm crit}$ for $p=2.5$ and choose $p=3.5$ as our fiducial case. We note, however, that the critical core mass may be, in fact, significantly lower that the results we present in \S\ref{critical}.

%So far we have assumed that the size distribution of dust grains is a power-law (\ref{eq:graindistr}) with $p=3.5$, which corresponds to a normal collisional cascade. If, however, coagulation is taken into account, the exponent $p$ can be approximated as $p=2.5$ \citep{dalessio01}. This results in a flatter and significantly lower opacity, which may substantially reduce the critical core mass. However, we have found that our model breaks down for low core masses ($M_{\rm c} \lesssim 3 M_{\oplus}$) under this assumption, i.e. the outer radiative zone becomes deep enough that the luminosity generated in this region can no longer be neglected. The crossover time is more than one order of magnitude lower for $p=2.5$, which implies that the critical core mass may be, in fact, significantly lower than presented in Figure \ref{fig:Mcritvsagg}.

%\section{Model Relevance in Planet Formation Theory}
\section{Effects of Planetesimal Accretion}
\label{acc}

% In this scenario, the energy the envelope receives from planetesimals balances its luminosity. The core's atmosphere is in steady state, and  $M_{\rm crit}$ is uniquely determined for a fixed planetesimal accretion rate, $\dot{M_{\rm c}}$, and a set of disk conditions. For high $\dot{M_{\rm c}}$, the resulting $M_{\rm crit}$ needed to grow an atmosphere during the disk lifetime is too large, which has led to the belief that core accretion does not work at large separations.

This study considers protoplanets with fully formed cores for which planetesimal accretion is negligible and KH contraction dominates the luminosity evolution of the atmosphere. This approach contrasts with that of models which assume high planetesimal accretion rates and find that the atmosphere is in steady state and solely heated due to accretion of solids. In both cases, as the envelope and core become comparable in mass, hydrostatic balance no longer holds and runaway gas accretion commences. For fast accretion, $M_{\rm crit}$ is uniquely defined as the maximum core mass for which the atmosphere is still in hydrostatic equilibrium for a fixed planetesimal accretion rate and a set of disk conditions. In this section we compare our results for $M_{\rm crit}$ to analogous results from steady-state fast planetesimal accretion calculations. We discuss the core accretion rates that are necessary for our regime to be valid in \S\ref{raf1}. In \S\ref{raf3}, we estimate core growth at the maximum rate for which the KH regime is valid, and show it is negligible over the timescale on which the atmosphere evolves. Finally, we compare our results with those assuming fast planetesimal accretion in \S\ref{raf2}.

 %In this section we investigate the core accretion rates that are necessary for our regime to be valid. We also discuss the conditions under which runaway gas accretion can be initiated due to the Kelvin-Helmholtz contraction of the atmosphere before it becomes critical due to planetesimal accretion.

\subsection{Planetesimal Accretion Rates}
\label{raf1}

Kelvin-Helmholtz contraction dominates an atmosphere's luminosity if  $L_{\mathrm{acc}} < L_{\rm{KH}}$, where $L_{\rm{acc}}$ is the accretion luminosity,

\begin{equation}
\label{eq:Lacc}
L_{acc}=G \frac{M_{\rm{c}} \dot{M_{\rm{c}}}}{R_{\rm{c}}},
\end{equation}
and $L_{\rm KH}$ is given by Equation (\ref{eq:coolingglobal}) with $L\co=\Gamma=0$. This condition is satisfied as long as the planetesimal accretion rate

\begin{equation}
\label{eq:McdotKH}
\dot{M\co}<\dot{M}_{\rm c, KH} \equiv \frac{L_{\rm KH} R\co}{G M\co}.
\end{equation} 
To illustrate the magnitude of $\dot{M}_{\rm c, KH}$, we choose as a fiducial case an atmosphere forming at 30 AU and with a core mass of $10 M_{\oplus}$. Since analytic studies of critical core masses at high planetesimal accretion rates assume an ideal gas EOS, for ease of comparison we choose an ideal gas polytropic EOS with constant adiabatic gradient $\delad=2/7$ and mean molecular weight $\mu=2.35$ (see also Paper I). For this choice of parameters, the runaway accretion time is $t _{\rm run}\sim$ 1.4 Myrs, which is within the typical lifetime of a protoplanetary disk. We also estimate two reference accretion rates. The first one is the core accretion rate $\dot{M}_{\rm c, acc}$ needed to grow the core to $M\co=10 M_{\oplus}$ on the same timescale as our model atmosphere, $\tau=1.4$ Myrs:

\begin{equation}
\label{eq:Mcdot}
\dot{M}_{\rm{c,acc}}(M_{\rm{c}}) \sim \frac{M_{\rm{c}}}{\tau}.
\end{equation}
The second reference planetesimal accretion rate is $\dot{M}_{\rm c, Hill}$, a typically assumed planetesimal accretion rate for which the random velocities of the planetesimals are of the order of the Hill velocity around the protoplanetary core (for a review, see \citealt{goldreich04}). Following R06 (equation A1),

%This is the accretion rate at the boundary between the dispersion dominated and shear dominated regimes.

\begin{equation}
\label{eq:MdotHill}
\dot{M}_{\rm{c,Hill}}=\Omega \Sigma_{\rm p} R\co R_{\rm H},
\end{equation}
where $\Sigma_{\rm p}$ is the surface density of solids, assumed to satisfy $\Sigma\di \approx 100 \Sigma_{\rm p}$ for a dust-to-gas ratio of 0.01.

Figure \ref{fig:accrates} shows that $\dot{M}_{\rm c,KH}$ is $\sim2-3$ orders of magnitude lower than $\dot{M}_{\rm c, acc}$. Had the core accreted planetesimals at the $\dot{M}_{\rm c, KH}$ rate since it started forming, it could not have grown large enough to attract an atmosphere within a typical disk lifetime. Our model requires that the planetesimal accretion rate is initially large during core growth, then significantly reduces as the gaseous envelope accumulates, as suggested by, e.g., \citet{pollack96}. This is a plausible scenario: the core's feeding zone may be depleted of solids if it is not refilled by radial drift of planetesimals through the nebula, or the core may form in the inner part of the disk and later be scattered outwards by other giants in the system \citep{ida13}.

%or the planet's feeding zone could have been depleted of solids due to a giant neighbor.

 \begin{figure}[h]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/acc_rates_paper.pdf}
\includegraphics[width=0.5\textwidth]{figures/acc_rates_paper.pdf}
%\vspace{-0.5in}
\caption{Various accretion rates for a planet forming at 30 AU and with a core mass $M\co=10 M_{\oplus}$, using a polyropic EOS and ISM power-law opacity. For this choice of parameters, the runaway accretion time is $t_{\rm run} \sim 1.4$ Myrs. The $\dot{M}_{\rm{atm}}$ (solid blue) curve represents the growth rate of the atmosphere as estimated by our model. The core accretion rate $\dot{M}_{\rm c, acc}$ (dashed green) necessary to grow the core on the timescale $t_{\rm run}$ is larger than $\dot{M}_{\rm c, KH}$ (dotted red), the maximum planetesimal accretion rate during KH contraction for which our regime is valid (see text). The frequently used planetesimal accretion rate $\dot{M}_{\rm c, Hill}$ (dashed-dotted light blue) for which the random velocity of the planetesimals is given by the Hill velocity due to the core (see text), also exceeds $\dot{M}_{\rm c, KH}$.This motivates our requirement  that planetesimal accretion must have slowed down after core growth for our model to be valid.}

%and $\dot{M}_{\rm{c,KH}}$ (dotted red) is the maximum planetesimal accretion rate during the gas contraction phase in order for our regime to be valid, i.e. $L_{\mathrm{acc}} < L_{\rm{KH}}$ (see text). For comparison, we plot the core accretion rate $\dot{M}_{\rm{c,acc}}$ (dashed green) necessary to grow the core on a timescale $\tau=t_{\rm run}$, and a frequently used planetesimal accretion rate $\dot{M}_{\rm{c,Hill}}$ (dash-dotted light blue) for which the random velocity of the planetesimals is given by the Hill velocity due to the core (see text). We note that $\dot{M}_{\rm c, KH}$ and $\dot{M}_{\rm c, Hill}$ are both lower than $\dot{M}_{\rm c, acc}$, motivating our requirement  that planetesimal accretion must have slowed down after core growth for our model to be valid.}
\label{fig:accrates}
\end{figure}

\subsection{Core Growth during KH Contraction}
\label{raf3}

Planetesimal accretion during the KH contraction phase of atmosphere growth at a rate $\dot{M}_{\rm c}<\dot{M}_{\rm{c,KH}}$  cannot alter the core mass enough to affect the time evolution of the atmosphere. We can quantitatively estimate the maximum increase in core mass as

\begin{equation}
\label{eq:cminc}
\Delta M_{\rm{c}} = \int_0^{t_{\rm{run}}} \dot{M_{\rm{c}}} dt \approx \sum_i \dot{M_{\rm{c}}}_i \Delta t_i,
\end{equation}

 \noindent where the accretion rate $ \dot{M_{\rm{c}}}_i $ is given by

 \begin{equation}
 \label{eq:Mdotexp}
 \dot{M_{\rm{c}}}_i =\frac{L_i R\co}{G M_{\rm{c}}} 
 \end{equation}

 \noindent from Equation (\ref{eq:Lacc}), with $L_i$ the luminosity of the atmosphere at time $t_i$ in our model. For $M_{\rm{c}}=10 M_{\oplus}$, we find $\Delta M_{\rm{c}} \approx 0.05 M_{\oplus} \ll 10 M_{\oplus}$. Core growth is negligible in our regime, and the time evolution of the atmosphere is thus insensitive to core mass changes at a rate imposed by the assumption that $L_{\rm{acc}}<L_{\rm{KH}}$.

 %It is easy to see that  $\dot{M}_{\rm{c,typical}}$ is more than one order of magnitude lower than the gas accretion rate of our model atmosphere $\dot{M}_{\rm{atm}}$, and lower than the core accretion rate $\dot{M}_{\rm{c,acc}}$ needed to grow the core and the atmosphere at the same time within the disk life time. As such, the formation of a giant planet by growing the core first, then letting the atmosphere cool is faster than growing the core and the atmosphere at the same time at a steady planetesimal accretion rate.

\subsection{Comparison with Steady-State Results}
\label{raf2}

We compare our results for $M_{\rm crit}$ with those of studies that assume large planetesimal accretion rates.  In principle, the disk lifetime could be short enough that our calculated $M_{\rm crit}$ 
%calculated using KH contraction in the absence of planetesimal accretion 
could exceed the $M_{\rm crit}$ estimated for a steady-state, accretion-heated core.  This prospect is not self-contradictory since at higher luminosity, an atmosphere evolves more quickly and hence reaches steady state on a timescale shorter than our calculated KH contraction time.   
%If the $M_{\rm crit}$ found by these studies were lower than the $M_{\rm crit}$ in our regime of negligible solids accretion, the atmosphere would undergo runaway gas accretion before KH contraction became dominant, and our regime would not be relevant. 
We show here that disk lifetimes are long enough that this is not the case.  Our model yields lower core masses than those found when fast planetesimal accretion is considered.

The critical core mass is larger for higher planetesimal accretion, as additional heating increases the core mass required for collapse. As such, if atmosphere collapse does not occur for the lowest value of $\dot{M}_{\rm c, KH}$ over the course of the atmosphere's growth, then it can only occur in the KH dominated regime. %In what follows we calculate the critical core mass $M_{\rm crit, KH}$ corresponding to $\min(\dot{M}_{\rm c, KH})$ and show that it is higher that the critical core mass we determined in the KH dominated regime.

%The maximum planetesimal accretion rate $\dot{M}_{\rm c, KH}$ that satisfies equation (\ref{eq:McdotKH})

%is denoted by the dotted line in Figure \ref{fig:accrates} for $a=60$ AU and $M\co=5M_{\oplus}$. If unstable atmosphere collapse does not occur due to planetesimal accretion for the lowest value of $M_{\rm c, KH}$, then runway gas accretion can only occur in the KH dominated regime. In what follows we calculate the critical core mass $M_{\rm c, KH}$ corresponding to $\dot{M}_{\rm c, KH}$ and show that it is higher than our calculated critical core mass.

%A core that forms on the same timescale as our model atmosphere accretes planetesimals at a rate given by equation (\ref{eq:Mcdot}).

%This accretion rate is dependent on the core mass, which is steadily increasing. We therefore compare the critical core mass due to planetesimal accretion at this rate $M_{\rm{crit,acc}}$ to the critical core mass as defined in our estimates $M_{\rm{c,crit}}$ (see section \ref{critical}). If $M_{\rm{crit,acc}}<M{\rm_{c, crit}}$, then the atmosphere has already initiated unstable gas accretion by the time Kelvin-Helmholtz contraction starts dominating.

%Specifically, we compare the critical core mass due to planetesimal accretion at the rate

%we compare the critical core mass $M_{\rm{c,acc}}$ due to planetesimal accretion, for an accretion rate that satisfies $L_{\rm{acc}}<L_{\rm{KH}}$ to the actual core mass $M_{\rm{c}}$ assumed fixed in our model. If $M_{\rm{c,acc}}<M{\rm{c}}$, then the atmosphere has already initiated unstable gas accretion by the time Kelvin-Helmholtz contraction starts dominating.

In order to estimate the critical core mass $M_{\rm crit, KH}$ corresponding to planetesimal accretion at the rate $\dot{M}_{\rm c, KH}$, we use the results of R06 for low luminosity atmospheres forming in the outer disk ($\gtrsim2-5$ AU). %, consistent with our region of interest. 
R06 assumes an ideal gas polytropic EOS and an opacity lower than that of the ISM (see Equation \ref{eq:opacitylaw}). For comparison, we calculate $M_{\rm crit}$ for an ideal gas polytrope and an opacity, $\kappa$, given by Equation \ref{eq:opacitylaw} with $F_{\kappa}$ reduced by a factor of 100.  This choice is comparable to the opacity law used by R06.\footnote{The power-law opacity of R06 is scaled to the (semimajor axis dependent) disk temperature, while our opacity is scaled to an absolute reference temperature. We thus cannot directly use the R06 opacities for our comparison.} %(see Paper I).

Following R06, we find that the critical core mass when accretion luminosity dominates the evolution of the atmosphere can be expressed as

\begin{equation}
\label{eq:critraf}
M_{\rm{crit, KH}} \sim \Big[\frac{\min[\dot{M}_{\rm c, KH}(M_{\rm{c}})]}{64 \pi^2 C} \frac{\kappa}{\sigma G^3} \frac{1}{R\co M\co^{1/3}} \Big(\frac{k_B}{\mu m_p}\Big)^4\Big]^{3/5},
\end{equation}
where $C$ is an order unity constant depending on the adiabatic gradient and disk properties (see R06, Equation B3). From Equation (\ref{eq:McdotKH}), the accretion rate $\dot{M}_{\rm c, KH}$ depends on the core mass $M\co$. We find $M_{\rm crit, KH}$ numerically by setting $M\co=M_{\rm crit, KH}$ on the right-hand side of Equation (\ref{eq:critraf}). The result is displayed in Figure \ref{fig:raf2}; the critical core mass corresponding to planetesimal accretion at the rates displayed in Figure \ref{fig:accrates} is displayed for comparison.

 \begin{figure}[h]
\centering
%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/Mc_vs_a_poly_comp.pdf}
\includegraphics[width=0.5\textwidth]{figures/Mc_vs_a_poly_comp.pdf}
%\vspace{-0.5in}
\caption{Comparison between the critical core mass $M_{\rm{crit, KH}}$ given significant planetesimal accretion and the critical core mass when gas contraction dominates, for a polytropic EOS and an ISM opacity reduced by a factor of 100. Our results yield lower core masses than in the fast planetesimal accretion case (e.g., \citealt{rafikov06}). The critical core mass corresponding to $\dot{M}_{\rm c, Hill}$ and $\dot{M}_{\rm c, acc}$ from Figure \ref{fig:accrates} is plotted for comparison.}
\label{fig:raf2}
\end{figure}

The critical core mass in the regime where KH contraction dominates is smaller than in the case in which planetesimal accretion dominates the evolution of the atmosphere. This leads to two conclusions:

\begin{enumerate}
\item Planetesimal accretion can be safely ignored in our regime.
\item Giant planets can form from smaller cores if planetesimal accretion significantly reduces during atmosphere growth. 
\end{enumerate}

%As additional heating due to planetesimals limits the atmosphere's ability to cool, our result represents a true minimum for the core mass needed to form a gas giant during the lifetime of the protoplanetary disk.

% First, we confirm that planetesimal accretion can be safely ignored in our regime of interest. Secondly, this comparison tells us that the core mass needed to form a giant planet is smaller if the core forms first, then accretes a massive envelope, than in the case where the core and atmosphere grown simultaneously in a high planetesimal accretion regime. Moreover, our result represents a true, absolute minimum on the core mass needed to form a giant planet during the lifetime of the protoplanetary disk, as our core no longer grows.

 \section{Summary}
 \label{conclusions}

 %OK, very interesting.  Clearly the text (including abstract and summary I think) needs to be modified to include this result.  The headline for this result seems to be: at very low temperatures, the choice of ortho/para ratio makes as big a difference as just using a simple polytrope vs. a real EOS!!  This is consistent with our knowledge that spin effects are more important than dissociation once the spin levels are no longer thermally populated ($T_disk \lesssim T_spin \sim 90$K).  And indeed the behavior for different ortho/para cases is similar at higher temperature, though we don?t go to high enough temp $T > few x T_spin$ for ortho/para ratio to become completely irrelevant.

 In this paper we study the formation of giant planets embedded in a gas disk. We consider atmospheric evolution around fully grown cores and determine the minimum (critical) core mass, $M_{\rm crit}$, required to form a gas giant during the typical lifetime of a protoplanetary disk. We improve the model developed in  \citet[hereafter Paper I]{piso14} by including realistic equation of state (EOS) tables and dust opacities.

 For a realistic EOS with the molecular hydrogen ($H_2$) spin isomers in thermal equilibrium, and grain growth opacity with maximum particle size $s_{\rm max}=1$ cm and a power-law size distribution (\ref{eq:graindistr}) with $p=3.5$, $M_{\rm crit}$ is $\sim$$8 M_{\oplus}$ at 5 AU in our fiducial disk and drops to $\sim$$5 M_{\oplus}$ at 100 AU. The realistic EOS and grain growth opacity have two competing effects on $M_{\rm crit}$:

 \begin{enumerate}
 \item Realistic EOS effects increase $M_{\rm crit}$ by a factor of $\sim$$2$ when compared to the ideal gas polytrope, for an equilibrium ratio of ortho- and parahydrogen. If the $H_2$ spin isomers are in a fixed 3:1 ratio, $M_{\rm crit}$ increases by a factor of $\sim$$2-4$ when compared to the polytrope. This increase is most significant at larger stellocentric distances, where disk temperatures are lower than the peak ortho-to-para conversion temperature of $\sim$$50$ K.

 \item Grain growth opacities decrease $M_{\rm crit}$ by a factor of $\sim$$3.5$ at 5 AU and by a factor of $\sim$$1.2$ at 100 AU, when compared to ISM opacities, for a particle distribution given by the power-law (\ref{eq:graindistr}) with $p=3.5$ and a maximum particle size of 1 cm. The critical core mass is less sensitive to the location in the disk when realistic opacities are used. If $p=2.5$, an approximation for coagulation,  $M_{\rm crit}$ further reduces by up to an order of magnitude. 
 \end{enumerate}

 %interstellar grain opacity, the critical core mass is $\sim$$20 M_{\oplus}$ at 5 AU and drops to $\sim$$6 M_{\oplus}$ in our fiducial disk. These results are more than twice as large than those calculated in Paper I for a polytropic EOS, and bring the critical core mass to $\sim$$10 M_{\oplus}$, the value typically quoted in many core accretion studies. When realistic opacities that include grain growth are used, the critical core mass is significantly lower, i.e. $\sim$$6 M_{\oplus}$ at 5 AU and $\sim$$4 M_{\oplus}$ at 100 AU. Different assumptions about the grain size distribution (e.g., that take into account coagulation) may further reduce the critical core mass.
 %\vspace{2in}

 In our core accretion models, the dissociation of molecular hydrogen slows the atmospheric cooling and gas accretion rate.  This finding is somewhat surprising because $H_2$ dissociation can trigger the gravitational collapse of protostellar or planetary mass gas clouds (\citealt{bodenheimer80}, \citealt{inutsuka12}).  The presence of a massive solid core is the key factor that prevents dissociation from inducing a global collapse in our models.

%%We note that before the onset of runaway gas accretion, hydrogen dissociation deep in the envelope slows  the contraction of a core's atmosphere. When runaway accretion has increased the planet's mass substantially, dissociation can have the opposite effect, triggering collapse \citep{bodenheimer80}.

%Dissociation is known to trigger the collapse of growing stars  \citep{larson69}. % on higher mass objects, such as  growing stars and collapsing planets forming stars can trigger the collapse of higher mass objects (\citealt{larson69}, \citealt{bodenheimer80}), .  deep in the envelope slows atmospheric contraction. This result contrasts with the effect of dissociation on star formation, where it can trigger dynamical instabilities and the collapse of higher mass objects (\citealt{larson69}, \citealt{bodenheimer80}).

Our results yield lower core masses than analogous results that consider high planetesimal accretion rates for which the core and atmosphere grow simultaneously. It is thus possible to form a giant planet from a smaller core if the core grows first, then the accretion rate of solids is reduced and a gaseous envelope is accumulated. Moreover, since additional heat sources such as planetesimal accretion limit the ability of the atmosphere to cool and undergo Kelvin-Helmholtz contraction, our results represent a true minimum on the core mass needed to form a giant planet during the typical lifetime of a protoplanetary disk.

% In this paper we study giant planet formation assuming that planetesimal accretion is negligible and the atmosphere evolution is dominated by KH gas contraction. We use the model developed in Paper I to determine the critical core mass to form a giant planet before disk dissipation assuming that the nebular gas obeys a realistic EOS that takes into account non-ideal effects. We find that variations in the adiabatic gradient due to thermodynamic effects such as dissociation and hydrogen spin isomers result in a critical core mass more than twice as large than in the case of a gas described by an ideal gas polytrope. This brings the critical core mass to $M_{\rm{c, crit}} \sim 10 M_{\oplus}$, which is the value typically quoted in many core accretion studies (see PY 13). We further make realistic assumptions about the dust opacity in the protoplanetary disk by taking into account grain growth. The resulting critical core mass is much less sensitive to outer boundary conditions (disk temperature and pressure) and is found to be $M_{\rm{c, crit}} \sim 5 M_{\oplus}$. Different grain size distribution assumptions (i.e., that take into account coagulation) may further reduce the critical core mass.

 %While for an ideal diatomic gas the minimum core mass to form a giant planet under the assumptions of our model is lower than the typically quoted value of $10 M_{\oplus}$ (see PY13), the inclusion of non-ideal effects brings this values back to around $10 M_{\oplus}$

% We also compare our results to studies that assume high planetesimal accretion rates, due to which the atmosphere is in steady state and entirely heated by planetesimals, and find that that our model yields lower critical core masses. This shows that it is easier to form a giant planet by growing the core first, then reducing the planetesimal accretion rate and letting the atmosphere evolve on a KH time scale. Moreover, since additional heat sources such as planetesimal accretion limit the ability of the atmosphere to cool and undergo KH contraction, our results represent a true minimum on the core mass needed to form a giant planet during the typical lifetime of a protoplanetary disk.

% In this paper we have studied the formation of giant planet atmospheres under the assumption that Kelvin-Helmoholtz gas contraction dominates the luminosity evolution of the atmosphere over planetesimal accretion. We built quasi-static two-layer atmosphere models with an inner convective region and an outer radiative region that matches smoothly onto the protoplanetary disk. We derived a cooling model to connect series of quasti-static atmospheres, and thus obtained an evolutionary history of the envelope. We defined the time at which unstable atmosphere collapse commences as $M_{\rm atm}(t)\sim M_{\rm c}$. From this we defined as ``critical core mass'' the minimum core mass for a protoplanet to initiate runaway gas accretion during the lifetime of the protoplanetary disk. We studied this minimum mass for a variety of disk conditions, nebular gas compositions and opacities. We found that the critical core mass decreases as we move further out in the disk, and is smaller for lower disk temperatures and opacities and for higher mean molecular weight of the gas. 
% 
% We find that the critical core mass to form a giant planet within the life time of the disk is smaller than the results yielded by studies that assume that the atmosphere evolution is dominated by the luminosity due to planetesimal accretion. We have showed that the planetesimal accretion rate needed to grow the core on a typical disk time scale is larger than the expected planetesimal accretion rates at large separations. As such, it is faster to form a planet by growing the core first in a fast planetesimal accretion regime (e.g., the core forms in the inner disk, then migrates outwards), then significantly reduce planetesimal accretion and allow a massive atmosphere to accumulate.

%Our study assumes that the protoplanetary core forms first, then it r

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/chapter3.tex
upd

--- 2015-11-17 tex/draft/ms_final_revised_2.tex
@@
-%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
-%\documentclass[12pt, preprint]{aastex}
-\documentclass[apj]{emulateapj}
-
-\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
-\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex
-
-%\usepackage{pdfsync}
-\usepackage{subeqnarray}
-\usepackage{natbib}
-\usepackage{color}
-\usepackage[utf8]{inputenc}
-
-\bibliographystyle{apj}
-
-\newcommand{\ie}{i.e.\ }
-\newcommand{\eg}{e.g.\ }
-\newcommand{\p}{\partial}
-\newcommand{\brak}[1]{\langle #1\rangle}
-
-\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
-\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
-\newcommand{\cm}{\; {\rm cm}}
-\newcommand{\mm}{\; {\rm mm}}
-%\newcommand{\ps}{\; {\rm s^{-1}}}
-\newcommand{\km}{\; {\rm km}}
-\newcommand{\au}{\; \varpi_{\rm AU}}
-\newcommand{\AU}{\; {\rm AU}}
-\def\K{\; {\rm K}}
-
-\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
-\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
-\newcommand{\nab}{\vc{\nabla}}
-\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
-\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
-\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}
-
-\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
-\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
-\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
-\newcommand{\App}[1]{Appendix~\ref{#1}}
-\newcommand{\Sec}[1]{Sect.~\ref{#1}}
-\newcommand{\Chap}[1]{Chapter~\ref{#1}}
-\newcommand{\Fig}[1]{Fig.~\ref{#1}}
-\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
-\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
-\newcommand{\Tab}[1]{Table \ref{#1}}
-
-\definecolor{gray}{gray}{0.5}
-\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}
-
-%\newenvironment{packed_item}{
-%\begin{itemize}
-%  \setlength{\itemsep}{1pt}
-%  \setlength{\parskip}{0pt}
-%  \setlength{\parsep}{0pt}
-%}{\end{itemize}}
-
-\begin{document}
-
-%\slugcomment{Draft Modified \today}
-
-\title{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}
-
-\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1, 3}, Ruth A. Murray-Clay\altaffilmark{2}}
-\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
-\altaffiltext{3}{Max Planck Institute for Astronomy, K\"onigstuhl 17, 69117 Heidelberg, Germany}
-
-\begin{abstract}
+%\begin{savequote}[75mm]
+%Nulla facilisi. In vel sem. Morbi id urna in diam dignissim feugiat. Proin molestie tortor eu velit. Aliquam erat volutpat. Nullam ultrices, diam tempus vulputate egestas, eros pede varius leo.
+%\qauthor{Quoteauthor Lastname}
+%\end{savequote}
+
+\chapter{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}
+
+\section*{abstract}
 %The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry. 
@@
 We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.
-\end{abstract}
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{drift_timescales_betaS1_gas_acc_new2.pdf}
+\includegraphics[width=0.5\textwidth]{figures/drift_timescales_betaS1_gas_acc_new2.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{desorption_distance_passive_active_colorbar_test2.pdf}
+\includegraphics[width=0.7\textwidth]{figures/desorption_distance_passive_active_colorbar_test2.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{desorption_distance_actual_vs_estimated_passive_active_new.pdf}
+\includegraphics[width=0.7\textwidth]{figures/desorption_distance_actual_vs_estimated_passive_active_new.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{s_t_a.pdf}
+\includegraphics[width=0.8\textwidth]{figures/s_t_a.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
+\includegraphics[width=0.55\textwidth]{figures/C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=\textwidth]{time_plots.pdf}
+\includegraphics[width=\textwidth]{figures/time_plots.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{desorption_distance_varying_Md.pdf}
+\includegraphics[width=0.5\textwidth]{figures/desorption_distance_varying_Md.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{desorption_distance_transition_disk_1000.pdf}
+\includegraphics[width=0.5\textwidth]{figures/desorption_distance_transition_disk_1000.pdf}
 %\vspace{-0.5in}
@@
 %Nevertheless,
-
-\acknowledgements{We thank the referee for a helpful and detailed report.
-We also thank Sean Andrews for useful conversations and suggestions.
-T.
-B.
-is grateful for support from the NASA Origins of Solar Systems grant NNX12AJ04G and the Smithsonian Institution Pell Grant program.}
-
-\appendix
-\section{Steady-state active disk solution} \label{app:steadystate}
-
-Following \citet{shakura73} and \citet{armitage10}, the steady-state solution for a geometrically thin, optically thick actively accreting disk with an $\alpha$-prescription for viscosity is governed by the following set of equations:
-
-\begin{subeqnarray}
-\label{eq:diskeq}
-\nu & = & \alpha c H \slabel{eq:nueq} \\
-c^2 & = & \frac{k_{\rm B} T_{\rm act}}{\mu m_{\rm p}} \slabel{eq:cdeq} \\
-\rho & = & \frac{1}{\sqrt{2 \pi}} \frac{\Sigma}{H} \slabel{eq:rhoeq} \\
-H & = & \frac{c}{\Omega_{k}} \slabel{eq:Heq} \\
-T_{\rm act}^4 & = & \frac{3}{4} \tau T_{\rm surf}^4 \slabel{eq:Teq} \\
-\tau & = & \frac{1}{2} \Sigma \kappa \slabel{eq:taueq} \\
-\nu \Sigma & = & \frac{\dot{M}}{3 \pi} \slabel{eq:Mdot} \\
-\sigma T_{\rm surf}^4 & = & \frac{9}{8} \nu \Sigma \Omega_{\rm k}^2 \slabel{eq:nusigeq} \\
-\kappa & = & \kappa_0 T_{\rm act}^2 \slabel{eq:kappaeq},
-\end{subeqnarray}
-where $T_{\rm surf}$ is the surface temperature of the disk and the other quantities are defined in the main text.
-This is a system of nine equations with nine unknowns ($\nu$, $c$, $H$, $T_{\rm act}$, $\rho$, $\Sigma$, $\tau$, $T_{\rm surf}$, $\kappa$) that can be solved numerically once $\alpha$ and $\kappa_0$ are specified.
-
-\section{Desorption distance analytic solution} \label{app:tdriftan}
-
-For a particle of size $s$ that desorbs and satisfies $\tau_{\rm s} \ll 1$ ($\tau_{\rm s}$ is the dimensionless stopping time, defined in Section \ref{sec:drift}), we can derive an explicit analytic solution for the particle's desorption distance in an irradiated disk.
-For $\tau_{\rm s} \ll 1$, a particle is in the Epstein drag regime (see Equation \ref{eq:ts}) and its drift velocity $\dot{r}$ (Equation \ref{eq:rdotpas}) can be approximated as
-\begin{equation}
-\label{eq:rdotapp}
-\dot{r} \approx -2 \eta \Omega_{\rm k} r \tau_{\rm s}.
-\end{equation}
-By using Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and setting $t_{\rm drift}=t_{\rm des}$, we can express a particle's desorption distance as
-
-\begin{equation}
-\label{eq:tan}
-r_{\rm des}=\Bigg(\frac{d}{q \,C} \,\, \mathcal{W}\Big[\frac{(B/A)^{-q/d} q\,C}{d}\Big]\Bigg)^{\frac{1}{q}},
-\end{equation}
-where $\mathcal{W}$ is the Lambert-W function, $q=3/7$ is the power-law coefficient in Equation (\ref{eq:diskT}), $d=-\frac{1}{2}+p-q$ with $p=1$ the power-law coefficient in Equation (\ref{eq:disksigma}), and
-\begin{subeqnarray}
-%\label{eq:tan}
-A & = &  \frac{\rho_0}{\rho_{\rm s}} \frac{r_0^2}{s c_0} r_0^d \\ 
-B & = & \frac{\rho_{\rm s} s}{3 \mu_{\rm x} N_{\rm x} \nu_{\rm x}} \\
-C & = & \frac{E_{\rm x}}{T_0} r_0^{-q}, \\
-\end{subeqnarray}
-where $r_0=1$ AU, $\rho_0=\rho(r_0)$ and $c_0=c(r_0)$.
-
-%
-%\emgr{Right now it's unclear to me what could go in an appendix, if anything.
-Maybe discuss a bit the algorithm to evolve $\Sigma_{\rm p}$ (although it is already explained in detail in the appendix of Birnstiel et al.
-2010).
-Maybe show some example profiles of $\Sigma_{\rm p}$ at different times and for different particle sizes.}
-
-\if\bibinc n
-\bibliography{refs}
-\fi
-
-\if\bibinc y
-\begin{thebibliography}
-\end{thebibliography}
-\fi
-
-\end{document}

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/chapter4.tex
upd

--- 2016-03-11 tex/draft2/ms_edited.tex
@@
-%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
-%\documentclass[12pt, preprint]{aastex}
-%\documentclass[dvips,12pt]{article}
-\documentclass[apj]{emulateapj}
+\chapter{The Role of Ice Compositions and Morphology for Snowlines and the C/N/O Ratios in Active Disks}
 
-\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
-\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex
-
-%\usepackage{pdfsync}
-\usepackage{subeqnarray}
-\usepackage{natbib}
-\usepackage{color}
-\usepackage[utf8]{inputenc}
-
-\bibliographystyle{apj}
-
-\newcommand{\ie}{i.e.\ }
-\newcommand{\eg}{e.g.\ }
-\newcommand{\p}{\partial}
-\newcommand{\brak}[1]{\langle #1\rangle}
-
-\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
-\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
-\newcommand{\cm}{\; {\rm cm}}
-\newcommand{\mm}{\; {\rm mm}}
-%\newcommand{\ps}{\; {\rm s^{-1}}}
-\newcommand{\km}{\; {\rm km}}
-\newcommand{\au}{\; \varpi_{\rm AU}}
-\newcommand{\AU}{\; {\rm AU}}
-\def\K{\; {\rm K}}
-
-\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
-\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
-\newcommand{\nab}{\vc{\nabla}}
-\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
-\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
-\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}
-
-\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
-\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
-\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
-\newcommand{\App}[1]{Appendix~\ref{#1}}
-\newcommand{\Sec}[1]{Sect.~\ref{#1}}
-\newcommand{\Chap}[1]{Chapter~\ref{#1}}
-\newcommand{\Fig}[1]{Fig.~\ref{#1}}
-\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
-\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
-\newcommand{\Tab}[1]{Table \ref{#1}}
-
-\definecolor{gray}{gray}{0.5}
-\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}
-
-%\newenvironment{packed_item}{
-%\begin{itemize}
-%  \setlength{\itemsep}{1pt}
-%  \setlength{\parskip}{0pt}
-%  \setlength{\parsep}{0pt}
-%}{\end{itemize}}
-
-\begin{document}
-
-%\slugcomment{Draft Modified \today}
-
-\title{The Role of Ice Compositions and Morphology for Snowlines and the C/N/O Ratios in Active Disks}
-
-\author{Ana-Maria A. Piso\altaffilmark{1}, Jamila Pegues\altaffilmark{2}, Karin I. \"Oberg\altaffilmark{1}}
-\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-\altaffiltext{2}{Department of Astrophysical Sciences, Princeton University, Princeton, NJ 08544}
-
-\begin{abstract}
+\section*{abstract}
 The elemental compositions of planets define their chemistry, and could potentially be used as beacons for their formation location if the elemental gas and grain ratios of planet birth environments, i.e.
@@
 I will discuss these results together with the effects of additional dynamical processes, and outline a path toward a coupled drift-desorption-chemistry model that will provide robust quantitative results for volatile snowline locations and C/N/O abundance ratios as the disk evolves in time.
-\end{abstract}
 
@@
 \centering
-\includegraphics[width=0.95\textwidth]{CNO_and_snowlines_2.eps}
+\includegraphics[width=0.95\textwidth]{figures/CNO_and_snowlines_2.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{C_O_ratio_CH4.eps}
+\includegraphics[width=0.5\textwidth]{figures/C_O_ratio_CH4.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{N_O_ratio.eps}
+\includegraphics[width=0.5\textwidth]{figures/N_O_ratio.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{C_O_water_ice.eps}
+\includegraphics[width=0.5\textwidth]{figures/C_O_water_ice.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{N_O_water_ice_many.eps}
+\includegraphics[width=0.5\textwidth]{figures/N_O_water_ice_many.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{CN_norm.eps}
+\includegraphics[width=0.5\textwidth]{figures/CN_norm.pdf}
 %\vspace{-0.5in}
@@
 %In future work, we plan to add new levels of complexity to our model in terms of disk chemistry, dynamics, and planetary dynamics, thus forming a solid framework for understanding the origins of gas giants.
-
-\acknowledgments{This work is supported by a Simons Collaboration on the Origins of Life (SCOL) investigator award to KIO. JP was supported by the Banneker institute.}
-
-%\emgr{Maybe we can include the summary in the discussion section?}
-
-%\section{Volatile Abundances and Binding Energies: Effect on C/O and N/O Ratios in a Static Disk}
-%
-%\subsection{Nitrogen and CH4}
-%
-%\emgr{Discuss that nitrogen is abundant in the solar system and should be abundant in disks as well, but that its dominant form is largely unknown.
-Discuss that the main carrier of nitrogen is N2, but some fraction of it can be NH3 as well.
-Present and motivate the choices that we make for NH3 abundances.
-Along the same lines, discuss CH4, and the choices that we make for CH4 abundances.}
-%
-%\subsection{Volatile Desorption Energies}
-%
-%\emgr{State that desorption energies for H2O and CO2 are well constrained experimentally, and that the CO2 binding energy is only weakly dependent on whether it's pure CO2 or combined with H2O, but that is not the case for CO, N2 (and perhaps CH4 and NH3?).
-Briefly discuss CO-CO, N2-N2, CO-H2O and N2-H2O (and perhaps the same for CH4 and NH3 if we find literature on that) experimental results for binding energies, and motivate the choices that we make.}
-%
-%\subsection{Results for C/O and N/O in a Static Disk}
-%
-%\emgr{For each of them (C/O and N/O), show a 3-panel plot as follows: each panel has a specific CH4/NH3 abundance (top: none, middle: median abundance, bottom: maximum abundance); for a given panel, have multiple curves for C/O or N/O, depending on the choice of binding energies, so that it's clear visually how the binding energy changes the snowline location.
-Discuss how different abundances  and binding energies affect snowline locations and C/O or N/O ratios.}
-%\section{Results}
-%
-%\subsection{N2, NH3 and CH4 Snowline Locations}
-%
-%\emgr{One multipanel 3x3 (or 3x2) rainbow plot similar to the snowline plots from Paper I, for \textit{one} choice for the binding energies (perhaps the largest ones, since we want a limit on how far in we can push the snowlines?).
-Rows: snowlines as a function or particle size for passive, active and (maybe) steady-state disk.
-Columns: N2, NH3 and CH4.
-Not entirely certain how necessary these plots/subsection actually are, since it is exactly the same qualitatively as in Paper I...}
-%
-%\subsection{C/O and N/O Ratios}
-%
-%\emgr{For each of them (C/O and N/O) show a 3x3 multipanel plot similar to the C/O plot from Paper I, for the same choice of binding energies as in the previous subsection.
-Columns: C/O or N/O in passive, active and steady-state disk.
-Rows: No CH4/NH3, median CH4/NH3, maximum CH4/NH3.
-Again, not sure if we need/want this for all disk choices, at least in the case of C/O since we already have that in Paper I.
-For N/O, quantify how the snowline location changes from a static disk due to drift and gas accretion.}
-
-\if\bibinc n
-\bibliography{refs}
-\fi
-
-\if\bibinc y
-\begin{thebibliography}
-\end{thebibliography}
-\fi
-
-\end{document}

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/introduction.tex
upd

\chapter{Introduction}
\label{introduction}

Within the last two decades, more than one thousand extrasolar planets (exoplanets)
have been discovered [1]. Their diversity in terms of mass, radius, location and composition
[2] provides an exciting field of research, with the eventual goal of finding planets that are
similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore
and understand how planets obtain their compositions. Observations of Earth-like planets
that can provide useful insight about their composition are challenging — the solid interior
structure of terrestrial planets cannot be detected, and their gaseous envelopes are small by
comparison (both in mass and radius), which makes it difficult to obtain atmospheric spectra
and find out what chemical compounds they are made of. We therefore turn to giant planets,
which have provided a rich and intriguing research area for decades. Gas giants contain most
of their mass in their atmosphere, hence their chemical composition is determined by that
of their envelopes. The last few years have seen a substantial increase in the number of
giant planets with observed atmospheric spectra (e.g., [3], [4]), which has enhanced our
understanding of these planets’ chemical structure, and has provided us with quantitative
information about the abundances of various compounds in their envelopes besides hydrogen
and helium. Finally, gas giants shape the architecture of planetary systems and affect the
delivery of volatile compunds to terrestrial planets, which has direct consequences for the
habitability of worlds similar to our own. Thus testing theories of planet formation against
gas giant compositions will help constrain planet formation theories more generally.
Both terrestrial and giant planets are born in protoplanetary disks, which implies that
their compositions are determined by and tightly linked to the structure and
composition of the disk. The chemical and dynamical evolution of disks, as well the
formation of giant planets have both been previously investigated in isolation. However, the
coupled chemo-dynamical disk evolution, planet compositions, and most importantly the
disk-planet connection have not yet been considered in detail. In this thesis, I uncover some of the answers to this issue from two standpoints: (1) by looking at the role of disk location in setting the conditions for the formation of wide-separation gas giants, and (2) by investigating how the structure and chemical composition of the protoplanetary disk at different radii affects the composition of nascent giant planets.

\subsection{The Role of Disk Location in Setting the Minimum Core Mass for Giant Planet formation}

\begin{itemize}

\item Gas giants are largely believed to form through core accretion (insert core accretion sketch from various talks). This process is particularly challenging in the outer disk due to long dynamical timescales. At the same time, wide separation gas giants have been discovered (HR 8799 plot). This poses an intriguing question: how do these planets form: I answer this question in the first part of my thesis; specifically I calculate the minimum Mcrit, which applies when cores no longer accrete solids. Standard core accretion studies are not built to properly explore this Mcrit. Here explain standard core accretion studies in more detail and insert e.g. one of the Mass vs time figures from Pollack et al. 1996. Explain how our study is different. Finalize by saying that the studies I performed clearly challenge previous claims that  core accretion cannot operate at wide separations, thus reopening the case for in-situ formation of wide separation gas giants.

\end{itemize}

\subsection{The Role of Disk Dynamics and Morphology in Setting Snowline locations and C/N/O Ratios}

\begin{itemize}

\item Disk Composition Regulates Planet Composition. It is thus essential to (1) predict what kinds of planet compositions result from planet location in different parts of the disk, and (2) conversely, bacl-track planet formation location based on planet composition.

\item Disks are complex. Show figure from Henning and Semenov with all processes that occur in disks and discuss which ones I tackle in the thesis.

\item However, we do know some things about disks. Volatile molecules have detected in disks (show figure with Spitzer IR spectrum of AA Tauri). Snowlines have also been detected (show CO snowline plot from Qi+13).

\item One important signature of atmospheric chemistry is the C/O ratio (explain why and show plot from Molliere+15 which shows the effect of varying C/O ratio on the abundance of other volatiles). Discuss claims that super stellar C/O ratios have been detected (show Wasp 12-b spectrum from Madhusudhan), but have since been refuted. Explain that this poses an intriguing question from a theoretical standpoint. Mention idea proposed by Oberg+11 (show plot). Discuss that disk dynamics have an important role, which is partly what we tackle in this part of the thesis.

\item There are other important volatiles besides H2O, CO2 and CO. Discuss the importance of N2 and show plot with abundances in comets that show that NH3 is the main carrier based on observations, but it doesn't account for all the nitrogen in the solar system.

\item Ice morphology is important. Discuss why. Discuss how binding energies depend on the ice environment and show plot from Edit's paper.

\item It follows that we have to take into account additional volatiles, abundances and ice morphologies besides disk dynamics to see their effect on snowline locations and the C/N/O ratios. We tackle this in Chaper xx.

\item We demonstrate in this part to the thesis that C/N/O ratios may be used to track a planet's formation origin, when combined with observations, but that more observations are needed.

\end{itemize}

\subsection{Disk-Planet Connection}

Our work in the two main parts of this thesis confirms that there is indeed a tight link between disks and planets. More importantly, as we will demonstrate in detail, the formation and composition of giant planets highly depends on disk location and properties.

This thesis is organized as follows....

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/conclusion.tex
upd

\chapter{Summary and Future Directions}
\label{conclusion}

In this thesis, we have calculated the minimum core mass required to form wide-sparation gas giants at different disk locations, and for different equations of state of the nebular gas and dust opacities. We have also determined how disk dynamics and volatile abundance and morphology affect volatile snowline locations, the C/N/O ratios across the disk, and therefore the compositions of nascent giant planets. Our results are summarized in Sections \ref{sec:core} and \ref{sec:volatile}, and we present directions for future work in Section \ref{sec:future}.

\section{Minimum Core Masses for Giant Planet Formation}
\label{sec:core}

In Chapters 2 and 3, we determined the minimum core mass $M_{\rm crit}$ required to form a gas giant at distances between 5 and 100 AU from the host star, motivated by the fact that standard core accretion models cannot explain the formation of giant planets in the outer disk. This minimum applies when the solid cores are no longer accreting solids, since any additional heating due to planetesimal accretion would heat up the core's atmosphere, inhibit its ability to cool and contract, and therefore increase the critical core mass. We thus considered atmospheres accreting around fully formed cores and undergoing Kelvin-Helmholtz contraction. In Chapter 2, we developed a quasi-static atmospheric evolution model by generating a series of static atmospheric profiles that are then connected temporally through a cooling equation. We considered an envelope structure consisting of an inner convective region and an outer radiative layer, and assumed a constant luminosity in the outer radiative region. We demonstrated that this is a valid approximation for our parameter space of interest. For an ideal gas polytrope and standard interstellar (ISM) opacities, we have found that $M_{\rm crit}$ decreases with semimajor axis from $\sim 8.5 M_{\oplus}$ at 5 AU to $\sim 3.5 M_{\oplus}$ at 100 AU. Our results are lower than the typically quoted value of $M_{\rm crit} \sim 10 M_{\oplus}$ (e.g., \citealt{rafikov06}), even in the more inner parts of the disk.

To obtain more robust quantitative results, in Chapter 3 we expanded the model developed in Chapter 2 by considering a realistic equation of state (EOS) for the nebular gas and realistic dust opacities that take into account grain growth. We parametrized the EOS through the adiabatic gradient $\delad$ (defined in Chapter 2), as $\delad$ relates the gas pressure, temperature and density. While for an ideal gas EOS $\delad$ is constant, non-ideal EOS effects cause features in the adiabatic gradient that change the atmospheric structure. At high temperatures, molecular hydrogen dissociates, while at low-temperature the rotational states of the hydrogen molecule are only partially excited so it no longer behaves like an ideal gas, due to the existence of H$_2$ in two spin isomeric forms, ortho- and parahydrogen. Both of these effects increase $M_{\rm crit}$ by a factor of $\sim$2 compared to the ideal gas. In contrast, grain growth opacities decrease $M_{\rm crit}$ dramatically. By taking these two competing effects together, we calculate $M_{\rm crit} \sim 8 M_{\oplus}$ at 5 AU, decreasing to $\sim 5 M_{\oplus}$ at 100 AU. While our atmospheric model is not equipped to calculate the critical core mass when grain coagulation is taken into account, we have demonstrated that grain growth with coagulation may decrease $M_{\rm crit}$ by up to one order of magnitude; $M_{\rm crit}$ may be as low as 1 $M_{\oplus}$. Our study thus clearly challenges previous claims that core accretion cannot operate in the outer disk, reopening the case for in situ formation of wide separation gas giants.

\section{The Role of Disk Dynamics and Ice Morphology on Snowline Locations and the C/N/O Ratios}
\label{sec:volatile}

In Chapters 4 and 5, we determined the effect of disk dynamics and ice morphology on the C/N/O ratio in active disks, which has direct implications on gas giant composition. This study was motivated by the fact that the locations of volatile snowlines in protoplanetary disks are a defining feature of both gas giant and disk chemistry, as they provide vital information about the abundance of these molecules in gas and dust throughout the disk. In Chapter 4, we expanded the model of \citet{oberg11} by considering the effect of disk dynamics on the snowline locations of the main carbon and oxygen carriers, i.e. H$_2$O, CO$_2$ and CO, and thus on the C/O ratio in dust and gas throughout the disk. We studied the effect of radial drift of solids and viscous gas accretion onto the central star on snowline locations by developing a semi-analytical model, which we applied to three different disks in increasing order of complexity. We applied our model for a range of initial particle sizes and calculated how drift and gas accretion affect their desorption location. We have found that there is a range of particle sizes, $0.05$ cm $\lesssim s \lesssim$ 7 m for our particular parameters, that desorb almost instantaneously and a fixed particle-size dependent from the central star, regardless of their initial location. Based on this information, we calculated the H$_2$O, CO$_2$ and CO snowline locations for different particle sizes in this range, and made estimates for the C/O ratio throughout the disk. We have found that the snowlines move inward as the particle size increases, and may differ by up to a factor of $\sim$2 due to drift and gas accretion compared to a static disk, which does not experience any dynamical processes. This variation in snowline locations in significant, and has important consequences for the compositions of gas giants forming in situ.

In Chapter 5, we expanded the model developed in Chapter 4 by considering additional volatiles, chemical abundances, and ice morphologies. As nitrogen is highly abundant in the Solar System and primarily found as N$_2$, we added nitrogen bearing species such as N$_2$ and NH$_3$ to our model, as well as hydrocarbons (specifically, CH$_4$). Motivated by laboratory experiments (e.g., \citealt{fayolle16}) that find significantly different binding energies for CO and N$_2$ depending on the ice environment in which they reside (pure versus water dominated ices), we calculated the CO and N$_2$ snowlines for both scenarios. We calculated the N/O ratio in static disks and found that it is highly enhanced compared to the stellar value: by a factor of $\sim$2 between the H$_2$O and CO$_2$ snowlines, by more than a factor of 3 between the CO$_2$ and CO snowlines, and by many orders of magnitude between the CO and N$_2$ snowlines, where oxygen gas is depleted. Thus we expect wide-separation giants to have an excess of nitrogen in their atmospheres, which may be used to trace their formation origins. We have also found that the binding environment has a large effect on the CO and N$_2$ snowline locations and may change them by factor of 3-4, both for static and dynamic disks. Thus by considering the combined effect of disk dynamics and ice morphology, the CO and N$_2$ snowline locations may change by up to a factor of $\sim$7, and may span 11-79 AU for N$_2$ in our disk model. This large uncertainty in snowline locations means that observations that anchor snowline locations at different stages of planet
formation are key to develop C/N/O ratios as a probe of planet formation zones.

\section{Future Directions}
\label{sec:future}

Our work has made important strides in our understudying of disk and planet compositions and the tight link between the two, as well as the effect of disk dynamics and chemistry on disk and planet compositions. As first step, we plan to expand our work by considering the effect of diffusion on the C/N/O ratios in viscous steady-state disks. Similarly to the model of \citet{owen14}, we will develop a simplified method to estimate the abundance of different volatiles at various disk locations, and thus their effect on the C/N/O ratio both in static and dynamic disks. We plan to explore further processes that affect snowline locations, such as variations in stelar luminosity and gas mass accretion rate, as well as particle growth and fragmentation. Our next step is to incorporate disk chemical processes, first by developing a simplified parametrized chemical  network that we will incorporate in our  drift-desorption model. By having a better understudying on how disk chemistry and dynamics affect the composition of nascent, and eventually mature planets, our work may provide essential context for characterizing the gas giants
that instruments such as JWST and the TESS will one day discover.

### 2016-04-13 intro.tex
upd

--- 2016-04-13 Dissertate-Harvard-LaTeX/chapters/introduction.tex
@@
-\chapter{Introduction}
-\label{introduction}
+\documentclass[apj, numberedappendix]{emulateapj}
 
+\usepackage{graphicx} 
+\usepackage{amsmath}
+\usepackage{hyperref}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{subeqnarray}
+\usepackage{epstopdf}
+%\usepackage{natbib}
+%\bibliographystyle{apj}
+
+\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}
+
+\newcommand{\ie}{i.e.\ }
+\newcommand{\eg}{e.g.\ }
+\newcommand{\p}{\partial}
+\newcommand{\xv}{\vc{x}}
+\newcommand{\kv}{\vc{k}}
+\newcommand{\brak}[1]{\langle #1\rangle}
+
+\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
+\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
+\newcommand{\cm}{\; {\rm cm}}
+\newcommand{\mm}{\; {\rm mm}}
+%\newcommand{\ps}{\; {\rm s^{-1}}}
+\newcommand{\km}{\; {\rm km}}
+
+\newcommand{\AU}{\; {\rm AU}}
+\newcommand{\yr}{\; {\rm yr}}
+\def\K{\; {\rm K}}
+
+\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
+\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
+\newcommand{\nab}{\vc{\nabla}}
+\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
+\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
+\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}
+
+\newcommand{\eq}[1]{equation\,(\ref{#1})}
+\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
+\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
+\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
+\newcommand{\Eqsss}[3]{Equations (\ref{#1}), (\ref{#2}) and~(\ref{#3})}
+\newcommand{\App}[1]{Appendix~\ref{#1}}
+\newcommand{\Sec}[1]{Sect.~\ref{#1}}
+\newcommand{\Chap}[1]{Chapter~\ref{#1}}
+\newcommand{\Fig}[1]{Figure~\ref{#1}}
+\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
+\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
+\newcommand{\Tab}[1]{Table \ref{#1}}
+
+\newenvironment{packed_item}{
+\begin{itemize}
+  \setlength{\itemsep}{1pt}
+  \setlength{\parskip}{0pt}
+  \setlength{\parsep}{0pt}
+}{\end{itemize}}
+
+\newcommand{\delad}{\nabla_{\rm ad}}
+\newcommand{\delrad}{\nabla_{\rm rad}}
+\newcommand{\Rg}{\mathcal{R}}
+\newcommand{\RB}{R_{\rm B}}
+\newcommand{\RH}{R_{\rm H}}
+\newcommand{\co}{_{\rm c}}
+\newcommand{\pla}{_{\rm p}}
+\newcommand{\di}{_{\rm d}}
+\newcommand{\cb}{_{\rm RCB}}
+\newcommand{\mc}{m_{\rm c \oplus}}
+\newcommand{\mcn}[1] { m_{ \rm c #1} }
+\newcommand{\mpn}[1] { m_{ \rm p #1} }
+\newcommand{\MC}{M_{\rm crit}}
+\newcommand{\au}{a_\oplus}
+\newcommand{\aun}[1]{ a_{#1} }
+\def\crit{_{\rm{c, crit}}}
+
+\begin{document}
+
+\section{Introduction}
+
+Two big questions. In the context of core accretion:
+- where in the disk can you form Gas Giants?
+- what compositions will the formed Gas Giants obtained?
+
+we can learn about this through a combination of studying planet formation in protoplanetary disks, and the end-results of planet formation, i.e.
+exo-planets and the architecture and composition of our own Solar System
+
+1. Protoplanetary disks
+1a. Density and temperature structures
+1b. Chemical processes and structures
+
+2. Observational constraints on Gas Giant formation and compositions
+
+3. Core accretion formation of planets and its challenges
+
+4. Interactions between disk physics and chemistry, and core accretion
+
+5. Connection between disk compositions and the compositions of nascent planets
+
+In recent years, the field of giant planet formation has aimed to answer two fundamental questions:
+
+\begin{enumerate}
+\item Where in the protoplanetary disk can gas giants form?
+\item What compositions will the formed giant planets have obtained?
+\end{enumerate}
+
+We can start uncovering the answers to these questions through a combination of studying planet formation in protoplanetary disks, the end-results of planet formation (i.e., exoplanets), and through comparisons with the architecture and composition of our own Solar System.
 Within the last two decades, more than one thousand extrasolar planets (exoplanets)
-have been discovered [1]. Their diversity in terms of mass, radius, location and composition
-[2] provides an exciting field of research, with the eventual goal of finding planets that are
+have been discovered \citet{batahla14}. Their diversity in terms of mass, radius, location and composition
+\citet{lissauer14} provides an exciting field of research, with the eventual goal of finding planets that are
 similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore
@@
 of their envelopes. The last few years have seen a substantial increase in the number of
-giant planets with observed atmospheric spectra (e.g., [3], [4]), which has enhanced our
+giant planets with observed atmospheric spectra (e.g., \citealt{debes13}, \citealt{kreidberg15}), which has enhanced our
 understanding of these planets’ chemical structure, and has provided us with quantitative
@@
 gas giant compositions will help constrain planet formation theories more generally.
+
 Both terrestrial and giant planets are born in protoplanetary disks, which implies that
@@
 In this thesis, I uncover some of the answers to this issue from two standpoints: (1) by looking at the role of disk location in setting the conditions for the formation of wide-separation gas giants, and (2) by investigating how the structure and chemical composition of the protoplanetary disk at different radii affects the composition of nascent giant planets.
+
+\subsection{Protoplanetary Disks}
+
+Chemical and dynamical processes in a protoplanetary disk affect the disk structure
+and composition, and thus the composition of nascent planets.
+Figure \ref{fig:disk_henning} shows various dynamical and chemical processes that occur in disks, many of which are not yet fully understood.
+Moreover, the timescales for various disk chemical and dynamical processes may be comparable at
+least in some parts of the disk (see Figure \ref{fig:timescales_semenov}).
 
@@
 This thesis is organized as follows....
+
+\end{document}

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/conclusion.tex
upd

@@
 
-In this thesis, we have calculated the minimum core mass required to form wide-sparation gas giants at different disk locations, and for different equations of state of the nebular gas and dust opacities.
+In this thesis, we have calculated the minimum core mass required to form wide-separation gas giants at different disk locations, and for different equations of state of the nebular gas and dust opacities.
 We have also determined how disk dynamics and volatile abundance and morphology affect volatile snowline locations, the C/N/O ratios across the disk, and therefore the compositions of nascent giant planets.
@@
 
-In Chapters 4 and 5, we determined the effect of disk dynamics and ice morphology on the C/N/O ratio in active disks, which has direct implications on gas giant composition.
+In Chapters 4 and 5, we determined the effect of disk dynamics and ice morphology on the C/N/O ratio in active disks, which has direct implications on gas giant compositions.
 This study was motivated by the fact that the locations of volatile snowlines in protoplanetary disks are a defining feature of both gas giant and disk chemistry, as they provide vital information about the abundance of these molecules in gas and dust throughout the disk.
@@
 We applied our model for a range of initial particle sizes and calculated how drift and gas accretion affect their desorption location.
-We have found that there is a range of particle sizes, $0.05$ cm $\lesssim s \lesssim$ 7 m for our particular parameters, that desorb almost instantaneously and a fixed particle-size dependent from the central star, regardless of their initial location.
+We have found that there is a range of particle sizes, $0.05$ cm $\lesssim s \lesssim$ 7 m for our particular parameters, that desorb almost instantaneously and a fixed particle-size dependent location from the central star, regardless of their initial position.
 Based on this information, we calculated the H$_2$O, CO$_2$ and CO snowline locations for different particle sizes in this range, and made estimates for the C/O ratio throughout the disk.
 We have found that the snowlines move inward as the particle size increases, and may differ by up to a factor of $\sim$2 due to drift and gas accretion compared to a static disk, which does not experience any dynamical processes.
-This variation in snowline locations in significant, and has important consequences for the compositions of gas giants forming in situ.
+This variation in snowline locations is significant, and has important consequences for the compositions of gas giants forming in situ.
 
@@
 Thus we expect wide-separation giants to have an excess of nitrogen in their atmospheres, which may be used to trace their formation origins.
-We have also found that the binding environment has a large effect on the CO and N$_2$ snowline locations and may change them by factor of 3-4, both for static and dynamic disks.
+We have also found that the binding environment has a large effect on the CO and N$_2$ snowline locations and may change them by factors of 3-4, both for static and dynamic disks.
 Thus by considering the combined effect of disk dynamics and ice morphology, the CO and N$_2$ snowline locations may change by up to a factor of $\sim$7, and may span 11-79 AU for N$_2$ in our disk model.

### 2016-04-13 Dissertate-Harvard-LaTeX/frontmatter/abstract.tex
upd

@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Morbi commodo, ipsum sed pharetra gravida, orci magna rhoncus neque, id pulvinar odio lorem non turpis.
-Nullam sit amet enim.
-Suspendisse id velit vitae ligula volutpat condimentum.
-Aliquam erat volutpat.
-Sed quis velit.
-Nulla facilisi.
-Nulla libero.
-Vivamus pharetra posuere sapien.
-Nam consectetuer.
-Sed aliquam, nunc eget euismod ullamcorper, lectus nunc ullamcorper orci, fermentum bibendum enim nibh eget ipsum.
-Donec porttitor ligula eu dolor.
-Maecenas vitae nulla consequat libero cursus venenatis.
-Nam magna enim, accumsan eu, blandit sed, blandit a, eros.
-
-Quisque facilisis erat a dui.
-Nam malesuada ornare dolor.
-Cras gravida, diam sit amet rhoncus ornare, erat elit consectetuer erat, id egestas pede nibh eget odio.
-Proin tincidunt, velit vel porta elementum, magna diam molestie sapien, non aliquet massa pede eu diam.
-Aliquam iaculis.
-Fusce et ipsum et nulla tristique facilisis.
-Donec eget sem sit amet ligula viverra gravida.
-Etiam vehicula urna vel turpis.
-Suspendisse sagittis ante a urna.
-Morbi a est quis orci consequat rutrum.
-Nullam egestas feugiat felis.
-Integer adipiscing semper ligula.
-Nunc molestie, nisl sit amet cursus convallis, sapien lectus pretium metus, vitae pretium enim wisi id lectus.
-Donec vestibulum.
-Etiam vel nibh.
-Nulla facilisi.
-Mauris pharetra.
-Donec augue.
-Fusce ultrices, neque id dignissim ultrices, tellus mauris dictum elit, vel lacinia enim metus eu nunc.
+The composition of planets is determined by and tightly linked to the composition of the protoplanetary disk in which they form.
+In the first part of my thesis, I study giant planet formation through core accretion.
+I show how the minimum core mass required to form a giant planet during the lifetime of the protoplanetary disk depends on the location in the disk, the equation of state of the nebular gas and dust opacity.
+This minimum applies when planetesimal accretion does not substantially heat the core's atmosphere.
+The minimum core mass decreases with semimajor axis, and may be significantly lower than the typically quoted value of 10 $M_{\oplus}$, thus challenging previous studies that core accretion cannot operate in the outer disk.
+In the second part, I explore how the composition and evolution of protoplanetary disks may affect the formation and chemical composition of giant planets.
+As the C/N/O ratios are an important signature of giant planet atmospheres, I show how the snowline locations of the main carbon, oxygen and nitrogen carriers, as well as the C/N/O ratios, are affected by disk dynamics and ice morphology.
+Compared to a static disk, disk dynamics and ice morphology combined may change the CO and N$_2$ snowline locations by a factor of 7.
+Moreover, the gas-phase N/O ratio is highly enhanced throughout most of the disk, meaning that wide-separation giants should have an excess of nitrogen in their atmospheres which may be used to trace their origins.
+The large range of possible CO and N$_2$ snowline locations, and hence of regions with highly enhanced N/O ratios, implies that snowline observations at various stages of planet formation are crucial in order to use C/N/O ratios as beacons for planet formation zones.
+ %. This affects the C/O ratio in gas and dust throughout the disk, and thus has direct implications in shaping the composition of nascent giant planets.

### 2016-04-13 intro.tex
upd

@@
-\documentclass[apj, numberedappendix]{emulateapj}
+%\documentclass[apj, numberedappendix]{emulateapj}
+\documentclass[12pt, preprint]{aastex}
 
@@
 
-Two big questions. In the context of core accretion:
-- where in the disk can you form Gas Giants?
-- what compositions will the formed Gas Giants obtained?
-
-we can learn about this through a combination of studying planet formation in protoplanetary disks, and the end-results of planet formation, i.e.
-exo-planets and the architecture and composition of our own Solar System
-
-1. Protoplanetary disks
-1a. Density and temperature structures
-1b. Chemical processes and structures
-
-2. Observational constraints on Gas Giant formation and compositions
-
-3. Core accretion formation of planets and its challenges
-
-4. Interactions between disk physics and chemistry, and core accretion
-
-5. Connection between disk compositions and the compositions of nascent planets
-
 In recent years, the field of giant planet formation has aimed to answer two fundamental questions:
@@
 Within the last two decades, more than one thousand extrasolar planets (exoplanets)
-have been discovered \citet{batahla14}. Their diversity in terms of mass, radius, location and composition
+have been discovered \citet{batalha14}. Their diversity in terms of mass, radius, location and composition
 \citet{lissauer14} provides an exciting field of research, with the eventual goal of finding planets that are
@@
 disk-planet connection have not yet been considered in detail.
-In this thesis, I uncover some of the answers to this issue from two standpoints: (1) by looking at the role of disk location in setting the conditions for the formation of wide-separation gas giants, and (2) by investigating how the structure and chemical composition of the protoplanetary disk at different radii affects the composition of nascent giant planets.
+In this thesis, we uncover some of the answers to this issue from two standpoints: (1) by looking at the role of disk location in setting the conditions for the formation of wide-separation gas giants, and (2) by investigating how the structure and chemical composition of the protoplanetary disk at different radii affects the composition of nascent giant planets.
 
-\subsection{Protoplanetary Disks}
+\section{Dynamics and Chemistry in Protoplanetary Disks}
 
-Chemical and dynamical processes in a protoplanetary disk affect the disk structure
-and composition, and thus the composition of nascent planets.
-Figure \ref{fig:disk_henning} shows various dynamical and chemical processes that occur in disks, many of which are not yet fully understood.
+Protoplanetary disks are the result of the collapse of a molecular cloud, and are structures composed of gas and dust that are rotationally supported and that represent the birth environment of planets (Figure \ref{fig:proto}).
+Disks are complex objects, whose structure and evolution are affected by a multitude of chemical and dynamical processes.
+Many of these processes are summarized in Figure \ref{fig:disk_henning}, and several of them are not fully understood.
 Moreover, the timescales for various disk chemical and dynamical processes may be comparable at
-least in some parts of the disk (see Figure \ref{fig:timescales_semenov}).
+least in some parts of the disk (see Figure \ref{fig:timescales_semenov}), which makes the coupling between chemistry and dynamics very challenging.
+As a first step it is thus easier to isolate the different dynamical and chemical processes that occur in the disk midplane where planets are born, and understand their role and relative importance in shaping disk and planet compositions.
+We explore some of these processes in Chapters 4 and 5, and investigate their consequences for the compositions of giant planets forming at different radii.
 
-\subsection{The Role of Disk Location in Setting the Minimum Core Mass for Giant Planet formation}
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/proto.png}
+%\vspace{-0.5in}
+\caption{The standard picture of molecular cloud to protostar to protoplanetary disk to planet formation. From Oberg et al. ?}
+\label{fig:proto}
+\end{figure}
 
-\begin{itemize}
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/disk.png}
+%\vspace{-0.5in}
+\caption{Cross section through a protoplanetary disk showing various chemical and dynamical processes that occur in disks. From \citet{henning13}.}
+\label{fig:disk_henning}
+\end{figure}
 
-\item Gas giants are largely believed to form through core accretion (insert core accretion sketch from various talks).
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/chemical_timescales.png}
+%\vspace{-0.5in}
+\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height.
+The numbered curves are $\log_10 \tau$ in million years.
+From \citet{semenov11}}
+\label{fig:timescales_semenov}
+\end{figure*}
+
+\subsection{Snowlines and Disk Dynamics}
+
+Volatile compounds, which have low sublimation temperatures, are of particular importance.
+Their relative abundance in gaseous and solid form determines their snowline locations, and they are essential in the planet formation process (e.g., \citealt{pontoppidan14}).
+Observational advancements in recent year have provided us with an increasing amount of information regarding volatiles in disks.
+As shown in Figure \ref{fig:spitzer}, volatile compounds have been detected in disks, which gives us clues about their abundance in different disks and at different disk locations.
+Moreover, with the advent of ALMA volatile snowlines have also been observed.
+This is essential as it provides us with knowledge regarding volatile abundances in gas and dust throughout the disk, and thus the compositions of giant planets.
+Figure \ref{fig:qi13} shows the detection of the CO snowline in TW Hya at 30 AU \citep{qi13}.
+This snowline has been detected indirectly through its tracer molecule N$_2$H$^+$, which becomes highly abundant as CO freezes out.
+
+While there are several volatiles in disks that can be further investigated, one important signature of atmospheric chemistry is the carbon-to-oxygen (C/O) ratio.
+This is demonstrated in Figure \ref{fig:molliere}, which shows a theoretical model of an exoplanet atmosphere: variations in the C/O ratio by factors of $\sim$3 change the abundance of other volatiles, such as H$_2$O and CH$_4$ by several orders of magnitude.
+From the observational perspective, an exciting potential discovery was that some planets, such as WASP-12 b, have a superstellar C/O ratio \citet{madhu11}.
+While this particular detection has since been unequivocally refuted (e.g., \citealt{krediberg15}), the advent of JWST in the near future will provide us with a significantly larger sample of atmospheric spectra which may be used to constrain C/O ratios in other giant planet atmospheres.
+The possibility of a superstellar C/O ratio in a disk or planet atmosphere is also intriguing from a theoretical standpoint.
+One likely explanation is that the main carbon and oxygen carriers, i.e.
+H$_2$O, CO$_2$ and CO, have different condensation temperatures.
+This will result in variations in the abundance of the volatiles in gas and solid form between their respective snowlines, which will subsequently change the C/O ratio at different disk radii.
+This idea was first postulated by \citet{oberg11}.
+Figure \ref{fig:oberg} shows the C/O ratio in gas and dust as a function of semimajor axis, assuming protostellar abundances for the volatiles.
+Indeed, the C/O ratio in gas is enhanced compared to the stellar value, particularly between the CO$_2$ and CO snowlines where it reaches unity.
+The pioneering work of \citet{oberg11} inspired Chapter 4 of this thesis.
+\citet{oberg11} consider a static disk, and thus do not account for dynamical processes such as redistribution of solids due to radial drift, the radial movement of the nebular gas, and accretion heating.
+We thus expand this model by considering the processes outlined above and their effect on snowline locations.
+In particular, we focus on how radial drift of solids and viscous gas accretion onto the central star affect the H$_2$O, CO$_2$ and CO snowline locations for particles of different sizes.
+We find that drift and accretion heating alone may move the snowlines inward by factors of $\sim$2 compared to a static disk, thus substantially changing the disk regions with enhanced gas-phase C/O ratios, which has direct consequences for planet formation.
+
+\subsection{Nitrogen and Ice Morphology}
+
+As we show above, the C/O ratio is highly important to study, but it only provides us with one piece of the puzzle that is constraining planet formation origins based on their compositions.
+In order to improve our theoretical models, we thus need to look at other molecules and volatile compounds.
+The first that comes to mind is nitrogen, as it highly abundant in the Solar System \citep{lodders03}, and primarily believed to exist as N$_2$ (e.g., \citealt{owen01}).
+Molecular nitrogen cannot be detected directly.
+However, an inventory of nitrogen cometary abundances shows that NH$_3$ is the most abundant observed nitrogen carrier, and yet its concentration is only a small fraction of the total nitrogen abundance, which leaves N$_2$ as the main nitrogen bearing species.
+In the context of planet compositions, N$_2$ is important due to its high volatility.
+The gas phase
+nitrogen-to-oxygen (N/O) ratio in the outer disk may thus be even more enhanced than the C/O ratio.
+Giant planets that form at wide separations should hence have an excess of nitrogen in their atmospheres,
+which could be used to trace their formation origin.
+We calculate the N/O ratio in gas and dust in Chapter 5, and confirm that it is indeed highly enhanced compared to the stellar value throughout most of the disk.
+
+The snowline locations of volatiles, including carbon, oxygen and nitrogen carriers, are highly dependent on the morphology of the icy particles.
+These can either form on a dust mantle as pure ices, or have a layered structure with the more volatile species closer to the outer layers (e.g., \citealt{pontoppidan03}).
+As H$_2$O is the least volatile species with high abundance in disks, one can typically expect the more volatile molecules to form on a water substrate (but see \citealt{bisschop06} for other scenarios).
+The ice environment in which the molecules reside, i.e.
+pure or water dominated ices, will determine their binding energies.
+This, in turn, will change the temperatures at which the ices desorb, and thus their snowline locations.
+This effect is particularly important in the case of CO and N$_2$, since their volatility is significantly higher than that of other carbon and nitrogen carriers, such as CO$_2$ or NH$_3$.
+Laboratory experiments \citet{fayolle16} have determined new values for the CO and N$_2$ binding energies, both as pure ices and water dominated.
+Figure \ref{fig:fayolle} shows the results of a temperature programmed desorption (TPD) experiment for the CO and N$_2$ desorption rate as a function of temperature, in water dominated and pure ice environments.
+The peak of the desorption curves shifts considerably between the different binding environments.
+This translates into a difference in binding energies by up to a factor of two between pure and water dominated CO or N$_2$.
+We show in Chapter 5 that this large difference in binding energies changes the CO and N$_2$ snowline locations by factors of 3-4 spending on the ice environment.
+By taking into account also the effect of disk dynamics, we find that the CO and N$_2$ snowlines may span several tens of AU.
+This uncertainty in snowline locations suggests yet again that more observations are needed to constrain disk volatile compositions at different locations, and thus trace the origins of gas giants based on their atmospheric compositions.
+
+\section{The Core Accretion Mechanism and Its Challenges}
+
+Gas giants are widely believed to form through core accretion (e.g., \citealt{pollack96}), a theory in which
+solid protoplanetary cores grow large enough to accumulate a massive atmosphere. In this
+model, planetesimals in a disk grow larger through collisions, eventually forming a planetary
+embryo, which continues to grow by attracting planetesimals in its neighborhood. Once an
+embryo becomes large enough so that its escape velocity exceeds the thermal velocity of the
+nebular gas in its vicinity, it starts accumulating a gaseous envelope. From this point on, the
+accretion of gas is regulated by the pressure support within the envelope — the amount of
+gas a core can accumulate is limited by the atmosphere’s ability to radiate away the energy
+due to the incoming planetesimals, as well as by envelope contraction.
+
+Protoplanetary disks dissipate on relatively short timescales of a few Myr.
+This implies that cores must grow fast in order to become large enough to attract massive atmospheres, and thus they need high planetesimal accretion rates.
+Studies that consider such high and constant rates (e.g., \citealt{stevenson82}, \citealt{boden86}, \citealt{wuchterl93}, \citealt{rafikov06}) find that the envelope accumulating around the planet is in steady state at all times, since all the energy due to the incoming planetesimals is radiated away by the atmosphere.
+It follows that the core and envelope grow simultaneously, and the mass of the atmosphere is a function of the core mass.
+Once the core and atmosphere have attained comparable masses, a rapid phase of runaway gas accretion starts and the gas giant can form.
+In this steady state scenario, there is therefore a uniquely determined core mass at which runaway accretion commences, called the critical core mass $M_{\rm crit}$.
+High planetesimal accretion rates cause the core to grow quickly, but this poses an additional challenge: fast accretion heats up the core's atmosphere, increase pressure support and inhibits the ability of the envelope to cool and contract, thus eventually increasing $M_{\rm crit}$.
+This is particularly challenging in the outer disk, where long dynamical times prevent large cores from forming before disk dissipation.
+\citet{rafikov11} shows that core accretion cannot operate beyond 40-50 AU at the accretion rates required to form the core and the atmosphere on the same timescale.
+
+All this seems to indicate that core accretion cannot work in the outer disk.
+This poses an intriguing question: could wide-seprataion gas giants, such as the HR 8799 system (Figure \ref{fig:HR8799}), have formed through core accretion?
+We can start answering this question by noting that planetesimal accretion does not need to be constant with time at a given location in the disk, which has been shown to be a viable scenario (e.g., \citealt{ikoma00}, \citealt{pollack96}).
+Figure \ref{fig:pollack} shows an example of this scenario.
+At early times, the core is in a high planetesimal accretion rate regime and grows fast, but the atmosphere remains small in comparison.
+As the planet depletes its feeding zone, planetesimal accretion is reduced.
+From this point on, the core no longer grows significantly, and thus the energy due to planetesimals can no longer balance radiative loses by the atmosphere; now the envelope accumulates gas while undergoing Kelvin-Helmholtz contraction, until its mass approximately exceeds the core mass and runaway gas accretion begins.
+Since now the envelope grows with time, there is no longer a uniquely determined $M_{\rm crit}$ --- in this scenario, the critical core mass is that of a core than can accrete n atmosphere with a mass equal to its own on a timescale equal to the disk lifetime.
+
+Such studies potentially provide a viable solution to the core accretion challenge at wide separations, but they are primarily focused on explaining the formation of Jupiter at 5.2 AU and thus do not explore a larger radii parameter space.
+Moreover, in order to truly understand whether core accretion can indeed work at stellocentric distances significantly larger than that of Jupiter, a more interesting question can be asked: what is the lowest possible core mass required to form a gas giant before disk dissipation?
+This minimum core mass is determined when planetesimal accretion has fully stopped, since additional accretion would actually increase this core mass (see above).
+Moreover, how does this minimum core mass depend on disk location, the properties of the nebular gas, and the opacity of the dust grains?
+We provide an answer to these questions in Chapters 2 and 3, where we find that the minimum core mass for giant planet formation is lower than the one typically quoted, 10 $M_{\oplus}$ (\citealt{stevenson82}, \citealt{rafikov06}), and may be as low as 1 $M_{\oplus}$.
+Our study thus reopens the case for in situ formation of wide-separation gas giants through core accretion.
+
+%\subsection{The Role of Disk Location in Setting the Minimum Core Mass for Giant Planet formation}
+%
+%\begin{itemize}
+%
+%\item Gas giants are largely believed to form through core accretion (insert core accretion sketch from various talks).
 This process is particularly challenging in the outer disk due to long dynamical timescales.
@@
 Finalize by saying that the studies I performed clearly challenge previous claims that  core accretion cannot operate at wide separations, thus reopening the case for in-situ formation of wide separation gas giants.
-
-\end{itemize}
-
-\subsection{The Role of Disk Dynamics and Morphology in Setting Snowline locations and C/N/O Ratios}
-
-\begin{itemize}
-
-\item Disk Composition Regulates Planet Composition.
+%
+%\end{itemize}
+%
+%\subsection{The Role of Disk Dynamics and Morphology in Setting Snowline locations and C/N/O Ratios}
+%
+%\begin{itemize}
+%
+%\item Disk Composition Regulates Planet Composition.
 It is thus essential to (1) predict what kinds of planet compositions result from planet location in different parts of the disk, and (2) conversely, bacl-track planet formation location based on planet composition.
-
-\item Disks are complex. Show figure from Henning and Semenov with all processes that occur in disks and discuss which ones I tackle in the thesis.
-
-\item However, we do know some things about disks.
+%
+%\item Disks are complex. Show figure from Henning and Semenov with all processes that occur in disks and discuss which ones I tackle in the thesis.
+%
+%\item However, we do know some things about disks.
 Volatile molecules have detected in disks (show figure with Spitzer IR spectrum of AA Tauri).
 Snowlines have also been detected (show CO snowline plot from Qi+13).
-
-\item One important signature of atmospheric chemistry is the C/O ratio (explain why and show plot from Molliere+15 which shows the effect of varying C/O ratio on the abundance of other volatiles).
+%
+%\item One important signature of atmospheric chemistry is the C/O ratio (explain why and show plot from Molliere+15 which shows the effect of varying C/O ratio on the abundance of other volatiles).
 Discuss claims that super stellar C/O ratios have been detected (show Wasp 12-b spectrum from Madhusudhan), but have since been refuted.
@@
 Discuss that disk dynamics have an important role, which is partly what we tackle in this part of the thesis.
+%
+%\item There are other important volatiles besides H2O, CO2 and CO.
+Discuss the importance of N2 and show plot with abundances in comets that show that NH3 is the main carrier based on observations, but it doesn't account for all the nitrogen in the solar system.
+%
+%\item Ice morphology is important. Discuss why. Discuss how binding energies depend on the ice environment and show plot from Edit's paper. 
+%
+%\item It follows that we have to take into account additional volatiles, abundances and ice morphologies besides disk dynamics to see their effect on snowline locations and the C/N/O ratios.
+We tackle this in Chaper xx.
+%
+%\item We demonstrate in this part to the thesis that C/N/O ratios may be used to track a planet's formation origin, when combined with observations, but that more observations are needed.
+%
+%\end{itemize}
 
-\item There are other important volatiles besides H2O, CO2 and CO.
-Discuss the importance of N2 and show plot with abundances in comets that show that NH3 is the main carrier based on observations, but it doesn't account for all the nitrogen in the solar system.
-
-\item Ice morphology is important. Discuss why. Discuss how binding energies depend on the ice environment and show plot from Edit's paper.
-
-\item It follows that we have to take into account additional volatiles, abundances and ice morphologies besides disk dynamics to see their effect on snowline locations and the C/N/O ratios.
-We tackle this in Chaper xx.
-
-\item We demonstrate in this part to the thesis that C/N/O ratios may be used to track a planet's formation origin, when combined with observations, but that more observations are needed.
-
-\end{itemize}
-
-\subsection{Disk-Planet Connection}
+\subsection{The Disk-Planet Connection}
 
 Our work in the two main parts of this thesis confirms that there is indeed a tight link between disks and planets.
-More importantly, as we will demonstrate in detail, the formation and composition of giant planets highly depends on disk location and properties.
-
-This thesis is organized as follows....
+Our study on core accretion shows that the core mass required to form a gas giant is highly dependent on the disk properties and composition, while our chapters on snowline locations and C/N/O ratios also demonstrate that the disk structure and chemical composition has direct effects on the volatile composition of giant planets atmospheres.
+This thesis thus sets the stage to uncovering the role of disk volatile composition, dynamics and chemistry in shaping the compositions of nascent giant planets.
 

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/introduction.tex
upd

@@
 \chapter{Introduction}
-\label{introduction}
 
+In recent years, the field of giant planet formation has aimed to answer two fundamental questions:
+
+\begin{enumerate}
+\item Where in the protoplanetary disk can gas giants form?
+\item What compositions will the formed giant planets have obtained?
+\end{enumerate}
+
+We can start uncovering the answers to these questions through a combination of studying planet formation in protoplanetary disks, the end-results of planet formation (i.e., exoplanets), and through comparisons with the architecture and composition of our own Solar System.
 Within the last two decades, more than one thousand extrasolar planets (exoplanets)
-have been discovered [1]. Their diversity in terms of mass, radius, location and composition
-[2] provides an exciting field of research, with the eventual goal of finding planets that are
+have been discovered \citet{batalha14}. Their diversity in terms of mass, radius, location and composition
+\citet{lissauer14} provides an exciting field of research, with the eventual goal of finding planets that are
 similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore
 and understand how planets obtain their compositions. Observations of Earth-like planets
-that can provide useful insight about their composition are challenging — the solid interior
+that can provide useful insight about their composition are challenging --- the solid interior
 structure of terrestrial planets cannot be detected, and their gaseous envelopes are small by
@@
 disk-planet connection have not yet been considered in detail.
-In this thesis, I uncover some of the answers to this issue from two standpoints: (1) by looking at the role of disk location in setting the conditions for the formation of wide-separation gas giants, and (2) by investigating how the structure and chemical composition of the protoplanetary disk at different radii affects the composition of nascent giant planets.
+In this thesis, we uncover some of the answers to this issue from two standpoints: (1) by looking at the role of disk location in setting the conditions for the formation of wide-separation gas giants, and (2) by investigating how the structure and chemical composition of the protoplanetary disk at different radii affects the composition of nascent giant planets.
 
-\subsection{The Role of Disk Location in Setting the Minimum Core Mass for Giant Planet formation}
+\section{Dynamics and Chemistry in Protoplanetary Disks}
 
-\begin{itemize}
+Protoplanetary disks are the result of the collapse of a molecular cloud, and are structures composed of gas and dust that are rotationally supported and that represent the birth environment of planets (Figure \ref{fig:proto}).
+Disks are complex objects, whose structure and evolution are affected by a multitude of chemical and dynamical processes.
+Many of these processes are summarized in Figure \ref{fig:disk_henning}, and several of them are not fully understood.
+Moreover, the timescales for various disk chemical and dynamical processes may be comparable at
+least in some parts of the disk (see Figure \ref{fig:timescales_semenov}), which makes the coupling between chemistry and dynamics very challenging.
+As a first step it is thus easier to isolate the different dynamical and chemical processes that occur in the disk midplane where planets are born, and understand their role and relative importance in shaping disk and planet compositions.
+We explore some of these processes in Chapters 4 and 5, and investigate their consequences for the compositions of giant planets forming at different radii.
 
-\item Gas giants are largely believed to form through core accretion (insert core accretion sketch from various talks).
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/proto.pdf}
+%\vspace{-0.5in}
+\caption{The standard picture of molecular cloud to protostar to protoplanetary disk to planet formation. From Oberg et al. ?}
+\label{fig:proto}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/disk.png}
+%\vspace{-0.5in}
+\caption{Cross section through a protoplanetary disk showing various chemical and dynamical processes that occur in disks. From \citet{henning13}.}
+\label{fig:disk_henning}
+\end{figure}
+
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/chemical_timescales.png}
+%\vspace{-0.5in}
+\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height.
+The numbered curves are $\log_{10} (\tau)$ in million years.
+From \citet{semenov11}}
+\label{fig:timescales_semenov}
+\end{figure*}
+
+\subsection{Snowlines and Disk Dynamics}
+
+Volatile compounds, which have low sublimation temperatures, are of particular importance.
+Their relative abundance in gaseous and solid form determines their snowline locations, and they are essential in the planet formation process (e.g., \citealt{pontoppidan14}).
+Observational advancements in recent year have provided us with an increasing amount of information regarding volatiles in disks.
+As shown in Figure \ref{fig:spitzer}, volatile compounds have been detected in disks, which gives us clues about their abundance in different disks and at different disk locations.
+Moreover, with the advent of ALMA volatile snowlines have also been observed.
+This is essential as it provides us with knowledge regarding volatile abundances in gas and dust throughout the disk, and thus the compositions of giant planets.
+Figure \ref{fig:qi13} shows the detection of the CO snowline in TW Hya at 30 AU \citep{qi13}.
+This snowline has been detected indirectly through its tracer molecule N$_2$H$^+$, which becomes highly abundant as CO freezes out.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/spitzer.png}
+%\vspace{-0.5in}
+\caption{Spitzer IR spectrum of the disk around AA Tauri showing detections of acetylene, hydrogen cyanide, carbon dioxide and water.}
+\label{fig:spitzer}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/CO.png}
+%\vspace{-0.5in}
+\caption{ALMA detection of the CO snowline in TW Hya at 30 AU. From \citet{qi13}.}
+\label{fig:qi13}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/molliere.png}
+%\vspace{-0.5in}
+\caption{Theoretical model of an exoplanet atmosphere showing that small variations in the C/O ratio produce very large variations in the abundance of other volatiles. From \citet{molliere15}.}
+\label{fig:molliere}
+\end{figure}
+
+While there are several volatiles in disks that can be further investigated, one important signature of atmospheric chemistry is the carbon-to-oxygen (C/O) ratio.
+This is demonstrated in Figure \ref{fig:molliere}, which shows a theoretical model of an exoplanet atmosphere: variations in the C/O ratio by factors of $\sim$3 change the abundance of other volatiles, such as H$_2$O and CH$_4$ by several orders of magnitude.
+From the observational perspective, an exciting potential discovery was that some planets, such as WASP-12 b, have a superstellar C/O ratio \citep{madhu11}.
+While this particular detection has since been unequivocally refuted (e.g., \citealt{kreidberg15}), the advent of JWST in the near future will provide us with a significantly larger sample of atmospheric spectra which may be used to constrain C/O ratios in other giant planet atmospheres.
+The possibility of a superstellar C/O ratio in a disk or planet atmosphere is also intriguing from a theoretical standpoint.
+One likely explanation is that the main carbon and oxygen carriers, i.e.
+H$_2$O, CO$_2$ and CO, have different condensation temperatures.
+This will result in variations in the abundance of the volatiles in gas and solid form between their respective snowlines, which will subsequently change the C/O ratio at different disk radii.
+This idea was first postulated by \citet{oberg11}.
+Figure \ref{fig:oberg} shows the C/O ratio in gas and dust as a function of semimajor axis, assuming protostellar abundances for the volatiles.
+Indeed, the C/O ratio in gas is enhanced compared to the stellar value, particularly between the CO$_2$ and CO snowlines where it reaches unity.
+The pioneering work of \citet{oberg11} inspired Chapter 4 of this thesis.
+\citet{oberg11} consider a static disk, and thus do not account for dynamical processes such as redistribution of solids due to radial drift, the radial movement of the nebular gas, and accretion heating.
+We thus expand this model by considering the processes outlined above and their effect on snowline locations.
+In particular, we focus on how radial drift of solids and viscous gas accretion onto the central star affect the H$_2$O, CO$_2$ and CO snowline locations for particles of different sizes.
+We find that drift and accretion heating alone may move the snowlines inward by factors of $\sim$2 compared to a static disk, thus substantially changing the disk regions with enhanced gas-phase C/O ratios, which has direct consequences for planet formation.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/oberg11.png}
+%\vspace{-0.5in}
+\caption{The C/O ratio in gas (solid line) and dust (dashed line) as a function of semimajor axis in a static disk.
+The dotted line shows the stellar value of 0.54.
+Gas-phase C/O ratios of order unity can be achieved in the outer disk.
+From \citet{oberg11}.
+}
+\label{fig:oberg}
+\end{figure}
+
+\subsection{Nitrogen and Ice Morphology}
+
+As we show above, the C/O ratio is highly important to study, but it only provides us with one piece of the puzzle that is constraining planet formation origins based on their compositions.
+In order to improve our theoretical models, we thus need to look at other molecules and volatile compounds.
+The first that comes to mind is nitrogen, as it highly abundant in the Solar System \citep{lodders03}, and primarily believed to exist as N$_2$ (e.g., \citealt{owen01}).
+Molecular nitrogen cannot be detected directly.
+However, an inventory of nitrogen cometary abundances shows that NH$_3$ is the most abundant observed nitrogen carrier, and yet its concentration is only a small fraction of the total nitrogen abundance, which leaves N$_2$ as the main nitrogen bearing species.
+In the context of planet compositions, N$_2$ is important due to its high volatility.
+The gas phase
+nitrogen-to-oxygen (N/O) ratio in the outer disk may thus be even more enhanced than the C/O ratio.
+Giant planets that form at wide separations should hence have an excess of nitrogen in their atmospheres,
+which could be used to trace their formation origin.
+We calculate the N/O ratio in gas and dust in Chapter 5, and confirm that it is indeed highly enhanced compared to the stellar value throughout most of the disk.
+
+The snowline locations of volatiles, including carbon, oxygen and nitrogen carriers, are highly dependent on the morphology of the icy particles.
+These can either form on a dust mantle as pure ices, or have a layered structure with the more volatile species closer to the outer layers (e.g., \citealt{pontoppidan03}).
+As H$_2$O is the least volatile species with high abundance in disks, one can typically expect the more volatile molecules to form on a water substrate (but see \citealt{bisschop06} for other scenarios).
+The ice environment in which the molecules reside, i.e.
+pure or water dominated ices, will determine their binding energies.
+This, in turn, will change the temperatures at which the ices desorb, and thus their snowline locations.
+This effect is particularly important in the case of CO and N$_2$, since their volatility is significantly higher than that of other carbon and nitrogen carriers, such as CO$_2$ or NH$_3$.
+Laboratory experiments \citet{fayolle16} have determined new values for the CO and N$_2$ binding energies, both as pure ices and water dominated.
+Figure \ref{fig:fayolle} shows the results of a temperature programmed desorption (TPD) experiment for the CO and N$_2$ desorption rate as a function of temperature, in water dominated and pure ice environments.
+The peak of the desorption curves shifts considerably between the different binding environments.
+This translates into a difference in binding energies by up to a factor of two between pure and water dominated CO or N$_2$.
+We show in Chapter 5 that this large difference in binding energies changes the CO and N$_2$ snowline locations by factors of 3-4 spending on the ice environment.
+By taking into account also the effect of disk dynamics, we find that the CO and N$_2$ snowlines may span several tens of AU.
+This uncertainty in snowline locations suggests yet again that more observations are needed to constrain disk volatile compositions at different locations, and thus trace the origins of gas giants based on their atmospheric compositions.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/fayolle.png}
+%\vspace{-0.5in}
+\caption{Desorption rate as a function of temperature in a TPD experiment, for CO and N$_2$ as pure ices (top panel), on a compact H$_2$O substrate (middle panel), and on porous H$_2$O substrate (bottom panel).
+The peaks in the desorption curves are at substantially higher temperatures for the water dominated ices.
+From \citet{fayolle16}.}
+\label{fig:fayolle}
+\end{figure}
+
+\section{The Core Accretion Mechanism and Its Challenges}
+
+Gas giants are widely believed to form through core accretion (e.g., \citealt{pollack96}), a theory in which
+solid protoplanetary cores grow large enough to accumulate a massive atmosphere. In this
+model, planetesimals in a disk grow larger through collisions, eventually forming a planetary
+embryo, which continues to grow by attracting planetesimals in its neighborhood. Once an
+embryo becomes large enough so that its escape velocity exceeds the thermal velocity of the
+nebular gas in its vicinity, it starts accumulating a gaseous envelope. From this point on, the
+accretion of gas is regulated by the pressure support within the envelope --- the amount of
+gas a core can accumulate is limited by the atmosphere's ability to radiate away the energy
+due to the incoming planetesimals, as well as by envelope contraction.
+
+Protoplanetary disks dissipate on relatively short timescales of a few Myr.
+This implies that cores must grow fast in order to become large enough to attract massive atmospheres, and thus they need high planetesimal accretion rates.
+Studies that consider such high and constant rates (e.g., \citealt{stevenson82}, \citealt{boden86}, \citealt{wuchterl93}, \citealt{rafikov06}) find that the envelope accumulating around the planet is in steady state at all times, since all the energy due to the incoming planetesimals is radiated away by the atmosphere.
+It follows that the core and envelope grow simultaneously, and the mass of the atmosphere is a function of the core mass.
+Once the core and atmosphere have attained comparable masses, a rapid phase of runaway gas accretion starts and the gas giant can form.
+In this steady state scenario, there is therefore a uniquely determined core mass at which runaway accretion commences, called the critical core mass $M_{\rm crit}$.
+High planetesimal accretion rates cause the core to grow quickly, but this poses an additional challenge: fast accretion heats up the core's atmosphere, increases pressure support and inhibits the ability of the envelope to cool and contract, thus eventually increasing $M_{\rm crit}$.
+This is particularly challenging in the outer disk, where long dynamical times prevent large cores from forming before disk dissipation.
+\citet{rafikov11} shows that core accretion cannot operate beyond 40-50 AU at the accretion rates required to form the core and the atmosphere on the same timescale.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/HR8799.png}
+%\vspace{-0.5in}
+\caption{The directly imaged HR 8799 system. Planets are located at 14.5 AU, 24 AU, 38 AU and 68 AU. From \citet{marois10}. }
+\label{fig:HR8799}
+\end{figure}
+
+All this seems to indicate that core accretion cannot work in the outer disk.
+This poses an intriguing question: could wide-separation gas giants, such as the HR 8799 system (Figure \ref{fig:HR8799}), have formed through core accretion?
+We can start answering this question by noting that planetesimal accretion does not need to be constant with time at a given location in the disk, which has been shown to be a viable scenario (e.g., \citealt{ikoma00}, \citealt{pollack96}).
+Figure \ref{fig:pollack} shows an example of this scenario.
+At early times, the core is in a high planetesimal accretion rate regime and grows fast, but the atmosphere remains small in comparison.
+As the planet depletes its feeding zone of solids, planetesimal accretion is reduced.
+From this point on, the core no longer grows significantly, and thus the energy due to planetesimals can no longer balance radiative loses by the atmosphere; now the envelope accumulates gas while undergoing Kelvin-Helmholtz contraction, until its mass approximately exceeds the core mass and runaway gas accretion begins.
+Since now the envelope grows with time, there is no longer a uniquely determined $M_{\rm crit}$ --- in this scenario, the critical core mass is that of a core than can accrete an atmosphere with a mass equal to its own on a timescale equal to the disk lifetime.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/pollack.png}
+%\vspace{-0.5in}
+\caption{The evolution of a nascent planet's core (denoted with the subscript Z), atmosphere (denoted with the subscript XY), and total mass (denoted with the subscript p) as a function of time, for a planet forming at 5.2 AU through core accretion.
+The plot depicts the three phases of the planet's evolution: (1) core growth while atmosphere remains small (until $\sim1.2$ Myr), (2) atmospheric growth while core growth is practically stalled (until $\sim$7 Myr), and (3) runaway gas accretion.}
+\label{fig:pollack}
+\end{figure}
+
+Such studies potentially provide a viable solution to the core accretion challenge at wide separations, but they are primarily focused on explaining the formation of Jupiter at 5.2 AU and thus do not explore a larger radii parameter space.
+Moreover, in order to truly understand whether core accretion can indeed work at stellocentric distances significantly larger than that of Jupiter, a more interesting question can be asked: what is the lowest possible core mass required to form a gas giant before disk dissipation?
+This minimum core mass is determined when planetesimal accretion has fully stopped, since additional accretion would actually increase this core mass (see above).
+Moreover, how does this minimum core mass depend on disk location, the properties of the nebular gas, and the opacity of the dust grains?
+We provide an answer to these questions in Chapters 2 and 3, where we find that the minimum core mass for giant planet formation is lower than the one typically quoted, 10 $M_{\oplus}$ (\citealt{stevenson82}, \citealt{rafikov06}), and may be as low as 1 $M_{\oplus}$.
+Our study thus reopens the case for in situ formation of wide-separation gas giants through core accretion.
+
+%\subsection{The Role of Disk Location in Setting the Minimum Core Mass for Giant Planet formation}
+%
+%\begin{itemize}
+%
+%\item Gas giants are largely believed to form through core accretion (insert core accretion sketch from various talks).
 This process is particularly challenging in the outer disk due to long dynamical timescales.
@@
 Discuss that disk dynamics have an important role, which is partly what we tackle in this part of the thesis.
+%
+%\item There are other important volatiles besides H2O, CO2 and CO.
+Discuss the importance of N2 and show plot with abundances in comets that show that NH3 is the main carrier based on observations, but it doesn't account for all the nitrogen in the solar system.
+%
+%\item Ice morphology is important. Discuss why. Discuss how binding energies depend on the ice environment and show plot from Edit's paper. 
+%
+%\item It follows that we have to take into account additional volatiles, abundances and ice morphologies besides disk dynamics to see their effect on snowline locations and the C/N/O ratios.
+We tackle this in Chaper xx.
+%
+%\item We demonstrate in this part to the thesis that C/N/O ratios may be used to track a planet's formation origin, when combined with observations, but that more observations are needed.
+%
+%\end{itemize}
 
-\item There are other important volatiles besides H2O, CO2 and CO.
-Discuss the importance of N2 and show plot with abundances in comets that show that NH3 is the main carrier based on observations, but it doesn't account for all the nitrogen in the solar system.
-
-\item Ice morphology is important. Discuss why. Discuss how binding energies depend on the ice environment and show plot from Edit's paper.
-
-\item It follows that we have to take into account additional volatiles, abundances and ice morphologies besides disk dynamics to see their effect on snowline locations and the C/N/O ratios.
-We tackle this in Chaper xx.
-
-\item We demonstrate in this part to the thesis that C/N/O ratios may be used to track a planet's formation origin, when combined with observations, but that more observations are needed.
-
-\end{itemize}
-
-\subsection{Disk-Planet Connection}
+\section{The Disk-Planet Connection}
 
 Our work in the two main parts of this thesis confirms that there is indeed a tight link between disks and planets.
-More importantly, as we will demonstrate in detail, the formation and composition of giant planets highly depends on disk location and properties.
-
-This thesis is organized as follows....
+Our study on core accretion shows that the core mass required to form a gas giant is highly dependent on the disk properties and composition, while our chapters on snowline locations and C/N/O ratios also demonstrate that the disk structure and chemical composition has direct effects on the volatile composition of giant planets atmospheres.
+This thesis thus sets the stage to uncovering the role of disk volatile composition, dynamics and chemistry in shaping the compositions of nascent giant planets.

### 2016-04-13 intro.tex
upd

@@
 
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/spitzer.png}
+%\vspace{-0.5in}
+\caption{Spitzer IR spectrum of the disk around AA Tauri showing detections of acetylene, hydrogen cyanide, carbon dioxide and water.}
+\label{fig:spitzer}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/CO.png}
+%\vspace{-0.5in}
+\caption{ALMA detection of the CO snowline in TW Hya at 30 AU. From \citet{qi13}.}
+\label{fig:qi13}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/molliere.png}
+%\vspace{-0.5in}
+\caption{Theoretical model of an exoplanet atmosphere showing that small variations in the C/O ratio produce very large variations in the abundance of other volatiles. From \citet{molliere15}.}
+\label{fig:molliere}
+\end{figure}
+
 While there are several volatiles in disks that can be further investigated, one important signature of atmospheric chemistry is the carbon-to-oxygen (C/O) ratio.
@@
 We find that drift and accretion heating alone may move the snowlines inward by factors of $\sim$2 compared to a static disk, thus substantially changing the disk regions with enhanced gas-phase C/O ratios, which has direct consequences for planet formation.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/oberg11.png}
+%\vspace{-0.5in}
+\caption{The C/O ratio in gas (solid line) and dust (dashed line) as a function of semimajor axis in a static disk.
+The dotted line shows the stellar value of 0.54.
+Gas-phase C/O ratios of order unity can be achieved in the outer disk.
+From \citet{oberg11}.
+}
+\label{fig:oberg}
+\end{figure}
 
@@
 
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/fayolle.png}
+%\vspace{-0.5in}
+\caption{Desorption rate as a function of temperature in a TPD experiment, for CO and N$_2$ as pure ices (top panel), on a compact H$_2$O substrate (middle panel), and on porous H$_2$O substrate (bottom panel).
+The peaks in the desorption curves are at substantially higher temperatures for the water dominated ices.
+From \citet{fayolle16}.}
+\label{fig:fayolle}
+\end{figure}
+
 \section{The Core Accretion Mechanism and Its Challenges}
@@
 
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/HR8799.png}
+%\vspace{-0.5in}
+\caption{The directly imaged HR 8799 system. Planets are located at 14.5 AU, 24 AU, 38 AU and 68 AU. From \citet{marois10}. }
+\label{fig:HR8799}
+\end{figure}
+
 All this seems to indicate that core accretion cannot work in the outer disk.
@@
 Since now the envelope grows with time, there is no longer a uniquely determined $M_{\rm crit}$ --- in this scenario, the critical core mass is that of a core than can accrete n atmosphere with a mass equal to its own on a timescale equal to the disk lifetime.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/pollack.png}
+%\vspace{-0.5in}
+\caption{The evolution of a nascent planet's core (denoted with the subscript Z), atmosphere (denoted with the subscript XY), and total mass (denoted with the subscript p) as a function of time, for a planet forming at 5.2 AU through core accretion.
+The plot depicts the three phases of the planet's evolution: (1) core growth while atmosphere remains small (until $\sim1.2$ Myr), (2) atmospheric growth while core growth is practically stalled (until $\sim$7 Myr), and (3) runaway gas accretion.}
+\label{fig:pollack}
+\end{figure}
 

### 2016-04-14 Dissertate-Harvard-LaTeX/chapters/conclusion.tex
upd

@@
 
-In this thesis, we have calculated the minimum core mass required to form wide-separation gas giants at different disk locations, and for different equations of state of the nebular gas and dust opacities.
-We have also determined how disk dynamics and volatile abundance and morphology affect volatile snowline locations, the C/N/O ratios across the disk, and therefore the compositions of nascent giant planets.
-Our results are summarized in Sections \ref{sec:core} and \ref{sec:volatile}, and we present directions for future work in Section \ref{sec:future}.
+In this thesis, I have calculated the minimum core mass required to form wide-separation gas giants at different disk locations, and for different equations of state of the nebular gas and dust opacities.
+I have also determined how disk dynamics and volatile abundance and morphology affect volatile snowline locations, the C/N/O ratios across the disk, and therefore the compositions of nascent giant planets.
+My results are summarized in Sections \ref{sec:core} and \ref{sec:volatile}, and I present directions for future work in Section \ref{sec:future}.
 
@@
 
-In Chapters 2 and 3, we determined the minimum core mass $M_{\rm crit}$ required to form a gas giant at distances between 5 and 100 AU from the host star, motivated by the fact that standard core accretion models cannot explain the formation of giant planets in the outer disk.
+In Chapters 2 and 3, I determined the minimum core mass $M_{\rm crit}$ required to form a gas giant at distances between 5 and 100 AU from the host star, motivated by the fact that standard core accretion models cannot explain the formation of giant planets in the outer disk.
 This minimum applies when the solid cores are no longer accreting solids, since any additional heating due to planetesimal accretion would heat up the core's atmosphere, inhibit its ability to cool and contract, and therefore increase the critical core mass.
-We thus considered atmospheres accreting around fully formed cores and undergoing Kelvin-Helmholtz contraction.
-In Chapter 2, we developed a quasi-static atmospheric evolution model by generating a series of static atmospheric profiles that are then connected temporally through a cooling equation.
-We considered an envelope structure consisting of an inner convective region and an outer radiative layer, and assumed a constant luminosity in the outer radiative region.
-We demonstrated that this is a valid approximation for our parameter space of interest.
-For an ideal gas polytrope and standard interstellar (ISM) opacities, we have found that $M_{\rm crit}$ decreases with semimajor axis from $\sim 8.5 M_{\oplus}$ at 5 AU to $\sim 3.5 M_{\oplus}$ at 100 AU.
-Our results are lower than the typically quoted value of $M_{\rm crit} \sim 10 M_{\oplus}$ (e.g., \citealt{rafikov06}), even in the more inner parts of the disk.
+I thus considered atmospheres accreting around fully formed cores and undergoing Kelvin-Helmholtz contraction.
+In Chapter 2, I developed a quasi-static atmospheric evolution model by generating a series of static atmospheric profiles that are then connected temporally through a cooling equation.
+I considered an envelope structure consisting of an inner convective region and an outer radiative layer, and assumed a constant luminosity in the outer radiative region.
+I demonstrated that this is a valid approximation for my parameter space of interest.
+For an ideal gas polytrope and standard interstellar (ISM) opacities, I have found that $M_{\rm crit}$ decreases with semimajor axis from $\sim 8.5 M_{\oplus}$ at 5 AU to $\sim 3.5 M_{\oplus}$ at 100 AU.
+My results are lower than the typically quoted value of $M_{\rm crit} \sim 10 M_{\oplus}$ (e.g., \citealt{rafikov06}), even in the more inner parts of the disk.
 
-To obtain more robust quantitative results, in Chapter 3 we expanded the model developed in Chapter 2 by considering a realistic equation of state (EOS) for the nebular gas and realistic dust opacities that take into account grain growth.
-We parametrized the EOS through the adiabatic gradient $\delad$ (defined in Chapter 2), as $\delad$ relates the gas pressure, temperature and density.
+To obtain more robust quantitative results, in Chapter 3 I expanded the model developed in Chapter 2 by considering a realistic equation of state (EOS) for the nebular gas and realistic dust opacities that take into account grain growth.
+I parametrized the EOS through the adiabatic gradient $\delad$ (defined in Chapter 2), as $\delad$ relates the gas pressure, temperature and density.
 While for an ideal gas EOS $\delad$ is constant, non-ideal EOS effects cause features in the adiabatic gradient that change the atmospheric structure.
-At high temperatures, molecular hydrogen dissociates, while at low-temperature the rotational states of the hydrogen molecule are only partially excited so it no longer behaves like an ideal gas, due to the existence of H$_2$ in two spin isomeric forms, ortho- and parahydrogen.
+At high temperatures, molecular hydrogen dissociates, while at low-temperatures the rotational states of the hydrogen molecule are only partially excited so it no longer behaves like an ideal gas, due to the existence of H$_2$ in two spin isomeric forms, ortho- and parahydrogen.
 Both of these effects increase $M_{\rm crit}$ by a factor of $\sim$2 compared to the ideal gas.
 In contrast, grain growth opacities decrease $M_{\rm crit}$ dramatically.
-By taking these two competing effects together, we calculate $M_{\rm crit} \sim 8 M_{\oplus}$ at 5 AU, decreasing to $\sim 5 M_{\oplus}$ at 100 AU.
-While our atmospheric model is not equipped to calculate the critical core mass when grain coagulation is taken into account, we have demonstrated that grain growth with coagulation may decrease $M_{\rm crit}$ by up to one order of magnitude; $M_{\rm crit}$ may be as low as 1 $M_{\oplus}$.
-Our study thus clearly challenges previous claims that core accretion cannot operate in the outer disk, reopening the case for in situ formation of wide separation gas giants.
+By taking these two competing effects together, I calculate $M_{\rm crit} \sim 8 M_{\oplus}$ at 5 AU, decreasing to $\sim 5 M_{\oplus}$ at 100 AU.
+While my atmospheric model is not equipped to calculate the critical core mass when grain coagulation is taken into account, I have demonstrated that grain growth with coagulation may decrease $M_{\rm crit}$ by up to one order of magnitude; $M_{\rm crit}$ may be as low as 1 $M_{\oplus}$.
+My study thus clearly challenges previous claims that core accretion cannot operate in the outer disk, reopening the case for in situ formation of wide separation gas giants.
 
@@
 
-In Chapters 4 and 5, we determined the effect of disk dynamics and ice morphology on the C/N/O ratio in active disks, which has direct implications on gas giant compositions.
+In Chapters 4 and 5, I determined the effect of disk dynamics and ice morphology on the C/N/O ratio in active disks, which has direct implications on gas giant compositions.
 This study was motivated by the fact that the locations of volatile snowlines in protoplanetary disks are a defining feature of both gas giant and disk chemistry, as they provide vital information about the abundance of these molecules in gas and dust throughout the disk.
-In Chapter 4, we expanded the model of \citet{oberg11} by considering the effect of disk dynamics on the snowline locations of the main carbon and oxygen carriers, i.e.
+In Chapter 4, I expanded the model of \citet{oberg11} by considering the effect of disk dynamics on the snowline locations of the main carbon and oxygen carriers, i.e.
 H$_2$O, CO$_2$ and CO, and thus on the C/O ratio in dust and gas throughout the disk.
-We studied the effect of radial drift of solids and viscous gas accretion onto the central star on snowline locations by developing a semi-analytical model, which we applied to three different disks in increasing order of complexity.
-We applied our model for a range of initial particle sizes and calculated how drift and gas accretion affect their desorption location.
-We have found that there is a range of particle sizes, $0.05$ cm $\lesssim s \lesssim$ 7 m for our particular parameters, that desorb almost instantaneously and a fixed particle-size dependent location from the central star, regardless of their initial position.
-Based on this information, we calculated the H$_2$O, CO$_2$ and CO snowline locations for different particle sizes in this range, and made estimates for the C/O ratio throughout the disk.
-We have found that the snowlines move inward as the particle size increases, and may differ by up to a factor of $\sim$2 due to drift and gas accretion compared to a static disk, which does not experience any dynamical processes.
+I studied the effect of radial drift of solids and viscous gas accretion onto the central star on snowline locations by developing a semi-analytical model, which I applied to three different disks in increasing order of complexity.
+I applied my model for a range of initial particle sizes and calculated how drift and gas accretion affect their desorption location.
+I have found that there is a range of particle sizes, $0.05$ cm $\lesssim s \lesssim$ 7 m for my particular parameters, that desorb almost instantaneously and at a fixed particle-size dependent location from the central star, regardless of their initial position.
+Based on this information, I calculated the H$_2$O, CO$_2$ and CO snowline locations for different particle sizes in this range, and made estimates for the C/O ratio throughout the disk.
+I have found that the snowlines move inward as the particle size increases, and may differ by up to a factor of $\sim$2 due to drift and gas accretion compared to a static disk, which does not experience any dynamical processes.
 This variation in snowline locations is significant, and has important consequences for the compositions of gas giants forming in situ.
 
-In Chapter 5, we expanded the model developed in Chapter 4 by considering additional volatiles, chemical abundances, and ice morphologies.
-As nitrogen is highly abundant in the Solar System and primarily found as N$_2$, we added nitrogen bearing species such as N$_2$ and NH$_3$ to our model, as well as hydrocarbons (specifically, CH$_4$).
-Motivated by laboratory experiments (e.g., \citealt{fayolle16}) that find significantly different binding energies for CO and N$_2$ depending on the ice environment in which they reside (pure versus water dominated ices), we calculated the CO and N$_2$ snowlines for both scenarios.
-We calculated the N/O ratio in static disks and found that it is highly enhanced compared to the stellar value: by a factor of $\sim$2 between the H$_2$O and CO$_2$ snowlines, by more than a factor of 3 between the CO$_2$ and CO snowlines, and by many orders of magnitude between the CO and N$_2$ snowlines, where oxygen gas is depleted.
-Thus we expect wide-separation giants to have an excess of nitrogen in their atmospheres, which may be used to trace their formation origins.
-We have also found that the binding environment has a large effect on the CO and N$_2$ snowline locations and may change them by factors of 3-4, both for static and dynamic disks.
-Thus by considering the combined effect of disk dynamics and ice morphology, the CO and N$_2$ snowline locations may change by up to a factor of $\sim$7, and may span 11-79 AU for N$_2$ in our disk model.
+In Chapter 5, I expanded the model developed in Chapter 4 by considering additional volatiles, chemical abundances, and ice morphologies.
+As nitrogen is highly abundant in the Solar System and primarily found as N$_2$, I added nitrogen bearing species such as N$_2$ and NH$_3$ to my model, as well as hydrocarbons (specifically, CH$_4$).
+Motivated by laboratory experiments (e.g., \citealt{fayolle16}) that find significantly different binding energies for CO and N$_2$ depending on the ice environment in which they reside (pure versus water dominated ices), I calculated the CO and N$_2$ snowlines for both scenarios.
+I calculated the N/O ratio in static disks and found that it is highly enhanced compared to the stellar value: by a factor of $\sim$2 between the H$_2$O and CO$_2$ snowlines, by more than a factor of 3 between the CO$_2$ and CO snowlines, and by many orders of magnitude between the CO and N$_2$ snowlines, where oxygen gas is depleted.
+Thus I expect wide-separation giants to have an excess of nitrogen in their atmospheres, which may be used to trace their formation origins.
+I have also found that the binding environment has a large effect on the CO and N$_2$ snowline locations and may change them by factors of 3-4, both for static and dynamic disks.
+Thus by considering the combined effect of disk dynamics and ice morphology, the CO and N$_2$ snowline locations may change by up to a factor of $\sim$7, and may span 11-79 AU for N$_2$ in my disk model.
 This large uncertainty in snowline locations means that observations that anchor snowline locations at different stages of planet
@@
 
-Our work has made important strides in our understudying of disk and planet compositions and the tight link between the two, as well as the effect of disk dynamics and chemistry on disk and planet compositions.
-As first step, we plan to expand our work by considering the effect of diffusion on the C/N/O ratios in viscous steady-state disks.
-Similarly to the model of \citet{owen14}, we will develop a simplified method to estimate the abundance of different volatiles at various disk locations, and thus their effect on the C/N/O ratio both in static and dynamic disks.
-We plan to explore further processes that affect snowline locations, such as variations in stelar luminosity and gas mass accretion rate, as well as particle growth and fragmentation.
-Our next step is to incorporate disk chemical processes, first by developing a simplified parametrized chemical  network that we will incorporate in our  drift-desorption model.
-By having a better understudying on how disk chemistry and dynamics affect the composition of nascent, and eventually mature planets, our work may provide essential context for characterizing the gas giants
+My work has made important strides in our understudying of disk and planet compositions and the tight link between the two, as well as the effect of disk dynamics and chemistry on disk and planet compositions.
+As a first step, I plan to expand my work by considering the effect of diffusion on the C/N/O ratios in viscous steady-state disks.
+Similarly to the model of \citet{owen14}, I will develop a simplified method to estimate the abundance of different volatiles at various disk locations, and thus their effect on the C/N/O ratio both in static and dynamic disks.
+
+Through analytical and numerical calculations, I will explore a range of dynamical processes that may affect snowline locations and the distribution of volatiles in disks, expanding and generalizing the framework developed in Chapter 4 and 5.
+Such effects include particle growth and fragmentation, as well as variations in the gas mass accretion rate and stellar luminosity.
+
+The complexity of disk chemistry means that coupling it with dynamical processes, while necessary, is non-trivial.
+As a next step I will thus couple the dynamical framework outlined above with time-dependent chemical models of increasing complexity, informed by results from state-of-the-art disk chemistry models (that can only be run on static disks).
+By having a better understudying on how disk chemistry and dynamics affect the composition of nascent, and eventually mature planets, my work may provide essential context for characterizing the gas giants
 that instruments such as JWST and the TESS will one day discover.

### 2016-04-14 Dissertate-Harvard-LaTeX/chapters/introduction.tex
upd

@@
 Disks are complex objects, whose structure and evolution are affected by a multitude of chemical and dynamical processes.
-Many of these processes are summarized in Figure \ref{fig:disk_henning}, and several of them are not fully understood.
-Moreover, the timescales for various disk chemical and dynamical processes may be comparable at
-least in some parts of the disk (see Figure \ref{fig:timescales_semenov}), which makes the coupling between chemistry and dynamics very challenging.
+The latter include transport of both dust through radial drift and gas through viscous gas accretion, as well as grain growth and dust settling to the disk midplane, which affect the disk structure.
+In terms of chemistry, the strongly irradiated disk surface is dominated by photochemical reactions, while the ionized lower layers have a rich ion-molecule chemistry.
+Finally, the cold disk midplane is mostly shielded from irradiation, and volatile compounds experience freeze-out.
+Many of these processes are summarized in \citet{henning13}, and several of them are not yet fully understood.
+Moreover, the timescales for various disk chemical and dynamical processes may be comparable at least in some parts of the disk \citep{semenov11}, which makes the coupling between chemistry and dynamics very challenging.
 As a first step it is thus easier to isolate the different dynamical and chemical processes that occur in the disk midplane where planets are born, and understand their role and relative importance in shaping disk and planet compositions.
@@
 %\vspace{-0.5in}
-\caption{The standard picture of molecular cloud to protostar to protoplanetary disk to planet formation. From Oberg et al. ?}
+\caption{The standard picture of molecular cloud to protostar to protoplanetary disk to planet formation. Reprinted by permission from Karin \"Oberg.}
 \label{fig:proto}
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.8\textwidth]{figures/disk.png}
-%\vspace{-0.5in}
-\caption{Cross section through a protoplanetary disk showing various chemical and dynamical processes that occur in disks. From \citet{henning13}.}
-\label{fig:disk_henning}
-\end{figure}
+%\subsection{Disk Temperature and Density Structure}
 
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=0.8\textwidth]{figures/chemical_timescales.png}
-%\vspace{-0.5in}
-\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height.
+%\begin{figure}[t!]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/disk.png}
+%%\vspace{-0.5in}
+%\caption{Cross section through a protoplanetary disk showing various chemical and dynamical processes that occur in disks. From \citet{henning13}.}
+%\label{fig:disk_henning}
+%\end{figure}
+
+%\begin{figure*}[t!]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/chemical_timescales.png}
+%%\vspace{-0.5in}
+%\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height.
 The numbered curves are $\log_{10} (\tau)$ in million years.
 From \citet{semenov11}}
-\label{fig:timescales_semenov}
-\end{figure*}
+%\label{fig:timescales_semenov}
+%\end{figure*}
 
@@
 Volatile compounds, which have low sublimation temperatures, are of particular importance.
-Their relative abundance in gaseous and solid form determines their snowline locations, and they are essential in the planet formation process (e.g., \citealt{pontoppidan14}).
+Their relative abundance in gaseous and solid form determines their snowline location, i.e.
+the distance in the disk where a volatile transitions from gaseous to solid form.
+Snowlines are essential in the planet formation process (e.g., \citealt{pontoppidan14}): they can enhance the amount of solid material just outside the snowline, due to dust sublimation, and they can produce pressure traps where dust particles pile up just inside the snowline.
 Observational advancements in recent year have provided us with an increasing amount of information regarding volatiles in disks.
-As shown in Figure \ref{fig:spitzer}, volatile compounds have been detected in disks, which gives us clues about their abundance in different disks and at different disk locations.
+Volatile compounds have been detected in disks (e.g., \citealt{henning13}), which gives us clues about their abundance in different disks and at different disk locations.
 Moreover, with the advent of ALMA volatile snowlines have also been observed.
@@
 This snowline has been detected indirectly through its tracer molecule N$_2$H$^+$, which becomes highly abundant as CO freezes out.
+Moreover, the H$_2$O snowline has also been detected in Tw Hya \citep{zhang13}, as well as the CO snowline in HD 163296 \citep{qi15}.
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.8\textwidth]{figures/spitzer.png}
-%\vspace{-0.5in}
-\caption{Spitzer IR spectrum of the disk around AA Tauri showing detections of acetylene, hydrogen cyanide, carbon dioxide and water.}
-\label{fig:spitzer}
-\end{figure}
+%\begin{figure}[t!]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/spitzer.png}
+%%\vspace{-0.5in}
+%\caption{Spitzer IR spectrum of the disk around AA Tauri showing detections of acetylene, hydrogen cyanide, carbon dioxide and water.}
+%\label{fig:spitzer}
+%\end{figure}
 
@@
 %\vspace{-0.5in}
-\caption{ALMA detection of the CO snowline in TW Hya at 30 AU. From \citet{qi13}.}
+\caption{ALMA detection of the CO snowline in TW Hya at 30 AU. Reprinted by permission from Karin \"Oberg.}
 \label{fig:qi13}
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.8\textwidth]{figures/molliere.png}
-%\vspace{-0.5in}
-\caption{Theoretical model of an exoplanet atmosphere showing that small variations in the C/O ratio produce very large variations in the abundance of other volatiles. From \citet{molliere15}.}
-\label{fig:molliere}
-\end{figure}
+%\begin{figure}[t!]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/molliere.png}
+%%\vspace{-0.5in}
+%\caption{Theoretical model of an exoplanet atmosphere showing that small variations in the C/O ratio produce very large variations in the abundance of other volatiles. From \citet{molliere15}.}
+%\label{fig:molliere}
+%\end{figure}
 
-While there are several volatiles in disks that can be further investigated, one important signature of atmospheric chemistry is the carbon-to-oxygen (C/O) ratio.
-This is demonstrated in Figure \ref{fig:molliere}, which shows a theoretical model of an exoplanet atmosphere: variations in the C/O ratio by factors of $\sim$3 change the abundance of other volatiles, such as H$_2$O and CH$_4$ by several orders of magnitude.
-From the observational perspective, an exciting potential discovery was that some planets, such as WASP-12 b, have a superstellar C/O ratio \citep{madhu11}.
-While this particular detection has since been unequivocally refuted (e.g., \citealt{kreidberg15}), the advent of JWST in the near future will provide us with a significantly larger sample of atmospheric spectra which may be used to constrain C/O ratios in other giant planet atmospheres.
-The possibility of a superstellar C/O ratio in a disk or planet atmosphere is also intriguing from a theoretical standpoint.
-One likely explanation is that the main carbon and oxygen carriers, i.e.
-H$_2$O, CO$_2$ and CO, have different condensation temperatures.
-This will result in variations in the abundance of the volatiles in gas and solid form between their respective snowlines, which will subsequently change the C/O ratio at different disk radii.
-This idea was first postulated by \citet{oberg11}.
+%\subsubsection{C/O Ratios in Disks}
+
+One important effect of the existence of snowlines is that disks are expected to contain different amounts of volatiles in gas and in dust at different locations.
+As the main carbon and oxygen carriers, i.e.
+H$_2$O, CO$_2$ and CO, are amongst the most abundant volatiles in comets and protostellar cores (\citealt{rodgers02}, \citealt{mumma11}, \citealt{henning13}), variations in the carbon-to-oxygen (C/O) ratio in gas and dust throughout a disk are particularly important.
+This issue was first addressed by \citet{oberg11}.
 Figure \ref{fig:oberg} shows the C/O ratio in gas and dust as a function of semimajor axis, assuming protostellar abundances for the volatiles.
-Indeed, the C/O ratio in gas is enhanced compared to the stellar value, particularly between the CO$_2$ and CO snowlines where it reaches unity.
+The C/O ratio in gas is enhanced compared to the stellar value, particularly between the CO$_2$ and CO snowlines where it reaches unity.
 The pioneering work of \citet{oberg11} inspired Chapter 4 of this thesis.
@@
 In particular, we focus on how radial drift of solids and viscous gas accretion onto the central star affect the H$_2$O, CO$_2$ and CO snowline locations for particles of different sizes.
-We find that drift and accretion heating alone may move the snowlines inward by factors of $\sim$2 compared to a static disk, thus substantially changing the disk regions with enhanced gas-phase C/O ratios, which has direct consequences for planet formation.
+We find that drift and accretion heating alone may move the snowlines inward by factors of $\sim$2 compared to a static disk, thus substantially changing the disk regions with enhanced gas-phase C/O ratios.
+%, which has direct consequences for planet formation.
 
@@
 Gas-phase C/O ratios of order unity can be achieved in the outer disk.
-From \citet{oberg11}.
+Reprinted by permission from Karin \"Oberg.
 }
@@
 
-\subsection{Nitrogen and Ice Morphology}
+%\subsubsection{The Importance of Nitrogen}
 
@@
 nitrogen-to-oxygen (N/O) ratio in the outer disk may thus be even more enhanced than the C/O ratio.
-Giant planets that form at wide separations should hence have an excess of nitrogen in their atmospheres,
-which could be used to trace their formation origin.
 We calculate the N/O ratio in gas and dust in Chapter 5, and confirm that it is indeed highly enhanced compared to the stellar value throughout most of the disk.
+
+%We note that this differential volatile condensation effect has already been used to explain potential claims hat some planets, such as WASP-12 b, have a superstellar C/O ratio \citep{madhu11}.
+While this particular detection has since been unequivocally refuted (e.g., \citealt{kreidberg15}), the advent of JWST in the near future will provide us with a significantly larger sample of atmospheric spectra which may be used to constrain C/O ratios in other giant planet atmospheres.
+%
+%
+%
+%While there are several volatiles in disks that can be further investigated, one important signature of atmospheric chemistry is the carbon-to-oxygen (C/O) ratio.
+This is demonstrated in Figure \ref{fig:molliere}, which shows a theoretical model of an exoplanet atmosphere: variations in the C/O ratio by factors of $\sim$3 change the abundance of other volatiles, such as H$_2$O and CH$_4$ by several orders of magnitude.
+From the observational perspective, an exciting potential discovery was that some planets, such as WASP-12 b, have a superstellar C/O ratio \citep{madhu11}.
+While this particular detection has since been unequivocally refuted (e.g., \citealt{kreidberg15}), the advent of JWST in the near future will provide us with a significantly larger sample of atmospheric spectra which may be used to constrain C/O ratios in other giant planet atmospheres.
+The possibility of a superstellar C/O ratio in a disk or planet atmosphere is also intriguing from a theoretical standpoint.
+One likely explanation is that the main carbon and oxygen carriers, i.e.
+H$_2$O, CO$_2$ and CO, have different condensation temperatures.
+This will result in variations in the abundance of the volatiles in gas and solid form between their respective snowlines, which will subsequently change the C/O ratio at different disk radii.
+This idea was first postulated by
+
+\subsection{Ice Morphology}
 
@@
 By taking into account also the effect of disk dynamics, we find that the CO and N$_2$ snowlines may span several tens of AU.
-This uncertainty in snowline locations suggests yet again that more observations are needed to constrain disk volatile compositions at different locations, and thus trace the origins of gas giants based on their atmospheric compositions.
+This uncertainty in snowline locations suggests yet again that more observations are needed to constrain disk volatile compositions at different locations.
+%, and thus trace the origins of gas giants based on their atmospheric compositions.
 
@@
 The peaks in the desorption curves are at substantially higher temperatures for the water dominated ices.
-From \citet{fayolle16}.}
+Reprinted by permission from Karin \"Oberg.}
 \label{fig:fayolle}
 \end{figure}
+
+\subsection{Consequences for the C/N/O Ratio in Planetary Atmospheres}
+
+Variations in the C/O ratio in gas and dust due to different volatile condensation temperatures have also been investigated in exoplanet atmospheres, both from a theoretical and observational standpoint.
+\citet{molliere15} show that varying the C/O ratio in a gas giant envelope by factors of $\sim$3 changes the abundance of other volatiles, such as H$_2$O and CH$_4$, by several orders of magnitude.
+From the observational perspective, an exciting potential discovery was that some planets, such as WASP-12 b, have a superstellar C/O ratio \citep{madhu11}.
+While this particular detection has since been unequivocally refuted (e.g., \citealt{kreidberg15}), the advent of JWST in the near future will provide us with a significantly larger sample of atmospheric spectra which may be used to constrain C/O ratios in other giant planet atmospheres.
+
+While observational constraints for the N/O ratio in planetary atmospheres do not currently exist, the high gas-phase N/O ratio enhancement at most disk radii suggests that giant planets that form at wide separations should have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
+Theoretical and observational knowledge of both C/O and N/O ratios could further help back-track the planet formation location based on planet composition.
 
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.8\textwidth]{figures/HR8799.png}
-%\vspace{-0.5in}
-\caption{The directly imaged HR 8799 system. Planets are located at 14.5 AU, 24 AU, 38 AU and 68 AU. From \citet{marois10}. }
-\label{fig:HR8799}
-\end{figure}
+%\begin{figure}[t!]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/HR8799.png}
+%%\vspace{-0.5in}
+%\caption{The directly imaged HR 8799 system. Planets are located at 14.5 AU, 24 AU, 38 AU and 68 AU. From \citet{marois10}. }
+%\label{fig:HR8799}
+%\end{figure}
 
 All this seems to indicate that core accretion cannot work in the outer disk.
-This poses an intriguing question: could wide-separation gas giants, such as the HR 8799 system (Figure \ref{fig:HR8799}), have formed through core accretion?
+This poses an intriguing question: could wide-separation gas giants, such as the HR 8799 system, have formed through core accretion?
 We can start answering this question by noting that planetesimal accretion does not need to be constant with time at a given location in the disk, which has been shown to be a viable scenario (e.g., \citealt{ikoma00}, \citealt{pollack96}).
-Figure \ref{fig:pollack} shows an example of this scenario.
+Figure 1 from \citet{pollack96} shows an example of this scenario.
 At early times, the core is in a high planetesimal accretion rate regime and grows fast, but the atmosphere remains small in comparison.
 As the planet depletes its feeding zone of solids, planetesimal accretion is reduced.
-From this point on, the core no longer grows significantly, and thus the energy due to planetesimals can no longer balance radiative loses by the atmosphere; now the envelope accumulates gas while undergoing Kelvin-Helmholtz contraction, until its mass approximately exceeds the core mass and runaway gas accretion begins.
+From this point on, the core no longer grows significantly, and thus the energy due to planetesimals can no longer balance radiative losses by the atmosphere; now the envelope accumulates gas while undergoing Kelvin-Helmholtz contraction, until its mass approximately exceeds the core mass and runaway gas accretion begins.
 Since now the envelope grows with time, there is no longer a uniquely determined $M_{\rm crit}$ --- in this scenario, the critical core mass is that of a core than can accrete an atmosphere with a mass equal to its own on a timescale equal to the disk lifetime.
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.8\textwidth]{figures/pollack.png}
-%\vspace{-0.5in}
-\caption{The evolution of a nascent planet's core (denoted with the subscript Z), atmosphere (denoted with the subscript XY), and total mass (denoted with the subscript p) as a function of time, for a planet forming at 5.2 AU through core accretion.
+%\begin{figure}[t!]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/pollack.png}
+%%\vspace{-0.5in}
+%\caption{The evolution of a nascent planet's core (denoted with the subscript Z), atmosphere (denoted with the subscript XY), and total mass (denoted with the subscript p) as a function of time, for a planet forming at 5.2 AU through core accretion.
 The plot depicts the three phases of the planet's evolution: (1) core growth while atmosphere remains small (until $\sim1.2$ Myr), (2) atmospheric growth while core growth is practically stalled (until $\sim$7 Myr), and (3) runaway gas accretion.}
-\label{fig:pollack}
-\end{figure}
+%\label{fig:pollack}
+%\end{figure}
 

### 2016-04-14 Dissertate-Harvard-LaTeX/frontmatter/abstract.tex
upd

@@
 In the second part, I explore how the composition and evolution of protoplanetary disks may affect the formation and chemical composition of giant planets.
-As the C/N/O ratios are an important signature of giant planet atmospheres, I show how the snowline locations of the main carbon, oxygen and nitrogen carriers, as well as the C/N/O ratios, are affected by disk dynamics and ice morphology.
+Volatile snowlines are highly important in the planet formation process.
+I thus show how the snowline locations of the main carbon, oxygen and nitrogen carriers, as well as the C/N/O ratios, are affected by disk dynamics and ice morphology.
 Compared to a static disk, disk dynamics and ice morphology combined may change the CO and N$_2$ snowline locations by a factor of 7.

### 2016-04-14 Dissertate-Harvard-LaTeX/chapters/introduction.tex
upd

@@
 
-\begin{figure}[t!]
+\begin{figure}[h]
 \centering
@@
 %\subsection{Disk Temperature and Density Structure}
+\subsection{Disk Temperature and Density Structure}
+
+The midplane temperature in a protoplanetary disk is set both by the irradiation from the host star and by accretion heating (e.g., \citealt{armitage10}).
+Accretion heating dominates in the inner disk (typically within a few AU) where the accretion flows are strongest, while the outer disk is dominated by stellar irradiation.
+While the irradiation component of the temperature simply follows a power-law in radius, the accretional component is determined by viscosity and the gas mass accretion rate, which in turn depend on the temperature itself, as well as the gas surface density (for a more complete review see \citealt{shakura73}).
+As the gas surface density changes with time in an active disk, it follows that determining the midplane temperature in an active disk in which both stellar irradiation and accretion heating contribute to the thermal evolution of the disk is non-trivial.
+We simplify this problem in Chapters 4 and 5 by assuming a steady-state disk with a constant mass accretion flow, and by solving the thin disk Shakura-Sunyaev equations.
+
+The gas surface density in a viscous disk follows the continuity equation, and is set by both the disk temperature and viscosity.
+Infall of material onto the disk or outflow may also contribute to the gas surface density evolution \citep{birnstiel10}.
+The dust surface density can be described by an advection-diffusion equation (e.g., \citealt{birnstiel12}), where both the radial movement of the dust and diffusion contribute to its evolution.
+The diffusion term is due to the fact that the dust is turbulently mixed by the gas, which causes a change in the dust-to-gas ratio and thus a radial gradient in the ratio between solid and gas surface densities.
+In the case of small particles that are well coupled to the gas, the surface density follows that of the gas and a constant dust-to-gas ratio is maintained.
+The movement of larger particles, however, is affected both by radial drift, gas drag and diffusion, which cause the dust surface density to "decouple" from that of the gas.
+This results in depletion of solids in some disk regions and a pile-up in others, depending on a particle's radial drift velocity and the level of turbulent mixing.
+The dust surface density is further affected by grain growth and fragmentation.
+We do not address all these effects in Chapter 4 and 5, and instead assume a constant inflow of particles such that the dust surface density remains spatially and temporally constant.
+We note, however, that these effects have to be taken into account for a more realistic modeling of the dust surface density, and thus of the surface density evolution of volatiles, both in solid and gaseous form.
 

### 2016-04-14 Dissertate-Harvard-LaTeX/chapters/chapter1.tex
upd

@@
 Many static studies reproduce a canonical $\MC \sim 10 M_\oplus$.
-However, \citet[hereafter R06]{Raf11, rafikov06} finds a wide range of values, $0.1 M_\oplus \lesssim \MC \lesssim 100 M_\oplus$, depending on the planetesimal accretion rate and disk parameters.
+However, \citet[hereafter R06]{rafikov11, rafikov06} finds a wide range of values, $0.1 M_\oplus \lesssim \MC \lesssim 100 M_\oplus$, depending on the planetesimal accretion rate and disk parameters.
 % \citep[see also][]{Raf11}.
@@
 Section \S\ref{sec:critical} gives our final results for growth timescales and critical core masses.
-We discuss some neglected effects in Section \S\ref{sec:neglected} and summarize our findings in Section \S\ref{sec:conclusions}.
+We discuss some neglected effects in Section \S\ref{sec:neglected1} and summarize our findings in Section \S\ref{sec:conclusions}.
 Some detailed derivations are presented in the appendices.
@@
 KH contraction.
-With a convective interior and radiative exterior, this model is motivated by similar models of hot Jupiters \citep{ab06, ym10}.
+With a convective interior and radiative exterior, this model is motivated by similar models of hot Jupiters \citep{arras06, youdin10}.
 
@@
 
-\section{Neglected Effects}\label{sec:neglected}
+\section{Neglected Effects}\label{sec:neglected1}
 Since one goal of this paper was to obtain a detailed understanding of atmospheric evolution with a simple model, we have necessarily ignored   effects of potential significance.

### 2016-04-14 Dissertate-Harvard-LaTeX/chapters/chapter1.tex
upd

@@
 These values, however, decrease to $\sim$$3$ Myrs and $\sim$$12 M_{\oplus}$ if planetesimal accretion is entirely shut off during the gas accretion phase.
-The core accretion model of \citet{HubBod05} for Jupiter's formation predicts an evolutionary time of 3.3 Myrs for a $10 M_{\oplus}$ core, for an interstellar opacity, which is also consistent with our result at 5 AU.
+The core accretion model of \citet{hubickyj05} for Jupiter's formation predicts an evolutionary time of 3.3 Myrs for a $10 M_{\oplus}$ core, for an interstellar opacity, which is also consistent with our result at 5 AU.
 
@@
 
-In terms of modern static studies, our results complement (and borrow some tools from) R06 and \citet{Raf11}.
+In terms of modern static studies, our results complement (and borrow some tools from) R06 and \citet{rafikov11}.
 These studies consider the disk radius dependence of core accretion for protoplanets that continuously accrete planetesimals.
-We show that the limits on core accretion at large radial distance claimed by \cite{Raf11} (and references therein) can be overcome if planetesimal accretion shuts off and the atmosphere is allowed to cool.
+We show that the limits on core accretion at large radial distance claimed by \cite{rafikov11} (and references therein) can be overcome if planetesimal accretion shuts off and the atmosphere is allowed to cool.
 Nevertheless, the plausibility of rapid core growth followed by negligible subsequent solid accretion admittedly remains uncertain, as discussed in more detail in \S\ref{sec:conclusions}.
@@
 In the low mass regime, $M\pla < M_{\rm th}/ \sqrt{3}$, where $\RB < \RH$, we assume that hydrostatic balance holds out to the outer boundary at $\RH$.
-In this low mass regime, \citet{Orm13} calculated the 2D (radial and azimuthal) flow patterns driven by stellar tides and disk headwinds.
+In this low mass regime, \citet{ormel13} calculated the 2D (radial and azimuthal) flow patterns driven by stellar tides and disk headwinds.
 On scales $\gtrsim \RB$ the flows no longer circulate the planet: they belong to the disk.
@@
 At $M\pla \gtrsim M_{\rm th}$ planets can open significant gaps \citep{zhu13}.
-At yet higher masses accretion instabilities could occur \citep{AylBat12}.
+At yet higher masses accretion instabilities could occur \citep{ayliffe12}.
 However, in this high mass regime, the spherically symmetric approximation has already broken down.
@@
 
-The importance of envelope opacity is an established factor in core accretion calculations (e.g., \citealt{stevenson82}, \citealt{ikoma00}, R06), and several studies explore the influence of opacity on the timescales for giant planet formation in detail (e.g., \citealt{HubBod05}).
+The importance of envelope opacity is an established factor in core accretion calculations (e.g., \citealt{stevenson82}, \citealt{ikoma00}, R06), and several studies explore the influence of opacity on the timescales for giant planet formation in detail (e.g., \citealt{hubickyj05}).
 Treating dust (and total) opacities as a power law in temperature is a simplification.
-Opacities drop by order unity when ice grains sublimate for $T \gtrsim 150$ K and they drop by orders of magnitude when silicate grains evaporate above $T \gtrsim 1500$ K \citep{semenov03, FerAle05}.
+Opacities drop by order unity when ice grains sublimate for $T \gtrsim 150$ K and they drop by orders of magnitude when silicate grains evaporate above $T \gtrsim 1500$ K \citep{semenov03, ferguson05}.
 Grain growth and composition also affect opacities.
@@
 Such ``radiative windows" arise if the opacity drop from ice and metal grain sublimation occurs at sufficiently low pressures.
-Our two-layer model ignores this possibility, but radiative windows are known to exist in hot Jupiter models \citep{burrows97, ab06} and have been seen in some core accretion models (Lissauer, personal communication).
+Our two-layer model ignores this possibility, but radiative windows are known to exist in hot Jupiter models \citep{burrows97, arras06} and have been seen in some core accretion models (Lissauer, personal communication).
 The role of radiative windows in core accretion models remains to be explored in detail.

### 2016-04-14 Dissertate-Harvard-LaTeX/chapters/chapter1.tex
upd

@@
 Dynamical models are needed to understand how runaway growth ends and final planet masses are determined.
-Relevant processes -- including gap opening in disks and the transition of accretion from spherical to planar -- can be simulated and then added to atmospheric evolution calculations \citep{LisHub09}.
+Relevant processes -- including gap opening in disks and the transition of accretion from spherical to planar -- can be simulated and then added to atmospheric evolution calculations \citep{lissauer09}.
 
@@
 Aside from this insight, the analytic model is useful because, despite the amplitude error, it explains the basic scaling of numerical cooling models.
-As is well known \citep{HubBod05}, a lower opacity, here scaling with $F_\kappa$, allows faster cooling and gives a lower critical core mass.
+As is well known \citep{hubickyj05}, a lower opacity, here scaling with $F_\kappa$, allows faster cooling and gives a lower critical core mass.
 The cooling timescale and critical core mass depend only weakly on the disk pressure, via the logarithmic factor $\xi$.
@@
 Lower opacities result in higher luminosities and faster evolution.
-Our model thus confirms a well-established result \citep{HubBod05}.
+Our model thus confirms a well-established result \citep{hubickyj05}.
 While clearly an important effect, atmospheric dust opacities are difficult to robustly predict.
@@
 
-\item A larger mean molecular weight reduces critical core masses, in agreement with \citet{HorIko11}.
+\item A larger mean molecular weight reduces critical core masses, in agreement with \citet{hori11}.
 If enrichment in heavy elements correlates with increased dust opacity, then the stronger opacity effect will dominate, increasing $\MC$.

### 2016-04-14 Dissertate-Harvard-LaTeX/chapters/chapter1.tex
upd

@@
 
-\begin{figure}[tb]
+\subsection{Atmospheric Structure}
+\label{sec:profiles}
+Figure \ref{fig:profiles} shows radial profiles at different stages of atmospheric growth around a $5 M_{\oplus}$ core at $60$ AU.
+Quoted mass values include the core plus atmosphere within the smaller of $\RB$ or $\RH$, which for these cases is $\RB$.
+The 8.99 $M_{\oplus}$ solution is the mass
+that satisfies our runaway growth criteria (described in \S\ref{sec:timeev}). % we can reach in our evolutionary sequence, as we explain further in \S\ref{sec:endoftime}.
+
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/PTm_profiles_v2_smallM.pdf}
+\includegraphics[width=0.6\textwidth]{figures/PTm_profiles_v2_smallM.eps}
 %\vspace{-0.5in}
@@
 
-\subsection{Atmospheric Structure}
-\label{sec:profiles}
-Figure \ref{fig:profiles} shows radial profiles at different stages of atmospheric growth around a $5 M_{\oplus}$ core at $60$ AU.
-Quoted mass values include the core plus atmosphere within the smaller of $\RB$ or $\RH$, which for these cases is $\RB$.
-The 8.99 $M_{\oplus}$ solution is the mass
-that satisfies our runaway growth criteria (described in \S\ref{sec:timeev}). % we can reach in our evolutionary sequence, as we explain further in \S\ref{sec:endoftime}.
+%\clearpage
 
@@
 \subsection{Time Evolution}\label{sec:timeev}
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/Lt_profiles_v2.pdf}
-%\vspace{-0.5in}
-\caption{Evolution of the luminosity and elapsed time during atmospheric growth around a $5 M_{\oplus}$ core at $60$ AU.
-The luminosity is initially high, then decreases as the atmosphere grows in mass and the radiative zone becomes optically thicker.
-Due to the neglect of self-gravity, the analytic model (\emph{dashed curve}) gives luminosities that are too low and evolution times that are too long.}
-\label{fig:Ltplot}
-\end{figure}
 
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/Lt_profiles_v2.eps}
+\vspace{-0.35in}
+\caption{Evolution of the luminosity and elapsed time during atmospheric growth around a $5 M_{\oplus}$ core at $60$ AU.
+The luminosity is initially high, then decreases as the atmosphere grows in mass and the radiative zone becomes optically thicker.
+Due to the neglect of self-gravity, the analytic model (\emph{dashed curve}) gives luminosities that are too low and evolution times that are too long.}
+\label{fig:Ltplot}
+\end{figure}
+
+%\clearpage
+
 The accelerated growth in the numerical model, as shown in the bottom panel of \Fig{fig:Ltplot}, is also a direct result of the higher cooling luminosities with self-gravity included.
@@
 Nevertheless, all our models show that self-gravity noticeably accelerates growth for $M_{\rm atm} > 0.1 M\co$.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/opacity_effect.eps}
+\vspace{-0.35in}
+\caption{The effect of dust abundance on atmospheric evolution.
+Reducing dust opacities by factors of 10 and 100 from standard Solar abundances gives higher luminosities and faster atmospheric growth.
+Plotted quantities are similar to \Fig{fig:Ltplot}, but for a $5 M_{\oplus}$ core at 10 AU.
+}  %The actual magnitude of dust opacities is difficult to robustly predict.}
+\label{fig:LtvsMopacity}
+\end{figure}
 
@@
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/opacity_effect.pdf}
-%\vspace{-0.5in}
-\caption{The effect of dust abundance on atmospheric evolution.
-Reducing dust opacities by factors of 10 and 100 from standard Solar abundances gives higher luminosities and faster atmospheric growth.
-Plotted quantities are similar to \Fig{fig:Ltplot}, but for a $5 M_{\oplus}$ core at 10 AU.
-}  %The actual magnitude of dust opacities is difficult to robustly predict.}
-\label{fig:LtvsMopacity}
-\end{figure}
+%\clearpage
 
@@
 Growth times are short both early -- when the radiative zone is transparent -- and late -- when self-gravity accelerates growth.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/Mt_profile_temp.eps}
+%\vspace{-0.5in}
+\caption{Evolution of the atmospheric growth timescale with mass around a $5 M_{\oplus}$ solid core  located at 10, 30 or 60 AU, for standard Solar opacities.
+Growth is slowest for $M_{\mathrm{atm}} \sim 1 - 3 M_{\oplus}$, i.e.\ before the crossover mass at $M_{\rm atm} = M\co$.}
+\label{fig:growthtime}
+\end{figure}
+
+%\clearpage
 
@@
 %The onset of runaway growth at high masses is also evident in \Figs{fig:Ltplot}{fig:LtvsMopacity} (bottom panels) as the flattening of the $t$ vs.\ $M_{\rm atm}$ curve.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/Mt_profile_temp.pdf}
-%\vspace{-0.5in}
-\caption{Evolution of the atmospheric growth timescale with mass around a $5 M_{\oplus}$ solid core  located at 10, 30 or 60 AU, for standard Solar opacities.
-Growth is slowest for $M_{\mathrm{atm}} \sim 1 - 3 M_{\oplus}$, i.e.\ before the crossover mass at $M_{\rm atm} = M\co$.}
-\label{fig:growthtime}
-\end{figure}
 
@@
 
-%At even higher masses than shown in \Fig{fig:coolingterms}, our cooling model breaks down.
-Mathematically, the breakdown manifests as a negative time step (due to an increase in internal entropy) when advancing to a higher atmospheric mass.
-This unphysical result has no deep significance, as the neglected luminosity overwhelms the model luminosity by this point.
-Effectively the model is telling us when to ignore it.
-Our results are unaffected by the breakdown, which occurs comfortably after $t_{\rm run}$, our threshold for runaway growth.
-
-\begin{figure}[tb]
+ \begin{figure}[H]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/cooling_a10_Mc5_rcb.pdf}
+\includegraphics[width=0.7\textwidth]{figures/cooling_a10_Mc5_rcb.eps}
 %\vspace{-0.5in}
@@
 
+%\clearpage
+
+%At even higher masses than shown in \Fig{fig:coolingterms}, our cooling model breaks down.
+Mathematically, the breakdown manifests as a negative time step (due to an increase in internal entropy) when advancing to a higher atmospheric mass.
+This unphysical result has no deep significance, as the neglected luminosity overwhelms the model luminosity by this point.
+Effectively the model is telling us when to ignore it.
+Our results are unaffected by the breakdown, which occurs comfortably after $t_{\rm run}$, our threshold for runaway growth.
+
 \section{Results for Giant Planet Formation}
@@
 
-\begin{figure}[tb]
+\begin{figure}[H]
 %\centering
-\hspace{-.1in}
-\includegraphics[width=0.5\textwidth]{figures/cumul_coolingtime_vs_Matm_10au_mu235.pdf}
-%\vspace{-0.5in}
+%\hspace{-.1in}
+\includegraphics[width=0.7\textwidth]{figures/cumul_coolingtime_vs_Matm_10au_mu235.eps}
+%\vspace{-0.35in}
 \caption{Time to grow an atmosphere of mass $M_{\rm{atm}}$ for cores with fixed masses between $5 M_{\oplus}$ and $14 M_{\oplus}$ (as labeled) at $10$ AU in our fiducial disk.
@@
 
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.48\textwidth]{figures/coolingtime_vs_Mc_10au.pdf}
-%\vspace{-0.5in}
-\caption{Runaway growth time, $t_{\rm{run}}$, vs.\ core mass at $10$ AU, for two values of the mean molecular weight.
-Our numerical model (\emph{solid curves}) is compared to our non-self-gravitating analytic model (\emph{dashed curves}, from Equation \ref{eq:tcoolanf}).
-A typical protoplanetary disk life time of $3$ Myr is plotted for comparison.
-The runaway growth time is larger for a lower mean molecular weight.}
-\label{fig:tvsMcomp}
-\end{figure}
-
 \subsubsection{Effects of Core Mass}
@@
 
+%\clearpage
+
 \Fig{fig:tvsMcomp} shows how $t_{\rm run}$ varies with core mass, also at 10 AU.
@@
 Moreover, in the analytic theory, crucial quantities like $M_{\rm atm}$ and $L$ (both roughly $\propto M\co^3$ near crossover) have non-linear dependence on core mass, offering plenty of opportunity for self-gravitational corrections.
+
+%\clearpage
 
@@
 
-\begin{figure*}[tb]
+\begin{figure}[H]
 \centering
-\includegraphics[width=1.\textwidth]{figures/TdPd_effect.pdf}
-\vspace{-0.3in}
+\includegraphics[width=0.7\textwidth]{figures/coolingtime_vs_Mc_10au.eps}
+%\vspace{-0.35in}
+\caption{Runaway growth time, $t_{\rm{run}}$, vs.\ core mass at $10$ AU, for two values of the mean molecular weight.
+Our numerical model (\emph{solid curves}) is compared to our non-self-gravitating analytic model (\emph{dashed curves}, from Equation \ref{eq:tcoolanf}).
+A typical protoplanetary disk life time of $3$ Myr is plotted for comparison.
+The runaway growth time is larger for a lower mean molecular weight.}
+\label{fig:tvsMcomp}
+\end{figure}
+
+\subsubsection{Effects of Disk Temperature and Pressure}
+\label{sec:TPeffects}
+
+\Fig{fig:TPeffects} shows how the runaway accretion time varies with disk temperature, $T\di$, or pressure, $P\di$, holding the other quantity fixed.
+The analytic model roughly reproduces the temperature and pressure scalings, again with some discrepancies due mainly to the neglect of self-gravity.
+Temperature variations are much more significant than pressure variations (note the difference in logarithmic and linear axes).
+Since midplane disk conditions depend only on temperature and pressure in our model,\footnote{See \Eq{eq:diskparam}.
+When gap opening is considered in models of later growth stages, the orbital frequency and effective viscosity become relevant as well.} the dominant effect of disk location is temperature.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=1.\textwidth]{figures/TdPd_effect.eps}
+%\vspace{-0.3in}
 \caption{Runaway growth time as a function of disk temperature (\emph{left}) and pressure (\emph{right}) around a $M_{\rm c} = 5 M_{\oplus}$ core.
@@
 \label{fig:TPeffects}
-\end{figure*}
-
-\subsubsection{Effects of Disk Temperature and Pressure}
-\label{sec:TPeffects}
-
-\Fig{fig:TPeffects} shows how the runaway accretion time varies with disk temperature, $T\di$, or pressure, $P\di$, holding the other quantity fixed.
-The analytic model roughly reproduces the temperature and pressure scalings, again with some discrepancies due mainly to the neglect of self-gravity.
-Temperature variations are much more significant than pressure variations (note the difference in logarithmic and linear axes).
-Since midplane disk conditions depend only on temperature and pressure in our model,\footnote{See \Eq{eq:diskparam}.
-When gap opening is considered in models of later growth stages, the orbital frequency and effective viscosity become relevant as well.} the dominant effect of disk location is temperature.
+\end{figure}
+
+%\clearpage
 
@@
 
-\begin{figure}[htb]
+\subsection{Critical Core Mass}
+\label{sec:critcore}
+
+The critical core mass declines with distance from the star, as shown in \Fig{fig:Mcvsa} for our standard disk model.
+The main reason for the decline, as explained above, is that atmospheres grow faster at lower temperatures.
+The lower densities and pressures in the outer disk have a much smaller effect.
+Since the vast majority of disk models have a disk temperature that declines with $a$, our qualitative result is robust.
+We also show that higher $\mu$ values give lower values of $\MC$.
+
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/Mcrit_vs_a_3Myrs_new.pdf}
+\includegraphics[width=0.7\textwidth]{figures/Mcrit_vs_a_3Myrs_new.eps}
 %\vspace{-0.5in}
@@
 
-\begin{figure}[htb]
+%\clearpage
+
+The average power-law decline from 1 to 100 AU (not plotted) is $\MC \propto a^{-0.3}$, for both choices of $\mu$.
+While not a drastic decline, the ability of distant low mass cores to accrete gas efficiently is significant for the interpretation of direct imaging surveys.
+However, our model offers no guarantee of copious giant planets at large distances.
+Many histories of solid accretion are possible, and solid cores may grow too slowly to allow the rapid gas accretion that we model.
+
+Our non-self-gravitating analytic model (dashed curves in \Fig{fig:Mcvsa}) over-predicts the steepness of the decline in $\MC$ with $a$.
+This discrepancy is not surprising, as we have shown that self-gravity strongly affects evolution, even before runaway growth begins and the crossover mass is reached.
+
+\Fig{fig:Mcritopacity} shows that reducing the opacity by an order of magnitude significantly reduces $\MC$.
+Furthermore, this opacity effect is stronger at larger distances.
+The opacity reduction by a factor of 10 lowers $\MC$ at 5 AU by a factor of $\sim$$2.5$ and at 100 AU by a factor of $\sim$$3.5$.
+Atmospheric opacity is a dominant uncertainty in core accretion modeling.
+However, unless atmospheric opacity varies significantly with disk radius, the general decline in $\MC$ with increasing $a$ should hold.
+For our scenario of negligible ongoing planetesimal accretion, it is tempting to think that dust could settle out of the radiative zone, lowering the opacity \citep{podolak03}.
+Nonetheless, since opacity near the RCB is a crucial factor, we speculate that convective overshoot may prevent very low opacities.
+
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/Mcrit_vs_a_3Myrs_opacity.pdf}
+\includegraphics[width=0.7\textwidth]{figures/Mcrit_vs_a_3Myrs_opacity.eps}
 %\vspace{-0.5in}
@@
 
-\subsection{Critical Core Mass}
-\label{sec:critcore}
-
-The critical core mass declines with distance from the star, as shown in \Fig{fig:Mcvsa} for our standard disk model.
-The main reason for the decline, as explained above, is that atmospheres grow faster at lower temperatures.
-The lower densities and pressures in the outer disk have a much smaller effect.
-Since the vast majority of disk models have a disk temperature that declines with $a$, our qualitative result is robust.
-We also show that higher $\mu$ values give lower values of $\MC$.
-
-The average power-law decline from 1 to 100 AU (not plotted) is $\MC \propto a^{-0.3}$, for both choices of $\mu$.
-While not a drastic decline, the ability of distant low mass cores to accrete gas efficiently is significant for the interpretation of direct imaging surveys.
-However, our model offers no guarantee of copious giant planets at large distances.
-Many histories of solid accretion are possible, and solid cores may grow too slowly to allow the rapid gas accretion that we model.
-
-Our non-self-gravitating analytic model (dashed curves in \Fig{fig:Mcvsa}) over-predicts the steepness of the decline in $\MC$ with $a$.
-This discrepancy is not surprising, as we have shown that self-gravity strongly affects evolution, even before runaway growth begins and the crossover mass is reached.
-
-\Fig{fig:Mcritopacity} shows that reducing the opacity by an order of magnitude significantly reduces $\MC$.
-Furthermore, this opacity effect is stronger at larger distances.
-The opacity reduction by a factor of 10 lowers $\MC$ at 5 AU by a factor of $\sim$$2.5$ and at 100 AU by a factor of $\sim$$3.5$.
-Atmospheric opacity is a dominant uncertainty in core accretion modeling.
-However, unless atmospheric opacity varies significantly with disk radius, the general decline in $\MC$ with increasing $a$ should hold.
-For our scenario of negligible ongoing planetesimal accretion, it is tempting to think that dust could settle out of the radiative zone, lowering the opacity \citep{podolak03}.
-Nonetheless, since opacity near the RCB is a crucial factor, we speculate that convective overshoot may prevent very low opacities.
+%\clearpage
 

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/chapter1.tex
upd

@@
 With the semimajor axis $a$ normalized to the outer disk as $\aun{10} = a/(10 \text{ AU})$, the gas surface density and mid-plane temperature are
-\begin{subeqnarray} \label{eq:diskparam}
+\begin{subeqnarray}  \label{1eq:diskparam}
 \varSigma\di  &=& 70 \,F_\varSigma \aun{10}^{-3/2} ~{\rm g~cm}^{-2} \\
@@
 \begin{equation}
-\label{eq:Pd}
+\label{1eq:Pd}
 P\di = 6.9 \times 10^{-3} F_\varSigma \sqrt{F_T} \, \aun{10}^{-45/14}~{\rm dyn~cm^{-2}}
@@
 \begin{equation}
-\label{eq:opacitylaw}
+\label{1eq:opacitylaw}
 \kappa= 2 F_\kappa  \left(\frac{T}{100\; \rm{K}}\right)^{\beta} \; \mathrm{cm^2 ~ g^{-1}},
@@
 \begin{equation}
-\label{eq:rc}
+\label{1eq:rc}
 R\co \equiv \left(\frac{3 M\co}{4 \pi \rho\co}\right)^{1/3} \approx 10^{-4} \mcn{10}^{1/3} ~\text{AU},
@@
 \begin{equation}
-\label{eq:RB}
+\label{1eq:RB}
 \RB \equiv \frac{G M\pla}{c\di^2} \approx 0.17 \, {\mpn{10}  \, \aun{10}^{3/7} \over F_T} ~\AU,
@@
 \begin{equation}
-\label{eq:RHill}
+\label{1eq:RHill}
 R_{\rm H} = \left(M\pla \over 3 M_\ast \right)^{1/3}a \approx 0.22 \, {\mpn{10}^{1/3} \, \aun{10} }~\AU ,
@@
 The relevant length scales of the atmosphere and disk satisfy the relation $\RB H\di^2 = 3 R_{\rm H}^3$.  The length scales are roughly equal at the ``thermal mass'' (e.g., \citealt{menou04})
-\begin{equation}\label{eq:Mth}
+\begin{equation}\label{1eq:Mth}
 M_{\rm th} > {c\di^{3} \over G \varOmega} \approx 25 \, {F_T^{3/2} \over \sqrt{m_\ast} } \, \aun{10}^{6/7}~ M_\oplus \, .
@@
 \subsection{Structure Equations and Boundary Conditions}
-\label{sec:struct}
+\label{1sec:struct}
 
@@
 \begin{subeqnarray}
-\label{eq:struct}
-\frac{dm}{dr}&=&4 \pi r^2 \rho\slabel{eq:structb} \\
-\frac{dP}{dr}&=&-\frac{G m}{r^2}\rho \slabel{eq:structa} \\
-\frac{dT}{dr}&=&\nabla \frac{T}{P}\frac{dP}{dr}\slabel{eq:structc} \\
-\frac{dL}{dr}&=&4 \pi r^2 \rho \left(\epsilon - \left. T {\partial S \over \partial t} \right|_m \right)\slabel{eq:structd}, 
+\label{1eq:struct}
+\frac{dm}{dr}&=&4 \pi r^2 \rho\slabel{1eq:structb} \\
+\frac{dP}{dr}&=&-\frac{G m}{r^2}\rho \slabel{1eq:structa} \\
+\frac{dT}{dr}&=&\nabla \frac{T}{P}\frac{dP}{dr}\slabel{1eq:structc} \\
+\frac{dL}{dr}&=&4 \pi r^2 \rho \left(\epsilon - \left. T {\partial S \over \partial t} \right|_m \right)\slabel{1eq:structd}, 
 \end{subeqnarray}
@@
 The enclosed mass  at radius $r$ is $m$.
-\Eq{eq:structc} simply defines the temperature gradient  $\nabla \equiv d \ln T/d \ln P$.
+\Eq{1eq:structc} simply defines the temperature gradient  $\nabla \equiv d \ln T/d \ln P$.
 
@@
 \begin{equation}
-\label{eq:delrad}
+\label{1eq:delrad}
 \delrad \equiv \frac{3 \kappa P}{64 \pi G m \sigma T^4} L\, ,
@@
 \begin{equation}
-\label{eq:delad}
+\label{1eq:delad}
 \delad \equiv \Big(\frac{d \ln T}{d \ln P}\Big)_{\rm{ad}}.
@@
 
-In the energy equation (\ref{eq:structd}), $\epsilon$ represents all local sources of heat input, except for the motion of the atmosphere itself.
+In the energy equation (\ref{1eq:structd}), $\epsilon$ represents all local sources of heat input, except for the motion of the atmosphere itself.
 From stellar structure, $\epsilon$ may be familiar as a nuclear burning term.
@@
 The partial time derivative is performed on shells of fixed mass.}  The partial time derivative would normally require our radial derivative to be partial as well.
-However, our subsequent developments will replace the local energy equation (\ref{eq:structd}) with global energy balance (see section \S\ref{cooling}), reverting the structure equations to time-independent ordinary differential equations (ODEs).
+However, our subsequent developments will replace the local energy equation (\ref{1eq:structd}) with global energy balance (see section \S\ref{cooling}), reverting the structure equations to time-independent ordinary differential equations (ODEs).
 %\textbf{}
 
-To solve the equation set (\ref{eq:struct}), an EOS is required for closure. In our study, we adopt an ideal gas law with a polytropic EOS 
-\begin{subeqnarray}\label{eq:idealEOS}
+To solve the equation set (\ref{1eq:struct}), an EOS is required for closure. In our study, we adopt an ideal gas law with a polytropic EOS 
+\begin{subeqnarray}\label{1eq:idealEOS}
 P &=& \rho \Rg T \, ,\slabel{eq:idealgas} \\
@@
 Our model atmospheres also obey $L(R\co) = 0$.
-Along with \Eq{eq:structd}, this boundary condition is incorporated in the global cooling model described below.
+Along with \Eq{1eq:structd}, this boundary condition is incorporated in the global cooling model described below.
 
@@
 \begin{subeqnarray}
-E_G&=&-\int_{M\co}^M \frac{G m}{r} dm \, , \label{eq:Eg} \\
-U&=&\int_{M\co}^M u dm \, .\slabel{eq:U}
+E_G&=&-\int_{M\co}^M \frac{G m}{r} dm \, , \label{1eq:Eg} \\
+U&=&\int_{M\co}^M u dm \, .\slabel{1eq:U}
 \end{subeqnarray}
@@
 \begin{equation}
-\label{eq:coolingstar}
+\label{1eq:coolingstar}
 L_M = L\co + \Gamma - \dot{E}.
@@
 \begin{equation}
-\label{eq:coolingglobal}
+\label{1eq:coolingglobal}
 L_M=L\co+\Gamma-\dot{E}+e_M \dot{M} - P_M \left. \frac{\partial V_M}{\partial t}\right|_M \, ,
@@
 
-In convective regions, the adiabatic structure is independent of luminosity and can be calculated without local energy balance, \Eq{eq:structd}.
-Thus, for fully convective objects, a cooling sequence can be established by connecting a series of adiabatic solutions using a global energy equation, $L_M = -\dot{E}$ or \Eq{eq:coolingstar}.
+In convective regions, the adiabatic structure is independent of luminosity and can be calculated without local energy balance, \Eq{1eq:structd}.
+Thus, for fully convective objects, a cooling sequence can be established by connecting a series of adiabatic solutions using a global energy equation, $L_M = -\dot{E}$ or \Eq{1eq:coolingstar}.
 Such methods are commonly used for their computational efficiency and are sometime referred to as ``following the adiabats," since the steady state solutions evolve in order of decreasing entropy \citep{marleau13}.
 
-In the radiative zone,  local energy balance, \Eq{eq:structd}, does affect the atmospheric structure.
+In the radiative zone,  local energy balance, \Eq{1eq:structd}, does affect the atmospheric structure.
 We proceed by assuming that the majority of energy is lost from the convective interior, and thus the luminosity can be treated as constant in the outer radiative zone, i.e.
-the RHS of \Eq{eq:structd} is set to zero.
+the RHS of \Eq{1eq:structd} is set to zero.
 This assumption greatly simplifies the numerical problem.
-Instead of solving time dependent partial differential equations, we can solve for  static solutions to a set of ODEs, Equations (\ref{eq:struct}a -- c), which we connect in a time series as described below.
+Instead of solving time dependent partial differential equations, we can solve for  static solutions to a set of ODEs, Equations (\ref{1eq:struct}a -- c), which we connect in a time series as described below.
 We show in \S\ref{sec:endoftime} that the approximation of constant luminosity in the outer layer holds for our regime of interest.
@@
 At the outer boundary, at $\RH(M_i)$, the temperature and pressure are set to the disk values.
-A luminosity value is required to compute $\delrad$ and integrate Equations (\ref{eq:struct}a--c).
+A luminosity value is required to compute $\delrad$ and integrate Equations (\ref{1eq:struct}a--c).
 The correct value of the luminosity is not known in advance, and is the eigenvalue of the problem.
@@
 
-To understand time evolution, we construct an array of solutions to Equations (\ref{eq:struct}a -- c) in order of increasing atmospheric mass.
-We then use global energy balance, \Eq{eq:coolingglobal}, to ``follow the mass" and place these solutions in a cooling sequence.
-To establish the time difference between neighboring solutions, we apply \Eq{eq:coolingglobal} at the RCB.
+To understand time evolution, we construct an array of solutions to Equations (\ref{1eq:struct}a -- c) in order of increasing atmospheric mass.
+We then use global energy balance, \Eq{1eq:coolingglobal}, to ``follow the mass" and place these solutions in a cooling sequence.
+To establish the time difference between neighboring solutions, we apply \Eq{1eq:coolingglobal} at the RCB.
 In principle, energy balance could be evaluated at any level.
@@
 \begin{equation}
-\label{eq:dti}
+\label{1eq:dti}
 \Delta t = %\left(
@@
 \end{equation} 
-using \Eq{eq:coolingglobal} with $\Gamma = L\co = 0$.
+using \Eq{1eq:coolingglobal} with $\Gamma = L\co = 0$.
 Brackets indicate an average of, and $\Delta$ indicates a difference between, the two states.
 All values are evaluated at the RCB.
-Due to the partial derivative in \Eq{eq:coolingglobal}, the volume difference $\Delta V_{<M>}$ is performed at fixed mass, here the average of the masses at the RCB.
+Due to the partial derivative in \Eq{1eq:coolingglobal}, the volume difference $\Delta V_{<M>}$ is performed at fixed mass, here the average of the masses at the RCB.
 
@@
 This approximation ignores the earliest stages of cooling, which are rapid enough to be of minor importance.
-The analytic model also ignores the surface terms in \Eq{eq:coolingglobal}, which we show to be a modest correction in \S\ref{sec:endoftime}.
+The analytic model also ignores the surface terms in \Eq{1eq:coolingglobal}, which we show to be a modest correction in \S\ref{sec:endoftime}.
 Finally, while the numerical models set the outer boundary at $\RH$, we simplify the analytic calculations by setting the outer boundary to infinity, effectively neglecting the finite scale-height of the disk.
@@
 \end{equation} 
-follows from \Eq{eq:delrad} and marginal convective stability, $\delrad = \delad$, where we define 
+follows from \Eq{1eq:delrad} and marginal convective stability, $\delrad = \delad$, where we define 
 \begin{equation} 
@@
 
-Our analytic cooling model uses $L = -\dot{E}$,  neglecting the surface terms in \Eq{eq:coolingglobal}.\footnote{\App{surfterms} shows that these terms are negligible for a non-self-gravitating model.}  %They do become relevant when self-gravity is important and included, as shown in \S\ref{sec:endoftime}.}
+Our analytic cooling model uses $L = -\dot{E}$,  neglecting the surface terms in \Eq{1eq:coolingglobal}.\footnote{\App{surfterms} shows that these terms are negligible for a non-self-gravitating model.}  %They do become relevant when self-gravity is important and included, as shown in \S\ref{sec:endoftime}.}
 Applying \Eqs{eq:Eus}{eq:Lcb}, the time it takes to cool the atmosphere until the RCB reaches a given pressure depth, $P\cb$, is 
@@
 
-In \Fig{fig:LtvsMopacity}, our standard dust opacity, \Eq{eq:opacitylaw}, is reduced by factors of 10 and 100.
+In \Fig{fig:LtvsMopacity}, our standard dust opacity, \Eq{1eq:opacitylaw}, is reduced by factors of 10 and 100.
 Lower opacities result in higher luminosities and faster evolution.
@@
 
- The individual terms in the global cooling model of \Eq{eq:coolingglobal}, evaluated at the RCB, are also plotted in \Fig{fig:coolingterms}.
+ The individual terms in the global cooling model of \Eq{1eq:coolingglobal}, evaluated at the RCB, are also plotted in \Fig{fig:coolingterms}.
 At low masses, the change in energy, $- \dot{E}$, makes the dominant contribution to luminosity.
@@
 %\vspace{-0.5in}
-\caption{Individual terms in the atmospheric cooling model of \Eq{eq:coolingglobal}, for a $5 M_\oplus$ core at 10 AU.
+\caption{Individual terms in the atmospheric cooling model of \Eq{1eq:coolingglobal}, for a $5 M_\oplus$ core at 10 AU.
 The dashed curve for accretion energy indicates a negative contribution.
@@
 The growth of atmospheres close to the star is also important, but spherical accretion models (including ours) are less applicable here.
-In the inner disk, critical core masses increase, yet lower mass planets start to open gaps and outgrow the disk scale height, see \Eq{eq:Mth}.
+In the inner disk, critical core masses increase, yet lower mass planets start to open gaps and outgrow the disk scale height, see \Eq{1eq:Mth}.
 These concerns prevent us from applying our model to the inner disk.
@@
 Temperature variations are much more significant than pressure variations (note the difference in logarithmic and linear axes).
-Since midplane disk conditions depend only on temperature and pressure in our model,\footnote{See \Eq{eq:diskparam}.
+Since midplane disk conditions depend only on temperature and pressure in our model,\footnote{See \Eq{1eq:diskparam}.
 When gap opening is considered in models of later growth stages, the orbital frequency and effective viscosity become relevant as well.} the dominant effect of disk location is temperature.
@@
 \end{equation}
-Our model reproduces the linear opacity dependence (see \S\ref{sec:timeev}), even though we use a temperature dependent opacity, \Eq{eq:opacitylaw}, instead of a constant $\kappa$.
+Our model reproduces the linear opacity dependence (see \S\ref{sec:timeev}), even though we use a temperature dependent opacity, \Eq{1eq:opacitylaw}, instead of a constant $\kappa$.
 Our growth times at 5 AU,
@@
 
-Our model uses an ideal gas law and a polytropic EOS, given by \Eq{eq:idealEOS}.
+Our model uses an ideal gas law and a polytropic EOS, given by \Eq{1eq:idealEOS}.
 However, non-ideal effects can affect atmospheric structure and evolution.

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/chapter2.tex
upd

@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centering
 %\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/delad_S_mixt.pdf}
-\includegraphics[width=0.62\textwidth]{figures/delad_S_mixt.pdf}
+\includegraphics[width=0.75\textwidth]{figures/delad_S_mixt.pdf}
 %%\vspace{-0.5in}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centering
 %\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/tplot.pdf}
-\includegraphics[width=0.5\textwidth]{figures/tplot.pdf}
+\includegraphics[width=0.7\textwidth]{figures/tplot.pdf}
 %%\vspace{-0.5in}
@@
 
-\begin{figure*}[tb]
+%In Figure \ref{fig:all_plot}, middle-right and upper-right panels, we show an example evolution with planet mass of $-dE/dM$ and $L$, respectively, for the EOSs described above.
+The realistic EOS increases $-dE/dM$ by a factor of $\sim$$3$ and decreases $L$ by a factor of $\sim$$1.5$, when compared to the ideal gas polytrope.
+%Atmospheric growth is therefore primarily slowed by changes in the energy profile $E(r)$, with the luminosity $L$ having a secondary effect. 
+%In what follows we explore each contribution separately.
+
+Throughout Section \ref{EOSeffects}, we use a power-law opacity given by
+\begin{equation}
+\label{eq:opacitylaw}
+\kappa=2 F_{\kappa} \Big(\frac{T}{T_{\rm ref}}\Big)^{\beta},
+\end{equation}  
+\noindent with $\beta =2$, $F_{\kappa} = 1$, and $T_{\rm{ref}}=100$K, appropriate for ice grains in the interstellar medium (ISM) \citep{bell94}.
+We improve our treatment of opacity in Section \ref{sec:opacity}.
+
+\begin{figure}[H]
 \centering
@@
 \label{fig:all_plot}
-\end{figure*}
-
-%In Figure \ref{fig:all_plot}, middle-right and upper-right panels, we show an example evolution with planet mass of $-dE/dM$ and $L$, respectively, for the EOSs described above.
-The realistic EOS increases $-dE/dM$ by a factor of $\sim$$3$ and decreases $L$ by a factor of $\sim$$1.5$, when compared to the ideal gas polytrope.
-%Atmospheric growth is therefore primarily slowed by changes in the energy profile $E(r)$, with the luminosity $L$ having a secondary effect. 
-%In what follows we explore each contribution separately.
-
-Throughout Section \ref{EOSeffects}, we use a power-law opacity given by
-\begin{equation}
-\label{eq:opacitylaw}
-\kappa=2 F_{\kappa} \Big(\frac{T}{T_{\rm ref}}\Big)^{\beta},
-\end{equation}  
-\noindent with $\beta =2$, $F_{\kappa} = 1$, and $T_{\rm{ref}}=100$K, appropriate for ice grains in the interstellar medium (ISM) \citep{bell94}.
-We improve our treatment of opacity in Section \ref{sec:opacity}.
+\end{figure}
 
@@
 
-\begin{figure*}[tb]
+\subsection{Role of $L$}\label{sec:roleofL}
+
+%The radiative and convective profiles must match at the RCB, constrained by the known total mass of the atmosphere
+
+%The EOS variations in the cold outer envelope affect $\dot{M}$ by decreasing the atmosphere's luminosity compared to an ideal gas polytrope (Figure \ref{fig:all_plot}, upper-right panel, for the same parameters as Figure \ref{fig:tplotall}).
+From Figures \ref{fig:deladmap} and \ref{fig:all_plot}, bottom-left panel, we see that the adiabatic gradient of the low-$T$ realistic EOS increases with radius (i.e., with decreasing temperature) throughout the convective zone when $T<500$ K, then becomes lower near the RCB.
+This results in a larger adiabatic gradient, on average, for the low-$T$ realistic EOS in the outer parts of the convective region of the atmosphere.
+
+%The variations in $\delad$ at low temperatures
+
+The increase in $\delad$ at low temperatures, where $H_2$ rotational states are not fully excited (Figure \ref{fig:all_plot}, upper-right) decreases the atmosphere's luminosity.
+%compared to an ideal gas polytrope 
+This result may be understood as consequence of matching between the atmosphere's interior convective and exterior radiative profiles.
+The profiles must match at the RCB, constrained by the known total mass.
+This match determines the pressure at the RCB, $P\cb \propto \exp(\RB/R\cb)$.
+Higher $P\cb$ implies lower luminosity (Equation \ref{eq:delrad}).
+
+The adiabatic gradient of EOS 2 is larger, on average, than for EOS 1 in the outer part of the convective region.\footnote {$\delad$ for EOS 2 increases with radius (i.e., with decreasing temperature) throughout the convective zone, then becomes lower near the RCB.} To understand how a variable, but overall larger $\delad$ in the outer atmosphere affects evolution, we study the simplified problem of atmospheres composed of an ideal gas with adiabatic gradient
+
+\begin{equation}
+\label{eq:deladvar}
+\delad = \left\{
+\begin{array}{l l}
+\nabla_{\rm ad, 1}, & \quad T > 500 \text{ K} \\
+\nabla_{\rm ad, 2}, & \quad T < 500 \text{ K}
+\end{array} 
+\right.
+\end{equation}   
+where $\nabla_{\rm ad, 1}$ and $\nabla_{\rm ad,2}$ are constant (Figure \ref{fig:varying_delad}).
+
+%We explore the independent effects of varying $\nabla_{\rm ad,1}$ and $\nabla_{\rm ad,2}$ by alternately fixing one and varying the other   (Figure \ref{fig:varying_delad}). 
+%The adiabatic gradient $\nabla_{\rm ad,1}$ dictates the temperature profile deep in the convective interior, while $\nabla_{\rm ad,2}$ sets the RCB temperature $T\cb$. 
+%Given these inputs, the RCB location and its pressure $P\cb \propto \exp(\RB/R\cb)$ (see Figure \ref{fig:varying_delad}, lower panels) are determined by matching the interior and exterior profiles.
+Higher $P\cb$ implies lower luminosity (Equation \ref{eq:delrad}).
+
+%the location of the RCB, is set by the matching between the convective interior and the outer radiative zone. As $P(r) \propto \exp(\RB/r)$ in the radiative region,
+
+%As $P(r) \propto \exp(\RB/r)$ in the outer radiative region,
+
+%Figure \ref{fig:varying_delad}, lower-left panel, shows the radial pressure profile for $\nabla_{\rm ad,1}=0.3$ and various $\nabla_{\rm ad,2}$, for the same parameters as Figures \ref{fig:tplotall} and \ref{fig:all_plot}, which are in agreement with the analytic expectation of an exponential profile.
+
+%The numerical solutions reproduce the analytic expectation of a nearly exponential pressure profile in the radiative region, $P(r) \propto \exp(\RB/r)$.
+Thus $L \propto 1/P\cb \propto 1/\exp(\RB/R\cb)$ is directly correlated to the physical depth of the RCB.
+
+%The location of the RCB is set by the matching between the convective interior and the outer radiative zone.
+
+The adiabatic gradient $\nabla_{\rm ad,1}$ dictates the temperature profile deep in the convective interior.
+Deep in the atmosphere, where $r \ll R\cb$, virial equilibrium yields a temperature profile
+\begin{equation}
+\label{eq:Tdeep}
+T(r) \approx \frac{\nabla_{\rm ad,1} G M\co}{\mathcal{R} r}.
+\end{equation}   
+Even though $\nabla_{\rm ad,1}$ is an order unity coefficient, the temperature scaling with $\nabla_{\rm ad,1}$ in \Eq{eq:Tdeep} is exact (Figure \ref{fig:varying_delad}, upper-right panel).
+Because the radial temperature profile in the atmospheric interior is set by $\nabla_{\rm ad,1}$, the radius $R_{\rm switch}$ at which the adiabatic gradient changes shows little variation with $\nabla_{\rm ad,2}$.
+%We can therefore match the fixed interior temperature profile to the exterior in order to determine $R\cb$ as a function of $\nabla_{\rm ad,2}$.
+Note, however, that \Eq{eq:Tdeep} is only valid to order of magnitude at $R_{\rm switch}$.
+
+%n contrast with the deep interior, however, we cannot rigorously apply \Eq{eq:Tdeep} at $R_{\rm switch}$, although the approximation is still valid to order unity.
+
+%We can use 
+The RCB temperature, $T\cb$, is an order unity correction from the disk temperature that depends on the adiabatic gradient in the outer envelope, $\nabla_{\rm ad,2}$.
+We have shown in Paper I that $T\cb \simeq T\di (1- 2 \nabla_{\rm ad, 2})^{-1/2}$ (for $\beta=2$ in Equation \ref{eq:opacitylaw}).
+This approximation agrees with the numerically calculated $T\cb$ shown in Figure \ref{fig:varying_delad} to within 5\%.
+Note that $T\cb$ modestly increases with $\nabla_{\rm ad,2}$ (upper-left panel), and is constant for fixed $\nabla_{\rm ad,2}$  regardless of $\nabla_{\rm ad,1}$ (upper-right panel).
+
+\begin{figure}[H]
 \centering
@@
 \label{fig:varying_delad}
-\end{figure*}
-
-\subsection{Role of $L$}\label{sec:roleofL}
-
-%The radiative and convective profiles must match at the RCB, constrained by the known total mass of the atmosphere
-
-%The EOS variations in the cold outer envelope affect $\dot{M}$ by decreasing the atmosphere's luminosity compared to an ideal gas polytrope (Figure \ref{fig:all_plot}, upper-right panel, for the same parameters as Figure \ref{fig:tplotall}).
-From Figures \ref{fig:deladmap} and \ref{fig:all_plot}, bottom-left panel, we see that the adiabatic gradient of the low-$T$ realistic EOS increases with radius (i.e., with decreasing temperature) throughout the convective zone when $T<500$ K, then becomes lower near the RCB.
-This results in a larger adiabatic gradient, on average, for the low-$T$ realistic EOS in the outer parts of the convective region of the atmosphere.
-
-%The variations in $\delad$ at low temperatures
-
-The increase in $\delad$ at low temperatures, where $H_2$ rotational states are not fully excited (Figure \ref{fig:all_plot}, upper-right) decreases the atmosphere's luminosity.
-%compared to an ideal gas polytrope 
-This result may be understood as consequence of matching between the atmosphere's interior convective and exterior radiative profiles.
-The profiles must match at the RCB, constrained by the known total mass.
-This match determines the pressure at the RCB, $P\cb \propto \exp(\RB/R\cb)$.
-Higher $P\cb$ implies lower luminosity (Equation \ref{eq:delrad}).
-
-The adiabatic gradient of EOS 2 is larger, on average, than for EOS 1 in the outer part of the convective region.\footnote {$\delad$ for EOS 2 increases with radius (i.e., with decreasing temperature) throughout the convective zone, then becomes lower near the RCB.} To understand how a variable, but overall larger $\delad$ in the outer atmosphere affects evolution, we study the simplified problem of atmospheres composed of an ideal gas with adiabatic gradient
-
-\begin{equation}
-\label{eq:deladvar}
-\delad = \left\{
-\begin{array}{l l}
-\nabla_{\rm ad, 1}, & \quad T > 500 \text{ K} \\
-\nabla_{\rm ad, 2}, & \quad T < 500 \text{ K}
-\end{array} 
-\right.
-\end{equation}   
-where $\nabla_{\rm ad, 1}$ and $\nabla_{\rm ad,2}$ are constant (Figure \ref{fig:varying_delad}).
-%We explore the independent effects of varying $\nabla_{\rm ad,1}$ and $\nabla_{\rm ad,2}$ by alternately fixing one and varying the other   (Figure \ref{fig:varying_delad}). 
-%The adiabatic gradient $\nabla_{\rm ad,1}$ dictates the temperature profile deep in the convective interior, while $\nabla_{\rm ad,2}$ sets the RCB temperature $T\cb$. 
-%Given these inputs, the RCB location and its pressure $P\cb \propto \exp(\RB/R\cb)$ (see Figure \ref{fig:varying_delad}, lower panels) are determined by matching the interior and exterior profiles.
-Higher $P\cb$ implies lower luminosity (Equation \ref{eq:delrad}).
-
-%the location of the RCB, is set by the matching between the convective interior and the outer radiative zone. As $P(r) \propto \exp(\RB/r)$ in the radiative region,
-
-%As $P(r) \propto \exp(\RB/r)$ in the outer radiative region,
-
-%Figure \ref{fig:varying_delad}, lower-left panel, shows the radial pressure profile for $\nabla_{\rm ad,1}=0.3$ and various $\nabla_{\rm ad,2}$, for the same parameters as Figures \ref{fig:tplotall} and \ref{fig:all_plot}, which are in agreement with the analytic expectation of an exponential profile.
-
-%The numerical solutions reproduce the analytic expectation of a nearly exponential pressure profile in the radiative region, $P(r) \propto \exp(\RB/r)$.
-Thus $L \propto 1/P\cb \propto 1/\exp(\RB/R\cb)$ is directly correlated to the physical depth of the RCB.
-
-%The location of the RCB is set by the matching between the convective interior and the outer radiative zone.
-
-The adiabatic gradient $\nabla_{\rm ad,1}$ dictates the temperature profile deep in the convective interior.
-Deep in the atmosphere, where $r \ll R\cb$, virial equilibrium yields a temperature profile
-\begin{equation}
-\label{eq:Tdeep}
-T(r) \approx \frac{\nabla_{\rm ad,1} G M\co}{\mathcal{R} r}.
-\end{equation}   
-Even though $\nabla_{\rm ad,1}$ is an order unity coefficient, the temperature scaling with $\nabla_{\rm ad,1}$ in \Eq{eq:Tdeep} is exact (Figure \ref{fig:varying_delad}, upper-right panel).
-Because the radial temperature profile in the atmospheric interior is set by $\nabla_{\rm ad,1}$, the radius $R_{\rm switch}$ at which the adiabatic gradient changes shows little variation with $\nabla_{\rm ad,2}$.
-%We can therefore match the fixed interior temperature profile to the exterior in order to determine $R\cb$ as a function of $\nabla_{\rm ad,2}$.
-Note, however, that \Eq{eq:Tdeep} is only valid to order of magnitude at $R_{\rm switch}$.
-
-%n contrast with the deep interior, however, we cannot rigorously apply \Eq{eq:Tdeep} at $R_{\rm switch}$, although the approximation is still valid to order unity.
-
-%We can use 
-The RCB temperature, $T\cb$, is an order unity correction from the disk temperature that depends on the adiabatic gradient in the outer envelope, $\nabla_{\rm ad,2}$.
-We have shown in Paper I that $T\cb \simeq T\di (1- 2 \nabla_{\rm ad, 2})^{-1/2}$ (for $\beta=2$ in Equation \ref{eq:opacitylaw}).
-This approximation agrees with the numerically calculated $T\cb$ shown in Figure \ref{fig:varying_delad} to within 5\%.
-Note that $T\cb$ modestly increases with $\nabla_{\rm ad,2}$ (upper-left panel), and is constant for fixed $\nabla_{\rm ad,2}$  regardless of $\nabla_{\rm ad,1}$ (upper-right panel).
+\end{figure}
 
@@
 %\textbf{t vs. M at fixed distance, similar to the plot from paper I. Compare scalings.}
-
-Figure \ref{fig:tvsMplot} displays the time evolution and the runaway growth time for atmospheres forming around cores with masses between 15 $M_{\oplus}$ and 25 $M_{\oplus}$ at $a=10$ AU in our fiducial disk, for a realistic EOS with an equilibrium ortho-to-para ratio and standard ISM opacity.
-Higher mass cores have shorter $t_{\rm run}$, consistent with the results of Paper I.
-We also note that $M_{\rm atm}$ at $t_{\rm run}$ is larger for the realistic EOS than for the polytropic EOS considered in Paper I, for the same core mass.
-
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
 %\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/t_vs_M_10au.pdf}
-\includegraphics[width=0.5\textwidth]{figures/t_vs_M_10au.pdf}
-%%\vspace{-0.5in}
+\includegraphics[width=0.65\textwidth]{figures/t_vs_M_10au.pdf}
+\vspace{-0.35in}
 \caption{Elapsed time as a function of atmosphere mass, for cores with fixed masses between $15 M_{\oplus}$ and $25 M_{\oplus}$ at $a=10$ AU in our fiducial disk, for a realistic EOS with an equilibrium ortho-to-para ratio and standard ISM opacity.
@@
 
-%\subsection{Critical Core Mass}
-%\label{Mcrit}
-
-%\textbf{Mcrit vs. a plot, realistic EOS and polytrope. Discuss the larger critical core mass for the real EOS in light of the effects from section 4.}
+Figure \ref{fig:tvsMplot} displays the time evolution and the runaway growth time for atmospheres forming around cores with masses between 15 $M_{\oplus}$ and 25 $M_{\oplus}$ at $a=10$ AU in our fiducial disk, for a realistic EOS with an equilibrium ortho-to-para ratio and standard ISM opacity.
+Higher mass cores have shorter $t_{\rm run}$, consistent with the results of Paper I.
+We also note that $M_{\rm atm}$ at $t_{\rm run}$ is larger for the realistic EOS than for the polytropic EOS considered in Paper I, for the same core mass.
 
@@
 
- %As such, non-ideal effects substantially increase the core mass needed to form a giant planet  before the dissipation of the protoplanetary disk.
-
-%Figure \ref{fig:Mvsaplot} shows the critical core mass for a massive atmosphere to form during a typical lifetime of a protoplanetary disk $t=3$ Myrs 
-%(e.g., \citealt{jay99}), 
-%for a gas described by a realistic EOS and an ISM dust opacity.
-The results of Paper I for an ideal diatomic gas are plotted for comparison.
-The inclusion of realistic EOS effects increases $M_{\rm crit}$ by more than a factor of two when compared to an ideal gas polytrope.
-%As such, non-ideal effects substantially increase the core mass needed to form a giant planet  before the dissipation of the protoplanetary disk.
-
-\begin{figure}[h!]
+Figure \ref{fig:Mcritvsagg} shows $M_{\rm crit}$ as a function of semimajor axis, for a realistic EOS with an equilibrium ortho-to-para ratio and grain growth opacity with a size distribution given by Equation (\ref{eq:graindistr}) with $p=3.5$ and maximum particle size $s_{\rm max}=1$ cm.
+The critical core mass is lower than in the standard interstellar opacity case, and less sensitive to location in the disk.
+Location primarily affects the atmosphere through the opacity in the outer envelope, which depends on disk temperature.
+For the simplified analytic model developed in Paper I, we approximated $t_{\rm run}$ by  $t_{\rm co}$, the time when $M_{\rm atm}=M\co$, and found that $t_{\rm co} \sim T_{\rm d}^{\beta+1/2}$, with $\beta$ the power-law exponent in Equation (\ref{eq:opacitylaw}).
+Opacity is less sensitive to temperature variations for larger grains and has an almost flat profile (see Figure \ref{fig:opacity}), which results in $\beta \ll 1$ and a much weaker temperature (and therefore semimajor axis) dependence of $M_{\rm crit}$, as seen in Figure \ref{fig:Mcritvsagg}.
+Moreover, grain growth reduces the absolute value of the opacity, which also lowers $M_{\rm crit}$.
+% As noted in Section \S\ref{sec:opacity}, the critical core mass may be significantly lower if coagulation is taken into account, i.e.
+$p=2.5$.
+
+%\subsection{Critical Core Mass}
+%\label{Mcrit}
+
+%\textbf{Mcrit vs. a plot, realistic EOS and polytrope. Discuss the larger critical core mass for the real EOS in light of the effects from section 4.}
+\begin{figure}[H]
 \centering
 %\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/Mc_vs_a_poly_real_paper.pdf}
-\includegraphics[width=0.5\textwidth]{figures/Mc_vs_a_poly_real_paper.pdf}
+\includegraphics[width=0.7\textwidth]{figures/Mc_vs_a_poly_real_paper.pdf}
 %%\vspace{-0.5in}
@@
 
-\begin{figure}[h!]
+ %As such, non-ideal effects substantially increase the core mass needed to form a giant planet  before the dissipation of the protoplanetary disk.
+
+%Figure \ref{fig:Mvsaplot} shows the critical core mass for a massive atmosphere to form during a typical lifetime of a protoplanetary disk $t=3$ Myrs 
+%(e.g., \citealt{jay99}), 
+%for a gas described by a realistic EOS and an ISM dust opacity.
+The results of Paper I for an ideal diatomic gas are plotted for comparison.
+The inclusion of realistic EOS effects increases $M_{\rm crit}$ by more than a factor of two when compared to an ideal gas polytrope.
+%As such, non-ideal effects substantially increase the core mass needed to form a giant planet  before the dissipation of the protoplanetary disk.
+
+\begin{figure}[H]
 \centering
 %\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/Mcrit_vs_a_gg.pdf}
-\includegraphics[width=0.5\textwidth]{figures/Mcrit_vs_a_gg.pdf}
+\includegraphics[width=0.7\textwidth]{figures/Mcrit_vs_a_gg.pdf}
 %%\vspace{-0.5in}
@@
 
-Figure \ref{fig:Mcritvsagg} shows $M_{\rm crit}$ as a function of semimajor axis, for a realistic EOS with an equilibrium ortho-to-para ratio and grain growth opacity with a size distribution given by Equation (\ref{eq:graindistr}) with $p=3.5$ and maximum particle size $s_{\rm max}=1$ cm.
-The critical core mass is lower than in the standard interstellar opacity case, and less sensitive to location in the disk.
-Location primarily affects the atmosphere through the opacity in the outer envelope, which depends on disk temperature.
-For the simplified analytic model developed in Paper I, we approximated $t_{\rm run}$ by  $t_{\rm co}$, the time when $M_{\rm atm}=M\co$, and found that $t_{\rm co} \sim T_{\rm d}^{\beta+1/2}$, with $\beta$ the power-law exponent in Equation (\ref{eq:opacitylaw}).
-Opacity is less sensitive to temperature variations for larger grains and has an almost flat profile (see Figure \ref{fig:opacity}), which results in $\beta \ll 1$ and a much weaker temperature (and therefore semimajor axis) dependence of $M_{\rm crit}$, as seen in Figure \ref{fig:Mcritvsagg}.
-Moreover, grain growth reduces the absolute value of the opacity, which also lowers $M_{\rm crit}$.
-% As noted in Section \S\ref{sec:opacity}, the critical core mass may be significantly lower if coagulation is taken into account, i.e.
-$p=2.5$.
-
-\begin{figure}[h!]
+In Equation (\ref{eq:graindistr}), the coefficient $p=3.5$ corresponds to a standard collisional cascade.
+If coagulation is taken into account, the exponent $p$ can be approximated as $p=2.5$ \citep{dalessio01}.
+This results in a flatter and significantly lower opacity (see \App{radwindow}), which may substantially reduce $M_{\rm crit}$.
+However, we have found that our model breaks down for low core masses ($M\co \lesssim 3 M_{\oplus}$) under our assumption of constant luminosity in the outer radiative layer.
+Figure \ref{fig:p25p35} shows the runaway accretion time $t_{\rm run}$ as a function of semimajor axis for the lowest core mass for which our model is valid, $M\co=4 M_{\oplus}$.
+The runaway accretion time is more than one order of magnitude lower for $p=2.5$, which implies that $M_{\rm crit}$ may be, in fact, significantly lower than presented in Figure \ref{fig:Mcritvsagg}.
+In other words, grain growth can yield critical core masses up to an order of magnitude lower than in the case where interstellar opacities are used.
+
+\begin{figure}[H]
 \centering
 %\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/tco_vs_a_Mc4_comp.pdf}
-\includegraphics[width=0.5\textwidth]{figures/tco_vs_a_Mc4_comp.pdf}
+\includegraphics[width=0.7\textwidth]{figures/tco_vs_a_Mc4_comp.pdf}
 %%\vspace{-0.5in}
@@
 \end{figure}
-
-In Equation (\ref{eq:graindistr}), the coefficient $p=3.5$ corresponds to a standard collisional cascade.
-If coagulation is taken into account, the exponent $p$ can be approximated as $p=2.5$ \citep{dalessio01}.
-This results in a flatter and significantly lower opacity (see \App{radwindow}), which may substantially reduce $M_{\rm crit}$.
-However, we have found that our model breaks down for low core masses ($M\co \lesssim 3 M_{\oplus}$) under our assumption of constant luminosity in the outer radiative layer.
-Figure \ref{fig:p25p35} shows the runaway accretion time $t_{\rm run}$ as a function of semimajor axis for the lowest core mass for which our model is valid, $M\co=4 M_{\oplus}$.
-The runaway accretion time is more than one order of magnitude lower for $p=2.5$, which implies that $M_{\rm crit}$ may be, in fact, significantly lower than presented in Figure \ref{fig:Mcritvsagg}.
-In other words, grain growth can yield critical core masses up to an order of magnitude lower than in the case where interstellar opacities are used.
 
@@
 
- \begin{figure}[h]
+ \begin{figure}[H]
 \centering
 %\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/acc_rates_paper.pdf}
-\includegraphics[width=0.5\textwidth]{figures/acc_rates_paper.pdf}
+\includegraphics[width=0.7\textwidth]{figures/acc_rates_paper.pdf}
 %\vspace{-0.5in}
@@
 
- \begin{figure}[h]
+ \begin{figure}[H]
 \centering
 %\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/Mc_vs_a_poly_comp.pdf}
-\includegraphics[width=0.5\textwidth]{figures/Mc_vs_a_poly_comp.pdf}
+\includegraphics[width=0.7\textwidth]{figures/Mc_vs_a_poly_comp.pdf}
 %\vspace{-0.5in}

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/chapter3.tex
upd

@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/drift_timescales_betaS1_gas_acc_new2.pdf}
+\includegraphics[width=0.7\textwidth]{figures/drift_timescales_betaS1_gas_acc_new2.pdf}
 %\vspace{-0.5in}
@@
 \label{sec:snowlines}
-
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=0.7\textwidth]{figures/desorption_distance_passive_active_colorbar_test2.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both an irradiated disk (left panels) and an evolving disk (right panels).
-The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The desorption distance for a static disk is shown for comparison (dashed vertical and horizontal lines).
-The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
-For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.}
-\label{fig:snowlines}
-\end{figure*}
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=0.7\textwidth]{figures/desorption_distance_actual_vs_estimated_passive_active_new.pdf}
-%\vspace{-0.5in}
-\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text).
-The estimate is performed for an irradiated disk (left panels) and an evolving disk (right panels).
-The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
-\label{fig:an_vs_actual}
-\end{figure*}
 
@@
 
-\begin{figure*}[t!]
+Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm drift} \sim t_{\rm des}$, given an initial particle size.
+We can calculate this location analytically by equating Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and solving for $r=r_{\rm des}(s)$ for a given particle size $s$.
+Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically.
+We display this result for the range of particle sizes that desorb at a fixed distance in an irradiated and an evolving disk (see Figure \ref{fig:snowlines}).
+We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it deviates for particles larger than $s \gtrsim10$ cm.
+For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the irradiated disk in the Epstein drag regime), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
+Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/desorption_distance_passive_active_colorbar_test2.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both an irradiated disk (left panels) and an evolving disk (right panels).
+The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+The desorption distance for a static disk is shown for comparison (dashed vertical and horizontal lines).
+The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar.
+For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.}
+\label{fig:snowlines}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{figures/desorption_distance_actual_vs_estimated_passive_active_new.pdf}
+%\vspace{-0.5in}
+\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text).
+The estimate is performed for an irradiated disk (left panels) and an evolving disk (right panels).
+The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
+\label{fig:an_vs_actual}
+\end{figure}
+
+%This is due to the fact that large bodies lose their mass as they desorb until they reach the same size as the smaller, meter-sized particles, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
+
+\begin{figure}[H]
 \centering
@@
 \label{fig:s_t_a}
-\end{figure*}
-
-Intuitively, this fixed $r_{\rm des}$ should be the location in the disk for which $t_{\rm drift} \sim t_{\rm des}$, given an initial particle size.
-We can calculate this location analytically by equating Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and solving for $r=r_{\rm des}(s)$ for a given particle size $s$.
-Figure \ref{fig:an_vs_actual} shows $r_{\rm des}$ calculated analytically using the prescription above as a function of the actual desorption distance calculated numerically.
-We display this result for the range of particle sizes that desorb at a fixed distance in an irradiated and an evolving disk (see Figure \ref{fig:snowlines}).
-We notice that the analytic approximation accurately reproduces the numerical result for most cases of interest, but it deviates for particles larger than $s \gtrsim10$ cm.
-For small particles with $\tau_{\rm s} \ll 1$, $t_{\rm drift}$ is a power-law in $r$ (for our parameters, $t_{\rm drift} \propto r^{-1/14}$ for the irradiated disk in the Epstein drag regime), and the Equation set (\ref{eq:ddt}) has an explicit analytic solution (see \App{app:tdriftan}).
-Once particles are large enough so that $\tau_{\rm s} \sim 1$, $t_{\rm drift}$ has a more complicated dependence on $r$ (see Equation \ref{eq:ts}), and the coupled drift-desorption differential equations have to be integrated numerically to obtain an accurate result.
-
-%This is due to the fact that large bodies lose their mass as they desorb until they reach the same size as the smaller, meter-sized particles, thus eventually following the same evolutionary track as the meter-sized bodies and evaporating at the same location.
+\end{figure}
 
@@
 Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.
-%, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance.
-%A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk.
-[not sure you need to explain this]
+
 The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started.
@@
 
-\begin{figure}[h!]
+Given our results in Section \ref{sec:snowlines}, a disk's C/O ratio in mainly affected by the snowline location for the particle size housing the most mass in ice.
+%We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
+Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then estimate the C/O ratio throughout the disk as a function of the particle size.
+Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
+In Figure \ref{1fig:CO_ratio}, we display the H$_2$O, CO$_2$, and CO snowline locations as a function of particle size for disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
+The minimum snowline distance for a disk is given by the curve corresponding to the maximum particle size it hosts.
+For grains that have grown to radii larger than $\sim$7 m and are able to drift and desorb, the $\sim$7 m snowline applies (see Section 3).
+%Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.
+
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.55\textwidth]{figures/C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
+\includegraphics[width=0.7\textwidth]{figures/C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
 %\vspace{-0.5in}
@@
 %, with the decrease more significant interior to the H$_2$O snowline.}
-\label{fig:CO_ratio}
+\label{1fig:CO_ratio}
 \end{figure}
-
-Given our results in Section \ref{sec:snowlines}, a disk's C/O ratio in mainly affected by the snowline location for the particle size housing the most mass in ice.
-%We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
-Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then estimate the C/O ratio throughout the disk as a function of the particle size.
-Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
-In Figure \ref{fig:CO_ratio}, we display the H$_2$O, CO$_2$, and CO snowline locations as a function of particle size for disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.
-The minimum snowline distance for a disk is given by the curve corresponding to the maximum particle size it hosts.
-For grains that have grown to radii larger than $\sim$7 m and are able to drift and desorb, the $\sim$7 m snowline applies (see Section 3).
-%Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.
 
@@
 
-Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for an irradiated disk, an evolving disk, and a viscous disk, under the simplifying assumption that the abundance relative to hydrogen for each volatile is fixed, so that drift and accretion affect only the locations of the snowlines.
+Figure \ref{1fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for an irradiated disk, an evolving disk, and a viscous disk, under the simplifying assumption that the abundance relative to hydrogen for each volatile is fixed, so that drift and accretion affect only the locations of the snowlines.
 We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.
@@
 As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines.
-For example, for the disk parameters and particle sizes displayed in Figure \ref{fig:CO_ratio}, water molecules in solid particles drift up to $\sim$1000 times faster across the H$_2$O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.
+For example, for the disk parameters and particle sizes displayed in Figure \ref{1fig:CO_ratio}, water molecules in solid particles drift up to $\sim$1000 times faster across the H$_2$O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.
 This differential inward motion will result in an increased oxygen gas abundance inside the H$_2$O snowline, and thus a (in some cases much) lower gaseous C/O ratio in this region.
@@
 
-Figure \ref{fig:CO_ratio} plots snowline curves for particle sizes $\sim0.5$ cm $\lesssim $s$ \lesssim$ 7m.
+Figure \ref{1fig:CO_ratio} plots snowline curves for particle sizes $\sim0.5$ cm $\lesssim $s$ \lesssim$ 7m.
 In the outermost disk, H$_2$O, CO$_2$, and CO all solidify.
@@
 We thus conclude that the C/O ratio interior to the H$_2$O and CO$_2$ snowlines in our evolving and viscous disks will be lower than in the static disk, due to the additional oxygen added to the gas by desorbing H$_2$O and CO$_2$.
-For these particle sizes, our calculated C/O ratio is an upper limit,  as indicated by the arrows in Figure \ref{fig:CO_ratio}.
+For these particle sizes, our calculated C/O ratio is an upper limit,  as indicated by the arrows in Figure \ref{1fig:CO_ratio}.
 
@@
 For the evolving disk, the gas-phase C/O ratio may decrease everywhere interior to the H$_2$O and CO$_2$ snowlines due to the decline in the surface density of solids with time at any given radius.
-For the viscous disk, the solid abundances at a fixed radius are constant, given that this model is not time-dependent, but the gas-to-solid ratio is not constant, which can result in a substantially lower C/O ratio interior to the H$_2$O and CO$_2$ snowlines compared to the static case (as indicated by the arrows in Figure \ref{fig:CO_ratio}).
-The main goal of Figure \ref{fig:CO_ratio}, however, is to show the different snowline radii in static and non-static disks, and therefore the locations in the disk where the gas-phase C/O ratio is reduced or increased, rather than provide a quantitative estimate of the magnitude of the C/O increase or decrease.
+For the viscous disk, the solid abundances at a fixed radius are constant, given that this model is not time-dependent, but the gas-to-solid ratio is not constant, which can result in a substantially lower C/O ratio interior to the H$_2$O and CO$_2$ snowlines compared to the static case (as indicated by the arrows in Figure \ref{1fig:CO_ratio}).
+The main goal of Figure \ref{1fig:CO_ratio}, however, is to show the different snowline radii in static and non-static disks, and therefore the locations in the disk where the gas-phase C/O ratio is reduced or increased, rather than provide a quantitative estimate of the magnitude of the C/O increase or decrease.
 
@@
 %only the kilometer-sized boulders that are very close to the snownlines will contribute to determining the snowline location. 
-Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
+Thus the innermost snowlines (depicted in red in Figure \ref{1fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star.
 For a grain size distribution with a maximum particle size different than our model, one can pick out the appropriate minimum snowline locations from this plot.%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).
@@
 
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=\textwidth]{figures/time_plots.pdf}
-%\vspace{-0.5in}
-\caption{Desorption or final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
-The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve).
-For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
-\label{fig:timeplots}
-\end{figure*}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/desorption_distance_varying_Md.pdf}
-%\vspace{-0.5in}
-\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines).
-The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve).
-A lower disk mass does not change the snowline location.}
-\label{fig:varMd}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/desorption_distance_transition_disk_1000.pdf}
-%\vspace{-0.5in}
-\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines).
-The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}.
-Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge, due to being trapped in a pressure maximum.}
-\label{fig:cavity}
-\end{figure}
-
 \subsection{Generality of Results: Dependence on Disk Parameters}
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{figures/time_plots.pdf}
+%\vspace{-0.5in}
+\caption{Desorption or final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels).
+The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve).
+For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
+\label{fig:timeplots}
+\end{figure}
+
 We choose as a fiducial model a total disk mass $M=0.1 M_{\odot}$.
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/desorption_distance_varying_Md.pdf}
+%\vspace{-0.5in}
+\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines).
+The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve).
+A lower disk mass does not change the snowline location.}
+\label{fig:varMd}
+\end{figure}
+
 %This result also has implications for transition disks, which have inner cavities significantly depleted of dust (e.g., \citealt{espaillat12}).
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/desorption_distance_transition_disk_1000.pdf}
+%\vspace{-0.5in}
+\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines).
+The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}.
+Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge, due to being trapped in a pressure maximum.}
+\label{fig:cavity}
+\end{figure}
+
 %[Need a sentence about how this makes the model framework also generally valid for more complicated disk structures.]
@@
 We note that our future work will address some of these issues.
+
+\begin{table}[H]
+\caption{The effects of dynamical and chemical processes on snowline shapes and locations}
+\begin{center}
+\begin{tabular}{|l|l|}\hline
+\textbf{Process} & \textbf{Effect} \\\hline
+Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star compared to the static snowline, $\rightarrow$ means that the snowline is pushed further from the host star compared to the static snowline.
+The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
+Gas accretion & $\leftarrow$ \footnote{Gas accretion pushes the snowlines inwards compared to the snowline locations in a static disk.
+However, accretional heating may push the snowline outwards compared to an evolving disk without accretion heating.} \\\hline
+Particle growth & $\leftarrow$ \footnote{As stated in the main text, if particles grow to km-sizes and above, the snowline remains the same as that of a static disk.} \\\hline
+Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
+Particle fragmentation & $\rightarrow$ $\leftarrow$ \\\hline
+Grain morphology & $\rightarrow$ \\\hline
+Particle composition & $\rightarrow$ $\leftarrow$ \\\hline
+Disk gaps and holes & $\rightarrow$ \\\hline
+Accretion rate evolution & $\rightarrow$ $\leftarrow$ \\\hline
+Stellar luminosity evolution & $\leftarrow$ \\\hline
+Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
+\end{tabular}
+
+\end{center}
+\end{table}
 
@@
 
-\begin{table}[t!]
-\caption{The effects of dynamical and chemical processes on snowline shapes and locations}
-\begin{center}
-\begin{tabular}{|l|l|}\hline
-\textbf{Process} & \textbf{Effect} \\\hline
-Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star compared to the static snowline, $\rightarrow$ means that the snowline is pushed further from the host star compared to the static snowline.
-The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
-Gas accretion & $\leftarrow$ \footnote{Gas accretion pushes the snowlines inwards compared to the snowline locations in a static disk.
-However, accretional heating may push the snowline outwards compared to an evolving disk without accretion heating.} \\\hline
-Particle growth & $\leftarrow$ \footnote{As stated in the main text, if particles grow to km-sizes and above, the snowline remains the same as that of a static disk.} \\\hline
-Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
-Particle fragmentation & $\rightarrow$ $\leftarrow$ \\\hline
-Grain morphology & $\rightarrow$ \\\hline
-Particle composition & $\rightarrow$ $\leftarrow$ \\\hline
-Disk gaps and holes & $\rightarrow$ \\\hline
-Accretion rate evolution & $\rightarrow$ $\leftarrow$ \\\hline
-Stellar luminosity evolution & $\leftarrow$ \\\hline
-Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
-\end{tabular}
-
-\end{center}
-\end{table}
 %causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.
@@
 The fragmentation of meter- to km-sized particles will move the snowlines outwards, as smaller particles desorb faster and further out from the host star (cf.
-Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}).
+Figures \ref{fig:snowlines} and \ref{1fig:CO_ratio}).
 Large boulders, which neither drift nor desorb, may become e.g.

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/chapter4.tex
upd

@@
 \section{Introduction}
-\label{sec:intro}
+\label{1sec:intro}
 
@@
 Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
-This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{sec:static}).
+This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{1sec:static}).
 
@@
 This paper is organized as follows.
-In Section \ref{sec:review}, we review the drift-desorption model developed in Paper I.
-We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain morphology and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{sec:results}.
-We address the implications of our results in Section \ref{sec:discussion} and summarize our findings in Section \ref{sec:summary}.
+In Section \ref{1sec:review}, we review the drift-desorption model developed in Paper I.
+We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain morphology and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{1sec:results}.
+We address the implications of our results in Section \ref{1sec:discussion} and summarize our findings in Section \ref{1sec:summary}.
 
@@
 \section{Coupled Drift-Desorption Model}
-\label{sec:review}
+\label{1sec:review}
 
 We begin with a brief review of Paper I's model for the effect of radial drift and viscous gas accretion on volatile snowline locations.
-We review our disk model in Section \ref{sec:disktime}, and summarize our numerical method and results in Section \ref{sec:driftdes}.
+We review our disk model in Section \ref{1sec:disktime}, and summarize our numerical method and results in Section \ref{1sec:driftdes}.
 
 \subsection{Disk Model}
-\label{sec:disktime}
+\label{1sec:disktime}
 
@@
 \begin{equation}
-\label{eq:diskT}
+\label{1eq:diskT}
 T = 120\, (r/\text{AU})^{-3/7} \,\,\text{K},
@@
 \begin{equation}
-\label{eq:activeT}
+\label{1eq:activeT}
 T^4 =\Big[\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}\Big]^4 + T_{\rm irr}^4,
 \end{equation}
-where $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}).
+where $T_{\rm irr}=T$ from Equation (\ref{1eq:diskT}).
 Here $G$ is the gravitational constant, $\kappa_0=2 \times 10^{-6}$ is a dimensionless opacity coefficient, $M_*=M_{\odot}$ is the mass of the central star, $\mu=2.35$ is the mean molecular weight of the nebular gas, $m_{\rm p}$ is the proton mass, $\Omega_{\rm k}=\sqrt{G M_{\odot}/r^3}$ is the Keplerian angular velocity, $\alpha=0.01$ is a dimensionless coefficient (see below for details), $k_{\rm B}$ is the Boltzmann constant, and $\sigma$ is the Stefan-Boltzmann constant.
@@
 The steady-state disk has an $\alpha$-viscosity prescription, where the kinematic viscosity is $\nu=\alpha c H$.
-Here $c \equiv \sqrt{k_{\rm B} T /(\mu m_{\rm p})}$ is the isothermal sound speed (with $T$ from Equation \ref{eq:activeT}), and $H \equiv c/\Omega_{\rm k}$ is the disk scale height.
+Here $c \equiv \sqrt{k_{\rm B} T /(\mu m_{\rm p})}$ is the isothermal sound speed (with $T$ from Equation \ref{1eq:activeT}), and $H \equiv c/\Omega_{\rm k}$ is the disk scale height.
 We can then determine the gas surface density for a viscous disk as (\citealt{shakura73}; see also Paper I for a more detailed explanation of these calculations):
 \begin{equation}
-\label{eq:Sigmaact}
+\label{1eq:Sigmaact}
 \Sigma=\frac{\dot{M}}{3 \pi \nu}.
@@
 \subsection{Desorption-Drift Equations and Results}
-\label{sec:driftdes}
+\label{1sec:driftdes}
 
@@
 \begin{subeqnarray}
-\label{eq:ddt}
-\frac{ds}{dt} &= & - \frac{3 \mu_x m_{\rm p}}{\rho_{\rm s}} N_x R_{\rm des, x}  \slabel{eq:dsdt} \\
-\frac{dr}{dt} &=& \dot{r} \slabel{eq:drdt},
+\label{1eq:ddt}
+\frac{ds}{dt} &= & - \frac{3 \mu_x m_{\rm p}}{\rho_{\rm s}} N_x R_{\rm des, x}  \slabel{1eq:dsdt} \\
+\frac{dr}{dt} &=& \dot{r} \slabel{1eq:drdt},
 \end{subeqnarray}
@@
 \begin{equation}
-\label{eq:Rdes}
+\label{1eq:Rdes}
 R_{\rm{des}, x} = \nu_x \exp{(-E_x/T_{\rm grain})},
@@
 where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}=T$ is the grain temperature (assumed to be the same as the disk temperature, see Paper I), and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well.
-We discuss our choices for $E_x$ for the different volatile species in Section \ref{sec:snowlines}.
+We discuss our choices for $E_x$ for the different volatile species in Section \ref{1sec:snowlines}.
 
@@
 \begin{equation}
-\label{eq:rdotact}
+\label{1eq:rdotact}
 \dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{\dot{r}_{\rm gas}}{1+\tau_{\rm s}^2},
@@
 \begin{equation}
-\label{eq:ts}
+\label{1eq:ts}
 t_{\rm s}= \left\{
@@
 where $\rho$ is the disk mid-plane density, $\lambda$ is the mean free path and Re is the Reynolds number.
-The gas accretion velocity $\dot{r}_{\rm gas}$ is determined from $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, for a fixed $\dot{M}$ and with $\Sigma$ given by Equation (\ref{eq:Sigmaact}).
+The gas accretion velocity $\dot{r}_{\rm gas}$ is determined from $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, for a fixed $\dot{M}$ and with $\Sigma$ given by Equation (\ref{1eq:Sigmaact}).
 
-For a particle of initial size $s_0$, we solve the Equation set (\ref{eq:ddt}) with the initial conditions $s(t_0)=s_0$ and $r(t_0)=r_0$, where $t_0$ is the time at which we start the integration and $r_0$ is the particle's initial location.
+For a particle of initial size $s_0$, we solve the Equation set (\ref{1eq:ddt}) with the initial conditions $s(t_0)=s_0$ and $r(t_0)=r_0$, where $t_0$ is the time at which we start the integration and $r_0$ is the particle's initial location.
 We stop our simulation after $t_{\rm d}=3$ Myr, the disk lifetime, since this is roughly the timescale on which planets form, and determine the desorption timescale $t_{\rm des}$ from $s(t_{\rm des})=0$, and thus a particle's desorption distance $r_{\rm des}=r(t_{\rm des})$.
@@
 \section{Results}
-\label{sec:results}
+\label{1sec:results}
 
 \subsection{Snowlines in a Static Disk: The Importance of Ice Morphology}
-\label{sec:snowlines}
+\label{1sec:snowlines}
 
-As we note in Section \ref{sec:intro}, the disk volatile composition and the ice morphology determine the location of important snowlines.
+As we note in Section \ref{1sec:intro}, the disk volatile composition and the ice morphology determine the location of important snowlines.
 In this work we focus on the primary carbon, oxygen and nitrogen carriers, i.e.
@@
 \subsection{C/N/O Ratios in Static Disks}
-\label{sec:static}
+\label{1sec:static}
 
@@
 \subsection{C/N/O Ratios in Dynamic Disks}
-\label{sec:dynamic}
+\label{1sec:dynamic}
 
@@
 
-Here we use the model of Section \ref{sec:review} to estimate the movement of the CO and N$_2$ snowlines for different grain morphologies in a viscous disk.
+Here we use the model of Section \ref{1sec:review} to estimate the movement of the CO and N$_2$ snowlines for different grain morphologies in a viscous disk.
 Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above.
@@
 \section{Discussion}
-\label{sec:discussion}
+\label{1sec:discussion}
 
@@
 It is thus possible, in principle, to use measurements of the N$_2$ abundance in Jupiter-family comets to determine where the N$_2$ snowline was located in our Solar system.
-However due to the uncertainty in the calculated location of the N$_2$ snowline (see Section \ref{sec:dynamic}), more detailed modeling is needed.
+However due to the uncertainty in the calculated location of the N$_2$ snowline (see Section \ref{1sec:dynamic}), more detailed modeling is needed.
 
@@
 \section{Summary}
-\label{sec:summary}
+\label{1sec:summary}
 

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/appendixA.tex
upd

@@
 
-To derive  the global energy equation (\ref{eq:coolingglobal}) for an embedded protoplanet, we generalize the analogous calculations in stellar structure theory, e.g.\ in \S4.3 of \citet{kippenhahn90}.
+To derive  the global energy equation (\ref{1eq:coolingglobal}) for an embedded protoplanet, we generalize the analogous calculations in stellar structure theory, e.g.\ in \S4.3 of \citet{kippenhahn90}.
 For our problem, we add the effects of finite core radius, surface pressure and mass accretion.
-We start with the local energy equation (\ref{eq:structd}), whose more natural form in Lagrangian (mass) coordinates is $\p L/ \p m = \epsilon - T \p S /\p t$.
+We start with the local energy equation (\ref{1eq:structd}), whose more natural form in Lagrangian (mass) coordinates is $\p L/ \p m = \epsilon - T \p S /\p t$.
 Integrating from the core to a higher shell with enclosed mass $M$ gives:
@@
 \end{equation}
-which follows from \Eqsss{eq:structb}{eq:structa}{eq:Eg} by integrating hydrostatic balance in Lagrangian coordinates.
+which follows from \Eqsss{1eq:structb}{1eq:structa}{1eq:Eg} by integrating hydrostatic balance in Lagrangian coordinates.
 As an aside, the integral in equation (\ref{eq:virial}) can be evaluated for a polytropic EOS to give simple expressions for the total energy:
@@
 %where the volumes, $V_M = 4 \pi R^3/3$ and $V\co = 4 \pi R\co^3/3$. 
-The first integral in \Eq{eq:EGdot} is the one we want, but the second one must be eliminated.  The time derivative of \Eq{eq:Eg} (times four) gives
+The first integral in \Eq{eq:EGdot} is the one we want, but the second one must be eliminated.  The time derivative of \Eq{1eq:Eg} (times four) gives
 \begin{subeqnarray}
@@
 \end{eqnarray} 
-Combining \Eqsss{eq:DLc}{eq:udot}{eq:PdVint}, we reproduce \Eq{eq:coolingglobal} with the accreted specific energy $e_M \equiv u_M - GM/R$.
+Combining \Eqsss{eq:DLc}{eq:udot}{eq:PdVint}, we reproduce \Eq{1eq:coolingglobal} with the accreted specific energy $e_M \equiv u_M - GM/R$.
 

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/appendixB.tex
upd

@@
 
-We consider the structure of a non-self-gravitating, isothermal atmosphere that extends outward from the radiative-convective boundary (RCB) and matches onto the disk density, $\rho_{\rm d}$, at a distance $r_{\rm fit} = n_{\rm fit} \RB$, where $R_{\rm B}$ is the Bondi radius defined in equation (\ref{eq:RB}).
-From equation (\ref{eq:structa}) the resulting density profile is
+We consider the structure of a non-self-gravitating, isothermal atmosphere that extends outward from the radiative-convective boundary (RCB) and matches onto the disk density, $\rho_{\rm d}$, at a distance $r_{\rm fit} = n_{\rm fit} \RB$, where $R_{\rm B}$ is the Bondi radius defined in equation (\ref{1eq:RB}).
+From equation (\ref{1eq:structa}) the resulting density profile is
 \begin{equation} \label{eq:rhoiso}
@@
 
-The mass of the atmosphere is determined by integrating equation (\ref{eq:structb}) from the RCB to the Bondi radius using the density profile (\ref{eq:rhoiso}) and can be approximated as
+The mass of the atmosphere is determined by integrating equation (\ref{1eq:structb}) from the RCB to the Bondi radius using the density profile (\ref{eq:rhoiso}) and can be approximated as
 \begin{equation} \label{eq:MatmISO}
@@
 %\subsection{Temperature Contrast at Convective Boundary}
-We estimate the temperature and pressure corrections at the RCB due to the fact that the radiative region is not purely isothermal. From equation (\ref{eq:delrad}), we express the radiative lapse rate
+We estimate the temperature and pressure corrections at the RCB due to the fact that the radiative region is not purely isothermal.
+From equation (\ref{1eq:delrad}), we express the radiative lapse rate
 \begin{equation}\label{eq:delradan}
@@
 
-\noindent where the second equality follows from the opacity law (\ref{eq:opacitylaw}) and $\nabla_{\rm d}$ is the radiative temperature gradient at the disk:
+\noindent where the second equality follows from the opacity law (\ref{1eq:opacitylaw}) and $\nabla_{\rm d}$ is the radiative temperature gradient at the disk:
 
@@
 \end{equation} 
- We can then determine the RCB radius $R\cb$ from \Eq{eq:structa} as 
+ We can then determine the RCB radius $R\cb$ from \Eq{1eq:structa} as 
 \begin{equation}\label{eq:RCBint}
@@
 As opposed from the temperature correction factor $\chi$, an analytic expression for  $\theta$ cannot be obtained.
-Estimates for $\chi$ and $\theta$ for different values of the exponent $\beta$ in the opacity law (\ref{eq:opacitylaw}) are presented in Table 1.
+Estimates for $\chi$ and $\theta$ for different values of the exponent $\beta$ in the opacity law (\ref{1eq:opacitylaw}) are presented in Table 1.
 
@@
 \label{surfterms}
-In this section we check the relevance of the neglected surface terms in \Eq{eq:coolingglobal}.
+In this section we check the relevance of the neglected surface terms in \Eq{1eq:coolingglobal}.
 We first show that accretion energy is only a small correction at the RCB, which is where we apply our cooling model.
@@
 
-\begin{table}{cccccc}  % <--- column justification (center/left/right)
+\begin{deluxetable}{cccccc}  % <--- column justification (center/left/right)
 \gdef \numcols {6}
@@
 \tablenotetext{a}{Since $\delad = \nabla_\infty$ there is no convective transition at depth for this case.}
-\end{table}
+\end{deluxetable}
 

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/chapter4.tex
upd

@@
 Observations that anchor snowline locations at different stages of planet formation are therefore key to develop C/N/O ratios as a probe of planet formation zones.
+
+\noindent The contents of this chapter have been submitted for publication in \textit{Piso, A.-M. A., Pegues, J., \& \"Oberg, K.I. 2016, ApJ, submitted}
 
@@
 
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=0.95\textwidth]{figures/CNO_and_snowlines_2.pdf}
-%\vspace{-0.5in}
-\caption{The total carbon, nitrogen and oxygen abundance in solids as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and water dominated ices (bottom panel).
-Relevant volatile snowlines are marked by the vertical dashed lines.
-The grain abundances are calculated as a function of the observed median CH$_4$ and NH$_3$ abundances in protostellar cores.
-The total grain abundance increases with semimajor axis as more and more species freeze out.}
-\label{fig:CNOstatic}
-\end{figure*}
-
 Following \citet{chiang10} and \citet{birnstiel12}, a particle's radial drift velocity can be approximated as 
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/CNO_and_snowlines_2.pdf}
+%\vspace{-0.5in}
+\caption{The total carbon, nitrogen and oxygen abundance in solids as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and water dominated ices (bottom panel).
+Relevant volatile snowlines are marked by the vertical dashed lines.
+The grain abundances are calculated as a function of the observed median CH$_4$ and NH$_3$ abundances in protostellar cores.
+The total grain abundance increases with semimajor axis as more and more species freeze out.}
+\label{fig:CNOstatic}
+\end{figure}
+
 We determine the location of the H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}.
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/C_O_ratio_CH4.pdf}
+\includegraphics[width=0.7\textwidth]{figures/C_O_ratio_CH4.pdf}
 %\vspace{-0.5in}
@@
 
-\begin{figure}[h!]
+Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk, for our three choices of the NH$_3$ abundance: no NH$_3$ (top panel), NH$_3$-mid (middle panel) and NH$_3$-max (bottom panel).
+For comparison, the horizontal dotted lines show the average N/O ratio in the disk.
+As expected, the gaseous N/O ratio generally exhibits an increasing trend towards the outer disk as more oxygen gas is depleted, with small decreases between the NH$_3$ and CO$_2$ snowlines (by 6\% for NH$_3$-mid and by 18\% for NH$_3$-max, respectively) due to NH$_3$ freeze-out.
+While the presence of NH$_3$ only moderately affects our results for the N/O ratio, NH$_3$ is important since otherwise the nitrogen content in solid bodies would be more depleted than is observed for comets and asteroids (\citealt{wyckoff91}, \citealt{mumma11}, \citealt{bergin15}).
+
+The gas-phase N/O ratio is enhanced by a factor of two outside the H$_2$O snowline compared to its average value, by more than a factor of three between the CO$_2$ and CO snowlines, and by orders of magnitude between the CO and N$_2$ snowlines.
+This latter region can span tens of AU depending on disk parameters and the relative CO and N$_2$ ice binding environment.
+This N/O  enhancement is more pronounced than the C/O gas phase enhancement of a factor of two in the outer disk (see Figure \ref{fig:COstatic}).
+%The N/O ratio reaches particularly high values between the CO and N$_2$ snowlines, where all the oxygen is now contained in grains.
+
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/N_O_ratio.pdf}
+\includegraphics[width=0.7\textwidth]{figures/N_O_ratio.pdf}
 %\vspace{-0.5in}
@@
 
-Figure \ref{fig:Nstatic} shows the snowline locations of the main oxygen and nitrogen carriers and the N/O ratio in gas and dust as a function of semimajor axis in a static disk, for our three choices of the NH$_3$ abundance: no NH$_3$ (top panel), NH$_3$-mid (middle panel) and NH$_3$-max (bottom panel).
-For comparison, the horizontal dotted lines show the average N/O ratio in the disk.
-As expected, the gaseous N/O ratio generally exhibits an increasing trend towards the outer disk as more oxygen gas is depleted, with small decreases between the NH$_3$ and CO$_2$ snowlines (by 6\% for NH$_3$-mid and by 18\% for NH$_3$-max, respectively) due to NH$_3$ freeze-out.
-While the presence of NH$_3$ only moderately affects our results for the N/O ratio, NH$_3$ is important since otherwise the nitrogen content in solid bodies would be more depleted than is observed for comets and asteroids (\citealt{wyckoff91}, \citealt{mumma11}, \citealt{bergin15}).
-
-The gas-phase N/O ratio is enhanced by a factor of two outside the H$_2$O snowline compared to its average value, by more than a factor of three between the CO$_2$ and CO snowlines, and by orders of magnitude between the CO and N$_2$ snowlines.
-This latter region can span tens of AU depending on disk parameters and the relative CO and N$_2$ ice binding environment.
-This N/O  enhancement is more pronounced than the C/O gas phase enhancement of a factor of two in the outer disk (see Figure \ref{fig:COstatic}).
-%The N/O ratio reaches particularly high values between the CO and N$_2$ snowlines, where all the oxygen is now contained in grains.
-
 \subsection{C/N/O Ratios in Dynamic Disks}
@@
 
-Here we use the model of Section \ref{1sec:review} to estimate the movement of the CO and N$_2$ snowlines for different grain morphologies in a viscous disk.
-Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above.
-We assume there is no carbon in the form of CH$_4$.
-The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.05$ cm do not form true snowlines.
-If the CO binding environment is known, the CO snowline moves inward by up to $\sim$50 \% compared to a static disk for each case (pure and water dominated ices) due to disk dynamics.
-The full range of potential CO snowlines taking into account both ice morphology and disk dynamics span $\sim$8.7 AU to $\sim$61 AU, which is a factor of $\sim$7 difference.
-This implies that gas phase C/O ratios of order unity may be reached in the giant planet forming zone, and the CO snowline may be inside 10 AU for certain disk parameters.
-
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/C_O_water_ice.pdf}
+\includegraphics[width=0.75\textwidth]{figures/C_O_water_ice.pdf}
 %\vspace{-0.5in}
@@
 
+Here we use the model of Section \ref{1sec:review} to estimate the movement of the CO and N$_2$ snowlines for different grain morphologies in a viscous disk.
+Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above.
+We assume there is no carbon in the form of CH$_4$.
+The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.05$ cm do not form true snowlines.
+If the CO binding environment is known, the CO snowline moves inward by up to $\sim$50 \% compared to a static disk for each case (pure and water dominated ices) due to disk dynamics.
+The full range of potential CO snowlines taking into account both ice morphology and disk dynamics span $\sim$8.7 AU to $\sim$61 AU, which is a factor of $\sim$7 difference.
+This implies that gas phase C/O ratios of order unity may be reached in the giant planet forming zone, and the CO snowline may be inside 10 AU for certain disk parameters.
+
 %As in the case for CO, the N$_2$ binding energy is also strongly dependent on the ice environment in which the grain resides.
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/N_O_water_ice_many.pdf}
+\includegraphics[width=0.75\textwidth]{figures/N_O_water_ice_many.pdf}
 %\vspace{-0.5in}
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/CN_norm.pdf}
+%\vspace{-0.5in}
+\caption{Gas phase C/O (blue curve) and N/O (red curve) ratios divided by the average C/O and N/O ratio in a static disk, assuming CO and N$_2$ are pure ices, and there is no CH$_4$ or NH$_3$.
+The dashed vertical lines mark the H$_2$O, CO$_2$, CO and N$_2$ snowlines.
+The arrow indicates that the N/O ratio is enhanced by orders of magnitude compared to its average value between the CO and N$_2$ snowlines.
+The gaseous N/O ratio is enhanced throughout most of the disk, and more enhanced than the C/O ratio.}
+\label{fig:enhance}
+\end{figure}
+
 Theoretical models of the magnitude and role of N/O (and N/C) ratios in exoplanet atmospheres are needed in order to use these ratios as probes for a planet's formation location.
@@
 Moreover, Figure \ref{fig:enhance} shows that giant planets that have formed in situ between the H$_2$O and CO snowlines are expected to present elevated both C/O and N/O ratios in their atmospheres, whereas planets between the CO and N$_2$ snowlines will have a highly enhanced N/O ratio in their atmospheres, but not C/O.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/CN_norm.pdf}
-%\vspace{-0.5in}
-\caption{Gas phase C/O (blue curve) and N/O (red curve) ratios divided by the average C/O and N/O ratio in a static disk, assuming CO and N$_2$ are pure ices, and there is no CH$_4$ or NH$_3$.
-The dashed vertical lines mark the H$_2$O, CO$_2$, CO and N$_2$ snowlines.
-The arrow indicates that the N/O ratio is enhanced by orders of magnitude compared to its average value between the CO and N$_2$ snowlines.
-The gaseous N/O ratio is enhanced throughout most of the disk, and more enhanced than the C/O ratio.}
-\label{fig:enhance}
-\end{figure}
 

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/appendixD.tex
upd

@@
-\section{Adiabatic Gradient Variations}
+\chapter{Adiabatic Gradient Variations}
 \label{alldelad}
 
-\subsection{Adiabatic Gradient during Partial Dissociation}\label{deladdiss}
+\section{Adiabatic Gradient during Partial Dissociation}\label{deladdiss}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centering
 %\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/delad_dissociation.pdf}
-\includegraphics[width=0.5\textwidth]{figures/delad_dissociation.pdf}
+\includegraphics[width=0.6\textwidth]{figures/delad_dissociation.pdf}
 %%\vspace{-0.5in}
@@
 
-\subsection{Adiabatic Gradient during Conversion of Spin Isomers}
+\section{Adiabatic Gradient during Conversion of Spin Isomers}
 \label{deladspin}
@@
 In contrast, the heat capacity for the 3:1 ortho-para ratio mixture will be a weighted average between the individual ortho- and para- components, and will hence have intermediate values between the two, as displayed in Figure \ref{fig:ucvr}.
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/ortho_para_energy.pdf}
+\includegraphics[width=0.6\textwidth]{figures/ortho_para_energy.pdf}
+%%\vspace{-0.5in}
+\caption{Internal energy per unit mass and specific heat capacity associated with rotation for parahydrogen (dashed blue), orthohydrogen (dotted green), the equilibrium mixture (solid red) and a fixed 3:1 ortho-to-para ratio (dash-dotted light blue) as a function of temperature.
+After \citet{farkas35}, Figure 1.}
+\label{fig:ucvr}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-\centering
-%\includegraphics[width=0.5\textwidth]{../../figs/ModelAtmospheres/RadSelfGravRealEOS/PaperFigs/ortho_para_energy.pdf}
-\includegraphics[width=0.5\textwidth]{figures/ortho_para_energy.pdf}
-%%\vspace{-0.5in}
-\caption{Internal energy per unit mass and specific heat capacity associated with rotation for parahydrogen (dashed blue), orthohydrogen (dotted green), the equilibrium mixture (solid red) and a fixed 3:1 ortho-to-para ratio (dash-dotted light blue) as a function of temperature.
-After \citet{farkas35}, Figure 1.}
-\label{fig:ucvr}
-\end{figure}
-
-\begin{figure}[h]
+\begin{figure}[H]
 \centering

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/conclusion.tex
upd

--- 2016-04-13 Dissertate-Harvard-LaTeX/chapters/conclusion.tex
@@
 While for an ideal gas EOS $\delad$ is constant, non-ideal EOS effects cause features in the adiabatic gradient that change the atmospheric structure.
-At high temperatures, molecular hydrogen dissociates, while at low-temperature the rotational states of the hydrogen molecule are only partially excited so it no longer behaves like an ideal gas, due to the existence of H$_2$ in two spin isomeric forms, ortho- and parahydrogen.
+At high temperatures, molecular hydrogen dissociates, while at low-temperatures the rotational states of the hydrogen molecule are only partially excited so it no longer behaves like an ideal gas, due to the existence of H$_2$ in two spin isomeric forms, ortho- and parahydrogen.
 Both of these effects increase $M_{\rm crit}$ by a factor of $\sim$2 compared to the ideal gas.
@@
 We applied our model for a range of initial particle sizes and calculated how drift and gas accretion affect their desorption location.
-We have found that there is a range of particle sizes, $0.05$ cm $\lesssim s \lesssim$ 7 m for our particular parameters, that desorb almost instantaneously and a fixed particle-size dependent location from the central star, regardless of their initial position.
+We have found that there is a range of particle sizes, $0.05$ cm $\lesssim s \lesssim$ 7 m for our particular parameters, that desorb almost instantaneously and at a fixed particle-size dependent location from the central star, regardless of their initial position.
 Based on this information, we calculated the H$_2$O, CO$_2$ and CO snowline locations for different particle sizes in this range, and made estimates for the C/O ratio throughout the disk.
@@
 Our work has made important strides in our understudying of disk and planet compositions and the tight link between the two, as well as the effect of disk dynamics and chemistry on disk and planet compositions.
-As first step, we plan to expand our work by considering the effect of diffusion on the C/N/O ratios in viscous steady-state disks.
+As a first step, we plan to expand this work by considering the effect of diffusion on the C/N/O ratios in viscous steady-state disks.
 Similarly to the model of \citet{owen14}, we will develop a simplified method to estimate the abundance of different volatiles at various disk locations, and thus their effect on the C/N/O ratio both in static and dynamic disks.
-We plan to explore further processes that affect snowline locations, such as variations in stelar luminosity and gas mass accretion rate, as well as particle growth and fragmentation.
-Our next step is to incorporate disk chemical processes, first by developing a simplified parametrized chemical  network that we will incorporate in our  drift-desorption model.
+
+Through analytical and numerical calculations, we will explore a range of dynamical processes that may affect snowline locations and the distribution of volatiles in disks, expanding and generalizing the framework developed in Chapter 4 and 5.
+Such effects include particle growth and fragmentation, as well as variations in the gas mass accretion rate and stellar luminosity.
+
+The complexity of disk chemistry means that coupling it with dynamical processes, while necessary, is non-trivial.
+As a next step we will thus couple the dynamical framework outlined above with time-dependent chemical models of increasing complexity, informed by results from state-of-the-art disk chemistry models (that can only be run on static disks).
 By having a better understudying on how disk chemistry and dynamics affect the composition of nascent, and eventually mature planets, our work may provide essential context for characterizing the gas giants

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/introduction.tex
upd

@@
 
-We can start uncovering the answers to these questions through a combination of studying planet formation in protoplanetary disks, the end-results of planet formation (i.e., exoplanets), and through comparisons with the architecture and composition of our own Solar System.
-Within the last two decades, more than one thousand extrasolar planets (exoplanets)
-have been discovered \citet{batalha14}. Their diversity in terms of mass, radius, location and composition
-\citet{lissauer14} provides an exciting field of research, with the eventual goal of finding planets that are
+We can start uncovering the answers to these questions through a combination of studying planet formation in protoplanetary disks, the end-results of planet formation (i.e., exoplanets), and the architecture and composition of our own Solar System.
+In the last two decades, more than one thousand extrasolar planets (exoplanets)
+have been discovered \citep{batalha14}. Their diversity in terms of mass, radius, location and composition
+\citep{lissauer14} provides an exciting field of research, with the eventual goal of finding planets that are
 similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore
@@
 comparison (both in mass and radius), which makes it difficult to obtain atmospheric spectra
-and find out what chemical compounds they are made of. We therefore turn to giant planets,
-which have provided a rich and intriguing research area for decades. Gas giants contain most
-of their mass in their atmosphere, hence their chemical composition is determined by that
-of their envelopes. The last few years have seen a substantial increase in the number of
+and find out what chemical compounds they are made of.
+
+To obtain constraints on how planets obtain their compositions, we instead turn to giant planets. Gas giants contain most
+of their mass in their atmosphere, hence their chemical composition is largely determined by that
+of their envelopes. These envelopes are observable with existing telescopes. The last few years have seen a substantial increase in the number of
 giant planets with observed atmospheric spectra (e.g., \citealt{debes13}, \citealt{kreidberg15}), which has enhanced our
-understanding of these planets’ chemical structure, and has provided us with quantitative
+understanding of these planets' chemical structure, and has provided us with quantitative
 information about the abundances of various compounds in their envelopes besides hydrogen
-and helium. Finally, gas giants shape the architecture of planetary systems and affect the
-delivery of volatile compunds to terrestrial planets, which has direct consequences for the
+and helium. Gas giant formation is also important to constrain since gas giants shape the architecture of planetary systems and affect the
+delivery of volatile compounds to terrestrial planets, which has direct consequences for the
 habitability of worlds similar to our own. Thus testing theories of planet formation against
@@
 
+\section{The Core Accretion Mechanism and Its Challenges}
+
+Gas giants are widely believed to form through core accretion (e.g., \citealt{pollack96}), a theory in which
+solid protoplanetary cores grow large enough to accumulate a massive atmosphere. In this
+model, planetesimals in a disk grow larger through collisions, eventually forming a planetary
+embryo, which continues to grow by attracting planetesimals in its neighborhood. Once an
+embryo becomes large enough so that its escape velocity exceeds the thermal velocity of the
+nebular gas in its vicinity, it starts accumulating a gaseous envelope. From this point on, the
+accretion of gas is regulated by the pressure support within the envelope --- the amount of
+gas a core can accumulate is limited by the atmosphere's ability to radiate away the energy
+due to the incoming planetesimals, as well as by envelope contraction.
+
+Protoplanetary disks dissipate on relatively short timescales of a few Myr \citep{williams11}.
+This implies that cores must grow fast in order to become large enough to attract massive atmospheres, and thus they need high planetesimal accretion rates.
+Studies that consider such high and constant rates (e.g., \citealt{stevenson82}, \citealt{boden86}, \citealt{wuchterl93}, \citealt{rafikov06}) find that the envelope accumulating around the planet is in steady state at all times, since all the energy due to the incoming planetesimals is radiated away by the atmosphere.
+It follows that the core and envelope grow simultaneously, and the mass of the atmosphere is a function of the core mass.
+Once the core and atmosphere have attained comparable masses, a rapid phase of runaway gas accretion starts and the gas giant can form.
+In this steady state scenario, there is therefore a uniquely determined core mass at which runaway accretion commences, called the critical core mass $M_{\rm crit}$.
+High planetesimal accretion rates cause the core to grow quickly, but this poses an additional challenge: fast accretion heats up the core's atmosphere, increases pressure support and inhibits the ability of the envelope to cool and contract, thus eventually increasing $M_{\rm crit}$.
+This is particularly challenging in the outer disk, where long dynamical times prevent large cores from forming before disk dissipation.
+\citet{rafikov11} shows that core accretion cannot operate beyond 40-50 AU at the accretion rates required to form the core and the atmosphere on the same timescale.
+
+%\begin{figure}[t!]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/HR8799.png}
+%%\vspace{-0.5in}
+%\caption{The directly imaged HR 8799 system. Planets are located at 14.5 AU, 24 AU, 38 AU and 68 AU. From \citet{marois10}. }
+%\label{fig:HR8799}
+%\end{figure}
+
+All this seems to indicate that core accretion cannot work in the outer disk.
+This poses an intriguing question: could wide-separation gas giants, such as the HR 8799 system, have formed through core accretion?
+We can start answering this question by noting that planetesimal accretion does not need to be constant with time at a given location in the disk, which has been shown to be a viable scenario (e.g., \citealt{ikoma00}, \citealt{pollack96}).
+Figure 1 from \citet{pollack96} shows an example of this scenario.
+At early times, the core is in a high planetesimal accretion rate regime and grows fast, but the atmosphere remains small in comparison.
+As the planet depletes its feeding zone of solids, planetesimal accretion is reduced.
+From this point on, the core no longer grows significantly, and thus the energy due to planetesimals can no longer balance radiative losses by the atmosphere; now the envelope accumulates gas while undergoing Kelvin-Helmholtz contraction, until its mass approximately exceeds the core mass and runaway gas accretion begins.
+Since now the envelope grows with time, there is no longer a uniquely determined $M_{\rm crit}$ --- in this scenario, the critical core mass is that of a core than can accrete an atmosphere with a mass equal to its own on a timescale equal to the disk lifetime.
+
+%\begin{figure}[t!]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/pollack.png}
+%%\vspace{-0.5in}
+%\caption{The evolution of a nascent planet's core (denoted with the subscript Z), atmosphere (denoted with the subscript XY), and total mass (denoted with the subscript p) as a function of time, for a planet forming at 5.2 AU through core accretion.
+The plot depicts the three phases of the planet's evolution: (1) core growth while atmosphere remains small (until $\sim1.2$ Myr), (2) atmospheric growth while core growth is practically stalled (until $\sim$7 Myr), and (3) runaway gas accretion.}
+%\label{fig:pollack}
+%\end{figure}
+
+Such studies potentially provide a viable solution to the core accretion challenge at wide separations, but they are primarily focused on explaining the formation of Jupiter at 5.2 AU and thus do not explore a larger radii parameter space.
+Moreover, in order to truly understand whether core accretion can indeed work at stellocentric distances significantly larger than that of Jupiter, a more interesting question can be asked: what is the lowest possible core mass required to form a gas giant before disk dissipation?
+This minimum core mass is determined when planetesimal accretion has fully stopped, since additional accretion would actually increase this core mass (see above).
+Moreover, how does this minimum core mass depend on disk location, the properties of the nebular gas, and the opacity of the dust grains?
+We provide an answer to these questions in Chapters 2 and 3.
+
 \section{Dynamics and Chemistry in Protoplanetary Disks}
 
-Protoplanetary disks are the result of the collapse of a molecular cloud, and are structures composed of gas and dust that are rotationally supported and that represent the birth environment of planets (Figure \ref{fig:proto}).
-Disks are complex objects, whose structure and evolution are affected by a multitude of chemical and dynamical processes.
+Protoplanetary disks are the result of the collapse of a molecular cloud.
+They are structures composed of gas and dust that are rotationally supported and that represent the birth environment of planets (Figure \ref{fig:proto}).
+Despite their simple geometry, disks are complex objects, whose structure and evolution are affected by a multitude of chemical and dynamical processes.
 The latter include transport of both dust through radial drift and gas through viscous gas accretion, as well as grain growth and dust settling to the disk midplane, which affect the disk structure.
@@
 Moreover, the timescales for various disk chemical and dynamical processes may be comparable at least in some parts of the disk \citep{semenov11}, which makes the coupling between chemistry and dynamics very challenging.
-As a first step it is thus easier to isolate the different dynamical and chemical processes that occur in the disk midplane where planets are born, and understand their role and relative importance in shaping disk and planet compositions.
-We explore some of these processes in Chapters 4 and 5, and investigate their consequences for the compositions of giant planets forming at different radii.
+
+Not all disk regions (and therefore all disk chemistry) are equally important to understand the disk-planet connection, however.
+Planets form in the midplane of disks, where the chemical structure is mainly set by condensation and sublimation of volatiles.
+As a first step we focus on the different dynamical and chemical processes that occur in the disk midplane, with the goal to understand their role and relative importance in shaping disk and planet compositions.
 
@@
 %\vspace{-0.5in}
-\caption{The standard picture of molecular cloud to protostar to protoplanetary disk to planet formation. Reprinted by permission from Karin \"Oberg.}
+\caption{The standard picture of molecular cloud to protostar to protoplanetary disk to planet formation. Reprinted with permission from \citet{oberg09}.}
 \label{fig:proto}
@@
 
-\subsection{Snowlines and Disk Dynamics}
+\subsection{Disk Snowlines}
 
-Volatile compounds, which have low sublimation temperatures, are of particular importance.
-Their relative abundance in gaseous and solid form determines their snowline location, i.e.
-the distance in the disk where a volatile transitions from gaseous to solid form.
-Snowlines are essential in the planet formation process (e.g., \citealt{pontoppidan14}): they can enhance the amount of solid material just outside the snowline, due to dust sublimation, and they can produce pressure traps where dust particles pile up just inside the snowline.
-Observational advancements in recent year have provided us with an increasing amount of information regarding volatiles in disks.
+Volatile compounds, which have low sublimation temperatures, are of particular importance for this thesis since they set the gas and dust compositions in the gas giant formation zone in disks.
+The relative sublimation and condensation rates of the volatiles determine where in disks they transition from being primarily gaseous to being primarily condensed out onto grains.
+These transitions turn out to be sharp and are referred to as snowlines.
+Snowlines are essential in the planet formation process (e.g., \citealt{pontoppidan14}).
+Observational advancements in recent years have provided us with an increasing amount of information regarding volatiles in disks.
 Volatile compounds have been detected in disks (e.g., \citealt{henning13}), which gives us clues about their abundance in different disks and at different disk locations.
@@
 This snowline has been detected indirectly through its tracer molecule N$_2$H$^+$, which becomes highly abundant as CO freezes out.
-Moreover, the H$_2$O snowline has also been detected in Tw Hya \citep{zhang13}, as well as the CO snowline in HD 163296 \citep{qi15}.
+The H$_2$O snowline has also been detected in TW Hya \citep{zhang13}, as well as the CO snowline in HD 163296 \citep{qi15}.
 
@@
 %\vspace{-0.5in}
-\caption{ALMA detection of the CO snowline in TW Hya at 30 AU. Reprinted by permission from Karin \"Oberg.}
+\caption{ALMA detection of the CO snowline in TW Hya at 30 AU. Reprinted with permission from \citet{qi13}.}
 \label{fig:qi13}
@@
 \citet{oberg11} consider a static disk, and thus do not account for dynamical processes such as redistribution of solids due to radial drift, the radial movement of the nebular gas, and accretion heating.
-We thus expand this model by considering the processes outlined above and their effect on snowline locations.
+We expand this model by considering the processes outlined above and their effect on snowline locations.
 In particular, we focus on how radial drift of solids and viscous gas accretion onto the central star affect the H$_2$O, CO$_2$ and CO snowline locations for particles of different sizes.
-We find that drift and accretion heating alone may move the snowlines inward by factors of $\sim$2 compared to a static disk, thus substantially changing the disk regions with enhanced gas-phase C/O ratios.
+%We find that drift and accretion heating alone may move the snowlines inward by factors of $\sim$2 compared to a static disk, thus substantially changing the disk regions with enhanced gas-phase C/O ratios.
 %, which has direct consequences for planet formation.
@@
 Gas-phase C/O ratios of order unity can be achieved in the outer disk.
-Reprinted by permission from Karin \"Oberg.
+Reprinted with permission from \citet{oberg11}.
 }
@@
 
-As we show above, the C/O ratio is highly important to study, but it only provides us with one piece of the puzzle that is constraining planet formation origins based on their compositions.
+As we show above, the C/O ratio is highly important to study, but it only provides us with one piece of the puzzle that is disk gas and grain elemental compositions in dynamic disks.
 In order to improve our theoretical models, we thus need to look at other molecules and volatile compounds.
@@
 nitrogen-to-oxygen (N/O) ratio in the outer disk may thus be even more enhanced than the C/O ratio.
-We calculate the N/O ratio in gas and dust in Chapter 5, and confirm that it is indeed highly enhanced compared to the stellar value throughout most of the disk.
+%We calculate the N/O ratio in gas and dust in Chapter 5, and confirm that it is indeed highly enhanced compared to the stellar value throughout most of the disk.
 
@@
 
-\subsection{Ice Morphology}
+%\subsection{Ice Morphology}
 
 The snowline locations of volatiles, including carbon, oxygen and nitrogen carriers, are highly dependent on the morphology of the icy particles.
-These can either form on a dust mantle as pure ices, or have a layered structure with the more volatile species closer to the outer layers (e.g., \citealt{pontoppidan03}).
+These can either be mixed ices or layered ices, where each ice layer is pure (e.g., \citealt{pontoppidan03}).
 As H$_2$O is the least volatile species with high abundance in disks, one can typically expect the more volatile molecules to form on a water substrate (but see \citealt{bisschop06} for other scenarios).
+In some environments the more volatile species may form a thick enough layer on top of the water ices that they essentially behave as pure ices, however.
 The ice environment in which the molecules reside, i.e.
@@
 This effect is particularly important in the case of CO and N$_2$, since their volatility is significantly higher than that of other carbon and nitrogen carriers, such as CO$_2$ or NH$_3$.
-Laboratory experiments \citet{fayolle16} have determined new values for the CO and N$_2$ binding energies, both as pure ices and water dominated.
+Laboratory experiments \citep{fayolle16} have determined new values for the CO and N$_2$ binding energies, both as pure ices and water dominated.
 Figure \ref{fig:fayolle} shows the results of a temperature programmed desorption (TPD) experiment for the CO and N$_2$ desorption rate as a function of temperature, in water dominated and pure ice environments.
@@
 This translates into a difference in binding energies by up to a factor of two between pure and water dominated CO or N$_2$.
-We show in Chapter 5 that this large difference in binding energies changes the CO and N$_2$ snowline locations by factors of 3-4 spending on the ice environment.
+%We show in Chapter 5 that this large difference in binding energies changes the CO and N$_2$ snowline locations by factors of 3-4 spending on the ice environment.
 By taking into account also the effect of disk dynamics, we find that the CO and N$_2$ snowlines may span several tens of AU.
@@
 The peaks in the desorption curves are at substantially higher temperatures for the water dominated ices.
-Reprinted by permission from Karin \"Oberg.}
+Reprinted with permission from \citet{fayolle16}.}
 \label{fig:fayolle}
 \end{figure}
-
-\subsection{Consequences for the C/N/O Ratio in Planetary Atmospheres}
-
-Variations in the C/O ratio in gas and dust due to different volatile condensation temperatures have also been investigated in exoplanet atmospheres, both from a theoretical and observational standpoint.
-\citet{molliere15} show that varying the C/O ratio in a gas giant envelope by factors of $\sim$3 changes the abundance of other volatiles, such as H$_2$O and CH$_4$, by several orders of magnitude.
-From the observational perspective, an exciting potential discovery was that some planets, such as WASP-12 b, have a superstellar C/O ratio \citep{madhu11}.
-While this particular detection has since been unequivocally refuted (e.g., \citealt{kreidberg15}), the advent of JWST in the near future will provide us with a significantly larger sample of atmospheric spectra which may be used to constrain C/O ratios in other giant planet atmospheres.
-
-While observational constraints for the N/O ratio in planetary atmospheres do not currently exist, the high gas-phase N/O ratio enhancement at most disk radii suggests that giant planets that form at wide separations should have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
-Theoretical and observational knowledge of both C/O and N/O ratios could further help back-track the planet formation location based on planet composition.
-
-\section{The Core Accretion Mechanism and Its Challenges}
-
-Gas giants are widely believed to form through core accretion (e.g., \citealt{pollack96}), a theory in which
-solid protoplanetary cores grow large enough to accumulate a massive atmosphere. In this
-model, planetesimals in a disk grow larger through collisions, eventually forming a planetary
-embryo, which continues to grow by attracting planetesimals in its neighborhood. Once an
-embryo becomes large enough so that its escape velocity exceeds the thermal velocity of the
-nebular gas in its vicinity, it starts accumulating a gaseous envelope. From this point on, the
-accretion of gas is regulated by the pressure support within the envelope --- the amount of
-gas a core can accumulate is limited by the atmosphere's ability to radiate away the energy
-due to the incoming planetesimals, as well as by envelope contraction.
-
-Protoplanetary disks dissipate on relatively short timescales of a few Myr.
-This implies that cores must grow fast in order to become large enough to attract massive atmospheres, and thus they need high planetesimal accretion rates.
-Studies that consider such high and constant rates (e.g., \citealt{stevenson82}, \citealt{boden86}, \citealt{wuchterl93}, \citealt{rafikov06}) find that the envelope accumulating around the planet is in steady state at all times, since all the energy due to the incoming planetesimals is radiated away by the atmosphere.
-It follows that the core and envelope grow simultaneously, and the mass of the atmosphere is a function of the core mass.
-Once the core and atmosphere have attained comparable masses, a rapid phase of runaway gas accretion starts and the gas giant can form.
-In this steady state scenario, there is therefore a uniquely determined core mass at which runaway accretion commences, called the critical core mass $M_{\rm crit}$.
-High planetesimal accretion rates cause the core to grow quickly, but this poses an additional challenge: fast accretion heats up the core's atmosphere, increases pressure support and inhibits the ability of the envelope to cool and contract, thus eventually increasing $M_{\rm crit}$.
-This is particularly challenging in the outer disk, where long dynamical times prevent large cores from forming before disk dissipation.
-\citet{rafikov11} shows that core accretion cannot operate beyond 40-50 AU at the accretion rates required to form the core and the atmosphere on the same timescale.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics[width=0.8\textwidth]{figures/HR8799.png}
-%%\vspace{-0.5in}
-%\caption{The directly imaged HR 8799 system. Planets are located at 14.5 AU, 24 AU, 38 AU and 68 AU. From \citet{marois10}. }
-%\label{fig:HR8799}
-%\end{figure}
-
-All this seems to indicate that core accretion cannot work in the outer disk.
-This poses an intriguing question: could wide-separation gas giants, such as the HR 8799 system, have formed through core accretion?
-We can start answering this question by noting that planetesimal accretion does not need to be constant with time at a given location in the disk, which has been shown to be a viable scenario (e.g., \citealt{ikoma00}, \citealt{pollack96}).
-Figure 1 from \citet{pollack96} shows an example of this scenario.
-At early times, the core is in a high planetesimal accretion rate regime and grows fast, but the atmosphere remains small in comparison.
-As the planet depletes its feeding zone of solids, planetesimal accretion is reduced.
-From this point on, the core no longer grows significantly, and thus the energy due to planetesimals can no longer balance radiative losses by the atmosphere; now the envelope accumulates gas while undergoing Kelvin-Helmholtz contraction, until its mass approximately exceeds the core mass and runaway gas accretion begins.
-Since now the envelope grows with time, there is no longer a uniquely determined $M_{\rm crit}$ --- in this scenario, the critical core mass is that of a core than can accrete an atmosphere with a mass equal to its own on a timescale equal to the disk lifetime.
-
-%\begin{figure}[t!]
-%\centering
-%\includegraphics[width=0.8\textwidth]{figures/pollack.png}
-%%\vspace{-0.5in}
-%\caption{The evolution of a nascent planet's core (denoted with the subscript Z), atmosphere (denoted with the subscript XY), and total mass (denoted with the subscript p) as a function of time, for a planet forming at 5.2 AU through core accretion.
-The plot depicts the three phases of the planet's evolution: (1) core growth while atmosphere remains small (until $\sim1.2$ Myr), (2) atmospheric growth while core growth is practically stalled (until $\sim$7 Myr), and (3) runaway gas accretion.}
-%\label{fig:pollack}
-%\end{figure}
-
-Such studies potentially provide a viable solution to the core accretion challenge at wide separations, but they are primarily focused on explaining the formation of Jupiter at 5.2 AU and thus do not explore a larger radii parameter space.
-Moreover, in order to truly understand whether core accretion can indeed work at stellocentric distances significantly larger than that of Jupiter, a more interesting question can be asked: what is the lowest possible core mass required to form a gas giant before disk dissipation?
-This minimum core mass is determined when planetesimal accretion has fully stopped, since additional accretion would actually increase this core mass (see above).
-Moreover, how does this minimum core mass depend on disk location, the properties of the nebular gas, and the opacity of the dust grains?
-We provide an answer to these questions in Chapters 2 and 3, where we find that the minimum core mass for giant planet formation is lower than the one typically quoted, 10 $M_{\oplus}$ (\citealt{stevenson82}, \citealt{rafikov06}), and may be as low as 1 $M_{\oplus}$.
-Our study thus reopens the case for in situ formation of wide-separation gas giants through core accretion.
 
@@
 
-Our work in the two main parts of this thesis confirms that there is indeed a tight link between disks and planets.
+\subsection{Consequences of Different Disk C/N/O Ratios on Planetary Atmospheres}
+
+Variations in the C/O ratio in gas and dust due to different volatile condensation temperatures have also been investigated in exoplanet atmospheres, both from a theoretical and observational standpoint.
+\citet{molliere15} show that varying the C/O ratio in a gas giant envelope by factors of $\sim$3 changes the abundance of other volatiles, such as H$_2$O and CH$_4$, by several orders of magnitude.
+From the observational perspective, an exciting potential discovery was that some planets, such as WASP-12 b, have a superstellar C/O ratio \citep{madhu11}.
+While this particular detection has since been unequivocally refuted (e.g., \citealt{kreidberg15}), the advent of JWST in the near future will provide us with a significantly larger sample of atmospheric spectra which may be used to constrain C/O ratios in other giant planet atmospheres.
+
+While observational constraints for the N/O ratio in planetary atmospheres do not currently exist, the high gas-phase N/O ratio enhancement at most disk radii suggests that giant planets that form at wide separations should have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
+Theoretical and observational knowledge of both C/O and N/O ratios could further help back-track the planet formation location based on planet composition.
+
+\subsection{Thesis Outline}
+
+This thesis is organized as follows.
+In Chapter 2, we determine the minimum core mass for giant planet formation for a range of disk radii, using idealized assumptions about the properties of the nebular gas and of the dust grains.
+We expand this model in Chapter 3, where we calculate this minimum core mass assuming a realistic gas equation of state and realistic opacities.
+In Chapter 4, we explore the role of disk dynamics in setting the snowline locations of the main carbon and oxygen carriers, and the consequences for the C/O ratio in gas and dust throughout the disk, as well as for the compositions of giant planets forming at different radii.
+Finally, we expand this model in Chapter 5 by considering additional volatile species and different ice morphologies.
+
+%where we find that the minimum core mass for giant planet formation is lower than the one typically quoted, 10 $M_{\oplus}$ (\citealt{stevenson82}, \citealt{rafikov06}), and may be as low as 1 $M_{\oplus}$.
+Our study thus reopens the case for in situ formation of wide-separation gas giants through core accretion.
+
+Our work in the two main parts of this thesis (Chapters 2 \& 3, and 4 \& 5, respectively) confirms that there is indeed a tight link between disks and planets.
 Our study on core accretion shows that the core mass required to form a gas giant is highly dependent on the disk properties and composition, while our chapters on snowline locations and C/N/O ratios also demonstrate that the disk structure and chemical composition has direct effects on the volatile composition of giant planets atmospheres.

### 2016-04-15 Dissertate-Harvard-LaTeX/frontmatter/thanks.tex
upd

% the acknowledgments section

First and foremost, I would like to thank Karin \"Oberg for all the support she has provided me with in the past three years, one as my mentor and the last two as my thesis advisor. None of this would have been possible without her constant guidance and encouragement. I cannot think of a moment when she hasn't been responsive, be it to bounce off ideas, to give me feedback on a paper, to counsel me regarding professional advancement, or to offer incredible support and understanding during the tougher times. I've learnt from her not only how to be a successful scientist, but also what qualities make one an exceptional advisor and mentor. I am grateful that Karin joined the Harvard faculty three years ago; it's been a privilege to work with her.

Secondly, I would like to offer my thanks to Ruth Murray-Clay for advising me during the first few years of my PhD. She guided me through my first steps as a professional astronomer, and she always encouraged me to become more independent as a scientist and to not be afraid to speak my mind. When I get to teach my first astronomy class, I will pleasantly remember the many hours Ruth and I spent doing order of magnitude calculations in her office or at group meetings.

I'd like to give a special thanks to Rosanne Di Stefano, my faculty mentor for the past two years. I've immensely enjoyed our weekly chats over breakfast at Sarah's or Hi-Rise. I am glad I could benefit from both her wisdom and her kindness when I needed either professional or personal advice. Even in the darkest moments, Rosanne has managed to put a smile on my face and truly make me believe that things always get better.

There are many, many other people at the CfA whom I'd like to thank and who made my graduate experience both productive and enjoyable. In particular, I would like to thank Sean Andrews, Matt Holman and Dimitar Sasselov for accepting to be part of my Thesis Advising Committee and for guiding me step by step towards a successful thesis completion. Although outside the CfA, I would like to thank Fred Ciesla for flying over from Chicago to be part of my Defense Committee, and for sharing excellent research ideas when I visited the University of Chicago last fall. I've also appreciated the occasional but insightful conversations with Dave Charbonneau, John Johnson and Alicia Soderberg. A special thanks goes to Til Birnstiel, with whom I've enjoyed collaborating and who would be almost instantly available when I had a question or I was stuck debugging my code. I would also like to thank the astrochemistry and disk groups for often offering me helpful suggestions and ideas, and for teaching me a thing or two about chemistry and radio astronomy.

My life in the Astronomy department has been better due to my fellow graduate students. Meredith MacGregor and Courtney Dressing have been wonderful office mates. I have often been inspired by their intelligence and hard working abilities, as much as I've enjoyed going to a movie or an after work drink with them. Anjali Tripathi has been an immense support throughout my time here; I cannot count the late nights we spent in one of our offices trying to help each other with some work problem or just talking about life.

This section would not be complete if I didn't thank my friends outside of astronomy, who have been there for me willing to give me a hug in the good times and especially during the bad ones. My (more frequent than I would like to admit) gchats with Andra about everything and nothing always make my day. The weekly Saturday morning breakfast with Carlo and the evenings of Walking Dead marathons have been the best, as were the frequent froyo runs with Luly. I will miss checking out new restaurants with Laura and Iulia, as I will miss John's smile and contagious optimism.

Finally, but perhaps most importantly, I would like to thank my family for loving and supporting me throughout these last ten years in Boston, even though they are so far away. My dad was the first to trigger my interest in astronomy by teaching me the names of the planets in our Solar System when I was around five. I don't have enough words to thank my mom. She is the most amazing person that I know; I would not have gotten anywhere close to here without her.

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/chapter1.tex
upd
update

@@
 
-\noindent The contents of this chapter are published in \textit{Piso, A.-M. A \& Youdin, A. N. Y}
+\noindent The contents of this chapter are published in \textit{Piso, A.-M. A \& Youdin, A. N. 2014, ApJ, 786, 21}
 
@@
 This paper is organized as follows.
-Section \S\ref{sec:model} describes our model of atmospheric growth.
+Section \S\ref{sec:model1} describes our model of atmospheric growth.
 In Section \S\ref{sec:coolingan}, we develop a simplified analytic version of our atmospheric model.
@@
 
-\section{Atmospheric Accretion Model} \label{sec:model}
+\section{Atmospheric Accretion Model} \label{sec:model1}
 
@@
 
-We examine the structure and evolution of our model atmospheres, calculated as described in \S\ref{sec:model}.
+We examine the structure and evolution of our model atmospheres, calculated as described in \S\ref{sec:model1}.
 For comparison, we also show results of the non-self-gravitating analytic model of \S\ref{sec:coolingan}.

### 2016-07-18 tex/draft2/ms_revised.tex
upd

--- 2016-03-11 tex/draft2/ms_edited.tex
@@
 
-\title{The Role of Ice Compositions and Morphology for Snowlines and the C/N/O Ratios in Active Disks}
+\title{The Role of Ice Compositions for Snowlines and the C/N/O Ratios in Active Disks}
 
@@
 In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, their ice binding environment, and the presence of snowlines of major volatiles at different distances from the central star.
-We explore the effects of disk dynamical processes, molecular compositions and abundances, and ice morphology on the snowline locations of the main C, O and N carriers, and the C/N/O ratios in gas and dust throughout the disk.
+We explore the effects of disk dynamical processes, molecular compositions and abundances, and ice \textbf{compositions} on the snowline locations of the main C, O and N carriers, and the C/N/O ratios in gas and dust throughout the disk.
 The gas-phase N/O ratio enhancement in the outer disk (exterior to the H$_2$O snowline) exceeds the C/O ratio enhancement for all reasonable volatile compositions.
-Ice morphology and disk dynamics individually change the snowline location of N$_2$, the main nitrogen carrier, by a factor of 2-3, and when considered together the range of possible N$_2$ snowline locations is $\sim$11-$\sim$79 AU in a standard disk model.
+Ice \textbf{compositions} and disk dynamics individually change the snowline location of N$_2$, the main nitrogen carrier, by a factor of 2-3, and when considered together the range of possible N$_2$ snowline locations is $\sim$11-$\sim$79 AU in a standard disk model.
 Observations that anchor snowline locations at different stages of planet formation are therefore key to develop C/N/O ratios as a probe of planet formation zones.
@@
 In the case of nitrogen, chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
-Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}).
-Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
-Because of the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio compared to its average value in the disk.
-Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
+Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}), \textbf{although Titan's N$_2$ atmospheric composition is not primordial but rather originates from primordial NH$_3$ accreted by the satellite (\citealt{atreya78}, \citealt{mandt14}).}  %Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
+%\textbf{While CO and N$_2$ have similar condensation temperatures (e.g., \citealt{fayolle16}), their snowline locations may }  
+Due to the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio compared to its average value in the disk.
+\textbf{This effect is most pronounced between the CO and N$_2$ snowlines, where the gas phase N/O ratio is enhanced by many orders of magnitude compared to the stellar value due to the depletion of oxygen gas in this region (see Section \ref{sec:results}).} Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
 In addition to N$_2$, a fraction of the nitrogen abundance may also be carried by less volatile species such as NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).
 
-The snowline locations of the main carbon, oxygen and nitrogen carriers strongly depend on the ice grain morphology.
-Very volatile species, such as CO and N$_2$, present binding energies, and therefore snowline locations, that are sensitive to the details of the morphology of the icy grain mantles.
-Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
+The snowline locations of the main carbon, oxygen and nitrogen carriers strongly depend on the ice grain \textbf{composition}.
+Very volatile species, such as CO and N$_2$, present binding energies, and therefore snowline locations, that are sensitive to the details of the \textbf{composition} of the icy grain mantles.
+\textbf{Spectroscopic observations suggest that up to 90\% of the CO ice is frozen in a layer that is thick enough and separated from the H$_2$O ice layer underneath so that it can be considered pure ice (e.g., \citealt{pontoppidan03}).
+However, if the CO (or N$_2$) ice layer is thin enough, then it will interact with the H$_2$O ice substrate (e.g., \citealt{collings03}).
+The ice binding energy will be different in this water dominated environment than in the pure ice case.} Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
 This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{sec:static}).
+\textbf{In protostellar cores, H$_2$O ice is primarily amorphous (e.g., \citealt{williams02}, \citealt{vandishoeck14}), and can either be compact or porous.
+When temperatures exceed $\sim$80-90 K, H$_2$O ice acquires a crystalline structure \citep{schegerer10}.
+The CO binding energy is larger in an amorphous porous water ice environment than in the compact case \citep{fayolle16} and slightly lower than in the crystalline case \citep{noble12}.
+Here we consider pure ices and ices dominated by amorphous porous H$_2$O ice.
+We refer to the latter simply as water dominated ices unless noted otherwise.}
 
@@
 
-In this work, we expand the coupled drift-desorption model developed in (\citealt{piso15b}; hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, as well as nitrogen-to-oxygen (N/O) ratios.
+In this work, we expand the coupled drift-desorption model developed in (\citealt{piso15b}; hereafter Paper I) by considering additional volatile molecules and abundances, ice \textbf{compositions}, as well as nitrogen-to-oxygen (N/O) ratios.
 This paper is organized as follows.
 In Section \ref{sec:review}, we review the drift-desorption model developed in Paper I.
-We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain morphology and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{sec:results}.
+We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain \textbf{compositions} and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{sec:results}.
 We address the implications of our results in Section \ref{sec:discussion} and summarize our findings in Section \ref{sec:summary}.
@@
 \centering
-\includegraphics[width=0.95\textwidth]{CNO_and_snowlines_2.eps}
+\includegraphics[width=0.95\textwidth]{CNO_and_snowlines_2.pdf}
 %\vspace{-0.5in}
@@
 
-\subsection{Snowlines in a Static Disk: The Importance of Ice Morphology}
+\subsection{Snowlines in a Static Disk: The Importance of Ice \textbf{Compositions}}
 \label{sec:snowlines}
 
-As we note in Section \ref{sec:intro}, the disk volatile composition and the ice morphology determine the location of important snowlines.
+As we note in Section \ref{sec:intro}, the disk volatile composition and the ice \textbf{composition} determine the location of important snowlines.
 In this work we focus on the primary carbon, oxygen and nitrogen carriers, i.e.
@@
 The binding energies of H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ as pure ices are 5800 K, 2000 K, 834 K, 1300 K, 767 K and 2965 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}, \citealt{martin14}).
-For CO and N$_2$ as water dominated ices, the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}.
+\textbf{For our fiducial disk model, these energies correspond to disk temperatures of 143 K, 48 K, 21 K, 30 K, 18 K and 68 K, respectively.} For CO and N$_2$ as water dominated ices, the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}, \textbf{corresponding to disk temperatures of 34 K and 31 K, respectively}.
 Figure \ref{fig:CNOstatic} shows the resulting snowline locations, assuming CO and N$_2$ pure ices (top panel), and CO and N$_2$ in water dominated ices (bottom panel).
@@
 This changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ.
+%\textbf{We note that the CO and N$_2$ snowlines will move inward if the ices are in an amorphous compact water environment and slightly shift outward in the case of a crystalline H$_2$O ice substrate.}
 
@@
 In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and to what extent they are affected by the presence of CH$_4$ and NH$_3$ over the full range of observed CH$_4$ and NH$_3$ abundances toward low-mass protostars.
-In this section we only consider pure ices.
+In this section we only consider pure ices, \textbf{i.e.
+ices that are layered on a silicate mantle.
+In reality, the icy grain will have a layered structure with volatiles residing on top of a H$_2$O ice substrate.
+However, if the volatile ice layer is thick enough and separated from the H$_2$O ice layer, the interaction between H$_2$O and the other volatile species will be minimal and thus the ices can be considered pure rather than water dominated (see also Section \ref{sec:intro}).}
+
  %Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{C_O_ratio_CH4.eps}
+\includegraphics[width=0.5\textwidth]{C_O_ratio_CH4.pdf}
 %\vspace{-0.5in}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{N_O_ratio.eps}
+\includegraphics[width=0.5\textwidth]{N_O_ratio.pdf}
 %\vspace{-0.5in}
@@
 If the CO binding environment is known, the CO snowline moves inward by up to $\sim$50 \% compared to a static disk for each case (pure and water dominated ices) due to disk dynamics.
-The full range of potential CO snowlines taking into account both ice morphology and disk dynamics span $\sim$8.7 AU to $\sim$61 AU, which is a factor of $\sim$7 difference.
+The full range of potential CO snowlines taking into account both ice \textbf{compositions} and disk dynamics span $\sim$8.7 AU to $\sim$61 AU, which is a factor of $\sim$7 difference.
 This implies that gas phase C/O ratios of order unity may be reached in the giant planet forming zone, and the CO snowline may be inside 10 AU for certain disk parameters.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{C_O_water_ice.eps}
+\includegraphics[width=0.5\textwidth]{C_O_water_ice.pdf}
 %\vspace{-0.5in}
@@
 The presence of CO in a water ice environment rather than as pure ice moves the CO snowline significantly inward by $\sim$70\%.
-Taken together, disk dynamics and ice morphology move the CO snowline inward by a factor of $\sim$7.}
+Taken together, disk dynamics and ice \textbf{compositions} move the CO snowline inward by a factor of $\sim$7.}
 \label{fig:CO_ratio}
@@
 Thus for each case (pure versus water dominated ices), the N$_2$ snowline moves inward by up to 50\% due to disk dynamics.
-By taking into account both ice morphology and disk dynamics, the full range of potential N$_2$ snowlines span $\sim$11 to $\sim$79 AU, which is a factor of $\sim$7 difference.
+By taking into account both ice \textbf{compositions} and disk dynamics, the full range of potential N$_2$ snowlines span $\sim$11 to $\sim$79 AU, which is a factor of $\sim$7 difference.
 Similarly to the case for CO, the N$_2$ snowline may be close to 10 AU for certain disk models.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{N_O_water_ice_many.eps}
+\includegraphics[width=0.5\textwidth]{N_O_water_ice_many.pdf}
 %\vspace{-0.5in}
@@
 The presence of N$_2$ in a water ice environment rather than as pure ice moves the N$_2$ snowline significantly inward by $\sim$70\%.
-Taken together, disk dynamics and ice morphology move the N$_2$ snowline inward by a factor of $\sim$7.
+Taken together, disk dynamics and ice \textbf{compositions} move the N$_2$ snowline inward by a factor of $\sim$7.
 The results of an enhanced gas-phase N/O ratio between the H$_2$O and CO snowlines compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{CN_norm.eps}
+\includegraphics[width=0.5\textwidth]{CN_norm.pdf}
 %\vspace{-0.5in}
@@
 
-Due to disk dynamics and ice morphology, the locations of the CO and N$_2$ snowlines, and thus the disk regions with highly elevated gas phase N/O and C/O ratios, are uncertain and may span tens of AU.
+Due to disk dynamics and ice \textbf{compositions}, the locations of the CO and N$_2$ snowlines, and thus the disk regions with highly elevated gas phase N/O and C/O ratios, are uncertain and may span tens of AU.
 Both ice morphologies discussed in this study, pure and water dominated ices, are plausible in protoplanetary disks and depend on whether H$_2$O and CO ices formed on similar timescales or successively (e.g., \citealt{garrod11}).
@@
 Changes in stellar luminosity (e.g., \citealt{kennedy06}) and gas mass accretion rate (e.g., \citealt{chambers09}), as well as the evolution of icy dust particles due to grain growth and fragmentation (e.g., \citealt{birnstiel12}), may introduce additional uncertainties in the snowline locations, and thus the C/N/O ratios.
-Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints on the relative importance of ice morphology and disk dynamics in setting snowline locations.
-The uncertainties in snowline locations caused by disk dynamics, ice morphology, and other effects outlined above can be resolved in extreme cases, such as a detection of a CO snowline at a temperature corresponding to pure CO ice desorption in a static disk (e.g., \citealt{qi13} at $\sim$ 17 K) or CO desorption from a water dominated ice in a dynamic disk.
-In intermediate cases it is more difficult to resolve the relative importance of ice morphology and disk dynamics.
+\textbf{Moreover, the diffusion of vapor across the snowlines following the cold finger effect (\citealt{stevenson88}, \citealt{cyr98}) will change the shape of the C/O and N/O curves and therefore the magnitude of the C/N/O ratios between different snowlines.} Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints on the relative importance of ice \textbf{compositions} and disk dynamics in setting snowline locations.
+The uncertainties in snowline locations caused by disk dynamics, ice \textbf{compositions}, and other effects outlined above can be resolved in extreme cases, such as a detection of a CO snowline at a temperature corresponding to pure CO ice desorption in a static disk (e.g., \citealt{qi13} at $\sim$ 17 K) or CO desorption from a water dominated ice in a dynamic disk.
+In intermediate cases it is more difficult to resolve the relative importance of ice \textbf{compositions} and disk dynamics.
 For example, the CO snowline in HD 163296 is at a higher temperature of $\sim$ 25 K \citep{qi15}, which could be caused either by CO being in a water dominated environment or by dynamical effects that push the CO snowline inward.
@@
 A low N$_2$/CO ratio is consistent with comets having formed inside the N$_2$ snowline where N$_2$ is still in the gas phase.
-Theoretical models suggest that Jupiter-family comets, such as 67P, originate from the Kuiper belt (\citealt{duncan97}; but see \citealt{rubin15} for alternative formation scenarios for 67P).
+\textbf{However, it is also possible that the measured N$_2$ abundance in 67P may be due to post-formation processes such as radiogenic heating \citep{rubin15}, and thus may not reflect the comet's primordial composition.} Theoretical models suggest that Jupiter-family comets, such as 67P, originate from the Kuiper belt (\citealt{duncan97}; but see \citealt{rubin15} for alternative formation scenarios for 67P).
 It is thus possible, in principle, to use measurements of the N$_2$ abundance in Jupiter-family comets to determine where the N$_2$ snowline was located in our Solar system.
-However due to the uncertainty in the calculated location of the N$_2$ snowline (see Section \ref{sec:dynamic}), more detailed modeling is needed.
+However due to the uncertainty in the calculated location of the N$_2$ snowline (see Section \ref{sec:dynamic}), \textbf{as well as the uncertainty of the formation zone of Jupiter-family comets (anywhere between 5 and $>$30 AU; \citealt{pontoppidan14}),} more detailed modeling is needed.
 
@@
 
-In this paper we explore the role of icy grain morphology and disk dynamics on the snowline locations of major volatile carrier molecules and the C/N/O ratios in protoplanetary disks.
+In this paper we explore the role of icy grain \textbf{compositions} and disk dynamics on the snowline locations of major volatile carrier molecules and the C/N/O ratios in protoplanetary disks.
 We enhance the coupled drift-desorption model developed in \citet{piso15b} by adding more carbon- and nitrogen-bearing species into our framework, and by considering different binding ice environments.

### 2016-07-20 Postdoc Applications/AnaMariaPisoResume.tex
update
upd

@@
 \vspace{-0.1in}
-\item[] Ph.D., Astronomy \& Astrophysics, May 2016 (expected)
-\item[] Advisor: Dr. Karin \"Oberg 
-\item[] Thesis Topic: ``Dynamics and Chemistry in Protoplanetary Disks''   
+\item[] Ph.D., Astronomy \& Astrophysics, May 2016
+\item[] Advisor: Prof. Karin \"Oberg 
+\item[] Thesis Title: ``Origins of Gas Giant Compositions: The Role of Disk Location and Dynamics''   
 \end{list1}
@@
 \begin{list2}
+
+\item[] {\bf Postdoctoral Fellow} \hfill {\bf September 2016 - }
+%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
+\begin{list1}
+%\vspace{-0.1in}
+\item[] UCLA, Earth, Planetary and Space Sciences Department \hfill \textit{Los Angeles, CA}
+%\item[] Project: The Magnetic Field Signature of Super Earths
+\item[] Advisor: Prof. Hilke Schlichting \\
+\end{list1}
+
+\item[] {\bf Postdoctoral Fellow} \hfill {\bf July 2016 - August 2016}
+%{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
+\begin{list1}
+%\vspace{-0.1in}
+\item[] Harvard College Observatory  \hfill \textit{Cambridge, MA}
+%\item[] Project: The Magnetic Field Signature of Super Earths
+\item[] Advisor: Prof. Karin \"Oberg \\
+\end{list1}
+
 \item[] {\bf Research assistant} \hfill {\bf August 2010 - July 2011}
@@
 \begin{list2}
+\item[] {\bf Piso, A.-M. A.}, Pegues, J., \& \"Oberg, K.I. \textit{The Role of Ice Compositions and Morphology For Snowlines and the C/N/O Ratios in Active Disks}. ApJ, submitted \\
+\end{list2}
+
+\begin{list2}
 \item[] {\bf Piso, A.-M.
@@
 \textit{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}.
-ApJ, resubmitted after referee report \\
+ApJ, 2015, 815, 109 \\
 \end{list2}
@@
 
-\section{\sc Publications in Preparation}
-
-\begin{list2}
-\item[] {\bf Piso, A.-M. A.}, \"Oberg, K.I., \& Pegues, J. \textit{The Role of Ice Compositions and Morphology For Snowlines and the C/N/O Ratios in Active Disks} \\
-\end{list2}
+%\section{\sc Publications in Preparation}
 
@@
 \section{\sc Conferences and Seminars}
+
+\begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+MIT Planetary Lunch Colloquium Series, Cambridge, MA, March 2016\\
+Invited talk \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets} \\
+AAS 227$^{th}$ Meeting, Kissimmee, FL, January 2016\\
+Dissertation talk \\
+\end{list2}
 
@@
 \item[] American Astronomical Society member
+\item[] Reviewer for ApJ
 %\item Romanian Student Association treasurer (2007-2010)

### 2016-07-21 tex/draft2/ms_revised.tex
upd

@@
 In the case of nitrogen, chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
-Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}), \textbf{although Titan's N$_2$ atmospheric composition is not primordial but rather originates from primordial NH$_3$ accreted by the satellite (\citealt{atreya78}, \citealt{mandt14}).}  %Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
+Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}), \textbf{although Titan's N$_2$ dominated atmospheric composition is not primordial but rather originates from primordial NH$_3$ accreted by the satellite (\citealt{atreya78}, \citealt{mandt14}).}  %Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
 %\textbf{While CO and N$_2$ have similar condensation temperatures (e.g., \citealt{fayolle16}), their snowline locations may }  
 Due to the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio compared to its average value in the disk.
-\textbf{This effect is most pronounced between the CO and N$_2$ snowlines, where the gas phase N/O ratio is enhanced by many orders of magnitude compared to the stellar value due to the depletion of oxygen gas in this region (see Section \ref{sec:results}).} Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
+\textbf{The larger N/O ratio enhancement is due to the fact that nitrogen is primarily carried by N$_2$, while carbon has other carriers besides CO.
+The N/O enhancement is most pronounced between the CO and N$_2$ snowlines, where the gas phase N/O ratio exceeds the stellar value by many orders of magnitude due to the depletion of oxygen gas in this region (see Section \ref{sec:results}).} Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
 In addition to N$_2$, a fraction of the nitrogen abundance may also be carried by less volatile species such as NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).
@@
 \textbf{Spectroscopic observations suggest that up to 90\% of the CO ice is frozen in a layer that is thick enough and separated from the H$_2$O ice layer underneath so that it can be considered pure ice (e.g., \citealt{pontoppidan03}).
-However, if the CO (or N$_2$) ice layer is thin enough, then it will interact with the H$_2$O ice substrate (e.g., \citealt{collings03}).
-The ice binding energy will be different in this water dominated environment than in the pure ice case.} Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
+However, if the CO (or N$_2$) ice layer is thin enough ($\sim$monolayer coverage), then it will interact with the H$_2$O ice substrate (e.g., \citealt{collings03}).
+The ice binding energy is significantly larger in this water dominated environment than in the pure ice case, as shown by laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}).} %have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
 This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{sec:static}).
@@
 When temperatures exceed $\sim$80-90 K, H$_2$O ice acquires a crystalline structure \citep{schegerer10}.
-The CO binding energy is larger in an amorphous porous water ice environment than in the compact case \citep{fayolle16} and slightly lower than in the crystalline case \citep{noble12}.
-Here we consider pure ices and ices dominated by amorphous porous H$_2$O ice.
+The CO binding energy is larger in an amorphous porous H$_2$O ice environment than in the amorphous compact or crystalline cases (\citealt{noble12}, \citealt{fayolle16}).
+The N$_2$ binding energy is also larger for a porous versus compact H$_2$O ice substrate.
+No equivalent studies have yet been performed for the deposition of N$_2$ on crystalline H$_2$O ice, but we expect the N$_2$ binding energy in this environment to follow the same trend as the CO binding energy, as CO and N$_2$ display a similar desorption behavior.
+To explore the range of distances at which CO and N$_2$ in different environments desorb, we consider the limiting cases: pure ices (lowest binding energy) and ices residing on an amorphous porous H$_2$O ice substrate (highest binding energy).
 We refer to the latter simply as water dominated ices unless noted otherwise.}
@@
 \label{eq:diskT}
-T = 120\, (r/\text{AU})^{-3/7} \,\,\text{K},
+T_{\rm irr} = 120\, (r/\text{AU})^{-3/7} \,\,\text{K},
 \end{equation}
@@
 \label{eq:activeT}
-T^4 =\Big[\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}\Big]^4 + T_{\rm irr}^4,
+T_{\rm visc}^4 =\Big[\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}\Big]^4 + T_{\rm irr}^4.
 \end{equation}
-where $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}).
+%where $T_{\rm irr} \textbf{} from Equation (\ref{eq:diskT}). 
 Here $G$ is the gravitational constant, $\kappa_0=2 \times 10^{-6}$ is a dimensionless opacity coefficient, $M_*=M_{\odot}$ is the mass of the central star, $\mu=2.35$ is the mean molecular weight of the nebular gas, $m_{\rm p}$ is the proton mass, $\Omega_{\rm k}=\sqrt{G M_{\odot}/r^3}$ is the Keplerian angular velocity, $\alpha=0.01$ is a dimensionless coefficient (see below for details), $k_{\rm B}$ is the Boltzmann constant, and $\sigma$ is the Stefan-Boltzmann constant.
@@
 The steady-state disk has an $\alpha$-viscosity prescription, where the kinematic viscosity is $\nu=\alpha c H$.
-Here $c \equiv \sqrt{k_{\rm B} T /(\mu m_{\rm p})}$ is the isothermal sound speed (with $T$ from Equation \ref{eq:activeT}), and $H \equiv c/\Omega_{\rm k}$ is the disk scale height.
+Here $c \equiv \sqrt{k_{\rm B} T_{\rm visc} /(\mu m_{\rm p})}$ is the isothermal sound speed and $H \equiv c/\Omega_{\rm k}$ is the disk scale height.
 We can then determine the gas surface density for a viscous disk as (\citealt{shakura73}; see also Paper I for a more detailed explanation of these calculations):
@@
 \end{equation}
-where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}=T$ is the grain temperature (assumed to be the same as the disk temperature, see Paper I), and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well.
+where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}$ is the grain temperature (assumed to be the same as the disk temperature, see Paper I), and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well.
 We discuss our choices for $E_x$ for the different volatile species in Section \ref{sec:snowlines}.
@@
 Freeze-out at the CO$_2$ and CO snowlines pulls more heavy elements into the grains than in the case of the H$_2$O snowline.
-The CO and N$_2$ snowlines move several tens of AU inward if the ices are water dominated rather than pure.
-This changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ.
+%The CO and N$_2$ snowlines move several tens of AU inward if the ices are water dominated rather than pure.
+
+\textbf{Figure \ref{fig:CNOstatic}, bottom panel, displays the snowline locations when CO and N$_2$ are in an amorphous porous water environment (see Section \ref{sec:intro}).
+The snowlines move outward by a factor of $\sim$3.3 if the ices are pure, and by up to a factor of $\sim$2 for an amorphous compact water substrate (not shown; our estimates are based on the results of \citealt{fayolle16}).
+The CO snowline moves outward by a factor of $\sim$1.5 if CO is in a crystalline rather than an amorphous porous water environment (not shown; our estimates are based on the results of \citealt{noble12}), and a similar trend is expected for the N$_2$ snowline (see Section \ref{sec:intro}).
+The results of Figure \ref{fig:CNOstatic} thus represent the limiting cases for the positions of the CO and N$_2$ snowlines, for different compositions of the icy grains and morphology of the water ice substrate.} This \textbf{variation in snowline locations} changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ.
 %\textbf{We note that the CO and N$_2$ snowlines will move inward if the ices are in an amorphous compact water environment and slightly shift outward in the case of a crystalline H$_2$O ice substrate.}
@@
 Changes in stellar luminosity (e.g., \citealt{kennedy06}) and gas mass accretion rate (e.g., \citealt{chambers09}), as well as the evolution of icy dust particles due to grain growth and fragmentation (e.g., \citealt{birnstiel12}), may introduce additional uncertainties in the snowline locations, and thus the C/N/O ratios.
-\textbf{Moreover, the diffusion of vapor across the snowlines following the cold finger effect (\citealt{stevenson88}, \citealt{cyr98}) will change the shape of the C/O and N/O curves and therefore the magnitude of the C/N/O ratios between different snowlines.} Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints on the relative importance of ice \textbf{compositions} and disk dynamics in setting snowline locations.
+\textbf{Moreover, the diffusion of vapor across the snowlines following the cold finger effect (\citealt{stevenson88}, \citealt{cyr98}) will change the shape of the C/O and N/O curves and therefore the magnitude of the C/N/O ratios between different snowlines.
+The effect of dynamical processes on snowline locations is discussed in more detail in Paper I, Section 5.2.} Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints on the relative importance of ice \textbf{compositions} and disk dynamics in setting snowline locations.
 The uncertainties in snowline locations caused by disk dynamics, ice \textbf{compositions}, and other effects outlined above can be resolved in extreme cases, such as a detection of a CO snowline at a temperature corresponding to pure CO ice desorption in a static disk (e.g., \citealt{qi13} at $\sim$ 17 K) or CO desorption from a water dominated ice in a dynamic disk.

### 2016-07-21 tex/draft2/ms_revised.tex
upd

@@
 In the case of nitrogen, chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}.
-Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}), \textbf{although Titan's N$_2$ dominated atmospheric composition is not primordial but rather originates from primordial NH$_3$ accreted by the satellite (\citealt{atreya78}, \citealt{mandt14}).}  %Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
+%Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}), \textbf{although Titan's N$_2$ dominated atmospheric composition is not primordial but rather originates from primordial NH$_3$ accreted by the satellite (\citealt{atreya78}, \citealt{mandt14}).}  %Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}.
 %\textbf{While CO and N$_2$ have similar condensation temperatures (e.g., \citealt{fayolle16}), their snowline locations may }  
-Due to the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio compared to its average value in the disk.
-\textbf{The larger N/O ratio enhancement is due to the fact that nitrogen is primarily carried by N$_2$, while carbon has other carriers besides CO.
-The N/O enhancement is most pronounced between the CO and N$_2$ snowlines, where the gas phase N/O ratio exceeds the stellar value by many orders of magnitude due to the depletion of oxygen gas in this region (see Section \ref{sec:results}).} Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
+Due to the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk \textbf{is expected to be high, perhaps} more enhanced than the C/O ratio compared to the Solar value.
+%\textbf{The larger N/O ratio enhancement is due to the fact that nitrogen is primarily carried by N$_2$, while carbon has other carriers besides CO.
+The N/O enhancement is most pronounced between the CO and N$_2$ snowlines, where the gas phase N/O ratio exceeds the stellar value by many orders of magnitude due to the depletion of oxygen gas in this region (see Section \ref{sec:results}).}
+\textbf{The C/O ratio in gas cannot exceed unity (i.e., a factor of $\sim$2 enhancement compared to the Solar value) since the major volatile carbon carrier is CO.
+In contrast, the N/O ratio mainly depends on the relative depletion of N$_2$ and oxygen carriers, and it will increase as each of the oxygen carrier snowlines (H$_2$O, CO$_2$, CO) is crossed.
+Beyond the CO snowline, there is no strict upper limit to the N/O ratio.
+The spatial extent of this latter region depends on the relative bond strengths of CO and N$_2$ to ice, but may be quite large (see Section \ref{sec:results}).} Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
 In addition to N$_2$, a fraction of the nitrogen abundance may also be carried by less volatile species such as NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).
+\textbf{The present day N$_2$ in Titan's atmosphere, for example, is thought to originate from accretion of primordial NH$_3$ (\citealt{atreya78}, \citealt{mandt14}).}
 
@@
 This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{sec:static}).
-\textbf{In protostellar cores, H$_2$O ice is primarily amorphous (e.g., \citealt{williams02}, \citealt{vandishoeck14}), and can either be compact or porous.
+\textbf{In protostellar cores, H$_2$O ice is primarily amorphous (e.g., \citealt{williams02}, \citealt{vandishoeck14}).
 When temperatures exceed $\sim$80-90 K, H$_2$O ice acquires a crystalline structure \citep{schegerer10}.

### 2016-09-14 tex/draft2/ms_revised_2.tex
upd

--- 2016-07-21 tex/draft2/ms_revised.tex
@@
 In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, their ice binding environment, and the presence of snowlines of major volatiles at different distances from the central star.
-We explore the effects of disk dynamical processes, molecular compositions and abundances, and ice \textbf{compositions} on the snowline locations of the main C, O and N carriers, and the C/N/O ratios in gas and dust throughout the disk.
+We explore the effects of disk dynamical processes, molecular compositions and abundances, and ice compositions on the snowline locations of the main C, O and N carriers, and the C/N/O ratios in gas and dust throughout the disk.
 The gas-phase N/O ratio enhancement in the outer disk (exterior to the H$_2$O snowline) exceeds the C/O ratio enhancement for all reasonable volatile compositions.
-Ice \textbf{compositions} and disk dynamics individually change the snowline location of N$_2$, the main nitrogen carrier, by a factor of 2-3, and when considered together the range of possible N$_2$ snowline locations is $\sim$11-$\sim$79 AU in a standard disk model.
+Ice compositions and disk dynamics individually change the snowline location of N$_2$, the main nitrogen carrier, by a factor of 2-3, and when considered together the range of possible N$_2$ snowline locations is $\sim$11-$\sim$79 AU in a standard disk model.
 Observations that anchor snowline locations at different stages of planet formation are therefore key to develop C/N/O ratios as a probe of planet formation zones.
@@
 %\textbf{While CO and N$_2$ have similar condensation temperatures (e.g., \citealt{fayolle16}), their snowline locations may }  
-Due to the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk \textbf{is expected to be high, perhaps} more enhanced than the C/O ratio compared to the Solar value.
+Due to the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk is expected to be high, perhaps more enhanced than the C/O ratio compared to the Solar value.
 %\textbf{The larger N/O ratio enhancement is due to the fact that nitrogen is primarily carried by N$_2$, while carbon has other carriers besides CO.
 The N/O enhancement is most pronounced between the CO and N$_2$ snowlines, where the gas phase N/O ratio exceeds the stellar value by many orders of magnitude due to the depletion of oxygen gas in this region (see Section \ref{sec:results}).}
-\textbf{The C/O ratio in gas cannot exceed unity (i.e., a factor of $\sim$2 enhancement compared to the Solar value) since the major volatile carbon carrier is CO.
+The C/O ratio in gas cannot exceed unity (i.e., a factor of $\sim$2 enhancement compared to the Solar value) since the major volatile carbon carrier is CO.
 In contrast, the N/O ratio mainly depends on the relative depletion of N$_2$ and oxygen carriers, and it will increase as each of the oxygen carrier snowlines (H$_2$O, CO$_2$, CO) is crossed.
 Beyond the CO snowline, there is no strict upper limit to the N/O ratio.
-The spatial extent of this latter region depends on the relative bond strengths of CO and N$_2$ to ice, but may be quite large (see Section \ref{sec:results}).} Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
+The spatial extent of this latter region depends on the relative bond strengths of CO and N$_2$ to ice, but may be quite large (see Section \ref{sec:results}).
+Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin.
 In addition to N$_2$, a fraction of the nitrogen abundance may also be carried by less volatile species such as NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).
-\textbf{The present day N$_2$ in Titan's atmosphere, for example, is thought to originate from accretion of primordial NH$_3$ (\citealt{atreya78}, \citealt{mandt14}).}
+The present day N$_2$ in Titan's atmosphere, for example, is thought to originate from accretion of primordial NH$_3$ (\citealt{atreya78}, \citealt{mandt14}).
 
-The snowline locations of the main carbon, oxygen and nitrogen carriers strongly depend on the ice grain \textbf{composition}.
-Very volatile species, such as CO and N$_2$, present binding energies, and therefore snowline locations, that are sensitive to the details of the \textbf{composition} of the icy grain mantles.
-\textbf{Spectroscopic observations suggest that up to 90\% of the CO ice is frozen in a layer that is thick enough and separated from the H$_2$O ice layer underneath so that it can be considered pure ice (e.g., \citealt{pontoppidan03}).
+The snowline locations of the main carbon, oxygen and nitrogen carriers strongly depend on the ice grain composition.
+Very volatile species, such as CO and N$_2$, present binding energies, and therefore snowline locations, that are sensitive to the details of the composition of the icy grain mantles.
+Spectroscopic observations suggest that up to 90\% of the CO ice is frozen in a layer that is thick enough and separated from the H$_2$O ice layer underneath so that it can be considered pure ice (e.g., \citealt{pontoppidan03}).
 However, if the CO (or N$_2$) ice layer is thin enough ($\sim$monolayer coverage), then it will interact with the H$_2$O ice substrate (e.g., \citealt{collings03}).
-The ice binding energy is significantly larger in this water dominated environment than in the pure ice case, as shown by laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}).} %have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
+The ice binding energy is significantly larger in this water dominated environment than in the pure ice case, as shown by laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}).
+%have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices.
 This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{sec:static}).
-\textbf{In protostellar cores, H$_2$O ice is primarily amorphous (e.g., \citealt{williams02}, \citealt{vandishoeck14}).
+In protostellar cores, H$_2$O ice is primarily amorphous (e.g., \citealt{williams02}, \citealt{vandishoeck14}).
 When temperatures exceed $\sim$80-90 K, H$_2$O ice acquires a crystalline structure \citep{schegerer10}.
@@
 To explore the range of distances at which CO and N$_2$ in different environments desorb, we consider the limiting cases: pure ices (lowest binding energy) and ices residing on an amorphous porous H$_2$O ice substrate (highest binding energy).
-We refer to the latter simply as water dominated ices unless noted otherwise.}
+We refer to the latter simply as water dominated ices unless noted otherwise.
 
@@
 
-In this work, we expand the coupled drift-desorption model developed in (\citealt{piso15b}; hereafter Paper I) by considering additional volatile molecules and abundances, ice \textbf{compositions}, as well as nitrogen-to-oxygen (N/O) ratios.
+In this work, we expand the coupled drift-desorption model developed in (\citealt{piso15b}; hereafter Paper I) by considering additional volatile molecules and abundances, ice compositions, as well as nitrogen-to-oxygen (N/O) ratios.
 This paper is organized as follows.
 In Section \ref{sec:review}, we review the drift-desorption model developed in Paper I.
-We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain \textbf{compositions} and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{sec:results}.
+We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain compositions and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{sec:results}.
 We address the implications of our results in Section \ref{sec:discussion} and summarize our findings in Section \ref{sec:summary}.
@@
 
-\subsection{Snowlines in a Static Disk: The Importance of Ice \textbf{Compositions}}
+\subsection{Snowlines in a Static Disk: The Importance of Ice Compositions}
 \label{sec:snowlines}
 
-As we note in Section \ref{sec:intro}, the disk volatile composition and the ice \textbf{composition} determine the location of important snowlines.
+As we note in Section \ref{sec:intro}, the disk volatile composition and the ice composition determine the location of important snowlines.
 In this work we focus on the primary carbon, oxygen and nitrogen carriers, i.e.
@@
 The binding energies of H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ as pure ices are 5800 K, 2000 K, 834 K, 1300 K, 767 K and 2965 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}, \citealt{martin14}).
-\textbf{For our fiducial disk model, these energies correspond to disk temperatures of 143 K, 48 K, 21 K, 30 K, 18 K and 68 K, respectively.} For CO and N$_2$ as water dominated ices, the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}, \textbf{corresponding to disk temperatures of 34 K and 31 K, respectively}.
+For our fiducial disk model, these energies correspond to disk temperatures of 143 K, 48 K, 21 K, 30 K, 18 K and 68 K, respectively.
+For CO and N$_2$ as water dominated ices, the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}, corresponding to disk temperatures of 34 K and 31 K, respectively.
 Figure \ref{fig:CNOstatic} shows the resulting snowline locations, assuming CO and N$_2$ pure ices (top panel), and CO and N$_2$ in water dominated ices (bottom panel).
@@
 
-\textbf{Figure \ref{fig:CNOstatic}, bottom panel, displays the snowline locations when CO and N$_2$ are in an amorphous porous water environment (see Section \ref{sec:intro}).
+Figure \ref{fig:CNOstatic}, bottom panel, displays the snowline locations when CO and N$_2$ are in an amorphous porous water environment (see Section \ref{sec:intro}).
 The snowlines move outward by a factor of $\sim$3.3 if the ices are pure, and by up to a factor of $\sim$2 for an amorphous compact water substrate (not shown; our estimates are based on the results of \citealt{fayolle16}).
 The CO snowline moves outward by a factor of $\sim$1.5 if CO is in a crystalline rather than an amorphous porous water environment (not shown; our estimates are based on the results of \citealt{noble12}), and a similar trend is expected for the N$_2$ snowline (see Section \ref{sec:intro}).
-The results of Figure \ref{fig:CNOstatic} thus represent the limiting cases for the positions of the CO and N$_2$ snowlines, for different compositions of the icy grains and morphology of the water ice substrate.} This \textbf{variation in snowline locations} changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ.
+The results of Figure \ref{fig:CNOstatic} thus represent the limiting cases for the positions of the CO and N$_2$ snowlines, for different compositions of the icy grains and morphology of the water ice substrate.
+This variation in snowline locations changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ.
 %\textbf{We note that the CO and N$_2$ snowlines will move inward if the ices are in an amorphous compact water environment and slightly shift outward in the case of a crystalline H$_2$O ice substrate.}
@@
 In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and to what extent they are affected by the presence of CH$_4$ and NH$_3$ over the full range of observed CH$_4$ and NH$_3$ abundances toward low-mass protostars.
-In this section we only consider pure ices, \textbf{i.e.
+In this section we only consider pure ices, i.e.
 ices that are layered on a silicate mantle.
 In reality, the icy grain will have a layered structure with volatiles residing on top of a H$_2$O ice substrate.
-However, if the volatile ice layer is thick enough and separated from the H$_2$O ice layer, the interaction between H$_2$O and the other volatile species will be minimal and thus the ices can be considered pure rather than water dominated (see also Section \ref{sec:intro}).}
+However, if the volatile ice layer is thick enough and separated from the H$_2$O ice layer, the interaction between H$_2$O and the other volatile species will be minimal and thus the ices can be considered pure rather than water dominated (see also Section \ref{sec:intro}).
 
@@
 If the CO binding environment is known, the CO snowline moves inward by up to $\sim$50 \% compared to a static disk for each case (pure and water dominated ices) due to disk dynamics.
-The full range of potential CO snowlines taking into account both ice \textbf{compositions} and disk dynamics span $\sim$8.7 AU to $\sim$61 AU, which is a factor of $\sim$7 difference.
+The full range of potential CO snowlines taking into account both ice compositions and disk dynamics span $\sim$8.7 AU to $\sim$61 AU, which is a factor of $\sim$7 difference.
 This implies that gas phase C/O ratios of order unity may be reached in the giant planet forming zone, and the CO snowline may be inside 10 AU for certain disk parameters.
@@
 The presence of CO in a water ice environment rather than as pure ice moves the CO snowline significantly inward by $\sim$70\%.
-Taken together, disk dynamics and ice \textbf{compositions} move the CO snowline inward by a factor of $\sim$7.}
+Taken together, disk dynamics and ice compositions move the CO snowline inward by a factor of $\sim$7.}
 \label{fig:CO_ratio}
@@
 Thus for each case (pure versus water dominated ices), the N$_2$ snowline moves inward by up to 50\% due to disk dynamics.
-By taking into account both ice \textbf{compositions} and disk dynamics, the full range of potential N$_2$ snowlines span $\sim$11 to $\sim$79 AU, which is a factor of $\sim$7 difference.
+By taking into account both ice compositions and disk dynamics, the full range of potential N$_2$ snowlines span $\sim$11 to $\sim$79 AU, which is a factor of $\sim$7 difference.
 Similarly to the case for CO, the N$_2$ snowline may be close to 10 AU for certain disk models.
@@
 The presence of N$_2$ in a water ice environment rather than as pure ice moves the N$_2$ snowline significantly inward by $\sim$70\%.
-Taken together, disk dynamics and ice \textbf{compositions} move the N$_2$ snowline inward by a factor of $\sim$7.
+Taken together, disk dynamics and ice compositions move the N$_2$ snowline inward by a factor of $\sim$7.
 The results of an enhanced gas-phase N/O ratio between the H$_2$O and CO snowlines compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}
@@
 
-Due to disk dynamics and ice \textbf{compositions}, the locations of the CO and N$_2$ snowlines, and thus the disk regions with highly elevated gas phase N/O and C/O ratios, are uncertain and may span tens of AU.
+Due to disk dynamics and ice compositions, the locations of the CO and N$_2$ snowlines, and thus the disk regions with highly elevated gas phase N/O and C/O ratios, are uncertain and may span tens of AU.
 Both ice morphologies discussed in this study, pure and water dominated ices, are plausible in protoplanetary disks and depend on whether H$_2$O and CO ices formed on similar timescales or successively (e.g., \citealt{garrod11}).
@@
 Changes in stellar luminosity (e.g., \citealt{kennedy06}) and gas mass accretion rate (e.g., \citealt{chambers09}), as well as the evolution of icy dust particles due to grain growth and fragmentation (e.g., \citealt{birnstiel12}), may introduce additional uncertainties in the snowline locations, and thus the C/N/O ratios.
-\textbf{Moreover, the diffusion of vapor across the snowlines following the cold finger effect (\citealt{stevenson88}, \citealt{cyr98}) will change the shape of the C/O and N/O curves and therefore the magnitude of the C/N/O ratios between different snowlines.
-The effect of dynamical processes on snowline locations is discussed in more detail in Paper I, Section 5.2.} Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints on the relative importance of ice \textbf{compositions} and disk dynamics in setting snowline locations.
-The uncertainties in snowline locations caused by disk dynamics, ice \textbf{compositions}, and other effects outlined above can be resolved in extreme cases, such as a detection of a CO snowline at a temperature corresponding to pure CO ice desorption in a static disk (e.g., \citealt{qi13} at $\sim$ 17 K) or CO desorption from a water dominated ice in a dynamic disk.
-In intermediate cases it is more difficult to resolve the relative importance of ice \textbf{compositions} and disk dynamics.
+Moreover, the diffusion of vapor across the snowlines following the cold finger effect (\citealt{stevenson88}, \citealt{cyr98}) will change the shape of the C/O and N/O curves and therefore the magnitude of the C/N/O ratios between different snowlines.
+The effect of dynamical processes on snowline locations is discussed in more detail in Paper I, Section 5.2.
+Given the number of uncertainties in snowline locations, detections of snowlines in a sample of disks at different evolutionary stages are needed to provide observational constraints on the relative importance of ice compositions and disk dynamics in setting snowline locations.
+The uncertainties in snowline locations caused by disk dynamics, ice compositions, and other effects outlined above can be resolved in extreme cases, such as a detection of a CO snowline at a temperature corresponding to pure CO ice desorption in a static disk (e.g., \citealt{qi13} at $\sim$ 17 K) or CO desorption from a water dominated ice in a dynamic disk.
+In intermediate cases it is more difficult to resolve the relative importance of ice compositions and disk dynamics.
 For example, the CO snowline in HD 163296 is at a higher temperature of $\sim$ 25 K \citep{qi15}, which could be caused either by CO being in a water dominated environment or by dynamical effects that push the CO snowline inward.
@@
 A low N$_2$/CO ratio is consistent with comets having formed inside the N$_2$ snowline where N$_2$ is still in the gas phase.
-\textbf{However, it is also possible that the measured N$_2$ abundance in 67P may be due to post-formation processes such as radiogenic heating \citep{rubin15}, and thus may not reflect the comet's primordial composition.} Theoretical models suggest that Jupiter-family comets, such as 67P, originate from the Kuiper belt (\citealt{duncan97}; but see \citealt{rubin15} for alternative formation scenarios for 67P).
+However, it is also possible that the measured N$_2$ abundance in 67P may be due to post-formation processes such as radiogenic heating \citep{rubin15}, and thus may not reflect the comet's primordial composition.
+Theoretical models suggest that Jupiter-family comets, such as 67P, originate from the Kuiper belt (\citealt{duncan97}; but see \citealt{rubin15} for alternative formation scenarios for 67P).
 It is thus possible, in principle, to use measurements of the N$_2$ abundance in Jupiter-family comets to determine where the N$_2$ snowline was located in our Solar system.
-However due to the uncertainty in the calculated location of the N$_2$ snowline (see Section \ref{sec:dynamic}), \textbf{as well as the uncertainty of the formation zone of Jupiter-family comets (anywhere between 5 and $>$30 AU; \citealt{pontoppidan14}),} more detailed modeling is needed.
+However due to the uncertainty in the calculated location of the N$_2$ snowline (see Section \ref{sec:dynamic}), as well as the uncertainty of the formation zone of Jupiter-family comets (anywhere between 5 and $>$30 AU; \citealt{pontoppidan14}), more detailed modeling is needed.
 
@@
 
-In this paper we explore the role of icy grain \textbf{compositions} and disk dynamics on the snowline locations of major volatile carrier molecules and the C/N/O ratios in protoplanetary disks.
+In this paper we explore the role of icy grain compositions and disk dynamics on the snowline locations of major volatile carrier molecules and the C/N/O ratios in protoplanetary disks.
 We enhance the coupled drift-desorption model developed in \citet{piso15b} by adding more carbon- and nitrogen-bearing species into our framework, and by considering different binding ice environments.

### 2016-10-10 tex/draft2/ms_revised_2.tex
upd

@@
 
-\author{Ana-Maria A. Piso\altaffilmark{1}, Jamila Pegues\altaffilmark{2}, Karin I. \"Oberg\altaffilmark{1}}
+\author{Ana-Maria A. Piso\altaffilmark{1,2}, Jamila Pegues\altaffilmark{1,3}, Karin I. \"Oberg\altaffilmark{1}}
 \altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-\altaffiltext{2}{Department of Astrophysical Sciences, Princeton University, Princeton, NJ 08544}
+\altaffiltext{2}{UCLA, 595 Charles E. Young Drive East, Los Angeles, CA 90095}
+\altaffiltext{3}{Department of Astrophysical Sciences, Princeton University, Princeton, NJ 08544}
 
@@
 The CO snowline moves outward by a factor of $\sim$1.5 if CO is in a crystalline rather than an amorphous porous water environment (not shown; our estimates are based on the results of \citealt{noble12}), and a similar trend is expected for the N$_2$ snowline (see Section \ref{sec:intro}).
+\textbf{Based on recent observations \citep{luspay16} and theoretical models (\citealt{lectez15}, \citealt{mousis16}) for the composition of comet 67P/Churyumov-Gerasimenko, a fraction of the disk CO (and possibly N$_2$) content may be trapped in clathrates.
+This would imply CO and N$_2$ snowline locations roughly corresponding to  CO and N$_2$ in a crystalline H$_2$O ice environment as described above.}
 The results of Figure \ref{fig:CNOstatic} thus represent the limiting cases for the positions of the CO and N$_2$ snowlines, for different compositions of the icy grains and morphology of the water ice substrate.

### 2016-10-21 tex/draft2/ms_revised_2.tex
upd

--- 2016-09-14 tex/draft2/ms_revised_2.tex
@@
 The CO snowline moves outward by a factor of $\sim$1.5 if CO is in a crystalline rather than an amorphous porous water environment (not shown; our estimates are based on the results of \citealt{noble12}), and a similar trend is expected for the N$_2$ snowline (see Section \ref{sec:intro}).
+%\textbf{Based on recent observations \citep{luspay16} and theoretical models (\citealt{lectez15}, \citealt{mousis16}) for the composition of comet 67P/Churyumov-Gerasimenko, a fraction of the disk CO (and possibly N$_2$) content may be trapped in clathrates.
+This would imply CO and N$_2$ snowline locations roughly corresponding to  CO and N$_2$ in a crystalline H$_2$O ice environment as described above.}
 The results of Figure \ref{fig:CNOstatic} thus represent the limiting cases for the positions of the CO and N$_2$ snowlines, for different compositions of the icy grains and morphology of the water ice substrate.
@@
 %\textbf{We note that the CO and N$_2$ snowlines will move inward if the ices are in an amorphous compact water environment and slightly shift outward in the case of a crystalline H$_2$O ice substrate.}
+\textbf{In our simple model, we ignore the effects of CO and N$_2$ entrapment in water ice through clathrate formation or other processes.
+Theoretical models aimed at explaining the composition of comet 67P/Churyumov-Gerasimenko suggest that a small fraction of the total CO and N$_2$ reservoir may be trapped in clathrates, and only released upon water sublimation (\citealt{lectez15}, \citealt{mousis16}).
+The importance of this process for CO and N$_2$ is not clear \citep{luspay16}, however, but would in either case not change the locations of the CO and N$_2$ snowlines, i.e.
+the locations where CO and N$_2$ transition from being mainly present as ice to being mainly present as vapor.}
 

### 2016-10-26 tex/draft2/ms_revised_2.tex
upd

@@
 Theoretical models aimed at explaining the composition of comet 67P/Churyumov-Gerasimenko suggest that a small fraction of the total CO and N$_2$ reservoir may be trapped in clathrates, and only released upon water sublimation (\citealt{lectez15}, \citealt{mousis16}).
-The importance of this process for CO and N$_2$ is not clear \citep{luspay16}, however, but would in either case not change the locations of the CO and N$_2$ snowlines, i.e.
+In this case, the CO and N$_2$ snowlines would be closer to the star than in the pure ice case.}
+%The importance of this process for CO and N$_2$ is not clear \citep{luspay16}, however, but would in either case not change the locations of the CO and N$_2$ snowlines, i.e.
 the locations where CO and N$_2$ transition from being mainly present as ice to being mainly present as vapor.}

### 2016-11-03 tex/draft2/ms_revised_2.tex
upd

@@
 %\textbf{We note that the CO and N$_2$ snowlines will move inward if the ices are in an amorphous compact water environment and slightly shift outward in the case of a crystalline H$_2$O ice substrate.}
-\textbf{In our simple model, we ignore the effects of CO and N$_2$ entrapment in water ice through clathrate formation or other processes.
+In our simple model, we ignore the effects of CO and N$_2$ entrapment in water ice through clathrate formation or other processes.
 Theoretical models aimed at explaining the composition of comet 67P/Churyumov-Gerasimenko suggest that a small fraction of the total CO and N$_2$ reservoir may be trapped in clathrates, and only released upon water sublimation (\citealt{lectez15}, \citealt{mousis16}).
-In this case, the CO and N$_2$ snowlines would be closer to the star than in the pure ice case.}
+In this case, the CO and N$_2$ snowlines would be closer to the star than in the pure ice case.
 %The importance of this process for CO and N$_2$ is not clear \citep{luspay16}, however, but would in either case not change the locations of the CO and N$_2$ snowlines, i.e.
@@
 
-\acknowledgments{This work is supported by a Simons Collaboration on the Origins of Life (SCOL) investigator award to KIO. JP was supported by the Banneker institute.}
+\acknowledgments{We thank the anonymous referee for helpful comments and suggestions.
+This work is supported by a Simons Collaboration on the Origins of Life (SCOL) investigator award to KIO.
+JP was supported by the Banneker institute.}
 

### 2016-11-04 python/exoplanet_read.py
upd

@@ -1,0 +1,17 @@
+import numpy as np
+
+f = open('../dat/ExoplanetData/transit.txt', 'r')
+Mtransit, atransit = [], []
+f.readline()
+f.readline()
+f.readline()
+f.readline()
+f.readline()
+f.readline()
+f.readline()
+f.readline()
+f.readline()
+f.readline()
+for line in f:
+    Mtransit = np.append(Mtransit, float(line.split()[1]))
+    atransit = np.append(atransit, float(line.split()[1]))

### 2016-11-05 Postdoc Applications/uc_background_statement.tex
upd

%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
%\documentclass[12pt, preprint]{aastex}
%\documentclass[apj]{emulateapj}
\documentclass[12pt, letterpaper]{article}

%\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
%\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex

%\usepackage{pdfsync}
%\usepackage{subeqnarray}
\usepackage[top=0.8in, bottom=0.7in, left=1in, right=1in]{geometry}
\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage{titling}
\usepackage{sectsty}
\usepackage{sidecap}
\usepackage{placeins}
\usepackage{indentfirst}
\setlength{\droptitle}{-7em}
\setlength{\abovecaptionskip}{-0.4ex}
\setlength{\belowcaptionskip}{-0.4ex}
%\pagenumbering{gobble}
\pagestyle{fancy}
\lhead{Ana-Maria Piso}
\rhead{Education and Background Statement}
\date{}
%\rhead{\thepage}

%\bibliographystyle{apj}

\title{\Large UC President's Postdoctoral Fellowship Education and Background Statement}
\author{Dr. Ana-Maria Piso}

%\newenvironment{packed_item}{
%\begin{itemize}
%  \setlength{\itemsep}{1pt}
%  \setlength{\parskip}{0pt}
%  \setlength{\parsep}{0pt}
%}{\end{itemize}}

\begin{document}
\maketitle

%\slugcomment{Draft Modified \today}

\vspace{-0.9cm}

I was born in Romania a couple of years before the 1989 revolution. As a woman growing up in a country that was not ready to promote values such as diversity, inclusion, tolerance, and equality between genders, I was not expected to pursue a scientific career. My family supported my choice, but society and the educational system did not. I moved to the United States to attend college with the expectation that the values and ethics that are often disregarded in my home country would be universally appreciated here. Unfortunately, I found out that this is not always the case, especially regarding gender equality. As I gradually advanced in my career in astrophysics, this disparity became more and more evident. As did the fact that girls and young women are often discouraged from considering scientific career paths, either due to the scarcity of opportunities and resources, or their own lack of self confidence, galvanized by little societal support or very few role models. I am therefore highly dedicated to mentoring, encouraging and empowering women, a group often deeply underrepresented in scientific fields.

Throughout my graduate studies, I have often been a mentor, in various capacities, to younger women already pursuing a scientific career, or just interested in and excited about the possibility. I was part of the Women In Science, Engineering, Technology and Mathematics (WISTEM) organization, a group that aims to foster a sense of community for women studying science and engineering at Harvard. The program pairs graduate and undergraduate students in a mentor-mentee relationship, with the aim of creating a network for women scientists. I served as a mentor for three years to a Harvard undergraduate. Through monthly coffees or lunches, we discussed graduate school opportunities, work/life balance, some of the challenges that she faced both as a woman and a minority, and the general prospect of a career in science. She had successful summer internships at Harvard Medical School and Harvard School of Public Health, and is now in her final year as a biostatistics major. In a similar capacity, I have been a mentor for a summer Research Experiences for Undergraduates (REU) student at the Center for Astrophysics (CfA). As part of Prof. Karin \"Oberg's group, I co-mentored two of her female undergraduate students for their senior thesis projects, one of whom received the award for best senior thesis, and the other who is now a graduate student at the CfA.

I am also committed to encouraging and empowering girls from minority groups or underprivileged communities. I was a Mentor Scientist at the Science Club For Girls (SCFG), a forum in which undergraduate and graduate women students in STEM fields teach elementary school girls from challenging social backgrounds some basic scientific concepts in a fun, interactive way. The goal is twofold --- to stimulate the girls' interest in science from a young age, and to empower them by serving as female role models. The format of the classes was usually a combination of discussion and hands-on experiments --- for example, second grade girls had to
make a prism and shine a flashlight through it to see the forming rainbow, and fourth grade girls had to filter dirty water using sand and cotton filters. As a SCFG Mentor Scientist, I learned how to explain very basic scientific concepts in a manner that is intuitive and understandable for a young audience, as well as resist the urge to jump in to help the girls in their experiments, but rather 
let them figure things out by themselves and thus boost their confidence.

At UCLA, I plan to join the Sociable Postdoc Society, a group that aims to create a community between postdoctoral scholars from different UCLA departments, as well as between postdocs at other local campuses. Among other things, I would like to organize regular social/academic lunches, in which scholars from different departments would give a short, widely accessible presentation about their research, while also having the chance to socialize and network with other postdocs. I believe this would be a great opportunity for people with similar career paths, and yet with different backgrounds, genders, nationalities, ethnicities, to interact, learn from each other's experiences, and embrace their diversity.

%\end{thebibliography}

%\bibliographystyle{abbrv}
%\bibliography{refs}

%\if\bibinc y
%\begin{thebibliography}
%\end{thebibliography}
%\fi

\end{document}

### 2016-11-05 Postdoc Applications/uc_research_statement.tex
upd

%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
%\documentclass[12pt, preprint]{aastex}
%\documentclass[apj]{emulateapj}
\documentclass[12pt, letterpaper]{article}

%\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
%\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex

%\usepackage{pdfsync}
%\usepackage{subeqnarray}
\usepackage[top=0.8in, bottom=0.7in, left=1in, right=1in]{geometry}
\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage{titling}
\usepackage{sectsty}
\usepackage{sidecap}
\usepackage{placeins}
\usepackage{indentfirst}
\setlength{\droptitle}{-7em}
\setlength{\abovecaptionskip}{-0.4ex}
\setlength{\belowcaptionskip}{-0.4ex}
%\pagenumbering{gobble}
\pagestyle{fancy}
\lhead{Ana-Maria Piso}
\rhead{UC President's Postdoctoral Fellowship Research Statement}
\date{}
%\rhead{\thepage}

%\bibliographystyle{apj}

\title{\Large UC President's Postdoctoral Fellowship Research Statement: \\
How Unique is the Solar System? Planetary Luminosities as Fingerprints for the Origins and Properties of Exoplanets}
\author{Dr. Ana-Maria Piso}

%\newenvironment{packed_item}{
%\begin{itemize}
%  \setlength{\itemsep}{1pt}
%  \setlength{\parskip}{0pt}
%  \setlength{\parsep}{0pt}
%}{\end{itemize}}

\begin{document}
\maketitle

%\slugcomment{Draft Modified \today}

\vspace{-0.9cm}

Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered [1]. Their diversity in terms of mass, radius, location and composition [2] provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life. The advent of the Kepler Space Telescope has resulted in an exponential increase in the number of exoplanets detected via the transit method. As shown in Figure 1, a large fraction of the planets detected by Kepler are the so-called ``Super Earths'' or ``Mini Neptunes'', planets with masses between 1 and 10 times that of the Earth, and which are likely to contain some of their mass in a gaseous envelope. The majority of these detected Super Earths are very close to their host stars, with smaller orbital periods than the Earth. These features imply that Super Earths do not have Solar System analogs. At the same time, they are ubiquitously found and predicted to be the most common types of exoplanets [3], which makes understanding their ages, formation and properties particularly relevant. Observational techniques such as gravitational microlensing and direct imaging have extended the orbital range at which planets have been detected from just outside the Earth's orbit to tens and even hundreds of astronomical units. As demonstrated in Figure 2, however, most of these planets are massive; wide-separation Super Earths or Neptune analogs have yet to be detected. Are they indeed rare at these large distances, or just not bright or young enough to be detected? The Gemini Planet Imager (GPI) exoplanet survey will be key in improving our knowledge on this matter. GPI is a direct imaging survey that aims to observe very young, bright planets. As planets become less bright with increasing age, GPI might thus be able to detect lower mass planets that are just at the end stages of their formation. A theoretical framework to interpret these observations is therefore crucial. In particular, a planet's luminosity is an essential feature because it can provide constraints on the planet's mass, age and formation mechanism. \textbf{I propose to develop a holistic framework to self consistently track a planet's luminosity from its formation and through its evolution after the dispersal of the protoplanetary disk. Such a model provides essential context for characterizing young planets that are and will be discovered via direct imaging surveys.}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{exo_solar}
%\vspace{-0.5in}
\caption{Planetary mass vs. orbital separation of the detected exoplanet population, via transit (red) and radial velocity (blue). Venus, Mars and Earth are displayed in the lower right corner. No Solar System Super Earth analogs have been yet detected.}
\label{fig:chemical}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{exo_solar_all}
%\vspace{-0.5in}
\caption{Same as above, but planets detected via gravitational microlensing (green) and direct imaging (yellow) are added. Besides the terrestrial planets, Jupiter, Saturn, Uranus and Neptune are also shown. There is a scarcity in Super Earth detections at wide separateions.}
\label{fig:chemical}
\end{figure}

\vspace{0.2in}

%\section{Coupled Chemical and Dynamical Disk Evolution} 
\textbf{1. Planet formation from the start.} Planet formation models largely fall in two categories: core accretion, where planetesimals collide and grow into a solid core which then accretes a gaseous envelope, and gravitational instability, where planets form due to an instability in the disk that causes the disk material to collapse under its own gravity. Core accretion models can track a planet's mass, radius and luminosity evolution from the very early stages of its formation. Current core accretion models focus primarily on gas giants ([4], [5]), but no equivalent numerical simulations exist for lower mass planets. \textbf{Using the expertise that I acquired during my dissertation research on the evolution of gas giants (see Figure 3), I will use and improve the model I developed in [6] and [7], and adapt it to Super Earth Formation.} This will provide us with valuable insight into a forming planet's luminosity and other physical properties from the beginning of the accretion of the gaseous envelope until the gas disk dissipation.

%this dynamical model  develop a (chemical models of increasing complexity) simplified time-dependent chemistry,

\vspace{0.2in}
%\section{Planet and Planetesimal Migration}
\textbf{2. Spontaneous mass loss.} After the end of the planet formation process and gas disk dispersal, a planet can lose its envelope either due to irradiation from its host star (photoevaporation), or due to its own cooling. Atmospheric mass loss, as well as the planet's cooling and contraction, will decrease a planet's luminosity in time. Most atmospheric escape models focus on mass loss due to photoevaporation ([8], [9]). However, [10] use an analytic model to consider both types of mass loss, and find that the mass loss due to internal cooling happens on a much faster timescale than due to photoevaporation. This implies that the former process is likely to dominate the cooling, mass and luminosity evolution of very young planets, which are the focus of this project. \textbf{Through numerical and analytic calculations, I will incorporate a planet's mass loss due to internal cooling in my model, and use this to track the planet's luminosity after disk dissipation.}

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{opacity_effect.pdf}
%\vspace{-0.5in}
\caption{Example of a planet's luminosity and time evolution with atmospheric mass, during the giant planet formation stage. From [6].}  %The actual magnitude of dust opacities is difficult to robustly predict.}
\label{fig:LtvsMopacity}
\end{figure}

\vspace{0.2in}

%\section{Model Planet Populations} 
\textbf{3. Luminosity evolution and connection to observations.} A planet's luminosity evolution has been analyzed before [11], but this was only done for gas giants. Moreover, these models consider arbitrary initial conditions, as they are only interested in studying the luminosities of mature, several billion years old exoplanets. \textbf{The combined results from parts 1 and 2 will represent the first time that a planet's luminosity evolution is tracked self consistently from its birth through after the disk dispersal.} These results will be crucial to understanding the properties of directly imaged young, bright planets. Furthermore, they may be able to help us distinguish between different formation scenarios Ð core accretion versus gravitational instability.

\vspace{0.2in}

\textbf{UCLA is the ideal place for me to continue my postdoctoral tenure as a President's Fellow.} I have tremendously enjoyed the time that I have spent here, and greatly benefitted from the diversity and academic excellence of the EPSS department. In particular, working with Prof. Hilke Schlichting is a privilege, and her expertise in planet formation and dynamics will be instrumental in achieving my research goals as a postdoctoral fellow and afterwards. Collaborating with Prof. David Jewitt, a leader in Solar System astrophysics, will be very useful for analyzing my findings in the context of the Solar System. As my research interests are at the intersection of astrophysics and planetary science, I would welcome collaborations with members of the Astronomy department faculty, such as Prof. Smadar Naoz. UCLA is also home to the Institute for Planets and Exoplanets (iPLEX), an interdisciplinary forum that ties together people with different scientific backgrounds. Outside research, I plan to be involved in the UCLA Social Postdoc Society. Through this, I will help foster connections between postdoctoral scholars from various departments across the campus, with the goal of sharing our diverse scientific knowledge and experience.
%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.7\textwidth]{CO_abundances}
%%\vspace{-0.5in}
%\caption{TBD}
%\label{fig:CO}
%\end{figure}

%\if\bibinc n
%\bibliography{refs}
%\fi
\FloatBarrier
%\def\bibfont{\footnotesize}
%\setlength{\bibsep}{0.0pt}

\section*{References}
\footnotesize
%\begin{thebibliography} {9}
\noindent $[1]$ Batalha, N. M. 2014, Proceedings of the National Academy of Science, 111, 12647 \\
$[2]$ Lissauer J. J., Dawson R. I., Tremaine S., 2014, Nature, 513, 336 \\
$[3]$ Fressin, F., Torres, G., Charbonneau, D., et al. 2013, ApJ, 766, 81\\
$[4]$ Pollack, J. B., Hubickyj, O., Bodenheimer, P., et al. 1996, Icarus, 124, 62 \\
$[5]$ Rafikov, R.R. 2006, ApJ, 648, 666 \\
$[6]$ \textbf{Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21} \\
$[7]$ \textbf{Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82} \\
$[8]$ Lopez, E. D., Fortney, J. J., \& Miller, N. 2012, ApJ, 761, 59 \\
$[9]$ Owen, J. E., \& Jackson, A. P. 2012, MNRAS, 425, 2931 \\
$[10]$ Ginzburg, S., Schlichting, H. E., \& Sari, R. 2016, ApJ, 825, 29 \\
$[11]$ Marley, M. S., Fortney, J. J., Hubickyj, O., Bodenheimer, P., \& Lissauer, J. J. 2007, ApJ, 655, 541

%\end{thebibliography}

%\bibliographystyle{abbrv}
%\bibliography{refs}

%\if\bibinc y
%\begin{thebibliography}
%\end{thebibliography}
%\fi

\end{document}

### 2016-11-05 Postdoc Applications/uc_research_statement.tex
upd

@@
 \title{\Large UC President's Postdoctoral Fellowship Research Statement: \\
-How Unique is the Solar System? Planetary Luminosities as Fingerprints for the Origins and Properties of Exoplanets}
+\textbf{How Unique is the Solar System? Planetary Luminosities as Fingerprints for the Origins and Properties of Exoplanets}}
 \author{Dr. Ana-Maria Piso}

### 2016-11-05 Postdoc Applications/AnaMariaPisoResume.tex
upd

@@
 
-Harvard-Smithsonian Center for Astrophysics & {\it Phone:}    (617) 818-6780 \\         
-60 Garden Street, MS-10 & {\it E-mail:}  apiso@cfa.harvard.edu\\       
-Cambridge, MA  02138  & {\it WWW:} www.cfa.harvard.edu/$\sim$apiso \\     
+Department of Earth, Planetary, and Space Sciences & {\it Phone:}    (617) 818-6780 \\         
+595 Charles E. Young Dr. East & {\it E-mail:}  apiso@ucla.edu\\       
+Los Angeles, CA 90095  & {\it WWW:} www.cfa.harvard.edu/$\sim$apiso \\     
 \end{tabular}
@@
 \vspace{-0.1in}
-\item[] Ph.D., Astronomy \& Astrophysics, May 2016
+\item[] Ph.D., Astronomy \& Astrophysics, April 2016
 \item[] Advisor: Prof. Karin \"Oberg 
@@
 \begin{list1}
-\item[] S.B., Physics,  June 2010 \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, Major GPA: 4.6/5.0
-\item[] S.B., Mathematics, June 2010\,\,\,\, Major GPA: 4.8/5.0
+\item[] S.B., Physics,  June 2010 % \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, Major GPA: 4.6/5.0
+\item[] S.B., Mathematics, June 2010% \,\,\,\, Major GPA: 4.8/5.0
 \end{list1}
@@
 
-\item[] {\bf Postdoctoral Fellow} \hfill {\bf September 2016 - }
+\item[] {\bf Postdoctoral Scholar} \hfill {\bf September 2016 - present}
 %{\bf Harvard-Smithsonian Center for Astrophysics} \hfill \textit{Cambridge, MA}
@@
 %\vspace{-0.1in}
-\item[] UCLA, Earth, Planetary and Space Sciences Department \hfill \textit{Los Angeles, CA}
+\item[] UCLA, Department of Earth, Planetary, and Space Sciences \hfill \textit{Los Angeles, CA}
 %\item[] Project: The Magnetic Field Signature of Super Earths
@@
 \item[] Neuron Group S.R.L. Software Company \hfill \textit{Bucharest, Romania}
-\item[] Digital map designer and database manager for the '112 Emergency Call Center' national project
+\item[] Digital map designer and database manager for the '112 Emergency Call Center' project
 %\item[] Advisors: Dr. Paola Rebusco \& Prof. Edmund Bertschinger
@@
 
-\item[] {\bf WISTEM Program Mentor} \hfill {\bf September 2013 - present}
+\item[] {\bf WISTEM Program Mentor} \hfill {\bf September 2013 - May 2016}
 \begin{list1}
@@
 \begin{list2}
-\item[] {\bf Piso, A.-M. A.}, Pegues, J., \& \"Oberg, K.I. \textit{The Role of Ice Compositions and Morphology For Snowlines and the C/N/O Ratios in Active Disks}. ApJ, submitted \\
+\item[] {\bf Piso, A.-M. A.}, Pegues, J., \& \"Oberg, K.I. \textit{The Role of Ice Compositions and Morphology For Snowlines and the C/N/O Ratios in Active Disks}. ApJ, accepted: arXiv:1611.00741  \\
 \end{list2}
@@
 \begin{list2}
+\item[] {\bf Origins of Gas Giant Compositions: The Role of Disk Location and Dynamics} \\
+iPLEX Lunch Seminar, Los Angeles, CA, September 2016\\
+Invited talk \\
+\end{list2}
+
+\begin{list2}
 \item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
 MIT Planetary Lunch Colloquium Series, Cambridge, MA, March 2016\\
+Invited talk \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+CfA Small Scale Seminar, Cambridge, MA, February 2016\\
 Invited talk \\

### 2016-11-06 Postdoc Applications/uc_background_statement.tex
upd

@@
 
-I was born in Romania a couple of years before the 1989 revolution.
+I was born in Romania a few years before the 1989 revolution.
 As a woman growing up in a country that was not ready to promote values such as diversity, inclusion, tolerance, and equality between genders, I was not expected to pursue a scientific career.
 My family supported my choice, but society and the educational system did not.
-I moved to the United States to attend college with the expectation that the values and ethics that are often disregarded in my home country would be universally appreciated here.
+I moved to the United States to study Physics at MIT with the expectation that the values and ethics that are often disregarded in my home country would be universally appreciated here.
 Unfortunately, I found out that this is not always the case, especially regarding gender equality.
@@
 
-Throughout my graduate studies, I have often been a mentor, in various capacities, to younger women already pursuing a scientific career, or just interested in and excited about the possibility.
+Throughout my graduate studies at Harvard University, I have often been a mentor, in various capacities, to younger women already pursuing a scientific career, or just interested in and excited about the possibility.
 I was part of the Women In Science, Engineering, Technology and Mathematics (WISTEM) organization, a group that aims to foster a sense of community for women studying science and engineering at Harvard.

### 2016-11-06 Postdoc Applications/uc_research_statement.tex
upd

@@
 Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered [1].
-Their diversity in terms of mass, radius, location and composition [2] provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life.
+Their diversity in terms of mass, radius, location and composition [2] provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and that may sustain life.
 The advent of the Kepler Space Telescope has resulted in an exponential increase in the number of exoplanets detected via the transit method.
@@
 The majority of these detected Super Earths are very close to their host stars, with smaller orbital periods than the Earth.
-These features imply that Super Earths do not have Solar System analogs.
-At the same time, they are ubiquitously found and predicted to be the most common types of exoplanets [3], which makes understanding their ages, formation and properties particularly relevant.
+These features imply that Super Earths are not Solar System analogs.
+At the same time, they are ubiquitously found and predicted to be the most common types of exoplanets known to date [3], which makes understanding their ages, formation and properties particularly relevant.
 Observational techniques such as gravitational microlensing and direct imaging have extended the orbital range at which planets have been detected from just outside the Earth's orbit to tens and even hundreds of astronomical units.
-As demonstrated in Figure 2, however, most of these planets are massive; wide-separation Super Earths or Neptune analogs have yet to be detected.
+As demonstrated in Figure 2, however, most of these planets have Jupiter-like masses; true Uranus or Neptune analogs have yet to be detected.
 Are they indeed rare at these large distances, or just not bright or young enough to be detected?
@@
 Venus, Mars and Earth are displayed in the lower right corner.
-No Solar System Super Earth analogs have been yet detected.}
+The detected Super Earths have no Solar System analogs.}
 \label{fig:chemical}
@@
 Besides the terrestrial planets, Jupiter, Saturn, Uranus and Neptune are also shown.
-There is a scarcity in Super Earth detections at wide separateions.}
+There is currently a scarcity of exoplanets with Earth to Saturn Masses outside $\sim$5 AU.
+New direct imaging surveys are expected to fill in this phase space.}
 \label{fig:chemical}
@@
 \caption{Example of a planet's luminosity and time evolution with atmospheric mass, during the giant planet formation stage.
+The different colored curves correspond to different assumptions about the atmospheric opacity.
 From [6].}  %The actual magnitude of dust opacities is difficult to robustly predict.}
@@
 Moreover, these models consider arbitrary initial conditions, as they are only interested in studying the luminosities of mature, several billion years old exoplanets.
-\textbf{The combined results from parts 1 and 2 will represent the first time that a planet's luminosity evolution is tracked self consistently from its birth through after the disk dispersal.} These results will be crucial to understanding the properties of directly imaged young, bright planets.
-Furthermore, they may be able to help us distinguish between different formation scenarios Ð core accretion versus gravitational instability.
+\textbf{The combined results from parts 1 and 2 will represent the first time that a planet's luminosity evolution is tracked self consistently from its birth through after the disk dispersal.} These results will be crucial for determining the masses of directly imaged exoplanets.
+They will be essential for correctly extrapolating from the observations of young, and hence bright systems, to the underlying population of exoplanets that is dominated by older, and hence significantly less luminous planets that will remain undetectable.
+Furthermore, these results may be able to help us distinguish between different formation scenarios --- core accretion versus gravitational instability.
 
@@
 UCLA is also home to the Institute for Planets and Exoplanets (iPLEX), an interdisciplinary forum that ties together people with different scientific backgrounds.
-Outside research, I plan to be involved in the UCLA Social Postdoc Society.
+%Outside research, I plan to be involved in the UCLA Social Postdoc Society.
 Through this, I will help foster connections between postdoctoral scholars from various departments across the campus, with the goal of sharing our diverse scientific knowledge and experience.

### 2017-04-22 Postdoc Applications/AnaMariaPisoResume.tex
upd

@@
 
+\item[] {\bf Organizer of UCLA iPLEX Lunch Talks } \hfill {\bf April 2017 - present}
+\begin{list1}
+\item[] Coordinator of weekly seminars in the Institute for Planets and Exoplanets \hfill \textit{Los Angeles, CA}
+\end{list1}
+
+\vspace{0.2in}
+
+\item[] {\bf Judge in the Los Angeles Basin EPS Student Research Symposium } \hfill {\bf April 2017}
+\begin{list1}
+\item[] Provided evaluations and feedback on science presentations by graduate and undergraduate students in Earth and Planetary Science \hfill \textit{Los Angeles, CA}
+\end{list1}
+
+\vspace{0.2in}
+
 \item[] {\bf WISTEM Program Mentor} \hfill {\bf September 2013 - May 2016}
@@
 \begin{list2}
-\item[] {\bf Piso, A.-M. A.}, Pegues, J., \& \"Oberg, K.I. \textit{The Role of Ice Compositions and Morphology For Snowlines and the C/N/O Ratios in Active Disks}. ApJ, accepted: arXiv:1611.00741  \\
+\item[] {\bf Piso, A.-M. A.}, Pegues, J., \& \"Oberg, K.I. \textit{The Role of Ice Compositions and Morphology For Snowlines and the C/N/O Ratios in Active Disks}. ApJ, 2016, 833, 203  \\
 \end{list2}
@@
 
-\section{\sc Conferences and Seminars}
+\section{\sc Invited Talks}
+
+\begin{list2}
+\item[] {\bf The Role of Ice Compositions and Disk Dynamics for Snowlines and the C/N/O Ratios in Active Disks} \\
+AAS Inner Solar Systems Meeting, Austin, TX, June 2017\\
+Invited talk \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf Origins of Gas Giant Compositions: The Role of Disk Location and Dynamics} \\
+Caltech Planetary Science Seminar, Pasadena, CA, February 2017\\
+\end{list2}
 
@@
 iPLEX Lunch Seminar, Los Angeles, CA, September 2016\\
-Invited talk \\
 \end{list2}
@@
 MIT Planetary Lunch Colloquium Series, Cambridge, MA, March 2016\\
-Invited talk \\
 \end{list2}
@@
 CfA Small Scale Seminar, Cambridge, MA, February 2016\\
-Invited talk \\
 \end{list2}
+
+\begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+University of Michigan Astronomy Lunch Talk, Ann Arbor, MI, November 2015 \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+University of Chicago Exoplanet Journal Club, Chicago, IL, November 2015 \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+MIT Exoplanet Tea, Cambridge, MA, October 2015 \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
+Center for Integrative Planetary Science Planet and Star Formation Seminar, Berkeley, CA, September 2015 \\
+\end{list2}
+
+\vspace*{-.13in}
+\begin{list2}
+\item[] {\bf The Solar Wind}\\ 
+Vienna Theory Lunch Club, TU Wien, Vienna, Austria, June 2009  \\
+\end{list2}
+
+\section{\sc Contributed Talks and Seminars}
 
@@
 \begin{list2}
-\item[] {\bf C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion} \\
-Extreme Solar Systems III, Waikoloa Village, HI, December 2015 \\
-Poster \\
-\end{list2}
-
-\begin{list2}
-\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
-University of Michigan Astronomy Lunch Talk, Ann Arbor, MI, November 2015 \\
-Seminar speaker \\
-\end{list2}
-
-\begin{list2}
-\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
-University of Chicago Exoplanet Journal Club, Chicago, IL, November 2015 \\
-Seminar speaker \\
-\end{list2}
-
-\begin{list2}
-\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
-MIT Exoplanet Tea, Cambridge, MA, October 2015 \\
-Seminar speaker \\
-\end{list2}
-
-\begin{list2}
-\item[] {\bf Giant Planet Formation and Snowlines in Protoplanetary Disks} \\
-Center for Integrative Planetary Science Planet and Star Formation Seminar, Berkeley, CA, September 2015 \\
-Invited talk \\
-\end{list2}
-
-\begin{list2}
 \item[] {\bf Minimum Core Masses for Giant Planet Formation} \\
 CfA Exoplanet Pizza Lunch, Cambridge, MA, May 2015  \\
-Internal department talk \\
 \end{list2}
@@
 Star and Planet Formation in the Southwest, Oracle, AZ, March 2015  \\
-Contributed talk \\
 \end{list2}
@@
 CfA Exoplanet Pizza Lunch, Cambridge, MA, November 2013  \\
-Internal department talk \\
+\end{list2}
+
+\begin{list2}
+\item[] {\bf On the Minimum Core Mass for Giant Planet Formation} \\
+IAUS 299: Exploring the Formation and Evolution of Planetary Systems, Victoria, BC, June 2013  \\
+\end{list2}
+
+\section{\sc Posters}
+
+\begin{list2}
+\item[] {\bf C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion} \\
+Extreme Solar Systems III, Waikoloa Village, HI, December 2015 \\
 \end{list2}
@@
 Protostars and Planets VI, Heidelberg, Germany, July 2013  \\
-Poster \\
-\end{list2}
-
-\begin{list2}
-\item[] {\bf On the Minimum Core Mass for Giant Planet Formation} \\
-IAUS 299: Exploring the Formation and Evolution of Planetary Systems, Victoria, BC, June 2013  \\
-Contributed talk \\
 \end{list2}
@@
 Exoplanets in Multi-body Systems in the Kepler Era, Aspen, CO, February 2013 \\
-Poster \\
 \end{list2}
@@
 APS April Meeting, Washington, DC, February 2010  \\
-Poster \\
-\end{list2}
-
-\vspace*{-.13in}
-\begin{list2}
-\item[] {\bf The Solar Wind}\\ 
-Vienna Theory Lunch Club, TU Wien, Vienna, Austria, June 2009  \\
-Invited talk \\
 \end{list2}

### 2017-06-15 Postdoc Applications/AnaMariaPisoResume.tex
upd

@@
 AAS Inner Solar Systems Meeting, Austin, TX, June 2017\\
-Invited talk \\
 \end{list2}

### 2017-06-15 Postdoc Applications/proposal_draft.tex
upd

--- 2015-12-30 Postdoc Applications/research_proposal_hilke.tex
@@
 \pagestyle{fancy}
-\lhead{Ana-Maria Piso}
+\lhead{Dr. Ana-Maria Piso}
 \rhead{Postdoctoral Position Research Proposal}
@@
 The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets\vspace{-2ex}}
-\author{Ana-Maria Piso \vspace{-2ex}}
+\author{Dr. Ana-Maria Piso \vspace{-2ex}}
 
@@
 It follows that these two effects cannot be decoupled, and that the chemo-dynamical evolution of the disk has to be studied simultaneously and self-consistently.
-Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research $[7]$.
+Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research ($[7]$, $[8]$).
 Such effects include particle growth and fragmentation, as well as accretion rate and stellar luminosity evolution.
@@
 \caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height.
-The numbered curves are $\log_{10} (\tau)$ in million years (from $[11]$).}
+The numbered curves are $\log_{10} (\tau)$ in million years (from $[12]$).}
 \label{fig:chemical}
@@
 %\vspace{-0.5in}
-\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[8]$).}
+\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[9]$).}
 \vspace{-0.25in}
@@
 This will change a planet's atmospheric composition since the disk chemical abundances are different at different disk locations.
-Additionally, giant planets may still accumulate planetesimals while accreting nebular gas $[9]$.
+Additionally, giant planets may still accumulate planetesimals while accreting nebular gas $[10]$.
 The final composition of a planet's atmosphere will thus depend on how much gas and solids are accreted in this stage.
@@
 The results from parts 1 and 2 will feed into a \textbf{large planet synthesis model}, in which I will use a grid of different \textbf{initial disk and planetary embryo conditions}.
-I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[10]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations.
+I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[11]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations.
 To minimize computational expenses, I will build a program from bottom-up that includes select physical and chemical processes based on the local simulations from steps 1 and 2.
@@
 
-\textbf{MIT is the ideal place for me to pursue my postdoctoral work.} Its research facilities and academic excellence are unparalleled.
+%\textbf{MIT is the ideal place for me to pursue my postdoctoral work.} Its research facilities and academic excellence are unparalleled.
 I would love the opportunity to work with \textbf{Prof.
@@
 $[4]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ApJ, 814, 66 \\
-$[5]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21 \\
-$[6]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
-$[7]$ Piso, A.-M. A., \"Oberg, K. I. , Birnstiel, T., \& Murray-Clay, R. A. 2015, ApJ, 815, 109 \\
-$[8]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
-$[9]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
-$[10]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
-$[11]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25
+\textbf{$[5]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21} \\
+\textbf{$[6]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82} \\
+\textbf{$[7]$ Piso, A.-M. A., \"Oberg, K. I. , Birnstiel, T., \& Murray-Clay, R. A. 2015, ApJ, 815, 109} \\
+\textbf{$[8]$ Piso, A.-M. A,. Pegues, J., \& \"Oberg, K. I.  2016, ApJ, 833, 203} \\
+$[9]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
+$[10]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
+$[11]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
+$[12]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25
 %\end{thebibliography}

### 2017-06-22 Postdoc Applications/NPP_proposal.tex
upd

%\documentclass[12pt]{amsart}
\documentclass[12pt, letterpaper]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\usepackage[]{natbib}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage{titling}
\usepackage{sectsty}
\usepackage{sidecap}
\usepackage{placeins}
\usepackage{indentfirst}
\usepackage{wrapfig}

\usepackage{setspace}  % Needed for Double Spacing
\usepackage[english]{babel}
\usepackage{lipsum}

\pagestyle{fancy}
\lhead{Dr. Ana-Maria A. Piso}
\rhead{NPP Research Proposal}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{THE ROLE OF DISK VOLATILE CHEMISTRY AND DYNAMICS IN SHAPING THE COMPOSITIONS OF NASCENT PLANETS}  % add title here
\author{DR. ANA-MARIA A. PISO}  % include Author's name here
%\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle

\doublespacing

\noindent
\textbf{Title of Research Opportunity}: STAR AND PLANET FORMATION (18361)\\
\smallskip
\noindent
\textbf{NASA Center}: JET PROPULSION LABORATORY\\
\smallskip
\noindent
\textbf{NASA Advisor}: DR. NEAL TURNER \\
\smallskip

\begin{abstract}

%Type your Abstract here. The abstract should be a brief summary of the scientific problem to be addressed, your proposed research strategy, your methodology, and expected results.

The compositions of giant planets are determined by and tightly connected to the composition of the protoplanetary disk in which they are born. In turn, disks are affected by a multitude of chemical and dynamical processes that affect their composition both spatially and temporally. While the evolution and chemical composition of disks and gas giant atmospheres have been explored separately from the dynamical or chemical perspective, studies that investigate the coupled effect of chemistry and dynamics on the disk and planet evolution do not exist to date. I will develop a self-consistent chemo-dynamical model to explore how the disk dynamics and chemistry, as well as the dynamics of planetesimals and planetary embryos, affect the compositions of mature giant planets that we see today. Through analytical and numerical calculations, I will investigate a range of chemical and dynamical processes that may affect a disk's, and hence a planet's, composition at different locations in the disk midplane. I will then generalize the above results by inputting them into a large planet synthesis model, in which I will explore a wide range of initial disk and planetary embryo conditions. This will enable me to (1) predict the kinds of giant planet compositions that can be obtained from planet formation at different disk locations, and (2) back-track a planet's formation location based on its composition. 
\end{abstract}

\section{STATEMENT OF PROBLEM}

%State the scientific problem you plan to address.  Put the problem in context and discuss the importance of the problem and its relevance to NASA?s goals and missions specifically, and its relevance to the field in general.

Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered $[1]$. Their diversity in terms of mass, radius, location and composition $[2]$ provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore and understand how planets obtain their compositions. Observations of Earth-like planets that can provide useful insight about their composition are challenging --- the solid interior structure of terrestrial planets cannot be detected, and their gaseous envelopes are small by comparison (both in mass and radius), which makes it difficult to obtain atmospheric spectra and find out what chemical compounds they are made of. We therefore turn to giant planets, which have provided a rich and intriguing research area for decades. Gas giants contain most of their mass in their atmosphere, hence their chemical composition is determined by that of their envelopes. The last few years have seen a substantial increase in the number of giant planets with observed atmospheric spectra (e.g., $[3]$, $[4]$), which has enhanced our understanding of these planets' chemical structure, and has provided us with quantitative information about the abundances of various compounds in their envelopes besides hydrogen and helium. Finally, gas giants shape the architecture of planetary systems and affect the delivery of volatile compounds to terrestrial planets, which has direct consequences for the habitability of worlds similar to our own. Thus testing theories of planet formation against gas giant compositions will help constrain planet formation theories more generally.

Both terrestrial and giant planets are born in protoplanetary disks, which implies that their \textbf{compositions are determined by and tightly linked to the structure and composition of the disk}. The chemical and dynamical evolution of disks, as well the formation of giant planets have both been previously investigated in isolation. However, the coupled chemo-dynamical disk evolution, planet compositions, and most importantly the disk-planet connection have not yet been considered in detail. As shown in my work on the minimum core mass of gas giants ($[5]$, $[6]$), planet formation depends sensitively on disk physics and chemistry. \textbf{I propose to develop a holistic chemo-dynamical framework to explore how disk dynamics and chemistry, as well as the dynamics of nascent planets and planetesimals, regulate the compositions of mature giant planets.} Such a model will enhance our understanding of planetary structures by enabling us to \textbf{predict what kind of planet compositions result from planet formation in different parts of the disk}, as well as \textbf{constrain a planet's formation location based on its chemical composition}. %Furthermore, this work provides essential context for characterizing the gas giants that instruments such as the James Webb Space Telescope (JWST) and the Transiting Exoplanet Survey Satellite (TESS) will one day discover. %\textbf{My proposed work has therefore direct relevance to NASA's Exoplanet Exploration Program.}

\subsection*{Relevance to NASA's Missions}

The research outlined above is directly related to NASA's Exoplanet Exploration Program. Applying my proposed theoretical framework to a wide range of disk and planetary embryos initial conditions will provide essential context for characterizing the atmospheric compositions of gas giants that instruments such as the \textbf{James Webb Space Telescope (JWST)} and the \textbf{Wide Field Infrared Survey Telescope (WFIRST)} will discover in the near future. Moreover, my results will be useful in selecting the most promising giant planets, from a compositional point of view, for eventual follow-up with JWST. With my current theoretical expertise and the skills that I will gain as a NASA Postdoctoral Fellow, I am in the position of playing a leading role in interpreting the compositions of newfound giant planets, and in using planet compositions as tracers of planet formation locations.

\section{SCIENCE BACKGROUND}

%Discuss the background of the problem that you propose to address.  Review the history of the problem, discuss progress in this field, and discuss what is needed to further advance the field.  Include citations and references as necessary (you should use an accepted citation format relevant for your field).

The chemical composition of protoplanetary disks is largely determined by the freeze-out of volatile species. The snowline locations of volatile molecules are thus crucial in determining the chemical abundances in gas and dust throughout the disk, as well as planet compositions.

One important effect of the existence of snowlines is that disks are expected to contain different amounts of volatiles in gas and in dust at different locations. As the main carbon and oxygen carriers, i.e. H$_2$O, CO$_2$ and CO, are amongst the most abundant volatiles in comets and protostellar cores ([7], [8], [9]), variations in the carbon-to-oxygen (C/O) ratio in gas and dust throughout a disk are particularly important. This issue was first addressed by [10]. Figure \ref{fig:oberg} shows the C/O ratio in gas and dust as a function of semimajor axis, assuming protostellar abundances for the volatiles. The C/O ratio in gas is enhanced compared to the stellar value, particularly between the CO$_2$ and CO snowlines where it reaches unity.

\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{oberg11.png}
%\vspace{-0.5in}
\caption{The C/O ratio in gas (solid line) and dust (dashed line) as a function of semimajor axis in a static disk. The dotted line shows the stellar value of 0.54. Gas-phase C/O ratios of order unity can be achieved in the outer disk. From [10]. }
\label{fig:oberg}
\end{figure}

The pioneering work of [10], as well as the purported detection of a superstellar gas-phase C/O ratio in the atmosphere of WASP 12-b [11], inspired further theoretical investigations into the C/O ratio in disk and planet atmospheres. The model of [11] assumes a static disk, thus ignoring disk chemistry or dynamics. In reality, however, there are many chemical and dynamical processes that occur in disks (see [9] for a review), which will change the chemical molecular and elemental abundances, and therefore C/O ratios. For example, [12] calculate the C/O ratio throughout the disk including dynamical effects such as radial drift, grain coagulation, and the diffusion of volatile vapors. [13] include planetary migration in their C/O ratio calculation, and use the C/O ratio to constrain migration mechanisms. Additional carbon volatile carriers, such as CH$_4$, are considered by [14],  and they find that the gas-phase C/O ratio may be enriched by up to a factor of four compared to the Solar value in the outer disk.  While the C/O ratio is an important signature of disk and planet atmospheric chemistry, many other compounds have been detected in disks [9], which may affect the composition of disks and of nascent planets even when only dynamical processes are included and the chemistry is static, such as in the studies described above.

From the chemistry perspective, several studies have modeled the chemical evolution of disk volatile abundances using chemical networks with hundreds to thousands of reactions (see., e.g., [15] and references in [9]); however, these works do not consider any dynamical processes.

In my recent work ([16], [17]), I explored the effect of radial drift of solids and viscous gas accretion onto the central star, as well as different ice compositions, on snowline locations and the C/N/O ratios in disks. Besides C and O carriers, I also considered nitrogen-bearing species, such as N$_2$ and NH$_3$. Figure \ref{fig:CO_ratio} shows that the processes described above can change the location of the CO snowline by up to a factor of $\sim$7, thus introducing a large uncertainty in the position of the CO snowline, and therefore the disk locations where the gas-phase C/O ratio is enhanced compared to the stellar value. The same effect was observed for the N$_2$ snowline location.

\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{C_O_water_ice.pdf}
%\vspace{-0.5in}
\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO as pure ice (top panel) or as water dominated ices (bottom panel). The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar. The C/O ratio in a static disk (black lines) is shown for comparison. Taken together, disk dynamics and ice compositions move the CO snowline inward by a factor of $\sim$7. From [17].} 
\label{fig:CO_ratio}
\end{figure}

It it thus clear that, in order to better understand and constrain disk and planet compositions, as well as use planet compositions as tracers of planet formation locations, the following are necessary: (1) the consideration of additional dynamical effects, (2) a self-consistent coupling of disk chemistry and dynamics, (3) the inclusion of planet dynamical effects in order to explore the disk-planet connection, and (4) applying (1), (2) and (3) to a wide parameter space of disk and planetary embryos initial conditions.

\section{GENERAL METHODOLOGY}

%Discuss the strategies and methods you will use. Describe why you chose your particular plan, and discuss how this plan improves upon previous work in this field.

\subsection{Coupled Chemical and Dynamical Disk Evolution} 
%\underline{\textbf{1. Coupled Chemical and Dynamical Disk Evolution}}

Chemical and dynamical processes in a protoplanetary disk affect the disk structure and composition, and thus the composition of nascent planets. As shown in Figure \ref{fig:chemical}, the timescales for various disk chemical and dynamical processes may be comparable at least in some parts of the disk. It follows that these two effects cannot be decoupled, and that the chemo-dynamical evolution of the disk has to be studied simultaneously and self-consistently. Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research ($[16]$, [17]). Such effects include particle growth and fragmentation, as well as accretion rate and stellar luminosity evolution.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chemical_timescales}
\vspace{-0.3in}
\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height. The numbered curves are $\log_{10} (\tau)$ in years (from $[20]$).}
\label{fig:chemical}
\end{figure}

From the chemistry perspective, molecular abundances vary significantly across a typical disk, due to steep gradients in temperature, density and radiation. 
%Figure 1 shows a theoretical example of how both changes in disk temperature, as well as time evolution, decrease the abundance of carbon monoxide (CO) by several orders of magnitude \cite{aikawa96}. 
The complexity of disk chemistry means that coupling it with dynamical processes, while necessary, is non-trivial.  I will couple the dynamical framework outlined above with \textbf{time-dependent chemical models of increasing complexity}, informed by results from state-of-the-art disk chemistry models (that can only be run on static disks). This will show how the snowline locations of volatiles, as well as the chemical composition of the disk gas and dust evolve, which has \textbf{direct implications on the compositions of young planets}. Moreover, since I am primarily interested in processes that affect the disk mid-plane where planets form, my results could be simplified by the fact that certain chemical processes may not be important in this disk parameter space. This step should be completed within 8 months.

%this dynamical model  develop a (chemical models of increasing complexity) simplified time-dependent chemistry,

%\vspace{0.2in}
\subsection{Planet and Planetesimal Migration}
%\underline{\textbf{2. Planet and Planetesimal Migration}}

The complex processes outlined in part 1 will directly impact the composition and dynamics of forming gas giants, as the latter are born and evolve simultaneously with the disk. 
\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{HLTau_nrao}
%\vspace{-0.5in}
\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[18]$).}
\label{fig:HLtau}
\end{figure}
The chemical evolution of the gas disk molecular abundances, as well as disk dynamics, determine how a young planet's atmospheric composition changes in time and with the planet's location, and therefore the final chemical structure of mature planets. \textbf{It is thus clear that disks and planets are deeply connected, and that this relation needs to be explored}. %I will add planetary embryos in the code developed in step 1, and

Giant planets can migrate through the disk while still accumulating gas. Figure \ref{fig:HLtau} shows an exciting ALMA observation of a planetary gap in the disk around HL Tau $[18]$. This will change a planet's atmospheric composition since the disk chemical abundances are different at different disk locations. Additionally, giant planets may still accumulate planetesimals while accreting nebular gas $[10]$. The final composition of a planet's atmosphere will thus depend on how much gas and solids are accreted in this stage. \textbf{I will add planet dynamical effects such as migration and planetesimal accretion in the chemical and dynamical model developed in part 1, and quantify how these processes affect the chemical composition of gas giant envelopes}. This step should be completed within 8 months.

%\vspace{0.2in}

\subsection{Model Planet Populations} 
%\underline{\textbf{3. Model Planet Populations}}

The results from parts 3.1 and 3.2 will feed into a \textbf{large planet synthesis model}, in which I will use a grid of different \textbf{initial disk and planetary embryo conditions}. I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[19]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations. To minimize computational expenses, I will build a program from bottom-up that includes select physical and chemical processes based on the local simulations from steps 1 and 2. This will allow me to predict the \textbf{range of planet compositions that can form under reasonable disk conditions}, as well as \textbf{constrain a planet's formation location based on its chemical composition}. %Comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[4]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets. 
This step should be completed within 8 months.

\section{EXPLANATION OF NEW TECHNIQUES}

%Discuss in detail how your methods, plans and techniques are new or novel, and what advantages and disadvantages they have compared to previous work.

\subsection{Chemical Modeling}

Current state-of-the-art chemical networks include hundreds of molecules and reactions. The majority of these models, however, are not specifically focused on the disk mid-plane, which is the relevant parameter space for planet formation and thus for my goals. Moreover, as discussed in Section 3.1, some or many of the processes that these networks model may not penetrate all the way to the cold disk mid-plane. \textbf{I will select only the chemical processes that I deem suitable for the disk mid-plane, thus potentially greatly reducing the number of reactions and reactants}. This will result in \textbf{significantly lower computational times}. Furthermore, I will create a table of reactions and compounds relevant only for the disk mid-plane \textbf{and make it publicly available} as a useful tool for other astronomers interested in this chemical parameter space.

\subsection{Disk Dynamical Processes}

Most studies that explore the effect of disk dynamics on disk/planet compositions and the C/O ratio (e.g., [12], [14]) study the various dynamical effects they include in their respective models simultaneously. Thus the relative importance and impact of each dynamical process on their final results is often not clear. In contrast, \textbf{I will incorporate dynamical processes in my model gradually, and at each step quantify the effect of a given process}. This method will enable me to (1) \textbf{include only the dynamical processes that I find important in my planet synthesis model} (see Section 3.3), and (2) \textbf{more easily trace a planet's formation location based on its composition} after the simulation described in Section 3.3, \textbf{as I will know the importance of each dynamical effect} that will have led to a given planet composition.

\subsection{Planet Synthesis Model}

As discussed in Section 3.3, I will build an \textbf{original and novel planet synthesis code}, which I will tailor to only include the chemical and dynamical processes that I find important for my science problem (see Sections 3.1., 3.2, 4.1., 4.2). In comparison with existing planet population synthesis programs, my model will have the advantage of \textbf{greatly decreasing computational costs}.

\section{EXPECTED RESULTS}

%Discuss the results you expect to achieve and how these results will influence the development of the field of research regarding your science problem.

My expected results can be summarized as follows:

1. A clear understanding of the chemical processes and molecular compounds that are relevant in the disk mid-plane (i.e. the planet formation region), as well as the expected molecular and elemental abundances. This will be important for further theoretical work in this field, as well as for both predicting and interpreting current and future observations of gas giant atmospheric spectra.

2. Further insight into disk and planetary dynamical processes that affect the locations of volatile snowlines and the elemental ratios in disks and planet atmospheres. This, and the result outlined in point 1, will substantially improve our current understanding of how giant planets obtain their compositions.

3. The results of my planet synthesis study will provide a robust framework of using a planet's composition as a tracer of its formation location. Furthermore, comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[4]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.

\begin{figure}[h!]
\centering
\vspace{-0.2in}
\includegraphics[width=0.9\textwidth]{water_spectrum}
\vspace{-0.2in}
\caption{The transmission spectrum of exoplanet WASP-12b measured with the Hubble Space Telescope/Wide Field Camera 3. The white dots are the measured transit depths, the blue squares and dark blue line are the best fit model, while the shaded regions represent 1- and 2 $\sigma$ intervals in the retrieved spectrum. The increase in transit depth around $\sim$1.4 $\mu$m is evidence for a water feature in the atmosphere (from $[4]$).}
\label{fig:HLTau}
\end{figure}

%\textbf{MIT is the ideal place for me to pursue my postdoctoral work.} Its research facilities and academic excellence are unparalleled. I would love the opportunity to work with \textbf{Prof. Hilke Schlichting}, an expert in planet formation theory and dynamics --- such a collaboration would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards. I would also like to collaborate with \textbf{Prof. Sara Seager}, a co-investigator of the TESS mission and a leading theorist in atmospheric chemistry, as well as with members of her group, who specialize in a broad range of exoplanet theory and computational topics. \textbf{Dr. Margaret Pan}, an excellent theorist in planetary dynamics and protoplanetary disks, will join the MIT EAPS department in early 2016 --- it would be a pleasure to work with her as well. Additionally, the MIT Physics department hosts leaders in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof. Joshua Winn}, which presents great prospects in connecting my theoretical research work with observations. \\

%Finally, in order to connect my theoretical work with observations, I would like to collaborate with \textbf{Prof. Joshua Winn}, who is an expert in discovering and characterizing exoplanets. \\

%\textbf{The University of Chicago is the ideal place for me to pursue my postdoctoral research}, due to its vibrant community of experts in protoplanetary disks and exoplanets, both in the Department of Astronomy and Department of Geophysical Sciences. I would love the opportunity to work with \textbf{Prof. Leslie Rogers}, an expert in planetary interiors and atmospheres --- such a collaboration would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards. In the Department of Geophysical Sciences, I would like to collaborate with \textbf{Prof. Fred Ciesla}, a leader in protoplanetary disk dynamics, chemical composition and evolution. Additionally, the Department of Astronomy hosts experts in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof. Dan Fabrycky} and \textbf{Prof. Jacob Bean}, which presents great prospects in connecting my theoretical research work with observations. \\

%\textbf{Fred Ciesla} (proposed faculty host) is an expert in protoplanetary disk dynamics, as well as chemical composition and evolution. Leslie Rogers, who will begin her faculty appointment in Fall 2016, is a leader in exoplanet theory, specifically planetary interiors and atmospheres, so I would love the opportunity to collaborate with her on my postdoctoral work. Additionally, the University of Chicago Department of Astronomy hosts leaders in detecting and characterizing worlds outside the Solar system, such as Dan Fabrycky and Jacob Bean, which presents great prospects in connecting my theoretical research work with observations.

%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.7\textwidth]{CO_abundances}
%%\vspace{-0.5in}
%\caption{TBD}
%\label{fig:CO}
%\end{figure}

%\if\bibinc n
%\bibliography{refs}
%\fi
\FloatBarrier
%\def\bibfont{\footnotesize}
%\setlength{\bibsep}{0.0pt}

%\section*{References}
\section{REFERENCES}
\singlespacing
%\begin{thebibliography} {9}
\noindent $[1]$ Batalha, N. M. 2014, Proceedings of the National Academy of Science, 111, 12647 \\
$[2]$ Lissauer J. J., Dawson R. I., \&Tremaine S., 2014, Nature, 513, 336 \\
$[3]$ Debes, J. H., Jang-Condell, H., Weinberger, A. J., Roberge, A., \& Schneider, G. 2013, ApJ, 771, 45 \\
$[4]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ApJ, 814, 66 \\
\textbf{$[5]$ Piso, A.-M. A., \& Youdin, A. N. Y. 2014, ApJ, 786, 21} \\
\textbf{$[6]$ Piso, A.-M. A., Youdin, A. N. Y., \& Murray-Clay, R. A. 2015, ApJ, 800, 82} \\
$[7]$ Rodgers, S. D., \& Charnley, S. B. 2002, MNRAS, 330, 660 \\
$[8]$ Mumma, M. J., \& Charnley, S. B. 2011, ARA\&A, 49, 471 \\
$[9]$ Henning, T., \& Semenov, D. 2013, Chemical Reviews, 113, 9016 \\
$[10]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
$[11]$ Madhusudhan, N., Harrington, J., Stevenson, K. B., et al. 2011, Nature, 469, 64 \\
$[12]$ Ali-Dib, M., Mousis, O., Petit, J.-M., \& Lunine, J. I. 2014, ApJ, 785, 125 \\
$[13]$ Madhusudhan, N., Amin, M. A., \& Kennedy, G. M. 2014, ApJL, 794, L12 \\
$[14]$ Thiabaud, A., Marboeuf, U., Alibert, Y., Leya, I., \& Mezger, K. 2015, A\&A, 574, A138 \\
$[15]$ Visser, R., van Dishoeck, E. F., Doty, S. D., \& Dullemond, C. P. 2009, A\&A, 495, 881 \\
\textbf{$[16]$ Piso, A.-M. A., \"Oberg, K. I. , Birnstiel, T., \& Murray-Clay, R. A. 2015, ApJ, 815, 109} \\
\textbf{$[17]$ Piso, A.-M. A,. Pegues, J., \& \"Oberg, K. I.  2016, ApJ, 833, 203} \\
$[18]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
$[19]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
$[20]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25

%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=2in]{example.jpg} 
%   \caption{example caption}
%   \label{fig:example}
%\end{figure}

\end{document}

### 2017-06-30 Postdoc Applications/NPP_proposal.tex
upd

@@
 \documentclass[12pt, letterpaper]{article}
-\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
-\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
+%\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
+%\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
 %\geometry{landscape}                % Activate for for rotated page geometry
@@
 
-\usepackage[]{natbib}
+\usepackage[top=1.1in, bottom=1.1in, left=1.4in, right=1.4in]{geometry}
+\usepackage{natbib}
 \usepackage{amssymb}
-\usepackage{natbib}
+%\usepackage{natbib}
 \usepackage{color}
@@
 \centering
-\includegraphics[width=0.6\textwidth]{oberg11.png}
+\includegraphics[width=0.7\textwidth]{oberg11.png}
 %\vspace{-0.5in}
@@
 
-From the chemistry perspective, several studies have modeled the chemical evolution of disk volatile abundances using chemical networks with hundreds to thousands of reactions (see., e.g., [15] and references in [9]); however, these works do not consider any dynamical processes.
+From the chemistry perspective, several studies have modeled the chemical evolution of disk volatile abundances using chemical networks with hundreds to thousands of reactions (see., e.g., [15], [16] and references in [9]); however, most of these works do not consider any dynamical processes.
 
-In my recent work ([16], [17]), I explored the effect of radial drift of solids and viscous gas accretion onto the central star, as well as different ice compositions, on snowline locations and the C/N/O ratios in disks.
+In my recent work ([17], [18]), I explored the effect of radial drift of solids and viscous gas accretion onto the central star, as well as different ice compositions, on snowline locations and the C/N/O ratios in disks.
 Besides C and O carriers, I also considered nitrogen-bearing species, such as N$_2$ and NH$_3$.
@@
 \centering
-\includegraphics[width=0.6\textwidth]{C_O_water_ice.pdf}
+\includegraphics[width=0.8\textwidth]{C_O_water_ice.pdf}
 %\vspace{-0.5in}
@@
 
-Chemical and dynamical processes in a protoplanetary disk affect the disk structure and composition, and thus the composition of nascent planets.
+Protoplanetary disks experience a multitude of chemical and dynamical processes, as shown schematically in Figure \ref{fig:disk}.
+These processes affect the disk structure and composition, and thus the composition of nascent planets.
 As shown in Figure \ref{fig:chemical}, the timescales for various disk chemical and dynamical processes may be comparable at least in some parts of the disk.
 It follows that these two effects cannot be decoupled, and that the chemo-dynamical evolution of the disk has to be studied simultaneously and self-consistently.
-Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research ($[16]$, [17]).
+Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research ($[17]$, [18]).
 Such effects include particle growth and fragmentation, as well as accretion rate and stellar luminosity evolution.
@@
 \centering
-\includegraphics[width=\textwidth]{chemical_timescales}
-\vspace{-0.3in}
-\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height.
-The numbered curves are $\log_{10} (\tau)$ in years (from $[20]$).}
-\label{fig:chemical}
+\includegraphics[width=\textwidth]{disk}
+\vspace{-0.1in}
+\caption{Sketch of chemical and dynamical processes in a protoplanetary disk around a Sun-like star (from $[9]$).}
+\label{fig:disk}
 \end{figure}
@@
 Moreover, since I am primarily interested in processes that affect the disk mid-plane where planets form, my results could be simplified by the fact that certain chemical processes may not be important in this disk parameter space.
-This step should be completed within 8 months.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chemical_timescales}
+\vspace{-0.3in}
+\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height.
+The numbered curves are $\log_{10} (\tau)$ in years (from $[28]$).}
+\label{fig:chemical}
+\end{figure}
+
+\textbf{My results from this step will be useful in interpreting JWST observations of volatile abundances and ratios in protoplanetary disk atmospheres.} The JWST \textbf{Mid-Infrared Instrument (MIRI)} (see, e.g., [19] for an overview) will enable abundance measurements of species such as H$_2$O, CO$_2$, C$_2$H$_2$, HCN in the atmospheres of a much broader range of disks that it was possible with the Spitzer Space Telescope (e.g., [20]).
+\textbf{Varying disk initial conditions, such as the temperature profile and initial species abundances}, and comparing them with the known properties of disks that JWST will observe, will help me in explaining \textbf{volatile ratios from the JWST data.} This step should be completed within 8 months.
 
@@
 %\vspace{-0.5in}
-\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[18]$).}
+\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[26]$).}
 \label{fig:HLtau}
@@
 Giant planets can migrate through the disk while still accumulating gas.
-Figure \ref{fig:HLtau} shows an exciting ALMA observation of a planetary gap in the disk around HL Tau $[18]$.
+Figure \ref{fig:HLtau} shows an exciting ALMA observation of a planetary gap in the disk around HL Tau $[26]$.
 This will change a planet's atmospheric composition since the disk chemical abundances are different at different disk locations.
@@
 The results from parts 3.1 and 3.2 will feed into a \textbf{large planet synthesis model}, in which I will use a grid of different \textbf{initial disk and planetary embryo conditions}.
-I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[19]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations.
+I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[27]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations.
 To minimize computational expenses, I will build a program from bottom-up that includes select physical and chemical processes based on the local simulations from steps 1 and 2.
-This will allow me to predict the \textbf{range of planet compositions that can form under reasonable disk conditions}, as well as \textbf{constrain a planet's formation location based on its chemical composition}.
+
+Based on the framework that I developed during my dissertation ([5], [6]), I will use the \textbf{core accretion model} for the formation of gas giants.
+The growth rate of giant nascent planets before the dispersal of the protoplanetary disk is highly dependent on the opacity of their envelopes.
+However, disk and atmosphere opacities are highly uncertain.
+I will therefore explore \textbf{a range of opacity models appropriate for different disk regions and temperatures, considering both dusty and grain-free envelopes}.
+(e.g., [21], [22], [23]).
+Since I am interested in the chemical composition of both nascent and mature giant planets, \textbf{I will carry my planet synthesis model forward through the dispersal of the gas disk}.
+Physical processes to be considered in this evolutionary stage are, for example, \textbf{spontaneous mass loss} due to loss of pressure support from the disk [24], or \textbf{dynamical interactions with forming terrestrial planets.}  Through numerical calculations, I will expand the analytic model of [24] for atmospheric mass loss.
+This will enable me to obtain \textbf{robust quantitative results} for the physical properties of gas giants that we see today, as well as \textbf{explore a broader range of disk and planet parameter space}.
+The latter will be particularly relevant for connecting my predictions with JWST detections, which will observe giant planets with a variety of masses, semimajor axes, and compositions.
+
+To conclude, my proposed planet synthesis model will allow me to predict the \textbf{range of planet compositions that can form under reasonable disk conditions}, as well as \textbf{constrain a planet's formation location based on its chemical composition}.
 %Comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[4]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.
@@
 In comparison with existing planet population synthesis programs, my model will have the advantage of \textbf{greatly decreasing computational costs}.
+Moreover, my complex chemo-dynamical planet synthesis model and code will serve as a useful \textbf{input of young model planets for subsequent atmospheric evolution codes}, such as is the case, for example, with the code developed by [25].
 
@@
 \vspace{-0.2in}
-\includegraphics[width=0.9\textwidth]{water_spectrum}
+\includegraphics[width=\textwidth]{water_spectrum}
 \vspace{-0.2in}
@@
 $[15]$ Visser, R., van Dishoeck, E. F., Doty, S. D., \& Dullemond, C. P. 2009, A\&A, 495, 881 \\
-\textbf{$[16]$ Piso, A.-M. A., \"Oberg, K. I. , Birnstiel, T., \& Murray-Clay, R. A. 2015, ApJ, 815, 109} \\
-\textbf{$[17]$ Piso, A.-M. A,. Pegues, J., \& \"Oberg, K. I.  2016, ApJ, 833, 203} \\
-$[18]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
-$[19]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
-$[20]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25
+$[16]$ Willacy, K., \& Woods, P. M. 2009, ApJ, 703, 479 \\
+\textbf{$[17]$ Piso, A.-M. A., \"Oberg, K. I. , Birnstiel, T., \& Murray-Clay, R. A. 2015, ApJ, 815, 109} \\
+\textbf{$[18]$ Piso, A.-M. A,. Pegues, J., \& \"Oberg, K. I.  2016, ApJ, 833, 203} \\
+$[19]$ Wright, G. S., Wright, D., Goodson, G. B., et al. 2015, PASP, 127, 595 \\
+$[20]$ Carr, J. S., \& Najita, J. R. 2011, ApJ, 733, 102 \\
+$[21]$ D`Alessio, P., Calvet, N., \& Hartmann, L. 2001, ApJ, 553, 321 \\
+$[22]$ Ferguson, J. W., Alexander, D. R., Allard, F., et al. 2005, ApJ, 623, 585 \\
+$[23]$ Freedman, R. S., Marley, M. S., \& Lodders, K. 2008, ApJS, 174, 504 \\
+$[24]$ Ginzburg, S., Schlichting, H. E., \& Sari, R. 2016, ApJ, 825, 29 \\
+$[25]$ Hu, R., Seager, S., \& Bains, W. 2012, ApJ, 761, 166 \\
+$[26]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
+$[27]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
+$[28]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25
 

### 2017-07-01 Postdoc Applications/NPP_proposal.tex
upd

@@
 The pioneering work of [10], as well as the purported detection of a superstellar gas-phase C/O ratio in the atmosphere of WASP 12-b [11], inspired further theoretical investigations into the C/O ratio in disk and planet atmospheres.
-The model of [11] assumes a static disk, thus ignoring disk chemistry or dynamics.
+The model of [10] assumes a static disk, thus ignoring disk chemistry or dynamics.
 In reality, however, there are many chemical and dynamical processes that occur in disks (see [9] for a review), which will change the chemical molecular and elemental abundances, and therefore C/O ratios.
@@
 Taken together, disk dynamics and ice compositions move the CO snowline inward by a factor of $\sim$7.
-From [17].}
+From [18].}
 \label{fig:CO_ratio}
@@
 The results of my planet synthesis study will provide a robust framework of using a planet's composition as a tracer of its formation location.
-Furthermore, comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[4]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.
+Furthermore, comparing my results with current observations of atmospheric spectra (e.g., Figure 6), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.
 
@@
 The white dots are the measured transit depths, the blue squares and dark blue line are the best fit model, while the shaded regions represent 1- and 2 $\sigma$ intervals in the retrieved spectrum.
-The increase in transit depth around $\sim$1.4 $\mu$m is evidence for a water feature in the atmosphere (from $[4]$).}
+The increase in transit depth around $\sim$1.4 $\mu$m is evidence for a water feature in the atmosphere.
+From $[4]$.}
 \label{fig:HLTau}
